### 2017-12-15 derivations.tex
start working on final derivation

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb}
\usepackage[
    backend=biber,
    style=numeric,
    citestyle=numeric
]{biblatex}
\addbibresource{derivations.bib}
\usepackage[breaklinks=true, linkcolor=blue, citecolor=blue, colorlinks=true]{hyperref}
% multiple eq references
\usepackage{cleverref}

\newcommand{\pluseq}{\mathrel{{+}{=}}}
\newcommand{\minuseq}{\mathrel{{-}{=}}}
\newcommand{\ns}{\ensuremath{{N_{sp}}}}
\newcommand{\conp}{constant-pressure}
\newcommand{\conv}{constant-volume}
\newcommand{\dconp}{\ensuremath{,\qquad\text{for \conp}}}
\newcommand{\dconv}{\ensuremath{,\qquad\text{for \conv}}}
\newcommand{\Ru}{\ensuremath{\mathcal{R}}}

\begin{document}
\section{Introduction}
This document contains the complete derivations for \texttt{pyJac} v2.0, which generates code to evaluate the chemical source term and analytical chemical kinetic Jacobians of \conp\slash\conv, fixed-mass, adiabatic reactors.

\section{State variables}

\begin{subequations}
\begin{align}
\Phi &= \left\{T, V, n_1, n_2 \ldots n_{\ns - 1}\right\}\dconp, \\
\Phi &= \left\{T, P, n_1, n_2 \ldots n_{\ns - 1}\right\}\dconv
\end{align}
\end{subequations}
where $T$ is the temperature, $P$ and $V$ the pressure and volume respectively, and $n_j$ the number of moles of the $j$th species in the model (containing $\ns$ total species).

From the ideal gas law, the number of moles of the final species is determined as:
\begin{subequations}
\begin{align}
n &= \frac{V P}{T \Ru} = \sum_{i=1}^{\ns}{n_i}, \label{source:moles}\\
n_{\ns} &= \frac{V P}{T \Ru} - \sum_{i=1}^{\ns - 1}{n_i}
\end{align}
\end{subequations}

\section{Chemical source terms}
The chemical source terms for this system are:
\begin{subequations}
\begin{align}
\frac{\text{d} \Phi }{\text{d} t } &= \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} V }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\}\dconp, \\
\frac{\text{d} \Phi }{\text{d} t } &= \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} P }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\}\dconv 
\end{align}
\end{subequations}
For both, the molar source terms are~\cite{TurnsStephenR2012Aitc}:
\begin{equation}
\frac{\text{d} n }{\text{d} t }_{k} = V \dot{\omega}_{k}
\label{source:spec}
\end{equation}
where $\dot{\omega}_k$ is the $kth$ species overall production rate, and the temperature production rates are~\cite{TurnsStephenR2012Aitc}:
\begin{subequations}
\label{source:temperature_incomplete}
\begin{align}
\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} H_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{p, k}}}\dconp, \\
\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} U_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{v, k}}}\dconv
\end{align}
\end{subequations}
where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure specific heat, constant-volume specific heat of species $k$ in molar units, while $[C]_{k}$ is the concentration.

From the ideal gas law, source terms for the volume and pressure variables may derived:
\begin{subequations}
\label{source:param_incomplete}
\begin{align}
\frac{\text{d} V }{\text{d} t } &= \frac{R_u}{P} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right)\dconp, \\
\frac{\text{d} P }{\text{d} t } &= \frac{R_u}{V} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right)\dconv
\end{align}
\end{subequations}

Finally, the conservation of mass is invoked:
\begin{equation*}
 m = \sum_{k=1}^{\ns} W_{k} n_{k},
\end{equation*}
to determine the source term of the last species:
\begin{equation}
 \frac{\text{d} n }{\text{d} t }_{\ns} = - \frac{1}{W_{\ns}} \sum_{k=1}^{\ns - 1} W_{k} \frac{\text{d} n }{\text{d} t }_{k}
 \label{source:spec_ns}
\end{equation}
where $W_{k}$ is the molecular weight of the $k$th species.
Using~\eqref{source:spec_ns}, the total molar rate of change can be determined:
\begin{align}
\frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns} \frac{\text{d} n }{\text{d} t }_{k}, \nonumber \\
\frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\text{d} n }{\text{d} t }_{k}
\label{source:total_molar}
\end{align}

Combining~\cref{source:param_incomplete,source:total_molar,source:spec} gives the final form of the pressure and volume terms:
\begin{subequations}
\label{source:param_complete}
\begin{align}
\frac{\text{d} V }{\text{d} t } &= V \left(\frac{T R_u}{P} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right)\dconp, \\
\frac{\text{d} P }{\text{d} t } &= \frac{P}{T} \frac{\text{d} T }{\text{d} t } + T R_u \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k}\dconv
\end{align}
\end{subequations}

\printbibliography

\end{document}

### 2017-12-15 derivations.tex
update temperature

@@
 \usepackage[breaklinks=true, linkcolor=blue, citecolor=blue, colorlinks=true]{hyperref}
-% multiple eq references
-\usepackage{cleverref}
 
@@
 \newcommand{\ns}{\ensuremath{{N_{sp}}}}
-\newcommand{\conp}{constant-pressure}
-\newcommand{\conv}{constant-volume}
+\newcommand{\conp}{CONP}
+\newcommand{\conv}{CONV}
 \newcommand{\dconp}{\ensuremath{,\qquad\text{for \conp}}}
@@
 This document contains the complete derivations for \texttt{pyJac} v2.0, which generates code to evaluate the chemical source term and analytical chemical kinetic Jacobians of \conp\slash\conv, fixed-mass, adiabatic reactors.
+Note that equations specific to the constant-pressure or constant-volume derivation will be marked with \conp or \conv respectively
 
@@
 
-Combining~\cref{source:param_incomplete,source:total_molar,source:spec} gives the final form of the pressure and volume terms:
+Combining~\eqref{source:spec},\eqref{source:param_incomplete},\eqref{source:total_molar} gives the final form of the pressure and volume terms:
 \begin{subequations}
@@
 \begin{align}
-\frac{\text{d} V }{\text{d} t } &= V \left(\frac{T R_u}{P} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right)\dconp, \\
-\frac{\text{d} P }{\text{d} t } &= \frac{P}{T} \frac{\text{d} T }{\text{d} t } + T R_u \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k}\dconv
+\frac{\text{d} V }{\text{d} t } &= V \left(\frac{T R_u}{P} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right)&\dconp, \\
+\frac{\text{d} P }{\text{d} t } &= \frac{P}{T} \frac{\text{d} T }{\text{d} t } + T R_u \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k}&\dconv
+\end{align}
+\end{subequations}
+Additionally, the temperature source terms may be expanded to remove references to the last species' source term and overall production rate:
+\begin{subequations}
+\label{source:temperature_complete}
+\begin{align}
+\frac{\text{d} T }{\text{d} t } = - \frac{\sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{p,\ns}} + \sum_{k=1}^{-1 + \ns} \left(- {C_{p,\ns}} + {C_{p, k}}\right) [C]_{k}}\dconp,\\
+\frac{\text{d} T }{\text{d} t } = - \frac{\sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{v,\ns}} + \sum_{k=1}^{-1 + \ns} \left(- {C_{v,\ns}} + {C_{v, k}}\right) [C]_{k}}\dconv
 \end{align}

### 2017-12-18 derivations.tex
define C and fix Ru
update

@@
 \newcommand{\ns}{\ensuremath{{N_{sp}}}}
+\newcommand{\nr}{\ensuremath{{N_{reac}}}}
 \newcommand{\conp}{CONP}
@@
 \section{Introduction}
-This document contains the complete derivations for \texttt{pyJac} v2.0, which generates code to evaluate the chemical source term and analytical chemical kinetic Jacobians of \conp\slash\conv, fixed-mass, adiabatic reactors.
-Note that equations specific to the constant-pressure or constant-volume derivation will be marked with \conp or \conv respectively
+This document contains the complete derivations for \texttt{pyJac} v2.0, which generates code to evaluate the chemical source term and analytical chemical kinetic Jacobians of constant-pressure\slash constant-volume, fixed-mass, adiabatic reactors.
+Note that equations specific to the constant-pressure or constant-volume derivation will be marked with \conp~or \conv~respectively.
+The derivations in this script based upon the output of the script \href{https://github.com/arghdos/SPyJac-paper/blob/master/derivations/scripts/derivations.py}{derivations.py} in the GitHub repository for this paper.
+This script depends on the symbolic mathematics library SymPy~\cite{sympy}, and was tested with versions 1.0 and 1.1.1
 
@@
 \begin{align}
-\frac{\text{d} V }{\text{d} t } &= \frac{R_u}{P} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right)\dconp, \\
-\frac{\text{d} P }{\text{d} t } &= \frac{R_u}{V} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right)\dconv
+\frac{\text{d} V }{\text{d} t } &= \frac{\Ru}{P} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right)\dconp, \\
+\frac{\text{d} P }{\text{d} t } &= \frac{\Ru}{V} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right)\dconv
 \end{align}
@@
 Finally, the conservation of mass is invoked:
-\begin{equation*}
- m = \sum_{k=1}^{\ns} W_{k} n_{k},
-\end{equation*}
+\begin{align*}
+ \frac{\text{d} m }{\text{d} t } &= 0 \\
+ &= \sum_{k=1}^{\ns}  W_{k} \frac{\text{d} n }{\text{d} t }_{k},
+\end{align*}
 to determine the source term of the last species:
@@
 \begin{align}
-\frac{\text{d} V }{\text{d} t } &= V \left(\frac{T R_u}{P} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right)&\dconp, \\
-\frac{\text{d} P }{\text{d} t } &= \frac{P}{T} \frac{\text{d} T }{\text{d} t } + T R_u \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k}&\dconv
+\frac{\text{d} V }{\text{d} t } &= V \left(\frac{T \Ru}{P} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right)&\dconp, \\
+\frac{\text{d} P }{\text{d} t } &= \frac{P}{T} \frac{\text{d} T }{\text{d} t } + T \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k}&\dconv
 \end{align}
 \end{subequations}
-Additionally, the temperature source terms may be expanded to remove references to the last species' source term and overall production rate:
+Additionally, the temperature source terms may be expanded to contain only state variables (i.e., by removal of the last species' source term and overall production rate):
 \begin{subequations}
@@
 \end{subequations}
+where $[C]$ is the total concentration:
+\begin{equation}
+ [C] = \frac{P}{T \Ru}
+\end{equation}
+This form is used for differentiation, but~\eqref{source:temperature_incomplete} will often be used as well due to its compactness.
+
+\section{Thermal properties}
+The standard-state thermodynamic properties (in molar units) for a gaseous species $k$ is defined using the standard seven-coefficient polynomial of Gordon and McBride~\cite{gordon1994computer}:
+\begin{align}
+\frac{C_{p,k}^{\circ}}{\mathcal{R}} &= a_{0,k} + T \left( a_{1,k} + T \left( a_{2,k} + T \left( a_{3,k} + a_{4,k} T \right) \right) \right) \label{e:cpk} \\
+\frac{H_k^{\circ}}{\mathcal{R}} &= T \left( a_{0,k} + T \left( \frac{a_{1,k}}{2} + T \left( \frac{a_{2,k}}{3} + T \left( \frac{a_{3,k}}{4} + \frac{a_{4,k}}{5} T \right) \right) \right) \right) + a_{5,k} \label{e:hk} \\
+\frac{S_k^{\circ}}{\mathcal{R}} &= a_{0,k} \ln T + T \left( a_{1,k} + T \left( \frac{a_{2,k}}{2} + T \left( \frac{a_{3,k}}{3} + \frac{a_{4,k}}{4} T \right) \right) \right) + a_{6,k} \label{e:sk}
+\end{align}
+where $C_{p,k}$ and $H_k$ are as described previously, $S_k$ is the entropy in molar units, and the ${}^{\circ}$ indicates a standard-state property at one atmosphere (equivalent to the property at any pressure for calorically perfect gases).
+
+\section{Reaction rate expressions}
+The net species rate of production is defined as:
+\begin{equation}
+ \dot{\omega}_{k} = \sum_{i=1}^{\nr} \nu_{k,i} q_{i}
+\end{equation}
+where $\nr$ is the number of reactions in the chemical kinetic model, $\nu_{k, i}$ the overall stoichionetric coefficient of species $k$ in reaction $i$ and $q_i$ the net rate-of-progress for reaction $i$:
+\begin{align}
+\nu_{k,i} &= \nu^{\prime}_{k,i} - \nu^{\prime\prime}_{k,i} \\
+q_{i} &= R_{i} c_{i}
+\end{align}
+with $\nu^{\prime}_{k,i}$ and $\nu^{\prime\prime}_{k,i}$ the
 

### 2017-12-18 derivations/scripts/derivations.py
fix signs

@@ -800,7 +800,7 @@ def _derivation(file, efile, conp=True, thermo_deriv=False):
     write_section('Definitions')
     nu_f = MyIndexedBase(r'\nu^{\prime}')
     nu_r = MyIndexedBase(r'\nu^{\prime\prime}')
-    nu = nu_f[k, i] - nu_r[k, i]
+    nu = nu_r[k, i] - nu_f[k, i]
     nu_sym = MyIndexedBase(r'\nu')
     write_eq(nu_sym[k, i], nu)
 

### 2017-12-18 derivations.tex
working

@@
 \usepackage[breaklinks=true, linkcolor=blue, citecolor=blue, colorlinks=true]{hyperref}
+% dcases
+\usepackage{mathtools}
 
@@
 
+%fix to dcases from here:http://tex.stackexchange.com/questions/252410/centering-in-dcases-environment/252414
+\MHInternalSyntaxOn
+\renewcommand{\dcases}
+ {
+  \MT_start_cases:nnnn
+    {\quad}
+    {$\m@th\displaystyle##$\hfil}
+    {$\m@th\displaystyle##$\hfil}
+    {\lbrace}
+ }
+\MHInternalSyntaxOff
+
 \begin{document}
@@
 
-\section{State variables}
+\section{Governing equations}
+\subsection{State variables}
 
@@
 
-\section{Chemical source terms}
+\subsection{Chemical source terms}
 The chemical source terms for this system are:
@@
 
-\section{Thermal properties}
+\subsection{Thermal properties}
 The standard-state thermodynamic properties (in molar units) for a gaseous species $k$ is defined using the standard seven-coefficient polynomial of Gordon and McBride~\cite{gordon1994computer}:
@@
 
-\section{Reaction rate expressions}
+\subsection{Reaction rate expressions}
 The net species rate of production is defined as:
@@
 \end{equation}
-where $\nr$ is the number of reactions in the chemical kinetic model, $\nu_{k, i}$ the overall stoichionetric coefficient of species $k$ in reaction $i$ and $q_i$ the net rate-of-progress for reaction $i$:
+where $\nr$ is the number of reactions in the chemical kinetic model, $\nu_{k, i}$ the overall stoichiometric coefficient of species $k$ in reaction $i$ and $q_i$ the net rate-of-progress for reaction $i$:
 \begin{align}
-\nu_{k,i} &= \nu^{\prime}_{k,i} - \nu^{\prime\prime}_{k,i} \\
+\nu_{k,i} &= \nu^{\prime\prime}_{k,i} - \nu^{\prime}_{k,i} \\
 q_{i} &= R_{i} c_{i}
 \end{align}
-with $\nu^{\prime}_{k,i}$ and $\nu^{\prime\prime}_{k,i}$ the
+with $\nu^{\prime}_{k,i}$ and $\nu^{\prime\prime}_{k,i}$ the product and reactant stoichiometric cofficients (respectively) of species $k$ in reaction $i$.
+The base rate-of-progress for the $i$th reversible reaction $R_{i}$ is given by:
+\begin{align}
+R_{i} &= {R_f}_{i} - {R_r}_{i} \\
+{R_f}_{i} &= {k_f}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} \\
+{R_r}_{i} &= {k_r}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
+\end{align}
+where ${k_f}_{i}$ and ${k_r}_{i}$ are the forward and reverse reaction coefficients (respectively) for the $i$th reaction, and the third-body\slash pressure modification $c_{i}$ is given by:
+\begin{equation}
+c_i = \begin{dcases}
+  1 &\text{for elementary reactions,} \\
+  [X]_i &\text{for third-body enhanced reactions,} \\
+  \frac{P_{r,i}}{1 + P_{r,i}} F_i &\text{for unimolecular/recombination falloff reactions, and} \\
+  \frac{1}{1 + P_{r,i}} F_i &\text{for chemically-activated bimolecular reactions,}
+  \end{dcases}
+\end{equation}
+where for the $i$th reaction $[X]_i$ is the third-body concentration, $P_{r,i}$ is the reduced pressure, and $F_i$ is the falloff blending factor.
+These terms are defined in the following sections.
+
+The forward reaction rate coefficient $k_{f, i}$ is given by the three-parameter Arrhenius expression:
+\begin{equation}
+  {k_f}_{i} = A_{i} T^{\beta_{i}} \operatorname{exp}\left({- \frac{{E_{a}}_{i}}{T R_u}}\right) \;,
+\end{equation}
+where $A_i$ is the pre-exponential factor, $\beta_i$ is the temperature exponent, and $T_{a, i}$ is the activation temperature given by $T_{a, i} = E_{a, i} / \mathcal{R}$.
+
+As given by Lu and Law~\cite{Lu:2009gh}, depending on the value of the Arrhenius parameters, $k_{f,i}$ can be calculated in different ways to minimize the computational cost:
+\begin{equation}
+  k_{f,i} =
+  \begin{dcases}
+  A_i & \text{if } \beta = 0 \text{ and } T_{a,i} = 0 \;, \\
+  \exp \left( \log A_i + \beta_i \log T \right)   & \text{if } \beta_i \neq 0 \text{ and } T_{a, i} = 0 \;, \\
+  \exp \left( \log A_i + \beta_i \log T - T_{a, i} / T \right) & \text{if } \beta_i \neq 0 \text{ and } T_{a, i} \neq 0 \;, \\
+  \exp \left( \log A_i - T_{a, i} / T \right)  & \text{if } \beta_i = 0 \text{ and } T_{a, i} \neq 0 \;, \text{ and} \\
+  A_i \prod^{\beta_i} T & \text{if } T_{a, i} = 0 \text{ and } \beta_i \in \mathbb{Z} \;,
+  \end{dcases}
+\end{equation}
+where $\mathbb{Z}$ is the set of integers; the extent of this specialization can be controlled when generating code via \texttt{pyjac}.
+
+\subsection{Reverse rate coefficient}
+By definition, the reverse rate coefficient of irreversible reactions ${k_r}_{i}$ is zero, while reversible reactions may have non-zero ${k_r}_{i}$.
+Note that in \texttt{pyJac}, reversible reactions with explicit reverse Arrhenius parameters are split into two irreversible reactions; this simplifies calculation inside the generated code, and eases comparison to Cantera~\cite{Goodwin:2015aa} which applies the same transformation.
+For reversible reactions without an explicit parameterization, the reverse rate coefficient is calculated from ratio of the forward rate coefficient and the equilibrium constant:
+\begin{align}
+ {k_r}_{i} &= \frac{{k_f}_{i}}{{K_c}_{i}}\; , \\
+ {K_c}_{i} &= \left(\left(\frac{P_{atm}}{T R_u}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}}\right) {K_p}_{i}\; , \\
+\end{align}
 

### 2017-12-18 derivations/scripts/derivations.py
add missing nu

@@ -871,7 +871,7 @@ def _derivation(file, efile, conp=True, thermo_deriv=False):
     write_dummy_eq(latex(Kp_sym[i]) + ' = ' +
                    r'\text{exp}(\frac{\Delta S^{\circ}_k}{R_u} - \frac{\Delta H^{\circ}_k}{R_u T})')
     write_dummy_eq(latex(Kp_sym[i]) + ' = ' +
-                   r'\text{exp}(\sum_{k=1}^{N_s}\frac{S^{\circ}_k}{R_u} - \frac{H^{\circ}_k}{R_u T})')
+                   r'\text{exp}\left(\sum_{k=1}^{N_s}\nu_{ki}\left(\frac{S^{\circ}_k}{R_u} - \frac{H^{\circ}_k}{R_u T}\right)\right)')
 
     B_sym = MyIndexedFunc('B', T)
     Kc = ((Patm / R)**Sum(nu_sym[k, i], (k, 1, Ns))) * \

### 2017-12-18 derivations.tex
Add plog/cheb & falloff parameterizations

@@
 \usepackage{mathtools}
+% for text right arrow
+\usepackage{textcomp}
 
@@
 
-\subsection{Chemical source terms}
-The chemical source terms for this system are:
+\subsection{Thermo-chemical source terms}
+The evolution of the thermo-chemical state variables of this system is described by a set of ordinary-differential equations:
 \begin{subequations}
@@
 \begin{equation}
+\label{e:rxn_pressure}
 c_i = \begin{dcases}
@@
 \begin{equation}
-  {k_f}_{i} = A_{i} T^{\beta_{i}} \operatorname{exp}\left({- \frac{{E_{a}}_{i}}{T R_u}}\right) \;,
+  \label{e:arrhenius}
+  {k_f}_{i} = A_{i} T^{\beta_{i}} \operatorname{exp}\left({- \frac{{E_{a}}_{i}}{T \Ru}}\right) \;,
 \end{equation}
@@
  {k_r}_{i} &= \frac{{k_f}_{i}}{{K_c}_{i}}\; , \\
- {K_c}_{i} &= \left(\left(\frac{P_{atm}}{T R_u}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}}\right) {K_p}_{i}\; , \\
+ {K_c}_{i} &= \left(\left(\frac{P_{atm}}{T \Ru}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}}\right) {K_p}_{i}\; ,\text{ and} \label{e:kc}\\
+ {K_p}_{i} &= \text{exp}\left(\frac{\Delta S^{\circ}_k}{\Ru} - \frac{\Delta H^{\circ}_k}{\Ru T}\right) = \text{exp}\left(\sum_{k=1}^{\ns}\nu_{ki}\left(\frac{S^{\circ}_k}{\Ru} - \frac{H^{\circ}_k}{\Ru T}\right)\right) \label{e:kp}
 \end{align}
+where $P_{atm}$ is the pressure of one standard atmosphere in appropriate units.
+
+By combining~\eqref{e:kc}~and~\eqref{e:kp}, we obtain:
+\begin{equation}
+ {K_c}_{i} = \left(\left(\frac{P_{atm}}{\Ru}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}}\right) \operatorname{exp}\left({\sum_{k=1}^{\ns} \nu_{k,i} B_{k}}\right)
+\end{equation}
+where, $B_k$ is:
+\begin{align}
+ B_{k} &= \frac{S^{\circ}_k}{\Ru} - \frac{H^{\circ}_k}{\Ru T} - \ln{T} \nonumber\; , \\
+ B_{k} &= T \left(T \left(T \left(\frac{T a_{k,4}}{20} + \frac{a_{k,3}}{12}\right) + \frac{a_{k,2}}{6}\right) + \frac{a_{k,1}}{2}\right) \nonumber \\
+       & \quad + \left(a_{k,0} - 1\right) \log{\left (T \right )} - a_{k,0} + a_{k,6} - \frac{a_{k,5}}{T}
+\end{align}
+from Eqs.~\eqref{e:hk} and~\eqref{e:sk}.
+
+\subsection{Third-body effects}
+\label{s:thdbody}
+
+For a reaction enhanced (or diminished) by the presence of a third body, the reaction rate is modified by the third-body concentration $[X]_i$ given by
+\begin{equation}
+[X]_{i} = \sum_{k=1}^{\ns} \alpha_{k,i} [C]_{k} \;,
+\end{equation}
+where $\alpha_{k,i}$ is the third-body efficiency of species $k$ in the $i$th reaction.
+For a default third-body efficiency of $\alpha_{k,i} = 1$, this may be rearranged to the compact-storage form:
+\begin{equation}
+ [X]_{i} = [C] + \sum_{k=1}^{\ns} \left(\alpha_{k,i} - 1\right) [C]_{k} \;,
+\end{equation}
+where only species with non-default third-body efficiencies must be stored in the model.
+Expanding the concentration of the last species gives:
+\begin{equation}
+\label{e:thd_mix}
+ [X]_{i}=[C] \alpha_{\ns,i} + \sum_{k=1}^{-1 + \ns} \left(- \alpha_{\ns,i} + \alpha_{k,i}\right) [C]_{k}\;.
+\end{equation}
+If all species in the mixture contribute equally as third bodies, then $\alpha_{k,i} = 1$ for all species:
+\begin{equation}
+\label{e:thd_unity}
+ [X]_{i} = [C] = \frac{P}{\Ru T} \;.
+\end{equation}
+In addition, a single species may act as the third body in which case
+\begin{equation}
+ [X]_{i} = [C_m]
+\end{equation}
+where the $m$th species is the third body.
+For evaluation of Jacobian entries, the expanded form:
+\begin{equation}
+\label{e:thd_spec}
+ [X]_{i}=\left([C] - \sum_{k=1}^{-1 + \ns} [C]_{k}\right) \delta_{\ns m} + \left(- \delta_{\ns m} + 1\right) [C]_{m}
+\end{equation}
+will be used, where the Kronecker delta $\delta_{\ns m}$ is unity if and only if the third body species $m$ is the last species in the model.
+
+\subsection{Falloff reactions}
+Unlike elementary and third-body reactions, falloff reactions exhibit a pressure dependence described as a blending of rates at low- and high-pressure limits; thus, the rate coefficients depend on a mixture of low-pressure- ($k_{0, i}$) and high-pressure-limit ($k_{\infty,i}$) coefficients, each with corresponding Arrhenius parameters and expressed using Eq.~\eqref{e:arrhenius}.
+The ratio of the coefficients $k_{0, i}$ and $k_{\infty, i}$, combined with the third-body concentration (based on either the mixture as a whole including any efficiencies $\alpha_{k,i}$, or a specific species), define a reduced pressure $P_{r,i}$ given by
+\begin{equation}
+ P_{r, i}=\frac{[X]_{i} k_{0, i}}{k_{\infty, i}}
+\end{equation}
+where $[X]_{i}$ is the appropriate third-body concentration as described in Sec.~\ref{s:thdbody}
+
+The falloff blending factor $F_i$ used in Eq.~\eqref{e:rxn_pressure} is determined based on one of three representations: the Lindemann~\cite{Lindemann:1922cz}, Troe~\cite{Gilbert:1983bb}, and SRI~\cite{Stewart:1989gj} falloff approaches
+\begin{equation}
+F_i = \begin{dcases}
+1 &\text{for Lindemann,} \\
+F_{\text{cent}}^{ \left( 1 + ( A_{\text{Troe}} / B_{\text{Troe}} )^2 \right)^{-1} } &\text{for Troe, or} \\
+d T^e \left( a \cdot \exp \left( -\frac{b}{T} \right) + \exp \left( -\frac{T}{c} \right) \right)^X &\text{for SRI.}
+\end{dcases}
+\end{equation}
+The Troe representation is described by the variables:
+\begin{align}
+ F_{cent} &= a \operatorname{exp}\left({- \frac{T}{T^{*}}}\right) + \left(- a + 1\right) \operatorname{exp}\left({- \frac{T}{T^{***}}}\right) + \operatorname{exp}\left({- \frac{T^{**}}{T}}\right) \;, \\
+ A_{Troe} &= - 0.67 \log_{10}{\left (F_{cent} \right )} + \log_{10}{\left (P_{r, i} \right )} - 0.4 \;\text{, and}\\
+ B_{Troe} &= - 1.1762 \log_{10}{\left (F_{cent} \right )} - 0.14 \log_{10}{\left (P_{r, i} \right )} + 0.806 
+\end{align}
+where $a$, $T^{***}$, $T^*$, and $T^{**}$ are specified parameters.
+The final parameter $T^{**}$ is optional, and, if it is not used, the final term of $F_{\text{cent}}$ is omitted.
+
+The exponent used in the SRI representation is given by:
+\begin{equation}
+ X = \frac{1}{\log_{10}^{2}{\left (P_{r, i} \right )} + 1}
+\end{equation}
+where $a$, $b$, and $c$ are required parameters.
+The parameters $d$ and $e$ are optional; if not specified, $d = 1$ and $e = 0$.
+
+\subsection{Pressure-dependent reactions}
+
+In addition to the falloff approach given previously, two additional formulations can be used to describe the pressure dependence of reactions that do not follow the modification factor $c_i$ approach.
+The first involves logarithmic interpolation between Arrhenius rates at two pressures~\cite{chemkin:2012,Goodwin:2015aa}, each evaluated using Eq.~\eqref{e:arrhenius}:
+\begin{align}
+k_1 (T) &= A_1 T^{\beta_1} \exp \left( -\frac{T_{a, 1}}{T} \right) \text{ at } P_1 \text{ and} \label{e:plog_k1} \\
+k_2 (T) &= A_2 T^{\beta_2} \exp \left( -\frac{T_{a, 2}}{T} \right) \text{ at } P_2 \;, \label{e:plog_k2}
+\end{align}
+where the Arrhenius coefficients are given for each pressure $p_1$ and $p_2$.
+Then, the reaction rate coefficient at a particular pressure $p$ between $p_1$ and $p_2$ can be determined through logarithmic interpolation:
+\begin{equation}
+\log{\left ({k_f}_{i} \right )} = \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) \left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \log{\left (k_{1} \right )}
+\end{equation}
+
+In addition, the pressure dependence of a reaction can be expressed through a bivariate Chebyshev polynomial fit~\cite{Venkatesh:1997hv,Venkatesh:1997ik,Venkatesh:2000gj,chemkin:2012,Goodwin:2015aa}:
+\begin{equation}
+\log_{10} k_f (T, p) = \sum_{i = 1}^{N_T} \sum_{j = 1}^{N_p} \eta_{ij} \phi_i (\tilde{T}) \phi_j \left(\tilde{p}\right) \label{e:cheb} \;,
+\end{equation}
+where $\eta_{ij}$ is the coefficient corresponding to the grid points $i$ and $j$, $N_T$ and $N_p$ are the numbers of grid points for temperature and pressure, respectively, and $\phi_n$ is the Chebyshev polynomial of the first kind of degree $n - 1$ typically expressed as
+\begin{equation}
+\phi_n (x) = \mathcal{T}_{n-1} (x) = \cos \left( (n - 1) \arccos (x) \right) \quad \text{for } |x| \leq 1 \;.
+\end{equation}
+The reduced temperature $\tilde{T}$ and pressure $\tilde{p}$ are given by
+\begin{align}
+\tilde{T} &\equiv \frac{2 T^{-1} - T^{-1}_{\min} - T^{-1}_{\max}}{T^{-1}_{\max} - T^{-1}_{\min}} \quad\text{and} \\
+\tilde{p} &\equiv \frac{2\log_{10} p - \log_{10} p_{\min} - \log_{10} p_{\max}}{\log_{10} p_{\max} - \log_{10} p_{\min}} \;,
+\end{align}
+where $T_{\min} \leq T \leq T_{\max}$ and $p_{\min} \leq p \leq p_{\max}$ describe the ranges of validity for temperature and pressure.
 

### 2017-12-19 derivations.tex
work on Jac

@@
 \begin{align}
-n &= \frac{V P}{T \Ru} = \sum_{i=1}^{\ns}{n_i}, \label{source:moles}\\
+n &= \frac{V P}{T \Ru} = \sum_{i=1}^{\ns}{n_i}, \label{e:source_moles}\\
 n_{\ns} &= \frac{V P}{T \Ru} - \sum_{i=1}^{\ns - 1}{n_i}
@@
 \begin{subequations}
+\label{e:source_terms}
 \begin{align}
-\frac{\text{d} \Phi }{\text{d} t } &= \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} V }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\}\dconp, \\
-\frac{\text{d} \Phi }{\text{d} t } &= \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} P }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\}\dconv
+f &= \frac{\text{d} \Phi }{\text{d} t } &= \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} V }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\}\dconp, \\
+f &= \frac{\text{d} \Phi }{\text{d} t } &= \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} P }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\}\dconv
 \end{align}
@@
 \end{subequations}
-Additionally, the temperature source terms may be expanded to contain only state variables (i.e., by removal of the last species' source term and overall production rate):
+
+Additionally, the concentration and moles of the last species in the model (which is not a state variable) may be expanded in terms of the state variables of the system:
+\begin{equation}
+ \label{e:last_spec_conc}
+   [C]_{\ns} = [C] - \sum_{k=1}^{-1 + \ns} [C]_{k}
+\end{equation}
+
+Using Eqs.~\eqref{e:last_spec_conc}, \eqref{source:spec} and \eqref{source:spec_ns}, the temperature source terms may be expanded to contain only state variables (i.e., by removal of the last species' source term and overall production rate):
 \begin{subequations}
@@
 \begin{equation}
+ \label{e:spec_rop}
  \dot{\omega}_{k} = \sum_{i=1}^{\nr} \nu_{k,i} q_{i}
@@
 \begin{align}
-R_{i} &= {R_f}_{i} - {R_r}_{i} \\
-{R_f}_{i} &= {k_f}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} \\
-{R_r}_{i} &= {k_r}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
+R_{i} &= {R_f}_{i} - {R_r}_{i} \label{e:ropnet}\\
+{R_f}_{i} &= {k_f}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} \label{e:ropf}\\
+{R_r}_{i} &= {k_r}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} \label{e:ropr}
 \end{align}
@@
 \begin{align}
- {k_r}_{i} &= \frac{{k_f}_{i}}{{K_c}_{i}}\; , \\
+ {k_r}_{i} &= \frac{{k_f}_{i}}{{K_c}_{i}}\; , \label{e:kr}\\
  {K_c}_{i} &= \left(\left(\frac{P_{atm}}{T \Ru}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}}\right) {K_p}_{i}\; ,\text{ and} \label{e:kc}\\
@@
 \begin{equation}
+ \label{e:kc_in_kp}
  {K_c}_{i} = \left(\left(\frac{P_{atm}}{\Ru}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}}\right) \operatorname{exp}\left({\sum_{k=1}^{\ns} \nu_{k,i} B_{k}}\right)
@@
 \begin{align}
+ \label{e:Bk}
  B_{k} &= \frac{S^{\circ}_k}{\Ru} - \frac{H^{\circ}_k}{\Ru T} - \ln{T} \nonumber\; , \\
@@
 
+\section{Jacobian derivation}
+
+Let $\mathcal{J}$ denote the Jacobian matrix corresponding to the set of ODEs defined in~\eqref{e:source_terms}.
+$\mathcal{J}$ is filled with the partial derivatives $\partial f / \partial \Phi$, such that:
+\begin{equation}
+ \mathcal{J}_{i,j} = \frac{\partial f_i}{\partial \Phi_j},\qquad j=1 \ldots \ns - 1
+\end{equation}
+We will begin by deriving subcomponents of the Jacobian matrix in the subsequent sections before presenting the final form of the Jacobian matrix in Sec.~\ref{s:jac_final}.
+
+\subsection{Species rate of production derivatives}
+First, we evaluate the derivatives of the net species rate of production equation~\eqref{e:spec_rop}, with respect to temperature:
+\begin{equation}
+ \frac{\partial \dot{\omega} }{\partial T }_{k} = \sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial T }_{i} + \nu_{k,i} \frac{\partial R }{\partial T }_{i} c_{i}\right) \;,
+\end{equation}
+other species:
+\begin{equation}
+ \frac{\partial q }{\partial n[k] }_{i} = R_{i} \frac{\partial c }{\partial n[j] }_{i} + \frac{\partial R }{\partial n[j] }_{i} c_{i}
+\end{equation}
+and the state parameter:
+\begin{subequations}
+ \begin{align}
+  \frac{\partial \dot{\omega} }{\partial V }_{k} &= \sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial V }_{i} + \nu_{k,i} \frac{\partial R }{\partial V }_{i} c_{i}\right)\dconp,\\
+  \frac{\partial \dot{\omega} }{\partial P }_{k} &= \sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial P }_{i} + \nu_{k,i} \frac{\partial R }{\partial P }_{i} c_{i}\right)\dconv
+ \end{align}
+\end{subequations}
+
+\subsection{Rate of progress derivatives}
+\subsubsection{Temperature derivative}
+Next, the rate of progress derivatives are evaluated, first with respect to temperature:
+\begin{equation}
+ \frac{\partial R }{\partial T }_{i} = \frac{\partial}{\partial T}\left({k_f}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right)
+\end{equation}
+the derivative of the forward reaction rate coefficient is:
+\begin{equation}
+ \label{e:dkf_dt}
+ \frac{\text{d} {k_f} }{\text{d} T }_{i} = \frac{{k_f}_{i}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
+\end{equation}
+While evaluating derivatives for the Jacobian, it is important to expand terms---e.g., concentration, rate of production, etc.---related to the last species to obtain the correct derivative, in this case using Eq.~\eqref{e:last_spec_conc}:
+\begin{equation}
+ {R_f} = \left(\left(- \sum_{k=1}^{-1 + \ns} [C]_{k} + \frac{P}{T \Ru}\right)^{\nu^{\prime}_{\ns,i}}\right) {k_f}_{i} \prod_{k=1}^{-1 + \ns} [C]_{k}^{\nu^{\prime}_{k,i}}
+\end{equation}
+The derivative of a species concentration with respect to temperature is zero for both \conp~and \conv:
+\begin{align}
+ \frac{\partial [C]_{k}}{\partial T} &= \frac{\partial}{\partial T} \left(\frac{n_k}{V}\right), \quad \left\{k \ne \ns\right\} \nonumber \\
+				     &= 0
+\end{align}
+as the volume is either constant (\conp) or a state variable and a function of time only (\conv).
+Hence, the full derivative of the forward rate of progress is:
+\begin{equation}
+ \label{e:dropfdt_1}
+ \frac{\partial {R_f} }{\partial T }_{i} = \frac{\text{d} {k_f} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{-1 + \ns} [C]_{k}^{\nu^{\prime}_{k,i}}
+\end{equation}
+By defining a temporary variable (for species $l = \left\{j, \ns\right\}$) $S^{\prime}_{l}$:
+\begin{equation}
+ \label{e:s_temp}
+ S^{\prime}_{l} = \nu^{\prime}_{l,i} [C]_{l}^{\nu^{\prime}_{l,i} - 1} \prod_{\substack{1 \leq l \leq l - 1\\l + 1 \leq l \leq \ns}} [C]_{l}^{\nu^{\prime}_{l,i}}
+\end{equation}
+and using~\eqref{e:dkf_dt}, Eq.~\eqref{e:dropfdt_1} can be simplified to:
+\begin{equation}
+ \label{e:dropf_dt}
+ \frac{\partial {R_f} }{\partial T }_{i} = - \frac{[C] S^{\prime}_{\ns}}{T} {k_f}_{i} + \frac{{R_f}_{i}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
+\end{equation}
+
+Starting from~\eqref{e:ropr} and applying the same expansion of the last species' concentration as previously, the temperature derivative of the reverse rate of progress is:
+\begin{equation}
+ \frac{\partial {R_r} }{\partial T }_{i} = \frac{\text{d} {k_r} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{-1 + \ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
+\end{equation}
+Next, Eq.~\eqref{e:kr} is considered to obtain the temperature derivative of the non-explicit reversible reaction rate coefficient:
+\begin{equation}
+ \frac{\text{d} {k_r} }{\text{d} T }_{i} = \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c} }{\text{d} T }_{i} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_r}_{i}
+\end{equation}
+The temperature derivative of the equilibrium constant is:
+\begin{equation}
+ \frac{\text{d} {K_c} }{\text{d} T }_{i} = {K_c}_{i} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} \;,
+\end{equation}
+with:
+\begin{equation}
+ \frac{\text{d} B }{\text{d} T }_{k} = T \left(T \left(\frac{T a_{k,4}}{5} + \frac{a_{k,3}}{4}\right) + \frac{a_{k,2}}{3}\right) + \frac{a_{k,1}}{2} + \frac{1}{T} \left(a_{k,0} - 1 + \frac{a_{k,5}}{T}\right)
+\end{equation}
+giving:
+\begin{equation}
+\label{e:dkr_dt}
+\frac{\text{d} {k_r} }{\text{d} T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_r}_{i} \;,
+\end{equation}
+and using an temporary value $S^{\prime\prime}_{\ns}$---defined analagously to Eq.~\eqref{e:s_temp} for the reverse direction---we obtain:
+\begin{equation}
+\label{e:dropr_dt}
+\frac{\partial {R_r} }{\partial T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_r}_{i} - \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_r}_{i}
+\end{equation}
+
+Finally, combining Eqs.~\eqref{e:dropf_dt} and~\eqref{e:dropr_dt} gives the total temperature derivative of the net rate of progress:
+\begin{align}
+ \label{e:drop_dt}
+ \frac{\partial R }{\partial T }_{i} =& \frac{{R_f}_{i}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right) - \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_r}_{i} \nonumber \\
+				      &\qquad + \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_r}_{i} - \frac{[C] S^{\prime}_{\ns}}{T} {k_f}_{i}
+\end{align}
+\subsubsection{Molar derivative}
+The derivative of the forward rate of progress with respect to the amount of moles of a species $j$ is:
+\begin{equation}
+ \frac{d}{d n_{k}} {R_f} = \left(\frac{\partial}{\partial n_{j}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right) {k_f}_{i} \;.
+\end{equation}
+The derivative of a species concentration with respect to moles of a species $j$ is given as:
+\begin{equation}
+ \frac{\partial [C_k]}{\partial n_j} =
+ \begin{dcases}
+ \frac{\delta_{j k}}{V} & k \ne \ns \\
+ -\frac{1}{V} & k = \ns
+ \end{dcases}
+\end{equation}
+Or put succinctly:
+\begin{equation}
+\label{e:dck_dnj}
+\frac{\partial [C_k]}{\partial n_j} = \frac{\delta_{j k} - \delta_{\ns k}}{V} \;.
+\end{equation}
+Using~\eqref{e:dck_dnj} the molar derivative of the forward rate of progress is found as:
+\begin{equation}
+ \frac{\partial {R_f} }{\partial n[j] }_{i} = {k_f}_{i} \sum_{k=1}^{\ns} \left(- \frac{\delta_{\ns k}}{V} + \frac{\delta_{j k}}{V}\right) \nu^{\prime}_{k,i} [C]_{k}^{\nu^{\prime}_{k,i} - 1} \prod_{\substack{1 \leq l \leq k - 1\\k + 1 \leq l \leq \ns}} [C]_{l}^{\nu^{\prime}_{l,i}}
+\end{equation}
+with the same temporary variables $S^{\prime}_{j}$ and $S^{\prime}_{\ns}$ as defined previously, this can be simplified to:
+\begin{equation}
+ \label{e:dropf_dnj}
+ \frac{\partial {R_f} }{\partial n[j] }_{i} = \frac{{k_f}_{i}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
+\end{equation}
+and by the same process, the reverse rate of progress derivative is found to be:
+\begin{equation}
+ \label{e:dropr_dnj}
+ \frac{\partial {R_r} }{\partial n[j] }_{i} = \frac{{k_r}_{i}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right)
+\end{equation}
+and the net rate of progress molar derivative:
+\begin{equation}
+ \frac{\partial R }{\partial n[j] }_{i} = - \frac{{k_r}_{i}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right) + \frac{{k_f}_{i}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
+\end{equation}
+
+\subsubsection{Parameter derivatives}
+The derivatives of the forward and reverse rates of progress with respect to the volume (\conp) or pressure (\conv) largely follow the same format of the previous sections.
+First, the derivative of the concentration of a species $k$ with respect to the parameter is:
+
+\begin{subequations}
+ % based on https://tex.stackexchange.com/a/239764/56227
+ % the idea is to take the longest case (1.b -- i.e., no hphantom) and use this
+ % to make a box around all cases such that the alignment works
+ \begin{align}
+ \frac{\partial [C] }{\partial V }_{k} &=
+ \begin{dcases}
+  \mathrlap{-\frac{[C]_{k}}{V}}\hphantom{\frac{1}{V} \sum_{k=1}^{-1 + \ns} [C]_{k}} & k \ne \ns \\
+  \frac{1}{V} \sum_{k=1}^{-1 + \ns} [C]_{k} & k = \ns
+ \end{dcases}&\dconp \\
+ \frac{\partial [C]_{k} }{\partial P } &=
+ \begin{dcases}
+  \mathrlap{\hphantom{0}0}\hphantom{\frac{1}{V} \sum_{k=1}^{-1 + \ns} [C]_{k}} & k \ne \ns \\
+  \mathrlap{\frac{1}{T \Ru}}\hphantom{\frac{1}{V} \sum_{k=1}^{-1 + \ns} [C]_{k}} & k = \ns
+ \end{dcases}&\dconv
+ \end{align}
+\end{subequations}
+
+\subsection{Final Jacobian entries}
+\label{s:jac_final}
+
 \printbibliography

### 2017-12-21 derivations.tex
drop/de

@@
 \subsubsection{Molar derivative}
+\label{s:rop_molar}
 The derivative of the forward rate of progress with respect to the amount of moles of a species $j$ is:
@@
 
+Following the outline of Sec.~\ref{s:rop_molar}, the forward rate of progress derivatives are:
+\begin{subequations}
+ \begin{align}
+  \frac{\partial {R_f} }{\partial V }_{i} &= \frac{[C] S^{\prime}_{\ns}}{V} {k_f}_{i} - \frac{{R_f}_{i}}{V} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}&\dconp\\
+  \frac{\partial {R_f} }{\partial P }_{i} &= \frac{S^{\prime}_{\ns} {k_f}_{i}}{T \Ru}&\dconv
+ \end{align}
+\end{subequations}
+and similarly, the reverse rate of progress derivatives:
+\begin{subequations}
+ \begin{align}
+  \frac{\partial {R_r} }{\partial V }_{i} &= \frac{[C] S^{\prime\prime}_{\ns}}{V} {k_r}_{i} - \frac{{R_r}_{i}}{V} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i}&\dconp\\
+  \frac{\partial {R_r} }{\partial P }_{i} &= \frac{S^{\prime\prime}_{\ns} {k_r}_{i}}{T \Ru}&\dconv
+ \end{align}
+\end{subequations}
+giving finally:
+\begin{subequations}
+ \label{e:dropi_de}
+ % linebreak and parenthesis sizing:
+ % https://tex.stackexchange.com/questions/49890/linebreak-between-left-and-right/49895
+ \begin{align}
+  \frac{\partial {R} }{\partial V }_{i} &= \frac{1}{V}\left(\vphantom{{R_r}_{i} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i}} [C]\left(S^{\prime}_{\ns} {k_f}_{i} - S^{\prime\prime}_{\ns} {k_r}_{i}\right) + \right.& \nonumber\\
+  &\left. \quad\left({R_r}_{i} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i} - {R_f}_{i} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}\right) \right)&\dconp\\
+  \frac{\partial {R} }{\partial P }_{i} &= \frac{1}{T \Ru} \left(S^{\prime}_{\ns} {k_f}_{i} - S^{\prime\prime}_{\ns} {k_r}_{i}\right)&\dconv
+ \end{align}
+\end{subequations}
+
+\subsection{Pressure-modification\slash Falloff function derivatives}
+
 \subsection{Final Jacobian entries}

### 2017-12-21 derivations.tex
cleveref updates and other fixes

@@
 \usepackage[utf8]{inputenc}
-\usepackage{amsmath, amssymb}
+\usepackage[english]{babel}
+\usepackage{csquotes}
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
 \usepackage[
@@
     style=numeric,
-    citestyle=numeric
+    citestyle=numeric,
+    sorting=none
 ]{biblatex}
 \addbibresource{derivations.bib}
-\usepackage[breaklinks=true, linkcolor=blue, citecolor=blue, colorlinks=true]{hyperref}
 % dcases
@@
 \usepackage{textcomp}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage[breaklinks=true, linkcolor=blue, citecolor=blue, colorlinks=true]{hyperref}
+% load after hyperref
+\usepackage[english,capitalise]{cleveref}
 
@@
 \subsection{State variables}
+The state vector for this derivation consists of the temperature, a thermodynamic parameter (pressure or volume) and the amount of moles of all species except the last in the model:
 
@@
 From the ideal gas law, the number of moles of the final species is determined as:
-\begin{subequations}
 \begin{align}
@@
 \end{align}
-\end{subequations}
 
@@
 \begin{align}
-f &= \frac{\text{d} \Phi }{\text{d} t } &= \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} V }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\}\dconp, \\
-f &= \frac{\text{d} \Phi }{\text{d} t } &= \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} P }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\}\dconv
+f &= \frac{\text{d} \Phi }{\text{d} t } = \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} V }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\}\dconp, \\
+f &= \frac{\text{d} \Phi }{\text{d} t } = \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} P }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\}\dconv
 \end{align}
@@
 \begin{equation}
-\frac{\text{d} n }{\text{d} t }_{k} = V \dot{\omega}_{k}
-\label{source:spec}
+\frac{\text{d} n }{\text{d} t }_{k} = V \dot{\omega}_{k} \label{e:spec}
 \end{equation}
@@
 \begin{subequations}
-\label{source:temperature_incomplete}
+\label{e:temperature_incomplete}
 \begin{align}
@@
 \end{subequations}
-where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure specific heat, constant-volume specific heat of species $k$ in molar units, while $[C]_{k}$ is the concentration.
+where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure specific heat, constant-volume specific heat of species $k$ in molar units, while $[C]_{k}$ is the concentration, given by:
+\begin{equation}
+ [C]_{k} = \frac{n_{k}}{V}
+\end{equation}
 
@@
 \begin{subequations}
-\label{source:param_incomplete}
+\label{e:param_incomplete}
 \begin{align}
@@
 
-Finally, the conservation of mass is invoked:
-\begin{align*}
- \frac{\text{d} m }{\text{d} t } &= 0 \\
- &= \sum_{k=1}^{\ns}  W_{k} \frac{\text{d} n }{\text{d} t }_{k},
-\end{align*}
-to determine the source term of the last species:
+To determine $\frac{\text{d} n }{\text{d} t }$, conservation of mass is invoked:
+\begin{equation}
+ \frac{\text{d} m }{\text{d} t } = 0 = \sum_{k=1}^{\ns}  W_{k} \frac{\text{d} n }{\text{d} t }_{k},
+\end{equation}
+from which, the source term of the last species may be written in terms of the species included in the state vector:
 \begin{equation}
  \frac{\text{d} n }{\text{d} t }_{\ns} = - \frac{1}{W_{\ns}} \sum_{k=1}^{\ns - 1} W_{k} \frac{\text{d} n }{\text{d} t }_{k}
- \label{source:spec_ns}
+ \label{e:spec_ns}
 \end{equation}
 where $W_{k}$ is the molecular weight of the $k$th species.
-Using~\eqref{source:spec_ns}, the total molar rate of change can be determined:
+Using~\cref{e:spec_ns}, the total molar rate of change can be determined:
 \begin{align}
@@
 \frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\text{d} n }{\text{d} t }_{k}
-\label{source:total_molar}
+\label{e:total_molar}
 \end{align}
 
-Combining~\eqref{source:spec},\eqref{source:param_incomplete},\eqref{source:total_molar} gives the final form of the pressure and volume terms:
+Combining~\cref{e:spec,e:param_incomplete,e:total_molar} gives the final form of the pressure and volume terms:
 \begin{subequations}
-\label{source:param_complete}
+\label{e:param_complete}
 \begin{align}
@@
 \end{equation}
+where $[C]$ is the total concentration:
+\begin{equation}
+ [C] = \frac{P}{T \Ru}
+\end{equation}
 
-Using Eqs.~\eqref{e:last_spec_conc}, \eqref{source:spec} and \eqref{source:spec_ns}, the temperature source terms may be expanded to contain only state variables (i.e., by removal of the last species' source term and overall production rate):
+Using~\cref{e:last_spec_conc,e:spec,e:spec_ns}, the temperature source terms may be expanded to contain only state variables (i.e., by removal of the last species' source term and overall production rate):
 \begin{subequations}
-\label{source:temperature_complete}
+\label{e:temperature_complete}
 \begin{align}
@@
 \end{subequations}
-where $[C]$ is the total concentration:
-\begin{equation}
- [C] = \frac{P}{T \Ru}
-\end{equation}
-This form is used for differentiation, but~\eqref{source:temperature_incomplete} will often be used as well due to its compactness.
+This form is used for differentiation, but~\cref{e:temperature_incomplete} will often be used as well due to its compactness.
 
@@
 
-By combining~\eqref{e:kc}~and~\eqref{e:kp}, we obtain:
+By combining~\cref{e:kc,e:kp}, we obtain:
 \begin{equation}
@@
 \end{align}
-from Eqs.~\eqref{e:hk} and~\eqref{e:sk}.
+from~\cref{e:hk,e:sk}.
 
@@
 \subsection{Falloff reactions}
-Unlike elementary and third-body reactions, falloff reactions exhibit a pressure dependence described as a blending of rates at low- and high-pressure limits; thus, the rate coefficients depend on a mixture of low-pressure- ($k_{0, i}$) and high-pressure-limit ($k_{\infty,i}$) coefficients, each with corresponding Arrhenius parameters and expressed using Eq.~\eqref{e:arrhenius}.
+Unlike elementary and third-body reactions, falloff reactions exhibit a pressure dependence described as a blending of rates at low- and high-pressure limits; thus, the rate coefficients depend on a mixture of low-pressure- ($k_{0, i}$) and high-pressure-limit ($k_{\infty,i}$) coefficients, each with corresponding Arrhenius parameters and expressed using Eq.~\cref{e:arrhenius}.
 The ratio of the coefficients $k_{0, i}$ and $k_{\infty, i}$, combined with the third-body concentration (based on either the mixture as a whole including any efficiencies $\alpha_{k,i}$, or a specific species), define a reduced pressure $P_{r,i}$ given by
@@
 \end{equation}
-where $[X]_{i}$ is the appropriate third-body concentration as described in Sec.~\ref{s:thdbody}
+where $[X]_{i}$ is the appropriate third-body concentration as described in~\cref{s:thdbody}
 
-The falloff blending factor $F_i$ used in Eq.~\eqref{e:rxn_pressure} is determined based on one of three representations: the Lindemann~\cite{Lindemann:1922cz}, Troe~\cite{Gilbert:1983bb}, and SRI~\cite{Stewart:1989gj} falloff approaches
+The falloff blending factor $F_i$ used in Eq.~\cref{e:rxn_pressure} is determined based on one of three representations: the Lindemann~\cite{Lindemann:1922cz}, Troe~\cite{Gilbert:1983bb}, and SRI~\cite{Stewart:1989gj} falloff approaches
 \begin{equation}
@@
 In addition to the falloff approach given previously, two additional formulations can be used to describe the pressure dependence of reactions that do not follow the modification factor $c_i$ approach.
-The first involves logarithmic interpolation between Arrhenius rates at two pressures~\cite{chemkin:2012,Goodwin:2015aa}, each evaluated using Eq.~\eqref{e:arrhenius}:
+The first involves logarithmic interpolation between Arrhenius rates at two pressures~\cite{chemkin:2012,Goodwin:2015aa}, each evaluated using Eq.~\cref{e:arrhenius}:
 \begin{align}
@@
 
-Let $\mathcal{J}$ denote the Jacobian matrix corresponding to the set of ODEs defined in~\eqref{e:source_terms}.
+Let $\mathcal{J}$ denote the Jacobian matrix corresponding to the set of ODEs defined in~\cref{e:source_terms}.
 $\mathcal{J}$ is filled with the partial derivatives $\partial f / \partial \Phi$, such that:
@@
 \end{equation}
-We will begin by deriving subcomponents of the Jacobian matrix in the subsequent sections before presenting the final form of the Jacobian matrix in Sec.~\ref{s:jac_final}.
+We will begin by deriving subcomponents of the Jacobian matrix in the subsequent sections before presenting the final form of the Jacobian matrix in~\cref{s:jac_final}.
 
 \subsection{Species rate of production derivatives}
-First, we evaluate the derivatives of the net species rate of production equation~\eqref{e:spec_rop}, with respect to temperature:
+First, we evaluate the derivatives of the net species rate of production equation~\cref{e:spec_rop}, with respect to temperature:
 \begin{equation}
+ \label{e:dwdot_dT}
  \frac{\partial \dot{\omega} }{\partial T }_{k} = \sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial T }_{i} + \nu_{k,i} \frac{\partial R }{\partial T }_{i} c_{i}\right) \;,
@@
 \begin{equation}
- \frac{\partial q }{\partial n[k] }_{i} = R_{i} \frac{\partial c }{\partial n[j] }_{i} + \frac{\partial R }{\partial n[j] }_{i} c_{i}
+ \label{e:dwdot_dnj}
+ \frac{\partial \dot{\omega} }{\partial n[j] }_{k} = \sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial n[j] }_{i} + \nu_{k,i} \frac{\partial R }{\partial n[j] }_{i} c_{i}\right)
 \end{equation}
@@
 \begin{subequations}
+ \label{e:dwdot_de}
  \begin{align}
@@
 \end{equation}
-While evaluating derivatives for the Jacobian, it is important to expand terms---e.g., concentration, rate of production, etc.---related to the last species to obtain the correct derivative, in this case using Eq.~\eqref{e:last_spec_conc}:
+While evaluating derivatives for the Jacobian, it is important to expand terms---e.g., concentration, rate of production, etc.---related to the last species to obtain the correct derivative, in this case using Eq.~\cref{e:last_spec_conc}:
 \begin{equation}
@@
 \end{equation}
-and using~\eqref{e:dkf_dt}, Eq.~\eqref{e:dropfdt_1} can be simplified to:
+and using~\cref{e:dkf_dt}, Eq.~\cref{e:dropfdt_1} can be simplified to:
 \begin{equation}
@@
 
-Starting from~\eqref{e:ropr} and applying the same expansion of the last species' concentration as previously, the temperature derivative of the reverse rate of progress is:
+Starting from~\cref{e:ropr} and applying the same expansion of the last species' concentration as previously, the temperature derivative of the reverse rate of progress is:
 \begin{equation}
@@
 \end{equation}
-Next, Eq.~\eqref{e:kr} is considered to obtain the temperature derivative of the non-explicit reversible reaction rate coefficient:
+Next, Eq.~\cref{e:kr} is considered to obtain the temperature derivative of the non-explicit reversible reaction rate coefficient:
 \begin{equation}
@@
 \end{equation}
-and using an temporary value $S^{\prime\prime}_{\ns}$---defined analagously to Eq.~\eqref{e:s_temp} for the reverse direction---we obtain:
+and using an temporary value $S^{\prime\prime}_{\ns}$---defined analagously to Eq.~\cref{e:s_temp} for the reverse direction---we obtain:
 \begin{equation}
@@
 
-Finally, combining Eqs.~\eqref{e:dropf_dt} and~\eqref{e:dropr_dt} gives the total temperature derivative of the net rate of progress:
+Finally, combining~\cref{e:dropf_dt,e:dropr_dt} gives the total temperature derivative of the net rate of progress:
 \begin{align}
@@
 \end{equation}
-Using~\eqref{e:dck_dnj} the molar derivative of the forward rate of progress is found as:
+Using~\cref{e:dck_dnj} the molar derivative of the forward rate of progress is found as:
 \begin{equation}
@@
 
-Following the outline of Sec.~\ref{s:rop_molar}, the forward rate of progress derivatives are:
+Following the outline of~\cref{s:rop_molar}, the forward rate of progress derivatives are:
 \begin{subequations}
@@
 
-\subsection{Pressure-modification\slash Falloff function derivatives}
+\subsection{Pressure modification\slash Falloff function derivatives}
 

### 2018-01-02 derivations.tex
updates

@@
 \addbibresource{derivations.bib}
+
 % dcases
@@
 \usepackage{amssymb}
-\usepackage{amsmath}
+\usepackage{amsmath,eqparbox,xparse}
 \usepackage[breaklinks=true, linkcolor=blue, citecolor=blue, colorlinks=true]{hyperref}
@@
 \usepackage[english,capitalise]{cleveref}
+
+% https://tex.stackexchange.com/a/34412/5764
+\makeatletter
+\NewDocumentCommand{\eqmathbox}{o O{c} m}{%
+  \IfValueTF{#1}
+    {\def\eqmathbox@##1##2{\eqmakebox[#1][#2]{$##1##2$}}}
+    {\def\eqmathbox@##1##2{\eqmakebox{$##1##2$}}}
+  \mathpalette\eqmathbox@{#3}
+}
+\makeatother
 
@@
 \newcommand{\conv}{CONV}
-\newcommand{\dconp}{\ensuremath{,\qquad\text{for \conp}}}
-\newcommand{\dconv}{\ensuremath{,\qquad\text{for \conv}}}
+\newcommand{\dconp}{\ensuremath{\text{for \conp,}}}
+\newcommand{\dconv}{\ensuremath{\text{for \conv,}}}
 \newcommand{\Ru}{\ensuremath{\mathcal{R}}}
+
+% treat subequations as equation for cleverref
+% https://tex.stackexchange.com/a/308627/56227
+\crefalias{subequation}{equation}
 
@@
   [X]_i &\text{for third-body enhanced reactions,} \\
-  \frac{P_{r,i}}{1 + P_{r,i}} F_i &\text{for unimolecular/recombination falloff reactions, and} \\
+  \frac{P_{r,i}}{1 + P_{r,i}} F_i &\text{for unimolecular\slash~recombination fall-off reactions, and} \\
   \frac{1}{1 + P_{r,i}} F_i &\text{for chemically-activated bimolecular reactions,}
@@
 \begin{equation}
+ \label{e:pr}
  P_{r, i}=\frac{[X]_{i} k_{0, i}}{k_{\infty, i}}
@@
  \label{e:dwdot_dnj}
- \frac{\partial \dot{\omega} }{\partial n[j] }_{k} = \sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial n[j] }_{i} + \nu_{k,i} \frac{\partial R }{\partial n[j] }_{i} c_{i}\right)
+ \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} = \sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial {n_j} }_{i} + \nu_{k,i} \frac{\partial R }{\partial {n_j} }_{i} c_{i}\right)
 \end{equation}
@@
 \begin{equation}
- \frac{d}{d n_{k}} {R_f} = \left(\frac{\partial}{\partial n_{j}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right) {k_f}_{i} \;.
+ \frac{d}{d n_{k}} {R_{f_i}} = \left(\frac{\partial}{\partial n_{j}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right) {k_f}_{i} \;.
 \end{equation}
@@
 \begin{equation}
- \frac{\partial {R_f} }{\partial n[j] }_{i} = {k_f}_{i} \sum_{k=1}^{\ns} \left(- \frac{\delta_{\ns k}}{V} + \frac{\delta_{j k}}{V}\right) \nu^{\prime}_{k,i} [C]_{k}^{\nu^{\prime}_{k,i} - 1} \prod_{\substack{1 \leq l \leq k - 1\\k + 1 \leq l \leq \ns}} [C]_{l}^{\nu^{\prime}_{l,i}}
+ \frac{\partial {R_f}_{i} }{\partial {n_j} } = {k_f}_{i} \sum_{k=1}^{\ns} \left(- \frac{\delta_{\ns k}}{V} + \frac{\delta_{j k}}{V}\right) \nu^{\prime}_{k,i} [C]_{k}^{\nu^{\prime}_{k,i} - 1} \prod_{\substack{1 \leq l \leq k - 1\\k + 1 \leq l \leq \ns}} [C]_{l}^{\nu^{\prime}_{l,i}}
 \end{equation}
@@
  \label{e:dropf_dnj}
- \frac{\partial {R_f} }{\partial n[j] }_{i} = \frac{{k_f}_{i}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
+ \frac{\partial {R_f}_{i} }{\partial {n_j} } = \frac{{k_f}_{i}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
 \end{equation}
@@
  \label{e:dropr_dnj}
- \frac{\partial {R_r} }{\partial n[j] }_{i} = \frac{{k_r}_{i}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right)
+ \frac{\partial {R_r}_{i} }{\partial {n_j} } = \frac{{k_r}_{i}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right)
 \end{equation}
@@
 \begin{equation}
- \frac{\partial R }{\partial n[j] }_{i} = - \frac{{k_r}_{i}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right) + \frac{{k_f}_{i}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
+ \frac{\partial R_{i} }{\partial {n_j} } = - \frac{{k_r}_{i}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right) + \frac{{k_f}_{i}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
 \end{equation}
@@
  \begin{align}
-  \frac{\partial {R_f} }{\partial V }_{i} &= \frac{[C] S^{\prime}_{\ns}}{V} {k_f}_{i} - \frac{{R_f}_{i}}{V} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}&\dconp\\
-  \frac{\partial {R_f} }{\partial P }_{i} &= \frac{S^{\prime}_{\ns} {k_f}_{i}}{T \Ru}&\dconv
+  \frac{\partial {R_f}_{i} }{\partial V } &= \frac{[C] S^{\prime}_{\ns}}{V} {k_f}_{i} - \frac{{R_f}_{i}}{V} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}&\dconp\\
+  \frac{\partial {R_f}_{i} }{\partial P } &= \frac{S^{\prime}_{\ns} {k_f}_{i}}{T \Ru}&\dconv
  \end{align}
@@
  \begin{align}
-  \frac{\partial {R_r} }{\partial V }_{i} &= \frac{[C] S^{\prime\prime}_{\ns}}{V} {k_r}_{i} - \frac{{R_r}_{i}}{V} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i}&\dconp\\
-  \frac{\partial {R_r} }{\partial P }_{i} &= \frac{S^{\prime\prime}_{\ns} {k_r}_{i}}{T \Ru}&\dconv
+  \frac{\partial {R_r}_{i} }{\partial V } &= \frac{[C] S^{\prime\prime}_{\ns}}{V} {k_r}_{i} - \frac{{R_r}_{i}}{V} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i}&\dconp\\
+  \frac{\partial {R_r}_{i} }{\partial P } &= \frac{S^{\prime\prime}_{\ns} {k_r}_{i}}{T \Ru}&\dconv
  \end{align}
@@
  \begin{align}
-  \frac{\partial {R} }{\partial V }_{i} &= \frac{1}{V}\left(\vphantom{{R_r}_{i} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i}} [C]\left(S^{\prime}_{\ns} {k_f}_{i} - S^{\prime\prime}_{\ns} {k_r}_{i}\right) + \right.& \nonumber\\
+  \frac{\partial {R}_{i} }{\partial V } &= \frac{1}{V}\left(\vphantom{{R_r}_{i} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i}} [C]\left(S^{\prime}_{\ns} {k_f}_{i} - S^{\prime\prime}_{\ns} {k_r}_{i}\right) + \right.& \nonumber\\
   &\left. \quad\left({R_r}_{i} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i} - {R_f}_{i} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}\right) \right)&\dconp\\
-  \frac{\partial {R} }{\partial P }_{i} &= \frac{1}{T \Ru} \left(S^{\prime}_{\ns} {k_f}_{i} - S^{\prime\prime}_{\ns} {k_r}_{i}\right)&\dconv
+  \frac{\partial {R}_{i} }{\partial P } &= \frac{1}{T \Ru} \left(S^{\prime}_{\ns} {k_f}_{i} - S^{\prime\prime}_{\ns} {k_r}_{i}\right)&\dconv
  \end{align}
@@
 \subsection{Pressure modification\slash Falloff function derivatives}
+\subsubsection{Elementary reactions}
+For elementary reactions, all derivatives of the pressure-modification term are zero:
+\begin{align}
+\eqmathbox[LHS_dci_elem][r]{\frac{\partial c_{i} }{\partial T }} &= 0 & \eqmakebox[cond_dci_elem][r]{,} \label{e:dci_elem_dt} \\
+\eqmathbox[LHS_dci_elem][r]{\frac{\partial c_{i} }{\partial {n_j} }} &= 0 & \eqmakebox[cond_dci_elem][r]{,} \label{e:dci_elem_dnj} \\
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dci_elem_de}
+ \begin{align}
+  \eqmathbox[LHS_dci_elem][r]{\frac{\partial c_{i} }{\partial V }} &= 0 & \eqmakebox[cond_dci_elem][r]{\dconp} \\
+  \eqmathbox[LHS_dci_elem][r]{\frac{\partial c_{i} }{\partial P }} &= 0 & \eqmakebox[cond_dci_elem][r]{\dconv}
+ \end{align}
+\end{subequations}
+
+\subsubsection{Third-body enhanced reactions}
+For a third-body enhanced reaction~\cref{e:thd_mix} is differentiated to give:
+\begin{align}
+ \eqmathbox[LHS_dci][r]{\frac{\partial [X]_i }{\partial T }} &= \eqmathbox[RHS_dci][l]{-\frac{[C] \alpha_{\ns,i}}{T}} & \eqmakebox[cond_dci_thd][r]{,} \label{e:dci_thd_dt} \\
+ \eqmathbox[LHS_dci][r]{\frac{\partial [X]_i }{\partial n_{j}}} &= \eqmathbox[RHS_dci][l]{\frac{1}{V} \left(- \alpha_{\ns,i} + \alpha_{j,i}\right)} & \eqmakebox[cond_dci_thd][r]{,} \label{e:dci_thd_dnj}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dci_thd_de}
+ \begin{align}
+  \eqmathbox[LHS_dci][r]{\frac{\partial [X]_i }{\partial V }} &= \eqmathbox[RHS_dci][l]{\frac{1}{V} \left([C] \alpha_{\ns,i} - [X]_{i}\right)} & \eqmakebox[cond_dci_thd][r]{\dconp}\\
+  \eqmathbox[LHS_dci][r]{\frac{\partial [X]_i }{\partial P }} &= \eqmathbox[RHS_dci][l]{\frac{\alpha_{\ns,i}}{T \Ru}} & \eqmakebox[cond_dci_thd][r]{\dconv}
+ \end{align}
+\end{subequations}
+
+If $\alpha_{j,i} = 1$ for all species $j$,~\cref{e:dci_thd_dt,e:dci_thd_dnj,e:dci_thd_de} may be simplified to:
+\begin{align}
+ \eqmathbox[lhs_dci_thd_unity][r]{\frac{\partial c_{i} }{\partial T }} &= \eqmathbox[rhs_dci_thd_unity][l]{- \frac{[C]}{T}}& \eqmakebox[cond_dci_thd_unity][r]{,} \label{e:dci_thd_unity_dt} \\
+ \eqmathbox[lhs_dci_thd_unity][r]{\frac{\partial c_{i} }{\partial n_{j} }} &= \eqmathbox[rhs_dci_thd_unity][l]{0}& \eqmakebox[cond_dci_thd_unity][r]{,}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dci_thd_unity_de}
+ \begin{align}
+  \eqmathbox[lhs_dci_thd_unity][r]{\frac{\partial c_{i} }{\partial V }} &= \eqmathbox[rhs_dci_thd_unity][l]{0} & \eqmakebox[cond_dci_thd_unity][r]{\dconp} \\
+  \eqmathbox[lhs_dci_thd_unity][r]{\frac{\partial c_{i} }{\partial P }} &= \eqmathbox[rhs_dci_thd_unity][l]{\frac{1}{T \Ru}}&\eqmakebox[cond_dci_thd_unity][r]{\dconv}
+ \end{align}
+\end{subequations}
+
+If the third-body is the $m$th species,~\cref{e:dci_thd_dt,e:dci_thd_dnj,e:dci_thd_de} become:
+\begin{align}
+ \eqmathbox[lhs_dci_thd_spec][r]{\frac{\partial c_{i} }{\partial T }} &= \eqmathbox[rhs_dci_thd_spec][l]{-\frac{\delta_{\ns m}}{T} [C]}&\eqmakebox[cond_dci_thd_spec][r]{,} \label{e:dci_thd_spec_dt} \\
+ \eqmathbox[lhs_dci_thd_spec][r]{\frac{\partial c_{i} }{\partial n_j }} &= \eqmathbox[rhs_dci_thd_spec][l]{\frac{1}{V} \left(- \delta_{\ns m} + \delta_{j m}\right)}&\eqmakebox[cond_dci_thd_spec][r]{,} \label{e:dci_thd_spec_dnj} \\
+\end{align}
+and:
+\begin{subequations}
+\label{e:dci_thd_spec_de}
+\begin{align}
+\eqmathbox[lhs_dci_thd_spec][r]{\frac{\partial c_{i} }{\partial V }} &= \eqmathbox[rhs_dci_thd_spec][l]{\frac{\delta_{\ns m}}{V} \left([C] - [C]_{\ns}\right) + \frac{[C]_{m}}{V} \left(\delta_{\ns m} - 1\right)} & \eqmakebox[cond_dci_thd_spec][r]{\dconp} \\
+\eqmathbox[lhs_dci_thd_spec][r]{\frac{\partial c_{i} }{\partial P }} &= \eqmathbox[rhs_dci_thd_spec][l]{\frac{\delta_{\ns m}}{T \Ru}} & \eqmakebox[cond_dci_thd_spec][r]{\dconv}
+\end{align}
+\end{subequations}
+
+\subsubsection{Unimolecular\slash~recombination fall-off reactions}
+\label{s:dfall}
+Derivatives of the third-body pressure modification term for unimolecular\slash~recombination fall-off reactions---given by~\cref{e:rxn_pressure}---are quite involved, and will be broken up over the next sections.
+The derivative of full third-body pressure modification term for fall-off reactions will be given in this section, while~\cref{s:dchem} will provide the same derivatives for chemically-activated bimolecular reactions.
+Subsequently, the derivatives of the reduced pressure and falloff blending factors will be explored in~\cref{s:dpr,s:dfi} respectively.
+
+The derivatives of the unimolecular\slash~recombination fall-off reaction third-body pressure modification term are:
+\begin{align}
+ \eqmathbox[lhs_dci_fall][r]{\frac{\partial c_{i} }{\partial T }} &= \eqmathbox[rhs_dci_fall][l]{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial T } + \frac{\partial P_{r, i} }{\partial T } \left(F_{i} - c_{i}\right)\right)} & \eqmakebox[cond_dci_fall][r]{,} \label{e:dci_fall_dt}\\
+ \eqmathbox[lhs_dci_fall][r]{\frac{\partial c_{i} }{\partial n_j}} &= \eqmathbox[rhs_dci_fall][l]{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial n[j] } + \frac{\partial P_{r, i} }{\partial n_j } \left(F_{i} - c_{i}\right)\right)} & \eqmakebox[cond_dci_fall][r]{,} \label{e:dci_fall_dnj}
+\end{align}
+and with respect to the thermodynamic parameter:
+\begin{subequations}
+ \label{e:dci_fall_de}
+ \begin{align}
+  \eqmathbox[lhs_dci_fall][r]{\frac{\partial c_i }{\partial V }} &= \eqmathbox[rhs_dci_fall][l]{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial V } + \frac{\partial P_{r, i} }{\partial V } \left(F_{i} - c_{i}\right)\right)} & \eqmakebox[cond_dci_fall][r]{\dconp} \\
+  \eqmathbox[lhs_dci_fall][r]{\frac{\partial c_i }{\partial P }} &= \eqmathbox[rhs_dci_fall][l]{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial P } + \frac{\partial P_{r, i} }{\partial P } \left(F_{i} - c_{i}\right)\right)} & \eqmakebox[cond_dci_fall][r]{\dconv}
+ \end{align}
+\end{subequations}
+
+\subsubsection{Chemically-activated bimolecular reactions}
+\label{s:dchem}
+
+Similarly, the chemically-activated bimolecular reactions are as follows:
+\begin{align}
+ \label{e:dci_chem_dt}
+ \eqmathbox[lhs_dci_chem][r]{\frac{\partial c_i }{\partial T }} &= \eqmathbox[rhs_dci_chem][l]{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial T } - \frac{\partial P_{r, i} }{\partial T } c_{i}\right)} & \eqmakebox[cond_dci_chem][r]{,} \\
+ \label{e:dci_chem_dnj}
+ \eqmathbox[lhs_dci_chem][r]{\frac{\partial c_i }{\partial n_j }} &= \eqmathbox[rhs_dci_chem][l]{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial n_j } - \frac{\partial P_{r, i} }{\partial n_j } c_{i}\right)} & \eqmakebox[cond_dci_chem][r]{,}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dci_chem_de}
+ \begin{align}
+  \eqmathbox[lhs_dci_chem][r]{\frac{\partial c_i }{\partial V }} &= \eqmathbox[rhs_dci_chem][l]{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial V } - \frac{\partial P_{r, i} }{\partial V } c_{i}\right)} & \eqmakebox[cond_dci_chem][r]{\dconp} \\
+  \eqmathbox[lhs_dci_chem][r]{\frac{\partial c_i }{\partial P }} &= \eqmathbox[rhs_dci_chem][l]{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial P } - \frac{\partial P_{r, i} }{\partial P } c_{i}\right)} & \eqmakebox[cond_dci_chem][r]{\dconv}
+ \end{align}
+\end{subequations}
+
+\subsubsection{Reduced pressure}
+\label{s:dpr}
+
+Next the reduced pressure derivatives in~\crefrange{e:dci_fall_dt}{e:dci_chem_de} will be evaluated; these depend on the form of the third-body efficiency used in~\cref{e:pr}, thus we will first work out the derivatives for each third-body efficiency form, and then develop a generalized form for compact representation.
+For a third-body concentration based upon the entire mixture, (i.e.,~\cref{e:thd_mix}) the derivatives are:
+\begin{align}
+ \eqmathbox[lhs_dpr_mix][r]{\frac{\partial P_{r, i} }{\partial T }} &= \eqmathbox[rhs_dpr_mix][l]{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) - \frac{[C] k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i}}} & \eqmakebox[cond_dpr_mix][r]{,} \label{e:dpr_mix_dt} \\
+ \eqmathbox[lhs_dpr_mix][r]{\frac{\partial P_{r, i} }{\partial n_j }} &= \eqmathbox[rhs_dpr_mix][l]{\frac{k_{0, i} \left(- \alpha_{\ns,i} + \alpha_{j,i}\right)}{k_{\infty, i} V}} & \eqmakebox[cond_dpr_mix][r]{,} \label{e:dpr_mix_dnj}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dpr_mix_de}
+ \begin{align}
+  \eqmathbox[lhs_dpr_mix][r]{\frac{\partial P_{r, i} }{\partial V }} &= \eqmathbox[rhs_dpr_mix][l]{- \frac{P_{r, i}}{V} + \frac{[C] k_{0, i} \alpha_{\ns,i}}{V k_{\infty, i}}} & \eqmakebox[cond_dpr_mix][r]{\dconp} \\
+  \eqmathbox[lhs_dpr_mix][r]{\frac{\partial P_{r, i} }{\partial P }} &= \eqmathbox[rhs_dpr_mix][l]{\frac{k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i} \Ru}} & \eqmakebox[cond_dpr_mix][r]{\dconv}
+ \end{align}
+\end{subequations}
+
+\Crefrange{e:dpr_mix_dt}{e:dpr_mix_de} may be recast in a more generalized form using temporary variables $\Theta_{\partial P_{r, i}, \ldots}$ and $\bar{\theta}_{\partial P_{r, i}, \ldots}$ to replace terms in the derivative:
+\begin{align}
+ \eqmathbox[lhs_dpr_gen][r]{\frac{\partial P_{r, i} }{\partial T } } &= \eqmathbox[rhs_dpr_gen][l]{P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}} & \eqmakebox[cond_dpr_gen][r]{,} \label{e:dpr_dt} \\
+ \eqmathbox[lhs_dpr_gen][r]{\frac{\partial P_{r, i} }{\partial n_j}} &= \eqmathbox[rhs_dpr_gen][l]{\frac{\bar{\theta}_{P_{r, i}, \partial n_j}}{V k_{\infty, i}} k_{0, i}} & \eqmakebox[cond_dpr_gen][r]{,} \label{e:dpr_dnj}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dpr_de}
+ \begin{align}
+  \eqmathbox[lhs_dpr_gen][r]{\frac{\partial P_{r, i} }{\partial V}} &= \eqmathbox[rhs_dpr_gen][l]{P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}} & \eqmakebox[cond_dpr_gen][r]{\dconp} \\
+  \eqmathbox[lhs_dpr_gen][r]{\frac{\partial P_{r, i} }{\partial P}} &= \eqmathbox[rhs_dpr_gen][l]{\bar{\theta}_{P_{r, i}, \partial P}} & \eqmakebox[cond_dpr_gen][r]{\dconv}
+ \end{align}
+\end{subequations}
+with:
+\begin{subequations}
+\label{e:dpr_theta_mix}
+\begin{align}
+ \Theta_{P_{r,i}, \partial T} &= \frac{1}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) \label{e:dtheta_pr_dt} \\
+ \bar{\theta}_{P_{r, i}, \partial T} &= - \frac{[C] k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i}} \\
+ \bar{\theta}_{P_{r, i}, \partial n_j} &= - \alpha_{\ns,i} + \alpha_{j,i} \\
+ \Theta_{P_{r,i}, \partial V} &= - \frac{1}{V} \\
+ \bar{\theta}_{P_{r, i}, \partial V} &= \frac{[C] k_{0, i} \alpha_{\ns,i}}{V k_{\infty, i}} \\
+ \bar{\theta}_{P_{r, i}, \partial P} &= \frac{k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i} \Ru}
+\end{align}
+\end{subequations}
+
+For a third-body efficiency with all $\alpha_{k,i} = 1$, the reduced pressures derivatives may be expressed as~\cref{e:dpr_de}, with:
+\begin{subequations}
+ \label{e:dpr_theta_unity}
+ \begin{align}
+  \Theta_{P_{r,i}, \partial T} &= \frac{1}{T} \left(\beta_{0} - \beta_{\infty} - 1 + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) \\
+  \bar{\theta}_{P_{r, i}, \partial T} &= 0 \\
+  \bar{\theta}_{P_{r, i}, \partial n_j} &= 0 \\
+  \Theta_{P_{r,i}, \partial V} &= 0 \\
+  \bar{\theta}_{P_{r, i}, \partial V} &= 0 \\
+  \bar{\theta}_{P_{r, i}, \partial P} &= \frac{k_{0, i}}{k_{\infty, i}} \frac{1}{\Ru T}
+ \end{align}
+\end{subequations}
+
+If species $m$ serves as the third-body concentration, the temporary terms are:
+\begin{subequations}
+ \label{e:dpr_theta_spec}
+ \begin{align}
+  \Theta_{P_{r,i}, \partial T} &= \frac{1}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) \\
+  \bar{\theta}_{P_{r, i}, \partial T} &= - \frac{[C] k_{0, i} \delta_{\ns m}}{T k_{\infty, i}} \\
+  \bar{\theta}_{P_{r, i}, \partial n_j} &= - \delta_{\ns m} + \delta_{j m} \\
+  \Theta_{P_{r,i}, \partial V} &= - \frac{1}{V} \\
+  \bar{\theta}_{P_{r, i}, \partial V} &= \frac{[C] k_{0, i} \delta_{\ns m}}{V k_{\infty, i}} \\
+  \bar{\theta}_{P_{r, i}, \partial P} &= \frac{k_{0, i} \delta_{\ns m}}{T k_{\infty, i} \Ru}
+ \end{align}
+\end{subequations}
+
+\subsubsection{Falloff blending factor}
+\label{s:dfi}
 

### 2018-01-02 derivations.tex
add pr derivatives

@@
 \newcommand{\conv}{CONV}
-\newcommand{\dconp}{\ensuremath{\text{for \conp,}}}
-\newcommand{\dconv}{\ensuremath{\text{for \conv,}}}
+\newcommand{\dconp}{\ifmmode\text{for \conp,}\else for \conp,\fi}
+\newcommand{\dconv}{\ifmmode\text{for \conv,}\else for \conv,\fi}
 \newcommand{\Ru}{\ensuremath{\mathcal{R}}}
@@
 \MHInternalSyntaxOff
+
+% allow page breaks for long equations:
+% https://tex.stackexchange.com/questions/51682/is-it-possible-to-pagebreak-aligned-equations
+\allowdisplaybreaks
 
@@
 
-\Crefrange{e:dpr_mix_dt}{e:dpr_mix_de} may be recast in a more generalized form using temporary variables $\Theta_{\partial P_{r, i}, \ldots}$ and $\bar{\theta}_{\partial P_{r, i}, \ldots}$ to replace terms in the derivative:
+For a third-body concentration where all $\alpha_{k,i}=1$:
+\begin{align}
+ \eqmathbox[lhs_dpr_unity][r]{\frac{\partial P_{r, i} }{\partial T }} &= \eqmathbox[rhs_dpr_unity][l]{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} - 1 + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right)} & \eqmakebox[cond_dpr_unity][r]{,} \label{e:dpr_unity_dt} \\
+ \eqmathbox[lhs_dpr_unity][r]{\frac{\partial P_{r, i} }{\partial n_j }} &= \eqmathbox[rhs_dpr_unity][l]{0} & \eqmakebox[cond_dpr_unity][r]{,} \label{e:dpr_unity_dnj}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dpr_unity_de}
+ \begin{align}
+  \eqmathbox[lhs_dpr_unity][r]{\frac{\partial P_{r, i} }{\partial V }} &= \eqmathbox[rhs_dpr_unity][l]{0} & \eqmakebox[cond_dpr_unity][r]{\dconp}\\
+  \eqmathbox[lhs_dpr_unity][r]{\frac{\partial P_{r, i} }{\partial P }} &= \eqmathbox[rhs_dpr_unity][l]{\frac{k_{0, i}}{k_{\infty, i}} \frac{1}{\Ru T}} & \eqmakebox[cond_dpr_unity][r]{\dconv}
+ \end{align}
+\end{subequations}
+
+Finally, for species $m$ as the third-body:
+\begin{align}
+ \eqmathbox[lhs_dpr_spec][r]{\frac{\partial P_{r, i} }{\partial T }} &= \eqmathbox[rhs_dpr_spec][l]{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) - \frac{[C] k_{0, i} \delta_{\ns m}}{T k_{\infty, i}}} & \eqmakebox[cond_dpr_spec][r]{,} \label{e:dpr_spec_dt} \\
+ \eqmathbox[lhs_dpr_spec][r]{\frac{\partial P_{r, i} }{\partial n_j }} &= \eqmathbox[rhs_dpr_spec][l]{\frac{k_{0, i}}{V k_{\infty, i}} \left(- \delta_{\ns m} + \delta_{j m}\right)} & \eqmakebox[cond_dpr_spec][r]{,} \label{e:dpr_spec_dnj}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dpr_spec_de}
+ \begin{align}
+  \eqmathbox[lhs_dpr_spec][r]{\frac{\partial P_{r, i} }{\partial V }} &= \eqmathbox[rhs_dpr_unity][l]{- \frac{P_{r, i}}{V} + \frac{[C] k_{0, i} \delta_{\ns m}}{V k_{\infty, i}}} & \eqmakebox[cond_dpr_spec][r]{\dconp}\\
+  \eqmathbox[lhs_dpr_spec][r]{\frac{\partial P_{r, i} }{\partial P }} &= \eqmathbox[rhs_dpr_unity][l]{\frac{k_{0, i} \delta_{\ns m}}{T k_{\infty, i} \Ru}} & \eqmakebox[cond_dpr_spec][r]{\dconv}
+ \end{align}
+\end{subequations}
+
+\Crefrange{e:dpr_mix_dt}{e:dpr_spec_de} may be recast in a more generalized form using temporary variables $\Theta_{\partial P_{r, i}, \ldots}$ and $\bar{\theta}_{\partial P_{r, i}, \ldots}$ to replace terms in the derivative:
 \begin{align}
@@
 with:
-\begin{subequations}
-\label{e:dpr_theta_mix}
 \begin{align}
- \Theta_{P_{r,i}, \partial T} &= \frac{1}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) \label{e:dtheta_pr_dt} \\
- \bar{\theta}_{P_{r, i}, \partial T} &= - \frac{[C] k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i}} \\
- \bar{\theta}_{P_{r, i}, \partial n_j} &= - \alpha_{\ns,i} + \alpha_{j,i} \\
- \Theta_{P_{r,i}, \partial V} &= - \frac{1}{V} \\
- \bar{\theta}_{P_{r, i}, \partial V} &= \frac{[C] k_{0, i} \alpha_{\ns,i}}{V k_{\infty, i}} \\
- \bar{\theta}_{P_{r, i}, \partial P} &= \frac{k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i} \Ru}
+ \label{e:dpr_theta}
+ \Theta_{P_{r,i}, \partial T} &=
+ \begin{dcases}
+  \frac{1}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) & \text{if mix,} \\
+  \frac{1}{T} \left(\beta_{0} - \beta_{\infty} - 1 + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) & \text{if all $\alpha_{k,i}$ = 1,} \\
+  \frac{1}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) & \text{for species $m$}
+ \end{dcases} \\
+ \bar{\theta}_{P_{r, i}, \partial T} &=
+ \begin{dcases}
+  \eqmathbox[theta_pr_bar_dt][l]{- \frac{[C] k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i}}} & \text{if mix,} \\
+  \eqmathbox[theta_pr_bar_dt][c]{0} & \text{if all $\alpha_{k,i}$ = 1,} \\
+  \eqmathbox[theta_pr_bar_dt][l]{- \frac{[C] k_{0, i} \delta_{\ns m}}{T k_{\infty, i}}} & \text{for species $m$}
+ \end{dcases} \\
+ \bar{\theta}_{P_{r, i}, \partial n_j} &=
+ \begin{dcases}
+  \eqmathbox[theta_pr_bar_dnj][l]{- \alpha_{\ns,i} + \alpha_{j,i}} & \text{if mix,} \\
+  \eqmathbox[theta_pr_bar_dnj][c]{0} & \text{if all $\alpha_{k,i}$ = 1,} \\
+  \eqmathbox[theta_pr_bar_dnj][l]{- \delta_{\ns m} + \delta_{j m}} & \text{for species $m$}
+ \end{dcases} \\
+ \Theta_{P_{r,i}, \partial V} &=
+ \begin{dcases}
+  \eqmathbox[Theta_pr_bar_dv][l]{- \frac{1}{V}} & \text{if mix,} \\
+  \eqmathbox[Theta_pr_bar_dv][c]{0} & \text{if all $\alpha_{k,i}$ = 1,} \\
+  \eqmathbox[Theta_pr_bar_dv][l]{- \frac{1}{V}} & \text{for species $m$}
+ \end{dcases} \\
+ \bar{\theta}_{P_{r, i}, \partial V} &=
+ \begin{dcases}
+  \eqmathbox[theta_pr_bar_dv][l]{\frac{[C] k_{0, i} \alpha_{\ns,i}}{V k_{\infty, i}}} & \text{if mix,} \\
+  \eqmathbox[theta_pr_bar_dv][c]{0} & \text{if all $\alpha_{k,i}$ = 1,} \\
+  \eqmathbox[theta_pr_bar_dv][l]{\frac{[C] k_{0, i} \delta_{\ns m}}{V k_{\infty, i}}} & \text{for species $m$}
+ \end{dcases} \\
+ \bar{\theta}_{P_{r, i}, \partial P} &=
+ \begin{dcases}
+  \eqmathbox[theta_pr_bar_dp][l]{\frac{k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i} \Ru}} & \text{if mix,} \\
+  \eqmathbox[theta_pr_bar_dp][l]{\frac{k_{0, i}}{T k_{\infty, i} \Ru}} & \text{if all $\alpha_{k,i}$ = 1,} \\
+  \eqmathbox[theta_pr_bar_dp][l]{\frac{k_{0, i} \delta_{\ns m}}{T k_{\infty, i} \Ru}} & \text{for species $m$}
+ \end{dcases} \\
 \end{align}
-\end{subequations}
-
-For a third-body efficiency with all $\alpha_{k,i} = 1$, the reduced pressures derivatives may be expressed as~\cref{e:dpr_de}, with:
-\begin{subequations}
- \label{e:dpr_theta_unity}
- \begin{align}
-  \Theta_{P_{r,i}, \partial T} &= \frac{1}{T} \left(\beta_{0} - \beta_{\infty} - 1 + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) \\
-  \bar{\theta}_{P_{r, i}, \partial T} &= 0 \\
-  \bar{\theta}_{P_{r, i}, \partial n_j} &= 0 \\
-  \Theta_{P_{r,i}, \partial V} &= 0 \\
-  \bar{\theta}_{P_{r, i}, \partial V} &= 0 \\
-  \bar{\theta}_{P_{r, i}, \partial P} &= \frac{k_{0, i}}{k_{\infty, i}} \frac{1}{\Ru T}
- \end{align}
-\end{subequations}
-
-If species $m$ serves as the third-body concentration, the temporary terms are:
-\begin{subequations}
- \label{e:dpr_theta_spec}
- \begin{align}
-  \Theta_{P_{r,i}, \partial T} &= \frac{1}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) \\
-  \bar{\theta}_{P_{r, i}, \partial T} &= - \frac{[C] k_{0, i} \delta_{\ns m}}{T k_{\infty, i}} \\
-  \bar{\theta}_{P_{r, i}, \partial n_j} &= - \delta_{\ns m} + \delta_{j m} \\
-  \Theta_{P_{r,i}, \partial V} &= - \frac{1}{V} \\
-  \bar{\theta}_{P_{r, i}, \partial V} &= \frac{[C] k_{0, i} \delta_{\ns m}}{V k_{\infty, i}} \\
-  \bar{\theta}_{P_{r, i}, \partial P} &= \frac{k_{0, i} \delta_{\ns m}}{T k_{\infty, i} \Ru}
- \end{align}
-\end{subequations}
 
@@
 \label{s:dfi}
+
+The falloff blending factor
 

### 2018-01-03 derivations.tex
use custom macros to make life easy

--- 2017-12-21 derivations.tex
@@
 
+% https://tex.stackexchange.com/a/34412/5764
+\makeatletter
+\NewDocumentCommand{\eqmathbox}{o O{c} m}{%
+  \IfValueTF{#1}
+    {\def\eqmathbox@##1##2{\eqmakebox[#1][#2]{$##1##2$}}}
+    {\def\eqmathbox@##1##2{\eqmakebox{$##1##2$}}}
+  \mathpalette\eqmathbox@{#3}
+}
+\makeatother
+
 \newcommand{\pluseq}{\mathrel{{+}{=}}}
 \newcommand{\minuseq}{\mathrel{{-}{=}}}
-\newcommand{\ns}{\ensuremath{{N_{sp}}}}
-\newcommand{\nr}{\ensuremath{{N_{reac}}}}
+\newcommand{\ns}{N_{sp}}
+\newcommand{\nr}{N_{reac}}
 \newcommand{\conp}{CONP}
 \newcommand{\conv}{CONV}
-\newcommand{\dconp}{\ensuremath{,\qquad\text{for \conp}}}
-\newcommand{\dconv}{\ensuremath{,\qquad\text{for \conv}}}
-\newcommand{\Ru}{\ensuremath{\mathcal{R}}}
+\newcommand{\dconp}{\ifmmode\text{for \conp,}\else for \conp,\fi}
+\newcommand{\dconv}{\ifmmode\text{for \conv,}\else for \conv,\fi}
+\newcommand{\Ru}{\mathcal{R}}
+
+% simple command for equal-sized eqmathbox's on the lhs / rhs and conditional
+% arguements are tag, text, alignment
+\NewDocumentCommand{\lhs}{mmO{r}}{\eqmathbox[LHS_#1][#3]{#2}}
+\NewDocumentCommand{\rhs}{mmO{l}}{\eqmathbox[RHS_#1][#3]{#2}}
+\NewDocumentCommand{\cond}{mmO{r}}{\eqmakebox[COND_#1][#3]{#2}}
+
+% treat subequations as equation for cleverref
+% https://tex.stackexchange.com/a/308627/56227
+\crefalias{subequation}{equation}
 
@@
 \end{equation}
+This form will be denoted as the \textbf{mixture}-based third-body efficiency.
+
 If all species in the mixture contribute equally as third bodies, then $\alpha_{k,i} = 1$ for all species:
@@
 \end{equation}
+This form will be called the \textbf{unity}-based third-body efficiency.
+
 In addition, a single species may act as the third body in which case
@@
 will be used, where the Kronecker delta $\delta_{\ns m}$ is unity if and only if the third body species $m$ is the last species in the model.
+\Cref{e:thd_spec} will be referred to as the \textbf{species}-based third-body efficiency.
 
@@
 \subsection{Pressure modification\slash Falloff function derivatives}
+\subsubsection{Elementary reactions}
+For elementary reactions, all derivatives of the pressure-modification term are zero:
+\begin{align}
+\lhs{dci_elem}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{,} \label{e:dci_elem_dt} \\
+\lhs{dci_elem}{\frac{\partial c_{i} }{\partial {n_j} }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{,} \label{e:dci_elem_dnj} \\
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dci_elem_de}
+ \begin{align}
+  \lhs{dci_elem}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{\dconp} \\
+  \lhs{dci_elem}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{\dconv}
+ \end{align}
+\end{subequations}
+
+\subsubsection{Third-body enhanced reactions}
+For a third-body enhanced reaction~\cref{e:thd_mix} is differentiated to give:
+\begin{align}
+ \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial T }} &= \rhs{dci_thd_mix}{-\frac{[C] \alpha_{\ns,i}}{T}} & \cond{dci_thd_mix}{,} \label{e:dci_thd_dt} \\
+ \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial n_{j}}} &= \rhs{dci_thd_mix}{\frac{1}{V} \left(- \alpha_{\ns,i} + \alpha_{j,i}\right)} & \cond{dci_thd_mix}{,} \label{e:dci_thd_dnj}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dci_thd_de}
+ \begin{align}
+  \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial V }} &= \rhs{dci_thd_mix}{\frac{1}{V} \left([C] \alpha_{\ns,i} - [X]_{i}\right)} & \cond{dci_thd_mix}{\dconp}\\
+  \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial P }} &= \rhs{dci_thd_mix}{\frac{\alpha_{\ns,i}}{T \Ru}} & \cond{dci_thd_mix}{\dconv}
+ \end{align}
+\end{subequations}
+
+If $\alpha_{j,i} = 1$ for all species $j$,~\cref{e:dci_thd_dt,e:dci_thd_dnj,e:dci_thd_de} may be simplified to:
+\begin{align}
+ \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_thd_unity}{- \frac{[C]}{T}}& \cond{dci_thd_unity}{,} \label{e:dci_thd_unity_dt} \\
+ \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial n_{j} }} &= \rhs{dci_thd_unity}{0}& \cond{dci_thd_unity}{,}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dci_thd_unity_de}
+ \begin{align}
+  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_thd_unity}{0} & \cond{dci_thd_unity}{\dconp} \\
+  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_thd_unity}{\frac{1}{T \Ru}}& \cond{dci_thd_unity}{\dconv}
+ \end{align}
+\end{subequations}
+
+If the third-body is the $m$th species,~\cref{e:dci_thd_dt,e:dci_thd_dnj,e:dci_thd_de} become:
+\begin{align}
+ \lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_thd_spec}{-\frac{\delta_{\ns m}}{T} [C]}&\cond{dci_thd_spec}{,} \label{e:dci_thd_spec_dt} \\
+ \lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial n_j }} &= \rhs{dci_thd_spec}{\frac{1}{V} \left(- \delta_{\ns m} + \delta_{j m}\right)}&\cond{dci_thd_spec}{,} \label{e:dci_thd_spec_dnj} \\
+\end{align}
+and:
+\begin{subequations}
+\label{e:dci_thd_spec_de}
+\begin{align}
+\lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_thd_spec}{\frac{\delta_{\ns m}}{V} \left([C] - [C]_{\ns}\right) + \frac{[C]_{m}}{V} \left(\delta_{\ns m} - 1\right)} & \cond{dci_thd_spec}{\dconp} \\
+\lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_thd_spec}{\frac{\delta_{\ns m}}{T \Ru}} & \cond{dci_thd_spec}{\dconv}
+\end{align}
+\end{subequations}
+
+\subsubsection{Unimolecular\slash~recombination fall-off reactions}
+\label{s:dfall}
+Derivatives of the third-body pressure modification term for unimolecular\slash~recombination fall-off reactions---given by~\cref{e:rxn_pressure}---are quite involved, and will be broken up over the next sections.
+The derivative of full third-body pressure modification term for fall-off reactions will be given in this section, while~\cref{s:dchem} will provide the same derivatives for chemically-activated bimolecular reactions.
+Subsequently, the derivatives of the reduced pressure and falloff blending factors will be explored in~\cref{s:dpr,s:dfi} respectively.
+
+The derivatives of the unimolecular\slash~recombination fall-off reaction third-body pressure modification term are:
+\begin{align}
+ \lhs{dci_fall}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial T } + \frac{\partial P_{r, i} }{\partial T } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{,} \label{e:dci_fall_dt}\\
+ \lhs{dci_fall}{\frac{\partial c_{i} }{\partial n_j}} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial n[j] } + \frac{\partial P_{r, i} }{\partial n_j } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{,} \label{e:dci_fall_dnj}
+\end{align}
+and with respect to the thermodynamic parameter:
+\begin{subequations}
+ \label{e:dci_fall_de}
+ \begin{align}
+  \lhs{dci_fall}{\frac{\partial c_i }{\partial V }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial V } + \frac{\partial P_{r, i} }{\partial V } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{\dconp} \\
+  \lhs{dci_fall}{\frac{\partial c_i }{\partial P }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial P } + \frac{\partial P_{r, i} }{\partial P } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{\dconv}
+ \end{align}
+\end{subequations}
+
+\subsubsection{Chemically-activated bimolecular reactions}
+\label{s:dchem}
+
+Similarly, the chemically-activated bimolecular reactions are as follows:
+\begin{align}
+ \label{e:dci_chem_dt}
+ \lhs{dci_chem}{\frac{\partial c_i }{\partial T }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial T } - \frac{\partial P_{r, i} }{\partial T } c_{i}\right)} & \cond{dci_chem}{,} \\
+ \label{e:dci_chem_dnj}
+ \lhs{dci_chem}{\frac{\partial c_i }{\partial n_j }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial n_j } - \frac{\partial P_{r, i} }{\partial n_j } c_{i}\right)} & \cond{dci_chem}{,}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dci_chem_de}
+ \begin{align}
+  \lhs{dci_chem}{\frac{\partial c_i }{\partial V }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial V } - \frac{\partial P_{r, i} }{\partial V } c_{i}\right)} & \cond{dci_chem}{\dconp} \\
+  \lhs{dci_chem}{\frac{\partial c_i }{\partial P }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial P } - \frac{\partial P_{r, i} }{\partial P } c_{i}\right)} & \cond{dci_chem}{\dconv}
+ \end{align}
+\end{subequations}
+
+\subsubsection{Reduced pressure}
+\label{s:dpr}
+
+Next the reduced pressure derivatives in~\crefrange{e:dci_fall_dt}{e:dci_chem_de} will be evaluated; these depend on the form of the third-body efficiency used in~\cref{e:pr}, thus we will first work out the derivatives for each third-body efficiency form, and then develop a generalized form for compact representation.
+For a third-body concentration based upon the entire mixture, (i.e.,~\cref{e:thd_mix}) the derivatives are:
+\begin{align}
+ \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial T }} &= \rhs{dpr_mix}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) - \frac{[C] k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i}}} & {,} \label{e:dpr_mix_dt} \\
+ \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial n_j }} &= \rhs{dpr_mix}{\frac{k_{0, i} \left(- \alpha_{\ns,i} + \alpha_{j,i}\right)}{k_{\infty, i} V}} & \cond{dpr_mix}{,} \label{e:dpr_mix_dnj}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dpr_mix_de}
+ \begin{align}
+  \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_mix}{- \frac{P_{r, i}}{V} + \frac{[C] k_{0, i} \alpha_{\ns,i}}{V k_{\infty, i}}} & \cond{dpr_mix}{\dconp} \\
+  \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_mix}{\frac{k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i} \Ru}} & \cond{dpr_mix}{\dconv}
+ \end{align}
+\end{subequations}
+
+For a unity third-body concentration (i.e., all $\alpha_{k,i}=1$):
+\begin{align}
+ \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial T }} &= \rhs{dpr_unity}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} - 1 + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right)} & \cond{dpr_unity}{,} \label{e:dpr_unity_dt} \\
+ \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial n_j }} &= \rhs{dpr_unity}{0} & \cond{dpr_unity}{,} \label{e:dpr_unity_dnj}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dpr_unity_de}
+ \begin{align}
+  \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_unity}{0} & \cond{dpr_unity}{\dconp}\\
+  \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_unity}{\frac{k_{0, i}}{k_{\infty, i}} \frac{1}{\Ru T}} & \cond{dpr_unity}{\dconv}
+ \end{align}
+\end{subequations}
+
+Finally, for species-based third-body (i.e., the species $m$ alone acts as the third body):
+\begin{align}
+ \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial T }} &=\rhs{dpr_spec}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) - \frac{[C] k_{0, i} \delta_{\ns m}}{T k_{\infty, i}}} & \cond{dpr_spec}{,} \label{e:dpr_spec_dt} \\
+ \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial n_j }} &=\rhs{dpr_spec}{\frac{k_{0, i}}{V k_{\infty, i}} \left(- \delta_{\ns m} + \delta_{j m}\right)} & \cond{dpr_spec}{,} \label{e:dpr_spec_dnj}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dpr_spec_de}
+ \begin{align}
+  \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_spec}{- \frac{P_{r, i}}{V} + \frac{[C] k_{0, i} \delta_{\ns m}}{V k_{\infty, i}}} & \cond{dpr_spec}{\dconp}\\
+  \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_spec}{\frac{k_{0, i} \delta_{\ns m}}{T k_{\infty, i} \Ru}} & \cond{dpr_spec}{\dconv}
+ \end{align}
+\end{subequations}
+
+\Crefrange{e:dpr_mix_dt}{e:dpr_spec_de} may be recast in a more generalized form using temporary variables $\Theta_{\partial P_{r, i}, \ldots}$ and $\bar{\theta}_{\partial P_{r, i}, \ldots}$ to replace terms in the derivative:
+\begin{align}
+ \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial T } } &= \rhs{dpr_gen}{P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}} & \cond{dpr_gen}{,} \label{e:dpr_dt} \\
+ \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial n_j}} &= \rhs{dpr_gen}{\frac{\bar{\theta}_{P_{r, i}, \partial n_j}}{V k_{\infty, i}} k_{0, i}} & \cond{dpr_gen}{,} \label{e:dpr_dnj}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dpr_de}
+ \begin{align}
+  \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial V}} &= \rhs{dpr_gen}{P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}} & \cond{dpr_gen}{\dconp} \\
+  \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial P}} &= \rhs{dpr_gen}{\bar{\theta}_{P_{r, i}, \partial P}} & \cond{dpr_gen}{\dconv}
+ \end{align}
+\end{subequations}
+with:
+\begin{align}
+ \label{e:dpr_theta}
+ \Theta_{P_{r,i}, \partial T} &=
+ \begin{dcases}
+  \frac{1}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) & \text{if mix,} \\
+  \frac{1}{T} \left(\beta_{0} - \beta_{\infty} - 1 + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) & \text{if unity,} \\
+  \frac{1}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) & \text{if species,}
+ \end{dcases} \\
+ \bar{\theta}_{P_{r, i}, \partial T} &=
+ \begin{dcases}
+  \rhs{theta_pr_bar_dt}{- \frac{[C] k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i}}} & \text{if mix,} \\
+  \rhs{theta_pr_bar_dt}{0}[c] & \text{if unity,} \\
+  \rhs{theta_pr_bar_dt}{- \frac{[C] k_{0, i} \delta_{\ns m}}{T k_{\infty, i}}} & \text{if species,}
+ \end{dcases} \\
+ \bar{\theta}_{P_{r, i}, \partial n_j} &=
+ \begin{dcases}
+  \rhs{theta_pr_bar_dnj}{- \alpha_{\ns,i} + \alpha_{j,i}} & \text{if mix,} \\
+  \rhs{theta_pr_bar_dnj}{0}[c] & \text{if unity,}  \\
+  \rhs{theta_pr_bar_dnj}{- \delta_{\ns m} + \delta_{j m}} & \text{if species,}
+ \end{dcases} \\
+ \Theta_{P_{r,i}, \partial V} &=
+ \begin{dcases}
+  \rhs{Theta_pr_bar_dv}{- \frac{1}{V}} & \text{if mix,} \\
+  \rhs{Theta_pr_bar_dv}{0}[c] & \text{if unity,}  \\
+  \rhs{Theta_pr_bar_dv}{- \frac{1}{V}} & \text{if species,}
+ \end{dcases} \\
+ \bar{\theta}_{P_{r, i}, \partial V} &=
+ \begin{dcases}
+  \rhs{theta_pr_bar_dv}{\frac{[C] k_{0, i} \alpha_{\ns,i}}{V k_{\infty, i}}} & \text{if mix,} \\
+  \rhs{theta_pr_bar_dv}{0}[c] & \text{if unity,}  \\
+  \rhs{theta_pr_bar_dv}{\frac{[C] k_{0, i} \delta_{\ns m}}{V k_{\infty, i}}} & \text{if species,}
+ \end{dcases} \\
+ \bar{\theta}_{P_{r, i}, \partial P} &=
+ \begin{dcases}
+  \rhs{theta_pr_bar_dp}{\frac{k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i} \Ru}} & \text{if mix,} \\
+  \rhs{theta_pr_bar_dp}{\frac{k_{0, i}}{T k_{\infty, i} \Ru}} & \text{if unity,}  \\
+  \rhs{theta_pr_bar_dp}{\frac{k_{0, i} \delta_{\ns m}}{T k_{\infty, i} \Ru}} & \text{if species,}
+ \end{dcases} \\
+\end{align}
+
+\subsubsection{Falloff blending factor}
+\label{s:dfi}
+
+The falloff blending factor
 

### 2018-01-03 derivations.tex
updates

@@
 \NewDocumentCommand{\cond}{mmO{r}}{\eqmakebox[COND_#1][#3]{#2}}
+
+% consistent multiline equation spacing
+\NewDocumentCommand{\mathindent}{O{space}}{\hphantom{\mathrel{#1}}}
+
+% bigger parenthesis
+% https://tex.stackexchange.com/questions/6794/about-big-parenthesis-larger-than-bigg
+\makeatletter
+\newcommand{\vast}{\bBigg@{4}}
+\newcommand{\Vast}{\bBigg@{5}}
+\makeatother
 
@@
 
-The falloff blending factor
+The falloff blending factor for all but Lindemann falloff reactions relies upon the reduced pressure derivatives developed in~\cref{s:dpr}.
+First, the blending factor for each falloff reaction type will be differentiated along with any subcomponents (e.g., the $F_{cent}$ term for Troe falloff reactions).
+Next, the reduced pressured derivatives given by~\cref{e:dpr_dt,e:dpr_dnj,e:dpr_de} will be subsituted in, before finally deriving an analagous generalized form for the falloff blending factor derivatives.
+
+For Lindemann falloff reactions:
+\begin{align}
+ \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial T }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{,} \label{e:dfi_lind_dt} \\
+ \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial n_j }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{,} \label{e:dfi_lind_dnj}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dfi_lind_de}
+ \begin{align}
+  \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{\dconp} \\
+  \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{\dconv}
+ \end{align}
+\end{subequations}
+
+For Troe falloff reactions:
+\begin{align}
+ \lhs{dfi_troe}{\frac{\partial F_{i} }{\partial T }} &= \rhs{dfi_troe}{\frac{\partial F_{i} }{\partial F_{cent} } \frac{\text{d} F_{cent} }{\text{d} T } + \frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial T }} & \cond{dfi_troe}{,} \label{e:dfi_troe_dt} \\
+ \lhs{dfi_troe}{\frac{\partial F_{i} }{\partial n_j }} &= \rhs{dfi_troe}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial n_j }} & \cond{dfi_troe}{,} \label{e:dfi_troe_dnj}
+\end{align}
+and:
+\begin{subequations}
+ \label{e:dfi_troe_de}
+ \begin{align}
+ \lhs{dfi_troe}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_troe}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial V }} & \cond{dfi_troe}{\dconp} \\
+ \lhs{dfi_troe}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_troe}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial P }} & \cond{dfi_troe}{\dconv}
+ \end{align}
+\end{subequations}
+where:
+\begin{align}
+ \frac{\partial F_{i} }{\partial F_{cent} } &= \frac{F_{i}}{\frac{A_{Troe}^{2}}{B_{Troe}^{2}} + 1} \left(\frac{2 A_{Troe} \log{\left (F_{cent} \right )}}{B_{Troe}^{2} \left(\frac{A_{Troe}^{2}}{B_{Troe}^{2}} + 1\right)} \left(\frac{A_{Troe}}{B_{Troe}} \frac{\partial B_{Troe} }{\partial F_{cent} } - \right.
+\right.
+& \nonumber \\
+ & \mathindent \left.
+\frac{\partial A_{Troe} }{\partial F_{cent} }\right) + \frac{1}{F_{cent}} \left.
+\vphantom{\frac{F_{i}}{\frac{A_{Troe}^{2}}{B_{Troe}^{2}} + 1}} \right) & \text{,} \label{e:dfi_troe_dfcent_partial} \\
+\frac{\text{d} F_{cent} }{\text{d} T } &= \frac{a}{T^{*}} \operatorname{exp}\left({- \frac{T}{T^{*}}}\right) - \frac{\operatorname{exp}\left({- \frac{T}{T^{***}}}\right)}{T^{***}} \left(- a + 1\right) + & \nonumber \\
+ & \mathindent \frac{T^{**}}{T^{2}} \operatorname{exp}\left({- \frac{T^{**}}{T}}\right) & \text{,} \label{e:dfcent_dt} \\
+\intertext{and:}
+\frac{\partial F_{i} }{\partial P_{r, i} } &= \frac{2 F_{i} A_{Troe} \log{\left (F_{cent} \right )}}{B_{Troe}^{2} \left(\frac{A_{Troe}^{2}}{B_{Troe}^{2}} + 1\right)^{2}} \left(\frac{A_{Troe}}{B_{Troe}} \frac{\partial B_{Troe} }{\partial P_{r, i} } - \frac{\partial A_{Troe} }{\partial P_{r, i} }\right) \label{e:dfi_troe_dpr_partial}
+\end{align}
+where the derivatives of $A_{Troe}$ and $B_{Troe}$ are:
+\begin{align}
+ \frac{\partial A_{Troe} }{\partial F_{cent} } &= - \frac{0.67}{F_{cent} \log{\left (10 \right )}} & \text{,} \label{e:datroe_dfcent} \\
+\frac{\partial B_{Troe} }{\partial F_{cent} } &= - \frac{1.1762}{F_{cent} \log{\left (10 \right )}} & \text{,} \label{e:dbtroe_dfcent} \\
+\frac{\partial A_{Troe} }{\partial P_{r, i} } &= \frac{1}{P_{r, i} \log{\left (10 \right )}} & \text{,} \label{e:datroe_dpr} \\
+\intertext{and:}
+\frac{\partial B_{Troe} }{\partial P_{r, i} } &= - \frac{0.14}{P_{r, i} \log{\left (10 \right )}} & \text{.} \label{e:dbtroe_dpr} \\
+\end{align}
+
+Combining~\cref{e:datroe_dfcent,e:dbtroe_dfcent,e:datroe_dpr,e:dbtroe_dpr} with~\cref{e:dfi_troe_dfcent_partial,e:dfi_troe_dpr_partial} yields:
+\begin{align}
+\frac{\partial F_{i} }{\partial F_{cent} } &= - \frac{F_{i} B_{Troe}}{F_{cent} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}} \left(2 A_{Troe} \left( 1.1762 A_{Troe} - \right.
+\right .
+& \nonumber \\
+& \mathindent[s] \left.
+\left.
+0.67 B_{Troe}\right) \log{\left (F_{cent} \right )} - B_{Troe} \left(A_{Troe}^{2} + B_{Troe}^{2}\right) \log{\left (10 \right )}\right) & \text{,} \label{e:dfi_troe_dfcent} \\
+\frac{\partial F_{i} }{\partial P_{r, i} } &= - \frac{2 F_{i} A_{Troe} \left(\frac{0.14 A_{Troe}}{B_{Troe}} + 1\right) \log{\left (F_{cent} \right )}}{B_{Troe}^{2} P_{r, i} \left(\frac{A_{Troe}^{2}}{B_{Troe}^{2}} + 1\right)^{2} \log{\left (10 \right )}} \label{e:dfi_troe_dpr} &
+\end{align}
+\Cref{e:dfcent_dt,e:dfi_troe_dfcent,e:dfi_troe_dpr} can then be substituted into~\cref{e:dfi_troe_dt,e:dfi_troe_dnj,e:dfi_troe_de} to obtain a final form for the Troe-falloff blending function derivatives, however this will be given as a generalized form at the end of this section.
+
+For SRI falloff reactions:
+\begin{align}
+ \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial T }} &= F_{i} \vast(\frac{X \left(- \frac{\operatorname{exp}\left({- \frac{T}{c}}\right)}{c} + \frac{a b}{T^{2}} \operatorname{exp}\left({- \frac{b}{T}}\right)\right)}{a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right)} + \nonumber \\
+ & \mathindent \frac{\partial P_{r, i} }{\partial T } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}+ \frac{e}{T} \vast) \; , \\
+ \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial {n_j} }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial {n_j} } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} \; ,
+\end{align}
+and:
+% TODO: not satistified with this yet...
+\begin{subequations}
+ \begin{align}
+ \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial V } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}
+ % note: no alignment as this causes eqn # to go down a line
+ \cond{dfi_sri}{\dconp} \\
+ \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial P } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}
+ % note: no alignment as this causes eqn # to go down a line
+ \cond{dfi_sri}{\dconv}
+ \end{align}
+
+\end{subequations}
 

### 2018-01-05 derivations.tex
updates

@@
 \subsection{Falloff reactions}
-Unlike elementary and third-body reactions, falloff reactions exhibit a pressure dependence described as a blending of rates at low- and high-pressure limits; thus, the rate coefficients depend on a mixture of low-pressure- ($k_{0, i}$) and high-pressure-limit ($k_{\infty,i}$) coefficients, each with corresponding Arrhenius parameters and expressed using Eq.~\cref{e:arrhenius}.
+Unlike elementary and third-body reactions, falloff reactions exhibit a pressure dependence described as a blending of rates at low- and high-pressure limits; thus, the rate coefficients depend on a mixture of low-pressure- ($k_{0, i}$) and high-pressure-limit ($k_{\infty,i}$) coefficients, each with corresponding Arrhenius parameters and expressed using \cref{e:arrhenius}.
 The ratio of the coefficients $k_{0, i}$ and $k_{\infty, i}$, combined with the third-body concentration (based on either the mixture as a whole including any efficiencies $\alpha_{k,i}$, or a specific species), define a reduced pressure $P_{r,i}$ given by
@@
 
-The falloff blending factor $F_i$ used in Eq.~\cref{e:rxn_pressure} is determined based on one of three representations: the Lindemann~\cite{Lindemann:1922cz}, Troe~\cite{Gilbert:1983bb}, and SRI~\cite{Stewart:1989gj} falloff approaches
+The falloff blending factor $F_i$ used in \cref{e:rxn_pressure} is determined based on one of three representations: the Lindemann~\cite{Lindemann:1922cz}, Troe~\cite{Gilbert:1983bb}, and SRI~\cite{Stewart:1989gj} falloff approaches
 \begin{equation}
@@
 In addition to the falloff approach given previously, two additional formulations can be used to describe the pressure dependence of reactions that do not follow the modification factor $c_i$ approach.
-The first involves logarithmic interpolation between Arrhenius rates at two pressures~\cite{chemkin:2012,Goodwin:2015aa}, each evaluated using Eq.~\cref{e:arrhenius}:
+The first involves logarithmic interpolation between Arrhenius rates at two pressures~\cite{chemkin:2012,Goodwin:2015aa}, each evaluated using \cref{e:arrhenius}:
 \begin{align}
@@
 \end{equation}
-While evaluating derivatives for the Jacobian, it is important to expand terms---e.g., concentration, rate of production, etc.---related to the last species to obtain the correct derivative, in this case using Eq.~\cref{e:last_spec_conc}:
+While evaluating derivatives for the Jacobian, it is important to expand terms---e.g., concentration, rate of production, etc.---related to the last species to obtain the correct derivative, in this case using \cref{e:last_spec_conc}:
 \begin{equation}
@@
 \end{equation}
-and using~\cref{e:dkf_dt}, Eq.~\cref{e:dropfdt_1} can be simplified to:
+and using~\cref{e:dkf_dt}, \cref{e:dropfdt_1} can be simplified to:
 \begin{equation}
@@
 \end{equation}
-Next, Eq.~\cref{e:kr} is considered to obtain the temperature derivative of the non-explicit reversible reaction rate coefficient:
+Next, \cref{e:kr} is considered to obtain the temperature derivative of the non-explicit reversible reaction rate coefficient:
 \begin{equation}
@@
 \end{equation}
-and using an temporary value $S^{\prime\prime}_{\ns}$---defined analagously to Eq.~\cref{e:s_temp} for the reverse direction---we obtain:
+and using an temporary value $S^{\prime\prime}_{\ns}$---defined analagously to \cref{e:s_temp} for the reverse direction---we obtain:
 \begin{equation}
@@
 
-\Crefrange{e:dpr_mix_dt}{e:dpr_spec_de} may be recast in a more generalized form using temporary variables $\Theta_{\partial P_{r, i}, \ldots}$ and $\bar{\theta}_{\partial P_{r, i}, \ldots}$ to replace terms in the derivative:
+\Crefrange{e:dpr_mix_dt}{e:dpr_spec_de} may be recast in a more generalized form using temporary variables $\Theta_{P_{r, i}, \partial \ldots}$ and $\bar{\theta}_{P_{r, i}, \partial \ldots}$---where, for example, $\Theta_{P_{r,i}, \partial T}$ corresponds to part of the reduced-pressure derivative with respect to temperature---to replace terms in each individual form of the reduced-pressure derivatives:
 \begin{align}
@@
   \rhs{theta_pr_bar_dp}{\frac{k_{0, i} \delta_{\ns m}}{T k_{\infty, i} \Ru}} & \text{if species,}
- \end{dcases} \\
+ \end{dcases}
 \end{align}
@@
  \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial T }} &= F_{i} \vast(\frac{X \left(- \frac{\operatorname{exp}\left({- \frac{T}{c}}\right)}{c} + \frac{a b}{T^{2}} \operatorname{exp}\left({- \frac{b}{T}}\right)\right)}{a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right)} + \nonumber \\
- & \mathindent \frac{\partial P_{r, i} }{\partial T } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}+ \frac{e}{T} \vast) \; , \\
- \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial {n_j} }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial {n_j} } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} \; ,
+ & \mathindent \frac{\partial P_{r, i} }{\partial T } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}+ \frac{e}{T} \vast) & , \label{e:dfi_sri_dt} \\
+ \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial {n_j} }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial {n_j} } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} & , \label{e:dfi_sri_dnj}
 \end{align}
 and:
-% TODO: not satistified with this yet...
+\begin{subequations}
+ \label{e:dfi_sri_de}
+ \begin{align}
+ \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial V } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} & \cond{dfi_sri}{\dconp} \\
+ \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial P } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} & \cond{dfi_sri}{\dconv}
+ \end{align}
+\end{subequations}
+where:
+\begin{align}
+ \frac{\text{d} X }{\text{d} P_{r, i} } &= - \frac{2 X^{2} \log{\left (P_{r, i} \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} &, \\
+ \frac{\partial X}{\partial n_j} &= \frac{\partial P_{r, i} }{\partial {n_j} } \frac{\text{d} X }{\text{d} P_{r, i} }
+\end{align}
+
+As with the reduced-pressure derivatives,~\cref{e:dfi_lind_dt,e:dfi_lind_dnj,e:dfi_lind_de,e:dfi_troe_dt,e:dfi_troe_dnj,e:dfi_troe_de,e:dfi_sri_dt,e:dfi_sri_dnj,e:dfi_sri_de} may be represented in a general form via introduction of temporary variables $\Theta_{F_i, \partial \ldots}$:
+\begin{align}
+\lhs{dfi_gen}{\frac{\partial F_{i} }{\partial T }} &= \rhs{dfi_gen}{F_{i} \Theta_{F_i, \partial T}} & \cond{dfi_gen}{,} \\
+\lhs{dfi_gen}{\frac{\partial F_{i} }{\partial {n_j}}} &= \rhs{dfi_gen}{\frac{F_{i} k_{0, i} \Theta_{F_i, \partial n_j}}{V k_{\infty, i}} \bar{\theta}_{P_{r, i}, \partial n_j}} & \cond{dfi_gen}{,}
+\end{align}
+and:
 \begin{subequations}
  \begin{align}
- \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial V } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}
- % note: no alignment as this causes eqn # to go down a line
- \cond{dfi_sri}{\dconp} \\
- \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial P } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}
- % note: no alignment as this causes eqn # to go down a line
- \cond{dfi_sri}{\dconv}
+  \lhs{dfi_gen}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_gen}{F_{i} \Theta_{F_i, \partial V}} & \cond{dfi_gen}{\dconp} \\
+  \lhs{dfi_gen}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_gen}{F_{i} \Theta_{F_i, \partial P}} & \cond{dfi_gen}{\dconv}
  \end{align}
-
 \end{subequations}
+where:
+\begin{align}
+\Theta_{F_i, \partial T} &=
+\begin{dcases}
+\rhs{dfi_dt}{0}[c] & \cond{dfi_dt}{if Lindemann}[l] \\
+\rhs{dfi_dt}{ - \frac{B_{Troe}}{F_{cent} P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}}
+	     \Biggl(} & \cond{dfi_dt}{} \nonumber \\
+\rhs{dfi_dt}{\mathindent 2 A_{Troe} F_{cent} \left(0.14 A_{Troe} + B_{Troe}\right)
+	     \left(P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right)
+	     \log{\left (F_{cent} \right )} + } & \cond{dfi_dt}{} \nonumber \\
+\rhs{dfi_dt}{\mathindent P_{r, i} \frac{\text{d} F_{cent} }{\text{d} T }
+	      \biggl( 2 A_{Troe} \left(1.1762 A_{Troe} - 0.67 B_{Troe}\right) \log{\left (F_{cent} \right )} - } & \cond{dfi_dt}{} \nonumber \\
+\rhs{dfi_dt}{\mathindent[plussomeextrastu] B_{Troe} \left(A_{Troe}^{2} + B_{Troe}^{2}\right) \log{\left (10 \right )}\biggr)\Biggr)} & \cond{dfi_dt}{if Troe}[l] \\
+\rhs{dfi_dt}{- \frac{X \left(\frac{\operatorname{exp}\left({- \frac{T}{c}}\right)}{c} - \frac{a b}{T^{2}} \operatorname{exp}\left({- \frac{b}{T}}\right)\right)}{a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right)} + \frac{e}{T} - \frac{2 X^{2} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \left(P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right) \log{\left (P_{r, i} \right )}} & \cond{dfi_dt}{if SRI}[l] \\
+\end{dcases}
+\end{align}
 

### 2018-01-10 derivations.tex
many edits

@@
 \NewDocumentCommand{\cond}{mmO{r}}{\eqmakebox[COND_#1][#3]{#2}}
+\NewDocumentCommand{\mathcond}{mmO{r}}{\eqmathbox[COND_#1][#3]{#2}}
 
@@
 First, we evaluate the derivatives of the net species rate of production equation~\cref{e:spec_rop}, with respect to temperature:
-\begin{equation}
+\begin{align}
  \label{e:dwdot_dT}
- \frac{\partial \dot{\omega} }{\partial T }_{k} = \sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial T }_{i} + \nu_{k,i} \frac{\partial R }{\partial T }_{i} c_{i}\right) \;,
-\end{equation}
+ \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial T }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial T }_{i} + \nu_{k,i} \frac{\partial R }{\partial T }_{i} c_{i}\right)} & \cond{dwdot}{,}
+\end{align}
 other species:
-\begin{equation}
+\begin{align}
  \label{e:dwdot_dnj}
- \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} = \sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial {n_j} }_{i} + \nu_{k,i} \frac{\partial R }{\partial {n_j} }_{i} c_{i}\right)
-\end{equation}
+ \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial {n_j} }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial {n_j} }_{i} + \nu_{k,i} \frac{\partial R }{\partial {n_j} }_{i} c_{i}\right)} & \cond{dwdot}{,}
+\end{align}
 and the state parameter:
@@
  \begin{align}
-  \frac{\partial \dot{\omega} }{\partial V }_{k} &= \sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial V }_{i} + \nu_{k,i} \frac{\partial R }{\partial V }_{i} c_{i}\right)\dconp,\\
-  \frac{\partial \dot{\omega} }{\partial P }_{k} &= \sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial P }_{i} + \nu_{k,i} \frac{\partial R }{\partial P }_{i} c_{i}\right)\dconv
+  \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial V }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial V }_{i} + \nu_{k,i} \frac{\partial R }{\partial V }_{i} c_{i}\right)} & \cond{dwdot}{\dconp}\\
+  \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial P }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial P }_{i} + \nu_{k,i} \frac{\partial R }{\partial P }_{i} c_{i}\right)} & \cond{dwdot}{\dconv}
  \end{align}
@@
 \begin{subequations}
- % based on https://tex.stackexchange.com/a/239764/56227
- % the idea is to take the longest case (1.b -- i.e., no hphantom) and use this
- % to make a box around all cases such that the alignment works
  \begin{align}
- \frac{\partial [C] }{\partial V }_{k} &=
+ \lhs{dck}{\frac{\partial [C] }{\partial V }_{k}} &=
   \begin{dcases}
-  \mathrlap{-\frac{[C]_{k}}{V}}\hphantom{\frac{1}{V} \sum_{k=1}^{-1 + \ns} [C]_{k}} & k \ne \ns \\
-  \frac{1}{V} \sum_{k=1}^{-1 + \ns} [C]_{k} & k = \ns
- \end{dcases}&\dconp \\
- \frac{\partial [C]_{k} }{\partial P } &=
- \begin{dcases}
-  \mathrlap{\hphantom{0}0}\hphantom{\frac{1}{V} \sum_{k=1}^{-1 + \ns} [C]_{k}} & k \ne \ns \\
-  \mathrlap{\frac{1}{T \Ru}}\hphantom{\frac{1}{V} \sum_{k=1}^{-1 + \ns} [C]_{k}} & k = \ns
- \end{dcases}&\dconv
+  \rhs{dck}{\frac{[C]_{k}}{V}} & \mathcond{dck}{k \ne \ns} \\
+  \rhs{dck}{\frac{1}{V} \sum_{k=1}^{-1 + \ns} [C]_{k}} & \mathcond{dck}{k = \ns}
+ \end{dcases}&\cond{dck}{\dconp}
+ \lhs{dck}{\frac{\partial [C]_{k} }{\partial P }} &=
+ \begin{dcases} \\
+   \rhs{dck}{0} & \mathcond{dck}{k \ne \ns} \\
+   \rhs{dck}{\frac{1}{T \Ru}} & \mathcond{dck}{k = \ns}
+ \end{dcases}&\cond{dck}{\dconv}
  \end{align}
@@
  \begin{align}
-  \frac{\partial {R_f}_{i} }{\partial V } &= \frac{[C] S^{\prime}_{\ns}}{V} {k_f}_{i} - \frac{{R_f}_{i}}{V} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}&\dconp\\
-  \frac{\partial {R_f}_{i} }{\partial P } &= \frac{S^{\prime}_{\ns} {k_f}_{i}}{T \Ru}&\dconv
+  \lhs{dri_de}{\frac{\partial {R_f}_{i} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime}_{\ns}}{V} {k_f}_{i} - \frac{{R_f}_{i}}{V} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}}&\cond{dri_de}{\dconp}\\
+  \lhs{dri_de}{\frac{\partial {R_f}_{i} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime}_{\ns} {k_f}_{i}}{T \Ru}}&\cond{dri_de}{\dconv}
  \end{align}
@@
  \begin{align}
-  \frac{\partial {R_r}_{i} }{\partial V } &= \frac{[C] S^{\prime\prime}_{\ns}}{V} {k_r}_{i} - \frac{{R_r}_{i}}{V} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i}&\dconp\\
-  \frac{\partial {R_r}_{i} }{\partial P } &= \frac{S^{\prime\prime}_{\ns} {k_r}_{i}}{T \Ru}&\dconv
+  \lhs{dri_de}{\frac{\partial {R_r}_{i} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime\prime}_{\ns}}{V} {k_r}_{i} - \frac{{R_r}_{i}}{V} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i}}&\cond{dri_de}{\dconp}\\
+  \lhs{dri_de}{\frac{\partial {R_r}_{i} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime\prime}_{\ns} {k_r}_{i}}{T \Ru}}&\cond{dri_de}{\dconv}
  \end{align}
@@
 \begin{align}
- \label{e:dpr_theta}
+ \label{e:dpr_theta_dt}
  \Theta_{P_{r,i}, \partial T} &=
@@
  \end{dcases} \\
+ \label{e:dpr_theta_bar_dt}
  \bar{\theta}_{P_{r, i}, \partial T} &=
@@
  \end{dcases} \\
+ \label{e:dpr_theta_bar_dnj}
  \bar{\theta}_{P_{r, i}, \partial n_j} &=
@@
  \end{dcases} \\
+ \label{e:dpr_theta_dv}
  \Theta_{P_{r,i}, \partial V} &=
@@
  \end{dcases} \\
+ \label{e:dpr_theta_bar_dv}
  \bar{\theta}_{P_{r, i}, \partial V} &=
@@
  \end{dcases} \\
+ \label{e:dpr_theta_dp}
  \bar{\theta}_{P_{r, i}, \partial P} &=
@@
 \end{subequations}
-where:
+where for Lindemann falloff reactions:
 \begin{align}
-\Theta_{F_i, \partial T} &=
-\begin{dcases}
-\rhs{dfi_dt}{0}[c] & \cond{dfi_dt}{if Lindemann}[l] \\
-\rhs{dfi_dt}{ - \frac{B_{Troe}}{F_{cent} P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}}
-	     \Biggl(} & \cond{dfi_dt}{} \nonumber \\
-\rhs{dfi_dt}{\mathindent 2 A_{Troe} F_{cent} \left(0.14 A_{Troe} + B_{Troe}\right)
-	     \left(P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right)
-	     \log{\left (F_{cent} \right )} + } & \cond{dfi_dt}{} \nonumber \\
-\rhs{dfi_dt}{\mathindent P_{r, i} \frac{\text{d} F_{cent} }{\text{d} T }
-	      \biggl( 2 A_{Troe} \left(1.1762 A_{Troe} - 0.67 B_{Troe}\right) \log{\left (F_{cent} \right )} - } & \cond{dfi_dt}{} \nonumber \\
-\rhs{dfi_dt}{\mathindent[plussomeextrastu] B_{Troe} \left(A_{Troe}^{2} + B_{Troe}^{2}\right) \log{\left (10 \right )}\biggr)\Biggr)} & \cond{dfi_dt}{if Troe}[l] \\
-\rhs{dfi_dt}{- \frac{X \left(\frac{\operatorname{exp}\left({- \frac{T}{c}}\right)}{c} - \frac{a b}{T^{2}} \operatorname{exp}\left({- \frac{b}{T}}\right)\right)}{a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right)} + \frac{e}{T} - \frac{2 X^{2} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \left(P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right) \log{\left (P_{r, i} \right )}} & \cond{dfi_dt}{if SRI}[l] \\
-\end{dcases}
+\lhs{dfi_lind}{\Theta_{F_i, \partial T}} &= \rhs{dfi_lind}{0} \\
+\lhs{dfi_lind}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_lind}{0}
 \end{align}
+and:
+\begin{subequations}
+ \begin{align}
+  \lhs{dfi_lind}{\Theta_{F_i, \partial V}} &= \rhs{dfi_lind}{0} \\
+  \lhs{dfi_lind}{\Theta_{F_i, \partial P}} &= \rhs{dfi_lind}{0}
+ \end{align}
+\end{subequations}
+For Troe falloff reactions:
+\begin{align}
+% troe
+\lhs{dfi_troe}{\Theta_{F_i, \partial T}} &= 
+\rhs{dfi_troe}{ - \frac{B_{Troe}}{F_{cent} P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}} \times \Biggl(} \nonumber \\
+&\rhs{dfi_troe}{ 2 A_{Troe} F_{cent} \left(0.14 A_{Troe} + B_{Troe}\right) \times } \nonumber \\
+&\rhs{dfi_troe}{\left( P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right) \log{\left (F_{cent} \right )} + P_{r, i} \frac{\text{d} F_{cent} }{\text{d} T } \times \biggl(} \nonumber \\
+&\rhs{dfi_troe}{\mathindent 2 A_{Troe} \left(1.1762 A_{Troe} - 0.67 B_{Troe}\right) \log{\left (F_{cent} \right )} - } \nonumber \\
+&\rhs{dfi_troe}{\mathindent B_{Troe} \left(A_{Troe}^{2} + B_{Troe}^{2}\right) \log{\left (10 \right )}\biggr)\Biggr)} \\
+\lhs{dfi_troe}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_troe}{- \frac{2 A_{Troe} B_{Troe} \left(0.14 A_{Troe} + B_{Troe}\right) \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}}}
+\end{align}
+and:
+\begin{subequations}
+ \begin{align}
+   \lhs{dfi_troe}{\Theta_{F_i, \partial V}} &= \rhs{dfi_troe}{- \frac{2 A_{Troe} B_{Troe} \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}} \times \bigl(} \nonumber \\
+					    & \rhs{dfi_troe}{\mathindent[alittlebit] 0.14 A_{Troe} + B_{Troe}\bigr) \left(P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}\right)} \\
+   \lhs{dfi_troe}{\Theta_{F_i, \partial P}} &= \rhs{dfi_troe}{- \frac{2 A_{Troe} B_{Troe} \bar{\theta}_{P_{r, i}, \partial P} \left(0.14 A_{Troe} + B_{Troe}\right) \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}}}
+ \end{align}
+\end{subequations}
+where $\Theta_{P_{r,i}, \partial \ldots}$ and $\bar{\theta}_{P_{r, i}, \partial \ldots}$ are given by~\cref{e:dpr_theta_dt,e:dpr_theta_bar_dt,e:dpr_theta_bar_dnj,e:dpr_theta_dv,e:dpr_theta_bar_dv,e:dpr_theta_dp} respectively.
+Finally, for SRI falloff reactions:
+\begin{align}
+ \lhs{dfi_sri}{\Theta_{F_i, \partial T}} &= 
+ \rhs{dfi_sri}{- \frac{X \left(\frac{\operatorname{exp}\left({- \frac{T}{c}}\right)}{c} - \frac{a b}{T^{2}} \operatorname{exp}\left({- \frac{b}{T}}\right)\right)}{a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right)} + \frac{e}{T} - \Biggl(} \nonumber \\
+ &\rhs{dfi_sri}{\mathindent \frac{2 X^{2} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \times } \nonumber \\
+ &\rhs{dfi_sri}{\mathindent \left(P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right) \log{\left (P_{r, i} \right ) \Biggr)}} \\
+ \lhs{dfi_sri}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_sri}{- \frac{2 X^{2} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \log{\left (P_{r, i} \right )}}
+\end{align}
+and:
+\begin{subequations}
+ \begin{align}
+  \lhs{dfi_sri}{\Theta_{F_i, \partial V}} &=
+  \rhs{dfi_sri}{- \frac{2 X^{2} \log{\left (P_{r, i} \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \left(P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}\right) \log \Biggl(} \nonumber \\
+& \rhs{dfi_sri}{\mathindent \left( a \operatorname{exp}\left({\frac{T}{c}}\right) + \operatorname{exp}\left({\frac{b}{T}}\right)\right) \operatorname{exp}\left({- \frac{T}{c} - \frac{b}{T}}\right) \Biggr)} \\
+  \lhs{dfi_sri}{\Theta_{F_i, \partial P}} &=
+  \rhs{dfi_sri}{- \frac{2 X^{2} \bar{\theta}_{P_{r, i}, \partial P} \log{\left (P_{r, i} \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}
+ \end{align}
+\end{subequations}
 

### 2018-01-11 derivations.tex
add temporary product based fall / chem forms?

@@
 \begin{align}
- \lhs{dci_fall}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial T } + \frac{\partial P_{r, i} }{\partial T } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{,} \label{e:dci_fall_dt}\\
- \lhs{dci_fall}{\frac{\partial c_{i} }{\partial n_j}} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial n[j] } + \frac{\partial P_{r, i} }{\partial n_j } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{,} \label{e:dci_fall_dnj}
+ \label{e:dci_fall_dt}
+ \lhs{dci_fall}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial T } + \frac{\partial P_{r, i} }{\partial T } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{,}\\
+ \label{e:dci_fall_dnj}
+ \lhs{dci_fall}{\frac{\partial c_{i} }{\partial n_j}} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial n[j] } + \frac{\partial P_{r, i} }{\partial n_j } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{,}
 \end{align}
@@
 \begin{align}
+\label{e:dfi_lind_dt_prod}
 \lhs{dfi_lind}{\Theta_{F_i, \partial T}} &= \rhs{dfi_lind}{0} \\
+\label{e:dfi_lind_dnj_prod}
 \lhs{dfi_lind}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_lind}{0}
@@
 \begin{subequations}
+ \label{e:dfi_lind_de_prod}
  \begin{align}
@@
 % troe
+\label{e:dfi_troe_dt_prod}
 \lhs{dfi_troe}{\Theta_{F_i, \partial T}} &= 
@@
 &\rhs{dfi_troe}{\mathindent B_{Troe} \left(A_{Troe}^{2} + B_{Troe}^{2}\right) \log{\left (10 \right )}\biggr)\Biggr)} \\
+\label{e:dfi_troe_dnj_prod}
 \lhs{dfi_troe}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_troe}{- \frac{2 A_{Troe} B_{Troe} \left(0.14 A_{Troe} + B_{Troe}\right) \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}}}
@@
 \begin{subequations}
+ \label{e:dfi_troe_de_prod}
  \begin{align}
@@
 \begin{align}
+ \label{e:dfi_sri_dt_prod}
  \lhs{dfi_sri}{\Theta_{F_i, \partial T}} &= 
@@
  &\rhs{dfi_sri}{\mathindent \left(P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right) \log{\left (P_{r, i} \right ) \Biggr)}} \\
+ \label{e:dfi_sri_dnj_prod}
  \lhs{dfi_sri}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_sri}{- \frac{2 X^{2} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \log{\left (P_{r, i} \right )}}
@@
 \begin{subequations}
+ \label{e:dfi_sri_de_prod}
  \begin{align}
@@
 
+\subsubsection{Unimolecular\slash~recombination fall-off reactions (temporary product form)}
+Using the temporary products developed in~\cref{s:dpr,s:dfi}, the fall-off reaction derivatives developed in~\cref{e:dci_fall_dt,e:dci_fall_dnj,e:dci_fall_de} may be simplified to:
+\begin{align}
+\lhs{dci_fall_complete}{\frac{\partial c }{\partial T }_{i}} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial T}}{P_{r, i} + 1} + \biggl(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial T}}{P_{r, i} + 1} + \Theta_{F_i, \partial T} + } & \nonumber \\
+							     &  \rhs{dci_fall_complete}{\mathindent[largerindentpleas] \Theta_{P_{r,i}, \partial T} - \frac{\bar{\theta}_{P_{r, i}, \partial T}}{P_{r, i} + 1}\biggr) c_{i}} & \cond{dci_fall_complete}{,} \\
+\lhs{dci_fall_complete}{\frac{\partial c }{\partial {n_j} }_{i}} &= \rhs{dci_fall_complete}{\frac{k_{0, i} \bar{\theta}_{P_{r, i}, \partial n_j}}{V k_{\infty, i} \left(P_{r, i} + 1\right)} \left(F_{i} \left(P_{r, i} \Theta_{F_i, \partial n_j} + 1\right) - c_{i}\right)} & \cond{dci_fall_complete}{,}
+\end{align}
+and:
+\begin{subequations}
+ \begin{align}
+  \lhs{dci_fall_complete}{\frac{\partial c }{\partial V }_{i}} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1} + \biggl(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial V}}{P_{r, i} + 1} + \Theta_{F_i, \partial V} + } & \nonumber \\
+							       &  \rhs{dci_fall_complete}{\mathindent[largerindentpleas] \Theta_{P_{r,i}, \partial V} - \frac{\bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1} \biggr) c_{i}} & \cond{dci_fall_complete}{\dconp} \\
+  \lhs{dci_fall_complete}{\frac{\partial c }{\partial P }_{i}} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1} + \left(\Theta_{F_i, \partial P} - \frac{\bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_fall_complete}{\dconv}
+ \end{align}
+\end{subequations}
+where the temporary products are given by~\cref{e:dpr_theta_dt,e:dpr_theta_bar_dt,e:dpr_theta_bar_dnj,e:dpr_theta_dv,e:dpr_theta_bar_dv,e:dpr_theta_dp}~and~\cref{e:dfi_lind_dt_prod,e:dfi_lind_dnj_prod,e:dfi_lind_de_prod,e:dfi_troe_dt_prod,e:dfi_troe_dnj_prod,e:dfi_troe_de_prod,e:dfi_sri_dt_prod,e:dfi_sri_dnj_prod,e:dfi_sri_de_prod}
+
+\subsubsection{Chemically-activated bimolecular reactions (temporary product form)}
+Similarly, the chemically-activated reaction derivatives in~\cref{e:dci_chem_dt,e:dci_chem_dnj,e:dci_chem_de} may be written as:
+\begin{align}
+\lhs{dci_chem_complete}{\frac{\partial c }{\partial T }_{i}} &= \rhs{dci_chem_complete}{\left(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial T}}{P_{r, i} + 1} + \Theta_{F_i, \partial T} - \frac{\bar{\theta}_{P_{r, i}, \partial T}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_chem_complete}{,} \\
+\lhs{dci_chem_complete}{\frac{\partial c }{\partial {n_j} }_{i}} &= \rhs{dci_chem_complete}{\frac{k_{0, i} \bar{\theta}_{P_{r, i}, \partial n_j} \left(F_{i} \Theta_{F_i, \partial n_j} - c_{i}\right)}{k_{\infty, i} V \left(P_{r, i} + 1\right)}} & \cond{dci_chem_complete}{,}
+\end{align}
+and:
+\begin{subequations}
+ \begin{align}
+  \lhs{dci_chem_complete}{\frac{\partial c }{\partial V }_{i}} &= \rhs{dci_fall_complete}{\left(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial V}}{P_{r, i} + 1} + \Theta_{F_i, \partial V} - \frac{\bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_chem_complete}{\dconp} \\
+  \lhs{dci_chem_complete}{\frac{\partial c }{\partial P }_{i}} &= \rhs{dci_fall_complete}{\left(\Theta_{F_i, \partial P} - \frac{\bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_chem_complete}{\dconv}
+ \end{align}
+\end{subequations}
+
 \subsection{Final Jacobian entries}

### 2018-01-12 derivations.tex
lots of fixes, finished falloff blending section & added PLOG / Cheb kf derivs

@@
 \subsection{State variables}
-The state vector for this derivation consists of the temperature, a thermodynamic parameter (pressure or volume) and the amount of moles of all species except the last in the model:
+The state vector for this derivation consists of the temperature, a thermodynamic state parameter (pressure or volume) and the amount of moles of all species except the last in the model:
 
@@
 \subsection{Pressure-dependent reactions}
+\label{s:pdep}
 
 In addition to the falloff approach given previously, two additional formulations can be used to describe the pressure dependence of reactions that do not follow the modification factor $c_i$ approach.
-The first involves logarithmic interpolation between Arrhenius rates at two pressures~\cite{chemkin:2012,Goodwin:2015aa}, each evaluated using \cref{e:arrhenius}:
+The first involves logarithmic interpolation between Arrhenius rates at two pressures~\cite{chemkin:2012,Goodwin:2015aa} (often termed P-Log reactions), each evaluated using~\cref{e:arrhenius}:
 \begin{align}
@@
 \begin{equation}
+\label{e:kf_plog}
 \log{\left ({k_f}_{i} \right )} = \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) \left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \log{\left (k_{1} \right )}
@@
 \begin{equation}
-\log_{10} k_f (T, p) = \sum_{i = 1}^{N_T} \sum_{j = 1}^{N_p} \eta_{ij} \phi_i (\tilde{T}) \phi_j \left(\tilde{p}\right) \label{e:cheb} \;,
+\label{e:kf_cheb}
+\log_{10} k_f (T, p) = \sum_{i = 1}^{N_T} \sum_{j = 1}^{N_p} \eta_{ij} \phi_i (\tilde{T}) \phi_j \left(\tilde{p}\right) \;,
 \end{equation}
@@
 \subsubsection{Temperature derivative}
+\label{s:dri_dt}
 Next, the rate of progress derivatives are evaluated, first with respect to temperature:
@@
 \end{equation}
-By defining a temporary variable (for species $l = \left\{j, \ns\right\}$) $S^{\prime}_{l}$:
+By defining a temporary variable $S^{\prime}_{l}$ (for species $l = \left\{j, \ns\right\}$):
 \begin{equation}
@@
 \begin{equation}
+ \label{e:droprdt_1}
  \frac{\partial {R_r} }{\partial T }_{i} = \frac{\text{d} {k_r} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{-1 + \ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
@@
 \subsubsection{Molar derivative}
-\label{s:rop_molar}
+\label{s:dri_dnj}
 The derivative of the forward rate of progress with respect to the amount of moles of a species $j$ is:
@@
 
-\subsubsection{Parameter derivatives}
+\subsubsection{State parameter derivatives}
+\label{s:dri_de}
 The derivatives of the forward and reverse rates of progress with respect to the volume (\conp) or pressure (\conv) largely follow the same format of the previous sections.
-First, the derivative of the concentration of a species $k$ with respect to the parameter is:
+First, the derivative of the concentration of a species $k$ with respect to the state parameter is:
 
@@
   \rhs{dck}{\frac{1}{V} \sum_{k=1}^{-1 + \ns} [C]_{k}} & \mathcond{dck}{k = \ns}
- \end{dcases}&\cond{dck}{\dconp}
+ \end{dcases}&\cond{dck}{\dconp} \\
  \lhs{dck}{\frac{\partial [C]_{k} }{\partial P }} &=
  \begin{dcases} \\
-   \rhs{dck}{0} & \mathcond{dck}{k \ne \ns} \\
+   \rhs{dck}{\mathindent[s]0} & \mathcond{dck}{k \ne \ns} \\
    \rhs{dck}{\frac{1}{T \Ru}} & \mathcond{dck}{k = \ns}
@@
 
-Following the outline of~\cref{s:rop_molar}, the forward rate of progress derivatives are:
+Following the outline of~\cref{s:dri_dnj}, the forward rate of progress derivatives are:
 \begin{subequations}
@@
  \label{e:dropi_de}
- % linebreak and parenthesis sizing:
- % https://tex.stackexchange.com/questions/49890/linebreak-between-left-and-right/49895
  \begin{align}
-  \frac{\partial {R}_{i} }{\partial V } &= \frac{1}{V}\left(\vphantom{{R_r}_{i} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i}} [C]\left(S^{\prime}_{\ns} {k_f}_{i} - S^{\prime\prime}_{\ns} {k_r}_{i}\right) + \right.& \nonumber\\
-  &\left. \quad\left({R_r}_{i} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i} - {R_f}_{i} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}\right) \right)&\dconp\\
-  \frac{\partial {R}_{i} }{\partial P } &= \frac{1}{T \Ru} \left(S^{\prime}_{\ns} {k_f}_{i} - S^{\prime\prime}_{\ns} {k_r}_{i}\right)&\dconv
+  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial V }} &= \rhs{dri_de}{\frac{1}{V}\Biggl( [C]\left(S^{\prime}_{\ns} {k_f}_{i} - S^{\prime\prime}_{\ns} {k_r}_{i}\right) + } & \nonumber\\
+  & \rhs{dri_de}{\mathindent[farth] \left({R_r}_{i} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i} - {R_f}_{i} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}\right) \Biggr)}&\cond{dri_de}{\dconp}\\
+  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial P }} &= \rhs{dri_de}{\frac{1}{T \Ru} \left(S^{\prime}_{\ns} {k_f}_{i} - S^{\prime\prime}_{\ns} {k_r}_{i}\right)}&\cond{dri_de}{\dconv}
  \end{align}
 \end{subequations}
+
+\subsubsection{Pressure-dependent reactions}
+The P-Log and Chebyshev pressure-dependent reactions, described by~\cref{e:kf_plog,e:kf_cheb} require separate treatment from the derivatives of strictly Arrhenius-based reaction rate formulations examined in~\cref{s:dri_dt,s:dri_dnj,s:dri_de}.
+Beginning with P-Log reactions, the derivative of the forward rate coefficient with respect to temperature is:
+\begin{align}
+ \frac{\partial {k_f}_{i} }{\partial T } &= \biggl( \frac{1}{k_{1}} \frac{\text{d} k_1 }{\text{d} T } + \frac{1}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} \left(- \frac{1}{k_{1}} \frac{\text{d} k_1 }{\text{d} T } + \frac{1}{k_{2}} \frac{\text{d} k_2 }{\text{d} T }\right) \times  \Bigl( \nonumber \\
+					 & \mathindent - \log{\left (P_{1} \right )} + \log{\left (P \right )}\Bigr)\biggr) {k_f}_{i}
+\end{align}
+while the derivatives of $k_{1}$ and $k_{2}$ are given by~\cref{e:dkf_dt}---with the corresponding Arrhenius parameters (see~\cref{e:plog_k1,e:plog_k2}) substituted in.
+Simplifying gives:
+\begin{align}
+ \label{e:dkf_plog_dt}
+ \frac{\partial {k_f}_{i} }{\partial T } &= \frac{{k_f}_{i}}{T} \Biggl(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \nonumber \\
+					 & \mathindent[morespac] \beta_1 + \frac{E_{a_1}}{T \Ru}\Biggr)
+\end{align}
+Similar to~\cref{s:dri_dt},~\cref{e:dkf_plog_dt} may be substituted into~\cref{e:dropfdt_1,e:droprdt_1} to give:
+\begin{align}
+ \frac{\partial R }{\partial T }_{i} &= \left(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \beta_1 + \frac{E_{a_1}}{T \Ru} \right) \times \nonumber \\
+				     & \mathindent \left(\frac{{R_f}_{i} - {R_r}_{i}}{T}\right) + {R_r}_{i} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_r}_{i} - S^{\prime}_{\ns} {k_f}_{i}\right)
+\end{align}
+
+For \conv~problems, the P-Log forward rate coefficient is also a function of the pressure:
+\begin{equation}
+ \frac{\partial {k_f}_{i} }{\partial P } = \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) {k_f}_{i}}{P \left(- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}\right)}
+\end{equation}
+substituting this into the net rate of progress derivative gives:
+\begin{align}
+ \frac{\partial R_{i} }{\partial P } &= \frac{1}{T \Ru} \left(- S^{\prime\prime}_{\ns} {k_r}_{i} + S^{\prime}_{\ns} {k_f}_{i}\right) + & \nonumber \\
+				     &  \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) \left({R_f}_{i} - {R_r}_{i}\right)}{P \left(- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}\right)} & \text{\dconv}
+\end{align}
+
+Next, the temperature derivative of Chevbyshev reactions will be considered:
+\begin{align}
+ \frac{\partial {k_f}_{i} }{\partial T } &= \log{\left (10 \right )} {k_f}_{i} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{\text{d} \tilde{T} }{\text{d} T } \left(j - 1\right) \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}
+\end{align}
+where $U_n$ is the Chebyshev polynomial of the second kind of degree n, expressed as:
+\begin{equation}
+ U_n \left(x\right) = \frac{\sin\left(\left(n + 1\right) \arccos x\right)}{\sin \left(\arccos x \right)}
+\end{equation}
+and:
+\begin{equation}
+ \frac{\text{d} \tilde{T}}{\text{d} T} = \frac{-2}{T^2 \left(-\frac{1}{T_{\text{min}}} + \frac{1}{T_{\text{max}}}\right)}
+\end{equation}
+giving:
+\begin{equation}
+ \frac{\partial {k_f}_{i} }{\partial T } = \log{\left (10 \right )} {k_f}_{i} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} - \frac{2 \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\text{min}}} + \frac{1}{T_{\text{max}}}\right)} \left(j - 1\right)
+\end{equation}
+Following the same process for Arrhenius-based reactions, the derivative of the net rate of progress with respect to temperature is:
+\begin{align}
+ \frac{\partial R }{\partial T }_{i} &= \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \left[ - \frac{2 \log{\left (10 \right )} \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\text{min}}} + \frac{1}{T_{\text{max}}}\right)} \left(j - 1\right) \right] \left({R_f}_{i} - {R_r}_{i}\right) + \nonumber \\
+				     & \mathindent \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} {R_r}_{i} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_r}_{i} - S^{\prime}_{\ns} {k_f}_{i}\right)
+\end{align}
+
+For \conv~problems, we again must consider the derivative of the forward rate coefficient with respect to pressure:
+\begin{equation}
+ \frac{\partial {k_f}_{i} }{\partial P } = \log{\left (10 \right )} {k_f}_{i} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{\text{d} \tilde{P} }{\text{d} P } \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}
+\end{equation}
+with:
+\begin{equation}
+ \frac{\text{d} \tilde{P} }{\text{d} P } = \frac{2}{P \left(\log{\left (P_{\text{max}} \right )} - \log{\left (P_{\text{min}} \right )}\right)}
+\end{equation}
+giving:
+\begin{equation}
+ \frac{\partial {k_f}_{i} }{\partial P } = \log{\left (10 \right )} {k_f}_{i} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\text{max}} \right )} - \log{\left (P_{\text{min}} \right )}\right)}
+\end{equation}
+and:
+\begin{align}
+ \frac{\partial R_{i} }{\partial P } &= \log{\left (10 \right )} \left({R_f}_{i} - {R_r}_{i}\right) \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\text{max}} \right )} - \log{\left (P_{\text{min}} \right )}\right)} + \nonumber \\
+				     & \mathindent \frac{1}{T \Ru} \left(- S^{\prime\prime}_{\ns} {k_r}_{i} + S^{\prime}_{\ns} {k_f}_{i}\right)
+\end{align}
 
@@
 \end{align}
-and with respect to the thermodynamic parameter:
+and with respect to the state parameter:
 \begin{subequations}

### 2018-01-12 derivations/scripts/derivations.py
fix incorrect display

@@ -1930,7 +1930,7 @@ def _derivation(file, efile, conp=True, thermo_deriv=False):
     write_eq(diff(Pri_sym, nk[j]), dPri_unitydnj)
 
     dPri_unityde = diff(Pri_unity, extra_var)
-    write_eq(diff(Pri_sym, nk[j]), dPri_unityde)
+    write_eq(diff(Pri_sym, extra_var), dPri_unityde)
 
     latexfile.write('Simplifying:\n')
     dPri_unitydT, dPri_unitydT_prifac, dPri_unitydT_prifac_sym,\

### 2018-01-12 derivations.tex
many edits and minor corrections

@@
 \section{Introduction}
-This document contains the complete derivations for \texttt{pyJac} v2.0, which generates code to evaluate the chemical source term and analytical chemical kinetic Jacobians of constant-pressure\slash constant-volume, fixed-mass, adiabatic reactors.
+This document contains the derivations for \texttt{pyJac} v2.0, which generates code to evaluate the chemical source terms and analytical chemical kinetic Jacobians of constant-pressure\slash constant-volume, fixed-mass, adiabatic reactors.
 Note that equations specific to the constant-pressure or constant-volume derivation will be marked with \conp~or \conv~respectively.
-The derivations in this script based upon the output of the script \href{https://github.com/arghdos/SPyJac-paper/blob/master/derivations/scripts/derivations.py}{derivations.py} in the GitHub repository for this paper.
+The derivations in this document based upon the output of the script \href{https://github.com/arghdos/SPyJac-paper/blob/master/derivations/scripts/derivations.py}{derivations.py} in the GitHub repository for this paper.
 This script depends on the symbolic mathematics library SymPy~\cite{sympy}, and was tested with versions 1.0 and 1.1.1
@@
 \subsection{State variables}
-The state vector for this derivation consists of the temperature, a thermodynamic state parameter (pressure or volume) and the amount of moles of all species except the last in the model:
-
-\begin{subequations}
-\begin{align}
-\Phi &= \left\{T, V, n_1, n_2 \ldots n_{\ns - 1}\right\}\dconp, \\
-\Phi &= \left\{T, P, n_1, n_2 \ldots n_{\ns - 1}\right\}\dconv
-\end{align}
-\end{subequations}
-where $T$ is the temperature, $P$ and $V$ the pressure and volume respectively, and $n_j$ the number of moles of the $j$th species in the model (containing $\ns$ total species).
+The state vector for this derivation consists of the temperature, a thermodynamic state parameter (pressure or volume) and the number of moles of all species except the last in the model:
+\begin{subequations}
+\begin{align}
+\Phi &= \left\{T, V, n_1, n_2 \ldots n_{\ns - 1}\right\}&\text{\dconp}\\
+\Phi &= \left\{T, P, n_1, n_2 \ldots n_{\ns - 1}\right\}&\text{\dconv}
+\end{align}
+\end{subequations}
+where $T$ is the temperature, $V$ and $P$ the volume and pressure respectively, and $n_j$ the number of moles of the $j$th species in the model (containing $\ns$ total species).
 
@@
 \begin{align}
-\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} H_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{p, k}}}\dconp, \\
-\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} U_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{v, k}}}\dconv
-\end{align}
-\end{subequations}
-where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure specific heat, constant-volume specific heat of species $k$ in molar units, while $[C]_{k}$ is the concentration, given by:
+\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} H_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{p, k}}} & \text{\dconp} \\
+\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} U_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{v, k}}} & \text{\dconv}
+\end{align}
+\end{subequations}
+where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure specific heat and constant-volume specific heat of species $k$ in molar units respectively, while $[C]_{k}$ is the concentration, given by:
 \begin{equation}
@@
 \begin{equation}
+ \label{e:mass_conservation}
  \frac{\text{d} m }{\text{d} t } = 0 = \sum_{k=1}^{\ns}  W_{k} \frac{\text{d} n }{\text{d} t }_{k},
 \end{equation}
-from which, the source term of the last species may be written in terms of the species included in the state vector:
+where $W_{k}$ is the molecular weight of the $k$th species.
+From~\cref{e:mass_conservation}, the source term of the last species may be written in terms of the species included in the state vector:
 \begin{equation}
@@
 \end{equation}
-where $W_{k}$ is the molecular weight of the $k$th species.
 Using~\cref{e:spec_ns}, the total molar rate of change can be determined:
@@
 \begin{align}
-\frac{\text{d} V }{\text{d} t } &= V \left(\frac{T \Ru}{P} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right)&\dconp, \\
-\frac{\text{d} P }{\text{d} t } &= \frac{P}{T} \frac{\text{d} T }{\text{d} t } + T \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k}&\dconv
+\frac{\text{d} V }{\text{d} t } &= V \left(\frac{T \Ru}{P} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right)& \text{\dconp} \\
+\frac{\text{d} P }{\text{d} t } &= \frac{P}{T} \frac{\text{d} T }{\text{d} t } + T \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k}&\text{\dconv}
 \end{align}
@@
 \begin{align}
-\frac{\text{d} T }{\text{d} t } = - \frac{\sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{p,\ns}} + \sum_{k=1}^{-1 + \ns} \left(- {C_{p,\ns}} + {C_{p, k}}\right) [C]_{k}}\dconp,\\
-\frac{\text{d} T }{\text{d} t } = - \frac{\sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{v,\ns}} + \sum_{k=1}^{-1 + \ns} \left(- {C_{v,\ns}} + {C_{v, k}}\right) [C]_{k}}\dconv
+\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{p,\ns}} + \sum_{k=1}^{-1 + \ns} \left(- {C_{p,\ns}} + {C_{p, k}}\right) [C]_{k}} &\text{\dconp}\\
+\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{v,\ns}} + \sum_{k=1}^{-1 + \ns} \left(- {C_{v,\ns}} + {C_{v, k}}\right) [C]_{k}} &\text{\dconv}
 \end{align}
@@
 \subsection{Reaction rate expressions}
-The net species rate of production is defined as:
+The net species production rate for species $k$ is defined as:
 \begin{equation}
@@
 c_i = \begin{dcases}
-  1 &\text{for elementary reactions,} \\
-  [X]_i &\text{for third-body enhanced reactions,} \\
-  \frac{P_{r,i}}{1 + P_{r,i}} F_i &\text{for unimolecular\slash~recombination fall-off reactions, and} \\
-  \frac{1}{1 + P_{r,i}} F_i &\text{for chemically-activated bimolecular reactions,}
+  \rhs{pmod}{1}[c] &\text{for elementary reactions,} \\
+  \rhs{pmod}{[X]_i}[c] &\text{for third-body enhanced reactions,} \\
+  \rhs{pmod}{\frac{P_{r,i}}{1 + P_{r,i}} F_i} &\text{for unimolecular\slash~recombination fall-off reactions, and} \\
+  \rhs{pmod}{\frac{1}{1 + P_{r,i}} F_i} &\text{for chemically-activated bimolecular reactions,}
   \end{dcases}
@@
 \subsection{Reverse rate coefficient}
-By definition, the reverse rate coefficient of irreversible reactions ${k_r}_{i}$ is zero, while reversible reactions may have non-zero ${k_r}_{i}$.
+By definition, the reverse rate coefficient---${k_r}_{i}$---of irreversible reactions is zero, while reversible reactions may have non-zero ${k_r}_{i}$.
 Note that in \texttt{pyJac}, reversible reactions with explicit reverse Arrhenius parameters are split into two irreversible reactions; this simplifies calculation inside the generated code, and eases comparison to Cantera~\cite{Goodwin:2015aa} which applies the same transformation.
@@
 \end{equation}
-This form will be denoted as the \textbf{mixture}-based third-body efficiency.
+This form will be denoted as the mixture-based (or \textbf{mix} for short) third-body efficiency.
 
@@
 \subsection{Falloff reactions}
-Unlike elementary and third-body reactions, falloff reactions exhibit a pressure dependence described as a blending of rates at low- and high-pressure limits; thus, the rate coefficients depend on a mixture of low-pressure- ($k_{0, i}$) and high-pressure-limit ($k_{\infty,i}$) coefficients, each with corresponding Arrhenius parameters and expressed using \cref{e:arrhenius}.
-The ratio of the coefficients $k_{0, i}$ and $k_{\infty, i}$, combined with the third-body concentration (based on either the mixture as a whole including any efficiencies $\alpha_{k,i}$, or a specific species), define a reduced pressure $P_{r,i}$ given by
+Unlike elementary and third-body reactions, falloff reactions exhibit a pressure dependence described as a blending of rates at low- and high-pressure limits; thus, the rate coefficients depend on a mixture of low-pressure ($k_{0, i}$) and high-pressure-limit ($k_{\infty,i}$) coefficients, each with corresponding Arrhenius parameters and expressed using \cref{e:arrhenius}.
+The ratio of the coefficients $k_{0, i}$ and $k_{\infty, i}$, combined with the third-body concentration, define a reduced pressure $P_{r,i}$ given by
 \begin{equation}
@@
 \end{equation}
-where $[X]_{i}$ is the appropriate third-body concentration as described in~\cref{s:thdbody}
+where $[X]_{i}$ is the appropriate third-body concentration as described in~\cref{s:thdbody}.
 
@@
 \begin{equation}
+\label{e:fi}
 F_i = \begin{dcases}
@@
 \end{equation}
-where $a$, $b$, and $c$ are required parameters.
-The parameters $d$ and $e$ are optional; if not specified, $d = 1$ and $e = 0$.
+The parameters $a$, $b$, and $c$ in the SRI falloff blending factor are required while $d$ and $e$ are optional; if not specified, $d = 1$ and $e = 0$.
 
@@
 \tilde{T} &\equiv \frac{2 T^{-1} - T^{-1}_{\min} - T^{-1}_{\max}}{T^{-1}_{\max} - T^{-1}_{\min}} \quad\text{and} \\
-\tilde{p} &\equiv \frac{2\log_{10} p - \log_{10} p_{\min} - \log_{10} p_{\max}}{\log_{10} p_{\max} - \log_{10} p_{\min}} \;,
+\tilde{P} &\equiv \frac{2\log_{10} P - \log_{10} P_{\min} - \log_{10} P_{\max}}{\log_{10} P_{\max} - \log_{10} P_{\min}} \;,
 \end{align}
@@
 
-\subsection{Species rate of production derivatives}
-First, we evaluate the derivatives of the net species rate of production equation~\cref{e:spec_rop}, with respect to temperature:
+\subsection{Species production rate derivatives}
+First, we evaluate the derivatives of the net species production rate equation~\cref{e:spec_rop}, with respect to temperature:
 \begin{align}
@@
 \end{equation}
-the derivative of the forward reaction rate coefficient is:
+The derivative of the forward reaction rate coefficient is:
 \begin{equation}
@@
 \end{equation}
-While evaluating derivatives for the Jacobian, it is important to expand terms---e.g., concentration, rate of production, etc.---related to the last species to obtain the correct derivative, in this case using \cref{e:last_spec_conc}:
+While evaluating derivatives for the Jacobian, it is important to expand terms---e.g., concentration, production rate, etc.---related to the last species to obtain the correct derivative, in this case using \cref{e:last_spec_conc}:
 \begin{equation}
@@
 Next, \cref{e:kr} is considered to obtain the temperature derivative of the non-explicit reversible reaction rate coefficient:
-\begin{equation}
- \frac{\text{d} {k_r} }{\text{d} T }_{i} = \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c} }{\text{d} T }_{i} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_r}_{i}
-\end{equation}
+\begin{align}
+ \frac{\text{d} {k_r} }{\text{d} T }_{i} &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } \frac{{k_f}_i}{{K_c}_i} + \frac{1}{{K_c}_{i}}\frac{\text{d} {k_f}_{i} }{\text{d} T } \right) \nonumber \\
+				         &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_r}_{i}
+\end{align}
 The temperature derivative of the equilibrium constant is:
 \begin{equation}
- \frac{\text{d} {K_c} }{\text{d} T }_{i} = {K_c}_{i} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} \;,
+ \frac{\text{d} {K_c}_{i} }{\text{d} T } = {K_c}_{i} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} \;,
 \end{equation}
@@
 \begin{equation}
+ \label{e:dropf_dnj_1}
  \frac{d}{d n_{k}} {R_{f_i}} = \left(\frac{\partial}{\partial n_{j}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right) {k_f}_{i} \;.
@@
 \end{equation}
-Using~\cref{e:dck_dnj} the molar derivative of the forward rate of progress is found as:
+Combining~\cref{e:dropf_dnj_1,e:dck_dnj} the molar derivative of the forward rate of progress is found as:
 \begin{equation}
@@
   \begin{dcases}
-  \rhs{dck}{\frac{[C]_{k}}{V}} & \mathcond{dck}{k \ne \ns} \\
+  \rhs{dck}{-\frac{[C]_{k}}{V}} & \mathcond{dck}{k \ne \ns} \\
   \rhs{dck}{\frac{1}{V} \sum_{k=1}^{-1 + \ns} [C]_{k}} & \mathcond{dck}{k = \ns}
@@
 \begin{align}
- \frac{\partial R }{\partial T }_{i} &= \left(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \beta_1 + \frac{E_{a_1}}{T \Ru} \right) \times \nonumber \\
+ \frac{\partial R_{i} }{\partial T } &= \left(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \beta_1 + \frac{E_{a_1}}{T \Ru} \right) \times \nonumber \\
 				     & \mathindent \left(\frac{{R_f}_{i} - {R_r}_{i}}{T}\right) + {R_r}_{i} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_r}_{i} - S^{\prime}_{\ns} {k_f}_{i}\right)
@@
 \begin{equation}
- \frac{\text{d} \tilde{T}}{\text{d} T} = \frac{-2}{T^2 \left(-\frac{1}{T_{\text{min}}} + \frac{1}{T_{\text{max}}}\right)}
+ \frac{\text{d} \tilde{T}}{\text{d} T} = \frac{-2}{T^2 \left(-\frac{1}{T_{\min}} + \frac{1}{T_{\text{max}}}\right)}
 \end{equation}
@@
 \begin{equation}
- \frac{\partial {k_f}_{i} }{\partial T } = \log{\left (10 \right )} {k_f}_{i} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} - \frac{2 \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\text{min}}} + \frac{1}{T_{\text{max}}}\right)} \left(j - 1\right)
-\end{equation}
-Following the same process for Arrhenius-based reactions, the derivative of the net rate of progress with respect to temperature is:
-\begin{align}
- \frac{\partial R }{\partial T }_{i} &= \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \left[ - \frac{2 \log{\left (10 \right )} \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\text{min}}} + \frac{1}{T_{\text{max}}}\right)} \left(j - 1\right) \right] \left({R_f}_{i} - {R_r}_{i}\right) + \nonumber \\
+ \frac{\partial {k_f}_{i} }{\partial T } = \log{\left (10 \right )} {k_f}_{i} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} - \frac{2 \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right)
+\end{equation}
+Following the same process for Arrhenius-based reactions (see~\cref{s:dri_dt}), the derivative of the net rate of progress with respect to temperature is:
+\begin{align}
+ \frac{\partial R }{\partial T }_{i} &= \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \left[ - \frac{2 \log{\left (10 \right )} \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right) \right] \left({R_f}_{i} - {R_r}_{i}\right) + \nonumber \\
 				     & \mathindent \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} {R_r}_{i} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_r}_{i} - S^{\prime}_{\ns} {k_f}_{i}\right)
@@
 \begin{equation}
- \frac{\text{d} \tilde{P} }{\text{d} P } = \frac{2}{P \left(\log{\left (P_{\text{max}} \right )} - \log{\left (P_{\text{min}} \right )}\right)}
+ \frac{\text{d} \tilde{P} }{\text{d} P } = \frac{2}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)}
 \end{equation}
@@
 \begin{equation}
- \frac{\partial {k_f}_{i} }{\partial P } = \log{\left (10 \right )} {k_f}_{i} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\text{max}} \right )} - \log{\left (P_{\text{min}} \right )}\right)}
-\end{equation}
-and:
-\begin{align}
- \frac{\partial R_{i} }{\partial P } &= \log{\left (10 \right )} \left({R_f}_{i} - {R_r}_{i}\right) \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\text{max}} \right )} - \log{\left (P_{\text{min}} \right )}\right)} + \nonumber \\
+ \frac{\partial {k_f}_{i} }{\partial P } = \log{\left (10 \right )} {k_f}_{i} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)}
+\end{equation}
+and:
+\begin{align}
+ \frac{\partial R_{i} }{\partial P } &= \log{\left (10 \right )} \left({R_f}_{i} - {R_r}_{i}\right) \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)} + \nonumber \\
 				     & \mathindent \frac{1}{T \Ru} \left(- S^{\prime\prime}_{\ns} {k_r}_{i} + S^{\prime}_{\ns} {k_f}_{i}\right)
@@
  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_thd_unity}{- \frac{[C]}{T}}& \cond{dci_thd_unity}{,} \label{e:dci_thd_unity_dt} \\
- \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial n_{j} }} &= \rhs{dci_thd_unity}{0}& \cond{dci_thd_unity}{,}
+ \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial n_{j} }} &= \rhs{dci_thd_unity}{0}[c] & \cond{dci_thd_unity}{,}
 \end{align}
@@
  \begin{align}
-  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_thd_unity}{0} & \cond{dci_thd_unity}{\dconp} \\
+  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_thd_unity}{0}[c] & \cond{dci_thd_unity}{\dconp} \\
   \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_thd_unity}{\frac{1}{T \Ru}}& \cond{dci_thd_unity}{\dconv}
@@
  \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial T } } &= \rhs{dpr_gen}{P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}} & \cond{dpr_gen}{,} \label{e:dpr_dt} \\
- \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial n_j}} &= \rhs{dpr_gen}{\frac{\bar{\theta}_{P_{r, i}, \partial n_j}}{V k_{\infty, i}} k_{0, i}} & \cond{dpr_gen}{,} \label{e:dpr_dnj}
+ \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial n_j}} &= \rhs{dpr_gen}{\frac{k_{0, i}\bar{\theta}_{P_{r, i}, \partial n_j}}{V k_{\infty, i}}} & \cond{dpr_gen}{,} \label{e:dpr_dnj}
 \end{align}
@@
 First, the blending factor for each falloff reaction type will be differentiated along with any subcomponents (e.g., the $F_{cent}$ term for Troe falloff reactions).
-Next, the reduced pressured derivatives given by~\cref{e:dpr_dt,e:dpr_dnj,e:dpr_de} will be subsituted in, before finally deriving an analagous generalized form for the falloff blending factor derivatives.
+Next, the reduced pressured derivatives given by~\cref{e:dpr_dt,e:dpr_dnj,e:dpr_de} will be substituted in, before finally deriving a generalized form for the falloff blending factor derivatives.
 
@@
 \begin{align}
- \lhs{dfi_troe}{\frac{\partial F_{i} }{\partial T }} &= \rhs{dfi_troe}{\frac{\partial F_{i} }{\partial F_{cent} } \frac{\text{d} F_{cent} }{\text{d} T } + \frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial T }} & \cond{dfi_troe}{,} \label{e:dfi_troe_dt} \\
- \lhs{dfi_troe}{\frac{\partial F_{i} }{\partial n_j }} &= \rhs{dfi_troe}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial n_j }} & \cond{dfi_troe}{,} \label{e:dfi_troe_dnj}
+ \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial T }} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial F_{cent} } \frac{\text{d} F_{cent} }{\text{d} T } + \frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial T }} & \cond{dfi_troe_1}{,} \label{e:dfi_troe_dt} \\
+ \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial n_j}} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial n_j }} & \cond{dfi_troe_1}{,} \label{e:dfi_troe_dnj}
 \end{align}
@@
  \begin{align}
- \lhs{dfi_troe}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_troe}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial V }} & \cond{dfi_troe}{\dconp} \\
- \lhs{dfi_troe}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_troe}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial P }} & \cond{dfi_troe}{\dconv}
+ \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial V }} & \cond{dfi_troe_1}{\dconp} \\
+ \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial P }} & \cond{dfi_troe_1}{\dconv}
  \end{align}
@@
 & \nonumber \\
- & \mathindent \left.
+ & \mathindent[morespaceplease] \left.
 \frac{\partial A_{Troe} }{\partial F_{cent} }\right) + \frac{1}{F_{cent}} \left.
@@
 \end{align}
-\Cref{e:dfcent_dt,e:dfi_troe_dfcent,e:dfi_troe_dpr} can then be substituted into~\cref{e:dfi_troe_dt,e:dfi_troe_dnj,e:dfi_troe_de} to obtain a final form for the Troe-falloff blending function derivatives, however this will be given as a generalized form at the end of this section.
+\Cref{e:dfcent_dt,e:dfi_troe_dfcent,e:dfi_troe_dpr} can then be substituted into~\cref{e:dfi_troe_dt,e:dfi_troe_dnj,e:dfi_troe_de} to obtain a final form for the Troe-falloff blending function derivatives, which will be given as a generalized form at the end of this section.
 
@@
 &\rhs{dfi_troe}{ 2 A_{Troe} F_{cent} \left(0.14 A_{Troe} + B_{Troe}\right) \times } \nonumber \\
-&\rhs{dfi_troe}{\left( P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right) \log{\left (F_{cent} \right )} + P_{r, i} \frac{\text{d} F_{cent} }{\text{d} T } \times \biggl(} \nonumber \\
+&\rhs{dfi_troe}{\left( P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right) \log{\left (F_{cent} \right )} + P_{r, i} \frac{\text{d} F_{cent} }{\text{d} T } \times \Bigl(} \nonumber \\
 &\rhs{dfi_troe}{\mathindent 2 A_{Troe} \left(1.1762 A_{Troe} - 0.67 B_{Troe}\right) \log{\left (F_{cent} \right )} - } \nonumber \\
-&\rhs{dfi_troe}{\mathindent B_{Troe} \left(A_{Troe}^{2} + B_{Troe}^{2}\right) \log{\left (10 \right )}\biggr)\Biggr)} \\
+&\rhs{dfi_troe}{\mathindent B_{Troe} \left(A_{Troe}^{2} + B_{Troe}^{2}\right) \log{\left (10 \right )}\Bigr)\Biggr)} \\
 \label{e:dfi_troe_dnj_prod}

### 2018-01-16 derivations/limit_test.py
add sympy limits detector to see if there's any better limit to give if Pr == 0

@@ -1,0 +1,155 @@
+from sympy import symbols, log, exp, limit, KroneckerDelta, diff
+from optionloop import OptionLoop
+
+
+def __get_dci(fall_type='chem', blend_type='troe', pr_type='mix', var='E'):
+    # create temperature
+    T = symbols('T')
+
+    # create kf's
+    kf, kinf = symbols('kf kinf', real=True, nonnegative=True)
+
+    # create third body efficiency & volume
+    V, C, P = symbols('V C P', real=True, nonnegative=True)
+    Xi, alphaNS, alphaj = symbols('Xi alphaNS alphaj', real=True)
+
+    # species
+    m, Ns, j = symbols('m Ns j', integer=True, nonnegative=True)
+
+    # create pr
+    Pr = kf * Xi / kinf
+    R = 8.314
+
+    # create Fi & Troe params
+    if blend_type == 'troe':
+        T1, T2, T3, a = symbols('T1 T2 T3 a', real=True)
+
+        Fcent = a * exp(-T / T1) + (1 - a) * exp(-T / T3) + exp(-T2 / T)
+        Atroe = -0.68 * log(Fcent, 10) + log(Pr, 10) - 0.4
+        Btroe = -1.1762 * log(Fcent, 10) - 0.14 * log(Pr, 10) + 0.806
+
+        Fi = Fcent ** (1 / (1 + (Atroe / Btroe)**2))
+    elif blend_type == 'sri':
+        a, b, c, d, e = symbols('a b c d e', real=True)
+        X = 1 / (log(Pr, 10) ** 2 + 1)
+        Fi = d * T**e * (a * exp(-b / T) + exp(-T / c)) ** X
+    elif blend_type == 'lind':
+        Fi = 1
+
+    # chemically activated form
+    if fall_type == 'chem':
+        ci = Fi / (1 + Pr)
+    elif fall_type == 'fall':
+        ci = Pr * Fi / (1 + Pr)
+
+    # now create derivative temporary products (assuming mixture)
+    if var == 'T':
+        b0, binf, e0, einf = symbols('b0 binf e0 einf', real=True)
+        if pr_type in ['mix', 'spec']:
+            Theta_Pr = (b0 - binf + e0 / (R * T) - einf / (R * T)) / T
+            if pr_type == 'mix':
+                theta_Pr = -C * kf * alphaNS / (T * kinf)
+            else:
[…]

### 2018-01-17 derivations/limit_test.py
broken limit

@@ -1,4 +1,6 @@
-from sympy import symbols, log, exp, limit, KroneckerDelta, diff
+from __future__ import print_function
+from sympy import symbols, log, exp, limit, KroneckerDelta, diff, \
+    Product, factor, Pow, Symbol, simplify, Limit
 from optionloop import OptionLoop
 
 
@@ -94,7 +96,7 @@ def __get_dci(fall_type='chem', blend_type='troe', pr_type='mix', var='E'):
                     Pr * Theta_Pr + theta_Pr) * log(Fcent) + Pr * diff(Fcent, T) * (
                         2 * Atroe * (1.1762 * Atroe - 0.67 * Btroe) * log(Fcent) -
                         Btroe * (Atroe**2 + Btroe**2) * log(10))
-                )
+            )
         elif var == 'nj':
             Theta_Fi = -2 * Atroe * Btroe * (0.14 * Atroe + Btroe) * log(Fcent) / (
                 Pr * (Atroe**2 + Btroe**2)**2 * log(10))
@@ -111,14 +113,17 @@ def __get_dci(fall_type='chem', blend_type='troe', pr_type='mix', var='E'):
                 a * exp(-b / T) + exp(-T / c)) + e / T - ((
                     2 * X**2 * log(a * exp(-b / T) + exp(-T / c))) / (Pr * log(10)**2) * (
                     (Theta_Pr * Pr + theta_Pr) * log(Pr))
-                )
+            )
         elif var == 'nj':
-            Theta_Fi = -2 * X**2 * log(a * exp(-b / T) + exp(-T / c)) * log(Pr) / (Pr * log(10)**2)
+            Theta_Fi = -2 * X**2 * \
+                log(a * exp(-b / T) + exp(-T / c)) * \
+                log(Pr) / (Pr * log(10)**2)
         elif var == 'V':
             Theta_Fi = (-2 * X**2 * log(Pr) / (Pr * log(10)**2)) * (Theta_Pr * Pr + theta_Pr) * log(
                 (a * exp(T / c) + exp(b / T)) * exp(-T / c - b / T))
         elif var == 'P':
-            Theta_Pr = (-2 * X**2 * theta_Pr * log(Pr) / (Pr * log(10)**2)) * log(a * exp(-b / T) + exp(-T / c))
+            Theta_Pr = (-2 * X**2 * theta_Pr * log(Pr) /
+                        (Pr * log(10)**2)) * log(a * exp(-b / T) + exp(-T / c))
 
     # and finally give dci
     if var == 'T':
[…]

### 2018-01-19 derivations.tex
edits

@@
 \newcommand{\dconp}{\ifmmode\text{for \conp,}\else for \conp,\fi}
-\newcommand{\dconv}{\ifmmode\text{for \conv,}\else for \conv,\fi}
+\newcommand{\dconv}{\ifmmode\text{for \conv,}\else for \conv.\fi}
 \newcommand{\Ru}{\mathcal{R}}
@@
 \end{equation}
-where $\dot{\omega}_k$ is the $kth$ species overall production rate, and the temperature production rates are~\cite{TurnsStephenR2012Aitc}:
+where $\dot{\omega}_k$ is the $kth$ species overall production rate, and the temperature production rate~\cite{TurnsStephenR2012Aitc} is:
 \begin{subequations}
@@
 \begin{align}
-\frac{\text{d} V }{\text{d} t } &= V \left(\frac{T \Ru}{P} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right)& \text{\dconp} \\
-\frac{\text{d} P }{\text{d} t } &= \frac{P}{T} \frac{\text{d} T }{\text{d} t } + T \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k}&\text{\dconv}
+\frac{\text{d} V }{\text{d} t } &= V \left(\frac{T \Ru}{P} \sum_{k=1}^{\ns  - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right)& \text{\dconp} \\
+\frac{\text{d} P }{\text{d} t } &= \frac{P}{T} \frac{\text{d} T }{\text{d} t } + T \Ru \sum_{k=1}^{\ns  - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k}&\text{\dconv}
 \end{align}
@@
  \label{e:last_spec_conc}
-   [C]_{\ns} = [C] - \sum_{k=1}^{-1 + \ns} [C]_{k}
+   [C]_{\ns} = [C] - \sum_{k=1}^{\ns  - 1} [C]_{k}
 \end{equation}
@@
 \begin{align}
-\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{p,\ns}} + \sum_{k=1}^{-1 + \ns} \left(- {C_{p,\ns}} + {C_{p, k}}\right) [C]_{k}} &\text{\dconp}\\
-\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{v,\ns}} + \sum_{k=1}^{-1 + \ns} \left(- {C_{v,\ns}} + {C_{v, k}}\right) [C]_{k}} &\text{\dconv}
+\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns  - 1} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{p,\ns}} + \sum_{k=1}^{\ns  - 1} \left(- {C_{p,\ns}} + {C_{p, k}}\right) [C]_{k}} &\text{\dconp}\\
+\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns  - 1} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{v,\ns}} + \sum_{k=1}^{\ns  - 1} \left(- {C_{v,\ns}} + {C_{v, k}}\right) [C]_{k}} &\text{\dconv}
 \end{align}
@@
 \label{e:thd_mix}
- [X]_{i}=[C] \alpha_{\ns,i} + \sum_{k=1}^{-1 + \ns} \left(- \alpha_{\ns,i} + \alpha_{k,i}\right) [C]_{k}\;.
+ [X]_{i}=[C] \alpha_{\ns,i} + \sum_{k=1}^{\ns  - 1} \left(- \alpha_{\ns,i} + \alpha_{k,i}\right) [C]_{k}\;.
 \end{equation}
@@
 \label{e:thd_spec}
- [X]_{i}=\left([C] - \sum_{k=1}^{-1 + \ns} [C]_{k}\right) \delta_{\ns m} + \left(- \delta_{\ns m} + 1\right) [C]_{m}
+ [X]_{i}=\left([C] - \sum_{k=1}^{\ns  - 1} [C]_{k}\right) \delta_{\ns m} + \left(- \delta_{\ns m} + 1\right) [C]_{m}
 \end{equation}
@@
 \begin{equation}
+ \label{e:jac_general}
  \mathcal{J}_{i,j} = \frac{\partial f_i}{\partial \Phi_j},\qquad j=1 \ldots \ns - 1
 \end{equation}
-We will begin by deriving subcomponents of the Jacobian matrix in the subsequent sections before presenting the final form of the Jacobian matrix in~\cref{s:jac_final}.
+The Jacobian entries resulting from~\cref{e:jac_general} are derived in~\cref{s:energy_derivs}, while various subcomponents of the Jacobian are derived in~\cref{s:dri_dt,s:dri_dnj,s:dri_de,s:dri_pdep,s:dpmod_fall} and the final form of the Jacobian matrix is presented in~\cref{s:jac_final}.
+
+We note that some parts of the Jacobian derivation---in particular derivatives of the energy equation---are quite complicated.
+The aim of this document is to provide the reader with an overview of the derivation process; some intermediate steps are left out due to brevity and typesetting constraints.
+For a complete derivation, the reader is directed to the output of the \href{https://github.com/arghdos/SPyJac-paper/blob/master/derivations/scripts/derivations.py}{derivations.py} script, from which this document was compiled.
+
+\subsection{Energy equation derivatives}
+\label{s:energy_derivs}
+\subsubsection{Temperature derivative}
+In this section, we will more closely follow the \conp~energy equation derivative with respect to temperature as the process is largely identical between \conp~and \conv.
+
+First, the derivative of full form of the energy source term~\cref{e:temperature_complete} with respect to temperature will be considered.
+To make the result more managable, the denominator of~\cref{e:temperature_complete} is first collapsed back into the form of~\cref{e:temperature_incomplete}, yielding for~\conp:
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{T}} &= - \frac{\left(- [C] \frac{\text{d} {C_p} }{\text{d} T }_{\ns} - \sum_{k=1}^{-1 + \ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns} + \frac{\text{d} {C_p} }{\text{d} T }_{k}\right) [C]_{k} + \frac{[C] {C_p}_{\ns}}{T}\right)}{\left(\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}\right)^{2}} \times \nonumber \\
+				     & \mathindent[largerindent] \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k} - \nonumber \\
+				     & \mathindent \frac{\sum_{k=1}^{-1 + \ns} \left(\left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial T }_{k} + \left(\frac{\text{d} H }{\text{d} T }_{k} - \frac{W_{k}}{W_{\ns}} \frac{\text{d} H }{\text{d} T }_{\ns}\right) \dot{\omega}_{k}\right)}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}}
+\end{align}
+Next, we may recognize that~\cref{e:temperature_incomplete} may be factored out of the first terms, yielding:
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \times \nonumber \\
+				     & \left( - [C] \frac{\text{d} {C_p} }{\text{d} T }_{\ns} - \sum_{k=1}^{-1 + \ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns} + \frac{\text{d} {C_p} }{\text{d} T }_{k}\right) [C]_{k} + \frac{[C] {C_p}_{\ns}}{T} \right) - \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns} \left(\left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial T }_{k} + \left(\frac{\text{d} H }{\text{d} T }_{k} - \frac{W_{k}}{W_{\ns}} \frac{\text{d} H }{\text{d} T }_{\ns}\right) \dot{\omega}_{k}\right) \Biggr]
+\end{align}
+Now, we invoke the identity:
+\begin{equation}
+ [C] = \sum_{k=1}^{\ns} [C]_{k}
+\end{equation}
+to simplify the specific heat terms:
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{k} + \frac{{C_p}_{\ns}}{T}\right) [C]_{k} + \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns} \biggl(\left(- H_{k} + \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial T }_{k} + \nonumber \\
+				     & \mathindent[larger] \left(- \frac{\text{d} H }{\text{d} T }_{k} + \frac{W_{k}}{W_{\ns}} \frac{\text{d} H }{\text{d} T }_{\ns}\right) \dot{\omega}_{k}\biggr)\Biggr]
+\end{align}
+and finally, we may substitute in the specific heat for the temperature derivative of enthalpy, giving:
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{k} + \frac{{C_p}_{\ns}}{T}\right) [C]_{k} + & \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns} \biggl( \left(- H_{k} + \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial T }_{k} + & \nonumber \\
+				     & \mathindent[larger] \left(- {C_p}_{k} + \frac{W_{k} {C_p}_{k} }{W_{\ns}} \right) \dot{\omega}_{k}\biggr)\Biggr] & \text{\dconp} \\
+\intertext{and similarly:}
+\frac{\partial\dot{T}}{\partial{T}}  &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns} \left(- \frac{\text{d} {C_v} }{\text{d} T }_{k} + \frac{{C_v}_{\ns}}{T}\right) [C]_{k} + & \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns} \biggl(\left(- U_{k} + \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial T }_{k} + & \nonumber \\
+				     & \mathindent[larger] \left(- {C_v}_{k} + \frac{W_{k} {C_v}_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}\biggr)\Biggr] & \text{\dconv}
+\end{align}
+\subsubsection{Molar derivative}
+The molar derivative of the energy equation is somewhat easier to work with,
 
@@
 \begin{equation}
- {R_f} = \left(\left(- \sum_{k=1}^{-1 + \ns} [C]_{k} + \frac{P}{T \Ru}\right)^{\nu^{\prime}_{\ns,i}}\right) {k_f}_{i} \prod_{k=1}^{-1 + \ns} [C]_{k}^{\nu^{\prime}_{k,i}}
+ {R_f} = \left(\left(- \sum_{k=1}^{\ns  - 1} [C]_{k} + \frac{P}{T \Ru}\right)^{\nu^{\prime}_{\ns,i}}\right) {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
 \end{equation}
@@
  \label{e:dropfdt_1}
- \frac{\partial {R_f} }{\partial T }_{i} = \frac{\text{d} {k_f} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{-1 + \ns} [C]_{k}^{\nu^{\prime}_{k,i}}
+ \frac{\partial {R_f} }{\partial T }_{i} = \frac{\text{d} {k_f} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
 \end{equation}
@@
  \label{e:droprdt_1}
- \frac{\partial {R_r} }{\partial T }_{i} = \frac{\text{d} {k_r} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{-1 + \ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
+ \frac{\partial {R_r} }{\partial T }_{i} = \frac{\text{d} {k_r} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
 \end{equation}
@@
   \rhs{dck}{-\frac{[C]_{k}}{V}} & \mathcond{dck}{k \ne \ns} \\
-  \rhs{dck}{\frac{1}{V} \sum_{k=1}^{-1 + \ns} [C]_{k}} & \mathcond{dck}{k = \ns}
+  \rhs{dck}{\frac{1}{V} \sum_{k=1}^{\ns  - 1} [C]_{k}} & \mathcond{dck}{k = \ns}
  \end{dcases}&\cond{dck}{\dconp} \\
@@
 \subsubsection{Pressure-dependent reactions}
+\label{s:dri_pdep}
 The P-Log and Chebyshev pressure-dependent reactions, described by~\cref{e:kf_plog,e:kf_cheb} require separate treatment from the derivatives of strictly Arrhenius-based reaction rate formulations examined in~\cref{s:dri_dt,s:dri_dnj,s:dri_de}.
@@
 \subsection{Pressure modification\slash Falloff function derivatives}
+\label{s:dpmod_fall}
 \subsubsection{Elementary reactions}

### 2018-01-19 derivations/limit_test.py

@@ -1,203 +1,0 @@
-from __future__ import print_function
-from sympy import symbols, log, exp, limit, KroneckerDelta, diff, \
-    Product, factor, Pow, Symbol, simplify, Limit
-from optionloop import OptionLoop
-
-
-def __get_dci(fall_type='chem', blend_type='troe', pr_type='mix', var='E'):
-    # create temperature
-    T = symbols('T')
-
-    # create kf's
-    kf, kinf = symbols('kf kinf', real=True, nonnegative=True)
-
-    # create third body efficiency & volume
-    V, C, P = symbols('V C P', real=True, nonnegative=True)
-    Xi, alphaNS, alphaj = symbols('Xi alphaNS alphaj', real=True)
-
-    # species
-    m, Ns, j = symbols('m Ns j', integer=True, nonnegative=True)
-
-    # create pr
-    Pr = kf * Xi / kinf
-    R = 8.314
-
-    # create Fi & Troe params
-    if blend_type == 'troe':
-        T1, T2, T3, a = symbols('T1 T2 T3 a', real=True)
-
-        Fcent = a * exp(-T / T1) + (1 - a) * exp(-T / T3) + exp(-T2 / T)
-        Atroe = -0.68 * log(Fcent, 10) + log(Pr, 10) - 0.4
-        Btroe = -1.1762 * log(Fcent, 10) - 0.14 * log(Pr, 10) + 0.806
-
-        Fi = Fcent ** (1 / (1 + (Atroe / Btroe)**2))
-    elif blend_type == 'sri':
-        a, b, c, d, e = symbols('a b c d e', real=True)
-        X = 1 / (log(Pr, 10) ** 2 + 1)
-        Fi = d * T**e * (a * exp(-b / T) + exp(-T / c)) ** X
-    elif blend_type == 'lind':
-        Fi = 1
-
-    # chemically activated form
-    if fall_type == 'chem':
-        ci = Fi / (1 + Pr)
-    elif fall_type == 'fall':
-        ci = Pr * Fi / (1 + Pr)
-
-    # now create derivative temporary products (assuming mixture)
-    if var == 'T':
-        b0, binf, e0, einf = symbols('b0 binf e0 einf', real=True)
-        if pr_type in ['mix', 'spec']:
-            Theta_Pr = (b0 - binf + e0 / (R * T) - einf / (R * T)) / T
-            if pr_type == 'mix':
-                theta_Pr = -C * kf * alphaNS / (T * kinf)
-            else:
[…]

### 2018-01-22 derivations.tex
edits

@@
 \end{equation}
-The Jacobian entries resulting from~\cref{e:jac_general} are derived in~\cref{s:energy_derivs}, while various subcomponents of the Jacobian are derived in~\cref{s:dri_dt,s:dri_dnj,s:dri_de,s:dri_pdep,s:dpmod_fall} and the final form of the Jacobian matrix is presented in~\cref{s:jac_final}.
+where $i$ and $j$ correspond to the row and column, respectively, of the entry in the Jacobian matrix.
+The Jacobian entries resulting from~\cref{e:jac_general} are derived in~\cref{s:energy_derivs,s:molar_source_derivatives,s:parameter_source_derivatives}, while various subcomponents of the Jacobian are derived in~\cref{s:dri_dt,s:dri_dnj,s:dri_de,s:dri_pdep,s:dpmod_fall}..
 
@@
 
-\subsection{Energy equation derivatives}
+\subsection{Temperature source term derivatives}
 \label{s:energy_derivs}
 \subsubsection{Temperature derivative}
-In this section, we will more closely follow the \conp~energy equation derivative with respect to temperature as the process is largely identical between \conp~and \conv.
-
-First, the derivative of full form of the energy source term~\cref{e:temperature_complete} with respect to temperature will be considered.
+\label{s:denergy_dt}
+First, we will derive the term:
+\begin{align}
+ \mathcal{J}_{1,1} &= \frac{\partial}{\partial T} \frac{\text{d} T}{\text{d} t} \nonumber \\
+ \intertext{or more simply:}
+ \mathcal{J}_{1,1} &= \frac{\partial\dot{T}}{\partial{T}}
+\end{align}
+To do so, we will more closely follow the \conp~energy equation derivative with respect to temperature as the process is largely identical between \conp~and \conv.
+
+The derivative of a species concentration with respect to temperature for both \conp~and \conv~is:
+\begin{align}
+ \frac{\partial [C]_{k}}{\partial T} = 
+ \begin{dcases}
+  \frac{\partial}{\partial T} \left(\frac{n_k}{V}\right) = 0 & \left\{k \ne \ns\right\} \\
+  \frac{\partial}{\partial T} \left([C] - \sum_{k=1}^{\ns - 1}{\frac{n_k}{V}}\right) = -\frac{[C]}{T} & \left\{k = \ns\right\}
+ \end{dcases}
+\end{align}
+as the pressure and volume are either constants (for \conp~and \conv~respectively) or a state variable and a function of time only.
+
+Next, the derivative of full form of the energy source term~\cref{e:temperature_complete} with respect to temperature will be considered.
 To make the result more managable, the denominator of~\cref{e:temperature_complete} is first collapsed back into the form of~\cref{e:temperature_incomplete}, yielding for~\conp:
 \begin{align}
- \frac{\partial\dot{T}}{\partial{T}} &= - \frac{\left(- [C] \frac{\text{d} {C_p} }{\text{d} T }_{\ns} - \sum_{k=1}^{-1 + \ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns} + \frac{\text{d} {C_p} }{\text{d} T }_{k}\right) [C]_{k} + \frac{[C] {C_p}_{\ns}}{T}\right)}{\left(\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}\right)^{2}} \times \nonumber \\
+ \frac{\partial\dot{T}}{\partial{T}} &= - \frac{\left( [C] \left( \frac{{C_p}_{\ns}}{T} - \frac{\text{d} {C_p} }{\text{d} T }_{\ns} \right) - \sum_{k=1}^{-1 + \ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns} + \frac{\text{d} {C_p} }{\text{d} T }_{k}\right) [C]_{k}\right)}{\left(\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}\right)^{2}} \times \nonumber \\
 				     & \mathindent[largerindent] \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k} - \nonumber \\
@@
  \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \times \nonumber \\
-				     & \left( - [C] \frac{\text{d} {C_p} }{\text{d} T }_{\ns} - \sum_{k=1}^{-1 + \ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns} + \frac{\text{d} {C_p} }{\text{d} T }_{k}\right) [C]_{k} + \frac{[C] {C_p}_{\ns}}{T} \right) - \nonumber \\
+				     & \left( [C] \left( \frac{{C_p}_{\ns}}{T} - \frac{\text{d} {C_p} }{\text{d} T }_{\ns} \right) - \sum_{k=1}^{-1 + \ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns} + \frac{\text{d} {C_p} }{\text{d} T }_{k}\right) [C]_{k}\right) - \nonumber \\
 				     & \sum_{k=1}^{-1 + \ns} \left(\left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial T }_{k} + \left(\frac{\text{d} H }{\text{d} T }_{k} - \frac{W_{k}}{W_{\ns}} \frac{\text{d} H }{\text{d} T }_{\ns}\right) \dot{\omega}_{k}\right) \Biggr]
@@
 and finally, we may substitute in the specific heat for the temperature derivative of enthalpy, giving:
+\begin{subequations}
 \begin{align}
@@
 \end{align}
+\end{subequations}
+The species production rate derivative with respect to temperature will be defined in~\cref{s:d_spec_prod}.
+
 \subsubsection{Molar derivative}
-The molar derivative of the energy equation is somewhat easier to work with,
-
-\subsection{Species production rate derivatives}
-First, we evaluate the derivatives of the net species production rate equation~\cref{e:spec_rop}, with respect to temperature:
-\begin{align}
- \label{e:dwdot_dT}
- \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial T }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial T }_{i} + \nu_{k,i} \frac{\partial R }{\partial T }_{i} c_{i}\right)} & \cond{dwdot}{,}
-\end{align}
-other species:
-\begin{align}
- \label{e:dwdot_dnj}
- \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial {n_j} }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial {n_j} }_{i} + \nu_{k,i} \frac{\partial R }{\partial {n_j} }_{i} c_{i}\right)} & \cond{dwdot}{,}
-\end{align}
-and the state parameter:
-\begin{subequations}
- \label{e:dwdot_de}
- \begin{align}
-  \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial V }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial V }_{i} + \nu_{k,i} \frac{\partial R }{\partial V }_{i} c_{i}\right)} & \cond{dwdot}{\dconp}\\
-  \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial P }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial P }_{i} + \nu_{k,i} \frac{\partial R }{\partial P }_{i} c_{i}\right)} & \cond{dwdot}{\dconv}
- \end{align}
-\end{subequations}
-
-\subsection{Rate of progress derivatives}
-\subsubsection{Temperature derivative}
-\label{s:dri_dt}
-Next, the rate of progress derivatives are evaluated, first with respect to temperature:
-\begin{equation}
- \frac{\partial R }{\partial T }_{i} = \frac{\partial}{\partial T}\left({k_f}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right)
-\end{equation}
-The derivative of the forward reaction rate coefficient is:
-\begin{equation}
- \label{e:dkf_dt}
- \frac{\text{d} {k_f} }{\text{d} T }_{i} = \frac{{k_f}_{i}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
-\end{equation}
-While evaluating derivatives for the Jacobian, it is important to expand terms---e.g., concentration, production rate, etc.---related to the last species to obtain the correct derivative, in this case using \cref{e:last_spec_conc}:
-\begin{equation}
- {R_f} = \left(\left(- \sum_{k=1}^{\ns  - 1} [C]_{k} + \frac{P}{T \Ru}\right)^{\nu^{\prime}_{\ns,i}}\right) {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
-\end{equation}
-The derivative of a species concentration with respect to temperature is zero for both \conp~and \conv:
-\begin{align}
- \frac{\partial [C]_{k}}{\partial T} &= \frac{\partial}{\partial T} \left(\frac{n_k}{V}\right), \quad \left\{k \ne \ns\right\} \nonumber \\
-				     &= 0
-\end{align}
-as the volume is either constant (\conp) or a state variable and a function of time only (\conv).
-Hence, the full derivative of the forward rate of progress is:
-\begin{equation}
- \label{e:dropfdt_1}
- \frac{\partial {R_f} }{\partial T }_{i} = \frac{\text{d} {k_f} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
-\end{equation}
-By defining a temporary variable $S^{\prime}_{l}$ (for species $l = \left\{j, \ns\right\}$):
-\begin{equation}
- \label{e:s_temp}
- S^{\prime}_{l} = \nu^{\prime}_{l,i} [C]_{l}^{\nu^{\prime}_{l,i} - 1} \prod_{\substack{1 \leq l \leq l - 1\\l + 1 \leq l \leq \ns}} [C]_{l}^{\nu^{\prime}_{l,i}}
-\end{equation}
-and using~\cref{e:dkf_dt}, \cref{e:dropfdt_1} can be simplified to:
-\begin{equation}
- \label{e:dropf_dt}
- \frac{\partial {R_f} }{\partial T }_{i} = - \frac{[C] S^{\prime}_{\ns}}{T} {k_f}_{i} + \frac{{R_f}_{i}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
-\end{equation}
-
-Starting from~\cref{e:ropr} and applying the same expansion of the last species' concentration as previously, the temperature derivative of the reverse rate of progress is:
-\begin{equation}
- \label{e:droprdt_1}
- \frac{\partial {R_r} }{\partial T }_{i} = \frac{\text{d} {k_r} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
-\end{equation}
-Next, \cref{e:kr} is considered to obtain the temperature derivative of the non-explicit reversible reaction rate coefficient:
-\begin{align}
- \frac{\text{d} {k_r} }{\text{d} T }_{i} &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } \frac{{k_f}_i}{{K_c}_i} + \frac{1}{{K_c}_{i}}\frac{\text{d} {k_f}_{i} }{\text{d} T } \right) \nonumber \\
-				         &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_r}_{i}
-\end{align}
-The temperature derivative of the equilibrium constant is:
-\begin{equation}
- \frac{\text{d} {K_c}_{i} }{\text{d} T } = {K_c}_{i} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} \;,
-\end{equation}
-with:
-\begin{equation}
- \frac{\text{d} B }{\text{d} T }_{k} = T \left(T \left(\frac{T a_{k,4}}{5} + \frac{a_{k,3}}{4}\right) + \frac{a_{k,2}}{3}\right) + \frac{a_{k,1}}{2} + \frac{1}{T} \left(a_{k,0} - 1 + \frac{a_{k,5}}{T}\right)
-\end{equation}
-giving:
-\begin{equation}
-\label{e:dkr_dt}
-\frac{\text{d} {k_r} }{\text{d} T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_r}_{i} \;,
-\end{equation}
-and using an temporary value $S^{\prime\prime}_{\ns}$---defined analagously to \cref{e:s_temp} for the reverse direction---we obtain:
-\begin{equation}
-\label{e:dropr_dt}
-\frac{\partial {R_r} }{\partial T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_r}_{i} - \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_r}_{i}
-\end{equation}
-
-Finally, combining~\cref{e:dropf_dt,e:dropr_dt} gives the total temperature derivative of the net rate of progress:
-\begin{align}
- \label{e:drop_dt}
- \frac{\partial R }{\partial T }_{i} =& \frac{{R_f}_{i}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right) - \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_r}_{i} \nonumber \\
-				      &\qquad + \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_r}_{i} - \frac{[C] S^{\prime}_{\ns}}{T} {k_f}_{i}
-\end{align}
-\subsubsection{Molar derivative}
-\label{s:dri_dnj}
-The derivative of the forward rate of progress with respect to the amount of moles of a species $j$ is:
-\begin{equation}
- \label{e:dropf_dnj_1}
- \frac{d}{d n_{k}} {R_{f_i}} = \left(\frac{\partial}{\partial n_{j}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right) {k_f}_{i} \;.
-\end{equation}
-The derivative of a species concentration with respect to moles of a species $j$ is given as:
-\begin{equation}
- \frac{\partial [C_k]}{\partial n_j} =
+Next the molar derivatives of the energy equation:
+\begin{align}
+ \mathcal{J}_{1,j} = \frac{\partial\dot{T}}{\partial{n_j}}
+\end{align}
+
+First, the derivative of a species concentration with respect to the moles of species $j$ is computed:
+\begin{align}
+ \frac{\partial [C_k]}{\partial n_j} &=
  \begin{dcases}
@@
  -\frac{1}{V} & k = \ns
- \end{dcases}
-\end{equation}
-Or put succinctly:
-\begin{equation}
+ \end{dcases} \\
+\intertext{Or put succinctly:}
 \label{e:dck_dnj}
-\frac{\partial [C_k]}{\partial n_j} = \frac{\delta_{j k} - \delta_{\ns k}}{V} \;.
-\end{equation}
-Combining~\cref{e:dropf_dnj_1,e:dck_dnj} the molar derivative of the forward rate of progress is found as:
-\begin{equation}
- \frac{\partial {R_f}_{i} }{\partial {n_j} } = {k_f}_{i} \sum_{k=1}^{\ns} \left(- \frac{\delta_{\ns k}}{V} + \frac{\delta_{j k}}{V}\right) \nu^{\prime}_{k,i} [C]_{k}^{\nu^{\prime}_{k,i} - 1} \prod_{\substack{1 \leq l \leq k - 1\\k + 1 \leq l \leq \ns}} [C]_{l}^{\nu^{\prime}_{l,i}}
-\end{equation}
-with the same temporary variables $S^{\prime}_{j}$ and $S^{\prime}_{\ns}$ as defined previously, this can be simplified to:
-\begin{equation}
- \label{e:dropf_dnj}
- \frac{\partial {R_f}_{i} }{\partial {n_j} } = \frac{{k_f}_{i}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
-\end{equation}
-and by the same process, the reverse rate of progress derivative is found to be:
-\begin{equation}
- \label{e:dropr_dnj}
- \frac{\partial {R_r}_{i} }{\partial {n_j} } = \frac{{k_r}_{i}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right)
-\end{equation}
-and the net rate of progress molar derivative:
-\begin{equation}
- \frac{\partial R_{i} }{\partial {n_j} } = - \frac{{k_r}_{i}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right) + \frac{{k_f}_{i}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
-\end{equation}
+\frac{\partial [C_k]}{\partial n_j} &= \frac{\delta_{j k} - \delta_{\ns k}}{V}
+\end{align}
+
+Similar to~\cref{s:denergy_dt} the derivative of the energy equation with respect to the moles of species $j$---after collapsing the specific-heat\slash~concentration sum---are given by:
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{\left(\sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}\right) \sum_{k=1}^{-1 + \ns} - \frac{\delta_{j k}}{V} \left({C_p}_{\ns} - {C_p}_{k}\right)}{\left(\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}\right)^{2}} - \nonumber \\
+				       & \mathindent \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k}
+\end{align}
+Next, the Kronecker delta summation is simplified, and the the compact form of the energy equation factored out yielding:
+\begin{subequations}
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} & \nonumber \\
+				       &  \mathindent[largerindentand] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_p}_{\ns} + {C_p}_{j}\right)\Biggr] & \text{\dconp} \\
+\intertext{and:}
+ \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} & \nonumber \\
+				       & \mathindent[largerindentand] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_v}_{\ns} + {C_v}_{j}\right) \Biggr] & \text{\dconv}
+\end{align}
+\end{subequations}
+Again, the species production rate derivative with respect to moles will be explored in~\cref{s:d_spec_prod}.
 
 \subsubsection{State parameter derivatives}
-\label{s:dri_de}
-The derivatives of the forward and reverse rates of progress with respect to the volume (\conp) or pressure (\conv) largely follow the same format of the previous sections.
+Next, the derivative of the energy equation with respect to the thermodynamic state parameter will be considered:
+\begin{subequations}
+\begin{align}
+ \mathcal{J}_{1, 2} & = \frac{\partial\dot{T}}{\partial{V}} & \text{\dconp} \\
+ \mathcal{J}_{1, 2} & = \frac{\partial\dot{T}}{\partial{P}} & \text{\dconv}
+\end{align}
+\end{subequations}
+
 First, the derivative of the concentration of a species $k$ with respect to the state parameter is:
-
-\begin{subequations}
- \begin{align}
- \lhs{dck}{\frac{\partial [C] }{\partial V }_{k}} &=
+\begin{subequations}
+ \label{e:dck_de}
+ \begin{align}
+ \lhs{dck}{\frac{\partial [C]_{k} }{\partial V }} &=
   \begin{dcases}
@@
  \lhs{dck}{\frac{\partial [C]_{k} }{\partial P }} &=
- \begin{dcases} \\
+ \begin{dcases}
    \rhs{dck}{\mathindent[s]0} & \mathcond{dck}{k \ne \ns} \\
@@
 \end{subequations}
-
-Following the outline of~\cref{s:dri_dnj}, the forward rate of progress derivatives are:
+while the specific heat and enthalpy\slash internal-energy are independent of the state parameter for both \conp~and \conv.
+Following the now familiar collapsing of the specific-heat\slash concentration sum, and substituting in the energy source term gives:
+\begin{subequations}
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{V}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ - \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial V }_{k} + & \nonumber \\
+				     & \mathindent[largerindentand] \frac{1}{V} \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{-1 + \ns} \left(- {C_p}_{\ns} + {C_p}_{k}\right) [C]_{k}\Biggr] & \text{\dconp} \\
+ \frac{\partial\dot{T}}{\partial{P}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial P }_{k} - & \nonumber \\
+				     & \mathindent[largerindentand] \frac{{C_v}_{\ns}}{T \Ru} \frac{\text{d} T }{\text{d} t } \Biggr] & \text{\dconv}
+\end{align}
+\end{subequations}
+Finally, the species production rate derivative with respect to the state parameter will be computed in~\cref{s:d_spec_prod}.
+
+\subsection{Molar source term derivatives}
+\label{s:molar_source_derivatives}
+\subsubsection{Temperature derivative}
+
+The derivative of the net molar rate of production of a species $k$ with respect to temperature is:
+\begin{align}
+ \mathcal{J}_{k + 2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial T } \nonumber \\
+ \intertext{or:}
+   \mathcal{J}_{k + 2, 1} & = V \frac{\partial \dot{\omega}_{k} }{\partial T }
+\end{align}
+
+\subsubsection{Molar derivative}
+
+Similarly, the derivative of the net molar rate of production of a species $k$ with respect to the moles of species $j$ is:
+\begin{align}
+ \mathcal{J}_{k + 2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial n_j } \nonumber \\
+  \intertext{or:}
+ \mathcal{J}_{k + 2, j + 2} & = V \frac{\partial \dot{\omega}_{k} }{\partial {n_j} }
+\end{align}
+
+\subsubsection{State parameter derivatives}
+
+Finally, the derivative of the net molar rate of production of a species $k$ with respect to the state parameter is:
+\begin{align}
+  \mathcal{J}_{k + 2, 2} &= \frac{\partial }{\partial V } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial V } & \nonumber \\
+			 &= V \frac{\partial \dot{\omega}_{k} }{\partial V } + \dot{\omega}_{k} & \text{\dconp}
+  \intertext{and:}
+  \mathcal{J}_{k + 2, 2} &= \frac{\partial }{\partial P } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial P } & \nonumber \\
+			 &=V \frac{\partial \dot{\omega}_{k} }{\partial P } & \text{\dconv}
+\end{align}
+
+\subsection{State parameter source term derivatives}
+\label{s:parameter_source_derivatives}
+\subsubsection{Temperature derivative}
+
+The derivatives of the thermodynamic state parameter source terms with respect to temperature are given by differentiating~\cref{e:param_complete}:
+\begin{align}
+ \mathcal{J}_{2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{V} }{\partial T} & \nonumber \\
+		    &= \frac{V}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(\frac{\partial \dot{\omega} }{\partial T }_{k} + \frac{\dot{\omega}_{k}}{T}\right) \nonumber \\
+		    &  \mathindent + \frac{V}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconp}
+\intertext{and:}
+ \mathcal{J}_{2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{P} }{\partial T} & \nonumber \\
+		    &= \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(T \frac{\partial \dot{\omega} }{\partial T }_{k} + \dot{\omega}_{k}\right) & \nonumber \\
+		    & \mathindent + \frac{P}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconv}
+\end{align}
+
+\subsubsection{Molar derivative}
+Similarly, the derivatives of the thermodynamic state parameter source terms with respect to the moles of species $j$ are:
+\begin{align}
+ \mathcal{J}_{2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{V} }{\partial n_j} & \nonumber \\
+		    &= V \left(\frac{1}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} + \frac{1}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} }\right) & \text{\dconp}
+\intertext{and:}
+ \mathcal{J}_{2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{P} }{\partial n_j} & \nonumber \\
+		    &=  \frac{P}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} } + T \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} & \text{\dconv}
+\end{align}
+
+\subsubsection{State parameter derivative}
+Finally, the derivative of the the thermodynamic state parameter source terms with respect to themselves are:
+\begin{align}
+ \mathcal{J}_{2, 2} &= \frac{\partial }{\partial V } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{V} }{\partial V} & \nonumber \\
+		    &= \frac{1}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(V \frac{\partial \dot{\omega} }{\partial V }_{k} + \dot{\omega}_{k}\right) \nonumber \\
+		    & \mathindent + \frac{1}{T} \left(V \frac{\text{d} \dot{T} }{\text{d} V } + \dot{T}\right) & \text{\dconp} \\
+ \intertext{and:}
+ \mathcal{J}_{2, 2} &= \frac{\partial }{\partial P } \frac{\text{d} P}{\text{d} t} = \frac{\partial \dot{P} }{\partial P} & \nonumber \\
+		    &= T \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial P }_{k} & \nonumber \\ 
+		    & \mathindent + \frac{1}{T} \left(P \frac{\text{d} \dot{T} }{\text{d} P } + \dot{T}\right) & \text{\dconv}
+\end{align}
+
+\subsection{Species production rate derivatives}
+\label{s:d_spec_prod}
+First, we evaluate the derivatives of the net species production rate equation~\cref{e:spec_rop}, with respect to temperature:
+\begin{align}
+ \label{e:dwdot_dT}
+ \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial T }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial T }_{i} + \nu_{k,i} \frac{\partial R }{\partial T }_{i} c_{i}\right)} & \cond{dwdot}{,}
+\end{align}
+other species:
+\begin{align}
+ \label{e:dwdot_dnj}
+ \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial {n_j} }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial {n_j} }_{i} + \nu_{k,i} \frac{\partial R }{\partial {n_j} }_{i} c_{i}\right)} & \cond{dwdot}{,}
+\end{align}
+and the state parameter:
+\begin{subequations}
+ \label{e:dwdot_de}
+ \begin{align}
+  \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial V }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial V }_{i} + \nu_{k,i} \frac{\partial R }{\partial V }_{i} c_{i}\right)} & \cond{dwdot}{\dconp}\\
+  \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial P }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial P }_{i} + \nu_{k,i} \frac{\partial R }{\partial P }_{i} c_{i}\right)} & \cond{dwdot}{\dconv}
+ \end{align}
+\end{subequations}
+
+\subsection{Rate of progress derivatives}
+\subsubsection{Temperature derivative}
+\label{s:dri_dt}
+Next, the rate of progress derivatives are evaluated, first with respect to temperature:
+\begin{equation}
+ \frac{\partial R }{\partial T }_{i} = \frac{\partial}{\partial T}\left({k_f}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right)
+\end{equation}
+The derivative of the forward reaction rate coefficient is:
+\begin{equation}
+ \label{e:dkf_dt}
+ \frac{\text{d} {k_f} }{\text{d} T }_{i} = \frac{{k_f}_{i}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
+\end{equation}
+While evaluating derivatives for the Jacobian, it is important to expand terms---e.g., concentration, production rate, etc.---related to the last species to obtain the correct derivative, in this case using \cref{e:last_spec_conc}:
+\begin{equation}
+ {R_f} = \left(\left(- \sum_{k=1}^{\ns  - 1} [C]_{k} + \frac{P}{T \Ru}\right)^{\nu^{\prime}_{\ns,i}}\right) {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
+\end{equation}
+
+Hence, the full derivative of the forward rate of progress is:
+\begin{equation}
+ \label{e:dropfdt_1}
+ \frac{\partial {R_f} }{\partial T }_{i} = \frac{\text{d} {k_f} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
+\end{equation}
+By defining a temporary variable $S^{\prime}_{l}$ (for species $l = \left\{j, \ns\right\}$):
+\begin{equation}
+ \label{e:s_temp}
+ S^{\prime}_{l} = \nu^{\prime}_{l,i} [C]_{l}^{\nu^{\prime}_{l,i} - 1} \prod_{\substack{1 \leq l \leq l - 1\\l + 1 \leq l \leq \ns}} [C]_{l}^{\nu^{\prime}_{l,i}}
+\end{equation}
+and using~\cref{e:dkf_dt}, \cref{e:dropfdt_1} can be simplified to:
+\begin{equation}
+ \label{e:dropf_dt}
+ \frac{\partial {R_f} }{\partial T }_{i} = - \frac{[C] S^{\prime}_{\ns}}{T} {k_f}_{i} + \frac{{R_f}_{i}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
+\end{equation}
+
+Starting from~\cref{e:ropr} and applying the same expansion of the last species' concentration as previously, the temperature derivative of the reverse rate of progress is:
+\begin{equation}
+ \label{e:droprdt_1}
+ \frac{\partial {R_r} }{\partial T }_{i} = \frac{\text{d} {k_r} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
+\end{equation}
+Next, \cref{e:kr} is considered to obtain the temperature derivative of the non-explicit reversible reaction rate coefficient:
+\begin{align}
+ \frac{\text{d} {k_r} }{\text{d} T }_{i} &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } \frac{{k_f}_i}{{K_c}_i} + \frac{1}{{K_c}_{i}}\frac{\text{d} {k_f}_{i} }{\text{d} T } \right) \nonumber \\
+				         &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_r}_{i}
+\end{align}
+The temperature derivative of the equilibrium constant is:
+\begin{equation}
+ \frac{\text{d} {K_c}_{i} }{\text{d} T } = {K_c}_{i} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} \;,
+\end{equation}
+with:
+\begin{equation}
+ \frac{\text{d} B }{\text{d} T }_{k} = T \left(T \left(\frac{T a_{k,4}}{5} + \frac{a_{k,3}}{4}\right) + \frac{a_{k,2}}{3}\right) + \frac{a_{k,1}}{2} + \frac{1}{T} \left(a_{k,0} - 1 + \frac{a_{k,5}}{T}\right)
+\end{equation}
+giving:
+\begin{equation}
+\label{e:dkr_dt}
+\frac{\text{d} {k_r} }{\text{d} T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_r}_{i} \;,
+\end{equation}
+and using an temporary value $S^{\prime\prime}_{\ns}$---defined analagously to \cref{e:s_temp} for the reverse direction---we obtain:
+\begin{equation}
+\label{e:dropr_dt}
+\frac{\partial {R_r} }{\partial T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_r}_{i} - \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_r}_{i}
+\end{equation}
+
+Finally, combining~\cref{e:dropf_dt,e:dropr_dt} gives the total temperature derivative of the net rate of progress:
+\begin{align}
+ \label{e:drop_dt}
+ \frac{\partial R }{\partial T }_{i} =& \frac{{R_f}_{i}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right) - \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_r}_{i} \nonumber \\
+				      &\qquad + \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_r}_{i} - \frac{[C] S^{\prime}_{\ns}}{T} {k_f}_{i}
+\end{align}
+\subsubsection{Molar derivative}
+\label{s:dri_dnj}
+The derivative of the forward rate of progress with respect to the amount of moles of a species $j$ is:
+\begin{equation}
+ \label{e:dropf_dnj_1}
+ \frac{d}{d n_{k}} {R_{f_i}} = \left(\frac{\partial}{\partial n_{j}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right) {k_f}_{i} \;.
+\end{equation}
+
+Combining~\cref{e:dropf_dnj_1,e:dck_dnj} the molar derivative of the forward rate of progress is found as:
+\begin{equation}
+ \frac{\partial {R_f}_{i} }{\partial {n_j} } = {k_f}_{i} \sum_{k=1}^{\ns} \left(- \frac{\delta_{\ns k}}{V} + \frac{\delta_{j k}}{V}\right) \nu^{\prime}_{k,i} [C]_{k}^{\nu^{\prime}_{k,i} - 1} \prod_{\substack{1 \leq l \leq k - 1\\k + 1 \leq l \leq \ns}} [C]_{l}^{\nu^{\prime}_{l,i}}
+\end{equation}
+with the same temporary variables $S^{\prime}_{j}$ and $S^{\prime}_{\ns}$ as defined previously, this can be simplified to:
+\begin{equation}
+ \label{e:dropf_dnj}
+ \frac{\partial {R_f}_{i} }{\partial {n_j} } = \frac{{k_f}_{i}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
+\end{equation}
+and by the same process, the reverse rate of progress derivative is found to be:
+\begin{equation}
+ \label{e:dropr_dnj}
+ \frac{\partial {R_r}_{i} }{\partial {n_j} } = \frac{{k_r}_{i}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right)
+\end{equation}
+and the net rate of progress molar derivative:
+\begin{equation}
+ \frac{\partial R_{i} }{\partial {n_j} } = - \frac{{k_r}_{i}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right) + \frac{{k_f}_{i}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
+\end{equation}
+
+\subsubsection{State parameter derivatives}
+\label{s:dri_de}
+Following the outline of~\cref{s:dri_dnj}---while utilising~\cref{e:dck_de} for the species concentration derivatives---the forward rate of progress derivatives are:
 \begin{subequations}
@@
 
-\subsection{Final Jacobian entries}
-\label{s:jac_final}
-
 \printbibliography

### 2018-01-22 derivations.tex
fixes

@@
 \begin{align}
-n &= \frac{V P}{T \Ru} = \sum_{i=1}^{\ns}{n_i}, \label{e:source_moles}\\
+n &= \frac{V P}{T \Ru} = \sum_{i=1}^{\ns}{n_i}\; , \nonumber \\
 n_{\ns} &= \frac{V P}{T \Ru} - \sum_{i=1}^{\ns - 1}{n_i}
@@
 \begin{align}
-f &= \frac{\text{d} \Phi }{\text{d} t } = \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} V }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\}\dconp, \\
-f &= \frac{\text{d} \Phi }{\text{d} t } = \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} P }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\}\dconv
+f &= \frac{\text{d} \Phi }{\text{d} t } = \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} V }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\} & \text{\dconp} \\
+f &= \frac{\text{d} \Phi }{\text{d} t } = \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} P }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\} & \text{\dconv}
 \end{align}
@@
 \begin{equation}
-\frac{\text{d} n }{\text{d} t }_{k} = V \dot{\omega}_{k} \label{e:spec}
+\frac{\text{d} n }{\text{d} t }_{k} = V \dot{\omega}_{k} \qquad k = 1, \ldots, \ns - 1, \label{e:spec}
 \end{equation}
@@
 \end{subequations}
-where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure specific heat and constant-volume specific heat of species $k$ in molar units respectively, while $[C]_{k}$ is the concentration, given by:
+where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure and constant-volume specific heat of species $k$ in molar units respectively, while $[C]_{k}$ is the concentration, given by:
 \begin{equation}
@@
 \begin{align}
-\frac{\text{d} V }{\text{d} t } &= \frac{\Ru}{P} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right)\dconp, \\
-\frac{\text{d} P }{\text{d} t } &= \frac{\Ru}{V} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right)\dconv
+\frac{\text{d} V }{\text{d} t } &= \frac{\Ru}{P} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right) & \text{\dconp} \\
+\frac{\text{d} P }{\text{d} t } &= \frac{\Ru}{V} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right) & \text{\dconv}
 \end{align}
@@
  \label{e:mass_conservation}
- \frac{\text{d} m }{\text{d} t } = 0 = \sum_{k=1}^{\ns}  W_{k} \frac{\text{d} n }{\text{d} t }_{k},
+ \frac{\text{d} m }{\text{d} t } = 0 = \sum_{k=1}^{\ns}  W_{k} \frac{\text{d} n_{k} }{\text{d} t },
 \end{equation}
@@
 \begin{equation}
- \frac{\text{d} n }{\text{d} t }_{\ns} = - \frac{1}{W_{\ns}} \sum_{k=1}^{\ns - 1} W_{k} \frac{\text{d} n }{\text{d} t }_{k}
+ \frac{\text{d} n }{\text{d} t }_{\ns} = \frac{-1}{W_{\ns}} \sum_{k=1}^{\ns - 1} W_{k} \frac{\text{d} n_{k} }{\text{d} t }
  \label{e:spec_ns}
@@
 \begin{align}
-\frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns} \frac{\text{d} n }{\text{d} t }_{k}, \nonumber \\
-\frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\text{d} n }{\text{d} t }_{k}
+\frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns} \frac{\text{d} n_{k} }{\text{d} t }, \nonumber \\
+\frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\text{d} n_{k} }{\text{d} t }
 \label{e:total_molar}
@@
 \begin{align}
-\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns  - 1} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{p,\ns}} + \sum_{k=1}^{\ns  - 1} \left(- {C_{p,\ns}} + {C_{p, k}}\right) [C]_{k}} &\text{\dconp}\\
-\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns  - 1} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{v,\ns}} + \sum_{k=1}^{\ns  - 1} \left(- {C_{v,\ns}} + {C_{v, k}}\right) [C]_{k}} &\text{\dconv}
-\end{align}
-\end{subequations}
-This form is used for differentiation, but~\cref{e:temperature_incomplete} will often be used as well due to its compactness.
+\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns  - 1} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{p,\ns}} + \sum_{k=1}^{\ns  - 1} \left({C_{p, k}} - {C_{p,\ns}}\right) [C]_{k}} &\text{\dconp}\\
+\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns  - 1} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{v,\ns}} + \sum_{k=1}^{\ns  - 1} \left({C_{v, k}} - {C_{v,\ns}}\right) [C]_{k}} &\text{\dconv}
+\end{align}
+\end{subequations}
+This form of the temperature source term will be used for derivation of Jacobian terms, but~\cref{e:temperature_incomplete} will often be used as well due to its compactness.
 
@@
 \begin{align}
-R_{i} &= {R_f}_{i} - {R_r}_{i} \label{e:ropnet}\\
-{R_f}_{i} &= {k_f}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} \label{e:ropf}\\
-{R_r}_{i} &= {k_r}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} \label{e:ropr}
-\end{align}
-where ${k_f}_{i}$ and ${k_r}_{i}$ are the forward and reverse reaction coefficients (respectively) for the $i$th reaction, and the third-body\slash pressure modification $c_{i}$ is given by:
+R_{i} &= {R_{f,i}} - {R_{r,i}} \label{e:ropnet}\\
+{R_{f, i}} &= {k_{f, i}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} \label{e:ropf}\\
+{R_{r, i}} &= {k_{r, i}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} \label{e:ropr}
+\end{align}
+where ${k_{f, i}}$ and ${k_{r, i}}$ are the forward and reverse reaction coefficients (respectively) for the $i$th reaction, and the third-body\slash pressure modification $c_{i}$ is given by:
 \begin{equation}
@@
   \label{e:arrhenius}
-  {k_f}_{i} = A_{i} T^{\beta_{i}} \operatorname{exp}\left({- \frac{{E_{a}}_{i}}{T \Ru}}\right) \;,
+  {k_{f, i}} = A_{i} T^{\beta_{i}} \operatorname{exp}\left({- \frac{{E_{a}}_{i}}{T \Ru}}\right) \;,
 \end{equation}
@@
 \subsection{Reverse rate coefficient}
-By definition, the reverse rate coefficient---${k_r}_{i}$---of irreversible reactions is zero, while reversible reactions may have non-zero ${k_r}_{i}$.
+By definition, the reverse rate coefficient---${k_{r, i}}$---of irreversible reactions is zero, while reversible reactions may have non-zero ${k_{r, i}}$.
 Note that in \texttt{pyJac}, reversible reactions with explicit reverse Arrhenius parameters are split into two irreversible reactions; this simplifies calculation inside the generated code, and eases comparison to Cantera~\cite{Goodwin:2015aa} which applies the same transformation.
@@
 \begin{align}
- {k_r}_{i} &= \frac{{k_f}_{i}}{{K_c}_{i}}\; , \label{e:kr}\\
+ {k_{r, i}} &= \frac{{k_{f, i}}}{{K_c}_{i}}\; , \label{e:kr}\\
  {K_c}_{i} &= \left(\left(\frac{P_{atm}}{T \Ru}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}}\right) {K_p}_{i}\; ,\text{ and} \label{e:kc}\\
@@
 \label{e:kf_plog}
-\log{\left ({k_f}_{i} \right )} = \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) \left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \log{\left (k_{1} \right )}
+\log{\left ({k_{f, i}} \right )} = \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) \left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \log{\left (k_{1} \right )}
 \end{equation}
@@
 \begin{equation}
- \frac{\partial R }{\partial T }_{i} = \frac{\partial}{\partial T}\left({k_f}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right)
+ \frac{\partial R }{\partial T }_{i} = \frac{\partial}{\partial T}\left({k_{f, i}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right)
 \end{equation}
@@
  \label{e:dkf_dt}
- \frac{\text{d} {k_f} }{\text{d} T }_{i} = \frac{{k_f}_{i}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
+ \frac{\text{d} {k_f} }{\text{d} T }_{i} = \frac{{k_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
 \end{equation}
@@
 \begin{equation}
- {R_f} = \left(\left(- \sum_{k=1}^{\ns  - 1} [C]_{k} + \frac{P}{T \Ru}\right)^{\nu^{\prime}_{\ns,i}}\right) {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
+ {R_f} = \left(\left(- \sum_{k=1}^{\ns  - 1} [C]_{k} + \frac{P}{T \Ru}\right)^{\nu^{\prime}_{\ns,i}}\right) {k_{f, i}} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
 \end{equation}
@@
  \label{e:dropfdt_1}
- \frac{\partial {R_f} }{\partial T }_{i} = \frac{\text{d} {k_f} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
+ \frac{\partial {R_f} }{\partial T }_{i} = \frac{\text{d} {k_f} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime}_{\ns,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
 \end{equation}
@@
  \label{e:dropf_dt}
- \frac{\partial {R_f} }{\partial T }_{i} = - \frac{[C] S^{\prime}_{\ns}}{T} {k_f}_{i} + \frac{{R_f}_{i}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
+ \frac{\partial {R_f} }{\partial T }_{i} = - \frac{[C] S^{\prime}_{\ns}}{T} {k_{f, i}} + \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
 \end{equation}
@@
  \label{e:droprdt_1}
- \frac{\partial {R_r} }{\partial T }_{i} = \frac{\text{d} {k_r} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
+ \frac{\partial {R_r} }{\partial T }_{i} = \frac{\text{d} {k_r} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime\prime}_{\ns,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
 \end{equation}
@@
 \begin{align}
- \frac{\text{d} {k_r} }{\text{d} T }_{i} &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } \frac{{k_f}_i}{{K_c}_i} + \frac{1}{{K_c}_{i}}\frac{\text{d} {k_f}_{i} }{\text{d} T } \right) \nonumber \\
-				         &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_r}_{i}
+ \frac{\text{d} {k_r} }{\text{d} T }_{i} &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } \frac{{k_f}_i}{{K_c}_i} + \frac{1}{{K_c}_{i}}\frac{\text{d} {k_{f, i}} }{\text{d} T } \right) \nonumber \\
+				         &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_{r, i}}
 \end{align}
@@
 \label{e:dkr_dt}
-\frac{\text{d} {k_r} }{\text{d} T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_r}_{i} \;,
+\frac{\text{d} {k_r} }{\text{d} T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_{r, i}} \;,
 \end{equation}
@@
 \label{e:dropr_dt}
-\frac{\partial {R_r} }{\partial T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_r}_{i} - \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_r}_{i}
+\frac{\partial {R_r} }{\partial T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_{r, i}} - \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_{r, i}}
 \end{equation}
@@
  \label{e:drop_dt}
- \frac{\partial R }{\partial T }_{i} =& \frac{{R_f}_{i}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right) - \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_r}_{i} \nonumber \\
-				      &\qquad + \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_r}_{i} - \frac{[C] S^{\prime}_{\ns}}{T} {k_f}_{i}
+ \frac{\partial R }{\partial T }_{i} =& \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right) - \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_{r, i}} \nonumber \\
+				      &\qquad + \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_{r, i}} - \frac{[C] S^{\prime}_{\ns}}{T} {k_{f, i}}
 \end{align}
@@
  \label{e:dropf_dnj_1}
- \frac{d}{d n_{k}} {R_{f_i}} = \left(\frac{\partial}{\partial n_{j}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right) {k_f}_{i} \;.
+ \frac{d}{d n_{k}} {R_{f_i}} = \left(\frac{\partial}{\partial n_{j}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right) {k_{f, i}} \;.
 \end{equation}
@@
 \begin{equation}
- \frac{\partial {R_f}_{i} }{\partial {n_j} } = {k_f}_{i} \sum_{k=1}^{\ns} \left(- \frac{\delta_{\ns k}}{V} + \frac{\delta_{j k}}{V}\right) \nu^{\prime}_{k,i} [C]_{k}^{\nu^{\prime}_{k,i} - 1} \prod_{\substack{1 \leq l \leq k - 1\\k + 1 \leq l \leq \ns}} [C]_{l}^{\nu^{\prime}_{l,i}}
+ \frac{\partial {R_{f, i}} }{\partial {n_j} } = {k_{f, i}} \sum_{k=1}^{\ns} \left(- \frac{\delta_{\ns k}}{V} + \frac{\delta_{j k}}{V}\right) \nu^{\prime}_{k,i} [C]_{k}^{\nu^{\prime}_{k,i} - 1} \prod_{\substack{1 \leq l \leq k - 1\\k + 1 \leq l \leq \ns}} [C]_{l}^{\nu^{\prime}_{l,i}}
 \end{equation}
@@
  \label{e:dropf_dnj}
- \frac{\partial {R_f}_{i} }{\partial {n_j} } = \frac{{k_f}_{i}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
+ \frac{\partial {R_{f, i}} }{\partial {n_j} } = \frac{{k_{f, i}}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
 \end{equation}
@@
  \label{e:dropr_dnj}
- \frac{\partial {R_r}_{i} }{\partial {n_j} } = \frac{{k_r}_{i}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right)
+ \frac{\partial {R_{r, i}} }{\partial {n_j} } = \frac{{k_{r, i}}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right)
 \end{equation}
@@
 \begin{equation}
- \frac{\partial R_{i} }{\partial {n_j} } = - \frac{{k_r}_{i}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right) + \frac{{k_f}_{i}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
+ \frac{\partial R_{i} }{\partial {n_j} } = - \frac{{k_{r, i}}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right) + \frac{{k_{f, i}}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
 \end{equation}
@@
  \begin{align}
-  \lhs{dri_de}{\frac{\partial {R_f}_{i} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime}_{\ns}}{V} {k_f}_{i} - \frac{{R_f}_{i}}{V} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}}&\cond{dri_de}{\dconp}\\
-  \lhs{dri_de}{\frac{\partial {R_f}_{i} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime}_{\ns} {k_f}_{i}}{T \Ru}}&\cond{dri_de}{\dconv}
+  \lhs{dri_de}{\frac{\partial {R_{f, i}} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime}_{\ns}}{V} {k_{f, i}} - \frac{{R_{f, i}}}{V} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}}&\cond{dri_de}{\dconp}\\
+  \lhs{dri_de}{\frac{\partial {R_{f, i}} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime}_{\ns} {k_{f, i}}}{T \Ru}}&\cond{dri_de}{\dconv}
  \end{align}
@@
  \begin{align}
-  \lhs{dri_de}{\frac{\partial {R_r}_{i} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime\prime}_{\ns}}{V} {k_r}_{i} - \frac{{R_r}_{i}}{V} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i}}&\cond{dri_de}{\dconp}\\
-  \lhs{dri_de}{\frac{\partial {R_r}_{i} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime\prime}_{\ns} {k_r}_{i}}{T \Ru}}&\cond{dri_de}{\dconv}
+  \lhs{dri_de}{\frac{\partial {R_{r, i}} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime\prime}_{\ns}}{V} {k_{r, i}} - \frac{{R_{r, i}}}{V} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i}}&\cond{dri_de}{\dconp}\\
+  \lhs{dri_de}{\frac{\partial {R_{r, i}} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime\prime}_{\ns} {k_{r, i}}}{T \Ru}}&\cond{dri_de}{\dconv}
  \end{align}
@@
  \begin{align}
-  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial V }} &= \rhs{dri_de}{\frac{1}{V}\Biggl( [C]\left(S^{\prime}_{\ns} {k_f}_{i} - S^{\prime\prime}_{\ns} {k_r}_{i}\right) + } & \nonumber\\
-  & \rhs{dri_de}{\mathindent[farth] \left({R_r}_{i} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i} - {R_f}_{i} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}\right) \Biggr)}&\cond{dri_de}{\dconp}\\
-  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial P }} &= \rhs{dri_de}{\frac{1}{T \Ru} \left(S^{\prime}_{\ns} {k_f}_{i} - S^{\prime\prime}_{\ns} {k_r}_{i}\right)}&\cond{dri_de}{\dconv}
+  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial V }} &= \rhs{dri_de}{\frac{1}{V}\Biggl( [C]\left(S^{\prime}_{\ns} {k_{f, i}} - S^{\prime\prime}_{\ns} {k_{r, i}}\right) + } & \nonumber\\
+  & \rhs{dri_de}{\mathindent[farth] \left({R_{r, i}} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i} - {R_{f, i}} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}\right) \Biggr)}&\cond{dri_de}{\dconp}\\
+  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial P }} &= \rhs{dri_de}{\frac{1}{T \Ru} \left(S^{\prime}_{\ns} {k_{f, i}} - S^{\prime\prime}_{\ns} {k_{r, i}}\right)}&\cond{dri_de}{\dconv}
  \end{align}
@@
 \begin{align}
- \frac{\partial {k_f}_{i} }{\partial T } &= \biggl( \frac{1}{k_{1}} \frac{\text{d} k_1 }{\text{d} T } + \frac{1}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} \left(- \frac{1}{k_{1}} \frac{\text{d} k_1 }{\text{d} T } + \frac{1}{k_{2}} \frac{\text{d} k_2 }{\text{d} T }\right) \times  \Bigl( \nonumber \\
-					 & \mathindent - \log{\left (P_{1} \right )} + \log{\left (P \right )}\Bigr)\biggr) {k_f}_{i}
+ \frac{\partial {k_{f, i}} }{\partial T } &= \biggl( \frac{1}{k_{1}} \frac{\text{d} k_1 }{\text{d} T } + \frac{1}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} \left(- \frac{1}{k_{1}} \frac{\text{d} k_1 }{\text{d} T } + \frac{1}{k_{2}} \frac{\text{d} k_2 }{\text{d} T }\right) \times  \Bigl( \nonumber \\
+					 & \mathindent - \log{\left (P_{1} \right )} + \log{\left (P \right )}\Bigr)\biggr) {k_{f, i}}
 \end{align}
@@
  \label{e:dkf_plog_dt}
- \frac{\partial {k_f}_{i} }{\partial T } &= \frac{{k_f}_{i}}{T} \Biggl(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \nonumber \\
+ \frac{\partial {k_{f, i}} }{\partial T } &= \frac{{k_{f, i}}}{T} \Biggl(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \nonumber \\
 					 & \mathindent[morespac] \beta_1 + \frac{E_{a_1}}{T \Ru}\Biggr)
@@
  \frac{\partial R_{i} }{\partial T } &= \left(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \beta_1 + \frac{E_{a_1}}{T \Ru} \right) \times \nonumber \\
-				     & \mathindent \left(\frac{{R_f}_{i} - {R_r}_{i}}{T}\right) + {R_r}_{i} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_r}_{i} - S^{\prime}_{\ns} {k_f}_{i}\right)
+				     & \mathindent \left(\frac{{R_{f, i}} - {R_{r, i}}}{T}\right) + {R_{r, i}} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_{r, i}} - S^{\prime}_{\ns} {k_{f, i}}\right)
 \end{align}
@@
 \begin{equation}
- \frac{\partial {k_f}_{i} }{\partial P } = \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) {k_f}_{i}}{P \left(- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}\right)}
+ \frac{\partial {k_{f, i}} }{\partial P } = \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) {k_{f, i}}}{P \left(- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}\right)}
 \end{equation}
@@
 \begin{align}
- \frac{\partial R_{i} }{\partial P } &= \frac{1}{T \Ru} \left(- S^{\prime\prime}_{\ns} {k_r}_{i} + S^{\prime}_{\ns} {k_f}_{i}\right) + & \nonumber \\
-				     &  \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) \left({R_f}_{i} - {R_r}_{i}\right)}{P \left(- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}\right)} & \text{\dconv}
+ \frac{\partial R_{i} }{\partial P } &= \frac{1}{T \Ru} \left(- S^{\prime\prime}_{\ns} {k_{r, i}} + S^{\prime}_{\ns} {k_{f, i}}\right) + & \nonumber \\
+				     &  \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) \left({R_{f, i}} - {R_{r, i}}\right)}{P \left(- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}\right)} & \text{\dconv}
 \end{align}
@@
 \begin{align}
- \frac{\partial {k_f}_{i} }{\partial T } &= \log{\left (10 \right )} {k_f}_{i} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{\text{d} \tilde{T} }{\text{d} T } \left(j - 1\right) \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}
+ \frac{\partial {k_{f, i}} }{\partial T } &= \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{\text{d} \tilde{T} }{\text{d} T } \left(j - 1\right) \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}
 \end{align}
@@
 \begin{equation}
- \frac{\partial {k_f}_{i} }{\partial T } = \log{\left (10 \right )} {k_f}_{i} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} - \frac{2 \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right)
+ \frac{\partial {k_{f, i}} }{\partial T } = \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} - \frac{2 \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right)
 \end{equation}
@@
 \begin{align}
- \frac{\partial R }{\partial T }_{i} &= \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \left[ - \frac{2 \log{\left (10 \right )} \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right) \right] \left({R_f}_{i} - {R_r}_{i}\right) + \nonumber \\
-				     & \mathindent \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} {R_r}_{i} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_r}_{i} - S^{\prime}_{\ns} {k_f}_{i}\right)
+ \frac{\partial R }{\partial T }_{i} &= \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \left[ - \frac{2 \log{\left (10 \right )} \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right) \right] \left({R_{f, i}} - {R_{r, i}}\right) + \nonumber \\
+				     & \mathindent \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} {R_{r, i}} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_{r, i}} - S^{\prime}_{\ns} {k_{f, i}}\right)
 \end{align}
@@
 \begin{equation}
- \frac{\partial {k_f}_{i} }{\partial P } = \log{\left (10 \right )} {k_f}_{i} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{\text{d} \tilde{P} }{\text{d} P } \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}
+ \frac{\partial {k_{f, i}} }{\partial P } = \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{\text{d} \tilde{P} }{\text{d} P } \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}
 \end{equation}
@@
 \begin{equation}
- \frac{\partial {k_f}_{i} }{\partial P } = \log{\left (10 \right )} {k_f}_{i} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)}
-\end{equation}
-and:
-\begin{align}
- \frac{\partial R_{i} }{\partial P } &= \log{\left (10 \right )} \left({R_f}_{i} - {R_r}_{i}\right) \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)} + \nonumber \\
-				     & \mathindent \frac{1}{T \Ru} \left(- S^{\prime\prime}_{\ns} {k_r}_{i} + S^{\prime}_{\ns} {k_f}_{i}\right)
+ \frac{\partial {k_{f, i}} }{\partial P } = \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)}
+\end{equation}
+and:
+\begin{align}
+ \frac{\partial R_{i} }{\partial P } &= \log{\left (10 \right )} \left({R_{f, i}} - {R_{r, i}}\right) \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)} + \nonumber \\
+				     & \mathindent \frac{1}{T \Ru} \left(- S^{\prime\prime}_{\ns} {k_{r, i}} + S^{\prime}_{\ns} {k_{f, i}}\right)
 \end{align}

### 2018-01-22 derivations.tex
fix

@@
 \subsection{Reverse rate coefficient}
-By definition, the reverse rate coefficient---${k_{r, i}}$---of irreversible reactions is zero, while reversible reactions may have non-zero ${k_{r, i}}$.
+By definition, the reverse rate coefficient, ${k_{r, i}}$, of irreversible reactions is zero, while reversible reactions may have non-zero ${k_{r, i}}$.
 Note that in \texttt{pyJac}, reversible reactions with explicit reverse Arrhenius parameters are split into two irreversible reactions; this simplifies calculation inside the generated code, and eases comparison to Cantera~\cite{Goodwin:2015aa} which applies the same transformation.
@@
 \begin{align}
- {k_{r, i}} &= \frac{{k_{f, i}}}{{K_c}_{i}}\; , \label{e:kr}\\
- {K_c}_{i} &= \left(\left(\frac{P_{atm}}{T \Ru}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}}\right) {K_p}_{i}\; ,\text{ and} \label{e:kc}\\
- {K_p}_{i} &= \text{exp}\left(\frac{\Delta S^{\circ}_k}{\Ru} - \frac{\Delta H^{\circ}_k}{\Ru T}\right) = \text{exp}\left(\sum_{k=1}^{\ns}\nu_{ki}\left(\frac{S^{\circ}_k}{\Ru} - \frac{H^{\circ}_k}{\Ru T}\right)\right) \label{e:kp}
+ {k_{r, i}} &= \frac{{k_{f, i}}}{{K_{c, i}}}\; , \label{e:kr}\\
+ {K_{c, i}} &= {K_{p,i}}\left(\frac{P_{atm}}{T \Ru}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}} \; ,\text{ and} \label{e:kc}\\
+ {K_{p,i}} &= \text{exp}\left(\frac{\Delta S^{\circ}_i}{\Ru} - \frac{\Delta H^{\circ}_i}{\Ru T}\right) = \text{exp}\left(\sum_{k=1}^{\ns}\nu_{ki}\left(\frac{S^{\circ}_k}{\Ru} - \frac{H^{\circ}_k}{\Ru T}\right)\right) \label{e:kp}
 \end{align}
@@
  \label{e:kc_in_kp}
- {K_c}_{i} = \left(\left(\frac{P_{atm}}{\Ru}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}}\right) \operatorname{exp}\left({\sum_{k=1}^{\ns} \nu_{k,i} B_{k}}\right)
+ {K_{c, i}} = \left(\left(\frac{P_{atm}}{\Ru}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}}\right) \operatorname{exp}\left({\sum_{k=1}^{\ns} \nu_{k,i} B_{k}}\right)
 \end{equation}
@@
 \begin{align}
- \frac{\text{d} {k_r} }{\text{d} T }_{i} &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } \frac{{k_f}_i}{{K_c}_i} + \frac{1}{{K_c}_{i}}\frac{\text{d} {k_{f, i}} }{\text{d} T } \right) \nonumber \\
-				         &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_{r, i}}
+ \frac{\text{d} {k_r} }{\text{d} T }_{i} &= \left(- \frac{1}{{K_{c, i}}} \frac{\text{d} {K_{c, i}} }{\text{d} T } \frac{{k_f}_i}{{K_c}_i} + \frac{1}{{K_{c, i}}}\frac{\text{d} {k_{f, i}} }{\text{d} T } \right) \nonumber \\
+				         &= \left(- \frac{1}{{K_{c, i}}} \frac{\text{d} {K_{c, i}} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_{r, i}}
 \end{align}
@@
 \begin{equation}
- \frac{\text{d} {K_c}_{i} }{\text{d} T } = {K_c}_{i} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} \;,
+ \frac{\text{d} {K_{c, i}} }{\text{d} T } = {K_{c, i}} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} \;,
 \end{equation}

### 2018-01-23 paper.tex
add first version of paper based on NCM

\documentclass[12pt,number,sort&compress]{elsarticle}

%======================================================================
\usepackage{todonotes}
\usepackage{graphicx}
\usepackage[binary-units]{siunitx}
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}, e.g., \ce{H2SO4}
\usepackage{booktabs,multicol} %better tables
\usepackage{subcaption} %subfigs
\sisetup{group-separator={,},
	detect-all,
	binary-units,
	list-units = single,
	range-units = single,
	range-phrase = --,
	per-mode = symbol-or-fraction,
	separate-uncertainty = true,
	multi-part-units = single,
	list-final-separator = {, and }
	%    scientific-notation = fixed
}
\usepackage[breaklinks=true, linkcolor=blue, citecolor=blue, colorlinks=true]{hyperref}
% load after hyperref
\usepackage[english,capitalise]{cleveref}

%======================================================================
% Add your bibliography file here, replace template.bib
\bibliographystyle{elsarticle-num}

\title{SIMD-vectorized Chemical Source Term Evaluation}

\author[1]{Nicholas Curtis\corref{corr}}
\ead{nicholas.curtis@uconn.edu}
\author[1]{Chih-Jen Sung}

\address[1]{Department of Mechanical Engineering, University of Connecticut, Storrs, CT 06269, USA}
\cortext[corr]{Corresponding author}

\begin{document}

\begin{frontmatter}

%====================================================================
\begin{abstract} % not to exceed 200 words
A code generation platform for single-instruction, multiple-data (SIMD) vectorized chemical source term evaluation was developed and validated against Cantera for a wide range of chemical kinetic models.
Speedups of up to \SIrange{1.90}{2.45}{$\times$} were observed over a baseline single-instruction, multiple-thread (SIMT) parallelized code.
Additionally, use of a row-major data layout was found to increase both the SIMD and SIMT performance by up to \SIrange{2.00}{2.38}{$\times$} over a column-major data layout.
Larger vector-widths were found to produce speedups up to \SI{1.8}{$\times$} as compared to a small vector-width for smaller models, e.g. GRI-Mech 3.0, but as the model size increased the performance of all vector-widths were similar.
Finally, the parallel scaling efficiency was found to be slightly better for larger models, and future extensions of this work were discussed.
\end{abstract}

% (Provide 2-4 keywords describing your research. Only abbreviations firmly
% established in the field may be used. These keywords will be used for
% sessioning/indexing purposes.)
\begin{keyword}
    Chemical Kinetics\sep SIMD\sep SIMT\sep Vector Processing\sep Jacobian
\end{keyword}

\end{frontmatter}

%====================================================================
\section{Introduction}
%

Single-Instruction, Multiple-Data (SIMD) and the related Single-Instruction, Multiple-Thread (SIMT) processing are two important vector-processing paradigms used increasingly in scientific computing.
Traditional multi-core parallelism is often used to increase central processing unit (CPU) performance, however SIMD processors---as well as SIMT processors, e.g. in the form of graphics processing units (GPUs)---have gained recognition due to their increased floating operation throughput.
The parallel programming standard OpenCL~\cite{stone2010opencl} has further enabled adoption of vector-processing based codes in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems (e.g. CPU, GPU, or Intel's Many Integrated Core architecture).

\begin{figure}[htb]
  \centering
  \begin{minipage}[t]{0.45\linewidth}
      \includegraphics[width=\textwidth]{SIMD.pdf}
      \caption{Schematic of SIMD processing.  A single processing element (e.g. a CPU core) contains a vector unit with $N_v$ lanes ($\text{L}_0,\ldots \text{L}_{N_v}$).  The vector unit executes a single instruction concurrently on multiple data.}
      \label{F:SIMD}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.45\linewidth}
      \includegraphics[width=\textwidth]{mem_access.pdf}
      \caption{Representation of an 2-D array with dimension $N \times K$ in CPU memory, with linear index marked.}
      \label{F:mem}
  \end{minipage}
\end{figure}

A SIMD instruction utilizes specialized vector processing hardware to execute the same floating point operation (e.g. multiplication, division, etc.) on a several different data concurrently (Fig.~\ref{F:SIMD}); the number of concurrent operations is known as the vector-width\footnote{OpenCL allows for use of vector-widths different from the actual hardware vector-width via implicit conversion, and may provide some performance benefit as will be studied in Sec.~\ref{S:results}}.
Nearly all modern CPUs have vector processing units capable of running SIMD instructions, typically capable of handling 2--4 concurrent double precision operations.
Additionally, specialized hardware accelerators, e.g. Intel's Xeon Phi co-processor or multi integrated core architecture (MIC), have been developed that have tens of cores with very wide vector-widths (e.g. 4--8 double precision operations); these very wide vector-widths are also expected on forthcoming Intel CPUs (the Skylake Xeon and Cannon Lake architectures).

\subsection{Data ordering and vectorization patterns}
\label{S:data}
When storing arrays for a chemical kinetic model, e.g. the concentration of species $k$ for the $j$-th of thermo-chemical state $[C_k]_j$---with $1 \le k \le N_{\text{sp}}$ the number of species in the model, and $1 \le j \le N_{\text{state}}$ the number of thermo-chemical states considered---there are two data-ordering choices.
In the ``F'' (Fortran column-major) format, the concentrations of a single species $k$ over all thermo-chemical states are adjacent.
In Fig.~\ref{F:mem} this would correspond to storing $[C_1]_1$ in index \num{0}, $[C_1]_2$ in index \num{1} and so on, with $N = N_{\text{state}}$, and $K = N_{\text{sp}}$.
In the ``C'' (C row-major) format, the storage order is transposed with $[C_1]_1$ in index \num{0}, $[C_2]_1$ in index \num{1} etc., and $N = N_{\text{sp}}$, $K = N_{\text{state}}$.
This ordering---along with the device (CPU, GPU, etc.) in question---have a large effect on the performance of SIMD\slash SIMT-vectorized algorithms.

In a \textit{shallow}-SIMD vectorization, each SIMD-lane evaluates chemical source terms for a different thermo-chemical state.
If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for e.g., states $1, 2,\ldots N_{v}$---the number of vector lanes---will load sequential locations in memory, however the $(k+1)$th species concentration will be $N_{\text{state}}$ memory locations away increasing the likelihood of cache-misses on the CPU.
In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state, thus SIMD-lanes loading species concentrations $k, k+1\ldots k + N_v$ for e.g., state 1, will access sequential memory locations if the data is stored in ``C''-order.
Further, in the ``C''-ordering the furthest difference between any two concentrations in a single thermo-chemical state is at most $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU.
However, a deep vectorization may result in SIMD-\textit{waste}---similar to thread-divergence in GPU-processing---caused by different SIMD-lanes executing different instructions (e.g. from differing for-loop bounds).

Finally, in a SIMT-vectorization, threads execute the same instruction over multiple data in lock-step without use of specialized vectorization instructions.
This is the paradigm used in GPU-processing, a more thorough review of which can be found in our previous work~\cite{CurtisGPU:2017}; one key-exception in this case is that the algorithm executed to evaluate the chemical source terms is identical for all thermo-chemical states, thus there is no thread-divergence for a shallow SIMT-vectorization. 
In Sec.~\ref{S:results} the effect of the data-ordering on the shallow-SIMD\slash SIMT vectorization performance will be studied.

\subsection{Previous works and goals of this study}
Several recent works, e.g.~\cite{CurtisGPU:2017}, have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
\cite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
The shallow-SIMD and SIMT-vectorization models were implemented in OpenCL and compared to an OpenMP baseline code that was deep SIMD-vectorized by simple compiler hints (a.k.a pragmas).
The shallow-SIMD vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance was \SIrange{1.4}{1.6}{$\times$} slower than the OpenMP baseline due to thread-divergence concerns.
\cite{kroshko2013efficient} implemented a shallow-SIMD vectorized third order stiff Rosenbrock integrator for atmospheric chemistry on a Cell Broadband Engine---a specially designed vector processor---finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code.

This work will study the performance of chemical source term evaluation of automatically generated shallow-SIMD vectorized codes for a wide range chemical kinetic models on the CPU.
The performance will be compared to a baseline shallow-SIMT vectorized code to determine the effective SIMD speedup.
Finally future extensions to this work will be detailed.

\section{Methodology}
\subsection{Chemical Kinetics Equations}
The current code evaluates the following chemical source terms:

\begin{equation}
\dot{\omega}_k = \sum_{i=1}^{N_{\text{reac}}} \nu_{ki} R_i c_i
\end{equation}
where $\dot{\omega}_k$, $N_{\text{reac}}$, $\nu_{ki}$, ${R_i}$, $c_i$ are the time rate of change of the concentration of species $k$, the number of reactions in the model, the net stoichiometric coefficient of species $k$ in reaction $i$, the net rate of progress of reaction $i$, and the pressure modification of reaction $i$, respectively.
For further detail on each term, the reader is referred to our previous work~\cite{Niemeyer:2016aa}; the code is capable of evaluating all modern reaction rate types (e.g. pressure-log, Chebyshev, etc.).

In addition, the temperature rate of change using the constant-pressure assumption\footnote{Note: in this context, the ``constant-pressure assumption'' refers to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure reactive-flow simulation.} is evaluated as:
\begin{equation}
\frac{\partial T}{\partial t} = -\frac{\sum_{k=1}^{N_{\text{sp}}} H^{\degree}_k \dot{\omega}_k}{\sum_{k=1}^{N_{\text{sp}}} [C_k] c_{p_k}^{\degree}}
\end{equation}
where $H^{\degree}_k$ and $c_{p_k}^{\degree}$ are the molar specific enthalpy and specific heat, respectively.
Although the code is equally capable of a evaluating the temperature rate using a constant-volume assumption, we omit this here for brevity.

\subsection{Code Generation}
Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates user specified psuedo-code and data to OpenCL, allowing for unit testing and easy changes of program structure, e.g. data ordering, vectorization, threading patterns etc.

\section{Results and Discussion}
\subsection{Comparison to Cantera}
The reaction rates of progress (ROP), species and temperature rates in this study are validated by comparison with Cantera~\cite{Cantera}, however special care must be taken due floating point arithmetic issues.

For a direct comparison, a relative error norm of a quantity $X_{ij}$ over all states $j$, and reactions (or species) $i$ was computed using the $L^{\infty}$ norm:
\begin{equation}
E_{X} = \left\lVert \frac{\left\lvert X_{ij,\text{CT}} - X_{ij}\right\rvert}{\num{e-10} + \num{e-6} * \left\lvert X_{ij,\text{CT}} \right\rvert} \right\rVert_{i,j,\infty}
\label{e:rel_err}
\end{equation}
where the \text{CT} subscript indicates values from Cantera~\cite{Cantera}

However, in computing the net ROP of reaction $i$ for state $j$ from the forward and reverse ROP: $R_{ij} = R_{ij}^{\prime} - R_{ij}^{\prime\prime}$, accuracy can easily be lost as the net ROP may be---particularly near chemical equilibrium---many orders of magnitude smaller than the forward or reverse rates.
To quantify this phenomena, the error in forward ROP is first defined as:
\begin{equation}
\varepsilon^{\prime}_{ij} = \left\lvert R_{ij}^{\prime} - R_{ij,\text{CT}}^{\prime} \right\rvert
\end{equation}
while the error in reverse ROP, $\varepsilon^{\prime\prime}_{ij}$, can be defined analogously.
Finally, for the reaction $i^{*}$ and state $j^{*}$ that result in the largest error in net ROP, i.e. $E_{R}$, an estimate of the error attributable to floating point error accumulation from the forward and reverse ROPs can be obtained as:
\begin{equation}
E_{\varepsilon} = \frac{\max(\varepsilon^{\prime}_{i^{*}j^{*}}\text{, }\varepsilon^{\prime\prime}_{i^{*}j^{*}})}{\num{e-10} + \num{e-6} * \left\lvert R_{i^{*}j^{*},\text{CT}} \right\rvert}
\end{equation}
This estimate allows for direct comparison of the error in forward or reverse ROPs to the value of the net ROP itself, if they are of similar magnitude the error in net ROP will be large.

\begin{table}[htb]
\sisetup{retain-zero-exponent=true}
\centering
\begin{tabular}{@{}S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] @{}}
\toprule
\multicolumn{1}{l}{Model} & \multicolumn{1}{c}{$E_{R^{\prime}}$} & \multicolumn{1}{c}{$E_{R^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R}$} & \multicolumn{1}{c}{$E_{\varepsilon}$} & \multicolumn{1}{c}{$E_{\dot{\omega}}$} & \multicolumn{1}{c}{$E_{\frac{\partial T}{\partial t}}$} \\
\midrule
\multicolumn{1}{l}{\ce{H2}\slash~\ce{CO}\cite{Burke:2011fh}} & 1.20e-8 & 6.33e-8 & 1.32e1 & 1.32e1 & 2.37e1 & 2.12e5 \\
\multicolumn{1}{l}{GRI-Mech.~3.0\cite{smith_gri-mech_30}}  & 3.11e-8 & 5.90e-8 & 1.21e0 & 1.29e0 & 2.84e0 & 2.64e4 \\
\multicolumn{1}{l}{USC-Mech II\cite{Wang:2007}}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 2.27e0 & 7.92e0 & 3.22e03 \\
\multicolumn{1}{l}{\ce{iC5H11OH}\cite{Sarathy:2013jr}} & 6.80e-8 &  1.48e-7 & 5.26e-1 & 4.87e-1 & 2.26e00 & 6.10e+01 \\
\bottomrule
\end{tabular}
\caption{Summary of rate of progress, species and temperature rate correctness.
Error statistics are based on the infinity-norm of the relative error detailed in Eq.~\eqref{e:rel_err} for each quantity.
}
\label{T:error}
\end{table}

In Table~\ref{T:error}, we see the results of this code as compared to Cantera~\cite{Cantera} on a library of partially stirred reaction conditions (PaSR) described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
Very close agreement is observed for the forward and reverse ROP for all models, however the net ROP error norm is \numrange{7}{9} orders of magnitude larger.
We see here that $E_{\varepsilon}$ is very similar in magnitude to $E_R$ in all cases, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
The species rate error norm is similar in magnitude to that of the net ROP, however the temperature rate error again increases as the error in species rates is amplified by multiplication with thermodynamic properties.
We note that the above discussion does not imply that computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.

\subsection{Performance}
\label{S:results}
The performance studies in this work were run on four ten-core \SI{2.2}{\giga\hertz} Intel Xeon E5-4640 v2 CPUs with \SI{20}{\mega\byte} of L3 cache memory, installed on an Ace Powerworks PW8027R-TRF+ with a Supermicro X9QR7-TF+/X9QRi-F+ baseboard.
Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
All code was compiled with \texttt{gcc 4.8.5} and ``\texttt{-O3}'' optimization, running on \texttt{v16.1.1} of the Intel OpenCL runtime supporting OpenCL \texttt{v1.2}.

\begin{figure}[htb]
  \centering
  \begin{minipage}[t]{0.48\linewidth}
    \includegraphics[width=\textwidth]{SIMD_SIMT_Order_comparison}
    \caption{Runtime in milliseconds per thermo-chemical state of shallow-SIMD and SIMT-vectorization on a single CPU core for both ``C'' (filled symbols) and ``F'' (empty symbols) data ordering.}
    \label{F:SIMDComp}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.48\linewidth}
    \includegraphics[width=\textwidth]{SIMD_Vecwidth_comparison}
    \caption{Speedup of a ``C''-ordered shallow-SIMD vectorization on a single CPU core when varying the supplied vector-width.}
    \label{F:Veccomp}
  \end{minipage}
\end{figure}
\begin{figure}[htb]
  \centering
  \begin{minipage}[t]{0.5\linewidth}
    \includegraphics[width=\textwidth]{SIMD_scaling}
    \caption{Strong parallel scaling efficiency of a ``C''-ordered shallow-SIMD vectorization using a vector-width of 16.}
    \label{F:SIMDscale}
  \end{minipage}
\end{figure}

In Fig.~\ref{F:SIMDComp}, the performance of the shallow-SIMD and SIMT vectorization are compared on a single CPU core for both ``C'' and ``F'' memory formats.
The ``C''-ordered SIMD\slash SIMT vectorizations show a speedup over the equivalent ``F''-ordered code by \SIrange{1.46}{2.00}{$\times$} and \SIrange{1.61}{2.38}{$\times$} respectively.
This demonstrates that enhanced data-locality of the ``C''-ordering for CPU caching---discussed in Sec.~\ref{S:data}---is key to high performance on the CPU, and suggests that a deep vectorization would see even greater accelerations.
The shallow vectorization is consistently faster than the SIMT vectorization, by \SIrange{1.73}{1.90}{$\times$} and \SIrange{2.00}{2.45}{$\times$} for the ``C'' and ``F'' memory formats respectively.
This is less than the theoretical maximum speedup on this CPU, which can process \SI{256}{\bit} AVX instructions (i.e. four double-precision operations) concurrently.
Most likely this is due to the non-sequential memory access patterns of the shallow vectorization discussed previously, but could also be caused in part by inefficiencies in translation of the OpenCL code to AVX code, a process handled by the Intel Opencl Runtime.

The effect of changing the vector-width supplied to a ``C''-ordered shallow-SIMD vectorization on a single CPU core is investigated in Fig.~\ref{F:Veccomp}.
Significant speedups are seen with larger vector-widths for the smaller models, however as the model size increases data must be loaded from increasingly distant memory locations and the performance of all three vector-widths becomes similar.

Finally, the strong parallel scaling efficiency is defined as:
\begin{equation}
S = \frac{t_1}{N t_N}
\end{equation}
where $t_1$ is the mean wall-clock runtime of chemical source terms evaluation on one core and $t_N$ the mean runtime on $N$ cores.
Fig.~\ref{F:SIMDscale} examines the strong parallel scaling efficiency for all four models with a ``C''-ordered shallow SIMD-vectorization and vector-width of 16.
Interestingly the efficiency of the hydrogen model decreases the fastest while the larger models have significantly larger scaling efficiencies for more than 4 cores.
Additionally, it appears that for a fixed number of cores, the scaling efficiency increases with model size.
This suggests that evaluation of the chemical-source terms for a single thermo-chemical state for the smaller models may not saturate the processing power of the CPU, some efficiency may be lost due to context switching within OpenCL itself.
Additionally, OpenCL allows assigning of work to devices by shared memory resources, e.g., a L3 cache, which was not pursued in this study for simplicity; this should be investigated in future work to determine the effect on scaling efficiency.

\section{Conclusions}
In this work, automatically generated OpenCL codes for shallow SIMD-vectorized chemical source term evaluation were developed and validated against Cantera~\cite{Cantera} for a wide range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
Significant speedups of up to \SIrange{1.90}{2.45}{$\times$} over a baseline SIMT-vectorized code were observed.
Two data-ordering schemes were investigated, showing a clear performance benefit for use of the ``C'' (row-major) ordering for the shallow vectorized code.
Further, this study suggests that a deep SIMD-vectorized code---currently under development---may see even greater accelerations due to expected increases in cache-hit rates.
Larger vector-widths were found to provide accelerations for the smaller models studied, but the speedup was similar over all vector-widths tested for the largest models investigated.
Finally, the strong parallel scaling efficiency was examined for the shallow vectorized code; the larger models exhibited higher scaling efficiency, possibly due to failure of the \ce{H2}\slash\ce{CO} model to saturate the CPU throughput.
Future extensions of this work will include development of a SIMD\slash SIMT-accelerated sparse-analytical Jacobian code targeted at CPUs, GPUs, and MICs.

\section{Acknowledgements}
This research was funded by the National Science Foundation under grant ACI-1534688.

\bibliography{paper}

\end{document}

### 2018-01-26 paper.tex
many updates, rewrote most of the data layout / vectorization sections

@@
 
-\title{SIMD-vectorized Chemical Source Term Evaluation}
+\newcommand{\pluseq}{\mathrel{{+}{=}}}
+\newcommand{\minuseq}{\mathrel{{-}{=}}}
+\newcommand{\ns}{N_{sp}}
+\newcommand{\nr}{N_{reac}}
+\newcommand{\conp}{CONP}
+\newcommand{\conv}{CONV}
+\newcommand{\dconp}{\ifmmode\text{for \conp,}\else for \conp,\fi}
+\newcommand{\dconv}{\ifmmode\text{for \conv,}\else for \conv.\fi}
+\newcommand{\Ru}{\mathcal{R}}
+
+\title{SIMD\slash SIMT-vectorized Thermo-chemical Source Term Evaluation}
 
@@
 \begin{abstract} % not to exceed 200 words
-A code generation platform for single-instruction, multiple-data (SIMD) vectorized chemical source term evaluation was developed and validated against Cantera for a wide range of chemical kinetic models.
+A code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized thermo-chemical source term evaluation was developed and validated against Cantera for a wide range of chemical kinetic models.
 Speedups of up to \SIrange{1.90}{2.45}{$\times$} were observed over a baseline single-instruction, multiple-thread (SIMT) parallelized code.
@@
 in the form of graphics processing units (GPUs)---have gained recognition due to their increased floating operation throughput.
-The parallel programming standard OpenCL~\cite{stone2010opencl} has further enabled adoption of vector-processing based codes in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems (e.g.
-CPU, GPU, or Intel's Many Integrated Core architecture).
+The parallel programming standard OpenCL~\cite{stone2010opencl} has further enabled adoption of vector-processing based codes in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems, e.g., the CPU, GPU, or Intel's Many Integrated Core architecture (MIC), etc.
+In this discussion, we will largely use OpenCL terminology to describe these processing paradigms, as it provides a convient way to classify otherwise disparate processor type (e.g., CPUs and GPUs).
 
@@
   \centering
-  \begin{minipage}[t]{0.45\linewidth}
+  \begin{subfigure}[htp]{0.45\linewidth}
       \includegraphics[width=\textwidth]{SIMD.pdf}
       \caption{Schematic of SIMD processing.
-A single processing element (e.g.
-a CPU core) contains a vector unit with $N_v$ lanes ($\text{L}_0,\ldots \text{L}_{N_v}$).
+A single compute unit (e.g., a CPU core) contains a vector unit with $N_v$ processing elements, together called a vector-lane.
 The vector unit executes a single instruction concurrently on multiple data.}
       \label{F:SIMD}
-  \end{minipage}
+  \end{subfigure}
   \hfill
-  \begin{minipage}[t]{0.45\linewidth}
-      \includegraphics[width=\textwidth]{mem_access.pdf}
-      \caption{Representation of an 2-D array with dimension $N \times K$ in CPU memory, with linear index marked.}
-      \label{F:mem}
-  \end{minipage}
+  \begin{subfigure}[htp]{0.45\linewidth}
+      \label{F:SIMT}
+      \includegraphics[width=\textwidth]{SIMT.pdf}
+      \caption{Schematic of SIMT processing.
+A single compute unit (e.g., a GPU streaming multiprocessor) contains many processing elements and hosts many threads, each with an instruction to execute (I1, I2).
+Threads with the same instruction execute concurrently on multiple data while the others must wait (leading to thread divergence).}
+  \end{subfigure}
 \end{figure}
 
-A SIMD instruction utilizes specialized vector processing hardware to execute the same floating point operation (e.g.
-multiplication, division, etc.) on a several different data concurrently (Fig.~\ref{F:SIMD}); the number of concurrent operations is known as the vector-width\footnote{OpenCL allows for use of vector-widths different from the actual hardware vector-width via implicit conversion, and may provide some performance benefit as will be studied in Sec.~\ref{S:results}}.
-Nearly all modern CPUs have vector processing units capable of running SIMD instructions, typically capable of handling 2--4 concurrent double precision operations.
+A typical modern CPU has many compute units (i.e., cores), each with specialized vector processing units capable of running SIMD instructions (\cref{F:SIMD}).
+A SIMD instruction utilizes the vector processor to execute the same floating point operation (e.g.
+multiplication, division, etc.) on a several different data concurrently; the number of concurrent operations possible is known as the vector-width\footnote{OpenCL allows for use of vector-widths different from the actual hardware vector-width via implicit conversion, and may provide some performance benefit as will be studied in Sec.~\ref{S:results}} and is typically around 2--4 double precision operations.
 Additionally, specialized hardware accelerators, e.g.
-Intel's Xeon Phi co-processor or multi integrated core architecture (MIC), have been developed that have tens of cores with very wide vector-widths (e.g.
-4--8 double precision operations); these very wide vector-widths are also expected on forthcoming Intel CPUs (the Skylake Xeon and Cannon Lake architectures).
+Intel's Xeon Phi co-processor---also known as the multi integrated core architecture (MIC)---have been developed that have tens of cores with very wide vector-widths (e.g.
+4--8 double precision operations); these very wide vector-widths are also expected on cutting-edge and forthcoming Intel CPUs (the Skylake Xeon and Cannon Lake architectures).
 
-\subsection{Data ordering and vectorization patterns}
-\label{S:data}
-When storing arrays for a chemical kinetic model, e.g.
-the concentration of species $k$ for the $j$-th of thermo-chemical state $[C_k]_j$---with $1 \le k \le N_{\text{sp}}$ the number of species in the model, and $1 \le j \le N_{\text{state}}$ the number of thermo-chemical states considered---there are two data-ordering choices.
-In the ``F'' (Fortran column-major) format, the concentrations of a single species $k$ over all thermo-chemical states are adjacent.
-In Fig.~\ref{F:mem} this would correspond to storing $[C_1]_1$ in index \num{0}, $[C_1]_2$ in index \num{1} and so on, with $N = N_{\text{state}}$, and $K = N_{\text{sp}}$.
-In the ``C'' (C row-major) format, the storage order is transposed with $[C_1]_1$ in index \num{0}, $[C_2]_1$ in index \num{1} etc., and $N = N_{\text{sp}}$, $K = N_{\text{state}}$.
-This ordering---along with the device (CPU, GPU, etc.) in question---have a large effect on the performance of SIMD\slash SIMT-vectorized algorithms.
-
-In a \textit{shallow}-SIMD vectorization, each SIMD-lane evaluates chemical source terms for a different thermo-chemical state.
-If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for e.g., states $1, 2,\ldots N_{v}$---the number of vector lanes---will load sequential locations in memory, however the $(k+1)$th species concentration will be $N_{\text{state}}$ memory locations away increasing the likelihood of cache-misses on the CPU.
-In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state, thus SIMD-lanes loading species concentrations $k, k+1\ldots k + N_v$ for e.g., state 1, will access sequential memory locations if the data is stored in ``C''-order.
-Further, in the ``C''-ordering the furthest difference between any two concentrations in a single thermo-chemical state is at most $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU.
-However, a deep vectorization may result in SIMD-\textit{waste}---similar to thread-divergence in GPU-processing---caused by different SIMD-lanes executing different instructions (e.g.
-from differing for-loop bounds).
-
-Finally, in a SIMT-vectorization, threads execute the same instruction over multiple data in lock-step without use of specialized vectorization instructions.
-This is the paradigm used in GPU-processing, a more thorough review of which can be found in our previous work~\cite{CurtisGPU:2017}; one key-exception in this case is that the algorithm executed to evaluate the chemical source terms is identical for all thermo-chemical states, thus there is no thread-divergence for a shallow SIMT-vectorization.
-In Sec.~\ref{S:results} the effect of the data-ordering on the shallow-SIMD\slash SIMT vectorization performance will be studied.
+SIMT processing---the foundation of modern GPUs~\cite{lindholm2008nvidia}---is a related computing paradigm that hosts large groups of threads on a single compute element (a streaming multiprocessor in NVIDIA terminology).
+A group of threads---typically \num{32} threads, known as a warp on NVIDIA GPUs----execute the same SIMD instruction on multiple data concurrently (\cref{F:SIMT}).
+If some threads must execute a different instruction---e.g., due to if\slash then branching or predication---they are forced to wait and execute later.
+This phenomena, known as thread-divergence, is a key consideration for SIMT-processing and can cause serious performance degradation for complicated algorithms~\cite{CurtisGPU:2017}.
 
@@
 Several recent works, e.g.~\cite{CurtisGPU:2017}, have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
-\cite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
+Stone et al., \cite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
 The shallow-SIMD and SIMT-vectorization models were implemented in OpenCL and compared to an OpenMP baseline code that was deep SIMD-vectorized by simple compiler hints (a.k.a pragmas).
@@
 
-This work will study the performance of chemical source term evaluation of automatically generated shallow-SIMD vectorized codes for a wide range chemical kinetic models on the CPU.
+This work will study the performance of thermo-chemical source term evaluation of automatically generated shallow-SIMD vectorized codes for a wide range chemical kinetic models on the CPU.
 The performance will be compared to a baseline shallow-SIMT vectorized code to determine the effective SIMD speedup.
@@
 \section{Methodology}
+\subsection{Data ordering and vectorization patterns}
+\label{S:data}
+
+\begin{figure}[htb]
+  \centering
+  \begin{minipage}{0.45\linewidth}
+    \begin{subfigure}[t]{\textwidth}
+      \includegraphics[width=\textwidth]{data_layouts.pdf}
+      \caption{A simple data array with K rows and N columns.}
+      \label{F:mem}
+    \end{subfigure}
+  \end{minipage}
+  \hfil
+  \begin{minipage}{0.45\linewidth}
+    \begin{subfigure}[t]{\textwidth}
+	\includegraphics[width=\textwidth]{row_major.pdf}
+	\caption{Row-major data ordering}
+	\label{F:row_major}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[t]{\textwidth}
+	\includegraphics[width=\textwidth]{column_major.pdf}
+	\caption{Column-major data ordering}
+	\label{F:column_major}
+    \end{subfigure}
+  \end{minipage}
+\end{figure}
+
+When storing arrays for a chemical kinetic model, the data-storage layout and vectorization patterns are critical to achieving high-performance code.
+\Cref{F:mem} depicts an example data array with K-rows and N-columns and index (i, j) corresponding to the $i$th row and $j$th column.
+For example the concentration of species $j$ for the $i$-th thermo-chemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{state}$, the number of thermo-chemical states considered, and $1 \le j \le \ns$, the number of species in the model.
+The stored concentrations would then have $K = N_{\text{state}}$ rows, and $N = N_{\text{sp}}$ columns.
+
+In the ``C'' (C row-major) format, the concentrations of all species $j$ for a single thermo-chemical condition $i$ would be stored sequentially in memory, i.e., with  $[C]_{1, 1}$ in index \num{0}, $[C]_{1, 2}$ in index \num{1} etc.
+as shown in \cref{F:row_major}.
+Conversely in the ``F'' (Fortran column-major) format, the concentrations of a single species $j$ over all thermo-chemical states are adjacent in memory, corresponding to storing $[C]_{1, 1}$ in index \num{0}, $[C]_{2, 1}$ in index \num{1} and so on, as shown in \cref{F:column_major}.
+This ordering---along with the device (CPU, GPU, etc.) and vectorization pattern in question---have a large effect on the performance of SIMD\slash SIMT-vectorized algorithms.
+
+In a \textit{shallow}-SIMD\slash SIMT vectorization, each SIMD-lane (or SIMT-thread) in a compute-unit evaluates thermo-chemical source terms for a different thermo-chemical state.
+If the data is stored in ``F''-order, the SIMD-lanes\slash SIMT-thread accessing $[C]_{1, j}\ldots[C]_{L, j})$, the concentration of species $j$ for states $1, 2,\ldots L$---the size of the vector lane for SIMD or the number of threads in a SIMT warp---will load sequential locations in memory, however the first $(j+1)$th species concentration, $[C]_{1, j+1}$ will be $N_{\text{state}}$ memory locations away; this increases the likelihood of cache-misses on the CPU~\cite{gray2000rules}, but conversely is well suited to coalesced memory accesses on the GPU~\cite{NVIDIA:2018}.
+In a \textit{deep}-SIMD\slash SIMT vectorization, a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermo-chemical source terms for a single thermo-chemical state, thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + L}$---the species concentrations for state 1, species $j \ldots j + L$---will access sequential memory locations if the data is stored in ``C''-order.
+Further, in the ``C''-ordering the furthest difference between any two species concentrations within the same thermo-chemical state is at most $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU, but leads to uncoalesced memory-accesses on the GPU.
+However, a deep vectorization requires both synchronization between SIMD-lanes\slash SIMT-threads, and may result in SIMD-waste\slash SIMT thread-divergence, caused by different SIMD-lanes\slash SIMT-threads executing different instructions (e.g.
+resulting from different if\slash then branches)
+
+\begin{figure}[htb]
+  \centering
+  \label{F:vector_data}
+  \begin{minipage}{\linewidth}
+    \begin{subfigure}[t]{\textwidth}
+	\includegraphics[width=\textwidth]{row_major_split.pdf}
+	\caption{Row-major data ordering}
+	\label{F:row_major_split}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[t]{\textwidth}
+	\includegraphics[width=\textwidth]{column_major_split.pdf}
+	\caption{Column-major data ordering}
+	\label{F:column_major_split}
+    \end{subfigure}
+  \end{minipage}
+\end{figure}
+
+Finally,~\cref{F:vector_data} shows a vectorized data-ordering that negates the disadvantages of a shallow, ``C''-ordered SIMD vectorization on the CPU (\cref{F:row_major_split}) and a deep, ``F''-ordered SIMT vectorization on the GPU (\cref{F:column_major_split}).
+In Sec.~\ref{S:results} the effect of the data-ordering on the shallow-SIMD\slash SIMT vectorization performance will be studied.
+
 \subsection{Chemical Kinetics Equations}
-The current code evaluates the following chemical source terms:
+The current code evaluates the following thermo-chemical source terms:
 
@@
 \end{equation}
-where $t_1$ is the mean wall-clock runtime of chemical source terms evaluation on one core and $t_N$ the mean runtime on $N$ cores.
+where $t_1$ is the mean wall-clock runtime of thermo-chemical source terms evaluation on one core and $t_N$ the mean runtime on $N$ cores.
 Fig.~\ref{F:SIMDscale} examines the strong parallel scaling efficiency for all four models with a ``C''-ordered shallow SIMD-vectorization and vector-width of 16.
@@
 \section{Conclusions}
-In this work, automatically generated OpenCL codes for shallow SIMD-vectorized chemical source term evaluation were developed and validated against Cantera~\cite{Cantera} for a wide range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
+In this work, automatically generated OpenCL codes for shallow SIMD-vectorized thermo-chemical source term evaluation were developed and validated against Cantera~\cite{Cantera} for a wide range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
 Significant speedups of up to \SIrange{1.90}{2.45}{$\times$} over a baseline SIMT-vectorized code were observed.

### 2018-01-26 derivations.tex
switch to elsarticle / natbib for derivs
start describing system in paper

@@
-\documentclass[12pt]{article}
+\documentclass[12pt,number,sort&compress]{elsarticle}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{lmodern}
-\usepackage[
-    backend=biber,
-    style=numeric,
-    citestyle=numeric,
-    sorting=none
-]{biblatex}
-\addbibresource{derivations.bib}
 
@@
 \usepackage[english,capitalise]{cleveref}
+\bibliographystyle{elsarticle-num}
 
@@
 \end{equation}
-where $\dot{\omega}_k$ is the $kth$ species overall production rate, and the temperature production rate~\cite{TurnsStephenR2012Aitc} is:
+where $\dot{\omega}_k$ is the $kth$ species' overall production rate, and the temperature production rate~\cite{TurnsStephenR2012Aitc} is:
 \begin{subequations}
@@
 \frac{\text{d} V }{\text{d} t } &= V \left(\frac{T \Ru}{P} \sum_{k=1}^{\ns  - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right)& \text{\dconp} \\
-\frac{\text{d} P }{\text{d} t } &= \frac{P}{T} \frac{\text{d} T }{\text{d} t } + T \Ru \sum_{k=1}^{\ns  - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k}&\text{\dconv}
+\frac{\text{d} P }{\text{d} t } &= T \Ru \sum_{k=1}^{\ns  - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{P}{T} \frac{\text{d} T }{\text{d} t }&\text{\dconv}
 \end{align}
@@
 
-\printbibliography
+\bibliography{derivations}
 

### 2018-01-26 paper.tex
start describing system in paper

@@
   \centering
-  \label{F:vector_data}
   \begin{minipage}{\linewidth}
@@
 	\includegraphics[width=\textwidth]{row_major_split.pdf}
-	\caption{Row-major data ordering}
+	\caption{Row-major, wide-vectorized data ordering}
 	\label{F:row_major_split}
@@
 	\includegraphics[width=\textwidth]{column_major_split.pdf}
-	\caption{Column-major data ordering}
+	\caption{Column-major, deep-vectorized data ordering}
 	\label{F:column_major_split}
@@
   \end{minipage}
+  \caption{Vectorized data ordering patterns}
+  \label{F:vector_data}
 \end{figure}
 
-Finally,~\cref{F:vector_data} shows a vectorized data-ordering that negates the disadvantages of a shallow, ``C''-ordered SIMD vectorization on the CPU (\cref{F:row_major_split}) and a deep, ``F''-ordered SIMT vectorization on the GPU (\cref{F:column_major_split}).
-In Sec.~\ref{S:results} the effect of the data-ordering on the shallow-SIMD\slash SIMT vectorization performance will be studied.
+Finally,~\cref{F:vector_data} shows a vectorized data-ordering that improves the caching patters of a shallow, ``C''-ordered SIMD vectorization on the CPU (\cref{F:row_major_split}) and a deep, ``F''-ordered SIMT vectorization on the GPU (\cref{F:column_major_split}).
+This is accomplished by splitting the slower-varying axis of the data array---rows for ``C''-ordering, and columns for ``F''-ordering---into chunks of size $L$ (the SIMD vector-width, or SIMT warp-size) and laying these data out contigously in memory.
+For example, using the data-ordering in~\cref{F:row_major_split}, the concentration of species $j$ for states $i\ldots i+L$, $[C]_{i, j} \ldots [C]_{i + L, j}$, are contiguous in memory and are followed by the concentration of species $j + 1$ for the same states, $[C]_{i, j + 1} \ldots [C]_{i + L, j + 1}$.
+This pattern, similar to OpenCL's native vector data-types---e.g., \texttt{double\num{8}} which treats \num{8} contiguous double precision floating point numbers as a single vector datum---ensures that any SIMD operation occurs on data that is contiguous in memory, greatly improving caching and SIMD-throughput.
+Conversely, the data-ordering in~\cref{F:column_major_split} is designed to enable coalesced memory accesses for ``F''-ordered, deep-SIMT vectorization on the GPU.
+The effects of these various data-ordering and vectorization patterns on performance will be studied in~\cref{S:results}.
 
-\subsection{Chemical Kinetics Equations}
-The current code evaluates the following thermo-chemical source terms:
+\subsection{Thermo-chemical Source Terms and Jacobian}
+This new version of \texttt{pyJac} is capable of evaluating the thermo-chemical source terms for using both the constant-pressure (\conp) or constant-volume (\conv) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
+In this section, we will only outline a brief summary of the system evaluated by \texttt{pyJac}; the interested reader is referred to our complete derivations, attached as supplimental material to this document.
 
+The thermo-chemical state vector derivation consists of the temperature, the non-constant thermodynamic state parameter (volume or pressure) and the number of moles of all species except the last in the chemical model which is typically taken to be the bath-gas (e.g., \ce{N2}):
+\begin{subequations}
+\begin{align}
+\Phi &= \left\{T, V, n_1, n_2 \ldots n_{\ns - 1}\right\}&\text{\dconp}\\
+\Phi &= \left\{T, P, n_1, n_2 \ldots n_{\ns - 1}\right\}&\text{\dconv}
+\end{align}
+\end{subequations}
+where $T$ is the temperature, $V$ and $P$ the volume and pressure respectively, and $n_j$ the number of moles of the $j$th species in the model (containing $\ns$ total species).
+
+For both \conp~and \conv, the molar source terms are~\cite{TurnsStephenR2012Aitc}:
 \begin{equation}
-\dot{\omega}_k = \sum_{i=1}^{N_{\text{reac}}} \nu_{ki} R_i c_i
+\frac{\text{d} n }{\text{d} t }_{k} = V \dot{\omega}_{k} \qquad k = 1, \ldots, \ns - 1, \label{e:spec}
 \end{equation}
-where $\dot{\omega}_k$, $N_{\text{reac}}$, $\nu_{ki}$, ${R_i}$, $c_i$ are the time rate of change of the concentration of species $k$, the number of reactions in the model, the net stoichiometric coefficient of species $k$ in reaction $i$, the net rate of progress of reaction $i$, and the pressure modification of reaction $i$, respectively.
-For further detail on each term, the reader is referred to our previous work~\cite{Niemeyer:2016aa}; the code is capable of evaluating all modern reaction rate types (e.g.
-pressure-log, Chebyshev, etc.).
-
-In addition, the temperature rate of change using the constant-pressure assumption\footnote{Note: in this context, the ``constant-pressure assumption'' refers to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure reactive-flow simulation.} is evaluated as:
+where $\dot{\omega}_k$ is the $kth$ species' overall production rate, and the temperature production rate~\cite{TurnsStephenR2012Aitc} is:
+\begin{subequations}
+\label{e:temperature_incomplete}
+\begin{align}
+\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} H_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{p, k}}} & \text{\dconp} \\
+\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} U_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{v, k}}} & \text{\dconv}
+\end{align}
+\end{subequations}
+where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure and constant-volume specific heat of species $k$ in molar units respectively, while $[C]_{k}$ is the concentration, given by:
 \begin{equation}
-\frac{\partial T}{\partial t} = -\frac{\sum_{k=1}^{N_{\text{sp}}} H^{\degree}_k \dot{\omega}_k}{\sum_{k=1}^{N_{\text{sp}}} [C_k] c_{p_k}^{\degree}}
+ [C]_{k} = \frac{n_{k}}{V}
 \end{equation}
-where $H^{\degree}_k$ and $c_{p_k}^{\degree}$ are the molar specific enthalpy and specific heat, respectively.
-Although the code is equally capable of a evaluating the temperature rate using a constant-volume assumption, we omit this here for brevity.
+By differentiating the ideal gas law, the volume and pressure source terms can be found:
+\begin{subequations}
+\label{e:param_complete}
+\begin{align}
+\frac{\text{d} V }{\text{d} t } &= V \left(\frac{T \Ru}{P} \sum_{k=1}^{\ns  - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right)& \text{\dconp} \\
+\frac{\text{d} P }{\text{d} t } &= T \Ru \sum_{k=1}^{\ns  - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{P}{T} \frac{\text{d} T }{\text{d} t }&\text{\dconv}
+\end{align}
+\end{subequations}
+where $\mathcal{R}$ is the ideal gas-constant.
 

### 2018-01-30 derivations.tex
move towards a unified equation dictionary

@@
 \usepackage{lmodern}
+
+% shared equations in separate file
+% https://yatb.giacomodrago.com/en/post/3/latex-loading-equations-from-an-external-file.html
+\usepackage{catchfilebetweentags}
+
+\newcommand{\loadeq}[1]{%
+   \ExecuteMetaData[eqn_dict.tex]{eq#1}%
+}
 
@@
 The state vector for this derivation consists of the temperature, a thermodynamic state parameter (pressure or volume) and the number of moles of all species except the last in the model:
-\begin{subequations}
-\begin{align}
-\Phi &= \left\{T, V, n_1, n_2 \ldots n_{\ns - 1}\right\}&\text{\dconp}\\
-\Phi &= \left\{T, P, n_1, n_2 \ldots n_{\ns - 1}\right\}&\text{\dconv}
-\end{align}
-\end{subequations}
+\loadeq{state}
 where $T$ is the temperature, $V$ and $P$ the volume and pressure respectively, and $n_j$ the number of moles of the $j$th species in the model (containing $\ns$ total species).
@@
 where $i$ and $j$ correspond to the row and column, respectively, of the entry in the Jacobian matrix.
-The Jacobian entries resulting from~\cref{e:jac_general} are derived in~\cref{s:energy_derivs,s:molar_source_derivatives,s:parameter_source_derivatives}, while various subcomponents of the Jacobian are derived in~\cref{s:dri_dt,s:dri_dnj,s:dri_de,s:dri_pdep,s:dpmod_fall}..
+The Jacobian entries resulting from~\cref{e:jac_general} are derived in~\cref{s:energy_derivs,s:molar_source_derivatives,s:parameter_source_derivatives}, while various subcomponents of the Jacobian are derived in~\cref{s:dri_dt,s:dri_dnj,s:dri_de,s:dri_pdep,s:dpmod_fall}.
+The final form of the Jacobian can be found in~\cref{s:jac_final}, which provides a useful summary for this document.
 
@@
 and finally, we may substitute in the specific heat for the temperature derivative of enthalpy, giving:
-\begin{subequations}
-\begin{align}
- \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{k} + \frac{{C_p}_{\ns}}{T}\right) [C]_{k} + & \nonumber \\
-				     & \sum_{k=1}^{-1 + \ns} \biggl( \left(- H_{k} + \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial T }_{k} + & \nonumber \\
-				     & \mathindent[larger] \left(- {C_p}_{k} + \frac{W_{k} {C_p}_{k} }{W_{\ns}} \right) \dot{\omega}_{k}\biggr)\Biggr] & \text{\dconp} \\
-\intertext{and similarly:}
-\frac{\partial\dot{T}}{\partial{T}}  &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns} \left(- \frac{\text{d} {C_v} }{\text{d} T }_{k} + \frac{{C_v}_{\ns}}{T}\right) [C]_{k} + & \nonumber \\
-				     & \sum_{k=1}^{-1 + \ns} \biggl(\left(- U_{k} + \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial T }_{k} + & \nonumber \\
-				     & \mathindent[larger] \left(- {C_v}_{k} + \frac{W_{k} {C_v}_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}\biggr)\Biggr] & \text{\dconv}
-\end{align}
-\end{subequations}
+\loadeq{dTdotdT}
 The species production rate derivative with respect to temperature will be defined in~\cref{s:d_spec_prod}.
@@
 
+\subsection{Final jacobian form}
+
 \bibliography{derivations}

### 2018-01-30 paper.tex
edits

@@
   \begin{subfigure}[htp]{0.45\linewidth}
-      \label{F:SIMT}
       \includegraphics[width=\textwidth]{SIMT.pdf}
@@
 Threads with the same instruction execute concurrently on multiple data while the others must wait (leading to thread divergence).}
+      \label{F:SIMT}
   \end{subfigure}
@@
 Additionally, specialized hardware accelerators, e.g.
-Intel's Xeon Phi co-processor---also known as the multi integrated core architecture (MIC)---have been developed that have tens of cores with very wide vector-widths (e.g.
+Intel's Xeon Phi co-processor (MIC) have been developed that have tens of cores with very wide vector-widths (e.g.
 4--8 double precision operations); these very wide vector-widths are also expected on cutting-edge and forthcoming Intel CPUs (the Skylake Xeon and Cannon Lake architectures).
@@
 SIMT processing---the foundation of modern GPUs~\cite{lindholm2008nvidia}---is a related computing paradigm that hosts large groups of threads on a single compute element (a streaming multiprocessor in NVIDIA terminology).
-A group of threads---typically \num{32} threads, known as a warp on NVIDIA GPUs----execute the same SIMD instruction on multiple data concurrently (\cref{F:SIMT}).
+A group of threads---typically \num{32} threads, known as a warp on NVIDIA GPUs---execute the same SIMD instruction on multiple data concurrently (\cref{F:SIMT}).
 If some threads must execute a different instruction---e.g., due to if\slash then branching or predication---they are forced to wait and execute later.
@@
       \includegraphics[width=\textwidth]{data_layouts.pdf}
-      \caption{A simple data array with K rows and N columns.}
+      \caption{A simple 2-D data array with K rows and N columns.}
       \label{F:mem}
@@
 
-In the ``C'' (C row-major) format, the concentrations of all species $j$ for a single thermo-chemical condition $i$ would be stored sequentially in memory, i.e., with  $[C]_{1, 1}$ in index \num{0}, $[C]_{1, 2}$ in index \num{1} etc.
+In the ``C'' (C row-major) format, the concentrations of all species for a single thermo-chemical condition $i$ would be stored sequentially in memory, i.e., with  $[C]_{1, 1}$ in index \num{1}, $[C]_{1, 2}$ in index \num{2} etc.
 as shown in \cref{F:row_major}.
-Conversely in the ``F'' (Fortran column-major) format, the concentrations of a single species $j$ over all thermo-chemical states are adjacent in memory, corresponding to storing $[C]_{1, 1}$ in index \num{0}, $[C]_{2, 1}$ in index \num{1} and so on, as shown in \cref{F:column_major}.
+Conversely in the ``F'' (Fortran column-major) format, the concentrations of a single species $j$ over all thermo-chemical states are adjacent in memory, corresponding to storing $[C]_{1, 1}$ in index \num{1}, $[C]_{2, 1}$ in index \num{2} and so on, as shown in \cref{F:column_major}.
 This ordering---along with the device (CPU, GPU, etc.) and vectorization pattern in question---have a large effect on the performance of SIMD\slash SIMT-vectorized algorithms.
@@
   \centering
-  \begin{minipage}{\linewidth}
+  \begin{minipage}{0.6\linewidth}
     \begin{subfigure}[t]{\textwidth}
@@
 
-Finally,~\cref{F:vector_data} shows a vectorized data-ordering that improves the caching patters of a shallow, ``C''-ordered SIMD vectorization on the CPU (\cref{F:row_major_split}) and a deep, ``F''-ordered SIMT vectorization on the GPU (\cref{F:column_major_split}).
+Finally,~\cref{F:vector_data} shows a vectorized data-ordering that improves the caching patterns of a shallow, ``C''-ordered SIMD vectorization on the CPU (\cref{F:row_major_split}) and a deep, ``F''-ordered SIMT vectorization on the GPU (\cref{F:column_major_split}).
 This is accomplished by splitting the slower-varying axis of the data array---rows for ``C''-ordering, and columns for ``F''-ordering---into chunks of size $L$ (the SIMD vector-width, or SIMT warp-size) and laying these data out contigously in memory.

### 2018-01-30 derivations.tex
move more eqns over to the dict

@@
 The evolution of the thermo-chemical state variables of this system is described by a set of ordinary-differential equations:
-\begin{subequations}
-\label{e:source_terms}
-\begin{align}
-f &= \frac{\text{d} \Phi }{\text{d} t } = \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} V }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\} & \text{\dconp} \\
-f &= \frac{\text{d} \Phi }{\text{d} t } = \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} P }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\} & \text{\dconv}
-\end{align}
-\end{subequations}
+\loadeq{dstate}
+
 For both, the molar source terms are~\cite{TurnsStephenR2012Aitc}:
-\begin{equation}
-\frac{\text{d} n }{\text{d} t }_{k} = V \dot{\omega}_{k} \qquad k = 1, \ldots, \ns - 1, \label{e:spec}
-\end{equation}
+\loadeq{dmolar}
 where $\dot{\omega}_k$ is the $kth$ species' overall production rate, and the temperature production rate~\cite{TurnsStephenR2012Aitc} is:
-\begin{subequations}
-\label{e:temperature_incomplete}
-\begin{align}
-\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} H_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{p, k}}} & \text{\dconp} \\
-\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} U_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{v, k}}} & \text{\dconv}
-\end{align}
-\end{subequations}
+\loadeq{dtemp}
 where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure and constant-volume specific heat of species $k$ in molar units respectively, while $[C]_{k}$ is the concentration, given by:
@@
 Combining~\cref{e:spec,e:param_incomplete,e:total_molar} gives the final form of the pressure and volume terms:
-\begin{subequations}
-\label{e:param_complete}
-\begin{align}
-\frac{\text{d} V }{\text{d} t } &= V \left(\frac{T \Ru}{P} \sum_{k=1}^{\ns  - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right)& \text{\dconp} \\
-\frac{\text{d} P }{\text{d} t } &= T \Ru \sum_{k=1}^{\ns  - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{P}{T} \frac{\text{d} T }{\text{d} t }&\text{\dconv}
-\end{align}
-\end{subequations}
+\loadeq{dparam}
 
@@
 \subsection{Final jacobian form}
+\label{s:jac_final}
 

### 2018-01-30 paper.tex
move more eqns over to the dict

@@
 \usepackage{subcaption} %subfigs
+
+% shared equations in separate file
+% https://yatb.giacomodrago.com/en/post/3/latex-loading-equations-from-an-external-file.html
+\usepackage{catchfilebetweentags}
+
+\newcommand{\loadeq}[1]{%
+   \ExecuteMetaData[eqn_dict.tex]{eq#1}%
+}
+
 \sisetup{group-separator={,},
@@
 The thermo-chemical state vector derivation consists of the temperature, the non-constant thermodynamic state parameter (volume or pressure) and the number of moles of all species except the last in the chemical model which is typically taken to be the bath-gas (e.g., \ce{N2}):
-\begin{subequations}
-\begin{align}
-\Phi &= \left\{T, V, n_1, n_2 \ldots n_{\ns - 1}\right\}&\text{\dconp}\\
-\Phi &= \left\{T, P, n_1, n_2 \ldots n_{\ns - 1}\right\}&\text{\dconv}
-\end{align}
-\end{subequations}
+\loadeq{state}
 where $T$ is the temperature, $V$ and $P$ the volume and pressure respectively, and $n_j$ the number of moles of the $j$th species in the model (containing $\ns$ total species).
@@
 For both \conp~and \conv, the molar source terms are~\cite{TurnsStephenR2012Aitc}:
-\begin{equation}
-\frac{\text{d} n }{\text{d} t }_{k} = V \dot{\omega}_{k} \qquad k = 1, \ldots, \ns - 1, \label{e:spec}
-\end{equation}
+\loadeq{dmolar}
 where $\dot{\omega}_k$ is the $kth$ species' overall production rate, and the temperature production rate~\cite{TurnsStephenR2012Aitc} is:
-\begin{subequations}
-\label{e:temperature_incomplete}
-\begin{align}
-\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} H_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{p, k}}} & \text{\dconp} \\
-\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} U_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{v, k}}} & \text{\dconv}
-\end{align}
-\end{subequations}
+\loadeq{dtemp}
 where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure and constant-volume specific heat of species $k$ in molar units respectively, while $[C]_{k}$ is the concentration, given by:
@@
 By differentiating the ideal gas law, the volume and pressure source terms can be found:
-\begin{subequations}
-\label{e:param_complete}
-\begin{align}
-\frac{\text{d} V }{\text{d} t } &= V \left(\frac{T \Ru}{P} \sum_{k=1}^{\ns  - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right)& \text{\dconp} \\
-\frac{\text{d} P }{\text{d} t } &= T \Ru \sum_{k=1}^{\ns  - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \dot{\omega}_{k} + \frac{P}{T} \frac{\text{d} T }{\text{d} t }&\text{\dconv}
-\end{align}
-\end{subequations}
+\loadeq{dparam}
 where $\mathcal{R}$ is the ideal gas-constant.

### 2018-01-31 derivations.tex
updates

--- 2018-01-26 derivations.tex
@@
 \begin{subequations}
+\label{e:dTdotdT_nojac}
 \begin{align}
@@
 Following the now familiar collapsing of the specific-heat\slash concentration sum, and substituting in the energy source term gives:
-\begin{subequations}
-\begin{align}
- \frac{\partial\dot{T}}{\partial{V}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ - \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial V }_{k} + & \nonumber \\
-				     & \mathindent[largerindentand] \frac{1}{V} \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{-1 + \ns} \left(- {C_p}_{\ns} + {C_p}_{k}\right) [C]_{k}\Biggr] & \text{\dconp} \\
- \frac{\partial\dot{T}}{\partial{P}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial P }_{k} - & \nonumber \\
-				     & \mathindent[largerindentand] \frac{{C_v}_{\ns}}{T \Ru} \frac{\text{d} T }{\text{d} t } \Biggr] & \text{\dconv}
-\end{align}
-\end{subequations}
+\loadeq{dTdParam}
 Finally, the species production rate derivative with respect to the state parameter will be computed in~\cref{s:d_spec_prod}.
@@
 \begin{align}
+ \label{e:dnkdotdT_compact}
  \mathcal{J}_{k + 2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial T } \nonumber \\
@@
 Finally, the derivative of the net molar rate of production of a species $k$ with respect to the state parameter is:
+\begin{subequations}
 \begin{align}
@@
 \end{align}
+\end{subequations}
 
@@
 The derivatives of the thermodynamic state parameter source terms with respect to temperature are given by differentiating~\cref{e:param_complete}:
+\begin{subequations}
+\label{e:dEdotdT_nojac}
 \begin{align}
@@
 Finally, combining~\cref{e:dropf_dt,e:dropr_dt} gives the total temperature derivative of the net rate of progress:
-\begin{align}
- \label{e:drop_dt}
- \frac{\partial R }{\partial T }_{i} =& \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right) - \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_{r, i}} \nonumber \\
-				      &\qquad + \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_{r, i}} - \frac{[C] S^{\prime}_{\ns}}{T} {k_{f, i}}
-\end{align}
+\loadeq{dropdT}
 \subsubsection{Molar derivative}
@@
 Similar to~\cref{s:dri_dt},~\cref{e:dkf_plog_dt} may be substituted into~\cref{e:dropfdt_1,e:droprdt_1} to give:
-\begin{align}
- \frac{\partial R_{i} }{\partial T } &= \left(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \beta_1 + \frac{E_{a_1}}{T \Ru} \right) \times \nonumber \\
-				     & \mathindent \left(\frac{{R_{f, i}} - {R_{r, i}}}{T}\right) + {R_{r, i}} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_{r, i}} - S^{\prime}_{\ns} {k_{f, i}}\right)
-\end{align}
+\loadeq{dropplogdT}
 
@@
 Following the same process for Arrhenius-based reactions (see~\cref{s:dri_dt}), the derivative of the net rate of progress with respect to temperature is:
-\begin{align}
- \frac{\partial R }{\partial T }_{i} &= \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \left[ - \frac{2 \log{\left (10 \right )} \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right) \right] \left({R_{f, i}} - {R_{r, i}}\right) + \nonumber \\
-				     & \mathindent \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} {R_{r, i}} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_{r, i}} - S^{\prime}_{\ns} {k_{f, i}}\right)
-\end{align}
+\loadeq{dropchebdT}
 
@@
 
+\section{Final jacobian form}
+\label{s:jac_final}
+
+Here we give a summary of the Jacobian derivations in the previous sections, as a complete reference guide to the equations evaluated by \texttt{pyJac}.
+
+\subsection{Temperature derivatives}
+In this section, the derivatives of all source terms with respect to the temperature are considered.
+
+\subsubsection{Temperature source term derivatives}
+Starting from~\cref{e:dTdotdT_nojac},~\cref{e:dnkdotdT_compact} is substituted in to yield:
+\loadeq{dTdotdT}
+
+\subsubsection{State parameter source term derivatives}
+Similarly,~\cref{e:dnkdotdT_compact} is substituted into~\cref{e:dEdotdT_nojac}, giving:
+\loadeq{dEdotdT}
+
+\subsubsection{Molar source term derivatives}
+From~\cref{e:dnkdotdT_compact} and~\cref{e:dwdot_dT}, a general form for the molar source term derivative with respect to temperature is obtained:
+\loadeq{dnkdotdT}
+As \texttt{pyJac} evaluates these derivative entries by looping over the reactions $i$ in the model, we convert~\cref{e:dnkdotdT} to an update form:
+\begin{align}
+ \lhs{dnkdotdT}{\mathcal{J}_{k + 2, 1}} &\pluseq  \rhs{dnkdotdT}{V \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial T }_{i} + \nu_{k,i} \frac{\partial R }{\partial T }_{i} c_{i}\right)} & \mathcond{dnkdotdT}{k = 1, \ldots, \ns}
+\end{align}
+the rate of progress temperature derivative for all reaction types---excluding Chebyshev and P-Log---is:
+\loadeq{dropdT}
+while for P-Log reactions:
+\loadeq{dropplogdT}
+and for Chebyshev reactions:
+\loadeq{dropchebdT}
+
 \bibliography{derivations}

### 2018-01-31 derivations.tex
reorder and cleanup

@@
 \newcommand{\dconp}{\ifmmode\text{for \conp,}\else for \conp,\fi}
-\newcommand{\dconv}{\ifmmode\text{for \conv,}\else for \conv.\fi}
+\newcommand{\dconv}{\ifmmode\text{for \conv,}\else for \conv,\fi}
 \newcommand{\Ru}{\mathcal{R}}
@@
 
-\subsubsection{Molar derivative}
-Next the molar derivatives of the energy equation:
-\begin{align}
- \mathcal{J}_{1,j} = \frac{\partial\dot{T}}{\partial{n_j}}
-\end{align}
-
-First, the derivative of a species concentration with respect to the moles of species $j$ is computed:
-\begin{align}
- \frac{\partial [C_k]}{\partial n_j} &=
- \begin{dcases}
- \frac{\delta_{j k}}{V} & k \ne \ns \\
- -\frac{1}{V} & k = \ns
- \end{dcases} \\
-\intertext{Or put succinctly:}
-\label{e:dck_dnj}
-\frac{\partial [C_k]}{\partial n_j} &= \frac{\delta_{j k} - \delta_{\ns k}}{V}
-\end{align}
-
-Similar to~\cref{s:denergy_dt} the derivative of the energy equation with respect to the moles of species $j$---after collapsing the specific-heat\slash~concentration sum---are given by:
-\begin{align}
- \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{\left(\sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}\right) \sum_{k=1}^{-1 + \ns} - \frac{\delta_{j k}}{V} \left({C_p}_{\ns} - {C_p}_{k}\right)}{\left(\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}\right)^{2}} - \nonumber \\
-				       & \mathindent \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k}
-\end{align}
-Next, the Kronecker delta summation is simplified, and the the compact form of the energy equation factored out yielding:
-\begin{subequations}
-\begin{align}
- \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} & \nonumber \\
-				       &  \mathindent[largerindentand] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_p}_{\ns} + {C_p}_{j}\right)\Biggr] & \text{\dconp} \\
-\intertext{and:}
- \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} & \nonumber \\
-				       & \mathindent[largerindentand] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_v}_{\ns} + {C_v}_{j}\right) \Biggr] & \text{\dconv}
-\end{align}
-\end{subequations}
-Again, the species production rate derivative with respect to moles will be explored in~\cref{s:d_spec_prod}.
-
 \subsubsection{State parameter derivatives}
@@
 \begin{subequations}
+\label{e:dTdotdE_initial}
 \begin{align}
@@
 while the specific heat and enthalpy\slash internal-energy are independent of the state parameter for both \conp~and \conv.
-Following the now familiar collapsing of the specific-heat\slash concentration sum, and substituting in the energy source term gives:
-\loadeq{dTdParam}
-Finally, the species production rate derivative with respect to the state parameter will be computed in~\cref{s:d_spec_prod}.
-
-\subsection{Molar source term derivatives}
-\label{s:molar_source_derivatives}
-\subsubsection{Temperature derivative}
-
-The derivative of the net molar rate of production of a species $k$ with respect to temperature is:
-\begin{align}
- \label{e:dnkdotdT_compact}
- \mathcal{J}_{k + 2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial T } \nonumber \\
- \intertext{or:}
- \mathcal{J}_{k + 2, 1} & = V \frac{\partial \dot{\omega}_{k} }{\partial T }
-\end{align}
+Following the same procedure as in~\cref{s:denergy_dt}, the result of~\cref{e:dTdotdE_initial} are simplified by collapsing the specific-heat\slash concentration sum, and substituting in the temperature source term, giving:
+\begin{subequations}
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{V}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ - \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial V }_{k} + & \nonumber \\
+				     & \mathindent[largerindentand] \frac{1}{V} \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{-1 + \ns} \left(- {C_p}_{\ns} + {C_p}_{k}\right) [C]_{k}\Biggr] & \text{\dconp} \\
+ \frac{\partial\dot{T}}{\partial{P}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial P }_{k} - & \nonumber \\
+				     & \mathindent[largerindentand] \frac{{C_v}_{\ns}}{T \Ru} \frac{\text{d} T }{\text{d} t } \Biggr] & \text{\dconv}
+\end{align}
+\end{subequations}
+The derivative of the species production rate with respect to the state parameter will be computed in~\cref{s:d_spec_prod}.
 
 \subsubsection{Molar derivative}
-
-Similarly, the derivative of the net molar rate of production of a species $k$ with respect to the moles of species $j$ is:
-\begin{align}
- \mathcal{J}_{k + 2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial n_j } \nonumber \\
-  \intertext{or:}
- \mathcal{J}_{k + 2, j + 2} & = V \frac{\partial \dot{\omega}_{k} }{\partial {n_j} }
-\end{align}
-
-\subsubsection{State parameter derivatives}
-
-Finally, the derivative of the net molar rate of production of a species $k$ with respect to the state parameter is:
-\begin{subequations}
-\begin{align}
-  \mathcal{J}_{k + 2, 2} &= \frac{\partial }{\partial V } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial V } & \nonumber \\
-			 &= V \frac{\partial \dot{\omega}_{k} }{\partial V } + \dot{\omega}_{k} & \text{\dconp}
+Finally, the molar derivative of the temperature source term is:
+\begin{align}
+ \mathcal{J}_{1,j} = \frac{\partial\dot{T}}{\partial{n_j}}
+\end{align}
+
+First, the derivative of a species concentration with respect to the moles of species $j$ is computed:
+\begin{align}
+ \frac{\partial [C_k]}{\partial n_j} &=
+ \begin{dcases}
+ \frac{\delta_{j k}}{V} & k \ne \ns \\
+ -\frac{1}{V} & k = \ns
+ \end{dcases} \\
+\intertext{Or put succinctly:}
+\label{e:dck_dnj}
+\frac{\partial [C_k]}{\partial n_j} &= \frac{\delta_{j k} - \delta_{\ns k}}{V}
+\end{align}
+
+Similar to~\cref{s:denergy_dt} the derivative of the energy equation with respect to the moles of species $j$---after collapsing the specific-heat\slash~concentration sum---are given by:
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{\left(\sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}\right) \sum_{k=1}^{-1 + \ns} - \frac{\delta_{j k}}{V} \left({C_p}_{\ns} - {C_p}_{k}\right)}{\left(\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}\right)^{2}} - \nonumber \\
+				       & \mathindent \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k}
+\end{align}
+Next, the Kronecker delta summation is simplified, and the the compact form of the energy equation factored out yielding:
+\begin{subequations}
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} & \nonumber \\
+				       &  \mathindent[largerindentand] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_p}_{\ns} + {C_p}_{j}\right)\Biggr] & \text{\dconp} \\
 \intertext{and:}
-  \mathcal{J}_{k + 2, 2} &= \frac{\partial }{\partial P } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial P } & \nonumber \\
-			 &=V \frac{\partial \dot{\omega}_{k} }{\partial P } & \text{\dconv}
-\end{align}
-\end{subequations}
+ \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} & \nonumber \\
+				       & \mathindent[largerindentand] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_v}_{\ns} + {C_v}_{j}\right) \Biggr] & \text{\dconv}
+\end{align}
+\end{subequations}
+Again, the derivative of the species production rate with respect to moles will be explored in~\cref{s:d_spec_prod}.
 
@@
 
-\subsubsection{Molar derivative}
-Similarly, the derivatives of the thermodynamic state parameter source terms with respect to the moles of species $j$ are:
-\begin{align}
- \mathcal{J}_{2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{V} }{\partial n_j} & \nonumber \\
-		    &= V \left(\frac{1}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} + \frac{1}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} }\right) & \text{\dconp}
-\intertext{and:}
- \mathcal{J}_{2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{P} }{\partial n_j} & \nonumber \\
-		    &=  \frac{P}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} } + T \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} & \text{\dconv}
-\end{align}
-
 \subsubsection{State parameter derivative}
-Finally, the derivative of the the thermodynamic state parameter source terms with respect to themselves are:
+Similarly, the derivative of the the thermodynamic state parameter source terms with respect to themselves are:
 \begin{align}
@@
 
+\subsubsection{Molar derivative}
+Finally, the derivatives of the thermodynamic state parameter source terms with respect to the moles of species $j$ are:
+\begin{align}
+ \mathcal{J}_{2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{V} }{\partial n_j} & \nonumber \\
+		    &= V \left(\frac{1}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} + \frac{1}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} }\right) & \text{\dconp}
+\intertext{and:}
+ \mathcal{J}_{2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{P} }{\partial n_j} & \nonumber \\
+		    &=  \frac{P}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} } + T \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} & \text{\dconv}
+\end{align}
+
+\subsection{Molar source term derivatives}
+\label{s:molar_source_derivatives}
+\subsubsection{Temperature derivative}
+
+The derivative of the net molar rate of production of a species $k$ with respect to temperature is:
+\begin{align}
+ \label{e:dnkdotdT_compact}
+ \mathcal{J}_{k + 2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial T } \nonumber \\
+ \intertext{or:}
+ \mathcal{J}_{k + 2, 1} & = V \frac{\partial \dot{\omega}_{k} }{\partial T }
+\end{align}
+
+\subsubsection{State parameter derivatives}
+
+Similarly, the derivative of the net molar rate of production of a species $k$ with respect to the state parameter is:
+\begin{subequations}
+\begin{align}
+  \mathcal{J}_{k + 2, 2} &= \frac{\partial }{\partial V } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial V } & \nonumber \\
+			 &= V \frac{\partial \dot{\omega}_{k} }{\partial V } + \dot{\omega}_{k} & \text{\dconp}
+  \intertext{and:}
+  \mathcal{J}_{k + 2, 2} &= \frac{\partial }{\partial P } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial P } & \nonumber \\
+			 &=V \frac{\partial \dot{\omega}_{k} }{\partial P } & \text{\dconv}
+\end{align}
+\end{subequations}
+
+\subsubsection{Molar derivative}
+
+Finally, the derivative of the net molar rate of production of a species $k$ with respect to the moles of species $j$ is:
+\begin{align}
+ \mathcal{J}_{k + 2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial n_j } \nonumber \\
+  \intertext{or:}
+ \mathcal{J}_{k + 2, j + 2} & = V \frac{\partial \dot{\omega}_{k} }{\partial {n_j} }
+\end{align}
+
 \subsection{Species production rate derivatives}
@@
  \label{e:dwdot_dT}
- \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial T }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial T }_{i} + \nu_{k,i} \frac{\partial R }{\partial T }_{i} c_{i}\right)} & \cond{dwdot}{,}
-\end{align}
-other species:
+ \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial T }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial T } + \nu_{k,i} \frac{\partial R_{i}}{\partial T } c_{i}\right)} & \cond{dwdot}{,}
+\end{align}
+the state parameter:
+\begin{subequations}
+ \label{e:dwdot_de}
+ \begin{align}
+  \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial V }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial V } + \nu_{k,i} \frac{\partial R_{i}}{\partial V } c_{i}\right)} & \cond{dwdot}{\dconp}\\
+  \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial P }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial P } + \nu_{k,i} \frac{\partial R_{i}}{\partial P } c_{i}\right)} & \cond{dwdot}{\dconv}
+ \end{align}
+\end{subequations}
+and other species:
 \begin{align}
@@
 \end{align}
-and the state parameter:
-\begin{subequations}
- \label{e:dwdot_de}
- \begin{align}
-  \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial V }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial V }_{i} + \nu_{k,i} \frac{\partial R }{\partial V }_{i} c_{i}\right)} & \cond{dwdot}{\dconp}\\
-  \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial P }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial P }_{i} + \nu_{k,i} \frac{\partial R }{\partial P }_{i} c_{i}\right)} & \cond{dwdot}{\dconv}
- \end{align}
-\end{subequations}
 
@@
 \begin{equation}
- \frac{\partial R }{\partial T }_{i} = \frac{\partial}{\partial T}\left({k_{f, i}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right)
+ \frac{\partial R_{i}}{\partial T } = \frac{\partial}{\partial T}\left({k_{f, i}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right)
 \end{equation}
@@
 Finally, combining~\cref{e:dropf_dt,e:dropr_dt} gives the total temperature derivative of the net rate of progress:
-\loadeq{dropdT}
+\begin{align}
+ \label{e:drop_dT}
+ \frac{\partial R_{i}}{\partial T } =& \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right) - \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_{r, i}} \nonumber \\
+				      &\qquad + \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_{r, i}} - \frac{[C] S^{\prime}_{\ns}}{T} {k_{f, i}}
+\end{align}
+
+\subsubsection{State parameter derivatives}
+\label{s:dri_de}
+Following the outline of~\cref{s:dri_dnj}---while utilising~\cref{e:dck_de} for the species concentration derivatives---the forward rate of progress derivatives are:
+\begin{subequations}
+ \begin{align}
+  \lhs{dri_de}{\frac{\partial {R_{f, i}} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime}_{\ns}}{V} {k_{f, i}} - \frac{{R_{f, i}}}{V} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}}&\cond{dri_de}{\dconp}\\
+  \lhs{dri_de}{\frac{\partial {R_{f, i}} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime}_{\ns} {k_{f, i}}}{T \Ru}}&\cond{dri_de}{\dconv}
+ \end{align}
+\end{subequations}
+and similarly, the reverse rate of progress derivatives:
+\begin{subequations}
+ \begin{align}
+  \lhs{dri_de}{\frac{\partial {R_{r, i}} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime\prime}_{\ns}}{V} {k_{r, i}} - \frac{{R_{r, i}}}{V} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i}}&\cond{dri_de}{\dconp}\\
+  \lhs{dri_de}{\frac{\partial {R_{r, i}} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime\prime}_{\ns} {k_{r, i}}}{T \Ru}}&\cond{dri_de}{\dconv}
+ \end{align}
+\end{subequations}
+giving finally:
+\begin{subequations}
+ \label{e:dropi_de}
+ \begin{align}
+  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial V }} &= \rhs{dri_de}{\frac{1}{V}\Biggl( [C]\left(S^{\prime}_{\ns} {k_{f, i}} - S^{\prime\prime}_{\ns} {k_{r, i}}\right) + } & \nonumber\\
+  & \rhs{dri_de}{\mathindent[farth] \left({R_{r, i}} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i} - {R_{f, i}} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}\right) \Biggr)}&\cond{dri_de}{\dconp}\\
+  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial P }} &= \rhs{dri_de}{\frac{1}{T \Ru} \left(S^{\prime}_{\ns} {k_{f, i}} - S^{\prime\prime}_{\ns} {k_{r, i}}\right)}&\cond{dri_de}{\dconv}
+ \end{align}
+\end{subequations}
+
 \subsubsection{Molar derivative}
@@
 \end{equation}
-
-\subsubsection{State parameter derivatives}
-\label{s:dri_de}
-Following the outline of~\cref{s:dri_dnj}---while utilising~\cref{e:dck_de} for the species concentration derivatives---the forward rate of progress derivatives are:
-\begin{subequations}
- \begin{align}
-  \lhs{dri_de}{\frac{\partial {R_{f, i}} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime}_{\ns}}{V} {k_{f, i}} - \frac{{R_{f, i}}}{V} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}}&\cond{dri_de}{\dconp}\\
-  \lhs{dri_de}{\frac{\partial {R_{f, i}} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime}_{\ns} {k_{f, i}}}{T \Ru}}&\cond{dri_de}{\dconv}
- \end{align}
-\end{subequations}
-and similarly, the reverse rate of progress derivatives:
-\begin{subequations}
- \begin{align}
-  \lhs{dri_de}{\frac{\partial {R_{r, i}} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime\prime}_{\ns}}{V} {k_{r, i}} - \frac{{R_{r, i}}}{V} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i}}&\cond{dri_de}{\dconp}\\
-  \lhs{dri_de}{\frac{\partial {R_{r, i}} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime\prime}_{\ns} {k_{r, i}}}{T \Ru}}&\cond{dri_de}{\dconv}
- \end{align}
-\end{subequations}
-giving finally:
-\begin{subequations}
- \label{e:dropi_de}
- \begin{align}
-  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial V }} &= \rhs{dri_de}{\frac{1}{V}\Biggl( [C]\left(S^{\prime}_{\ns} {k_{f, i}} - S^{\prime\prime}_{\ns} {k_{r, i}}\right) + } & \nonumber\\
-  & \rhs{dri_de}{\mathindent[farth] \left({R_{r, i}} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i} - {R_{f, i}} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}\right) \Biggr)}&\cond{dri_de}{\dconp}\\
-  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial P }} &= \rhs{dri_de}{\frac{1}{T \Ru} \left(S^{\prime}_{\ns} {k_{f, i}} - S^{\prime\prime}_{\ns} {k_{r, i}}\right)}&\cond{dri_de}{\dconv}
- \end{align}
-\end{subequations}
 
@@
 Similar to~\cref{s:dri_dt},~\cref{e:dkf_plog_dt} may be substituted into~\cref{e:dropfdt_1,e:droprdt_1} to give:
-\loadeq{dropplogdT}
+\begin{align}
+ \label{e:drop_plog_dT}
+ \frac{\partial R_{i} }{\partial T } &= \left(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \beta_1 + \frac{E_{a_1}}{T \Ru} \right) \times \nonumber \\
+				     & \mathindent \left(\frac{{R_{f, i}} - {R_{r, i}}}{T}\right) + {R_{r, i}} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_{r, i}} - S^{\prime}_{\ns} {k_{f, i}}\right)
+\end{align}
 
@@
 Following the same process for Arrhenius-based reactions (see~\cref{s:dri_dt}), the derivative of the net rate of progress with respect to temperature is:
-\loadeq{dropchebdT}
+\begin{align}
+ \label{e:drop_cheb_dT}
+ \frac{\partial R_{i}}{\partial T } &= \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \left[ - \frac{2 \log{\left (10 \right )} \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right) \right] \left({R_{f, i}} - {R_{r, i}}\right) + \nonumber \\
+				     & \mathindent \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} {R_{r, i}} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_{r, i}} - S^{\prime}_{\ns} {k_{f, i}}\right)
+\end{align}
 
@@
 \subsubsection{Elementary reactions}
-For elementary reactions, all derivatives of the pressure-modification term are zero:
-\begin{align}
-\lhs{dci_elem}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{,} \label{e:dci_elem_dt} \\
-\lhs{dci_elem}{\frac{\partial c_{i} }{\partial {n_j} }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{,} \label{e:dci_elem_dnj} \\
-\end{align}
-and:
+For elementary reactions, the derivative of the pressure-modification term with respect to temperature is:
+\begin{align}
+\lhs{dci_elem}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{,} \label{e:dci_elem_dt}
+\end{align}
+the state parameter:
 \begin{subequations}
@@
 \end{subequations}
+and the moles:
+\begin{align}
+ \lhs{dci_elem}{\frac{\partial c_{i} }{\partial {n_j} }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{.} \label{e:dci_elem_dnj}
+\end{align}
 
 \subsubsection{Third-body enhanced reactions}
-For a third-body enhanced reaction~\cref{e:thd_mix} is differentiated to give:
-\begin{align}
- \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial T }} &= \rhs{dci_thd_mix}{-\frac{[C] \alpha_{\ns,i}}{T}} & \cond{dci_thd_mix}{,} \label{e:dci_thd_dt} \\
- \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial n_{j}}} &= \rhs{dci_thd_mix}{\frac{1}{V} \left(- \alpha_{\ns,i} + \alpha_{j,i}\right)} & \cond{dci_thd_mix}{,} \label{e:dci_thd_dnj}
-\end{align}
-and:
+For a third-body enhanced reaction~\cref{e:thd_mix} is differentiated with respect to temperature to give:
+\begin{align}
+ \label{e:dci_thd_dt}
+ \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial T }} &= \rhs{dci_thd_mix}{-\frac{[C] \alpha_{\ns,i}}{T}} & \cond{dci_thd_mix}{,}
+\end{align}
+and with respect to the state parameter:
 \begin{subequations}
@@
 \end{subequations}
+and finally, with respect to the moles of species $j$:
+\begin{align}
+\label{e:dci_thd_dnj}
+\lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial n_{j}}} &= \rhs{dci_thd_mix}{\frac{1}{V} \left(- \alpha_{\ns,i} + \alpha_{j,i}\right)} & \cond{dci_thd_mix}{.}
+\end{align}
 
@@
 \begin{align}
- \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_thd_unity}{- \frac{[C]}{T}}& \cond{dci_thd_unity}{,} \label{e:dci_thd_unity_dt} \\
- \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial n_{j} }} &= \rhs{dci_thd_unity}{0}[c] & \cond{dci_thd_unity}{,}
-\end{align}
-and:
+ \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_thd_unity}{- \frac{[C]}{T}}& \cond{dci_thd_unity}{,} \label{e:dci_thd_unity_dt}
+\end{align}
+and for the state parameter:
 \begin{subequations}
@@
 \end{subequations}
+and finally for the moles of species $j$:
+\begin{align}
+  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial n_{j} }} &= \rhs{dci_thd_unity}{0}[c] & \cond{dci_thd_unity}{,}
+\end{align}
 
@@
 \begin{align}
- \lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_thd_spec}{-\frac{\delta_{\ns m}}{T} [C]}&\cond{dci_thd_spec}{,} \label{e:dci_thd_spec_dt} \\
- \lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial n_j }} &= \rhs{dci_thd_spec}{\frac{1}{V} \left(- \delta_{\ns m} + \delta_{j m}\right)}&\cond{dci_thd_spec}{,} \label{e:dci_thd_spec_dnj} \\
+ \lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_thd_spec}{-\frac{\delta_{\ns m}}{T} [C]}&\cond{dci_thd_spec}{,} \label{e:dci_thd_spec_dt}
 \end{align}
@@
 \end{subequations}
+and for the molar derivative:
+\begin{align}
+ \lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial n_j }} &= \rhs{dci_thd_spec}{\frac{1}{V} \left(- \delta_{\ns m} + \delta_{j m}\right)}&\cond{dci_thd_spec}{,} \label{e:dci_thd_spec_dnj}
+\end{align}
 
@@
  \label{e:dci_fall_dt}
- \lhs{dci_fall}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial T } + \frac{\partial P_{r, i} }{\partial T } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{,}\\
+ \lhs{dci_fall}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial T } + \frac{\partial P_{r, i} }{\partial T } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{,}
+\end{align}
+and with respect to the state parameter:
+\begin{subequations}
+ \label{e:dci_fall_de}
+ \begin{align}
+  \lhs{dci_fall}{\frac{\partial c_i }{\partial V }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial V } + \frac{\partial P_{r, i} }{\partial V } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{\dconp} \\
+  \lhs{dci_fall}{\frac{\partial c_i }{\partial P }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial P } + \frac{\partial P_{r, i} }{\partial P } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{\dconv}
+ \end{align}
+\end{subequations}
+and finally the molar derivative:
+\begin{align}
  \label{e:dci_fall_dnj}
@@
 \end{align}
-and with respect to the state parameter:
-\begin{subequations}
- \label{e:dci_fall_de}
- \begin{align}
-  \lhs{dci_fall}{\frac{\partial c_i }{\partial V }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial V } + \frac{\partial P_{r, i} }{\partial V } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{\dconp} \\
-  \lhs{dci_fall}{\frac{\partial c_i }{\partial P }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial P } + \frac{\partial P_{r, i} }{\partial P } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{\dconv}
- \end{align}
-\end{subequations}
 
@@
  \lhs{dci_chem}{\frac{\partial c_i }{\partial T }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial T } - \frac{\partial P_{r, i} }{\partial T } c_{i}\right)} & \cond{dci_chem}{,} \\
- \label{e:dci_chem_dnj}
- \lhs{dci_chem}{\frac{\partial c_i }{\partial n_j }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial n_j } - \frac{\partial P_{r, i} }{\partial n_j } c_{i}\right)} & \cond{dci_chem}{,}
-\end{align}
-and:
+ \end{align}
+and for the state parameter derivative:
 \begin{subequations}
@@
 \end{subequations}
+and finally, the molar derivative:
+\begin{align}
+ \label{e:dci_chem_dnj}
+ \lhs{dci_chem}{\frac{\partial c_i }{\partial n_j }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial n_j } - \frac{\partial P_{r, i} }{\partial n_j } c_{i}\right)} & \cond{dci_chem}{.}
+\end{align}
 
@@
 \begin{align}
- \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial T }} &= \rhs{dpr_mix}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) - \frac{[C] k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i}}} & {,} \label{e:dpr_mix_dt} \\
- \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial n_j }} &= \rhs{dpr_mix}{\frac{k_{0, i} \left(- \alpha_{\ns,i} + \alpha_{j,i}\right)}{k_{\infty, i} V}} & \cond{dpr_mix}{,} \label{e:dpr_mix_dnj}
-\end{align}
-and:
+ \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial T }} &= \rhs{dpr_mix}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) - \frac{[C] k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i}}} & \cond{dpr_mix}{,} \label{e:dpr_mix_dt}
+\end{align}
+and for the state parameter derivative:
 \begin{subequations}
@@
 \end{subequations}
+and:
+\begin{align}
+ \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial n_j }} &= \rhs{dpr_mix}{\frac{k_{0, i} \left(- \alpha_{\ns,i} + \alpha_{j,i}\right)}{k_{\infty, i} V}} & \cond{dpr_mix}{.} \label{e:dpr_mix_dnj}
+\end{align}
 
@@
 \begin{align}
- \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial T }} &= \rhs{dpr_unity}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} - 1 + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right)} & \cond{dpr_unity}{,} \label{e:dpr_unity_dt} \\
- \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial n_j }} &= \rhs{dpr_unity}{0} & \cond{dpr_unity}{,} \label{e:dpr_unity_dnj}
-\end{align}
-and:
+ \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial T }} &= \rhs{dpr_unity}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} - 1 + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right)} & \cond{dpr_unity}{,} \label{e:dpr_unity_dt}
+\end{align}
+and for the state parameter:
 \begin{subequations}
@@
 \end{subequations}
+and the molar derivative:
+\begin{align}
+ \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial n_j }} &= \rhs{dpr_unity}{0} & \cond{dpr_unity}{.} \label{e:dpr_unity_dnj}
+\end{align}
 
@@
 \begin{align}
- \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial T }} &=\rhs{dpr_spec}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) - \frac{[C] k_{0, i} \delta_{\ns m}}{T k_{\infty, i}}} & \cond{dpr_spec}{,} \label{e:dpr_spec_dt} \\
- \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial n_j }} &=\rhs{dpr_spec}{\frac{k_{0, i}}{V k_{\infty, i}} \left(- \delta_{\ns m} + \delta_{j m}\right)} & \cond{dpr_spec}{,} \label{e:dpr_spec_dnj}
-\end{align}
-and:
+ \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial T }} &=\rhs{dpr_spec}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) - \frac{[C] k_{0, i} \delta_{\ns m}}{T k_{\infty, i}}} & \cond{dpr_spec}{,} \label{e:dpr_spec_dt}
+\end{align}
+for the state paramter derivative:
 \begin{subequations}
@@
 \end{subequations}
+and the molar derivative:
+\begin{align}
+ \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial n_j }} &=\rhs{dpr_spec}{\frac{k_{0, i}}{V k_{\infty, i}} \left(- \delta_{\ns m} + \delta_{j m}\right)} & \cond{dpr_spec}{.} \label{e:dpr_spec_dnj}
+\end{align}
 
@@
 \begin{align}
- \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial T } } &= \rhs{dpr_gen}{P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}} & \cond{dpr_gen}{,} \label{e:dpr_dt} \\
- \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial n_j}} &= \rhs{dpr_gen}{\frac{k_{0, i}\bar{\theta}_{P_{r, i}, \partial n_j}}{V k_{\infty, i}}} & \cond{dpr_gen}{,} \label{e:dpr_dnj}
-\end{align}
-and:
+ \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial T } } &= \rhs{dpr_gen}{P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}} & \cond{dpr_gen}{,} \label{e:dpr_dt}
+\end{align}
+and the state parameter:
 \begin{subequations}
@@
 \end{subequations}
+and the molar derivative:
+\begin{align}
+ \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial n_j}} &= \rhs{dpr_gen}{\frac{k_{0, i}\bar{\theta}_{P_{r, i}, \partial n_j}}{V k_{\infty, i}}} & \cond{dpr_gen}{.} \label{e:dpr_dnj}
+\end{align}
+
 with:
@@
   \rhs{theta_pr_bar_dt}{- \frac{[C] k_{0, i} \delta_{\ns m}}{T k_{\infty, i}}} & \text{if species,}
- \end{dcases} \\
- \label{e:dpr_theta_bar_dnj}
- \bar{\theta}_{P_{r, i}, \partial n_j} &=
- \begin{dcases}
-  \rhs{theta_pr_bar_dnj}{- \alpha_{\ns,i} + \alpha_{j,i}} & \text{if mix,} \\
-  \rhs{theta_pr_bar_dnj}{0}[c] & \text{if unity,}  \\
-  \rhs{theta_pr_bar_dnj}{- \delta_{\ns m} + \delta_{j m}} & \text{if species,}
  \end{dcases} \\
@@
   \rhs{theta_pr_bar_dp}{\frac{k_{0, i} \delta_{\ns m}}{T k_{\infty, i} \Ru}} & \text{if species,}
+ \end{dcases} \\
+ \label{e:dpr_theta_bar_dnj}
+ \bar{\theta}_{P_{r, i}, \partial n_j} &=
+ \begin{dcases}
+  \rhs{theta_pr_bar_dnj}{- \alpha_{\ns,i} + \alpha_{j,i}} & \text{if mix,} \\
+  \rhs{theta_pr_bar_dnj}{0}[c] & \text{if unity,}  \\
+  \rhs{theta_pr_bar_dnj}{- \delta_{\ns m} + \delta_{j m}} & \text{if species,}
  \end{dcases}
@@
  \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial T }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{,} \label{e:dfi_lind_dt} \\
- \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial n_j }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{,} \label{e:dfi_lind_dnj}
-\end{align}
-and:
+\end{align}
+and, for the state parameter:
 \begin{subequations}
@@
 \end{subequations}
+and the molar derivative:
+\begin{align}
+ \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial n_j }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{,} \label{e:dfi_lind_dnj}
+\end{align}
 
@@
 \begin{align}
- \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial T }} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial F_{cent} } \frac{\text{d} F_{cent} }{\text{d} T } + \frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial T }} & \cond{dfi_troe_1}{,} \label{e:dfi_troe_dt} \\
- \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial n_j}} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial n_j }} & \cond{dfi_troe_1}{,} \label{e:dfi_troe_dnj}
-\end{align}
-and:
+ \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial T }} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial F_{cent} } \frac{\text{d} F_{cent} }{\text{d} T } + \frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial T }} & \cond{dfi_troe_1}{,} \label{e:dfi_troe_dt}
+\end{align}
+and the state parameter derivative:
 \begin{subequations}
@@
 \end{subequations}
+and:
+\begin{align}
+ \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial n_j}} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial n_j }} & \cond{dfi_troe_1}{.} \label{e:dfi_troe_dnj}
+\end{align}
 where:
@@
 \begin{align}
- \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial T }} &= F_{i} \vast(\frac{X \left(- \frac{\operatorname{exp}\left({- \frac{T}{c}}\right)}{c} + \frac{a b}{T^{2}} \operatorname{exp}\left({- \frac{b}{T}}\right)\right)}{a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right)} + \nonumber \\
- & \mathindent \frac{\partial P_{r, i} }{\partial T } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}+ \frac{e}{T} \vast) & , \label{e:dfi_sri_dt} \\
- \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial {n_j} }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial {n_j} } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} & , \label{e:dfi_sri_dnj}
-\end{align}
-and:
+ \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial T }} &= \rhs{dfi_sri}{F_{i} \vast(\frac{X \left(- \frac{\operatorname{exp}\left({- \frac{T}{c}}\right)}{c} + \frac{a b}{T^{2}} \operatorname{exp}\left({- \frac{b}{T}}\right)\right)}{a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right)} + } & \nonumber \\
+ & \rhs{dfi_sri}{\mathindent \frac{\partial P_{r, i} }{\partial T } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\biggl[}} & \nonumber \\
+ & \rhs{dfi_sri}{\mathindent[largerstill] a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \biggr] + \frac{e}{T} \vast)} & \cond{dfi_sri}{,} \label{e:dfi_sri_dt}
+\end{align}
+the state parameter derivative:
 \begin{subequations}
@@
 \end{subequations}
+and:
+\begin{align} 
+ \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial {n_j} }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial {n_j} } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} & \cond{dfi_sri}{.} \label{e:dfi_sri_dnj}
+\end{align}
+
 where:
@@
 \begin{align}
-\lhs{dfi_gen}{\frac{\partial F_{i} }{\partial T }} &= \rhs{dfi_gen}{F_{i} \Theta_{F_i, \partial T}} & \cond{dfi_gen}{,} \\
-\lhs{dfi_gen}{\frac{\partial F_{i} }{\partial {n_j}}} &= \rhs{dfi_gen}{\frac{F_{i} k_{0, i} \Theta_{F_i, \partial n_j}}{V k_{\infty, i}} \bar{\theta}_{P_{r, i}, \partial n_j}} & \cond{dfi_gen}{,}
-\end{align}
-and:
+\lhs{dfi_gen}{\frac{\partial F_{i} }{\partial T }} &= \rhs{dfi_gen}{F_{i} \Theta_{F_i, \partial T}} & \cond{dfi_gen}{,}
+\end{align}
+for the state parameter derivative:
 \begin{subequations}
@@
 \end{subequations}
+and the molar derivative:
+\begin{align}
+ \lhs{dfi_gen}{\frac{\partial F_{i} }{\partial {n_j}}} &= \rhs{dfi_gen}{\frac{F_{i} k_{0, i} \Theta_{F_i, \partial n_j}}{V k_{\infty, i}} \bar{\theta}_{P_{r, i}, \partial n_j}} & \cond{dfi_gen}{.}
+\end{align}
 where for Lindemann falloff reactions:
@@
 \label{e:dfi_lind_dt_prod}
-\lhs{dfi_lind}{\Theta_{F_i, \partial T}} &= \rhs{dfi_lind}{0} \\
+\lhs{dfi_lind_gen}{\Theta_{F_i, \partial T}} &= \rhs{dfi_lind_gen}{0}
+\end{align}
+and, for the state parameter derivative:
+\begin{subequations}
+ \label{e:dfi_lind_de_prod}
+ \begin{align}
+  \lhs{dfi_lind_gen}{\Theta_{F_i, \partial V}} &= \rhs{dfi_lind_gen}{0} \\
+  \lhs{dfi_lind_gen}{\Theta_{F_i, \partial P}} &= \rhs{dfi_lind_gen}{0}
+ \end{align}
+\end{subequations}
+and:
+\begin{align}
 \label{e:dfi_lind_dnj_prod}
-\lhs{dfi_lind}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_lind}{0}
-\end{align}
-and:
-\begin{subequations}
- \label{e:dfi_lind_de_prod}
- \begin{align}
-  \lhs{dfi_lind}{\Theta_{F_i, \partial V}} &= \rhs{dfi_lind}{0} \\
-  \lhs{dfi_lind}{\Theta_{F_i, \partial P}} &= \rhs{dfi_lind}{0}
- \end{align}
-\end{subequations}
+\lhs{dfi_lind_gen}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_lind_gen}{0}
+\end{align}
+
 For Troe falloff reactions:
@@
 \label{e:dfi_troe_dt_prod}
-\lhs{dfi_troe}{\Theta_{F_i, \partial T}} &= 
-\rhs{dfi_troe}{ - \frac{B_{Troe}}{F_{cent} P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}} \times \Biggl(} \nonumber \\
-&\rhs{dfi_troe}{ 2 A_{Troe} F_{cent} \left(0.14 A_{Troe} + B_{Troe}\right) \times } \nonumber \\
-&\rhs{dfi_troe}{\left( P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right) \log{\left (F_{cent} \right )} + P_{r, i} \frac{\text{d} F_{cent} }{\text{d} T } \times \Bigl(} \nonumber \\
-&\rhs{dfi_troe}{\mathindent 2 A_{Troe} \left(1.1762 A_{Troe} - 0.67 B_{Troe}\right) \log{\left (F_{cent} \right )} - } \nonumber \\
-&\rhs{dfi_troe}{\mathindent B_{Troe} \left(A_{Troe}^{2} + B_{Troe}^{2}\right) \log{\left (10 \right )}\Bigr)\Biggr)} \\
+\lhs{dfi_troe_gen}{\Theta_{F_i, \partial T}} &=
+\rhs{dfi_troe_gen}{ - \frac{B_{Troe}}{F_{cent} P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}} \times \Biggl(} \nonumber \\
+&\rhs{dfi_troe_gen}{ 2 A_{Troe} F_{cent} \left(0.14 A_{Troe} + B_{Troe}\right) \times } \nonumber \\
+&\rhs{dfi_troe_gen}{\left( P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right) \log{\left (F_{cent} \right )} + P_{r, i} \frac{\text{d} F_{cent} }{\text{d} T } \times \Bigl(} \nonumber \\
+&\rhs{dfi_troe_gen}{\mathindent 2 A_{Troe} \left(1.1762 A_{Troe} - 0.67 B_{Troe}\right) \log{\left (F_{cent} \right )} - } \nonumber \\
+&\rhs{dfi_troe_gen}{\mathindent B_{Troe} \left(A_{Troe}^{2} + B_{Troe}^{2}\right) \log{\left (10 \right )}\Bigr)\Biggr)}
+\end{align}
+and for the state parameter derivative:
+\begin{subequations}
+ \label{e:dfi_troe_de_prod}
+ \begin{align}
+   \lhs{dfi_troe_gen}{\Theta_{F_i, \partial V}} &= \rhs{dfi_troe_gen}{- \frac{2 A_{Troe} B_{Troe} \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}} \times \bigl(} \nonumber \\
+					    & \rhs{dfi_troe_gen}{\mathindent[alittlebit] 0.14 A_{Troe} + B_{Troe}\bigr) \left(P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}\right)} \\
+   \lhs{dfi_troe_gen}{\Theta_{F_i, \partial P}} &= \rhs{dfi_troe_gen}{- \frac{2 A_{Troe} B_{Troe} \bar{\theta}_{P_{r, i}, \partial P} \left(0.14 A_{Troe} + B_{Troe}\right) \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}}}
+ \end{align}
+\end{subequations}
+and:
+\begin{align}
  \label{e:dfi_troe_dnj_prod}
-\lhs{dfi_troe}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_troe}{- \frac{2 A_{Troe} B_{Troe} \left(0.14 A_{Troe} + B_{Troe}\right) \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}}}
+\lhs{dfi_troe_gen}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_troe_gen}{- \frac{2 A_{Troe} B_{Troe} \left(0.14 A_{Troe} + B_{Troe}\right) \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}}}
+\end{align}
+where $\Theta_{P_{r,i}, \partial \ldots}$ and $\bar{\theta}_{P_{r, i}, \partial \ldots}$ are given by~\cref{e:dpr_theta_dt,e:dpr_theta_bar_dt,e:dpr_theta_bar_dnj,e:dpr_theta_dv,e:dpr_theta_bar_dv,e:dpr_theta_dp} respectively.
+
+Finally, for SRI falloff reactions:
+\begin{align}
+ \label{e:dfi_sri_dt_prod}
+ \lhs{dfi_sri_gen}{\Theta_{F_i, \partial T}} &=
+ \rhs{dfi_sri_gen}{- \frac{X \left(\frac{\operatorname{exp}\left({- \frac{T}{c}}\right)}{c} - \frac{a b}{T^{2}} \operatorname{exp}\left({- \frac{b}{T}}\right)\right)}{a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right)} + \frac{e}{T} - \Biggl(} \nonumber \\
+ &\rhs{dfi_sri_gen}{\mathindent \frac{2 X^{2} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \times } \nonumber \\
+ &\rhs{dfi_sri_gen}{\mathindent \left(P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right) \log{\left (P_{r, i} \right ) \Biggr)}}
 \end{align}
@@
 \begin{subequations}
- \label{e:dfi_troe_de_prod}
- \begin{align}
-   \lhs{dfi_troe}{\Theta_{F_i, \partial V}} &= \rhs{dfi_troe}{- \frac{2 A_{Troe} B_{Troe} \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}} \times \bigl(} \nonumber \\
-					    & \rhs{dfi_troe}{\mathindent[alittlebit] 0.14 A_{Troe} + B_{Troe}\bigr) \left(P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}\right)} \\
-   \lhs{dfi_troe}{\Theta_{F_i, \partial P}} &= \rhs{dfi_troe}{- \frac{2 A_{Troe} B_{Troe} \bar{\theta}_{P_{r, i}, \partial P} \left(0.14 A_{Troe} + B_{Troe}\right) \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}}}
- \end{align}
-\end{subequations}
-where $\Theta_{P_{r,i}, \partial \ldots}$ and $\bar{\theta}_{P_{r, i}, \partial \ldots}$ are given by~\cref{e:dpr_theta_dt,e:dpr_theta_bar_dt,e:dpr_theta_bar_dnj,e:dpr_theta_dv,e:dpr_theta_bar_dv,e:dpr_theta_dp} respectively.
-Finally, for SRI falloff reactions:
-\begin{align}
- \label{e:dfi_sri_dt_prod}
- \lhs{dfi_sri}{\Theta_{F_i, \partial T}} &= 
- \rhs{dfi_sri}{- \frac{X \left(\frac{\operatorname{exp}\left({- \frac{T}{c}}\right)}{c} - \frac{a b}{T^{2}} \operatorname{exp}\left({- \frac{b}{T}}\right)\right)}{a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right)} + \frac{e}{T} - \Biggl(} \nonumber \\
- &\rhs{dfi_sri}{\mathindent \frac{2 X^{2} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \times } \nonumber \\
- &\rhs{dfi_sri}{\mathindent \left(P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right) \log{\left (P_{r, i} \right ) \Biggr)}} \\
+ \label{e:dfi_sri_de_prod}
+ \begin{align}
+  \lhs{dfi_sri_gen}{\Theta_{F_i, \partial V}} &=
+  \rhs{dfi_sri_gen}{- \frac{2 X^{2} \log{\left (P_{r, i} \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \left(P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}\right) \log \Biggl(} \nonumber \\
+& \rhs{dfi_sri_gen}{\mathindent \left( a \operatorname{exp}\left({\frac{T}{c}}\right) + \operatorname{exp}\left({\frac{b}{T}}\right)\right) \operatorname{exp}\left({- \frac{T}{c} - \frac{b}{T}}\right) \Biggr)} \\
+  \lhs{dfi_sri_gen}{\Theta_{F_i, \partial P}} &=
+  \rhs{dfi_sri_gen}{- \frac{2 X^{2} \bar{\theta}_{P_{r, i}, \partial P} \log{\left (P_{r, i} \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}
+ \end{align}
+\end{subequations}
+and finally, for the molar derivative:
+\begin{align}
  \label{e:dfi_sri_dnj_prod}
- \lhs{dfi_sri}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_sri}{- \frac{2 X^{2} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \log{\left (P_{r, i} \right )}}
-\end{align}
-and:
-\begin{subequations}
- \label{e:dfi_sri_de_prod}
- \begin{align}
-  \lhs{dfi_sri}{\Theta_{F_i, \partial V}} &=
-  \rhs{dfi_sri}{- \frac{2 X^{2} \log{\left (P_{r, i} \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \left(P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}\right) \log \Biggl(} \nonumber \\
-& \rhs{dfi_sri}{\mathindent \left( a \operatorname{exp}\left({\frac{T}{c}}\right) + \operatorname{exp}\left({\frac{b}{T}}\right)\right) \operatorname{exp}\left({- \frac{T}{c} - \frac{b}{T}}\right) \Biggr)} \\
-  \lhs{dfi_sri}{\Theta_{F_i, \partial P}} &=
-  \rhs{dfi_sri}{- \frac{2 X^{2} \bar{\theta}_{P_{r, i}, \partial P} \log{\left (P_{r, i} \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}
- \end{align}
-\end{subequations}
+ \lhs{dfi_sri_gen}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_sri_gen}{- \frac{2 X^{2} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \log{\left (P_{r, i} \right )}}
+\end{align}
 
 \subsubsection{Unimolecular\slash~recombination fall-off reactions (temporary product form)}
+\label{s:dfall_final}
 Using the temporary products developed in~\cref{s:dpr,s:dfi}, the fall-off reaction derivatives developed in~\cref{e:dci_fall_dt,e:dci_fall_dnj,e:dci_fall_de} may be simplified to:
 \begin{align}
-\lhs{dci_fall_complete}{\frac{\partial c }{\partial T }_{i}} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial T}}{P_{r, i} + 1} + \biggl(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial T}}{P_{r, i} + 1} + \Theta_{F_i, \partial T} + } & \nonumber \\
-							     &  \rhs{dci_fall_complete}{\mathindent[largerindentpleas] \Theta_{P_{r,i}, \partial T} - \frac{\bar{\theta}_{P_{r, i}, \partial T}}{P_{r, i} + 1}\biggr) c_{i}} & \cond{dci_fall_complete}{,} \\
-\lhs{dci_fall_complete}{\frac{\partial c }{\partial {n_j} }_{i}} &= \rhs{dci_fall_complete}{\frac{k_{0, i} \bar{\theta}_{P_{r, i}, \partial n_j}}{V k_{\infty, i} \left(P_{r, i} + 1\right)} \left(F_{i} \left(P_{r, i} \Theta_{F_i, \partial n_j} + 1\right) - c_{i}\right)} & \cond{dci_fall_complete}{,}
-\end{align}
-and:
-\begin{subequations}
- \begin{align}
-  \lhs{dci_fall_complete}{\frac{\partial c }{\partial V }_{i}} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1} + \biggl(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial V}}{P_{r, i} + 1} + \Theta_{F_i, \partial V} + } & \nonumber \\
+\label{e:dci_fall_dt_final}
+\lhs{dci_fall_complete}{\frac{\partial c_{i}}{\partial T }} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial T}}{P_{r, i} + 1} + \biggl(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial T}}{P_{r, i} + 1} + \Theta_{F_i, \partial T} + } & \nonumber \\
+							     &  \rhs{dci_fall_complete}{\mathindent[largerindentpleas] \Theta_{P_{r,i}, \partial T} - \frac{\bar{\theta}_{P_{r, i}, \partial T}}{P_{r, i} + 1}\biggr) c_{i}} & \cond{dci_fall_complete}{,}
+\end{align}
+for the state parameter derivative:
+\begin{subequations}
+ \label{e:dci_fall_de_final}
+ \begin{align}
+  \lhs{dci_fall_complete}{\frac{\partial c_{i}}{\partial V }} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1} + \biggl(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial V}}{P_{r, i} + 1} + \Theta_{F_i, \partial V} + } & \nonumber \\
 							       &  \rhs{dci_fall_complete}{\mathindent[largerindentpleas] \Theta_{P_{r,i}, \partial V} - \frac{\bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1} \biggr) c_{i}} & \cond{dci_fall_complete}{\dconp} \\
-  \lhs{dci_fall_complete}{\frac{\partial c }{\partial P }_{i}} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1} + \left(\Theta_{F_i, \partial P} - \frac{\bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_fall_complete}{\dconv}
- \end{align}
-\end{subequations}
-where the temporary products are given by~\cref{e:dpr_theta_dt,e:dpr_theta_bar_dt,e:dpr_theta_bar_dnj,e:dpr_theta_dv,e:dpr_theta_bar_dv,e:dpr_theta_dp}~and~\cref{e:dfi_lind_dt_prod,e:dfi_lind_dnj_prod,e:dfi_lind_de_prod,e:dfi_troe_dt_prod,e:dfi_troe_dnj_prod,e:dfi_troe_de_prod,e:dfi_sri_dt_prod,e:dfi_sri_dnj_prod,e:dfi_sri_de_prod}
+  \lhs{dci_fall_complete}{\frac{\partial c_{i}}{\partial P }} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1} + \left(\Theta_{F_i, \partial P} - \frac{\bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_fall_complete}{\dconv}
+ \end{align}
+\end{subequations}
+and the molar derivative:
+\begin{align}
+ \label{e:dci_fall_dnj_final}
+\lhs{dci_fall_complete}{\frac{\partial c }{\partial {n_j} }_{i}} &= \rhs{dci_fall_complete}{\frac{k_{0, i} \bar{\theta}_{P_{r, i}, \partial n_j}}{V k_{\infty, i} \left(P_{r, i} + 1\right)} \left(F_{i} \left(P_{r, i} \Theta_{F_i, \partial n_j} + 1\right) - c_{i}\right)} & \cond{dci_fall_complete}{.}
+\end{align}
+The temporary products in~\cref{e:dci_fall_dt_final,e:dci_fall_de_final,e:dci_fall_dnj_final} are given by~\cref{e:dpr_theta_dt,e:dpr_theta_bar_dt,e:dpr_theta_bar_dnj,e:dpr_theta_dv,e:dpr_theta_bar_dv,e:dpr_theta_dp}~and~\cref{e:dfi_lind_dt_prod,e:dfi_lind_dnj_prod,e:dfi_lind_de_prod,e:dfi_troe_dt_prod,e:dfi_troe_dnj_prod,e:dfi_troe_de_prod,e:dfi_sri_dt_prod,e:dfi_sri_dnj_prod,e:dfi_sri_de_prod}.
 
@@
 \begin{align}
-\lhs{dci_chem_complete}{\frac{\partial c }{\partial T }_{i}} &= \rhs{dci_chem_complete}{\left(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial T}}{P_{r, i} + 1} + \Theta_{F_i, \partial T} - \frac{\bar{\theta}_{P_{r, i}, \partial T}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_chem_complete}{,} \\
-\lhs{dci_chem_complete}{\frac{\partial c }{\partial {n_j} }_{i}} &= \rhs{dci_chem_complete}{\frac{k_{0, i} \bar{\theta}_{P_{r, i}, \partial n_j} \left(F_{i} \Theta_{F_i, \partial n_j} - c_{i}\right)}{k_{\infty, i} V \left(P_{r, i} + 1\right)}} & \cond{dci_chem_complete}{,}
-\end{align}
-and:
-\begin{subequations}
- \begin{align}
-  \lhs{dci_chem_complete}{\frac{\partial c }{\partial V }_{i}} &= \rhs{dci_fall_complete}{\left(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial V}}{P_{r, i} + 1} + \Theta_{F_i, \partial V} - \frac{\bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_chem_complete}{\dconp} \\
-  \lhs{dci_chem_complete}{\frac{\partial c }{\partial P }_{i}} &= \rhs{dci_fall_complete}{\left(\Theta_{F_i, \partial P} - \frac{\bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_chem_complete}{\dconv}
- \end{align}
-\end{subequations}
+\label{e:dci_chem_dt_final}
+\lhs{dci_chem_complete}{\frac{\partial c_{i}}{\partial T }} &= \rhs{dci_chem_complete}{\left(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial T}}{P_{r, i} + 1} + \Theta_{F_i, \partial T} - \frac{\bar{\theta}_{P_{r, i}, \partial T}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_chem_complete}{,}
+\end{align}
+and the state paramter derivative:
+\begin{subequations}
+ \begin{align}
+  \lhs{dci_chem_complete}{\frac{\partial c_{i}}{\partial V }} &= \rhs{dci_fall_complete}{\left(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial V}}{P_{r, i} + 1} + \Theta_{F_i, \partial V} - \frac{\bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_chem_complete}{\dconp} \\
+  \lhs{dci_chem_complete}{\frac{\partial c_{i}}{\partial P }} &= \rhs{dci_fall_complete}{\left(\Theta_{F_i, \partial P} - \frac{\bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_chem_complete}{\dconv}
+ \end{align}
+\end{subequations}
+and the molar derivative:
+\begin{align}
+\lhs{dci_chem_complete}{\frac{\partial c }{\partial {n_j} }_{i}} &= \rhs{dci_chem_complete}{\frac{k_{0, i} \bar{\theta}_{P_{r, i}, \partial n_j} \left(F_{i} \Theta_{F_i, \partial n_j} - c_{i}\right)}{k_{\infty, i} V \left(P_{r, i} + 1\right)}} & \cond{dci_chem_complete}{.}
+\end{align}
+The temporary products are the same as those detailed in~\cref{s:dfall_final}.
 
@@
 \begin{align}
- \lhs{dnkdotdT}{\mathcal{J}_{k + 2, 1}} &\pluseq  \rhs{dnkdotdT}{V \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial T }_{i} + \nu_{k,i} \frac{\partial R }{\partial T }_{i} c_{i}\right)} & \mathcond{dnkdotdT}{k = 1, \ldots, \ns}
-\end{align}
-the rate of progress temperature derivative for all reaction types---excluding Chebyshev and P-Log---is:
-\loadeq{dropdT}
-while for P-Log reactions:
-\loadeq{dropplogdT}
-and for Chebyshev reactions:
-\loadeq{dropchebdT}
+ \lhs{dnkdotdT}{\mathcal{J}_{k + 2, 1}} &\pluseq  \rhs{dnkdotdT}{V \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial T } + \nu_{k,i} \frac{\partial R_{i}}{\partial T } c_{i}\right)} & \mathcond{dnkdotdT}{k = 1, \ldots, \ns}
+\end{align}
+where the rate of progress derivatives are given by~\cref{e:drop_dT,e:drop_plog_dT,e:drop_cheb_dT} for arrhenius-based---i.e., elementary, third-body and falloff\slash chemically-activated---reactions, P-Log reactions and Chebyshev reactions respectively.
+The temperature derivatives of the pressure modification term are found in~\cref{e:dci_thd_dt,e:dci_thd_unity_dt,e:dci_thd_spec_dt} for third-body reactions, and~\cref{e:dci_fall_dt_final,e:dci_chem_dt_final} for falloff \slash chemically-activated reactions respectively.
+
+\subsection{State parameter derivatives}
+In this section, the derivatives of all source terms with respect to the state parameter are considered.
+
+\subsubsection{Temperature source term derivatives}
+Starting from~\cref{e:dTdotdT_nojac},~\cref{e:dnkdotdT_compact} is substituted in to yield:
 

### 2018-02-01 derivations.tex
fix lots of indicies outside of derivative from sympy

@@
 \begin{align}
- \frac{\partial\dot{T}}{\partial{T}} &= - \frac{\left( [C] \left( \frac{{C_p}_{\ns}}{T} - \frac{\text{d} {C_p} }{\text{d} T }_{\ns} \right) - \sum_{k=1}^{-1 + \ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns} + \frac{\text{d} {C_p} }{\text{d} T }_{k}\right) [C]_{k}\right)}{\left(\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}\right)^{2}} \times \nonumber \\
+ \frac{\partial\dot{T}}{\partial{T}} &= - \frac{\left( [C] \left( \frac{{C_p}_{\ns}}{T} - \frac{\text{d} {C_p} }{\text{d} T }_{\ns} \right) - \sum_{k=1}^{-1 + \ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns} + \frac{\text{d} C_{p, k} }{\text{d} T }\right) [C]_{k}\right)}{\left(\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}\right)^{2}} \times \nonumber \\
 				     & \mathindent[largerindent] \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k} - \nonumber \\
-				     & \mathindent \frac{\sum_{k=1}^{-1 + \ns} \left(\left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial T }_{k} + \left(\frac{\text{d} H }{\text{d} T }_{k} - \frac{W_{k}}{W_{\ns}} \frac{\text{d} H }{\text{d} T }_{\ns}\right) \dot{\omega}_{k}\right)}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}}
+				     & \mathindent \frac{\sum_{k=1}^{-1 + \ns} \left(\left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \left(\frac{\text{d} H_{k} }{\text{d} T } - \frac{W_{k}}{W_{\ns}} \frac{\text{d} H }{\text{d} T }_{\ns}\right) \dot{\omega}_{k}\right)}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}}
 \end{align}
@@
  \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \times \nonumber \\
-				     & \left( [C] \left( \frac{{C_p}_{\ns}}{T} - \frac{\text{d} {C_p} }{\text{d} T }_{\ns} \right) - \sum_{k=1}^{-1 + \ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns} + \frac{\text{d} {C_p} }{\text{d} T }_{k}\right) [C]_{k}\right) - \nonumber \\
-				     & \sum_{k=1}^{-1 + \ns} \left(\left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial T }_{k} + \left(\frac{\text{d} H }{\text{d} T }_{k} - \frac{W_{k}}{W_{\ns}} \frac{\text{d} H }{\text{d} T }_{\ns}\right) \dot{\omega}_{k}\right) \Biggr]
+				     & \left( [C] \left( \frac{{C_p}_{\ns}}{T} - \frac{\text{d} {C_p} }{\text{d} T }_{\ns} \right) - \sum_{k=1}^{-1 + \ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns} + \frac{\text{d} {C_p}_{k} }{\text{d} T }\right) [C]_{k}\right) - \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns} \left(\left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \left(\frac{\text{d} H_{k} }{\text{d} T } - \frac{W_{k}}{W_{\ns}} \frac{\text{d} H }{\text{d} T }_{\ns}\right) \dot{\omega}_{k}\right) \Biggr]
 \end{align}
@@
 \begin{align}
- \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{k} + \frac{{C_p}_{\ns}}{T}\right) [C]_{k} + \nonumber \\
-				     & \sum_{k=1}^{-1 + \ns} \biggl(\left(- H_{k} + \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial T }_{k} + \nonumber \\
-				     & \mathindent[larger] \left(- \frac{\text{d} H }{\text{d} T }_{k} + \frac{W_{k}}{W_{\ns}} \frac{\text{d} H }{\text{d} T }_{\ns}\right) \dot{\omega}_{k}\biggr)\Biggr]
+ \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns} \left(- \frac{\text{d} C_{p, k} }{\text{d} T } + \frac{{C_p}_{\ns}}{T}\right) [C]_{k} + \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns} \biggl(\left(- H_{k} + \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \nonumber \\
+				     & \mathindent[larger] \left(- \frac{\text{d} H_{k} }{\text{d} T } + \frac{W_{k}}{W_{\ns}} \frac{\text{d} H }{\text{d} T }_{\ns}\right) \dot{\omega}_{k}\biggr)\Biggr]
 \end{align}
@@
 \begin{align}
- \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{k} + \frac{{C_p}_{\ns}}{T}\right) [C]_{k} + & \nonumber \\
-				     & \sum_{k=1}^{-1 + \ns} \biggl( \left(- H_{k} + \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial T }_{k} + & \nonumber \\
+ \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns} \left(- \frac{\text{d} C_{p, k} }{\text{d} T } + \frac{{C_p}_{\ns}}{T}\right) [C]_{k} + & \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns} \biggl( \left(- H_{k} + \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + & \nonumber \\
 				     & \mathindent[larger] \left(- {C_p}_{k} + \frac{W_{k} {C_p}_{k} }{W_{\ns}} \right) \dot{\omega}_{k}\biggr)\Biggr] & \text{\dconp} \\
 \intertext{and similarly:}
-\frac{\partial\dot{T}}{\partial{T}}  &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns} \left(- \frac{\text{d} {C_v} }{\text{d} T }_{k} + \frac{{C_v}_{\ns}}{T}\right) [C]_{k} + & \nonumber \\
-				     & \sum_{k=1}^{-1 + \ns} \biggl(\left(- U_{k} + \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial T }_{k} + & \nonumber \\
+\frac{\partial\dot{T}}{\partial{T}}  &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns} \left(- \frac{\text{d} C_{v, k} }{\text{d} T } + \frac{{C_v}_{\ns}}{T}\right) [C]_{k} + & \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns} \biggl(\left(- U_{k} + \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + & \nonumber \\
 				     & \mathindent[larger] \left(- {C_v}_{k} + \frac{W_{k} {C_v}_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}\biggr)\Biggr] & \text{\dconv}
@@
 \begin{subequations}
-\begin{align}
- \frac{\partial\dot{T}}{\partial{V}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ - \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial V }_{k} + & \nonumber \\
+\label{e:dTdotdE_nojac}
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{V}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ - \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial V } + & \nonumber \\
 				     & \mathindent[largerindentand] \frac{1}{V} \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{-1 + \ns} \left(- {C_p}_{\ns} + {C_p}_{k}\right) [C]_{k}\Biggr] & \text{\dconp} \\
- \frac{\partial\dot{T}}{\partial{P}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial P }_{k} - & \nonumber \\
+ \frac{\partial\dot{T}}{\partial{P}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial P } - & \nonumber \\
 				     & \mathindent[largerindentand] \frac{{C_v}_{\ns}}{T \Ru} \frac{\text{d} T }{\text{d} t } \Biggr] & \text{\dconv}
@@
  \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{\left(\sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}\right) \sum_{k=1}^{-1 + \ns} - \frac{\delta_{j k}}{V} \left({C_p}_{\ns} - {C_p}_{k}\right)}{\left(\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}\right)^{2}} - \nonumber \\
-				       & \mathindent \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k}
-\end{align}
-Next, the Kronecker delta summation is simplified, and the the compact form of the energy equation factored out yielding:
-\begin{subequations}
-\begin{align}
- \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} & \nonumber \\
+				       & \mathindent \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} }
+\end{align}
+Next, the Kronecker delta summation is simplified, and the compact form of the energy equation factored out yielding:
+\begin{subequations}
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \nonumber \\
 				       &  \mathindent[largerindentand] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_p}_{\ns} + {C_p}_{j}\right)\Biggr] & \text{\dconp} \\
 \intertext{and:}
- \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} & \nonumber \\
+ \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \nonumber \\
 				       & \mathindent[largerindentand] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_v}_{\ns} + {C_v}_{j}\right) \Biggr] & \text{\dconv}
@@
  \mathcal{J}_{2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{V} }{\partial T} & \nonumber \\
-		    &= \frac{V}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(\frac{\partial \dot{\omega} }{\partial T }_{k} + \frac{\dot{\omega}_{k}}{T}\right) \nonumber \\
+		    &= \frac{V}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(\frac{\partial \dot{\omega}_{k} }{\partial T } + \frac{\dot{\omega}_{k}}{T}\right) \nonumber \\
 		    &  \mathindent + \frac{V}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconp}
@@
  \mathcal{J}_{2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{P} }{\partial T} & \nonumber \\
-		    &= \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(T \frac{\partial \dot{\omega} }{\partial T }_{k} + \dot{\omega}_{k}\right) & \nonumber \\
+		    &= \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(T \frac{\partial \dot{\omega}_{k} }{\partial T } + \dot{\omega}_{k}\right) & \nonumber \\
 		    & \mathindent + \frac{P}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconv}
@@
 \subsubsection{State parameter derivative}
-Similarly, the derivative of the the thermodynamic state parameter source terms with respect to themselves are:
+Similarly, the derivative of the thermodynamic state parameter source terms with respect to themselves are:
 \begin{align}
  \mathcal{J}_{2, 2} &= \frac{\partial }{\partial V } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{V} }{\partial V} & \nonumber \\
-		    &= \frac{1}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(V \frac{\partial \dot{\omega} }{\partial V }_{k} + \dot{\omega}_{k}\right) \nonumber \\
+		    &= \frac{1}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(V \frac{\partial \dot{\omega}_{k} }{\partial V } + \dot{\omega}_{k}\right) \nonumber \\
 		    & \mathindent + \frac{1}{T} \left(V \frac{\text{d} \dot{T} }{\text{d} V } + \dot{T}\right) & \text{\dconp} \\
@@
  \mathcal{J}_{2, 2} &= \frac{\partial }{\partial P } \frac{\text{d} P}{\text{d} t} = \frac{\partial \dot{P} }{\partial P} & \nonumber \\
-		    &= T \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial P }_{k} & \nonumber \\
+		    &= T \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial P } & \nonumber \\
 		    & \mathindent + \frac{1}{T} \left(P \frac{\text{d} \dot{T} }{\text{d} P } + \dot{T}\right) & \text{\dconv}
@@
  \mathcal{J}_{2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{V} }{\partial n_j} & \nonumber \\
-		    &= V \left(\frac{1}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} + \frac{1}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} }\right) & \text{\dconp}
+		    &= V \left(\frac{1}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } + \frac{1}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} }\right) & \text{\dconp}
 \intertext{and:}
  \mathcal{J}_{2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{P} }{\partial n_j} & \nonumber \\
-		    &=  \frac{P}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} } + T \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial {n_j} }_{k} & \text{\dconv}
+		    &=  \frac{P}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} } + T \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \text{\dconv}
 \end{align}
@@
 \begin{subequations}
+\label{e:dnkdotdE_compact}
 \begin{align}
@@
  \label{e:dwdot_dT}
- \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial T }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial T } + \nu_{k,i} \frac{\partial R_{i}}{\partial T } c_{i}\right)} & \cond{dwdot}{,}
+ \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial T }} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial T } + \nu_{k,i} \frac{\partial R_{i}}{\partial T } c_{i}\right)} & \cond{dwdot}{,}
 \end{align}
@@
  \begin{align}
-  \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial V }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial V } + \nu_{k,i} \frac{\partial R_{i}}{\partial V } c_{i}\right)} & \cond{dwdot}{\dconp}\\
-  \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial P }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial P } + \nu_{k,i} \frac{\partial R_{i}}{\partial P } c_{i}\right)} & \cond{dwdot}{\dconv}
+  \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial V }} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial V } + \nu_{k,i} \frac{\partial R_{i}}{\partial V } c_{i}\right)} & \cond{dwdot}{\dconp}\\
+  \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial P }} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial P } + \nu_{k,i} \frac{\partial R_{i}}{\partial P } c_{i}\right)} & \cond{dwdot}{\dconv}
  \end{align}
@@
  \label{e:dwdot_dnj}
- \lhs{dwdot}{\frac{\partial \dot{\omega} }{\partial {n_j} }_{k}} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial {n_j} }_{i} + \nu_{k,i} \frac{\partial R }{\partial {n_j} }_{i} c_{i}\right)} & \cond{dwdot}{,}
+ \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial {n_j} }} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c_{i} }{\partial {n_j} } + \nu_{k,i} \frac{\partial R_{i} }{\partial {n_j} } c_{i}\right)} & \cond{dwdot}{,}
 \end{align}
@@
  \label{e:dkf_dt}
- \frac{\text{d} {k_f} }{\text{d} T }_{i} = \frac{{k_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
+ \frac{\text{d} k_{f, i} }{\text{d} T } = \frac{{k_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
 \end{equation}
@@
  \label{e:dropfdt_1}
- \frac{\partial {R_f} }{\partial T }_{i} = \frac{\text{d} {k_f} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime}_{\ns,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
+ \frac{\partial {R_f} _{i}}{\partial T } = \frac{\text{d} k_{f, i} }{\text{d} T } \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime}_{\ns,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
 \end{equation}
@@
  \label{e:dropf_dt}
- \frac{\partial {R_f} }{\partial T }_{i} = - \frac{[C] S^{\prime}_{\ns}}{T} {k_{f, i}} + \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
+ \frac{\partial {R_f}_{i}}{\partial T } = - \frac{[C] S^{\prime}_{\ns}}{T} {k_{f, i}} + \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
 \end{equation}
@@
  \label{e:droprdt_1}
- \frac{\partial {R_r} }{\partial T }_{i} = \frac{\text{d} {k_r} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime\prime}_{\ns,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
+ \frac{\partial R_{r, i}}{\partial T } = \frac{\text{d} k_{r, i} }{\text{d} T } \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime\prime}_{\ns,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
 \end{equation}
@@
 \begin{align}
- \frac{\text{d} {k_r} }{\text{d} T }_{i} &= \left(- \frac{1}{{K_{c, i}}} \frac{\text{d} {K_{c, i}} }{\text{d} T } \frac{{k_f}_i}{{K_c}_i} + \frac{1}{{K_{c, i}}}\frac{\text{d} {k_{f, i}} }{\text{d} T } \right) \nonumber \\
+ \frac{\text{d} k_{r, i} }{\text{d} T } &= \left(- \frac{1}{{K_{c, i}}} \frac{\text{d} {K_{c, i}} }{\text{d} T } \frac{{k_f}_i}{{K_c}_i} + \frac{1}{{K_{c, i}}}\frac{\text{d} {k_{f, i}} }{\text{d} T } \right) \nonumber \\
 				         &= \left(- \frac{1}{{K_{c, i}}} \frac{\text{d} {K_{c, i}} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_{r, i}}
@@
 \begin{equation}
- \frac{\text{d} {K_{c, i}} }{\text{d} T } = {K_{c, i}} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} \;,
+ \frac{\text{d} {K_{c, i}} }{\text{d} T } = {K_{c, i}} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } \;,
 \end{equation}
@@
 \begin{equation}
- \frac{\text{d} B }{\text{d} T }_{k} = T \left(T \left(\frac{T a_{k,4}}{5} + \frac{a_{k,3}}{4}\right) + \frac{a_{k,2}}{3}\right) + \frac{a_{k,1}}{2} + \frac{1}{T} \left(a_{k,0} - 1 + \frac{a_{k,5}}{T}\right)
+ \frac{\text{d} B_{k} }{\text{d} T } = T \left(T \left(\frac{T a_{k,4}}{5} + \frac{a_{k,3}}{4}\right) + \frac{a_{k,2}}{3}\right) + \frac{a_{k,1}}{2} + \frac{1}{T} \left(a_{k,0} - 1 + \frac{a_{k,5}}{T}\right)
 \end{equation}
@@
 \label{e:dkr_dt}
-\frac{\text{d} {k_r} }{\text{d} T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_{r, i}} \;,
+\frac{\text{d} k_{r, i} }{\text{d} T } = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B_{k}}{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_{r, i}} \;,
 \end{equation}
@@
 \label{e:dropr_dt}
-\frac{\partial {R_r} }{\partial T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_{r, i}} - \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_{r, i}}
+\frac{\partial R_{r, i}}{\partial T } = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B_{k}}{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_{r, i}} - \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_{r, i}}
 \end{equation}
@@
  \label{e:drop_dT}
- \frac{\partial R_{i}}{\partial T } =& \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right) - \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_{r, i}} \nonumber \\
+ \frac{\partial R_{i}}{\partial T } =& \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right) - \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_{r, i}} \nonumber \\
 				      &\qquad + \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_{r, i}} - \frac{[C] S^{\prime}_{\ns}}{T} {k_{f, i}}
@@
  \frac{\partial R_{i} }{\partial T } &= \left(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \beta_1 + \frac{E_{a_1}}{T \Ru} \right) \times \nonumber \\
-				     & \mathindent \left(\frac{{R_{f, i}} - {R_{r, i}}}{T}\right) + {R_{r, i}} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_{r, i}} - S^{\prime}_{\ns} {k_{f, i}}\right)
+				     & \mathindent \left(\frac{{R_{f, i}} - {R_{r, i}}}{T}\right) + {R_{r, i}} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_{r, i}} - S^{\prime}_{\ns} {k_{f, i}}\right)
 \end{align}
@@
  \frac{\partial R_{i}}{\partial T } &= \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \left[ - \frac{2 \log{\left (10 \right )} \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right) \right] \left({R_{f, i}} - {R_{r, i}}\right) + \nonumber \\
-				     & \mathindent \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} {R_{r, i}} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_{r, i}} - S^{\prime}_{\ns} {k_{f, i}}\right)
+				     & \mathindent \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } {R_{r, i}} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_{r, i}} - S^{\prime}_{\ns} {k_{f, i}}\right)
 \end{align}
@@
  \label{e:dci_fall_dnj_final}
-\lhs{dci_fall_complete}{\frac{\partial c }{\partial {n_j} }_{i}} &= \rhs{dci_fall_complete}{\frac{k_{0, i} \bar{\theta}_{P_{r, i}, \partial n_j}}{V k_{\infty, i} \left(P_{r, i} + 1\right)} \left(F_{i} \left(P_{r, i} \Theta_{F_i, \partial n_j} + 1\right) - c_{i}\right)} & \cond{dci_fall_complete}{.}
+\lhs{dci_fall_complete}{\frac{\partial c_{i} }{\partial {n_j} }} &= \rhs{dci_fall_complete}{\frac{k_{0, i} \bar{\theta}_{P_{r, i}, \partial n_j}}{V k_{\infty, i} \left(P_{r, i} + 1\right)} \left(F_{i} \left(P_{r, i} \Theta_{F_i, \partial n_j} + 1\right) - c_{i}\right)} & \cond{dci_fall_complete}{.}
 \end{align}
@@
 \begin{align}
-\lhs{dci_chem_complete}{\frac{\partial c }{\partial {n_j} }_{i}} &= \rhs{dci_chem_complete}{\frac{k_{0, i} \bar{\theta}_{P_{r, i}, \partial n_j} \left(F_{i} \Theta_{F_i, \partial n_j} - c_{i}\right)}{k_{\infty, i} V \left(P_{r, i} + 1\right)}} & \cond{dci_chem_complete}{.}
+\lhs{dci_chem_complete}{\frac{\partial c_{i} }{\partial {n_j} }} &= \rhs{dci_chem_complete}{\frac{k_{0, i} \bar{\theta}_{P_{r, i}, \partial n_j} \left(F_{i} \Theta_{F_i, \partial n_j} - c_{i}\right)}{k_{\infty, i} V \left(P_{r, i} + 1\right)}} & \cond{dci_chem_complete}{.}
 \end{align}
@@
 \subsubsection{Temperature source term derivatives}
+\label{s:dtdotdt_final}
 Starting from~\cref{e:dTdotdT_nojac},~\cref{e:dnkdotdT_compact} is substituted in to yield:

### 2018-02-01 derivations.tex
update

@@
 \subsubsection{Temperature source term derivatives}
-Starting from~\cref{e:dTdotdT_nojac},~\cref{e:dnkdotdT_compact} is substituted in to yield:
+Similarly to~\cref{s:dtdotdt_final},~\cref{e:dnkdotdE_compact} is substituted into~\cref{e:dTdotdE_nojac} to yield:
+\loadeq{dTdotdE}
 

### 2018-02-01 derivations.tex
finish final jacobian section

--- 2018-02-01 derivations.tex
@@
 \begin{align}
- \mathcal{J}_{1,j} = \frac{\partial\dot{T}}{\partial{n_j}}
+ \label{e:dTdotdnj_init}
+ \mathcal{J}_{1,j + 2} = \frac{\partial\dot{T}}{\partial{n_j}}
 \end{align}
@@
 \begin{subequations}
+\label{e:dTdotdnj_nojac}
 \begin{align}
@@
 Similarly, the derivative of the thermodynamic state parameter source terms with respect to themselves are:
+\begin{subequations}
+\label{e:dEdotdE_nojac}
 \begin{align}
@@
 Finally, the derivatives of the thermodynamic state parameter source terms with respect to the moles of species $j$ are:
+\begin{subequations}
+\label{e:dEdotdnj_nojac}
 \begin{align}
@@
   \intertext{or:}
+ \label{e:dnkdotdj_nojac}
  \mathcal{J}_{k + 2, j + 2} & = V \frac{\partial \dot{\omega}_{k} }{\partial {n_j} }
@@
 \begin{align}
+ \label{e:dropi_plog_de}
  \frac{\partial R_{i} }{\partial P } &= \frac{1}{T \Ru} \left(- S^{\prime\prime}_{\ns} {k_{r, i}} + S^{\prime}_{\ns} {k_{f, i}}\right) + & \nonumber \\
@@
 \begin{align}
+ \label{e:dropi_cheb_de}
  \frac{\partial R_{i} }{\partial P } &= \log{\left (10 \right )} \left({R_{f, i}} - {R_{r, i}}\right) \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)} + \nonumber \\
@@
 \begin{subequations}
+ \label{e:dci_chem_de_final}
  \begin{align}
@@
 \subsubsection{Molar source term derivatives}
+\label{s:dtdotdnj}
 From~\cref{e:dnkdotdT_compact} and~\cref{e:dwdot_dT}, a general form for the molar source term derivative with respect to temperature is obtained:
@@
 \begin{align}
- \lhs{dnkdotdT}{\mathcal{J}_{k + 2, 1}} &\pluseq  \rhs{dnkdotdT}{V \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial T } + \nu_{k,i} \frac{\partial R_{i}}{\partial T } c_{i}\right)} & \mathcond{dnkdotdT}{k = 1, \ldots, \ns}
+ \lhs{dnkdotdT}{\mathcal{J}_{k + 2, 1}} &\pluseq  \rhs{dnkdotdT}{V \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial T } + \nu_{k,i} \frac{\partial R_{i}}{\partial T } c_{i}\right)} & \mathcond{dnkdotdT}{k = 1, \ldots, \ns - 1}
 \end{align}
@@
 \subsubsection{Temperature source term derivatives}
-Starting from~\cref{e:dTdotdT_nojac},~\cref{e:dnkdotdT_compact} is substituted in to yield:
+Similar to~\cref{s:dtdotdt_final},~\cref{e:dnkdotdE_compact} is substituted into~\cref{e:dTdotdE_nojac} to yield:
+\loadeq{dTdotdE}
+
+\subsubsection{State parameter source term derivatives}
+Next,~\cref{e:dEdotdE_nojac} is simplified via~\cref{e:dnkdotdE_compact,e:dTdotdE_initial} to give:
+\loadeq{dEdotdE}
+
+\subsubsection{Molar source term derivatives}
+To obtain a final form for the molar source term derivative with respect to the state paramter,~\cref{e:dnkdotdE_compact,e:dwdot_de} are combined:
+\loadeq{dnkdotdE}
+
+The derivative of the net rate of progress with respect to the state paramter is given by~\cref{e:dropi_de} for arrhenius-based reactions,~\cref{e:dropi_plog_de,e:dropi_cheb_de} for P-Log and Chebyshev reactions for a \conv~system, and~\cref{e:dropi_de} for a \conp~system.
+The pressure modification term derivatives are given by~\cref{e:dci_thd_de,e:dci_thd_unity_de,e:dci_thd_spec_de} for third-body enhanced reactions, and~\cref{e:dci_fall_de_final,e:dci_chem_de_final} for falloff\slash chemically-activated reactions respectively.
+
+\subsection{Molar derivatives}
+Finally, this section details the Jacobian entries corresponding to derivatives with respect to the species moles.
+
+\subsubsection{Temperature source term derivatives}
+The temperature source term derivative with respect to the moles of species $j$, \cref{e:dTdotdnj_nojac}, is combined with~\cref{e:dnkdotdj_nojac}, giving:
+\loadeq{dTdotdnj}
+
+\subsubsection{State parameter derivatives}
+From~\cref{e:dEdotdnj_nojac,e:dTdotdnj_init,e:dnkdotdj_nojac}, the final form of the state parameter source term derivative with respect to the moles of species $j$ is:
+\loadeq{dEdotdnj}
+
+\subsubsection{Molar source term derivatives}
+Finally,~\cref{e:dnkdotdj_nojac,e:dwdot_dnj} are combined yielding:
+\loadeq{dnkdotdnj}
+As in~\cref{s:dtdotdnj}, this is converted to an form that may be updated per-reaction $i$:
+\begin{align}
+ \lhs{dnkdotdnj}{\mathcal{J}_{k + 2, j + 2}} &\pluseq \rhs{dnkdotdnj}{V \left(\nu_{k,i} R_{i} \frac{\partial c_{i} }{\partial {n_j} } + \nu_{k,i} \frac{\partial R_{i} }{\partial {n_j} } c_{i}\right)} & \mathcond{dnkdotdnj}{k = 1, \ldots, \ns - 1}
+\end{align}
+
+\pagebreak
+\section{References}
 

### 2018-02-02 derivations.tex
clarify k0 vs kinf for fall / chem, update form for last jac

@@
 
-\subsection{Falloff reactions}
-Unlike elementary and third-body reactions, falloff reactions exhibit a pressure dependence described as a blending of rates at low- and high-pressure limits; thus, the rate coefficients depend on a mixture of low-pressure ($k_{0, i}$) and high-pressure-limit ($k_{\infty,i}$) coefficients, each with corresponding Arrhenius parameters and expressed using \cref{e:arrhenius}.
+\subsection{Falloff and chemically-activated reactions}
+Unlike elementary and third-body reactions, falloff\slash chemically-activated reactions exhibit a pressure dependence described as a blending of rates at low- and high-pressure limits; thus, the rate coefficients depend on a mixture of low-pressure ($k_{0, i}$) and high-pressure-limit ($k_{\infty,i}$) coefficients, each with corresponding Arrhenius parameters and expressed using \cref{e:arrhenius}.
+It is noted that when the forward rate coefficient is used---e.g., as in~\cref{e:ropf}---$k_{f, i}$ is taken to be the high-pressure reaction coefficient ($k_{\infty,i}$) for falloff reactions and the low-pressure reaction coefficent ($k_{0,i}$) for chemically-activated reactions.
 The ratio of the coefficients $k_{0, i}$ and $k_{\infty, i}$, combined with the third-body concentration, define a reduced pressure $P_{r,i}$ given by
@@
 \begin{equation}
+ \label{e:dropi_dnj}
  \frac{\partial R_{i} }{\partial {n_j} } = - \frac{{k_{r, i}}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right) + \frac{{k_{f, i}}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
@@
 \begin{align}
+  \label{e:dci_thd_unity_dnj}
   \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial n_{j} }} &= \rhs{dci_thd_unity}{0}[c] & \cond{dci_thd_unity}{,}
@@
 \begin{align}
+\label{e:dci_chem_dnj_final}
 \lhs{dci_chem_complete}{\frac{\partial c_{i} }{\partial {n_j} }} &= \rhs{dci_chem_complete}{\frac{k_{0, i} \bar{\theta}_{P_{r, i}, \partial n_j} \left(F_{i} \Theta_{F_i, \partial n_j} - c_{i}\right)}{k_{\infty, i} V \left(P_{r, i} + 1\right)}} & \cond{dci_chem_complete}{.}
@@
 \loadeq{dnkdotdT}
-As \texttt{pyJac} evaluates these derivative entries by looping over the reactions $i$ in the model, we convert~\cref{e:dnkdotdT} to an update form:
-\begin{align}
- \lhs{dnkdotdT}{\mathcal{J}_{k + 2, 1}} &\pluseq  \rhs{dnkdotdT}{V \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial T } + \nu_{k,i} \frac{\partial R_{i}}{\partial T } c_{i}\right)} & \mathcond{dnkdotdT}{k = 1, \ldots, \ns - 1}
-\end{align}
 where the rate of progress derivatives are given by~\cref{e:drop_dT,e:drop_plog_dT,e:drop_cheb_dT} for arrhenius-based---i.e., elementary, third-body and falloff\slash chemically-activated---reactions, P-Log reactions and Chebyshev reactions respectively.
@@
 \loadeq{dnkdotdnj}
-As in~\cref{s:dtdotdnj}, this is converted to an form that may be updated per-reaction $i$:
+As \texttt{pyJac} evaluates these derivative entries by looping over the reactions $i$ in the model, \cref{e:dnkdotdnj} is converted to an update form:
 \begin{align}
@@
 \end{align}
+The molar derivatives of the net rate of progress are given by~\cref{e:dropi_dnj}, and the pressure modification term by~\cref{e:dci_thd_dnj,e:dci_thd_spec_dnj,e:dci_thd_unity_dnj} for third-body enhanced reactions, and~\cref{e:dci_fall_dnj_final,e:dci_chem_dnj_final} for falloff\slash chemically-activated reactions, respectively.
 

### 2018-02-06 derivations.tex
update

@@
 \usepackage{amsmath,eqparbox,xparse}
+
 \usepackage[breaklinks=true, linkcolor=blue, citecolor=blue, colorlinks=true]{hyperref}
@@
 \subsection{Reaction rate expressions}
+\label{s:rop_net}
 The net species production rate for species $k$ is defined as:
@@
 \subsection{Falloff and chemically-activated reactions}
+\label{s:falloff}
 Unlike elementary and third-body reactions, falloff\slash chemically-activated reactions exhibit a pressure dependence described as a blending of rates at low- and high-pressure limits; thus, the rate coefficients depend on a mixture of low-pressure ($k_{0, i}$) and high-pressure-limit ($k_{\infty,i}$) coefficients, each with corresponding Arrhenius parameters and expressed using \cref{e:arrhenius}.
-It is noted that when the forward rate coefficient is used---e.g., as in~\cref{e:ropf}---$k_{f, i}$ is taken to be the high-pressure reaction coefficient ($k_{\infty,i}$) for falloff reactions and the low-pressure reaction coefficent ($k_{0,i}$) for chemically-activated reactions.
+It is noted that when the forward rate coefficient is used---e.g., as in~\cref{e:ropf}---$k_{f, i}$ is taken to be the high-pressure reaction coefficient ($k_{\infty,i}$) for falloff reactions and the low-pressure reaction coefficent ($k_{0,i}$) for chemically-activated reactions~\cite{Goodwin:2015aa}.
 The ratio of the coefficients $k_{0, i}$ and $k_{\infty, i}$, combined with the third-body concentration, define a reduced pressure $P_{r,i}$ given by
@@
 $\mathcal{J}$ is filled with the partial derivatives $\partial f / \partial \Phi$, such that:
-\begin{equation}
- \label{e:jac_general}
- \mathcal{J}_{i,j} = \frac{\partial f_i}{\partial \Phi_j},\qquad j=1 \ldots \ns - 1
-\end{equation}
+\loadeq{jacgeneral}
 where $i$ and $j$ correspond to the row and column, respectively, of the entry in the Jacobian matrix.
-The Jacobian entries resulting from~\cref{e:jac_general} are derived in~\cref{s:energy_derivs,s:molar_source_derivatives,s:parameter_source_derivatives}, while various subcomponents of the Jacobian are derived in~\cref{s:dri_dt,s:dri_dnj,s:dri_de,s:dri_pdep,s:dpmod_fall}.
+The Jacobian entries resulting from~\cref{e:jac_general} are derived in~\cref{s:energy_derivs,s:molar_source_derivatives,s:parameter_source_derivatives}, while various subcomponents of the Jacobian are derived in~\cref{s:dri_dt,s:dri_dnj,s:dri_de,s:dri_pdep,s:dci}.
 The final form of the Jacobian can be found in~\cref{s:jac_final}, which provides a useful summary for this document.
@@
  \intertext{or more simply:}
+ \label{e:dTdotdT_initial}
  \mathcal{J}_{1,1} &= \frac{\partial\dot{T}}{\partial{T}}
@@
 \subsection{Rate of progress derivatives}
+\label{s:dropi}
 \subsubsection{Temperature derivative}
@@
 \subsection{Pressure modification\slash Falloff function derivatives}
-\label{s:dpmod_fall}
+\label{s:dci}
 \subsubsection{Elementary reactions}
@@
 \subsubsection{State parameter source term derivatives}
-Similarly,~\cref{e:dnkdotdT_compact} is substituted into~\cref{e:dEdotdT_nojac}, giving:
+Similarly,~\cref{e:dnkdotdT_compact,e:dTdotdT_initial} are substituted into~\cref{e:dEdotdT_nojac}, giving:
 \loadeq{dEdotdT}

### 2018-02-06 paper.tex
update

@@
 \usepackage{gensymb}
-\usepackage{amsmath}
 \usepackage[version=3]{mhchem} % Formula subscripts using \ce{}, e.g., \ce{H2SO4}
@@
 }
+
+% external references https://tex.stackexchange.com/a/14365/56227
+\usepackage{xr-hyper}
+
+\usepackage{amsmath,eqparbox,xparse}
+
+\usepackage[breaklinks=true, linkcolor=blue, citecolor=blue, colorlinks=true]{hyperref}
+% load after hyperref
+\usepackage[english,capitalise]{cleveref}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%      Custom Commands     %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% https://tex.stackexchange.com/a/34412/5764
+\makeatletter
+\NewDocumentCommand{\eqmathbox}{o O{c} m}{%
+  \IfValueTF{#1}
+    {\def\eqmathbox@##1##2{\eqmakebox[#1][#2]{$##1##2$}}}
+    {\def\eqmathbox@##1##2{\eqmakebox{$##1##2$}}}
+  \mathpalette\eqmathbox@{#3}
+}
+\makeatother
 
@@
 }
-\usepackage[breaklinks=true, linkcolor=blue, citecolor=blue, colorlinks=true]{hyperref}
-% load after hyperref
-\usepackage[english,capitalise]{cleveref}
+
+% load external document after hyperref https://tex.stackexchange.com/a/281417/56227
+\externaldocument[deriv-]{derivations}
 
@@
 \newcommand{\Ru}{\mathcal{R}}
+
+% simple command for equal-sized eqmathbox's on the lhs / rhs and conditional
+% arguements are tag, text, alignment
+\NewDocumentCommand{\lhs}{mmO{r}}{\eqmathbox[LHS_#1][#3]{#2}}
+\NewDocumentCommand{\rhs}{mmO{l}}{\eqmathbox[RHS_#1][#3]{#2}}
+\NewDocumentCommand{\cond}{mmO{r}}{\eqmakebox[COND_#1][#3]{#2}}
+\NewDocumentCommand{\mathcond}{mmO{r}}{\eqmathbox[COND_#1][#3]{#2}}
+
+% consistent multiline equation spacing
+\NewDocumentCommand{\mathindent}{O{space}}{\hphantom{\mathrel{#1}}}
 
@@
 When storing arrays for a chemical kinetic model, the data-storage layout and vectorization patterns are critical to achieving high-performance code.
-\Cref{F:mem} depicts an example data array with K-rows and N-columns and index (i, j) corresponding to the $i$th row and $j$th column.
+\Cref{F:mem} depicts an example data array with K-rows and N-columns and index ($i$, $j$) corresponding to the $i$th row and $j$th column.
 For example the concentration of species $j$ for the $i$-th thermo-chemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{state}$, the number of thermo-chemical states considered, and $1 \le j \le \ns$, the number of species in the model.
@@
 where $T$ is the temperature, $V$ and $P$ the volume and pressure respectively, and $n_j$ the number of moles of the $j$th species in the model (containing $\ns$ total species).
+The evolution of the thermo-chemical state vector is described by a set of chemical kinetic ordinary differential equations:
+\loadeq{dstate}
 
@@
 \loadeq{dmolar}
-where $\dot{\omega}_k$ is the $kth$ species' overall production rate, and the temperature production rate~\cite{TurnsStephenR2012Aitc} is:
+where $\dot{\omega}_k$ is the $kth$ species' overall production rate:
+\begin{equation}
+ \dot{\omega}_{k} = \sum_{i=1}^{\nr} \nu_{k, i} R_{i} c_{i}
+\end{equation}
+where $\nu_{k, i}$ is the net stoichiometic coefficient of species $k$ in reaction $i$, $R_{i}$ the net rate of progress of reaction $i$ (described in~\cref{deriv-s:rop_net} of the supplimentary material), and $c_{i}$ the pressure modification term i.e., for third-body (see~\cref{deriv-s:thdbody} of the supplimental), or fall-off\slash chemically-activated reactions (\cref{deriv-s:falloff} of the same).
+
+The temperature source term~\cite{TurnsStephenR2012Aitc} is:
 \loadeq{dtemp}
@@
 
-\subsection{Code Generation}
-Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates user specified psuedo-code and data to OpenCL, allowing for unit testing and easy changes of program structure, e.g.
-data ordering, vectorization, threading patterns etc.
+The Jacobian entries computed by \texttt{pyJac} are arranged such that entry ($i$, $j$) corresponds to the partial derivative of the $i$-th source term in~\cref{e:source_terms} by the $j$-th state variable in~\cref{e:state}:
+\loadeq{jacgeneral}
+
+The derivative of the temperature source term with respect to itself is:
+\loadeq{dTdotdT}
+and with respect to the thermodynamic state parameter (the volume \slash pressure for \conp\slash\conv~respectively):
+\loadeq{dTdotdE}
+and the moles of species $j$:
+\loadeq{dTdotdnj}
+
+The derivative of the thermodynamic state parameter with respect to the temperature is:
+\loadeq{dEdotdT}
+and with respect to itself:
+\loadeq{dEdotdE}
+and finally, by the moles of species $j$:
+\loadeq{dEdotdnj}
+
+The molar source terms are more complicated due to the derivatives of the net rate of progress---see~\cref{deriv-s:dropi}---and pressure modification terms---\cref{deriv-s:dci}---for the reactions.
+The reader is again encouraged to read the supplimental material covering the exact derivations and forms of the reactions solved by \texttt{pyJac}.
+The derivative of the molar source term with respect to temperature is:
+\loadeq{dnkdotdT}
+while, similarly the derivative with respect to the state parameter is:
+\loadeq{dnkdotdE}
+and the moles of species $j$:
+\loadeq{dnkdotdnj}
+
+\subsection{Code Generation and Testing Infrastructure}
+Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates user specified psuedo-code and data to OpenCL\slash C-code.
+This allows for easy changes of program structure, e.g.
+data ordering, vectorization, threading patterns and easy switching of target-languages (and simpler extension to additional languages, e.g., CUDA).
+Additionally, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, and via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source terms.
+The source-terms and subcomponents thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the autodifferentiation code Adept~\cite{adept-v11} is used to obtain reference answers for Jacobian subcomponents.
+Validation of the complete source-terms and Jacobian codes will be discussed in detail in~\cref{s:validation}.
 
 \section{Results and Discussion}
-\subsection{Comparison to Cantera}
+\subsection{Source term validation}
+\label{s:validation}
 The reaction rates of progress (ROP), species and temperature rates in this study are validated by comparison with Cantera~\cite{Cantera}, however special care must be taken due floating point arithmetic issues.

### 2018-02-06 plotting_scripts/general_plotting.py
first pass at using NCM scripts & new data

@@ -1,125 +1,0 @@
-"""Module with
-"""
-import numpy as np
-import matplotlib.pyplot as plt
-
-font_size = 'large'
-
-legend_key = {'H2':r'H$_2$/CO',
-              'CH4':r'GRI-Mech 3.0',
-              'C2H4':r'USC-Mech II',
-              'IC5H11OH':r'IC$_5$H$_{11}$OH'
-              }
-
-def plot(plotdata, marker, name, minx=None, miny=None, plot_std=True,
-         norm=None, return_y=False, reacs_as_x=True, color=None, hollow=False
-         ):
-    """Plot performance as a function of reaction count.
-    """
-    if reacs_as_x:
-        plotdata = sorted(plotdata, key=lambda x: x.num_reacs)
-        x_vals = [x.num_reacs for x in plotdata]
-    else:
-        plotdata = sorted(plotdata, key=lambda x: x.num_specs)
-        x_vals = [x.num_specs for x in plotdata]
-    if norm is None:
-        y_vals = [np.array(x.y) / x.x for x in plotdata]
-    else:
-        y_vals = [norm(x) for x in plotdata]
-
-    err_vals = [np.std(y) for y in y_vals]
-    y_vals = [np.mean(y) for y in y_vals]
-    argdict = {'x':x_vals,
-               'y':y_vals,
-               'linestyle':'',
-               'marker':marker,
-               'label':name
-               }
-    if color:
-        argdict['markeredgecolor'] = color
-        argdict['color'] = color
-    elif hollow:
-        argdict['markerfacecolor'] = 'None'
-        argdict['label'] += ' (smem)'
-    if plot_std:
-        argdict['yerr'] = err_vals
-    if plot_std:
-        line = plt.errorbar(**argdict)
-    else:
-        line = plt.plot(**argdict)
-
-    minx = (x_vals[0] if minx is None
-            else x_vals[0] if x_vals[0] < minx
-            else minx
-            )
-    miny = (y_vals[0] if miny is None
-            else y_vals[0] if y_vals[0] < miny
-            else miny
-            )
-    retval = minx, miny
-    if return_y:
-        retval = (retval, y_vals, err_vals)
-    return retval
-
-
-def plot_scaling(plotdata, markerlist, colorlist, minx=None, miny=None,
-                 label_locs=None, plot_std=True, hollow=False
-                 ):
-    """Plots performance data with varying number of conditions.
-    """
[…]

### 2018-02-06 plotting_scripts/performance_comparison.py

@@ -1,238 +1,0 @@
-"""Module for performance comparison plotting.
-"""
-from __future__ import print_function
-
-# Standard library
-import sys
-import os
-
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib.backends.backend_pdf import PdfPages
-from scipy import optimize
-
-# Local imports
-from performance_extractor import get_data
-from general_plotting import legend_key, plot
-
-font_size = 'large'
-
-FD_marker = '>'
-pj_marker = 'o'
-tc_marker = 's'
-
-def fr_exp10(x):
-    """Returns base 10 mantissa and exponent of float.
-    """
-    # ensure number
-    assert isinstance(x, (int, float))
-    exp = int(np.log10(x))
-    mantissa = x / 10**exp
-    if mantissa < 1.0:
-        mantissa *= 10.
-        exp -= 1
-    return mantissa, exp
-
-
-def nice_names(x):
-    """Returns data series names for printing.
-    """
-    if x.finite_difference:
-        name = 'Finite Difference'
-        marker = FD_marker
-    elif x.lang == 'tchem':
-        name = 'TChem'
-        marker = tc_marker
-    else:
-        name = 'pyJac'
-        marker = pj_marker
-    if x.lang == 'cuda':
-        name += ' (GPU)'
-    return name, marker
-
-
-def get_fullscale(data):
-    mechanisms = set([x.mechanism for x in data])
-    #need to filter out runs on subsets
-    for mech in mechanisms:
-        max_x = max(x.x for x in [y for y in data if y.mechanism == mech])
-        data = [x for x in data if x.mechanism != mech or
-                (x.x == max_x)
-                ]
-    return data
-
-
-def fit_order(plotdata, y_vals, std, order=None, color='k', text_loc=None):
-    """Gets and plots best-fit line for performance data.
-    """
-    x_vals = sorted([x.num_reacs for x in plotdata])
-    maxx = int(x_vals[-1])
-
-    def fit_func(p, x):
-        n = order
-        const = p[0]
-        if order is None:
-            n = p[0]
-            const = p[1]
-
-        # Need to handle when arguments of log may be >= 0
-        if const > 0.0:
-            return n * np.log10(np.array(x).clip(min=1.e-20)) + np.log10(const)
-        else:
-            return np.inf
-
-    def errfunc(p, x, y):
-        return (fit_func(p, x) - np.log10(y)) # Distance to the target function
-
-    # use specified order of polynomial; otherwise, fit should find this
-    if order is None:
[…]

### 2018-02-06 plotting_scripts/performance_extractor.py

@@ -1,86 +1,0 @@
-"""Module for extracting Jacobian matrix eval performance data.
-"""
-
-# Standard library
-import sys
-import os
-import os.path
-import json
-
-import numpy as np
-
-def parse_file(directory, mech, filename, num_specs, num_reacs):
-    data_arr = []
-    with open(os.path.join(directory, filename)) as file:
-        lines = [line.strip() for line in file.readlines()]
-    data = {}
-    for line in lines:
-        try:
-            x1, y1 = [float(f) for f in line.split(',')]
-            if not x1 in data:
-                data[x1] = []
-            data[x1].append(y1)
-        except:
-            assert line == 'SUCCESS'
-            pass
-    for x in data:
-        data_arr.append(data_point(mech, filename, num_specs,
-                        num_reacs, x, data[x])
-                        )
-    return data_arr
-
-
-class data_point(object):
-    def __init__(self, mechanism, filename, num_specs, num_reacs, x, y):
-        self.mechanism = mechanism
-
-        lang, cache_opt, smem, fd, dummy = filename.split('_')
-        self.lang = lang
-        self.cache_opt = cache_opt == 'co'
-        self.smem = smem == 'smem'
-        self.finite_difference = fd == 'fd'
-
-        self.num_specs = num_specs
-        self.num_reacs = num_reacs
-        self.x = x
-        self.y = y
-
-    def __repr__(self):
-        return self.__str__()
-
-    def __str__(self):
-        return "lang={},\tmech={},\tFD={},\tsmem={}\tco={}\n".format(
-            self.lang, self.mechanism, self.finite_difference,
-            self.smem, self.cache_opt
-            )
-
-
-def get_data(home_dir=None):
-    if home_dir is None:
-        home_dir = os.path.join(sys.path[0], '../')
-        home_dir = os.path.realpath(home_dir)
-    d = os.path.join(home_dir, 'data')
-    dirs = [o for o in os.listdir(d) if os.path.isdir(os.path.join(d,o))]
-
-    #ok, now we have a list of data directories, open each and look for the text files
-    data = []
-    for directory in dirs:
-        thedir = os.path.join(d, directory)
-
-        #get text files with performance data
-        files = [o for o in os.listdir(thedir) if
-                 os.path.isfile(os.path.join(thedir,o)) and o.endswith('.txt')
-                 ]
-
-        # read mechanism info
[…]

### 2018-02-06 plotting_scripts/plot_ch4_pasr_data.py

@@ -1,61 +1,0 @@
-#! /usr/bin/env python
-"""Plots partially stirred reactor results for methane-air.
-"""
-import sys
-import os.path
-
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib.backends.backend_pdf import PdfPages
-
-font_size = 'large'
-
-home_dir = os.path.join(sys.path[0], '../')
-home_dir = os.path.realpath(home_dir)
-fig_dir = os.path.join(home_dir, 'figures')
-
-# Load CH4/air, 600 K results
-state_data = np.load(os.path.join(home_dir, 'data', 'CH4', 'pasr_out_ch4_1.npy'))
-
-# First plot the mean temperature
-times = state_data[:, 0, 0]
-mean_temp = np.zeros(state_data.shape[0])
-for i in range(len(mean_temp)):
-    mean_temp[i] = np.mean(state_data[i, :, 1])
-
-plt.plot(times, mean_temp)
-plt.xlim([0, 0.05])
-plt.ylim([1900, 2400])
-plt.xlabel('Time (s)', fontsize=font_size)
-plt.ylabel('Mean temperature (K)', fontsize=font_size)
-
-pp = PdfPages(os.path.join(fig_dir, 'CH4_600K_1atm_mean_temperature.pdf'))
-pp.savefig()
-pp.close()
-
-plt.close()
-
-# Now plot particle data
-state_data = state_data.reshape(state_data.shape[0] * state_data.shape[1],
-                                state_data.shape[2]
-                                )
-
-
-# Time vs temperature
-
-plt.subplot(211)
-plt.scatter(state_data[:,1], state_data[:,0], s=10, marker='.', c='k')
-plt.ylabel('Time (s)', fontsize=font_size)
-plt.xlim([600, 2400])
-plt.ylim([0, 0.05])
-
-plt.subplot(212)
-n, bins, patches = plt.hist(state_data[:,1], 100,
-                            normed=1, histtype='stepfilled'
-                            )
-plt.ylabel('PDF', fontsize=font_size)
-plt.xlabel('Temperature (K)', fontsize=font_size)
-
-pp = PdfPages(os.path.join(fig_dir, 'CH4_600K_1atm_particle_temperature.pdf'))
-pp.savefig()
-pp.close()

### 2018-02-06 plotting_scripts/plot_cpu_comparison.py

@@ -1,30 +1,0 @@
-#! /usr/bin/env python
-"""Plots comparison of CPU performance for Jacobian matrix evaluations.
-"""
-from __future__ import print_function
-
-import numpy as np
-
-from performance_comparison import fullscale_comp
-
-# Location of label: (data index, x position multiplier, y position multiplier)
-text_loc = [(3, 0.85, 0.8), #fd
-            (3, 1.0, 0.35), #pyjac
-            (2, 1.4, 1.1)  #tchem
-            ]
-fit_params, data = fullscale_comp('c', text_loc=text_loc, loc_override=4)
-
-for x in fit_params:
-    print(x[0] + '\tR^2 : {}'.format(x[1]))
-
-pJ = next(val for val in fit_params if 'pyJac' in val[0])
-tC = next(val for val in fit_params if 'TChem' in val[0])
-fD = next(val for val in fit_params if 'Finite' in val[0])
-
-print('TChem / pyJac performance ratios: ')
-for i in range(len(data[fit_params.index(tC)])):
-    print(data[fit_params.index(tC)][i] / data[fit_params.index(pJ)][i])
-
-print('Finite difference / pyJac performance ratios: ')
-for i in range(len(data[fit_params.index(fD)])):
-    print(data[fit_params.index(fD)][i] / data[fit_params.index(pJ)][i])

### 2018-02-06 plotting_scripts/plot_gpu_comparison.py

@@ -1,25 +1,0 @@
-#! /usr/bin/env python
-"""Plots comparison of GPU performance for Jacobian matrix evaluations.
-"""
-from __future__ import print_function
-
-import numpy as np
-
-from performance_comparison import fullscale_comp
-
-# Location of label: (data index, x position multiplier, y position multiplier)
-text_loc = [(3, 0.85, 0.9), #fd
-            (3, 1.0, 0.55), #pyjac
-            ]
-
-fit_params, data = fullscale_comp('cuda', text_loc=text_loc, loc_override=4)
-
-for x in fit_params:
-    print(x[0] + '\tR^2 : {}'.format(x[1]))
-
-pJ = next(val for val in fit_params if 'pyJac' in val[0])
-fD = next(val for val in fit_params if 'Finite' in val[0])
-
-print('Finite difference / pyJac performance ratios: ')
-for i in range(len(data[fit_params.index(fD)])):
-    print(data[fit_params.index(fD)][i] / data[fit_params.index(pJ)][i])

### 2018-02-06 plotting_scripts/plot_gpu_scaling.py

@@ -1,88 +1,0 @@
-#! /usr/bin/env python
-"""Plots GPU performance scaling data for pyJac Jacobian matrix evaluation.
-"""
-
-import sys
-import os.path
-
-import matplotlib.pyplot as plt
-from matplotlib.backends.backend_pdf import PdfPages
-import numpy as np
-
-# Local imports
-from performance_extractor import get_data
-from general_plotting import plot_scaling
-
-home_dir = None
-font_size = 'large'
-
-legend_markers = ['o', 'v', 's', '>']
-legend_colors = ['b', 'g', 'r', 'c']
-
-# x position of text label, and multiplier for y position
-label_locs = [(10, 1.75),
-              (10, 0.6),
-              (10, 1.75),
-              (10, 1.75)
-              ]
-
-
-if home_dir is None:
-    home_dir = os.path.join(sys.path[0], '../')
-    home_dir = os.path.realpath(home_dir)
-d = os.path.join(home_dir, 'figures')
-
-# Get CUDA pyJac datapoints, without cache optimization or shared memory
-data = get_data()
-plotdata = [x for x in data if x.lang == 'cuda'
-            and not x.cache_opt
-            and not x.smem
-            and not x.finite_difference
-            ]
-
-fig, ax = plt.subplots()
-minx, miny = plot_scaling(plotdata, legend_markers, legend_colors,
-                          label_locs=label_locs
-                          )
-ax.set_yscale('log')
-ax.set_xscale('log')
-ax.set_ylim(ymin=miny*0.85)
-ax.set_xlim(xmin=minx*0.85)
-#ax.legend(loc=0, numpoints=1, frameon=False)
-# add some text for labels, title and axes ticks
-ax.set_ylabel('Mean evaluation time', fontsize=font_size)
-ax.set_xlabel('Number of conditions', fontsize=font_size)
-
-pp = PdfPages(os.path.join(d, 'gpu_performance_scaling.pdf'))
-pp.savefig()
-pp.close()
-
-plt.close()
-
-
-# Get CUDA pyJac datapoints with shared memory
-plotdata = [x for x in data if x.lang == 'cuda'
-            and not x.cache_opt
-            and x.smem
-            and not x.finite_difference
-            ]
-
-fig, ax = plt.subplots()
-minx, miny = plot_scaling(plotdata, legend_markers, legend_colors,
[…]

### 2018-02-06 scripts/data_parser.py

@@ -1,0 +1,47 @@
+import collections
+import os
+import yaml
+
+rundata = collections.namedtuple(
+    'data', 'num_conditions comptime overhead runtime')
+mechdata = collections.namedtuple('mechdata', 'mech num_species num_reactions')
+run = collections.namedtuple(
+    'run',
+    'type sparse lang vecwidth order vectype platform rates kernel cores conp'
+    'rundata mechdata')
+
+
+def parse_data(directory='performance'):
+    # find all mechs
+    mechs = [x for x in os.listdir(directory)
+             if os.path.isdir(os.path.join(directory, x))]
+    perf_data = {}
+    for mech in mechs:
+        perf_data[mech] = []
+        path = os.path.join(directory, mech)
+        # get files
+        files = [os.path.join(path, x) for x in os.listdir(path)]
+        files = [x for x in files if os.path.isfile(x)]
+        # read mechanism info
+        mech_meta_file = next(x for x in files if x.endswith('.yaml'))
+        with open(mech_meta_file, 'r') as f:
+            mech_info = mechdata(**yaml.load(f))
+        files.remove(mech_meta_file)
+        for file in files:
+            # get name
+            name = os.path.basename(file)
+            # get run parameters
+            params = name[:name.index('.txt')].split('_')
+            if params[0] == 'spec':
+                # insert 'sparse' type
+                params.insert('sparse', 1)
+            # get data
+            with open(file, 'r') as f:
+                lines = [[float(x) for x in l.strip().split(',')]
+                         for l in f.readlines() if l.strip()]
+            datalist = [rundata(*x) for x in lines]
+            # add to run
+            params.extend([datalist, mech_info])
+            perf_data[mech].append(run(*params))
+
+    return perf_data

### 2018-02-06 scripts/error.py

@@ -1,0 +1,102 @@
+import os
+import numpy as np
+import yaml
+
+script_dir = os.path.dirname(os.path.abspath(__file__))
+error_path = os.path.join(script_dir, 'error_checking')
+mechs = [os.path.join(error_path, x) for x in os.listdir(
+    error_path) if os.path.isdir(os.path.join(error_path, x))]
+
+rtol = 1e-6
+atol = 1e-10
+
+err_dicts = {}
+for mech in mechs:
+    mech_name = os.path.basename(os.path.normpath(mech))
+    # get file list
+    files = [os.path.join(mech, x) for x in os.listdir(
+        mech) if os.path.isfile(os.path.join(mech, x)) if x.endswith('.npz')
+        and 'spec' in x]
+    err_dicts[mech_name] = {}
+
+    def __get_size(name):
+        if 'rop' in name:
+            if 'net' in name or 'fwd' in name:
+                return mech_info['n_reactions']
+            else:
+                return mech_info['n_reversible']
+        elif 'wdot' in name:
+            return mech_info['n_species']
+        elif 'phi' in name:
+            return mech_info['n_species'] + 1
+
+    # get mech info
+    mech_info = os.path.join(mech, mech_name.lower() + '.yaml')
+    with open(mech_info, 'r') as file:
+        mech_info = yaml.load(file.read())
+
+    for file in files:
+        arrs = np.load(file)
+        for name in arrs:
+            if 'value' in name or 'component' in name or 'store' in name:
+                continue
+            errs = arrs[name]
+            values = arrs[name + '_value']
+            errs = errs / (atol + rtol * np.abs(values))
+
+            precs = None
+            if 'rop_net' in name:
+                # calculate the precision norms
+                precs = arrs['rop_component'] / (atol + rtol * np.abs(values))
+
+            if name not in err_dicts[mech_name]:
+                err_dicts[mech_name][name] = np.zeros((__get_size(name)))
+
+                if precs is not None:
+                    err_dicts[mech_name]['rop_component'] = np.zeros((__get_size(
+                        'rop_net')))
+
+            if name in err_dicts[mech_name]:
+                if errs.shape != err_dicts[mech_name][name].shape:
+                    # check that we have a split
+                    assert 'C_w' in file or 'F_d' in file
+                    # discard extra zeros resulting from split padding
+                    errs = errs[:__get_size(name)]
+
[…]

### 2018-02-06 scripts/general_plotting.py

@@ -1,0 +1,154 @@
+"""Module with
+"""
+from __future__ import division
+import numpy as np
+import matplotlib.pyplot as plt
+import plot_styles as ps
+
+font_size = 'large'
+
+def process_data(plotdata, plot, reacs_as_x=True,
+    plot_cores=False):
+    """
+    Process the data into an easily usable form
+    """
+    if reacs_as_x:
+        plotdata = sorted(plotdata, key=lambda x: x.mechdata.num_reactions)
+        x_vals = [x.mechdata.num_reactions for x in plotdata]
+    elif plot_cores:
+        plotdata = sorted(plotdata, key=lambda x: float(x.cores))
+        x_vals = [float(x.cores) for x in plotdata]
+    else:
+        plotdata = sorted(plotdata, key=lambda x: x.mechdata.num_species)
+        x_vals = [x.mechdata.num_species for x in plotdata]
+
+    y_vals = []
+    err_vals = []
+    for run in plotdata:
+        ys = np.array([getattr(data, plot) for data in run.rundata])
+        zs = np.std(ys)
+        if plot == 'runtime':
+            ys = np.array([getattr(data, plot) / data.num_conditions for data in run.rundata])
+            zs /= run.rundata[0].num_conditions
+        err_vals.append(zs)
+        y_vals.append(np.mean(ys))
+    return x_vals, y_vals, err_vals
+
+
+def plot(plot, x_vals, y_vals, err_vals, minx=None, miny=None, maxx=None, maxy=None, plot_std=True,
+        return_y=False, labels=[], plot_ind=None, marker_func=None):
+    """Plot performance as a function of reaction count.
+    """
+
+    if marker_func is not None:
+        name = labels[plot_ind]
+        marker, hollow, color = marker_func(name)
+    elif plot_ind is not None:
+        assert labels
+        marker, hollow = ps.marker_wheel[plot_ind]
+        color = ps.color_wheel[plot_ind]
+        name = labels[plot_ind]
+    else:
+        marker, hollow = ps.marker_dict[plot]
+        color = ps.color_dict[plot]
+        name = plot
+
+    argdict = {'x':x_vals,
+               'y':y_vals,
+               'linestyle':'',
+               'marker':marker,
+               'label':ps.pretty_names(name)
+               }
+    argdict['color'] = color
+    argdict['markeredgecolor'] = color
+    if not hollow:
+        argdict['markersize'] = ps.marker_style['size']
+    else:
[…]

### 2018-02-06 scripts/perf_plotter.py

@@ -1,0 +1,300 @@
+from __future__ import division
+
+import data_parser
+import argparse
+import plot_styles as ps
+import matplotlib.pyplot as plt
+import general_plotting as gp
+import numpy as np
+import itertools
+import functools
+import copy
+
+run = data_parser.run
+rundata = data_parser.rundata
+data_clean = data_parser.parse_data()
+reacs_as_x = False
+norm = None
+
+#from http://stackoverflow.com/a/31174427
+sentinel = object()
+def rgetattr(obj, attr, default=sentinel):
+    if default is sentinel:
+        _getattr = getattr
+    else:
+        def _getattr(obj, name):
+            return getattr(obj, name, default)
+    return functools.reduce(_getattr, [obj]+attr.split('.'))
+
+def __compare(r, name, compare_value, plot_cores=False):
+    """
+    Specialty comparison function to account for differences
+    in runtypes
+    """
+    if name == 'vecwidth' and r.vectype == 'par':
+        return True
+    if plot_cores and name == 'cores':
+        return True
+    return rgetattr(r, name) == compare_value
+
+def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
+        legend_handler=None, marker_func=None,
+        minx=None, miny=None, maxx=None, maxy=None, ylog=False,
+        return_vals=False, **filters):
+    data = copy.copy(data_clean)
+    #create fig, ax
+    fig = plt.figure()
+    ax = plt.subplot(1,1,1)
+    #apply filters
+    for f in filters:
+        if filters[f] is None:
+            continue
+        assert any(data[x] for x in data), 'No data matching all filters'
+        if f in data[list(data.keys())[0]][0]._asdict():
+            for mech in data:
+                data[mech] = [x for x in data[mech] if __compare(x, f, filters[f])]
+    #compute data
+    for mech in data:
+        for run in data[mech]:
+            rund = run.rundata
+    #now plot data
+    to_plot = ['runtime']
+    if filters.pop('plot_compilation', False):
+        to_plot.append('comptime')
+    if filters.pop('plot_overhead', False):
+        to_plot.append('overhead')
+    #get data
+    plot_data = [x for mech in data for x in data[mech]]
+    #get labels
+    diff_check = [x for x in run._fields if x not in ['mechdata', 'rundata']]
+    diffs = [set([getattr(x, check) for x in plot_data]) for check in diff_check]
+    #get # with more than 1 option
[…]

### 2018-02-06 scripts/plot_styles.py

@@ -1,0 +1,88 @@
+#a single consolidated place to import
+#such that all figures have identical styling (when possible)
+
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import os.path
+
+#setup latex
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
+plt.rc('text.latex',
+    preamble=r'\usepackage{amsmath},\usepackage{siunitx},\usepackage[version=3]{mhchem}')
+plt.rc('font', family='serif')
+
+legend_style = {'loc':0,
+    'fontsize':18,
+    'numpoints':1,
+    'shadow':True,
+    'fancybox':True
+}
+
+tick_font_size = 20
+
+marker_style = {
+    'size' : 15
+}
+
+clear_marker_style = {
+    'size' : 17
+}
+
+default_keys = ['runtime', 'comptime', 'overhead']
+marker_wheel = [('o', True), ('>', True), ('s', True),
+    ('D', True)]
+marker_dict = {x : marker_wheel[i] for i, x in enumerate(default_keys)}
+
+class dummy_formatter(object):
+    def __init__(self, choicedict):
+        self.choices = choicedict.keys()
+        self.values = [choicedict[x] for x in self.choices]
+
+    def __hash__(self):
+        return hash(repr(self))
+
+    def format(self, val):
+        return next(self.values[i] for i, x in enumerate(self.choices)
+            if x == val)
+
+legend_key = {'H2':r'H$_2$/CO',
+              'CH4':r'GRI-Mech 3.0',
+              'C2H4':r'USC-Mech II',
+              'IC5H11OH':r'IC$_5$H$_{11}$OH'
+              }
+
+def pretty_names(pname, short=False):
+    pname_dict = {'runtime' : 'Runtime',
+        'comptime' : 'Compilation time',
+        'overhead' : 'Kernel Construction Overhead',
+        'vecwidth' : 'Vector Width = {}',
+        'vectype' : dummy_formatter({'w' : 'Shallow SIMD',
+            'par' : 'SIMT',
+            'd' : 'Deep SIMD'}),
+        'order' : dummy_formatter({'C' : 'C-order',
+            'F' : 'F-order'}),
+        'kernel' : dummy_formatter({'single' : 'Single Rate Kernel',
+            'split' : 'Split Rate Kernels'}),
+        'rates' : dummy_formatter({'fixed' : 'Fixed Rate Specialization',
[…]

### 2018-02-06 scripts/scaling_plotter.py

@@ -1,0 +1,9 @@
+import perf_plotter as pp
+
+pp.plotter('../SIMD_scaling.pdf', show=False,
+    order='C', vectype='w', vecwidth='16', rates='hybrid', kernel='split',
+    minx=0)
+
+pp.plotter('../SIMT_scaling.pdf', show=False,
+    order='C', vectype='par', vecwidth='16', rates='hybrid', kernel='split',
+    minx=0)

### 2018-02-06 scripts/simd_and_order_comp.py

@@ -1,0 +1,60 @@
+import perf_plotter as pp
+import plot_styles as ps
+import matplotlib.patches as mpatches
+import matplotlib.lines as mlines
+
+SIMD_color = ps.color_wheel[0]
+SIMT_color = ps.color_wheel[1]
+SIMT_sym = ps.marker_wheel[0][0]
+SIMD_sym = ps.marker_wheel[1][0]
+C_hollow = False
+F_hollow = True
+
+def marker_func(name):
+    retval = []
+    if 'SIMT' in name:
+        retval.append(SIMT_sym)
+    else:
+        retval.append(SIMD_sym)
+    if 'C-order' in name:
+        retval.append(C_hollow)
+    else:
+        retval.append(F_hollow)
+    if 'SIMT' in name:
+        retval.append(SIMT_color)
+    else:
+        retval.append(SIMD_color)
+    return tuple(retval)
+
+labels = ['SIMT', 'SIMD', 'C-order (Filled)', 'F-order (Hollow)']
+handles = [mlines.Line2D([], [], color=SIMT_color, marker=SIMT_sym, linestyle='',
+                          markersize=ps.marker_style['size'], label='SIMT'),
+           mlines.Line2D([], [], color=SIMD_color, marker=SIMD_sym, linestyle='',
+                          markersize=ps.marker_style['size'], label='SIMD'),
+           mlines.Line2D([], [], linestyle='', marker=None, label='C-order (Filled)'),
+           mlines.Line2D([], [], linestyle='', marker=None, label='F-order (Hollow)')]
+legend = (handles, labels)
+
+xval, yval, zval, label = pp.plotter('../SIMD_SIMT_Order_comparison.pdf', show=False,
+    cores='1', vecwidth='8', rates='hybrid', kernel='split',
+    legend_handler=legend, marker_func=marker_func, return_vals=True,
+    norm=False, ylog=True)
+
+#print stats
+C_order = [i for i,l in enumerate(label) if 'C-order' in l]
+F_order = [i for i in range(len(label)) if i not in C_order]
+
+C_SIMT = next(i for i in C_order if 'SIMT' in label[i])
+C_SIMD = next(i for i in C_order if i != C_SIMT)
+F_SIMT = next(i for i in F_order if 'SIMT' in label[i])
[…]

### 2018-02-06 scripts/vecwidth_plotter.py

@@ -1,0 +1,4 @@
+import perf_plotter as pp
+
+pp.plotter('../SIMD_Vecwidth_comparison.pdf', show=False,
+    cores='1', order='C', vectype='w', rates='hybrid', kernel='split')

### 2018-02-06 paper.tex
update table

@@
 
+\todo[inline]{update species error with new data -- in particular the \num{e-4} for i-pentanol are all from november. preliminary data on the GPU shows ~\num{4e-8}}
 \begin{table}[htb]
@@
 \centering
-\begin{tabular}{@{}S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] @{}}
+\begin{tabular}{@{}l S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1]@{}}
 \toprule
-\multicolumn{1}{l}{Model} & \multicolumn{1}{c}{$E_{R^{\prime}}$} & \multicolumn{1}{c}{$E_{R^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R}$} & \multicolumn{1}{c}{$E_{\varepsilon}$} & \multicolumn{1}{c}{$E_{\dot{\omega}}$} & \multicolumn{1}{c}{$E_{\frac{\partial T}{\partial t}}$} \\
+\multicolumn{1}{l}{Model} & \multicolumn{1}{l}{\ce{H2}\slash~\ce{CO}\cite{Burke:2011fh}} & \multicolumn{1}{l}{GRI-Mech.~3.0\cite{smith_gri-mech_30}} & \multicolumn{1}{l}{USC-Mech II\cite{Wang:2007}} & \multicolumn{1}{l}{\ce{iC5H11OH}\cite{Sarathy:2013jr}} \\
 \midrule
-\multicolumn{1}{l}{\ce{H2}\slash~\ce{CO}\cite{Burke:2011fh}} & 1.20e-8 & 6.33e-8 & 1.32e1 & 1.32e1 & 2.37e1 & 2.12e5 \\
-\multicolumn{1}{l}{GRI-Mech.~3.0\cite{smith_gri-mech_30}}  & 3.11e-8 & 5.90e-8 & 1.21e0 & 1.29e0 & 2.84e0 & 2.64e4 \\
-\multicolumn{1}{l}{USC-Mech II\cite{Wang:2007}}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 2.27e0 & 7.92e0 & 3.22e03 \\
-\multicolumn{1}{l}{\ce{iC5H11OH}\cite{Sarathy:2013jr}} & 6.80e-8 &  1.48e-7 & 5.26e-1 & 4.87e-1 & 2.26e00 & 6.10e+01 \\
+\multicolumn{1}{c}{$E_{R^{\prime}}$} & 1.61e-8 & 2.95e-8 & 1.04e-7 & 4.86e-4 \\
+\multicolumn{1}{c}{$E_{R^{\prime\prime}}$} & 6.93e-8 & 6.43e-8 & 1.20e-7 & 5.07e-4 \\
+\multicolumn{1}{c}{$E_{R}$}  & 1.49e1 & 1.17e0 & 2.79e0 & 5.07e-1 \\
+\multicolumn{1}{c}{$E_{\varepsilon}$} & 1.52e1 & 1.09e0 & 2.91e0 & 4.43e-1 \\
+\multicolumn{1}{c}{$E_{\frac{\text{d} n}{\text{d} t}}$} & 2.53e1 & 2.64e0 & 7.44e0 & 1.58e1 \\
+\multicolumn{1}{c}{$E_{\frac{\text{d} T}{\text{d} t}}$} & 3.94e5 & 3.58e8 & 3.95e6 & 7.11e7 \\
+\multicolumn{1}{c}{$E_{\frac{\text{d} S}{\text{d} t}}$} & 3.52e12 & 3.46e12 & 3.44e12 & 3.38e12 \\
 \bottomrule

### 2018-02-06 scripts/source_term_error.py

@@ -43,6 +43,9 @@ for mech in mechs:
             errs = arrs[name]
             values = arrs[name + '_value']
             errs = errs / (atol + rtol * np.abs(values))
+            if 'IC5H11OH' in file and 'rop_fwd' in name and np.any(errs > 1e-4):
+                from time import ctime
+                print(os.path.basename(file), ctime(os.path.getmtime(file)))
 
             precs = None
             if 'rop_net' in name:

### 2018-02-06 paper.tex
add cpu / gpu data

@@
 Error statistics are based on the infinity-norm of the relative error detailed in Eq.~\eqref{e:rel_err} for each quantity.
+The ``S'' in $E_{\frac{\text{d} S}{\text{d} t}}$ refers to the thermodynamic state parameter, either $V$ or $P$ for \conp~and \conv~respectively.
 }
@@
 We see here that $E_{\varepsilon}$ is very similar in magnitude to $E_R$ in all cases, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
-The species rate error norm is similar in magnitude to that of the net ROP, however the temperature rate error again increases as the error in species rates is amplified by multiplication with thermodynamic properties.
+The molar species production rate error norm is similar in magnitude to that of the net ROP, however the temperature and state parameter rate errors again increases as the error in species rates is amplified by multiplication with thermodynamic properties.
 We note that the above discussion does not imply that computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
@@
 \label{S:results}
-The performance studies in this work were run on four ten-core \SI{2.2}{\giga\hertz} Intel Xeon E5-4640 v2 CPUs with \SI{20}{\mega\byte} of L3 cache memory, installed on an Ace Powerworks PW8027R-TRF+ with a Supermicro X9QR7-TF+/X9QRi-F+ baseboard.
+The performance studies in this work were run on a variety of different machines:
+
+%16.1.1 & 1.0 \slash 5.0.0 \\
+
+\begin{table}[htb]
+\centering
+\begin{tabular}{@{}l l l l l@{}}
+\toprule
+Model & Instruction Set & Vector Width & Cores & Identifier \\
+\midrule
+Intel Xeon X5650 & SSE4.2 & \SI{128}{\bit} & $2 \times 6$ & \texttt{sse} \\
+Intel E5-2690 V3 & AVX2 & \SI{256}{\bit} & $2 \times 12$ & \texttt{haswell} \\
+\bottomrule
+\end{tabular}
+\label{t:cpus}
+\caption{The CPU configurations used in this study.}
+\end{table}
+\begin{table}
+\centering
+\begin{tabular}{@{}l l l l l@{}}
+\toprule
+Model & CUDA Version & CUDA Cores & Identifier \\
+\midrule
+Tesla C2075 & 9.0.176 & \num{448} & \texttt{C2075} \\
+Tesla K40m & 8.0.61 & 2880 & \texttt{K40m} \\
+\bottomrule
+\end{tabular}
+\label{t:gpus}
+\caption{The GPU configurations used in this study.}
+\end{table}
+
+four ten-core \SI{2.2}{\giga\hertz} Intel Xeon E5-4640 v2 CPUs with \SI{20}{\mega\byte} of L3 cache memory, installed on an Ace Powerworks PW8027R-TRF+ with a Supermicro X9QR7-TF+/X9QRi-F+ baseboard.
 Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.

### 2018-02-06 scripts/jac_error.py

@@ -1,0 +1,26 @@
+import os
+import numpy as np
+from load_error_files import print_error
+
+rtol = 1e-6
+atol = 1e-10
+
+
+def update(err_dict, err, filename=None, **kwargs):
+    for name in err:
+        if 'amax' in name or 'value' in name:
+            continue
+        if err[name] > 1e20:
+            from time import ctime
+            print(os.path.basename(filename), ctime(os.path.getmtime(filename)))
+            print('Bad data detected...')
+            continue
+        err_dict[name] = np.maximum(err_dict[name], err[name])
+
+
+def printer(err_dict):
+    print('\n'.join(
+        ['{}: {:.15e}'.format(k, v) for k, v in err_dict.items()]))
+
+
+print_error('jac', update, printer)

### 2018-02-06 scripts/load_error_files.py

@@ -1,0 +1,48 @@
+import os
+import numpy as np
+import yaml
+from collections import defaultdict
+
+script_dir = os.path.dirname(os.path.abspath(__file__))
+error_path = os.path.join(script_dir, 'error_checking')
+
+
+def get_error_files(ftype):
+    mechs = [os.path.join(error_path, x) for x in os.listdir(
+        error_path) if os.path.isdir(os.path.join(error_path, x))]
+
+    # get file list
+    files = {}
+    for mech in mechs:
+        files[mech] = [os.path.join(mech, x) for x in os.listdir(
+            mech) if os.path.isfile(os.path.join(mech, x)) if x.endswith('.npz')
+            and ftype in x]
+
+    return files
+
+
+def run_error_calcs(ftype, updater):
+    files = get_error_files(ftype)
+    err_dicts = {}
+    for mech in files:
+        mech_name = os.path.basename(os.path.normpath(mech))
+        err_dicts[mech_name] = defaultdict(lambda: 0)
+
+        # get mech info
+        mech_info = os.path.join(mech, mech_name.lower() + '.yaml')
+        with open(mech_info, 'r') as file:
+            mech_info = yaml.load(file.read())
+
+        for file in files[mech]:
+            updater(err_dicts[mech_name], np.load(file), filename=file,
+                    mech_info=mech_info)
+
+    return err_dicts
+
+
+def print_error(ftype, updater, printer):
+    err_dicts = run_error_calcs(ftype, updater)
+    mech_arr = ['H2', 'CH4', 'C2H4', 'IC5H11OH']
+    for mech in mech_arr:
+        print(mech)
+        printer(err_dicts[mech])

### 2018-02-06 scripts/source_term_error.py

@@ -1,24 +1,12 @@
 import os
 import numpy as np
-import yaml
-
-script_dir = os.path.dirname(os.path.abspath(__file__))
-error_path = os.path.join(script_dir, 'error_checking')
-mechs = [os.path.join(error_path, x) for x in os.listdir(
-    error_path) if os.path.isdir(os.path.join(error_path, x))]
+from load_error_files import print_error
 
 rtol = 1e-6
 atol = 1e-10
 
-err_dicts = {}
-for mech in mechs:
-    mech_name = os.path.basename(os.path.normpath(mech))
-    # get file list
-    files = [os.path.join(mech, x) for x in os.listdir(
-        mech) if os.path.isfile(os.path.join(mech, x)) if x.endswith('.npz')
-        and 'spec' in x]
-    err_dicts[mech_name] = {}
 
+def updater(err_dict, err, filename=None, mech_info={}):
     def __get_size(name):
         if 'rop' in name:
             if 'net' in name or 'fwd' in name:
@@ -30,66 +18,54 @@ for mech in mechs:
         elif 'phi' in name:
             return mech_info['n_species'] + 1
 
-    # get mech info
-    mech_info = os.path.join(mech, mech_name.lower() + '.yaml')
-    with open(mech_info, 'r') as file:
-        mech_info = yaml.load(file.read())
+    for name in err:
+        if 'value' in name or 'component' in name or 'store' in name:
+            continue
+        errs = err[name]
+        values = err[name + '_value']
+        errs = errs / (atol + rtol * np.abs(values))
+        if 'rop_fwd' == name or 'rop_rev' == name and np.any(errs > 1e-4):
+            from time import ctime
+            print(filename, ctime(os.path.getmtime(filename)))
+            print('Bad data detected...')
 
-    for file in files:
-        arrs = np.load(file)
-        for name in arrs:
-            if 'value' in name or 'component' in name or 'store' in name:
-                continue
-            errs = arrs[name]
-            values = arrs[name + '_value']
-            errs = errs / (atol + rtol * np.abs(values))
-            if 'IC5H11OH' in file and 'rop_fwd' in name and np.any(errs > 1e-4):
-                from time import ctime
-                print(os.path.basename(file), ctime(os.path.getmtime(file)))
+        precs = None
+        if 'rop_net' in name:
+            # calculate the precision norms
+            precs = err['rop_component'] / (atol + rtol * np.abs(values))
 
-            precs = None
-            if 'rop_net' in name:
-                # calculate the precision norms
[…]

### 2018-02-07 scripts/data_parser.py
lots of plotting updateS

@@ -1,17 +1,46 @@
 import collections
 import os
 import yaml
+import hashlib
+import pickle
 
 rundata = collections.namedtuple(
-    'data', 'num_conditions comptime overhead runtime')
-mechdata = collections.namedtuple('mechdata', 'mech num_species num_reactions')
+    'rundata', 'num_conditions comptime overhead runtime')
+mechdata = collections.namedtuple(
+    'mechdata', 'mech n_species n_reactions n_reversible')
 run = collections.namedtuple(
     'run',
-    'type sparse lang vecwidth order vectype platform rates kernel cores conp'
-    'rundata mechdata')
+    tuple('rtype sparse lang vecwidth order vectype platform rates kernel cores '
+          'conp descriptor rundata mechdata'.split()))
+
+script_dir = os.path.dirname(os.path.normpath(__file__))
 
 
-def parse_data(directory='performance'):
+def md5sum(filename, blocksize=65536):
+    hash = hashlib.md5()
+    with open(filename, "rb") as f:
+        for block in iter(lambda: f.read(blocksize), b""):
+            hash.update(block)
+    return hash.hexdigest()
+
+
+def __listdir(path, files=True, endswith=''):
+    plist = [os.path.join(path, x) for x in os.listdir(path)]
+    plist = [x for x in plist if os.path.isfile(x) == files]
+    if endswith:
+        plist = [x for x in plist if x.endswith(endswith)]
+    return plist
+
+
+def files_in(path, endswith='.txt'):
+    return __listdir(path, endswith=endswith)
+
+
+def dirs_in(path):
+    return __listdir(path, False)
+
+
+def parse_data(directory=os.path.join(script_dir, 'performance')):
     # find all mechs
     mechs = [x for x in os.listdir(directory)
              if os.path.isdir(os.path.join(directory, x))]
@@ -19,29 +48,57 @@ def parse_data(directory='performance'):
     for mech in mechs:
         perf_data[mech] = []
         path = os.path.join(directory, mech)
-        # get files
-        files = [os.path.join(path, x) for x in os.listdir(path)]
-        files = [x for x in files if os.path.isfile(x)]
+        mech_meta_file = next(x for x in files_in(path, '.yaml'))
         # read mechanism info
-        mech_meta_file = next(x for x in files if x.endswith('.yaml'))
         with open(mech_meta_file, 'r') as f:
             mech_info = mechdata(**yaml.load(f))
-        files.remove(mech_meta_file)
-        for file in files:
-            # get name
[…]

### 2018-02-07 scripts/general_plotting.py

@@ -7,20 +7,21 @@ import plot_styles as ps
 
 font_size = 'large'
 
+
 def process_data(plotdata, plot, reacs_as_x=True,
-    plot_cores=False):
+                 plot_cores=False):
     """
     Process the data into an easily usable form
     """
     if reacs_as_x:
-        plotdata = sorted(plotdata, key=lambda x: x.mechdata.num_reactions)
-        x_vals = [x.mechdata.num_reactions for x in plotdata]
+        plotdata = sorted(plotdata, key=lambda x: x.mechdata.n_reactions)
+        x_vals = [x.mechdata.n_reactions for x in plotdata]
     elif plot_cores:
         plotdata = sorted(plotdata, key=lambda x: float(x.cores))
         x_vals = [float(x.cores) for x in plotdata]
     else:
-        plotdata = sorted(plotdata, key=lambda x: x.mechdata.num_species)
-        x_vals = [x.mechdata.num_species for x in plotdata]
+        plotdata = sorted(plotdata, key=lambda x: x.mechdata.n_species)
+        x_vals = [x.mechdata.n_species for x in plotdata]
 
     y_vals = []
     err_vals = []
@@ -28,15 +29,16 @@ def process_data(plotdata, plot, reacs_as_x=True,
         ys = np.array([getattr(data, plot) for data in run.rundata])
         zs = np.std(ys)
         if plot == 'runtime':
-            ys = np.array([getattr(data, plot) / data.num_conditions for data in run.rundata])
+            ys = np.array(
+                [getattr(data, plot) / data.num_conditions for data in run.rundata])
             zs /= run.rundata[0].num_conditions
         err_vals.append(zs)
         y_vals.append(np.mean(ys))
     return x_vals, y_vals, err_vals
 
 
-def plot(plot, x_vals, y_vals, err_vals, minx=None, miny=None, maxx=None, maxy=None, plot_std=True,
-        return_y=False, labels=[], plot_ind=None, marker_func=None):
+def plot(plot, x_vals, y_vals, err_vals, minx=None, miny=None, maxx=None, maxy=None,
+         plot_std=True, return_y=False, labels=[], plot_ind=None, marker_func=None):
     """Plot performance as a function of reaction count.
     """
 
@@ -45,19 +47,19 @@ def plot(plot, x_vals, y_vals, err_vals, minx=None, miny=None, maxx=None, maxy=N
         marker, hollow, color = marker_func(name)
[…]

### 2018-02-07 scripts/jac_error.py

@@ -12,8 +12,7 @@ def update(err_dict, err, filename=None, **kwargs):
             continue
         if err[name] > 1e20:
             from time import ctime
-            print(os.path.basename(filename), ctime(os.path.getmtime(filename)))
-            print('Bad data detected...')
+            print(filename, ctime(os.path.getmtime(filename)))
             continue
         err_dict[name] = np.maximum(err_dict[name], err[name])
 

### 2018-02-07 scripts/perf_plotter.py

@@ -16,15 +16,18 @@ data_clean = data_parser.parse_data()
 reacs_as_x = False
 norm = None
 
-#from http://stackoverflow.com/a/31174427
+# from http://stackoverflow.com/a/31174427
 sentinel = object()
+
+
 def rgetattr(obj, attr, default=sentinel):
     if default is sentinel:
         _getattr = getattr
     else:
         def _getattr(obj, name):
             return getattr(obj, name, default)
-    return functools.reduce(_getattr, [obj]+attr.split('.'))
+    return functools.reduce(_getattr, [obj] + attr.split('.'))
+
 
 def __compare(r, name, compare_value, plot_cores=False):
     """
@@ -37,44 +40,44 @@ def __compare(r, name, compare_value, plot_cores=False):
         return True
     return rgetattr(r, name) == compare_value
 
+
 def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
-        legend_handler=None, marker_func=None,
-        minx=None, miny=None, maxx=None, maxy=None, ylog=False,
-        return_vals=False, **filters):
+            legend_handler=None, marker_func=None,
+            minx=None, miny=None, maxx=None, maxy=None, ylog=False,
+            return_vals=False, **filters):
     data = copy.copy(data_clean)
-    #create fig, ax
-    fig = plt.figure()
-    ax = plt.subplot(1,1,1)
-    #apply filters
+    # create fig, ax
+    plt.figure()
+    ax = plt.subplot(1, 1, 1)
+    # apply filters
     for f in filters:
         if filters[f] is None:
             continue
         assert any(data[x] for x in data), 'No data matching all filters'
-        if f in data[list(data.keys())[0]][0]._asdict():
+        if f in data[list(data.keys())[0]][0]._fields:
             for mech in data:
-                data[mech] = [x for x in data[mech] if __compare(x, f, filters[f])]
-    #compute data
-    for mech in data:
-        for run in data[mech]:
-            rund = run.rundata
-    #now plot data
+                data[mech] = [x for x in data[mech]
+                              if __compare(x, f, filters[f])]
+    # now plot data
     to_plot = ['runtime']
     if filters.pop('plot_compilation', False):
         to_plot.append('comptime')
     if filters.pop('plot_overhead', False):
         to_plot.append('overhead')
-    #get data
+    # get data
     plot_data = [x for mech in data for x in data[mech]]
-    #get labels
+    # get labels
+    import pdb; pdb.set_trace()
[…]

### 2018-02-07 scripts/plot_styles.py

@@ -1,38 +1,40 @@
-#a single consolidated place to import
-#such that all figures have identical styling (when possible)
+# a single consolidated place to import
+# such that all figures have identical styling (when possible)
 
 import matplotlib as mpl
 import matplotlib.pyplot as plt
 import os.path
 
-#setup latex
+# setup latex
 plt.rc('text', usetex=True)
 plt.rc('font', family='serif')
 plt.rc('text.latex',
-    preamble=r'\usepackage{amsmath},\usepackage{siunitx},\usepackage[version=3]{mhchem}')
+       preamble=r'\usepackage{amsmath},\usepackage{siunitx},'
+                r'\usepackage[version=3]{mhchem}')
 plt.rc('font', family='serif')
 
-legend_style = {'loc':0,
-    'fontsize':18,
-    'numpoints':1,
-    'shadow':True,
-    'fancybox':True
-}
+legend_style = {'loc': 0,
+                'fontsize': 18,
+                'numpoints': 1,
+                'shadow': True,
+                'fancybox': True
+                }
 
 tick_font_size = 20
 
 marker_style = {
-    'size' : 15
+    'size': 15
 }
 
 clear_marker_style = {
-    'size' : 17
+    'size': 17
 }
 
 default_keys = ['runtime', 'comptime', 'overhead']
 marker_wheel = [('o', True), ('>', True), ('s', True),
-    ('D', True)]
-marker_dict = {x : marker_wheel[i] for i, x in enumerate(default_keys)}
+                ('D', True)]
+marker_dict = {x: marker_wheel[i] for i, x in enumerate(default_keys)}
+
 
 class dummy_formatter(object):
     def __init__(self, choicedict):
@@ -44,44 +46,48 @@ class dummy_formatter(object):
 
     def format(self, val):
         return next(self.values[i] for i, x in enumerate(self.choices)
-            if x == val)
+                    if x == val)
 
-legend_key = {'H2':r'H$_2$/CO',
-              'CH4':r'GRI-Mech 3.0',
-              'C2H4':r'USC-Mech II',
-              'IC5H11OH':r'IC$_5$H$_{11}$OH'
+
+legend_key = {'H2': r'H$_2$/CO',
+              'CH4': r'GRI-Mech 3.0',
+              'C2H4': r'USC-Mech II',
+              'IC5H11OH': r'IC$_5$H$_{11}$OH'
               }
 
 def pretty_names(pname, short=False):
[…]

### 2018-02-08 scripts/data_parser.py
plotting in python notebook

@@ -70,6 +70,7 @@ def parse_data(directory=os.path.join(script_dir, 'performance')):
                                 continue
                         except EOFError:
                             pass
+                    print('Loading pickled data failed')
 
                 # get name
                 name = os.path.basename(file)

### 2018-02-08 scripts/general_plotting.py

@@ -38,11 +38,15 @@ def process_data(plotdata, plot, reacs_as_x=True,
 
 
 def plot(plot, x_vals, y_vals, err_vals, minx=None, miny=None, maxx=None, maxy=None,
-         plot_std=True, return_y=False, labels=[], plot_ind=None, marker_func=None):
+         plot_std=True, return_y=False, labels=[], plot_ind=None, marker_func=None,
+         label=None, marker=None):
     """Plot performance as a function of reaction count.
     """
 
-    if marker_func is not None:
+    if label is not None and marker is not None:
+        name = label
+        marker, hollow, color = marker
+    elif marker_func is not None:
         name = labels[plot_ind]
         marker, hollow, color = marker_func(name)
     elif plot_ind is not None:

### 2018-02-08 scripts/perf_plotter.py

@@ -9,10 +9,12 @@ import numpy as np
 import itertools
 import functools
 import copy
+import six
+import os
+import pickle
 
 run = data_parser.run
 rundata = data_parser.rundata
-data_clean = data_parser.parse_data()
 reacs_as_x = False
 norm = None
 
@@ -29,35 +31,61 @@ def rgetattr(obj, attr, default=sentinel):
     return functools.reduce(_getattr, [obj] + attr.split('.'))
 
 
-def __compare(r, name, compare_value, plot_cores=False):
+def __compare(r, name, compare_value, plot_cores=False, strict=False):
     """
     Specialty comparison function to account for differences
     in runtypes
     """
-    if name == 'vecwidth' and r.vectype == 'par':
+    if six.callable(compare_value):
+        return compare_value(rgetattr(r, name))
+    if name == 'vecwidth' and r.vectype == 'par' and not strict:
         return True
     if plot_cores and name == 'cores':
         return True
     return rgetattr(r, name) == compare_value
 
 
-def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
-            legend_handler=None, marker_func=None,
-            minx=None, miny=None, maxx=None, maxy=None, ylog=False,
-            return_vals=False, **filters):
+def flatten(data):
+    return [x for mech in data for x in data[mech] if x.rundata]
+
+
+def get_filtered_data(data_clean, warn=True, strict=False, **filters):
     data = copy.copy(data_clean)
-    # create fig, ax
-    plt.figure()
-    ax = plt.subplot(1, 1, 1)
     # apply filters
     for f in filters:
         if filters[f] is None:
             continue
-        assert any(data[x] for x in data), 'No data matching all filters'
-        if f in data[list(data.keys())[0]][0]._fields:
+        if warn:
+            assert any(data[x] for x in data), 'No data matching all filters'
+        if f in run._fields:
             for mech in data:
                 data[mech] = [x for x in data[mech]
-                              if __compare(x, f, filters[f])]
+                              if __compare(x, f, filters[f], strict=strict)]
+
+    return data
+
+
+def get_diffs(data):
+    diff_check = [x for x in run._fields if x not in ['mechdata', 'rundata']]
+    diffs = [set([getattr(x, check) for x in data])
+             for check in diff_check]
+    # get # with more than 1 option
+    diff_locs = [i for i in range(len(diffs)) if len(diffs[i]) > 1]
+    diffs = [x for x in diffs if len(x) > 1]
[…]

### 2018-02-08 scripts/plot_styles.py

@@ -1,9 +1,7 @@
 # a single consolidated place to import
 # such that all figures have identical styling (when possible)
 
-import matplotlib as mpl
 import matplotlib.pyplot as plt
-import os.path
 
 # setup latex
 plt.rc('text', usetex=True)
@@ -55,14 +53,15 @@ legend_key = {'H2': r'H$_2$/CO',
               'IC5H11OH': r'IC$_5$H$_{11}$OH'
               }
 
+
 def pretty_names(pname, short=False):
     pname_dict = {'runtime': 'Runtime',
                   'comptime': 'Compilation time',
                   'overhead': 'Kernel Construction Overhead',
                   'vecwidth': 'Vector Width = {}',
-                  'vectype': dummy_formatter({'w': 'Shallow SIMD',
-                                              'par': 'SIMT',
-                                              'd': 'Deep SIMD'}),
+                  'vectype': dummy_formatter({'w': 'Shallow',
+                                              'par': 'Parallel',
+                                              'd': 'Deep'}),
                   'order': dummy_formatter({'C': 'C-order',
                                             'F': 'F-order'}),
                   'kernel': dummy_formatter({'single': 'Single Rate Kernel',
@@ -70,9 +69,10 @@ def pretty_names(pname, short=False):
                   'rates': dummy_formatter({'fixed': 'Fixed Rate Specialization',
                                             'hybrid': 'Hybrid Rate Specialization',
                                             'full': 'Full Rate Specialization'}),
-                  'mechdata.mech': dummy_formatter(legend_key)
-                  'descriptor': dummy_formatter({'srv2': 'sse',
-                                                 'haswell': 'haswell'})
+                  'mechdata.mech': dummy_formatter(legend_key),
+                  'descriptor': dummy_formatter({'srv2': r'\texttt{sse4.2}',
+                                                 'haswell': r'\texttt{avx2}'}),
+                  'lang': dummy_formatter({'c': r'OpenMP'})
                   }
     if pname in pname_dict:
         return pname_dict[pname]

### 2018-02-09 paper.tex
update

@@
 \midrule
-Intel Xeon X5650 & SSE4.2 & \SI{128}{\bit} & $2 \times 6$ & \texttt{sse} \\
-Intel E5-2690 V3 & AVX2 & \SI{256}{\bit} & $2 \times 12$ & \texttt{haswell} \\
+Intel Xeon X5650 & SSE4.2 & \SI{128}{\bit} & $2 \times 6$ & \texttt{sse4.2} \\
+Intel E5-2690 V3 & AVX2 & \SI{256}{\bit} & $2 \times 12$ & \texttt{avx2} \\
 \bottomrule

### 2018-02-09 scripts/data_parser.py
update plits

@@ -3,15 +3,30 @@ import os
 import yaml
 import hashlib
 import pickle
+import copy
 
 rundata = collections.namedtuple(
     'rundata', 'num_conditions comptime overhead runtime')
 mechdata = collections.namedtuple(
-    'mechdata', 'mech n_species n_reactions n_reversible')
-run = collections.namedtuple(
-    'run',
-    tuple('rtype sparse lang vecwidth order vectype platform rates kernel cores '
-          'conp descriptor rundata mechdata'.split()))
+    'mechdata', 'name mech n_species n_reactions n_reversible')
+
+
+class run (
+        collections.namedtuple('run', tuple(
+            'rtype sparse lang vecwidth order '
+            'vectype platform rates kernel cores '
+            'conp descriptor rundata mechdata'.split()))):
+    __slots__ = ()
+
+    def __repr__(self):
+        return self.__class__.__name__ + '({})'.format(', '.join([
+            '{}={}'.format(self._fields[i], self[i])
+            for i in range(len(self._fields)) if self._fields[i] not in
+            ['rundata', 'mechdata']]))
+
+    def copy(self):
+        return copy.deepcopy(self)
+
 
 script_dir = os.path.dirname(os.path.normpath(__file__))
 
@@ -40,7 +55,7 @@ def dirs_in(path):
     return __listdir(path, False)
 
 
-def parse_data(directory=os.path.join(script_dir, 'performance')):
+def parse_data(rebuild=False, directory=os.path.join(script_dir, 'performance')):
     # find all mechs
     mechs = [x for x in os.listdir(directory)
              if os.path.isdir(os.path.join(directory, x))]
@@ -51,7 +66,7 @@ def parse_data(directory=os.path.join(script_dir, 'performance')):
         mech_meta_file = next(x for x in files_in(path, '.yaml'))
         # read mechanism info
         with open(mech_meta_file, 'r') as f:
-            mech_info = mechdata(**yaml.load(f))
+            mech_info = mechdata(name=mech, **yaml.load(f))
         for desc in dirs_in(path):
             # get decriptor
             descriptor = os.path.basename(os.path.normpath(desc))
@@ -59,7 +74,8 @@ def parse_data(directory=os.path.join(script_dir, 'performance')):
                 # check for pickled data
                 pickle_file = file.replace('.txt', '.pickle')
                 md5_file = file.replace('.txt', '.md5')
-                if os.path.isfile(pickle_file) and os.path.isfile(md5_file):
[…]

### 2018-02-09 scripts/general_plotting.py

@@ -9,19 +9,46 @@ font_size = 'large'
 
 
 def process_data(plotdata, plot, reacs_as_x=True,
-                 plot_cores=False):
+                 plot_cores=False, plot_scaling=False):
     """
     Process the data into an easily usable form
     """
-    if reacs_as_x:
-        plotdata = sorted(plotdata, key=lambda x: x.mechdata.n_reactions)
-        x_vals = [x.mechdata.n_reactions for x in plotdata]
-    elif plot_cores:
+    def __mechsort(data):
+        if reacs_as_x:
+            data = sorted(data, key=lambda x: x.mechdata.n_reactions)
+            return data, [x.mechdata.n_reactions for x in data]
+        else:
+            data = sorted(data, key=lambda x: x.mechdata.n_species)
+            return data, [x.mechdata.n_species for x in data]
+
+    if plot_cores:
         plotdata = sorted(plotdata, key=lambda x: float(x.cores))
         x_vals = [float(x.cores) for x in plotdata]
+    elif plot_scaling:
+        # first sort by # mech
+        plotdata, _ = __mechsort(plotdata)
+        # next sort within mechanism by # of conditions
+        for mech in set([x.mechdata.mech for x in plotdata]):
+            subdata = [x for x in plotdata if x.mechdata.mech == mech]
+            x_vals = []
+            y_vals = []
+            err_vals = []
+            for data in subdata:
+                for xv in set(run.num_conditions for run in data.rundata):
+                    subruns = [run for run in data.rundata
+                               if run.num_conditions == xv]
+                    # take mean & std
+                    x_vals.append(xv)
+                    yv = np.array([getattr(run, plot) for run in subruns])
+                    if plot == 'runtime':
+                        yv /= xv
+                    y_vals.append(np.mean(yv))
+                    err_vals.append(np.std(yv))
+
+        return np.array(x_vals), np.array(y_vals), np.array(err_vals)
+
     else:
-        plotdata = sorted(plotdata, key=lambda x: x.mechdata.n_species)
-        x_vals = [x.mechdata.n_species for x in plotdata]
+        plotdata, x_vals = __mechsort(plotdata)
 
     y_vals = []
     err_vals = []
@@ -37,12 +64,8 @@ def process_data(plotdata, plot, reacs_as_x=True,
     return x_vals, y_vals, err_vals
 
 
-def plot(plot, x_vals, y_vals, err_vals, minx=None, miny=None, maxx=None, maxy=None,
[…]

### 2018-02-09 scripts/perf_plotter.py

@@ -15,6 +15,7 @@ import pickle
 
 run = data_parser.run
 rundata = data_parser.rundata
+mechdata = data_parser.mechdata
 reacs_as_x = False
 norm = None
 
@@ -57,10 +58,12 @@ def get_filtered_data(data_clean, warn=True, strict=False, **filters):
             continue
         if warn:
             assert any(data[x] for x in data), 'No data matching all filters'
-        if f in run._fields:
+        try:
             for mech in data:
                 data[mech] = [x for x in data[mech]
                               if __compare(x, f, filters[f], strict=strict)]
+        except AttributeError:
+            pass
 
     return data
 
@@ -122,8 +125,7 @@ def plotter(data_clean, plot_name='', show=True, plot_reacs=True, norm=True,
     if not diff_locs:
         # regular plot
         for plot in to_plot:
-            gp.plot(plot, *gp.process_data(plot_data,
-                                           plot, reacs_as_x=plot_reacs))
+            gp.plot(*gp.process_data(plot_data, plot, reacs_as_x=plot_reacs))
     else:
         # create map dict
         loc_map = {}
@@ -242,7 +244,7 @@ def plotter(data_clean, plot_name='', show=True, plot_reacs=True, norm=True,
 
         # and finally plot
         for i in range(len(y_vals)):
-            gp.plot('', x_vals[i], y_vals[i], z_vals[i],
+            gp.plot(x_vals[i], y_vals[i], z_vals[i],
                     labels=labels, plot_ind=i, marker_func=marker_func)
         ax.set_xlim([minx, maxx])
         ax.set_ylim([miny, maxy])
@@ -366,7 +368,7 @@ if __name__ == '__main__':
         pass
     finally:
         if data_clean is None:
-            data_clean = data_parser.parse_data()
+            data_clean = data_parser.parse_data(rebuild=opts['rebuild'])
             with open(os.path.join(script_dir, 'data.pickle'), 'wb') as file:
                 pickle.dump(data_clean, file)
 

### 2018-02-09 scripts/plot_styles.py

@@ -69,9 +69,10 @@ def pretty_names(pname, short=False):
                   'rates': dummy_formatter({'fixed': 'Fixed Rate Specialization',
                                             'hybrid': 'Hybrid Rate Specialization',
                                             'full': 'Full Rate Specialization'}),
-                  'mechdata.mech': dummy_formatter(legend_key),
+                  'mechdata.name': dummy_formatter(legend_key),
                   'descriptor': dummy_formatter({'srv2': r'\texttt{sse4.2}',
-                                                 'haswell': r'\texttt{avx2}'}),
+                                                 'haswell': r'\texttt{avx2}',
+                                                 'gpu': r'\texttt{K40m}'}),
                   'lang': dummy_formatter({'c': r'OpenMP'})
                   }
     if pname in pname_dict:

### 2018-02-10 paper.tex
update error discussion w/ new data

@@
 \usepackage{subcaption} %subfigs
+\usepackage{lmodern}
+\usepackage[T1]{fontenc}
+\usepackage{textcomp}
 
@@
 
-\todo[inline]{update species error with new data -- in particular the \num{e-4} for i-pentanol are all from november. preliminary data on the GPU shows ~\num{4e-8}}
 \begin{table}[htb]
@@
 \midrule
-\multicolumn{1}{c}{$E_{R^{\prime}}$} & 1.61e-8 & 2.95e-8 & 1.04e-7 & 4.86e-4 \\
-\multicolumn{1}{c}{$E_{R^{\prime\prime}}$} & 6.93e-8 & 6.43e-8 & 1.20e-7 & 5.07e-4 \\
-\multicolumn{1}{c}{$E_{R}$}  & 1.49e1 & 1.17e0 & 2.79e0 & 5.07e-1 \\
-\multicolumn{1}{c}{$E_{\varepsilon}$} & 1.52e1 & 1.09e0 & 2.91e0 & 4.43e-1 \\
-\multicolumn{1}{c}{$E_{\frac{\text{d} n}{\text{d} t}}$} & 2.53e1 & 2.64e0 & 7.44e0 & 1.58e1 \\
-\multicolumn{1}{c}{$E_{\frac{\text{d} T}{\text{d} t}}$} & 3.94e5 & 3.58e8 & 3.95e6 & 7.11e7 \\
+\multicolumn{1}{c}{$E_{R^{\prime}}$} & 1.56e-8 & 2.95e-8 & 9.42e-8 & 4.86e-4 \\
+\multicolumn{1}{c}{$E_{R^{\prime\prime}}$} & 6.92e-8 & 6.53e-8 & 1.20e-7 & 5.07e-4 \\
+\multicolumn{1}{c}{$E_{R}$}  & 1.49e1 & 1.11e0 & 2.80e0 & 4.82e-1 \\
+\multicolumn{1}{c}{$E_{\varepsilon}$} & 1.48e1 & 1.13e0 & 2.93e0 & 5.03e-1 \\
+\multicolumn{1}{c}{$E_{\frac{\text{d} n}{\text{d} t}}$} & 2.53e1 & 2.60e0 & 7.62e0 & 1.58e1 \\
+\multicolumn{1}{c}{$E_{\frac{\text{d} T}{\text{d} t}}$} & 3.94e5 & 3.35e8 & 3.95e6 & 7.11e7 \\
 \multicolumn{1}{c}{$E_{\frac{\text{d} S}{\text{d} t}}$} & 3.52e12 & 3.46e12 & 3.44e12 & 3.38e12 \\
@@
 }
-\label{T:error}
+\label{T:source_error}
 \end{table}
 
-In Table~\ref{T:error}, we see the results of this code as compared to Cantera~\cite{Cantera} on a library of partially stirred reaction conditions (PaSR) described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
-Very close agreement is observed for the forward and reverse ROP for all models, however the net ROP error norm is \numrange{7}{9} orders of magnitude larger.
-We see here that $E_{\varepsilon}$ is very similar in magnitude to $E_R$ in all cases, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
+In Table~\ref{T:source_error}, we see the results of this code as compared to Cantera~\cite{Cantera} on a library of partially stirred reaction conditions (PaSR) described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
+Very close agreement is observed for the forward and reverse ROP for all models, however the error norm is \textasciitilde\numrange{3}{4} orders of magnitude larger for the isopentanol model.
+This results from differences in evaluation of P-Log reactions between \texttt{pyJac} and Cantera; namely \texttt{pyJac} computes the logarithm of the upper and lower reaction Arrhenius rates (see~\cref{deriv-s:pdep}) analytically while Cantera evaluates this term numerically.
+If the error from P-Log reactions are not considered in~\cref{e:rel_err}, the forward and reverse ROP falls to \num{5.44e-08} and \num{1.59e-07} respectively.
+We note that this discrepency does not imply any incorrectness in either \texttt{pyJac} or Cantera---in fact, the error is still well within the proscribed tolerances in~\cref{e:rel_err}---but merely emphasizes the effect small code-changes can have on the accumulation of floating point errors.
+
+This point is further underscored by the error in the net ROP, which is \textasciitilde\numrange{3}{9} (or \numrange{7}{9} for the P-Log excluded error) orders of magnitude larger than the error in either the forward or reverse ROP.
+In~\cref{T:source_error}, we see that $E_{\varepsilon}$ is very similar in magnitude to $E_R$ in all cases, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
 The molar species production rate error norm is similar in magnitude to that of the net ROP, however the temperature and state parameter rate errors again increases as the error in species rates is amplified by multiplication with thermodynamic properties.
-We note that the above discussion does not imply that computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
+Again, we note that the above discussion does not imply that discrepencies in computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
 

### 2018-02-10 scripts/load_error_files.py
print plog / cheb excluded error if in mech

@@ -1,6 +1,7 @@
 import os
 import numpy as np
 import yaml
+import cantera as ct
 from collections import defaultdict
 
 script_dir = os.path.dirname(os.path.abspath(__file__))
@@ -21,6 +22,19 @@ def get_error_files(ftype):
     return files
 
 
+def add_rxn_specific_inds(mech_info, rxn_type, rxn_str, path):
+    # load gas
+    gas = ct.Solution(os.path.join(path, mech_info['mech']))
+    # get fwd reaction indicies
+    mech_info['rop_fwd_' + rxn_str] = np.array([i for i, r in enumerate(
+        gas.reactions()) if not isinstance(r, rxn_type)])
+    # get reversible indicies
+    rev_rxns = [i for i, r in enumerate(gas.reactions()) if r.reversible]
+    mech_info['rop_rev_' + rxn_str] = np.array([
+        i for i, fi in enumerate(rev_rxns)
+        if not isinstance(gas.reaction(fi), rxn_type)])
+
+
 def run_error_calcs(ftype, updater):
     files = get_error_files(ftype)
     err_dicts = {}
@@ -29,9 +43,15 @@ def run_error_calcs(ftype, updater):
         err_dicts[mech_name] = defaultdict(lambda: 0)
 
         # get mech info
-        mech_info = os.path.join(mech, mech_name.lower() + '.yaml')
+        mech_info = os.path.join(mech, mech_name + '.yaml')
         with open(mech_info, 'r') as file:
             mech_info = yaml.load(file.read())
+            if mech_info['n_cheb']:
+                add_rxn_specific_inds(mech_info, ct.ChebyshevReaction, 'cheb_inds',
+                                      mech)
+            if mech_info['n_plog']:
+                add_rxn_specific_inds(mech_info, ct.PlogReaction, 'plog_inds',
+                                      mech)
 
         for file in files[mech]:
             updater(err_dicts[mech_name], np.load(file), filename=file,

### 2018-02-10 scripts/source_term_error.py

@@ -24,7 +24,8 @@ def updater(err_dict, err, filename=None, mech_info={}):
         errs = err[name]
         values = err[name + '_value']
         errs = errs / (atol + rtol * np.abs(values))
-        if 'rop_fwd' == name or 'rop_rev' == name and np.any(errs > 1e-4):
+        if ('rop_fwd' == name or 'rop_rev' == name) and 'store' in name and np.any(
+                errs > 1e-4):
             from time import ctime
             print(filename, ctime(os.path.getmtime(filename)))
             print('Bad data detected...')
@@ -36,6 +37,12 @@ def updater(err_dict, err, filename=None, mech_info={}):
 
         if name not in err_dict:
             err_dict[name] = np.zeros((__get_size(name)))
+
+            if ('rop_fwd' in name or 'rop_rev' in name):
+                if mech_info['n_cheb']:
+                    err_dict[name + '_nocheb'] = np.zeros((__get_size(name)))
+                if mech_info['n_plog']:
+                    err_dict[name + '_noplog'] = np.zeros((__get_size(name)))
 
             if precs is not None:
                 err_dict['rop_component'] = np.zeros((__get_size(
@@ -49,6 +56,17 @@ def updater(err_dict, err, filename=None, mech_info={}):
 
         err_dict[name] = np.maximum(
             err_dict[name], errs)
+
+        def __update_rxn_type(rxn_str):
+            if ('rop_fwd' == name or 'rop_rev' == name) and \
+                    mech_info['n_' + rxn_str]:
+                inds = mech_info[name + '_' + rxn_str + '_inds']
+                err_dict[name + '_no' + rxn_str][inds] = np.maximum(
+                    err_dict[name + '_no' + rxn_str][inds], errs[inds])
+
+        __update_rxn_type('cheb')
+        __update_rxn_type('plog')
+
         if 'rop_net' in name:
             update_locs = np.where(err_dict[name] == errs)
             # update the precision norms at these locations

### 2018-02-13 scripts/data_parser.py
many updates

@@ -8,7 +8,7 @@ import copy
 rundata = collections.namedtuple(
     'rundata', 'num_conditions comptime overhead runtime')
 mechdata = collections.namedtuple(
-    'mechdata', 'name mech n_species n_reactions n_reversible')
+    'mechdata', 'name mech n_species n_reactions n_reversible n_cheb n_plog')
 
 
 class run (
@@ -66,7 +66,7 @@ def parse_data(rebuild=False, directory=os.path.join(script_dir, 'performance'))
         mech_meta_file = next(x for x in files_in(path, '.yaml'))
         # read mechanism info
         with open(mech_meta_file, 'r') as f:
-            mech_info = mechdata(name=mech, **yaml.load(f))
+            mech_info = mechdata(**yaml.load(f))
         for desc in dirs_in(path):
             # get decriptor
             descriptor = os.path.basename(os.path.normpath(desc))
@@ -109,6 +109,13 @@ def parse_data(rebuild=False, directory=os.path.join(script_dir, 'performance'))
                 # add to run
                 params.extend([descriptor, data, mech_info])
                 perf_data[mech].append(run(*params))
+
+                if descriptor == 'srv2' and perf_data[mech][-1].platform == 'nvidia':
+                    perf_data[mech][-1] = perf_data[mech][-1]._replace(
+                        descriptor='srv2-gpu')
+                if perf_data[mech][-1].lang == 'c':
+                    perf_data[mech][-1] = perf_data[mech][-1]._replace(
+                        vectype='openmp')
 
                 # get file checksum and pickle
                 checksum = md5sum(file)

### 2018-02-13 scripts/jac_error.py

@@ -10,9 +10,13 @@ def update(err_dict, err, filename=None, **kwargs):
     for name in err:
         if 'amax' in name or 'value' in name:
             continue
-        if err[name] > 1e20:
+        if any(x in name for x in ['thresholded']) and err[name] > 1e1:
             from time import ctime
-            print(filename, ctime(os.path.getmtime(filename)))
+            file = filename[filename.index('error_checking'):].replace(
+                'error_checking', 'performance')
+            path = os.path.dirname(file)
+            name = os.path.basename(file)
+            print(os.path.join(path, 'gpu', name))
             continue
         err_dict[name] = np.maximum(err_dict[name], err[name])
 

### 2018-02-13 scripts/perf_plotter.py

@@ -50,12 +50,14 @@ def flatten(data):
     return [x for mech in data for x in data[mech] if x.rundata]
 
 
-def get_filtered_data(data_clean, warn=True, strict=False, **filters):
+def get_filtered_data(data_clean, warn=True, strict=False, printf=False, **filters):
     data = copy.copy(data_clean)
     # apply filters
     for f in filters:
         if filters[f] is None:
             continue
+        if printf:
+            print(f, filters[f], sum(len(data[x]) for x in data))
         if warn:
             assert any(data[x] for x in data), 'No data matching all filters'
         try:
@@ -367,7 +369,7 @@ if __name__ == '__main__':
     except:
         pass
     finally:
-        if data_clean is None:
+        if data_clean is None or opts['rebuild']:
             data_clean = data_parser.parse_data(rebuild=opts['rebuild'])
             with open(os.path.join(script_dir, 'data.pickle'), 'wb') as file:
                 pickle.dump(data_clean, file)

### 2018-02-13 scripts/plot_styles.py

@@ -31,6 +31,19 @@ clear_marker_style = {
 default_keys = ['runtime', 'comptime', 'overhead']
 marker_wheel = [('o', True), ('>', True), ('s', True),
                 ('D', True)]
+
+
+class wheeler(object):
+    def __init__(self, wheel, raise_on_oob=True):
+        self.raise_on_oob = raise_on_oob
+        self.wheel = wheel[:]
+
+    def __getitem__(self, key):
+        if not self.raise_on_oob and key >= len(self.wheel):
+            key = key % len(self.wheel)
+        return self.wheel[key]
+
+
 marker_dict = {x: marker_wheel[i] for i, x in enumerate(default_keys)}
 
 
@@ -61,7 +74,8 @@ def pretty_names(pname, short=False):
                   'vecwidth': 'Vector Width = {}',
                   'vectype': dummy_formatter({'w': 'Shallow',
                                               'par': 'Parallel',
-                                              'd': 'Deep'}),
+                                              'd': 'Deep',
+                                              'openmp': r'OpenMP'}),
                   'order': dummy_formatter({'C': 'C-order',
                                             'F': 'F-order'}),
                   'kernel': dummy_formatter({'single': 'Single Rate Kernel',
@@ -69,11 +83,12 @@ def pretty_names(pname, short=False):
                   'rates': dummy_formatter({'fixed': 'Fixed Rate Specialization',
                                             'hybrid': 'Hybrid Rate Specialization',
                                             'full': 'Full Rate Specialization'}),
+                  'conp': '{}',
                   'mechdata.name': dummy_formatter(legend_key),
                   'descriptor': dummy_formatter({'srv2': r'\texttt{sse4.2}',
+                                                 'srv2-gpu': r'\texttt{C2075}',
                                                  'haswell': r'\texttt{avx2}',
-                                                 'gpu': r'\texttt{K40m}'}),
-                  'lang': dummy_formatter({'c': r'OpenMP'})
+                                                 'gpu': r'\texttt{K40m}'})
                   }
     if pname in pname_dict:
         return pname_dict[pname]

### 2018-02-13 scripts/plot_styles.py
plot updates

@@ -83,12 +83,16 @@ def pretty_names(pname, short=False):
                   'rates': dummy_formatter({'fixed': 'Fixed Rate Specialization',
                                             'hybrid': 'Hybrid Rate Specialization',
                                             'full': 'Full Rate Specialization'}),
+                  'sparse': dummy_formatter({'sparse': 'Sparse',
+                                             'full': 'Dense'}),
                   'conp': '{}',
                   'mechdata.name': dummy_formatter(legend_key),
                   'descriptor': dummy_formatter({'srv2': r'\texttt{sse4.2}',
                                                  'srv2-gpu': r'\texttt{C2075}',
                                                  'haswell': r'\texttt{avx2}',
-                                                 'gpu': r'\texttt{K40m}'})
+                                                 'gpu': r'\texttt{K40m}'}),
+                  'rtype': dummy_formatter({'jac': 'Analytical',
+                                            'fdjac': 'Finite Difference'})
                   }
     if pname in pname_dict:
         return pname_dict[pname]

### 2018-02-14 paper.tex
updates

@@
 Additionally, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, and via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source terms.
-The source-terms and subcomponents thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the autodifferentiation code Adept~\cite{adept-v11} is used to obtain reference answers for Jacobian subcomponents.
+The source-terms and subcomponents thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{adept-v11} is used to obtain reference answers for Jacobian subcomponents.
 Validation of the complete source-terms and Jacobian codes will be discussed in detail in~\cref{s:validation}.
@@
 
-This point is further underscored by the error in the net ROP, which is \textasciitilde\numrange{3}{9} (or \numrange{7}{9} for the P-Log excluded error) orders of magnitude larger than the error in either the forward or reverse ROP.
+This point is further underscored by the error in the net ROP, which is \textasciitilde\numrange{3}{9} orders of magnitude (or \numrange{7}{9} orders of magnitude for the P-Log excluded error) larger than the error in either the forward or reverse ROP.
 In~\cref{T:source_error}, we see that $E_{\varepsilon}$ is very similar in magnitude to $E_R$ in all cases, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
@@
 
+\subsection{Jacobian validation}
+
+As in our previous work~\cite{Niemeyer:2016aa}, the correctness of the Jacobian codes generated by \texttt{pyJac} was determined by comparison to Jacobians obtained by automatic differentiation of the source rate subroutines via the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
+The reasons for this choice are explained fully in~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with relatively programming little effort.
+The discrepency between the analytical and automatic differentiated Jacobians for thermochemical state $k$, denoted by $\mathcal{J}_k$ and $\hat{\mathcal{J}}_k$ respectively, is determined by the relative error Frobenius norm over all Jacobian indicies $i, j$:
+\begin{equation}
+ \label{e:jac_error_base}
+ E_{\text{rel}, k} = \frac{\left\lVert \hat{\mathcal{J}}_{ijk} - \mathcal{J}_{ijk} \right\rVert_{i,j,F}}{\left\lVert \hat{\mathcal{J}}_{ijk} \right\rVert_{i,j,F}}
+\end{equation}
+
+To avoid large relative errors of small nonzero Jacobian elements due to accumulation of floating point error, the Frobenius norm of the automatically differentiated Jacobian is calculated over all thermochemical states $k$:
+\begin{equation}
+ \label{e:thresh}
+ \mathcal{T} = \left\lVert \mathcal{\hat{J}} \right\rVert_{k, F}
+\end{equation}
+The error statistics reported in this section are then based only on matrix elements where $\mathcal{J}_{ijk} \ge \frac{\mathcal{T}}{\mathcal{C}}$; this filtered form of~\cref{e:jac_error_base} is denoted $E_{\mathcal{C}}$.
+
+\todo{rerun, especially H2 error}
+\begin{table}[tbp]
+\centering
+\begin{tabular}{@{}l S S[table-format=1.3e1] S[table-format=1.3e1] S[table-format=1.3e1]@{}}
+\toprule
+Model                 & \multicolumn{1}{c}{Sample size}   & \multicolumn{1}{c}{$\bar{\mathcal{T}}$}  & \multicolumn{1}{c}{$E_{10^{20}}$}   & \multicolumn{1}{c}{$E_{10^{15}}$} \\
+\midrule
+\ce{H2}\slash \ce{CO} & \num{900900}                      & \num{6.431e+18}      & \num{8.083e-1}  & \num{1.907e-5} \\
+GRI-Mech 3.0          & \num{450900}                      & \num{7.783e+19}      & \num{1.890e-7}  & \num{1.877e-7} \\
+USC-Mech II           & \num{91800}                       & \num{2.830e+21}      & \num{5.567e-3}  & \num{1.704e-7} \\
+\ce{iC5H11OH}         & \num{450900}                      & \num{2.733e+26}      & \num{1.136e-3}  & \num{2.764e-5} \\
+\bottomrule
+\end{tabular}
+\caption{Summary of Jacobian matrix validation results.
+Error statistics are the largest filtered relative error $E_\mathcal{C}$ for over all samples.
+It is noted that the threshold described in~\cref{e:thresh} varies slightly between the \conp~and \conv~cases; the reported $\bar{\mathcal{T}}$ is the average of the two, however the appropriate value was used during calculations of the error statistics..
+}
+\label{T:error}
+\end{table}
+
+\Cref{T:error} presents the error statistics of \texttt{pyJac} over the tested PaSR thermochemical condition database.
+For all tested cases, \todo{need srv2 nvidia tests to really confirm here} the maximum filtered error norm is less than unity.
+The filtered error norm is largest for the \ce{H2}\slash\ce{CO} model with the more stringent tolerance ($\mathcal{C} = 10^{20}$).
+We note that $\mathcal{T}$ for this model is smaller than the tolerance in this case, and hence we are comparing numbers smaller than $\mathcal{O}(1)$; relaxing the tolerance to $\mathcal{C} = 10^{15}$ significantly lowers the filtered error norm to $\mathcal{O}(10^{-5})$.
+Similar error characteristics are seen for all models except GRI-Mech 3.0---in which the error norms for both tolerances are $\sim\mathcal{O}(10^{-7})$---that is a larger error norm for the more stringent tolerance that drops 2--4 orders of magnitude for the relaxed tolerance.
+
 \subsection{Performance}
@@
 The performance studies in this work were run on a variety of different machines:
-
-%16.1.1 & 1.0 \slash 5.0.0 \\
 
@@
 \toprule
-Model & CUDA Version & CUDA Cores & Identifier \\
+Model & NVIDIA Driver Version & CUDA Cores & Identifier \\
 \midrule
-Tesla C2075 & 9.0.176 & \num{448} & \texttt{C2075} \\
-Tesla K40m & 8.0.61 & 2880 & \texttt{K40m} \\
+Tesla C2075 & 384.81 & \num{448} & \texttt{C2075} \\
+Tesla K40m & 375.26 & \num{2880} & \texttt{K40m} \\
 \bottomrule
@@
 \label{t:gpus}
-\caption{The GPU configurations used in this study.}
+\caption{The GPU configurations used in this study.  NVIDIA's OpenCL runtime is provided with the graphic driver, rather than any specific version of CUDA.}
 \end{table}

### 2018-02-14 scripts/jac_error.py

@@ -8,22 +8,43 @@ atol = 1e-10
 
 def update(err_dict, err, filename=None, **kwargs):
     for name in err:
-        if 'amax' in name or 'value' in name:
+        if ('amax' in name or 'value' in name) and name != 'jac_threshold_value':
             continue
-        if any(x in name for x in ['thresholded']) and err[name] > 1e1:
+        assert 'fdjac' not in filename
+        if 'nvidia' in filename:
             from time import ctime
-            file = filename[filename.index('error_checking'):].replace(
-                'error_checking', 'performance')
-            path = os.path.dirname(file)
-            name = os.path.basename(file)
-            print(os.path.join(path, 'gpu', name))
+            print('Known driver issue:', filename, ctime(os.path.getmtime(filename)))
+            continue
+        # special handling for threshold
+        if name == 'jac_threshold_value':
+            sparse = 'sparse' if ('sparse' in filename) else 'dense'
+            conp = 'conp' if ('conp' in filename) else 'conv'
+            val = err[name]
+            name += '_{}_{}'.format(sparse, conp)
+            if name in err_dict:
+                assert np.isclose(err_dict[name], np.asscalar(val))
+            err_dict[name] = np.asscalar(val)
             continue
         err_dict[name] = np.maximum(err_dict[name], err[name])
+        if name + '_mean' not in err_dict:
+            err_dict[name + '_mean'] = (0., 0.)
+        err_dict[name + '_mean'] = (err_dict[name + '_mean'][0] + err[name],
+                                    err_dict[name + '_mean'][1] + 1)
 
 
 def printer(err_dict):
+    mean_thr = (0., 0.)
+    for name in err_dict:
+        if name.endswith('_mean'):
+            err_dict[name] = err_dict[name][0] / err_dict[name][1]
+        elif 'jac_threshold_value' in name:
+            mean_thr = (mean_thr[0] + err_dict[name], mean_thr[1] + 1)
+    # print mean threshold
+    mean_thr = mean_thr[0] / mean_thr[1]
+
     print('\n'.join(
         ['{}: {:.15e}'.format(k, v) for k, v in err_dict.items()]))
+    print('mean_threshold: {:.15e}'.format(mean_thr))
 
 
 print_error('jac', update, printer)

### 2018-02-14 paper.tex
flesh out lit. review section

@@
 \subsection{Previous works and goals of this study}
-Several recent works, e.g.~\cite{CurtisGPU:2017}, have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
-Stone et al., \cite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
+A number of recent works have investigated the use of high-performance SIMT-devices to accelerate reactive-flow and chemical kinetics simulations.
+Spafford et al.~\cite{Spafford:2010aa} investigated an implementation of an explicit direct numerical simulation code for compressible turbulent combustion on a Tesla C1060 GPU, and found an order of magnitude speedup in evaluation fo species production rates as compared to a serial CPU implementation on an AMD-Opteron processor.
+Shi et al.~\cite{Shi:2011aa} developed a code to evaluate species rates and factorize the chemical kinetic Jacobian on a Tesla C2050 GPU, and found an order of magnitude speedup during integration of large chemical models when these operations were offloaded to the GPU as compared to the standard \texttt{CHEMKIN}~\cite{kee1989chemkin} and \texttt{LAPACK}~\cite{Anderson:1999aa} libraries on a quad-core Intel i7 930 CPU; it was not clear how\slash if the CPU code was parallelized.
+Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit fourth-order Runge--Kutta integrator for non-stiff chemical kinetic integration on a Tesla C2075 GPU, and found a speedup of nearly two orders of magnitude when compared with a sequential CPU-code on an Intel Xeon X5650 CPU.
+Shi et al.~\cite{Shi:2012aa} later extended their work to develop a stabilized-explicit solver on a Tesla C2050; paired with a traditional implicit integrator on quad-core Intel i7 930 which handled the stiffest computational cells, an \SIrange{11}{46}{$\times$} speedup was achieved over the base implicit CPU integrator during simulation of a three-dimensional premixed diesel engine simulation.
+Le et al.~\cite{Le2013596} implemented two high-order shock-capturing reactive-flow codes on a Tesla C2070, and found a \numrange{30}{50}$\times$ speedup over a sequential CPU version of the same code on a Intel Xeon X5650.
+Stone and Davis~\cite{Stone:2013aa} implemented the implicit VODE~\cite{Brown:1989vl} integrator on a Fermi M2050 GPU, achieving an order of magnitude speedup over the baseline CPU version on a single core of an AMD Opteron 6134 Magny-Cours.
+Niemeyer and Sung~\cite{Niemeyer:2014aa} later developed a stabilized explicit second-order Runge--Kutta--Chebyshev algorithm on a Tesla C2075, demonstrating an order of magnitude speedup for a over a CPU implementation of VODE on a six-core Intel X5650 for moderately stiff chemical kinetics.
+Sewerin and Rigopoulos~\cite{Sewerin20151375} implemented a three-stage\slash fifth-order implicit Runge--Kutta method~\cite{wanner1991solving} on a both high-end (Nvidia Quadro 6000) and consumer-grade (Nvidia Quadro 600) GPUs, as well as a standard CPU (two-core, four-thread Intel i5-520M) and a scientific workstation (eight-core, 16-thread Intel Xeon E5-2687W); the high-end GPU was at best \SI{1.8}{$\times$} slower than the workstation CPU (16 threads), while the consumer level GPU was at best \SI{5.5}{$\times$} slower than the corresponding standard CPU (four threads).
+Niemeyer et al.~\cite{Niemeyer:2016aa} additionally developed and validated an analytical chemical kinetic Jacobian code for the CPU and GPU.
+Yonkee and Sutherland~\cite{Yonkee2016} implemented SIMT-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations on an Intel Xeon E5-2680 CPU and an Tesla K20 GPU.
+In evaluation of the thermochemical properties, speedups between \SIrange{8}{13}{$\times$} were on 16 CPU cores, and \SI{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU respectively.
+Curtis et al.~\cite{CurtisGPU:2017} implemented a different formulation of the same fifth-order implicit Runge--Kutta method, as well as two-fourth order exponential integration techniques paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on a Tesla C2075; the implicit Runge--Kutta method was roughly equivalent to VODE running on 12--38 Intel Xeon E5-4640 v2 CPU cores for two smaller chemical models with an integration time step of \SI{e-6}{$\sec$}.
+
+In contrast, SIMD-based chemical kinetics evaluation\slash integration have been far less studied.
+Linford et al.~\cite{Linford:2011} implemented a three-stage, second-order Rosenbrock atmospheric chemical kinetics solver on the CPU, GPU and cell croadband engine (CBE)---a specially designed vector processor---and found speedups regularly exceeding~\SI{25}{$\times$} over a serial CPU implementation.
+Kroshko and Spiteri~\cite{kroshko2013efficient} implemented a shallow-SIMD vectorized third order stiff Rosenbrock integrator for atmospheric chemistry on the CBE finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code.
+Stone et al.,~\cite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
 The shallow-SIMD and SIMT-vectorization models were implemented in OpenCL and compared to an OpenMP baseline code that was deep SIMD-vectorized by simple compiler hints (a.k.a pragmas).
 The shallow-SIMD vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance was \SIrange{1.4}{1.6}{$\times$} slower than the OpenMP baseline due to thread-divergence concerns.
-\cite{kroshko2013efficient} implemented a shallow-SIMD vectorized third order stiff Rosenbrock integrator for atmospheric chemistry on a Cell Broadband Engine---a specially designed vector processor---finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code.
 
@@
 The source-terms and subcomponents thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{adept-v11} is used to obtain reference answers for Jacobian subcomponents.
-Validation of the complete source-terms and Jacobian codes will be discussed in detail in~\cref{s:validation}.
+The Portable OpenCL (POCL) implementation~\cite{poclIJPP} and OpenMP~\cite{dagum1998openmp} are used to run OpenCL and C unit-testing, respectively, on the continous integration framework Travis CI~\cite{travis:2018}.
+Validation of the complete (as opposed to the subcomponent testing discussed here) generated source-terms and Jacobian codes will be discussed in detail in~\cref{s:validation}.
 
 \section{Results and Discussion}
+\subsection{Validation and performance study platforms}
+
+The performance and validation studies for this work were run on a variety of CPU and GPU platforms.
+\Cref{t:cpus} shows the number of cores, vector instruction set and model of the CPUs used in this work; each CPU had both \texttt{v16.1.1} of the Intel OpenCL runtime~\cite{intelopencl:2018} and \texttt{v1.0} of the POCL~\cite{poclIJPP} runtime installed, each enabling OpenCL \texttt{v1.2} execution.
+In addition,~\cref{t:gpus} lists the model, number of CUDA cores and NVIDIA driver of each of the GPUs used in this work.
+NVIDIA's OpenCL runtime is bundled with the NVIDIA driver~\cite{NVIDIA:2018}, hence we use driver version to specify the OpenCL runtime version.
+
+\begin{table}[htb]
+\centering
+\begin{tabular}{@{}l l l l l@{}}
+\toprule
+CPU Model & Instruction Set & Vector Width & Cores & Identifier \\
+\midrule
+Xeon X5650 & SSE4.2 & \SI{128}{\bit} & $2 \times 6$ & \texttt{sse4.2} \\
+E5-2690 V3 & AVX2 & \SI{256}{\bit} & $2 \times 12$ & \texttt{avx2} \\
+\bottomrule
+\end{tabular}
+\label{t:cpus}
+\caption{The Intel CPU configurations used in this study. The identifier field will be used as a shorthand descriptor in the performance plots to quickly identify the CPU type.}
+\end{table}
+\begin{table}
+\centering
+\begin{tabular}{@{}l l l l@{}}
+\toprule
+NVIDIA Model & Driver Version & CUDA Cores & Identifier \\
+\midrule
+Tesla C2075 & \num{384.81} & \num{448} & \texttt{C2075} \\
+Tesla K40m & \num{375.26} & \num{2880} & \texttt{K40m} \\
+\bottomrule
+\end{tabular}
+\label{t:gpus}
+\caption{The NVIDIA GPU configurations used in this study.  NVIDIA's OpenCL runtime is provided with the graphic driver, rather than any specific version of CUDA.
+Again, the identifier field will be used as a shorthand descriptor during analysis of the performance results.}
+\end{table}
+
 \subsection{Source term validation}
@@
 \label{S:results}
-The performance studies in this work were run on a variety of different machines:
-
-\begin{table}[htb]
-\centering
-\begin{tabular}{@{}l l l l l@{}}
-\toprule
-Model & Instruction Set & Vector Width & Cores & Identifier \\
-\midrule
-Intel Xeon X5650 & SSE4.2 & \SI{128}{\bit} & $2 \times 6$ & \texttt{sse4.2} \\
-Intel E5-2690 V3 & AVX2 & \SI{256}{\bit} & $2 \times 12$ & \texttt{avx2} \\
-\bottomrule
-\end{tabular}
-\label{t:cpus}
-\caption{The CPU configurations used in this study.}
-\end{table}
-\begin{table}
-\centering
-\begin{tabular}{@{}l l l l l@{}}
-\toprule
-Model & NVIDIA Driver Version & CUDA Cores & Identifier \\
-\midrule
-Tesla C2075 & 384.81 & \num{448} & \texttt{C2075} \\
-Tesla K40m & 375.26 & \num{2880} & \texttt{K40m} \\
-\bottomrule
-\end{tabular}
-\label{t:gpus}
-\caption{The GPU configurations used in this study.  NVIDIA's OpenCL runtime is provided with the graphic driver, rather than any specific version of CUDA.}
-\end{table}
+The performance studies in this work were run on a variety of different machines.
 
@@
 Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
-All code was compiled with \texttt{gcc 4.8.5} and ``\texttt{-O3}'' optimization, running on \texttt{v16.1.1} of the Intel OpenCL runtime supporting OpenCL \texttt{v1.2}.
+All code was compiled with \texttt{gcc 4.8.5} and ``\texttt{-O3}'' optimization, running on  of the
 

### 2018-02-14 paper.tex
goals

@@
 
-This work will study the performance of thermo-chemical source term evaluation of automatically generated shallow-SIMD vectorized codes for a wide range chemical kinetic models on the CPU.
-The performance will be compared to a baseline shallow-SIMT vectorized code to determine the effective SIMD speedup.
-Finally future extensions to this work will be detailed.
+This work will build upon our previous analytical chemical kinetic Jacobian code,~\cite{Niemeyer:2016aa} in order to:
+\begin{itemize}
+ \item Derive and validate a new Jacobian formulation that greatly increases sparsity
+ \item Enable cross-platform SIMD\slash SIMT vectorization for the CPU, GPU and other accelerators
+ \item Investigate the performance for a wide range chemical kinetic models, specifically looking at speedups due to SIMD\slash SIMT-vectorization as well as relative to the previous version of \texttt{pyJac}, and finally
+ \item Discuss future extensions to this work as well as the most promising directions for SIMD\slash SIMT vectorization in reactive flow simulations
+\end{itemize}
+.
 

### 2018-02-15 paper.tex
edits

@@
-\documentclass[12pt,number,sort&compress]{elsarticle}
+\documentclass[12pt,number,sort&compress,preprint]{elsarticle}
 
@@
 
-\title{SIMD\slash SIMT-vectorized Thermo-chemical Source Term Evaluation}
+\title{SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source Term Evaluation}
 
@@
 \begin{abstract} % not to exceed 200 words
-A code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized thermo-chemical source term evaluation was developed and validated against Cantera for a wide range of chemical kinetic models.
-Speedups of up to \SIrange{1.90}{2.45}{$\times$} were observed over a baseline single-instruction, multiple-thread (SIMT) parallelized code.
-Additionally, use of a row-major data layout was found to increase both the SIMD and SIMT performance by up to \SIrange{2.00}{2.38}{$\times$} over a column-major data layout.
-Larger vector-widths were found to produce speedups up to \SI{1.8}{$\times$} as compared to a small vector-width for smaller models, e.g.
-GRI-Mech 3.0, but as the model size increased the performance of all vector-widths were similar.
-Finally, the parallel scaling efficiency was found to be slightly better for larger models, and future extensions of this work were discussed.
+A code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized thermo-chemical source term and sparse chemical kinetic Jacobian evaluation was developed and validated for a wide range of chemical kinetic models.
+\todo[inline]{this}
 \end{abstract}
@@
 Single-Instruction, Multiple-Data (SIMD) and the related Single-Instruction, Multiple-Thread (SIMT) processing are two important vector-processing paradigms used increasingly in scientific computing.
-Traditional multi-core parallelism is often used to increase central processing unit (CPU) performance, however SIMD processors---as well as SIMT processors, e.g.
-in the form of graphics processing units (GPUs)---have gained recognition due to their increased floating operation throughput.
+Traditional multi-core parallelism is often used to increase central processing unit (CPU) performance, however SIMD processors---as well as SIMT processors, e.g., in the form of graphics processing units (GPUs)---have gained recognition due to their increased floating operation throughput.
 The parallel programming standard OpenCL~\cite{stone2010opencl} has further enabled adoption of vector-processing based codes in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems, e.g., the CPU, GPU, or Intel's Many Integrated Core architecture (MIC), etc.
-In this discussion, we will largely use OpenCL terminology to describe these processing paradigms, as it provides a convient way to classify otherwise disparate processor type (e.g., CPUs and GPUs).
+In this discussion, we will largely use OpenCL terminology to describe these processing paradigms, as it provides a convenient way to classify otherwise disparate processor type (e.g., CPUs and GPUs), however the concepts discussed are broadly applicable to SIMD\slash SIMT processing and not tied to OpenCL specifically.
 
@@
 Intel's Xeon Phi co-processor (MIC) have been developed that have tens of cores with very wide vector-widths (e.g.
-4--8 double precision operations); these very wide vector-widths are also expected on cutting-edge and forthcoming Intel CPUs (the Skylake Xeon and Cannon Lake architectures).
+4--8 double precision operations); these very wide vector-widths are also found on cutting-edge and expected on forthcoming Intel CPUs (the Skylake Xeon and Cannon Lake architectures).
 
@@
 A number of recent works have investigated the use of high-performance SIMT-devices to accelerate reactive-flow and chemical kinetics simulations.
-Spafford et al.~\cite{Spafford:2010aa} investigated an implementation of an explicit direct numerical simulation code for compressible turbulent combustion on a Tesla C1060 GPU, and found an order of magnitude speedup in evaluation fo species production rates as compared to a serial CPU implementation on an AMD-Opteron processor.
-Shi et al.~\cite{Shi:2011aa} developed a code to evaluate species rates and factorize the chemical kinetic Jacobian on a Tesla C2050 GPU, and found an order of magnitude speedup during integration of large chemical models when these operations were offloaded to the GPU as compared to the standard \texttt{CHEMKIN}~\cite{kee1989chemkin} and \texttt{LAPACK}~\cite{Anderson:1999aa} libraries on a quad-core Intel i7 930 CPU; it was not clear how\slash if the CPU code was parallelized.
+Spafford et al.~\cite{Spafford:2010aa} investigated an implementation of an explicit direct numerical simulation code for compressible turbulent combustion on a Tesla C1060 GPU, and found an order of magnitude speedup in evaluation of species production rates as compared to a serial CPU implementation on an AMD-Opteron processor.
+Shi et al.~\cite{Shi:2011aa} developed a code to evaluate species production rates and factorize the chemical kinetic Jacobian on a Tesla C2050 GPU, and found an order of magnitude speedup during integration of large chemical models when these operations were offloaded to the GPU as compared to the standard \texttt{CHEMKIN}~\cite{kee1989chemkin} and \texttt{LAPACK}~\cite{Anderson:1999aa} libraries on a quad-core Intel i7 930 CPU; it was not clear how\slash if the CPU code was parallelized.
 Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit fourth-order Runge--Kutta integrator for non-stiff chemical kinetic integration on a Tesla C2075 GPU, and found a speedup of nearly two orders of magnitude when compared with a sequential CPU-code on an Intel Xeon X5650 CPU.
@@
 Niemeyer and Sung~\cite{Niemeyer:2014aa} later developed a stabilized explicit second-order Runge--Kutta--Chebyshev algorithm on a Tesla C2075, demonstrating an order of magnitude speedup for a over a CPU implementation of VODE on a six-core Intel X5650 for moderately stiff chemical kinetics.
-Sewerin and Rigopoulos~\cite{Sewerin20151375} implemented a three-stage\slash fifth-order implicit Runge--Kutta method~\cite{wanner1991solving} on a both high-end (Nvidia Quadro 6000) and consumer-grade (Nvidia Quadro 600) GPUs, as well as a standard CPU (two-core, four-thread Intel i5-520M) and a scientific workstation (eight-core, 16-thread Intel Xeon E5-2687W); the high-end GPU was at best \SI{1.8}{$\times$} slower than the workstation CPU (16 threads), while the consumer level GPU was at best \SI{5.5}{$\times$} slower than the corresponding standard CPU (four threads).
+Sewerin and Rigopoulos~\cite{Sewerin20151375} implemented a three-stage\slash fifth-order implicit Runge--Kutta method~\cite{wanner1991solving} on both a high-end (Nvidia Quadro 6000) and consumer-grade (Nvidia Quadro 600) GPUs, as well as a standard CPU (two-core, four-thread Intel i5-520M) and a scientific workstation (eight-core, 16-thread Intel Xeon E5-2687W); the high-end GPU was at best \SI{1.8}{$\times$} slower than the workstation CPU (16 threads), while the consumer level GPU was at best \SI{5.5}{$\times$} slower than the corresponding standard CPU (four threads).
 Niemeyer et al.~\cite{Niemeyer:2016aa} additionally developed and validated an analytical chemical kinetic Jacobian code for the CPU and GPU.
-Yonkee and Sutherland~\cite{Yonkee2016} implemented SIMT-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations on an Intel Xeon E5-2680 CPU and an Tesla K20 GPU.
-In evaluation of the thermochemical properties, speedups between \SIrange{8}{13}{$\times$} were on 16 CPU cores, and \SI{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU respectively.
-Curtis et al.~\cite{CurtisGPU:2017} implemented a different formulation of the same fifth-order implicit Runge--Kutta method, as well as two-fourth order exponential integration techniques paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on a Tesla C2075; the implicit Runge--Kutta method was roughly equivalent to VODE running on 12--38 Intel Xeon E5-4640 v2 CPU cores for two smaller chemical models with an integration time step of \SI{e-6}{$\sec$}.
+Yonkee and Sutherland~\cite{Yonkee2016} implemented SIMT-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs) on an Intel Xeon E5-2680 CPU and an Tesla K20 GPU.
+In evaluation of the thermo-chemical properties, speedups between \SIrange{8}{13}{$\times$} were on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU respectively.
+Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two-fourth order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on a Tesla C2075; the implicit Runge--Kutta method was roughly equivalent to VODE running on 12--38 Intel Xeon E5-4640 v2 CPU cores for two smaller chemical models with an integration time step of \SI{e-6}{$\sec$}.
 
@@
 Stone et al.,~\cite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
-The shallow-SIMD and SIMT-vectorization models were implemented in OpenCL and compared to an OpenMP baseline code that was deep SIMD-vectorized by simple compiler hints (a.k.a pragmas).
-The shallow-SIMD vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance was \SIrange{1.4}{1.6}{$\times$} slower than the OpenMP baseline due to thread-divergence concerns.
+The shallow-SIMD and SIMT-vectorization models were implemented in OpenCL and compared to an OpenMP baseline code that was deep SIMD-vectorized by simple compiler hints (i.e., \texttt{\#pragmas}).
+The shallow-SIMD vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance was only \SIrange{1.4}{1.6}{$\times$} faster than the OpenMP baseline due to thread-divergence concerns.
 
-This work will build upon our previous analytical chemical kinetic Jacobian code,~\cite{Niemeyer:2016aa} in order to:
+This work will build upon our previous analytical chemical kinetic Jacobian code, \texttt{pyJac}~\cite{Niemeyer:2016aa}, in order to:
 \begin{itemize}
@@
  \item Investigate the performance for a wide range chemical kinetic models, specifically looking at speedups due to SIMD\slash SIMT-vectorization as well as relative to the previous version of \texttt{pyJac}, and finally
- \item Discuss future extensions to this work as well as the most promising directions for SIMD\slash SIMT vectorization in reactive flow simulations
+ \item Discuss future extensions to this work as well as the most promising directions for SIMD\slash SIMT vectorization in reactive-flow simulations
 \end{itemize}
@@
 
-In a \textit{shallow}-SIMD\slash SIMT vectorization, each SIMD-lane (or SIMT-thread) in a compute-unit evaluates thermo-chemical source terms for a different thermo-chemical state.
-If the data is stored in ``F''-order, the SIMD-lanes\slash SIMT-thread accessing $[C]_{1, j}\ldots[C]_{L, j})$, the concentration of species $j$ for states $1, 2,\ldots L$---the size of the vector lane for SIMD or the number of threads in a SIMT warp---will load sequential locations in memory, however the first $(j+1)$th species concentration, $[C]_{1, j+1}$ will be $N_{\text{state}}$ memory locations away; this increases the likelihood of cache-misses on the CPU~\cite{gray2000rules}, but conversely is well suited to coalesced memory accesses on the GPU~\cite{NVIDIA:2018}.
-In a \textit{deep}-SIMD\slash SIMT vectorization, a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermo-chemical source terms for a single thermo-chemical state, thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + L}$---the species concentrations for state 1, species $j \ldots j + L$---will access sequential memory locations if the data is stored in ``C''-order.
+In a \textit{shallow}-SIMD\slash SIMT vectorization (also referred to as ``per-thread'' in previous works~\cite{Stone:2013aa} using GPUs), each SIMD-lane (or SIMT-thread) in a compute-unit evaluates thermo-chemical source terms\slash chemical kinetic Jacobian for a different thermo-chemical state.
+If the data is stored in ``F''-order, the SIMD-lanes\slash SIMT-thread accessing $[C]_{1, j}\ldots[C]_{L, j}$, the concentration of species $j$ for states $1, 2,\ldots L$---the size of the vector lane for SIMD or the number of threads in a SIMT warp---will load sequential locations in memory.
+The first $(j+1)$th species concentration however, $[C]_{1, j+1}$, will be $N_{\text{state}}$ memory locations away; this increases the likelihood of cache-misses on the CPU~\cite{gray2000rules}, but conversely is well suited to coalesced memory accesses on the GPU~\cite{NVIDIA:2018}.
+
+In a \textit{deep}-SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works), a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermo-chemical source terms for a single thermo-chemical state, thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + L}$---the species concentrations for state 1, species $j \ldots j + L$---will access sequential memory locations if the data is stored in ``C''-order.
 Further, in the ``C''-ordering the furthest difference between any two species concentrations within the same thermo-chemical state is at most $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU, but leads to uncoalesced memory-accesses on the GPU.
@@
 \Cref{t:cpus} shows the number of cores, vector instruction set and model of the CPUs used in this work; each CPU had both \texttt{v16.1.1} of the Intel OpenCL runtime~\cite{intelopencl:2018} and \texttt{v1.0} of the POCL~\cite{poclIJPP} runtime installed, each enabling OpenCL \texttt{v1.2} execution.
+Additionally, \texttt{v5.0} of the LLVM\slash\texttt{clang}~\cite{Lattner:2004:LCF:977395.977673} compiler chain was installed on all machines to enable use of POCL.
 In addition,~\cref{t:gpus} lists the model, number of CUDA cores and NVIDIA driver of each of the GPUs used in this work.
@@
 \end{tabular}
+\caption{The Intel CPU configurations used in this study. The identifier field will be used as a shorthand descriptor in the performance plots to quickly identify the CPU type.}
 \label{t:cpus}
-\caption{The Intel CPU configurations used in this study. The identifier field will be used as a shorthand descriptor in the performance plots to quickly identify the CPU type.}
 \end{table}
@@
 \end{tabular}
+\caption{The NVIDIA GPU configurations used in this study.  NVIDIA's OpenCL runtime is provided with the graphic driver, rather than any specific version of CUDA.
 \label{t:gpus}
-\caption{The NVIDIA GPU configurations used in this study.  NVIDIA's OpenCL runtime is provided with the graphic driver, rather than any specific version of CUDA.
 Again, the identifier field will be used as a shorthand descriptor during analysis of the performance results.}
@@
 If the error from P-Log reactions are not considered in~\cref{e:rel_err}, the forward and reverse ROP falls to \num{5.44e-08} and \num{1.59e-07} respectively.
-We note that this discrepency does not imply any incorrectness in either \texttt{pyJac} or Cantera---in fact, the error is still well within the proscribed tolerances in~\cref{e:rel_err}---but merely emphasizes the effect small code-changes can have on the accumulation of floating point errors.
+We note that this discrepancy does not imply any incorrectness in either \texttt{pyJac} or Cantera---in fact, the error is still well within the proscribed tolerances in~\cref{e:rel_err}---but merely emphasizes the effect small code-changes can have on the accumulation of floating point errors.
 
@@
 In~\cref{T:source_error}, we see that $E_{\varepsilon}$ is very similar in magnitude to $E_R$ in all cases, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
-The molar species production rate error norm is similar in magnitude to that of the net ROP, however the temperature and state parameter rate errors again increases as the error in species rates is amplified by multiplication with thermodynamic properties.
+The molar species production rate error norm is similar in magnitude to that of the net ROP, however the temperature and state parameter rate errors again increases as the error in net species production rates is amplified by multiplication with thermodynamic properties.
 Again, we note that the above discussion does not imply that discrepencies in computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
@@
 
-As in our previous work~\cite{Niemeyer:2016aa}, the correctness of the Jacobian codes generated by \texttt{pyJac} was determined by comparison to Jacobians obtained by automatic differentiation of the source rate subroutines via the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
-The reasons for this choice are explained fully in~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with relatively programming little effort.
-The discrepency between the analytical and automatic differentiated Jacobians for thermochemical state $k$, denoted by $\mathcal{J}_k$ and $\hat{\mathcal{J}}_k$ respectively, is determined by the relative error Frobenius norm over all Jacobian indicies $i, j$:
+As in our previous work~\cite{Niemeyer:2016aa}, the correctness of the Jacobian codes generated by \texttt{pyJac} was determined by comparison to the Jacobian obtained by automatic differentiation of the source rate subroutines generated by \texttt{pyJac} via the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
+The reasons for this choice are explained fully in~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with minimal additional programming little effort.
+The discrepancy between the analytical and automatic differentiated Jacobians for thermo-chemical state $k$, denoted by $\mathcal{J}_k$ and $\hat{\mathcal{J}}_k$ respectively, is determined by the relative error Frobenius norm over all Jacobian indicies $i, j$:
 \begin{equation}
@@
 
-To avoid large relative errors of small nonzero Jacobian elements due to accumulation of floating point error, the Frobenius norm of the automatically differentiated Jacobian is calculated over all thermochemical states $k$:
+To avoid large relative errors of small nonzero Jacobian elements due to accumulation of floating point error, the Frobenius norm of the automatically differentiated Jacobian is calculated over all thermo-chemical states $k$:
 \begin{equation}
@@
 
-\Cref{T:error} presents the error statistics of \texttt{pyJac} over the tested PaSR thermochemical condition database.
+\Cref{T:error} presents the error statistics of \texttt{pyJac} over the tested PaSR thermo-chemical condition database.
 For all tested cases, \todo{need srv2 nvidia tests to really confirm here} the maximum filtered error norm is less than unity.

### 2018-02-16 derivations.tex
add sparsity characterization

@@
 \subsection{State variables}
+\label{s:state}
 The state vector for this derivation consists of the temperature, a thermodynamic state parameter (pressure or volume) and the number of moles of all species except the last in the model:
@@
 \begin{align}
+\label{e:last_species}
 n &= \frac{V P}{T \Ru} = \sum_{i=1}^{\ns}{n_i}\; , \nonumber \\
@@
 \subsection{Thermo-chemical source terms}
+\label{s:source}
 The evolution of the thermo-chemical state variables of this system is described by a set of ordinary-differential equations:
@@
 \subsubsection{Third-body enhanced reactions}
+\label{s:dci_thd}
 For a third-body enhanced reaction~\cref{e:thd_mix} is differentiated with respect to temperature to give:

### 2018-02-16 paper.tex
add sparsity characterization

@@
 \NewDocumentCommand{\mathindent}{O{space}}{\hphantom{\mathrel{#1}}}
+
+% set figure path
+\graphicspath{{./figures/}}
 
@@
 This new version of \texttt{pyJac} is capable of evaluating the thermo-chemical source terms for using both the constant-pressure (\conp) or constant-volume (\conv) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
-In this section, we will only outline a brief summary of the system evaluated by \texttt{pyJac}; the interested reader is referred to our complete derivations, attached as supplimental material to this document.
+In this section, we will only outline a brief summary of the system evaluated by \texttt{pyJac}; the interested reader is referred to our complete derivations, attached as supplemental material to this document.
 
@@
 where $T$ is the temperature, $V$ and $P$ the volume and pressure respectively, and $n_j$ the number of moles of the $j$th species in the model (containing $\ns$ total species).
+
+This state vector---inspired by~\cite{SCHWER2002270}---has a number of features that recommend its use.
+First, the state vector results in highly sparse chemical kinetic Jacobians as will be detailed in~\cref{S:sparsity}.
+Second, mass is explicitly conserved in this formulation owing to the fact that the moles of the final species is calculated from the ideal gas law, and the rate of change of the final species from conservation of mass (see \cref{deriv-s:state,deriv-s:source} of the supplemental material respectively); i.e., system is not over-constrained.
+Finally, the chemical kinetic Jacobian changes relatively little between the \conp~and \conv~forms, making maintaining the code-base much simpler.
+Although most current combustion codes do not use species moles as a state variable, to\slash from more common mass\slash mole-fractions and moles is straightforward, and once inside integration of an chemical kinetic initial value problem, the choice of variables no longer matter.
+
 The evolution of the thermo-chemical state vector is described by a set of chemical kinetic ordinary differential equations:
@@
 \end{equation}
-where $\nu_{k, i}$ is the net stoichiometic coefficient of species $k$ in reaction $i$, $R_{i}$ the net rate of progress of reaction $i$ (described in~\cref{deriv-s:rop_net} of the supplimentary material), and $c_{i}$ the pressure modification term i.e., for third-body (see~\cref{deriv-s:thdbody} of the supplimental), or fall-off\slash chemically-activated reactions (\cref{deriv-s:falloff} of the same).
+where $\nu_{k, i}$ is the net stoichiometic coefficient of species $k$ in reaction $i$, $R_{i}$ the net rate of progress of reaction $i$ (described in~\cref{deriv-s:rop_net} of the supplemental), and $c_{i}$ the pressure modification term i.e., for third-body (see~\cref{deriv-s:thdbody} of the supplimental), or fall-off\slash chemically-activated reactions (\cref{deriv-s:falloff} of the same).
 
@@
 
+\subsection{Sparsity patterns}
+\label{S:sparsity}
+
+In general, the Jacobian matrices generated by \texttt{pyJac} are largely sparse with non-zero entries mainly corresponding to species that participate in the same reaction (or non-default efficiency third-body species in a reaction), and mostly dense rows \slash columns corresponding to the temperature and thermodynamic state parameter.
+However, the explicit-mass conservation formulation of \texttt{pyJac} can introduce additional non-zero entries in two ways.
+First, if the last-species in the model (i.e., the bath-gas) participates directly in a reaction, the derivative of the forward or reverse rate of progress with respect to any other species is non-zero (see~\cref{deriv-s:dri_dnj}).
+Similarly, if the last-species has a third-body efficiency not equal to the default (unity), this will again force all the derivative of the pressure-modification term with respect to other species to be non-zero (see~\cref{deriv-s:dci_thd}).
+Either case will result in an undesirable fully dense Jacobian row for all species that have a non-zero net stoichiometic coefficient in such a reaction.
+
+However, \texttt{pyJac} allows the user to simply ignore these derivatives (via a command-line switch) to and avoid any negative adverse affects on the Jacobian sparsity.
+The rational behind this choice is that many common implicit integration techniques (e.g., CVODE~\cite{cvode:2.8.2}) used in solution of chemical kinetic initial value problems are formulated around the assumption that the supplied Jacobian is approximate; this allows the Jacobian and its LU factorization to be reused for multiple internal integration time-steps, accelerating the solution process.
+For such solvers, there is no need for the exact form of the Jacobian and the so-called ``appropriate'' form is preferable.
+Hence, in this section we will detail the sparsity of both forms of the Jacobian for the chemical models tested.
+
+In~\cref{F:jac_sparsity}, a graphical representation of the sparsity pattern for GRI-Mech 3.0 is displayed.
+In particular we note that~\cref{F:jac_sparse_approx} is has several rows that are no longer fully-dense, as result of its approximate form -- these rows correspond to species directly interacting with \ce{N2}, largely in GRI-Mech 3.0's \ce{NOx} chemistry reactions.
+Table~\cref{T:jac_sparsity} shows the density of the exact and approximate Jacobians for all mechanisms tested in this work.
+The smallest model, \ce{H2}\slash\ce{CO} is very dense with \textasciitilde\SI{70}{\percent} of the Jacobian entries non-zero however this drops to just \textasciitilde\SI{50}{\percent} for GRI-Mech 3.0, and continues to decrease to \textasciitilde\SI{27}{\percent} for USC-Mech II and just \textasciitilde\SI{10}{\percent} for the isopentanol model.
+The approximate Jacobian assumption drops the density of Jacobian by \textasciitilde\SIrange{3}{7}{\percent} for all models.
+
+\begin{figure}[htb]
+  \centering
+  \begin{subfigure}{0.45\linewidth}
+      \includegraphics[width=\textwidth]{ch4_sparsity_exact.pdf}
+      \caption{The ``exact'' Jacobian.}
+      \label{F:jac_sparse_exact}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{0.45\linewidth}
+      \includegraphics[width=\textwidth]{ch4_sparsity_approximate.pdf}
+      \caption{The ``approximate'' Jacobian.}
+      \label{F:jac_sparse_approx}
+  \end{subfigure}
+  \caption{A graphical representation of the sparsity pattern of the chemical kinetic Jacobian generated by \texttt{pyJac} for GRI-Mech 3.0.
+	   Black squares indicate a non-zero Jacobian entry, while white square correspond to an empty entry.
+	   The numbers indicate the index of the entry in the state vector.}
+  \label{F:jac_sparsity}
+\end{figure}
+
+\begin{table}[tbp]
+\centering
+\begin{tabular}{@{}l S[table-format=2.1] S[table-format=2.1]@{}}
+\toprule
+Model                 & \multicolumn{1}{c}{Exact Jacobian Density} & \multicolumn{1}{c}{Approximate Jacobian Density} \\
+\midrule
+\ce{H2}\slash \ce{CO} & \SI{71.4}{\percent} & \SI{68.4}{\percent} \\
+GRI-Mech 3.0          & \SI{56.7}{\percent} & \SI{49.8}{\percent} \\
+USC-Mech II           & \SI{28.2}{\percent} & \SI{26.4}{\percent} \\
+\ce{iC5H11OH}         & \SI{11.5}{\percent} & \SI{7.98}{\percent} \\
+\bottomrule
+\end{tabular}
+\caption{The density of the exact and approximate Jacobians generated by \texttt{pyJac} for the various models studied.}
+\label{T:jac_sparsity}
+\end{table}
+
 \subsection{Performance}

### 2018-02-16 paper.tex
clarify testing platforms, add OpenCL discussion & link MWE

@@
 \section{Results and Discussion}
-\subsection{Validation and performance study platforms}
+\subsection{Testing platforms and practical notes on OpenCL usage}
 
@@
 NVIDIA's OpenCL runtime is bundled with the NVIDIA driver~\cite{NVIDIA:2018}, hence we use driver version to specify the OpenCL runtime version.
+
+Finally,~\cref{t:platforms} lists the platforms and vectorization\slash execution patterns they are capable of running.
+It is important to note, that while OpenCL provides an easy way to enable cross-platform execution, there are some serious potential pit-falls in its use.
+Typically speaking, the closed source OpenCL runtimes tested in this work (Intel and NVIDIA) contain more bugs that result in compilation failures or simply incorrect vectorized machine code; further, these runtimes (in our experience) tend to be less responsive to fixing said bugs, with relatively infrequent new releases (or in NVIDIAs case, change-logs of bugfixes).
+As demonstration, we've created a minimum working example~\cite{nvidia_mwe} that demonstrates a failure of NVIDIA's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla K40m GPU; by simply changing to another runtime (e.g., Intel), with no other code changes, the correct result can be found.
+This provides a particularly vexing problem for the programmer, as there is often little that can be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
+
+Further, the Intel and NVIDIA OpenCL runtimes lack a key feature---implementations of atomic operations on double precision variables---that are currently necessary for \texttt{pyJac} to run deep-vectorized code.
+On the other hand, POCL is an open source OpenCL runtime that works on all CPU types tested in this work, and does implement these atomic operations; it is our experience that the POCL community is very responsive to bug reports and user outreach.
+However, POCL's implicit vectorization module---which utilizes the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to attempt to translate OpenCL code to vectorized machine code---typically fails to achieve much speedup, if any.
+Thus POCL is a particularly useful tool for validation, but not necessarily for performance studies.
+Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the only significant code change is building the wrapper that initializes\slash transfers memory and calls the kernel.
+Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py}) is a key priority going forwards.
+This discussion will be expanded upon in~\cref{s:future} to highlight the most promising future directions.
 
@@
 \end{table}
+
 \begin{table}
@@
 Tesla C2075 & \num{384.81} & \num{448} & \texttt{C2075} \\
-Tesla K40m & \num{375.26} & \num{2880} & \texttt{K40m} \\
+Tesla K40m & \num{387.26} & \num{2880} & \texttt{K40m} \\
 \bottomrule
@@
 \caption{The NVIDIA GPU configurations used in this study.  NVIDIA's OpenCL runtime is provided with the graphic driver, rather than any specific version of CUDA.
+Again, the identifier field will be used as a shorthand descriptor during analysis of the performance results.}
 \label{t:gpus}
-Again, the identifier field will be used as a shorthand descriptor during analysis of the performance results.}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{@{}l c c c@{}}
+\toprule
+Platform & Parallel & Shallow Vectorization & Deep Vectorization \\
+\midrule
+OpenMP & X & -- & -- \\
+POCL OpenCL & X & X & X \\
+Intel OpenCL & X & X & -- \\
+NVIDIA OpenCL & -- & X & -- \\
+\bottomrule
+\end{tabular}
+\caption{The various platforms used in this study and the execution \slash vectorization patterns they are capable of running}
+\label{t:platforms}
 \end{table}
@@
 In particular we note that~\cref{F:jac_sparse_approx} is has several rows that are no longer fully-dense, as result of its approximate form -- these rows correspond to species directly interacting with \ce{N2}, largely in GRI-Mech 3.0's \ce{NOx} chemistry reactions.
-Table~\cref{T:jac_sparsity} shows the density of the exact and approximate Jacobians for all mechanisms tested in this work.
+\cref{T:jac_sparsity} shows the density of the exact and approximate Jacobians for all mechanisms tested in this work.
 The smallest model, \ce{H2}\slash\ce{CO} is very dense with \textasciitilde\SI{70}{\percent} of the Jacobian entries non-zero however this drops to just \textasciitilde\SI{50}{\percent} for GRI-Mech 3.0, and continues to decrease to \textasciitilde\SI{27}{\percent} for USC-Mech II and just \textasciitilde\SI{10}{\percent} for the isopentanol model.
@@
 
+\section{Future directions}
+\label{s:future}
+NVIDIA 375.26 is a bad driver, yo.
+
 \section{Conclusions}

### 2018-02-17 paper.tex
update error stats

@@
 
-\todo{rerun, especially H2 error}
 \begin{table}[tbp]
@@
 \midrule
-\ce{H2}\slash \ce{CO} & \num{900900}                      & \num{6.431e+18}      & \num{8.083e-1}  & \num{1.907e-5} \\
+\ce{H2}\slash \ce{CO} & \num{900900}                      & \num{6.431e+18}      & \num{8.084e-1}  & \num{1.907e-5} \\
 GRI-Mech 3.0          & \num{450900}                      & \num{7.783e+19}      & \num{1.890e-7}  & \num{1.877e-7} \\
-USC-Mech II           & \num{91800}                       & \num{2.830e+21}      & \num{5.567e-3}  & \num{1.704e-7} \\
-\ce{iC5H11OH}         & \num{450900}                      & \num{2.733e+26}      & \num{1.136e-3}  & \num{2.764e-5} \\
+USC-Mech II           & \num{91800}                       & \num{2.830e+21}      & \num{1.540e-3}  & \num{1.704e-7} \\
+\ce{iC5H11OH}         & \num{450900}                      & \num{2.733e+26}      & \num{1.363e-3}  & \num{2.764e-5} \\
 \bottomrule
@@
 \Cref{T:error} presents the error statistics of \texttt{pyJac} over the tested PaSR thermo-chemical condition database.
-For all tested cases, \todo{need srv2 nvidia tests to really confirm here} the maximum filtered error norm is less than unity.
+For all tested cases, the maximum filtered error norm is less than unity.
 The filtered error norm is largest for the \ce{H2}\slash\ce{CO} model with the more stringent tolerance ($\mathcal{C} = 10^{20}$).

### 2018-02-17 scripts/jac_error.py

@@ -1,4 +1,3 @@
-import os
 import numpy as np
 from load_error_files import print_error
 
@@ -11,10 +10,8 @@ def update(err_dict, err, filename=None, **kwargs):
         if ('amax' in name or 'value' in name) and name != 'jac_threshold_value':
             continue
         assert 'fdjac' not in filename
-        if 'nvidia' in filename:
-            from time import ctime
-            print('Known driver issue:', filename, ctime(os.path.getmtime(filename)))
-            continue
+        if err_dict['jac_thresholded_15'] > 1e0:
+            print(filename)
         # special handling for threshold
         if name == 'jac_threshold_value':
             sparse = 'sparse' if ('sparse' in filename) else 'dense'
@@ -22,7 +19,7 @@ def update(err_dict, err, filename=None, **kwargs):
             val = err[name]
             name += '_{}_{}'.format(sparse, conp)
             if name in err_dict:
-                assert np.isclose(err_dict[name], np.asscalar(val))
+                assert np.isclose(err_dict[name], np.asscalar(val), rtol=5e-4)
             err_dict[name] = np.asscalar(val)
             continue
         err_dict[name] = np.maximum(err_dict[name], err[name])
@@ -43,7 +40,8 @@ def printer(err_dict):
     mean_thr = mean_thr[0] / mean_thr[1]
 
     print('\n'.join(
-        ['{}: {:.15e}'.format(k, v) for k, v in err_dict.items()]))
+        ['{}: {:.15e}'.format(k, v) for k, v in err_dict.items()
+         if 'thresholded' in k]))
     print('mean_threshold: {:.15e}'.format(mean_thr))
 
 

### 2018-02-17 scripts/load_error_files.py

@@ -3,28 +3,34 @@ import numpy as np
 import yaml
 import cantera as ct
 from collections import defaultdict
+from os.path import join, dirname, abspath, isdir, isfile
 
-script_dir = os.path.dirname(os.path.abspath(__file__))
-error_path = os.path.join(script_dir, 'error_checking')
+script_dir = dirname(abspath(__file__))
+error_path = join(script_dir, 'error_checking')
 
 
 def get_error_files(ftype):
-    mechs = [os.path.join(error_path, x) for x in os.listdir(
-        error_path) if os.path.isdir(os.path.join(error_path, x))]
+    mechs = [join(error_path, x) for x in os.listdir(
+        error_path) if os.path.isdir(join(error_path, x))]
+
+    def __load_from_path(path):
+        return [join(path, x) for x in os.listdir(path) if isfile(join(path, x))
+                and x.endswith('.npz') and ftype in x]
 
     # get file list
     files = {}
     for mech in mechs:
-        files[mech] = [os.path.join(mech, x) for x in os.listdir(
-            mech) if os.path.isfile(os.path.join(mech, x)) if x.endswith('.npz')
-            and ftype in x]
+        files[mech] = __load_from_path(mech)
+        # check for subdirs
+        for d in [join(mech, x) for x in os.listdir(mech) if isdir(join(mech, x))]:
+            files[mech].extend(__load_from_path(d))
 
     return files
 
 
 def add_rxn_specific_inds(mech_info, rxn_type, rxn_str, path):
     # load gas
-    gas = ct.Solution(os.path.join(path, mech_info['mech']))
+    gas = ct.Solution(join(path, mech_info['mech']))
     # get fwd reaction indicies
     mech_info['rop_fwd_' + rxn_str] = np.array([i for i, r in enumerate(
         gas.reactions()) if not isinstance(r, rxn_type)])
@@ -43,7 +49,7 @@ def run_error_calcs(ftype, updater):
         err_dicts[mech_name] = defaultdict(lambda: 0)
 
         # get mech info
-        mech_info = os.path.join(mech, mech_name + '.yaml')
+        mech_info = join(mech, mech_name + '.yaml')
         with open(mech_info, 'r') as file:
             mech_info = yaml.load(file.read())
             if mech_info['n_cheb']:

### 2018-02-17 paper.tex
update

@@
 \centering
-\begin{tabular}{@{}l l l l l@{}}
+\begin{tabular}{@{}l l l@{}}
 \toprule
-CPU Model & Instruction Set & Vector Width & Cores & Identifier \\
+CPU Model        & Xeon X5650      & E5-2690 V3     \\
 \midrule
-Xeon X5650 & SSE4.2 & \SI{128}{\bit} & $2 \times 6$ & \texttt{sse4.2} \\
-E5-2690 V3 & AVX2 & \SI{256}{\bit} & $2 \times 12$ & \texttt{avx2} \\
+Instruction Set  & SSE4.2 	   & AVX2 	    \\
+Vector Width     & \SI{128}{\bit}  & \SI{256}{\bit} \\
+Cores            & $2 \times 6$    & $2 \times 12$  \\
+Identifier       & \texttt{sse4.2} & \texttt{avx2}  \\
+OpenCL Version   & \num{1.2}       & \num{1.2}      \\
 \bottomrule
@@
 \centering
-\begin{tabular}{@{}l l l l@{}}
+\begin{tabular}{@{}l l l l l@{}}
 \toprule
-NVIDIA Model & Driver Version & CUDA Cores & Identifier \\
+NVIDIA Model   & Tesla C2075    & Tesla K40m    \\
 \midrule
-Tesla C2075 & \num{384.81} & \num{448} & \texttt{C2075} \\
-Tesla K40m & \num{387.26} & \num{2880} & \texttt{K40m} \\
+Driver Version & \num{384.81}   & \num{387.26}  \\
+CUDA Cores     & \num{448}      & \num{2880}    \\
+Identifier     & \texttt{C2075} & \texttt{K40m} \\
+OpenCL Version & \num{1.1}	& \num{1.2}	\\
 \bottomrule
@@
 The rational behind this choice is that many common implicit integration techniques (e.g., CVODE~\cite{cvode:2.8.2}) used in solution of chemical kinetic initial value problems are formulated around the assumption that the supplied Jacobian is approximate; this allows the Jacobian and its LU factorization to be reused for multiple internal integration time-steps, accelerating the solution process.
-For such solvers, there is no need for the exact form of the Jacobian and the so-called ``appropriate'' form is preferable.
+For such solvers, there is no need for the exact form of the Jacobian and the so-called ``approximate'' form is preferable.
 Hence, in this section we will detail the sparsity of both forms of the Jacobian for the chemical models tested.
@@
 \label{S:results}
-The performance studies in this work were run on a variety of different machines.
+The performance studies in this work were run on the platforms listed in~\cref{t:cpus,t:gpus}.
+Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
+The wrapping code---responsible for initializing\slash transfering memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \texttt{avx2}\slash\texttt{K40m} platforms, and \texttt{gcc 4.8.5} on the \texttt{sse4.2}\slash\texttt{C2075} machines.
+The optimization level ``\texttt{-O3}'' was used in all cases, and no ``fast-math'' OpenCL optimizations were enabled.
+Additionally, the ``exact'' form of the Jacobian (as opposed the the ``appropriate'' form discussed in~\cref{S:sparsity}) was used in all cases.
 
-four ten-core \SI{2.2}{\giga\hertz} Intel Xeon E5-4640 v2 CPUs with \SI{20}{\mega\byte} of L3 cache memory, installed on an Ace Powerworks PW8027R-TRF+ with a Supermicro X9QR7-TF+/X9QRi-F+ baseboard.
-Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
-All code was compiled with \texttt{gcc 4.8.5} and ``\texttt{-O3}'' optimization, running on  of the
+\Cref{f:source} shows the runtime of \texttt{pyJac} on the test platforms.
 
@@
    \centering
-  \begin{minipage}[t]{0.48\linewidth}
-    \includegraphics[width=\textwidth]{SIMD_SIMT_Order_comparison}
-    \caption{Runtime in milliseconds per thermo-chemical state of shallow-SIMD and SIMT-vectorization on a single CPU core for both ``C'' (filled symbols) and ``F'' (empty symbols) data ordering.}
-    \label{F:SIMDComp}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[t]{0.48\linewidth}
-    \includegraphics[width=\textwidth]{SIMD_Vecwidth_comparison}
-    \caption{Speedup of a ``C''-ordered shallow-SIMD vectorization on a single CPU core when varying the supplied vector-width.}
-    \label{F:Veccomp}
-  \end{minipage}
+  \begin{subfigure}{0.45\linewidth}
+      \includegraphics[width=\textwidth]{intel_source.pdf}
+      \caption{The ``exact'' Jacobian.}
+      \label{F:intel_source_noorm}
+  \end{subfigure}
+ \label{f:source}
 \end{figure}
-\begin{figure}[htb]
-  \centering
-  \begin{minipage}[t]{0.5\linewidth}
-    \includegraphics[width=\textwidth]{SIMD_scaling}
-    \caption{Strong parallel scaling efficiency of a ``C''-ordered shallow-SIMD vectorization using a vector-width of 16.}
-    \label{F:SIMDscale}
-  \end{minipage}
-\end{figure}
-
-In Fig.~\ref{F:SIMDComp}, the performance of the shallow-SIMD and SIMT vectorization are compared on a single CPU core for both ``C'' and ``F'' memory formats.
-The ``C''-ordered SIMD\slash SIMT vectorizations show a speedup over the equivalent ``F''-ordered code by \SIrange{1.46}{2.00}{$\times$} and \SIrange{1.61}{2.38}{$\times$} respectively.
-This demonstrates that enhanced data-locality of the ``C''-ordering for CPU caching---discussed in Sec.~\ref{S:data}---is key to high performance on the CPU, and suggests that a deep vectorization would see even greater accelerations.
-The shallow vectorization is consistently faster than the SIMT vectorization, by \SIrange{1.73}{1.90}{$\times$} and \SIrange{2.00}{2.45}{$\times$} for the ``C'' and ``F'' memory formats respectively.
-This is less than the theoretical maximum speedup on this CPU, which can process \SI{256}{\bit} AVX instructions (i.e. four double-precision operations) concurrently.
-Most likely this is due to the non-sequential memory access patterns of the shallow vectorization discussed previously, but could also be caused in part by inefficiencies in translation of the OpenCL code to AVX code, a process handled by the Intel Opencl Runtime.
-
-The effect of changing the vector-width supplied to a ``C''-ordered shallow-SIMD vectorization on a single CPU core is investigated in Fig.~\ref{F:Veccomp}.
-Significant speedups are seen with larger vector-widths for the smaller models, however as the model size increases data must be loaded from increasingly distant memory locations and the performance of all three vector-widths becomes similar.
-
-Finally, the strong parallel scaling efficiency is defined as:
-\begin{equation}
-S = \frac{t_1}{N t_N}
-\end{equation}
-where $t_1$ is the mean wall-clock runtime of thermo-chemical source terms evaluation on one core and $t_N$ the mean runtime on $N$ cores.
-Fig.~\ref{F:SIMDscale} examines the strong parallel scaling efficiency for all four models with a ``C''-ordered shallow SIMD-vectorization and vector-width of 16.
-Interestingly the efficiency of the hydrogen model decreases the fastest while the larger models have significantly larger scaling efficiencies for more than 4 cores.
-Additionally, it appears that for a fixed number of cores, the scaling efficiency increases with model size.
-This suggests that evaluation of the chemical-source terms for a single thermo-chemical state for the smaller models may not saturate the processing power of the CPU, some efficiency may be lost due to context switching within OpenCL itself.
-Additionally, OpenCL allows assigning of work to devices by shared memory resources, e.g., a L3 cache, which was not pursued in this study for simplicity; this should be investigated in future work to determine the effect on scaling efficiency.
 

### 2018-02-17 scripts/general_plotting.py

@@ -66,12 +66,13 @@ def process_data(plotdata, plot, reacs_as_x=True,
 
 def get_marker_and_label(plot_ind=None, marker_func=None, label=None, marker=None,
                          labels=[]):
+    size = None
     if label is not None and marker is not None:
         name = label
-        marker, hollow, color = marker
+        marker, hollow, color, size = marker
     elif marker_func is not None:
         name = labels[plot_ind]
-        marker, hollow, color = marker_func(name)
+        marker, hollow, color, size = marker_func(name)
     elif plot_ind is not None:
         assert labels
         marker, hollow = ps.marker_wheel[plot_ind % len(ps.marker_wheel)]
@@ -81,7 +82,11 @@ def get_marker_and_label(plot_ind=None, marker_func=None, label=None, marker=Non
         marker, hollow = ps.marker_dict[plot]
         color = ps.color_dict[plot]
         name = plot
-    return marker, hollow, color, name
+
+    if size is None:
+        size = ps.marker_style['size'] if not hollow else ps.clear_marker_style[
+            'size']
+    return marker, hollow, color, name, size
 
 
 def plot(x_vals, y_vals, err_vals, minx=None, miny=None, maxx=None, maxy=None,
@@ -89,7 +94,7 @@ def plot(x_vals, y_vals, err_vals, minx=None, miny=None, maxx=None, maxy=None,
          label=None, marker=None):
     """Plot performance as a function of reaction count."""
 
-    marker, hollow, color, name = get_marker_and_label(
+    marker, hollow, color, name, size = get_marker_and_label(
         plot_ind, marker_func, label, marker, labels)
 
     argdict = {'x': x_vals,
@@ -100,11 +105,9 @@ def plot(x_vals, y_vals, err_vals, minx=None, miny=None, maxx=None, maxy=None,
                }
     argdict['color'] = color
     argdict['markeredgecolor'] = color
-    if not hollow:
-        argdict['markersize'] = ps.marker_style['size']
-    else:
+    argdict['markersize'] = size
+    if hollow:
         argdict['markerfacecolor'] = 'None'
-        argdict['markersize'] = ps.clear_marker_style['size']
 
     if plot_std:
         argdict['yerr'] = err_vals

### 2018-02-17 scripts/make_mech_data.py

@@ -1,0 +1,40 @@
+import argparse
+import cantera as ct
+import yaml
+import os
+import ntpath
+
+parser = argparse.ArgumentParser()
+parser.add_argument('-m', '--mech',
+                    type=str,
+                    help='The mechanism to generate the data file for')
+parser.add_argument('-n', '--name',
+                    type=str,
+                    default=None,
+                    required=False,
+                    help='The desired mechanism name for plotting. '
+                         'If not specified, the mechanism filename is used')
+parser.add_argument('-o', '--outpath',
+                    type=str,
+                    default='.',
+                    required=False,
+                    help='The path to place the generated mechanism data file in.')
+
+opts = parser.parse_args()
+gas = ct.Solution(opts.mech)
+mech = ntpath.basename(opts.mech)
+name = opts.name if opts.name is not None else mech[:mech.index('.cti')]
+
+outfile = os.path.join(opts.outpath, name + '.yaml')
+with open(outfile, 'w') as file:
+    yaml.dump(dict(
+        mech=mech,
+        name=name,
+        n_reactions=gas.n_reactions,
+        n_species=gas.n_species,
+        n_reversible=len([x for x in gas.reactions() if x.reversible]),
+        n_cheb=len([x for x in gas.reactions()
+                    if isinstance(x, ct.ChebyshevReaction)]),
+        n_plog=len([x for x in gas.reactions()
+                    if isinstance(x, ct.PlogReaction)])
+        ), file, default_flow_style=False)

### 2018-02-17 scripts/sparse_plotter.py

@@ -1,0 +1,58 @@
+from __future__ import print_function, division
+
+from argparse import ArgumentParser
+import numpy as np
+import cantera as ct
+from pyjac.core.mech_interpret import read_mech_ct
+from pyjac.core.create_jacobian import determine_jac_inds
+from pyjac.loopy_utils.loopy_utils import JacobianType, RateSpecialization
+import matplotlib.pyplot as plt
+
+
+def plot(mech, jac_type, out=''):
+    # plot the sparsity patterns of our tested mechanisms
+
+    # get gas
+    gas = ct.Solution(mech)
+    _, specs, reacs = read_mech_ct(gas=gas)
+
+    # get indicies
+    info = determine_jac_inds(reacs, specs, RateSpecialization.fixed, jac_type)
+    inds = info['jac_inds']
+
+    # set grid for plotting
+    grid = np.zeros((gas.n_species + 1, gas.n_species + 1))
+    grid[inds['flat_C'][:, 0], inds['flat_C'][:, 1]] = 1
+
+    print('{}, Density: {:.3}%'.format(
+        str(jac_type), 100. * np.where(grid)[0].size / grid.size))
+
+    plt.imshow(grid, cmap='Greys',  interpolation='none')
+
+    if out:
+        if out.endswith('.pdf'):
+            out = out[:out.index('.pdf')]
+        out += '_' + str(jac_type)[str(jac_type).index('.') + 1:] + '.pdf'
+        plt.tight_layout()
+        plt.savefig(out)
+    else:
+        plt.show()
+
+
+if __name__ == '__main__':
+    parser = ArgumentParser()
+    parser.add_argument('-m', '--mech',
+                        type=str,
+                        required=True,
+                        help='The mechanism to plot')
+    parser.add_argument('-o', '--out',
+                        type=str,
+                        required=False,
+                        default='',
+                        help='The filename to store the generated plots in. '
+                             'If not supplied, the images will only be shown to the '
+                             'screen.')
+    opts = parser.parse_args()
+
+    plot(opts.mech, JacobianType.exact, opts.out)
+    plot(opts.mech, JacobianType.approximate, opts.out)

### 2018-02-18 paper.tex
updates

@@
 \newcommand{\Ru}{\mathcal{R}}
+\newcommand{\avx}{\texttt{avx2}}
+\newcommand{\sse}{\texttt{sse4.2}}
+\newcommand{\gpuold}{\texttt{C2075}}
+\newcommand{\gpunew}{\texttt{K40m}}
 
@@
 Cores            & $2 \times 6$    & $2 \times 12$  \\
-Identifier       & \texttt{sse4.2} & \texttt{avx2}  \\
+Identifier       & \sse & \avx  \\
 OpenCL Version   & \num{1.2}       & \num{1.2}      \\
@@
 CUDA Cores     & \num{448}      & \num{2880}    \\
-Identifier     & \texttt{C2075} & \texttt{K40m} \\
+Identifier     & \gpuold & \gpunew \\
 OpenCL Version & \num{1.1}	& \num{1.2}	\\
@@
 Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
-The wrapping code---responsible for initializing\slash transfering memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \texttt{avx2}\slash\texttt{K40m} platforms, and \texttt{gcc 4.8.5} on the \texttt{sse4.2}\slash\texttt{C2075} machines.
+The wrapping code---responsible for initializing\slash transfering memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx\slash\gpunew platforms, and \texttt{gcc 4.8.5} on the \sse\slash\gpuold machines.
 The optimization level ``\texttt{-O3}'' was used in all cases, and no ``fast-math'' OpenCL optimizations were enabled.
@@
 
-\Cref{f:source} shows the runtime of \texttt{pyJac} on the test platforms.
+\Cref{f:source} explores the performance of the source term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
+In~\cref{F:intel_source_noorm}, the mean runtime per initial condition is shown for both the~\avx and~\sse CPUs, using both Intel OpenCL and OpenMP.
+This normalization is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{T:error}}).
+For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the parallel (non-vectorized) Intel OpenCL code is slightly faster than OpenMP in all cases.\todo{rerun OpenMP w/ mtune=native}
+As expected, the \avx machine is moderately faster than the \sse CPU; additionally, the wide-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.
+In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is normalized per-machine, such that the speedup of the wide-vectorized OpenCL code can be readily assessed.
+On both machines, the unvectorized OpenCL code is \textasciitilde\SIrange{1.25}{1.5}{$\times$} faster than the baseline parallel OpenMP code.
+Additionally, the wide-vectorized OpenCL code is \textasciitilde\SIrange{2.5}{3}{$\times$} and \textasciitilde\SIrange{3.5}{4}{$\times$} faster than the OpenMP code for the \sse and \avx machines respectively.
+In contrast,~\cref{F:pocl_source} shows the runtime of deep, wide and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
+No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP is faster in both cases.
+The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work, however briefly speaking it is possible vectorization might be achieved on POCL via combination of relaxed math (``-cl-fast-relaxed-math''), use of explicit OpenCL vector types and future improvements to LLVM's vectorization module.
+This will be explored more in future work, but for the moment POCL is still quite useful as a validation tool.
 
@@
   \begin{subfigure}{0.45\linewidth}
+      \includegraphics[width=\textwidth]{intel_source_nonorm.pdf}
+      \caption{The mean runtime per condition for each chemical model using the Intel OpenCL and OpenMP on both CPUs studied; the names in the legend correspond to the identifiers listed in~\cref{t:cpus}}
+      \label{F:intel_source_noorm}
+  \end{subfigure}
+  \begin{subfigure}{0.45\linewidth}
       \includegraphics[width=\textwidth]{intel_source.pdf}
-      \caption{The ``exact'' Jacobian.}
-      \label{F:intel_source_noorm}
+      \caption{The speedup achieved over the baseline OpenMP parallelization by both the unvectorized and wide-vectorized Intel OpenCL codes for each chemical model.}
+      \label{F:intel_source}
+  \end{subfigure}
+  \\
+  \begin{subfigure}{0.45\linewidth}
+      \includegraphics[width=\textwidth]{pocl_source_nonorm.pdf}
+      \caption{The mean runtime per condition of the Portable OpenCL runtime compared to OpenMP parallelization for each chemical model.}
+      \label{F:pocl_source}
   \end{subfigure}

### 2018-02-18 paper.tex
start on performance

@@
   \centering
-  \begin{subfigure}[htp]{0.45\linewidth}
+  \begin{subfigure}[t]{0.45\linewidth}
       \includegraphics[width=\textwidth]{SIMD.pdf}
@@
   \hfill
-  \begin{subfigure}[htp]{0.45\linewidth}
+  \begin{subfigure}[t]{0.45\linewidth}
       \includegraphics[width=\textwidth]{SIMT.pdf}
@@
   \centering
-  \begin{subfigure}{0.45\linewidth}
+  \begin{subfigure}[t]{0.45\linewidth}
       \includegraphics[width=\textwidth]{ch4_sparsity_exact.pdf}
@@
   \hfill
-  \begin{subfigure}{0.45\linewidth}
+  \begin{subfigure}[t]{0.45\linewidth}
       \includegraphics[width=\textwidth]{ch4_sparsity_approximate.pdf}
@@
 No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP is faster in both cases.
-The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work, however briefly speaking it is possible vectorization might be achieved on POCL via combination of relaxed math (``-cl-fast-relaxed-math''), use of explicit OpenCL vector types and future improvements to LLVM's vectorization module.
+The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work, however briefly speaking vectorization might be achieved on POCL via a combination of relaxed math (``-cl-fast-relaxed-math''), use of explicit OpenCL vector types and newer versions of LLVM (specifically, potential improvments to the vectorization module).
 This will be explored more in future work, but for the moment POCL is still quite useful as a validation tool.
@@
    \centering
-  \begin{subfigure}{0.45\linewidth}
+  \begin{subfigure}[t]{0.45\linewidth}
       \includegraphics[width=\textwidth]{intel_source_nonorm.pdf}
-      \caption{The mean runtime per condition for each chemical model using the Intel OpenCL and OpenMP on both CPUs studied; the names in the legend correspond to the identifiers listed in~\cref{t:cpus}}
+      \caption{The mean runtime per condition for each chemical model using the Intel OpenCL and OpenMP on both CPUs studied.}
       \label{F:intel_source_noorm}
   \end{subfigure}
-  \begin{subfigure}{0.45\linewidth}
+  \hfill
+  \begin{subfigure}[t]{0.45\linewidth}
       \includegraphics[width=\textwidth]{intel_source.pdf}
-      \caption{The speedup achieved over the baseline OpenMP parallelization by both the unvectorized and wide-vectorized Intel OpenCL codes for each chemical model.}
+      \caption{The speedup achieved over the baseline OpenMP parallelization by both the unvectorized and wide-vectorized Intel OpenCL codes.}
       \label{F:intel_source}
@@
   \\
-  \begin{subfigure}{0.45\linewidth}
+  \begin{subfigure}[t]{0.45\linewidth}
       \includegraphics[width=\textwidth]{pocl_source_nonorm.pdf}
-      \caption{The mean runtime per condition of the Portable OpenCL runtime compared to OpenMP parallelization for each chemical model.}
+      \caption{The mean runtime per condition of the Portable OpenCL runtime compared to OpenMP parallelization.}
       \label{F:pocl_source}
   \end{subfigure}
+ \caption{Mean runtimes per condition and speedups achieved by the various CPU OpenCL runtimes compared to OpenMP parallelization for each chemical model studied.
+The names in the legends correspond to the identifiers listed in~\cref{t:cpus}}
  \label{f:source}
+\end{figure}
+
+\Cref{f:source_permutate} shows the effect of changing data-ordering patterns as well as the \conp~or \conv~formulation on the performance of source term evaluation in \texttt{pyJac}.
+In~\cref{F:source_conpvsconv}, we see that choice of the constant-pressure or constant-volume formulation has little to no effect on runtime for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx machine.
+Generally speaking, the difference between the \conp~and \conv~formulations has little effect on performance regardless of CPU\slash GPU choice, or even for Jacobian evaluation which will be explored later.
+In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx machine; the speedup presented is normalized per language, i.e., the roughly \SIrange{1.5}{2}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-order OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
+Additionally, we note that the ``C''-ordered wide-vectorization in~\cref{F:source_cvsf} utilizes the vectorized-data ordering described in~\cref{S:data}.
+In general, the data-ordering has a much more profound affect on the vectorized-OpenCL codes; this is due to the nature of SIMD-operations, which are much more efficient when operating on contiguous (in memory) data rather than incurring an expensive memory ``gather'' operation.
+The ``C''-ordered unvectorized OpenCL code (which uses the default unvectorized row-major data ordering) is \textasciitilde\SIrange{1.5}{2.5}{$\times$} faster than the unvectorized ``F''-ordered code, and is in some cases almost as fast as the ``F''-ordered wide-vectorized OpenCL implementation.
+However, the ``C''-ordered wide-vectorization achieves speedups of \textasciitilde\SIrange{4}{7}{$\times$} demonstrating the value of the vectorized data ordering for CPU execution.
+
+\begin{figure}[htb]
+   \centering
+  \begin{subfigure}[t]{0.45\linewidth}
+      \includegraphics[width=\textwidth]{source_conpvsconv.pdf}
+      \caption{The mean runtime per condition for each chemical model using both the \conp~and \conv~formulations on Intel OpenCL and OpenMP for \avx.}
+      \label{F:source_conpvsconv}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.45\linewidth}
+      \includegraphics[width=\textwidth]{source_cvsf.pdf}
+      \caption{The speedup achieved by ``C'' ordering over ``F'' ordering for Intel OpenCL and OpenMP on the \avx CPU.
+The speedup presented is normalized per-language (OpenMP and OpenCL) to better assess the effect of the data-ordering.}
+      \label{F:source_cvsf}
+  \end{subfigure}
+  \caption{The effect of the \conp~and \conv~formulations, as well as ``C'' and ``F'' data-ordering on source term evaluation performance in \texttt{pyJac}.
+	   It is noted the wide-vectorized ``C'' ordered OpenCL case corresponds to the vectorized data ordering described in~\cref{S:data}}
+  \label{f:source_permutate}
 \end{figure}

### 2018-02-18 scripts/plot_styles.py

@@ -73,7 +73,7 @@ def pretty_names(pname, short=False):
                   'overhead': 'Kernel Construction Overhead',
                   'vecwidth': 'Vector Width = {}',
                   'vectype': dummy_formatter({'w': 'Shallow',
-                                              'par': 'Parallel',
+                                              'par': 'Unvectorized',
                                               'd': 'Deep',
                                               'openmp': r'OpenMP'}),
                   'order': dummy_formatter({'C': 'C-order',

### 2018-02-18 derivations.tex
move custom commands to style file

@@
 
-\newcommand{\loadeq}[1]{%
-   \ExecuteMetaData[eqn_dict.tex]{eq#1}%
-}
-
 % dcases
@@
 \usepackage{amsmath,eqparbox,xparse}
+\usepackage{macros}
 
@@
 \bibliographystyle{elsarticle-num}
-
-% https://tex.stackexchange.com/a/34412/5764
-\makeatletter
-\NewDocumentCommand{\eqmathbox}{o O{c} m}{%
-  \IfValueTF{#1}
-    {\def\eqmathbox@##1##2{\eqmakebox[#1][#2]{$##1##2$}}}
-    {\def\eqmathbox@##1##2{\eqmakebox{$##1##2$}}}
-  \mathpalette\eqmathbox@{#3}
-}
-\makeatother
-
-\newcommand{\pluseq}{\mathrel{{+}{=}}}
-\newcommand{\minuseq}{\mathrel{{-}{=}}}
-\newcommand{\ns}{N_{sp}}
-\newcommand{\nr}{N_{reac}}
-\newcommand{\conp}{CONP}
-\newcommand{\conv}{CONV}
-\newcommand{\dconp}{\ifmmode\text{for \conp,}\else for \conp,\fi}
-\newcommand{\dconv}{\ifmmode\text{for \conv,}\else for \conv,\fi}
-\newcommand{\Ru}{\mathcal{R}}
-
-% simple command for equal-sized eqmathbox's on the lhs / rhs and conditional
-% arguements are tag, text, alignment
-\NewDocumentCommand{\lhs}{mmO{r}}{\eqmathbox[LHS_#1][#3]{#2}}
-\NewDocumentCommand{\rhs}{mmO{l}}{\eqmathbox[RHS_#1][#3]{#2}}
-\NewDocumentCommand{\cond}{mmO{r}}{\eqmakebox[COND_#1][#3]{#2}}
-\NewDocumentCommand{\mathcond}{mmO{r}}{\eqmathbox[COND_#1][#3]{#2}}
-
-% consistent multiline equation spacing
-\NewDocumentCommand{\mathindent}{O{space}}{\hphantom{\mathrel{#1}}}
 
@@
 
-%fix to dcases from here:http://tex.stackexchange.com/questions/252410/centering-in-dcases-environment/252414
-\MHInternalSyntaxOn
-\renewcommand{\dcases}
- {
-  \MT_start_cases:nnnn
-    {\quad}
-    {$\m@th\displaystyle##$\hfil}
-    {$\m@th\displaystyle##$\hfil}
-    {\lbrace}
- }
-\MHInternalSyntaxOff
-
 % allow page breaks for long equations:
@@
 This document contains the derivations for \texttt{pyJac} v2.0, which generates code to evaluate the chemical source terms and analytical chemical kinetic Jacobians of constant-pressure\slash constant-volume, fixed-mass, adiabatic reactors.
-Note that equations specific to the constant-pressure or constant-volume derivation will be marked with \conp~or \conv~respectively.
+Note that equations specific to the constant-pressure or constant-volume derivation will be marked with \conp/ or \conv/ respectively.
 The derivations in this document based upon the output of the script \href{https://github.com/arghdos/SPyJac-paper/blob/master/derivations/scripts/derivations.py}{derivations.py} in the GitHub repository for this paper.
@@
 \loadeq{state}
-where $T$ is the temperature, $V$ and $P$ the volume and pressure respectively, and $n_j$ the number of moles of the $j$th species in the model (containing $\ns$ total species).
+where $T$ is the temperature, $V$ and $P$ the volume and pressure respectively, and $n_j$ the number of moles of the $j$th species in the model (containing $\ns/$ total species).
 
@@
 \label{e:last_species}
-n &= \frac{V P}{T \Ru} = \sum_{i=1}^{\ns}{n_i}\; , \nonumber \\
-n_{\ns} &= \frac{V P}{T \Ru} - \sum_{i=1}^{\ns - 1}{n_i}
+n &= \frac{V P}{T \Ru/} = \sum_{i=1}^{\ns/}{n_i}\; , \nonumber \\
+n_{\ns/} &= \frac{V P}{T \Ru/} - \sum_{i=1}^{\ns/ - 1}{n_i}
 \end{align}
@@
 \begin{align}
-\frac{\text{d} V }{\text{d} t } &= \frac{\Ru}{P} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right) & \text{\dconp} \\
-\frac{\text{d} P }{\text{d} t } &= \frac{\Ru}{V} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right) & \text{\dconv}
+\frac{\text{d} V }{\text{d} t } &= \frac{\Ru/}{P} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right) & \text{\dconp/} \\
+\frac{\text{d} P }{\text{d} t } &= \frac{\Ru/}{V} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right) & \text{\dconv/}
 \end{align}
@@
  \label{e:mass_conservation}
- \frac{\text{d} m }{\text{d} t } = 0 = \sum_{k=1}^{\ns}  W_{k} \frac{\text{d} n_{k} }{\text{d} t },
+ \frac{\text{d} m }{\text{d} t } = 0 = \sum_{k=1}^{\ns/}  W_{k} \frac{\text{d} n_{k} }{\text{d} t },
 \end{equation}
@@
 \begin{equation}
- \frac{\text{d} n }{\text{d} t }_{\ns} = \frac{-1}{W_{\ns}} \sum_{k=1}^{\ns - 1} W_{k} \frac{\text{d} n_{k} }{\text{d} t }
+ \frac{\text{d} n }{\text{d} t }_{\ns/} = \frac{-1}{W_{\ns/}} \sum_{k=1}^{\ns/ - 1} W_{k} \frac{\text{d} n_{k} }{\text{d} t }
  \label{e:spec_ns}
@@
 \begin{align}
-\frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns} \frac{\text{d} n_{k} }{\text{d} t }, \nonumber \\
-\frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\text{d} n_{k} }{\text{d} t }
+\frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns/} \frac{\text{d} n_{k} }{\text{d} t }, \nonumber \\
+\frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns/ - 1} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \frac{\text{d} n_{k} }{\text{d} t }
 \label{e:total_molar}
@@
  \label{e:last_spec_conc}
-   [C]_{\ns} = [C] - \sum_{k=1}^{\ns  - 1} [C]_{k}
+   [C]_{\ns/} = [C] - \sum_{k=1}^{\ns/  - 1} [C]_{k}
 \end{equation}
@@
 \begin{equation}
- [C] = \frac{P}{T \Ru}
+ [C] = \frac{P}{T \Ru/}
 \end{equation}
@@
 \begin{align}
-\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns  - 1} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{p,\ns}} + \sum_{k=1}^{\ns  - 1} \left({C_{p, k}} - {C_{p,\ns}}\right) [C]_{k}} &\text{\dconp}\\
-\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns  - 1} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{v,\ns}} + \sum_{k=1}^{\ns  - 1} \left({C_{v, k}} - {C_{v,\ns}}\right) [C]_{k}} &\text{\dconv}
+\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns/  - 1} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \dot{\omega}_{k}}{[C] {C_{p,\ns/}} + \sum_{k=1}^{\ns/  - 1} \left({C_{p, k}} - {C_{p,\ns/}}\right) [C]_{k}} &\text{\dconp/}\\
+\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns/  - 1} \left(U_{k} - \frac{W_{k} U_{\ns/}}{W_{\ns/}}\right) \dot{\omega}_{k}}{[C] {C_{v,\ns/}} + \sum_{k=1}^{\ns/  - 1} \left({C_{v, k}} - {C_{v,\ns/}}\right) [C]_{k}} &\text{\dconv/}
 \end{align}
@@
  \label{e:spec_rop}
- \dot{\omega}_{k} = \sum_{i=1}^{\nr} \nu_{k,i} q_{i}
-\end{equation}
-where $\nr$ is the number of reactions in the chemical kinetic model, $\nu_{k, i}$ the overall stoichiometric coefficient of species $k$ in reaction $i$ and $q_i$ the net rate-of-progress for reaction $i$:
+ \dot{\omega}_{k} = \sum_{i=1}^{\nr//} \nu_{k,i} q_{i}
+\end{equation}
+where $\nr//$ is the number of reactions in the chemical kinetic model, $\nu_{k, i}$ the overall stoichiometric coefficient of species $k$ in reaction $i$ and $q_i$ the net rate-of-progress for reaction $i$:
 \begin{align}
@@
 R_{i} &= {R_{f,i}} - {R_{r,i}} \label{e:ropnet}\\
-{R_{f, i}} &= {k_{f, i}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} \label{e:ropf}\\
-{R_{r, i}} &= {k_{r, i}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} \label{e:ropr}
+{R_{f, i}} &= {k_{f, i}} \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime}_{k,i}} \label{e:ropf}\\
+{R_{r, i}} &= {k_{r, i}} \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime\prime}_{k,i}} \label{e:ropr}
 \end{align}
@@
   \label{e:arrhenius}
-  {k_{f, i}} = A_{i} T^{\beta_{i}} \operatorname{exp}\left({- \frac{{E_{a}}_{i}}{T \Ru}}\right) \;,
+  {k_{f, i}} = A_{i} T^{\beta_{i}} \operatorname{exp}\left({- \frac{{E_{a}}_{i}}{T \Ru/}}\right) \;,
 \end{equation}
@@
  {k_{r, i}} &= \frac{{k_{f, i}}}{{K_{c, i}}}\; , \label{e:kr}\\
- {K_{c, i}} &= {K_{p,i}}\left(\frac{P_{atm}}{T \Ru}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}} \; ,\text{ and} \label{e:kc}\\
- {K_{p,i}} &= \text{exp}\left(\frac{\Delta S^{\circ}_i}{\Ru} - \frac{\Delta H^{\circ}_i}{\Ru T}\right) = \text{exp}\left(\sum_{k=1}^{\ns}\nu_{ki}\left(\frac{S^{\circ}_k}{\Ru} - \frac{H^{\circ}_k}{\Ru T}\right)\right) \label{e:kp}
+ {K_{c, i}} &= {K_{p,i}}\left(\frac{P_{atm}}{T \Ru/}\right)^{\sum_{k=1}^{\ns/} \nu_{k,i}} \; ,\text{ and} \label{e:kc}\\
+ {K_{p,i}} &= \text{exp}\left(\frac{\Delta S^{\circ}_i}{\Ru/} - \frac{\Delta H^{\circ}_i}{\Ru/ T}\right) = \text{exp}\left(\sum_{k=1}^{\ns/}\nu_{ki}\left(\frac{S^{\circ}_k}{\Ru/} - \frac{H^{\circ}_k}{\Ru/ T}\right)\right) \label{e:kp}
 \end{align}
@@
  \label{e:kc_in_kp}
- {K_{c, i}} = \left(\left(\frac{P_{atm}}{\Ru}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}}\right) \operatorname{exp}\left({\sum_{k=1}^{\ns} \nu_{k,i} B_{k}}\right)
+ {K_{c, i}} = \left(\left(\frac{P_{atm}}{\Ru/}\right)^{\sum_{k=1}^{\ns/} \nu_{k,i}}\right) \operatorname{exp}\left({\sum_{k=1}^{\ns/} \nu_{k,i} B_{k}}\right)
 \end{equation}
@@
  \label{e:Bk}
- B_{k} &= \frac{S^{\circ}_k}{\Ru} - \frac{H^{\circ}_k}{\Ru T} - \ln{T} \nonumber\; , \\
+ B_{k} &= \frac{S^{\circ}_k}{\Ru/} - \frac{H^{\circ}_k}{\Ru/ T} - \ln{T} \nonumber\; , \\
  B_{k} &= T \left(T \left(T \left(\frac{T a_{k,4}}{20} + \frac{a_{k,3}}{12}\right) + \frac{a_{k,2}}{6}\right) + \frac{a_{k,1}}{2}\right) \nonumber \\
@@
 \begin{equation}
-[X]_{i} = \sum_{k=1}^{\ns} \alpha_{k,i} [C]_{k} \;,
+[X]_{i} = \sum_{k=1}^{\ns/} \alpha_{k,i} [C]_{k} \;,
 \end{equation}
@@
 \begin{equation}
- [X]_{i} = [C] + \sum_{k=1}^{\ns} \left(\alpha_{k,i} - 1\right) [C]_{k} \;,
+ [X]_{i} = [C] + \sum_{k=1}^{\ns/} \left(\alpha_{k,i} - 1\right) [C]_{k} \;,
 \end{equation}
@@
 \label{e:thd_mix}
- [X]_{i}=[C] \alpha_{\ns,i} + \sum_{k=1}^{\ns  - 1} \left(- \alpha_{\ns,i} + \alpha_{k,i}\right) [C]_{k}\;.
+ [X]_{i}=[C] \alpha_{\ns/,i} + \sum_{k=1}^{\ns/  - 1} \left(- \alpha_{\ns/,i} + \alpha_{k,i}\right) [C]_{k}\;.
 \end{equation}
@@
 \label{e:thd_unity}
- [X]_{i} = [C] = \frac{P}{\Ru T} \;.
+ [X]_{i} = [C] = \frac{P}{\Ru/ T} \;.
 \end{equation}
@@
 \label{e:thd_spec}
- [X]_{i}=\left([C] - \sum_{k=1}^{\ns  - 1} [C]_{k}\right) \delta_{\ns m} + \left(- \delta_{\ns m} + 1\right) [C]_{m}
-\end{equation}
-will be used, where the Kronecker delta $\delta_{\ns m}$ is unity if and only if the third body species $m$ is the last species in the model.
+ [X]_{i}=\left([C] - \sum_{k=1}^{\ns/  - 1} [C]_{k}\right) \delta_{\ns/ m} + \left(- \delta_{\ns/ m} + 1\right) [C]_{m}
+\end{equation}
+will be used, where the Kronecker delta $\delta_{\ns/ m}$ is unity if and only if the third body species $m$ is the last species in the model.
 \Cref{e:thd_spec} will be referred to as the \textbf{species}-based third-body efficiency.
@@
 \end{align}
-To do so, we will more closely follow the \conp~energy equation derivative with respect to temperature as the process is largely identical between \conp~and \conv.
-
-The derivative of a species concentration with respect to temperature for both \conp~and \conv~is:
+To do so, we will more closely follow the \conp/ energy equation derivative with respect to temperature as the process is largely identical between \conp/ and \conv/.
+
+The derivative of a species concentration with respect to temperature for both \conp/ and \conv/ is:
 \begin{align}
@@
  \begin{dcases}
-  \frac{\partial}{\partial T} \left(\frac{n_k}{V}\right) = 0 & \left\{k \ne \ns\right\} \\
-  \frac{\partial}{\partial T} \left([C] - \sum_{k=1}^{\ns - 1}{\frac{n_k}{V}}\right) = -\frac{[C]}{T} & \left\{k = \ns\right\}
+  \frac{\partial}{\partial T} \left(\frac{n_k}{V}\right) = 0 & \left\{k \ne \ns/\right\} \\
+  \frac{\partial}{\partial T} \left([C] - \sum_{k=1}^{\ns/ - 1}{\frac{n_k}{V}}\right) = -\frac{[C]}{T} & \left\{k = \ns/\right\}
  \end{dcases}
 \end{align}
-as the pressure and volume are either constants (for \conp~and \conv~respectively) or a state variable and a function of time only.
+as the pressure and volume are either constants (for \conp/ and \conv/ respectively) or a state variable and a function of time only.
 
 Next, the derivative of full form of the energy source term~\cref{e:temperature_complete} with respect to temperature will be considered.
-To make the result more managable, the denominator of~\cref{e:temperature_complete} is first collapsed back into the form of~\cref{e:temperature_incomplete}, yielding for~\conp:
-\begin{align}
- \frac{\partial\dot{T}}{\partial{T}} &= - \frac{\left( [C] \left( \frac{{C_p}_{\ns}}{T} - \frac{\text{d} {C_p} }{\text{d} T }_{\ns} \right) - \sum_{k=1}^{-1 + \ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns} + \frac{\text{d} C_{p, k} }{\text{d} T }\right) [C]_{k}\right)}{\left(\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}\right)^{2}} \times \nonumber \\
-				     & \mathindent[largerindent] \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k} - \nonumber \\
-				     & \mathindent \frac{\sum_{k=1}^{-1 + \ns} \left(\left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \left(\frac{\text{d} H_{k} }{\text{d} T } - \frac{W_{k}}{W_{\ns}} \frac{\text{d} H }{\text{d} T }_{\ns}\right) \dot{\omega}_{k}\right)}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}}
+To make the result more managable, the denominator of~\cref{e:temperature_complete} is first collapsed back into the form of~\cref{e:temperature_incomplete}, yielding for \conp/:
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{T}} &= - \frac{\left( [C] \left( \frac{{C_p}_{\ns/}}{T} - \frac{\text{d} {C_p} }{\text{d} T }_{\ns/} \right) - \sum_{k=1}^{-1 + \ns/} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns/} + \frac{\text{d} C_{p, k} }{\text{d} T }\right) [C]_{k}\right)}{\left(\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}\right)^{2}} \times \nonumber \\
+				     & \mathindent[largerindent] \sum_{k=1}^{-1 + \ns/} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \dot{\omega}_{k} - \nonumber \\
+				     & \mathindent \frac{\sum_{k=1}^{-1 + \ns/} \left(\left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \left(\frac{\text{d} H_{k} }{\text{d} T } - \frac{W_{k}}{W_{\ns/}} \frac{\text{d} H }{\text{d} T }_{\ns/}\right) \dot{\omega}_{k}\right)}{\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}}
 \end{align}
@@
 \begin{align}
- \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \times \nonumber \\
-				     & \left( [C] \left( \frac{{C_p}_{\ns}}{T} - \frac{\text{d} {C_p} }{\text{d} T }_{\ns} \right) - \sum_{k=1}^{-1 + \ns} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns} + \frac{\text{d} {C_p}_{k} }{\text{d} T }\right) [C]_{k}\right) - \nonumber \\
-				     & \sum_{k=1}^{-1 + \ns} \left(\left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \left(\frac{\text{d} H_{k} }{\text{d} T } - \frac{W_{k}}{W_{\ns}} \frac{\text{d} H }{\text{d} T }_{\ns}\right) \dot{\omega}_{k}\right) \Biggr]
+ \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \times \nonumber \\
+				     & \left( [C] \left( \frac{{C_p}_{\ns/}}{T} - \frac{\text{d} {C_p} }{\text{d} T }_{\ns/} \right) - \sum_{k=1}^{-1 + \ns/} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns/} + \frac{\text{d} {C_p}_{k} }{\text{d} T }\right) [C]_{k}\right) - \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns/} \left(\left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \left(\frac{\text{d} H_{k} }{\text{d} T } - \frac{W_{k}}{W_{\ns/}} \frac{\text{d} H }{\text{d} T }_{\ns/}\right) \dot{\omega}_{k}\right) \Biggr]
 \end{align}
@@
 \begin{equation}
- [C] = \sum_{k=1}^{\ns} [C]_{k}
+ [C] = \sum_{k=1}^{\ns/} [C]_{k}
 \end{equation}
@@
 \begin{align}
- \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns} \left(- \frac{\text{d} C_{p, k} }{\text{d} T } + \frac{{C_p}_{\ns}}{T}\right) [C]_{k} + \nonumber \\
-				     & \sum_{k=1}^{-1 + \ns} \biggl(\left(- H_{k} + \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \nonumber \\
-				     & \mathindent[larger] \left(- \frac{\text{d} H_{k} }{\text{d} T } + \frac{W_{k}}{W_{\ns}} \frac{\text{d} H }{\text{d} T }_{\ns}\right) \dot{\omega}_{k}\biggr)\Biggr]
+ \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns/} \left(- \frac{\text{d} C_{p, k} }{\text{d} T } + \frac{{C_p}_{\ns/}}{T}\right) [C]_{k} + \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns/} \biggl(\left(- H_{k} + \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \nonumber \\
+				     & \mathindent[larger] \left(- \frac{\text{d} H_{k} }{\text{d} T } + \frac{W_{k}}{W_{\ns/}} \frac{\text{d} H }{\text{d} T }_{\ns/}\right) \dot{\omega}_{k}\biggr)\Biggr]
 \end{align}
@@
 \begin{align}
- \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns} \left(- \frac{\text{d} C_{p, k} }{\text{d} T } + \frac{{C_p}_{\ns}}{T}\right) [C]_{k} + & \nonumber \\
-				     & \sum_{k=1}^{-1 + \ns} \biggl( \left(- H_{k} + \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + & \nonumber \\
-				     & \mathindent[larger] \left(- {C_p}_{k} + \frac{W_{k} {C_p}_{k} }{W_{\ns}} \right) \dot{\omega}_{k}\biggr)\Biggr] & \text{\dconp} \\
+ \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns/} \left(- \frac{\text{d} C_{p, k} }{\text{d} T } + \frac{{C_p}_{\ns/}}{T}\right) [C]_{k} + & \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns/} \biggl( \left(- H_{k} + \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + & \nonumber \\
+				     & \mathindent[larger] \left(- {C_p}_{k} + \frac{W_{k} {C_p}_{k} }{W_{\ns/}} \right) \dot{\omega}_{k}\biggr)\Biggr] & \text{\dconp/} \\
 \intertext{and similarly:}
-\frac{\partial\dot{T}}{\partial{T}}  &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns} \left(- \frac{\text{d} C_{v, k} }{\text{d} T } + \frac{{C_v}_{\ns}}{T}\right) [C]_{k} + & \nonumber \\
-				     & \sum_{k=1}^{-1 + \ns} \biggl(\left(- U_{k} + \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + & \nonumber \\
-				     & \mathindent[larger] \left(- {C_v}_{k} + \frac{W_{k} {C_v}_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}\biggr)\Biggr] & \text{\dconv}
+\frac{\partial\dot{T}}{\partial{T}}  &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_v}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns/} \left(- \frac{\text{d} C_{v, k} }{\text{d} T } + \frac{{C_v}_{\ns/}}{T}\right) [C]_{k} + & \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns/} \biggl(\left(- U_{k} + \frac{W_{k} U_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + & \nonumber \\
+				     & \mathindent[larger] \left(- {C_v}_{k} + \frac{W_{k} {C_v}_{\ns/}}{W_{\ns/}}\right) \dot{\omega}_{k}\biggr)\Biggr] & \text{\dconv/}
 \end{align}
@@
 \begin{align}
- \mathcal{J}_{1, 2} & = \frac{\partial\dot{T}}{\partial{V}} & \text{\dconp} \\
- \mathcal{J}_{1, 2} & = \frac{\partial\dot{T}}{\partial{P}} & \text{\dconv}
+ \mathcal{J}_{1, 2} & = \frac{\partial\dot{T}}{\partial{V}} & \text{\dconp/} \\
+ \mathcal{J}_{1, 2} & = \frac{\partial\dot{T}}{\partial{P}} & \text{\dconv/}
 \end{align}
@@
   \begin{dcases}
-  \rhs{dck}{-\frac{[C]_{k}}{V}} & \mathcond{dck}{k \ne \ns} \\
-  \rhs{dck}{\frac{1}{V} \sum_{k=1}^{\ns  - 1} [C]_{k}} & \mathcond{dck}{k = \ns}
- \end{dcases}&\cond{dck}{\dconp} \\
+  \rhs{dck}{-\frac{[C]_{k}}{V}} & \mathcond{dck}{k \ne \ns/} \\
+  \rhs{dck}{\frac{1}{V} \sum_{k=1}^{\ns/  - 1} [C]_{k}} & \mathcond{dck}{k = \ns/}
+ \end{dcases}&\cond{dck}{\dconp/} \\
  \lhs{dck}{\frac{\partial [C]_{k} }{\partial P }} &=
  \begin{dcases}
-   \rhs{dck}{\mathindent[s]0} & \mathcond{dck}{k \ne \ns} \\
-   \rhs{dck}{\frac{1}{T \Ru}} & \mathcond{dck}{k = \ns}
- \end{dcases}&\cond{dck}{\dconv}
- \end{align}
-\end{subequations}
-while the specific heat and enthalpy\slash internal-energy are independent of the state parameter for both \conp~and \conv.
+   \rhs{dck}{\mathindent[s]0} & \mathcond{dck}{k \ne \ns/} \\
+   \rhs{dck}{\frac{1}{T \Ru/}} & \mathcond{dck}{k = \ns/}
+ \end{dcases}&\cond{dck}{\dconv/}
+ \end{align}
+\end{subequations}
+while the specific heat and enthalpy\slash internal-energy are independent of the state parameter for both \conp/ and \conv/.
 Following the same procedure as in~\cref{s:denergy_dt}, the result of~\cref{e:dTdotdE_initial} are simplified by collapsing the specific-heat\slash concentration sum, and substituting in the temperature source term, giving:
@@
 \begin{align}
- \frac{\partial\dot{T}}{\partial{V}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ - \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial V } + & \nonumber \\
-				     & \mathindent[largerindentand] \frac{1}{V} \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{-1 + \ns} \left(- {C_p}_{\ns} + {C_p}_{k}\right) [C]_{k}\Biggr] & \text{\dconp} \\
- \frac{\partial\dot{T}}{\partial{P}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial P } - & \nonumber \\
-				     & \mathindent[largerindentand] \frac{{C_v}_{\ns}}{T \Ru} \frac{\text{d} T }{\text{d} t } \Biggr] & \text{\dconv}
+ \frac{\partial\dot{T}}{\partial{V}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}} \Biggl[ - \sum_{k=1}^{-1 + \ns/} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial V } + & \nonumber \\
+				     & \mathindent[largerindentand] \frac{1}{V} \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{-1 + \ns/} \left(- {C_p}_{\ns/} + {C_p}_{k}\right) [C]_{k}\Biggr] & \text{\dconp/} \\
+ \frac{\partial\dot{T}}{\partial{P}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns/} \left(U_{k} - \frac{W_{k} U_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial P } - & \nonumber \\
+				     & \mathindent[largerindentand] \frac{{C_v}_{\ns/}}{T \Ru/} \frac{\text{d} T }{\text{d} t } \Biggr] & \text{\dconv/}
 \end{align}
@@
  \begin{dcases}
- \frac{\delta_{j k}}{V} & k \ne \ns \\
- -\frac{1}{V} & k = \ns
+ \frac{\delta_{j k}}{V} & k \ne \ns/ \\
+ -\frac{1}{V} & k = \ns/
  \end{dcases} \\
@@
 \label{e:dck_dnj}
-\frac{\partial [C_k]}{\partial n_j} &= \frac{\delta_{j k} - \delta_{\ns k}}{V}
+\frac{\partial [C_k]}{\partial n_j} &= \frac{\delta_{j k} - \delta_{\ns/ k}}{V}
 \end{align}
@@
 \begin{align}
- \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{\left(\sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}\right) \sum_{k=1}^{-1 + \ns} - \frac{\delta_{j k}}{V} \left({C_p}_{\ns} - {C_p}_{k}\right)}{\left(\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}\right)^{2}} - \nonumber \\
-				       & \mathindent \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} }
+ \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{\left(\sum_{k=1}^{-1 + \ns/} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \dot{\omega}_{k}\right) \sum_{k=1}^{-1 + \ns/} - \frac{\delta_{j k}}{V} \left({C_p}_{\ns/} - {C_p}_{k}\right)}{\left(\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}\right)^{2}} - \nonumber \\
+				       & \mathindent \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}} \sum_{k=1}^{-1 + \ns/} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} }
 \end{align}
@@
 \begin{align}
- \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \nonumber \\
-				       &  \mathindent[largerindentand] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_p}_{\ns} + {C_p}_{j}\right)\Biggr] & \text{\dconp} \\
+ \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns/} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \nonumber \\
+				       &  \mathindent[largerindentand] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_p}_{\ns/} + {C_p}_{j}\right)\Biggr] & \text{\dconp/} \\
 \intertext{and:}
- \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \nonumber \\
-				       & \mathindent[largerindentand] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_v}_{\ns} + {C_v}_{j}\right) \Biggr] & \text{\dconv}
+ \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns/} \left(U_{k} - \frac{W_{k} U_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \nonumber \\
+				       & \mathindent[largerindentand] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_v}_{\ns/} + {C_v}_{j}\right) \Biggr] & \text{\dconv/}
 \end{align}
@@
  \mathcal{J}_{2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{V} }{\partial T} & \nonumber \\
-		    &= \frac{V}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(\frac{\partial \dot{\omega}_{k} }{\partial T } + \frac{\dot{\omega}_{k}}{T}\right) \nonumber \\
-		    &  \mathindent + \frac{V}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconp}
+		    &= \frac{V}{[C]} \sum_{k=1}^{-1 + \ns/} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \left(\frac{\partial \dot{\omega}_{k} }{\partial T } + \frac{\dot{\omega}_{k}}{T}\right) \nonumber \\
+		    &  \mathindent + \frac{V}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconp/}
 \intertext{and:}
  \mathcal{J}_{2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{P} }{\partial T} & \nonumber \\
-		    &= \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(T \frac{\partial \dot{\omega}_{k} }{\partial T } + \dot{\omega}_{k}\right) & \nonumber \\
-		    & \mathindent + \frac{P}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconv}
+		    &= \Ru/ \sum_{k=1}^{-1 + \ns/} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \left(T \frac{\partial \dot{\omega}_{k} }{\partial T } + \dot{\omega}_{k}\right) & \nonumber \\
+		    & \mathindent + \frac{P}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconv/}
 \end{align}
@@
  \mathcal{J}_{2, 2} &= \frac{\partial }{\partial V } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{V} }{\partial V} & \nonumber \\
-		    &= \frac{1}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(V \frac{\partial \dot{\omega}_{k} }{\partial V } + \dot{\omega}_{k}\right) \nonumber \\
-		    & \mathindent + \frac{1}{T} \left(V \frac{\text{d} \dot{T} }{\text{d} V } + \dot{T}\right) & \text{\dconp} \\
+		    &= \frac{1}{[C]} \sum_{k=1}^{-1 + \ns/} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \left(V \frac{\partial \dot{\omega}_{k} }{\partial V } + \dot{\omega}_{k}\right) \nonumber \\
+		    & \mathindent + \frac{1}{T} \left(V \frac{\text{d} \dot{T} }{\text{d} V } + \dot{T}\right) & \text{\dconp/} \\
  \intertext{and:}
  \mathcal{J}_{2, 2} &= \frac{\partial }{\partial P } \frac{\text{d} P}{\text{d} t} = \frac{\partial \dot{P} }{\partial P} & \nonumber \\
-		    &= T \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial P } & \nonumber \\
-		    & \mathindent + \frac{1}{T} \left(P \frac{\text{d} \dot{T} }{\text{d} P } + \dot{T}\right) & \text{\dconv}
+		    &= T \Ru/ \sum_{k=1}^{-1 + \ns/} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial P } & \nonumber \\
+		    & \mathindent + \frac{1}{T} \left(P \frac{\text{d} \dot{T} }{\text{d} P } + \dot{T}\right) & \text{\dconv/}
 \end{align}
@@
  \mathcal{J}_{2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{V} }{\partial n_j} & \nonumber \\
-		    &= V \left(\frac{1}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } + \frac{1}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} }\right) & \text{\dconp}
+		    &= V \left(\frac{1}{[C]} \sum_{k=1}^{-1 + \ns/} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } + \frac{1}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} }\right) & \text{\dconp/}
 \intertext{and:}
  \mathcal{J}_{2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{P} }{\partial n_j} & \nonumber \\
-		    &=  \frac{P}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} } + T \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \text{\dconv}
+		    &=  \frac{P}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} } + T \Ru/ \sum_{k=1}^{-1 + \ns/} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \text{\dconv/}
 \end{align}
@@
   \mathcal{J}_{k + 2, 2} &= \frac{\partial }{\partial V } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial V } & \nonumber \\
-			 &= V \frac{\partial \dot{\omega}_{k} }{\partial V } + \dot{\omega}_{k} & \text{\dconp}
+			 &= V \frac{\partial \dot{\omega}_{k} }{\partial V } + \dot{\omega}_{k} & \text{\dconp/}
   \intertext{and:}
   \mathcal{J}_{k + 2, 2} &= \frac{\partial }{\partial P } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial P } & \nonumber \\
-			 &=V \frac{\partial \dot{\omega}_{k} }{\partial P } & \text{\dconv}
+			 &=V \frac{\partial \dot{\omega}_{k} }{\partial P } & \text{\dconv/}
 \end{align}
@@
  \label{e:dwdot_dT}
- \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial T }} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial T } + \nu_{k,i} \frac{\partial R_{i}}{\partial T } c_{i}\right)} & \cond{dwdot}{,}
+ \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial T }} &= \rhs{dwdot}{\sum_{i=1}^{\nr//} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial T } + \nu_{k,i} \frac{\partial R_{i}}{\partial T } c_{i}\right)} & \cond{dwdot}{,}
 \end{align}
@@
  \begin{align}
-  \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial V }} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial V } + \nu_{k,i} \frac{\partial R_{i}}{\partial V } c_{i}\right)} & \cond{dwdot}{\dconp}\\
-  \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial P }} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial P } + \nu_{k,i} \frac{\partial R_{i}}{\partial P } c_{i}\right)} & \cond{dwdot}{\dconv}
+  \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial V }} &= \rhs{dwdot}{\sum_{i=1}^{\nr//} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial V } + \nu_{k,i} \frac{\partial R_{i}}{\partial V } c_{i}\right)} & \cond{dwdot}{\dconp/}\\
+  \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial P }} &= \rhs{dwdot}{\sum_{i=1}^{\nr//} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial P } + \nu_{k,i} \frac{\partial R_{i}}{\partial P } c_{i}\right)} & \cond{dwdot}{\dconv/}
  \end{align}
@@
  \label{e:dwdot_dnj}
- \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial {n_j} }} &= \rhs{dwdot}{\sum_{i=1}^{\nr} \left(\nu_{k,i} R_{i} \frac{\partial c_{i} }{\partial {n_j} } + \nu_{k,i} \frac{\partial R_{i} }{\partial {n_j} } c_{i}\right)} & \cond{dwdot}{,}
+ \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial {n_j} }} &= \rhs{dwdot}{\sum_{i=1}^{\nr//} \left(\nu_{k,i} R_{i} \frac{\partial c_{i} }{\partial {n_j} } + \nu_{k,i} \frac{\partial R_{i} }{\partial {n_j} } c_{i}\right)} & \cond{dwdot}{,}
 \end{align}
@@
 \begin{equation}
- \frac{\partial R_{i}}{\partial T } = \frac{\partial}{\partial T}\left({k_{f, i}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right)
+ \frac{\partial R_{i}}{\partial T } = \frac{\partial}{\partial T}\left({k_{f, i}} \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime}_{k,i}}\right)
 \end{equation}
@@
  \label{e:dkf_dt}
- \frac{\text{d} k_{f, i} }{\text{d} T } = \frac{{k_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
+ \frac{\text{d} k_{f, i} }{\text{d} T } = \frac{{k_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)
 \end{equation}
@@
 \begin{equation}
- {R_f} = \left(\left(- \sum_{k=1}^{\ns  - 1} [C]_{k} + \frac{P}{T \Ru}\right)^{\nu^{\prime}_{\ns,i}}\right) {k_{f, i}} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
+ {R_f} = \left(\left(- \sum_{k=1}^{\ns/  - 1} [C]_{k} + \frac{P}{T \Ru/}\right)^{\nu^{\prime}_{\ns/,i}}\right) {k_{f, i}} \prod_{k=1}^{\ns/  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
 \end{equation}
@@
  \label{e:dropfdt_1}
- \frac{\partial {R_f} _{i}}{\partial T } = \frac{\text{d} k_{f, i} }{\text{d} T } \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime}_{\ns,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
-\end{equation}
-By defining a temporary variable $S^{\prime}_{l}$ (for species $l = \left\{j, \ns\right\}$):
+ \frac{\partial {R_f} _{i}}{\partial T } = \frac{\text{d} k_{f, i} }{\text{d} T } \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns/,i}}{T} [C]_{\ns/}^{\nu^{\prime}_{\ns/,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns/  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
+\end{equation}
+By defining a temporary variable $S^{\prime}_{l}$ (for species $l = \left\{j, \ns/\right\}$):
 \begin{equation}
  \label{e:s_temp}
- S^{\prime}_{l} = \nu^{\prime}_{l,i} [C]_{l}^{\nu^{\prime}_{l,i} - 1} \prod_{\substack{1 \leq l \leq l - 1\\l + 1 \leq l \leq \ns}} [C]_{l}^{\nu^{\prime}_{l,i}}
+ S^{\prime}_{l} = \nu^{\prime}_{l,i} [C]_{l}^{\nu^{\prime}_{l,i} - 1} \prod_{\substack{1 \leq l \leq l - 1\\l + 1 \leq l \leq \ns/}} [C]_{l}^{\nu^{\prime}_{l,i}}
 \end{equation}
@@
  \label{e:dropf_dt}
- \frac{\partial {R_f}_{i}}{\partial T } = - \frac{[C] S^{\prime}_{\ns}}{T} {k_{f, i}} + \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
+ \frac{\partial {R_f}_{i}}{\partial T } = - \frac{[C] S^{\prime}_{\ns/}}{T} {k_{f, i}} + \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)
 \end{equation}
@@
  \label{e:droprdt_1}
- \frac{\partial R_{r, i}}{\partial T } = \frac{\text{d} k_{r, i} }{\text{d} T } \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime\prime}_{\ns,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
+ \frac{\partial R_{r, i}}{\partial T } = \frac{\text{d} k_{r, i} }{\text{d} T } \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns/,i}}{T} [C]_{\ns/}^{\nu^{\prime\prime}_{\ns/,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns/  - 1} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
 \end{equation}
@@
  \frac{\text{d} k_{r, i} }{\text{d} T } &= \left(- \frac{1}{{K_{c, i}}} \frac{\text{d} {K_{c, i}} }{\text{d} T } \frac{{k_f}_i}{{K_c}_i} + \frac{1}{{K_{c, i}}}\frac{\text{d} {k_{f, i}} }{\text{d} T } \right) \nonumber \\
-				         &= \left(- \frac{1}{{K_{c, i}}} \frac{\text{d} {K_{c, i}} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_{r, i}}
+				         &= \left(- \frac{1}{{K_{c, i}}} \frac{\text{d} {K_{c, i}} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)\right) {k_{r, i}}
 \end{align}
@@
 \begin{equation}
- \frac{\text{d} {K_{c, i}} }{\text{d} T } = {K_{c, i}} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } \;,
+ \frac{\text{d} {K_{c, i}} }{\text{d} T } = {K_{c, i}} \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } \;,
 \end{equation}
@@
 \label{e:dkr_dt}
-\frac{\text{d} k_{r, i} }{\text{d} T } = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B_{k}}{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_{r, i}} \;,
-\end{equation}
-and using an temporary value $S^{\prime\prime}_{\ns}$---defined analagously to \cref{e:s_temp} for the reverse direction---we obtain:
+\frac{\text{d} k_{r, i} }{\text{d} T } = \left(- \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k}}{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)\right) {k_{r, i}} \;,
+\end{equation}
+and using an temporary value $S^{\prime\prime}_{\ns/}$---defined analagously to \cref{e:s_temp} for the reverse direction---we obtain:
 \begin{equation}
 \label{e:dropr_dt}
-\frac{\partial R_{r, i}}{\partial T } = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B_{k}}{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_{r, i}} - \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_{r, i}}
+\frac{\partial R_{r, i}}{\partial T } = \left(- \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k}}{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)\right) {R_{r, i}} - \frac{[C] S^{\prime\prime}_{\ns/}}{T} {k_{r, i}}
 \end{equation}
@@
  \label{e:drop_dT}
- \frac{\partial R_{i}}{\partial T } =& \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right) - \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_{r, i}} \nonumber \\
-				      &\qquad + \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_{r, i}} - \frac{[C] S^{\prime}_{\ns}}{T} {k_{f, i}}
+ \frac{\partial R_{i}}{\partial T } =& \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right) - \left(- \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)\right) {R_{r, i}} \nonumber \\
+				      &\qquad + \frac{[C] S^{\prime\prime}_{\ns/}}{T} {k_{r, i}} - \frac{[C] S^{\prime}_{\ns/}}{T} {k_{f, i}}
 \end{align}
@@
  \begin{align}
-  \lhs{dri_de}{\frac{\partial {R_{f, i}} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime}_{\ns}}{V} {k_{f, i}} - \frac{{R_{f, i}}}{V} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}}&\cond{dri_de}{\dconp}\\
-  \lhs{dri_de}{\frac{\partial {R_{f, i}} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime}_{\ns} {k_{f, i}}}{T \Ru}}&\cond{dri_de}{\dconv}
+  \lhs{dri_de}{\frac{\partial {R_{f, i}} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime}_{\ns/}}{V} {k_{f, i}} - \frac{{R_{f, i}}}{V} \sum_{k=1}^{\ns/} \nu^{\prime}_{k,i}}&\cond{dri_de}{\dconp/}\\
+  \lhs{dri_de}{\frac{\partial {R_{f, i}} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime}_{\ns/} {k_{f, i}}}{T \Ru/}}&\cond{dri_de}{\dconv/}
  \end{align}
@@
  \begin{align}
-  \lhs{dri_de}{\frac{\partial {R_{r, i}} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime\prime}_{\ns}}{V} {k_{r, i}} - \frac{{R_{r, i}}}{V} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i}}&\cond{dri_de}{\dconp}\\
-  \lhs{dri_de}{\frac{\partial {R_{r, i}} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime\prime}_{\ns} {k_{r, i}}}{T \Ru}}&\cond{dri_de}{\dconv}
+  \lhs{dri_de}{\frac{\partial {R_{r, i}} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime\prime}_{\ns/}}{V} {k_{r, i}} - \frac{{R_{r, i}}}{V} \sum_{k=1}^{\ns/} \nu^{\prime\prime}_{k,i}}&\cond{dri_de}{\dconp/}\\
+  \lhs{dri_de}{\frac{\partial {R_{r, i}} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime\prime}_{\ns/} {k_{r, i}}}{T \Ru/}}&\cond{dri_de}{\dconv/}
  \end{align}
@@
  \begin{align}
-  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial V }} &= \rhs{dri_de}{\frac{1}{V}\Biggl( [C]\left(S^{\prime}_{\ns} {k_{f, i}} - S^{\prime\prime}_{\ns} {k_{r, i}}\right) + } & \nonumber\\
-  & \rhs{dri_de}{\mathindent[farth] \left({R_{r, i}} \sum_{k=1}^{\ns} \nu^{\prime\prime}_{k,i} - {R_{f, i}} \sum_{k=1}^{\ns} \nu^{\prime}_{k,i}\right) \Biggr)}&\cond{dri_de}{\dconp}\\
-  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial P }} &= \rhs{dri_de}{\frac{1}{T \Ru} \left(S^{\prime}_{\ns} {k_{f, i}} - S^{\prime\prime}_{\ns} {k_{r, i}}\right)}&\cond{dri_de}{\dconv}
+  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial V }} &= \rhs{dri_de}{\frac{1}{V}\Biggl( [C]\left(S^{\prime}_{\ns/} {k_{f, i}} - S^{\prime\prime}_{\ns/} {k_{r, i}}\right) + } & \nonumber\\
+  & \rhs{dri_de}{\mathindent[farth] \left({R_{r, i}} \sum_{k=1}^{\ns/} \nu^{\prime\prime}_{k,i} - {R_{f, i}} \sum_{k=1}^{\ns/} \nu^{\prime}_{k,i}\right) \Biggr)}&\cond{dri_de}{\dconp/}\\
+  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial P }} &= \rhs{dri_de}{\frac{1}{T \Ru/} \left(S^{\prime}_{\ns/} {k_{f, i}} - S^{\prime\prime}_{\ns/} {k_{r, i}}\right)}&\cond{dri_de}{\dconv/}
  \end{align}
@@
  \label{e:dropf_dnj_1}
- \frac{d}{d n_{k}} {R_{f_i}} = \left(\frac{\partial}{\partial n_{j}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right) {k_{f, i}} \;.
+ \frac{d}{d n_{k}} {R_{f_i}} = \left(\frac{\partial}{\partial n_{j}} \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime}_{k,i}}\right) {k_{f, i}} \;.
 \end{equation}
@@
 \begin{equation}
- \frac{\partial {R_{f, i}} }{\partial {n_j} } = {k_{f, i}} \sum_{k=1}^{\ns} \left(- \frac{\delta_{\ns k}}{V} + \frac{\delta_{j k}}{V}\right) \nu^{\prime}_{k,i} [C]_{k}^{\nu^{\prime}_{k,i} - 1} \prod_{\substack{1 \leq l \leq k - 1\\k + 1 \leq l \leq \ns}} [C]_{l}^{\nu^{\prime}_{l,i}}
-\end{equation}
-with the same temporary variables $S^{\prime}_{j}$ and $S^{\prime}_{\ns}$ as defined previously, this can be simplified to:
+ \frac{\partial {R_{f, i}} }{\partial {n_j} } = {k_{f, i}} \sum_{k=1}^{\ns/} \left(- \frac{\delta_{\ns/ k}}{V} + \frac{\delta_{j k}}{V}\right) \nu^{\prime}_{k,i} [C]_{k}^{\nu^{\prime}_{k,i} - 1} \prod_{\substack{1 \leq l \leq k - 1\\k + 1 \leq l \leq \ns/}} [C]_{l}^{\nu^{\prime}_{l,i}}
+\end{equation}
+with the same temporary variables $S^{\prime}_{j}$ and $S^{\prime}_{\ns/}$ as defined previously, this can be simplified to:
 \begin{equation}
  \label{e:dropf_dnj}
- \frac{\partial {R_{f, i}} }{\partial {n_j} } = \frac{{k_{f, i}}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
+ \frac{\partial {R_{f, i}} }{\partial {n_j} } = \frac{{k_{f, i}}}{V} \left(- S^{\prime}_{\ns/} + S^{\prime}_{j}\right)
 \end{equation}
@@
  \label{e:dropr_dnj}
- \frac{\partial {R_{r, i}} }{\partial {n_j} } = \frac{{k_{r, i}}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right)
+ \frac{\partial {R_{r, i}} }{\partial {n_j} } = \frac{{k_{r, i}}}{V} \left(- S^{\prime\prime}_{\ns/} + S^{\prime\prime}_{j}\right)
 \end{equation}
@@
  \label{e:dropi_dnj}
- \frac{\partial R_{i} }{\partial {n_j} } = - \frac{{k_{r, i}}}{V} \left(- S^{\prime\prime}_{\ns} + S^{\prime\prime}_{j}\right) + \frac{{k_{f, i}}}{V} \left(- S^{\prime}_{\ns} + S^{\prime}_{j}\right)
+ \frac{\partial R_{i} }{\partial {n_j} } = - \frac{{k_{r, i}}}{V} \left(- S^{\prime\prime}_{\ns/} + S^{\prime\prime}_{j}\right) + \frac{{k_{f, i}}}{V} \left(- S^{\prime}_{\ns/} + S^{\prime}_{j}\right)
 \end{equation}
@@
  \label{e:dkf_plog_dt}
- \frac{\partial {k_{f, i}} }{\partial T } &= \frac{{k_{f, i}}}{T} \Biggl(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \nonumber \\
-					 & \mathindent[morespac] \beta_1 + \frac{E_{a_1}}{T \Ru}\Biggr)
+ \frac{\partial {k_{f, i}} }{\partial T } &= \frac{{k_{f, i}}}{T} \Biggl(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru/} + \frac{E_{a_2}}{T \Ru/}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \nonumber \\
+					 & \mathindent[morespac] \beta_1 + \frac{E_{a_1}}{T \Ru/}\Biggr)
 \end{align}
@@
  \label{e:drop_plog_dT}
- \frac{\partial R_{i} }{\partial T } &= \left(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \beta_1 + \frac{E_{a_1}}{T \Ru} \right) \times \nonumber \\
-				     & \mathindent \left(\frac{{R_{f, i}} - {R_{r, i}}}{T}\right) + {R_{r, i}} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_{r, i}} - S^{\prime}_{\ns} {k_{f, i}}\right)
-\end{align}
-
-For \conv~problems, the P-Log forward rate coefficient is also a function of the pressure:
+ \frac{\partial R_{i} }{\partial T } &= \left(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru/} + \frac{E_{a_2}}{T \Ru/}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \beta_1 + \frac{E_{a_1}}{T \Ru/} \right) \times \nonumber \\
+				     & \mathindent \left(\frac{{R_{f, i}} - {R_{r, i}}}{T}\right) + {R_{r, i}} \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns/} {k_{r, i}} - S^{\prime}_{\ns/} {k_{f, i}}\right)
+\end{align}
+
+For \conv/ problems, the P-Log forward rate coefficient is also a function of the pressure:
 \begin{equation}
@@
  \label{e:dropi_plog_de}
- \frac{\partial R_{i} }{\partial P } &= \frac{1}{T \Ru} \left(- S^{\prime\prime}_{\ns} {k_{r, i}} + S^{\prime}_{\ns} {k_{f, i}}\right) + & \nonumber \\
-				     &  \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) \left({R_{f, i}} - {R_{r, i}}\right)}{P \left(- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}\right)} & \text{\dconv}
+ \frac{\partial R_{i} }{\partial P } &= \frac{1}{T \Ru/} \left(- S^{\prime\prime}_{\ns/} {k_{r, i}} + S^{\prime}_{\ns/} {k_{f, i}}\right) + & \nonumber \\
+				     &  \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) \left({R_{f, i}} - {R_{r, i}}\right)}{P \left(- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}\right)} & \text{\dconv/}
 \end{align}
@@
  \frac{\partial R_{i}}{\partial T } &= \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \left[ - \frac{2 \log{\left (10 \right )} \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right) \right] \left({R_{f, i}} - {R_{r, i}}\right) + \nonumber \\
-				     & \mathindent \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } {R_{r, i}} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_{r, i}} - S^{\prime}_{\ns} {k_{f, i}}\right)
-\end{align}
-
-For \conv~problems, we again must consider the derivative of the forward rate coefficient with respect to pressure:
+				     & \mathindent \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } {R_{r, i}} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns/} {k_{r, i}} - S^{\prime}_{\ns/} {k_{f, i}}\right)
+\end{align}
+
+For \conv/ problems, we again must consider the derivative of the forward rate coefficient with respect to pressure:
 \begin{equation}
@@
  \frac{\partial R_{i} }{\partial P } &= \log{\left (10 \right )} \left({R_{f, i}} - {R_{r, i}}\right) \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)} + \nonumber \\
-				     & \mathindent \frac{1}{T \Ru} \left(- S^{\prime\prime}_{\ns} {k_{r, i}} + S^{\prime}_{\ns} {k_{f, i}}\right)
+				     & \mathindent \frac{1}{T \Ru/} \left(- S^{\prime\prime}_{\ns/} {k_{r, i}} + S^{\prime}_{\ns/} {k_{f, i}}\right)
 \end{align}
@@
  \begin{align}
-  \lhs{dci_elem}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{\dconp} \\
-  \lhs{dci_elem}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{\dconv}
+  \lhs{dci_elem}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{\dconp/} \\
+  \lhs{dci_elem}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{\dconv/}
  \end{align}
@@
  \label{e:dci_thd_dt}
- \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial T }} &= \rhs{dci_thd_mix}{-\frac{[C] \alpha_{\ns,i}}{T}} & \cond{dci_thd_mix}{,}
+ \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial T }} &= \rhs{dci_thd_mix}{-\frac{[C] \alpha_{\ns/,i}}{T}} & \cond{dci_thd_mix}{,}
 \end{align}
@@
  \begin{align}
- \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial V }} &= \rhs{dci_thd_mix}{\frac{1}{V} \left([C] \alpha_{\ns,i} - [X]_{i}\right)} & \cond{dci_thd_mix}{\dconp}\\
- \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial P }} &= \rhs{dci_thd_mix}{\frac{\alpha_{\ns,i}}{T \Ru}} & \cond{dci_thd_mix}{\dconv}
+ \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial V }} &= \rhs{dci_thd_mix}{\frac{1}{V} \left([C] \alpha_{\ns/,i} - [X]_{i}\right)} & \cond{dci_thd_mix}{\dconp/}\\
+ \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial P }} &= \rhs{dci_thd_mix}{\frac{\alpha_{\ns/,i}}{T \Ru/}} & \cond{dci_thd_mix}{\dconv/}
  \end{align}
@@
 \label{e:dci_thd_dnj}
-\lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial n_{j}}} &= \rhs{dci_thd_mix}{\frac{1}{V} \left(- \alpha_{\ns,i} + \alpha_{j,i}\right)} & \cond{dci_thd_mix}{.}
+\lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial n_{j}}} &= \rhs{dci_thd_mix}{\frac{1}{V} \left(- \alpha_{\ns/,i} + \alpha_{j,i}\right)} & \cond{dci_thd_mix}{.}
 \end{align}
@@
  \begin{align}
-  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_thd_unity}{0}[c] & \cond{dci_thd_unity}{\dconp} \\
-  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_thd_unity}{\frac{1}{T \Ru}}& \cond{dci_thd_unity}{\dconv}
+  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_thd_unity}{0}[c] & \cond{dci_thd_unity}{\dconp/} \\
+  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_thd_unity}{\frac{1}{T \Ru/}}& \cond{dci_thd_unity}{\dconv/}
  \end{align}
@@
 \begin{align}
- \lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_thd_spec}{-\frac{\delta_{\ns m}}{T} [C]}&\cond{dci_thd_spec}{,} \label{e:dci_thd_spec_dt}
+ \lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial T }} &= \rhs{dci_thd_spec}{-\frac{\delta_{\ns/ m}}{T} [C]}&\cond{dci_thd_spec}{,} \label{e:dci_thd_spec_dt}
 \end{align}
@@
 \begin{align}
-\lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_thd_spec}{\frac{\delta_{\ns m}}{V} \left([C] - [C]_{\ns}\right) + \frac{[C]_{m}}{V} \left(\delta_{\ns m} - 1\right)} & \cond{dci_thd_spec}{\dconp} \\
-\lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_thd_spec}{\frac{\delta_{\ns m}}{T \Ru}} & \cond{dci_thd_spec}{\dconv}
+\lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_thd_spec}{\frac{\delta_{\ns/ m}}{V} \left([C] - [C]_{\ns/}\right) + \frac{[C]_{m}}{V} \left(\delta_{\ns/ m} - 1\right)} & \cond{dci_thd_spec}{\dconp/} \\
+\lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_thd_spec}{\frac{\delta_{\ns/ m}}{T \Ru/}} & \cond{dci_thd_spec}{\dconv/}
 \end{align}
@@
 \begin{align}
- \lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial n_j }} &= \rhs{dci_thd_spec}{\frac{1}{V} \left(- \delta_{\ns m} + \delta_{j m}\right)}&\cond{dci_thd_spec}{,} \label{e:dci_thd_spec_dnj}
+ \lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial n_j }} &= \rhs{dci_thd_spec}{\frac{1}{V} \left(- \delta_{\ns/ m} + \delta_{j m}\right)}&\cond{dci_thd_spec}{,} \label{e:dci_thd_spec_dnj}
 \end{align}
@@
  \begin{align}
-  \lhs{dci_fall}{\frac{\partial c_i }{\partial V }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial V } + \frac{\partial P_{r, i} }{\partial V } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{\dconp} \\
-  \lhs{dci_fall}{\frac{\partial c_i }{\partial P }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial P } + \frac{\partial P_{r, i} }{\partial P } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{\dconv}
+  \lhs{dci_fall}{\frac{\partial c_i }{\partial V }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial V } + \frac{\partial P_{r, i} }{\partial V } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{\dconp/} \\
+  \lhs{dci_fall}{\frac{\partial c_i }{\partial P }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial P } + \frac{\partial P_{r, i} }{\partial P } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{\dconv/}
  \end{align}
@@
  \begin{align}
-  \lhs{dci_chem}{\frac{\partial c_i }{\partial V }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial V } - \frac{\partial P_{r, i} }{\partial V } c_{i}\right)} & \cond{dci_chem}{\dconp} \\
-  \lhs{dci_chem}{\frac{\partial c_i }{\partial P }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial P } - \frac{\partial P_{r, i} }{\partial P } c_{i}\right)} & \cond{dci_chem}{\dconv}
+  \lhs{dci_chem}{\frac{\partial c_i }{\partial V }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial V } - \frac{\partial P_{r, i} }{\partial V } c_{i}\right)} & \cond{dci_chem}{\dconp/} \\
+  \lhs{dci_chem}{\frac{\partial c_i }{\partial P }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial P } - \frac{\partial P_{r, i} }{\partial P } c_{i}\right)} & \cond{dci_chem}{\dconv/}
  \end{align}
@@
 \begin{align}
- \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial T }} &= \rhs{dpr_mix}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) - \frac{[C] k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i}}} & \cond{dpr_mix}{,} \label{e:dpr_mix_dt}
+ \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial T }} &= \rhs{dpr_mix}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru/} - \frac{E_{a, \infty}}{T \Ru/}\right) - \frac{[C] k_{0, i} \alpha_{\ns/,i}}{T k_{\infty, i}}} & \cond{dpr_mix}{,} \label{e:dpr_mix_dt}
 \end{align}
@@
  \begin{align}
-  \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_mix}{- \frac{P_{r, i}}{V} + \frac{[C] k_{0, i} \alpha_{\ns,i}}{V k_{\infty, i}}} & \cond{dpr_mix}{\dconp} \\
-  \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_mix}{\frac{k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i} \Ru}} & \cond{dpr_mix}{\dconv}
+  \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_mix}{- \frac{P_{r, i}}{V} + \frac{[C] k_{0, i} \alpha_{\ns/,i}}{V k_{\infty, i}}} & \cond{dpr_mix}{\dconp/} \\
+  \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_mix}{\frac{k_{0, i} \alpha_{\ns/,i}}{T k_{\infty, i} \Ru/}} & \cond{dpr_mix}{\dconv/}
  \end{align}
@@
 \begin{align}
- \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial n_j }} &= \rhs{dpr_mix}{\frac{k_{0, i} \left(- \alpha_{\ns,i} + \alpha_{j,i}\right)}{k_{\infty, i} V}} & \cond{dpr_mix}{.} \label{e:dpr_mix_dnj}
+ \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial n_j }} &= \rhs{dpr_mix}{\frac{k_{0, i} \left(- \alpha_{\ns/,i} + \alpha_{j,i}\right)}{k_{\infty, i} V}} & \cond{dpr_mix}{.} \label{e:dpr_mix_dnj}
 \end{align}
@@
 \begin{align}
- \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial T }} &= \rhs{dpr_unity}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} - 1 + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right)} & \cond{dpr_unity}{,} \label{e:dpr_unity_dt}
+ \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial T }} &= \rhs{dpr_unity}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} - 1 + \frac{E_{a, 0}}{T \Ru/} - \frac{E_{a, \infty}}{T \Ru/}\right)} & \cond{dpr_unity}{,} \label{e:dpr_unity_dt}
 \end{align}
@@
  \begin{align}
-  \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_unity}{0} & \cond{dpr_unity}{\dconp}\\
-  \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_unity}{\frac{k_{0, i}}{k_{\infty, i}} \frac{1}{\Ru T}} & \cond{dpr_unity}{\dconv}
+  \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_unity}{0} & \cond{dpr_unity}{\dconp/}\\
+  \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_unity}{\frac{k_{0, i}}{k_{\infty, i}} \frac{1}{\Ru/ T}} & \cond{dpr_unity}{\dconv/}
  \end{align}
@@
 \begin{align}
- \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial T }} &=\rhs{dpr_spec}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) - \frac{[C] k_{0, i} \delta_{\ns m}}{T k_{\infty, i}}} & \cond{dpr_spec}{,} \label{e:dpr_spec_dt}
+ \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial T }} &=\rhs{dpr_spec}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru/} - \frac{E_{a, \infty}}{T \Ru/}\right) - \frac{[C] k_{0, i} \delta_{\ns/ m}}{T k_{\infty, i}}} & \cond{dpr_spec}{,} \label{e:dpr_spec_dt}
 \end{align}
@@
  \begin{align}
-  \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_spec}{- \frac{P_{r, i}}{V} + \frac{[C] k_{0, i} \delta_{\ns m}}{V k_{\infty, i}}} & \cond{dpr_spec}{\dconp}\\
-  \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_spec}{\frac{k_{0, i} \delta_{\ns m}}{T k_{\infty, i} \Ru}} & \cond{dpr_spec}{\dconv}
+  \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_spec}{- \frac{P_{r, i}}{V} + \frac{[C] k_{0, i} \delta_{\ns/ m}}{V k_{\infty, i}}} & \cond{dpr_spec}{\dconp/}\\
+  \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_spec}{\frac{k_{0, i} \delta_{\ns/ m}}{T k_{\infty, i} \Ru/}} & \cond{dpr_spec}{\dconv/}
  \end{align}
@@
 \begin{align}
- \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial n_j }} &=\rhs{dpr_spec}{\frac{k_{0, i}}{V k_{\infty, i}} \left(- \delta_{\ns m} + \delta_{j m}\right)} & \cond{dpr_spec}{.} \label{e:dpr_spec_dnj}
+ \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial n_j }} &=\rhs{dpr_spec}{\frac{k_{0, i}}{V k_{\infty, i}} \left(- \delta_{\ns/ m} + \delta_{j m}\right)} & \cond{dpr_spec}{.} \label{e:dpr_spec_dnj}
 \end{align}
@@
  \begin{align}
-  \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial V}} &= \rhs{dpr_gen}{P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}} & \cond{dpr_gen}{\dconp} \\
-  \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial P}} &= \rhs{dpr_gen}{\bar{\theta}_{P_{r, i}, \partial P}} & \cond{dpr_gen}{\dconv}
+  \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial V}} &= \rhs{dpr_gen}{P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}} & \cond{dpr_gen}{\dconp/} \\
+  \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial P}} &= \rhs{dpr_gen}{\bar{\theta}_{P_{r, i}, \partial P}} & \cond{dpr_gen}{\dconv/}
  \end{align}
@@
  \begin{dcases}
-  \frac{1}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) & \text{if mix,} \\
-  \frac{1}{T} \left(\beta_{0} - \beta_{\infty} - 1 + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) & \text{if unity,} \\
-  \frac{1}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru} - \frac{E_{a, \infty}}{T \Ru}\right) & \text{if species,}
+  \frac{1}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru/} - \frac{E_{a, \infty}}{T \Ru/}\right) & \text{if mix,} \\
+  \frac{1}{T} \left(\beta_{0} - \beta_{\infty} - 1 + \frac{E_{a, 0}}{T \Ru/} - \frac{E_{a, \infty}}{T \Ru/}\right) & \text{if unity,} \\
+  \frac{1}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru/} - \frac{E_{a, \infty}}{T \Ru/}\right) & \text{if species,}
  \end{dcases} \\
@@
  \begin{dcases}
-  \rhs{theta_pr_bar_dt}{- \frac{[C] k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i}}} & \text{if mix,} \\
+  \rhs{theta_pr_bar_dt}{- \frac{[C] k_{0, i} \alpha_{\ns/,i}}{T k_{\infty, i}}} & \text{if mix,} \\
   \rhs{theta_pr_bar_dt}{0}[c] & \text{if unity,} \\
-  \rhs{theta_pr_bar_dt}{- \frac{[C] k_{0, i} \delta_{\ns m}}{T k_{\infty, i}}} & \text{if species,}
+  \rhs{theta_pr_bar_dt}{- \frac{[C] k_{0, i} \delta_{\ns/ m}}{T k_{\infty, i}}} & \text{if species,}
  \end{dcases} \\
@@
  \begin{dcases}
-  \rhs{theta_pr_bar_dv}{\frac{[C] k_{0, i} \alpha_{\ns,i}}{V k_{\infty, i}}} & \text{if mix,} \\
+  \rhs{theta_pr_bar_dv}{\frac{[C] k_{0, i} \alpha_{\ns/,i}}{V k_{\infty, i}}} & \text{if mix,} \\
   \rhs{theta_pr_bar_dv}{0}[c] & \text{if unity,}  \\
-  \rhs{theta_pr_bar_dv}{\frac{[C] k_{0, i} \delta_{\ns m}}{V k_{\infty, i}}} & \text{if species,}
+  \rhs{theta_pr_bar_dv}{\frac{[C] k_{0, i} \delta_{\ns/ m}}{V k_{\infty, i}}} & \text{if species,}
  \end{dcases} \\
@@
  \begin{dcases}
-  \rhs{theta_pr_bar_dp}{\frac{k_{0, i} \alpha_{\ns,i}}{T k_{\infty, i} \Ru}} & \text{if mix,} \\
-  \rhs{theta_pr_bar_dp}{\frac{k_{0, i}}{T k_{\infty, i} \Ru}} & \text{if unity,}  \\
-  \rhs{theta_pr_bar_dp}{\frac{k_{0, i} \delta_{\ns m}}{T k_{\infty, i} \Ru}} & \text{if species,}
+  \rhs{theta_pr_bar_dp}{\frac{k_{0, i} \alpha_{\ns/,i}}{T k_{\infty, i} \Ru/}} & \text{if mix,} \\
+  \rhs{theta_pr_bar_dp}{\frac{k_{0, i}}{T k_{\infty, i} \Ru/}} & \text{if unity,}  \\
+  \rhs{theta_pr_bar_dp}{\frac{k_{0, i} \delta_{\ns/ m}}{T k_{\infty, i} \Ru/}} & \text{if species,}
  \end{dcases} \\
@@
  \begin{dcases}
-  \rhs{theta_pr_bar_dnj}{- \alpha_{\ns,i} + \alpha_{j,i}} & \text{if mix,} \\
+  \rhs{theta_pr_bar_dnj}{- \alpha_{\ns/,i} + \alpha_{j,i}} & \text{if mix,} \\
   \rhs{theta_pr_bar_dnj}{0}[c] & \text{if unity,}  \\
-  \rhs{theta_pr_bar_dnj}{- \delta_{\ns m} + \delta_{j m}} & \text{if species,}
+  \rhs{theta_pr_bar_dnj}{- \delta_{\ns/ m} + \delta_{j m}} & \text{if species,}
  \end{dcases}
@@
  \begin{align}
-  \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{\dconp} \\
-  \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{\dconv}
+  \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{\dconp/} \\
+  \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{\dconv/}
  \end{align}
@@
  \begin{align}
- \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial V }} & \cond{dfi_troe_1}{\dconp} \\
- \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial P }} & \cond{dfi_troe_1}{\dconv}
+ \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial V }} & \cond{dfi_troe_1}{\dconp/} \\
+ \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial P }} & \cond{dfi_troe_1}{\dconv/}
  \end{align}
@@
  \begin{align}
- \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial V } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} & \cond{dfi_sri}{\dconp} \\
- \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial P } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} & \cond{dfi_sri}{\dconv}
+ \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial V } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} & \cond{dfi_sri}{\dconp/} \\
+ \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial P } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} & \cond{dfi_sri}{\dconv/}
  \end{align}
@@
  \begin{align}
-  \lhs{dfi_gen}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_gen}{F_{i} \Theta_{F_i, \partial V}} & \cond{dfi_gen}{\dconp} \\
-  \lhs{dfi_gen}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_gen}{F_{i} \Theta_{F_i, \partial P}} & \cond{dfi_gen}{\dconv}
+  \lhs{dfi_gen}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_gen}{F_{i} \Theta_{F_i, \partial V}} & \cond{dfi_gen}{\dconp/} \\
+  \lhs{dfi_gen}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_gen}{F_{i} \Theta_{F_i, \partial P}} & \cond{dfi_gen}{\dconv/}
  \end{align}
@@
   \lhs{dci_fall_complete}{\frac{\partial c_{i}}{\partial V }} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1} + \biggl(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial V}}{P_{r, i} + 1} + \Theta_{F_i, \partial V} + } & \nonumber \\
-							       &  \rhs{dci_fall_complete}{\mathindent[largerindentpleas] \Theta_{P_{r,i}, \partial V} - \frac{\bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1} \biggr) c_{i}} & \cond{dci_fall_complete}{\dconp} \\
-  \lhs{dci_fall_complete}{\frac{\partial c_{i}}{\partial P }} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1} + \left(\Theta_{F_i, \partial P} - \frac{\bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_fall_complete}{\dconv}
+							       &  \rhs{dci_fall_complete}{\mathindent[largerindentpleas] \Theta_{P_{r,i}, \partial V} - \frac{\bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1} \biggr) c_{i}} & \cond{dci_fall_complete}{\dconp/} \\
+  \lhs{dci_fall_complete}{\frac{\partial c_{i}}{\partial P }} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1} + \left(\Theta_{F_i, \partial P} - \frac{\bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_fall_complete}{\dconv/}
  \end{align}
@@
  \begin{align}
-  \lhs{dci_chem_complete}{\frac{\partial c_{i}}{\partial V }} &= \rhs{dci_fall_complete}{\left(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial V}}{P_{r, i} + 1} + \Theta_{F_i, \partial V} - \frac{\bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_chem_complete}{\dconp} \\
-  \lhs{dci_chem_complete}{\frac{\partial c_{i}}{\partial P }} &= \rhs{dci_fall_complete}{\left(\Theta_{F_i, \partial P} - \frac{\bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_chem_complete}{\dconv}
+  \lhs{dci_chem_complete}{\frac{\partial c_{i}}{\partial V }} &= \rhs{dci_fall_complete}{\left(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial V}}{P_{r, i} + 1} + \Theta_{F_i, \partial V} - \frac{\bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_chem_complete}{\dconp/} \\
+  \lhs{dci_chem_complete}{\frac{\partial c_{i}}{\partial P }} &= \rhs{dci_fall_complete}{\left(\Theta_{F_i, \partial P} - \frac{\bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_chem_complete}{\dconv/}
  \end{align}
@@
 
-The derivative of the net rate of progress with respect to the state paramter is given by~\cref{e:dropi_de} for arrhenius-based reactions,~\cref{e:dropi_plog_de,e:dropi_cheb_de} for P-Log and Chebyshev reactions for a \conv~system, and~\cref{e:dropi_de} for a \conp~system.
+The derivative of the net rate of progress with respect to the state paramter is given by~\cref{e:dropi_de} for arrhenius-based reactions,~\cref{e:dropi_plog_de,e:dropi_cheb_de} for P-Log and Chebyshev reactions for a \conv/ system, and~\cref{e:dropi_de} for a \conp/ system.
 The pressure modification term derivatives are given by~\cref{e:dci_thd_de,e:dci_thd_unity_de,e:dci_thd_spec_de} for third-body enhanced reactions, and~\cref{e:dci_fall_de_final,e:dci_chem_de_final} for falloff\slash chemically-activated reactions respectively.
@@
 \begin{align}
- \lhs{dnkdotdnj}{\mathcal{J}_{k + 2, j + 2}} &\pluseq \rhs{dnkdotdnj}{V \left(\nu_{k,i} R_{i} \frac{\partial c_{i} }{\partial {n_j} } + \nu_{k,i} \frac{\partial R_{i} }{\partial {n_j} } c_{i}\right)} & \mathcond{dnkdotdnj}{k = 1, \ldots, \ns - 1}
+ \lhs{dnkdotdnj}{\mathcal{J}_{k + 2, j + 2}} &\pluseq/ \rhs{dnkdotdnj}{V \left(\nu_{k,i} R_{i} \frac{\partial c_{i} }{\partial {n_j} } + \nu_{k,i} \frac{\partial R_{i} }{\partial {n_j} } c_{i}\right)} & \mathcond{dnkdotdnj}{k = 1, \ldots, \ns/ - 1}
 \end{align}

### 2018-02-18 paper.tex
move custom commands to style file

@@
 
-\newcommand{\loadeq}[1]{%
-   \ExecuteMetaData[eqn_dict.tex]{eq#1}%
-}
-
 % external references https://tex.stackexchange.com/a/14365/56227
@@
 
-\usepackage{amsmath,eqparbox,xparse}
+\usepackage{amsmath,eqparbox,xparse,mathtools}
+
+% load common macros / commands
+\usepackage{macros}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% https://tex.stackexchange.com/a/34412/5764
-\makeatletter
-\NewDocumentCommand{\eqmathbox}{o O{c} m}{%
-  \IfValueTF{#1}
-    {\def\eqmathbox@##1##2{\eqmakebox[#1][#2]{$##1##2$}}}
-    {\def\eqmathbox@##1##2{\eqmakebox{$##1##2$}}}
-  \mathpalette\eqmathbox@{#3}
-}
-\makeatother
 
@@
 \bibliographystyle{elsarticle-num}
-
-\newcommand{\pluseq}{\mathrel{{+}{=}}}
-\newcommand{\minuseq}{\mathrel{{-}{=}}}
-\newcommand{\ns}{N_{sp}}
-\newcommand{\nr}{N_{reac}}
-\newcommand{\conp}{CONP}
-\newcommand{\conv}{CONV}
-\newcommand{\dconp}{\ifmmode\text{for \conp,}\else for \conp,\fi}
-\newcommand{\dconv}{\ifmmode\text{for \conv,}\else for \conv.\fi}
-\newcommand{\Ru}{\mathcal{R}}
-\newcommand{\avx}{\texttt{avx2}}
-\newcommand{\sse}{\texttt{sse4.2}}
-\newcommand{\gpuold}{\texttt{C2075}}
-\newcommand{\gpunew}{\texttt{K40m}}
-
-% simple command for equal-sized eqmathbox's on the lhs / rhs and conditional
-% arguements are tag, text, alignment
-\NewDocumentCommand{\lhs}{mmO{r}}{\eqmathbox[LHS_#1][#3]{#2}}
-\NewDocumentCommand{\rhs}{mmO{l}}{\eqmathbox[RHS_#1][#3]{#2}}
-\NewDocumentCommand{\cond}{mmO{r}}{\eqmakebox[COND_#1][#3]{#2}}
-\NewDocumentCommand{\mathcond}{mmO{r}}{\eqmathbox[COND_#1][#3]{#2}}
-
-% consistent multiline equation spacing
-\NewDocumentCommand{\mathindent}{O{space}}{\hphantom{\mathrel{#1}}}
 
@@
 \Cref{F:mem} depicts an example data array with K-rows and N-columns and index ($i$, $j$) corresponding to the $i$th row and $j$th column.
-For example the concentration of species $j$ for the $i$-th thermo-chemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{state}$, the number of thermo-chemical states considered, and $1 \le j \le \ns$, the number of species in the model.
+For example the concentration of species $j$ for the $i$-th thermo-chemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{state}$, the number of thermo-chemical states considered, and $1 \le j \le \ns/$, the number of species in the model.
 The stored concentrations would then have $K = N_{\text{state}}$ rows, and $N = N_{\text{sp}}$ columns.
@@
 \subsection{Thermo-chemical Source Terms and Jacobian}
-This new version of \texttt{pyJac} is capable of evaluating the thermo-chemical source terms for using both the constant-pressure (\conp) or constant-volume (\conv) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
+This new version of \texttt{pyJac} is capable of evaluating the thermo-chemical source terms for using both the constant-pressure (\conp/) or constant-volume (\conv/) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
 In this section, we will only outline a brief summary of the system evaluated by \texttt{pyJac}; the interested reader is referred to our complete derivations, attached as supplemental material to this document.
@@
 Second, mass is explicitly conserved in this formulation owing to the fact that the moles of the final species is calculated from the ideal gas law, and the rate of change of the final species from conservation of mass (see \cref{deriv-s:state,deriv-s:source} of the supplemental material respectively); i.e., system is not over-constrained.
-Finally, the chemical kinetic Jacobian changes relatively little between the \conp~and \conv~forms, making maintaining the code-base much simpler.
+Finally, the chemical kinetic Jacobian changes relatively little between the \conp/ and \conv/ forms, making maintaining the code-base much simpler.
 Although most current combustion codes do not use species moles as a state variable, to\slash from more common mass\slash mole-fractions and moles is straightforward, and once inside integration of an chemical kinetic initial value problem, the choice of variables no longer matter.
@@
 
-For both \conp~and \conv, the molar source terms are~\cite{TurnsStephenR2012Aitc}:
+For both \conp/ and \conv/, the molar source terms are~\cite{TurnsStephenR2012Aitc}:
 \loadeq{dmolar}
@@
 \begin{equation}
- \dot{\omega}_{k} = \sum_{i=1}^{\nr} \nu_{k, i} R_{i} c_{i}
+ \dot{\omega}_{k} = \sum_{i=1}^{\nr/} \nu_{k, i} R_{i} c_{i}
 \end{equation}
@@
 \loadeq{dTdotdT}
-and with respect to the thermodynamic state parameter (the volume \slash pressure for \conp\slash\conv~respectively):
+and with respect to the thermodynamic state parameter (the volume \slash pressure for \conp/\slash\conv/ respectively):
 \loadeq{dTdotdE}
@@
 Cores            & $2 \times 6$    & $2 \times 12$  \\
-Identifier       & \sse & \avx  \\
+Identifier       & \sse/ 	   & \avx/  	    \\
 OpenCL Version   & \num{1.2}       & \num{1.2}      \\
@@
 CUDA Cores     & \num{448}      & \num{2880}    \\
-Identifier     & \gpuold & \gpunew \\
+Identifier     & \gpuold/ 	& \gpunew/	\\
 OpenCL Version & \num{1.1}	& \num{1.2}	\\
@@
 Error statistics are based on the infinity-norm of the relative error detailed in Eq.~\eqref{e:rel_err} for each quantity.
-The ``S'' in $E_{\frac{\text{d} S}{\text{d} t}}$ refers to the thermodynamic state parameter, either $V$ or $P$ for \conp~and \conv~respectively.
+The ``S'' in $E_{\frac{\text{d} S}{\text{d} t}}$ refers to the thermodynamic state parameter, either $V$ or $P$ for \conp/ and \conv/ respectively.
 }
@@
 Error statistics are the largest filtered relative error $E_\mathcal{C}$ for over all samples.
-It is noted that the threshold described in~\cref{e:thresh} varies slightly between the \conp~and \conv~cases; the reported $\bar{\mathcal{T}}$ is the average of the two, however the appropriate value was used during calculations of the error statistics..
+It is noted that the threshold described in~\cref{e:thresh} varies slightly between the \conp/ and \conv/ cases; the reported $\bar{\mathcal{T}}$ is the average of the two, however the appropriate value was used during calculations of the error statistics..
 }
@@
 Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
-The wrapping code---responsible for initializing\slash transfering memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx\slash\gpunew platforms, and \texttt{gcc 4.8.5} on the \sse\slash\gpuold machines.
+The wrapping code---responsible for initializing\slash transfering memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms, and \texttt{gcc 4.8.5} on the \sse/\slash\gpuold/ machines.
 The optimization level ``\texttt{-O3}'' was used in all cases, and no ``fast-math'' OpenCL optimizations were enabled.
@@
 \Cref{f:source} explores the performance of the source term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
-In~\cref{F:intel_source_noorm}, the mean runtime per initial condition is shown for both the~\avx and~\sse CPUs, using both Intel OpenCL and OpenMP.
+In~\cref{F:intel_source_noorm}, the mean runtime per initial condition is shown for both the~\avx/ and~\sse/ CPUs, using both Intel OpenCL and OpenMP.
 This normalization is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{T:error}}).
 For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the parallel (non-vectorized) Intel OpenCL code is slightly faster than OpenMP in all cases.\todo{rerun OpenMP w/ mtune=native}
-As expected, the \avx machine is moderately faster than the \sse CPU; additionally, the wide-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.
+As expected, the \avx/ machine is moderately faster than the \sse/ CPU; additionally, the wide-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.
 In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is normalized per-machine, such that the speedup of the wide-vectorized OpenCL code can be readily assessed.
 On both machines, the unvectorized OpenCL code is \textasciitilde\SIrange{1.25}{1.5}{$\times$} faster than the baseline parallel OpenMP code.
-Additionally, the wide-vectorized OpenCL code is \textasciitilde\SIrange{2.5}{3}{$\times$} and \textasciitilde\SIrange{3.5}{4}{$\times$} faster than the OpenMP code for the \sse and \avx machines respectively.
+Additionally, the wide-vectorized OpenCL code is \textasciitilde\SIrange{2.5}{3}{$\times$} and \textasciitilde\SIrange{3.5}{4}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
 In contrast,~\cref{F:pocl_source} shows the runtime of deep, wide and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
@@
 
-\Cref{f:source_permutate} shows the effect of changing data-ordering patterns as well as the \conp~or \conv~formulation on the performance of source term evaluation in \texttt{pyJac}.
-In~\cref{F:source_conpvsconv}, we see that choice of the constant-pressure or constant-volume formulation has little to no effect on runtime for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx machine.
-Generally speaking, the difference between the \conp~and \conv~formulations has little effect on performance regardless of CPU\slash GPU choice, or even for Jacobian evaluation which will be explored later.
-In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx machine; the speedup presented is normalized per language, i.e., the roughly \SIrange{1.5}{2}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-order OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
+\Cref{f:source_permutate} shows the effect of changing data-ordering patterns as well as the \conp/ or \conv/ formulation on the performance of source term evaluation in \texttt{pyJac}.
+In~\cref{F:source_conpvsconv}, we see that choice of the constant-pressure or constant-volume formulation has little to no effect on runtime for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx/ machine.
+Generally speaking, the difference between the \conp/ and \conv/ formulations has little effect on performance regardless of CPU\slash GPU choice, or even for Jacobian evaluation which will be explored later.
+In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is normalized per language, i.e., the roughly \SIrange{1.5}{2}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-order OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
 Additionally, we note that the ``C''-ordered wide-vectorization in~\cref{F:source_cvsf} utilizes the vectorized-data ordering described in~\cref{S:data}.
@@
       \includegraphics[width=\textwidth]{source_conpvsconv.pdf}
-      \caption{The mean runtime per condition for each chemical model using both the \conp~and \conv~formulations on Intel OpenCL and OpenMP for \avx.}
+      \caption{The mean runtime per condition for each chemical model using both the \conp/ and \conv/ formulations on Intel OpenCL and OpenMP for \avx/.}
       \label{F:source_conpvsconv}
@@
       \includegraphics[width=\textwidth]{source_cvsf.pdf}
-      \caption{The speedup achieved by ``C'' ordering over ``F'' ordering for Intel OpenCL and OpenMP on the \avx CPU.
+      \caption{The speedup achieved by ``C'' ordering over ``F'' ordering for Intel OpenCL and OpenMP on the \avx/ CPU.
 The speedup presented is normalized per-language (OpenMP and OpenCL) to better assess the effect of the data-ordering.}
@@
   \end{subfigure}
-  \caption{The effect of the \conp~and \conv~formulations, as well as ``C'' and ``F'' data-ordering on source term evaluation performance in \texttt{pyJac}.
+  \caption{The effect of the \conp/ and \conv/ formulations, as well as ``C'' and ``F'' data-ordering on source term evaluation performance in \texttt{pyJac}.
 	   It is noted the wide-vectorized ``C'' ordered OpenCL case corresponds to the vectorized data ordering described in~\cref{S:data}}

### 2018-02-18 paper.tex
update plots & performance section

@@
   \centering
-  \begin{subfigure}[t]{0.45\linewidth}
+  \begin{subfigure}[t]{0.48\linewidth}
       \includegraphics[width=\textwidth]{ch4_sparsity_exact.pdf}
@@
   \hfill
-  \begin{subfigure}[t]{0.45\linewidth}
+  \begin{subfigure}[t]{0.48\linewidth}
       \includegraphics[width=\textwidth]{ch4_sparsity_approximate.pdf}
@@
 Additionally, the ``exact'' form of the Jacobian (as opposed the the ``appropriate'' form discussed in~\cref{S:sparsity}) was used in all cases.
+Finally, unless otherwise stated the \conp/ assumption was used (the differences between \conp/ and \conv/ evaluations will be explored) and the runtimes reported are for the number of conditions specified in~\cref{T:error}.
 
-\Cref{f:source} explores the performance of the source term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
+\Cref{F:cpu_source} explores the performance of the source term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
 In~\cref{F:intel_source_noorm}, the mean runtime per initial condition is shown for both the~\avx/ and~\sse/ CPUs, using both Intel OpenCL and OpenMP.
-This normalization is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{T:error}}).
+This normalization is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{T:error}).
 For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the parallel (non-vectorized) Intel OpenCL code is slightly faster than OpenMP in all cases.\todo{rerun OpenMP w/ mtune=native}
 As expected, the \avx/ machine is moderately faster than the \sse/ CPU; additionally, the wide-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.
+
 In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is normalized per-machine, such that the speedup of the wide-vectorized OpenCL code can be readily assessed.
@@
    \centering
-  \begin{subfigure}[t]{0.45\linewidth}
+  \begin{subfigure}[t]{0.48\linewidth}
       \includegraphics[width=\textwidth]{intel_source_nonorm.pdf}
@@
   \hfill
-  \begin{subfigure}[t]{0.45\linewidth}
+  \begin{subfigure}[t]{0.48\linewidth}
       \includegraphics[width=\textwidth]{intel_source.pdf}
@@
   \\
-  \begin{subfigure}[t]{0.45\linewidth}
+  \begin{subfigure}[t]{0.48\linewidth}
       \includegraphics[width=\textwidth]{pocl_source_nonorm.pdf}
@@
 The names in the legends correspond to the identifiers listed in~\cref{t:cpus}}
- \label{f:source}
+ \label{F:cpu_source}
+\end{figure}
+
+\Cref{F:gpu_source} shows the performance of the source term evaluation on the GPUs listed in~\cref{t:gpus}.
+In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean runtime per condition; the mean runtime continues to decrease for all chemical models until \textasciitilde\num{e4} conditions, at which point the GPU becomes saturated and performance levels off.
+\Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling}; the speedup ranges from \textasciitilde\SIrange{1.75}{2.5}{$\times$} depending on the chemical model in question.
+Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector width size (a.k.a., the block size in NVIDIA terminology) has minimal effect on performance; it is possible that use of larger vector sizes could have an effect on performance, but this was not considered in this work.
+
+\todo{rerun NVIDIA on srv2 -- see if that high variability in isopentanol persists}
+\begin{figure}[htb]
+   \centering
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{gpu_source_scaling.pdf}
+      \caption{The mean runtime per condition for each chemical model on the \gpunew/ GPU as a function of the number of initial conditions tested.}
+      \label{F:gpu_source_scaling}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{gpu_source_speedup.pdf}
+      \caption{The speedup achieved on the \gpunew/ versus the \gpuold/ GPU; the names correspond to the identifiers listed in~\cref{t:gpus}}
+      \label{F:gpu_source_speedup}
+  \end{subfigure}
+  \caption{\texttt{pyJac} source term evaluation performance on the NVIDIA GPUs}
+  \label{F:gpu_source}
 \end{figure}
@@
    \centering
-  \begin{subfigure}[t]{0.45\linewidth}
+  \begin{subfigure}[t]{0.48\linewidth}
       \includegraphics[width=\textwidth]{source_conpvsconv.pdf}
@@
   \hfill
-  \begin{subfigure}[t]{0.45\linewidth}
+  \begin{subfigure}[t]{0.48\linewidth}
       \includegraphics[width=\textwidth]{source_cvsf.pdf}
@@
   \end{subfigure}
+  \\
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{source_gpu_conpvsconv.pdf}
+      \caption{The affect on performance of``C'' vs ``F''-ordering for the wide-vectorized NVIDIA OpenCL code on both GPUs.
+The speedup presented is normalized per-GPU to better assess the effect of the data-ordering on each GPU.}
+      \label{F:source_conpvsconv}
+  \end{subfigure}
   \caption{The effect of the \conp/ and \conv/ formulations, as well as ``C'' and ``F'' data-ordering on source term evaluation performance in \texttt{pyJac}.
@@
   \label{f:source_permutate}
+
 \end{figure}

### 2018-02-18 scripts/general_plotting.py

@@ -106,6 +106,7 @@ def plot(x_vals, y_vals, err_vals, minx=None, miny=None, maxx=None, maxy=None,
     argdict['color'] = color
     argdict['markeredgecolor'] = color
     argdict['markersize'] = size
+    argdict['markeredgewidth'] = 1.25
     if hollow:
         argdict['markerfacecolor'] = 'None'
 

### 2018-02-18 scripts/plot_styles.py

@@ -12,13 +12,14 @@ plt.rc('text.latex',
 plt.rc('font', family='serif')
 
 legend_style = {'loc': 0,
-                'fontsize': 18,
+                'fontsize': 16,
                 'numpoints': 1,
                 'shadow': True,
                 'fancybox': True
                 }
 
 tick_font_size = 20
+title_font_size = 24
 
 marker_style = {
     'size': 15
@@ -38,7 +39,7 @@ class wheeler(object):
         self.raise_on_oob = raise_on_oob
         self.wheel = wheel[:]
 
-    def __getitem__(self, key):
+    def __call__(self, key):
         if not self.raise_on_oob and key >= len(self.wheel):
             key = key % len(self.wheel)
         return self.wheel[key]
@@ -106,9 +107,11 @@ color_dict = {x: color_wheel[i] for i, x in enumerate(default_keys)}
 
 def finalize(tight=True):
     ax = plt.gca()
-    for item in ([ax.title, ax.xaxis.label, ax.yaxis.label] +
-                 ax.get_xticklabels() + ax.get_yticklabels()):
+    for item in (ax.get_xticklabels() + ax.get_yticklabels()):
         item.set_fontsize(tick_font_size)
+
+    for item in (ax.title, ax.xaxis.label, ax.yaxis.label):
+        item.set_fontsize(title_font_size)
 
     if tight:
         plt.tight_layout()

### 2018-02-19 paper.tex
paper updates & first crack at parallel scaling

@@
 
+\subsubsection{Source term evaluation}
+\label{S:source_results}
+
 \Cref{F:cpu_source} explores the performance of the source term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
@@
 
-In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is normalized per-machine, such that the speedup of the wide-vectorized OpenCL code can be readily assessed.
+In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is calculated per-machine, such that the speedup of the wide-vectorized OpenCL code can be readily assessed.
 On both machines, the unvectorized OpenCL code is \textasciitilde\SIrange{1.25}{1.5}{$\times$} faster than the baseline parallel OpenMP code.
@@
 Generally speaking, the difference between the \conp/ and \conv/ formulations has little effect on performance regardless of CPU\slash GPU choice, or even for Jacobian evaluation which will be explored later.
-In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is normalized per language, i.e., the roughly \SIrange{1.5}{2}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-order OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
+
+In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, i.e., the roughly \SIrange{1.5}{2}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-order OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
 Additionally, we note that the ``C''-ordered wide-vectorization in~\cref{F:source_cvsf} utilizes the vectorized-data ordering described in~\cref{S:data}.
@@
 However, the ``C''-ordered wide-vectorization achieves speedups of \textasciitilde\SIrange{4}{7}{$\times$} demonstrating the value of the vectorized data ordering for CPU execution.
+
+\Cref{F:source_gpu_cvsf} shows the effect of the ``C'' versus ``F''-ordering on source term evaluation performance on both GPUs.
+The speedup reported is calculated per-GPU, such that the effect of data ordering can be seen on each individual GPU.
+In general, the ``C''-ordered wide-vectorization is \textasciitilde\SI{1.05}{1.2}{$\times$} faster on the \newgpu/ GPU while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, except for the isopentanol model where the ``F''-ordered data is \textasciitilde\SI{1.4}{$\times$} faster (though, we note that there was high variablity in this case).
+\todo{again, pending rerun on \gpuold/}
+The ``C''-ordered speedup on the \gpunew/ GPU is counter to what one might expect, as typically speaking coalesced memory access in a wide-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
+However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector width size (and thus, likely coalesced) but the improved data-locality; coupled with the larger cache sizes of the \gpunew/ GPU result in this modest speedup for ``C''-data ordering.
 
@@
       \caption{The speedup achieved by ``C'' ordering over ``F'' ordering for Intel OpenCL and OpenMP on the \avx/ CPU.
-The speedup presented is normalized per-language (OpenMP and OpenCL) to better assess the effect of the data-ordering.}
+The speedup presented is calculated per-language (OpenMP and OpenCL) to better assess the effect of the data-ordering.}
       \label{F:source_cvsf}
@@
   \begin{subfigure}[t]{0.48\linewidth}
-      \includegraphics[width=\textwidth]{source_gpu_conpvsconv.pdf}
-      \caption{The affect on performance of``C'' vs ``F''-ordering for the wide-vectorized NVIDIA OpenCL code on both GPUs.
-The speedup presented is normalized per-GPU to better assess the effect of the data-ordering on each GPU.}
-      \label{F:source_conpvsconv}
+      \includegraphics[width=\textwidth]{source_gpu_cvsf.pdf}
+      \caption{The affect on performance of ``C'' vs ``F''-ordering for the wide-vectorized NVIDIA OpenCL code on both GPUs with a vector width of 128.
+The speedup presented is calculated per-GPU to better assess the effect of the data-ordering on each GPU.}
+      \label{F:source_gpu_cvsf}
   \end{subfigure}

### 2018-02-19 scripts/general_plotting.py

@@ -9,7 +9,7 @@ font_size = 'large'
 
 
 def process_data(plotdata, plot, reacs_as_x=True,
-                 plot_cores=False, plot_scaling=False):
+                 plot_cores=False, plot_condition_scaling=False):
     """
     Process the data into an easily usable form
     """
@@ -24,7 +24,7 @@ def process_data(plotdata, plot, reacs_as_x=True,
     if plot_cores:
         plotdata = sorted(plotdata, key=lambda x: float(x.cores))
         x_vals = [float(x.cores) for x in plotdata]
-    elif plot_scaling:
+    elif plot_condition_scaling:
         # first sort by # mech
         plotdata, _ = __mechsort(plotdata)
         # next sort within mechanism by # of conditions

### 2018-02-19 scripts/plot_styles.py

@@ -86,7 +86,8 @@ def pretty_names(pname, short=False):
                                             'full': 'Full Rate Specialization'}),
                   'sparse': dummy_formatter({'sparse': 'Sparse',
                                              'full': 'Dense'}),
-                  'conp': '{}',
+                  'conp': dummy_formatter({'conp': 'CONP',
+                                           'conv': 'CONV'}),
                   'mechdata.name': dummy_formatter(legend_key),
                   'descriptor': dummy_formatter({'srv2': r'\texttt{sse4.2}',
                                                  'srv2-gpu': r'\texttt{C2075}',

### 2018-02-19 paper.tex
lots of changes

@@
 }
+
+\DeclareSIUnit\doubles{doubles}
 
@@
 Instruction Set  & SSE4.2 	   & AVX2 	    \\
-Vector Width     & \SI{128}{\bit}  & \SI{256}{\bit} \\
+Vector Width     & \SI{2}{\doubles}&\SI{4}{\doubles}\\
 Cores            & $2 \times 6$    & $2 \times 12$  \\
@@
 \end{tabular}
-\caption{The Intel CPU configurations used in this study. The identifier field will be used as a shorthand descriptor in the performance plots to quickly identify the CPU type.}
+\caption{The Intel CPU configurations used in this study.
+	 The vector width reported in this table is reported in (ideal) number of double operations per SIMD instruction, as this will be used in measuring SIMD efficiency; for reference, the vector width in bits of the \sse/ and \avx/ machines are \SI{128}{\bit} and \SI{256}{\bit} respectively.
+	 The identifier field will be used as a shorthand descriptor in the performance plots to quickly identify the CPU type.}
 \label{t:cpus}
@@
 
-\begin{table}
+\begin{table}[htb]
 \centering
@@
 
-\begin{table}[htb]
+\begin{table}[htbp]
 \sisetup{retain-zero-exponent=true}
@@
 
-\begin{table}[tbp]
+\begin{table}[htbp]
 \centering
@@
 Additionally, the ``exact'' form of the Jacobian (as opposed the the ``appropriate'' form discussed in~\cref{S:sparsity}) was used in all cases.
-Finally, unless otherwise stated the \conp/ assumption was used (the differences between \conp/ and \conv/ evaluations will be explored) and the runtimes reported are for the number of conditions specified in~\cref{T:error}.
+Finally, unless otherwise stated the \conp/ assumption was used, a vector width of \num{8}\slash\num{128} was used for all CPU\slash GPU cases respectively, ``C''\slash ``F''-ordered data was used on the CPU\slash GPU, and the runtimes reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
+The effects of choice of vector-width on the CPU\slash GPU, as well as the differences between \conp/ and \conv/ evaluations will be explored in~\cref{S:source_results}, and briefly discussed in~\cref{S:jacobian_results}.
 
@@
 \Cref{F:cpu_source} explores the performance of the source term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
+Source term evaluations, critial in their own right for direct numerical simulations~\cite{Spafford:2010aa} among other applications, also give us a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation peformance.
+
 In~\cref{F:intel_source_noorm}, the mean runtime per initial condition is shown for both the~\avx/ and~\sse/ CPUs, using both Intel OpenCL and OpenMP.
@@
 
-\begin{figure}[htb]
+\begin{figure}[htbp]
    \centering
@@
 \Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling}; the speedup ranges from \textasciitilde\SIrange{1.75}{2.5}{$\times$} depending on the chemical model in question.
-Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector width size (a.k.a., the block size in NVIDIA terminology) has minimal effect on performance; it is possible that use of larger vector sizes could have an effect on performance, but this was not considered in this work.
+Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector width size (a.k.a., the block size in NVIDIA terminology) has minimal effect on performance; it is possible that use of larger vector sizes could have an effect on performance, but this was not considered in this work.\todo{rerun NVIDIA on srv2 -- see if that high variability in isopentanol persists}
 
-\todo{rerun NVIDIA on srv2 -- see if that high variability in isopentanol persists}
-\begin{figure}[htb]
+\begin{figure}[htbp]
    \centering
@@
 The speedup reported is calculated per-GPU, such that the effect of data ordering can be seen on each individual GPU.
-In general, the ``C''-ordered wide-vectorization is \textasciitilde\SI{1.05}{1.2}{$\times$} faster on the \newgpu/ GPU while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, except for the isopentanol model where the ``F''-ordered data is \textasciitilde\SI{1.4}{$\times$} faster (though, we note that there was high variablity in this case).
+In general, the ``C''-ordered wide-vectorization is \textasciitilde\SIrange{1.05}{1.2}{$\times$} faster on the \gpunew/ GPU while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, except for the isopentanol model where the ``F''-ordered data is \textasciitilde\SI{1.4}{$\times$} faster (though, we note that there was high variablity in this case).
 \todo{again, pending rerun on \gpuold/}
@@
   \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{source_vector_width.pdf}
+      \caption{The effect of vector-width on ``C''-ordered wide-vectorized Intel OpenCL source term evaluations in the \avx/ CPU.}
+      \label{F:source_vector_width}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
       \includegraphics[width=\textwidth]{source_gpu_cvsf.pdf}
@@
   \label{f:source_permutate}
+\end{figure}
 
+Finally,~\cref{F:source_scaling} displays the (strong) parallel scaling efficiency and SIMD efficiency for the CPU platforms.
+The strong parallel scaling efficiency is defined as:
+\begin{equation}
+ \label{e:strong_scaling}
+ \varepsilon = \frac{\bar{t}_{1}}{N \bar{t}_{N}}
+\end{equation}
+where $\bar{t}_{N}$ is the mean runtime per condition on $N$ CPU cores, and $\bar{t}_{1}$ the same on a single CPU core.
+The strong parallel scaling efficiency measures the speedup due to the use of additional CPU cores, as a fraction of linear speedup; strong scaling tends to decrease with the number of processors used due to memory bandwidth limitations and decreasing computation work allocated per CPU core~\cite{strong_scaling}.
+
+\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source rate evaluation in \texttt{pyJac} on the \avx/ machine for both the wide-vectorized Intel OpenCL and OpenMP codes.
+In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for Intel OpenCL; although the mechanism scalings vary somewhat for OpenMP, there is no clear trend.
+It is likely the poor scaling of the \ce{H2}\slash\ce{CO} for Intel OpenCL results from both its relatively small size, and few falloff\slash chemically activated reaction (and in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD intructions.
+Additionally, OpenMP tends to have higher scalings then the wide-vectorized OpenCL code, e.g., \textasciitilde\num{0.8} and \textasciitilde\num{0.7} for \num{4} and \num{8} CPU cores respectively, compared to just \textasciitilde\num{0.7} and \textasciitilde\num{0.5} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the SIMD-vectorization itself the cause of this poorer scaling.
+This is again likely due to relatively small amounts of work associated with source rate evaluation; the longest runtime (the isopentanol model on a single core) takes only \textasciitilde$\SI{2}{\sec}$ for OpenCL, and \textasciitilde$\SI{9}{\sec}$ for OpenMP.
+By the time 4--8 CPU cores are utilized, the runtime of wide-vectorized OpenCL code can be measured in hundreds of milliseconds, and parallel overhead begins to dominate, decreasing efficiency.
+In contrast, the OpenMP code requires \textasciitilde\num{4} cores to run as fast as the wide-vectorized OpenCL code in this case, hence parallel overhead concerns become increasingly important for 8--16 cores.
+Indeed, we will see that the strong scalings for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.
+
+The SIMD efficiency is defined as:
+\begin{equation}
+ \label{e:simd_efficiency}
+ \varepsilon_{\text{SIMD}} = \frac{\bar{t}_{\text{unvec}}}{W \bar{t}_{\text{wide}}}
+\end{equation}
+where $\bar{t}_{\text{unvec}}$ is the mean runtime per condition of the unvectorized OpenCL code, $\bar{t}_{\text{wide}}$ the same for the wide-vectorized OpenCL code, and $W$, the vector width reported in number of double operations in~\cref{t:cpus}
+This measure compares the actual speedup due to wide-vectorization compared to the nominal vector-width of the machine in question.
+\Cref{F:source_simd_scaling} shows the SIMD efficiency of source term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models have higher SIMD-efficiency, \textasciitilde\num{0.75}, and the smaller models have lower SIMD-efficiency, \textasciitilde\num{0.625}.
+This again demonstrates that the SIMD-vectorization becomes efficient as the amount of work to perform increases (with increasing model size).
+Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one; this is likely caused by a combination of use of an OpenCL vector-width greater than the native CPU vector-width (i.e., \num{8} v.s., \num{2}), and the improved data-locality for the vectorized-data ordering, as discussed in~\cref{S:data}, and results a modest \textasciitilde\SIrange{10}{25}{\percent} improvement over the nominal vector width.
+
+\begin{figure}[htbp]
+   \centering
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{source_parallel_scaling.pdf}
+      \caption{The strong parallel scaling efficiency (as defined in~\cref{e:strong_scaling}) of source term evaluation for Intel OpenCL\slash OpenMP on the \avx/ machine.}
+      \label{F:source_parallel_scaling}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{source_simd_efficiency.pdf}
+      \caption{The SIMD efficiency (\cref{e:simd_efficiency}) of source term evaluation for the Intel OpenCL runtime on a single core of the \avx/ CPU.}
+      \label{F:source_simd_scaling}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{source_sse_simd_efficiency.pdf}
+      \caption{The SIMD efficiency (\cref{e:simd_efficiency}) of source term evaluation for the Intel OpenCL runtime on a single core of the \sse/ CPU.}
+      \label{F:source_sse_simd_scaling}
+  \end{subfigure}
+  \caption{The parallel scaling efficiency and SIMD efficiency of source term evaluation for Intel OpenCL on the \avx/ and \sse/ CPUs.}
+  \label{F:source_scaling}
 \end{figure}
+
+\subsubsection{Jacobian evaluation}
+\label{S:jacobian_results}
 

### 2018-02-19 scripts/general_plotting.py

@@ -9,21 +9,26 @@ font_size = 'large'
 
 
 def process_data(plotdata, plot, reacs_as_x=True,
-                 plot_cores=False, plot_condition_scaling=False):
+                 plot_cores=False, plot_condition_scaling=False,
+                 plot_simd=False):
     """
     Process the data into an easily usable form
     """
+    def __mechx(datum):
+        return datum.mechdata.n_reactions if reacs_as_x else datum.mechdata.n_species
+
     def __mechsort(data):
-        if reacs_as_x:
-            data = sorted(data, key=lambda x: x.mechdata.n_reactions)
-            return data, [x.mechdata.n_reactions for x in data]
-        else:
-            data = sorted(data, key=lambda x: x.mechdata.n_species)
-            return data, [x.mechdata.n_species for x in data]
+        data = sorted(data, key=lambda x: __mechx(x))
+        return data, [__mechx(x) for x in data]
 
     if plot_cores:
         plotdata = sorted(plotdata, key=lambda x: float(x.cores))
         x_vals = [float(x.cores) for x in plotdata]
+    elif plot_simd:
+        plotdata = sorted(plotdata, key=lambda x: (
+            __mechx(x),
+            ['par', 'w', 'd'].index(x.vectype))),
+        x_vals = [__mechx(x) for x in plotdata]
     elif plot_condition_scaling:
         # first sort by # mech
         plotdata, _ = __mechsort(plotdata)

### 2018-02-19 scripts/plot_styles.py

@@ -19,6 +19,7 @@ legend_style = {'loc': 0,
                 }
 
 tick_font_size = 20
+minor_tick_font_size = 16
 title_font_size = 24
 
 marker_style = {
@@ -107,12 +108,12 @@ color_dict = {x: color_wheel[i] for i, x in enumerate(default_keys)}
 
 
 def finalize(tight=True):
+    # triger tick positioning
     ax = plt.gca()
-    for item in (ax.get_xticklabels() + ax.get_yticklabels()):
-        item.set_fontsize(tick_font_size)
+    if tight:
+        plt.tight_layout()
+
+    plt.tick_params(axis='both', which='major', labelsize=tick_font_size)
 
     for item in (ax.title, ax.xaxis.label, ax.yaxis.label):
         item.set_fontsize(title_font_size)
-
-    if tight:
-        plt.tight_layout()

### 2018-02-19 paper.tex
wide->shallow

@@
 \begin{abstract} % not to exceed 200 words
-A code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized thermo-chemical source term and sparse chemical kinetic Jacobian evaluation was developed and validated for a wide range of chemical kinetic models.
+A code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized thermo-chemical source term and sparse chemical kinetic Jacobian evaluation was developed and validated for a shallow range of chemical kinetic models.
 \todo[inline]{this}
@@
 	\includegraphics[width=\textwidth]{row_major_split.pdf}
-	\caption{Row-major, wide-vectorized data ordering}
+	\caption{Row-major, shallow-vectorized data ordering}
 	\label{F:row_major_split}
@@
 For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the parallel (non-vectorized) Intel OpenCL code is slightly faster than OpenMP in all cases.\todo{rerun OpenMP w/ mtune=native}
-As expected, the \avx/ machine is moderately faster than the \sse/ CPU; additionally, the wide-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.
+As expected, the \avx/ machine is moderately faster than the \sse/ CPU; additionally, the shallow-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.
 
-In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is calculated per-machine, such that the speedup of the wide-vectorized OpenCL code can be readily assessed.
+In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is calculated per-machine, such that the speedup of the shallow-vectorized OpenCL code can be readily assessed.
 On both machines, the unvectorized OpenCL code is \textasciitilde\SIrange{1.25}{1.5}{$\times$} faster than the baseline parallel OpenMP code.
-Additionally, the wide-vectorized OpenCL code is \textasciitilde\SIrange{2.5}{3}{$\times$} and \textasciitilde\SIrange{3.5}{4}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
-In contrast,~\cref{F:pocl_source} shows the runtime of deep, wide and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
+Additionally, the shallow-vectorized OpenCL code is \textasciitilde\SIrange{2.5}{3}{$\times$} and \textasciitilde\SIrange{3.5}{4}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
+In contrast,~\cref{F:pocl_source} shows the runtime of deep, shallow and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
 No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP is faster in both cases.
@@
 
-\begin{figure}[htbp]
+\begin{figure}[htbp!]
    \centering
@@
       \includegraphics[width=\textwidth]{intel_source.pdf}
-      \caption{The speedup achieved over the baseline OpenMP parallelization by both the unvectorized and wide-vectorized Intel OpenCL codes.}
+      \caption{The speedup achieved over the baseline OpenMP parallelization by both the unvectorized and shallow-vectorized Intel OpenCL codes; the speedup is presented on per-machine basis, e.g., the \avx2/ shallow-vectorized code speedup is relative to OpenMP on the same CPU.}
       \label{F:intel_source}
@@
  \caption{Mean runtimes per condition and speedups achieved by the various CPU OpenCL runtimes compared to OpenMP parallelization for each chemical model studied.
-The names in the legends correspond to the identifiers listed in~\cref{t:cpus}}
+The names in the legends correspond to the identifiers listed in~\cref{t:cpus}.}
  \label{F:cpu_source}
@@
 In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, i.e., the roughly \SIrange{1.5}{2}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-order OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
-Additionally, we note that the ``C''-ordered wide-vectorization in~\cref{F:source_cvsf} utilizes the vectorized-data ordering described in~\cref{S:data}.
+Additionally, we note that the ``C''-ordered shallow-vectorization in~\cref{F:source_cvsf} utilizes the vectorized-data ordering described in~\cref{S:data}.
 In general, the data-ordering has a much more profound affect on the vectorized-OpenCL codes; this is due to the nature of SIMD-operations, which are much more efficient when operating on contiguous (in memory) data rather than incurring an expensive memory ``gather'' operation.
-The ``C''-ordered unvectorized OpenCL code (which uses the default unvectorized row-major data ordering) is \textasciitilde\SIrange{1.5}{2.5}{$\times$} faster than the unvectorized ``F''-ordered code, and is in some cases almost as fast as the ``F''-ordered wide-vectorized OpenCL implementation.
-However, the ``C''-ordered wide-vectorization achieves speedups of \textasciitilde\SIrange{4}{7}{$\times$} demonstrating the value of the vectorized data ordering for CPU execution.
+The ``C''-ordered unvectorized OpenCL code (which uses the default unvectorized row-major data ordering) is \textasciitilde\SIrange{1.5}{2.5}{$\times$} faster than the unvectorized ``F''-ordered code, and is in some cases almost as fast as the ``F''-ordered shallow-vectorized OpenCL implementation.
+However, the ``C''-ordered shallow-vectorization achieves speedups of \textasciitilde\SIrange{4}{7}{$\times$} demonstrating the value of the vectorized data ordering for CPU execution.
 
@@
 The speedup reported is calculated per-GPU, such that the effect of data ordering can be seen on each individual GPU.
-In general, the ``C''-ordered wide-vectorization is \textasciitilde\SIrange{1.05}{1.2}{$\times$} faster on the \gpunew/ GPU while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, except for the isopentanol model where the ``F''-ordered data is \textasciitilde\SI{1.4}{$\times$} faster (though, we note that there was high variablity in this case).
+In general, the ``C''-ordered shallow-vectorization is \textasciitilde\SIrange{1.05}{1.2}{$\times$} faster on the \gpunew/ GPU while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, except for the isopentanol model where the ``F''-ordered data is \textasciitilde\SI{1.4}{$\times$} faster (though, we note that there was high variablity in this case).
 \todo{again, pending rerun on \gpuold/}
-The ``C''-ordered speedup on the \gpunew/ GPU is counter to what one might expect, as typically speaking coalesced memory access in a wide-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
+The ``C''-ordered speedup on the \gpunew/ GPU is counter to what one might expect, as typically speaking coalesced memory access in a shallow-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
 However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector width size (and thus, likely coalesced) but the improved data-locality; coupled with the larger cache sizes of the \gpunew/ GPU result in this modest speedup for ``C''-data ordering.
@@
       \includegraphics[width=\textwidth]{source_vector_width.pdf}
-      \caption{The effect of vector-width on ``C''-ordered wide-vectorized Intel OpenCL source term evaluations in the \avx/ CPU.}
+      \caption{The effect of vector-width on ``C''-ordered shallow-vectorized Intel OpenCL source term evaluations in the \avx/ CPU.}
       \label{F:source_vector_width}
@@
       \includegraphics[width=\textwidth]{source_gpu_cvsf.pdf}
-      \caption{The affect on performance of ``C'' vs ``F''-ordering for the wide-vectorized NVIDIA OpenCL code on both GPUs with a vector width of 128.
+      \caption{The affect on performance of ``C'' vs ``F''-ordering for the shallow-vectorized NVIDIA OpenCL code on both GPUs with a vector width of 128.
 The speedup presented is calculated per-GPU to better assess the effect of the data-ordering on each GPU.}
@@
   \caption{The effect of the \conp/ and \conv/ formulations, as well as ``C'' and ``F'' data-ordering on source term evaluation performance in \texttt{pyJac}.
-	   It is noted the wide-vectorized ``C'' ordered OpenCL case corresponds to the vectorized data ordering described in~\cref{S:data}}
+	   It is noted the shallow-vectorized ``C'' ordered OpenCL case corresponds to the vectorized data ordering described in~\cref{S:data}}
   \label{f:source_permutate}
@@
 
-\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source rate evaluation in \texttt{pyJac} on the \avx/ machine for both the wide-vectorized Intel OpenCL and OpenMP codes.
+\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source rate evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
 In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for Intel OpenCL; although the mechanism scalings vary somewhat for OpenMP, there is no clear trend.
 It is likely the poor scaling of the \ce{H2}\slash\ce{CO} for Intel OpenCL results from both its relatively small size, and few falloff\slash chemically activated reaction (and in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD intructions.
-Additionally, OpenMP tends to have higher scalings then the wide-vectorized OpenCL code, e.g., \textasciitilde\num{0.8} and \textasciitilde\num{0.7} for \num{4} and \num{8} CPU cores respectively, compared to just \textasciitilde\num{0.7} and \textasciitilde\num{0.5} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the SIMD-vectorization itself the cause of this poorer scaling.
+Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.8} and \textasciitilde\num{0.7} for \num{4} and \num{8} CPU cores respectively, compared to just \textasciitilde\num{0.7} and \textasciitilde\num{0.5} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the SIMD-vectorization itself the cause of this poorer scaling.
 This is again likely due to relatively small amounts of work associated with source rate evaluation; the longest runtime (the isopentanol model on a single core) takes only \textasciitilde$\SI{2}{\sec}$ for OpenCL, and \textasciitilde$\SI{9}{\sec}$ for OpenMP.
-By the time 4--8 CPU cores are utilized, the runtime of wide-vectorized OpenCL code can be measured in hundreds of milliseconds, and parallel overhead begins to dominate, decreasing efficiency.
-In contrast, the OpenMP code requires \textasciitilde\num{4} cores to run as fast as the wide-vectorized OpenCL code in this case, hence parallel overhead concerns become increasingly important for 8--16 cores.
+By the time 4--8 CPU cores are utilized, the runtime of shallow-vectorized OpenCL code can be measured in hundreds of milliseconds, and parallel overhead begins to dominate, decreasing efficiency.
+In contrast, the OpenMP code requires \textasciitilde\num{4} cores to run as fast as the shallow-vectorized OpenCL code in this case, hence parallel overhead concerns become increasingly important for 8--16 cores.
 Indeed, we will see that the strong scalings for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.
@@
  \label{e:simd_efficiency}
- \varepsilon_{\text{SIMD}} = \frac{\bar{t}_{\text{unvec}}}{W \bar{t}_{\text{wide}}}
+ \varepsilon_{\text{SIMD}} = \frac{\bar{t}_{\text{unvec}}}{W \bar{t}_{\text{shallow}}}
 \end{equation}
-where $\bar{t}_{\text{unvec}}$ is the mean runtime per condition of the unvectorized OpenCL code, $\bar{t}_{\text{wide}}$ the same for the wide-vectorized OpenCL code, and $W$, the vector width reported in number of double operations in~\cref{t:cpus}
-This measure compares the actual speedup due to wide-vectorization compared to the nominal vector-width of the machine in question.
+where $\bar{t}_{\text{unvec}}$ is the mean runtime per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$, the vector width reported in number of double operations in~\cref{t:cpus}
+This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.
 \Cref{F:source_simd_scaling} shows the SIMD efficiency of source term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models have higher SIMD-efficiency, \textasciitilde\num{0.75}, and the smaller models have lower SIMD-efficiency, \textasciitilde\num{0.625}.
@@
 \section{Conclusions}
-In this work, automatically generated OpenCL codes for shallow SIMD-vectorized thermo-chemical source term evaluation were developed and validated against Cantera~\cite{Cantera} for a wide range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
+In this work, automatically generated OpenCL codes for shallow SIMD-vectorized thermo-chemical source term evaluation were developed and validated against Cantera~\cite{Cantera} for a shallow range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
 Significant speedups of up to \SIrange{1.90}{2.45}{$\times$} over a baseline SIMT-vectorized code were observed.

### 2018-02-20 paper.tex
edits

@@
 Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
-The wrapping code---responsible for initializing\slash transfering memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms, and \texttt{gcc 4.8.5} on the \sse/\slash\gpuold/ machines.
-The optimization level ``\texttt{-O3}'' was used in all cases, and no ``fast-math'' OpenCL optimizations were enabled.
+The wrapping code---responsible for initializing\slash transfering memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms, and \texttt{gcc v4.8.5} on the \sse/\slash\gpuold/ machines.
+The optimization level ``\texttt{-O3 -mtune=native}''\todo[inline]{again, rerun OpenMP w/ this} was used, and no ``fast-math'' OpenCL optimizations were enabled.
 Additionally, the ``exact'' form of the Jacobian (as opposed the the ``appropriate'' form discussed in~\cref{S:sparsity}) was used in all cases.
-Finally, unless otherwise stated the \conp/ assumption was used, a vector width of \num{8}\slash\num{128} was used for all CPU\slash GPU cases respectively, ``C''\slash ``F''-ordered data was used on the CPU\slash GPU, and the runtimes reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
+Finally, unless stated otherwise, the \conp/ assumption was used, along with a vector width of \num{8}\slash\num{128} for all CPU\slash GPU cases respectively; ``C''\slash ``F''-ordered data was used on the CPU\slash GPU, and the runtimes reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
 The effects of choice of vector-width on the CPU\slash GPU, as well as the differences between \conp/ and \conv/ evaluations will be explored in~\cref{S:source_results}, and briefly discussed in~\cref{S:jacobian_results}.
@@
 \Cref{F:cpu_source} explores the performance of the source term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
-Source term evaluations, critial in their own right for direct numerical simulations~\cite{Spafford:2010aa} among other applications, also give us a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation peformance.
+Source term evaluations---critial in their own right for direct numerical simulations of reacting flow~\cite{Spafford:2010aa}, among other applications---also give us a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation peformance.
 
 In~\cref{F:intel_source_noorm}, the mean runtime per initial condition is shown for both the~\avx/ and~\sse/ CPUs, using both Intel OpenCL and OpenMP.
-This normalization is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{T:error}).
-For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the parallel (non-vectorized) Intel OpenCL code is slightly faster than OpenMP in all cases.\todo{rerun OpenMP w/ mtune=native}
-As expected, the \avx/ machine is moderately faster than the \sse/ CPU; additionally, the shallow-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.
+This normalization of the runtime by the number of initial conditions tested is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{T:error}).
+For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the unvectorized (strictly parallely) Intel OpenCL code is slightly faster than OpenMP in all cases.\todo{rerun OpenMP w/ mtune=native}
+As expected, the \avx/ machine is moderately faster than the \sse/ CPU for these strictly parallely cases; additionally, the shallow-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.
 
-In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is calculated per-machine, such that the speedup of the shallow-vectorized OpenCL code can be readily assessed.
+In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is calculated per-machine, e.g., the \avx/ shallow-vectorized code speedup is relative to OpenMP on the same CPU such that the speedup of the shallow-vectorized OpenCL code can be readily assessed.
 On both machines, the unvectorized OpenCL code is \textasciitilde\SIrange{1.25}{1.5}{$\times$} faster than the baseline parallel OpenMP code.
@@
 No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP is faster in both cases.
-The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work, however briefly speaking vectorization might be achieved on POCL via a combination of relaxed math (``-cl-fast-relaxed-math''), use of explicit OpenCL vector types and newer versions of LLVM (specifically, potential improvments to the vectorization module).
+The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work; however, briefly speaking vectorization might be achieved on POCL via a combination of relaxed math (``\texttt{-cl-fast-relaxed-math}''), use of explicit OpenCL vector types and newer versions of LLVM\slash POCL (specifically, potential improvements to the vectorization module).
 This will be explored more in future work, but for the moment POCL is still quite useful as a validation tool.
@@
       \includegraphics[width=\textwidth]{intel_source.pdf}
-      \caption{The speedup achieved over the baseline OpenMP parallelization by both the unvectorized and shallow-vectorized Intel OpenCL codes; the speedup is presented on per-machine basis, e.g., the \avx2/ shallow-vectorized code speedup is relative to OpenMP on the same CPU.}
+      \caption{The speedup achieved over the baseline OpenMP parallelization by both the unvectorized and shallow-vectorized Intel OpenCL codes; the speedup is presented on per-machine basis.}
       \label{F:intel_source}
@@
 \Cref{F:gpu_source} shows the performance of the source term evaluation on the GPUs listed in~\cref{t:gpus}.
-In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean runtime per condition; the mean runtime continues to decrease for all chemical models until \textasciitilde\num{e4} conditions, at which point the GPU becomes saturated and performance levels off.
+In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean runtime per condition; the mean runtime continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
 \Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling}; the speedup ranges from \textasciitilde\SIrange{1.75}{2.5}{$\times$} depending on the chemical model in question.
@@
 
-\Cref{f:source_permutate} shows the effect of changing data-ordering patterns as well as the \conp/ or \conv/ formulation on the performance of source term evaluation in \texttt{pyJac}.
+\Cref{f:source_permutate} shows the effect of changing data-ordering patterns, the \conp/ or \conv/ formulation, and the CPU vector-width size on the performance of source term evaluation in \texttt{pyJac}.
 In~\cref{F:source_conpvsconv}, we see that choice of the constant-pressure or constant-volume formulation has little to no effect on runtime for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx/ machine.
@@
 
-In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, i.e., the roughly \SIrange{1.5}{2}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-order OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
+In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, i.e., the roughly \SIrange{1.5}{2}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-ordered OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
 Additionally, we note that the ``C''-ordered shallow-vectorization in~\cref{F:source_cvsf} utilizes the vectorized-data ordering described in~\cref{S:data}.
@@
 
-\Cref{F:source_gpu_cvsf} shows the effect of the ``C'' versus ``F''-ordering on source term evaluation performance on both GPUs.
-The speedup reported is calculated per-GPU, such that the effect of data ordering can be seen on each individual GPU.
+\Cref{F:source_gpu_cvsf} shows the mean runtime per-condition of the ``C'' versus ``F''-ordering on source term evaluation performance on both GPUs.
 In general, the ``C''-ordered shallow-vectorization is \textasciitilde\SIrange{1.05}{1.2}{$\times$} faster on the \gpunew/ GPU while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, except for the isopentanol model where the ``F''-ordered data is \textasciitilde\SI{1.4}{$\times$} faster (though, we note that there was high variablity in this case).
@@
 The ``C''-ordered speedup on the \gpunew/ GPU is counter to what one might expect, as typically speaking coalesced memory access in a shallow-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
-However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector width size (and thus, likely coalesced) but the improved data-locality; coupled with the larger cache sizes of the \gpunew/ GPU result in this modest speedup for ``C''-data ordering.
+However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector width size (and thus, likely coalesced) but the improved data-locality, coupled with the larger cache sizes of the \gpunew/ GPU result in this modest speedup for ``C''-data ordering.
+
+\Cref{F:source_vector_width} shows the effect of changing vector-width size on source rate evaluation performance the \avx/ CPU.
+The vector-width of size \num{8} was the fastest tested, while the larger vector-width of \num{16} is slightly slower due to increased register pressure~\cite{intel_vectypes}.
+It is unclear why the vector-width of \num{4} results in no speedup at all (it is in fact, the slowest case); Intel's vectorization guide~\cite{intel_vecknobs} mentions that a heuristic determines the optimal vector-width (in this case, it appears from compiler output to be \num{8}), so it is possible using vector-width smaller than the heuristic breaks the implicit vectorizer.
+It is noted that this issue does not occur for a vector width of \num{4} on the \sse/ CPU.
 
@@
   \begin{subfigure}[t]{0.48\linewidth}
-      \includegraphics[width=\textwidth]{source_vector_width.pdf}
-      \caption{The effect of vector-width on ``C''-ordered shallow-vectorized Intel OpenCL source term evaluations in the \avx/ CPU.}
-      \label{F:source_vector_width}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.48\linewidth}
       \includegraphics[width=\textwidth]{source_gpu_cvsf.pdf}
@@
   \end{subfigure}
-  \caption{The effect of the \conp/ and \conv/ formulations, as well as ``C'' and ``F'' data-ordering on source term evaluation performance in \texttt{pyJac}.
-	   It is noted the shallow-vectorized ``C'' ordered OpenCL case corresponds to the vectorized data ordering described in~\cref{S:data}}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{source_vector_width.pdf}
+      \caption{The effect of vector-width on ``C''-ordered shallow-vectorized Intel OpenCL source term evaluations in the \avx/ CPU.}
+      \label{F:source_vector_width}
+  \end{subfigure}
+  \caption{The effect of the \conp/ and \conv/ formulations, ``C'' and ``F'' data-ordering, and CPU vector-width on source term evaluation performance in \texttt{pyJac}.
+	   It is noted the shallow-vectorized ``C'' ordered OpenCL cases corresponds to the vectorized data ordering described in~\cref{S:data}}
   \label{f:source_permutate}
@@
 Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.8} and \textasciitilde\num{0.7} for \num{4} and \num{8} CPU cores respectively, compared to just \textasciitilde\num{0.7} and \textasciitilde\num{0.5} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the SIMD-vectorization itself the cause of this poorer scaling.
-This is again likely due to relatively small amounts of work associated with source rate evaluation; the longest runtime (the isopentanol model on a single core) takes only \textasciitilde$\SI{2}{\sec}$ for OpenCL, and \textasciitilde$\SI{9}{\sec}$ for OpenMP.
+This is again likely due to relatively small amounts of work associated with source rate evaluation; the longest runtime (the isopentanol model on a single core) takes only \textasciitilde\SI{2}{\second} for OpenCL, and \textasciitilde\SI{9}{\second} for OpenMP.
 By the time 4--8 CPU cores are utilized, the runtime of shallow-vectorized OpenCL code can be measured in hundreds of milliseconds, and parallel overhead begins to dominate, decreasing efficiency.
-In contrast, the OpenMP code requires \textasciitilde\num{4} cores to run as fast as the shallow-vectorized OpenCL code in this case, hence parallel overhead concerns become increasingly important for 8--16 cores.
-Indeed, we will see that the strong scalings for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.
+In contrast, the OpenMP code requires \textasciitilde\num{4} cores to match the performance of the shallow-vectorized OpenCL code in this case, hence parallel overhead concerns become increasingly important for 8--16 cores.
+Indeed, as will be investigated in~\cref{s:jacobian_results} the strong scalings for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.
 
@@
 \end{equation}
-where $\bar{t}_{\text{unvec}}$ is the mean runtime per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$, the vector width reported in number of double operations in~\cref{t:cpus}
+where $\bar{t}_{\text{unvec}}$ is the mean runtime per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$, the vector width reported in number of double operations in~\cref{t:cpus}.
 This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.
-\Cref{F:source_simd_scaling} shows the SIMD efficiency of source term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models have higher SIMD-efficiency, \textasciitilde\num{0.75}, and the smaller models have lower SIMD-efficiency, \textasciitilde\num{0.625}.
+\Cref{F:source_simd_scaling} shows the SIMD efficiency of source term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models have higher SIMD-efficiency of \textasciitilde\num{0.75}, and the smaller models have lower SIMD-efficiency, \textasciitilde\num{0.625}.
 This again demonstrates that the SIMD-vectorization becomes efficient as the amount of work to perform increases (with increasing model size).

### 2018-02-21 paper.tex
add CSC/CSR matrix format discussion, start on Jacobian data

@@
 Error statistics are the largest filtered relative error $E_\mathcal{C}$ for over all samples.
-It is noted that the threshold described in~\cref{e:thresh} varies slightly between the \conp/ and \conv/ cases; the reported $\bar{\mathcal{T}}$ is the average of the two, however the appropriate value was used during calculations of the error statistics..
+It is noted that the threshold described in~\cref{e:thresh} varies slightly between the \conp/ and \conv/ cases; the reported $\bar{\mathcal{T}}$ is the average of the two, however the appropriate value was used during calculations of the error statistics.
 }
@@
 \end{figure}
+
+Currently, \texttt{pyJac} can utilize two common sparse matrix storage formats~\cite{netlib_templates}, compressed row\slash column storage (CSR\slash CSC), used for ``C'' and ``F''-ordered data respectively.
+These formats contain three vectors, for convenience we will outline only the CSR format but the CSC format is very similar~\cite{netlib_templates}.
+An $N\times N$ CSC matrix is composed of three vectors, a value vector of length $NNZ$---the number of non-zero elements in the Jacobian---which stores the elements of the Jacobian, a row pointer vector---of length $N$---which stores the locations in the value vector that begin a row, and a column index vector---of length $NNZ$---which stores the column indicies of the elements in the value vector.
+
+The Jacobian access pattern used by \texttt{pyJac} is fairly irregular currently; for simplicity we will only discuss looping-structure of species derivatives calculations as these form the bulk of the computation, and have the most challenging Jacobian access patterns.
+In general, an outer loop iterates over all reactions of a certain type (e.g., falloff reactions) and calculates the relevant Jacobian subproducts (e.g., the derivative of the falloff pressure modification term with respect to a species $j$).
+Two inner loops then iterate over the species in a reaction, updating the Jacobian entries for these species as appropriate.
+This pattern leads to fairly easily vectorizable code, and efficient Jacobian evaluation as the bulk of the computation is dependent only on the reaction in question, as discussed in our previous work~\cite{Niemeyer:2016aa}.
+Generally, this means that a lookup operation is required to find the sparse Jacobian index for any pair of state variables; in some cases this can be avoided, e.g., the rows corresponding to derivatives of the temperature and thermodynamic state parameter source terms are fully dense in \texttt{pyJac}, and hence no lookup is necessary.
+This lookup operation is currently implemented as a simple for-loop, e.g., for a sparse lookup of a pair of indicies $(i, j)$ in a compressed row storage matrix, the lookup function searches the column index vector between the values $\texttt{row\_pointer}[i] \ldots \texttt{row\_pointer}[i + 1]$ for $j$, and returns the offset from $\texttt{row\_pointer}[i]$ (or $\num{-1}$ if not found).
+As will be explored in~\cref{S:jacobian_results}, this slows down sparse Jacobian evaluation, and might be improved---at the cost of increased constant data usage, a limitation for OpenCL---by a static mapping of the full Jacobian indicies to the sparse index (or some null value if the entry is empty).
+Additionally, this might be an excellent usage of OpenCL's Image memory type (similar to texture memory in CUDA terminology) and both of these sparse indexing techniques merit future investigation.
 
@@
 In contrast, the OpenMP code requires \textasciitilde\num{4} cores to match the performance of the shallow-vectorized OpenCL code in this case, hence parallel overhead concerns become increasingly important for 8--16 cores.
-Indeed, as will be investigated in~\cref{s:jacobian_results} the strong scalings for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.
+Indeed, as will be investigated in~\cref{S:jacobian_results} the strong scalings for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.
 
@@
 \label{S:jacobian_results}
+\Cref{F:jacobian_perfomance} shows the performance of the sparse and dense Jacobian evaluations in \texttt{pyJac} on the CPU platforms.
+In~\cref{F:sparse_vs_dense}, the mean runtime per condition is presented for the shallow-vectorized Intel OpenCL and OpenMP codes on the \avx/ CPU.
+The sparse Jacobian evaluations are slower on both Intel OpenCL and OpenMP due to indirect lookup indexing requirements, as discussed in~\cref{S:sparsity}.
+Interestingly the shallow-vectorized OpenCL code is less negatively impacted by the indirect lookup; the sparse OpenMP code is \SIrange{2.5}{7}{$\times$} slower than the dense OpenMP evaluation, while the sparse shallow-vectorized OpenCL code is just \SIrange{1.4}{3.3}{$\times$} slower than its dense counterpart.
+Indeed, the shallow-vectorized sparse OpenCL code is faster than the dense OpenMP code in all cases.
+\Cref{F:sparse_vs_dense_speedup} shows the speedup of the sparse\slash dense shallow-vectorized OpenCL the over the same sparse\slash dense Jacobian format on OpenMP; a speedup of \SIrange{3.7}{4.2}{$\times$} is exhibited for the dense OpenCL code, increasing to \SIrange{6.6}{8.4}{$\times$} for the sparse Jacobian.
+
+\begin{figure}[htbp]
+   \centering
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{sparse_vs_dense.pdf}
+      \caption{The mean runtime per-condition of sparse and dense Jacobian evaluation for Intel OpenCL\slash OpenMP on the \avx/ machine.}
+      \label{F:sparse_vs_dense}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{sparse_vs_dense_speedup.pdf}
+      \caption{The speedup of the sparse and dense Jacobian evaluation for Intel OpenCL over the sparse\slash dense OpenMP baseline on the \avx/ machine.}
+      \label{F:sparse_vs_dense_speedup}
+  \end{subfigure}
+  \caption{Performance of sparse and dense Jacobian evaluation in \texttt{pyJac}.}
+  \label{F:jacobian_perfomance}
+\end{figure}
 

### 2018-02-22 paper.tex
updates

@@
 The sparse Jacobian evaluations are slower on both Intel OpenCL and OpenMP due to indirect lookup indexing requirements, as discussed in~\cref{S:sparsity}.
-Interestingly the shallow-vectorized OpenCL code is less negatively impacted by the indirect lookup; the sparse OpenMP code is \SIrange{2.5}{7}{$\times$} slower than the dense OpenMP evaluation, while the sparse shallow-vectorized OpenCL code is just \SIrange{1.4}{3.3}{$\times$} slower than its dense counterpart.
-Indeed, the shallow-vectorized sparse OpenCL code is faster than the dense OpenMP code in all cases.
-\Cref{F:sparse_vs_dense_speedup} shows the speedup of the sparse\slash dense shallow-vectorized OpenCL the over the same sparse\slash dense Jacobian format on OpenMP; a speedup of \SIrange{3.7}{4.2}{$\times$} is exhibited for the dense OpenCL code, increasing to \SIrange{6.6}{8.4}{$\times$} for the sparse Jacobian.
+Interestingly the shallow-vectorized OpenCL code is less negatively impacted by the indirect lookup; the sparse OpenMP code is \SIrange{2.5}{7}{$\times$} slower than the dense OpenMP evaluation, while the sparse shallow-vectorized OpenCL code is just \SIrange{1.4}{3.3}{$\times$} slower than its dense counterpart
+As a result, the shallow-vectorized sparse OpenCL code is faster than the dense OpenMP code in all cases on the \avx/ machine (\cref{F:sparse_vs_dense}).
+\Cref{F:sparse_vs_dense_speedup} shows the speedup of the sparse\slash dense shallow-vectorized OpenCL the over the same sparse\slash dense Jacobian format on OpenMP; the dense OpenCL is \SIrange{3.7}{4.2}{$\times$} faster than the corresponding dense OpenMP code, this speedup increases to \SIrange{6.6}{8.4}{$\times$} for the sparse Jacobian.
+On the \sse/ machine,~\cref{F:sparse_vs_dense_sse} shows similar results; the sparse OpenMP code is the slowest in all cases, and the shallow-vectorized OpenCL code is nearly as fast as the dense OpenMP code.
+Once again, the sparse OpenCL code is far less negatively impacted by the indirect lookup, and is only \SIrange{1.8}{3.2}{$\times$} slower than its dense counterpart, while the sparse OpenMP code is significantly (\SIrange{2.25}{6.2}{$\times$}) slower than the dense version.
+In~\cref{F:sparse_vs_dense_sse_speedup}, the speedup of the sparse and dense shallow-vectorized OpenCL codes are compared to their OpenMP version; the dense OpenCL code is \SIrange{3.8}{4.2}{$\times$} faster, while the sparse shallow-vectorization achieves a speedup of \SIrange{6.6}{9.1}{$\times$}.
 
@@
   \end{subfigure}
-  \caption{Performance of sparse and dense Jacobian evaluation in \texttt{pyJac}.}
+  \\
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{sparse_vs_dense_sse.pdf}
+      \caption{The mean runtime per-condition of sparse and dense Jacobian evaluation for Intel OpenCL\slash OpenMP on the \sse/ machine.}
+      \label{F:sparse_vs_dense_sse}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{sparse_vs_dense_sse_speedup.pdf}
+      \caption{The speedup of the sparse and dense Jacobian evaluation for Intel OpenCL over the sparse\slash dense OpenMP baseline on the \sse/ machine.}
+      \label{F:sparse_vs_dense_sse_speedup}
+  \end{subfigure}
+  \caption{Performance of sparse and dense Jacobian evaluation on the CPU platforms in \texttt{pyJac}.}
   \label{F:jacobian_perfomance}
+\end{figure}
+
+In~\cref{F:sparse_jac_scaling}, the parallel scaling efficiency of the sparse shallow-vectorized OpenCL is compared to the sparse OpenMP code.
+The plot is a bit difficult to read, as most of the data-points are clustered together however the key takeaway is that the parallel scaling efficiency of the shallow-vectorized OpenCL code is much closer in value to that of the OpenMP code as compared to the source rate evaluations (\cref{F:source_parallel_scaling}).
+The \ce{H2}\slash\ce{CO} model has the worst scaling for both codes, ranging from \numrange{0.93}{0.54} and \numrange{0.99}{0.86} for OpenCL and OpenMP respectively on \numrange{2}{16} cores.
+As the model size increases, the efficiency of the OpenCL code improves dramatically, reaching \numrange{0.997}{0.84} for the isopentanol model.
+It is worth noting that the OpenCL code's scaling is more negatively affected by the switch from a single 12-core CPU (\cref{t:cpus}) to two CPUs (i.e., the 16 core data-point) than the OpenMP code.
+In contrast,~\cref{F:dense_jac_scaling} shows scaling for the shallow-vectorized dense Jacobian OpenCL code that is higher than for source-rate evaluation, but not as good as for sparse Jacobian evaluation.
+
+\begin{figure}[htbp]
+   \centering
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{sparse_jac_scaling.pdf}
+      \caption{Sparse Jacobian evaluation parallel scaling efficiency for Intel OpenCL\slash OpenMP on the \avx/ machine.}
+      \label{F:sparse_jac_scaling}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{dense_jac_scaling.pdf}
+      \caption{Dense Jacobian evaluation parallel scaling efficiency for Intel OpenCL\slash OpenMP on the \avx/ machine.}
+      \label{F:dense_jac_scaling}
+  \end{subfigure}
+  \caption{The parallel scaling efficiencies of sparse and dense Jacobian evaluation for the shallow-vectorized Intel OpenCL and OpenMP codes on the \avx/ CPU.}
 \end{figure}

### 2018-02-22 paper.tex
done for tonight

@@
 In contrast,~\cref{F:dense_jac_scaling} shows scaling for the shallow-vectorized dense Jacobian OpenCL code that is higher than for source-rate evaluation, but not as good as for sparse Jacobian evaluation.
+In this case, the USC-Mech II model\todo{note: don't have data for ipentanol yet} has the worst scaling for both codes; ranging from \numrange{0.96}{0.63} and \numrange{0.91}{0.34} for OpenMP and OpenCL respectively.
+As in~\cref{S:source_results}, we note that the shallow-vectorized OpenCL code on a single core of the \avx/ machine is roughly as fast than the OpenMP code on four cores of the same; hence, the OpenMP code has more room to improve performance before parallel-overhead and memory-bandwidth issues become a limiting concern.
 
@@
 \end{figure}
+
+The GPU performance
 

### 2018-02-23 paper.tex
updates from last night

@@
 \midrule
-\ce{H2}\slash \ce{CO} & \num{900900}                      & \num{6.431e+18}      & \num{8.084e-1}  & \num{1.907e-5} \\
-GRI-Mech 3.0          & \num{450900}                      & \num{7.783e+19}      & \num{1.890e-7}  & \num{1.877e-7} \\
-USC-Mech II           & \num{91800}                       & \num{2.830e+21}      & \num{1.540e-3}  & \num{1.704e-7} \\
+\ce{H2}\slash \ce{CO} & \num{900900}                      & \num{6.431e+18}      & \num{1.742e00}  & \num{4.508e-5} \\
+GRI-Mech 3.0          & \num{450900}                      & \num{7.783e+19}      & \num{3.842e-7}  & \num{3.687e-7} \\
+USC-Mech II           & \num{91800}                       & \num{2.830e+21}      & \num{1.119e-2}  & \num{1.983e-7} \\
 \ce{iC5H11OH}         & \num{450900}                      & \num{2.733e+26}      & \num{1.363e-3}  & \num{2.764e-5} \\
@@
 
-In~\cref{F:intel_source_noorm}, the mean runtime per initial condition is shown for both the~\avx/ and~\sse/ CPUs, using both Intel OpenCL and OpenMP.
+In~\cref{F:intel_source_nonorm}, the mean runtime per initial condition is shown for both the~\avx/ and~\sse/ CPUs, using both Intel OpenCL and OpenMP.
 This normalization of the runtime by the number of initial conditions tested is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{T:error}).
@@
 In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is calculated per-machine, e.g., the \avx/ shallow-vectorized code speedup is relative to OpenMP on the same CPU such that the speedup of the shallow-vectorized OpenCL code can be readily assessed.
-On both machines, the unvectorized OpenCL code is \textasciitilde\SIrange{1.25}{1.5}{$\times$} faster than the baseline parallel OpenMP code.
-Additionally, the shallow-vectorized OpenCL code is \textasciitilde\SIrange{2.5}{3}{$\times$} and \textasciitilde\SIrange{3.5}{4}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
+On both machines, the unvectorized OpenCL code is faster than the baseline parallel OpenMP code; by \SIrange{1.25}{1.51}{$\times$} on the \avx/ CPU, and \textasciitilde\SI{1.25}{$\times$} on the \sse/ machine.
+Additionally, the shallow-vectorized OpenCL code is \textasciitilde\SIrange{2.5}{2.9}{$\times$} and \textasciitilde\SIrange{3.4}{4.1}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
 In contrast,~\cref{F:pocl_source} shows the runtime of deep, shallow and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
@@
       \caption{The mean runtime per condition for each chemical model using the Intel OpenCL and OpenMP on both CPUs studied.}
-      \label{F:intel_source_noorm}
+      \label{F:intel_source_nonorm}
   \end{subfigure}
@@
 In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean runtime per condition; the mean runtime continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
-\Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling}; the speedup ranges from \textasciitilde\SIrange{1.75}{2.5}{$\times$} depending on the chemical model in question.
+\Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling}; the speedup ranges from \textasciitilde\SIrange{1.7}{2.6}{$\times$} depending on the chemical model in question.
 Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector width size (a.k.a., the block size in NVIDIA terminology) has minimal effect on performance; it is possible that use of larger vector sizes could have an effect on performance, but this was not considered in this work.\todo{rerun NVIDIA on srv2 -- see if that high variability in isopentanol persists}
 
-\begin{figure}[htbp!]
+\begin{figure}[htbp]
    \centering
@@
 
-In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, i.e., the roughly \SIrange{1.5}{2}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-ordered OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
+In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, i.e., the \SIrange{1.35}{2.1}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-ordered OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
 Additionally, we note that the ``C''-ordered shallow-vectorization in~\cref{F:source_cvsf} utilizes the vectorized-data ordering described in~\cref{S:data}.
 In general, the data-ordering has a much more profound affect on the vectorized-OpenCL codes; this is due to the nature of SIMD-operations, which are much more efficient when operating on contiguous (in memory) data rather than incurring an expensive memory ``gather'' operation.
-The ``C''-ordered unvectorized OpenCL code (which uses the default unvectorized row-major data ordering) is \textasciitilde\SIrange{1.5}{2.5}{$\times$} faster than the unvectorized ``F''-ordered code, and is in some cases almost as fast as the ``F''-ordered shallow-vectorized OpenCL implementation.
-However, the ``C''-ordered shallow-vectorization achieves speedups of \textasciitilde\SIrange{4}{7}{$\times$} demonstrating the value of the vectorized data ordering for CPU execution.
+The ``C''-ordered unvectorized OpenCL code (which uses the default unvectorized row-major data ordering) is \SIrange{1.6}{2.3}{$\times$} faster than the unvectorized ``F''-ordered code, and is in some cases almost as fast as the ``F''-ordered shallow-vectorized OpenCL implementation.
+However, the ``C''-ordered shallow-vectorization achieves speedups of \SIrange{3.9}{7.2}{$\times$} demonstrating the value of the vectorized data ordering for CPU execution.
 
 \Cref{F:source_gpu_cvsf} shows the mean runtime per-condition of the ``C'' versus ``F''-ordering on source term evaluation performance on both GPUs.
-In general, the ``C''-ordered shallow-vectorization is \textasciitilde\SIrange{1.05}{1.2}{$\times$} faster on the \gpunew/ GPU while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, except for the isopentanol model where the ``F''-ordered data is \textasciitilde\SI{1.4}{$\times$} faster (though, we note that there was high variablity in this case).
+In general, the ``C'' and ``F''-ordered shallow-vectorizations are nearly equivalent on both GPUs while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, with the exception of the isopentanol model where the ``C''-ordered data is up to \SI{1.1}{$\times$} faster on the \gpunew/ (we note that there was high variablity for the \gpuold/ in this case).
 \todo{again, pending rerun on \gpuold/}
-The ``C''-ordered speedup on the \gpunew/ GPU is counter to what one might expect, as typically speaking coalesced memory access in a shallow-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
-However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector width size (and thus, likely coalesced) but the improved data-locality, coupled with the larger cache sizes of the \gpunew/ GPU result in this modest speedup for ``C''-data ordering.
+The rough equivalence between the ``C'' and ``F''-ordered codes GPUs is counter to what one might expect, as typically speaking coalesced memory access in a shallow-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
+However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector width size and thus, likely coalesced.
+% but the improved data-locality, coupled with the larger cache sizes of the \gpunew/ GPU result in this modest speedup for ``C''-data ordering.
 
@@
 In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for Intel OpenCL; although the mechanism scalings vary somewhat for OpenMP, there is no clear trend.
-It is likely the poor scaling of the \ce{H2}\slash\ce{CO} for Intel OpenCL results from both its relatively small size, and few falloff\slash chemically activated reaction (and in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD intructions.
-Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.8} and \textasciitilde\num{0.7} for \num{4} and \num{8} CPU cores respectively, compared to just \textasciitilde\num{0.7} and \textasciitilde\num{0.5} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the SIMD-vectorization itself the cause of this poorer scaling.
+It is likely the poor scaling of the \ce{H2}\slash\ce{CO} for Intel OpenCL results from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD intructions.
+Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.
 This is again likely due to relatively small amounts of work associated with source rate evaluation; the longest runtime (the isopentanol model on a single core) takes only \textasciitilde\SI{2}{\second} for OpenCL, and \textasciitilde\SI{9}{\second} for OpenMP.
-By the time 4--8 CPU cores are utilized, the runtime of shallow-vectorized OpenCL code can be measured in hundreds of milliseconds, and parallel overhead begins to dominate, decreasing efficiency.
-In contrast, the OpenMP code requires \textasciitilde\num{4} cores to match the performance of the shallow-vectorized OpenCL code in this case, hence parallel overhead concerns become increasingly important for 8--16 cores.
-Indeed, as will be investigated in~\cref{S:jacobian_results} the strong scalings for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.
+By the time \numrange{4}{8} CPU cores are utilized in the shallow-vectorized OpenCL code, the runtime can be measured in tens to hundreds of milliseconds and parallel overhead begins to dominate, decreasing efficiency.
+In contrast, the OpenMP code running on \num{4} cores is as fast (\ce{H2}\slash\ce{CO}) or slightly slower than the shallow-vectorized OpenCL code on a single core, hence for \numrange{8}{16} OpenMP cores the parallel overhead concerns become increasingly important; correspondingly, the OpenMP code's parallel efficiency drops to \num{0.64} for \num{16} cores, similar to the OpenCL code at \num{4} cores.
+Indeed, as will be investigated in~\cref{S:jacobian_results} the strong parallel scaling efficiency for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.
 
@@
 This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.
-\Cref{F:source_simd_scaling} shows the SIMD efficiency of source term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models have higher SIMD-efficiency of \textasciitilde\num{0.75}, and the smaller models have lower SIMD-efficiency, \textasciitilde\num{0.625}.
+\Cref{F:source_simd_scaling} shows the SIMD efficiency of source term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models (isopentanol and USC-Mech II) have higher SIMD-efficiency of \numrange{0.76}{0.78}, and the smaller models (\ce{H2}\slash\ce{CO}, GRI-Mech \num{3.0}) have lower SIMD-efficiency, \numrange{0.6}{0.66}.
 This again demonstrates that the SIMD-vectorization becomes efficient as the amount of work to perform increases (with increasing model size).
-Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one; this is likely caused by a combination of use of an OpenCL vector-width greater than the native CPU vector-width (i.e., \num{8} v.s., \num{2}), and the improved data-locality for the vectorized-data ordering, as discussed in~\cref{S:data}, and results a modest \textasciitilde\SIrange{10}{25}{\percent} improvement over the nominal vector width.
+Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one; this is likely caused by a combination of use of an OpenCL vector-width greater than the native CPU vector-width (i.e., \num{8} v.s., \num{2}), and the improved data-locality for the vectorized-data ordering, as discussed in~\cref{S:data}, and results a modest \SIrange{7}{17}{\percent} improvement over the nominal vector width.
 
@@
 As a result, the shallow-vectorized sparse OpenCL code is faster than the dense OpenMP code in all cases on the \avx/ machine (\cref{F:sparse_vs_dense}).
-\Cref{F:sparse_vs_dense_speedup} shows the speedup of the sparse\slash dense shallow-vectorized OpenCL the over the same sparse\slash dense Jacobian format on OpenMP; the dense OpenCL is \SIrange{3.7}{4.2}{$\times$} faster than the corresponding dense OpenMP code, this speedup increases to \SIrange{6.6}{8.4}{$\times$} for the sparse Jacobian.
+\Cref{F:sparse_vs_dense_speedup} shows the speedup of the sparse\slash dense shallow-vectorized OpenCL the over the same sparse\slash dense Jacobian format on OpenMP; the dense OpenCL is \SIrange{3.7}{4.2}{$\times$} faster than the corresponding dense OpenMP code, this speedup increases to \SIrange{6.6}{9.1}{$\times$} for the sparse Jacobian.
 On the \sse/ machine,~\cref{F:sparse_vs_dense_sse} shows similar results; the sparse OpenMP code is the slowest in all cases, and the shallow-vectorized OpenCL code is nearly as fast as the dense OpenMP code.
-Once again, the sparse OpenCL code is far less negatively impacted by the indirect lookup, and is only \SIrange{1.8}{3.2}{$\times$} slower than its dense counterpart, while the sparse OpenMP code is significantly (\SIrange{2.25}{6.2}{$\times$}) slower than the dense version.
-In~\cref{F:sparse_vs_dense_sse_speedup}, the speedup of the sparse and dense shallow-vectorized OpenCL codes are compared to their OpenMP version; the dense OpenCL code is \SIrange{3.8}{4.2}{$\times$} faster, while the sparse shallow-vectorization achieves a speedup of \SIrange{6.6}{9.1}{$\times$}.
+Once again, the sparse OpenCL code is far less negatively impacted by the indirect lookup, and is only \SIrange{1.8}{3.2}{$\times$} slower than its dense counterpart, while the sparse OpenMP code is significantly (\SIrange{2.25}{6.1}{$\times$}) slower than the dense version.
+In~\cref{F:sparse_vs_dense_sse_speedup}, the speedup of the sparse and dense shallow-vectorized OpenCL codes are compared to their OpenMP version; the dense OpenCL code is \SIrange{2}{2.5}{$\times$} faster, while the sparse shallow-vectorization achieves a speedup of \SIrange{3.1}{4.3}{$\times$}.
 
@@
 
-In~\cref{F:sparse_jac_scaling}, the parallel scaling efficiency of the sparse shallow-vectorized OpenCL is compared to the sparse OpenMP code.
+In~\cref{F:sparse_jac_scaling}, the strong parallel scaling efficiency of the sparse shallow-vectorized OpenCL is compared to the sparse OpenMP code.
 The plot is a bit difficult to read, as most of the data-points are clustered together however the key takeaway is that the parallel scaling efficiency of the shallow-vectorized OpenCL code is much closer in value to that of the OpenMP code as compared to the source rate evaluations (\cref{F:source_parallel_scaling}).
@@
 It is worth noting that the OpenCL code's scaling is more negatively affected by the switch from a single 12-core CPU (\cref{t:cpus}) to two CPUs (i.e., the 16 core data-point) than the OpenMP code.
-In contrast,~\cref{F:dense_jac_scaling} shows scaling for the shallow-vectorized dense Jacobian OpenCL code that is higher than for source-rate evaluation, but not as good as for sparse Jacobian evaluation.
-In this case, the USC-Mech II model\todo{note: don't have data for ipentanol yet} has the worst scaling for both codes; ranging from \numrange{0.96}{0.63} and \numrange{0.91}{0.34} for OpenMP and OpenCL respectively.
+\Cref{F:dense_jac_scaling} shows scaling for the shallow-vectorized dense Jacobian OpenCL code.
+In this case, the isopentanol model\todo{note: don't have data for openmp yet} has the worst scaling for both codes, we note that due to the sheer size of the dense isopentanol Jacobian---the Jacobian for a single thermo-chemical state takes over \SI{1}{\mega\byte} of data---the total number of states for the dense isopentanol Jacobian was limited to \num{50000} (which takes over \SI{50}{\giga\byte} of memory); this limiting of the number of initial states greatly drops the computation cost, and adversely affects the scaling efficiency as has been discussed previously.
+Excluding the isopentanol model, the dense shallow-vectorized OpenCL scalings are similar (if slightly higher) to that of the source rate evaluation, e.g., \numrange{0.7}{0.77} and \numrange{0.52}{0.61} for \num{4} and \num{8} cores respectively.
 As in~\cref{S:source_results}, we note that the shallow-vectorized OpenCL code on a single core of the \avx/ machine is roughly as fast than the OpenMP code on four cores of the same; hence, the OpenMP code has more room to improve performance before parallel-overhead and memory-bandwidth issues become a limiting concern.
@@
 
-The GPU performance
+The mean runtime per-condition for the sparse and dense Jacobian evaluation on the \gpunew/ GPU is plotted in~\cref{F:gpu_sparse_vs_dense}.
+As in the species evaluation case, the mean runtime per-condition drops steadily, for both cases becoming constant after just \num{3e3} states (except for \ce{H2}\slash\ce{CO}, which levels off near \num{e4} conditions).
+\Cref{F:gpu_jacobian_speedup} presents the speedup of the \gpunew/ over the \gpuold/ GPU for sparse\slash dense Jacobian evaluation; sparse evaluation on the \gpunew/ is \SIrange{1.22}{1.63}{$\times$} faster, while dense evaluation is \SIrange{1.62}{2.1}{$\times$} faster.
+The speedup on the \gpunew/ appears to decrease with increasing mechanism size for the sparse evalution, however we note that the \gpuold/ GPU has high variability in runtimes for the larger models.\todo{missing srv2 data for ipentanol, and need to rerun srv2 data}.
+
+\begin{figure}[htbp]
+   \centering
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{gpu_sparse_vs_dense.pdf}
+      \caption{Mean runtime per-condition of sparse and dense Jacobian evaluation on the \gpunew/ GPU.}
+      \label{F:gpu_sparse_vs_dense}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{gpu_jacobian_speedup.pdf}
+      \caption{Speedup of the \gpunew/ over \gpuold/ GPUs for sparse and dence Jacobian evaluation.}
+      \label{F:gpu_jacobian_speedup}
+  \end{subfigure}
+  \caption{The performance of sparse\slash dense Jacobian evaluation on the \gpunew/ and \gpuold/ GPUs.}
+\end{figure}
 

### 2018-02-23 scripts/jac_error.py

@@ -40,7 +40,7 @@ def printer(err_dict):
     mean_thr = mean_thr[0] / mean_thr[1]
 
     print('\n'.join(
-        ['{}: {:.15e}'.format(k, v) for k, v in err_dict.items()
+        ['{}: {:.3e}'.format(k, v) for k, v in sorted(err_dict.items())
          if 'thresholded' in k]))
     print('mean_threshold: {:.15e}'.format(mean_thr))
 

### 2018-02-23 paper.tex
add initial finite difference jacobian stuff

@@
 \Cref{F:gpu_jacobian_speedup} presents the speedup of the \gpunew/ over the \gpuold/ GPU for sparse\slash dense Jacobian evaluation; sparse evaluation on the \gpunew/ is \SIrange{1.22}{1.63}{$\times$} faster, while dense evaluation is \SIrange{1.62}{2.1}{$\times$} faster.
-The speedup on the \gpunew/ appears to decrease with increasing mechanism size for the sparse evalution, however we note that the \gpuold/ GPU has high variability in runtimes for the larger models.\todo{missing srv2 data for ipentanol, and need to rerun srv2 data}.
+The speedup on the \gpunew/ appears to decrease with increasing mechanism size for the sparse evalution, however we note that the \gpuold/ GPU has high variability in runtimes for the larger models.\todo{missing srv2 data for ipentanol, and need to rerun srv2 data}
 
@@
   \caption{The performance of sparse\slash dense Jacobian evaluation on the \gpunew/ and \gpuold/ GPUs.}
+\end{figure}
+
+In~\cref{F:fd_vs_analytical}, the performance of the sparse analytical-Jacobian is compared to a sparse first-order finite-difference Jacobian on both the \avx/ CPU and the \gpuold/ \gpunew/ GPUs.\todo{missing data}
+\Cref{F:fd_vs_analytical_cpu} shows large speedups for both the OpenMP code and shallow-vectorized OpenCL code; the analytical OpenMP Jacobian is \SIrange{3.92}{-11111}{$\times$} faster, while the analytical OpenCL Jacobian achives speedups of \SIrange{17}{55}{$\times$}.
+It is noted that the current finite-difference formulation breaks Intel's auto-vectorizer, hence the OpenCL speedup comparison is made against the unvectorized OpenCL code; it is not a high priority to fix this issue, as the finite-difference Jacobian is implemented for comparison purposes only.
+Although we do not display the dense finite-difference Jacobian speedup in~\cref{F:fd_vs_analytical_cpu}, the dense OpenCL and OpenMP analytical Jacobian codes out-perform the finite-difference variants by even larger margins; \SIrange{24}{-155}{$\times$} for OpenCL, and \SIrange{10}{-11111}{$\times$} for OpenMP.
+A comparison of the sparse analytical and finite-difference Jacobians is presented in~\cref{F:fd_vs_analytical_gpu}.
+The analytical Jacobian on the \gpunew/ shows large speedups, increasing with chemical model size, in the range of \SIrange{5}{16}{$\times$}; however, the \gpuold/ GPU shows almost no speedup, it is not clear why this is the case.\todo{reaaaally need to re-run this data}
+Although not pictured, the dense analytical-Jacobian on the \gpunew/ GPU has larger speedups compared with the dense finite-difference Jacobian, in the range of \SIrange{5}{42}{$\times$}.
+
+\begin{figure}[htbp]
+   \centering
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{finite_difference_vs_analytical.pdf}
+      \caption{Mean runtime per-condition of sparse and dense Jacobian evaluation on the \gpunew/ GPU.}
+      \label{F:fd_vs_analytical_cpu}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{finite_difference_vs_analytical_gpu.pdf}
+      \caption{Speedup of the \gpunew/ over \gpuold/ GPUs for sparse and dence Jacobian evaluation.}
+      \label{F:fd_vs_analytical_gpu}
+  \end{subfigure}
+  \caption{The performance of a sparse, first-order forward finite-difference Jacobian compared to the analytical Jacobian on the \avx/ CPU and both GPUs.}
+  \label{F:fd_vs_analytical}
 \end{figure}

### 2018-02-23 scripts/plot_styles.py
start trying to incorporate old pyjac data

@@ -95,7 +95,10 @@ def pretty_names(pname, short=False):
                                                  'haswell': r'\texttt{avx2}',
                                                  'gpu': r'\texttt{K40m}'}),
                   'rtype': dummy_formatter({'jac': 'Analytical',
-                                            'fdjac': 'Finite Difference'})
+                                            'fdjac': 'Finite Difference'}),
+                  'platform': dummy_formatter({'srv2': r'\texttt{pyJac}-v2',
+                                               'srv2-gpu': r'\texttt{pyJac}-v2',
+                                               'v1': r'\texttt{pyJac}-v1'})
                   }
     if pname in pname_dict:
         return pname_dict[pname]

### 2018-02-23 scripts/scaling_plotter.py

@@ -1,9 +1,0 @@
-import perf_plotter as pp
-
-pp.plotter('../SIMD_scaling.pdf', show=False,
-    order='C', vectype='w', vecwidth='16', rates='hybrid', kernel='split',
-    minx=0)
-
-pp.plotter('../SIMT_scaling.pdf', show=False,
-    order='C', vectype='par', vecwidth='16', rates='hybrid', kernel='split',
-    minx=0)

### 2018-02-23 scripts/simd_and_order_comp.py

@@ -1,60 +1,0 @@
-import perf_plotter as pp
-import plot_styles as ps
-import matplotlib.patches as mpatches
-import matplotlib.lines as mlines
-
-SIMD_color = ps.color_wheel[0]
-SIMT_color = ps.color_wheel[1]
-SIMT_sym = ps.marker_wheel[0][0]
-SIMD_sym = ps.marker_wheel[1][0]
-C_hollow = False
-F_hollow = True
-
-def marker_func(name):
-    retval = []
-    if 'SIMT' in name:
-        retval.append(SIMT_sym)
-    else:
-        retval.append(SIMD_sym)
-    if 'C-order' in name:
-        retval.append(C_hollow)
-    else:
-        retval.append(F_hollow)
-    if 'SIMT' in name:
-        retval.append(SIMT_color)
-    else:
-        retval.append(SIMD_color)
-    return tuple(retval)
-
-labels = ['SIMT', 'SIMD', 'C-order (Filled)', 'F-order (Hollow)']
-handles = [mlines.Line2D([], [], color=SIMT_color, marker=SIMT_sym, linestyle='',
-                          markersize=ps.marker_style['size'], label='SIMT'),
-           mlines.Line2D([], [], color=SIMD_color, marker=SIMD_sym, linestyle='',
-                          markersize=ps.marker_style['size'], label='SIMD'),
-           mlines.Line2D([], [], linestyle='', marker=None, label='C-order (Filled)'),
-           mlines.Line2D([], [], linestyle='', marker=None, label='F-order (Hollow)')]
-legend = (handles, labels)
-
-xval, yval, zval, label = pp.plotter('../SIMD_SIMT_Order_comparison.pdf', show=False,
-    cores='1', vecwidth='8', rates='hybrid', kernel='split',
-    legend_handler=legend, marker_func=marker_func, return_vals=True,
-    norm=False, ylog=True)
-
-#print stats
-C_order = [i for i,l in enumerate(label) if 'C-order' in l]
-F_order = [i for i in range(len(label)) if i not in C_order]
-
-C_SIMT = next(i for i in C_order if 'SIMT' in label[i])
-C_SIMD = next(i for i in C_order if i != C_SIMT)
-F_SIMT = next(i for i in F_order if 'SIMT' in label[i])
[…]

### 2018-02-23 scripts/vecwidth_plotter.py

@@ -1,4 +1,0 @@
-import perf_plotter as pp
-
-pp.plotter('../SIMD_Vecwidth_comparison.pdf', show=False,
-    cores='1', order='C', vectype='w', rates='hybrid', kernel='split')

### 2018-02-27 scripts/plot_styles.py
first pass at v1 / v2 comparison

@@ -93,12 +93,11 @@ def pretty_names(pname, short=False):
                   'descriptor': dummy_formatter({'srv2': r'\texttt{sse4.2}',
                                                  'srv2-gpu': r'\texttt{C2075}',
                                                  'haswell': r'\texttt{avx2}',
-                                                 'gpu': r'\texttt{K40m}'}),
+                                                 'gpu': r'\texttt{K40m}',
+                                                 'v1': r'\texttt{sse4.2-v1}',
+                                                 'v1-gpu': r'\texttt{C2075.2-v1}'}),
                   'rtype': dummy_formatter({'jac': 'Analytical',
-                                            'fdjac': 'Finite Difference'}),
-                  'platform': dummy_formatter({'srv2': r'\texttt{pyJac}-v2',
-                                               'srv2-gpu': r'\texttt{pyJac}-v2',
-                                               'v1': r'\texttt{pyJac}-v1'})
+                                            'fdjac': 'Finite Difference'})
                   }
     if pname in pname_dict:
         return pname_dict[pname]

### 2018-03-01 paper.tex
first pass at v1 / v2 comparison
update

@@
       \includegraphics[width=\textwidth]{finite_difference_vs_analytical.pdf}
-      \caption{Mean runtime per-condition of sparse and dense Jacobian evaluation on the \gpunew/ GPU.}
+      \caption{Speedup of the analytical versus finite-difference Jacobian evaluation on the \avx/ CPU, normalized per-language}
       \label{F:fd_vs_analytical_cpu}
@@
       \includegraphics[width=\textwidth]{finite_difference_vs_analytical_gpu.pdf}
-      \caption{Speedup of the \gpunew/ over \gpuold/ GPUs for sparse and dence Jacobian evaluation.}
+      \caption{Speedup of the analytical versus finite-difference Jacobian evaluation on both the \gpunew/ and \gpuold/ GPUs, normalized per-GPU}
       \label{F:fd_vs_analytical_gpu}
@@
   \label{F:fd_vs_analytical}
+\end{figure}
+
+Finally, in~\cref{F:v1_vs_v2} compares the performance for evaluation of the dense-analytical Jacobian of the new version of \texttt{pyJac} to the previous version~\cite{pyjac16} on the \sse/ CPU and \gpuold/ GPU; dense Jacobian evaluation was selected as it was the only type implemented in the previous verison.
+
+\begin{figure}[htbp]
+   \centering
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{v1_vs_v2.pdf}
+      \caption{Speedup of dense Jacobian evaluation in \texttt{pyJac-v2} over \texttt{pyJac-v1} on the \sse/ CPU}
+      \label{F:v1_vs_v2_cpu}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{v1_vs_v2_gpu.pdf}
+      \caption{Speedup of dense Jacobian evaluation in \texttt{pyJac-v2} over \texttt{pyJac-v1} on the \gpuold/ GPU}
+      \label{F:v1_vs_v2_gpu}
+  \end{subfigure}
+  \caption{Performance comparison of the new version of \texttt{pyJac} (\texttt{v2}) and the previous verison, \texttt{v1}~\cite{pyjac16}.}
+  \label{F:v1_vs_v2}
 \end{figure}

### 2018-03-02 derivations.tex
edits towards finalizing derivations

@@
 n &= \frac{V P}{T \Ru/} = \sum_{i=1}^{\ns/}{n_i}\; , \nonumber \\
-n_{\ns/} &= \frac{V P}{T \Ru/} - \sum_{i=1}^{\ns/ - 1}{n_i}
-\end{align}
+n_{\ns/} &= \frac{V P}{T \Ru/} - \sum_{i=1}^{\ns/ - 1}{n_i}\;,
+\end{align}
+where $\mathcal{R}$ is the ideal gas constant in molar units.
 
@@
 
-For both, the molar source terms are~\cite{TurnsStephenR2012Aitc}:
+For both \conp/ and \conv/, the molar source terms are~\cite{TurnsStephenR2012Aitc}:
 \loadeq{dmolar}
@@
 \begin{equation}
- [C]_{k} = \frac{n_{k}}{V}
+ [C]_{k} = \frac{n_{k}}{V}\;\text{.}
 \end{equation}
@@
 \begin{align}
-\frac{\text{d} V }{\text{d} t } &= \frac{\Ru/}{P} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right) & \text{\dconp/} \\
-\frac{\text{d} P }{\text{d} t } &= \frac{\Ru/}{V} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right) & \text{\dconv/}
+\frac{\text{d} V }{\text{d} t } &= \frac{\Ru/}{P} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right) & \text{\dconp/,} \\
+\frac{\text{d} P }{\text{d} t } &= \frac{\Ru/}{V} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right) & \text{\dconv/.}
 \end{align}
@@
 \begin{equation}
- \frac{\text{d} n }{\text{d} t }_{\ns/} = \frac{-1}{W_{\ns/}} \sum_{k=1}^{\ns/ - 1} W_{k} \frac{\text{d} n_{k} }{\text{d} t }
+ \frac{\text{d} n_{\ns/} }{\text{d} t } = - \sum_{k=1}^{\ns/ - 1} \frac{W_{k}}{W_{\ns/}} \frac{\text{d} n_{k} }{\text{d} t }
  \label{e:spec_ns}
@@
  \label{e:last_spec_conc}
-   [C]_{\ns/} = [C] - \sum_{k=1}^{\ns/  - 1} [C]_{k}
+   [C]_{\ns/} = [C] - \sum_{k=1}^{\ns/  - 1} [C]_{k}\;,
 \end{equation}
@@
 \begin{equation}
- [C] = \frac{P}{T \Ru/}
+ [C] = \frac{P}{T \Ru/}\;.
 \end{equation}
@@
 \begin{align}
-\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns/  - 1} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \dot{\omega}_{k}}{[C] {C_{p,\ns/}} + \sum_{k=1}^{\ns/  - 1} \left({C_{p, k}} - {C_{p,\ns/}}\right) [C]_{k}} &\text{\dconp/}\\
-\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns/  - 1} \left(U_{k} - \frac{W_{k} U_{\ns/}}{W_{\ns/}}\right) \dot{\omega}_{k}}{[C] {C_{v,\ns/}} + \sum_{k=1}^{\ns/  - 1} \left({C_{v, k}} - {C_{v,\ns/}}\right) [C]_{k}} &\text{\dconv/}
+\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns/  - 1} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \dot{\omega}_{k}}{[C] {C_{p,\ns/}} + \sum_{k=1}^{\ns/  - 1} \left({C_{p, k}} - {C_{p,\ns/}}\right) [C]_{k}} &\text{\dconp/,}\\
+\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns/  - 1} \left(U_{k} - \frac{W_{k} U_{\ns/}}{W_{\ns/}}\right) \dot{\omega}_{k}}{[C] {C_{v,\ns/}} + \sum_{k=1}^{\ns/  - 1} \left({C_{v, k}} - {C_{v,\ns/}}\right) [C]_{k}} &\text{\dconv/.}
 \end{align}
@@
 \begin{align}
-\frac{C_{p,k}^{\circ}}{\mathcal{R}} &= a_{0,k} + T \left( a_{1,k} + T \left( a_{2,k} + T \left( a_{3,k} + a_{4,k} T \right) \right) \right) \label{e:cpk} \\
-\frac{H_k^{\circ}}{\mathcal{R}} &= T \left( a_{0,k} + T \left( \frac{a_{1,k}}{2} + T \left( \frac{a_{2,k}}{3} + T \left( \frac{a_{3,k}}{4} + \frac{a_{4,k}}{5} T \right) \right) \right) \right) + a_{5,k} \label{e:hk} \\
-\frac{S_k^{\circ}}{\mathcal{R}} &= a_{0,k} \ln T + T \left( a_{1,k} + T \left( \frac{a_{2,k}}{2} + T \left( \frac{a_{3,k}}{3} + \frac{a_{4,k}}{4} T \right) \right) \right) + a_{6,k} \label{e:sk}
+\frac{C_{p,k}^{\circ}}{\mathcal{R}} &= a_{0,k} + T \left( a_{1,k} + T \left( a_{2,k} + T \left( a_{3,k} + a_{4,k} T \right) \right) \right) \;, \label{e:cpk} \\
+\frac{H_k^{\circ}}{\mathcal{R}} &= T \left( a_{0,k} + T \left( \frac{a_{1,k}}{2} + T \left( \frac{a_{2,k}}{3} + T \left( \frac{a_{3,k}}{4} + \frac{a_{4,k}}{5} T \right) \right) \right) \right) + a_{5,k} \;, \label{e:hk} \\
+\frac{S_k^{\circ}}{\mathcal{R}} &= a_{0,k} \ln T + T \left( a_{1,k} + T \left( \frac{a_{2,k}}{2} + T \left( \frac{a_{3,k}}{3} + \frac{a_{4,k}}{4} T \right) \right) \right) + a_{6,k} \;, \label{e:sk}
 \end{align}
@@
  \label{e:spec_rop}
- \dot{\omega}_{k} = \sum_{i=1}^{\nr//} \nu_{k,i} q_{i}
-\end{equation}
-where $\nr//$ is the number of reactions in the chemical kinetic model, $\nu_{k, i}$ the overall stoichiometric coefficient of species $k$ in reaction $i$ and $q_i$ the net rate-of-progress for reaction $i$:
-\begin{align}
-\nu_{k,i} &= \nu^{\prime\prime}_{k,i} - \nu^{\prime}_{k,i} \\
-q_{i} &= R_{i} c_{i}
-\end{align}
-with $\nu^{\prime}_{k,i}$ and $\nu^{\prime\prime}_{k,i}$ the product and reactant stoichiometric cofficients (respectively) of species $k$ in reaction $i$.
+ \dot{\omega}_{k} = \sum_{i=1}^{\nr/} \nu_{k,i} q_{i}\;,
+\end{equation}
+where $\nr/$ is the number of reactions in the chemical kinetic model, $\nu_{k, i}$ the overall stoichiometric coefficient of species $k$ in reaction $i$ and $q_i$ the net rate-of-progress for reaction $i$:
+\begin{align}
+\nu_{k,i} &= \nu^{\prime\prime}_{k,i} - \nu^{\prime}_{k,i} \;, \\
+q_{i} &= R_{i} c_{i} \;,
+\end{align}
+with $\nu^{\prime\prime}_{k,i}$ and $\nu^{\prime}_{k,i}$ the product and reactant stoichiometric cofficients (respectively) of species $k$ in reaction $i$.
 The base rate-of-progress for the $i$th reversible reaction $R_{i}$ is given by:
 \begin{align}
-R_{i} &= {R_{f,i}} - {R_{r,i}} \label{e:ropnet}\\
-{R_{f, i}} &= {k_{f, i}} \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime}_{k,i}} \label{e:ropf}\\
-{R_{r, i}} &= {k_{r, i}} \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime\prime}_{k,i}} \label{e:ropr}
+R_{i} &= {R_{f,i}} - {R_{r,i}}\;, \label{e:ropnet}\\
+{R_{f, i}} &= {k_{f, i}} \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime}_{k,i}} \label{e:ropf}\;,\\
+{R_{r, i}} &= {k_{r, i}} \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime\prime}_{k,i}} \label{e:ropr}\;,
 \end{align}
@@
  {K_{c, i}} &= {K_{p,i}}\left(\frac{P_{atm}}{T \Ru/}\right)^{\sum_{k=1}^{\ns/} \nu_{k,i}} \; ,\text{ and} \label{e:kc}\\
- {K_{p,i}} &= \text{exp}\left(\frac{\Delta S^{\circ}_i}{\Ru/} - \frac{\Delta H^{\circ}_i}{\Ru/ T}\right) = \text{exp}\left(\sum_{k=1}^{\ns/}\nu_{ki}\left(\frac{S^{\circ}_k}{\Ru/} - \frac{H^{\circ}_k}{\Ru/ T}\right)\right) \label{e:kp}
+ {K_{p,i}} &= \text{exp}\left(\frac{\Delta S^{\circ}_i}{\Ru/} - \frac{\Delta H^{\circ}_i}{\Ru/ T}\right) = \text{exp}\left(\sum_{k=1}^{\ns/}\nu_{ki}\left(\frac{S^{\circ}_k}{\Ru/} - \frac{H^{\circ}_k}{\Ru/ T}\right)\right) \;, \label{e:kp}
 \end{align}
@@
  \label{e:kc_in_kp}
- {K_{c, i}} = \left(\left(\frac{P_{atm}}{\Ru/}\right)^{\sum_{k=1}^{\ns/} \nu_{k,i}}\right) \operatorname{exp}\left({\sum_{k=1}^{\ns/} \nu_{k,i} B_{k}}\right)
-\end{equation}
-where, $B_k$ is:
+ {K_{c, i}} = \left(\left(\frac{P_{atm}}{\Ru/}\right)^{\sum_{k=1}^{\ns/} \nu_{k,i}}\right) \operatorname{exp}\left({\sum_{k=1}^{\ns/} \nu_{k,i} B_{k}}\right)\;,
+\end{equation}
+where from~\cref{e:hk,e:sk}, $B_k$ is:
 \begin{align}
@@
  B_{k} &= T \left(T \left(T \left(\frac{T a_{k,4}}{20} + \frac{a_{k,3}}{12}\right) + \frac{a_{k,2}}{6}\right) + \frac{a_{k,1}}{2}\right) \nonumber \\
-       & \quad + \left(a_{k,0} - 1\right) \log{\left (T \right )} - a_{k,0} + a_{k,6} - \frac{a_{k,5}}{T}
-\end{align}
-from~\cref{e:hk,e:sk}.
+       & \quad + \left(a_{k,0} - 1\right) \log{\left (T \right )} - a_{k,0} + a_{k,6} - \frac{a_{k,5}}{T} \;.
+\end{align}
 
@@
 
-For a reaction enhanced (or diminished) by the presence of a third body, the reaction rate is modified by the third-body concentration $[X]_i$ given by
+For a reaction enhanced (or diminished) by the presence of a third body, the reaction rate is modified by the third-body concentration $[X]_i$ given by:
 \begin{equation}
@@
 
-If all species in the mixture contribute equally as third bodies, then $\alpha_{k,i} = 1$ for all species:
+If all species in the mixture contribute equally as third bodies, i.e., $\alpha_{k,i} = 1$ for all species, then:
 \begin{equation}
@@
 \begin{equation}
- [X]_{i} = [C_m]
+ [X]_{i} = [C]_m\;,
 \end{equation}
@@
 \label{e:thd_spec}
- [X]_{i}=\left([C] - \sum_{k=1}^{\ns/  - 1} [C]_{k}\right) \delta_{\ns/ m} + \left(- \delta_{\ns/ m} + 1\right) [C]_{m}
-\end{equation}
-will be used, where the Kronecker delta $\delta_{\ns/ m}$ is unity if and only if the third body species $m$ is the last species in the model.
+ [X]_{i}=\left([C] - \sum_{k=1}^{\ns/  - 1} [C]_{k}\right) \delta_{\ns/, m} + \left(- \delta_{\ns/, m} + 1\right) [C]_{m}
+\end{equation}
+will be used, where the Kronecker delta $\delta_{\ns/, m}$ is unity if and only if the third body species $m$ is the last species in the model.
 \Cref{e:thd_spec} will be referred to as the \textbf{species}-based third-body efficiency.
@@
  A_{Troe} &= - 0.67 \log_{10}{\left (F_{cent} \right )} + \log_{10}{\left (P_{r, i} \right )} - 0.4 \;\text{, and}\\
- B_{Troe} &= - 1.1762 \log_{10}{\left (F_{cent} \right )} - 0.14 \log_{10}{\left (P_{r, i} \right )} + 0.806
+ B_{Troe} &= - 1.1762 \log_{10}{\left (F_{cent} \right )} - 0.14 \log_{10}{\left (P_{r, i} \right )} + 0.806 \;,
 \end{align}
@@
 \begin{equation}
- X = \frac{1}{\log_{10}^{2}{\left (P_{r, i} \right )} + 1}
+ X = \frac{1}{\log_{10}^{2}{\left (P_{r, i} \right )} + 1}\;.
 \end{equation}
@@
 \begin{align}
-k_1 (T) &= A_1 T^{\beta_1} \exp \left( -\frac{T_{a, 1}}{T} \right) \text{ at } P_1 \text{ and} \label{e:plog_k1} \\
+k_1 (T) &= A_1 T^{\beta_1} \exp \left( -\frac{T_{a, 1}}{T} \right) \text{ at } P_1 \text{ and,} \label{e:plog_k1} \\
 k_2 (T) &= A_2 T^{\beta_2} \exp \left( -\frac{T_{a, 2}}{T} \right) \text{ at } P_2 \;, \label{e:plog_k2}
 \end{align}
-where the Arrhenius coefficients are given for each pressure $p_1$ and $p_2$.
-Then, the reaction rate coefficient at a particular pressure $p$ between $p_1$ and $p_2$ can be determined through logarithmic interpolation:
+where the Arrhenius coefficients are given for each pressure $P_1$ and $P_2$.
+Then, the reaction rate coefficient at a particular pressure $P$ between $P_1$ and $P_2$ can be determined through logarithmic interpolation:
 \begin{equation}
@@
 \label{e:kf_cheb}
-\log_{10} k_f (T, p) = \sum_{i = 1}^{N_T} \sum_{j = 1}^{N_p} \eta_{ij} \phi_i (\tilde{T}) \phi_j \left(\tilde{p}\right) \;,
+\log_{10} k_f (T, P) = \sum_{i = 1}^{N_T} \sum_{j = 1}^{N_p} \eta_{ij} \phi_i (\tilde{T}) \phi_j \left(\tilde{P}\right) \;,
 \end{equation}
@@
 \begin{align}
-\tilde{T} &\equiv \frac{2 T^{-1} - T^{-1}_{\min} - T^{-1}_{\max}}{T^{-1}_{\max} - T^{-1}_{\min}} \quad\text{and} \\
+\tilde{T} &\equiv \frac{2 T^{-1} - T^{-1}_{\min} - T^{-1}_{\max}}{T^{-1}_{\max} - T^{-1}_{\min}} \quad\text{and,} \\
 \tilde{P} &\equiv \frac{2\log_{10} P - \log_{10} P_{\min} - \log_{10} P_{\max}}{\log_{10} P_{\max} - \log_{10} P_{\min}} \;,
 \end{align}
-where $T_{\min} \leq T \leq T_{\max}$ and $p_{\min} \leq p \leq p_{\max}$ describe the ranges of validity for temperature and pressure.
+where $T_{\min} \leq T \leq T_{\max}$ and $P_{\min} \leq P \leq P_{\max}$ describe the ranges of validity for temperature and pressure.
 
@@
 We note that some parts of the Jacobian derivation---in particular derivatives of the energy equation---are quite complicated.
-The aim of this document is to provide the reader with an overview of the derivation process; some intermediate steps are left out due to brevity and typesetting constraints.
+The aim of this document is to provide the reader with an overview of the derivation process; some intermediate steps are left out due to length and typesetting constraints.
 For a complete derivation, the reader is directed to the output of the \href{https://github.com/arghdos/SPyJac-paper/blob/master/derivations/scripts/derivations.py}{derivations.py} script, from which this document was compiled.
@@
 \label{s:denergy_dt}
-First, we will derive the term:
-\begin{align}
- \mathcal{J}_{1,1} &= \frac{\partial}{\partial T} \frac{\text{d} T}{\text{d} t} \nonumber \\
+First, the temperature source term's derivative with respect to the temperature:
+\begin{align}
+ \mathcal{J}_{1,1} &= \frac{\partial}{\partial T} \frac{\text{d} T}{\text{d} t} \;, \nonumber \\
  \intertext{or more simply:}
  \label{e:dTdotdT_initial}
- \mathcal{J}_{1,1} &= \frac{\partial\dot{T}}{\partial{T}}
-\end{align}
-To do so, we will more closely follow the \conp/ energy equation derivative with respect to temperature as the process is largely identical between \conp/ and \conv/.
+ \mathcal{J}_{1,1} &= \frac{\partial\dot{T}}{\partial{T}} \;,
+\end{align}
+will be derived.
+As the derivation process for this Jacobian entry is very similar between \conp/ and \conv/ energy equations, we will focus on the Jacobian entry for the \conp/ temperature source term and give the result for \conv/ at the end of this section.
 
@@
  \begin{dcases}
-  \frac{\partial}{\partial T} \left(\frac{n_k}{V}\right) = 0 & \left\{k \ne \ns/\right\} \\
-  \frac{\partial}{\partial T} \left([C] - \sum_{k=1}^{\ns/ - 1}{\frac{n_k}{V}}\right) = -\frac{[C]}{T} & \left\{k = \ns/\right\}
+  \frac{\partial}{\partial T} \left(\frac{n_k}{V}\right) = 0 & \text{for } k \ne \ns/\;\text{ and,} \\
+  \frac{\partial}{\partial T} \left([C] - \sum_{k=1}^{\ns/ - 1}{\frac{n_k}{V}}\right) = -\frac{[C]}{T} & \text{for } k = \ns/\;,
  \end{dcases}
 \end{align}
-as the pressure and volume are either constants (for \conp/ and \conv/ respectively) or a state variable and a function of time only.
-
-Next, the derivative of full form of the energy source term~\cref{e:temperature_complete} with respect to temperature will be considered.
+as the pressure and volume are either constants (for \conp/ and \conv/ respectively) or a state variable that is a function of time only.
+
+Next, the derivative of full form of the temperature source term~\cref{e:temperature_complete} with respect to temperature will be considered.
 To make the result more managable, the denominator of~\cref{e:temperature_complete} is first collapsed back into the form of~\cref{e:temperature_incomplete}, yielding for \conp/:
 \begin{align}
- \frac{\partial\dot{T}}{\partial{T}} &= - \frac{\left( [C] \left( \frac{{C_p}_{\ns/}}{T} - \frac{\text{d} {C_p} }{\text{d} T }_{\ns/} \right) - \sum_{k=1}^{-1 + \ns/} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns/} + \frac{\text{d} C_{p, k} }{\text{d} T }\right) [C]_{k}\right)}{\left(\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}\right)^{2}} \times \nonumber \\
+ \frac{\partial\dot{T}}{\partial{T}} &= - \frac{\left( [C] \left( \frac{{C_{p, \ns/}}}{T} - \frac{\text{d} {C_{p, \ns/}} }{\text{d} T } \right) - \sum_{k=1}^{-1 + \ns/} \left(- \frac{\text{d} {C_{p, \ns/}} }{\text{d} T } + \frac{\text{d} C_{p, k} }{\text{d} T }\right) [C]_{k}\right)}{\left(\sum_{k=1}^{\ns/} [C]_{k} {C_{p, k}}\right)^{2}} \times \nonumber \\
 				     & \mathindent[largerindent] \sum_{k=1}^{-1 + \ns/} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \dot{\omega}_{k} - \nonumber \\
-				     & \mathindent \frac{\sum_{k=1}^{-1 + \ns/} \left(\left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \left(\frac{\text{d} H_{k} }{\text{d} T } - \frac{W_{k}}{W_{\ns/}} \frac{\text{d} H }{\text{d} T }_{\ns/}\right) \dot{\omega}_{k}\right)}{\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}}
-\end{align}
-Next, we may recognize that~\cref{e:temperature_incomplete} may be factored out of the first terms, yielding:
-\begin{align}
- \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \times \nonumber \\
-				     & \left( [C] \left( \frac{{C_p}_{\ns/}}{T} - \frac{\text{d} {C_p} }{\text{d} T }_{\ns/} \right) - \sum_{k=1}^{-1 + \ns/} \left(- \frac{\text{d} {C_p} }{\text{d} T }_{\ns/} + \frac{\text{d} {C_p}_{k} }{\text{d} T }\right) [C]_{k}\right) - \nonumber \\
-				     & \sum_{k=1}^{-1 + \ns/} \left(\left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \left(\frac{\text{d} H_{k} }{\text{d} T } - \frac{W_{k}}{W_{\ns/}} \frac{\text{d} H }{\text{d} T }_{\ns/}\right) \dot{\omega}_{k}\right) \Biggr]
-\end{align}
-Now, we invoke the identity:
+				     & \mathindent \frac{\sum_{k=1}^{-1 + \ns/} \left(\left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \left(\frac{\text{d} H_{k} }{\text{d} T } - \frac{W_{k}}{W_{\ns/}} \frac{\text{d} H_{\ns/} }{\text{d} T }\right) \dot{\omega}_{k}\right)}{\sum_{k=1}^{\ns/} [C]_{k} {C_{p, k}}} \;.
+\end{align}
+Next,~\cref{e:temperature_incomplete} may be factored out, yielding:
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_{p, k}}} \Biggl[ \frac{\text{d} T }{\text{d} t } \times \nonumber \\
+				     & \left( [C] \left( \frac{{C_{p, \ns/}}}{T} - \frac{\text{d} {C_{p, \ns/}} }{\text{d} T } \right) - \sum_{k=1}^{-1 + \ns/} \left(- \frac{\text{d} {C_{p, \ns/}} }{\text{d} T } + \frac{\text{d} {C_{p, k}} }{\text{d} T }\right) [C]_{k}\right) - \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns/} \left(\left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \left(\frac{\text{d} H_{k} }{\text{d} T } - \frac{W_{k}}{W_{\ns/}} \frac{\text{d} H_{\ns/} }{\text{d} T }\right) \dot{\omega}_{k}\right) \Biggr]\;.
+\end{align}
+Now, the identity:
 \begin{equation}
@@
 \end{equation}
-to simplify the specific heat terms:
-\begin{align}
- \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns/} \left(- \frac{\text{d} C_{p, k} }{\text{d} T } + \frac{{C_p}_{\ns/}}{T}\right) [C]_{k} + \nonumber \\
-				     & \sum_{k=1}^{-1 + \ns/} \biggl(\left(- H_{k} + \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \nonumber \\
-				     & \mathindent[larger] \left(- \frac{\text{d} H_{k} }{\text{d} T } + \frac{W_{k}}{W_{\ns/}} \frac{\text{d} H }{\text{d} T }_{\ns/}\right) \dot{\omega}_{k}\biggr)\Biggr]
-\end{align}
-and finally, we may substitute in the specific heat for the temperature derivative of enthalpy, giving:
+is invoked to simplify the specific heat terms:
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_{p, k}}} \vast[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns/} \left(- \frac{\text{d} C_{p, k} }{\text{d} T } + \frac{{C_{p, \ns/}}}{T}\right) [C]_{k} + \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns/} \Biggl(\left(- H_{k} + \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + \nonumber \\
+				     & \mathindent[largerst] \left(- \frac{\text{d} H_{k} }{\text{d} T } + \frac{W_{k}}{W_{\ns/}} \frac{\text{d} H_{\ns/} }{\text{d} T }\right) \dot{\omega}_{k}\Biggr)\vast]\;.
+\end{align}
+Finally, the derivative of species enthalpy with respect to temperature:
+\begin{equation}
+ C_{p,k} = \frac{\text{d}H_k}{\text{d}T}\;,
+\end{equation}
+is substituted in, giving:
 \begin{subequations}
@@
 \begin{align}
- \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns/} \left(- \frac{\text{d} C_{p, k} }{\text{d} T } + \frac{{C_p}_{\ns/}}{T}\right) [C]_{k} + & \nonumber \\
-				     & \sum_{k=1}^{-1 + \ns/} \biggl( \left(- H_{k} + \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + & \nonumber \\
-				     & \mathindent[larger] \left(- {C_p}_{k} + \frac{W_{k} {C_p}_{k} }{W_{\ns/}} \right) \dot{\omega}_{k}\biggr)\Biggr] & \text{\dconp/} \\
+ \frac{\partial\dot{T}}{\partial{T}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_{p, k}}} \vast[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns/} \left(- \frac{\text{d} C_{p, k} }{\text{d} T } + \frac{{C_{p, \ns/}}}{T}\right) [C]_{k} + & \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns/} \Biggl( \left(- H_{k} + \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + & \nonumber \\
+				     & \mathindent[largerst] \left(- {C_{p, k}} + \frac{W_{k} {C_{p, k}} }{W_{\ns/}} \right) \dot{\omega}_{k}\Biggr)\vast] & \text{\dconp/,} \\
 \intertext{and similarly:}
-\frac{\partial\dot{T}}{\partial{T}}  &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_v}_{k}} \Biggl[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns/} \left(- \frac{\text{d} C_{v, k} }{\text{d} T } + \frac{{C_v}_{\ns/}}{T}\right) [C]_{k} + & \nonumber \\
-				     & \sum_{k=1}^{-1 + \ns/} \biggl(\left(- U_{k} + \frac{W_{k} U_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + & \nonumber \\
-				     & \mathindent[larger] \left(- {C_v}_{k} + \frac{W_{k} {C_v}_{\ns/}}{W_{\ns/}}\right) \dot{\omega}_{k}\biggr)\Biggr] & \text{\dconv/}
+\frac{\partial\dot{T}}{\partial{T}}  &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_{v, k}}} \vast[ \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{\ns/} \left(- \frac{\text{d} C_{v, k} }{\text{d} T } + \frac{{C_{v, \ns/}}}{T}\right) [C]_{k} + & \nonumber \\
+				     & \sum_{k=1}^{-1 + \ns/} \Biggl(\left(- U_{k} + \frac{W_{k} U_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial T } + & \nonumber \\
+				     & \mathindent[largerst] \left(- {C_{v, k}} + \frac{W_{k} {C_{v, \ns/}}}{W_{\ns/}}\right) \dot{\omega}_{k}\Biggr)\vast] & \text{\dconv/.}
 \end{align}
@@
 \begin{align}
- \mathcal{J}_{1, 2} & = \frac{\partial\dot{T}}{\partial{V}} & \text{\dconp/} \\
- \mathcal{J}_{1, 2} & = \frac{\partial\dot{T}}{\partial{P}} & \text{\dconv/}
+ \mathcal{J}_{1, 2} & = \frac{\partial\dot{T}}{\partial{V}} & \text{\dconp/,} \\
+ \mathcal{J}_{1, 2} & = \frac{\partial\dot{T}}{\partial{P}} & \text{\dconv/.}
 \end{align}
@@
   \begin{dcases}
-  \rhs{dck}{-\frac{[C]_{k}}{V}} & \mathcond{dck}{k \ne \ns/} \\
-  \rhs{dck}{\frac{1}{V} \sum_{k=1}^{\ns/  - 1} [C]_{k}} & \mathcond{dck}{k = \ns/}
- \end{dcases}&\cond{dck}{\dconp/} \\
+  \rhs{dck}{-\frac{[C]_{k}}{V}} & \mathcond{dck}{\text{with }k \ne \ns/} \\
+  \rhs{dck}{\frac{1}{V} \sum_{k=1}^{\ns/  - 1} [C]_{k}} & \mathcond{dck}{\text{with }k = \ns/}
+ \end{dcases}&\cond{dck}{\dconp/,} \\
  \lhs{dck}{\frac{\partial [C]_{k} }{\partial P }} &=
  \begin{dcases}
-   \rhs{dck}{\mathindent[s]0} & \mathcond{dck}{k \ne \ns/} \\
-   \rhs{dck}{\frac{1}{T \Ru/}} & \mathcond{dck}{k = \ns/}
- \end{dcases}&\cond{dck}{\dconv/}
+   \rhs{dck}{\mathindent[s]0} & \mathcond{dck}{\text{with }k \ne \ns/} \\
+   \rhs{dck}{\frac{1}{T \Ru/}} & \mathcond{dck}{\text{with }k = \ns/}
+ \end{dcases}&\cond{dck}{\dconv/,}
  \end{align}
@@
 while the specific heat and enthalpy\slash internal-energy are independent of the state parameter for both \conp/ and \conv/.
-Following the same procedure as in~\cref{s:denergy_dt}, the result of~\cref{e:dTdotdE_initial} are simplified by collapsing the specific-heat\slash concentration sum, and substituting in the temperature source term, giving:
+Following the same procedure as in~\cref{s:denergy_dt}, the result of~\cref{e:dTdotdE_initial} is simplified by collapsing the specific-heat\slash concentration sum, and substituting in the temperature source term, giving:
 \begin{subequations}
@@
 \begin{align}
- \frac{\partial\dot{T}}{\partial{V}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}} \Biggl[ - \sum_{k=1}^{-1 + \ns/} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial V } + & \nonumber \\
-				     & \mathindent[largerindentand] \frac{1}{V} \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{-1 + \ns/} \left(- {C_p}_{\ns/} + {C_p}_{k}\right) [C]_{k}\Biggr] & \text{\dconp/} \\
- \frac{\partial\dot{T}}{\partial{P}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns/} \left(U_{k} - \frac{W_{k} U_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial P } - & \nonumber \\
-				     & \mathindent[largerindentand] \frac{{C_v}_{\ns/}}{T \Ru/} \frac{\text{d} T }{\text{d} t } \Biggr] & \text{\dconv/}
+ \frac{\partial\dot{T}}{\partial{V}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_{p, k}}} \Biggl[ \nonumber \\
+				     & \mathindent - \sum_{k=1}^{-1 + \ns/} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial V } + & \nonumber \\
+				     & \mathindent  \frac{1}{V} \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{-1 + \ns/} \left(- {C_{p, \ns/}} + {C_{p, k}}\right) [C]_{k}\Biggr] & \text{\dconp/,} \\
+ \frac{\partial\dot{T}}{\partial{P}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_{v, k}}} \Biggl[ \nonumber \\
+				     & \mathindent - \sum_{k=1}^{-1 + \ns/} \left(U_{k} - \frac{W_{k} U_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial P } - \frac{{C_{v, \ns/}}}{T \Ru/} \frac{\text{d} T }{\text{d} t } \Biggr] & \text{\dconv/.}
 \end{align}
@@
 \subsubsection{Molar derivative}
-Finally, the molar derivative of the temperature source term is:
+Finally, the derivative of the temperature source term with respect to the moles of species $j$ is:
 \begin{align}
@@
 
-First, the derivative of a species concentration with respect to the moles of species $j$ is computed:
-\begin{align}
- \frac{\partial [C_k]}{\partial n_j} &=
+First, the molar derivative of a species concentration is computed as:
+\begin{align}
+ \frac{\partial [C]_k}{\partial n_j} &=
  \begin{dcases}
- \frac{\delta_{j k}}{V} & k \ne \ns/ \\
- -\frac{1}{V} & k = \ns/
+ \frac{\delta_{j k}}{V} & \text{for }k \ne \ns/ \\
+ -\frac{1}{V} & \text{for }k = \ns/
  \end{dcases} \\
@@
 \label{e:dck_dnj}
-\frac{\partial [C_k]}{\partial n_j} &= \frac{\delta_{j k} - \delta_{\ns/ k}}{V}
-\end{align}
-
-Similar to~\cref{s:denergy_dt} the derivative of the energy equation with respect to the moles of species $j$---after collapsing the specific-heat\slash~concentration sum---are given by:
-\begin{align}
- \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{\left(\sum_{k=1}^{-1 + \ns/} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \dot{\omega}_{k}\right) \sum_{k=1}^{-1 + \ns/} - \frac{\delta_{j k}}{V} \left({C_p}_{\ns/} - {C_p}_{k}\right)}{\left(\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}\right)^{2}} - \nonumber \\
-				       & \mathindent \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}} \sum_{k=1}^{-1 + \ns/} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} }
-\end{align}
-Next, the Kronecker delta summation is simplified, and the compact form of the energy equation factored out yielding:
+\frac{\partial [C]_k}{\partial n_j} &= \frac{\delta_{j k} - \delta_{\ns/ k}}{V}
+\end{align}
+
+The derivative of the energy equation with respect to the moles of species $j$---after collapsing the specific-heat\slash~concentration sum, as in Similar to~\cref{s:denergy_dt}---is given by:
+\begin{align}
+ \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{\left(\sum_{k=1}^{-1 + \ns/} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \dot{\omega}_{k}\right) \sum_{k=1}^{-1 + \ns/} - \frac{\delta_{j k}}{V} \left({C_{p, \ns/}} - {C_{p, k}}\right)}{\left(\sum_{k=1}^{\ns/} [C]_{k} {C_{p, k}}\right)^{2}} \nonumber \\
+				       & \mathindent - \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_{p, k}}} \sum_{k=1}^{-1 + \ns/} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} }\;.
+\end{align}
+Next, the Kronecker delta summation is simplified, and the compact form of the energy equation (\cref{e:temperature_incomplete}) factored out yielding:
 \begin{subequations}
@@
 \begin{align}
- \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_p}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns/} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \nonumber \\
-				       &  \mathindent[largerindentand] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_p}_{\ns/} + {C_p}_{j}\right)\Biggr] & \text{\dconp/} \\
+ \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_{p, k}}} \Biggl[- \sum_{k=1}^{-1 + \ns/} \left(H_{k} - \frac{W_{k} H_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \nonumber \\
+				       &  \mathindent[largerindentands] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_{p, \ns/}} + {C_{p, j}}\right)\Biggr] & \text{\dconp/,} \\
 \intertext{and:}
- \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns/} \left(U_{k} - \frac{W_{k} U_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \nonumber \\
-				       & \mathindent[largerindentand] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_v}_{\ns/} + {C_v}_{j}\right) \Biggr] & \text{\dconv/}
+ \frac{\partial\dot{T}}{\partial{n_j}} &= \frac{1}{\sum_{k=1}^{\ns/} [C]_{k} {C_{v, k}}} \Biggl[- \sum_{k=1}^{-1 + \ns/} \left(U_{k} - \frac{W_{k} U_{\ns/}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \nonumber \\
+				       & \mathindent[largerindentands] - \frac{1}{V} \frac{\text{d} T }{\text{d} t } \left(- {C_{v, \ns/}} + {C_{v, j}}\right) \Biggr] & \text{\dconv/.}
 \end{align}
@@
 		    &= \frac{V}{[C]} \sum_{k=1}^{-1 + \ns/} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \left(\frac{\partial \dot{\omega}_{k} }{\partial T } + \frac{\dot{\omega}_{k}}{T}\right) \nonumber \\
-		    &  \mathindent + \frac{V}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconp/}
+		    &  \mathindent + \frac{V}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconp/,}
 \intertext{and:}
@@
 		    &= \Ru/ \sum_{k=1}^{-1 + \ns/} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \left(T \frac{\partial \dot{\omega}_{k} }{\partial T } + \dot{\omega}_{k}\right) & \nonumber \\
-		    & \mathindent + \frac{P}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconv/}
+		    & \mathindent + \frac{P}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconv/.}
 \end{align}
@@
 \subsubsection{State parameter derivative}
-Similarly, the derivative of the thermodynamic state parameter source terms with respect to themselves are:
+Similarly, the derivatives of the thermodynamic state parameter source terms with respect to themselves are:
 \begin{subequations}
@@
 		    &= \frac{1}{[C]} \sum_{k=1}^{-1 + \ns/} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \left(V \frac{\partial \dot{\omega}_{k} }{\partial V } + \dot{\omega}_{k}\right) \nonumber \\
-		    & \mathindent + \frac{1}{T} \left(V \frac{\text{d} \dot{T} }{\text{d} V } + \dot{T}\right) & \text{\dconp/} \\
+		    & \mathindent + \frac{1}{T} \left(V \frac{\text{d} \dot{T} }{\text{d} V } + \dot{T}\right) & \text{\dconp/,} \\
  \intertext{and:}
@@
 		    &= T \Ru/ \sum_{k=1}^{-1 + \ns/} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial P } & \nonumber \\
-		    & \mathindent + \frac{1}{T} \left(P \frac{\text{d} \dot{T} }{\text{d} P } + \dot{T}\right) & \text{\dconv/}
+		    & \mathindent + \frac{1}{T} \left(P \frac{\text{d} \dot{T} }{\text{d} P } + \dot{T}\right) & \text{\dconv/.}
 \end{align}
@@
  \mathcal{J}_{2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{V} }{\partial n_j} & \nonumber \\
-		    &= V \left(\frac{1}{[C]} \sum_{k=1}^{-1 + \ns/} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } + \frac{1}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} }\right) & \text{\dconp/}
+		    &= V \left(\frac{1}{[C]} \sum_{k=1}^{-1 + \ns/} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } + \frac{1}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} }\right) & \text{\dconp/,}
 \intertext{and:}
  \mathcal{J}_{2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{P} }{\partial n_j} & \nonumber \\
-		    &=  \frac{P}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} } + T \Ru/ \sum_{k=1}^{-1 + \ns/} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \text{\dconv/}
+		    &=  \frac{P}{T} \frac{\text{d} \dot{T} }{\text{d} {n_j} } + T \Ru/ \sum_{k=1}^{-1 + \ns/} \left(1 - \frac{W_{k}}{W_{\ns/}}\right) \frac{\partial \dot{\omega}_{k} }{\partial {n_j} } & \text{\dconv/.}
 \end{align}
@@
  \label{e:dnkdotdT_compact}
- \mathcal{J}_{k + 2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial T } \nonumber \\
+ \mathcal{J}_{k + 2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial T }\;, \nonumber \\
  \intertext{or:}
- \mathcal{J}_{k + 2, 1} & = V \frac{\partial \dot{\omega}_{k} }{\partial T }
+ \mathcal{J}_{k + 2, 1} & = V \frac{\partial \dot{\omega}_{k} }{\partial T }\;.
 \end{align}
@@
   \mathcal{J}_{k + 2, 2} &= \frac{\partial }{\partial V } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial V } & \nonumber \\
-			 &= V \frac{\partial \dot{\omega}_{k} }{\partial V } + \dot{\omega}_{k} & \text{\dconp/}
+			 &= V \frac{\partial \dot{\omega}_{k} }{\partial V } + \dot{\omega}_{k} & \text{\dconp/,}
   \intertext{and:}
   \mathcal{J}_{k + 2, 2} &= \frac{\partial }{\partial P } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial P } & \nonumber \\
-			 &=V \frac{\partial \dot{\omega}_{k} }{\partial P } & \text{\dconv/}
+			 &=V \frac{\partial \dot{\omega}_{k} }{\partial P } & \text{\dconv/.}
 \end{align}
@@
 \begin{align}
- \mathcal{J}_{k + 2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial n_j } \nonumber \\
+ \mathcal{J}_{k + 2, j + 2} &= \frac{\partial }{\partial n_j } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial n_j } \;, \nonumber \\
   \intertext{or:}
  \label{e:dnkdotdj_nojac}
- \mathcal{J}_{k + 2, j + 2} & = V \frac{\partial \dot{\omega}_{k} }{\partial {n_j} }
+ \mathcal{J}_{k + 2, j + 2} & = V \frac{\partial \dot{\omega}_{k} }{\partial {n_j} }\;.
 \end{align}
@@
 \label{s:d_spec_prod}
-First, we evaluate the derivatives of the net species production rate equation~\cref{e:spec_rop}, with respect to temperature:
+The derivatives of the net species production rate equation~\cref{e:spec_rop}, with respect to temperature are:
 \begin{align}
  \label{e:dwdot_dT}
- \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial T }} &= \rhs{dwdot}{\sum_{i=1}^{\nr//} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial T } + \nu_{k,i} \frac{\partial R_{i}}{\partial T } c_{i}\right)} & \cond{dwdot}{,}
+ \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial T }} &= \rhs{dwdot}{\sum_{i=1}^{\nr/} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial T } + \nu_{k,i} \frac{\partial R_{i}}{\partial T } c_{i}\right)} & \cond{dwdot}{,}
 \end{align}
@@
  \begin{align}
-  \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial V }} &= \rhs{dwdot}{\sum_{i=1}^{\nr//} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial V } + \nu_{k,i} \frac{\partial R_{i}}{\partial V } c_{i}\right)} & \cond{dwdot}{\dconp/}\\
-  \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial P }} &= \rhs{dwdot}{\sum_{i=1}^{\nr//} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial P } + \nu_{k,i} \frac{\partial R_{i}}{\partial P } c_{i}\right)} & \cond{dwdot}{\dconv/}
+  \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial V }} &= \rhs{dwdot}{\sum_{i=1}^{\nr/} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial V } + \nu_{k,i} \frac{\partial R_{i}}{\partial V } c_{i}\right)} & \cond{dwdot}{\dconp/,}\\
+  \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial P }} &= \rhs{dwdot}{\sum_{i=1}^{\nr/} \left(\nu_{k,i} R_{i} \frac{\partial c_{i}}{\partial P } + \nu_{k,i} \frac{\partial R_{i}}{\partial P } c_{i}\right)} & \cond{dwdot}{\dconv/,}
  \end{align}
@@
  \label{e:dwdot_dnj}
- \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial {n_j} }} &= \rhs{dwdot}{\sum_{i=1}^{\nr//} \left(\nu_{k,i} R_{i} \frac{\partial c_{i} }{\partial {n_j} } + \nu_{k,i} \frac{\partial R_{i} }{\partial {n_j} } c_{i}\right)} & \cond{dwdot}{,}
+ \lhs{dwdot}{\frac{\partial \dot{\omega}_{k} }{\partial {n_j} }} &= \rhs{dwdot}{\sum_{i=1}^{\nr/} \left(\nu_{k,i} R_{i} \frac{\partial c_{i} }{\partial {n_j} } + \nu_{k,i} \frac{\partial R_{i} }{\partial {n_j} } c_{i}\right)} & \cond{dwdot}{.}
 \end{align}
@@
 \begin{equation}
- \frac{\partial R_{i}}{\partial T } = \frac{\partial}{\partial T}\left({k_{f, i}} \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime}_{k,i}}\right)
+ \frac{\partial R_{i}}{\partial T } = \frac{\partial}{\partial T}\left({k_{f, i}} \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime}_{k,i}}\right)\;.
 \end{equation}
@@
  \label{e:dkf_dt}
- \frac{\text{d} k_{f, i} }{\text{d} T } = \frac{{k_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)
+ \frac{\text{d} k_{f, i} }{\text{d} T } = \frac{{k_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)\;.
 \end{equation}
@@
 \begin{equation}
- {R_f} = \left(\left(- \sum_{k=1}^{\ns/  - 1} [C]_{k} + \frac{P}{T \Ru/}\right)^{\nu^{\prime}_{\ns/,i}}\right) {k_{f, i}} \prod_{k=1}^{\ns/  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
+ {R_f} = \left(\left(- \sum_{k=1}^{\ns/  - 1} [C]_{k} + \frac{P}{T \Ru/}\right)^{\nu^{\prime}_{\ns/,i}}\right) {k_{f, i}} \prod_{k=1}^{\ns/  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}\;.
 \end{equation}
@@
  \label{e:dropfdt_1}
- \frac{\partial {R_f} _{i}}{\partial T } = \frac{\text{d} k_{f, i} }{\text{d} T } \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns/,i}}{T} [C]_{\ns/}^{\nu^{\prime}_{\ns/,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns/  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
+ \frac{\partial {R_f} _{i}}{\partial T } = \frac{\text{d} k_{f, i} }{\text{d} T } \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns/,i}}{T} [C]_{\ns/}^{\nu^{\prime}_{\ns/,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns/  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}\;.
 \end{equation}
@@
  \label{e:s_temp}
- S^{\prime}_{l} = \nu^{\prime}_{l,i} [C]_{l}^{\nu^{\prime}_{l,i} - 1} \prod_{\substack{1 \leq l \leq l - 1\\l + 1 \leq l \leq \ns/}} [C]_{l}^{\nu^{\prime}_{l,i}}
+ S^{\prime}_{l} = \nu^{\prime}_{l,i} [C]_{l}^{\nu^{\prime}_{l,i} - 1} \prod_{\substack{1 \leq l \leq l - 1\\l + 1 \leq l \leq \ns/}} [C]_{l}^{\nu^{\prime}_{l,i}}\;,
 \end{equation}
@@
  \label{e:dropf_dt}
- \frac{\partial {R_f}_{i}}{\partial T } = - \frac{[C] S^{\prime}_{\ns/}}{T} {k_{f, i}} + \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)
-\end{equation}
-
-Starting from~\cref{e:ropr} and applying the same expansion of the last species' concentration as previously, the temperature derivative of the reverse rate of progress is:
+ \frac{\partial {R_f}_{i}}{\partial T } = - \frac{[C] S^{\prime}_{\ns/}}{T} {k_{f, i}} + \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)\;.
+\end{equation}
+
+Starting from~\cref{e:ropr} and applying the same expansion of the $\ns/$-th species' concentration as previously, the temperature derivative of the reverse rate of progress is:
 \begin{equation}
  \label{e:droprdt_1}
- \frac{\partial R_{r, i}}{\partial T } = \frac{\text{d} k_{r, i} }{\text{d} T } \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns/,i}}{T} [C]_{\ns/}^{\nu^{\prime\prime}_{\ns/,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns/  - 1} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
+ \frac{\partial R_{r, i}}{\partial T } = \frac{\text{d} k_{r, i} }{\text{d} T } \prod_{k=1}^{\ns/} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns/,i}}{T} [C]_{\ns/}^{\nu^{\prime\prime}_{\ns/,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns/  - 1} [C]_{k}^{\nu^{\prime\prime}_{k,i}}\;.
 \end{equation}
@@
 \begin{align}
- \frac{\text{d} k_{r, i} }{\text{d} T } &= \left(- \frac{1}{{K_{c, i}}} \frac{\text{d} {K_{c, i}} }{\text{d} T } \frac{{k_f}_i}{{K_c}_i} + \frac{1}{{K_{c, i}}}\frac{\text{d} {k_{f, i}} }{\text{d} T } \right) \nonumber \\
-				         &= \left(- \frac{1}{{K_{c, i}}} \frac{\text{d} {K_{c, i}} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)\right) {k_{r, i}}
+ \frac{\text{d} k_{r, i} }{\text{d} T } &= \left(- \frac{1}{{K_{c, i}}} \frac{\text{d} {K_{c, i}} }{\text{d} T } \frac{{k_f}_i}{{K_c}_i} + \frac{1}{{K_{c, i}}}\frac{\text{d} {k_{f, i}} }{\text{d} T } \right) \;, \nonumber \\
+				         &= \left(- \frac{1}{{K_{c, i}}} \frac{\text{d} {K_{c, i}} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)\right) {k_{r, i}}\;.
 \end{align}
@@
 \begin{equation}
- \frac{\text{d} B_{k} }{\text{d} T } = T \left(T \left(\frac{T a_{k,4}}{5} + \frac{a_{k,3}}{4}\right) + \frac{a_{k,2}}{3}\right) + \frac{a_{k,1}}{2} + \frac{1}{T} \left(a_{k,0} - 1 + \frac{a_{k,5}}{T}\right)
+ \frac{\text{d} B_{k} }{\text{d} T } = T \left(T \left(\frac{T a_{k,4}}{5} + \frac{a_{k,3}}{4}\right) + \frac{a_{k,2}}{3}\right) + \frac{a_{k,1}}{2} + \frac{1}{T} \left(a_{k,0} - 1 + \frac{a_{k,5}}{T}\right)\;,
 \end{equation}
@@
 \label{e:dkr_dt}
-\frac{\text{d} k_{r, i} }{\text{d} T } = \left(- \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k}}{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)\right) {k_{r, i}} \;,
-\end{equation}
-and using an temporary value $S^{\prime\prime}_{\ns/}$---defined analagously to \cref{e:s_temp} for the reverse direction---we obtain:
+\frac{\text{d} k_{r, i} }{\text{d} T } = \left(- \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k}}{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)\right) {k_{r, i}} \;.
+\end{equation}
+Using a temporary value $S^{\prime\prime}_{\ns/}$---defined analagously to \cref{e:s_temp} for the reverse reaction rate---the full deriviative of the reverse rate of progress with respect to temperature is obtained:
 \begin{equation}
 \label{e:dropr_dt}
-\frac{\partial R_{r, i}}{\partial T } = \left(- \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k}}{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)\right) {R_{r, i}} - \frac{[C] S^{\prime\prime}_{\ns/}}{T} {k_{r, i}}
+\frac{\partial R_{r, i}}{\partial T } = \left(- \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k}}{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)\right) {R_{r, i}} - \frac{[C] S^{\prime\prime}_{\ns/}}{T} {k_{r, i}}\;.
 \end{equation}
@@
  \frac{\partial R_{i}}{\partial T } =& \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right) - \left(- \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru/}\right)\right) {R_{r, i}} \nonumber \\
-				      &\qquad + \frac{[C] S^{\prime\prime}_{\ns/}}{T} {k_{r, i}} - \frac{[C] S^{\prime}_{\ns/}}{T} {k_{f, i}}
+				      &\qquad + \frac{[C] S^{\prime\prime}_{\ns/}}{T} {k_{r, i}} - \frac{[C] S^{\prime}_{\ns/}}{T} {k_{f, i}}\;.
 \end{align}
@@
 \label{s:dri_de}
-Following the outline of~\cref{s:dri_dnj}---while utilising~\cref{e:dck_de} for the species concentration derivatives---the forward rate of progress derivatives are:
-\begin{subequations}
- \begin{align}
-  \lhs{dri_de}{\frac{\partial {R_{f, i}} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime}_{\ns/}}{V} {k_{f, i}} - \frac{{R_{f, i}}}{V} \sum_{k=1}^{\ns/} \nu^{\prime}_{k,i}}&\cond{dri_de}{\dconp/}\\
-  \lhs{dri_de}{\frac{\partial {R_{f, i}} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime}_{\ns/} {k_{f, i}}}{T \Ru/}}&\cond{dri_de}{\dconv/}
- \end{align}
-\end{subequations}
-and similarly, the reverse rate of progress derivatives:
-\begin{subequations}
- \begin{align}
-  \lhs{dri_de}{\frac{\partial {R_{r, i}} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime\prime}_{\ns/}}{V} {k_{r, i}} - \frac{{R_{r, i}}}{V} \sum_{k=1}^{\ns/} \nu^{\prime\prime}_{k,i}}&\cond{dri_de}{\dconp/}\\
-  \lhs{dri_de}{\frac{\partial {R_{r, i}} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime\prime}_{\ns/} {k_{r, i}}}{T \Ru/}}&\cond{dri_de}{\dconv/}
- \end{align}
-\end{subequations}
-giving finally:
+Following the outline of~\cref{s:dri_dt}---while utilising~\cref{e:dck_de} for the species concentration derivatives---the forward rate of progress derivatives are:
+\begin{subequations}
+ \label{e:dropf_de}
+ \begin{align}
+  \lhs{dri_de}{\frac{\partial {R_{f, i}} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime}_{\ns/}}{V} {k_{f, i}} - \frac{{R_{f, i}}}{V} \sum_{k=1}^{\ns/} \nu^{\prime}_{k,i}}&\cond{dri_de}{\dconp/,}\\
+  \lhs{dri_de}{\frac{\partial {R_{f, i}} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime}_{\ns/} {k_{f, i}}}{T \Ru/}}&\cond{dri_de}{\dconv/.}
+ \end{align}
+\end{subequations}
+Similarly, the reverse rate of progress derivatives are:
+\begin{subequations}
+ \label{e:dropr_de}
+ \begin{align}
+  \lhs{dri_de}{\frac{\partial {R_{r, i}} }{\partial V }} &= \rhs{dri_de}{\frac{[C] S^{\prime\prime}_{\ns/}}{V} {k_{r, i}} - \frac{{R_{r, i}}}{V} \sum_{k=1}^{\ns/} \nu^{\prime\prime}_{k,i}}&\cond{dri_de}{\dconp/,}\\
+  \intertext{and:}
+  \lhs{dri_de}{\frac{\partial {R_{r, i}} }{\partial P }} &= \rhs{dri_de}{\frac{S^{\prime\prime}_{\ns/} {k_{r, i}}}{T \Ru/}}&\cond{dri_de}{\dconv/.}
+ \end{align}
+\end{subequations}
+Combining~\cref{e:dropf_de,e:dropr_de} gives:
 \begin{subequations}
@@
   \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial V }} &= \rhs{dri_de}{\frac{1}{V}\Biggl( [C]\left(S^{\prime}_{\ns/} {k_{f, i}} - S^{\prime\prime}_{\ns/} {k_{r, i}}\right) + } & \nonumber\\
-  & \rhs{dri_de}{\mathindent[farth] \left({R_{r, i}} \sum_{k=1}^{\ns/} \nu^{\prime\prime}_{k,i} - {R_{f, i}} \sum_{k=1}^{\ns/} \nu^{\prime}_{k,i}\right) \Biggr)}&\cond{dri_de}{\dconp/}\\
-  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial P }} &= \rhs{dri_de}{\frac{1}{T \Ru/} \left(S^{\prime}_{\ns/} {k_{f, i}} - S^{\prime\prime}_{\ns/} {k_{r, i}}\right)}&\cond{dri_de}{\dconv/}
+  & \rhs{dri_de}{\mathindent[farth] \left({R_{r, i}} \sum_{k=1}^{\ns/} \nu^{\prime\prime}_{k,i} - {R_{f, i}} \sum_{k=1}^{\ns/} \nu^{\prime}_{k,i}\right) \Biggr)}&\cond{dri_de}{\dconp/,}\\
+  \intertext{and:}
+  \lhs{dri_de}{\frac{\partial {R}_{i} }{\partial P }} &= \rhs{dri_de}{\frac{1}{T \Ru/} \left(S^{\prime}_{\ns/} {k_{f, i}} - S^{\prime\prime}_{\ns/} {k_{r, i}}\right)}&\cond{dri_de}{\dconv/.}
  \end{align}
@@
 
-Combining~\cref{e:dropf_dnj_1,e:dck_dnj} the molar derivative of the forward rate of progress is found as:
+Combining~\cref{e:dropf_dnj_1,e:dck_dnj} the molar derivative of the forward rate of progress is:
 \begin{equation}
@@
  \label{e:dropf_dnj}
- \frac{\partial {R_{f, i}} }{\partial {n_j} } = \frac{{k_{f, i}}}{V} \left(- S^{\prime}_{\ns/} + S^{\prime}_{j}\right)
-\end{equation}
-and by the same process, the reverse rate of progress derivative is found to be:
+ \frac{\partial {R_{f, i}} }{\partial {n_j} } = \frac{{k_{f, i}}}{V} \left(- S^{\prime}_{\ns/} + S^{\prime}_{j}\right)\;.
+\end{equation}
+By the same process, the reverse rate of progress derivative is calculated as:
 \begin{equation}
  \label{e:dropr_dnj}
- \frac{\partial {R_{r, i}} }{\partial {n_j} } = \frac{{k_{r, i}}}{V} \left(- S^{\prime\prime}_{\ns/} + S^{\prime\prime}_{j}\right)
-\end{equation}
-and the net rate of progress molar derivative:
+ \frac{\partial {R_{r, i}} }{\partial {n_j} } = \frac{{k_{r, i}}}{V} \left(- S^{\prime\prime}_{\ns/} + S^{\prime\prime}_{j}\right)\;,
+\end{equation}
+and from~\cref{e:dropf_dnj,e:dropr_dnj} the net rate of progress molar derivative is found:
 \begin{equation}
  \label{e:dropi_dnj}
- \frac{\partial R_{i} }{\partial {n_j} } = - \frac{{k_{r, i}}}{V} \left(- S^{\prime\prime}_{\ns/} + S^{\prime\prime}_{j}\right) + \frac{{k_{f, i}}}{V} \left(- S^{\prime}_{\ns/} + S^{\prime}_{j}\right)
+ \frac{\partial R_{i} }{\partial {n_j} } = - \frac{{k_{r, i}}}{V} \left(- S^{\prime\prime}_{\ns/} + S^{\prime\prime}_{j}\right) + \frac{{k_{f, i}}}{V} \left(- S^{\prime}_{\ns/} + S^{\prime}_{j}\right)\;.
 \end{equation}
@@
 \label{s:dri_pdep}
-The P-Log and Chebyshev pressure-dependent reactions, described by~\cref{e:kf_plog,e:kf_cheb} require separate treatment from the derivatives of strictly Arrhenius-based reaction rate formulations examined in~\cref{s:dri_dt,s:dri_dnj,s:dri_de}.
+The P-Log and Chebyshev pressure-dependent reactions, described by \cref{e:kf_plog,e:kf_cheb} require separate treatment from the derivatives of strictly Arrhenius-based reaction rate formulations examined in \cref{s:dri_dt,s:dri_dnj,s:dri_de}.
 Beginning with P-Log reactions, the derivative of the forward rate coefficient with respect to temperature is:
 \begin{align}
+ \label{e:dkf_plog_dt_initial}
  \frac{\partial {k_{f, i}} }{\partial T } &= \biggl( \frac{1}{k_{1}} \frac{\text{d} k_1 }{\text{d} T } + \frac{1}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} \left(- \frac{1}{k_{1}} \frac{\text{d} k_1 }{\text{d} T } + \frac{1}{k_{2}} \frac{\text{d} k_2 }{\text{d} T }\right) \times  \Bigl( \nonumber \\
-					 & \mathindent - \log{\left (P_{1} \right )} + \log{\left (P \right )}\Bigr)\biggr) {k_{f, i}}
-\end{align}
-while the derivatives of $k_{1}$ and $k_{2}$ are given by~\cref{e:dkf_dt}---with the corresponding Arrhenius parameters (see~\cref{e:plog_k1,e:plog_k2}) substituted in.
-Simplifying gives:
+					 & \mathindent - \log{\left (P_{1} \right )} + \log{\left (P \right )}\Bigr)\biggr) {k_{f, i}}\;,
+\end{align}
+and the derivatives of $k_{1}$ and $k_{2}$ are given by~\cref{e:dkf_dt} with the corresponding Arrhenius parameters (see \cref{e:plog_k1,e:plog_k2}) substituted in.
+
+Simplifying \cref{e:dkf_plog_dt_initial} gives:
 \begin{align}
  \label{e:dkf_plog_dt}
- \frac{\partial {k_{f, i}} }{\partial T } &= \frac{{k_{f, i}}}{T} \Biggl(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru/} + \frac{E_{a_2}}{T \Ru/}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \nonumber \\
-					 & \mathindent[morespac] \beta_1 + \frac{E_{a_1}}{T \Ru/}\Biggr)
+ \frac{\partial {k_{f, i}} }{\partial T } &= \frac{{k_{f, i}}}{T} \Biggl(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} \times \nonumber \\
+					 & \mathindent[morespac] \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru/} + \frac{E_{a_2}}{T \Ru/}\right) + \beta_1 + \frac{E_{a_1}}{T \Ru/}\Biggr)\;.
 \end{align}
@@
  \frac{\partial R_{i} }{\partial T } &= \left(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru/} + \frac{E_{a_2}}{T \Ru/}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \beta_1 + \frac{E_{a_1}}{T \Ru/} \right) \times \nonumber \\
-				     & \mathindent \left(\frac{{R_{f, i}} - {R_{r, i}}}{T}\right) + {R_{r, i}} \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns/} {k_{r, i}} - S^{\prime}_{\ns/} {k_{f, i}}\right)
-\end{align}
-
-For \conv/ problems, the P-Log forward rate coefficient is also a function of the pressure:
-\begin{equation}
- \frac{\partial {k_{f, i}} }{\partial P } = \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) {k_{f, i}}}{P \left(- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}\right)}
+				     & \mathindent[less] \left(\frac{{R_{f, i}} - {R_{r, i}}}{T}\right) + {R_{r, i}} \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns/} {k_{r, i}} - S^{\prime}_{\ns/} {k_{f, i}}\right)\;.
+\end{align}
+
+For \conv/ problems, the P-Log forward reaction rate coefficient is also a function of the pressure:
+\begin{equation}
+ \frac{\partial {k_{f, i}} }{\partial P } = \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) {k_{f, i}}}{P \left(- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}\right)}\;,
 \end{equation}
@@
  \frac{\partial R_{i} }{\partial P } &= \frac{1}{T \Ru/} \left(- S^{\prime\prime}_{\ns/} {k_{r, i}} + S^{\prime}_{\ns/} {k_{f, i}}\right) + & \nonumber \\
-				     &  \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) \left({R_{f, i}} - {R_{r, i}}\right)}{P \left(- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}\right)} & \text{\dconv/}
+				     &  \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) \left({R_{f, i}} - {R_{r, i}}\right)}{P \left(- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}\right)} & \text{\dconv/.}
 \end{align}
@@
 \begin{align}
- \frac{\partial {k_{f, i}} }{\partial T } &= \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{\text{d} \tilde{T} }{\text{d} T } \left(j - 1\right) \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}
+ \frac{\partial {k_{f, i}} }{\partial T } &= \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{\text{d} \tilde{T} }{\text{d} T } \left(j - 1\right) \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j} \;,
 \end{align}
@@
 \begin{equation}
- U_n \left(x\right) = \frac{\sin\left(\left(n + 1\right) \arccos x\right)}{\sin \left(\arccos x \right)}
+ U_n \left(x\right) = \frac{\sin\left(\left(n + 1\right) \arccos x\right)}{\sin \left(\arccos x \right)}\,
 \end{equation}
@@
 \begin{equation}
- \frac{\text{d} \tilde{T}}{\text{d} T} = \frac{-2}{T^2 \left(-\frac{1}{T_{\min}} + \frac{1}{T_{\text{max}}}\right)}
+ \frac{\text{d} \tilde{T}}{\text{d} T} = \frac{-2}{T^2 \left(-\frac{1}{T_{\min}} + \frac{1}{T_{\text{max}}}\right)}\;,
 \end{equation}
@@
 \begin{equation}
- \frac{\partial {k_{f, i}} }{\partial T } = \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} - \frac{2 \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right)
+ \frac{\partial {k_{f, i}} }{\partial T } = \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} - \frac{2 \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right)\;.
 \end{equation}
@@
  \frac{\partial R_{i}}{\partial T } &= \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \left[ - \frac{2 \log{\left (10 \right )} \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right) \right] \left({R_{f, i}} - {R_{r, i}}\right) + \nonumber \\
-				     & \mathindent \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } {R_{r, i}} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns/} {k_{r, i}} - S^{\prime}_{\ns/} {k_{f, i}}\right)
-\end{align}
-
-For \conv/ problems, we again must consider the derivative of the forward rate coefficient with respect to pressure:
-\begin{equation}
- \frac{\partial {k_{f, i}} }{\partial P } = \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{\text{d} \tilde{P} }{\text{d} P } \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}
+				     & \mathindent \sum_{k=1}^{\ns/} \nu_{k,i} \frac{\text{d} B_{k} }{\text{d} T } {R_{r, i}} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns/} {k_{r, i}} - S^{\prime}_{\ns/} {k_{f, i}}\right)\;.
+\end{align}
+
+For \conv/ problems, the derivative of the forward rate coefficient with respect to pressure must again be considered:
+\begin{equation}
+ \frac{\partial {k_{f, i}} }{\partial P } = \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{\text{d} \tilde{P} }{\text{d} P } \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}\;,
 \end{equation}
@@
 \begin{equation}
- \frac{\text{d} \tilde{P} }{\text{d} P } = \frac{2}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)}
+ \frac{\text{d} \tilde{P} }{\text{d} P } = \frac{2}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)}\;,
 \end{equation}
@@
 \begin{equation}
- \frac{\partial {k_{f, i}} }{\partial P } = \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)}
-\end{equation}
-and:
+ \frac{\partial {k_{f, i}} }{\partial P } = \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)}\;.
+\end{equation}
+Finally, the net rate of progress derivative for a Chebyshev reaction with respect to pressure is:
 \begin{align}
  \label{e:dropi_cheb_de}
- \frac{\partial R_{i} }{\partial P } &= \log{\left (10 \right )} \left({R_{f, i}} - {R_{r, i}}\right) \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)} + \nonumber \\
-				     & \mathindent \frac{1}{T \Ru/} \left(- S^{\prime\prime}_{\ns/} {k_{r, i}} + S^{\prime}_{\ns/} {k_{f, i}}\right)
+ \frac{\partial R_{i} }{\partial P } &= \log{\left (10 \right )} \left({R_{f, i}} - {R_{r, i}}\right) \times &\nonumber \\
+				     & \mathindent \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{2 \left(l - 1\right) \mathcal{T}_{j - 1}\left(\tilde{T}\right) U_{l - 2}\left(\tilde{P}\right) \eta_{l,j}}{P \left(\log{\left (P_{\max} \right )} - \log{\left (P_{\min} \right )}\right)} + & \nonumber \\
+				     & \mathindent \frac{1}{T \Ru/} \left(- S^{\prime\prime}_{\ns/} {k_{r, i}} + S^{\prime}_{\ns/} {k_{f, i}}\right) & \text{\dconv/.}
 \end{align}

### 2018-03-02 paper.tex
edits towards finalizing derivations

@@
 \end{equation}
-By differentiating the ideal gas law, the volume and pressure source terms can be found:
+By differentiating the ideal gas law:
+\begin{equation}
+ PV = n\mathcal{R}T \;,
+\end{equation}
+where $\mathcal{R}$ is the ideal gas-constant, the volume and pressure source terms can be found:
 \loadeq{dparam}
-where $\mathcal{R}$ is the ideal gas-constant.
 

### 2018-03-02 derivations.tex
hopefully final edits for derivations

@@
 \end{align}
-the state parameter:
+and with respect to the state parameter:
 \begin{subequations}
@@
  \begin{align}
-  \lhs{dci_elem}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{\dconp/} \\
-  \lhs{dci_elem}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{\dconv/}
- \end{align}
-\end{subequations}
-and the moles:
+  \lhs{dci_elem}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{\dconp/,} \\
+  \lhs{dci_elem}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_elem}{0} & \cond{dci_elem}{\dconv/,}
+ \end{align}
+\end{subequations}
+and finally, with respect to the moles of species $j$:
 \begin{align}
@@
 \label{s:dci_thd}
-For a third-body enhanced reaction~\cref{e:thd_mix} is differentiated with respect to temperature to give:
+For a \textbf{mixture} based third-body enhanced reaction~\cref{e:thd_mix} is differentiated with respect to temperature to give:
 \begin{align}
@@
  \begin{align}
- \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial V }} &= \rhs{dci_thd_mix}{\frac{1}{V} \left([C] \alpha_{\ns/,i} - [X]_{i}\right)} & \cond{dci_thd_mix}{\dconp/}\\
- \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial P }} &= \rhs{dci_thd_mix}{\frac{\alpha_{\ns/,i}}{T \Ru/}} & \cond{dci_thd_mix}{\dconv/}
+ \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial V }} &= \rhs{dci_thd_mix}{\frac{1}{V} \left([C] \alpha_{\ns/,i} - [X]_{i}\right)} & \cond{dci_thd_mix}{\dconp/,}\\
+ \lhs{dci_thd_mix}{\frac{\partial [X]_i }{\partial P }} &= \rhs{dci_thd_mix}{\frac{\alpha_{\ns/,i}}{T \Ru/}} & \cond{dci_thd_mix}{\dconv/,}
  \end{align}
@@
 
-If $\alpha_{j,i} = 1$ for all species $j$,~\cref{e:dci_thd_dt,e:dci_thd_dnj,e:dci_thd_de} may be simplified to:
+If $\alpha_{j,i} = 1$ for all species $j$ (i.e., a \textbf{unity} third-body), \cref{e:dci_thd_dt,e:dci_thd_dnj,e:dci_thd_de} may be simplified to:
 \begin{align}
@@
  \begin{align}
-  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_thd_unity}{0}[c] & \cond{dci_thd_unity}{\dconp/} \\
-  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_thd_unity}{\frac{1}{T \Ru/}}& \cond{dci_thd_unity}{\dconv/}
+  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_thd_unity}{0}[c] & \cond{dci_thd_unity}{\dconp/,} \\
+  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_thd_unity}{\frac{1}{T \Ru/}}& \cond{dci_thd_unity}{\dconv/,}
  \end{align}
@@
   \label{e:dci_thd_unity_dnj}
-  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial n_{j} }} &= \rhs{dci_thd_unity}{0}[c] & \cond{dci_thd_unity}{,}
-\end{align}
-
-If the third-body is the $m$th species,~\cref{e:dci_thd_dt,e:dci_thd_dnj,e:dci_thd_de} become:
+  \lhs{dci_thd_unity}{\frac{\partial c_{i} }{\partial n_{j} }} &= \rhs{dci_thd_unity}{0}[c] & \cond{dci_thd_unity}{.}
+\end{align}
+
+If the third-body is the $m$th \textbf{species},~\cref{e:dci_thd_dt,e:dci_thd_dnj,e:dci_thd_de} become:
 \begin{align}
@@
 \begin{align}
-\lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_thd_spec}{\frac{\delta_{\ns/ m}}{V} \left([C] - [C]_{\ns/}\right) + \frac{[C]_{m}}{V} \left(\delta_{\ns/ m} - 1\right)} & \cond{dci_thd_spec}{\dconp/} \\
-\lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_thd_spec}{\frac{\delta_{\ns/ m}}{T \Ru/}} & \cond{dci_thd_spec}{\dconv/}
+\lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial V }} &= \rhs{dci_thd_spec}{\frac{\delta_{\ns/ m}}{V} \left([C] - [C]_{\ns/}\right) + \frac{[C]_{m}}{V} \left(\delta_{\ns/ m} - 1\right)} & \cond{dci_thd_spec}{\dconp/,} \\
+\lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial P }} &= \rhs{dci_thd_spec}{\frac{\delta_{\ns/ m}}{T \Ru/}} & \cond{dci_thd_spec}{\dconv/,}
 \end{align}
@@
 \begin{align}
- \lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial n_j }} &= \rhs{dci_thd_spec}{\frac{1}{V} \left(- \delta_{\ns/ m} + \delta_{j m}\right)}&\cond{dci_thd_spec}{,} \label{e:dci_thd_spec_dnj}
+ \lhs{dci_thd_spec}{\frac{\partial c_{i} }{\partial n_j }} &= \rhs{dci_thd_spec}{\frac{1}{V} \left(- \delta_{\ns/ m} + \delta_{j m}\right)}&\cond{dci_thd_spec}{.} \label{e:dci_thd_spec_dnj}
 \end{align}
@@
  \begin{align}
-  \lhs{dci_fall}{\frac{\partial c_i }{\partial V }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial V } + \frac{\partial P_{r, i} }{\partial V } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{\dconp/} \\
-  \lhs{dci_fall}{\frac{\partial c_i }{\partial P }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial P } + \frac{\partial P_{r, i} }{\partial P } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{\dconv/}
+  \lhs{dci_fall}{\frac{\partial c_i }{\partial V }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial V } + \frac{\partial P_{r, i} }{\partial V } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{\dconp/,} \\
+  \lhs{dci_fall}{\frac{\partial c_i }{\partial P }} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial P } + \frac{\partial P_{r, i} }{\partial P } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{\dconv/,}
  \end{align}
@@
  \label{e:dci_fall_dnj}
- \lhs{dci_fall}{\frac{\partial c_{i} }{\partial n_j}} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial n[j] } + \frac{\partial P_{r, i} }{\partial n_j } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{,}
+ \lhs{dci_fall}{\frac{\partial c_{i} }{\partial n_j}} &= \rhs{dci_fall}{\frac{1}{P_{r, i} + 1} \left(P_{r, i} \frac{\partial F_{i} }{\partial n[j] } + \frac{\partial P_{r, i} }{\partial n_j } \left(F_{i} - c_{i}\right)\right)} & \cond{dci_fall}{.}
 \end{align}
@@
  \begin{align}
-  \lhs{dci_chem}{\frac{\partial c_i }{\partial V }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial V } - \frac{\partial P_{r, i} }{\partial V } c_{i}\right)} & \cond{dci_chem}{\dconp/} \\
-  \lhs{dci_chem}{\frac{\partial c_i }{\partial P }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial P } - \frac{\partial P_{r, i} }{\partial P } c_{i}\right)} & \cond{dci_chem}{\dconv/}
+  \lhs{dci_chem}{\frac{\partial c_i }{\partial V }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial V } - \frac{\partial P_{r, i} }{\partial V } c_{i}\right)} & \cond{dci_chem}{\dconp/,} \\
+  \lhs{dci_chem}{\frac{\partial c_i }{\partial P }} &= \rhs{dci_chem}{\frac{1}{P_{r, i} + 1} \left(\frac{\partial F_{i} }{\partial P } - \frac{\partial P_{r, i} }{\partial P } c_{i}\right)} & \cond{dci_chem}{\dconv/,}
  \end{align}
@@
 Next the reduced pressure derivatives in~\crefrange{e:dci_fall_dt}{e:dci_chem_de} will be evaluated; these depend on the form of the third-body efficiency used in~\cref{e:pr}, thus we will first work out the derivatives for each third-body efficiency form, and then develop a generalized form for compact representation.
-For a third-body concentration based upon the entire mixture, (i.e.,~\cref{e:thd_mix}) the derivatives are:
+For a third-body concentration based upon the entire mixture, (i.e.,~\cref{e:thd_mix}) the temperature derivative is:
 \begin{align}
@@
 \end{align}
-and for the state parameter derivative:
+and the state parameter derivatives are:
 \begin{subequations}
@@
  \begin{align}
-  \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_mix}{- \frac{P_{r, i}}{V} + \frac{[C] k_{0, i} \alpha_{\ns/,i}}{V k_{\infty, i}}} & \cond{dpr_mix}{\dconp/} \\
-  \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_mix}{\frac{k_{0, i} \alpha_{\ns/,i}}{T k_{\infty, i} \Ru/}} & \cond{dpr_mix}{\dconv/}
- \end{align}
-\end{subequations}
-and:
+  \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_mix}{- \frac{P_{r, i}}{V} + \frac{[C] k_{0, i} \alpha_{\ns/,i}}{V k_{\infty, i}}} & \cond{dpr_mix}{\dconp/,} \\
+  \lhs{dpr_mix}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_mix}{\frac{k_{0, i} \alpha_{\ns/,i}}{T k_{\infty, i} \Ru/}} & \cond{dpr_mix}{\dconv/,}
+ \end{align}
+\end{subequations}
+and finally, the molar derivative is given by:
 \begin{align}
@@
 \end{align}
-and for the state parameter:
+and the state parameter derivatives are:
 \begin{subequations}
@@
  \begin{align}
-  \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_unity}{0} & \cond{dpr_unity}{\dconp/}\\
-  \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_unity}{\frac{k_{0, i}}{k_{\infty, i}} \frac{1}{\Ru/ T}} & \cond{dpr_unity}{\dconv/}
+  \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_unity}{0} & \cond{dpr_unity}{\dconp/,}\\
+  \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_unity}{\frac{k_{0, i}}{k_{\infty, i}} \frac{1}{\Ru/ T}} & \cond{dpr_unity}{\dconv/,}
+ \end{align}
+\end{subequations}
+and the molar derivative is:
+\begin{align}
+ \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial n_j }} &= \rhs{dpr_unity}{0} & \cond{dpr_unity}{.} \label{e:dpr_unity_dnj}
+\end{align}
+
+Finally, for species-based third-body (i.e., the species $m$ alone acts as the third body):
+\begin{align}
+ \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial T }} &=\rhs{dpr_spec}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru/} - \frac{E_{a, \infty}}{T \Ru/}\right) - \frac{[C] k_{0, i} \delta_{\ns/ m}}{T k_{\infty, i}}} & \cond{dpr_spec}{,} \label{e:dpr_spec_dt}
+\end{align}
+for the state parameter derivative:
+\begin{subequations}
+ \label{e:dpr_spec_de}
+ \begin{align}
+  \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_spec}{- \frac{P_{r, i}}{V} + \frac{[C] k_{0, i} \delta_{\ns/ m}}{V k_{\infty, i}}} & \cond{dpr_spec}{\dconp/,}\\
+  \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_spec}{\frac{k_{0, i} \delta_{\ns/ m}}{T k_{\infty, i} \Ru/}} & \cond{dpr_spec}{\dconv/,}
  \end{align}
@@
 \begin{align}
- \lhs{dpr_unity}{\frac{\partial P_{r, i} }{\partial n_j }} &= \rhs{dpr_unity}{0} & \cond{dpr_unity}{.} \label{e:dpr_unity_dnj}
-\end{align}
-
-Finally, for species-based third-body (i.e., the species $m$ alone acts as the third body):
-\begin{align}
- \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial T }} &=\rhs{dpr_spec}{\frac{P_{r, i}}{T} \left(\beta_{0} - \beta_{\infty} + \frac{E_{a, 0}}{T \Ru/} - \frac{E_{a, \infty}}{T \Ru/}\right) - \frac{[C] k_{0, i} \delta_{\ns/ m}}{T k_{\infty, i}}} & \cond{dpr_spec}{,} \label{e:dpr_spec_dt}
-\end{align}
-for the state paramter derivative:
-\begin{subequations}
- \label{e:dpr_spec_de}
- \begin{align}
-  \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial V }} &= \rhs{dpr_spec}{- \frac{P_{r, i}}{V} + \frac{[C] k_{0, i} \delta_{\ns/ m}}{V k_{\infty, i}}} & \cond{dpr_spec}{\dconp/}\\
-  \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial P }} &= \rhs{dpr_spec}{\frac{k_{0, i} \delta_{\ns/ m}}{T k_{\infty, i} \Ru/}} & \cond{dpr_spec}{\dconv/}
- \end{align}
-\end{subequations}
-and the molar derivative:
-\begin{align}
  \lhs{dpr_spec}{\frac{\partial P_{r, i} }{\partial n_j }} &=\rhs{dpr_spec}{\frac{k_{0, i}}{V k_{\infty, i}} \left(- \delta_{\ns/ m} + \delta_{j m}\right)} & \cond{dpr_spec}{.} \label{e:dpr_spec_dnj}
@@
  \begin{align}
-  \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial V}} &= \rhs{dpr_gen}{P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}} & \cond{dpr_gen}{\dconp/} \\
-  \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial P}} &= \rhs{dpr_gen}{\bar{\theta}_{P_{r, i}, \partial P}} & \cond{dpr_gen}{\dconv/}
+  \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial V}} &= \rhs{dpr_gen}{P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}} & \cond{dpr_gen}{\dconp/,} \\
+  \lhs{dpr_gen}{\frac{\partial P_{r, i} }{\partial P}} &= \rhs{dpr_gen}{\bar{\theta}_{P_{r, i}, \partial P}} & \cond{dpr_gen}{\dconv/,}
  \end{align}
@@
   \rhs{theta_pr_bar_dnj}{0}[c] & \text{if unity,}  \\
-  \rhs{theta_pr_bar_dnj}{- \delta_{\ns/ m} + \delta_{j m}} & \text{if species,}
+  \rhs{theta_pr_bar_dnj}{- \delta_{\ns/ m} + \delta_{j m}} & \text{if species.}
  \end{dcases}
@@
 
-The falloff blending factor for all but Lindemann falloff reactions relies upon the reduced pressure derivatives developed in~\cref{s:dpr}.
+Derivatives of the falloff blending factor---except for for all but the Lindemann falloff blending factor, which is not a function of the reduced pressure---rely upon the reduced pressure derivatives developed in~\cref{s:dpr}.
 First, the blending factor for each falloff reaction type will be differentiated along with any subcomponents (e.g., the $F_{cent}$ term for Troe falloff reactions).
-Next, the reduced pressured derivatives given by~\cref{e:dpr_dt,e:dpr_dnj,e:dpr_de} will be substituted in, before finally deriving a generalized form for the falloff blending factor derivatives.
+Next, the reduced pressured derivatives given by~\cref{e:dpr_dt,e:dpr_dnj,e:dpr_de} will be substituted in and finally a generalized form for the falloff blending factor derivatives will be determined.
 
@@
  \begin{align}
-  \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{\dconp/} \\
-  \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{\dconv/}
+  \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{\dconp/,} \\
+  \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{\dconv/,}
  \end{align}
@@
 \begin{align}
- \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial n_j }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{,} \label{e:dfi_lind_dnj}
+ \lhs{dfi_lind}{\frac{\partial F_{i} }{\partial n_j }} &= \rhs{dfi_lind}{0} & \cond{dfi_lind}{.} \label{e:dfi_lind_dnj}
 \end{align}
@@
  \begin{align}
- \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial V }} & \cond{dfi_troe_1}{\dconp/} \\
- \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial P }} & \cond{dfi_troe_1}{\dconv/}
+ \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial V }} & \cond{dfi_troe_1}{\dconp/,} \\
+ \lhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_troe_1}{\frac{\partial F_{i} }{\partial P_{r, i} } \frac{\partial P_{r, i} }{\partial P }} & \cond{dfi_troe_1}{\dconv/,}
  \end{align}
@@
 \end{align}
-where:
+The Troe falloff blending factor derivative with respect to $F_{cent}$ is:
 \begin{align}
@@
 \vphantom{\frac{F_{i}}{\frac{A_{Troe}^{2}}{B_{Troe}^{2}} + 1}} \right) & \text{,} \label{e:dfi_troe_dfcent_partial} \\
+\intertext{and:}
 \frac{\text{d} F_{cent} }{\text{d} T } &= \frac{a}{T^{*}} \operatorname{exp}\left({- \frac{T}{T^{*}}}\right) - \frac{\operatorname{exp}\left({- \frac{T}{T^{***}}}\right)}{T^{***}} \left(- a + 1\right) + & \nonumber \\
@@
 \intertext{and:}
-\frac{\partial F_{i} }{\partial P_{r, i} } &= \frac{2 F_{i} A_{Troe} \log{\left (F_{cent} \right )}}{B_{Troe}^{2} \left(\frac{A_{Troe}^{2}}{B_{Troe}^{2}} + 1\right)^{2}} \left(\frac{A_{Troe}}{B_{Troe}} \frac{\partial B_{Troe} }{\partial P_{r, i} } - \frac{\partial A_{Troe} }{\partial P_{r, i} }\right) \label{e:dfi_troe_dpr_partial}
-\end{align}
-where the derivatives of $A_{Troe}$ and $B_{Troe}$ are:
+\frac{\partial F_{i} }{\partial P_{r, i} } &= \frac{2 F_{i} A_{Troe} \log{\left (F_{cent} \right )}}{B_{Troe}^{2} \left(\frac{A_{Troe}^{2}}{B_{Troe}^{2}} + 1\right)^{2}} \left(\frac{A_{Troe}}{B_{Troe}} \frac{\partial B_{Troe} }{\partial P_{r, i} } - \frac{\partial A_{Troe} }{\partial P_{r, i} }\right) \;.
+\label{e:dfi_troe_dpr_partial}
+\end{align}
+Finally, the derivatives of $A_{Troe}$ and $B_{Troe}$ are:
 \begin{align}
@@
 0.67 B_{Troe}\right) \log{\left (F_{cent} \right )} - B_{Troe} \left(A_{Troe}^{2} + B_{Troe}^{2}\right) \log{\left (10 \right )}\right) & \text{,} \label{e:dfi_troe_dfcent} \\
+\intertext{and:}
 \frac{\partial F_{i} }{\partial P_{r, i} } &= - \frac{2 F_{i} A_{Troe} \left(\frac{0.14 A_{Troe}}{B_{Troe}} + 1\right) \log{\left (F_{cent} \right )}}{B_{Troe}^{2} P_{r, i} \left(\frac{A_{Troe}^{2}}{B_{Troe}^{2}} + 1\right)^{2} \log{\left (10 \right )}} \label{e:dfi_troe_dpr} &
@@
 
-For SRI falloff reactions:
+For SRI falloff reactions, the temperature derivative of the falloff blending factor is:
 \begin{align}
  \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial T }} &= \rhs{dfi_sri}{F_{i} \vast(\frac{X \left(- \frac{\operatorname{exp}\left({- \frac{T}{c}}\right)}{c} + \frac{a b}{T^{2}} \operatorname{exp}\left({- \frac{b}{T}}\right)\right)}{a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right)} + } & \nonumber \\
- & \rhs{dfi_sri}{\mathindent \frac{\partial P_{r, i} }{\partial T } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\biggl[}} & \nonumber \\
+ & \rhs{dfi_sri}{\mathindent[extras] \frac{\partial P_{r, i} }{\partial T } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\biggl[}} & \nonumber \\
  & \rhs{dfi_sri}{\mathindent[largerstill] a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \biggr] + \frac{e}{T} \vast)} & \cond{dfi_sri}{,} \label{e:dfi_sri_dt}
 \end{align}
-the state parameter derivative:
+and the state parameter derivatives:
 \begin{subequations}
@@
  \begin{align}
- \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial V } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} & \cond{dfi_sri}{\dconp/} \\
- \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial P } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} & \cond{dfi_sri}{\dconv/}
- \end{align}
-\end{subequations}
-and:
+ \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial V } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} & \cond{dfi_sri}{\dconp/,} \\
+ \lhs{dfi_sri}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_sri}{F_{i} \frac{\partial P_{r, i} }{\partial P } \frac{\text{d} X }{\text{d} P_{r, i} } \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}} & \cond{dfi_sri}{\dconv/,}
+ \end{align}
+\end{subequations}
+and finally, with respect to species moles:
 \begin{align}
@@
 
-where:
+The derivatives of the SRI exponent $X$ are:
 \begin{align}
  \frac{\text{d} X }{\text{d} P_{r, i} } &= - \frac{2 X^{2} \log{\left (P_{r, i} \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} &, \\
- \frac{\partial X}{\partial n_j} &= \frac{\partial P_{r, i} }{\partial {n_j} } \frac{\text{d} X }{\text{d} P_{r, i} }
-\end{align}
-
-As with the reduced-pressure derivatives,~\cref{e:dfi_lind_dt,e:dfi_lind_dnj,e:dfi_lind_de,e:dfi_troe_dt,e:dfi_troe_dnj,e:dfi_troe_de,e:dfi_sri_dt,e:dfi_sri_dnj,e:dfi_sri_de} may be represented in a general form via introduction of temporary variables $\Theta_{F_i, \partial \ldots}$:
+ \frac{\partial X}{\partial n_j} &= \frac{\partial P_{r, i} }{\partial {n_j} } \frac{\text{d} X }{\text{d} P_{r, i} } &.
+\end{align}
+
+As with the reduced-pressure derivatives,~\cref{e:dfi_lind_dt,e:dfi_lind_dnj,e:dfi_lind_de,e:dfi_troe_dt,e:dfi_troe_dnj,e:dfi_troe_de,e:dfi_sri_dt,e:dfi_sri_dnj,e:dfi_sri_de} may be represented in a general form via introduction of temporary variables $\Theta_{F_i, \partial \ldots}$, giving for the temperature derivative:
 \begin{align}
@@
 \end{align}
-for the state parameter derivative:
-\begin{subequations}
- \begin{align}
-  \lhs{dfi_gen}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_gen}{F_{i} \Theta_{F_i, \partial V}} & \cond{dfi_gen}{\dconp/} \\
-  \lhs{dfi_gen}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_gen}{F_{i} \Theta_{F_i, \partial P}} & \cond{dfi_gen}{\dconv/}
- \end{align}
-\end{subequations}
-and the molar derivative:
+for the state parameter derivatives:
+\begin{subequations}
+ \begin{align}
+  \lhs{dfi_gen}{\frac{\partial F_{i} }{\partial V }} &= \rhs{dfi_gen}{F_{i} \Theta_{F_i, \partial V}} & \cond{dfi_gen}{\dconp/,} \\
+  \lhs{dfi_gen}{\frac{\partial F_{i} }{\partial P }} &= \rhs{dfi_gen}{F_{i} \Theta_{F_i, \partial P}} & \cond{dfi_gen}{\dconv/,}
+ \end{align}
+\end{subequations}
+and finally, for the molar derivative:
 \begin{align}
@@
 \end{align}
-where for Lindemann falloff reactions:
+
+For Lindemann falloff reactions:
 \begin{align}
 \label{e:dfi_lind_dt_prod}
-\lhs{dfi_lind_gen}{\Theta_{F_i, \partial T}} &= \rhs{dfi_lind_gen}{0}
-\end{align}
-and, for the state parameter derivative:
+\lhs{dfi_lind_gen}{\Theta_{F_i, \partial T}} &= \rhs{dfi_lind_gen}{0} &,
+\end{align}
+and, for the state parameter derivatives:
 \begin{subequations}
@@
  \begin{align}
-  \lhs{dfi_lind_gen}{\Theta_{F_i, \partial V}} &= \rhs{dfi_lind_gen}{0} \\
-  \lhs{dfi_lind_gen}{\Theta_{F_i, \partial P}} &= \rhs{dfi_lind_gen}{0}
- \end{align}
-\end{subequations}
-and:
+  \lhs{dfi_lind_gen}{\Theta_{F_i, \partial V}} &= \rhs{dfi_lind_gen}{0} &, \\
+  \lhs{dfi_lind_gen}{\Theta_{F_i, \partial P}} &= \rhs{dfi_lind_gen}{0} &,
+ \end{align}
+\end{subequations}
+and for the molar derivative:
 \begin{align}
 \label{e:dfi_lind_dnj_prod}
-\lhs{dfi_lind_gen}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_lind_gen}{0}
+\lhs{dfi_lind_gen}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_lind_gen}{0} &.
 \end{align}
@@
 &\rhs{dfi_troe_gen}{\mathindent 2 A_{Troe} \left(1.1762 A_{Troe} - 0.67 B_{Troe}\right) \log{\left (F_{cent} \right )} - } \nonumber \\
-&\rhs{dfi_troe_gen}{\mathindent B_{Troe} \left(A_{Troe}^{2} + B_{Troe}^{2}\right) \log{\left (10 \right )}\Bigr)\Biggr)}
-\end{align}
-and for the state parameter derivative:
+&\rhs{dfi_troe_gen}{\mathindent B_{Troe} \left(A_{Troe}^{2} + B_{Troe}^{2}\right) \log{\left (10 \right )}\Bigr)\Biggr)} \;,
+\end{align}
+and for the state parameter derivatives:
 \begin{subequations}
@@
    \lhs{dfi_troe_gen}{\Theta_{F_i, \partial V}} &= \rhs{dfi_troe_gen}{- \frac{2 A_{Troe} B_{Troe} \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}} \times \bigl(} \nonumber \\
-					    & \rhs{dfi_troe_gen}{\mathindent[alittlebit] 0.14 A_{Troe} + B_{Troe}\bigr) \left(P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}\right)} \\
-   \lhs{dfi_troe_gen}{\Theta_{F_i, \partial P}} &= \rhs{dfi_troe_gen}{- \frac{2 A_{Troe} B_{Troe} \bar{\theta}_{P_{r, i}, \partial P} \left(0.14 A_{Troe} + B_{Troe}\right) \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}}}
- \end{align}
-\end{subequations}
-and:
+					    & \rhs{dfi_troe_gen}{\mathindent[alittlebit] 0.14 A_{Troe} + B_{Troe}\bigr) \left(P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}\right)} \;, \\
+   \lhs{dfi_troe_gen}{\Theta_{F_i, \partial P}} &= \rhs{dfi_troe_gen}{- \frac{2 A_{Troe} B_{Troe} \bar{\theta}_{P_{r, i}, \partial P} \left(0.14 A_{Troe} + B_{Troe}\right) \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}}}\;,
+ \end{align}
+\end{subequations}
+and the molar derivative:
 \begin{align}
  \label{e:dfi_troe_dnj_prod}
-\lhs{dfi_troe_gen}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_troe_gen}{- \frac{2 A_{Troe} B_{Troe} \left(0.14 A_{Troe} + B_{Troe}\right) \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}}}
+\lhs{dfi_troe_gen}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_troe_gen}{- \frac{2 A_{Troe} B_{Troe} \left(0.14 A_{Troe} + B_{Troe}\right) \log{\left (F_{cent} \right )}}{P_{r, i} \left(A_{Troe}^{2} + B_{Troe}^{2}\right)^{2} \log{\left (10 \right )}}}\;.
 \end{align}
@@
  &\rhs{dfi_sri_gen}{\mathindent \frac{2 X^{2} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \times } \nonumber \\
- &\rhs{dfi_sri_gen}{\mathindent \left(P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right) \log{\left (P_{r, i} \right ) \Biggr)}}
+ &\rhs{dfi_sri_gen}{\mathindent \left(P_{r, i} \Theta_{P_{r,i}, \partial T} + \bar{\theta}_{P_{r, i}, \partial T}\right) \log{\left (P_{r, i} \right ) \Biggr)}}\;,
 \end{align}
@@
   \rhs{dfi_sri_gen}{- \frac{2 X^{2} \log{\left (P_{r, i} \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \left(P_{r, i} \Theta_{P_{r,i}, \partial V} + \bar{\theta}_{P_{r, i}, \partial V}\right) \log \Biggl(} \nonumber \\
-& \rhs{dfi_sri_gen}{\mathindent \left( a \operatorname{exp}\left({\frac{T}{c}}\right) + \operatorname{exp}\left({\frac{b}{T}}\right)\right) \operatorname{exp}\left({- \frac{T}{c} - \frac{b}{T}}\right) \Biggr)} \\
+& \rhs{dfi_sri_gen}{\mathindent \left( a \operatorname{exp}\left({\frac{T}{c}}\right) + \operatorname{exp}\left({\frac{b}{T}}\right)\right) \operatorname{exp}\left({- \frac{T}{c} - \frac{b}{T}}\right) \Biggr)}\;, \\
   \lhs{dfi_sri_gen}{\Theta_{F_i, \partial P}} &=
-  \rhs{dfi_sri_gen}{- \frac{2 X^{2} \bar{\theta}_{P_{r, i}, \partial P} \log{\left (P_{r, i} \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}
+  \rhs{dfi_sri_gen}{- \frac{2 X^{2} \bar{\theta}_{P_{r, i}, \partial P} \log{\left (P_{r, i} \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}\;,
  \end{align}
@@
  \label{e:dfi_sri_dnj_prod}
- \lhs{dfi_sri_gen}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_sri_gen}{- \frac{2 X^{2} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \log{\left (P_{r, i} \right )}}
+ \lhs{dfi_sri_gen}{\Theta_{F_i, \partial n_j}} &= \rhs{dfi_sri_gen}{- \frac{2 X^{2} \log{\left (a \operatorname{exp}\left({- \frac{b}{T}}\right) + \operatorname{exp}\left({- \frac{T}{c}}\right) \right )}}{P_{r, i} \log^{2}{\left (10 \right )}} \log{\left (P_{r, i} \right )}}\;.
 \end{align}
@@
 \end{align}
-for the state parameter derivative:
+and, for the state parameter derivatives:
 \begin{subequations}
@@
   \lhs{dci_fall_complete}{\frac{\partial c_{i}}{\partial V }} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1} + \biggl(- \frac{P_{r, i} \Theta_{P_{r,i}, \partial V}}{P_{r, i} + 1} + \Theta_{F_i, \partial V} + } & \nonumber \\
-							       &  \rhs{dci_fall_complete}{\mathindent[largerindentpleas] \Theta_{P_{r,i}, \partial V} - \frac{\bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1} \biggr) c_{i}} & \cond{dci_fall_complete}{\dconp/} \\
-  \lhs{dci_fall_complete}{\frac{\partial c_{i}}{\partial P }} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1} + \left(\Theta_{F_i, \partial P} - \frac{\bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_fall_complete}{\dconv/}
+							       &  \rhs{dci_fall_complete}{\mathindent[largerindentpleas] \Theta_{P_{r,i}, \partial V} - \frac{\bar{\theta}_{P_{r, i}, \partial V}}{P_{r, i} + 1} \biggr) c_{i}} & \cond{dci_fall_complete}{\dconp/,} \\
+  \lhs{dci_fall_complete}{\frac{\partial c_{i}}{\partial P }} &= \rhs{dci_fall_complete}{\frac{F_{i} \bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1} + \left(\Theta_{F_i, \partial P} - \frac{\bar{\theta}_{P_{r, i}, \partial P}}{P_{r, i} + 1}\right) c_{i}} & \cond{dci_fall_complete}{\dconv/,}
  \end{align}
@@
 \end{align}
-and the state paramter derivative:
+and the state parameter derivatives:
 \begin{subequations}
@@
 \subsubsection{Molar source term derivatives}
-To obtain a final form for the molar source term derivative with respect to the state paramter,~\cref{e:dnkdotdE_compact,e:dwdot_de} are combined:
+To obtain a final form for the molar source term derivative with respect to the state parameter,~\cref{e:dnkdotdE_compact,e:dwdot_de} are combined:
 \loadeq{dnkdotdE}
 
-The derivative of the net rate of progress with respect to the state paramter is given by~\cref{e:dropi_de} for arrhenius-based reactions,~\cref{e:dropi_plog_de,e:dropi_cheb_de} for P-Log and Chebyshev reactions for a \conv/ system, and~\cref{e:dropi_de} for a \conp/ system.
+The derivative of the net rate of progress with respect to the state parameter is given by~\cref{e:dropi_de} for arrhenius-based reactions.
+For P-Log and Chebyshev reactions in a \conv/ system~\cref{e:dropi_plog_de,e:dropi_cheb_de} are used, while~\cref{e:dropi_de} is used for a \conp/ system.
 The pressure modification term derivatives are given by~\cref{e:dci_thd_de,e:dci_thd_unity_de,e:dci_thd_spec_de} for third-body enhanced reactions, and~\cref{e:dci_fall_de_final,e:dci_chem_de_final} for falloff\slash chemically-activated reactions respectively.

### 2018-03-02 paper.tex
put more about jacobian / model size in intro

@@
 
+As the importance of detailed chemical kinetics for predictive reactive-flow simulations has become recognized~\cref{LU2009192}, the size and complexity of chemical kinetic models describing current and next-generation fuels relevant to transportation and power generation have increased greatly; e.g., a recent biodiesel model consists of \textasciitilde\num{3500} chemical species, and over \num{17000} reactions.
+Moreover, the cost of evaluating the chemical source terms scales linearly with the size of the model, while evaluation and factorization of the chemical kinetic Jacobian, if naively implemented, scale quadratically and cubically with the number of species in the model~\cref{LU2009192}.
+These factors often make the use of detailed chemical kinetics prohibitive in practice; for direct numerical simulations, evaluation of reaction rates can consume around half of the total runtime~\cite{Spafford:2010aa}.
+In addition, chemical kinetic Jacobian evaluation and factorization---required for most implicit integration techniques utilized to deal with numerical stiffness---becomes a bottleneck for even modestly sized chemical models in realistic reactive flow simulations~\cite{LU2009192}, and necessitates use of other cost-reduction strategies.
+
 Single-Instruction, Multiple-Data (SIMD) and the related Single-Instruction, Multiple-Thread (SIMT) processing are two important vector-processing paradigms used increasingly in scientific computing.
-Traditional multi-core parallelism is often used to increase central processing unit (CPU) performance, however SIMD processors---as well as SIMT processors, e.g., in the form of graphics processing units (GPUs)---have gained recognition due to their increased floating operation throughput.
+Traditional multi-core parallelism is often used to increase central processing unit (CPU) performance, however as exponential growth in processing power---colloquially known as Moore's law---has slowed~\cite{khan2018science}, SIMD processors as well as SIMT processors, e.g., in the form of graphics processing units (GPUs), have gained recognition due to their increased floating operation throughput.
 The parallel programming standard OpenCL~\cite{stone2010opencl} has further enabled adoption of vector-processing based codes in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems, e.g., the CPU, GPU, or Intel's Many Integrated Core architecture (MIC), etc.
@@
 
-\begin{figure}[htb]
+\begin{figure}[phtb]
   \centering
@@
 SIMT processing---the foundation of modern GPUs~\cite{lindholm2008nvidia}---is a related computing paradigm that hosts large groups of threads on a single compute element (a streaming multiprocessor in NVIDIA terminology).
-A group of threads---typically \num{32} threads, known as a warp on NVIDIA GPUs---execute the same SIMD instruction on multiple data concurrently (\cref{F:SIMT}).
+A group of threads---typically \num{32}, known as a warp on NVIDIA GPUs---execute the same SIMD instruction on multiple data concurrently (\cref{F:SIMT}).
 If some threads must execute a different instruction---e.g., due to if\slash then branching or predication---they are forced to wait and execute later.
@@
 Niemeyer et al.~\cite{Niemeyer:2016aa} additionally developed and validated an analytical chemical kinetic Jacobian code for the CPU and GPU.
-Yonkee and Sutherland~\cite{Yonkee2016} implemented SIMT-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs) on an Intel Xeon E5-2680 CPU and an Tesla K20 GPU.
+Yonkee and Sutherland~\cite{Yonkee2016} implemented SIMT-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs) on an Intel Xeon E5-2680 CPU and a Tesla K20 GPU.
 In evaluation of the thermo-chemical properties, speedups between \SIrange{8}{13}{$\times$} were on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU respectively.
@@
 \end{itemize}
-.
 

### 2018-03-05 paper.tex
update

@@
 
-As the importance of detailed chemical kinetics for predictive reactive-flow simulations has become recognized~\cref{LU2009192}, the size and complexity of chemical kinetic models describing current and next-generation fuels relevant to transportation and power generation have increased greatly; e.g., a recent biodiesel model consists of \textasciitilde\num{3500} chemical species, and over \num{17000} reactions.
-Moreover, the cost of evaluating the chemical source terms scales linearly with the size of the model, while evaluation and factorization of the chemical kinetic Jacobian, if naively implemented, scale quadratically and cubically with the number of species in the model~\cref{LU2009192}.
+As the importance of detailed chemical kinetics for predictive reactive-flow simulations has become recognized~\cite{LU2009192}, the size and complexity of chemical kinetic models describing current and next-generation fuels relevant to transportation and power generation have increased greatly; e.g., a recent biodiesel model consists of \textasciitilde\num{3500} chemical species, and over \num{17000} reactions.
+Moreover, the cost of evaluating the chemical source terms scales linearly with the size of the model, while evaluation and factorization of the chemical kinetic Jacobian, if naively implemented, scale quadratically and cubically with the number of species in the model~\cite{LU2009192}.
 These factors often make the use of detailed chemical kinetics prohibitive in practice; for direct numerical simulations, evaluation of reaction rates can consume around half of the total runtime~\cite{Spafford:2010aa}.
-In addition, chemical kinetic Jacobian evaluation and factorization---required for most implicit integration techniques utilized to deal with numerical stiffness---becomes a bottleneck for even modestly sized chemical models in realistic reactive flow simulations~\cite{LU2009192}, and necessitates use of other cost-reduction strategies.
+In addition, chemical kinetic Jacobian evaluation and factorization---required for most common implicit integration techniques utilized to deal with numerical stiffness---becomes a bottleneck for even modestly-sized chemical models in realistic reactive flow simulations, and necessitates use of other cost-reduction strategies~\cite{LU2009192}.
 

### 2018-03-05 paper.tex
add discussion of analytical jacobian

@@
 
-As the importance of detailed chemical kinetics for predictive reactive-flow simulations has become recognized~\cite{LU2009192}, the size and complexity of chemical kinetic models describing current and next-generation fuels relevant to transportation and power generation have increased greatly; e.g., a recent biodiesel model consists of \textasciitilde\num{3500} chemical species, and over \num{17000} reactions.
+As the importance of detailed chemical kinetics for predictive reactive-flow simulations has become recognized~\cite{LU2009192}, the size and complexity of chemical kinetic models describing current and next-generation fuels relevant to transportation and power generation have increased greatly; e.g., a recent biodiesel model~\cite{WESTBROOK2011742} consists of \textasciitilde\num{3500} chemical species, and over \num{17000} reactions.
 Moreover, the cost of evaluating the chemical source terms scales linearly with the size of the model, while evaluation and factorization of the chemical kinetic Jacobian, if naively implemented, scale quadratically and cubically with the number of species in the model~\cite{LU2009192}.
-These factors often make the use of detailed chemical kinetics prohibitive in practice; for direct numerical simulations, evaluation of reaction rates can consume around half of the total runtime~\cite{Spafford:2010aa}.
+These factors often make the use of detailed chemical kinetics prohibitive in practice; e.g., for direct numerical simulations, evaluation of reaction rates with a 22-species model can consume around half of the total runtime~\cite{Spafford:2010aa}.
 In addition, chemical kinetic Jacobian evaluation and factorization---required for most common implicit integration techniques utilized to deal with numerical stiffness---becomes a bottleneck for even modestly-sized chemical models in realistic reactive flow simulations, and necessitates use of other cost-reduction strategies~\cite{LU2009192}.
 
-Single-Instruction, Multiple-Data (SIMD) and the related Single-Instruction, Multiple-Thread (SIMT) processing are two important vector-processing paradigms used increasingly in scientific computing.
+A host of techniques exist to lessen the computational demand of chemical kinetic calculations while maintaining fidelity, falling broadly into a few categories; removal of unimportant species and reactions unimportant species and reactions~\cite{Lu:2006bb,Pepiot-Desjardins:2008,Hiremath:2010jw,Niemeyer:2010bt,Curtis:2015}, lumping of species with similar thermo-chemical properties~\cite{Lu:2007,Ahmed:2007fa,Pepiot:2008kq}, and time-scale methods that reduce chemical stiffness~\cite{Maas:1992ws,Lam:1994ws,Lu:2001ve,Gou:2010}.
+The interested reader is referred to the recent review of Tur\'anyi and Tomlin~\cite{turanyi2016analysis} for a comprehensive overview.
+In addition to the previously mentioned cost reduction methods, effort has gone into improving the integration algorithms and codes that evaluate the chemical kinetics~\cite{Gou:2010,SCHWER2002270,Niemeyer:2016aa,GAO2015287}.
+In particular, a carefully derived analytical formulation of the Jacobian can greatly increase the sparsity of the matrix~\cite{SCHWER2002270} and also drop the cost of Jacobian evaluation from scaling quadratically with the number of species in the model---for a naive finite-difference implementation---to a linear dependence~\cite{LU2009192}, while Jacobian factorization can be accelerated via sparse-matrix techniques~\cite{superlu99}.
+Use of Single-Instruction, Multiple-Data (SIMD) and the related Single-Instruction, Multiple-Thread (SIMT) processors to accelerate chemical kinetic simulations has also been studied, e.g., as in~\cite{CurtisGPU:2017,Sewerin20151375,Shi:2012aa,Niemeyer:2016aa}.
+
+SIMD and SIMT programming are two important vector-processing paradigms used increasingly in scientific computing.
 Traditional multi-core parallelism is often used to increase central processing unit (CPU) performance, however as exponential growth in processing power---colloquially known as Moore's law---has slowed~\cite{khan2018science}, SIMD processors as well as SIMT processors, e.g., in the form of graphics processing units (GPUs), have gained recognition due to their increased floating operation throughput.
@@
 \subsection{Previous works and goals of this study}
+
+Recognizing the bottleneck caused by evaluation and factorization of the chemical kinetic Jacobian, a number of recent works have developed analytical chemical kinetic Jacobian codes, although as will be discussed at the end of this section this work offers several improvements over past efforts.
+The \texttt{TChem} software package~\cite{Safta:2011vn} was one of the first packages developed providing this functionality, but has several limitations including incompatibilities with modern reaction types (i.e., pressure dependent Arrhenius and Chebyshev reactions), and is not safely run in parallel~\cite{Curtis2017:tchem}.
+Youssefi~\cite{Youssefi:2011tm} discussed the accuracy of Jacobian matricies evaluated with finite-differences compared to analytical derivations and their affect on computational efficiency as well as on the accuracy of time-scale analysis techniques such as CSP~\cite{Lam:1994ws}.
+Bisetti~\cite{Bisetti:2012jw} developed an isothermal, isobaric analytical Jacobian code-generation utility; although incompatible with many combustion codes, a significant increase in Jacobian sparsity was exhibited.
+Additionally, in the same work Bisetti provided a novel way to compute dense matrix-vector multiplications resulting from a change of system variables without need for storage of the full dense Jacobian.
+Perini et al.~\cite{Perini:2012gy} developed an analytical Jacobian code for constant-volume combustion, with additional options to increase sparsity (at expect of strict correctness), and enable tabulation of temperature dependent properties; an \SI{80}{\percent} speedup over a finite-difference-based Jacobian was reported when used in a multidimensional reactive-flow simulation.
+Dijkmans et al.~\cite{Dijkmans:2014bb} developed a GPU-based analytical Jacobian code with optional tabulation of temperature-dependent properties, and found speedups up to \SI{120}{$\times$} for zero-dimensional chemical kinetic integration with large chemical models (\textasciitilde\num{3000} species) with a Tesla C2075 GPU.
+Niemeyer et al.~\cite{Niemeyer:2016aa} created and validated the open-source analytical chemical kinetic Jacobian code-generator, supporting both the CPU and GPU, finding speedups of \SIrange{3}{7.5}{$\times$} over a finite-difference Jacobian on the CPU.
+Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but it is not publicly available nor was it validated (outside the context of use with an implicit-integration technique); in addition the Jacobian was based on an over-constrained system, and a differential algebraic solver should be used for strict correctness.
+
 A number of recent works have investigated the use of high-performance SIMT-devices to accelerate reactive-flow and chemical kinetics simulations.

### 2018-03-05 paper.tex
move to future

@@
 \end{itemize}
+To our knowledge, this is the first work that provides vectorized source term and Jacobian evaluation for any chemical model on the CPU, GPU and other accelerators.
 
@@
 \section{Results and Discussion}
-\subsection{Testing platforms and practical notes on OpenCL usage}
+\subsection{Testing platforms}
 
@@
 NVIDIA's OpenCL runtime is bundled with the NVIDIA driver~\cite{NVIDIA:2018}, hence we use driver version to specify the OpenCL runtime version.
-
 Finally,~\cref{t:platforms} lists the platforms and vectorization\slash execution patterns they are capable of running.
-It is important to note, that while OpenCL provides an easy way to enable cross-platform execution, there are some serious potential pit-falls in its use.
-Typically speaking, the closed source OpenCL runtimes tested in this work (Intel and NVIDIA) contain more bugs that result in compilation failures or simply incorrect vectorized machine code; further, these runtimes (in our experience) tend to be less responsive to fixing said bugs, with relatively infrequent new releases (or in NVIDIAs case, change-logs of bugfixes).
-As demonstration, we've created a minimum working example~\cite{nvidia_mwe} that demonstrates a failure of NVIDIA's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla K40m GPU; by simply changing to another runtime (e.g., Intel), with no other code changes, the correct result can be found.
-This provides a particularly vexing problem for the programmer, as there is often little that can be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
-
-Further, the Intel and NVIDIA OpenCL runtimes lack a key feature---implementations of atomic operations on double precision variables---that are currently necessary for \texttt{pyJac} to run deep-vectorized code.
+The Intel and NVIDIA OpenCL runtimes lack implementations of atomic operations on double precision variables; these are currently necessary for \texttt{pyJac} to run deep-vectorized code.
 On the other hand, POCL is an open source OpenCL runtime that works on all CPU types tested in this work, and does implement these atomic operations; it is our experience that the POCL community is very responsive to bug reports and user outreach.
@@
 Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the only significant code change is building the wrapper that initializes\slash transfers memory and calls the kernel.
-Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py}) is a key priority going forwards.
 This discussion will be expanded upon in~\cref{s:future} to highlight the most promising future directions.
@@
 No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP is faster in both cases.
-The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work; however, briefly speaking vectorization might be achieved on POCL via a combination of relaxed math (``\texttt{-cl-fast-relaxed-math}''), use of explicit OpenCL vector types and newer versions of LLVM\slash POCL (specifically, potential improvements to the vectorization module).
+The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work; however, briefly speaking vectorization might be achieved on POCL via a combination of fixed-array strides (rather than potentially varying with problem-size), use of relaxed math (``\texttt{-cl-fast-relaxed-math}''), explicit OpenCL vector types and newer versions of LLVM\slash POCL (specifically, potential improvements to the vectorization module).
 This will be explored more in future work, but for the moment POCL is still quite useful as a validation tool.
@@
 
-\section{Future directions}
+\section{Future directions and practical notes on OpenCL usage}
 \label{s:future}
-NVIDIA 375.26 is a bad driver, yo.
+It is important to note, that while OpenCL provides an easy way to enable cross-platform execution, there are some serious potential pit-falls in its use.
+Typically speaking, the closed source OpenCL runtimes tested in this work (Intel and NVIDIA) contain more bugs that result in compilation failures or simply incorrect vectorized machine code; further, these runtimes (in our experience) tend to be less responsive to fixing said bugs, with relatively infrequent new releases (or in NVIDIAs case, change-logs of bugfixes).
+As demonstration, we've created a minimum working example~\cite{nvidia_mwe} that demonstrates a failure of NVIDIA's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla K40m GPU; by simply changing to another runtime (e.g., Intel), with no other code changes, the correct result can be found.
+This provides a particularly vexing problem for the programmer, as there is often little that can be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
+
+Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py}) is a key priority going forwards.
 

### 2018-03-05 paper.tex
edits

@@
 As the importance of detailed chemical kinetics for predictive reactive-flow simulations has become recognized~\cite{LU2009192}, the size and complexity of chemical kinetic models describing current and next-generation fuels relevant to transportation and power generation have increased greatly; e.g., a recent biodiesel model~\cite{WESTBROOK2011742} consists of \textasciitilde\num{3500} chemical species, and over \num{17000} reactions.
-Moreover, the cost of evaluating the chemical source terms scales linearly with the size of the model, while evaluation and factorization of the chemical kinetic Jacobian, if naively implemented, scale quadratically and cubically with the number of species in the model~\cite{LU2009192}.
-These factors often make the use of detailed chemical kinetics prohibitive in practice; e.g., for direct numerical simulations, evaluation of reaction rates with a 22-species model can consume around half of the total runtime~\cite{Spafford:2010aa}.
-In addition, chemical kinetic Jacobian evaluation and factorization---required for most common implicit integration techniques utilized to deal with numerical stiffness---becomes a bottleneck for even modestly-sized chemical models in realistic reactive flow simulations, and necessitates use of other cost-reduction strategies~\cite{LU2009192}.
+Moreover, the cost of evaluating the chemical source terms scales linearly with the size of the model, while evaluation and factorization of the chemical kinetic Jacobian---if naively implemented---scale quadratically and cubically with the number of species in the model~\cite{LU2009192}.
+These factors often make the use of detailed chemical kinetics prohibitive in practice; e.g., in a recent direct numerical simulation~\cite{Spafford:2010aa} utilizing a 22-species model, evaluation of reaction rates consumed around half of the total runtime.
+In addition, chemical kinetic Jacobian evaluation and factorization---required for most common implicit integration techniques utilized to deal with numerical stiffness---is a bottleneck for even modestly-sized chemical models in realistic reactive flow simulations, and necessitates use of other cost-reduction strategies~\cite{LU2009192}.
 
@@
 In addition to the previously mentioned cost reduction methods, effort has gone into improving the integration algorithms and codes that evaluate the chemical kinetics~\cite{Gou:2010,SCHWER2002270,Niemeyer:2016aa,GAO2015287}.
-In particular, a carefully derived analytical formulation of the Jacobian can greatly increase the sparsity of the matrix~\cite{SCHWER2002270} and also drop the cost of Jacobian evaluation from scaling quadratically with the number of species in the model---for a naive finite-difference implementation---to a linear dependence~\cite{LU2009192}, while Jacobian factorization can be accelerated via sparse-matrix techniques~\cite{superlu99}.
+In particular, a carefully derived analytical formulation of the Jacobian can greatly increase the sparsity of the matrix~\cite{SCHWER2002270} and drop the cost of Jacobian evaluation from scaling quadratically with the number of species in the model---for a naive finite-difference implementation---to a linear dependence~\cite{LU2009192}, while Jacobian factorization can be accelerated via sparse-matrix techniques~\cite{superlu99}.
 Use of Single-Instruction, Multiple-Data (SIMD) and the related Single-Instruction, Multiple-Thread (SIMT) processors to accelerate chemical kinetic simulations has also been studied, e.g., as in~\cite{CurtisGPU:2017,Sewerin20151375,Shi:2012aa,Niemeyer:2016aa}.
@@
 Intel's Xeon Phi co-processor (MIC) have been developed that have tens of cores with very wide vector-widths (e.g.
-4--8 double precision operations); these very wide vector-widths are also found on cutting-edge and expected on forthcoming Intel CPUs (the Skylake Xeon and Cannon Lake architectures).
+4--8 double precision operations); these very wide vector-widths are also found on cutting-edge and forthcoming Intel CPUs; i.e., the Skylake Xeon and Cannon Lake architectures, respectively.
 
@@
 
-Recognizing the bottleneck caused by evaluation and factorization of the chemical kinetic Jacobian, a number of recent works have developed analytical chemical kinetic Jacobian codes, although as will be discussed at the end of this section this work offers several improvements over past efforts.
-The \texttt{TChem} software package~\cite{Safta:2011vn} was one of the first packages developed providing this functionality, but has several limitations including incompatibilities with modern reaction types (i.e., pressure dependent Arrhenius and Chebyshev reactions), and is not safely run in parallel~\cite{Curtis2017:tchem}.
-Youssefi~\cite{Youssefi:2011tm} discussed the accuracy of Jacobian matricies evaluated with finite-differences compared to analytical derivations and their affect on computational efficiency as well as on the accuracy of time-scale analysis techniques such as CSP~\cite{Lam:1994ws}.
+Recognizing the need to accelerate Jacobian evaluation and factorization, a number of recent works have been published on analytical chemical kinetic Jacobian evaluation; although as will be discussed at the end of this section, this work offers several key improvements over past efforts.
+The \texttt{TChem} software package~\cite{Safta:2011vn} was one of the first packages developed providing this functionality, but has several limitations including: incompatibilities with modern reaction types (i.e., pressure dependent Arrhenius and Chebyshev reactions), and lack of thread-safety to enable parallel execution~\cite{Curtis2017:tchem}.
+Youssefi~\cite{Youssefi:2011tm} explored the importance of analytical Jacobian matrices for time-scale analysis techniques as well as their effect on computational efficiency in zero-dimensional homogeneous reactor simulations.
 Bisetti~\cite{Bisetti:2012jw} developed an isothermal, isobaric analytical Jacobian code-generation utility; although incompatible with many combustion codes, a significant increase in Jacobian sparsity was exhibited.
 Additionally, in the same work Bisetti provided a novel way to compute dense matrix-vector multiplications resulting from a change of system variables without need for storage of the full dense Jacobian.
-Perini et al.~\cite{Perini:2012gy} developed an analytical Jacobian code for constant-volume combustion, with additional options to increase sparsity (at expect of strict correctness), and enable tabulation of temperature dependent properties; an \SI{80}{\percent} speedup over a finite-difference-based Jacobian was reported when used in a multidimensional reactive-flow simulation.
+Perini et al.~\cite{Perini:2012gy} developed an analytical Jacobian code for constant-volume combustion, with additional options to increase sparsity (at expect of strict correctness), and enable tabulation of temperature dependent properties; a \SI{80}{\percent} speedup over a finite-difference-based Jacobian was reported when used in a multidimensional reactive-flow simulation.
 Dijkmans et al.~\cite{Dijkmans:2014bb} developed a GPU-based analytical Jacobian code with optional tabulation of temperature-dependent properties, and found speedups up to \SI{120}{$\times$} for zero-dimensional chemical kinetic integration with large chemical models (\textasciitilde\num{3000} species) with a Tesla C2075 GPU.
-Niemeyer et al.~\cite{Niemeyer:2016aa} created and validated the open-source analytical chemical kinetic Jacobian code-generator, supporting both the CPU and GPU, finding speedups of \SIrange{3}{7.5}{$\times$} over a finite-difference Jacobian on the CPU.
-Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but it is not publicly available nor was it validated (outside the context of use with an implicit-integration technique); in addition the Jacobian was based on an over-constrained system, and a differential algebraic solver should be used for strict correctness.
+Niemeyer et al.~\cite{Niemeyer:2016aa} created and validated the open-source analytical chemical kinetic Jacobian code-generator, \texttt{pyJac}, supporting parallel execution on the CPU and SIMT execution on the GPU; a speedup of \SIrange{3}{7.5}{$\times$} over a finite-difference Jacobian was found on the CPU.
+Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but it was not validated outside the context of use with an implicit-integration technique; in addition the Jacobian was based on an over-constrained system, and a differential algebraic equation (DAE) solver should be used for strict correctness\slash conservation.
 
@@
 Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit fourth-order Runge--Kutta integrator for non-stiff chemical kinetic integration on a Tesla C2075 GPU, and found a speedup of nearly two orders of magnitude when compared with a sequential CPU-code on an Intel Xeon X5650 CPU.
-Shi et al.~\cite{Shi:2012aa} later extended their work to develop a stabilized-explicit solver on a Tesla C2050; paired with a traditional implicit integrator on quad-core Intel i7 930 which handled the stiffest computational cells, an \SIrange{11}{46}{$\times$} speedup was achieved over the base implicit CPU integrator during simulation of a three-dimensional premixed diesel engine simulation.
+Shi et al.~\cite{Shi:2012aa} later extended their work to develop a stabilized-explicit solver on a Tesla C2050; paired with a traditional implicit integrator on quad-core Intel i7 930 which handled the stiffest computational cells; a \SIrange{11}{46}{$\times$} speedup was achieved over the base implicit CPU integrator during simulation of a three-dimensional premixed diesel engine simulation.
 Le et al.~\cite{Le2013596} implemented two high-order shock-capturing reactive-flow codes on a Tesla C2070, and found a \numrange{30}{50}$\times$ speedup over a sequential CPU version of the same code on a Intel Xeon X5650.
@@
 Sewerin and Rigopoulos~\cite{Sewerin20151375} implemented a three-stage\slash fifth-order implicit Runge--Kutta method~\cite{wanner1991solving} on both a high-end (Nvidia Quadro 6000) and consumer-grade (Nvidia Quadro 600) GPUs, as well as a standard CPU (two-core, four-thread Intel i5-520M) and a scientific workstation (eight-core, 16-thread Intel Xeon E5-2687W); the high-end GPU was at best \SI{1.8}{$\times$} slower than the workstation CPU (16 threads), while the consumer level GPU was at best \SI{5.5}{$\times$} slower than the corresponding standard CPU (four threads).
-Niemeyer et al.~\cite{Niemeyer:2016aa} additionally developed and validated an analytical chemical kinetic Jacobian code for the CPU and GPU.
-Yonkee and Sutherland~\cite{Yonkee2016} implemented SIMT-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs) on an Intel Xeon E5-2680 CPU and a Tesla K20 GPU.
+Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs) on an Intel Xeon E5-2680 CPU and a Tesla K20 GPU.
 In evaluation of the thermo-chemical properties, speedups between \SIrange{8}{13}{$\times$} were on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU respectively.
-Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two-fourth order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on a Tesla C2075; the implicit Runge--Kutta method was roughly equivalent to VODE running on 12--38 Intel Xeon E5-4640 v2 CPU cores for two smaller chemical models with an integration time step of \SI{e-6}{$\sec$}.
+Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two-fourth order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on a Tesla C2075; the implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{cvode:2.8.2} running on 12--38 Intel Xeon E5-4640 v2 CPU cores for two smaller chemical models with an integration time step of \SI{e-6}{$\sec$}.
 
 In contrast, SIMD-based chemical kinetics evaluation\slash integration have been far less studied.
-Linford et al.~\cite{Linford:2011} implemented a three-stage, second-order Rosenbrock atmospheric chemical kinetics solver on the CPU, GPU and cell croadband engine (CBE)---a specially designed vector processor---and found speedups regularly exceeding~\SI{25}{$\times$} over a serial CPU implementation.
-Kroshko and Spiteri~\cite{kroshko2013efficient} implemented a shallow-SIMD vectorized third order stiff Rosenbrock integrator for atmospheric chemistry on the CBE finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code.
+Linford et al.~\cite{Linford:2011} implemented a three-stage, second-order Rosenbrock integrator for atmospheric chemical kinetics on the CPU, GPU and cell croadband engine (CBE)---a specially designed vector processor---and found speedups regularly exceeding~\SI{25}{$\times$} over a serial CPU implementation.
+Kroshko and Spiteri~\cite{kroshko2013efficient} implemented a SIMD-vectorized third-order stiff Rosenbrock integrator for atmospheric chemistry on the CBE finding a speedup of \SI{1.89}{$\times$} (a parallel scaling efficiency of \SI{0.94}{$\percent$}) over a serial version of the same code.
 Stone et al.,~\cite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
-The shallow-SIMD and SIMT-vectorization models were implemented in OpenCL and compared to an OpenMP baseline code that was deep SIMD-vectorized by simple compiler hints (i.e., \texttt{\#pragmas}).
-The shallow-SIMD vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance was only \SIrange{1.4}{1.6}{$\times$} faster than the OpenMP baseline due to thread-divergence concerns.
+SIMD-vectorization improved the integrator performance over an OpenMP baseline (which was vectorized by simple compiler hints (i.e., \texttt{\#pragmas}) by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance was only \SIrange{1.4}{1.6}{$\times$} faster than the OpenMP baseline due to thread-divergence concerns.
 
@@
 \end{itemize}
-To our knowledge, this is the first work that provides vectorized source term and Jacobian evaluation for any chemical model on the CPU, GPU and other accelerators.
+To our knowledge, this is the first effort that enables vectorized chemical-kinetic source-term and Jacobian evaluation for any chemical model on the CPU, GPU and other accelerators.
 
@@
 \Cref{F:mem} depicts an example data array with K-rows and N-columns and index ($i$, $j$) corresponding to the $i$th row and $j$th column.
-For example the concentration of species $j$ for the $i$-th thermo-chemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{state}$, the number of thermo-chemical states considered, and $1 \le j \le \ns/$, the number of species in the model.
+For example the concentration of species $j$ for the $i$-th thermo-chemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{state}$, the number of thermo-chemical states considered for evaluation, and $1 \le j \le \ns/$, the number of species in the model.
 The stored concentrations would then have $K = N_{\text{state}}$ rows, and $N = N_{\text{sp}}$ columns.
 
-In the ``C'' (C row-major) format, the concentrations of all species for a single thermo-chemical condition $i$ would be stored sequentially in memory, i.e., with  $[C]_{1, 1}$ in index \num{1}, $[C]_{1, 2}$ in index \num{2} etc.
-as shown in \cref{F:row_major}.
+In the ``C'' (C row-major) format, the concentrations of all species for a single thermo-chemical condition $i$ would be stored sequentially in memory, i.e., with  $[C]_{1, 1}$ in index \num{1}, $[C]_{1, 2}$ in index \num{2} etc., as shown in \cref{F:row_major}.
 Conversely in the ``F'' (Fortran column-major) format, the concentrations of a single species $j$ over all thermo-chemical states are adjacent in memory, corresponding to storing $[C]_{1, 1}$ in index \num{1}, $[C]_{2, 1}$ in index \num{2} and so on, as shown in \cref{F:column_major}.
@@
 In a \textit{shallow}-SIMD\slash SIMT vectorization (also referred to as ``per-thread'' in previous works~\cite{Stone:2013aa} using GPUs), each SIMD-lane (or SIMT-thread) in a compute-unit evaluates thermo-chemical source terms\slash chemical kinetic Jacobian for a different thermo-chemical state.
-If the data is stored in ``F''-order, the SIMD-lanes\slash SIMT-thread accessing $[C]_{1, j}\ldots[C]_{L, j}$, the concentration of species $j$ for states $1, 2,\ldots L$---the size of the vector lane for SIMD or the number of threads in a SIMT warp---will load sequential locations in memory.
-The first $(j+1)$th species concentration however, $[C]_{1, j+1}$, will be $N_{\text{state}}$ memory locations away; this increases the likelihood of cache-misses on the CPU~\cite{gray2000rules}, but conversely is well suited to coalesced memory accesses on the GPU~\cite{NVIDIA:2018}.
+If the data is stored in ``F''-order, the SIMD-lanes\slash SIMT-thread accessing $[C]_{1, j}\ldots[C]_{N_v, j}$---the concentration of species $j$ for states $1, 2,\ldots N_v$, where $N_v$ is the size of the vector lane for SIMD or the number of threads in a SIMT warp---will load sequential locations in memory.
+The first $(j+1)$th species concentration, $[C]_{1, j+1}$, will be $N_{\text{state}}$ memory locations away; this increases the likelihood of cache-misses on the CPU~\cite{gray2000rules}, but conversely is well suited to coalesced memory accesses on the GPU~\cite{NVIDIA:2018}.
 
-In a \textit{deep}-SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works), a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermo-chemical source terms for a single thermo-chemical state, thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + L}$---the species concentrations for state 1, species $j \ldots j + L$---will access sequential memory locations if the data is stored in ``C''-order.
-Further, in the ``C''-ordering the furthest difference between any two species concentrations within the same thermo-chemical state is at most $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU, but leads to uncoalesced memory-accesses on the GPU.
-However, a deep vectorization requires both synchronization between SIMD-lanes\slash SIMT-threads, and may result in SIMD-waste\slash SIMT thread-divergence, caused by different SIMD-lanes\slash SIMT-threads executing different instructions (e.g.
-resulting from different if\slash then branches)
+In a \textit{deep}-SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works), a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermo-chemical source terms for a single thermo-chemical state; thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + N_v}$---the species concentrations for state 1, species $j \ldots j + L$---will access sequential memory locations if the data is stored in ``C''-order.
+Further, in the ``C''-ordering the furthest difference between any two species concentrations within the same thermo-chemical state is at most $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU, but may lead to uncoalesced memory-accesses on the GPU.
+In addition, a deep vectorization requires both synchronization between SIMD-lanes\slash SIMT-threads via memory-fences\slash barriers (an expensive operation), and may result in SIMD-waste\slash SIMT thread-divergence, caused by different SIMD-lanes\slash SIMT-threads executing different instructions (e.g., resulting from different if\slash then branches)
 
@@
 Finally,~\cref{F:vector_data} shows a vectorized data-ordering that improves the caching patterns of a shallow, ``C''-ordered SIMD vectorization on the CPU (\cref{F:row_major_split}) and a deep, ``F''-ordered SIMT vectorization on the GPU (\cref{F:column_major_split}).
-This is accomplished by splitting the slower-varying axis of the data array---rows for ``C''-ordering, and columns for ``F''-ordering---into chunks of size $L$ (the SIMD vector-width, or SIMT warp-size) and laying these data out contigously in memory.
-For example, using the data-ordering in~\cref{F:row_major_split}, the concentration of species $j$ for states $i\ldots i+L$, $[C]_{i, j} \ldots [C]_{i + L, j}$, are contiguous in memory and are followed by the concentration of species $j + 1$ for the same states, $[C]_{i, j + 1} \ldots [C]_{i + L, j + 1}$.
+This is accomplished by splitting the slower-varying axis of the data array---columns for ``C''-ordering, and rows for ``F''-ordering---into chunks of size $N_v$ (the SIMD vector-width, or SIMT warp-size) and laying these data out contiguously in memory.
+For example, using the data-ordering in~\cref{F:row_major_split}, the concentration of species $j$ for states $i\ldots i+N_v$, $[C]_{i, j} \ldots [C]_{i + N_v, j}$, are contiguous in memory and are followed by the concentration of species $j + 1$ for the same states, $[C]_{i, j + 1} \ldots [C]_{i + N_v, j + 1}$.
 This pattern, similar to OpenCL's native vector data-types---e.g., \texttt{double\num{8}} which treats \num{8} contiguous double precision floating point numbers as a single vector datum---ensures that any SIMD operation occurs on data that is contiguous in memory, greatly improving caching and SIMD-throughput.
@@
 This new version of \texttt{pyJac} is capable of evaluating the thermo-chemical source terms for using both the constant-pressure (\conp/) or constant-volume (\conv/) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
-In this section, we will only outline a brief summary of the system evaluated by \texttt{pyJac}; the interested reader is referred to our complete derivations, attached as supplemental material to this document.
+In this section, we will only outline a brief summary of the system evaluated by \texttt{pyJac}; the interested reader is referred to the complete derivations, attached as supplemental material to this document.
 
-The thermo-chemical state vector derivation consists of the temperature, the non-constant thermodynamic state parameter (volume or pressure) and the number of moles of all species except the last in the chemical model which is typically taken to be the bath-gas (e.g., \ce{N2}):
+The thermo-chemical state vector derivation consists of the temperature, a non-constant thermodynamic state parameter (volume or pressure for \conp/ and \conv/ respectively), and the number of moles of all species except the last species in the chemical model which is typically taken to be the bath-gas (e.g., \ce{N2}):
 \loadeq{state}
@@
 First, the state vector results in highly sparse chemical kinetic Jacobians as will be detailed in~\cref{S:sparsity}.
-Second, mass is explicitly conserved in this formulation owing to the fact that the moles of the final species is calculated from the ideal gas law, and the rate of change of the final species from conservation of mass (see \cref{deriv-s:state,deriv-s:source} of the supplemental material respectively); i.e., system is not over-constrained.
+Second, mass is explicitly conserved in this formulation owing to the fact that the moles of the final species is calculated from the ideal gas law, and the rate of change of the final species from conservation of mass (see \cref{deriv-s:state,deriv-s:source} of the supplemental material respectively); additionally, system is not over-constrained and does not require use of a more complicated DAE solver (as compared to an regular ODE solver) for integration\todo{hansen citation here}.
 Finally, the chemical kinetic Jacobian changes relatively little between the \conp/ and \conv/ forms, making maintaining the code-base much simpler.
-Although most current combustion codes do not use species moles as a state variable, to\slash from more common mass\slash mole-fractions and moles is straightforward, and once inside integration of an chemical kinetic initial value problem, the choice of variables no longer matter.
+Although most current combustion codes do not use species moles as a state variable, conversion to\slash from more common mass\slash mole-fractions and moles is straightforward, and once inside integration of an chemical kinetic initial value problem (IVP), the choice of variables no longer matter.
 
@@
 \end{equation}
-where $\nu_{k, i}$ is the net stoichiometic coefficient of species $k$ in reaction $i$, $R_{i}$ the net rate of progress of reaction $i$ (described in~\cref{deriv-s:rop_net} of the supplemental), and $c_{i}$ the pressure modification term i.e., for third-body (see~\cref{deriv-s:thdbody} of the supplimental), or fall-off\slash chemically-activated reactions (\cref{deriv-s:falloff} of the same).
+with $\nu_{k, i}$ the net stoichiometic coefficient of species $k$ in reaction $i$, $R_{i}$ the net rate of progress of reaction $i$ (described in~\cref{deriv-s:rop_net} of the supplemental), and $c_{i}$ the pressure modification term i.e., for third-body (see~\cref{deriv-s:thdbody} of the supplemental), or fall-off\slash chemically-activated reactions (\cref{deriv-s:falloff} of the same).
 
@@
 
-The molar source terms are more complicated due to the derivatives of the net rate of progress---see~\cref{deriv-s:dropi}---and pressure modification terms---\cref{deriv-s:dci}---for the reactions.
-The reader is again encouraged to read the supplimental material covering the exact derivations and forms of the reactions solved by \texttt{pyJac}.
+The molar source terms are more complicated due to the derivatives of the net rate of progress---see~\cref{deriv-s:dropi}---and pressure modification terms---\cref{deriv-s:dci}---of the reactions.
+The reader is again encouraged to read the supplemental material covering the exact derivations and forms of the reactions solved by \texttt{pyJac}.
 The derivative of the molar source term with respect to temperature is:
@@
 Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates user specified psuedo-code and data to OpenCL\slash C-code.
-This allows for easy changes of program structure, e.g.
-data ordering, vectorization, threading patterns and easy switching of target-languages (and simpler extension to additional languages, e.g., CUDA).
-Additionally, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, and via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source terms.
+This allows for easy changes of program structure, e.g. data ordering, vectorization, threading patterns and switching of target-languages (and simpler extension to additional languages, e.g., CUDA).
+Additionally, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source terms.
 The source-terms and subcomponents thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{adept-v11} is used to obtain reference answers for Jacobian subcomponents.
@@
 Additionally, \texttt{v5.0} of the LLVM\slash\texttt{clang}~\cite{Lattner:2004:LCF:977395.977673} compiler chain was installed on all machines to enable use of POCL.
-In addition,~\cref{t:gpus} lists the model, number of CUDA cores and NVIDIA driver of each of the GPUs used in this work.
+\Cref{t:gpus} lists the model, number of CUDA cores and NVIDIA driver of each of the GPUs used in this work.
 NVIDIA's OpenCL runtime is bundled with the NVIDIA driver~\cite{NVIDIA:2018}, hence we use driver version to specify the OpenCL runtime version.
+
 Finally,~\cref{t:platforms} lists the platforms and vectorization\slash execution patterns they are capable of running.
@@
 On the other hand, POCL is an open source OpenCL runtime that works on all CPU types tested in this work, and does implement these atomic operations; it is our experience that the POCL community is very responsive to bug reports and user outreach.
-However, POCL's implicit vectorization module---which utilizes the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to attempt to translate OpenCL code to vectorized machine code---typically fails to achieve much speedup, if any.
+However, POCL's implicit vectorization module---which utilizes the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to translate OpenCL code to vectorized machine code---typically fails to achieve much speedup, if any.
 Thus POCL is a particularly useful tool for validation, but not necessarily for performance studies.
-Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the only significant code change is building the wrapper that initializes\slash transfers memory and calls the kernel.
 This discussion will be expanded upon in~\cref{s:future} to highlight the most promising future directions.
@@
 
+Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the only significant code change is building the wrapper that initializes\slash transfers memory and calls the kernel.
 Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py}) is a key priority going forwards.

### 2018-03-05 paper.tex
edits & allow per-platform error reporting

@@
 Recognizing the need to accelerate Jacobian evaluation and factorization, a number of recent works have been published on analytical chemical kinetic Jacobian evaluation; although as will be discussed at the end of this section, this work offers several key improvements over past efforts.
-The \texttt{TChem} software package~\cite{Safta:2011vn} was one of the first packages developed providing this functionality, but has several limitations including: incompatibilities with modern reaction types (i.e., pressure dependent Arrhenius and Chebyshev reactions), and lack of thread-safety to enable parallel execution~\cite{Curtis2017:tchem}.
+The \texttt{TChem} software package~\cite{Safta:2011vn} was one of the first packages developed providing this functionality, but has several limitations including: incompatibilities with modern reaction types---i.e., pressure dependent Arrhenius (or P-Log) and Chebyshev reactions---and lack of thread-safety to enable parallel execution~\cite{Curtis2017:tchem}.
 Youssefi~\cite{Youssefi:2011tm} explored the importance of analytical Jacobian matrices for time-scale analysis techniques as well as their effect on computational efficiency in zero-dimensional homogeneous reactor simulations.
@@
 \end{tabular}
-\caption{Summary of rate of progress, species and temperature rate correctness.
+\caption{Summary of rate of progress, species, temperature and thermodynamic state parameter rate correctness.
 Error statistics are based on the infinity-norm of the relative error detailed in Eq.~\eqref{e:rel_err} for each quantity.
@@
 As in our previous work~\cite{Niemeyer:2016aa}, the correctness of the Jacobian codes generated by \texttt{pyJac} was determined by comparison to the Jacobian obtained by automatic differentiation of the source rate subroutines generated by \texttt{pyJac} via the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
-The reasons for this choice are explained fully in~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with minimal additional programming little effort.
+The reasons for this choice are explained fully in~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with minimal additional programming effort.
 The discrepancy between the analytical and automatic differentiated Jacobians for thermo-chemical state $k$, denoted by $\mathcal{J}_k$ and $\hat{\mathcal{J}}_k$ respectively, is determined by the relative error Frobenius norm over all Jacobian indicies $i, j$:
@@
 \begin{table}[htbp]
+\sisetup{retain-zero-exponent=true}
 \centering

### 2018-03-05 scripts/jac_error.py

@@ -45,4 +45,6 @@ def printer(err_dict):
     print('mean_threshold: {:.15e}'.format(mean_thr))
 
 
+
+print_error('jac', update, printer, True)
 print_error('jac', update, printer)

### 2018-03-05 scripts/load_error_files.py

@@ -41,9 +41,22 @@ def add_rxn_specific_inds(mech_info, rxn_type, rxn_str, path):
         if not isinstance(gas.reaction(fi), rxn_type)])
 
 
-def run_error_calcs(ftype, updater):
+def get_platform(file):
+    if 'nvidia' in file:
+        return 'nvidia'
+    elif 'portable' in file:
+        return 'pocl'
+    elif 'intel' in file:
+        return 'intel'
+    else:
+        assert '_c_' in file
+        return 'openmp'
+
+
+def run_error_calcs(ftype, updater, per_platform=False):
     files = get_error_files(ftype)
     err_dicts = {}
+    platforms = set()
     for mech in files:
         mech_name = os.path.basename(os.path.normpath(mech))
         err_dicts[mech_name] = defaultdict(lambda: 0)
@@ -60,15 +73,33 @@ def run_error_calcs(ftype, updater):
                                       mech)
 
         for file in files[mech]:
+            if per_platform:
+                platform = get_platform(file)
+                platforms.add(platform)
+                if platform not in err_dicts[mech_name]:
+                    err_dicts[mech_name][platform] = defaultdict(lambda: 0)
+                updater(err_dicts[mech_name][platform], np.load(file), filename=file,
+                    mech_info=mech_info)
+                continue
+
             updater(err_dicts[mech_name], np.load(file), filename=file,
                     mech_info=mech_info)
 
+    if per_platform:
+        return platforms, err_dicts
     return err_dicts
 
 
-def print_error(ftype, updater, printer):
-    err_dicts = run_error_calcs(ftype, updater)
+def print_error(ftype, updater, printer, per_platform=False):
+    err_dicts = run_error_calcs(ftype, updater, per_platform=per_platform)
+    if per_platform:
+        platforms, err_dicts = err_dicts
     mech_arr = ['H2', 'CH4', 'C2H4', 'IC5H11OH']
     for mech in mech_arr:
         print(mech)
+        if per_platform:
+            for platform in err_dicts:
+                print(platform)
+                printer(err_dicts[mech][platform])
+                continue
         printer(err_dicts[mech])

### 2018-03-06 scripts/jac_error.py
update

@@ -38,7 +38,6 @@ def printer(err_dict):
             mean_thr = (mean_thr[0] + err_dict[name], mean_thr[1] + 1)
     # print mean threshold
     mean_thr = mean_thr[0] / mean_thr[1]
-
     print('\n'.join(
         ['{}: {:.3e}'.format(k, v) for k, v in sorted(err_dict.items())
          if 'thresholded' in k]))

### 2018-03-06 scripts/load_error_files.py

@@ -98,8 +98,8 @@ def print_error(ftype, updater, printer, per_platform=False):
     for mech in mech_arr:
         print(mech)
         if per_platform:
-            for platform in err_dicts:
+            for platform in err_dicts[mech]:
                 print(platform)
                 printer(err_dicts[mech][platform])
-                continue
+            continue
         printer(err_dicts[mech])

### 2018-03-07 paper.tex
revise; move opencl neg to end, rework Jac error discussion, oether edits

@@
 \subsection{Code Generation and Testing Infrastructure}
+\label{s:unittest}
 Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates user specified psuedo-code and data to OpenCL\slash C-code.
@@
 \subsection{Testing platforms}
-
-The performance and validation studies for this work were run on a variety of CPU and GPU platforms.
-\Cref{t:cpus} shows the number of cores, vector instruction set and model of the CPUs used in this work; each CPU had both \texttt{v16.1.1} of the Intel OpenCL runtime~\cite{intelopencl:2018} and \texttt{v1.0} of the POCL~\cite{poclIJPP} runtime installed, each enabling OpenCL \texttt{v1.2} execution.
-Additionally, \texttt{v5.0} of the LLVM\slash\texttt{clang}~\cite{Lattner:2004:LCF:977395.977673} compiler chain was installed on all machines to enable use of POCL.
-\Cref{t:gpus} lists the model, number of CUDA cores and NVIDIA driver of each of the GPUs used in this work.
-NVIDIA's OpenCL runtime is bundled with the NVIDIA driver~\cite{NVIDIA:2018}, hence we use driver version to specify the OpenCL runtime version.
-
-Finally,~\cref{t:platforms} lists the platforms and vectorization\slash execution patterns they are capable of running.
-The Intel and NVIDIA OpenCL runtimes lack implementations of atomic operations on double precision variables; these are currently necessary for \texttt{pyJac} to run deep-vectorized code.
-On the other hand, POCL is an open source OpenCL runtime that works on all CPU types tested in this work, and does implement these atomic operations; it is our experience that the POCL community is very responsive to bug reports and user outreach.
-However, POCL's implicit vectorization module---which utilizes the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to translate OpenCL code to vectorized machine code---typically fails to achieve much speedup, if any.
-Thus POCL is a particularly useful tool for validation, but not necessarily for performance studies.
-This discussion will be expanded upon in~\cref{s:future} to highlight the most promising future directions.
 
@@
 
+The performance and validation studies for this work were run on a variety of CPU and GPU platforms.
+\Cref{t:cpus} shows the number of cores, vector instruction set and model of the CPUs used in this work; each CPU had both \texttt{v16.1.1} of the Intel OpenCL runtime~\cite{intelopencl:2018} and \texttt{v1.0} of the POCL~\cite{poclIJPP} runtime installed, each enabling OpenCL \texttt{v1.2} execution.
+Additionally, \texttt{v5.0} of the LLVM\slash\texttt{clang}~\cite{Lattner:2004:LCF:977395.977673} compiler chain was installed on all machines to enable use of POCL.
+\Cref{t:gpus} lists the model, number of CUDA cores and NVIDIA driver of each of the GPUs used in this work.
+NVIDIA's OpenCL runtime is bundled with the NVIDIA driver~\cite{NVIDIA:2018}, hence we use driver version to specify the OpenCL runtime version.
+
 \begin{table}[htb]
@@
 
+\Cref{t:platforms} lists the platforms and vectorization\slash execution patterns they are capable of running.
+The Intel and NVIDIA OpenCL runtimes lack implementations of atomic operations on double precision variables; these are currently necessary for \texttt{pyJac} to run deep-vectorized code.
+On the other hand, POCL is an open source OpenCL runtime that works on all CPU types tested in this work, and does implement these atomic operations; it is our experience that the POCL community is very responsive to bug reports and user outreach.
+However, POCL's implicit vectorization module---which utilizes the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to translate OpenCL code to vectorized machine code---typically fails to achieve much speedup, if any.
+Thus POCL is a particularly useful tool for validation, but not necessarily for performance studies.
+This discussion will be expanded upon in~\cref{s:future} to highlight the most promising future directions.
+
 \begin{table}[htb]
@@
 \label{t:platforms}
+\end{table}
+
+Finally, \cref{t:models} displays the chemical kinetic models used in this work, as well as number of partially stirred reaction conditions (PaSR) in the condition database for each.
+Creation of the PaSR databases are described in detail our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
+
+\begin{table}[htb]
+\centering
+\begin{tabular}{@{}l c c @{}}
+\toprule
+Model &  Number of Conditions & Reference \\
+\midrule
+\ce{H2}\slash~\ce{CO} & \num{900900} & \cite{Burke:2011fh} \\
+GRI-Mech.~3.0         & \num{450900} & \cite{smith_gri-mech_30} \\
+USC-Mech II           & \num{91800}  & \cite{Wang:2007} \\
+\ce{iC5H11OH}         & \num{450900} & \cite{Sarathy:2013jr} \\
+\bottomrule
+\end{tabular}
+\caption{The chemical kinetic models used in this study and number of conditions in the partially stirred reactor database for each.}
+\label{t:models}
 \end{table}
@@
 \toprule
-\multicolumn{1}{l}{Model} & \multicolumn{1}{l}{\ce{H2}\slash~\ce{CO}\cite{Burke:2011fh}} & \multicolumn{1}{l}{GRI-Mech.~3.0\cite{smith_gri-mech_30}} & \multicolumn{1}{l}{USC-Mech II\cite{Wang:2007}} & \multicolumn{1}{l}{\ce{iC5H11OH}\cite{Sarathy:2013jr}} \\
+\multicolumn{1}{l}{Model} & \multicolumn{1}{l}{\ce{H2}\slash~\ce{CO}} & \multicolumn{1}{l}{GRI-Mech.~3.0} & \multicolumn{1}{l}{USC-Mech II} & \multicolumn{1}{l}{\ce{iC5H11OH}} \\
 \midrule
@@
 
-In Table~\ref{T:source_error}, we see the results of this code as compared to Cantera~\cite{Cantera} on a library of partially stirred reaction conditions (PaSR) described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
+In Table~\ref{T:source_error}, we see the results of this code as compared to Cantera~\cite{Cantera} on the library of PaSR conditions (\cref{t:models}).
 Very close agreement is observed for the forward and reverse ROP for all models, however the error norm is \textasciitilde\numrange{3}{4} orders of magnitude larger for the isopentanol model.
@@
  \label{e:jac_error_base}
- E_{\text{rel}, k} = \frac{\left\lVert \hat{\mathcal{J}}_{ijk} - \mathcal{J}_{ijk} \right\rVert_{i,j,F}}{\left\lVert \hat{\mathcal{J}}_{ijk} \right\rVert_{i,j,F}}
+ E_{\text{rel}, k} = \left\lVert \frac{\hat{\mathcal{J}}_{ijk} - \mathcal{J}_{ijk}}{\hat{\mathcal{J}}_{ijk}} \right\rVert_{i,j,F}
 \end{equation}
@@
 \end{equation}
-The error statistics reported in this section are then based only on matrix elements where $\mathcal{J}_{ijk} \ge \frac{\mathcal{T}}{\mathcal{C}}$; this filtered form of~\cref{e:jac_error_base} is denoted $E_{\mathcal{C}}$.
+The error statistics reported in this section are then based only on matrix elements where $\mathcal{J}_{ijk} \ge \frac{\mathcal{T}}{\mathcal{C}}$; this filtered form of~\cref{e:jac_error_base} is denoted $E_{\mathcal{C},k}$.
+Finally, the Frobenius norm is calculated over all the states in the PaSR thermo-chemical condition database:
+\begin{equation}
+ \label{e:thresholded_error}
+ E_{\mathcal{C}} = \left\lVert E_{\mathcal{C},k} \right\rVert_{k, F}
+\end{equation}
+
+It is noted that this error norm is quite different from the relative error Frobenius norm suggested by Anderson et al.~\cite{Anderson:1999aa} for quantifying the error of matrices in LAPACK:
+\begin{align}
+ \label{e:error_lapack}
+ \lhs{lapack}{E_{\mathcal{L}, k}} &=  \rhs{lapack}{\frac{\left\lVert \hat{\mathcal{J}}_{ijk} - \mathcal{J}_{ijk} \right\rVert_{i,j,F}}{\left\lVert \hat{\mathcal{J}}_{ijk} \right\rVert_{i,j,F}}} \nonumber \\
+ \lhs{lapack}{E_{\mathcal{L}}} &= \rhs{lapack}{\left\lVert  E_{\mathcal{L}, k} \right\rVert_{k, F}}[c]
+\end{align}
+In our experience, the LAPACK error norm is often dominated by the accuracy of the larger elements in the Jacobian, while our filtered error norm (while significantly more stringent) is useful for identifying errors in smaller elements.
+For reference, both our own error norm and the LAPACK error norm will be reported.
 
@@
 \centering
-\begin{tabular}{@{}l S S[table-format=1.3e1] S[table-format=1.3e1] S[table-format=1.3e1]@{}}
+\begin{tabular}{@{}l S[table-format=.0] S[table-format=1.3e1] S[table-format=1.3e1] S[table-format=1.3e1] S[table-format=1.3e1] @{}}
 \toprule
-Model                 & \multicolumn{1}{c}{Sample size}   & \multicolumn{1}{c}{$\bar{\mathcal{T}}$}  & \multicolumn{1}{c}{$E_{10^{20}}$}   & \multicolumn{1}{c}{$E_{10^{15}}$} \\
+Model                 & \multicolumn{1}{c}{$E_{\mathcal{L}}$} & \multicolumn{1}{c}{$\bar{\mathcal{T}}$} & \multicolumn{1}{c}{$E_{10^{20}}$}   & \multicolumn{1}{c}{$E_{10^{15}}$} \\
 \midrule
-\ce{H2}\slash \ce{CO} & \num{900900}                      & \num{6.431e+18}      & \num{1.742e00}  & \num{4.508e-5} \\
-GRI-Mech 3.0          & \num{450900}                      & \num{7.783e+19}      & \num{3.842e-7}  & \num{3.687e-7} \\
-USC-Mech II           & \num{91800}                       & \num{2.830e+21}      & \num{1.119e-2}  & \num{1.983e-7} \\
-\ce{iC5H11OH}         & \num{450900}                      & \num{2.733e+26}      & \num{1.363e-3}  & \num{2.764e-5} \\
+\ce{H2}\slash \ce{CO} & \num{1.048e-11}      & \num{6.431e+18}      & \num{1.741e+0}  & \num{4.508e-5} \\
+GRI-Mech 3.0          & \num{1.124e-14}      & \num{7.783e+19}      & \num{3.842e-7}  & \num{3.687e-7} \\
+USC-Mech II           & \num{9.632e-15}      & \num{2.830e+21}      & \num{1.119e-2}  & \num{1.983e-7} \\
+\ce{iC5H11OH}         & \num{1.227e-10}      & \num{2.733e+26}      & \num{1.363e-3}  & \num{2.764e-5} \\
 \bottomrule
@@
 
-\Cref{T:error} presents the error statistics of \texttt{pyJac} over the tested PaSR thermo-chemical condition database.
-For all tested cases, the maximum filtered error norm is less than unity.
-The filtered error norm is largest for the \ce{H2}\slash\ce{CO} model with the more stringent tolerance ($\mathcal{C} = 10^{20}$).
-We note that $\mathcal{T}$ for this model is smaller than the tolerance in this case, and hence we are comparing numbers smaller than $\mathcal{O}(1)$; relaxing the tolerance to $\mathcal{C} = 10^{15}$ significantly lowers the filtered error norm to $\mathcal{O}(10^{-5})$.
-Similar error characteristics are seen for all models except GRI-Mech 3.0---in which the error norms for both tolerances are $\sim\mathcal{O}(10^{-7})$---that is a larger error norm for the more stringent tolerance that drops 2--4 orders of magnitude for the relaxed tolerance.
+\Cref{T:error} reports the maximum $E_{\mathcal{C}}$ and $E_{\mathcal{L}}$ values over all test platforms\slash vectorization patterns (see~\cref{t:platforms}) and the sparse and dense Jacobians (see~\cref{S:sparsity}).
+The most stringent filtered error norm ($\mathcal{C} = 10^{20}$) ranges from \numrange[retain-zero-exponent=true]{e-7}{e0}; the largest error is for the \ce{H2}\slash\ce{CO} model.
+For this model, $\mathcal{T}$  is smaller than the tolerance of \num{e20}, and hence the error norm considers smaller than $\mathcal{O}(1)$.
+It is noted that for GRI-Mech 3.0, which has a $\mathcal{T}$ roughly an order of magnitude larger, the stringent error norm is significantly smaller: $\mathcal{O}(\num{e-7})$.
+Given the intricacy of floating point error evaluation (see~\cref{s:validation}), the use of different languages and OpenCL platforms (the effect of these differences will be explored in Appendix \textbf{???}\todo{this}), and the general complexity of \texttt{pyJac} it would be exceedingly difficult to pinpoint an exact cause for this phenomena.
+Indeed, the stringent filtered error norm was chosen specifically for its ability to detect errors in relatively smaller Jacobian entries.
+To ensure there is no bug or error present in \texttt{pyJac}, a relaxed filtered error norm ($\mathcal{C} = 10^{15}$) is also presented for each model in~\cref{T:error}.
+This relaxed norm is smaller by \numrange{2}{5} orders of magnitude for all models---except GRI-Mech 3.0, where the stringent case has very small error as previously discussed---indicating that accuracy is higher when the smaller Jacobian entries are excluded.
+
+The relative LAPACK error norm---ranging from \textasciitilde\numrange{e-11}{e-14}---re-enforces this finding, as it indicates roughly \numrange{11}{14} digits of accuracy~\cite{Anderson:1999aa}.
+Interestingly, the LAPACK error norm does not correlate well with the more stringent filtered error norm, e.g., USC-Mech II has the lowest LAPACK error norm (\num{9.632e-15}), but the second largest stringent filtered error norm (\num{1.119e-2}).
+Conversely, the LAPACK error norm does seem to correlate with the relaxed filtered error norm; the two models with the largest LAPACK error, \ce{H2}\slash\ce{CO} and \ce{iC5H11OH}, have the largest relaxed filtered error norms ($\mathcal{O}(\num{e-5})$).
+This study, along with the individual unit-testing of all chemical source terms and Jacobian sub-components described in~\cref{s:unittest}, gives us high confidence in the correctness of this new version of \texttt{pyJac}.
 
@@
 In~\cref{F:intel_source_nonorm}, the mean runtime per initial condition is shown for both the~\avx/ and~\sse/ CPUs, using both Intel OpenCL and OpenMP.
-This normalization of the runtime by the number of initial conditions tested is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{T:error}).
+This normalization of the runtime by the number of initial conditions tested is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{t:models}).
 For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the unvectorized (strictly parallely) Intel OpenCL code is slightly faster than OpenMP in all cases.\todo{rerun OpenMP w/ mtune=native}

### 2018-03-07 scripts/jac_error.py

@@ -7,7 +7,8 @@ atol = 1e-10
 
 def update(err_dict, err, filename=None, **kwargs):
     for name in err:
-        if ('amax' in name or 'value' in name) and name != 'jac_threshold_value':
+        if ('amax' in name or 'value' in name) and \
+                name != 'jac_threshold_value':
             continue
         assert 'fdjac' not in filename
         if err_dict['jac_thresholded_15'] > 1e0:
@@ -41,8 +42,8 @@ def printer(err_dict):
     print('\n'.join(
         ['{}: {:.3e}'.format(k, v) for k, v in sorted(err_dict.items())
          if 'thresholded' in k]))
+    print('jac_lapack', err_dict['jac_lapack'])
     print('mean_threshold: {:.15e}'.format(mean_thr))
-
 
 
 print_error('jac', update, printer, True)

### 2018-03-07 paper.tex
revise error discussion

@@
 
-As in our previous work~\cite{Niemeyer:2016aa}, the correctness of the Jacobian codes generated by \texttt{pyJac} was determined by comparison to the Jacobian obtained by automatic differentiation of the source rate subroutines generated by \texttt{pyJac} via the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
+As in our previous work~\cite{Niemeyer:2016aa}, the correctness of the Jacobian was determined by comparison to the Jacobian obtained by automatic differentiation of the source rate subroutines generated by \texttt{pyJac} via the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
 The reasons for this choice are explained fully in~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with minimal additional programming effort.
@@
 
-To avoid large relative errors of small nonzero Jacobian elements due to accumulation of floating point error, the Frobenius norm of the automatically differentiated Jacobian is calculated over all thermo-chemical states $k$:
+To avoid large relative errors of very small nonzero Jacobian elements due to accumulation of floating point error, the Frobenius norm of the automatically differentiated Jacobian is calculated over all thermo-chemical states $k$:
 \begin{equation}
@@
 \end{align}
-In our experience, the LAPACK error norm is often dominated by the accuracy of the larger elements in the Jacobian, while our filtered error norm (while significantly more stringent) is useful for identifying errors in smaller elements.
+In our experience, the LAPACK error norm is often dominated by the accuracy of the larger elements in the Jacobian, while the filtered error norm (although significantly more stringent) is useful for identifying errors in both large and small Jacobian entries.
+Further, with a tunable threshold parameter, $\mathcal{C}$, the error of different ranges of of element sizes can be assessed, and the effects of floating point error can be isolated.
 For reference, both our own error norm and the LAPACK error norm will be reported.
@@
 \toprule
-Model                 & \multicolumn{1}{c}{$E_{\mathcal{L}}$} & \multicolumn{1}{c}{$\bar{\mathcal{T}}$} & \multicolumn{1}{c}{$E_{10^{20}}$}   & \multicolumn{1}{c}{$E_{10^{15}}$} \\
+Model                 & \multicolumn{1}{c}{$E_{\mathcal{L}}$} & \multicolumn{1}{c}{$\bar{\mathcal{T}}$} & \multicolumn{1}{c}{$E_{\mathcal{C} = 10^{20}}$}   & \multicolumn{1}{c}{$E_{\mathcal{C} = 10^{15}}$} \\
 \midrule
@@
 \caption{Summary of Jacobian matrix validation results.
-Error statistics are the largest filtered relative error $E_\mathcal{C}$ for over all samples.
+Error statistics are the largest filtered relative error $E_\mathcal{C}$ for over all test platforms, vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/ and sparse\slash dense Jacobians.
 It is noted that the threshold described in~\cref{e:thresh} varies slightly between the \conp/ and \conv/ cases; the reported $\bar{\mathcal{T}}$ is the average of the two, however the appropriate value was used during calculations of the error statistics.
@@
 
-\Cref{T:error} reports the maximum $E_{\mathcal{C}}$ and $E_{\mathcal{L}}$ values over all test platforms\slash vectorization patterns (see~\cref{t:platforms}) and the sparse and dense Jacobians (see~\cref{S:sparsity}).
+\Cref{T:error} reports the maximum $E_{\mathcal{C}}$ and $E_{\mathcal{L}}$ values over all test platforms\slash vectorization patterns (see~\cref{t:platforms}), sparse\slash dense (see~\cref{S:sparsity}) and \conp/\slash\conv/ Jacobians.
 The most stringent filtered error norm ($\mathcal{C} = 10^{20}$) ranges from \numrange[retain-zero-exponent=true]{e-7}{e0}; the largest error is for the \ce{H2}\slash\ce{CO} model.
@@
 It is noted that for GRI-Mech 3.0, which has a $\mathcal{T}$ roughly an order of magnitude larger, the stringent error norm is significantly smaller: $\mathcal{O}(\num{e-7})$.
-Given the intricacy of floating point error evaluation (see~\cref{s:validation}), the use of different languages and OpenCL platforms (the effect of these differences will be explored in Appendix \textbf{???}\todo{this}), and the general complexity of \texttt{pyJac} it would be exceedingly difficult to pinpoint an exact cause for this phenomena.
-Indeed, the stringent filtered error norm was chosen specifically for its ability to detect errors in relatively smaller Jacobian entries.
-To ensure there is no bug or error present in \texttt{pyJac}, a relaxed filtered error norm ($\mathcal{C} = 10^{15}$) is also presented for each model in~\cref{T:error}.
-This relaxed norm is smaller by \numrange{2}{5} orders of magnitude for all models---except GRI-Mech 3.0, where the stringent case has very small error as previously discussed---indicating that accuracy is higher when the smaller Jacobian entries are excluded.
+Given the intricacy of floating point error evaluation, the use of different languages and OpenCL platforms (the effect of these differences will be explored in Appendix \textbf{???}\todo{this}), and the general complexity of \texttt{pyJac} it would be exceedingly difficult to pinpoint an exact cause for this phenomena, as was done in~\cref{s:validation}.
+To ensure there is no bug or error present in the Jacobians generated by \texttt{pyJac}, a relaxed filtered error norm ($\mathcal{C} = 10^{15}$) is also presented for each model in~\cref{T:error}.
+This relaxed norm is smaller by \numrange{2}{5} orders of magnitude for all models---except GRI-Mech 3.0, where the stringent case has very small error as previously discussed---indicating that accuracy is higher when the smaller Jacobian entries are excluded; this suggests floating point error accumulation is the primary driver of the stringent filtered error norm.
 
-The relative LAPACK error norm---ranging from \textasciitilde\numrange{e-11}{e-14}---re-enforces this finding, as it indicates roughly \numrange{11}{14} digits of accuracy~\cite{Anderson:1999aa}.
-Interestingly, the LAPACK error norm does not correlate well with the more stringent filtered error norm, e.g., USC-Mech II has the lowest LAPACK error norm (\num{9.632e-15}), but the second largest stringent filtered error norm (\num{1.119e-2}).
-Conversely, the LAPACK error norm does seem to correlate with the relaxed filtered error norm; the two models with the largest LAPACK error, \ce{H2}\slash\ce{CO} and \ce{iC5H11OH}, have the largest relaxed filtered error norms ($\mathcal{O}(\num{e-5})$).
-This study, along with the individual unit-testing of all chemical source terms and Jacobian sub-components described in~\cref{s:unittest}, gives us high confidence in the correctness of this new version of \texttt{pyJac}.
+The relative LAPACK error norm---ranging from \textasciitilde\numrange{e-10}{e-14}---re-enforces this finding, as it indicates roughly \numrange{10}{14} digits of accuracy~\cite{Anderson:1999aa}.
+Further, the LAPACK error norm does not correlate well with the stringent filtered error norm, e.g., USC-Mech II has the lowest LAPACK error norm (\num{9.632e-15}), but the second largest stringent filtered error norm (\num{1.119e-2}).
+Conversely, the LAPACK error norm does seem to correlate with the relaxed filtered error norm; the two models with the largest LAPACK error, \ce{H2}\slash\ce{CO} and \ce{iC5H11OH}, have the largest relaxed filtered error norms ($\mathcal{O}(\num{e-5})$), again suggesting that the stringent error norm is influenced by floating point error accumulation.
+These findings, along with the individual unit-testing of all chemical source terms and Jacobian sub-components described in~\cref{s:unittest}, gives high confidence in the correctness of this new version of \texttt{pyJac}.
 

### 2018-03-08 paper.tex
source performance edits

@@
 \end{align}
-In our experience, the LAPACK error norm is often dominated by the accuracy of the larger elements in the Jacobian, while the filtered error norm (although significantly more stringent) is useful for identifying errors in both large and small Jacobian entries.
+In our experience, the LAPACK error norm is often dominated by the accuracy of the larger elements in the Jacobian, while the filtered error norm is useful for identifying errors in both large and small Jacobian entries.
 Further, with a tunable threshold parameter, $\mathcal{C}$, the error of different ranges of of element sizes can be assessed, and the effects of floating point error can be isolated.
@@
 
-In general, the Jacobian matrices generated by \texttt{pyJac} are largely sparse with non-zero entries mainly corresponding to species that participate in the same reaction (or non-default efficiency third-body species in a reaction), and mostly dense rows \slash columns corresponding to the temperature and thermodynamic state parameter.
+In general, the Jacobian matrices generated by \texttt{pyJac} are largely sparse with non-zero entries corresponding mainly to species that participate in the same reaction (or non-default efficiency third-body species in a reaction), and mostly dense rows\slash columns corresponding to the temperature and thermodynamic state parameter.
 However, the explicit-mass conservation formulation of \texttt{pyJac} can introduce additional non-zero entries in two ways.
-First, if the last-species in the model (i.e., the bath-gas) participates directly in a reaction, the derivative of the forward or reverse rate of progress with respect to any other species is non-zero (see~\cref{deriv-s:dri_dnj}).
-Similarly, if the last-species has a third-body efficiency not equal to the default (unity), this will again force all the derivative of the pressure-modification term with respect to other species to be non-zero (see~\cref{deriv-s:dci_thd}).
+First, if the last-species in the model (i.e., the bath-gas) participates directly in a reaction, the derivative of the forward or reverse rate of progress with respect to any other species is non-zero (see~\cref{deriv-s:dri_dnj} of the supplemental).
+Similarly, if the last-species has a third-body efficiency not equal to the default (unity), this will again force all the derivative of the pressure-modification term with respect to other species to be non-zero (see~\cref{deriv-s:dci_thd} of the same).
 Either case will result in an undesirable fully dense Jacobian row for all species that have a non-zero net stoichiometic coefficient in such a reaction.
 
-However, \texttt{pyJac} allows the user to simply ignore these derivatives (via a command-line switch) to and avoid any negative adverse affects on the Jacobian sparsity.
+However, \texttt{pyJac} allows the user to simply ignore these derivatives (via a command-line switch) and avoid any negative adverse affects on the Jacobian sparsity.
 The rational behind this choice is that many common implicit integration techniques (e.g., CVODE~\cite{cvode:2.8.2}) used in solution of chemical kinetic initial value problems are formulated around the assumption that the supplied Jacobian is approximate; this allows the Jacobian and its LU factorization to be reused for multiple internal integration time-steps, accelerating the solution process.
@@
 In~\cref{F:jac_sparsity}, a graphical representation of the sparsity pattern for GRI-Mech 3.0 is displayed.
-In particular we note that~\cref{F:jac_sparse_approx} is has several rows that are no longer fully-dense, as result of its approximate form -- these rows correspond to species directly interacting with \ce{N2}, largely in GRI-Mech 3.0's \ce{NOx} chemistry reactions.
+In particular we note that~\cref{F:jac_sparse_approx} is has several rows that are no longer fully-dense, as result of its approximate form; these rows correspond to species directly interacting with \ce{N2}, largely in GRI-Mech 3.0's \ce{NOx} chemistry reactions.
 \cref{T:jac_sparsity} shows the density of the exact and approximate Jacobians for all mechanisms tested in this work.
-The smallest model, \ce{H2}\slash\ce{CO} is very dense with \textasciitilde\SI{70}{\percent} of the Jacobian entries non-zero however this drops to just \textasciitilde\SI{50}{\percent} for GRI-Mech 3.0, and continues to decrease to \textasciitilde\SI{27}{\percent} for USC-Mech II and just \textasciitilde\SI{10}{\percent} for the isopentanol model.
+The smallest model, \ce{H2}\slash\ce{CO} is very dense with \textasciitilde\SI{70}{\percent} of the Jacobian entries non-zero, however this drops to \textasciitilde\SI{50}{\percent} for GRI-Mech 3.0, and continues to decrease to \textasciitilde\SI{27}{\percent} for USC-Mech II and just \textasciitilde\SI{10}{\percent} for the isopentanol model.
 The approximate Jacobian assumption drops the density of Jacobian by \textasciitilde\SIrange{3}{7}{\percent} for all models.
@@
 Currently, \texttt{pyJac} can utilize two common sparse matrix storage formats~\cite{netlib_templates}, compressed row\slash column storage (CSR\slash CSC), used for ``C'' and ``F''-ordered data respectively.
-These formats contain three vectors, for convenience we will outline only the CSR format but the CSC format is very similar~\cite{netlib_templates}.
-An $N\times N$ CSC matrix is composed of three vectors, a value vector of length $NNZ$---the number of non-zero elements in the Jacobian---which stores the elements of the Jacobian, a row pointer vector---of length $N$---which stores the locations in the value vector that begin a row, and a column index vector---of length $NNZ$---which stores the column indicies of the elements in the value vector.
-
-The Jacobian access pattern used by \texttt{pyJac} is fairly irregular currently; for simplicity we will only discuss looping-structure of species derivatives calculations as these form the bulk of the computation, and have the most challenging Jacobian access patterns.
-In general, an outer loop iterates over all reactions of a certain type (e.g., falloff reactions) and calculates the relevant Jacobian subproducts (e.g., the derivative of the falloff pressure modification term with respect to a species $j$).
-Two inner loops then iterate over the species in a reaction, updating the Jacobian entries for these species as appropriate.
-This pattern leads to fairly easily vectorizable code, and efficient Jacobian evaluation as the bulk of the computation is dependent only on the reaction in question, as discussed in our previous work~\cite{Niemeyer:2016aa}.
-Generally, this means that a lookup operation is required to find the sparse Jacobian index for any pair of state variables; in some cases this can be avoided, e.g., the rows corresponding to derivatives of the temperature and thermodynamic state parameter source terms are fully dense in \texttt{pyJac}, and hence no lookup is necessary.
-This lookup operation is currently implemented as a simple for-loop, e.g., for a sparse lookup of a pair of indicies $(i, j)$ in a compressed row storage matrix, the lookup function searches the column index vector between the values $\texttt{row\_pointer}[i] \ldots \texttt{row\_pointer}[i + 1]$ for $j$, and returns the offset from $\texttt{row\_pointer}[i]$ (or $\num{-1}$ if not found).
-As will be explored in~\cref{S:jacobian_results}, this slows down sparse Jacobian evaluation, and might be improved---at the cost of increased constant data usage, a limitation for OpenCL---by a static mapping of the full Jacobian indicies to the sparse index (or some null value if the entry is empty).
-Additionally, this might be an excellent usage of OpenCL's Image memory type (similar to texture memory in CUDA terminology) and both of these sparse indexing techniques merit future investigation.
+Tor convenience we will outline only the CSR format but the CSC format is very similar~\cite{netlib_templates}.
+An $N\times N$ CSR matrix is composed of three vectors, a value vector of length $NNZ$---the number of non-zero elements in the Jacobian---which stores the elements of the Jacobian, a row pointer vector---of length $N$---which stores the locations in the value vector that begin a row, and a column index vector---of length $NNZ$---which stores the column indicies of the elements in the value vector.
 
@@
 
+The Jacobian access pattern used by \texttt{pyJac} is fairly irregular; for simplicity we will only discuss looping-structure of species derivatives calculations as these form the bulk of the computation, and have the most challenging Jacobian access patterns.
+In general, an outer loop iterates over all reactions of a certain type (e.g., falloff reactions) and calculates the relevant Jacobian subproducts---independent of any particular species---for the reaction (e.g., the the derivative of the falloff pressure modification term).
+Two inner loops then iterate over the species in a reaction, updating the Jacobian entries for these species as appropriate.
+This pattern leads to fairly easily vectorizable code, and efficient Jacobian evaluation as the bulk of the computation is dependent only on the reaction in question, as discussed in our previous work~\cite{Niemeyer:2016aa}.
+Generally, this means that a lookup operation is required to find the sparse Jacobian index for any pair of state variables; in some cases this can be avoided, e.g., the rows corresponding to derivatives of the temperature and thermodynamic state parameter source terms are fully dense in \texttt{pyJac}, and hence no lookup is necessary.
+This lookup operation is currently implemented as a simple for-loop, e.g., for a sparse lookup of a pair of indicies $(i, j)$ in a compressed row storage matrix, the lookup function searches the column index vector between the values $\texttt{row\_pointer}[i] \ldots \texttt{row\_pointer}[i + 1]$ for $j$, and returns the offset from $\texttt{row\_pointer}[i]$ (or $\num{-1}$ if not found).
+As will be explored in~\cref{S:jacobian_results}, this slows down sparse Jacobian evaluation, and might be improved---at the cost of increased constant data usage, a limitation for OpenCL---by a static mapping of the full Jacobian indicies to the sparse index (or some null value if the entry is empty).
+Additionally, this might be an excellent usage of OpenCL's Image memory type (similar to texture memory in CUDA terminology) and both of these sparse indexing techniques merit future investigation.
+
 \subsection{Performance}
@@
 Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
-The wrapping code---responsible for initializing\slash transfering memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms, and \texttt{gcc v4.8.5} on the \sse/\slash\gpuold/ machines.
-The optimization level ``\texttt{-O3 -mtune=native}''\todo[inline]{again, rerun OpenMP w/ this} was used, and no ``fast-math'' OpenCL optimizations were enabled.
-Additionally, the ``exact'' form of the Jacobian (as opposed the the ``appropriate'' form discussed in~\cref{S:sparsity}) was used in all cases.
-Finally, unless stated otherwise, the \conp/ assumption was used, along with a vector width of \num{8}\slash\num{128} for all CPU\slash GPU cases respectively; ``C''\slash ``F''-ordered data was used on the CPU\slash GPU, and the runtimes reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
-The effects of choice of vector-width on the CPU\slash GPU, as well as the differences between \conp/ and \conv/ evaluations will be explored in~\cref{S:source_results}, and briefly discussed in~\cref{S:jacobian_results}.
+The OpenMP Jacobian\slash source rate evaluation kernels, as well as the OpenMP\slash OpenCL wrapping code---responsible for initializing\slash transferring memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms, and \texttt{gcc v4.8.5} on the \sse/\slash\gpuold/ machines.
+The optimization level ``\texttt{-O3 -mtune=native}'' was used, and no ``fast-math'' OpenCL optimizations were enabled.
+Additionally, the exact form of the Jacobian (as opposed the the ``approximate'' form discussed in~\cref{S:sparsity}) was used in all cases.
+Finally, unless stated otherwise: the performance results utilized the \conp/ assumption, a vector width of \num{8}\slash\num{128} and ``C''\slash ``F''-ordered data for the CPU\slash GPU cases respectively; the runtimes reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
+The effects of choice of vector-width, data-ordering and differences between \conp/ and \conv/ evaluations on the CPU\slash GPU will be explored in~\cref{S:source_results}.
 
@@
 \Cref{F:cpu_source} explores the performance of the source term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
-Source term evaluations---critial in their own right for direct numerical simulations of reacting flow~\cite{Spafford:2010aa}, among other applications---also give us a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation peformance.
+Source term evaluations---critical in their own right for direct numerical simulations of reacting flow~\cite{Spafford:2010aa}, among other applications---also give a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation performance.
 
@@
 This normalization of the runtime by the number of initial conditions tested is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{t:models}).
-For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the unvectorized (strictly parallely) Intel OpenCL code is slightly faster than OpenMP in all cases.\todo{rerun OpenMP w/ mtune=native}
-As expected, the \avx/ machine is moderately faster than the \sse/ CPU for these strictly parallely cases; additionally, the shallow-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.
+For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the unvectorized (strictly parallely) Intel OpenCL code is slightly faster than OpenMP in all cases.
+As expected, the \avx/ machine is faster than the \sse/ CPU for the strictly parallel cases, \SIrange{1.82}{2.13}{$\times$} for the unvectorized OpenCL case, and \SIrange{1.72}{1.85}{$\times$} for OpenMP.
+Additionally, the shallow-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.
 
-In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is calculated per-machine, e.g., the \avx/ shallow-vectorized code speedup is relative to OpenMP on the same CPU such that the speedup of the shallow-vectorized OpenCL code can be readily assessed.
-On both machines, the unvectorized OpenCL code is faster than the baseline parallel OpenMP code; by \SIrange{1.25}{1.51}{$\times$} on the \avx/ CPU, and \textasciitilde\SI{1.25}{$\times$} on the \sse/ machine.
-Additionally, the shallow-vectorized OpenCL code is \textasciitilde\SIrange{2.5}{2.9}{$\times$} and \textasciitilde\SIrange{3.4}{4.1}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
-In contrast,~\cref{F:pocl_source} shows the runtime of deep, shallow and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
-No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP is faster in both cases.
-The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work; however, briefly speaking vectorization might be achieved on POCL via a combination of fixed-array strides (rather than potentially varying with problem-size), use of relaxed math (``\texttt{-cl-fast-relaxed-math}''), explicit OpenCL vector types and newer versions of LLVM\slash POCL (specifically, potential improvements to the vectorization module).
-This will be explored more in future work, but for the moment POCL is still quite useful as a validation tool.
-
-\begin{figure}[htbp!]
+\begin{figure}[htbp]
    \centering
@@
 
+In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is calculated per-machine, e.g., the \avx/ shallow-vectorized code speedup is relative to OpenMP on the same CPU such that the speedup of the shallow-vectorized OpenCL code can be readily assessed.
+On both machines, the unvectorized OpenCL code is faster than the baseline parallel OpenMP code; by \SIrange{1.30}{1.51}{$\times$} on the \avx/ CPU, and \SIrange{1.25}{1.27}{$\times$} on the \sse/ machine.
+Additionally, the shallow-vectorized OpenCL code is \textasciitilde\SIrange{2.53}{2.92}{$\times$} and \SIrange{3.40}{4.08}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
+In contrast,~\cref{F:pocl_source} shows the mean runtime per condition of deep, shallow and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
+No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP case is faster in both cases.
+The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work; however, briefly speaking vectorization might be achieved on POCL via a combination of fixed-array strides---rather than (potentially) varying with the number of initial conditions being evaluated, as currently is implemented---use of relaxed math (``\texttt{-cl-fast-relaxed-math}''), explicit OpenCL vector types and newer versions of LLVM\slash POCL (specifically, potential improvements to the vectorization module).
+This will be explored more in future work, but for the moment POCL is still quite useful as a validation tool.
+
 \Cref{F:gpu_source} shows the performance of the source term evaluation on the GPUs listed in~\cref{t:gpus}.
 In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean runtime per condition; the mean runtime continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
-\Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling}; the speedup ranges from \textasciitilde\SIrange{1.7}{2.6}{$\times$} depending on the chemical model in question.
-Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector width size (a.k.a., the block size in NVIDIA terminology) has minimal effect on performance; it is possible that use of larger vector sizes could have an effect on performance, but this was not considered in this work.\todo{rerun NVIDIA on srv2 -- see if that high variability in isopentanol persists}
-
+Interestingly, the last data-point for the GRI-Mech 3.0 shows an increase in runtime per condition, this is likely due to multiple data transfers to\slash from the GPU required to complete the source rate evaluations in this case; similar (if smaller) increases in runtime per condition can be seen for the \ce{iC5H11OH} model near \num{e4} and \num{e5} conditions.\todo{rerunning}
+\Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} for two different vector widths (a.k.a., the block size in NVIDIA terminology), \num{64} and \num{128}; the fastest \gpunew/ case (a vector width of 128) is \SIrange{1.34}{4.89}{$\times$} faster than the slowest case (either \gpunew/ or \gpuold/ with a vector width of \num{64}) depending on the chemical model in question.
+Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector width has minimal effect on performance for most \gpuold/ cases (excepting \ce{H2}\slash\ce{CO}), but a large effect on the performance of the \gpunew/ GPU; this is likely caused by higher occupancy~\cite{occupancy} on the \gpunew/ GPU, however it is unclear exactly how NVIDIA's OpenCL runtime balances the registers\slash warps per streaming-multiprocessor, as controls occupancy in CUDA~\cite{occupancy}; this is a topic that merits future investigation.
 \begin{figure}[htbp]
@@
 In~\cref{F:source_conpvsconv}, we see that choice of the constant-pressure or constant-volume formulation has little to no effect on runtime for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx/ machine.
-Generally speaking, the difference between the \conp/ and \conv/ formulations has little effect on performance regardless of CPU\slash GPU choice, or even for Jacobian evaluation which will be explored later.
-
-In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, i.e., the \SIrange{1.35}{2.1}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-ordered OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
-Additionally, we note that the ``C''-ordered shallow-vectorization in~\cref{F:source_cvsf} utilizes the vectorized-data ordering described in~\cref{S:data}.
-In general, the data-ordering has a much more profound affect on the vectorized-OpenCL codes; this is due to the nature of SIMD-operations, which are much more efficient when operating on contiguous (in memory) data rather than incurring an expensive memory ``gather'' operation.
-The ``C''-ordered unvectorized OpenCL code (which uses the default unvectorized row-major data ordering) is \SIrange{1.6}{2.3}{$\times$} faster than the unvectorized ``F''-ordered code, and is in some cases almost as fast as the ``F''-ordered shallow-vectorized OpenCL implementation.
-However, the ``C''-ordered shallow-vectorization achieves speedups of \SIrange{3.9}{7.2}{$\times$} demonstrating the value of the vectorized data ordering for CPU execution.
-
-\Cref{F:source_gpu_cvsf} shows the mean runtime per-condition of the ``C'' versus ``F''-ordering on source term evaluation performance on both GPUs.
-In general, the ``C'' and ``F''-ordered shallow-vectorizations are nearly equivalent on both GPUs while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, with the exception of the isopentanol model where the ``C''-ordered data is up to \SI{1.1}{$\times$} faster on the \gpunew/ (we note that there was high variablity for the \gpuold/ in this case).
-\todo{again, pending rerun on \gpuold/}
-The rough equivalence between the ``C'' and ``F''-ordered codes GPUs is counter to what one might expect, as typically speaking coalesced memory access in a shallow-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
-However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector width size and thus, likely coalesced.
-% but the improved data-locality, coupled with the larger cache sizes of the \gpunew/ GPU result in this modest speedup for ``C''-data ordering.
-
-\Cref{F:source_vector_width} shows the effect of changing vector-width size on source rate evaluation performance the \avx/ CPU.
-The vector-width of size \num{8} was the fastest tested, while the larger vector-width of \num{16} is slightly slower due to increased register pressure~\cite{intel_vectypes}.
-It is unclear why the vector-width of \num{4} results in no speedup at all (it is in fact, the slowest case); Intel's vectorization guide~\cite{intel_vecknobs} mentions that a heuristic determines the optimal vector-width (in this case, it appears from compiler output to be \num{8}), so it is possible using vector-width smaller than the heuristic breaks the implicit vectorizer.
-It is noted that this issue does not occur for a vector width of \num{4} on the \sse/ CPU.
+Generally speaking, the difference between the \conp/ and \conv/ formulations has only a marginal effect on performance regardless of CPU\slash GPU choice.
 
@@
       \includegraphics[width=\textwidth]{source_conpvsconv.pdf}
-      \caption{The mean runtime per condition for each chemical model using both the \conp/ and \conv/ formulations on Intel OpenCL and OpenMP for \avx/.}
+      \caption{The mean runtime per condition for each chemical model using both the \conp/ and \conv/ formulations on Intel OpenCL and OpenMP on the \avx/ CPU.}
       \label{F:source_conpvsconv}
@@
 
-Finally,~\cref{F:source_scaling} displays the (strong) parallel scaling efficiency and SIMD efficiency for the CPU platforms.
-The strong parallel scaling efficiency is defined as:
-\begin{equation}
- \label{e:strong_scaling}
- \varepsilon = \frac{\bar{t}_{1}}{N \bar{t}_{N}}
-\end{equation}
-where $\bar{t}_{N}$ is the mean runtime per condition on $N$ CPU cores, and $\bar{t}_{1}$ the same on a single CPU core.
-The strong parallel scaling efficiency measures the speedup due to the use of additional CPU cores, as a fraction of linear speedup; strong scaling tends to decrease with the number of processors used due to memory bandwidth limitations and decreasing computation work allocated per CPU core~\cite{strong_scaling}.
+In contrast,~\Cref{F:source_cvsf} shows significant speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, e.g., the \SIrange{1.35}{2.09}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-ordered OpenMP baseline.
+Additionally, it is noted that the ``C''-ordered shallow-vectorization in~\cref{F:source_cvsf} utilizes the vectorized-data ordering described in~\cref{S:data}; this case achieves speedups of \SIrange{2.14}{2.58}{$\times$} over the ``F''-ordered shallow-vectorization, demonstrating the value of the vectorized data ordering for CPU execution.
 
-\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source rate evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
-In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for Intel OpenCL; although the mechanism scalings vary somewhat for OpenMP, there is no clear trend.
-It is likely the poor scaling of the \ce{H2}\slash\ce{CO} for Intel OpenCL results from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD intructions.
-Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.
-This is again likely due to relatively small amounts of work associated with source rate evaluation; the longest runtime (the isopentanol model on a single core) takes only \textasciitilde\SI{2}{\second} for OpenCL, and \textasciitilde\SI{9}{\second} for OpenMP.
-By the time \numrange{4}{8} CPU cores are utilized in the shallow-vectorized OpenCL code, the runtime can be measured in tens to hundreds of milliseconds and parallel overhead begins to dominate, decreasing efficiency.
-In contrast, the OpenMP code running on \num{4} cores is as fast (\ce{H2}\slash\ce{CO}) or slightly slower than the shallow-vectorized OpenCL code on a single core, hence for \numrange{8}{16} OpenMP cores the parallel overhead concerns become increasingly important; correspondingly, the OpenMP code's parallel efficiency drops to \num{0.64} for \num{16} cores, similar to the OpenCL code at \num{4} cores.
-Indeed, as will be investigated in~\cref{S:jacobian_results} the strong parallel scaling efficiency for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.
+\Cref{F:source_gpu_cvsf} shows the performance effect of the ``C'' versus ``F''-ordering on source term evaluation performance on both GPUs, with the speedup presented per-GPU.
+\todo[inline]{rewrite this once I get the re-run performance data for the source terms}
+In general, the ``C'' and ``F''-ordered shallow-vectorizations are nearly equivalent on both GPUs while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, with the exception of the isopentanol model where the ``C''-ordered data is up to \SI{1.1}{$\times$} faster on the \gpunew/ (we note that there was high variablity for the \gpuold/ in this case).
+\todo{again, pending rerun on \gpuold/}
+The rough equivalence between the ``C'' and ``F''-ordered codes GPUs is counter to what one might expect, as typically speaking coalesced memory access in a shallow-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
+However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector width size and thus, likely coalesced.
+% but the improved data-locality, coupled with the larger cache sizes of the \gpunew/ GPU result in this modest speedup for ``C''-data ordering.
 
-The SIMD efficiency is defined as:
-\begin{equation}
- \label{e:simd_efficiency}
- \varepsilon_{\text{SIMD}} = \frac{\bar{t}_{\text{unvec}}}{W \bar{t}_{\text{shallow}}}
-\end{equation}
-where $\bar{t}_{\text{unvec}}$ is the mean runtime per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$, the vector width reported in number of double operations in~\cref{t:cpus}.
-This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.
-\Cref{F:source_simd_scaling} shows the SIMD efficiency of source term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models (isopentanol and USC-Mech II) have higher SIMD-efficiency of \numrange{0.76}{0.78}, and the smaller models (\ce{H2}\slash\ce{CO}, GRI-Mech \num{3.0}) have lower SIMD-efficiency, \numrange{0.6}{0.66}.
-This again demonstrates that the SIMD-vectorization becomes efficient as the amount of work to perform increases (with increasing model size).
-Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one; this is likely caused by a combination of use of an OpenCL vector-width greater than the native CPU vector-width (i.e., \num{8} v.s., \num{2}), and the improved data-locality for the vectorized-data ordering, as discussed in~\cref{S:data}, and results a modest \SIrange{7}{17}{\percent} improvement over the nominal vector width.
+\Cref{F:source_vector_width} shows the effect of changing vector-width size on source rate evaluation performance the \avx/ CPU.
+The vector-width of size \num{8} was the fastest tested, while the larger vector-width of \num{16} is slightly slower due to increased register pressure~\cite{intel_vectypes}.
+It is unclear why the vector-width of \num{4} results in no speedup at all (it is in fact, the slowest case); Intel's vectorization guide~\cite{intel_vecknobs} mentions that a heuristic determines the optimal vector-width (in this case, it appears from compiler output to be \num{8}), so it is possible using vector-width smaller than the heuristic breaks the implicit vectorizer.
+It is noted that this issue does not occur for a vector width of \num{4} on the \sse/ CPU.
 
@@
 \end{figure}
+
+Finally,~\cref{F:source_scaling} displays the (strong) parallel scaling efficiency and SIMD efficiency for the CPU platforms.
+The strong parallel scaling efficiency is defined as:
+\begin{equation}
+ \label{e:strong_scaling}
+ \varepsilon = \frac{\bar{t}_{1}}{N \bar{t}_{N}}
+\end{equation}
+where $\bar{t}_{N}$ is the mean runtime per condition on $N$ CPU cores, and $\bar{t}_{1}$ the same on a single CPU core.
+The strong parallel scaling efficiency measures the speedup due to the use of additional CPU cores, as a fraction of linear speedup; strong scaling tends to decrease with the number of processors used due to memory bandwidth limitations and decreasing computation work allocated per CPU core~\cite{strong_scaling}.
+
+\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source rate evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
+In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for Intel OpenCL; although the mechanism scalings vary somewhat for OpenMP, there is no clear trend.
+It is likely the poor scaling of the \ce{H2}\slash\ce{CO} for Intel OpenCL results from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD intructions.
+Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.
+This is again likely due to relatively small amounts of work associated with source rate evaluation; the longest runtime (the isopentanol model on a single core) takes only \textasciitilde\SI{22}{\second} for OpenCL, and \textasciitilde\SI{90}{\second} for OpenMP.
+By the time \numrange{4}{8} CPU cores are utilized in the shallow-vectorized OpenCL code, the runtime is just \textasciitilde\SIrange{5}{8}{\sec} and parallel overhead begins to dominate, decreasing efficiency.
+In contrast, the OpenMP code running on \num{4} cores is as fast (\ce{H2}\slash\ce{CO}) or slightly slower than the shallow-vectorized OpenCL code on a single core, hence parallel overhead concerns become increasingly important for \numrange{8}{16} OpenMP cores; correspondingly, the OpenMP code's parallel efficiency drops to \num{0.64} for \num{16} cores, similar to the OpenCL code at \num{4} cores.
+Indeed, as will be investigated in~\cref{S:jacobian_results} the strong parallel scaling efficiency for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.
+
+The SIMD efficiency is defined as:
+\begin{equation}
+ \label{e:simd_efficiency}
+ \varepsilon_{\text{SIMD}} = \frac{\bar{t}_{\text{unvec}}}{W \bar{t}_{\text{shallow}}}
+\end{equation}
+where $\bar{t}_{\text{unvec}}$ is the mean runtime per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$, the vector width reported in number of double operations in~\cref{t:cpus}.
+This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.
+\Cref{F:source_simd_scaling} shows the SIMD efficiency of source term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models (isopentanol and USC-Mech II) have higher SIMD-efficiency of \numrange{0.76}{0.78}, and the smaller models (\ce{H2}\slash\ce{CO}, GRI-Mech \num{3.0}) have lower SIMD-efficiency, \numrange{0.6}{0.66}.
+This again demonstrates that the SIMD-vectorization becomes efficient as the amount of work to perform increases (with increasing model size).
+Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one; this is likely caused by a combination of use of an OpenCL vector-width greater than the native CPU vector-width (i.e., \num{8} v.s., \num{2}), and the improved data-locality for the vectorized-data ordering, as discussed in~\cref{S:data}, and results a modest \SIrange{7}{17}{\percent} improvement over the nominal vector width.
 

### 2018-03-08 paper.tex
edits

@@
   \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{source_crossover.pdf}
+      \caption{The mean runtime per-condition of OpenMP source term evaluation on different cores compared to a shallow-vectorized Intel OpenCL evaluation on \num{1} (solid lines) and \num{4} (dashed lines) cores on \avx/ machine.}
+      \label{F:source_crossover}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[t]{0.48\linewidth}
       \includegraphics[width=\textwidth]{source_simd_efficiency.pdf}
@@
   \end{subfigure}
-  \\
+  \hfill
   \begin{subfigure}[t]{0.48\linewidth}
@@
 Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.
-This is again likely due to relatively small amounts of work associated with source rate evaluation; the longest runtime (the isopentanol model on a single core) takes only \textasciitilde\SI{22}{\second} for OpenCL, and \textasciitilde\SI{90}{\second} for OpenMP.
-By the time \numrange{4}{8} CPU cores are utilized in the shallow-vectorized OpenCL code, the runtime is just \textasciitilde\SIrange{5}{8}{\sec} and parallel overhead begins to dominate, decreasing efficiency.
-In contrast, the OpenMP code running on \num{4} cores is as fast (\ce{H2}\slash\ce{CO}) or slightly slower than the shallow-vectorized OpenCL code on a single core, hence parallel overhead concerns become increasingly important for \numrange{8}{16} OpenMP cores; correspondingly, the OpenMP code's parallel efficiency drops to \num{0.64} for \num{16} cores, similar to the OpenCL code at \num{4} cores.
-Indeed, as will be investigated in~\cref{S:jacobian_results} the strong parallel scaling efficiency for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.
+This is due to better performance of the shallow-vectorized OpenCL code, coupled with the relatively small amounts of work associated with source rate evaluation.
+To illustrate this,~\cref{F:source_crossover} shows the mean runtime per-condition of the OpenMP source term evaluations for \numrange{2}{16} cores, compared to the shallow-vectorized OpenCL code on \num{1} (solid lines) and \num{4} (dashed lines) four cores on the \avx/ machine.
+For all chemical models, the mean runtime per-condition (and hence the computational work allocated per-core, one of the key-drivers of parallel scaling efficiency~\cite{strong_scaling}) of OpenMP running on \num{4} cores is roughly equal to that of the shallow-vectorized OpenCL code on a single core.
+Similarly, OpenMP running on \num{16} cores is roughly equivalent to the OpenCL code on \num{4} cores.
+Therefore, a more fair comparison of parallel scaling efficiencies is to compare OpenCL running on \num{4} cores, to OpenMP on \num{16}; the OpenMP code's parallel efficiency drops to \textasciitilde\num{0.64} for \num{16} cores, similar to OpenCL's parallel scaling efficiency of \numrange{0.66}{0.72} at \num{4} cores.
+Indeed, as will be investigated in~\cref{S:jacobian_results} the strong parallel scaling efficiency for sparse Jacobian evaluation---the most computationally intensive task in this work---are far more similar between Intel OpenCL and OpenMP.
 

### 2018-03-08 paper.tex
updates

@@
 OpenCL Version & \num{1.1}	& \num{1.2}	\\
+\addtocounter{footnote}{1}
+Memory\footnotemark[\thefootnote] & \SI{6}{\giga\byte} & \SI{12}{\giga\byte} \\
 \bottomrule
@@
 \caption{The NVIDIA GPU configurations used in this study.  NVIDIA's OpenCL runtime is provided with the graphic driver, rather than any specific version of CUDA.
-Again, the identifier field will be used as a shorthand descriptor during analysis of the performance results.}
+Again, the identifier field will be used as a shorthand descriptor during analysis of the performance results.
+}
 \label{t:gpus}
 \end{table}
+\footnotetext{Due to an driver implementation issue~\cite{nvidia_memory}, the total memory used on the \gpuold/ and \gpunew/ GPUs were to limited to \SI{4}{\giga\byte} and \SI{10}{\giga\byte} respectively.\addtocounter{footnote}{1}}
 
@@
 
-As in our previous work~\cite{Niemeyer:2016aa}, the correctness of the Jacobian was determined by comparison to the Jacobian obtained by automatic differentiation of the source rate subroutines generated by \texttt{pyJac} via the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
+As in our previous work~\cite{Niemeyer:2016aa}, the correctness of the Jacobian was determined by comparison to the Jacobian obtained by automatic differentiation of the source-rate subroutines generated by \texttt{pyJac} via the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
 The reasons for this choice are explained fully in~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with minimal additional programming effort.
@@
 Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
-The OpenMP Jacobian\slash source rate evaluation kernels, as well as the OpenMP\slash OpenCL wrapping code---responsible for initializing\slash transferring memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms, and \texttt{gcc v4.8.5} on the \sse/\slash\gpuold/ machines.
+The OpenMP Jacobian\slash source-rate kernels, as well as the OpenMP\slash OpenCL wrapping code---responsible for initializing\slash transferring memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms, and \texttt{gcc v4.8.5} on the \sse/\slash\gpuold/ machines.
 The optimization level ``\texttt{-O3 -mtune=native}'' was used, and no ``fast-math'' OpenCL optimizations were enabled.
@@
 In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean runtime per condition; the mean runtime continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
-Interestingly, the last data-point for the GRI-Mech 3.0 shows an increase in runtime per condition, this is likely due to multiple data transfers to\slash from the GPU required to complete the source rate evaluations in this case; similar (if smaller) increases in runtime per condition can be seen for the \ce{iC5H11OH} model near \num{e4} and \num{e5} conditions.\todo{rerunning}
+Interestingly, the last data-point for the GRI-Mech 3.0 shows an increase in runtime per condition, this is likely due to multiple data transfers to\slash from the GPU required to complete the source-rates in this case; similar (if smaller) increases in runtime per condition can be seen for the \ce{iC5H11OH} model near \num{e4} and \num{e5} conditions.\todo{rerunning}
 \Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} for two different vector widths (a.k.a., the block size in NVIDIA terminology), \num{64} and \num{128}; the fastest \gpunew/ case (a vector width of 128) is \SIrange{1.34}{4.89}{$\times$} faster than the slowest case (either \gpunew/ or \gpuold/ with a vector width of \num{64}) depending on the chemical model in question.
@@
 
-\Cref{F:source_vector_width} shows the effect of changing vector-width size on source rate evaluation performance the \avx/ CPU.
+\Cref{F:source_vector_width} shows the effect of changing vector-width size on source-rate performance the \avx/ CPU.
 The vector-width of size \num{8} was the fastest tested, while the larger vector-width of \num{16} is slightly slower due to increased register pressure~\cite{intel_vectypes}.
@@
 
-\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source rate evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
+\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source-rate in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
 In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for Intel OpenCL; although the mechanism scalings vary somewhat for OpenMP, there is no clear trend.
@@
 Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.
-This is due to better performance of the shallow-vectorized OpenCL code, coupled with the relatively small amounts of work associated with source rate evaluation.
+This is due to better performance of the shallow-vectorized OpenCL code, coupled with the relatively small amounts of work associated with source-rate.
 To illustrate this,~\cref{F:source_crossover} shows the mean runtime per-condition of the OpenMP source term evaluations for \numrange{2}{16} cores, compared to the shallow-vectorized OpenCL code on \num{1} (solid lines) and \num{4} (dashed lines) four cores on the \avx/ machine.
@@
 \label{S:jacobian_results}
-\Cref{F:jacobian_perfomance} shows the performance of the sparse and dense Jacobian evaluations in \texttt{pyJac} on the CPU platforms.
-In~\cref{F:sparse_vs_dense}, the mean runtime per condition is presented for the shallow-vectorized Intel OpenCL and OpenMP codes on the \avx/ CPU.
-The sparse Jacobian evaluations are slower on both Intel OpenCL and OpenMP due to indirect lookup indexing requirements, as discussed in~\cref{S:sparsity}.
-Interestingly the shallow-vectorized OpenCL code is less negatively impacted by the indirect lookup; the sparse OpenMP code is \SIrange{2.5}{7}{$\times$} slower than the dense OpenMP evaluation, while the sparse shallow-vectorized OpenCL code is just \SIrange{1.4}{3.3}{$\times$} slower than its dense counterpart
-As a result, the shallow-vectorized sparse OpenCL code is faster than the dense OpenMP code in all cases on the \avx/ machine (\cref{F:sparse_vs_dense}).
-\Cref{F:sparse_vs_dense_speedup} shows the speedup of the sparse\slash dense shallow-vectorized OpenCL the over the same sparse\slash dense Jacobian format on OpenMP; the dense OpenCL is \SIrange{3.7}{4.2}{$\times$} faster than the corresponding dense OpenMP code, this speedup increases to \SIrange{6.6}{9.1}{$\times$} for the sparse Jacobian.
-On the \sse/ machine,~\cref{F:sparse_vs_dense_sse} shows similar results; the sparse OpenMP code is the slowest in all cases, and the shallow-vectorized OpenCL code is nearly as fast as the dense OpenMP code.
-Once again, the sparse OpenCL code is far less negatively impacted by the indirect lookup, and is only \SIrange{1.8}{3.2}{$\times$} slower than its dense counterpart, while the sparse OpenMP code is significantly (\SIrange{2.25}{6.1}{$\times$}) slower than the dense version.
-In~\cref{F:sparse_vs_dense_sse_speedup}, the speedup of the sparse and dense shallow-vectorized OpenCL codes are compared to their OpenMP version; the dense OpenCL code is \SIrange{2}{2.5}{$\times$} faster, while the sparse shallow-vectorization achieves a speedup of \SIrange{3.1}{4.3}{$\times$}.
 
@@
   \begin{subfigure}[t]{0.48\linewidth}
-      \includegraphics[width=\textwidth]{sparse_vs_dense_sse_speedup.pdf}
+      \includegraphics[width=\textwidth]{sparse_vs_dense_see_speedup.pdf}
       \caption{The speedup of the sparse and dense Jacobian evaluation for Intel OpenCL over the sparse\slash dense OpenMP baseline on the \sse/ machine.}
@@
 
-In~\cref{F:sparse_jac_scaling}, the strong parallel scaling efficiency of the sparse shallow-vectorized OpenCL is compared to the sparse OpenMP code.
-The plot is a bit difficult to read, as most of the data-points are clustered together however the key takeaway is that the parallel scaling efficiency of the shallow-vectorized OpenCL code is much closer in value to that of the OpenMP code as compared to the source rate evaluations (\cref{F:source_parallel_scaling}).
-The \ce{H2}\slash\ce{CO} model has the worst scaling for both codes, ranging from \numrange{0.93}{0.54} and \numrange{0.99}{0.86} for OpenCL and OpenMP respectively on \numrange{2}{16} cores.
-As the model size increases, the efficiency of the OpenCL code improves dramatically, reaching \numrange{0.997}{0.84} for the isopentanol model.
-It is worth noting that the OpenCL code's scaling is more negatively affected by the switch from a single 12-core CPU (\cref{t:cpus}) to two CPUs (i.e., the 16 core data-point) than the OpenMP code.
-\Cref{F:dense_jac_scaling} shows scaling for the shallow-vectorized dense Jacobian OpenCL code.
-In this case, the isopentanol model\todo{note: don't have data for openmp yet} has the worst scaling for both codes, we note that due to the sheer size of the dense isopentanol Jacobian---the Jacobian for a single thermo-chemical state takes over \SI{1}{\mega\byte} of data---the total number of states for the dense isopentanol Jacobian was limited to \num{50000} (which takes over \SI{50}{\giga\byte} of memory); this limiting of the number of initial states greatly drops the computation cost, and adversely affects the scaling efficiency as has been discussed previously.
-Excluding the isopentanol model, the dense shallow-vectorized OpenCL scalings are similar (if slightly higher) to that of the source rate evaluation, e.g., \numrange{0.7}{0.77} and \numrange{0.52}{0.61} for \num{4} and \num{8} cores respectively.
-As in~\cref{S:source_results}, we note that the shallow-vectorized OpenCL code on a single core of the \avx/ machine is roughly as fast than the OpenMP code on four cores of the same; hence, the OpenMP code has more room to improve performance before parallel-overhead and memory-bandwidth issues become a limiting concern.
+\Cref{F:jacobian_perfomance} shows the performance of the sparse and dense Jacobian evaluations in \texttt{pyJac} on the CPU platforms.
+In~\cref{F:sparse_vs_dense}, the mean runtime per condition is presented for the shallow-vectorized Intel OpenCL and OpenMP codes on the \avx/ CPU.
+The sparse Jacobian evaluations are slower on both Intel OpenCL and OpenMP due to indirect lookup indexing requirements, as discussed in~\cref{S:sparsity}.
+Interestingly the shallow-vectorized OpenCL code is less negatively impacted by the indirect lookup; the sparse OpenMP code is \SIrange{2.47}{10.42}{$\times$} slower than the dense OpenMP evaluation, while the sparse shallow-vectorized OpenCL code is just \SIrange{1.41}{3.34}{$\times$} slower than its dense counterpart
+As a result, the shallow-vectorized sparse OpenCL code is as fast, or faster than the dense OpenMP code in all cases on the \avx/ machine (\cref{F:sparse_vs_dense}).
+\Cref{F:sparse_vs_dense_speedup} shows the speedup of the sparse\slash dense shallow-vectorized OpenCL the over the same sparse\slash dense Jacobian format on OpenMP; the dense OpenCL code is \SIrange{3.03}{4.23}{$\times$} faster than the corresponding dense OpenMP code, this speedup increases to \SIrange{6.63}{9.44}{$\times$} for the sparse Jacobian.
+
+On the \sse/ machine,~\cref{F:sparse_vs_dense_sse} shows similar results; the sparse OpenMP code is the slowest in all cases, and the shallow-vectorized OpenCL code is nearly as fast as the dense OpenMP code.
+Once again, the sparse OpenCL code is far less negatively impacted by the indirect lookup, and is only \SIrange{1.76}{3.33}{$\times$} slower than its dense counterpart, while the sparse OpenMP code is significantly (\SIrange{2.25}{8.72}{$\times$}) slower than the dense version.
+In~\cref{F:sparse_vs_dense_sse_speedup}, the speedup of the sparse and dense shallow-vectorized OpenCL codes are compared to their OpenMP version; the dense OpenCL code is \SIrange{1.92}{2.47}{$\times$} faster, while the sparse shallow-vectorization achieves a speedup of \SIrange{3.14}{5.03}{$\times$}.
 
@@
 
-The mean runtime per-condition for the sparse and dense Jacobian evaluation on the \gpunew/ GPU is plotted in~\cref{F:gpu_sparse_vs_dense}.
-As in the species evaluation case, the mean runtime per-condition drops steadily, for both cases becoming constant after just \num{3e3} states (except for \ce{H2}\slash\ce{CO}, which levels off near \num{e4} conditions).
-\Cref{F:gpu_jacobian_speedup} presents the speedup of the \gpunew/ over the \gpuold/ GPU for sparse\slash dense Jacobian evaluation; sparse evaluation on the \gpunew/ is \SIrange{1.22}{1.63}{$\times$} faster, while dense evaluation is \SIrange{1.62}{2.1}{$\times$} faster.
-The speedup on the \gpunew/ appears to decrease with increasing mechanism size for the sparse evalution, however we note that the \gpuold/ GPU has high variability in runtimes for the larger models.\todo{missing srv2 data for ipentanol, and need to rerun srv2 data}
+In~\cref{F:sparse_jac_scaling}, the strong parallel scaling efficiency of the sparse shallow-vectorized OpenCL is compared to the sparse OpenMP code.
+The plot is a bit difficult to read, as most of the data-points are clustered together, however the key takeaway is that the parallel scaling efficiency of the shallow-vectorized OpenCL code is very similar to that of the OpenMP code, in contrast to the source-rates parallel scaling efficiencies (\cref{F:source_parallel_scaling}).
+The \ce{H2}\slash\ce{CO} model has the worst scaling for both codes, ranging from \numrange{0.94}{0.54} and \numrange{0.99}{0.82} for OpenCL and OpenMP respectively on \numrange{2}{16} cores.
+As the model size increases, the efficiency of the OpenCL code improves dramatically, reaching \numrange{0.997}{0.84} for the isopentanol model.
+\Cref{F:dense_jac_scaling} shows scaling for the shallow-vectorized dense Jacobian OpenCL code.
+In this case, the isopentanol model has the worst scaling for both codes, it is noted that due to the sheer size of the dense isopentanol Jacobian---storing the dense matrix for a single thermo-chemical state takes over \SI{1}{\mega\byte} of data---the total number of states for the dense isopentanol Jacobian evaluation was limited to \num{50000} (which requires over \SI{50}{\giga\byte} of memory); this greatly drops the computation cost for this case, and adversely affects the scaling efficiency as was discussed in~\cref{S:source_results}.
+Excluding the isopentanol model, the dense shallow-vectorized OpenCL scalings are slightly higher than for source-rate evaluation, e.g., \numrange{0.70}{0.78} and \numrange{0.52}{0.61} for \num{4} and \num{8} cores respectively (compared to \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for shallow-vectorized OpenCL source-rate on \num{4} and \num{8} cores respectively); this is due to the higher computational cost of Jacobian evaluation, and hence more available work per CPU-core.
+As in~\cref{S:source_results}, we note that the shallow-vectorized dense Jacobian code running on \num{1} and \num{4} cores of the \avx/ machine is roughly as fast as the OpenMP code on \num{4} and \num{16} cores respectively; the parallel scaling efficiency of OpenMP at \num{16} cores (excluding isopentanol as previously) is \numrange{0.63}{0.71}, similar to the shallow-vectorization's efficiency of \numrange{0.70}{0.78} for \num{4} cores.
 
@@
       \includegraphics[width=\textwidth]{gpu_jacobian_speedup.pdf}
-      \caption{Speedup of the \gpunew/ over \gpuold/ GPUs for sparse and dence Jacobian evaluation.}
+      \caption{Speedup of the \gpunew/ over \gpuold/ GPUs for sparse and dense Jacobian evaluation, normalized per-Jacobian type (i.e., sparse versus dense).}
       \label{F:gpu_jacobian_speedup}
@@
 
-In~\cref{F:fd_vs_analytical}, the performance of the sparse analytical-Jacobian is compared to a sparse first-order finite-difference Jacobian on both the \avx/ CPU and the \gpuold/ \gpunew/ GPUs.\todo{missing data}
-\Cref{F:fd_vs_analytical_cpu} shows large speedups for both the OpenMP code and shallow-vectorized OpenCL code; the analytical OpenMP Jacobian is \SIrange{3.92}{-11111}{$\times$} faster, while the analytical OpenCL Jacobian achives speedups of \SIrange{17}{55}{$\times$}.
-It is noted that the current finite-difference formulation breaks Intel's auto-vectorizer, hence the OpenCL speedup comparison is made against the unvectorized OpenCL code; it is not a high priority to fix this issue, as the finite-difference Jacobian is implemented for comparison purposes only.
-Although we do not display the dense finite-difference Jacobian speedup in~\cref{F:fd_vs_analytical_cpu}, the dense OpenCL and OpenMP analytical Jacobian codes out-perform the finite-difference variants by even larger margins; \SIrange{24}{-155}{$\times$} for OpenCL, and \SIrange{10}{-11111}{$\times$} for OpenMP.
-A comparison of the sparse analytical and finite-difference Jacobians is presented in~\cref{F:fd_vs_analytical_gpu}.
-The analytical Jacobian on the \gpunew/ shows large speedups, increasing with chemical model size, in the range of \SIrange{5}{16}{$\times$}; however, the \gpuold/ GPU shows almost no speedup, it is not clear why this is the case.\todo{reaaaally need to re-run this data}
-Although not pictured, the dense analytical-Jacobian on the \gpunew/ GPU has larger speedups compared with the dense finite-difference Jacobian, in the range of \SIrange{5}{42}{$\times$}.
+The mean runtime per-condition for the sparse and dense Jacobian evaluation on the \gpunew/ GPU is plotted in~\cref{F:gpu_sparse_vs_dense}.
+As in the species evaluation case, the mean runtime per-condition drops steadily, for both cases becoming constant after just \num{3e3} states (except for \ce{H2}\slash\ce{CO}, which levels off near \num{e4} conditions).
+Sparse jacobian evaluation is significantly slower for all models before the GPU becomes saturated (due to the indirect indexing lookup), however past the saturation point the performance gap between the sparse and dense evaluations narrows.
+This is likely caused by the combination of more sparse Jacobian matrices fitting in the \gpunew/'s memory, as well as improved data-locality\slash caching.
+\Cref{F:gpu_jacobian_speedup} presents the speedup of the \gpunew/ over the \gpuold/ GPU for sparse\slash dense Jacobian evaluation; sparse evaluation on the \gpunew/ is \SIrange{1.10}{1.59}{$\times$} faster than on the \gpuold/, while dense evaluation is \SIrange{1.36}{3.0}{$\times$} faster.
+The speedup on the \gpunew/ appears to decrease with increasing mechanism size for the sparse evalution, but increases for the larger models in dense evaluation; this is likely due to the larger memory size of the \gpunew/, and hence fewer data-transferring operations.\todo{missing srv2 data for full conditions}
 
@@
       \includegraphics[width=\textwidth]{finite_difference_vs_analytical.pdf}
-      \caption{Speedup of the analytical versus finite-difference Jacobian evaluation on the \avx/ CPU, normalized per-language}
+      \caption{Speedup of the sparse analytical Jacobian versus finite-difference Jacobian evaluation on the \avx/ CPU, normalized per-language}
       \label{F:fd_vs_analytical_cpu}
@@
 
-Finally, in~\cref{F:v1_vs_v2} compares the performance for evaluation of the dense-analytical Jacobian of the new version of \texttt{pyJac} to the previous version~\cite{pyjac16} on the \sse/ CPU and \gpuold/ GPU; dense Jacobian evaluation was selected as it was the only type implemented in the previous verison.
+In~\cref{F:fd_vs_analytical}, the performance of the sparse analytical-Jacobian is compared to a sparse first-order finite-difference Jacobian on both the \avx/ CPU and the \gpuold/ \gpunew/ GPUs.
+\Cref{F:fd_vs_analytical_cpu} shows large speedups for both the OpenMP code and shallow-vectorized OpenCL code; the analytical OpenMP Jacobian is \SIrange{3.92}{8.67}{$\times$} faster, while the analytical OpenCL Jacobian achives speedups of \SIrange{17.22}{55.11}{$\times$}.
+It is noted that the isopentanol case was excluded here, as a single run of the sparse finite-difference Jacobian using either OpenCL or OpenMP took over \num{12} hours of runtime, in addition the current finite-difference formulation breaks Intel's auto-vectorizer, hence the OpenCL speedup comparison is made against the unvectorized OpenCL code; it is not a high priority to fix this issue, as the finite-difference Jacobian is implemented for comparison purposes only.
+Although we do not display the dense finite-difference Jacobian speedup in~\cref{F:fd_vs_analytical_cpu}, the dense OpenCL and OpenMP analytical Jacobian codes out-perform the finite-difference variants by even larger margins; \SIrange{24.44}{245.63}{$\times$} for OpenCL, and \SIrange{9.68}{112.73}{$\times$} for OpenMP (these data do include the isopentanol model, though limited to \num{50000} conditions as discussed previously).
+A comparison of the sparse analytical and finite-difference Jacobians is presented in~\cref{F:fd_vs_analytical_gpu}.
+The analytical Jacobian on the \gpunew/ shows large speedups, increasing with chemical model size, in the range of \SIrange{4.66}{17.43}{$\times$}; however, the \gpuold/ GPU shows almost no speedup, it is not clear why this is the case, it is possible that the finite-difference Jacobian evaluation does not function correctly on the \gpuold/ GPU.
+Although not pictured, the dense analytical-Jacobian on the \gpunew/ GPU has larger speedups compared with the dense finite-difference Jacobian: \SIrange{5.32}{44.87}{$\times$}.
 
@@
 \end{figure}
+
+\Cref{F:v1_vs_v2} compares the performance for evaluation of the dense-analytical Jacobian of the new version of \texttt{pyJac} to the previous version~\cite{pyjac16} on the \sse/ CPU and \gpuold/ GPU; dense Jacobian evaluation was selected as it was the only type implemented in the previous version of \texttt{pyJac}.
+On the \sse/ CPU, the \texttt{pyJac-v2} is faster than \texttt{pyJac-v1} for OpenMP evaluation for the larger chemical models; the static code generated by \texttt{pyJac-v1} is \SI{1.79}{$\times$} faster for the \ce{H2}\slash\ce{CO} model, and only \SI{1.09}{$\times$} slower for the GRI-Mech 3.0 model.
+In contrast, the loop-based OpenMP code of \texttt{pyJac-v2} is \SIrange{3.37}{10.19}{$\times$} faster than \texttt{pyJac-v1} for the USC-Mech II and isopentanol models.
+The shallow-vectorized OpenCL code is faster than \texttt{pyJac-v1} in all cases, achieving speedups of \SI{1.37}{19.56}{$\times$} increasing with model size.
 

### 2018-03-09 paper.tex
revise V1 vs V2 discussion & future directions

@@
 Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates user specified psuedo-code and data to OpenCL\slash C-code.
-This allows for easy changes of program structure, e.g. data ordering, vectorization, threading patterns and switching of target-languages (and simpler extension to additional languages, e.g., CUDA).
-Additionally, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source terms.
-The source-terms and subcomponents thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{adept-v11} is used to obtain reference answers for Jacobian subcomponents.
-The Portable OpenCL (POCL) implementation~\cite{poclIJPP} and OpenMP~\cite{dagum1998openmp} are used to run OpenCL and C unit-testing, respectively, on the continous integration framework Travis CI~\cite{travis:2018}.
-Validation of the complete (as opposed to the subcomponent testing discussed here) generated source-terms and Jacobian codes will be discussed in detail in~\cref{s:validation}.
+As the name implies, \texttt{loo.py} generates code based on for-loops, this differs from the previous version of \texttt{pyJac}\cite{pyjac16} which generates static code (i.e., fully unrolled loops, with thermodynamic\slash reaction parameters written directly in code rather than stored in arrays).
+This caused some issues with large file sizes, long compilation times and even occasionally breaking the \texttt{gcc} and \texttt{nvcc} compilers (the latter issue necessitated breaking the Jacobian\slash source-term evaluations into separate files)\cite{Niemeyer:2016aa}.
+The implications of this change will be studied in~\cref{S:jacobian_results}, where the performance of the new version of~\texttt{pyJac} will be compared to the previous versions.
+
+In addition, \texttt{loo.py} allows the user to more easily make changes the structure of the generated program, e.g., data ordering, vectorization, threading patterns, as well as switching the target-language (and simpler extension to additional languages, e.g., CUDA).
+Further, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source terms.
+The source-terms and subcomponents thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{adept-v11} is used to obtain reference answers for Jacobian sub-components.
+The Portable OpenCL (POCL) implementation~\cite{poclIJPP} and OpenMP~\cite{dagum1998openmp} are used to run OpenCL and C unit-testing, respectively, on the continuous integration framework Travis CI~\cite{travis:2018}.
+Validation of the complete (as opposed to the sub-component testing discussed here) generated source-terms and Jacobian codes will be discussed in detail in~\cref{s:validation}.
 
@@
 The Intel and NVIDIA OpenCL runtimes lack implementations of atomic operations on double precision variables; these are currently necessary for \texttt{pyJac} to run deep-vectorized code.
-On the other hand, POCL is an open source OpenCL runtime that works on all CPU types tested in this work, and does implement these atomic operations; it is our experience that the POCL community is very responsive to bug reports and user outreach.
+On the other hand, POCL is an open source OpenCL runtime that works on all CPU types tested in this work, and does implement these atomic operations.
 However, POCL's implicit vectorization module---which utilizes the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to translate OpenCL code to vectorized machine code---typically fails to achieve much speedup, if any.
@@
 
+\Cref{F:v1_vs_v2} compares the performance for evaluation of the dense-analytical Jacobian of the new version of \texttt{pyJac} to the previous version~\cite{pyjac16} on the \sse/ CPU and \gpuold/ GPU; dense Jacobian evaluation was selected as it was the only type implemented in the previous version of \texttt{pyJac}.
+On the \sse/ CPU, the \texttt{pyJac-v2} is faster than \texttt{pyJac-v1} for OpenMP evaluation for the larger chemical models; the static code generated by \texttt{pyJac-v1} (see~\cref{s:unittest}) is \SI{1.79}{$\times$} faster for the \ce{H2}\slash\ce{CO} model, and only \SI{1.09}{$\times$} slower for the GRI-Mech 3.0 model.
+In contrast, the loop-based OpenMP code of \texttt{pyJac-v2} is \SIrange{3.37}{10.19}{$\times$} faster than \texttt{pyJac-v1} for the USC-Mech II and isopentanol models.
+The shallow-vectorized OpenCL code is faster than \texttt{pyJac-v1} in all cases, achieving speedups of \SI{1.37}{19.56}{$\times$} increasing with model size.
+In~\cref{F:v1_vs_v2_gpu}, the performance of the \texttt{pyJac-v2} dense-analytical Jacobian on the \gpuold/ GPU is compared to \texttt{pyJac-v1} on the same.
+As with the CPU, the static code of \texttt{pyJac-v1} is slightly faster for the \ce{H2}\slash\ce{CO} model, however \texttt{pyJac-v2} outperforms \texttt{pyJac-v1} by \SIrange{1.25}{2.84}{$\times$} for the other models; it is noted that the decrease in speedup for the isopentanol model is likely due to the limited number of conditions in dense-evaluation for \texttt{pyJac-v2}, as noted earlier.
+
 \begin{figure}[htbp]
@@
   \end{subfigure}
-  \caption{Performance comparison of the new version of \texttt{pyJac} (\texttt{v2}) and the previous verison, \texttt{v1}~\cite{pyjac16}.}
+  \caption{Performance comparison of the new version of \texttt{pyJac} (\texttt{v2}) and the previous version, \texttt{v1}~\cite{pyjac16}.}
   \label{F:v1_vs_v2}
@@
 
-\Cref{F:v1_vs_v2} compares the performance for evaluation of the dense-analytical Jacobian of the new version of \texttt{pyJac} to the previous version~\cite{pyjac16} on the \sse/ CPU and \gpuold/ GPU; dense Jacobian evaluation was selected as it was the only type implemented in the previous version of \texttt{pyJac}.
-On the \sse/ CPU, the \texttt{pyJac-v2} is faster than \texttt{pyJac-v1} for OpenMP evaluation for the larger chemical models; the static code generated by \texttt{pyJac-v1} is \SI{1.79}{$\times$} faster for the \ce{H2}\slash\ce{CO} model, and only \SI{1.09}{$\times$} slower for the GRI-Mech 3.0 model.
-In contrast, the loop-based OpenMP code of \texttt{pyJac-v2} is \SIrange{3.37}{10.19}{$\times$} faster than \texttt{pyJac-v1} for the USC-Mech II and isopentanol models.
-The shallow-vectorized OpenCL code is faster than \texttt{pyJac-v1} in all cases, achieving speedups of \SI{1.37}{19.56}{$\times$} increasing with model size.
-
 \section{Future directions and practical notes on OpenCL usage}
 \label{s:future}
-It is important to note, that while OpenCL provides an easy way to enable cross-platform execution, there are some serious potential pit-falls in its use.
-Typically speaking, the closed source OpenCL runtimes tested in this work (Intel and NVIDIA) contain more bugs that result in compilation failures or simply incorrect vectorized machine code; further, these runtimes (in our experience) tend to be less responsive to fixing said bugs, with relatively infrequent new releases (or in NVIDIAs case, change-logs of bugfixes).
-As demonstration, we've created a minimum working example~\cite{nvidia_mwe} that demonstrates a failure of NVIDIA's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla K40m GPU; by simply changing to another runtime (e.g., Intel), with no other code changes, the correct result can be found.
+It is important to note that while OpenCL provides an easy way to enable cross-platform execution, and significant speedups were achieved via shallow-vectorized OpenCL code in this work, there are some serious potential pit-falls in its use.
+Typically speaking, the closed source OpenCL runtimes tested in this work (Intel and NVIDIA) contain bugs that result in compilation failures, simply incorrect vectorized machine code or even segmentation faults; further, these runtimes (in our experience) tend to be less responsive to fixing said bugs, with relatively infrequent new releases (or in NVIDIA's case, changelogs\slash public records of bug-fixes).
+On the other hand, the open-source OpenCL runtime used in this work, POCL, has (in our experience) far fewer implementation bugs, and when issues arise the community is very responsive to bug reports and user outreach; however, POCL fails to achieve vectorization as noted in~\cref{S:source_results}.
+
+As demonstration of the type of issue we're discussing, a minimum working example has been created~\cite{nvidia_mwe} that demonstrates a failure of NVIDIA's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla K40m GPU; by simply changing to another runtime (e.g., Intel), with no other code changes, the correct result can be found.
 This provides a particularly vexing problem for the programmer, as there is often little that can be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
+Further, as noted throughout this work certain transformations or code-generation patterns can break OpenCL execution\slash vectorization; e.g., the failure of POCL to achieve vectorization, Intel OpenCL's failure to vectorize the finite-difference Jacobian, the lack of speedup for the finite-difference Jacobian on the NVIDIA runtime for the \gpuold/ GPU.
+Indeed the vectorization attempted by the OpenCL runtimes (or reasons for incorrectly vectorized code) are not always obvious to the user.
+It is possible, particularly for the Intel OpenCL \slash POCL runtimes, that better performance might be achieved using so called ``explicit'' vectorization, i.e., through use of inbuilt vector types such as the \texttt{double8}.
+This will be investigated in future works.
 
-Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the only significant code change is building the wrapper that initializes\slash transfers memory and calls the kernel.
-Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py}) is a key priority going forwards.
+Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change is building the wrapper that initializes\slash transfers memory and calls the kernel.
+Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler pragmas) is a key priority going forwards.
+In addition, CUDA~\cite{NVIDIA:2018} and Intel's open-source OpenCL alternative ISPC~\cite{pharr2012ispc} have been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017} and (in the case of ISPC) initial testing.
+The current deep-vectorization formulation would be executable for a CUDA target---as CUDA implements atomic double-precision floating-point operations---further recommending its use.
+Extending execution to these targets to make CPU\slash GPU more reliable vectorization is also a future goal.
 

### 2018-03-09 paper.tex
bibl cleanup

@@
 In evaluation of the thermo-chemical properties, speedups between \SIrange{8}{13}{$\times$} were on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU respectively.
-Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two-fourth order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on a Tesla C2075; the implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{cvode:2.8.2} running on 12--38 Intel Xeon E5-4640 v2 CPU cores for two smaller chemical models with an integration time step of \SI{e-6}{$\sec$}.
+Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two-fourth order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on a Tesla C2075; the implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on 12--38 Intel Xeon E5-4640 v2 CPU cores for two smaller chemical models with an integration time step of \SI{e-6}{$\sec$}.
 
@@
 Further, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source terms.
-The source-terms and subcomponents thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{adept-v11} is used to obtain reference answers for Jacobian sub-components.
+The source-terms and subcomponents thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{hogan2014fast,adept-v11} is used to obtain reference answers for Jacobian sub-components.
 The Portable OpenCL (POCL) implementation~\cite{poclIJPP} and OpenMP~\cite{dagum1998openmp} are used to run OpenCL and C unit-testing, respectively, on the continuous integration framework Travis CI~\cite{travis:2018}.
@@
 However, \texttt{pyJac} allows the user to simply ignore these derivatives (via a command-line switch) and avoid any negative adverse affects on the Jacobian sparsity.
-The rational behind this choice is that many common implicit integration techniques (e.g., CVODE~\cite{cvode:2.8.2}) used in solution of chemical kinetic initial value problems are formulated around the assumption that the supplied Jacobian is approximate; this allows the Jacobian and its LU factorization to be reused for multiple internal integration time-steps, accelerating the solution process.
+The rational behind this choice is that many common implicit integration techniques (e.g., CVODE~\cite{Hindmarsh:2005}) used in solution of chemical kinetic initial value problems are formulated around the assumption that the supplied Jacobian is approximate; this allows the Jacobian and its LU factorization to be reused for multiple internal integration time-steps, accelerating the solution process.
 For such solvers, there is no need for the exact form of the Jacobian and the so-called ``approximate'' form is preferable.

### 2018-03-09 paper.tex
s/runtime/run time for describing execution times vs e.g., the Intel runtime"

@@
 \begin{abstract} % not to exceed 200 words
-A code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized thermo-chemical source term and sparse chemical kinetic Jacobian evaluation was developed and validated for a shallow range of chemical kinetic models.
+A code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized thermo-chemical source term and sparse\slash dense chemical kinetic Jacobian evaluation was developed and validated for a wide range of chemical kinetic models.
+
 \todo[inline]{this}
@@
 Moreover, the cost of evaluating the chemical source terms scales linearly with the size of the model, while evaluation and factorization of the chemical kinetic Jacobian---if naively implemented---scale quadratically and cubically with the number of species in the model~\cite{LU2009192}.
-These factors often make the use of detailed chemical kinetics prohibitive in practice; e.g., in a recent direct numerical simulation~\cite{Spafford:2010aa} utilizing a 22-species model, evaluation of reaction rates consumed around half of the total runtime.
-In addition, chemical kinetic Jacobian evaluation and factorization---required for most common implicit integration techniques utilized to deal with numerical stiffness---is a bottleneck for even modestly-sized chemical models in realistic reactive flow simulations, and necessitates use of other cost-reduction strategies~\cite{LU2009192}.
+These factors often make the use of detailed chemical kinetics prohibitive in practice; e.g., in a recent direct numerical simulation~\cite{Spafford:2010aa} utilizing a 22-species model, evaluation of reaction rates consumed around half of the total run time.
+In addition, chemical kinetic Jacobian evaluation and factorization---required for most common implicit integration techniques utilized to deal with numerical stiffness---is a bottleneck for even modestly-sized chemical models in realistic reactive flow simulations, necessitating use of other cost-reduction strategies~\cite{LU2009192}.
 
@@
 Additionally, the exact form of the Jacobian (as opposed the the ``approximate'' form discussed in~\cref{S:sparsity}) was used in all cases.
-Finally, unless stated otherwise: the performance results utilized the \conp/ assumption, a vector width of \num{8}\slash\num{128} and ``C''\slash ``F''-ordered data for the CPU\slash GPU cases respectively; the runtimes reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
+Finally, unless stated otherwise: the performance results utilized the \conp/ assumption, a vector width of \num{8}\slash\num{128} and ``C''\slash ``F''-ordered data for the CPU\slash GPU cases respectively; the run times reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
 The effects of choice of vector-width, data-ordering and differences between \conp/ and \conv/ evaluations on the CPU\slash GPU will be explored in~\cref{S:source_results}.
@@
 
-In~\cref{F:intel_source_nonorm}, the mean runtime per initial condition is shown for both the~\avx/ and~\sse/ CPUs, using both Intel OpenCL and OpenMP.
-This normalization of the runtime by the number of initial conditions tested is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{t:models}).
+In~\cref{F:intel_source_nonorm}, the mean run time per initial condition is shown for both the~\avx/ and~\sse/ CPUs, using both Intel OpenCL and OpenMP.
+This normalization of the run time by the number of initial conditions tested is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{t:models}).
 For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the unvectorized (strictly parallely) Intel OpenCL code is slightly faster than OpenMP in all cases.
@@
       \includegraphics[width=\textwidth]{intel_source_nonorm.pdf}
-      \caption{The mean runtime per condition for each chemical model using the Intel OpenCL and OpenMP on both CPUs studied.}
+      \caption{The mean run time per condition for each chemical model using the Intel OpenCL and OpenMP on both CPUs studied.}
       \label{F:intel_source_nonorm}
@@
       \includegraphics[width=\textwidth]{pocl_source_nonorm.pdf}
-      \caption{The mean runtime per condition of the Portable OpenCL runtime compared to OpenMP parallelization.}
+      \caption{The mean run time per condition of the Portable OpenCL runtime compared to OpenMP parallelization.}
       \label{F:pocl_source}
   \end{subfigure}
- \caption{Mean runtimes per condition and speedups achieved by the various CPU OpenCL runtimes compared to OpenMP parallelization for each chemical model studied.
+ \caption{Mean run times per condition and speedups achieved by the various CPU OpenCL runtimes compared to OpenMP parallelization for each chemical model studied.
 The names in the legends correspond to the identifiers listed in~\cref{t:cpus}.}
@@
 Additionally, the shallow-vectorized OpenCL code is \textasciitilde\SIrange{2.53}{2.92}{$\times$} and \SIrange{3.40}{4.08}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
-In contrast,~\cref{F:pocl_source} shows the mean runtime per condition of deep, shallow and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
+In contrast,~\cref{F:pocl_source} shows the mean run time per condition of deep, shallow and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
 No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP case is faster in both cases.
@@
 \Cref{F:gpu_source} shows the performance of the source term evaluation on the GPUs listed in~\cref{t:gpus}.
-In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean runtime per condition; the mean runtime continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
-Interestingly, the last data-point for the GRI-Mech 3.0 shows an increase in runtime per condition, this is likely due to multiple data transfers to\slash from the GPU required to complete the source-rates in this case; similar (if smaller) increases in runtime per condition can be seen for the \ce{iC5H11OH} model near \num{e4} and \num{e5} conditions.\todo{rerunning}
+In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean run time per condition; the mean run time continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
+Interestingly, the last data-point for the GRI-Mech 3.0 shows an increase in run time per condition, this is likely due to multiple data transfers to\slash from the GPU required to complete the source-rates in this case; similar (if smaller) increases in run time per condition can be seen for the \ce{iC5H11OH} model near \num{e4} and \num{e5} conditions.\todo{rerunning}
 \Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} for two different vector widths (a.k.a., the block size in NVIDIA terminology), \num{64} and \num{128}; the fastest \gpunew/ case (a vector width of 128) is \SIrange{1.34}{4.89}{$\times$} faster than the slowest case (either \gpunew/ or \gpuold/ with a vector width of \num{64}) depending on the chemical model in question.
@@
       \includegraphics[width=\textwidth]{gpu_source_scaling.pdf}
-      \caption{The mean runtime per condition for each chemical model on the \gpunew/ GPU as a function of the number of initial conditions tested.}
+      \caption{The mean run time per condition for each chemical model on the \gpunew/ GPU as a function of the number of initial conditions tested.}
       \label{F:gpu_source_scaling}
@@
 \Cref{f:source_permutate} shows the effect of changing data-ordering patterns, the \conp/ or \conv/ formulation, and the CPU vector-width size on the performance of source term evaluation in \texttt{pyJac}.
-In~\cref{F:source_conpvsconv}, we see that choice of the constant-pressure or constant-volume formulation has little to no effect on runtime for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx/ machine.
+In~\cref{F:source_conpvsconv}, we see that choice of the constant-pressure or constant-volume formulation has little to no effect on run time for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx/ machine.
 Generally speaking, the difference between the \conp/ and \conv/ formulations has only a marginal effect on performance regardless of CPU\slash GPU choice.
@@
       \includegraphics[width=\textwidth]{source_conpvsconv.pdf}
-      \caption{The mean runtime per condition for each chemical model using both the \conp/ and \conv/ formulations on Intel OpenCL and OpenMP on the \avx/ CPU.}
+      \caption{The mean run time per condition for each chemical model using both the \conp/ and \conv/ formulations on Intel OpenCL and OpenMP on the \avx/ CPU.}
       \label{F:source_conpvsconv}
@@
 \Cref{F:source_gpu_cvsf} shows the performance effect of the ``C'' versus ``F''-ordering on source term evaluation performance on both GPUs, with the speedup presented per-GPU.
-\todo[inline]{rewrite this once I get the re-run performance data for the source terms}
 In general, the ``C'' and ``F''-ordered shallow-vectorizations are nearly equivalent on both GPUs while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, with the exception of the isopentanol model where the ``C''-ordered data is up to \SI{1.1}{$\times$} faster on the \gpunew/ (we note that there was high variablity for the \gpuold/ in this case).
@@
       \includegraphics[width=\textwidth]{source_crossover.pdf}
-      \caption{The mean runtime per-condition of OpenMP source term evaluation on different cores compared to a shallow-vectorized Intel OpenCL evaluation on \num{1} (solid lines) and \num{4} (dashed lines) cores on \avx/ machine.}
+      \caption{The mean run time per-condition of OpenMP source term evaluation on different cores compared to a shallow-vectorized Intel OpenCL evaluation on \num{1} (solid lines) and \num{4} (dashed lines) cores on \avx/ machine.}
       \label{F:source_crossover}
@@
 \end{equation}
-where $\bar{t}_{N}$ is the mean runtime per condition on $N$ CPU cores, and $\bar{t}_{1}$ the same on a single CPU core.
+where $\bar{t}_{N}$ is the mean run time per condition on $N$ CPU cores, and $\bar{t}_{1}$ the same on a single CPU core.
 The strong parallel scaling efficiency measures the speedup due to the use of additional CPU cores, as a fraction of linear speedup; strong scaling tends to decrease with the number of processors used due to memory bandwidth limitations and decreasing computation work allocated per CPU core~\cite{strong_scaling}.
@@
 This is due to better performance of the shallow-vectorized OpenCL code, coupled with the relatively small amounts of work associated with source-rate.
-To illustrate this,~\cref{F:source_crossover} shows the mean runtime per-condition of the OpenMP source term evaluations for \numrange{2}{16} cores, compared to the shallow-vectorized OpenCL code on \num{1} (solid lines) and \num{4} (dashed lines) four cores on the \avx/ machine.
-For all chemical models, the mean runtime per-condition (and hence the computational work allocated per-core, one of the key-drivers of parallel scaling efficiency~\cite{strong_scaling}) of OpenMP running on \num{4} cores is roughly equal to that of the shallow-vectorized OpenCL code on a single core.
+To illustrate this,~\cref{F:source_crossover} shows the mean run time per-condition of the OpenMP source term evaluations for \numrange{2}{16} cores, compared to the shallow-vectorized OpenCL code on \num{1} (solid lines) and \num{4} (dashed lines) four cores on the \avx/ machine.
+For all chemical models, the mean run time per-condition (and hence the computational work allocated per-core, one of the key-drivers of parallel scaling efficiency~\cite{strong_scaling}) of OpenMP running on \num{4} cores is roughly equal to that of the shallow-vectorized OpenCL code on a single core.
 Similarly, OpenMP running on \num{16} cores is roughly equivalent to the OpenCL code on \num{4} cores.
@@
 \end{equation}
-where $\bar{t}_{\text{unvec}}$ is the mean runtime per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$, the vector width reported in number of double operations in~\cref{t:cpus}.
+where $\bar{t}_{\text{unvec}}$ is the mean run time per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$, the vector width reported in number of double operations in~\cref{t:cpus}.
 This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.
@@
       \includegraphics[width=\textwidth]{sparse_vs_dense.pdf}
-      \caption{The mean runtime per-condition of sparse and dense Jacobian evaluation for Intel OpenCL\slash OpenMP on the \avx/ machine.}
+      \caption{The mean run time per-condition of sparse and dense Jacobian evaluation for Intel OpenCL\slash OpenMP on the \avx/ machine.}
       \label{F:sparse_vs_dense}
@@
       \includegraphics[width=\textwidth]{sparse_vs_dense_sse.pdf}
-      \caption{The mean runtime per-condition of sparse and dense Jacobian evaluation for Intel OpenCL\slash OpenMP on the \sse/ machine.}
+      \caption{The mean run time per-condition of sparse and dense Jacobian evaluation for Intel OpenCL\slash OpenMP on the \sse/ machine.}
       \label{F:sparse_vs_dense_sse}
@@
 \Cref{F:jacobian_perfomance} shows the performance of the sparse and dense Jacobian evaluations in \texttt{pyJac} on the CPU platforms.
-In~\cref{F:sparse_vs_dense}, the mean runtime per condition is presented for the shallow-vectorized Intel OpenCL and OpenMP codes on the \avx/ CPU.
+In~\cref{F:sparse_vs_dense}, the mean run time per condition is presented for the shallow-vectorized Intel OpenCL and OpenMP codes on the \avx/ CPU.
 The sparse Jacobian evaluations are slower on both Intel OpenCL and OpenMP due to indirect lookup indexing requirements, as discussed in~\cref{S:sparsity}.
@@
       \includegraphics[width=\textwidth]{gpu_sparse_vs_dense.pdf}
-      \caption{Mean runtime per-condition of sparse and dense Jacobian evaluation on the \gpunew/ GPU.}
+      \caption{Mean run time per-condition of sparse and dense Jacobian evaluation on the \gpunew/ GPU.}
       \label{F:gpu_sparse_vs_dense}
@@
 
-The mean runtime per-condition for the sparse and dense Jacobian evaluation on the \gpunew/ GPU is plotted in~\cref{F:gpu_sparse_vs_dense}.
-As in the species evaluation case, the mean runtime per-condition drops steadily, for both cases becoming constant after just \num{3e3} states (except for \ce{H2}\slash\ce{CO}, which levels off near \num{e4} conditions).
+The mean run time per-condition for the sparse and dense Jacobian evaluation on the \gpunew/ GPU is plotted in~\cref{F:gpu_sparse_vs_dense}.
+As in the species evaluation case, the mean run time per-condition drops steadily, for both cases becoming constant after just \num{3e3} states (except for \ce{H2}\slash\ce{CO}, which levels off near \num{e4} conditions).
 Sparse jacobian evaluation is significantly slower for all models before the GPU becomes saturated (due to the indirect indexing lookup), however past the saturation point the performance gap between the sparse and dense evaluations narrows.
@@
 \Cref{F:fd_vs_analytical_cpu} shows large speedups for both the OpenMP code and shallow-vectorized OpenCL code; the analytical OpenMP Jacobian is \SIrange{3.92}{8.67}{$\times$} faster, while the analytical OpenCL Jacobian achives speedups of \SIrange{17.22}{55.11}{$\times$}.
-It is noted that the isopentanol case was excluded here, as a single run of the sparse finite-difference Jacobian using either OpenCL or OpenMP took over \num{12} hours of runtime, in addition the current finite-difference formulation breaks Intel's auto-vectorizer, hence the OpenCL speedup comparison is made against the unvectorized OpenCL code; it is not a high priority to fix this issue, as the finite-difference Jacobian is implemented for comparison purposes only.
+It is noted that the isopentanol case was excluded here, as a single run of the sparse finite-difference Jacobian using either OpenCL or OpenMP took over \num{12} hours of run time, in addition the current finite-difference formulation breaks Intel's auto-vectorizer, hence the OpenCL speedup comparison is made against the unvectorized OpenCL code; it is not a high priority to fix this issue, as the finite-difference Jacobian is implemented for comparison purposes only.
 Although we do not display the dense finite-difference Jacobian speedup in~\cref{F:fd_vs_analytical_cpu}, the dense OpenCL and OpenMP analytical Jacobian codes out-perform the finite-difference variants by even larger margins; \SIrange{24.44}{245.63}{$\times$} for OpenCL, and \SIrange{9.68}{112.73}{$\times$} for OpenMP (these data do include the isopentanol model, though limited to \num{50000} conditions as discussed previously).

### 2018-03-10 paper.tex
edits

@@
 These factors often make the use of detailed chemical kinetics prohibitive in practice; e.g., in a recent direct numerical simulation~\cite{Spafford:2010aa} utilizing a 22-species model, evaluation of reaction rates consumed around half of the total run time.
-In addition, chemical kinetic Jacobian evaluation and factorization---required for most common implicit integration techniques utilized to deal with numerical stiffness---is a bottleneck for even modestly-sized chemical models in realistic reactive flow simulations, necessitating use of other cost-reduction strategies~\cite{LU2009192}.
+In addition, chemical kinetic Jacobian evaluation and factorization---utilized by most common implicit integration techniques to deal with numerical stiffness---is a bottleneck for even modestly-sized chemical models in realistic reactive flow simulations, necessitating use of other cost-reduction strategies~\cite{LU2009192}.
 
-A host of techniques exist to lessen the computational demand of chemical kinetic calculations while maintaining fidelity, falling broadly into a few categories; removal of unimportant species and reactions unimportant species and reactions~\cite{Lu:2006bb,Pepiot-Desjardins:2008,Hiremath:2010jw,Niemeyer:2010bt,Curtis:2015}, lumping of species with similar thermo-chemical properties~\cite{Lu:2007,Ahmed:2007fa,Pepiot:2008kq}, and time-scale methods that reduce chemical stiffness~\cite{Maas:1992ws,Lam:1994ws,Lu:2001ve,Gou:2010}.
+A host of techniques have been developed to lessen the computational demand of chemical kinetic calculations while maintaining fidelity, these fall broadly into a few categories: removal of unimportant species and reactions unimportant species and reactions~\cite{Lu:2006bb,Pepiot-Desjardins:2008,Hiremath:2010jw,Niemeyer:2010bt,Curtis:2015}, lumping of species with similar thermo-chemical properties~\cite{Lu:2007,Ahmed:2007fa,Pepiot:2008kq}, and time-scale methods that reduce chemical stiffness~\cite{Maas:1992ws,Lam:1994ws,Lu:2001ve,Gou:2010}.
 The interested reader is referred to the recent review of Tur\'anyi and Tomlin~\cite{turanyi2016analysis} for a comprehensive overview.
@@
 
-\begin{figure}[phtb]
+\begin{figure}[htb]
   \centering
@@
 
-SIMT processing---the foundation of modern GPUs~\cite{lindholm2008nvidia}---is a related computing paradigm that hosts large groups of threads on a single compute element (a streaming multiprocessor in NVIDIA terminology).
+SIMT processing---the foundation of modern GPUs~\cite{lindholm2008nvidia}---is a related computing paradigm that hosts large numbers of threads on a single compute element (a streaming multiprocessor in NVIDIA terminology).
 A group of threads---typically \num{32}, known as a warp on NVIDIA GPUs---execute the same SIMD instruction on multiple data concurrently (\cref{F:SIMT}).
@@
 Niemeyer et al.~\cite{Niemeyer:2016aa} created and validated the open-source analytical chemical kinetic Jacobian code-generator, \texttt{pyJac}, supporting parallel execution on the CPU and SIMT execution on the GPU; a speedup of \SIrange{3}{7.5}{$\times$} over a finite-difference Jacobian was found on the CPU.
-Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but it was not validated outside the context of use with an implicit-integration technique; in addition the Jacobian was based on an over-constrained system, and a differential algebraic equation (DAE) solver should be used for strict correctness\slash conservation.
+Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but it was not validated outside the context of use with an implicit-integration technique; in addition the Jacobian was based on an over-constrained system, the effect on strict conservation of mass\slash energy was not studied.
 
@@
 Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit fourth-order Runge--Kutta integrator for non-stiff chemical kinetic integration on a Tesla C2075 GPU, and found a speedup of nearly two orders of magnitude when compared with a sequential CPU-code on an Intel Xeon X5650 CPU.
-Shi et al.~\cite{Shi:2012aa} later extended their work to develop a stabilized-explicit solver on a Tesla C2050; paired with a traditional implicit integrator on quad-core Intel i7 930 which handled the stiffest computational cells; a \SIrange{11}{46}{$\times$} speedup was achieved over the base implicit CPU integrator during simulation of a three-dimensional premixed diesel engine simulation.
+Shi et al.~\cite{Shi:2012aa} later extended their work to develop a stabilized-explicit solver on a Tesla C2050, paired with a traditional implicit integrator on quad-core Intel i7 930 which handled the stiffest computational cells; a \SIrange{11}{46}{$\times$} speedup was achieved over the base implicit CPU integrator during simulation of a three-dimensional premixed diesel engine simulation.
 Le et al.~\cite{Le2013596} implemented two high-order shock-capturing reactive-flow codes on a Tesla C2070, and found a \numrange{30}{50}$\times$ speedup over a sequential CPU version of the same code on a Intel Xeon X5650.
@@
 Niemeyer and Sung~\cite{Niemeyer:2014aa} later developed a stabilized explicit second-order Runge--Kutta--Chebyshev algorithm on a Tesla C2075, demonstrating an order of magnitude speedup for a over a CPU implementation of VODE on a six-core Intel X5650 for moderately stiff chemical kinetics.
-Sewerin and Rigopoulos~\cite{Sewerin20151375} implemented a three-stage\slash fifth-order implicit Runge--Kutta method~\cite{wanner1991solving} on both a high-end (Nvidia Quadro 6000) and consumer-grade (Nvidia Quadro 600) GPUs, as well as a standard CPU (two-core, four-thread Intel i5-520M) and a scientific workstation (eight-core, 16-thread Intel Xeon E5-2687W); the high-end GPU was at best \SI{1.8}{$\times$} slower than the workstation CPU (16 threads), while the consumer level GPU was at best \SI{5.5}{$\times$} slower than the corresponding standard CPU (four threads).
+Sewerin and Rigopoulos~\cite{Sewerin20151375} implemented a three-stage\slash fifth-order implicit Runge--Kutta method~\cite{wanner1991solving} on both a high-end (NVIDIA Quadro 6000) and consumer-grade (Nvidia Quadro 600) GPUs, as well as a standard CPU (two-core, four-thread Intel i5-520M) and a scientific workstation (eight-core, 16-thread Intel Xeon E5-2687W); the high-end GPU was at best \SI{1.8}{$\times$} slower than the workstation CPU (16 threads), while the consumer level GPU was at best \SI{5.5}{$\times$} slower than the corresponding standard CPU (four threads).
 Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs) on an Intel Xeon E5-2680 CPU and a Tesla K20 GPU.
-In evaluation of the thermo-chemical properties, speedups between \SIrange{8}{13}{$\times$} were on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU respectively.
+In evaluation of the thermo-chemical properties, speedups between \SIrange{8}{13}{$\times$} were found (over a serial version of the same code) while running on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU respectively.
 Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two-fourth order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on a Tesla C2075; the implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on 12--38 Intel Xeon E5-4640 v2 CPU cores for two smaller chemical models with an integration time step of \SI{e-6}{$\sec$}.
@@
 Linford et al.~\cite{Linford:2011} implemented a three-stage, second-order Rosenbrock integrator for atmospheric chemical kinetics on the CPU, GPU and cell croadband engine (CBE)---a specially designed vector processor---and found speedups regularly exceeding~\SI{25}{$\times$} over a serial CPU implementation.
-Kroshko and Spiteri~\cite{kroshko2013efficient} implemented a SIMD-vectorized third-order stiff Rosenbrock integrator for atmospheric chemistry on the CBE finding a speedup of \SI{1.89}{$\times$} (a parallel scaling efficiency of \SI{0.94}{$\percent$}) over a serial version of the same code.
+Kroshko and Spiteri~\cite{kroshko2013efficient} implemented a SIMD-vectorized third-order stiff Rosenbrock integrator for atmospheric chemistry on the CBE and found a speedup of \SI{1.89}{$\times$} (a parallel scaling efficiency of \SI{0.94}{$\percent$}) over a serial version of the same code.
 Stone et al.,~\cite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
@@
 \end{itemize}
-To our knowledge, this is the first effort that enables vectorized chemical-kinetic source-term and Jacobian evaluation for any chemical model on the CPU, GPU and other accelerators.
+To our knowledge, this is the first effort that enables vectorized chemical-kinetic source-term and Jacobian evaluation for any chemical model on a wide-selection of platforms, i.e., the CPU, GPU and other accelerators.
 
@@
 When storing arrays for a chemical kinetic model, the data-storage layout and vectorization patterns are critical to achieving high-performance code.
-\Cref{F:mem} depicts an example data array with K-rows and N-columns and index ($i$, $j$) corresponding to the $i$th row and $j$th column.
-For example the concentration of species $j$ for the $i$-th thermo-chemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{state}$, the number of thermo-chemical states considered for evaluation, and $1 \le j \le \ns/$, the number of species in the model.
+\Cref{F:mem} depicts an example data array with K-rows and N-columns, with index ($i$, $j$) corresponding to the $i$th row and $j$th column.
+For example the concentration of species $j$ for the $i$th thermo-chemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{state}$, the number of thermo-chemical states considered for evaluation, and $1 \le j \le \ns/$, the number of species in the model.
 The stored concentrations would then have $K = N_{\text{state}}$ rows, and $N = N_{\text{sp}}$ columns.
@@
 Conversely in the ``F'' (Fortran column-major) format, the concentrations of a single species $j$ over all thermo-chemical states are adjacent in memory, corresponding to storing $[C]_{1, 1}$ in index \num{1}, $[C]_{2, 1}$ in index \num{2} and so on, as shown in \cref{F:column_major}.
-This ordering---along with the device (CPU, GPU, etc.) and vectorization pattern in question---have a large effect on the performance of SIMD\slash SIMT-vectorized algorithms.
+This ordering---along with the device (CPU, GPU, etc.\@) and vectorization pattern in question---have a large effect on the performance of SIMD\slash SIMT-vectorized algorithms.
 
-In a \textit{shallow}-SIMD\slash SIMT vectorization (also referred to as ``per-thread'' in previous works~\cite{Stone:2013aa} using GPUs), each SIMD-lane (or SIMT-thread) in a compute-unit evaluates thermo-chemical source terms\slash chemical kinetic Jacobian for a different thermo-chemical state.
-If the data is stored in ``F''-order, the SIMD-lanes\slash SIMT-thread accessing $[C]_{1, j}\ldots[C]_{N_v, j}$---the concentration of species $j$ for states $1, 2,\ldots N_v$, where $N_v$ is the size of the vector lane for SIMD or the number of threads in a SIMT warp---will load sequential locations in memory.
+In a \textit{shallow}-SIMD\slash SIMT vectorization (also referred to as ``per-thread'' in previous works~\cite{Stone:2013aa} using GPUs), each SIMD-lane (or SIMT-thread) in a compute-unit evaluates the thermo-chemical source terms\slash chemical kinetic Jacobian for a different thermo-chemical state.
+If the data is stored in ``F''-order, the SIMD-lanes\slash SIMT-thread accessing $[C]_{1, j}\ldots[C]_{N_v, j}$---the concentration of species $j$ for states $1, 2,\ldots N_v$, where $N_v$ is the SIMD vector-width or the number of threads in a SIMT warp---will load sequential locations in memory.
 The first $(j+1)$th species concentration, $[C]_{1, j+1}$, will be $N_{\text{state}}$ memory locations away; this increases the likelihood of cache-misses on the CPU~\cite{gray2000rules}, but conversely is well suited to coalesced memory accesses on the GPU~\cite{NVIDIA:2018}.
 
-In a \textit{deep}-SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works), a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermo-chemical source terms for a single thermo-chemical state; thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + N_v}$---the species concentrations for state 1, species $j \ldots j + L$---will access sequential memory locations if the data is stored in ``C''-order.
+In a \textit{deep}-SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works), a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermo-chemical source terms for a single thermo-chemical state; thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + N_v}$---the species concentrations for state 1, species $j \ldots j + N_v$---will access sequential memory locations if the data is stored in ``C''-order.
 Further, in the ``C''-ordering the furthest difference between any two species concentrations within the same thermo-chemical state is at most $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU, but may lead to uncoalesced memory-accesses on the GPU.
@@
 This is accomplished by splitting the slower-varying axis of the data array---columns for ``C''-ordering, and rows for ``F''-ordering---into chunks of size $N_v$ (the SIMD vector-width, or SIMT warp-size) and laying these data out contiguously in memory.
-For example, using the data-ordering in~\cref{F:row_major_split}, the concentration of species $j$ for states $i\ldots i+N_v$, $[C]_{i, j} \ldots [C]_{i + N_v, j}$, are contiguous in memory and are followed by the concentration of species $j + 1$ for the same states, $[C]_{i, j + 1} \ldots [C]_{i + N_v, j + 1}$.
+For example, using the shallow-vectorized ``C''-ordering pictured in~\cref{F:row_major_split}, the concentration of species $j$ for states $i\ldots i+N_v$, $[C]_{i, j} \ldots [C]_{i + N_v, j}$, are contiguous in memory and are followed by the concentration of species $j + 1$ for the same states, $[C]_{i, j + 1} \ldots [C]_{i + N_v, j + 1}$.
 This pattern, similar to OpenCL's native vector data-types---e.g., \texttt{double\num{8}} which treats \num{8} contiguous double precision floating point numbers as a single vector datum---ensures that any SIMD operation occurs on data that is contiguous in memory, greatly improving caching and SIMD-throughput.
@@
 First, the state vector results in highly sparse chemical kinetic Jacobians as will be detailed in~\cref{S:sparsity}.
-Second, mass is explicitly conserved in this formulation owing to the fact that the moles of the final species is calculated from the ideal gas law, and the rate of change of the final species from conservation of mass (see \cref{deriv-s:state,deriv-s:source} of the supplemental material respectively); additionally, system is not over-constrained and does not require use of a more complicated DAE solver (as compared to an regular ODE solver) for integration\todo{hansen citation here}.
+Second, mass is explicitly conserved in this formulation owing to the fact that the moles of the final species is calculated from the ideal gas law, and the rate of change of the final species from conservation of mass (see \cref{deriv-s:state,deriv-s:source} of the supplemental material respectively); additionally, system is not over-constrained and does not require use of a more complicated DAE solver (as compared to an ODE integrator) for integration\todo{hansen citation here}.
 Finally, the chemical kinetic Jacobian changes relatively little between the \conp/ and \conv/ forms, making maintaining the code-base much simpler.
@@
 \end{equation}
-where $\mathcal{R}$ is the ideal gas-constant, the volume and pressure source terms can be found:
+where $\mathcal{R}$ is the ideal gas-constant in molar units, the volume and pressure source terms can be found:
 \loadeq{dparam}
@@
 The molar source terms are more complicated due to the derivatives of the net rate of progress---see~\cref{deriv-s:dropi}---and pressure modification terms---\cref{deriv-s:dci}---of the reactions.
-The reader is again encouraged to read the supplemental material covering the exact derivations and forms of the reactions solved by \texttt{pyJac}.
+The reader is again encouraged to read the supplemental material covering the exact derivations and equations solved by \texttt{pyJac}.
 The derivative of the molar source term with respect to temperature is:
@@
 Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates user specified psuedo-code and data to OpenCL\slash C-code.
-As the name implies, \texttt{loo.py} generates code based on for-loops, this differs from the previous version of \texttt{pyJac}\cite{pyjac16} which generates static code (i.e., fully unrolled loops, with thermodynamic\slash reaction parameters written directly in code rather than stored in arrays).
-This caused some issues with large file sizes, long compilation times and even occasionally breaking the \texttt{gcc} and \texttt{nvcc} compilers (the latter issue necessitated breaking the Jacobian\slash source-term evaluations into separate files)\cite{Niemeyer:2016aa}.
+As the name implies, \texttt{loo.py} generates code based on for-loops, this differs from the previous version of \texttt{pyJac}~\cite{pyjac16} which generates static code (i.e., fully unrolled loops, with thermodynamic\slash reaction parameters written directly in code rather than stored in arrays).
+In our previous work,~\cite{Niemeyer:2016aa} this static code generation caused some issues with large file sizes, long compilation times and even occasionally breaking the \texttt{gcc} and \texttt{nvcc} compilers (the latter issue necessitated breaking the Jacobian\slash source-term evaluations into separate files).
 The implications of this change will be studied in~\cref{S:jacobian_results}, where the performance of the new version of~\texttt{pyJac} will be compared to the previous versions.
 
-In addition, \texttt{loo.py} allows the user to more easily make changes the structure of the generated program, e.g., data ordering, vectorization, threading patterns, as well as switching the target-language (and simpler extension to additional languages, e.g., CUDA).
+In addition, \texttt{loo.py} allows the user to more easily make changes the structure of the generated program, e.g., data ordering, vectorization, and threading patterns, as well as switching the target-language for code-generation (and simpler extension to additional languages, e.g., CUDA).
 Further, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source terms.
@@
 \caption{The Intel CPU configurations used in this study.
-	 The vector width reported in this table is reported in (ideal) number of double operations per SIMD instruction, as this will be used in measuring SIMD efficiency; for reference, the vector width in bits of the \sse/ and \avx/ machines are \SI{128}{\bit} and \SI{256}{\bit} respectively.
+	 The vector-width reported in this table is reported in (ideal) number of double operations per SIMD instruction, as this will be used in measuring SIMD efficiency; for reference, the vector-width in bits of the \sse/ and \avx/ machines are \SI{128}{\bit} and \SI{256}{\bit} respectively.
 	 The identifier field will be used as a shorthand descriptor in the performance plots to quickly identify the CPU type.}
@@
 \Cref{t:gpus} lists the model, number of CUDA cores and NVIDIA driver of each of the GPUs used in this work.
-NVIDIA's OpenCL runtime is bundled with the NVIDIA driver~\cite{NVIDIA:2018}, hence we use driver version to specify the OpenCL runtime version.
+NVIDIA's OpenCL runtime is bundled with the NVIDIA driver~\cite{NVIDIA:2018}, hence we the driver version is used to specify the OpenCL runtime version.
 
@@
 
-In Table~\ref{T:source_error}, we see the results of this code as compared to Cantera~\cite{Cantera} on the library of PaSR conditions (\cref{t:models}).
+In Table~\ref{T:source_error}, the results of this code's source-rate evaluations are compared to Cantera~\cite{Cantera} on the library of PaSR conditions (\cref{t:models}).
 Very close agreement is observed for the forward and reverse ROP for all models, however the error norm is \textasciitilde\numrange{3}{4} orders of magnitude larger for the isopentanol model.
-This results from differences in evaluation of P-Log reactions between \texttt{pyJac} and Cantera; namely \texttt{pyJac} computes the logarithm of the upper and lower reaction Arrhenius rates (see~\cref{deriv-s:pdep}) analytically while Cantera evaluates this term numerically.
+This results from differences in evaluation of P-Log reactions between \texttt{pyJac} and Cantera; namely \texttt{pyJac} computes the logarithm of the upper and lower reaction Arrhenius rates (see supplemental~\cref{deriv-s:pdep}) analytically while Cantera evaluates this term numerically.
 If the error from P-Log reactions are not considered in~\cref{e:rel_err}, the forward and reverse ROP falls to \num{5.44e-08} and \num{1.59e-07} respectively.
@@
 In~\cref{T:source_error}, we see that $E_{\varepsilon}$ is very similar in magnitude to $E_R$ in all cases, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
-The molar species production rate error norm is similar in magnitude to that of the net ROP, however the temperature and state parameter rate errors again increases as the error in net species production rates is amplified by multiplication with thermodynamic properties.
+The molar species production rate error norm is similar in magnitude to that of the net ROP, however the temperature and state parameter rate errors again increase as the error in net species production rates is amplified by multiplication with thermodynamic properties.
 Again, we note that the above discussion does not imply that discrepencies in computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
@@
 The most stringent filtered error norm ($\mathcal{C} = 10^{20}$) ranges from \numrange[retain-zero-exponent=true]{e-7}{e0}; the largest error is for the \ce{H2}\slash\ce{CO} model.
-For this model, $\mathcal{T}$  is smaller than the tolerance of \num{e20}, and hence the error norm considers smaller than $\mathcal{O}(1)$.
+For this model, $\mathcal{T}$  is smaller than the tolerance of \num{e20}, and hence the error norm considers Jacobian entries smaller than $\mathcal{O}(1)$.
 It is noted that for GRI-Mech 3.0, which has a $\mathcal{T}$ roughly an order of magnitude larger, the stringent error norm is significantly smaller: $\mathcal{O}(\num{e-7})$.
@@
 Further, the LAPACK error norm does not correlate well with the stringent filtered error norm, e.g., USC-Mech II has the lowest LAPACK error norm (\num{9.632e-15}), but the second largest stringent filtered error norm (\num{1.119e-2}).
-Conversely, the LAPACK error norm does seem to correlate with the relaxed filtered error norm; the two models with the largest LAPACK error, \ce{H2}\slash\ce{CO} and \ce{iC5H11OH}, have the largest relaxed filtered error norms ($\mathcal{O}(\num{e-5})$), again suggesting that the stringent error norm is influenced by floating point error accumulation.
+Conversely, the LAPACK error norm does seem to correlate with the relaxed filtered error norm; the two models with the largest LAPACK error, \ce{H2}\slash\ce{CO} and \ce{iC5H11OH}, have the largest relaxed filtered error norms, $\mathcal{O}(\num{e-5})$, again suggesting that the stringent error norm is influenced by floating point error accumulation.
 These findings, along with the individual unit-testing of all chemical source terms and Jacobian sub-components described in~\cref{s:unittest}, gives high confidence in the correctness of this new version of \texttt{pyJac}.
@@
 Currently, \texttt{pyJac} can utilize two common sparse matrix storage formats~\cite{netlib_templates}, compressed row\slash column storage (CSR\slash CSC), used for ``C'' and ``F''-ordered data respectively.
-Tor convenience we will outline only the CSR format but the CSC format is very similar~\cite{netlib_templates}.
+For convenience we will outline only the CSR format but the CSC format is very similar~\cite{netlib_templates}.
 An $N\times N$ CSR matrix is composed of three vectors, a value vector of length $NNZ$---the number of non-zero elements in the Jacobian---which stores the elements of the Jacobian, a row pointer vector---of length $N$---which stores the locations in the value vector that begin a row, and a column index vector---of length $NNZ$---which stores the column indicies of the elements in the value vector.
@@
 The performance studies in this work were run on the platforms listed in~\cref{t:cpus,t:gpus}.
-Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
+Run times in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
 The OpenMP Jacobian\slash source-rate kernels, as well as the OpenMP\slash OpenCL wrapping code---responsible for initializing\slash transferring memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms, and \texttt{gcc v4.8.5} on the \sse/\slash\gpuold/ machines.
@@
 Additionally, the exact form of the Jacobian (as opposed the the ``approximate'' form discussed in~\cref{S:sparsity}) was used in all cases.
-Finally, unless stated otherwise: the performance results utilized the \conp/ assumption, a vector width of \num{8}\slash\num{128} and ``C''\slash ``F''-ordered data for the CPU\slash GPU cases respectively; the run times reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
+Finally, unless stated otherwise: the performance results utilized the \conp/ assumption, a vector-width of \num{8}\slash\num{128} and ``C''\slash ``F''-ordered data for the CPU\slash GPU cases respectively; the run times reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
 The effects of choice of vector-width, data-ordering and differences between \conp/ and \conv/ evaluations on the CPU\slash GPU will be explored in~\cref{S:source_results}.
@@
 On both machines, the unvectorized OpenCL code is faster than the baseline parallel OpenMP code; by \SIrange{1.30}{1.51}{$\times$} on the \avx/ CPU, and \SIrange{1.25}{1.27}{$\times$} on the \sse/ machine.
-Additionally, the shallow-vectorized OpenCL code is \textasciitilde\SIrange{2.53}{2.92}{$\times$} and \SIrange{3.40}{4.08}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
+Additionally, the shallow-vectorized OpenCL code is \SIrange{2.53}{2.92}{$\times$} and \SIrange{3.40}{4.08}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
+
 In contrast,~\cref{F:pocl_source} shows the mean run time per condition of deep, shallow and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
-No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP case is faster in both cases.
+No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP case is faster on both CPUs.
 The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work; however, briefly speaking vectorization might be achieved on POCL via a combination of fixed-array strides---rather than (potentially) varying with the number of initial conditions being evaluated, as currently is implemented---use of relaxed math (``\texttt{-cl-fast-relaxed-math}''), explicit OpenCL vector types and newer versions of LLVM\slash POCL (specifically, potential improvements to the vectorization module).
@@
 In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean run time per condition; the mean run time continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
-Interestingly, the last data-point for the GRI-Mech 3.0 shows an increase in run time per condition, this is likely due to multiple data transfers to\slash from the GPU required to complete the source-rates in this case; similar (if smaller) increases in run time per condition can be seen for the \ce{iC5H11OH} model near \num{e4} and \num{e5} conditions.\todo{rerunning}
-\Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} for two different vector widths (a.k.a., the block size in NVIDIA terminology), \num{64} and \num{128}; the fastest \gpunew/ case (a vector width of 128) is \SIrange{1.34}{4.89}{$\times$} faster than the slowest case (either \gpunew/ or \gpuold/ with a vector width of \num{64}) depending on the chemical model in question.
-Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector width has minimal effect on performance for most \gpuold/ cases (excepting \ce{H2}\slash\ce{CO}), but a large effect on the performance of the \gpunew/ GPU; this is likely caused by higher occupancy~\cite{occupancy} on the \gpunew/ GPU, however it is unclear exactly how NVIDIA's OpenCL runtime balances the registers\slash warps per streaming-multiprocessor, as controls occupancy in CUDA~\cite{occupancy}; this is a topic that merits future investigation.
+Interestingly, the last data-point for the GRI-Mech 3.0 shows an increase in run time per condition, this is likely due to multiple data transfers to\slash from the GPU required to complete the source-rates in this case; similar (if smaller) increases in run time per condition can be seen for the \ce{iC5H11OH} model near \num{e4} and \num{e5} conditions.\todo{Update all this after I get rerun w/ less dumb copies}
+\Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} for two different vector-widths (a.k.a., the block size in NVIDIA terminology), \num{64} and \num{128}; the fastest \gpunew/ case (a vector-width of 128) is \SIrange{1.34}{4.89}{$\times$} faster than the slowest case (either \gpunew/ or \gpuold/ with a vector-width of \num{64}) depending on the chemical model in question.
+Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector-width has minimal effect on performance for most \gpuold/ cases (excepting \ce{H2}\slash\ce{CO}), but a large effect on the performance of the \gpunew/ GPU; this is likely caused by higher occupancy~\cite{occupancy} on the \gpunew/ GPU, however it is unclear exactly how NVIDIA's OpenCL runtime balances the registers\slash warps per streaming-multiprocessor, as controls occupancy in CUDA~\cite{occupancy}; this is a topic that merits future investigation.
 \begin{figure}[htbp]
@@
       \includegraphics[width=\textwidth]{source_gpu_cvsf.pdf}
-      \caption{The affect on performance of ``C'' vs ``F''-ordering for the shallow-vectorized NVIDIA OpenCL code on both GPUs with a vector width of 128.
+      \caption{The affect on performance of ``C'' vs ``F''-ordering for the shallow-vectorized NVIDIA OpenCL code on both GPUs with a vector-width of 128.
 The speedup presented is calculated per-GPU to better assess the effect of the data-ordering on each GPU.}
@@
 In general, the ``C'' and ``F''-ordered shallow-vectorizations are nearly equivalent on both GPUs while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, with the exception of the isopentanol model where the ``C''-ordered data is up to \SI{1.1}{$\times$} faster on the \gpunew/ (we note that there was high variablity for the \gpuold/ in this case).
-\todo{again, pending rerun on \gpuold/}
+\todo{again, pending rerun on with less copying}
 The rough equivalence between the ``C'' and ``F''-ordered codes GPUs is counter to what one might expect, as typically speaking coalesced memory access in a shallow-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
-However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector width size and thus, likely coalesced.
+However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector-width size and thus, likely coalesced.
 % but the improved data-locality, coupled with the larger cache sizes of the \gpunew/ GPU result in this modest speedup for ``C''-data ordering.
@@
 It is unclear why the vector-width of \num{4} results in no speedup at all (it is in fact, the slowest case); Intel's vectorization guide~\cite{intel_vecknobs} mentions that a heuristic determines the optimal vector-width (in this case, it appears from compiler output to be \num{8}), so it is possible using vector-width smaller than the heuristic breaks the implicit vectorizer.
-It is noted that this issue does not occur for a vector width of \num{4} on the \sse/ CPU.
+It is noted that this issue does not occur for a vector-width of \num{4} on the \sse/ CPU.
+
+Finally,~\cref{F:source_scaling} displays the (strong) parallel scaling efficiency and SIMD efficiency for the CPU platforms.
+The strong parallel scaling efficiency is defined as:
+\begin{equation}
+ \label{e:strong_scaling}
+ \varepsilon = \frac{\bar{t}_{1}}{N \bar{t}_{N}}
+\end{equation}
+where $\bar{t}_{N}$ is the mean run time per condition on $N$ CPU cores, and $\bar{t}_{1}$ the same on a single CPU core.
+The strong parallel scaling efficiency measures the speedup due to the use of additional CPU cores, as a fraction of linear speedup; strong scaling tends to decrease with the number of processors used due to memory bandwidth limitations and decreasing computation work allocated per CPU core~\cite{strong_scaling}.
 
@@
 
-Finally,~\cref{F:source_scaling} displays the (strong) parallel scaling efficiency and SIMD efficiency for the CPU platforms.
-The strong parallel scaling efficiency is defined as:
-\begin{equation}
- \label{e:strong_scaling}
- \varepsilon = \frac{\bar{t}_{1}}{N \bar{t}_{N}}
-\end{equation}
-where $\bar{t}_{N}$ is the mean run time per condition on $N$ CPU cores, and $\bar{t}_{1}$ the same on a single CPU core.
-The strong parallel scaling efficiency measures the speedup due to the use of additional CPU cores, as a fraction of linear speedup; strong scaling tends to decrease with the number of processors used due to memory bandwidth limitations and decreasing computation work allocated per CPU core~\cite{strong_scaling}.
-
 \Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source-rate in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
@@
 Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.
-This is due to better performance of the shallow-vectorized OpenCL code, coupled with the relatively small amounts of work associated with source-rate.
-To illustrate this,~\cref{F:source_crossover} shows the mean run time per-condition of the OpenMP source term evaluations for \numrange{2}{16} cores, compared to the shallow-vectorized OpenCL code on \num{1} (solid lines) and \num{4} (dashed lines) four cores on the \avx/ machine.
+This is due in large part to the superior performance of the shallow-vectorized OpenCL code, coupled with the relatively small amounts of work associated with source-rate evaluation.
+To illustrate this,~\cref{F:source_crossover} shows the mean run time per-condition of the OpenMP source term evaluations for \numrange{2}{16} cores, compared to the shallow-vectorized OpenCL code on \num{1} (solid line) and \num{4} (dashed line) four cores on the \avx/ machine.
 For all chemical models, the mean run time per-condition (and hence the computational work allocated per-core, one of the key-drivers of parallel scaling efficiency~\cite{strong_scaling}) of OpenMP running on \num{4} cores is roughly equal to that of the shallow-vectorized OpenCL code on a single core.
@@
 \end{equation}
-where $\bar{t}_{\text{unvec}}$ is the mean run time per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$, the vector width reported in number of double operations in~\cref{t:cpus}.
+where $\bar{t}_{\text{unvec}}$ is the mean run time per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$, the vector-width reported in number of double operations in~\cref{t:cpus}.
 This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.
@@
 This again demonstrates that the SIMD-vectorization becomes efficient as the amount of work to perform increases (with increasing model size).
-Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one; this is likely caused by a combination of use of an OpenCL vector-width greater than the native CPU vector-width (i.e., \num{8} v.s., \num{2}), and the improved data-locality for the vectorized-data ordering, as discussed in~\cref{S:data}, and results a modest \SIrange{7}{17}{\percent} improvement over the nominal vector width.
+Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one; this is likely caused by a combination of use of an OpenCL vector-width greater than the native CPU vector-width (i.e., \num{8} v.s., \num{2}), and the improved data-locality for the vectorized-data ordering, as discussed in~\cref{S:data}, and results a modest \SIrange{7}{17}{\percent} improvement over the nominal vector-width.
 
@@
 In this case, the isopentanol model has the worst scaling for both codes, it is noted that due to the sheer size of the dense isopentanol Jacobian---storing the dense matrix for a single thermo-chemical state takes over \SI{1}{\mega\byte} of data---the total number of states for the dense isopentanol Jacobian evaluation was limited to \num{50000} (which requires over \SI{50}{\giga\byte} of memory); this greatly drops the computation cost for this case, and adversely affects the scaling efficiency as was discussed in~\cref{S:source_results}.
-Excluding the isopentanol model, the dense shallow-vectorized OpenCL scalings are slightly higher than for source-rate evaluation, e.g., \numrange{0.70}{0.78} and \numrange{0.52}{0.61} for \num{4} and \num{8} cores respectively (compared to \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for shallow-vectorized OpenCL source-rate on \num{4} and \num{8} cores respectively); this is due to the higher computational cost of Jacobian evaluation, and hence more available work per CPU-core.
+Excluding the isopentanol model, the dense shallow-vectorized OpenCL scalings are slightly higher than for source-rate evaluation, e.g., \numrange{0.70}{0.78} and \numrange{0.52}{0.61} for \num{4} and \num{8} cores respectively (compared to \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for shallow-vectorized OpenCL source-rate on the same \num{4} and \num{8} cores); this is due to the higher computational cost of Jacobian evaluation, and hence more available work per CPU-core.
 As in~\cref{S:source_results}, we note that the shallow-vectorized dense Jacobian code running on \num{1} and \num{4} cores of the \avx/ machine is roughly as fast as the OpenMP code on \num{4} and \num{16} cores respectively; the parallel scaling efficiency of OpenMP at \num{16} cores (excluding isopentanol as previously) is \numrange{0.63}{0.71}, similar to the shallow-vectorization's efficiency of \numrange{0.70}{0.78} for \num{4} cores.
@@
 \Cref{F:fd_vs_analytical_cpu} shows large speedups for both the OpenMP code and shallow-vectorized OpenCL code; the analytical OpenMP Jacobian is \SIrange{3.92}{8.67}{$\times$} faster, while the analytical OpenCL Jacobian achives speedups of \SIrange{17.22}{55.11}{$\times$}.
-It is noted that the isopentanol case was excluded here, as a single run of the sparse finite-difference Jacobian using either OpenCL or OpenMP took over \num{12} hours of run time, in addition the current finite-difference formulation breaks Intel's auto-vectorizer, hence the OpenCL speedup comparison is made against the unvectorized OpenCL code; it is not a high priority to fix this issue, as the finite-difference Jacobian is implemented for comparison purposes only.
+It is noted that the isopentanol case was excluded here, as a single run of the sparse finite-difference Jacobian using either OpenCL or OpenMP took over \num{12} hours of run time.
+In addition the current finite-difference formulation breaks Intel's auto-vectorizer, hence the OpenCL speedup comparison is made against the unvectorized OpenCL code; it is not a high priority to fix this issue, as the finite-difference Jacobian is implemented for comparison purposes only.
 Although we do not display the dense finite-difference Jacobian speedup in~\cref{F:fd_vs_analytical_cpu}, the dense OpenCL and OpenMP analytical Jacobian codes out-perform the finite-difference variants by even larger margins; \SIrange{24.44}{245.63}{$\times$} for OpenCL, and \SIrange{9.68}{112.73}{$\times$} for OpenMP (these data do include the isopentanol model, though limited to \num{50000} conditions as discussed previously).
 A comparison of the sparse analytical and finite-difference Jacobians is presented in~\cref{F:fd_vs_analytical_gpu}.
-The analytical Jacobian on the \gpunew/ shows large speedups, increasing with chemical model size, in the range of \SIrange{4.66}{17.43}{$\times$}; however, the \gpuold/ GPU shows almost no speedup, it is not clear why this is the case, it is possible that the finite-difference Jacobian evaluation does not function correctly on the \gpuold/ GPU.
+The analytical Jacobian on the \gpunew/ and \gpuold/ shows large (and very similar) speedups, increasing with chemical model size, in the range of \SIrange{4.66}{17.43}{$\times$}.
 Although not pictured, the dense analytical-Jacobian on the \gpunew/ GPU has larger speedups compared with the dense finite-difference Jacobian: \SIrange{5.32}{44.87}{$\times$}.
@@
 \Cref{F:v1_vs_v2} compares the performance for evaluation of the dense-analytical Jacobian of the new version of \texttt{pyJac} to the previous version~\cite{pyjac16} on the \sse/ CPU and \gpuold/ GPU; dense Jacobian evaluation was selected as it was the only type implemented in the previous version of \texttt{pyJac}.
-On the \sse/ CPU, the \texttt{pyJac-v2} is faster than \texttt{pyJac-v1} for OpenMP evaluation for the larger chemical models; the static code generated by \texttt{pyJac-v1} (see~\cref{s:unittest}) is \SI{1.79}{$\times$} faster for the \ce{H2}\slash\ce{CO} model, and only \SI{1.09}{$\times$} slower for the GRI-Mech 3.0 model.
+On the \sse/ CPU, the \texttt{pyJac-v2} is faster than \texttt{pyJac-v1} for OpenMP evaluation for the larger chemical models; the static OpenMP code generated by \texttt{pyJac-v1} (see~\cref{s:unittest}) is \SI{1.79}{$\times$} faster for the \ce{H2}\slash\ce{CO} model, and only \SI{1.09}{$\times$} slower for the GRI-Mech 3.0 model.
 In contrast, the loop-based OpenMP code of \texttt{pyJac-v2} is \SIrange{3.37}{10.19}{$\times$} faster than \texttt{pyJac-v1} for the USC-Mech II and isopentanol models.
-The shallow-vectorized OpenCL code is faster than \texttt{pyJac-v1} in all cases, achieving speedups of \SI{1.37}{19.56}{$\times$} increasing with model size.
+The shallow-vectorized OpenCL code is faster than \texttt{pyJac-v1} in all cases, achieving speedups of \SIrange{1.37}{19.56}{$\times$} increasing with model size.
 In~\cref{F:v1_vs_v2_gpu}, the performance of the \texttt{pyJac-v2} dense-analytical Jacobian on the \gpuold/ GPU is compared to \texttt{pyJac-v1} on the same.

### 2018-03-10 paper.tex
fix

@@
 This provides a particularly vexing problem for the programmer, as there is often little that can be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
-Further, as noted throughout this work certain transformations or code-generation patterns can break OpenCL execution\slash vectorization; e.g., the failure of POCL to achieve vectorization, Intel OpenCL's failure to vectorize the finite-difference Jacobian, the lack of speedup for the finite-difference Jacobian on the NVIDIA runtime for the \gpuold/ GPU.
+Further, as noted throughout this work certain transformations or code-generation patterns can break OpenCL execution\slash vectorization; e.g., the failure of POCL to achieve vectorization, Intel OpenCL's failure to vectorize the finite-difference Jacobian, the lack of speedup for the finite-difference Jacobian on the NVIDIA runtime for the \gpuold/ GPU\todo{this went away?}.
 Indeed the vectorization attempted by the OpenCL runtimes (or reasons for incorrectly vectorized code) are not always obvious to the user.
-It is possible, particularly for the Intel OpenCL \slash POCL runtimes, that better performance might be achieved using so called ``explicit'' vectorization, i.e., through use of inbuilt vector types such as the \texttt{double8}.
+It is possible, particularly for the Intel OpenCL \slash POCL runtimes, that better performance might be achieved using so called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}.
 This will be investigated in future works.
@@
 In addition, CUDA~\cite{NVIDIA:2018} and Intel's open-source OpenCL alternative ISPC~\cite{pharr2012ispc} have been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017} and (in the case of ISPC) initial testing.
-The current deep-vectorization formulation would be executable for a CUDA target---as CUDA implements atomic double-precision floating-point operations---further recommending its use.
-Extending execution to these targets to make CPU\slash GPU more reliable vectorization is also a future goal.
+The current deep-vectorization formulation would be executable for both CUDA and ISPC targets, as these languages implement double-precision atomic operations, further recommending their use.
+Extending execution to these targets to make CPU\slash GPU more reliable vectorization is a future goal.
 

### 2018-03-12 paper.tex
spelling

@@
 Niemeyer and Sung~\cite{Niemeyer:2014aa} later developed a stabilized explicit second-order Runge--Kutta--Chebyshev algorithm on a Tesla C2075, demonstrating an order of magnitude speedup for a over a CPU implementation of VODE on a six-core Intel X5650 for moderately stiff chemical kinetics.
-Sewerin and Rigopoulos~\cite{Sewerin20151375} implemented a three-stage\slash fifth-order implicit Runge--Kutta method~\cite{wanner1991solving} on both a high-end (NVIDIA Quadro 6000) and consumer-grade (Nvidia Quadro 600) GPUs, as well as a standard CPU (two-core, four-thread Intel i5-520M) and a scientific workstation (eight-core, 16-thread Intel Xeon E5-2687W); the high-end GPU was at best \SI{1.8}{$\times$} slower than the workstation CPU (16 threads), while the consumer level GPU was at best \SI{5.5}{$\times$} slower than the corresponding standard CPU (four threads).
+Sewerin and Rigopoulos~\cite{Sewerin20151375} implemented a three-stage\slash fifth-order implicit Runge--Kutta method~\cite{wanner1991solving} on both a high-end (NVIDIA Quadro 6000) and consumer-grade (NVIDIA Quadro 600) GPUs, as well as a standard CPU (two-core, four-thread Intel i5-520M) and a scientific workstation (eight-core, 16-thread Intel Xeon E5-2687W); the high-end GPU was at best \SI{1.8}{$\times$} slower than the workstation CPU (16 threads), while the consumer level GPU was at best \SI{5.5}{$\times$} slower than the corresponding standard CPU (four threads).
 Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs) on an Intel Xeon E5-2680 CPU and a Tesla K20 GPU.
@@
 
-The Jacobian entries computed by \texttt{pyJac} are arranged such that entry ($i$, $j$) corresponds to the partial derivative of the $i$-th source term in~\cref{e:source_terms} by the $j$-th state variable in~\cref{e:state}:
+The Jacobian entries computed by \texttt{pyJac} are arranged such that entry ($i$, $j$) corresponds to the partial derivative of the $i$th source term in~\cref{e:source_terms} by the $j$th state variable in~\cref{e:state}:
 \loadeq{jacgeneral}
@@
 \label{s:unittest}
-Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates user specified psuedo-code and data to OpenCL\slash C-code.
+Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates user specified pseudo-code and data to OpenCL\slash C-code.
 As the name implies, \texttt{loo.py} generates code based on for-loops, this differs from the previous version of \texttt{pyJac}~\cite{pyjac16} which generates static code (i.e., fully unrolled loops, with thermodynamic\slash reaction parameters written directly in code rather than stored in arrays).
@@
 Further, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source terms.
-The source-terms and subcomponents thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{hogan2014fast,adept-v11} is used to obtain reference answers for Jacobian sub-components.
+The source-terms and sub-components thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{hogan2014fast,adept-v11} is used to obtain reference answers for Jacobian sub-components.
 The Portable OpenCL (POCL) implementation~\cite{poclIJPP} and OpenMP~\cite{dagum1998openmp} are used to run OpenCL and C unit-testing, respectively, on the continuous integration framework Travis CI~\cite{travis:2018}.
@@
 \caption{Summary of rate of progress, species, temperature and thermodynamic state parameter rate correctness.
-Error statistics are based on the infinity-norm of the relative error detailed in Eq.~\eqref{e:rel_err} for each quantity.
+Error statistics are based on the infinity-norm of the relative error detailed in~\cref{e:rel_err} for each quantity.
 The ``S'' in $E_{\frac{\text{d} S}{\text{d} t}}$ refers to the thermodynamic state parameter, either $V$ or $P$ for \conp/ and \conv/ respectively.
@@
 This point is further underscored by the error in the net ROP, which is \textasciitilde\numrange{3}{9} orders of magnitude (or \numrange{7}{9} orders of magnitude for the P-Log excluded error) larger than the error in either the forward or reverse ROP.
-In~\cref{T:source_error}, we see that $E_{\varepsilon}$ is very similar in magnitude to $E_R$ in all cases, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
+In~\cref{T:source_error}, it is seen that $E_{\varepsilon}$ is very similar in magnitude to $E_R$ in all cases, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
 The molar species production rate error norm is similar in magnitude to that of the net ROP, however the temperature and state parameter rate errors again increase as the error in net species production rates is amplified by multiplication with thermodynamic properties.
-Again, we note that the above discussion does not imply that discrepencies in computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
+Again, we note that the above discussion does not imply that discrepancies in computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
 
@@
 The reasons for this choice are explained fully in~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with minimal additional programming effort.
-The discrepancy between the analytical and automatic differentiated Jacobians for thermo-chemical state $k$, denoted by $\mathcal{J}_k$ and $\hat{\mathcal{J}}_k$ respectively, is determined by the relative error Frobenius norm over all Jacobian indicies $i, j$:
+The discrepancy between the analytical and automatic differentiated Jacobians for thermo-chemical state $k$, denoted by $\mathcal{J}_k$ and $\hat{\mathcal{J}}_k$ respectively, is determined by the relative error Frobenius norm over all Jacobian indices $i, j$:
 \begin{equation}
@@
 For convenience we will outline only the CSR format but the CSC format is very similar~\cite{netlib_templates}.
-An $N\times N$ CSR matrix is composed of three vectors, a value vector of length $NNZ$---the number of non-zero elements in the Jacobian---which stores the elements of the Jacobian, a row pointer vector---of length $N$---which stores the locations in the value vector that begin a row, and a column index vector---of length $NNZ$---which stores the column indicies of the elements in the value vector.
+An $N\times N$ CSR matrix is composed of three vectors, a value vector of length $NNZ$---the number of non-zero elements in the Jacobian---which stores the elements of the Jacobian, a row pointer vector---of length $N$---which stores the locations in the value vector that begin a row, and a column index vector---of length $NNZ$---which stores the column indices of the elements in the value vector.
 
@@
 Generally, this means that a lookup operation is required to find the sparse Jacobian index for any pair of state variables; in some cases this can be avoided, e.g., the rows corresponding to derivatives of the temperature and thermodynamic state parameter source terms are fully dense in \texttt{pyJac}, and hence no lookup is necessary.
-This lookup operation is currently implemented as a simple for-loop, e.g., for a sparse lookup of a pair of indicies $(i, j)$ in a compressed row storage matrix, the lookup function searches the column index vector between the values $\texttt{row\_pointer}[i] \ldots \texttt{row\_pointer}[i + 1]$ for $j$, and returns the offset from $\texttt{row\_pointer}[i]$ (or $\num{-1}$ if not found).
-As will be explored in~\cref{S:jacobian_results}, this slows down sparse Jacobian evaluation, and might be improved---at the cost of increased constant data usage, a limitation for OpenCL---by a static mapping of the full Jacobian indicies to the sparse index (or some null value if the entry is empty).
+This lookup operation is currently implemented as a simple for-loop, e.g., for a sparse lookup of a pair of indices $(i, j)$ in a compressed row storage matrix, the lookup function searches the column index vector between the values $\texttt{row\_pointer}[i] \ldots \texttt{row\_pointer}[i + 1]$ for $j$, and returns the offset from $\texttt{row\_pointer}[i]$ (or $\num{-1}$ if not found).
+As will be explored in~\cref{S:jacobian_results}, this slows down sparse Jacobian evaluation, and might be improved---at the cost of increased constant data usage, a limitation for OpenCL---by a static mapping of the full Jacobian indices to the sparse index (or some null value if the entry is empty).
 Additionally, this might be an excellent usage of OpenCL's Image memory type (similar to texture memory in CUDA terminology) and both of these sparse indexing techniques merit future investigation.
@@
 This normalization of the run time by the number of initial conditions tested is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{t:models}).
-For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the unvectorized (strictly parallely) Intel OpenCL code is slightly faster than OpenMP in all cases.
+For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the unvectorized (strictly parallel) Intel OpenCL code is slightly faster than OpenMP in all cases.
 As expected, the \avx/ machine is faster than the \sse/ CPU for the strictly parallel cases, \SIrange{1.82}{2.13}{$\times$} for the unvectorized OpenCL case, and \SIrange{1.72}{1.85}{$\times$} for OpenMP.
@@
 \Cref{F:gpu_source} shows the performance of the source term evaluation on the GPUs listed in~\cref{t:gpus}.
-In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean run time per condition; the mean run time continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
-Interestingly, the last data-point for the GRI-Mech 3.0 shows an increase in run time per condition, this is likely due to multiple data transfers to\slash from the GPU required to complete the source-rates in this case; similar (if smaller) increases in run time per condition can be seen for the \ce{iC5H11OH} model near \num{e4} and \num{e5} conditions.\todo{Update all this after I get rerun w/ less dumb copies}
-\Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} for two different vector-widths (a.k.a., the block size in NVIDIA terminology), \num{64} and \num{128}; the fastest \gpunew/ case (a vector-width of 128) is \SIrange{1.34}{4.89}{$\times$} faster than the slowest case (either \gpunew/ or \gpuold/ with a vector-width of \num{64}) depending on the chemical model in question.
-Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector-width has minimal effect on performance for most \gpuold/ cases (excepting \ce{H2}\slash\ce{CO}), but a large effect on the performance of the \gpunew/ GPU; this is likely caused by higher occupancy~\cite{occupancy} on the \gpunew/ GPU, however it is unclear exactly how NVIDIA's OpenCL runtime balances the registers\slash warps per streaming-multiprocessor, as controls occupancy in CUDA~\cite{occupancy}; this is a topic that merits future investigation.
+In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean run time per condition; the run time continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
+\Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} for two different vector-widths (a.k.a., the block size in NVIDIA terminology), \num{64} and \num{128}; the fastest \gpunew/ case with a vector-width of 128 is \SIrange{1.40}{1.88}{$\times$} faster than the slowest case (\gpuold/ with a vector-width of \num{64}) depending on the chemical model in question.
+Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector-width has minimal effect on performance for most of the \gpunew/ and \gpuold/ cases; the GRI-Mech 3.0 and USC-Mech II models show the largest improvements with a vector-width of \num{128}, \textasciitilde\SIrange{10}{18}{\percent} for both GPUs.
+This is likely caused by higher occupancy on the GPUs, however it is unclear exactly how NVIDIA's OpenCL runtime balances the registers\slash warps per streaming-multiprocessor, as controls occupancy in CUDA~\cite{occupancy}; use of even larger vector-widths (i.e., \num{256}\texttt{+}) on the GPU is a topic that merits future investigation.
 \begin{figure}[htbp]
@@
 \Cref{f:source_permutate} shows the effect of changing data-ordering patterns, the \conp/ or \conv/ formulation, and the CPU vector-width size on the performance of source term evaluation in \texttt{pyJac}.
-In~\cref{F:source_conpvsconv}, we see that choice of the constant-pressure or constant-volume formulation has little to no effect on run time for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx/ machine.
+In~\cref{F:source_conpvsconv}, it is seen that choice of the constant-pressure or constant-volume formulation has little to no effect on run time for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx/ machine.
 Generally speaking, the difference between the \conp/ and \conv/ formulations has only a marginal effect on performance regardless of CPU\slash GPU choice.
@@
       \includegraphics[width=\textwidth]{source_gpu_cvsf.pdf}
-      \caption{The affect on performance of ``C'' vs ``F''-ordering for the shallow-vectorized NVIDIA OpenCL code on both GPUs with a vector-width of 128.
-The speedup presented is calculated per-GPU to better assess the effect of the data-ordering on each GPU.}
+      \caption{The affect on performance of ``C'' vs ``F''-ordering for the shallow-vectorized NVIDIA OpenCL code on both GPUs with a vector-width of \num{128}.}
       \label{F:source_gpu_cvsf}
@@
 \Cref{F:source_gpu_cvsf} shows the performance effect of the ``C'' versus ``F''-ordering on source term evaluation performance on both GPUs, with the speedup presented per-GPU.
-In general, the ``C'' and ``F''-ordered shallow-vectorizations are nearly equivalent on both GPUs while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, with the exception of the isopentanol model where the ``C''-ordered data is up to \SI{1.1}{$\times$} faster on the \gpunew/ (we note that there was high variablity for the \gpuold/ in this case).
-\todo{again, pending rerun on with less copying}
+In general, the ``C'' and ``F''-ordered shallow-vectorizations are nearly equivalent on both GPUs, with less than a \SI{10}{\percent} difference in run time between data-orderings. 
+It is noted that for the isopentanol model, ``C''-ordered data is \textasciitilde\SI{1.08}{$\times$} faster on both GPUs (while the trend is less clear for the other models).
 The rough equivalence between the ``C'' and ``F''-ordered codes GPUs is counter to what one might expect, as typically speaking coalesced memory access in a shallow-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
-However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector-width size and thus, likely coalesced.
-% but the improved data-locality, coupled with the larger cache sizes of the \gpunew/ GPU result in this modest speedup for ``C''-data ordering.
+However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector-width size and thus, likely well-coalesced.
 
@@
 
-\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source-rate in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
+\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source-rate evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
 In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for Intel OpenCL; although the mechanism scalings vary somewhat for OpenMP, there is no clear trend.
-It is likely the poor scaling of the \ce{H2}\slash\ce{CO} for Intel OpenCL results from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD intructions.
+It is likely the poor scaling of the \ce{H2}\slash\ce{CO} for Intel OpenCL results from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD instructions.
 Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.
@@
 The mean run time per-condition for the sparse and dense Jacobian evaluation on the \gpunew/ GPU is plotted in~\cref{F:gpu_sparse_vs_dense}.
-As in the species evaluation case, the mean run time per-condition drops steadily, for both cases becoming constant after just \num{3e3} states (except for \ce{H2}\slash\ce{CO}, which levels off near \num{e4} conditions).
-Sparse jacobian evaluation is significantly slower for all models before the GPU becomes saturated (due to the indirect indexing lookup), however past the saturation point the performance gap between the sparse and dense evaluations narrows.
-This is likely caused by the combination of more sparse Jacobian matrices fitting in the \gpunew/'s memory, as well as improved data-locality\slash caching.
+As in the species evaluation case, the mean run time per-condition drops steadily, for both cases becoming roughly constant after just \num{3e3} states (except for \ce{H2}\slash\ce{CO}, which levels off near \num{e4} conditions).
+Sparse Jacobian evaluation is significantly slower for all models before the GPU becomes saturated (due to the indirect indexing lookup), however past the saturation point the performance gap between the sparse and dense evaluations narrows.
+This is likely caused by the combination of being able to fit many more sparse Jacobian matrices in the \gpunew/'s memory, as well as improved data-locality\slash caching due to the smaller size of the sparse Jacobian.
 \Cref{F:gpu_jacobian_speedup} presents the speedup of the \gpunew/ over the \gpuold/ GPU for sparse\slash dense Jacobian evaluation; sparse evaluation on the \gpunew/ is \SIrange{1.10}{1.59}{$\times$} faster than on the \gpuold/, while dense evaluation is \SIrange{1.36}{3.0}{$\times$} faster.
-The speedup on the \gpunew/ appears to decrease with increasing mechanism size for the sparse evalution, but increases for the larger models in dense evaluation; this is likely due to the larger memory size of the \gpunew/, and hence fewer data-transferring operations.\todo{missing srv2 data for full conditions}
+The speedup on the \gpunew/ appears to decrease with increasing mechanism size for the sparse evaluation, but increases for the larger models in dense evaluation; this is likely due to the larger memory size of the \gpunew/, and hence fewer data-transfer operations.
+
+In~\cref{F:fd_vs_analytical}, the performance of the sparse analytical-Jacobian is compared to a sparse first-order finite-difference Jacobian on both the \avx/ CPU and the \gpuold/ \gpunew/ GPUs.
+\Cref{F:fd_vs_analytical_cpu} shows large speedups for both the OpenMP code and shallow-vectorized OpenCL code; the analytical OpenMP Jacobian is \SIrange{3.92}{8.67}{$\times$} faster, while the analytical OpenCL Jacobian achieves speedups of \SIrange{17.22}{55.11}{$\times$}.
+It is noted that the isopentanol case was excluded here, as a single run of the sparse finite-difference Jacobian using either OpenCL or OpenMP took over \num{12} hours of run time.
+In addition the current finite-difference formulation breaks Intel's auto-vectorizer, hence the OpenCL speedup comparison is made against the unvectorized OpenCL code; it is not a high priority to fix this issue, as the finite-difference Jacobian is implemented for comparison purposes only.
+Although we do not display the dense finite-difference Jacobian speedup in~\cref{F:fd_vs_analytical_cpu}, the dense OpenCL and OpenMP analytical Jacobian codes out-perform the finite-difference variants by even larger margins; \SIrange{24.44}{245.63}{$\times$} for OpenCL, and \SIrange{9.68}{112.73}{$\times$} for OpenMP (these data do include the isopentanol model, though limited to \num{50000} conditions as discussed previously).
+A comparison of the sparse analytical and finite-difference Jacobians on the GPUs is presented in~\cref{F:fd_vs_analytical_gpu}.
+The analytical Jacobian on the \gpunew/ and \gpuold/ shows large speedups, increasing with chemical model size, in the range of \SIrange{3.81}{17.60}{$\times$}; the \gpunew/ has a larger speedup than the \gpuold/ (\SI{17.60}{$\times$} v.s., \SI{14.75}{$\times$}) for the isopentanol model due to it's larger memory size.
+Although not pictured, the dense analytical-Jacobian on the \gpunew/ GPU has larger speedups compared with the dense finite-difference Jacobian: \SIrange{4.04}{45.13}{$\times$}; the \gpunew/ GPU again has a significantly larger speedup than the \gpuold/ for the isopentanol model (\SI{45.13}{$\times$} vs \SI{23.85}{$\times$}), further underscoring the effect of the larger available memory on the \gpunew/.
 
@@
 \end{figure}
-
-In~\cref{F:fd_vs_analytical}, the performance of the sparse analytical-Jacobian is compared to a sparse first-order finite-difference Jacobian on both the \avx/ CPU and the \gpuold/ \gpunew/ GPUs.
-\Cref{F:fd_vs_analytical_cpu} shows large speedups for both the OpenMP code and shallow-vectorized OpenCL code; the analytical OpenMP Jacobian is \SIrange{3.92}{8.67}{$\times$} faster, while the analytical OpenCL Jacobian achives speedups of \SIrange{17.22}{55.11}{$\times$}.
-It is noted that the isopentanol case was excluded here, as a single run of the sparse finite-difference Jacobian using either OpenCL or OpenMP took over \num{12} hours of run time.
-In addition the current finite-difference formulation breaks Intel's auto-vectorizer, hence the OpenCL speedup comparison is made against the unvectorized OpenCL code; it is not a high priority to fix this issue, as the finite-difference Jacobian is implemented for comparison purposes only.
-Although we do not display the dense finite-difference Jacobian speedup in~\cref{F:fd_vs_analytical_cpu}, the dense OpenCL and OpenMP analytical Jacobian codes out-perform the finite-difference variants by even larger margins; \SIrange{24.44}{245.63}{$\times$} for OpenCL, and \SIrange{9.68}{112.73}{$\times$} for OpenMP (these data do include the isopentanol model, though limited to \num{50000} conditions as discussed previously).
-A comparison of the sparse analytical and finite-difference Jacobians is presented in~\cref{F:fd_vs_analytical_gpu}.
-The analytical Jacobian on the \gpunew/ and \gpuold/ shows large (and very similar) speedups, increasing with chemical model size, in the range of \SIrange{4.66}{17.43}{$\times$}.
-Although not pictured, the dense analytical-Jacobian on the \gpunew/ GPU has larger speedups compared with the dense finite-difference Jacobian: \SIrange{5.32}{44.87}{$\times$}.
 
@@
 This provides a particularly vexing problem for the programmer, as there is often little that can be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
-Further, as noted throughout this work certain transformations or code-generation patterns can break OpenCL execution\slash vectorization; e.g., the failure of POCL to achieve vectorization, Intel OpenCL's failure to vectorize the finite-difference Jacobian, the lack of speedup for the finite-difference Jacobian on the NVIDIA runtime for the \gpuold/ GPU\todo{this went away?}.
+Further, as noted throughout this work certain transformations or code-generation patterns can break OpenCL execution\slash vectorization; e.g., the failure of POCL to achieve vectorization or Intel OpenCL's failure to vectorize the finite-difference Jacobian.
 Indeed the vectorization attempted by the OpenCL runtimes (or reasons for incorrectly vectorized code) are not always obvious to the user.
@@
 
-\section{Acknowledgements}
+\section{Acknowledgments}
 This research was funded by the National Science Foundation under grant ACI-1534688.

### 2018-03-12 paper.tex
s/source-rate/source-term

@@
 \begin{abstract} % not to exceed 200 words
-A code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized thermo-chemical source term and sparse\slash dense chemical kinetic Jacobian evaluation was developed and validated for a wide range of chemical kinetic models.
+A code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation was developed and validated for a wide range of chemical kinetic models.
 
@@
 As the importance of detailed chemical kinetics for predictive reactive-flow simulations has become recognized~\cite{LU2009192}, the size and complexity of chemical kinetic models describing current and next-generation fuels relevant to transportation and power generation have increased greatly; e.g., a recent biodiesel model~\cite{WESTBROOK2011742} consists of \textasciitilde\num{3500} chemical species, and over \num{17000} reactions.
-Moreover, the cost of evaluating the chemical source terms scales linearly with the size of the model, while evaluation and factorization of the chemical kinetic Jacobian---if naively implemented---scale quadratically and cubically with the number of species in the model~\cite{LU2009192}.
+Moreover, the cost of evaluating the chemical source-terms scales linearly with the size of the model, while evaluation and factorization of the chemical kinetic Jacobian---if naively implemented---scale quadratically and cubically with the number of species in the model~\cite{LU2009192}.
 These factors often make the use of detailed chemical kinetics prohibitive in practice; e.g., in a recent direct numerical simulation~\cite{Spafford:2010aa} utilizing a 22-species model, evaluation of reaction rates consumed around half of the total run time.
@@
 
-In a \textit{shallow}-SIMD\slash SIMT vectorization (also referred to as ``per-thread'' in previous works~\cite{Stone:2013aa} using GPUs), each SIMD-lane (or SIMT-thread) in a compute-unit evaluates the thermo-chemical source terms\slash chemical kinetic Jacobian for a different thermo-chemical state.
+In a \textit{shallow}-SIMD\slash SIMT vectorization (also referred to as ``per-thread'' in previous works~\cite{Stone:2013aa} using GPUs), each SIMD-lane (or SIMT-thread) in a compute-unit evaluates the thermo-chemical source-terms\slash chemical kinetic Jacobian for a different thermo-chemical state.
 If the data is stored in ``F''-order, the SIMD-lanes\slash SIMT-thread accessing $[C]_{1, j}\ldots[C]_{N_v, j}$---the concentration of species $j$ for states $1, 2,\ldots N_v$, where $N_v$ is the SIMD vector-width or the number of threads in a SIMT warp---will load sequential locations in memory.
@@
 
-In a \textit{deep}-SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works), a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermo-chemical source terms for a single thermo-chemical state; thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + N_v}$---the species concentrations for state 1, species $j \ldots j + N_v$---will access sequential memory locations if the data is stored in ``C''-order.
+In a \textit{deep}-SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works), a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermo-chemical source-terms for a single thermo-chemical state; thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + N_v}$---the species concentrations for state 1, species $j \ldots j + N_v$---will access sequential memory locations if the data is stored in ``C''-order.
 Further, in the ``C''-ordering the furthest difference between any two species concentrations within the same thermo-chemical state is at most $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU, but may lead to uncoalesced memory-accesses on the GPU.
@@
 \subsection{Thermo-chemical Source Terms and Jacobian}
-This new version of \texttt{pyJac} is capable of evaluating the thermo-chemical source terms for using both the constant-pressure (\conp/) or constant-volume (\conv/) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
+This new version of \texttt{pyJac} is capable of evaluating the thermo-chemical source-terms for using both the constant-pressure (\conp/) or constant-volume (\conv/) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
 In this section, we will only outline a brief summary of the system evaluated by \texttt{pyJac}; the interested reader is referred to the complete derivations, attached as supplemental material to this document.
@@
 
-For both \conp/ and \conv/, the molar source terms are~\cite{TurnsStephenR2012Aitc}:
+For both \conp/ and \conv/, the molar source-terms are~\cite{TurnsStephenR2012Aitc}:
 \loadeq{dmolar}
@@
 
-The temperature source term~\cite{TurnsStephenR2012Aitc} is:
+The temperature source-term~\cite{TurnsStephenR2012Aitc} is:
 \loadeq{dtemp}
@@
 \end{equation}
-where $\mathcal{R}$ is the ideal gas-constant in molar units, the volume and pressure source terms can be found:
+where $\mathcal{R}$ is the ideal gas-constant in molar units, the volume and pressure source-terms can be found:
 \loadeq{dparam}
 
-The Jacobian entries computed by \texttt{pyJac} are arranged such that entry ($i$, $j$) corresponds to the partial derivative of the $i$th source term in~\cref{e:source_terms} by the $j$th state variable in~\cref{e:state}:
+The Jacobian entries computed by \texttt{pyJac} are arranged such that entry ($i$, $j$) corresponds to the partial derivative of the $i$th source-term in~\cref{e:source_terms} by the $j$th state variable in~\cref{e:state}:
 \loadeq{jacgeneral}
 
-The derivative of the temperature source term with respect to itself is:
+The derivative of the temperature source-term with respect to itself is:
 \loadeq{dTdotdT}
@@
 
-The molar source terms are more complicated due to the derivatives of the net rate of progress---see~\cref{deriv-s:dropi}---and pressure modification terms---\cref{deriv-s:dci}---of the reactions.
+The molar source-terms are more complicated due to the derivatives of the net rate of progress---see~\cref{deriv-s:dropi}---and pressure modification terms---\cref{deriv-s:dci}---of the reactions.
 The reader is again encouraged to read the supplemental material covering the exact derivations and equations solved by \texttt{pyJac}.
-The derivative of the molar source term with respect to temperature is:
+The derivative of the molar source-term with respect to temperature is:
 \loadeq{dnkdotdT}
@@
 In addition, \texttt{loo.py} allows the user to more easily make changes the structure of the generated program, e.g., data ordering, vectorization, and threading patterns, as well as switching the target-language for code-generation (and simpler extension to additional languages, e.g., CUDA).
-Further, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source terms.
+Further, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source-terms.
 The source-terms and sub-components thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{hogan2014fast,adept-v11} is used to obtain reference answers for Jacobian sub-components.
@@
 
-\subsection{Source term validation}
+\subsection{Source-term validation}
 \label{s:validation}
@@
 
-In Table~\ref{T:source_error}, the results of this code's source-rate evaluations are compared to Cantera~\cite{Cantera} on the library of PaSR conditions (\cref{t:models}).
+In Table~\ref{T:source_error}, the results of this code's source-term evaluations are compared to Cantera~\cite{Cantera} on the library of PaSR conditions (\cref{t:models}).
 Very close agreement is observed for the forward and reverse ROP for all models, however the error norm is \textasciitilde\numrange{3}{4} orders of magnitude larger for the isopentanol model.
@@
 
-As in our previous work~\cite{Niemeyer:2016aa}, the correctness of the Jacobian was determined by comparison to the Jacobian obtained by automatic differentiation of the source-rate subroutines generated by \texttt{pyJac} via the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
+As in our previous work~\cite{Niemeyer:2016aa}, the correctness of the Jacobian was determined by comparison to the Jacobian obtained by automatic differentiation of the source-term subroutines generated by \texttt{pyJac} via the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
 The reasons for this choice are explained fully in~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with minimal additional programming effort.
@@
 Conversely, the LAPACK error norm does seem to correlate with the relaxed filtered error norm; the two models with the largest LAPACK error, \ce{H2}\slash\ce{CO} and \ce{iC5H11OH}, have the largest relaxed filtered error norms, $\mathcal{O}(\num{e-5})$, again suggesting that the stringent error norm is influenced by floating point error accumulation.
-These findings, along with the individual unit-testing of all chemical source terms and Jacobian sub-components described in~\cref{s:unittest}, gives high confidence in the correctness of this new version of \texttt{pyJac}.
+These findings, along with the individual unit-testing of all chemical source-terms and Jacobian sub-components described in~\cref{s:unittest}, gives high confidence in the correctness of this new version of \texttt{pyJac}.
 
@@
 This pattern leads to fairly easily vectorizable code, and efficient Jacobian evaluation as the bulk of the computation is dependent only on the reaction in question, as discussed in our previous work~\cite{Niemeyer:2016aa}.
-Generally, this means that a lookup operation is required to find the sparse Jacobian index for any pair of state variables; in some cases this can be avoided, e.g., the rows corresponding to derivatives of the temperature and thermodynamic state parameter source terms are fully dense in \texttt{pyJac}, and hence no lookup is necessary.
+Generally, this means that a lookup operation is required to find the sparse Jacobian index for any pair of state variables; in some cases this can be avoided, e.g., the rows corresponding to derivatives of the temperature and thermodynamic state parameter source-terms are fully dense in \texttt{pyJac}, and hence no lookup is necessary.
 This lookup operation is currently implemented as a simple for-loop, e.g., for a sparse lookup of a pair of indices $(i, j)$ in a compressed row storage matrix, the lookup function searches the column index vector between the values $\texttt{row\_pointer}[i] \ldots \texttt{row\_pointer}[i + 1]$ for $j$, and returns the offset from $\texttt{row\_pointer}[i]$ (or $\num{-1}$ if not found).
@@
 Run times in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
-The OpenMP Jacobian\slash source-rate kernels, as well as the OpenMP\slash OpenCL wrapping code---responsible for initializing\slash transferring memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms, and \texttt{gcc v4.8.5} on the \sse/\slash\gpuold/ machines.
+The OpenMP Jacobian\slash source-term kernels, as well as the OpenMP\slash OpenCL wrapping code---responsible for initializing\slash transferring memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms, and \texttt{gcc v4.8.5} on the \sse/\slash\gpuold/ machines.
 The optimization level ``\texttt{-O3 -mtune=native}'' was used, and no ``fast-math'' OpenCL optimizations were enabled.
@@
 
-\Cref{F:cpu_source} explores the performance of the source term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
+\Cref{F:cpu_source} explores the performance of the source-term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
 Source term evaluations---critical in their own right for direct numerical simulations of reacting flow~\cite{Spafford:2010aa}, among other applications---also give a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation performance.
@@
 
-\Cref{F:gpu_source} shows the performance of the source term evaluation on the GPUs listed in~\cref{t:gpus}.
-In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean run time per condition; the run time continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
-\Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} for two different vector-widths (a.k.a., the block size in NVIDIA terminology), \num{64} and \num{128}; the fastest \gpunew/ case with a vector-width of 128 is \SIrange{1.40}{1.88}{$\times$} faster than the slowest case (\gpuold/ with a vector-width of \num{64}) depending on the chemical model in question.
+\Cref{F:gpu_source} shows the performance of the source-term evaluation on the GPUs listed in~\cref{t:gpus}.
+In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source-terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean run time per condition; the run time continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
+\Cref{F:gpu_source_speedup} shows the speedup in source-term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} for two different vector-widths (a.k.a., the block size in NVIDIA terminology), \num{64} and \num{128}; the fastest \gpunew/ case with a vector-width of 128 is \SIrange{1.40}{1.88}{$\times$} faster than the slowest case (\gpuold/ with a vector-width of \num{64}) depending on the chemical model in question.
 Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector-width has minimal effect on performance for most of the \gpunew/ and \gpuold/ cases; the GRI-Mech 3.0 and USC-Mech II models show the largest improvements with a vector-width of \num{128}, \textasciitilde\SIrange{10}{18}{\percent} for both GPUs.
@@
   \end{subfigure}
-  \caption{\texttt{pyJac} source term evaluation performance on the NVIDIA GPUs}
+  \caption{\texttt{pyJac} source-term evaluation performance on the NVIDIA GPUs}
   \label{F:gpu_source}
@@
 
-\Cref{f:source_permutate} shows the effect of changing data-ordering patterns, the \conp/ or \conv/ formulation, and the CPU vector-width size on the performance of source term evaluation in \texttt{pyJac}.
+\Cref{f:source_permutate} shows the effect of changing data-ordering patterns, the \conp/ or \conv/ formulation, and the CPU vector-width size on the performance of source-term evaluation in \texttt{pyJac}.
 In~\cref{F:source_conpvsconv}, it is seen that choice of the constant-pressure or constant-volume formulation has little to no effect on run time for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx/ machine.
@@
       \includegraphics[width=\textwidth]{source_vector_width.pdf}
-      \caption{The effect of vector-width on ``C''-ordered shallow-vectorized Intel OpenCL source term evaluations in the \avx/ CPU.}
+      \caption{The effect of vector-width on ``C''-ordered shallow-vectorized Intel OpenCL source-term evaluations in the \avx/ CPU.}
       \label{F:source_vector_width}
   \end{subfigure}
-  \caption{The effect of the \conp/ and \conv/ formulations, ``C'' and ``F'' data-ordering, and CPU vector-width on source term evaluation performance in \texttt{pyJac}.
+  \caption{The effect of the \conp/ and \conv/ formulations, ``C'' and ``F'' data-ordering, and CPU vector-width on source-term evaluation performance in \texttt{pyJac}.
 	   It is noted the shallow-vectorized ``C'' ordered OpenCL cases corresponds to the vectorized data ordering described in~\cref{S:data}}
@@
 
-\Cref{F:source_gpu_cvsf} shows the performance effect of the ``C'' versus ``F''-ordering on source term evaluation performance on both GPUs, with the speedup presented per-GPU.
+\Cref{F:source_gpu_cvsf} shows the performance effect of the ``C'' versus ``F''-ordering on source-term evaluation performance on both GPUs, with the speedup presented per-GPU.
 In general, the ``C'' and ``F''-ordered shallow-vectorizations are nearly equivalent on both GPUs, with less than a \SI{10}{\percent} difference in run time between data-orderings. 
@@
 
-\Cref{F:source_vector_width} shows the effect of changing vector-width size on source-rate performance the \avx/ CPU.
+\Cref{F:source_vector_width} shows the effect of changing vector-width size on source-term performance the \avx/ CPU.
 The vector-width of size \num{8} was the fastest tested, while the larger vector-width of \num{16} is slightly slower due to increased register pressure~\cite{intel_vectypes}.
@@
       \includegraphics[width=\textwidth]{source_parallel_scaling.pdf}
-      \caption{The strong parallel scaling efficiency (as defined in~\cref{e:strong_scaling}) of source term evaluation for Intel OpenCL\slash OpenMP on the \avx/ machine.}
+      \caption{The strong parallel scaling efficiency (as defined in~\cref{e:strong_scaling}) of source-term evaluation for Intel OpenCL\slash OpenMP on the \avx/ machine.}
       \label{F:source_parallel_scaling}
@@
       \includegraphics[width=\textwidth]{source_crossover.pdf}
-      \caption{The mean run time per-condition of OpenMP source term evaluation on different cores compared to a shallow-vectorized Intel OpenCL evaluation on \num{1} (solid lines) and \num{4} (dashed lines) cores on \avx/ machine.}
+      \caption{The mean run time per-condition of OpenMP source-term evaluation on different cores compared to a shallow-vectorized Intel OpenCL evaluation on \num{1} (solid lines) and \num{4} (dashed lines) cores on \avx/ machine.}
       \label{F:source_crossover}
@@
       \includegraphics[width=\textwidth]{source_simd_efficiency.pdf}
-      \caption{The SIMD efficiency (\cref{e:simd_efficiency}) of source term evaluation for the Intel OpenCL runtime on a single core of the \avx/ CPU.}
+      \caption{The SIMD efficiency (\cref{e:simd_efficiency}) of source-term evaluation for the Intel OpenCL runtime on a single core of the \avx/ CPU.}
       \label{F:source_simd_scaling}
@@
       \includegraphics[width=\textwidth]{source_sse_simd_efficiency.pdf}
-      \caption{The SIMD efficiency (\cref{e:simd_efficiency}) of source term evaluation for the Intel OpenCL runtime on a single core of the \sse/ CPU.}
+      \caption{The SIMD efficiency (\cref{e:simd_efficiency}) of source-term evaluation for the Intel OpenCL runtime on a single core of the \sse/ CPU.}
       \label{F:source_sse_simd_scaling}
   \end{subfigure}
-  \caption{The parallel scaling efficiency and SIMD efficiency of source term evaluation for Intel OpenCL on the \avx/ and \sse/ CPUs.}
+  \caption{The parallel scaling efficiency and SIMD efficiency of source-term evaluation for Intel OpenCL on the \avx/ and \sse/ CPUs.}
   \label{F:source_scaling}
@@
 
-\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source-rate evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
+\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source-term evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
 In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for Intel OpenCL; although the mechanism scalings vary somewhat for OpenMP, there is no clear trend.
@@
 Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.
-This is due in large part to the superior performance of the shallow-vectorized OpenCL code, coupled with the relatively small amounts of work associated with source-rate evaluation.
-To illustrate this,~\cref{F:source_crossover} shows the mean run time per-condition of the OpenMP source term evaluations for \numrange{2}{16} cores, compared to the shallow-vectorized OpenCL code on \num{1} (solid line) and \num{4} (dashed line) four cores on the \avx/ machine.
+This is due in large part to the superior performance of the shallow-vectorized OpenCL code, coupled with the relatively small amounts of work associated with source-term evaluation.
+To illustrate this,~\cref{F:source_crossover} shows the mean run time per-condition of the OpenMP source-term evaluations for \numrange{2}{16} cores, compared to the shallow-vectorized OpenCL code on \num{1} (solid line) and \num{4} (dashed line) four cores on the \avx/ machine.
 For all chemical models, the mean run time per-condition (and hence the computational work allocated per-core, one of the key-drivers of parallel scaling efficiency~\cite{strong_scaling}) of OpenMP running on \num{4} cores is roughly equal to that of the shallow-vectorized OpenCL code on a single core.
@@
 This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.
-\Cref{F:source_simd_scaling} shows the SIMD efficiency of source term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models (isopentanol and USC-Mech II) have higher SIMD-efficiency of \numrange{0.76}{0.78}, and the smaller models (\ce{H2}\slash\ce{CO}, GRI-Mech \num{3.0}) have lower SIMD-efficiency, \numrange{0.6}{0.66}.
+\Cref{F:source_simd_scaling} shows the SIMD efficiency of source-term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models (isopentanol and USC-Mech II) have higher SIMD-efficiency of \numrange{0.76}{0.78}, and the smaller models (\ce{H2}\slash\ce{CO}, GRI-Mech \num{3.0}) have lower SIMD-efficiency, \numrange{0.6}{0.66}.
 This again demonstrates that the SIMD-vectorization becomes efficient as the amount of work to perform increases (with increasing model size).
@@
 In~\cref{F:sparse_jac_scaling}, the strong parallel scaling efficiency of the sparse shallow-vectorized OpenCL is compared to the sparse OpenMP code.
-The plot is a bit difficult to read, as most of the data-points are clustered together, however the key takeaway is that the parallel scaling efficiency of the shallow-vectorized OpenCL code is very similar to that of the OpenMP code, in contrast to the source-rates parallel scaling efficiencies (\cref{F:source_parallel_scaling}).
+The plot is a bit difficult to read, as most of the data-points are clustered together, however the key takeaway is that the parallel scaling efficiency of the shallow-vectorized OpenCL code is very similar to that of the OpenMP code, in contrast to the source-terms parallel scaling efficiencies (\cref{F:source_parallel_scaling}).
 The \ce{H2}\slash\ce{CO} model has the worst scaling for both codes, ranging from \numrange{0.94}{0.54} and \numrange{0.99}{0.82} for OpenCL and OpenMP respectively on \numrange{2}{16} cores.
@@
 In this case, the isopentanol model has the worst scaling for both codes, it is noted that due to the sheer size of the dense isopentanol Jacobian---storing the dense matrix for a single thermo-chemical state takes over \SI{1}{\mega\byte} of data---the total number of states for the dense isopentanol Jacobian evaluation was limited to \num{50000} (which requires over \SI{50}{\giga\byte} of memory); this greatly drops the computation cost for this case, and adversely affects the scaling efficiency as was discussed in~\cref{S:source_results}.
-Excluding the isopentanol model, the dense shallow-vectorized OpenCL scalings are slightly higher than for source-rate evaluation, e.g., \numrange{0.70}{0.78} and \numrange{0.52}{0.61} for \num{4} and \num{8} cores respectively (compared to \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for shallow-vectorized OpenCL source-rate on the same \num{4} and \num{8} cores); this is due to the higher computational cost of Jacobian evaluation, and hence more available work per CPU-core.
+Excluding the isopentanol model, the dense shallow-vectorized OpenCL scalings are slightly higher than for source-term evaluation, e.g., \numrange{0.70}{0.78} and \numrange{0.52}{0.61} for \num{4} and \num{8} cores respectively (compared to \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for shallow-vectorized OpenCL source-term on the same \num{4} and \num{8} cores); this is due to the higher computational cost of Jacobian evaluation, and hence more available work per CPU-core.
 As in~\cref{S:source_results}, we note that the shallow-vectorized dense Jacobian code running on \num{1} and \num{4} cores of the \avx/ machine is roughly as fast as the OpenMP code on \num{4} and \num{16} cores respectively; the parallel scaling efficiency of OpenMP at \num{16} cores (excluding isopentanol as previously) is \numrange{0.63}{0.71}, similar to the shallow-vectorization's efficiency of \numrange{0.70}{0.78} for \num{4} cores.
@@
 \section{Conclusions}
-In this work, automatically generated OpenCL codes for shallow SIMD-vectorized thermo-chemical source term evaluation were developed and validated against Cantera~\cite{Cantera} for a shallow range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
+In this work, automatically generated OpenCL codes for shallow SIMD-vectorized thermo-chemical source-term evaluation were developed and validated against Cantera~\cite{Cantera} for a shallow range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
 Significant speedups of up to \SIrange{1.90}{2.45}{$\times$} over a baseline SIMT-vectorized code were observed.

### 2018-03-12 paper.tex
edits

@@
 
-\title{SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source Term Evaluation}
+\title{SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source-Term Evaluation}
 
@@
 \begin{keyword}
-    Chemical Kinetics\sep SIMD\sep SIMT\sep Vector Processing\sep Jacobian
+    Chemical Kinetics\sep SIMD\sep SIMT\sep Sparse\sep Jacobian
 \end{keyword}
@@
 
-\subsection{Thermo-chemical Source Terms and Jacobian}
+\subsection{Thermo-chemical Source-Terms and Jacobian}
 This new version of \texttt{pyJac} is capable of evaluating the thermo-chemical source-terms for using both the constant-pressure (\conp/) or constant-volume (\conv/) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
@@
 \midrule
-\ce{H2}\slash~\ce{CO} & \num{900900} & \cite{Burke:2011fh} \\
+\ce{H2}\slash\ce{CO} & \num{900900} & \cite{Burke:2011fh} \\
 GRI-Mech.~3.0         & \num{450900} & \cite{smith_gri-mech_30} \\
@@
 \toprule
-\multicolumn{1}{l}{Model} & \multicolumn{1}{l}{\ce{H2}\slash~\ce{CO}} & \multicolumn{1}{l}{GRI-Mech.~3.0} & \multicolumn{1}{l}{USC-Mech II} & \multicolumn{1}{l}{\ce{iC5H11OH}} \\
+\multicolumn{1}{l}{Model} & \multicolumn{1}{l}{\ce{H2}\slash\ce{CO}} & \multicolumn{1}{l}{GRI-Mech.~3.0} & \multicolumn{1}{l}{USC-Mech II} & \multicolumn{1}{l}{\ce{iC5H11OH}} \\
 \midrule
@@
 
-\subsubsection{Source term evaluation}
+\subsubsection{Source-term evaluation}
 \label{S:source_results}
@@
 \Cref{F:cpu_source} explores the performance of the source-term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
-Source term evaluations---critical in their own right for direct numerical simulations of reacting flow~\cite{Spafford:2010aa}, among other applications---also give a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation performance.
+Source-term evaluations---critical in their own right for direct numerical simulations of reacting flow~\cite{Spafford:2010aa}, among other applications---also give a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation performance.
 

### 2018-03-13 paper.tex
edits and conclusion

@@
 As the importance of detailed chemical kinetics for predictive reactive-flow simulations has become recognized~\cite{LU2009192}, the size and complexity of chemical kinetic models describing current and next-generation fuels relevant to transportation and power generation have increased greatly; e.g., a recent biodiesel model~\cite{WESTBROOK2011742} consists of \textasciitilde\num{3500} chemical species, and over \num{17000} reactions.
-Moreover, the cost of evaluating the chemical source-terms scales linearly with the size of the model, while evaluation and factorization of the chemical kinetic Jacobian---if naively implemented---scale quadratically and cubically with the number of species in the model~\cite{LU2009192}.
+Moreover, the cost of evaluating the chemical source-terms scales linearly with the size of the model, while evaluation and factorization of the chemical kinetic Jacobian---if naively implemented via finite-difference method---scale quadratically and cubically with the number of species in the model~\cite{LU2009192}.
 These factors often make the use of detailed chemical kinetics prohibitive in practice; e.g., in a recent direct numerical simulation~\cite{Spafford:2010aa} utilizing a 22-species model, evaluation of reaction rates consumed around half of the total run time.
@@
 
-A host of techniques have been developed to lessen the computational demand of chemical kinetic calculations while maintaining fidelity, these fall broadly into a few categories: removal of unimportant species and reactions unimportant species and reactions~\cite{Lu:2006bb,Pepiot-Desjardins:2008,Hiremath:2010jw,Niemeyer:2010bt,Curtis:2015}, lumping of species with similar thermo-chemical properties~\cite{Lu:2007,Ahmed:2007fa,Pepiot:2008kq}, and time-scale methods that reduce chemical stiffness~\cite{Maas:1992ws,Lam:1994ws,Lu:2001ve,Gou:2010}.
+A host of techniques have been developed to lessen the computational demand of chemical kinetic calculations while maintaining fidelity, these fall broadly into a few categories: removal of unimportant species and reactions unimportant species and reactions~\cite{Lu:2006bb,Pepiot-Desjardins:2008,Hiremath:2010jw,Niemeyer:2010bt,Curtis:2015}, lumping of species with similar thermo-chemical properties~\cite{Lu:2007,Ahmed:2007fa,Pepiot:2008kq}, and time-scale methods that reduce numerical stiffness~\cite{Maas:1992ws,Lam:1994ws,Lu:2001ve,Gou:2010}.
 The interested reader is referred to the recent review of Tur\'anyi and Tomlin~\cite{turanyi2016analysis} for a comprehensive overview.
+
 In addition to the previously mentioned cost reduction methods, effort has gone into improving the integration algorithms and codes that evaluate the chemical kinetics~\cite{Gou:2010,SCHWER2002270,Niemeyer:2016aa,GAO2015287}.
-In particular, a carefully derived analytical formulation of the Jacobian can greatly increase the sparsity of the matrix~\cite{SCHWER2002270} and drop the cost of Jacobian evaluation from scaling quadratically with the number of species in the model---for a naive finite-difference implementation---to a linear dependence~\cite{LU2009192}, while Jacobian factorization can be accelerated via sparse-matrix techniques~\cite{superlu99}.
+In particular, a carefully derived analytical formulation of the Jacobian can greatly increase the sparsity of the matrix~\cite{SCHWER2002270} and drop the cost of Jacobian evaluation to a to a linear dependence on the number of species in the model~\cite{LU2009192}; Jacobian factorization can be accelerated via sparse-matrix techniques~\cite{superlu99}.
 Use of Single-Instruction, Multiple-Data (SIMD) and the related Single-Instruction, Multiple-Thread (SIMT) processors to accelerate chemical kinetic simulations has also been studied, e.g., as in~\cite{CurtisGPU:2017,Sewerin20151375,Shi:2012aa,Niemeyer:2016aa}.
@@
 \begin{itemize}
- \item Derive and validate a new Jacobian formulation that greatly increases sparsity
+ \item Derive and validate a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity
  \item Enable cross-platform SIMD\slash SIMT vectorization for the CPU, GPU and other accelerators
@@
 When storing arrays for a chemical kinetic model, the data-storage layout and vectorization patterns are critical to achieving high-performance code.
-\Cref{F:mem} depicts an example data array with K-rows and N-columns, with index ($i$, $j$) corresponding to the $i$th row and $j$th column.
+\Cref{F:mem} depicts an example data array with $K$-rows and $N$-columns, and index ($i$, $j$) corresponding to the $i$th row and $j$th column.
 For example the concentration of species $j$ for the $i$th thermo-chemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{state}$, the number of thermo-chemical states considered for evaluation, and $1 \le j \le \ns/$, the number of species in the model.
@@
 In a \textit{deep}-SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works), a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermo-chemical source-terms for a single thermo-chemical state; thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + N_v}$---the species concentrations for state 1, species $j \ldots j + N_v$---will access sequential memory locations if the data is stored in ``C''-order.
-Further, in the ``C''-ordering the furthest difference between any two species concentrations within the same thermo-chemical state is at most $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU, but may lead to uncoalesced memory-accesses on the GPU.
-In addition, a deep vectorization requires both synchronization between SIMD-lanes\slash SIMT-threads via memory-fences\slash barriers (an expensive operation), and may result in SIMD-waste\slash SIMT thread-divergence, caused by different SIMD-lanes\slash SIMT-threads executing different instructions (e.g., resulting from different if\slash then branches)
+Further, in ``C''-ordering the maximum distance between any two species concentrations within the same thermo-chemical state is $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU, but may lead to uncoalesced memory-accesses on the GPU.
+In addition, a deep vectorization requires both synchronization between SIMD-lanes\slash SIMT-threads via memory-fences\slash barriers (an expensive operation), and may result in SIMD-waste\slash SIMT thread-divergence, caused by different SIMD-lanes\slash SIMT-threads executing different instructions (e.g., resulting from different if\slash then branches).
+It is noted that generally speaking, shallow-vectorizations may also experience SIMD-waste\slash SIMT thread-divergence, e.g., in chemical kinetic integration due to varying internal solver time-step sizes~\cite{CurtisGPU:2017}; however, in this work shallow-vectorizations are largely unaffected by this concern.
 
@@
 \end{equation}
-with $\nu_{k, i}$ the net stoichiometic coefficient of species $k$ in reaction $i$, $R_{i}$ the net rate of progress of reaction $i$ (described in~\cref{deriv-s:rop_net} of the supplemental), and $c_{i}$ the pressure modification term i.e., for third-body (see~\cref{deriv-s:thdbody} of the supplemental), or fall-off\slash chemically-activated reactions (\cref{deriv-s:falloff} of the same).
+with $\nu_{k, i}$ the net stoichiometic coefficient of species $k$ in reaction $i$, $R_{i}$ the net rate of progress of reaction $i$ (described in~\cref{deriv-s:rop_net} of the supplemental), and $c_{i}$ the pressure modification term i.e., for third-body (see~\cref{deriv-s:thdbody} of the supplemental), or fall-off\slash chemically-activated reactions (\cref{deriv-s:falloff} of the same); \texttt{pyJac} is capable of evaluating all modern reaction types, e.g., P-Log and Chebyshev reactions.
 
@@
 \loadeq{dTdotdT}
-and with respect to the thermodynamic state parameter (the volume \slash pressure for \conp/\slash\conv/ respectively):
+and with respect to the thermodynamic state-parameter source-term (the volume \slash pressure for \conp/\slash\conv/ respectively):
 \loadeq{dTdotdE}
@@
 \loadeq{dEdotdE}
-and finally, by the moles of species $j$:
+and finally, with respect to the moles of species $j$:
 \loadeq{dEdotdnj}
@@
 \label{s:unittest}
-Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates user specified pseudo-code and data to OpenCL\slash C-code.
+Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates pseudo-code and data to OpenCL\slash C-code.
 As the name implies, \texttt{loo.py} generates code based on for-loops, this differs from the previous version of \texttt{pyJac}~\cite{pyjac16} which generates static code (i.e., fully unrolled loops, with thermodynamic\slash reaction parameters written directly in code rather than stored in arrays).
-In our previous work,~\cite{Niemeyer:2016aa} this static code generation caused some issues with large file sizes, long compilation times and even occasionally breaking the \texttt{gcc} and \texttt{nvcc} compilers (the latter issue necessitated breaking the Jacobian\slash source-term evaluations into separate files).
-The implications of this change will be studied in~\cref{S:jacobian_results}, where the performance of the new version of~\texttt{pyJac} will be compared to the previous versions.
+In our previous work~\cite{Niemeyer:2016aa}, this static code generation caused some issues with large file sizes, long compilation times and even occasionally breaking the \texttt{gcc} and \texttt{nvcc} compilers (the latter issue necessitated breaking the Jacobian\slash source-term evaluations into separate files).
+The implications of this change will be studied in~\cref{S:jacobian_results}, where the performance of the new version of~\texttt{pyJac} will be compared to the previous version.
 
-In addition, \texttt{loo.py} allows the user to more easily make changes the structure of the generated program, e.g., data ordering, vectorization, and threading patterns, as well as switching the target-language for code-generation (and simpler extension to additional languages, e.g., CUDA).
+In addition, \texttt{loo.py} allows the user to more easily make changes to the structure of the generated program, e.g., the data ordering, vectorization, and threading patterns, as well as switching the target-language for code-generation (and simpler extension to additional languages, e.g., CUDA).
 Further, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source-terms.
@@
 \caption{The Intel CPU configurations used in this study.
-	 The vector-width reported in this table is reported in (ideal) number of double operations per SIMD instruction, as this will be used in measuring SIMD efficiency; for reference, the vector-width in bits of the \sse/ and \avx/ machines are \SI{128}{\bit} and \SI{256}{\bit} respectively.
+	 The vector-width reported in this table is reported in (ideal) number of double operations per SIMD instruction, as this will be used in measuring SIMD-efficiency; for reference, the vector-width in bits of the \sse/ and \avx/ machines are \SI{128}{\bit} and \SI{256}{\bit} respectively.
 	 The identifier field will be used as a shorthand descriptor in the performance plots to quickly identify the CPU type.}
@@
 \Cref{t:gpus} lists the model, number of CUDA cores and NVIDIA driver of each of the GPUs used in this work.
-NVIDIA's OpenCL runtime is bundled with the NVIDIA driver~\cite{NVIDIA:2018}, hence we the driver version is used to specify the OpenCL runtime version.
+NVIDIA's OpenCL runtime is bundled with the NVIDIA driver~\cite{NVIDIA:2018}, hence the driver version is used to specify the OpenCL runtime version.
 
@@
 \end{tabular}
-\caption{The NVIDIA GPU configurations used in this study.  NVIDIA's OpenCL runtime is provided with the graphic driver, rather than any specific version of CUDA.
+\caption{The NVIDIA GPU configurations used in this study.  NVIDIA's OpenCL runtime is provided with the graphics driver, rather than any specific version of CUDA.
 Again, the identifier field will be used as a shorthand descriptor during analysis of the performance results.
@@
 Thus POCL is a particularly useful tool for validation, but not necessarily for performance studies.
-This discussion will be expanded upon in~\cref{s:future} to highlight the most promising future directions.
+This discussion will be expanded upon in~\cref{s:future} to highlight future directions.
 
@@
 \end{tabular}
-\caption{Summary of rate of progress, species, temperature and thermodynamic state parameter rate correctness.
+\caption{Summary of rate of progress, species, temperature and thermodynamic state-parameter rate correctness.
 Error statistics are based on the infinity-norm of the relative error detailed in~\cref{e:rel_err} for each quantity.
@@
 In~\cref{T:source_error}, it is seen that $E_{\varepsilon}$ is very similar in magnitude to $E_R$ in all cases, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
-The molar species production rate error norm is similar in magnitude to that of the net ROP, however the temperature and state parameter rate errors again increase as the error in net species production rates is amplified by multiplication with thermodynamic properties.
+The molar species production rate error norm is similar in magnitude to that of the net ROP, however the temperature and state-parameter rate errors again increase as the error in net species production rates is amplified by multiplication with thermodynamic properties.
 Again, we note that the above discussion does not imply that discrepancies in computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
@@
 
-As in our previous work~\cite{Niemeyer:2016aa}, the correctness of the Jacobian was determined by comparison to the Jacobian obtained by automatic differentiation of the source-term subroutines generated by \texttt{pyJac} via the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
+As in our previous work~\cite{Niemeyer:2016aa}, the correctness of the Jacobian was determined by comparison to the Jacobian obtained by automatic differentiation of the source-terms kernel generated by \texttt{pyJac} via the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
 The reasons for this choice are explained fully in~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with minimal additional programming effort.
@@
 \caption{Summary of Jacobian matrix validation results.
-Error statistics are the largest filtered relative error $E_\mathcal{C}$ for over all test platforms, vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/ and sparse\slash dense Jacobians.
+The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all test platforms, vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/ and sparse\slash dense Jacobians.
 It is noted that the threshold described in~\cref{e:thresh} varies slightly between the \conp/ and \conv/ cases; the reported $\bar{\mathcal{T}}$ is the average of the two, however the appropriate value was used during calculations of the error statistics.
@@
 Further, the LAPACK error norm does not correlate well with the stringent filtered error norm, e.g., USC-Mech II has the lowest LAPACK error norm (\num{9.632e-15}), but the second largest stringent filtered error norm (\num{1.119e-2}).
-Conversely, the LAPACK error norm does seem to correlate with the relaxed filtered error norm; the two models with the largest LAPACK error, \ce{H2}\slash\ce{CO} and \ce{iC5H11OH}, have the largest relaxed filtered error norms, $\mathcal{O}(\num{e-5})$, again suggesting that the stringent error norm is influenced by floating point error accumulation.
+Conversely, the LAPACK error norm does seem to correlate with the relaxed filtered error norm.
+The two models with the largest LAPACK error, \ce{H2}\slash\ce{CO} and \ce{iC5H11OH}, have the largest relaxed filtered error norms, $\mathcal{O}(\num{e-5})$, again suggesting that the stringent error norm is influenced by floating point error accumulation.
 These findings, along with the individual unit-testing of all chemical source-terms and Jacobian sub-components described in~\cref{s:unittest}, gives high confidence in the correctness of this new version of \texttt{pyJac}.
@@
 
-However, \texttt{pyJac} allows the user to simply ignore these derivatives (via a command-line switch) and avoid any negative adverse affects on the Jacobian sparsity.
+However, \texttt{pyJac} allows the user to simply ignore these derivatives (via a command-line switch) and avoid any adverse effects on the Jacobian sparsity.
 The rational behind this choice is that many common implicit integration techniques (e.g., CVODE~\cite{Hindmarsh:2005}) used in solution of chemical kinetic initial value problems are formulated around the assumption that the supplied Jacobian is approximate; this allows the Jacobian and its LU factorization to be reused for multiple internal integration time-steps, accelerating the solution process.
@@
 \cref{T:jac_sparsity} shows the density of the exact and approximate Jacobians for all mechanisms tested in this work.
-The smallest model, \ce{H2}\slash\ce{CO} is very dense with \textasciitilde\SI{70}{\percent} of the Jacobian entries non-zero, however this drops to \textasciitilde\SI{50}{\percent} for GRI-Mech 3.0, and continues to decrease to \textasciitilde\SI{27}{\percent} for USC-Mech II and just \textasciitilde\SI{10}{\percent} for the isopentanol model.
+The smallest model, \ce{H2}\slash\ce{CO} is very dense with \SI{71.4}{\percent} of the Jacobian entries non-zero, however this drops to \SI{56.7}{\percent} for GRI-Mech 3.0, and continues to decrease to \SI{28.7}{\percent} for USC-Mech II and just \SI{11.5}{\percent} for the isopentanol model.
 The approximate Jacobian assumption drops the density of Jacobian by \textasciitilde\SIrange{3}{7}{\percent} for all models.
@@
 The Jacobian access pattern used by \texttt{pyJac} is fairly irregular; for simplicity we will only discuss looping-structure of species derivatives calculations as these form the bulk of the computation, and have the most challenging Jacobian access patterns.
-In general, an outer loop iterates over all reactions of a certain type (e.g., falloff reactions) and calculates the relevant Jacobian subproducts---independent of any particular species---for the reaction (e.g., the the derivative of the falloff pressure modification term).
+In general, an outer loop iterates over all reactions of a certain type (e.g., falloff reactions) and calculates the relevant Jacobian subproducts---independent of any particular species---for the reaction (e.g., the derivative of the falloff pressure modification term).
 Two inner loops then iterate over the species in a reaction, updating the Jacobian entries for these species as appropriate.
@@
 Generally, this means that a lookup operation is required to find the sparse Jacobian index for any pair of state variables; in some cases this can be avoided, e.g., the rows corresponding to derivatives of the temperature and thermodynamic state parameter source-terms are fully dense in \texttt{pyJac}, and hence no lookup is necessary.
-This lookup operation is currently implemented as a simple for-loop, e.g., for a sparse lookup of a pair of indices $(i, j)$ in a compressed row storage matrix, the lookup function searches the column index vector between the values $\texttt{row\_pointer}[i] \ldots \texttt{row\_pointer}[i + 1]$ for $j$, and returns the offset from $\texttt{row\_pointer}[i]$ (or $\num{-1}$ if not found).
+This lookup operation is currently implemented as a simple for-loop, e.g., for a sparse lookup of a pair of indices $(i, j)$ in a CSR matrix, the lookup function searches the column index vector between the values $\texttt{row\_pointer}[i] \ldots \texttt{row\_pointer}[i + 1]$ for $j$, and returns the offset from $\texttt{row\_pointer}[i]$ (or $\num{-1}$ if not found).
 As will be explored in~\cref{S:jacobian_results}, this slows down sparse Jacobian evaluation, and might be improved---at the cost of increased constant data usage, a limitation for OpenCL---by a static mapping of the full Jacobian indices to the sparse index (or some null value if the entry is empty).
@@
 Additionally, the exact form of the Jacobian (as opposed the the ``approximate'' form discussed in~\cref{S:sparsity}) was used in all cases.
-Finally, unless stated otherwise: the performance results utilized the \conp/ assumption, a vector-width of \num{8}\slash\num{128} and ``C''\slash ``F''-ordered data for the CPU\slash GPU cases respectively; the run times reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
-The effects of choice of vector-width, data-ordering and differences between \conp/ and \conv/ evaluations on the CPU\slash GPU will be explored in~\cref{S:source_results}.
+Finally, unless stated otherwise: the performance results utilized a single CPU core, the \conp/ assumption, a vector-width of \num{8}\slash\num{128} and ``C''\slash ``F''-ordered data for the CPU\slash GPU cases respectively; the run times reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
+The effects of choice of vector-width, data-ordering and differences between \conp/ and \conv/ evaluations on the CPU\slash GPU will be explored in~\cref{S:source_results}, while parallel scaling for multiple CPU cores will be exampled in~\cref{S:source_results,S:jacobian_results}.
 
@@
 In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source-terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean run time per condition; the run time continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
-\Cref{F:gpu_source_speedup} shows the speedup in source-term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} for two different vector-widths (a.k.a., the block size in NVIDIA terminology), \num{64} and \num{128}; the fastest \gpunew/ case with a vector-width of 128 is \SIrange{1.40}{1.88}{$\times$} faster than the slowest case (\gpuold/ with a vector-width of \num{64}) depending on the chemical model in question.
+\Cref{F:gpu_source_speedup} shows the speedup in source-term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} with two different vector-widths (a.k.a., the block size in NVIDIA terminology), \num{64} and \num{128}; the fastest \gpunew/ case with a vector-width of 128 is \SIrange{1.40}{1.88}{$\times$} faster than the slowest case (\gpuold/ with a vector-width of \num{64}) depending on the chemical model in question.
 Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector-width has minimal effect on performance for most of the \gpunew/ and \gpuold/ cases; the GRI-Mech 3.0 and USC-Mech II models show the largest improvements with a vector-width of \num{128}, \textasciitilde\SIrange{10}{18}{\percent} for both GPUs.
@@
 In contrast,~\Cref{F:source_cvsf} shows significant speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, e.g., the \SIrange{1.35}{2.09}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-ordered OpenMP baseline.
-Additionally, it is noted that the ``C''-ordered shallow-vectorization in~\cref{F:source_cvsf} utilizes the vectorized-data ordering described in~\cref{S:data}; this case achieves speedups of \SIrange{2.14}{2.58}{$\times$} over the ``F''-ordered shallow-vectorization, demonstrating the value of the vectorized data ordering for CPU execution.
+Additionally, it is noted that the ``C''-ordered shallow-vectorization in~\cref{F:source_cvsf} (as well as the other shallow-vectorized CPU data shown in \cref{S:source_results,S:jacobian_results}) utilizes the vectorized-data ordering described in~\cref{S:data}; this case achieves speedups of \SIrange{2.14}{2.58}{$\times$} over the ``F''-ordered shallow-vectorization, demonstrating the value of the vectorized data ordering for CPU execution.
 
-\Cref{F:source_gpu_cvsf} shows the performance effect of the ``C'' versus ``F''-ordering on source-term evaluation performance on both GPUs, with the speedup presented per-GPU.
-In general, the ``C'' and ``F''-ordered shallow-vectorizations are nearly equivalent on both GPUs, with less than a \SI{10}{\percent} difference in run time between data-orderings. 
+\Cref{F:source_gpu_cvsf} shows the performance effect of the ``C'' versus ``F''-ordering on source-term evaluation performance on both GPUs, with the speedup presented per-GPU; the ``C'' and ``F''-ordered shallow-vectorizations are nearly equivalent on both GPUs, with less than a \SI{10}{\percent} difference in run time between data-orderings.
 It is noted that for the isopentanol model, ``C''-ordered data is \textasciitilde\SI{1.08}{$\times$} faster on both GPUs (while the trend is less clear for the other models).
@@
 \Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source-term evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
-In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for Intel OpenCL; although the mechanism scalings vary somewhat for OpenMP, there is no clear trend.
-It is likely the poor scaling of the \ce{H2}\slash\ce{CO} for Intel OpenCL results from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD instructions.
+In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for both Intel OpenCL and OpenMP, likely resulting from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD instructions\slash multiple threads.
 Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.
 This is due in large part to the superior performance of the shallow-vectorized OpenCL code, coupled with the relatively small amounts of work associated with source-term evaluation.
-To illustrate this,~\cref{F:source_crossover} shows the mean run time per-condition of the OpenMP source-term evaluations for \numrange{2}{16} cores, compared to the shallow-vectorized OpenCL code on \num{1} (solid line) and \num{4} (dashed line) four cores on the \avx/ machine.
+To illustrate this,~\cref{F:source_crossover} shows the mean run time per-condition of the OpenMP source-term evaluations for \numrange{2}{16} cores, compared to the shallow-vectorized OpenCL code on \num{1} (solid line) and \num{4} (dashed line) cores on the \avx/ machine.
 For all chemical models, the mean run time per-condition (and hence the computational work allocated per-core, one of the key-drivers of parallel scaling efficiency~\cite{strong_scaling}) of OpenMP running on \num{4} cores is roughly equal to that of the shallow-vectorized OpenCL code on a single core.
@@
 Therefore, a more fair comparison of parallel scaling efficiencies is to compare OpenCL running on \num{4} cores, to OpenMP on \num{16}; the OpenMP code's parallel efficiency drops to \textasciitilde\num{0.64} for \num{16} cores, similar to OpenCL's parallel scaling efficiency of \numrange{0.66}{0.72} at \num{4} cores.
-Indeed, as will be investigated in~\cref{S:jacobian_results} the strong parallel scaling efficiency for sparse Jacobian evaluation---the most computationally intensive task in this work---are far more similar between Intel OpenCL and OpenMP.
+Indeed, as will be seen in~\cref{S:jacobian_results} the strong parallel scaling efficiency for sparse Jacobian evaluation---the most computationally intensive task in this work---are far more similar between Intel OpenCL and OpenMP.
 
@@
 The sparse Jacobian evaluations are slower on both Intel OpenCL and OpenMP due to indirect lookup indexing requirements, as discussed in~\cref{S:sparsity}.
-Interestingly the shallow-vectorized OpenCL code is less negatively impacted by the indirect lookup; the sparse OpenMP code is \SIrange{2.47}{10.42}{$\times$} slower than the dense OpenMP evaluation, while the sparse shallow-vectorized OpenCL code is just \SIrange{1.41}{3.34}{$\times$} slower than its dense counterpart
+Interestingly the shallow-vectorized OpenCL code is less negatively impacted by the indirect lookup; the sparse OpenMP code is \SIrange{2.47}{10.42}{$\times$} slower than the dense OpenMP evaluation, while the sparse shallow-vectorized OpenCL code is just \SIrange{1.41}{3.34}{$\times$} slower than its dense counterpart.
 As a result, the shallow-vectorized sparse OpenCL code is as fast, or faster than the dense OpenMP code in all cases on the \avx/ machine (\cref{F:sparse_vs_dense}).
@@
 In~\cref{F:sparse_jac_scaling}, the strong parallel scaling efficiency of the sparse shallow-vectorized OpenCL is compared to the sparse OpenMP code.
-The plot is a bit difficult to read, as most of the data-points are clustered together, however the key takeaway is that the parallel scaling efficiency of the shallow-vectorized OpenCL code is very similar to that of the OpenMP code, in contrast to the source-terms parallel scaling efficiencies (\cref{F:source_parallel_scaling}).
+The plot is a bit difficult to read, as most of the data-points are clustered together, however the key takeaway is that the parallel scaling efficiency of the shallow-vectorized OpenCL code is very similar to that of the OpenMP code, in contrast to the parallel scaling efficiency of source-term evaluation (\cref{F:source_parallel_scaling}).
 The \ce{H2}\slash\ce{CO} model has the worst scaling for both codes, ranging from \numrange{0.94}{0.54} and \numrange{0.99}{0.82} for OpenCL and OpenMP respectively on \numrange{2}{16} cores.
@@
 Sparse Jacobian evaluation is significantly slower for all models before the GPU becomes saturated (due to the indirect indexing lookup), however past the saturation point the performance gap between the sparse and dense evaluations narrows.
-This is likely caused by the combination of being able to fit many more sparse Jacobian matrices in the \gpunew/'s memory, as well as improved data-locality\slash caching due to the smaller size of the sparse Jacobian.
+This is likely due to the ability to fit many more sparse Jacobian matrices in the \gpunew/'s memory, as well as improved data-locality\slash caching due to the smaller size of the sparse Jacobian.
 \Cref{F:gpu_jacobian_speedup} presents the speedup of the \gpunew/ over the \gpuold/ GPU for sparse\slash dense Jacobian evaluation; sparse evaluation on the \gpunew/ is \SIrange{1.10}{1.59}{$\times$} faster than on the \gpuold/, while dense evaluation is \SIrange{1.36}{3.0}{$\times$} faster.
-The speedup on the \gpunew/ appears to decrease with increasing mechanism size for the sparse evaluation, but increases for the larger models in dense evaluation; this is likely due to the larger memory size of the \gpunew/, and hence fewer data-transfer operations.
+The speedup on the \gpunew/ appears to decrease with increasing mechanism size for the sparse evaluation, but increases for the larger models in dense evaluation; this is likely due to the larger available memory of the \gpunew/, and hence fewer data-transfer operations to\slash from the GPU.
 
@@
 It is noted that the isopentanol case was excluded here, as a single run of the sparse finite-difference Jacobian using either OpenCL or OpenMP took over \num{12} hours of run time.
-In addition the current finite-difference formulation breaks Intel's auto-vectorizer, hence the OpenCL speedup comparison is made against the unvectorized OpenCL code; it is not a high priority to fix this issue, as the finite-difference Jacobian is implemented for comparison purposes only.
+In addition the current finite-difference formulation breaks Intel's auto-vectorizer, hence the OpenCL speedup comparison is made against the unvectorized OpenCL code; it is not a high priority to fix this issue, as the finite-difference Jacobian was implemented for comparison purposes only.
 Although we do not display the dense finite-difference Jacobian speedup in~\cref{F:fd_vs_analytical_cpu}, the dense OpenCL and OpenMP analytical Jacobian codes out-perform the finite-difference variants by even larger margins; \SIrange{24.44}{245.63}{$\times$} for OpenCL, and \SIrange{9.68}{112.73}{$\times$} for OpenMP (these data do include the isopentanol model, though limited to \num{50000} conditions as discussed previously).
 A comparison of the sparse analytical and finite-difference Jacobians on the GPUs is presented in~\cref{F:fd_vs_analytical_gpu}.
-The analytical Jacobian on the \gpunew/ and \gpuold/ shows large speedups, increasing with chemical model size, in the range of \SIrange{3.81}{17.60}{$\times$}; the \gpunew/ has a larger speedup than the \gpuold/ (\SI{17.60}{$\times$} v.s., \SI{14.75}{$\times$}) for the isopentanol model due to it's larger memory size.
+The analytical Jacobian on the \gpunew/ and \gpuold/ shows large speedups, increasing with chemical model size, in the range of \SIrange{3.81}{17.60}{$\times$}; the \gpunew/ has a larger speedup than the \gpuold/ (\SI{17.60}{$\times$} v.s., \SI{14.75}{$\times$}) for the isopentanol model due to its larger available memory.
 Although not pictured, the dense analytical-Jacobian on the \gpunew/ GPU has larger speedups compared with the dense finite-difference Jacobian: \SIrange{4.04}{45.13}{$\times$}; the \gpunew/ GPU again has a significantly larger speedup than the \gpuold/ for the isopentanol model (\SI{45.13}{$\times$} vs \SI{23.85}{$\times$}), further underscoring the effect of the larger available memory on the \gpunew/.
@@
 Indeed the vectorization attempted by the OpenCL runtimes (or reasons for incorrectly vectorized code) are not always obvious to the user.
-It is possible, particularly for the Intel OpenCL \slash POCL runtimes, that better performance might be achieved using so called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}.
+It is possible, particularly for the Intel OpenCL\slash POCL runtimes, that better performance might be achieved using so called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}.
 This will be investigated in future works.
@@
 
+Finally, one key-component missing in this current version of \texttt{pyJac} is vectorized sparse\slash dense linear-algebra subroutines to maximize the performance of LU-factorization and matrix-vector multiplication (commonly used in implicit-integration techniques).
+Third-party\slash open-source options exist for some target-languages, e.g., cuBLAS~\cite{cublas},clBLAS\slash clSPARSE~\cite{clmath} or SuperLU~\cite{superlu99}, but these do not necessarily cover all targets\slash necessary linear-algebra operations and, in the case of the CUDA\slash OpenCL libraries, are often optimized operation on one-large matrix instead of many (relatively) smaller matrices.
+The extent to which these existing programs may be used needs to be assessed and missing operations should be implemented in \texttt{loo.py} to ensure easy switching between target-languages\slash vectorization types.
+
 \section{Conclusions}
-In this work, automatically generated OpenCL codes for shallow SIMD-vectorized thermo-chemical source-term evaluation were developed and validated against Cantera~\cite{Cantera} for a shallow range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
-Significant speedups of up to \SIrange{1.90}{2.45}{$\times$} over a baseline SIMT-vectorized code were observed.
-Two data-ordering schemes were investigated, showing a clear performance benefit for use of the ``C'' (row-major) ordering for the shallow vectorized code.
-Further, this study suggests that a deep SIMD-vectorized code---currently under development---may see even greater accelerations due to expected increases in cache-hit rates.
-Larger vector-widths were found to provide accelerations for the smaller models studied, but the speedup was similar over all vector-widths tested for the largest models investigated.
-Finally, the strong parallel scaling efficiency was examined for the shallow vectorized code; the larger models exhibited higher scaling efficiency, possibly due to failure of the \ce{H2}\slash\ce{CO} model to saturate the CPU throughput.
-Future extensions of this work will include development of a SIMD\slash SIMT-accelerated sparse-analytical Jacobian code targeted at CPUs, GPUs, and MICs.
+In this work, automatically generated OpenCL codes for SIMD\slash SIMT-vectorized thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation were developed and validated against for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
+Significant speedups in shallow SIMD-vectorized source-term evaluation over a parallel OpenMP code were demonstrated, up to \SIrange{2.53}{2.92}{$\times$} and \SIrange{3.40}{4.08}{$\times$} for \sse/ and \avx/ capable CPUs; in addition, source-term evaluation was \SIrange{1.40}{1.88}{$\times$} faster on a \gpunew/ GPU compared to a \gpuold GPU; deep-vectorization, though possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, did not yield any performance benefit as POCL did not achieve vectorizations for any execution pattern studied.
+Vectorized data-orderings specifically formulated for shallow\slash deep SIMD\slash SIMT vectorizations were proposed, resulting in a shallow-vectorized ``C'' (row major) speedup of \SI{2.14}{2.58}{$\times$} over a ``F'' (column major) ordered shallow-vectorization, compared to a speedup of \SI{1.35}{2.09}{$\times$} for the ``C''-ordered OpenMP code (over an ``F''-ordered equivalent).
+In addition, the effect of the \conp/\slash\conv/ formulation was determined to be negligible on source-term evaluations, while varying the vector-width size had minimal (\textasciitilde\SI{10}{$\percent$}) performance effects on both the CPU and GPU.
+The strong parallel scaling efficiency was examined for source-term evaluation; the parallel OpenMP code generally had better efficiencies, but when compared to a shallow-vectorized code with roughly equivalent performance---e.g., source-term shallow-vectorized evaluation on \num{4} cores to \num{16} OpenMP cores---the scalings were similar.
+
+The performance of sparse and dense Jacobian evaluation was also characterized on the CPU\slash GPU platforms.
+The sparse shallow-vectorized OpenCL code was \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.14}{5.03}{$\times$} faster than OpenMP evaluation on the \avx/ and \sse/ CPUs respectively, while dense shallow-vectorized evaluation was \SIrange{3.03}{4.23}{$\times$} and \SIrange{1.92}{2.47}{$\times$} faster on the same.
+Sparse Jacobian evaluation was slower than dense evaluation on all CPU\slash GPU platforms due to indirect lookup requirements in array indexing, however the shallow-vectorized OpenCL code was less adversely affected---a slowdown of \SIrange{1.41}{3.34}{$\times$} on the \avx/ CPU---than OpenMP, which was \SIrange{2.47}{10.42}{$\times$} slower on the same.
+On the \gpunew/ GPU, sparse evaluation was \SIrange{1.10}{1.59}{$\times$} faster than on the \gpuold/ GPU, while dense evaluation was \SIrange{1.36}{3.0}{$\times$} faster; the use of OpenCL Image\slash CUDA Texture memory to accelerate the indirect lookup for sparse matrix evaluation should be investigated in the future.
+Analytical Jacobian evaluation was significantly faster than a first-order finite-difference based approach on all platforms, achieving speedups of \SIrange{3.81}{17.60}{$\times$} and \SIrange{4.04}{45.13}{$\times$} for sparse and dense evaluation (respectively) on the GPUs, while on the \avx/ CPU analytical evaluation was \SIrange{3.92}{55.11}{$\times$} and \SIrange{9.68}{245.63}{$\times$} for sparse and dense (respectively) OpenMP\slash shallow-vectorized OpenCL evaluation.
+Finally, the performance of dense analytical-Jacobian evaluation in this new version of \texttt{pyJac} was compared to the previous version~\cite{pyjac16} on the \sse/\slash\gpuold/ CPU\slash GPUs; OpenMP evaluation was \SI{1.79}{$\times$} slower for the smallest chemical model, but \SIrange{3.37}{10.19}{$\times$} faster for the larger models, while the shallow-vectorized OpenCL code was \SIrange{1.37}{19.56}{$\times$} faster than the previous version over all chemical models;  similarly, evaluation on the \gpuold/ GPU was \SI{1.89}{$\times$} slower for the smallest model but \SIrange{1.25}{2.84}{$\times$} faster for the larger models.
+
+Future extensions of this work will include additional target-languages (vectorized OpenMP, CUDA, ISPC) to improve ease of use and reliability, as well as to enable meaningful deep-vectorization.
+Additionally, existing vectorized sparse\slash dense-matrix libraries will be assessed and missing operations for LU-factorization\slash matrix-vector multiplication implemented.
 
@@
 
+\pagebreak
 \bibliography{paper}

### 2018-03-13 paper.tex
first draft -- only needs final revisions

@@
 \begin{abstract} % not to exceed 200 words
-A code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation was developed and validated for a wide range of chemical kinetic models.
-
-\todo[inline]{this}
+A code generation platform, \texttt{pyJac}, for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized constant-pressure\slash constant-volume thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation was developed and validated for a wide range of chemical kinetic models.
+Speedups of \SIrange{3.40}{4.08}{$\times$} were found for shallow-vectorized OpenCL source-rate evaluation compared a parallel OpenMP code on an \avx/ CPU, increasing to \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.03}{4.23}{$\times$} for sparse and dense chemical kinetic Jacobian evaluation; in addition, the performance of a \gpunew/ GPU was compared to a \gpuold/ finding speedups of \SIrange{1.40}{1.88}{$\times$}, \SIrange{1.10}{1.59}{$\times$}, and \SIrange{1.36}{3.0}{$\times$} for source-rate, sparse Jacobian and dense Jacobian evaluation, respectively.
+Additionally, the effect of data-ordering was investigated and an ordering specifically formulated for vectorized evaluation proposed; as well, the effect of the \conp/\slash\conv/ assumptions and varying vector widths were studied on source-term evaluation performance.
+Speedups over a first-order finite-difference approach reached \SIrange{17.60}{45.13}{$\times$} and \SIrange{55.11}{245.63}{$\times$} for sparse\slash dense evaluation on the CPU\slash GPU respectively, while dense Jacobian evaluation was up to \SI{19.56}{$\times$} and \SI{2.84}{$\times$} times faster than a previous version of \texttt{pyJac} on a \sse/ CPU and \gpuold/ GPU respectively.
+Finally, future directions for vectorized chemical kinetic evaluation and sparse linear-algebra techniques were discussed.
 \end{abstract}
@@
 Finally, one key-component missing in this current version of \texttt{pyJac} is vectorized sparse\slash dense linear-algebra subroutines to maximize the performance of LU-factorization and matrix-vector multiplication (commonly used in implicit-integration techniques).
-Third-party\slash open-source options exist for some target-languages, e.g., cuBLAS~\cite{cublas},clBLAS\slash clSPARSE~\cite{clmath} or SuperLU~\cite{superlu99}, but these do not necessarily cover all targets\slash necessary linear-algebra operations and, in the case of the CUDA\slash OpenCL libraries, are often optimized operation on one-large matrix instead of many (relatively) smaller matrices.
+Third-party\slash open-source options exist for some target-languages, e.g., cuBLAS~\cite{cublas},clBLAS\slash clSPARSE~\cite{clmath} or SuperLU~\cite{superlu99}, but these do not necessarily cover all targets\slash required linear-algebra operations and, in the case of the CUDA\slash OpenCL libraries, are often optimized operation on one-large matrix instead of many (relatively) smaller matrices.
 The extent to which these existing programs may be used needs to be assessed and missing operations should be implemented in \texttt{loo.py} to ensure easy switching between target-languages\slash vectorization types.
@@
 In this work, automatically generated OpenCL codes for SIMD\slash SIMT-vectorized thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation were developed and validated against for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
-Significant speedups in shallow SIMD-vectorized source-term evaluation over a parallel OpenMP code were demonstrated, up to \SIrange{2.53}{2.92}{$\times$} and \SIrange{3.40}{4.08}{$\times$} for \sse/ and \avx/ capable CPUs; in addition, source-term evaluation was \SIrange{1.40}{1.88}{$\times$} faster on a \gpunew/ GPU compared to a \gpuold GPU; deep-vectorization, though possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, did not yield any performance benefit as POCL did not achieve vectorizations for any execution pattern studied.
-Vectorized data-orderings specifically formulated for shallow\slash deep SIMD\slash SIMT vectorizations were proposed, resulting in a shallow-vectorized ``C'' (row major) speedup of \SI{2.14}{2.58}{$\times$} over a ``F'' (column major) ordered shallow-vectorization, compared to a speedup of \SI{1.35}{2.09}{$\times$} for the ``C''-ordered OpenMP code (over an ``F''-ordered equivalent).
+Significant speedups in shallow SIMD-vectorized source-term evaluation over a parallel OpenMP code were demonstrated, up to \SIrange{2.53}{2.92}{$\times$} and \SIrange{3.40}{4.08}{$\times$} for \sse/ and \avx/ capable CPUs; in addition, source-term evaluation was \SIrange{1.40}{1.88}{$\times$} faster on a \gpunew/ GPU compared to a \gpuold/ GPU; deep-vectorization, though possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, did not yield any performance benefit as POCL did not achieve vectorizations for any execution pattern studied.
+Vectorized data-orderings specifically formulated for shallow\slash deep SIMD\slash SIMT vectorizations were proposed, resulting in a shallow-vectorized ``C'' (row major) speedup of \SIrange{2.14}{2.58}{$\times$} over a ``F'' (column major) ordered shallow-vectorization, compared to a speedup of \SIrange{1.35}{2.09}{$\times$} for the ``C''-ordered OpenMP code (over an ``F''-ordered equivalent).
 In addition, the effect of the \conp/\slash\conv/ formulation was determined to be negligible on source-term evaluations, while varying the vector-width size had minimal (\textasciitilde\SI{10}{$\percent$}) performance effects on both the CPU and GPU.

### 2018-03-13 paper.tex
finalize error disc. & add appendix of per-platform error

--- 2018-03-13 paper.tex
@@
 \usepackage[english,capitalise]{cleveref}
+
+% https://tex.stackexchange.com/a/146707/56227
+\crefname{appendix}{}{}
 
@@
 \begin{abstract} % not to exceed 200 words
-A code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation was developed and validated for a wide range of chemical kinetic models.
-
-\todo[inline]{this}
+A code generation platform, \texttt{pyJac}, for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized constant-pressure\slash constant-volume thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation was studied for a selection of chemical kinetic models covering the range of sizes typically used in reactive-flow simulations.
+A new formulation of the chemical kinetic governing equations was derived and validated, resulting Jacobian sparsities of \SIrange{28.6}{92.0}{$\percent$} for the tested models.
+Speedups of \SIrange{3.40}{4.08}{$\times$} were found for shallow-vectorized OpenCL source-rate evaluation compared a parallel OpenMP code on an \avx/ central processing unit (CPU), increasing to \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.03}{4.23}{$\times$} for sparse and dense chemical kinetic Jacobian evaluation; in addition, the performance of a NVIDIA \gpunew/ graphics processing unit (GPU) was compared to a NVIDIA \gpuold/ GPU finding speedups of \SIrange{1.40}{1.88}{$\times$}, \SIrange{1.10}{1.59}{$\times$}, and \SIrange{1.36}{3.0}{$\times$} for source-rate, sparse and dense Jacobian evaluation, respectively.
+Additionally, the effect of data-ordering was investigated and an storage pattern specifically formulated for vectorized evaluation proposed; as well, the effect of the \conp/\slash\conv/ assumptions and varying vector widths were studied on source-term evaluation performance.
+Speedups over a first-order finite-difference approach reached \SIrange{17.60}{45.13}{$\times$} and \SIrange{55.11}{245.63}{$\times$} for sparse\slash dense evaluation on the CPU\slash GPU respectively, while dense Jacobian evaluation was up to \SI{19.56}{$\times$} and \SI{2.84}{$\times$} times faster than a previous version of \texttt{pyJac} on a \sse/ CPU and \gpuold/ GPU respectively.
+Finally, future directions for vectorized chemical kinetic evaluation and sparse linear-algebra techniques were discussed.
 \end{abstract}
@@
 First, the state vector results in highly sparse chemical kinetic Jacobians as will be detailed in~\cref{S:sparsity}.
-Second, mass is explicitly conserved in this formulation owing to the fact that the moles of the final species is calculated from the ideal gas law, and the rate of change of the final species from conservation of mass (see \cref{deriv-s:state,deriv-s:source} of the supplemental material respectively); additionally, system is not over-constrained and does not require use of a more complicated DAE solver (as compared to an ODE integrator) for integration\todo{hansen citation here}.
+Second, mass is explicitly conserved in this formulation owing to the fact that the moles of the final species is calculated from the ideal gas law, and the rate of change of the final species from conservation of mass (see \cref{deriv-s:state,deriv-s:source} of the supplemental material respectively); additionally, system is not over-constrained and does not require use of a more complicated DAE solver (as compared to an ODE integrator) for integration.
 Finally, the chemical kinetic Jacobian changes relatively little between the \conp/ and \conv/ forms, making maintaining the code-base much simpler.
@@
 \subsection{Jacobian validation}
+\label{S:jac_valid}
 
@@
 \midrule
-\ce{H2}\slash \ce{CO} & \num{1.048e-11}      & \num{6.431e+18}      & \num{1.741e+0}  & \num{4.508e-5} \\
-GRI-Mech 3.0          & \num{1.124e-14}      & \num{7.783e+19}      & \num{3.842e-7}  & \num{3.687e-7} \\
-USC-Mech II           & \num{9.632e-15}      & \num{2.830e+21}      & \num{1.119e-2}  & \num{1.983e-7} \\
+\ce{H2}\slash \ce{CO} & \num{1.862e-14}      & \num{6.431e+18}      & \num{1.741e+0}  & \num{4.508e-5} \\
+GRI-Mech 3.0          & \num{1.567e-14}      & \num{7.783e+19}      & \num{3.842e-7}  & \num{3.687e-7} \\
+USC-Mech II           & \num{1.137e-14}      & \num{2.830e+21}      & \num{1.199e-2}  & \num{1.983e-7} \\
 \ce{iC5H11OH}         & \num{1.227e-10}      & \num{2.733e+26}      & \num{1.363e-3}  & \num{2.764e-5} \\
@@
 It is noted that for GRI-Mech 3.0, which has a $\mathcal{T}$ roughly an order of magnitude larger, the stringent error norm is significantly smaller: $\mathcal{O}(\num{e-7})$.
-Given the intricacy of floating point error evaluation, the use of different languages and OpenCL platforms (the effect of these differences will be explored in Appendix \textbf{???}\todo{this}), and the general complexity of \texttt{pyJac} it would be exceedingly difficult to pinpoint an exact cause for this phenomena, as was done in~\cref{s:validation}.
+Given the intricacy of floating point error evaluation, the use of different languages and OpenCL platforms (the effect of these differences will be explored in \cref{A:per_platform}), and the general complexity of \texttt{pyJac} it would be exceedingly difficult to pinpoint an exact cause for this phenomena, as was done in~\cref{s:validation}.
 To ensure there is no bug or error present in the Jacobians generated by \texttt{pyJac}, a relaxed filtered error norm ($\mathcal{C} = 10^{15}$) is also presented for each model in~\cref{T:error}.
@@
 The relative LAPACK error norm---ranging from \textasciitilde\numrange{e-10}{e-14}---re-enforces this finding, as it indicates roughly \numrange{10}{14} digits of accuracy~\cite{Anderson:1999aa}.
-Further, the LAPACK error norm does not correlate well with the stringent filtered error norm, e.g., USC-Mech II has the lowest LAPACK error norm (\num{9.632e-15}), but the second largest stringent filtered error norm (\num{1.119e-2}).
-Conversely, the LAPACK error norm does seem to correlate with the relaxed filtered error norm.
-The two models with the largest LAPACK error, \ce{H2}\slash\ce{CO} and \ce{iC5H11OH}, have the largest relaxed filtered error norms, $\mathcal{O}(\num{e-5})$, again suggesting that the stringent error norm is influenced by floating point error accumulation.
+The \ce{iC5H11OH} model has the largest LAPACK error norm, likely due to the presence of P-Log\slash Chebyshev reactions and associated complicated derivatives with many logarithms, exponetiations and summations.
+Further, the LAPACK error norm does not correlate well with the stringent filtered error norm, e.g., USC-Mech II has the smallest LAPACK error norm (\num{1.137e-14}), but the second largest stringent filtered error norm (\num{1.119e-2}).
+Conversely, the model with the largest LAPACK error norm, \ce{iC5H11OH} (with $E_{\mathcal{L}} = \num{1.227e-10}$), has the second smallest stringent filtered error norm (\num{1.363e-3}) , again suggesting that the stringent error norm is influenced by floating point error accumulation.
 These findings, along with the individual unit-testing of all chemical source-terms and Jacobian sub-components described in~\cref{s:unittest}, gives high confidence in the correctness of this new version of \texttt{pyJac}.
@@
 \label{s:future}
-It is important to note that while OpenCL provides an easy way to enable cross-platform execution, and significant speedups were achieved via shallow-vectorized OpenCL code in this work, there are some serious potential pit-falls in its use.
-Typically speaking, the closed source OpenCL runtimes tested in this work (Intel and NVIDIA) contain bugs that result in compilation failures, simply incorrect vectorized machine code or even segmentation faults; further, these runtimes (in our experience) tend to be less responsive to fixing said bugs, with relatively infrequent new releases (or in NVIDIA's case, changelogs\slash public records of bug-fixes).
+It is important to note that while OpenCL provides an simple interface to enable cross-platform execution, and significant speedups were achieved via shallow-vectorized OpenCL code in this work, there are some serious potential pit-falls in its use.
+The closed source OpenCL runtimes tested in this work (Intel and NVIDIA) contain several bugs that result in compilation failures, simply incorrect vectorized machine code or even segmentation faults.
+Further, these runtimes (in our experience) tend to be less responsive to fixing said bugs, with relatively infrequent new releases (or in NVIDIA's case, lack of changelogs\slash public records of bug-fixes).
 On the other hand, the open-source OpenCL runtime used in this work, POCL, has (in our experience) far fewer implementation bugs, and when issues arise the community is very responsive to bug reports and user outreach; however, POCL fails to achieve vectorization as noted in~\cref{S:source_results}.
 
-As demonstration of the type of issue we're discussing, a minimum working example has been created~\cite{nvidia_mwe} that demonstrates a failure of NVIDIA's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla K40m GPU; by simply changing to another runtime (e.g., Intel), with no other code changes, the correct result can be found.
+As demonstration of the type of issue we're discussing, a minimum working example has been created~\cite{nvidia_mwe} that demonstrates a failure of NVIDIA's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla \gpunew/ GPU; by simply changing to another runtime (e.g., Intel)---with no code changes or recompilation---the correct result can be found.
 This provides a particularly vexing problem for the programmer, as there is often little that can be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
-Further, as noted throughout this work certain transformations or code-generation patterns can break OpenCL execution\slash vectorization; e.g., the failure of POCL to achieve vectorization or Intel OpenCL's failure to vectorize the finite-difference Jacobian.
-Indeed the vectorization attempted by the OpenCL runtimes (or reasons for incorrectly vectorized code) are not always obvious to the user.
-It is possible, particularly for the Intel OpenCL\slash POCL runtimes, that better performance might be achieved using so called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}.
-This will be investigated in future works.
-
-Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change is building the wrapper that initializes\slash transfers memory and calls the kernel.
-Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler pragmas) is a key priority going forwards.
-In addition, CUDA~\cite{NVIDIA:2018} and Intel's open-source OpenCL alternative ISPC~\cite{pharr2012ispc} have been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017} and (in the case of ISPC) initial testing.
+Further, as noted throughout this work certain code-generation patterns can break OpenCL execution\slash vectorization, e.g., the failure of POCL to achieve vectorization or Intel OpenCL's failure to vectorize the finite-difference Jacobian.
+Indeed the vectorization process attempted by most OpenCL runtimes (and thus, reasons for incorrect\slash unvectorized code-output) are obscured from the user, making reasoning about errors or performance trends quite difficult.
+
+Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change requires building of the wrapper that initializes\slash transfers memory and calls the source-rate\slash Jacobian kernel.
+Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py}) is a key priority going forwards as OpenMP is a standard library on most machines.
+In addition, CUDA~\cite{NVIDIA:2018}, and Intel's open-source OpenCL alternative, ISPC~\cite{pharr2012ispc}, have been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017} and (in the case of ISPC) initial testing.
 The current deep-vectorization formulation would be executable for both CUDA and ISPC targets, as these languages implement double-precision atomic operations, further recommending their use.
-Extending execution to these targets to make CPU\slash GPU more reliable vectorization is a future goal.
+It is also possible, particularly for the Intel OpenCL\slash POCL runtimes, that better performance\slash stability might be achieved using so-called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}; this may also enable deep-vectorization on these runtimes.
+Extending execution to new targets and improvements to the existing OpenCL targets is a future goal to improve performance, reliability and ease of use for CPU\slash GPU vectorizated execution.
 
 Finally, one key-component missing in this current version of \texttt{pyJac} is vectorized sparse\slash dense linear-algebra subroutines to maximize the performance of LU-factorization and matrix-vector multiplication (commonly used in implicit-integration techniques).
-Third-party\slash open-source options exist for some target-languages, e.g., cuBLAS~\cite{cublas},clBLAS\slash clSPARSE~\cite{clmath} or SuperLU~\cite{superlu99}, but these do not necessarily cover all targets\slash necessary linear-algebra operations and, in the case of the CUDA\slash OpenCL libraries, are often optimized operation on one-large matrix instead of many (relatively) smaller matrices.
+Third-party\slash open-source options exist for some target-languages, e.g., cuBLAS~\cite{cublas}, clBLAS\slash clSPARSE~\cite{clmath} or SuperLU~\cite{superlu99}, but these do not necessarily cover all targets\slash required linear-algebra operations and, in the case of the CUDA\slash OpenCL libraries, are often optimized operation on one large matrix instead of many (relatively) smaller matrices.
 The extent to which these existing programs may be used needs to be assessed and missing operations should be implemented in \texttt{loo.py} to ensure easy switching between target-languages\slash vectorization types.
@@
 \section{Conclusions}
-In this work, automatically generated OpenCL codes for SIMD\slash SIMT-vectorized thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation were developed and validated against for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
-Significant speedups in shallow SIMD-vectorized source-term evaluation over a parallel OpenMP code were demonstrated, up to \SIrange{2.53}{2.92}{$\times$} and \SIrange{3.40}{4.08}{$\times$} for \sse/ and \avx/ capable CPUs; in addition, source-term evaluation was \SIrange{1.40}{1.88}{$\times$} faster on a \gpunew/ GPU compared to a \gpuold GPU; deep-vectorization, though possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, did not yield any performance benefit as POCL did not achieve vectorizations for any execution pattern studied.
-Vectorized data-orderings specifically formulated for shallow\slash deep SIMD\slash SIMT vectorizations were proposed, resulting in a shallow-vectorized ``C'' (row major) speedup of \SI{2.14}{2.58}{$\times$} over a ``F'' (column major) ordered shallow-vectorization, compared to a speedup of \SI{1.35}{2.09}{$\times$} for the ``C''-ordered OpenMP code (over an ``F''-ordered equivalent).
-In addition, the effect of the \conp/\slash\conv/ formulation was determined to be negligible on source-term evaluations, while varying the vector-width size had minimal (\textasciitilde\SI{10}{$\percent$}) performance effects on both the CPU and GPU.
+In this work, automatically generated OpenCL codes for SIMD\slash SIMT-vectorized thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation were developed and validated for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
+A new formulation resulted in highly sparse (\SIrange{28.6}{88.5}{\percent}) chemical kinetic Jacobians, with options to futher increase sparsity (to \SIrange{31.66}{92.0}{\percent}) at the expense of strict correctness.
+Significant speedups in shallow SIMD-vectorized source-term evaluation over a parallel OpenMP code were demonstrated, up to \SIrange{2.53}{2.92}{$\times$} and \SIrange{3.40}{4.08}{$\times$} for \sse/ and \avx/ capable CPUs; in addition, source-term evaluation was \SIrange{1.40}{1.88}{$\times$} faster on a \gpunew/ GPU compared to a \gpuold/ GPU.
+Deep-vectorization, though possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, did not yield any performance benefit as POCL did not achieve vectorizations for any execution pattern studied; deep-vectorization deserves further study with other platforms (e.g., CUDA).
+Vectorized data-orderings specifically formulated for shallow\slash deep SIMD\slash SIMT vectorizations were proposed, resulting in a shallow-vectorized speedup of \SIrange{2.14}{2.58}{$\times$} over a more standard ``F''-ordered (column major) ordered shallow-vectorization; when compared to the corresponding speedup of \SIrange{1.35}{2.09}{$\times$} for the ``C''-ordered (row major) OpenMP code over an ``F''-ordered equivalent, the value of the vectorized data-ordering is evident.
+In addition, the choice of \conp/\slash\conv/ formulation was determined to be negligible on source-term evaluations, while varying the OpenCL vector-width had only modest (\textasciitilde\SI{10}{$\percent$}) performance effects on both the CPU and GPU.
 The strong parallel scaling efficiency was examined for source-term evaluation; the parallel OpenMP code generally had better efficiencies, but when compared to a shallow-vectorized code with roughly equivalent performance---e.g., source-term shallow-vectorized evaluation on \num{4} cores to \num{16} OpenMP cores---the scalings were similar.
+Examination of SIMD-efficiency revealed increasing speedups with chemical model-size for source-term evaluation on the \avx/ CPU, while on the \sse/ CPU speedups larger than the nominal vector-width were achieved.
 
@@
 On the \gpunew/ GPU, sparse evaluation was \SIrange{1.10}{1.59}{$\times$} faster than on the \gpuold/ GPU, while dense evaluation was \SIrange{1.36}{3.0}{$\times$} faster; the use of OpenCL Image\slash CUDA Texture memory to accelerate the indirect lookup for sparse matrix evaluation should be investigated in the future.
-Analytical Jacobian evaluation was significantly faster than a first-order finite-difference based approach on all platforms, achieving speedups of \SIrange{3.81}{17.60}{$\times$} and \SIrange{4.04}{45.13}{$\times$} for sparse and dense evaluation (respectively) on the GPUs, while on the \avx/ CPU analytical evaluation was \SIrange{3.92}{55.11}{$\times$} and \SIrange{9.68}{245.63}{$\times$} for sparse and dense (respectively) OpenMP\slash shallow-vectorized OpenCL evaluation.
-Finally, the performance of dense analytical-Jacobian evaluation in this new version of \texttt{pyJac} was compared to the previous version~\cite{pyjac16} on the \sse/\slash\gpuold/ CPU\slash GPUs; OpenMP evaluation was \SI{1.79}{$\times$} slower for the smallest chemical model, but \SIrange{3.37}{10.19}{$\times$} faster for the larger models, while the shallow-vectorized OpenCL code was \SIrange{1.37}{19.56}{$\times$} faster than the previous version over all chemical models;  similarly, evaluation on the \gpuold/ GPU was \SI{1.89}{$\times$} slower for the smallest model but \SIrange{1.25}{2.84}{$\times$} faster for the larger models.
+Analytical Jacobian evaluation was significantly faster than a first-order finite-difference based approach on all platforms, achieving speedups of \SIrange{3.81}{17.60}{$\times$} and \SIrange{4.04}{45.13}{$\times$} for sparse and dense evaluation (respectively) on the GPUs; on the \avx/ CPU analytical evaluation was \SIrange{3.92}{55.11}{$\times$} and \SIrange{9.68}{245.63}{$\times$} for sparse and dense Jacobian evaluation (respectively) with OpenMP\slash shallow-vectorized OpenCL.
+Finally, the performance of dense analytical-Jacobian evaluation in this new version of \texttt{pyJac} was compared to the previous version~\cite{pyjac16} on the \sse/\slash\gpuold/ CPU\slash GPUs; OpenMP evaluation on the \sse/ CPU was \SI{1.79}{$\times$} slower for the smallest chemical model, but \SIrange{3.37}{10.19}{$\times$} faster for the larger models, while the shallow-vectorized OpenCL code was \SIrange{1.37}{19.56}{$\times$} faster than the previous version over all chemical models.
+Similarly, evaluation on the \gpuold/ GPU was \SI{1.89}{$\times$} slower for the smallest model but \SIrange{1.25}{2.84}{$\times$} faster for the larger models.
 
@@
 \pagebreak
+
+%% The Appendices part starts with the command \appendix;
+%% appendix sections are then done as normal sections
+\appendix
+\setcounter{figure}{0}
+\setcounter{table}{0}
+
+% Fix for missing space between "Appendix" and letter
+\renewcommand*{\thesection}{\appendixname~\Alph{section}}
+
+\section{Jacobian error statistics per test-platform}
+\label{A:per_platform}
+
+This section gives more detail on the results presented in~\cref{S:jac_valid}, breaking down the reported error statistics per test-platform\slash language.
+The error of the Intel OpenCL runtime is presented in~\cref{T:intel_error}, the Portable OpenCL (POCL) runtime in~\cref{T:pocl_error}, OpenMP in~\cref{T:omp_error} and the NVIDIA OpenCL runtime in~\cref{T:nv_error}.
+POCL and OpenMP tend to have the smallest error norms, while NVIDIA tends to have the largest; in particular the stringent filtered error norm $E_{\mathcal{C} = 10^{20}}$ is two orders of magnitude larger for the NVIDIA runtime than the other test platforms with the \ce{H2}\slash\ce{CO} and USC-Mech II models.
+
+\begin{table}[htbp]
+\sisetup{retain-zero-exponent=true}
+\centering
+\begin{tabular}{@{}l S[table-format=.0] S[table-format=1.3e1] S[table-format=1.3e1] S[table-format=1.3e1] S[table-format=1.3e1] @{}}
+\toprule
+Model                 & \multicolumn{1}{c}{$E_{\mathcal{L}}$} & \multicolumn{1}{c}{$E_{\mathcal{C} = 10^{20}}$}   & \multicolumn{1}{c}{$E_{\mathcal{C} = 10^{15}}$} \\
+\midrule
+\ce{H2}\slash \ce{CO} & \num{1.455e-14}      & \num{8.084e-01}  & \num{1.907e-05} \\
+GRI-Mech 3.0          & \num{1.567e-14}      & \num{1.469e-07}  & \num{1.316e-07} \\
+USC-Mech II           & \num{9.632e-15}      & \num{5.567e-03}  & \num{1.704e-07} \\
+\ce{iC5H11OH}         & \num{1.227e-10}      & \num{1.363e-03}  & \num{2.864e-05} \\
+\bottomrule
+\end{tabular}
+\caption{Summary of Jacobian matrix validation results for the Intel OpenCL runtime.
+The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/ and sparse\slash dense Jacobians.
+The threshold for the filtered relative error is the same as reported in~\cref{S:jac_valid}.
+}
+\label{T:intel_error}
+\end{table}
+
+\begin{table}[htbp]
+\sisetup{retain-zero-exponent=true}
+\centering
+\begin{tabular}{@{}l S[table-format=.0] S[table-format=1.3e1] S[table-format=1.3e1] S[table-format=1.3e1] S[table-format=1.3e1] @{}}
+\toprule
+Model                 & \multicolumn{1}{c}{$E_{\mathcal{L}}$} & \multicolumn{1}{c}{$E_{\mathcal{C} = 10^{20}}$}   & \multicolumn{1}{c}{$E_{\mathcal{C} = 10^{15}}$} \\
+\midrule
+\ce{H2}\slash \ce{CO} & \num{1.456e-14}      & \num{1.230e-01}  & \num{3.951e-06} \\
+GRI-Mech 3.0          & \num{1.014e-14}      & \num{1.890e-07}  & \num{1.877e-07} \\
+USC-Mech II           & \num{9.632e-15}      & \num{8.998e-04}  & \num{1.201e-08} \\
+\ce{iC5H11OH}         & \num{9.133e-15}      & \num{1.723e-05}  & \num{5.108e-07} \\
+\bottomrule
+\end{tabular}
+\caption{Summary of Jacobian matrix validation results for the Portable OpenCL (POCL) runtime.
+The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/ and sparse\slash dense Jacobians.
+The threshold for the filtered relative error is the same as reported in~\cref{S:jac_valid}.
+}
+\label{T:pocl_error}
+\end{table}
+
+\begin{table}[htbp]
+\sisetup{retain-zero-exponent=true}
+\centering
+\begin{tabular}{@{}l S[table-format=.0] S[table-format=1.3e1] S[table-format=1.3e1] S[table-format=1.3e1] S[table-format=1.3e1] @{}}
+\toprule
+Model                 & \multicolumn{1}{c}{$E_{\mathcal{L}}$} & \multicolumn{1}{c}{$E_{\mathcal{C} = 10^{20}}$}   & \multicolumn{1}{c}{$E_{\mathcal{C} = 10^{15}}$} \\
+\midrule
+\ce{H2}\slash \ce{CO} & \num{5.962e-15}      & \num{3.614e-02}  & \num{1.657e-06} \\
+GRI-Mech 3.0          & \num{1.297e-15}      & \num{1.321e-07}  & \num{1.316e-07} \\
+USC-Mech II           & \num{9.630e-15}      & \num{4.185e-04}  & \num{6.746e-09} \\
+\ce{iC5H11OH}         & \num{6.131e-15}      & \num{1.721e-05}  & \num{5.108e-07} \\
+\bottomrule
+\end{tabular}
+\caption{Summary of Jacobian matrix validation results for OpenMP execution.
+The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/ and sparse\slash dense Jacobians.
+The threshold for the filtered relative error is the same as reported in~\cref{S:jac_valid}.
+}
+\label{T:omp_error}
+\end{table}
+
+\begin{table}[htbp]
+\sisetup{retain-zero-exponent=true}
+\centering
+\begin{tabular}{@{}l S[table-format=.0] S[table-format=1.3e1] S[table-format=1.3e1] S[table-format=1.3e1] S[table-format=1.3e1] @{}}
+\toprule
+Model                 & \multicolumn{1}{c}{$E_{\mathcal{L}}$} & \multicolumn{1}{c}{$E_{\mathcal{C} = 10^{20}}$}   & \multicolumn{1}{c}{$E_{\mathcal{C} = 10^{15}}$} \\
+\midrule
+\ce{H2}\slash \ce{CO} & \num{1.862e-14}      & \num{1.741e+00}  & \num{4.508e-05} \\
+GRI-Mech 3.0          & \num{1.489e-14}      & \num{3.842e-07}  & \num{3.687e-07} \\
+USC-Mech II           & \num{1.174e-14}      & \num{1.119e-02}  & \num{1.983e-07} \\
+\ce{iC5H11OH}         & \num{8.602e-15}      & \num{1.748e-05}  & \num{5.109e-07} \\
+\bottomrule
+\end{tabular}
+\caption{Summary of Jacobian matrix validation results for NVIDIA OpenCL execution.
+The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/ and sparse\slash dense Jacobians.
+The threshold for the filtered relative error is the same as reported in~\cref{S:jac_valid}.
+}
+\label{T:nv_error}
+\end{table}
+
+\pagebreak
 \bibliography{paper}

### 2018-03-13 scripts/load_error_files.py

@@ -66,8 +66,8 @@ def run_error_calcs(ftype, updater, per_platform=False):
         with open(mech_info, 'r') as file:
             mech_info = yaml.load(file.read())
             if mech_info['n_cheb']:
-                add_rxn_specific_inds(mech_info, ct.ChebyshevReaction, 'cheb_inds',
-                                      mech)
+                add_rxn_specific_inds(mech_info, ct.ChebyshevReaction,
+                                      'cheb_inds', mech)
             if mech_info['n_plog']:
                 add_rxn_specific_inds(mech_info, ct.PlogReaction, 'plog_inds',
                                       mech)
@@ -78,8 +78,9 @@ def run_error_calcs(ftype, updater, per_platform=False):
                 platforms.add(platform)
                 if platform not in err_dicts[mech_name]:
                     err_dicts[mech_name][platform] = defaultdict(lambda: 0)
-                updater(err_dicts[mech_name][platform], np.load(file), filename=file,
-                    mech_info=mech_info)
+                updater(err_dicts[mech_name][platform], np.load(file),
+                        filename=file,
+                        mech_info=mech_info)
                 continue
 
             updater(err_dicts[mech_name], np.load(file), filename=file,

### 2018-03-21 scripts/plot_styles.py
plot fixes

@@ -95,7 +95,7 @@ def pretty_names(pname, short=False):
                                                  'haswell': r'\texttt{avx2}',
                                                  'gpu': r'\texttt{K40m}',
                                                  'v1': r'\texttt{sse4.2-v1}',
-                                                 'v1-gpu': r'\texttt{C2075.2-v1}'}),
+                                                 'v1-gpu': r'\texttt{C2075-v1}'}),
                   'rtype': dummy_formatter({'jac': 'Analytical',
                                             'fdjac': 'Finite Difference'})
                   }

### 2018-03-21 paper.tex
first pass at dr. sung edits

@@
 \begin{abstract} % not to exceed 200 words
-A code generation platform, \texttt{pyJac}, for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized constant-pressure\slash constant-volume thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation was studied for a selection of chemical kinetic models covering the range of sizes typically used in reactive-flow simulations.
+A code generation platform, \texttt{pyJac}, for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized constant-pressure (\conp/)\slash constant-volume (\conv/) thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation has been developed.
+Selected chemical kinetic models covering the range of sizes typically used in reactive-flow simulations were used for demonstration.
 A new formulation of the chemical kinetic governing equations was derived and validated, resulting Jacobian sparsities of \SIrange{28.6}{92.0}{$\percent$} for the tested models.
-Speedups of \SIrange{3.40}{4.08}{$\times$} were found for shallow-vectorized OpenCL source-rate evaluation compared a parallel OpenMP code on an \avx/ central processing unit (CPU), increasing to \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.03}{4.23}{$\times$} for sparse and dense chemical kinetic Jacobian evaluation; in addition, the performance of a NVIDIA \gpunew/ graphics processing unit (GPU) was compared to a NVIDIA \gpuold/ GPU finding speedups of \SIrange{1.40}{1.88}{$\times$}, \SIrange{1.10}{1.59}{$\times$}, and \SIrange{1.36}{3.0}{$\times$} for source-rate, sparse and dense Jacobian evaluation, respectively.
-Additionally, the effect of data-ordering was investigated and an storage pattern specifically formulated for vectorized evaluation proposed; as well, the effect of the \conp/\slash\conv/ assumptions and varying vector widths were studied on source-term evaluation performance.
-Speedups over a first-order finite-difference approach reached \SIrange{17.60}{45.13}{$\times$} and \SIrange{55.11}{245.63}{$\times$} for sparse\slash dense evaluation on the CPU\slash GPU respectively, while dense Jacobian evaluation was up to \SI{19.56}{$\times$} and \SI{2.84}{$\times$} times faster than a previous version of \texttt{pyJac} on a \sse/ CPU and \gpuold/ GPU respectively.
+Speedups of \SIrange{3.40}{4.08}{$\times$} were found for shallow-vectorized OpenCL source-rate evaluation compared a parallel OpenMP code on an \avx/ central processing unit (CPU), increasing to \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.03}{4.23}{$\times$} for sparse and dense chemical kinetic Jacobian evaluation, respectively.
+In addition, the performance of a NVIDIA \gpunew/ graphics processing unit (GPU) was compared to a NVIDIA \gpuold/ GPU finding speedups of \SIrange{1.40}{1.88}{$\times$}, \SIrange{1.10}{1.59}{$\times$}, and \SIrange{1.36}{3.0}{$\times$} for source-rate, sparse and dense Jacobian evaluation, respectively.
+Furthermore, the effect of data-ordering was investigated and an storage pattern specifically formulated for vectorized evaluation was proposed; as well, the effect of the \conp/\slash\conv/ assumptions and varying vector widths were studied on source-term evaluation performance.
+Speedups over a first-order finite-difference approach reached \SIrange{17.60}{45.13}{$\times$} and \SIrange{55.11}{245.63}{$\times$} for sparse\slash dense evaluation on the CPU\slash GPU respectively, while dense Jacobian evaluation was up to \SI{19.56}{$\times$} and \SI{2.84}{$\times$} times faster than a previous version of \texttt{pyJac} on a \sse/ CPU and \gpuold/ GPU, respectively.
 Finally, future directions for vectorized chemical kinetic evaluation and sparse linear-algebra techniques were discussed.
@@
 
-As the importance of detailed chemical kinetics for predictive reactive-flow simulations has become recognized~\cite{LU2009192}, the size and complexity of chemical kinetic models describing current and next-generation fuels relevant to transportation and power generation have increased greatly; e.g., a recent biodiesel model~\cite{WESTBROOK2011742} consists of \textasciitilde\num{3500} chemical species, and over \num{17000} reactions.
+As the importance of detailed chemical kinetics for predictive reactive-flow simulations has become recognized~\cite{LU2009192}, the size and complexity of chemical kinetic models describing current and next-generation fuels relevant to transportation and power generation have increased greatly; e.g., a biodiesel model~\cite{WESTBROOK2011742} consists of \textasciitilde\num{3500} chemical species, and over \num{17000} reactions.
 Moreover, the cost of evaluating the chemical source-terms scales linearly with the size of the model, while evaluation and factorization of the chemical kinetic Jacobian---if naively implemented via finite-difference method---scale quadratically and cubically with the number of species in the model~\cite{LU2009192}.
-These factors often make the use of detailed chemical kinetics prohibitive in practice; e.g., in a recent direct numerical simulation~\cite{Spafford:2010aa} utilizing a 22-species model, evaluation of reaction rates consumed around half of the total run time.
+These factors often make the use of detailed chemical kinetics prohibitive in practice; e.g., in a direct numerical simulation~\cite{Spafford:2010aa} utilizing a 22-species model, evaluation of reaction rates consumed around half of the total run time.
 In addition, chemical kinetic Jacobian evaluation and factorization---utilized by most common implicit integration techniques to deal with numerical stiffness---is a bottleneck for even modestly-sized chemical models in realistic reactive flow simulations, necessitating use of other cost-reduction strategies~\cite{LU2009192}.
 
-A host of techniques have been developed to lessen the computational demand of chemical kinetic calculations while maintaining fidelity, these fall broadly into a few categories: removal of unimportant species and reactions unimportant species and reactions~\cite{Lu:2006bb,Pepiot-Desjardins:2008,Hiremath:2010jw,Niemeyer:2010bt,Curtis:2015}, lumping of species with similar thermo-chemical properties~\cite{Lu:2007,Ahmed:2007fa,Pepiot:2008kq}, and time-scale methods that reduce numerical stiffness~\cite{Maas:1992ws,Lam:1994ws,Lu:2001ve,Gou:2010}.
+A host of techniques have been developed to lessen the computational demand of chemical kinetic calculations while maintaining fidelity, these fall broadly into a few categories: removal of unimportant species and reactions~\cite{Lu:2006bb,Pepiot-Desjardins:2008,Hiremath:2010jw,Niemeyer:2010bt,Curtis:2015}, lumping of species with similar thermo-chemical properties~\cite{Lu:2007,Ahmed:2007fa,Pepiot:2008kq}, and time-scale methods that reduce numerical stiffness~\cite{Maas:1992ws,Lam:1994ws,Lu:2001ve,Gou:2010}.
 The interested reader is referred to the recent review of Tur\'anyi and Tomlin~\cite{turanyi2016analysis} for a comprehensive overview.
@@
 Traditional multi-core parallelism is often used to increase central processing unit (CPU) performance, however as exponential growth in processing power---colloquially known as Moore's law---has slowed~\cite{khan2018science}, SIMD processors as well as SIMT processors, e.g., in the form of graphics processing units (GPUs), have gained recognition due to their increased floating operation throughput.
-The parallel programming standard OpenCL~\cite{stone2010opencl} has further enabled adoption of vector-processing based codes in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems, e.g., the CPU, GPU, or Intel's Many Integrated Core architecture (MIC), etc.
-In this discussion, we will largely use OpenCL terminology to describe these processing paradigms, as it provides a convenient way to classify otherwise disparate processor type (e.g., CPUs and GPUs), however the concepts discussed are broadly applicable to SIMD\slash SIMT processing and not tied to OpenCL specifically.
+The parallel programming standard OpenCL~\cite{stone2010opencl} has further enabled adoption of vector-processing based codes in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems, e.g., CPU, GPU, or Intel's Many Integrated Core (MIC) architecture, etc.
+In this discussion, we will largely use OpenCL terminology to describe these processing paradigms, as it provides a convenient way to classify otherwise disparate processor type (e.g., CPUs and GPUs), however the concepts discussed herein are broadly applicable to SIMD\slash SIMT processing and not tied to OpenCL specifically.
 
@@
       \caption{Schematic of SIMD processing.
-A single compute unit (e.g., a CPU core) contains a vector unit with $N_v$ processing elements, together called a vector-lane.
+A single compute unit (e.g., a CPU core) contains a vector unit with $N_v$ processing elements (PEs), together called a vector-lane.
 The vector unit executes a single instruction concurrently on multiple data.}
@@
       \caption{Schematic of SIMT processing.
-A single compute unit (e.g., a GPU streaming multiprocessor) contains many processing elements and hosts many threads, each with an instruction to execute (I1, I2).
+A single compute unit (e.g., a GPU streaming multiprocessor) contains many processing elements (PEs) and hosts many threads, each with an instruction to execute (I1, I2).
 Threads with the same instruction execute concurrently on multiple data while the others must wait (leading to thread divergence).}
@@
   \end{subfigure}
+  \caption{Simple diagrams explaining the fundamentals of the SIMD and SIMT vector-processing paradigms.}
 \end{figure}
@@
 A typical modern CPU has many compute units (i.e., cores), each with specialized vector processing units capable of running SIMD instructions (\cref{F:SIMD}).
-A SIMD instruction utilizes the vector processor to execute the same floating point operation (e.g.
-multiplication, division, etc.) on a several different data concurrently; the number of concurrent operations possible is known as the vector-width\footnote{OpenCL allows for use of vector-widths different from the actual hardware vector-width via implicit conversion, and may provide some performance benefit as will be studied in Sec.~\ref{S:results}} and is typically around 2--4 double precision operations.
-Additionally, specialized hardware accelerators, e.g.
-Intel's Xeon Phi co-processor (MIC) have been developed that have tens of cores with very wide vector-widths (e.g.
-4--8 double precision operations); these very wide vector-widths are also found on cutting-edge and forthcoming Intel CPUs; i.e., the Skylake Xeon and Cannon Lake architectures, respectively.
+A SIMD instruction utilizes the vector processor to execute the same floating point operation (e.g., multiplication, division, etc.) on a several different data concurrently; the number of concurrent operations possible is known as the vector-width\footnote{OpenCL allows for use of vector-widths different from the actual hardware vector-width via implicit conversion, and may provide some performance benefit as will be studied in Sec.~\ref{S:results}} and is typically around 2--4 double precision operations.
+Additionally, specialized hardware accelerators, e.g., Intel's Xeon Phi co-processor (MIC architecture) have been developed that have tens of cores with very wide vector-widths (e.g., \numrange{4}{8} double precision operations); these very wide vector-widths are also found on cutting-edge and forthcoming Intel CPUs; i.e., the Skylake Xeon and Cannon Lake architectures, respectively.
 
@@
 If some threads must execute a different instruction---e.g., due to if\slash then branching or predication---they are forced to wait and execute later.
-This phenomena, known as thread-divergence, is a key consideration for SIMT-processing and can cause serious performance degradation for complicated algorithms~\cite{CurtisGPU:2017}.
+This phenomenon, known as thread-divergence, is a key consideration for SIMT-processing and can cause serious performance degradation for complicated algorithms~\cite{CurtisGPU:2017}.
 
@@
 Youssefi~\cite{Youssefi:2011tm} explored the importance of analytical Jacobian matrices for time-scale analysis techniques as well as their effect on computational efficiency in zero-dimensional homogeneous reactor simulations.
-Bisetti~\cite{Bisetti:2012jw} developed an isothermal, isobaric analytical Jacobian code-generation utility; although incompatible with many combustion codes, a significant increase in Jacobian sparsity was exhibited.
+Bisetti~\cite{Bisetti:2012jw} developed an isothermal, isobaric analytical Jacobian code-generation utility; although this set of assumptions (particularly, the choice of an isothermal system) are incompatible with many combustion simulations, a significant increase in Jacobian sparsity was exhibited.
 Additionally, in the same work Bisetti provided a novel way to compute dense matrix-vector multiplications resulting from a change of system variables without need for storage of the full dense Jacobian.
-Perini et al.~\cite{Perini:2012gy} developed an analytical Jacobian code for constant-volume combustion, with additional options to increase sparsity (at expect of strict correctness), and enable tabulation of temperature dependent properties; a \SI{80}{\percent} speedup over a finite-difference-based Jacobian was reported when used in a multidimensional reactive-flow simulation.
-Dijkmans et al.~\cite{Dijkmans:2014bb} developed a GPU-based analytical Jacobian code with optional tabulation of temperature-dependent properties, and found speedups up to \SI{120}{$\times$} for zero-dimensional chemical kinetic integration with large chemical models (\textasciitilde\num{3000} species) with a Tesla C2075 GPU.
+Perini et al.~\cite{Perini:2012gy} developed an analytical Jacobian code for constant-volume combustion, with additional options to increase sparsity (at expense of strict correctness), and enable tabulation of temperature dependent properties; a \SI{80}{\percent} speedup over a finite-difference-based Jacobian was reported when used in a multidimensional reactive-flow simulation.
+Dijkmans et al.~\cite{Dijkmans:2014bb} developed a GPU-based analytical Jacobian code with optional tabulation of temperature-dependent properties, and found speedups up to \SI{120}{$\times$} for zero-dimensional chemical kinetic integration with large chemical models (\textasciitilde\num{3000} species) using a Tesla C2075 GPU.
 Niemeyer et al.~\cite{Niemeyer:2016aa} created and validated the open-source analytical chemical kinetic Jacobian code-generator, \texttt{pyJac}, supporting parallel execution on the CPU and SIMT execution on the GPU; a speedup of \SIrange{3}{7.5}{$\times$} over a finite-difference Jacobian was found on the CPU.
-Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but it was not validated outside the context of use with an implicit-integration technique; in addition the Jacobian was based on an over-constrained system, the effect on strict conservation of mass\slash energy was not studied.
+Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but it was not validated outside the context of use with an implicit-integration technique; in addition, since the Jacobian was based on an over-constrained system, the effect on strict conservation of mass\slash energy was not studied.
 
@@
 Stone and Davis~\cite{Stone:2013aa} implemented the implicit VODE~\cite{Brown:1989vl} integrator on a Fermi M2050 GPU, achieving an order of magnitude speedup over the baseline CPU version on a single core of an AMD Opteron 6134 Magny-Cours.
-Niemeyer and Sung~\cite{Niemeyer:2014aa} later developed a stabilized explicit second-order Runge--Kutta--Chebyshev algorithm on a Tesla C2075, demonstrating an order of magnitude speedup for a over a CPU implementation of VODE on a six-core Intel X5650 for moderately stiff chemical kinetics.
+Niemeyer and Sung~\cite{Niemeyer:2014aa} later developed a stabilized explicit second-order Runge--Kutta--Chebyshev algorithm on a Tesla C2075, demonstrating an order of magnitude speedup for a CPU implementation of VODE on a six-core Intel X5650 for moderately stiff chemical kinetics.
 Sewerin and Rigopoulos~\cite{Sewerin20151375} implemented a three-stage\slash fifth-order implicit Runge--Kutta method~\cite{wanner1991solving} on both a high-end (NVIDIA Quadro 6000) and consumer-grade (NVIDIA Quadro 600) GPUs, as well as a standard CPU (two-core, four-thread Intel i5-520M) and a scientific workstation (eight-core, 16-thread Intel Xeon E5-2687W); the high-end GPU was at best \SI{1.8}{$\times$} slower than the workstation CPU (16 threads), while the consumer level GPU was at best \SI{5.5}{$\times$} slower than the corresponding standard CPU (four threads).
 Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs) on an Intel Xeon E5-2680 CPU and a Tesla K20 GPU.
-In evaluation of the thermo-chemical properties, speedups between \SIrange{8}{13}{$\times$} were found (over a serial version of the same code) while running on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU respectively.
-Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two-fourth order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on a Tesla C2075; the implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on 12--38 Intel Xeon E5-4640 v2 CPU cores for two smaller chemical models with an integration time step of \SI{e-6}{$\sec$}.
+In evaluation of the thermo-chemical properties, speedups between \SIrange{8}{13}{$\times$} were found (over a serial version of the same code) while running on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU, respectively.
+Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on a Tesla C2075; the implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on 12--38 Intel Xeon E5-4640 v2 CPU cores for two relatively small chemical models with an integration time step of \SI{e-6}{$\sec$}.
 
@@
 Linford et al.~\cite{Linford:2011} implemented a three-stage, second-order Rosenbrock integrator for atmospheric chemical kinetics on the CPU, GPU and cell croadband engine (CBE)---a specially designed vector processor---and found speedups regularly exceeding~\SI{25}{$\times$} over a serial CPU implementation.
-Kroshko and Spiteri~\cite{kroshko2013efficient} implemented a SIMD-vectorized third-order stiff Rosenbrock integrator for atmospheric chemistry on the CBE and found a speedup of \SI{1.89}{$\times$} (a parallel scaling efficiency of \SI{0.94}{$\percent$}) over a serial version of the same code.
-Stone et al.,~\cite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
-SIMD-vectorization improved the integrator performance over an OpenMP baseline (which was vectorized by simple compiler hints (i.e., \texttt{\#pragmas}) by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance was only \SIrange{1.4}{1.6}{$\times$} faster than the OpenMP baseline due to thread-divergence concerns.
+Kroshko and Spiteri~\cite{kroshko2013efficient} implemented a SIMD-vectorized third-order stiff Rosenbrock integrator for atmospheric chemistry on the CBE and found a speedup of \SI{1.89}{$\times$} (a parallel scaling efficiency of \SI{94}{$\percent$}) over a serial version of the same code.
+Stone and Niemeyer~\cite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
+SIMD-vectorization improved the integrator performance over an OpenMP baseline (which was vectorized by simple compiler hints (i.e., \texttt{\#pragmas}) by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance was only \SIrange{1.4}{1.6}{$\times$} faster than the OpenMP baseline due to thread-divergence concerns~\cite{stone2016}.
 
@@
  \item Investigate the performance for a wide range chemical kinetic models, specifically looking at speedups due to SIMD\slash SIMT-vectorization as well as relative to the previous version of \texttt{pyJac}, and finally
- \item Discuss future extensions to this work as well as the most promising directions for SIMD\slash SIMT vectorization in reactive-flow simulations
+ \item Discuss future extensions to this work as well as several promising directions for SIMD\slash SIMT vectorization in reactive-flow simulations.
 \end{itemize}
@@
     \\
+    \\
+    \\
+    \\
     \begin{subfigure}[t]{\textwidth}
@@
 When storing arrays for a chemical kinetic model, the data-storage layout and vectorization patterns are critical to achieving high-performance code.
-\Cref{F:mem} depicts an example data array with $K$-rows and $N$-columns, and index ($i$, $j$) corresponding to the $i$th row and $j$th column.
-For example the concentration of species $j$ for the $i$th thermo-chemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{state}$, the number of thermo-chemical states considered for evaluation, and $1 \le j \le \ns/$, the number of species in the model.
+\Cref{F:mem} depicts an example data array with $K$ rows and $N$ columns, and index ($i$, $j$) corresponding to the $i$th row and $j$th column.
+For example, the concentration of species $j$ for the $i$th thermo-chemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{state}$ (the number of thermo-chemical states considered for evaluation), and $1 \le j \le \ns/$ (the number of species in the model).
 The stored concentrations would then have $K = N_{\text{state}}$ rows, and $N = N_{\text{sp}}$ columns.
 
-In the ``C'' (C row-major) format, the concentrations of all species for a single thermo-chemical condition $i$ would be stored sequentially in memory, i.e., with  $[C]_{1, 1}$ in index \num{1}, $[C]_{1, 2}$ in index \num{2} etc., as shown in \cref{F:row_major}.
+In the ``C'' (C row-major) format, the concentrations of all species for a single thermo-chemical condition $i$ would be stored sequentially in memory, i.e., with  $[C]_{1, 1}$ in index \num{1}, $[C]_{1, 2}$ in index \num{2}, etc., as shown in \cref{F:row_major}.
 Conversely in the ``F'' (Fortran column-major) format, the concentrations of a single species $j$ over all thermo-chemical states are adjacent in memory, corresponding to storing $[C]_{1, 1}$ in index \num{1}, $[C]_{2, 1}$ in index \num{2} and so on, as shown in \cref{F:column_major}.
@@
 
-In a \textit{deep}-SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works), a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermo-chemical source-terms for a single thermo-chemical state; thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + N_v}$---the species concentrations for state 1, species $j \ldots j + N_v$---will access sequential memory locations if the data is stored in ``C''-order.
+In a \textit{deep}-SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works (e.g.,~\cite{Stone:2013aa,CurtisGPU:2017}), a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermo-chemical source-terms for a single thermo-chemical state; thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + N_v}$---the species concentrations for state 1, species $j \ldots j + N_v$---will access sequential memory locations if the data is stored in ``C''-order.
 Further, in ``C''-ordering the maximum distance between any two species concentrations within the same thermo-chemical state is $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU, but may lead to uncoalesced memory-accesses on the GPU.
@@
     \end{subfigure}
+    \\
+    \\
+    \\
     \\
@@
 This new version of \texttt{pyJac} is capable of evaluating the thermo-chemical source-terms for using both the constant-pressure (\conp/) or constant-volume (\conv/) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
-In this section, we will only outline a brief summary of the system evaluated by \texttt{pyJac}; the interested reader is referred to the complete derivations, attached as supplemental material to this document.
-
-The thermo-chemical state vector derivation consists of the temperature, a non-constant thermodynamic state parameter (volume or pressure for \conp/ and \conv/ respectively), and the number of moles of all species except the last species in the chemical model which is typically taken to be the bath-gas (e.g., \ce{N2}):
+In this section, we will only outline a brief summary of the system evaluated by \texttt{pyJac}; the interested reader is referred to the supplemental material for the complete derivations.
+
+The thermo-chemical state vector derivation consists of the temperature, a non-constant thermodynamic state parameter (volume or pressure for \conp/ and \conv/, respectively), and the number of moles of all species except the last species in the chemical model which is typically taken to be the bath-gas (e.g., \ce{N2}):
 \loadeq{state}
-where $T$ is the temperature, $V$ and $P$ the volume and pressure respectively, and $n_j$ the number of moles of the $j$th species in the model (containing $\ns/$ total species).
+where $T$ is the temperature, $V$ and $P$ the volume and pressure, respectively, and $n_j$ the number of moles of the $j$th species in the model (containing $\ns/$ total species).
 
@@
 First, the state vector results in highly sparse chemical kinetic Jacobians as will be detailed in~\cref{S:sparsity}.
-Second, mass is explicitly conserved in this formulation owing to the fact that the moles of the final species is calculated from the ideal gas law, and the rate of change of the final species from conservation of mass (see \cref{deriv-s:state,deriv-s:source} of the supplemental material respectively); additionally, system is not over-constrained and does not require use of a more complicated DAE solver (as compared to an ODE integrator) for integration.
+Second, mass is explicitly conserved in this formulation owing to the fact that the moles of the final species is calculated from the ideal gas law, and the rate of change of the final species from conservation of mass (see \cref{deriv-s:state,deriv-s:source} of the supplemental material, respectively); additionally, system is not over-constrained and does not require use of a more complicated differential algebraic equation solver (as compared to an ODE integrator) for integration.
 Finally, the chemical kinetic Jacobian changes relatively little between the \conp/ and \conv/ forms, making maintaining the code-base much simpler.
-Although most current combustion codes do not use species moles as a state variable, conversion to\slash from more common mass\slash mole-fractions and moles is straightforward, and once inside integration of an chemical kinetic initial value problem (IVP), the choice of variables no longer matter.
+Although most current combustion codes do not use species moles as a state variable, conversion to\slash from more common mass\slash mole-fractions and moles is straightforward, and once inside integration of an chemical kinetic initial value problem (IVP), the choice of variables no longer matters.
 
@@
 \loadeq{dmolar}
-where $\dot{\omega}_k$ is the $kth$ species' overall production rate:
+where $\dot{\omega}_k$ is the $kth$ species' overall molar production rate:
 \begin{equation}
@@
 \end{equation}
-with $\nu_{k, i}$ the net stoichiometic coefficient of species $k$ in reaction $i$, $R_{i}$ the net rate of progress of reaction $i$ (described in~\cref{deriv-s:rop_net} of the supplemental), and $c_{i}$ the pressure modification term i.e., for third-body (see~\cref{deriv-s:thdbody} of the supplemental), or fall-off\slash chemically-activated reactions (\cref{deriv-s:falloff} of the same); \texttt{pyJac} is capable of evaluating all modern reaction types, e.g., P-Log and Chebyshev reactions.
+with $\nu_{k, i}$ the net stoichiometic coefficient of species $k$ in reaction $i$, $R_{i}$ the net rate of progress of reaction $i$ (described in~\cref{deriv-s:rop_net} of the supplemental material), and $c_{i}$ the pressure-dependent modification term, i.e., for third-body (see~\cref{deriv-s:thdbody} of the supplemental material), or falloff\slash chemically-activated reactions (\cref{deriv-s:falloff} of the same); \texttt{pyJac} is capable of evaluating all modern reaction types, e.g., P-Log and Chebyshev reactions.
+In addition, $N_{reac}$ is the total number of reactions in the chemical kinetic model.
 
@@
 \loadeq{dtemp}
-where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure and constant-volume specific heat of species $k$ in molar units respectively, while $[C]_{k}$ is the concentration, given by:
+where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure specific heat and constant-volume specific heat of species $k$ in molar units respectively, while $[C]_{k}$ is the concentration, given by:
 \begin{equation}
@@
 \end{equation}
-where $\mathcal{R}$ is the ideal gas-constant in molar units, the volume and pressure source-terms can be found:
+where $\mathcal{R}$ is the ideal gas-constant in molar units, the volume and pressure source-terms can be found (where $W_k$ and $W_{\ns/}$ are the molecular weights of species $k$ and $\ns/$, respectively):
 \loadeq{dparam}
@@
 \loadeq{dTdotdT}
-and with respect to the thermodynamic state-parameter source-term (the volume \slash pressure for \conp/\slash\conv/ respectively):
+with respect to the thermodynamic state-parameter source-term (the volume \slash pressure for \conp/\slash\conv/, respectively) it's derivative:
 \loadeq{dTdotdE}
-and the moles of species $j$:
+The derivative for the moles of species $j$:
 \loadeq{dTdotdnj}
@@
 \loadeq{dEdotdE}
-and finally, with respect to the moles of species $j$:
+where $[C]$ is the total molar concentration:
+\begin{equation}
+ [C] = \frac{P}{\mathcal{R} T}\;.
+\end{equation}
+Finally, the derivative with respect to the moles of species $j$:
 \loadeq{dEdotdnj}
@@
 Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates pseudo-code and data to OpenCL\slash C-code.
-As the name implies, \texttt{loo.py} generates code based on for-loops, this differs from the previous version of \texttt{pyJac}~\cite{pyjac16} which generates static code (i.e., fully unrolled loops, with thermodynamic\slash reaction parameters written directly in code rather than stored in arrays).
+As the name implies, \texttt{loo.py} generates code based on for-loops; this differs from the previous version of \texttt{pyJac}~\cite{pyjac16} which generates static code (i.e., fully unrolled loops, with thermodynamic\slash reaction parameters written directly in code rather than stored in arrays).
 In our previous work~\cite{Niemeyer:2016aa}, this static code generation caused some issues with large file sizes, long compilation times and even occasionally breaking the \texttt{gcc} and \texttt{nvcc} compilers (the latter issue necessitated breaking the Jacobian\slash source-term evaluations into separate files).
@@
 \caption{The Intel CPU configurations used in this study.
-	 The vector-width reported in this table is reported in (ideal) number of double operations per SIMD instruction, as this will be used in measuring SIMD-efficiency; for reference, the vector-width in bits of the \sse/ and \avx/ machines are \SI{128}{\bit} and \SI{256}{\bit} respectively.
+	 The vector-width reported in this table is reported in (ideal) number of double operations per SIMD instruction, as this will be used in measuring SIMD-efficiency; for reference, the vector-width in bits of the \sse/ and \avx/ machines are \SI{128}{\bit} and \SI{256}{\bit}, respectively.
 	 The identifier field will be used as a shorthand descriptor in the performance plots to quickly identify the CPU type.}
@@
 \end{table}
-\footnotetext{Due to an driver implementation issue~\cite{nvidia_memory}, the total memory used on the \gpuold/ and \gpunew/ GPUs were to limited to \SI{4}{\giga\byte} and \SI{10}{\giga\byte} respectively.\addtocounter{footnote}{1}}
-
-\Cref{t:platforms} lists the platforms and vectorization\slash execution patterns they are capable of running.
+\footnotetext{Due to a driver implementation issue~\cite{nvidia_memory}, the total memory used on the \gpuold/ and \gpunew/ GPUs were to limited to \SI{4}{\giga\byte} and \SI{10}{\giga\byte}, respectively.\addtocounter{footnote}{1}}
+
+\Cref{t:platforms} lists the platforms and vectorization\slash execution patterns that they are capable of running.
 The Intel and NVIDIA OpenCL runtimes lack implementations of atomic operations on double precision variables; these are currently necessary for \texttt{pyJac} to run deep-vectorized code.
@@
 \end{tabular}
-\caption{The various platforms used in this study and the execution \slash vectorization patterns they are capable of running}
+\caption{The various platforms used in this study and the execution \slash vectorization patterns that they are capable of running}
 \label{t:platforms}
@@
 
-Finally, \cref{t:models} displays the chemical kinetic models used in this work, as well as number of partially stirred reaction conditions (PaSR) in the condition database for each.
-Creation of the PaSR databases are described in detail our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
+Finally, \cref{t:models} displays the chemical kinetic models used in this work, as well as number of partially stirred reaction conditions (PaSR) used in the condition database for each.
+Creation of the PaSR databases is described in detail in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
 
@@
 \label{s:validation}
-The reaction rates of progress (ROP), species and temperature rates in this study are validated by comparison with Cantera~\cite{Cantera}, however special care must be taken due floating point arithmetic issues.
+The reaction rates of progress (ROP), species and temperature rates in this study are validated by comparison with Cantera~\cite{Cantera}. 
+However special care must be taken due floating point arithmetic issues.
 
@@
 
-However, in computing the net ROP of reaction $i$ for state $j$ from the forward and reverse ROP: $R_{ij} = R_{ij}^{\prime} - R_{ij}^{\prime\prime}$, accuracy can easily be lost as the net ROP may be---particularly near chemical equilibrium---many orders of magnitude smaller than the forward or reverse rates.
+However, in computing the net ROP of reaction $i$ for state $j$ from the forward and reverse ROP: $R_{ij} = R_{ij}^{\prime} - R_{ij}^{\prime\prime}$, accuracy can easily be lost as the net ROP may be---particularly near chemical equilibrium---many orders of magnitude smaller than the forward and\slash or reverse rates.
 To quantify this phenomena, the error in forward ROP is first defined as:
@@
 while the error in reverse ROP, $\varepsilon^{\prime\prime}_{ij}$, can be defined analogously.
-Finally, for the reaction $i^{*}$ and state $j^{*}$ that result in the largest error in net ROP, i.e.
+Finally, for the reaction $i^{*}$ and the state $j^{*}$ that result in the largest error in net ROP, i.e.
 $E_{R}$, an estimate of the error attributable to floating point error accumulation from the forward and reverse ROPs can be obtained as:
@@
 Error statistics are based on the infinity-norm of the relative error detailed in~\cref{e:rel_err} for each quantity.
-The ``S'' in $E_{\frac{\text{d} S}{\text{d} t}}$ refers to the thermodynamic state parameter, either $V$ or $P$ for \conp/ and \conv/ respectively.
+The ``S'' in $E_{\frac{\text{d} S}{\text{d} t}}$ refers to the thermodynamic state parameter, either $V$ or $P$ for \conp/ and \conv/, respectively.
 }
@@
 In Table~\ref{T:source_error}, the results of this code's source-term evaluations are compared to Cantera~\cite{Cantera} on the library of PaSR conditions (\cref{t:models}).
-Very close agreement is observed for the forward and reverse ROP for all models, however the error norm is \textasciitilde\numrange{3}{4} orders of magnitude larger for the isopentanol model.
+Very close agreement is observed for the forward and reverse ROPs for all models, however the error norm is \textasciitilde\numrange{3}{4} orders of magnitude larger for the isopentanol model.
 This results from differences in evaluation of P-Log reactions between \texttt{pyJac} and Cantera; namely \texttt{pyJac} computes the logarithm of the upper and lower reaction Arrhenius rates (see supplemental~\cref{deriv-s:pdep}) analytically while Cantera evaluates this term numerically.
-If the error from P-Log reactions are not considered in~\cref{e:rel_err}, the forward and reverse ROP falls to \num{5.44e-08} and \num{1.59e-07} respectively.
+If the error from P-Log reactions are not considered in~\cref{e:rel_err}, the errors for the forward and reverse ROPs fall to \num{5.44e-08} and \num{1.59e-07}, respectively.
 We note that this discrepancy does not imply any incorrectness in either \texttt{pyJac} or Cantera---in fact, the error is still well within the proscribed tolerances in~\cref{e:rel_err}---but merely emphasizes the effect small code-changes can have on the accumulation of floating point errors.
@@
 The molar species production rate error norm is similar in magnitude to that of the net ROP, however the temperature and state-parameter rate errors again increase as the error in net species production rates is amplified by multiplication with thermodynamic properties.
-Again, we note that the above discussion does not imply that discrepancies in computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
+Again, we note that the above discussion does not imply that discrepancies in computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROPs are nearly equal, implying the reaction is in near-equilibrium.
 
@@
 \end{equation}
-The error statistics reported in this section are then based only on matrix elements where $\mathcal{J}_{ijk} \ge \frac{\mathcal{T}}{\mathcal{C}}$; this filtered form of~\cref{e:jac_error_base} is denoted $E_{\mathcal{C},k}$.
+The error statistics reported in this section are then based only on matrix elements where $\mathcal{J}_{ijk} \ge \frac{\mathcal{T}}{\mathcal{C}}$, where $\mathcal{C}$ is the threshold parameter; this filtered form of~\cref{e:jac_error_base} is denoted $E_{\mathcal{C},k}$.
 Finally, the Frobenius norm is calculated over all the states in the PaSR thermo-chemical condition database:
@@
 The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all test platforms, vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/ and sparse\slash dense Jacobians.
-It is noted that the threshold described in~\cref{e:thresh} varies slightly between the \conp/ and \conv/ cases; the reported $\bar{\mathcal{T}}$ is the average of the two, however the appropriate value was used during calculations of the error statistics.
+It is noted that the Frobenius norm described in~\cref{e:thresh} varies slightly between the \conp/ and \conv/ cases; the reported $\bar{\mathcal{T}}$ is the average of the two, however the appropriate value was used during calculations of the error statistics.
 }
@@
 It is noted that for GRI-Mech 3.0, which has a $\mathcal{T}$ roughly an order of magnitude larger, the stringent error norm is significantly smaller: $\mathcal{O}(\num{e-7})$.
-Given the intricacy of floating point error evaluation, the use of different languages and OpenCL platforms (the effect of these differences will be explored in \cref{A:per_platform}), and the general complexity of \texttt{pyJac} it would be exceedingly difficult to pinpoint an exact cause for this phenomena, as was done in~\cref{s:validation}.
+Given the intricacy of floating point error evaluation, the use of different languages and OpenCL platforms (the effect of these differences will be explored in \cref{A:per_platform}), and for the general complexity of \texttt{pyJac} it would be exceedingly difficult to pinpoint an exact cause for this phenomenon, as was done in~\cref{s:validation}.
 To ensure there is no bug or error present in the Jacobians generated by \texttt{pyJac}, a relaxed filtered error norm ($\mathcal{C} = 10^{15}$) is also presented for each model in~\cref{T:error}.
@@
 However, the explicit-mass conservation formulation of \texttt{pyJac} can introduce additional non-zero entries in two ways.
-First, if the last-species in the model (i.e., the bath-gas) participates directly in a reaction, the derivative of the forward or reverse rate of progress with respect to any other species is non-zero (see~\cref{deriv-s:dri_dnj} of the supplemental).
+First, if the last-species in the model (i.e., the bath-gas) participates directly in a reaction, the derivative of the forward or reverse rate of progress with respect to any other species is non-zero (see~\cref{deriv-s:dri_dnj} of the supplemental material).
 Similarly, if the last-species has a third-body efficiency not equal to the default (unity), this will again force all the derivative of the pressure-modification term with respect to other species to be non-zero (see~\cref{deriv-s:dci_thd} of the same).
@@
 
-However, \texttt{pyJac} allows the user to simply ignore these derivatives (via a command-line switch) and avoid any adverse effects on the Jacobian sparsity.
+However, the current \texttt{pyJac} allows the user to simply ignore these derivatives (via a command-line switch) and avoid any adverse effects on the Jacobian sparsity.
 The rational behind this choice is that many common implicit integration techniques (e.g., CVODE~\cite{Hindmarsh:2005}) used in solution of chemical kinetic initial value problems are formulated around the assumption that the supplied Jacobian is approximate; this allows the Jacobian and its LU factorization to be reused for multiple internal integration time-steps, accelerating the solution process.
@@
 In~\cref{F:jac_sparsity}, a graphical representation of the sparsity pattern for GRI-Mech 3.0 is displayed.
-In particular we note that~\cref{F:jac_sparse_approx} is has several rows that are no longer fully-dense, as result of its approximate form; these rows correspond to species directly interacting with \ce{N2}, largely in GRI-Mech 3.0's \ce{NOx} chemistry reactions.
-\cref{T:jac_sparsity} shows the density of the exact and approximate Jacobians for all mechanisms tested in this work.
-The smallest model, \ce{H2}\slash\ce{CO} is very dense with \SI{71.4}{\percent} of the Jacobian entries non-zero, however this drops to \SI{56.7}{\percent} for GRI-Mech 3.0, and continues to decrease to \SI{28.7}{\percent} for USC-Mech II and just \SI{11.5}{\percent} for the isopentanol model.
+In particular we note that~\cref{F:jac_sparse_approx} is has several rows that are no longer fully-dense, as result of its approximate form; these rows correspond to species directly interacting with \ce{N2}, largely in GRI-Mech 3.0's nitrogen chemistry reactions.
+\cref{T:jac_sparsity} shows the density of the exact and approximate Jacobians for all chemical kinetic models tested in this work.
+The smallest model, \ce{H2}\slash\ce{CO} is very dense with \SI{71.4}{\percent} of the exact Jacobian entries non-zero, however this drops to \SI{56.7}{\percent} for GRI-Mech 3.0, and continues to decrease to \SI{28.2}{\percent} for USC-Mech II and just \SI{11.5}{\percent} for the isopentanol model.
 The approximate Jacobian assumption drops the density of Jacobian by \textasciitilde\SIrange{3}{7}{\percent} for all models.
@@
 
-Currently, \texttt{pyJac} can utilize two common sparse matrix storage formats~\cite{netlib_templates}, compressed row\slash column storage (CSR\slash CSC), used for ``C'' and ``F''-ordered data respectively.
-For convenience we will outline only the CSR format but the CSC format is very similar~\cite{netlib_templates}.
-An $N\times N$ CSR matrix is composed of three vectors, a value vector of length $NNZ$---the number of non-zero elements in the Jacobian---which stores the elements of the Jacobian, a row pointer vector---of length $N$---which stores the locations in the value vector that begin a row, and a column index vector---of length $NNZ$---which stores the column indices of the elements in the value vector.
+Currently, \texttt{pyJac} can utilize two common sparse matrix storage formats~\cite{netlib_templates}, compressed row\slash column storage (CRS\slash CCS), used for ``C'' and ``F''-ordered data respectively.
+For convenience we will outline only the CRS format but the CCS format is very similar~\cite{netlib_templates}.
+An $N\times N$ CRS matrix is composed of three vectors, a value vector of length $NNZ$---the number of non-zero elements in the Jacobian---which stores the elements of the Jacobian, a row pointer vector---of length $N$---which stores the locations in the value vector that begin a row, and a column index vector---of length $NNZ$---which stores the column indices of the elements in the value vector.
 
@@
 Generally, this means that a lookup operation is required to find the sparse Jacobian index for any pair of state variables; in some cases this can be avoided, e.g., the rows corresponding to derivatives of the temperature and thermodynamic state parameter source-terms are fully dense in \texttt{pyJac}, and hence no lookup is necessary.
-This lookup operation is currently implemented as a simple for-loop, e.g., for a sparse lookup of a pair of indices $(i, j)$ in a CSR matrix, the lookup function searches the column index vector between the values $\texttt{row\_pointer}[i] \ldots \texttt{row\_pointer}[i + 1]$ for $j$, and returns the offset from $\texttt{row\_pointer}[i]$ (or $\num{-1}$ if not found).
+This lookup operation is currently implemented as a simple for-loop, e.g., for a sparse lookup of a pair of indices $(i, j)$ in a CRS matrix, the lookup function searches the column index vector between the values $\texttt{row\_pointer}[i] \ldots \texttt{row\_pointer}[i + 1]$ for $j$, and returns the offset from $\texttt{row\_pointer}[i]$ (or $\num{-1}$ if not found).
 As will be explored in~\cref{S:jacobian_results}, this slows down sparse Jacobian evaluation, and might be improved---at the cost of increased constant data usage, a limitation for OpenCL---by a static mapping of the full Jacobian indices to the sparse index (or some null value if the entry is empty).
@@
 Additionally, the exact form of the Jacobian (as opposed the the ``approximate'' form discussed in~\cref{S:sparsity}) was used in all cases.
-Finally, unless stated otherwise: the performance results utilized a single CPU core, the \conp/ assumption, a vector-width of \num{8}\slash\num{128} and ``C''\slash ``F''-ordered data for the CPU\slash GPU cases respectively; the run times reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
+Finally, unless stated otherwise: the performance results utilized a single CPU core, the \conp/ assumption, a vector-width of \num{8}\slash\num{128} and ``C''\slash ``F''-ordered data for the CPU\slash GPU cases, respectively; the run times reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
 The effects of choice of vector-width, data-ordering and differences between \conp/ and \conv/ evaluations on the CPU\slash GPU will be explored in~\cref{S:source_results}, while parallel scaling for multiple CPU cores will be exampled in~\cref{S:source_results,S:jacobian_results}.
@@
 \Cref{F:cpu_source} explores the performance of the source-term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
-Source-term evaluations---critical in their own right for direct numerical simulations of reacting flow~\cite{Spafford:2010aa}, among other applications---also give a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation performance.
+Source-term evaluations---critical in their own right for direct numerical simulations of reacting flows~\cite{Spafford:2010aa}, among other applications---also give a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation performance.
 
@@
 On both machines, the unvectorized OpenCL code is faster than the baseline parallel OpenMP code; by \SIrange{1.30}{1.51}{$\times$} on the \avx/ CPU, and \SIrange{1.25}{1.27}{$\times$} on the \sse/ machine.
-Additionally, the shallow-vectorized OpenCL code is \SIrange{2.53}{2.92}{$\times$} and \SIrange{3.40}{4.08}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
+Additionally, the shallow-vectorized OpenCL code is \SIrange{2.53}{2.92}{$\times$} and \SIrange{3.40}{4.08}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines, respectively.
 
@@
 In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source-terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean run time per condition; the run time continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
-\Cref{F:gpu_source_speedup} shows the speedup in source-term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} with two different vector-widths (a.k.a., the block size in NVIDIA terminology), \num{64} and \num{128}; the fastest \gpunew/ case with a vector-width of 128 is \SIrange{1.40}{1.88}{$\times$} faster than the slowest case (\gpuold/ with a vector-width of \num{64}) depending on the chemical model in question.
+\Cref{F:gpu_source_speedup} shows the speedup in source-term evaluations that the \gpunew/ GPU achieves over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} with two different vector-widths (a.k.a., the block size in NVIDIA terminology), \num{64} and \num{128}; the fastest \gpunew/ case with a vector-width of 128 is \SIrange{1.40}{1.88}{$\times$} faster than the slowest case (\gpuold/ with a vector-width of \num{64}) depending on the chemical model in question.
 Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector-width has minimal effect on performance for most of the \gpunew/ and \gpuold/ cases; the GRI-Mech 3.0 and USC-Mech II models show the largest improvements with a vector-width of \num{128}, \textasciitilde\SIrange{10}{18}{\percent} for both GPUs.
@@
 \Cref{f:source_permutate} shows the effect of changing data-ordering patterns, the \conp/ or \conv/ formulation, and the CPU vector-width size on the performance of source-term evaluation in \texttt{pyJac}.
-In~\cref{F:source_conpvsconv}, it is seen that choice of the constant-pressure or constant-volume formulation has little to no effect on run time for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx/ machine.
+In~\cref{F:source_conpvsconv}, it is seen that choice of the \conp/ or \conv/ formulation has little to no effect on run time for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx/ machine.
 Generally speaking, the difference between the \conp/ and \conv/ formulations has only a marginal effect on performance regardless of CPU\slash GPU choice.
@@
       \includegraphics[width=\textwidth]{source_vector_width.pdf}
-      \caption{The effect of vector-width on ``C''-ordered shallow-vectorized Intel OpenCL source-term evaluations in the \avx/ CPU.}
+      \caption{The effect of vector-width on ``C''-ordered shallow-vectorized Intel OpenCL source-term evaluations on the \avx/ CPU.}
       \label{F:source_vector_width}
@@
   \caption{The effect of the \conp/ and \conv/ formulations, ``C'' and ``F'' data-ordering, and CPU vector-width on source-term evaluation performance in \texttt{pyJac}.
-	   It is noted the shallow-vectorized ``C'' ordered OpenCL cases corresponds to the vectorized data ordering described in~\cref{S:data}}
+	   It is noted the shallow-vectorized ``C'' ordered OpenCL cases correspond to the vectorized data ordering described in~\cref{S:data}}
   \label{f:source_permutate}
@@
 
-In contrast,~\Cref{F:source_cvsf} shows significant speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, e.g., the \SIrange{1.35}{2.09}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-ordered OpenMP baseline.
+In contrast,~\cref{F:source_cvsf} shows significant speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, e.g., the \SIrange{1.35}{2.09}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-ordered OpenMP baseline.
 Additionally, it is noted that the ``C''-ordered shallow-vectorization in~\cref{F:source_cvsf} (as well as the other shallow-vectorized CPU data shown in \cref{S:source_results,S:jacobian_results}) utilizes the vectorized-data ordering described in~\cref{S:data}; this case achieves speedups of \SIrange{2.14}{2.58}{$\times$} over the ``F''-ordered shallow-vectorization, demonstrating the value of the vectorized data ordering for CPU execution.
@@
 It is noted that for the isopentanol model, ``C''-ordered data is \textasciitilde\SI{1.08}{$\times$} faster on both GPUs (while the trend is less clear for the other models).
-The rough equivalence between the ``C'' and ``F''-ordered codes GPUs is counter to what one might expect, as typically speaking coalesced memory access in a shallow-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
-However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector-width size and thus, likely well-coalesced.
+The rough equivalence between the ``C'' and ``F''-ordered codes on GPUs is counter to what one might expect, as typically speaking coalesced memory access in a shallow-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
+However, the vectorized-data ordering here ensures that memory accesses are aligned to the vector-width size and thus, likely well-coalesced.
 
@@
       \includegraphics[width=\textwidth]{source_crossover.pdf}
-      \caption{The mean run time per-condition of OpenMP source-term evaluation on different cores compared to a shallow-vectorized Intel OpenCL evaluation on \num{1} (solid lines) and \num{4} (dashed lines) cores on \avx/ machine.}
+      \caption{The mean run time per condition of OpenMP source-term evaluation on different cores compared to a shallow-vectorized Intel OpenCL evaluation on \num{1} (solid lines) and \num{4} (dashed lines) cores on \avx/ machine.}
       \label{F:source_crossover}
@@
 In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for both Intel OpenCL and OpenMP, likely resulting from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD instructions\slash multiple threads.
-Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.
+Additionally, OpenMP tends to have higher scalings than the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores, respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.
 This is due in large part to the superior performance of the shallow-vectorized OpenCL code, coupled with the relatively small amounts of work associated with source-term evaluation.
@@
 This again demonstrates that the SIMD-vectorization becomes efficient as the amount of work to perform increases (with increasing model size).
-Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one; this is likely caused by a combination of use of an OpenCL vector-width greater than the native CPU vector-width (i.e., \num{8} v.s., \num{2}), and the improved data-locality for the vectorized-data ordering, as discussed in~\cref{S:data}, and results a modest \SIrange{7}{17}{\percent} improvement over the nominal vector-width.
+Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one; this is likely caused by a combination of use of an OpenCL vector-width greater than the native CPU vector-width (i.e., \num{8} versus \num{2}), and the improved data-locality for the vectorized-data ordering, as discussed in~\cref{S:data}, and results in a modest \SIrange{7}{17}{\percent} improvement over the nominal vector-width.
 
@@
       \includegraphics[width=\textwidth]{sparse_vs_dense.pdf}
-      \caption{The mean run time per-condition of sparse and dense Jacobian evaluation for Intel OpenCL\slash OpenMP on the \avx/ machine.}
+      \caption{The mean run time per condition of sparse and dense Jacobian evaluations for Intel OpenCL\slash OpenMP on the \avx/ machine.}
       \label{F:sparse_vs_dense}
@@
       \includegraphics[width=\textwidth]{sparse_vs_dense_speedup.pdf}
-      \caption{The speedup of the sparse and dense Jacobian evaluation for Intel OpenCL over the sparse\slash dense OpenMP baseline on the \avx/ machine.}
+      \caption{The speedup of the sparse and dense Jacobian evaluations for Intel OpenCL over the sparse\slash dense OpenMP baseline on the \avx/ machine.}
       \label{F:sparse_vs_dense_speedup}
@@
       \includegraphics[width=\textwidth]{sparse_vs_dense_sse.pdf}
-      \caption{The mean run time per-condition of sparse and dense Jacobian evaluation for Intel OpenCL\slash OpenMP on the \sse/ machine.}
+      \caption{The mean run time per condition of sparse and dense Jacobian evaluations for Intel OpenCL\slash OpenMP on the \sse/ machine.}
       \label{F:sparse_vs_dense_sse}
@@
       \includegraphics[width=\textwidth]{sparse_vs_dense_see_speedup.pdf}
-      \caption{The speedup of the sparse and dense Jacobian evaluation for Intel OpenCL over the sparse\slash dense OpenMP baseline on the \sse/ machine.}
+      \caption{The speedup of the sparse and dense Jacobian evaluations for Intel OpenCL over the sparse\slash dense OpenMP baseline on the \sse/ machine.}
       \label{F:sparse_vs_dense_sse_speedup}
   \end{subfigure}
-  \caption{Performance of sparse and dense Jacobian evaluation on the CPU platforms in \texttt{pyJac}.}
+  \caption{Performance of sparse and dense Jacobian evaluations on the CPU platforms in \texttt{pyJac}.}
   \label{F:jacobian_perfomance}
@@
 As a result, the shallow-vectorized sparse OpenCL code is as fast, or faster than the dense OpenMP code in all cases on the \avx/ machine (\cref{F:sparse_vs_dense}).
-\Cref{F:sparse_vs_dense_speedup} shows the speedup of the sparse\slash dense shallow-vectorized OpenCL the over the same sparse\slash dense Jacobian format on OpenMP; the dense OpenCL code is \SIrange{3.03}{4.23}{$\times$} faster than the corresponding dense OpenMP code, this speedup increases to \SIrange{6.63}{9.44}{$\times$} for the sparse Jacobian.
+\Cref{F:sparse_vs_dense_speedup} shows the speedup of the sparse\slash dense shallow-vectorized OpenCL over the same sparse\slash dense Jacobian format on OpenMP; the dense OpenCL code is \SIrange{3.03}{4.23}{$\times$} faster than the corresponding dense OpenMP code.
+This speedup increases to \SIrange{6.63}{9.44}{$\times$} for the sparse Jacobian.
 
@@
   \end{subfigure}
-  \caption{The parallel scaling efficiencies of sparse and dense Jacobian evaluation for the shallow-vectorized Intel OpenCL and OpenMP codes on the \avx/ CPU.}
+  \caption{The parallel scaling efficiencies of sparse and dense Jacobian evaluations for the shallow-vectorized Intel OpenCL and OpenMP codes on the \avx/ CPU.}
 \end{figure}
@@
 The plot is a bit difficult to read, as most of the data-points are clustered together, however the key takeaway is that the parallel scaling efficiency of the shallow-vectorized OpenCL code is very similar to that of the OpenMP code, in contrast to the parallel scaling efficiency of source-term evaluation (\cref{F:source_parallel_scaling}).
-The \ce{H2}\slash\ce{CO} model has the worst scaling for both codes, ranging from \numrange{0.94}{0.54} and \numrange{0.99}{0.82} for OpenCL and OpenMP respectively on \numrange{2}{16} cores.
+The \ce{H2}\slash\ce{CO} model has the worst scaling for both codes, ranging from \numrange{0.94}{0.54} and \numrange{0.99}{0.82} for OpenCL and OpenMP, respectively on \numrange{2}{16} cores.
 As the model size increases, the efficiency of the OpenCL code improves dramatically, reaching \numrange{0.997}{0.84} for the isopentanol model.
 \Cref{F:dense_jac_scaling} shows scaling for the shallow-vectorized dense Jacobian OpenCL code.
-In this case, the isopentanol model has the worst scaling for both codes, it is noted that due to the sheer size of the dense isopentanol Jacobian---storing the dense matrix for a single thermo-chemical state takes over \SI{1}{\mega\byte} of data---the total number of states for the dense isopentanol Jacobian evaluation was limited to \num{50000} (which requires over \SI{50}{\giga\byte} of memory); this greatly drops the computation cost for this case, and adversely affects the scaling efficiency as was discussed in~\cref{S:source_results}.
-Excluding the isopentanol model, the dense shallow-vectorized OpenCL scalings are slightly higher than for source-term evaluation, e.g., \numrange{0.70}{0.78} and \numrange{0.52}{0.61} for \num{4} and \num{8} cores respectively (compared to \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for shallow-vectorized OpenCL source-term on the same \num{4} and \num{8} cores); this is due to the higher computational cost of Jacobian evaluation, and hence more available work per CPU-core.
-As in~\cref{S:source_results}, we note that the shallow-vectorized dense Jacobian code running on \num{1} and \num{4} cores of the \avx/ machine is roughly as fast as the OpenMP code on \num{4} and \num{16} cores respectively; the parallel scaling efficiency of OpenMP at \num{16} cores (excluding isopentanol as previously) is \numrange{0.63}{0.71}, similar to the shallow-vectorization's efficiency of \numrange{0.70}{0.78} for \num{4} cores.
+In this case, the isopentanol model has the worst scaling for both codes.
+Tt is noted that due to the sheer size of the dense isopentanol Jacobian---storing the dense matrix for a single thermo-chemical state takes over \SI{1}{\mega\byte} of data---the total number of states for the dense isopentanol Jacobian evaluation was limited to \num{50000} (which requires over \SI{50}{\giga\byte} of memory); this greatly drops the computation cost for this case, and adversely affects the scaling efficiency as was discussed in~\cref{S:source_results}.
+Excluding the isopentanol model, the dense shallow-vectorized OpenCL scalings are slightly higher than for source-term evaluation, e.g., \numrange{0.70}{0.78} and \numrange{0.52}{0.61} for \num{4} and \num{8} cores, respectively (compared to \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for shallow-vectorized OpenCL source-term on the same \num{4} and \num{8} cores); this is due to the higher computational cost of Jacobian evaluation, and hence more available work per CPU-core.
+As in~\cref{S:source_results}, we note that the shallow-vectorized dense Jacobian code running on \num{1} and \num{4} cores of the \avx/ machine is roughly as fast as the OpenMP code on \num{4} and \num{16} cores, respectively; the parallel scaling efficiency of OpenMP at \num{16} cores (excluding isopentanol as previously) is \numrange{0.63}{0.71}, similar to the shallow-vectorization's efficiency of \numrange{0.70}{0.78} for \num{4} cores.
 
@@
       \includegraphics[width=\textwidth]{gpu_sparse_vs_dense.pdf}
-      \caption{Mean run time per-condition of sparse and dense Jacobian evaluation on the \gpunew/ GPU.}
+      \caption{Mean run time per condition of sparse and dense Jacobian evaluations on the \gpunew/ GPU.}
       \label{F:gpu_sparse_vs_dense}
@@
       \includegraphics[width=\textwidth]{gpu_jacobian_speedup.pdf}
-      \caption{Speedup of the \gpunew/ over \gpuold/ GPUs for sparse and dense Jacobian evaluation, normalized per-Jacobian type (i.e., sparse versus dense).}
+      \caption{Speedup of the \gpunew/ over \gpuold/ GPUs for sparse and dense Jacobian evaluations, normalized per-Jacobian type (i.e., sparse versus dense).}
       \label{F:gpu_jacobian_speedup}
@@
 
-The mean run time per-condition for the sparse and dense Jacobian evaluation on the \gpunew/ GPU is plotted in~\cref{F:gpu_sparse_vs_dense}.
-As in the species evaluation case, the mean run time per-condition drops steadily, for both cases becoming roughly constant after just \num{3e3} states (except for \ce{H2}\slash\ce{CO}, which levels off near \num{e4} conditions).
+The mean run time per condition for the sparse and dense Jacobian evaluations on the \gpunew/ GPU is plotted in~\cref{F:gpu_sparse_vs_dense}.
+As in the species evaluation case, the mean run time per condition drops steadily, for both cases becoming roughly constant after just \num{3e3} states (except for \ce{H2}\slash\ce{CO}, which levels off near \num{e4} conditions).
 Sparse Jacobian evaluation is significantly slower for all models before the GPU becomes saturated (due to the indirect indexing lookup), however past the saturation point the performance gap between the sparse and dense evaluations narrows.
@@
 The analytical Jacobian on the \gpunew/ and \gpuold/ shows large speedups, increasing with chemical model size, in the range of \SIrange{3.81}{17.60}{$\times$}; the \gpunew/ has a larger speedup than the \gpuold/ (\SI{17.60}{$\times$} v.s., \SI{14.75}{$\times$}) for the isopentanol model due to its larger available memory.
-Although not pictured, the dense analytical-Jacobian on the \gpunew/ GPU has larger speedups compared with the dense finite-difference Jacobian: \SIrange{4.04}{45.13}{$\times$}; the \gpunew/ GPU again has a significantly larger speedup than the \gpuold/ for the isopentanol model (\SI{45.13}{$\times$} vs \SI{23.85}{$\times$}), further underscoring the effect of the larger available memory on the \gpunew/.
+Although not pictured, the dense analytical-Jacobian on the \gpunew/ GPU has larger speedups compared with the dense finite-difference Jacobian: \SIrange{4.04}{45.13}{$\times$}; the \gpunew/ GPU again has a significantly larger speedup than the \gpuold/ for the isopentanol model (\SI{45.13}{$\times$} versus \SI{23.85}{$\times$}), further underscoring the effect of the larger available memory on the \gpunew/.
 
@@
 
-\Cref{F:v1_vs_v2} compares the performance for evaluation of the dense-analytical Jacobian of the new version of \texttt{pyJac} to the previous version~\cite{pyjac16} on the \sse/ CPU and \gpuold/ GPU; dense Jacobian evaluation was selected as it was the only type implemented in the previous version of \texttt{pyJac}.
+\Cref{F:v1_vs_v2} compares the performance for evaluation of the dense-analytical Jacobian of the new version of \texttt{pyJac}, \texttt{pyJac-v2}, to the previous version, \texttt{pyJac-v1} \cite{pyjac16}, on the \sse/ CPU and \gpuold/ GPU; dense Jacobian evaluation was selected as it was the only type implemented in the previous version of \texttt{pyJac}.
 On the \sse/ CPU, the \texttt{pyJac-v2} is faster than \texttt{pyJac-v1} for OpenMP evaluation for the larger chemical models; the static OpenMP code generated by \texttt{pyJac-v1} (see~\cref{s:unittest}) is \SI{1.79}{$\times$} faster for the \ce{H2}\slash\ce{CO} model, and only \SI{1.09}{$\times$} slower for the GRI-Mech 3.0 model.
@@
 
-As demonstration of the type of issue we're discussing, a minimum working example has been created~\cite{nvidia_mwe} that demonstrates a failure of NVIDIA's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla \gpunew/ GPU; by simply changing to another runtime (e.g., Intel)---with no code changes or recompilation---the correct result can be found.
+As demonstration of the type of issue discussed, a minimum working example has been created~\cite{nvidia_mwe} that demonstrates a failure of NVIDIA's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla \gpunew/ GPU; by simply changing to another runtime (e.g., Intel)---with no code changes or recompilation---the correct result can be found.
 This provides a particularly vexing problem for the programmer, as there is often little that can be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
 Further, as noted throughout this work certain code-generation patterns can break OpenCL execution\slash vectorization, e.g., the failure of POCL to achieve vectorization or Intel OpenCL's failure to vectorize the finite-difference Jacobian.
-Indeed the vectorization process attempted by most OpenCL runtimes (and thus, reasons for incorrect\slash unvectorized code-output) are obscured from the user, making reasoning about errors or performance trends quite difficult.
+Indeed the vectorization process attempted by most OpenCL runtimes (and thus, reasons for incorrect\slash unvectorized code-output) is obscured from the user, making reasoning about errors or performance trends quite difficult.
 
@@
 Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py}) is a key priority going forwards as OpenMP is a standard library on most machines.
-In addition, CUDA~\cite{NVIDIA:2018}, and Intel's open-source OpenCL alternative, ISPC~\cite{pharr2012ispc}, have been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017} and (in the case of ISPC) initial testing.
+In addition, CUDA~\cite{NVIDIA:2018} has been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017}, while Intel's open-source OpenCL alternative, ISPC~\cite{pharr2012ispc}, has been relatively stable and easy to work with during preliminary efforts with the unit-testing discussed in~\cref{s:unittest}.
 The current deep-vectorization formulation would be executable for both CUDA and ISPC targets, as these languages implement double-precision atomic operations, further recommending their use.
@@
 Deep-vectorization, though possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, did not yield any performance benefit as POCL did not achieve vectorizations for any execution pattern studied; deep-vectorization deserves further study with other platforms (e.g., CUDA).
-Vectorized data-orderings specifically formulated for shallow\slash deep SIMD\slash SIMT vectorizations were proposed, resulting in a shallow-vectorized speedup of \SIrange{2.14}{2.58}{$\times$} over a more standard ``F''-ordered (column major) ordered shallow-vectorization; when compared to the corresponding speedup of \SIrange{1.35}{2.09}{$\times$} for the ``C''-ordered (row major) OpenMP code over an ``F''-ordered equivalent, the value of the vectorized data-ordering is evident.
+Vectorized data-orderings specifically formulated for shallow\slash deep SIMD\slash SIMT vectorizations were proposed, resulting in a shallow-vectorized speedup of \SIrange{2.14}{2.58}{$\times$} over a more standard ``F''-ordered (column major) shallow-vectorization; when compared to the corresponding speedup of \SIrange{1.35}{2.09}{$\times$} for the ``C''-ordered (row major) OpenMP code over an ``F''-ordered equivalent, the value of the vectorized data-ordering is evident.
 In addition, the choice of \conp/\slash\conv/ formulation was determined to be negligible on source-term evaluations, while varying the OpenCL vector-width had only modest (\textasciitilde\SI{10}{$\percent$}) performance effects on both the CPU and GPU.
-The strong parallel scaling efficiency was examined for source-term evaluation; the parallel OpenMP code generally had better efficiencies, but when compared to a shallow-vectorized code with roughly equivalent performance---e.g., source-term shallow-vectorized evaluation on \num{4} cores to \num{16} OpenMP cores---the scalings were similar.
+The strong parallel scaling efficiency was also examined for source-term evaluation; the parallel OpenMP code generally had better efficiencies, but when compared to a shallow-vectorized code with roughly equivalent performance---e.g., source-term shallow-vectorized evaluation on \num{4} cores to \num{16} OpenMP cores---the scalings were similar.
 Examination of SIMD-efficiency revealed increasing speedups with chemical model-size for source-term evaluation on the \avx/ CPU, while on the \sse/ CPU speedups larger than the nominal vector-width were achieved.
@@
 The performance of sparse and dense Jacobian evaluation was also characterized on the CPU\slash GPU platforms.
-The sparse shallow-vectorized OpenCL code was \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.14}{5.03}{$\times$} faster than OpenMP evaluation on the \avx/ and \sse/ CPUs respectively, while dense shallow-vectorized evaluation was \SIrange{3.03}{4.23}{$\times$} and \SIrange{1.92}{2.47}{$\times$} faster on the same.
+The sparse shallow-vectorized OpenCL code was \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.14}{5.03}{$\times$} faster than OpenMP evaluation on the \avx/ and \sse/ CPUs, respectively, while dense shallow-vectorized evaluation was \SIrange{3.03}{4.23}{$\times$} and \SIrange{1.92}{2.47}{$\times$} faster on the same.
 Sparse Jacobian evaluation was slower than dense evaluation on all CPU\slash GPU platforms due to indirect lookup requirements in array indexing, however the shallow-vectorized OpenCL code was less adversely affected---a slowdown of \SIrange{1.41}{3.34}{$\times$} on the \avx/ CPU---than OpenMP, which was \SIrange{2.47}{10.42}{$\times$} slower on the same.
@@
 Analytical Jacobian evaluation was significantly faster than a first-order finite-difference based approach on all platforms, achieving speedups of \SIrange{3.81}{17.60}{$\times$} and \SIrange{4.04}{45.13}{$\times$} for sparse and dense evaluation (respectively) on the GPUs; on the \avx/ CPU analytical evaluation was \SIrange{3.92}{55.11}{$\times$} and \SIrange{9.68}{245.63}{$\times$} for sparse and dense Jacobian evaluation (respectively) with OpenMP\slash shallow-vectorized OpenCL.
-Finally, the performance of dense analytical-Jacobian evaluation in this new version of \texttt{pyJac} was compared to the previous version~\cite{pyjac16} on the \sse/\slash\gpuold/ CPU\slash GPUs; OpenMP evaluation on the \sse/ CPU was \SI{1.79}{$\times$} slower for the smallest chemical model, but \SIrange{3.37}{10.19}{$\times$} faster for the larger models, while the shallow-vectorized OpenCL code was \SIrange{1.37}{19.56}{$\times$} faster than the previous version over all chemical models.
+Finally, the performance of dense analytical-Jacobian evaluation in this new version of \texttt{pyJac} was compared to the previous version~\cite{pyjac16} on the \sse/\slash\gpuold/ CPU\slash GPU; OpenMP evaluation on the \sse/ CPU was \SI{1.79}{$\times$} slower for the smallest chemical model, but \SIrange{3.37}{10.19}{$\times$} faster for the larger models, while the shallow-vectorized OpenCL code was \SIrange{1.37}{19.56}{$\times$} faster than the previous version over all chemical models.
 Similarly, evaluation on the \gpuold/ GPU was \SI{1.89}{$\times$} slower for the smallest model but \SIrange{1.25}{2.84}{$\times$} faster for the larger models.
 
-Future extensions of this work will include additional target-languages (vectorized OpenMP, CUDA, ISPC) to improve ease of use and reliability, as well as to enable meaningful deep-vectorization.
-Additionally, existing vectorized sparse\slash dense-matrix libraries will be assessed and missing operations for LU-factorization\slash matrix-vector multiplication implemented.
+Future extensions of this work will include additional target-languages (vectorized OpenMP, CUDA, ISPC, etc.) to improve ease of use and reliability, as well as to enable meaningful deep-vectorization.
+Additionally, existing vectorized sparse\slash dense-matrix libraries will be assessed and missing operations for LU-factorization\slash matrix-vector multiplication should be implemented.
 

### 2018-03-21 derivations/limittest/limit_test.py

@@ -1,0 +1,204 @@
+from __future__ import print_function
+from sympy import symbols, log, exp, limit, KroneckerDelta, diff, \
+    Product, factor, Pow, Symbol, simplify, Limit, Mul, expand
+from optionloop import OptionLoop
+
+
+def __get_dci(fall_type='chem', blend_type='troe', pr_type='mix', var='E'):
+    # create temperature
+    T = symbols('T')
+
+    # create kf's
+    kf, kinf = symbols('kf kinf', real=True, nonnegative=True)
+
+    # create third body efficiency & volume
+    V, C, P = symbols('V C P', real=True, nonnegative=True)
+    Xi, alphaNS, alphaj = symbols('Xi alphaNS alphaj', real=True)
+
+    # species
+    m, Ns, j = symbols('m Ns j', integer=True, nonnegative=True)
+
+    # create pr
+    Pr = kf * Xi / kinf
+    R = 8.314
+
+    # create Fi & Troe params
+    if blend_type == 'troe':
+        T1, T2, T3, a = symbols('T1 T2 T3 a', real=True)
+
+        Fcent = a * exp(-T / T1) + (1 - a) * exp(-T / T3) + exp(-T2 / T)
+        Atroe = -0.68 * log(Fcent, 10) + log(Pr, 10) - 0.4
+        Btroe = -1.1762 * log(Fcent, 10) - 0.14 * log(Pr, 10) + 0.806
+
+        Fi = Fcent ** (1 / (1 + (Atroe / Btroe)**2))
+    elif blend_type == 'sri':
+        a, b, c, d, e = symbols('a b c d e', real=True)
+        X = 1 / (log(Pr, 10) ** 2 + 1)
+        Fi = d * T**e * (a * exp(-b / T) + exp(-T / c)) ** X
+    elif blend_type == 'lind':
+        Fi = 1
+
+    # chemically activated form
+    if fall_type == 'chem':
+        ci = Fi / (1 + Pr)
+    elif fall_type == 'fall':
+        ci = Pr * Fi / (1 + Pr)
+
+    # now create derivative temporary products (assuming mixture)
+    if var == 'T':
+        b0, binf, e0, einf = symbols('b0 binf e0 einf', real=True)
+        if pr_type in ['mix', 'spec']:
+            Theta_Pr = (b0 - binf + e0 / (R * T) - einf / (R * T)) / T
+            if pr_type == 'mix':
+                theta_Pr = -C * kf * alphaNS / (T * kinf)
+            else:
[…]

### 2018-03-21 derivations.tex
edit
derivation frontmatter

@@
 
+\title{A System of Equations and Derivation of Sparse Constant-Pressure\slash Constant-Volume Chemical Kinetic Jacobians for \texttt{pyJac}}
+
+\author[1]{Nicholas Curtis\corref{corr}}
+\ead{nicholas.curtis@uconn.edu}
+\author[1]{Chih-Jen Sung}
+
+\address[1]{Department of Mechanical Engineering, University of Connecticut, Storrs, CT 06269, USA}
+\cortext[corr]{Corresponding author}
+
 \begin{document}
+
+\begin{frontmatter}
+
+%====================================================================
+\begin{abstract} % not to exceed 200 words
+This document presents the system of chemical kinetic equations evaluated by the new version of code-generation platform, \texttt{pyJac-v2}.
+First, the thermo-chemical source terms for fixed-mass, adiabatic constant-pressure\slash constant-volume (\conp/\slash\conv/) reactors are given, followed by the derivation of the sparse analytical chemical kinetic Jacobian for both reactors.
+Finally, the resulting Jacobian equations are summarized using a generalized notation for a quick-reference.
+\end{abstract}
+
+% (Provide 2-4 keywords describing your research. Only abbreviations firmly
+% established in the field may be used. These keywords will be used for
+% sessioning/indexing purposes.)
+\begin{keyword}
+    Chemical Kinetics\sep SIMD\sep SIMT\sep Sparse\sep Jacobian
+\end{keyword}
+
+\end{frontmatter}
+
 \section{Introduction}
@@
 \loadeq{dparam}
+$W_{k}$ and $W_{\ns/}$ are the molecular weights of species $k$ and $\ns/$, respectively.
 

### 2018-03-21 paper.tex
fix

@@
 Youssefi~\cite{Youssefi:2011tm} explored the importance of analytical Jacobian matrices for time-scale analysis techniques as well as their effect on computational efficiency in zero-dimensional homogeneous reactor simulations.
-Bisetti~\cite{Bisetti:2012jw} developed an isothermal, isobaric analytical Jacobian code-generation utility; although this set of assumptions (particularly, the choice of an isothermal system) are incompatible with many combustion simulations, a significant increase in Jacobian sparsity was exhibited.
+Bisetti~\cite{Bisetti:2012jw} developed an isothermal, isobaric analytical Jacobian code-generation utility; although the chosen isothermal assumption is not typical in most combustion simulations, a significant increase in Jacobian sparsity was exhibited.
 Additionally, in the same work Bisetti provided a novel way to compute dense matrix-vector multiplications resulting from a change of system variables without need for storage of the full dense Jacobian.

### 2018-03-22 paper.tex
rewrite lit review to make it less similar to past works

@@
 Recognizing the need to accelerate Jacobian evaluation and factorization, a number of recent works have been published on analytical chemical kinetic Jacobian evaluation; although as will be discussed at the end of this section, this work offers several key improvements over past efforts.
-The \texttt{TChem} software package~\cite{Safta:2011vn} was one of the first packages developed providing this functionality, but has several limitations including: incompatibilities with modern reaction types---i.e., pressure dependent Arrhenius (or P-Log) and Chebyshev reactions---and lack of thread-safety to enable parallel execution~\cite{Curtis2017:tchem}.
+The \texttt{TChem} software package~\cite{Safta:2011vn} was one of the first codes developed providing analytical Jacobian evaluation, but has several limitations including: incompatibilities with modern reaction types---i.e., pressure dependent Arrhenius (or P-Log) and Chebyshev reactions---and lack of thread-safety to enable parallel execution~\cite{Curtis2017:tchem}.
 Youssefi~\cite{Youssefi:2011tm} explored the importance of analytical Jacobian matrices for time-scale analysis techniques as well as their effect on computational efficiency in zero-dimensional homogeneous reactor simulations.
@@
 Additionally, in the same work Bisetti provided a novel way to compute dense matrix-vector multiplications resulting from a change of system variables without need for storage of the full dense Jacobian.
-Perini et al.~\cite{Perini:2012gy} developed an analytical Jacobian code for constant-volume combustion, with additional options to increase sparsity (at expense of strict correctness), and enable tabulation of temperature dependent properties; a \SI{80}{\percent} speedup over a finite-difference-based Jacobian was reported when used in a multidimensional reactive-flow simulation.
+Perini et al.~\cite{Perini:2012gy} developed an analytical Jacobian code for constant-volume combustion, with additional options to increase sparsity (at expense of strict correctness), and enable tabulation of temperature dependent properties; an \SI{80}{\percent} speedup over a finite-difference-based Jacobian was reported when used in a multidimensional reactive-flow simulation.
 Dijkmans et al.~\cite{Dijkmans:2014bb} developed a GPU-based analytical Jacobian code with optional tabulation of temperature-dependent properties, and found speedups up to \SI{120}{$\times$} for zero-dimensional chemical kinetic integration with large chemical models (\textasciitilde\num{3000} species) using a Tesla C2075 GPU.
@@
 A number of recent works have investigated the use of high-performance SIMT-devices to accelerate reactive-flow and chemical kinetics simulations.
-Spafford et al.~\cite{Spafford:2010aa} investigated an implementation of an explicit direct numerical simulation code for compressible turbulent combustion on a Tesla C1060 GPU, and found an order of magnitude speedup in evaluation of species production rates as compared to a serial CPU implementation on an AMD-Opteron processor.
-Shi et al.~\cite{Shi:2011aa} developed a code to evaluate species production rates and factorize the chemical kinetic Jacobian on a Tesla C2050 GPU, and found an order of magnitude speedup during integration of large chemical models when these operations were offloaded to the GPU as compared to the standard \texttt{CHEMKIN}~\cite{kee1989chemkin} and \texttt{LAPACK}~\cite{Anderson:1999aa} libraries on a quad-core Intel i7 930 CPU; it was not clear how\slash if the CPU code was parallelized.
-Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit fourth-order Runge--Kutta integrator for non-stiff chemical kinetic integration on a Tesla C2075 GPU, and found a speedup of nearly two orders of magnitude when compared with a sequential CPU-code on an Intel Xeon X5650 CPU.
-Shi et al.~\cite{Shi:2012aa} later extended their work to develop a stabilized-explicit solver on a Tesla C2050, paired with a traditional implicit integrator on quad-core Intel i7 930 which handled the stiffest computational cells; a \SIrange{11}{46}{$\times$} speedup was achieved over the base implicit CPU integrator during simulation of a three-dimensional premixed diesel engine simulation.
-Le et al.~\cite{Le2013596} implemented two high-order shock-capturing reactive-flow codes on a Tesla C2070, and found a \numrange{30}{50}$\times$ speedup over a sequential CPU version of the same code on a Intel Xeon X5650.
-Stone and Davis~\cite{Stone:2013aa} implemented the implicit VODE~\cite{Brown:1989vl} integrator on a Fermi M2050 GPU, achieving an order of magnitude speedup over the baseline CPU version on a single core of an AMD Opteron 6134 Magny-Cours.
-Niemeyer and Sung~\cite{Niemeyer:2014aa} later developed a stabilized explicit second-order Runge--Kutta--Chebyshev algorithm on a Tesla C2075, demonstrating an order of magnitude speedup for a CPU implementation of VODE on a six-core Intel X5650 for moderately stiff chemical kinetics.
-Sewerin and Rigopoulos~\cite{Sewerin20151375} implemented a three-stage\slash fifth-order implicit Runge--Kutta method~\cite{wanner1991solving} on both a high-end (NVIDIA Quadro 6000) and consumer-grade (NVIDIA Quadro 600) GPUs, as well as a standard CPU (two-core, four-thread Intel i5-520M) and a scientific workstation (eight-core, 16-thread Intel Xeon E5-2687W); the high-end GPU was at best \SI{1.8}{$\times$} slower than the workstation CPU (16 threads), while the consumer level GPU was at best \SI{5.5}{$\times$} slower than the corresponding standard CPU (four threads).
+Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup on a Tesla C1060 compared to a serial implementation on a AMD-Opteron processor.
+Shi et al.~\cite{Shi:2011aa} paired GPU-based chemical source-term and chemical kinetic Jacobian factorization with two implicit solvers, achieving an order of magnitude speedup for homogeneous reactor simulations of large chemical models on a Tesla 2050 GPU, as compared to a serial Intel i7 930 CPU implementation.
+Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential code on an Intel Xeon X5650 CPU.
+Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, a traditional CPU-based implicit integrator handled the stiffest computational cells while a stabilized-explicit solver on a Tesla C2050 GPU solved the less stiff cells; this hybrid solution technique was \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
+Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code on a Tesla C2070 as compared to a sequential version of the same on a Intel Xeon X5650 CPU.
+Stone and Davis~\cite{Stone:2013aa} investigated a GPU-version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order of magnitude speedup over a serial CPU implementation.
+Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics; an order of magnitude was achieved on a Tesla C2075 GPU over a multithreaded VODE solver on a six-core Intel X5650 CPU.
+Sewerin and Rigopoulos~\cite{Sewerin20151375} studied an fifth-order implicit Runge-Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU was at best \SI{1.8}{$\times$} slower than the corresponding CPU.
 Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs) on an Intel Xeon E5-2680 CPU and a Tesla K20 GPU.

### 2018-03-22 paper.tex
preprint

@@
 Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup on a Tesla C1060 compared to a serial implementation on a AMD-Opteron processor.
-Shi et al.~\cite{Shi:2011aa} paired GPU-based chemical source-term and chemical kinetic Jacobian factorization with two implicit solvers, achieving an order of magnitude speedup for homogeneous reactor simulations of large chemical models on a Tesla 2050 GPU, as compared to a serial Intel i7 930 CPU implementation.
+Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order of magnitude speedup for homogeneous reactor simulations of large chemical models on a Tesla 2050 GPU over a serial Intel i7 930 CPU implementation.
 Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential code on an Intel Xeon X5650 CPU.
 Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, a traditional CPU-based implicit integrator handled the stiffest computational cells while a stabilized-explicit solver on a Tesla C2050 GPU solved the less stiff cells; this hybrid solution technique was \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
-Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code on a Tesla C2070 as compared to a sequential version of the same on a Intel Xeon X5650 CPU.
+Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code on a Tesla C2070 as compared to a sequential version of the same on an Intel Xeon X5650 CPU.
 Stone and Davis~\cite{Stone:2013aa} investigated a GPU-version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order of magnitude speedup over a serial CPU implementation.
-Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics; an order of magnitude was achieved on a Tesla C2075 GPU over a multithreaded VODE solver on a six-core Intel X5650 CPU.
-Sewerin and Rigopoulos~\cite{Sewerin20151375} studied an fifth-order implicit Runge-Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU was at best \SI{1.8}{$\times$} slower than the corresponding CPU.
+Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup on a Tesla C2075 GPU over a multithreaded VODE solver on a six-core Intel X5650 CPU.
+Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge-Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
 Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs) on an Intel Xeon E5-2680 CPU and a Tesla K20 GPU.

### 2018-05-18 paper.tex
first pass at response to reviewers

@@
 \usepackage{textcomp}
+\usepackage{ulem}
+
+\usepackage{xcolor}
+\definecolor{darkgreen}{rgb}{0,0.5,0}
+
+% \usepackage[xcolor]{changebar}
+% \newcommand{\add}[1]{{\sloppy\cbcolor{teal}\textcolor{teal}{\cbstart {#1}\cbend}}}  % add
+% \newcommand{\delete}[1]{\sloppy\cbcolor{red}\textcolor{red}{\cbdelete\sout{#1}}}
+
+\newcommand{\add}[1]{{\sloppy\textcolor{teal}{#1}}}  % add
+\newcommand{\delete}[1]{\sloppy\textcolor{red}{\sout{#1}}} %delete with no strikeout
 
@@
 Recognizing the need to accelerate Jacobian evaluation and factorization, a number of recent works have been published on analytical chemical kinetic Jacobian evaluation; although as will be discussed at the end of this section, this work offers several key improvements over past efforts.
+\add{Schwer et al.~\cite{SCHWER2002270} were among the first to recognize the critical importance of a sparse analytical Jacobian to accelerate chemical kinetic simulations.}
 The \texttt{TChem} software package~\cite{Safta:2011vn} was one of the first codes developed providing analytical Jacobian evaluation, but has several limitations including: incompatibilities with modern reaction types---i.e., pressure dependent Arrhenius (or P-Log) and Chebyshev reactions---and lack of thread-safety to enable parallel execution~\cite{Curtis2017:tchem}.
@@
 Dijkmans et al.~\cite{Dijkmans:2014bb} developed a GPU-based analytical Jacobian code with optional tabulation of temperature-dependent properties, and found speedups up to \SI{120}{$\times$} for zero-dimensional chemical kinetic integration with large chemical models (\textasciitilde\num{3000} species) using a Tesla C2075 GPU.
+\add{Bauer et al.~\cite{Bauer:2014} used warp-specialization to improve GPU-vectorization over a standard data-parallel vectorization approach; speedups up to \SIrange{2.81}{3.75}{$\times$}, \SIrange{1.91}{2.58}{$\times$}, and \SIrange{1.4}{1.5}{$\times$} were achieved on a Telsa K20c GPU for evaluation of viscosity, species diffusion and chemical source terms, respectively}.
 Niemeyer et al.~\cite{Niemeyer:2016aa} created and validated the open-source analytical chemical kinetic Jacobian code-generator, \texttt{pyJac}, supporting parallel execution on the CPU and SIMT execution on the GPU; a speedup of \SIrange{3}{7.5}{$\times$} over a finite-difference Jacobian was found on the CPU.
@@
 \end{itemize}
-To our knowledge, this is the first effort that enables vectorized chemical-kinetic source-term and Jacobian evaluation for any chemical model on a wide-selection of platforms, i.e., the CPU, GPU and other accelerators.
+To our knowledge, this is the first open-source, validated effort that enables vectorized chemical-kinetic source-term and Jacobian evaluation for any chemical model on a wide-selection of platforms, i.e., the CPU, GPU and other accelerators.
 
@@
   \end{minipage}
+  \caption{Simple data-layout patterns for 2-D arrays}
 \end{figure}
@@
 In addition, a deep vectorization requires both synchronization between SIMD-lanes\slash SIMT-threads via memory-fences\slash barriers (an expensive operation), and may result in SIMD-waste\slash SIMT thread-divergence, caused by different SIMD-lanes\slash SIMT-threads executing different instructions (e.g., resulting from different if\slash then branches).
-It is noted that generally speaking, shallow-vectorizations may also experience SIMD-waste\slash SIMT thread-divergence, e.g., in chemical kinetic integration due to varying internal solver time-step sizes~\cite{CurtisGPU:2017}; however, in this work shallow-vectorizations are largely unaffected by this concern.
+It is noted that generally speaking, shallow-vectorizations may also experience SIMD-waste\slash SIMT thread-divergence, e.g., in chemical kinetic integration due to varying internal solver time-step sizes~\cite{CurtisGPU:2017}; however, in this work shallow-vectorizations are largely unaffected by this concern \add{as the only major code-paths that differ between vector-lanes are high\slash low-temperature polynomial evaluations and differing pressures for P-Log reactions, a relatively small concern as compared to differing internal ODE integration time-steps~\cite{CurtisGPU:2017}}.
 
@@
 However, the current \texttt{pyJac} allows the user to simply ignore these derivatives (via a command-line switch) and avoid any adverse effects on the Jacobian sparsity.
-The rational behind this choice is that many common implicit integration techniques (e.g., CVODE~\cite{Hindmarsh:2005}) used in solution of chemical kinetic initial value problems are formulated around the assumption that the supplied Jacobian is approximate; this allows the Jacobian and its LU factorization to be reused for multiple internal integration time-steps, accelerating the solution process.
+The rationale behind this choice is that many common implicit integration techniques (e.g., CVODE~\cite{Hindmarsh:2005}) used in solution of chemical kinetic initial value problems are formulated around the assumption that the supplied Jacobian is approximate; this allows the Jacobian and its LU factorization to be reused for multiple internal integration time-steps, accelerating the solution process.
 For such solvers, there is no need for the exact form of the Jacobian and the so-called ``approximate'' form is preferable.
+\add{Though this might be used as a crude form of preconditioning for such solvers, the primary purpose is merely to increase the resulting sparsity of the Jacobian; the reader is directed to~\cite{MCNENLY2015581} for a more thorough discussion of pre-conditioners.}
 Hence, in this section we will detail the sparsity of both forms of the Jacobian for the chemical models tested.
@@
 In contrast,~\cref{F:pocl_source} shows the mean run time per condition of deep, shallow and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
-No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP case is faster on both CPUs.
-The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work; however, briefly speaking vectorization might be achieved on POCL via a combination of fixed-array strides---rather than (potentially) varying with the number of initial conditions being evaluated, as currently is implemented---use of relaxed math (``\texttt{-cl-fast-relaxed-math}''), explicit OpenCL vector types and newer versions of LLVM\slash POCL (specifically, potential improvements to the vectorization module).
-This will be explored more in future work, but for the moment POCL is still quite useful as a validation tool.
+No speedup is achieved for either vectorization type on either CPU---\add{though we stress that Intel OpenCL runtime achieved vectorization using the exact same code and hardware}---indeed, the OpenMP case is faster on both CPUs.
+The reasons for this lack of vectorization are quite technical,~\delete{and somewhat outside the scope of this work}however \add{personal communications~\cite{pocl_communication} with the developers of POCL revealed that, briefly speaking, substantial changes are required to the way POCL prepares LLVM intermediate-representation code (specifically, to ensure LLVM properly recognizes uniform vectorization loop-bounds), use of relaxed math (i.e., ``\texttt{-cl-fast-relaxed-math}'') to allow LLVM's loop-vectorizer to prove a vector instruction is identical to the corresponding scalar version, LLVM's handling of branches (potentially within POCL instructions themselves) and memory access patterns in order to achieve vectorization.}
+As will be discussed in~\cref{s:future}, \add{it is hoped that use of explicit vector-types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a validation tool.
 

### 2018-05-21 paper.tex
mark add / delete

@@
 \subsection{Previous works and goals of this study}
+\label{S:Goals}
 
@@
 \end{itemize}
-To our knowledge, this is the first open-source, validated effort that enables vectorized chemical-kinetic source-term and Jacobian evaluation for any chemical model on a wide-selection of platforms, i.e., the CPU, GPU and other accelerators.
+To our knowledge, this is the first \add{open-source, validated} effort that enables vectorized chemical-kinetic source-term and Jacobian evaluation for any chemical model on a wide-selection of platforms, i.e., the CPU, GPU and other accelerators.
 
@@
   \end{minipage}
-  \caption{Simple data-layout patterns for 2-D arrays}
+  \caption{\add{Simple data-layout patterns for 2-D arrays}}
 \end{figure}

### 2018-05-21 paper.tex
move to derivations

@@
 
-The derivative of the temperature source-term with respect to itself is:
-\loadeq{dTdotdT}
-with respect to the thermodynamic state-parameter source-term (the volume \slash pressure for \conp/\slash\conv/, respectively) it's derivative:
-\loadeq{dTdotdE}
-The derivative for the moles of species $j$:
-\loadeq{dTdotdnj}
-
-The derivative of the thermodynamic state parameter with respect to the temperature is:
-\loadeq{dEdotdT}
-and with respect to itself:
-\loadeq{dEdotdE}
-where $[C]$ is the total molar concentration:
-\begin{equation}
- [C] = \frac{P}{\mathcal{R} T}\;.
-\end{equation}
-Finally, the derivative with respect to the moles of species $j$:
-\loadeq{dEdotdnj}
-
-The molar source-terms are more complicated due to the derivatives of the net rate of progress---see~\cref{deriv-s:dropi}---and pressure modification terms---\cref{deriv-s:dci}---of the reactions.
-The reader is again encouraged to read the supplemental material covering the exact derivations and equations solved by \texttt{pyJac}.
-The derivative of the molar source-term with respect to temperature is:
-\loadeq{dnkdotdT}
-while, similarly the derivative with respect to the state parameter is:
-\loadeq{dnkdotdE}
-and the moles of species $j$:
-\loadeq{dnkdotdnj}
+\delete{The derivative of the temperature source-term with respect to itself is:\ldots}
+\add{The complete derivation of the Jacobian for \texttt{pyJac} is attached as supplemental material for the interested reader.}
 

### 2018-05-21 paper.tex
reformulate conclusions to highlight goals

@@
 
-\section{Future directions and practical notes on OpenCL usage}
-\label{s:future}
-It is important to note that while OpenCL provides an simple interface to enable cross-platform execution, and significant speedups were achieved via shallow-vectorized OpenCL code in this work, there are some serious potential pit-falls in its use.
-The closed source OpenCL runtimes tested in this work (Intel and NVIDIA) contain several bugs that result in compilation failures, simply incorrect vectorized machine code or even segmentation faults.
-Further, these runtimes (in our experience) tend to be less responsive to fixing said bugs, with relatively infrequent new releases (or in NVIDIA's case, lack of changelogs\slash public records of bug-fixes).
-On the other hand, the open-source OpenCL runtime used in this work, POCL, has (in our experience) far fewer implementation bugs, and when issues arise the community is very responsive to bug reports and user outreach; however, POCL fails to achieve vectorization as noted in~\cref{S:source_results}.
-
-As demonstration of the type of issue discussed, a minimum working example has been created~\cite{nvidia_mwe} that demonstrates a failure of NVIDIA's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla \gpunew/ GPU; by simply changing to another runtime (e.g., Intel)---with no code changes or recompilation---the correct result can be found.
-This provides a particularly vexing problem for the programmer, as there is often little that can be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
-Further, as noted throughout this work certain code-generation patterns can break OpenCL execution\slash vectorization, e.g., the failure of POCL to achieve vectorization or Intel OpenCL's failure to vectorize the finite-difference Jacobian.
-Indeed the vectorization process attempted by most OpenCL runtimes (and thus, reasons for incorrect\slash unvectorized code-output) is obscured from the user, making reasoning about errors or performance trends quite difficult.
-
-Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change requires building of the wrapper that initializes\slash transfers memory and calls the source-rate\slash Jacobian kernel.
-Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py}) is a key priority going forwards as OpenMP is a standard library on most machines.
-In addition, CUDA~\cite{NVIDIA:2018} has been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017}, while Intel's open-source OpenCL alternative, ISPC~\cite{pharr2012ispc}, has been relatively stable and easy to work with during preliminary efforts with the unit-testing discussed in~\cref{s:unittest}.
-The current deep-vectorization formulation would be executable for both CUDA and ISPC targets, as these languages implement double-precision atomic operations, further recommending their use.
-It is also possible, particularly for the Intel OpenCL\slash POCL runtimes, that better performance\slash stability might be achieved using so-called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}; this may also enable deep-vectorization on these runtimes.
-Extending execution to new targets and improvements to the existing OpenCL targets is a future goal to improve performance, reliability and ease of use for CPU\slash GPU vectorizated execution.
-
-Finally, one key-component missing in this current version of \texttt{pyJac} is vectorized sparse\slash dense linear-algebra subroutines to maximize the performance of LU-factorization and matrix-vector multiplication (commonly used in implicit-integration techniques).
-Third-party\slash open-source options exist for some target-languages, e.g., cuBLAS~\cite{cublas}, clBLAS\slash clSPARSE~\cite{clmath} or SuperLU~\cite{superlu99}, but these do not necessarily cover all targets\slash required linear-algebra operations and, in the case of the CUDA\slash OpenCL libraries, are often optimized operation on one large matrix instead of many (relatively) smaller matrices.
-The extent to which these existing programs may be used needs to be assessed and missing operations should be implemented in \texttt{loo.py} to ensure easy switching between target-languages\slash vectorization types.
-
 \section{Conclusions}
 In this work, automatically generated OpenCL codes for SIMD\slash SIMT-vectorized thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation were developed and validated for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
+\add{Reviewing the goals of this work, as discussed in~\cref{S:Goals}}:
+
+\subsection{\add{Derive and validate a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity}}
 A new formulation resulted in highly sparse (\SIrange{28.6}{88.5}{\percent}) chemical kinetic Jacobians, with options to futher increase sparsity (to \SIrange{31.66}{92.0}{\percent}) at the expense of strict correctness.
+
+\subsection{\add{Enable cross-platform SIMD\slash SIMT vectorization for the CPU, GPU and other accelerators}}
+\add{Shallow-vectorized evaluation of \texttt{pyJac}'s chemical kinetic source-term and analytical Jacobian was demonstrated on both the CPU and GPU via OpenCL.}
+Deep-vectorization, though possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, did not yield any performance benefit as POCL did not achieve vectorizations for any execution pattern studied; deep-vectorization deserves further study with other platforms (e.g., CUDA).
+\add{In addition, parallel OpenMP evaluation was demonstrated on the CPU.}
+
+\subsection{\add{Investigate the performance for a wide range chemical kinetic models, specifically looking at speedups due to SIMD\slash SIMT-vectorization as well as relative to the previous version of \texttt{pyJac}}}
+
 Significant speedups in shallow SIMD-vectorized source-term evaluation over a parallel OpenMP code were demonstrated, up to \SIrange{2.53}{2.92}{$\times$} and \SIrange{3.40}{4.08}{$\times$} for \sse/ and \avx/ capable CPUs; in addition, source-term evaluation was \SIrange{1.40}{1.88}{$\times$} faster on a \gpunew/ GPU compared to a \gpuold/ GPU.
-Deep-vectorization, though possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, did not yield any performance benefit as POCL did not achieve vectorizations for any execution pattern studied; deep-vectorization deserves further study with other platforms (e.g., CUDA).
 Vectorized data-orderings specifically formulated for shallow\slash deep SIMD\slash SIMT vectorizations were proposed, resulting in a shallow-vectorized speedup of \SIrange{2.14}{2.58}{$\times$} over a more standard ``F''-ordered (column major) shallow-vectorization; when compared to the corresponding speedup of \SIrange{1.35}{2.09}{$\times$} for the ``C''-ordered (row major) OpenMP code over an ``F''-ordered equivalent, the value of the vectorized data-ordering is evident.
@@
 
-Future extensions of this work will include additional target-languages (vectorized OpenMP, CUDA, ISPC, etc.) to improve ease of use and reliability, as well as to enable meaningful deep-vectorization.
-Additionally, existing vectorized sparse\slash dense-matrix libraries will be assessed and missing operations for LU-factorization\slash matrix-vector multiplication should be implemented.
+\subsection{\add{Future extensions to this work and practical notes on OpenCL usage}}
+
+It is important to note that while OpenCL provides an simple interface to enable cross-platform execution, and significant speedups were achieved via shallow-vectorized OpenCL code in this work, there are some serious potential pit-falls in its use.
+The closed source OpenCL runtimes tested in this work (Intel and NVIDIA) contain several bugs that result in compilation failures, simply incorrect vectorized machine code or even segmentation faults.
+Further, these runtimes (in our experience) tend to be less responsive to fixing said bugs, with relatively infrequent new releases (or in NVIDIA's case, lack of changelogs\slash public records of bug-fixes).
+On the other hand, the open-source OpenCL runtime used in this work, POCL, has (in our experience) far fewer implementation bugs, and when issues arise the community is very responsive to bug reports and user outreach; however, POCL fails to achieve vectorization as noted in~\cref{S:source_results}.
+
+As demonstration of the type of issue discussed, a minimum working example has been created~\cite{nvidia_mwe} that demonstrates a failure of NVIDIA's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla \gpunew/ GPU; by simply changing to another runtime (e.g., Intel)---with no code changes or recompilation---the correct result can be found.
+This provides a particularly vexing problem for the programmer, as there is often little that can be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
+Further, as noted throughout this work certain code-generation patterns can break OpenCL execution\slash vectorization, e.g., the failure of POCL to achieve vectorization or Intel OpenCL's failure to vectorize the finite-difference Jacobian.
+Indeed the vectorization process attempted by most OpenCL runtimes (and thus, reasons for incorrect\slash unvectorized code-output) is obscured from the user, making reasoning about errors or performance trends quite difficult.
+
+Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change requires building of the wrapper that initializes\slash transfers memory and calls the source-rate\slash Jacobian kernel.
+Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler \texttt{\#pragma}s) is a key priority going forwards as OpenMP is a standard library on most machines.
+In addition, CUDA~\cite{NVIDIA:2018} has been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017}, while Intel's open-source OpenCL alternative, ISPC~\cite{pharr2012ispc}, has been relatively stable and easy to work with during preliminary efforts with the unit-testing discussed in~\cref{s:unittest}.
+The current deep-vectorization formulation would be executable for both CUDA and ISPC targets, as these languages implement double-precision atomic operations, further recommending their use.
+It is also possible, particularly for the Intel OpenCL\slash POCL runtimes, that better performance\slash stability might be achieved using so-called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}; \add{specifically this change is hoped to enable vectorization on the POCL runtime} and might also enable deep-vectorization on the Intel OpenCL runtime.
+\add{Finally, future extensions of this work will include additional target-languages (vectorized OpenMP, CUDA, ISPC, etc.) to improve ease of use and reliability, as well as improvements to the existing OpenCL targets, e.g., to enable meaningful deep-vectorization.}
+
+One key-component missing in this current version of \texttt{pyJac} is vectorized sparse\slash dense linear-algebra subroutines to maximize the performance of LU-factorization and matrix-vector multiplication (commonly used in implicit-integration techniques).
+Third-party\slash open-source options exist for some target-languages, e.g., cuBLAS~\cite{cublas}, clBLAS\slash clSPARSE~\cite{clmath} or SuperLU~\cite{superlu99}, but these do not necessarily cover all targets\slash required linear-algebra operations and, in the case of the CUDA\slash OpenCL libraries, are often optimized operation on one large matrix instead of many (relatively) smaller matrices.
+The extent to which these existing programs may be used needs to be assessed and missing operations should be implemented in \texttt{loo.py} to ensure easy switching between target-languages\slash vectorization types.
 

### 2018-05-21 paper.tex
add sutherland refs

@@
 Niemeyer et al.~\cite{Niemeyer:2016aa} created and validated the open-source analytical chemical kinetic Jacobian code-generator, \texttt{pyJac}, supporting parallel execution on the CPU and SIMT execution on the GPU; a speedup of \SIrange{3}{7.5}{$\times$} over a finite-difference Jacobian was found on the CPU.
-Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but it was not validated outside the context of use with an implicit-integration technique; in addition, since the Jacobian was based on an over-constrained system, the effect on strict conservation of mass\slash energy was not studied.
+Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but it was not validated outside the context of use with an implicit-integration technique; in addition, since the Jacobian was based on an over-constrained system~\cite{HANSEN2018257}, the effect on strict conservation of mass\slash energy was not studied.
+\add{Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermochemical state-vectors and their resulting effect on consistency and errors in conserved properties such as mass and energy.}
 
@@
 First, the state vector results in highly sparse chemical kinetic Jacobians as will be detailed in~\cref{S:sparsity}.
-Second, mass is explicitly conserved in this formulation owing to the fact that the moles of the final species is calculated from the ideal gas law, and the rate of change of the final species from conservation of mass (see \cref{deriv-s:state,deriv-s:source} of the supplemental material, respectively); additionally, system is not over-constrained and does not require use of a more complicated differential algebraic equation solver (as compared to an ODE integrator) for integration.
+Second, mass is explicitly conserved in this formulation owing to the fact that the moles of the final species is calculated from the ideal gas law, and the rate of change of the final species from conservation of mass (see \cref{deriv-s:state,deriv-s:source} of the supplemental material, respectively); additionally, system is not over-constrained~\cite{HANSEN2018257} and does not require use of a more complicated differential algebraic equation solver (as compared to an ODE integrator) for integration.
 Finally, the chemical kinetic Jacobian changes relatively little between the \conp/ and \conv/ forms, making maintaining the code-base much simpler.

### 2018-05-21 paper.tex
Add bit about implicit / linearly implicit & cema

@@
 Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but it was not validated outside the context of use with an implicit-integration technique; in addition, since the Jacobian was based on an over-constrained system~\cite{HANSEN2018257}, the effect on strict conservation of mass\slash energy was not studied.
-\add{Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermochemical state-vectors and their resulting effect on consistency and errors in conserved properties such as mass and energy.}
+\add{Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermochemical state-vectors and their resulting effect on consistency and errors in conserved properties such as mass and energy, and the effect on implicit and linearly-implicit integration techniques an analysis techniques such as chemical explosive mode analysis~\cite{lu_yoo_chen_law_2010}.}
 

### 2018-05-21 paper.tex
cleanup

--- 2018-05-21 paper.tex
@@
 Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but it was not validated outside the context of use with an implicit-integration technique; in addition, since the Jacobian was based on an over-constrained system~\cite{HANSEN2018257}, the effect on strict conservation of mass\slash energy was not studied.
-\add{Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermochemical state-vectors and their resulting effect on consistency and errors in conserved properties such as mass and energy.}
+\add{Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermochemical state-vectors and their resulting effect on consistency and errors in conserved properties such as mass and energy, and the effect on implicit\slash linearly-implicit integration techniques, as well as analysis techniques such as chemical explosive mode analysis~\cite{lu_yoo_chen_law_2010}.}
 

### 2018-05-21 paper.tex
remove specifics of hardward in litreview

@@
 Dijkmans et al.~\cite{Dijkmans:2014bb} developed a GPU-based analytical Jacobian code with optional tabulation of temperature-dependent properties, and found speedups up to \SI{120}{$\times$} for zero-dimensional chemical kinetic integration with large chemical models (\textasciitilde\num{3000} species) using a Tesla C2075 GPU.
-\add{Bauer et al.~\cite{Bauer:2014} used warp-specialization to improve GPU-vectorization over a standard data-parallel vectorization approach; speedups up to \SIrange{2.81}{3.75}{$\times$}, \SIrange{1.91}{2.58}{$\times$}, and \SIrange{1.4}{1.5}{$\times$} were achieved on a Telsa K20c GPU for evaluation of viscosity, species diffusion and chemical source terms, respectively}.
+\add{Bauer et al.~\cite{Bauer:2014} used warp-specialization to improve GPU-vectorization over a standard data-parallel vectorization approach; speedups up to \SIrange{2.81}{3.75}{$\times$}, \SIrange{1.91}{2.58}{$\times$}, and \SIrange{1.4}{1.5}{$\times$} were achieved on a Telsa K20c GPU\todo{keep hardware spec?} for evaluation of viscosity, species diffusion and chemical source terms, respectively}.
 Niemeyer et al.~\cite{Niemeyer:2016aa} created and validated the open-source analytical chemical kinetic Jacobian code-generator, \texttt{pyJac}, supporting parallel execution on the CPU and SIMT execution on the GPU; a speedup of \SIrange{3}{7.5}{$\times$} over a finite-difference Jacobian was found on the CPU.
@@
 
+\todo[inline]{I don't know that I actually want to remove the processor information alone here -- it doesn't save us much space.
+It would be more valuable to re-write this entirely, but I'm not sure that's necessary either}
 A number of recent works have investigated the use of high-performance SIMT-devices to accelerate reactive-flow and chemical kinetics simulations.
-Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup on a Tesla C1060 compared to a serial implementation on a AMD-Opteron processor.
-Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order of magnitude speedup for homogeneous reactor simulations of large chemical models on a Tesla 2050 GPU over a serial Intel i7 930 CPU implementation.
-Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential code on an Intel Xeon X5650 CPU.
-Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, a traditional CPU-based implicit integrator handled the stiffest computational cells while a stabilized-explicit solver on a Tesla C2050 GPU solved the less stiff cells; this hybrid solution technique was \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
-Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code on a Tesla C2070 as compared to a sequential version of the same on an Intel Xeon X5650 CPU.
+Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup \delete{on a Tesla C1060} compared to a serial implementation\add{ CPU}\delete{on a AMD-Opteron processor}.
+Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order of magnitude speedup for homogeneous reactor simulations of large chemical models\delete{on a Tesla 2050 GPU} over a serial\delete{Intel i7 930} CPU implementation.
+Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential \add{CPU} code\delete{ on an Intel Xeon X5650 CPU}.
+Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, a traditional CPU-based implicit integrator handled the stiffest computational cells while a stabilized-explicit GPU-solver\delete{on a Tesla C2050 GPU} solved the less stiff cells; this hybrid solution technique was \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
+Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code\delete{ on a Tesla C2070} as compared to a sequential\add{ CPU} version of the same\delete{ on an Intel Xeon X5650 CPU}.
 Stone and Davis~\cite{Stone:2013aa} investigated a GPU-version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order of magnitude speedup over a serial CPU implementation.
-Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup on a Tesla C2075 GPU over a multithreaded VODE solver on a six-core Intel X5650 CPU.
+Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup\delete{on a Tesla C2075 GPU} over a multithreaded VODE solver on a six-core\delete{Intel X5650} CPU.
 Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge-Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
-Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs) on an Intel Xeon E5-2680 CPU and a Tesla K20 GPU.
+Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs)\delete{ on an Intel Xeon E5-2680}\add{both the CPU} and\delete{a Tesla K20} GPU.
 In evaluation of the thermo-chemical properties, speedups between \SIrange{8}{13}{$\times$} were found (over a serial version of the same code) while running on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU, respectively.
-Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on a Tesla C2075; the implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on 12--38 Intel Xeon E5-4640 v2 CPU cores for two relatively small chemical models with an integration time step of \SI{e-6}{$\sec$}.
+Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa}\delete{ on a Tesla C2075}\add{ on the GPU and CPU}; the\add{ GPU-based} implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38}\delete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models with an integration time step of \SI{e-6}{$\sec$}.
 

### 2018-05-21 paper.tex
fixes

@@
 A number of recent works have investigated the use of high-performance SIMT-devices to accelerate reactive-flow and chemical kinetics simulations.
-Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup \delete{on a Tesla C1060} compared to a serial implementation\add{ CPU}\delete{on a AMD-Opteron processor}.
+Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup \delete{on a Tesla C1060} compared to a serial implementation \add{CPU}\delete{on a AMD-Opteron processor}.
 Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order of magnitude speedup for homogeneous reactor simulations of large chemical models\delete{on a Tesla 2050 GPU} over a serial\delete{Intel i7 930} CPU implementation.
-Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential \add{CPU} code\delete{ on an Intel Xeon X5650 CPU}.
+Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential \add{CPU} code\delete{on an Intel Xeon X5650 CPU}.
 Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, a traditional CPU-based implicit integrator handled the stiffest computational cells while a stabilized-explicit GPU-solver\delete{on a Tesla C2050 GPU} solved the less stiff cells; this hybrid solution technique was \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
-Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code\delete{ on a Tesla C2070} as compared to a sequential\add{ CPU} version of the same\delete{ on an Intel Xeon X5650 CPU}.
+Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code\delete{ on a Tesla C2070} as compared to a sequential \add{CPU} version of the same\delete{on an Intel Xeon X5650 CPU}.
 Stone and Davis~\cite{Stone:2013aa} investigated a GPU-version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order of magnitude speedup over a serial CPU implementation.
@@
 Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge-Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
-Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs)\delete{ on an Intel Xeon E5-2680}\add{both the CPU} and\delete{a Tesla K20} GPU.
+Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs)\delete{on an Intel Xeon E5-2680} \add{both the CPU} and\delete{a Tesla K20} GPU.
 In evaluation of the thermo-chemical properties, speedups between \SIrange{8}{13}{$\times$} were found (over a serial version of the same code) while running on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU, respectively.
-Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa}\delete{ on a Tesla C2075}\add{ on the GPU and CPU}; the\add{ GPU-based} implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38}\delete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models with an integration time step of \SI{e-6}{$\sec$}.
+Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa}\delete{on a Tesla C2075} \add{on the GPU and CPU}; the \add{GPU-based} implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38}\delete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models with an integration time step of \SI{e-6}{$\sec$}.
 
@@
 Thus POCL is a particularly useful tool for validation, but not necessarily for performance studies.
-This discussion will be expanded upon in~\cref{s:future} to highlight future directions.
+This discussion will be expanded upon in~\cref{S:future} to highlight future directions.
 

### 2018-05-21 paper.tex
reword

@@
 Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but it was not validated outside the context of use with an implicit-integration technique; in addition, since the Jacobian was based on an over-constrained system~\cite{HANSEN2018257}, the effect on strict conservation of mass\slash energy was not studied.
-\add{Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermochemical state-vectors and their resulting effect on consistency and errors in conserved properties such as mass and energy, and the effect on implicit\slash linearly-implicit integration techniques, as well as analysis techniques such as chemical explosive mode analysis~\cite{lu_yoo_chen_law_2010}.}
+\add{Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermochemical state-vectors and their resulting effect on consistency and errors in conserved properties such as mass and energy; futher, the effect on implicit\slash linearly-implicit integration techniques and analysis techniques such as chemical explosive mode analysis~\cite{lu_yoo_chen_law_2010} were characterized.}
 

### 2018-05-21 paper.tex
spelling

--- 2018-05-21 paper.tex
@@
 Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but it was not validated outside the context of use with an implicit-integration technique; in addition, since the Jacobian was based on an over-constrained system~\cite{HANSEN2018257}, the effect on strict conservation of mass\slash energy was not studied.
-\add{Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermochemical state-vectors and their resulting effect on consistency and errors in conserved properties such as mass and energy, and the effect on implicit\slash linearly-implicit integration techniques, as well as analysis techniques such as chemical explosive mode analysis~\cite{lu_yoo_chen_law_2010}.}
+\add{Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermo-chemical state-vectors and their resulting effect on consistency and errors in conserved properties such as mass and energy; further, the effect on implicit\slash linearly-implicit integration techniques and analysis techniques such as chemical explosive mode analysis~\cite{lu_yoo_chen_law_2010} were characterized.}
 

### 2018-05-21 paper.tex
fixes

@@
 For such solvers, there is no need for the exact form of the Jacobian and the so-called ``approximate'' form is preferable.
-\add{Though this might be used as a crude form of preconditioning for such solvers, the primary purpose is merely to increase the resulting sparsity of the Jacobian; the reader is directed to~\cite{MCNENLY2015581} for a more thorough discussion of pre-conditioners.}
+\add{Though this might be used as a crude form of preconditioning for such solvers, the primary purpose is merely to increase the resulting sparsity of the Jacobian; the reader is directed to~\cite{MCNENLY2015581} for a more thorough discussion of preconditioners.}
 Hence, in this section we will detail the sparsity of both forms of the Jacobian for the chemical models tested.
@@
 The reasons for this lack of vectorization are quite technical,~\delete{and somewhat outside the scope of this work}however \add{personal communications~\cite{pocl_communication} with the developers of POCL revealed that, briefly speaking, substantial changes are required to the way POCL prepares LLVM intermediate-representation code (specifically, to ensure LLVM properly recognizes uniform vectorization loop-bounds), use of relaxed math (i.e., ``\texttt{-cl-fast-relaxed-math}'') to allow LLVM's loop-vectorizer to prove a vector instruction is identical to the corresponding scalar version, LLVM's handling of branches (potentially within POCL instructions themselves) and memory access patterns in order to achieve vectorization.}
-As will be discussed in~\cref{s:future}, \add{it is hoped that use of explicit vector-types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a validation tool.
+As will be discussed in~\cref{S:future}, \add{it is hoped that use of explicit vector-types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a validation tool.
 
@@
 \subsection{\add{Future extensions to this work and practical notes on OpenCL usage}}
+\label{S:future}
 

### 2018-05-22 scripts/minimum_spanning_set.py
add utility to reduce # of reactions in mechanism for scaling test

@@ -1,0 +1,164 @@
+import os
+
+import cantera as ct
+import numpy as np
+from scipy.optimize import curve_fit
+import matplotlib.pyplot as plt
+
+file = os.path.abspath(os.path.dirname(__file__))
+gas = ct.Solution(os.path.join(file, os.pardir, 'mech_data',
+                               'Sarathy_ic5_mech_rev.cti'))
+
+
+def arrhenius(T, A, b, Ea, print_stats=False):
+    vals = A * np.power(T, b) * np.exp(-Ea / (ct.gas_constant * T))
+    if print_stats:
+        print(A, b, Ea, np.max(np.abs(vals - kf) / kf))
+    return vals
+
+
+# first, convert all reactions to a single rate form
+reacs = gas.reactions()[:]
+for i, reac in enumerate(reacs):
+    if isinstance(reac, ct.ChemicallyActivatedReaction):
+        reacs[i] = ct.ElementaryReaction(reac.reactants, reac.products)
+        reacs[i].rate = reac.high_rate
+    elif isinstance(reac, ct.FalloffReaction):
+        reacs[i] = ct.ElementaryReaction(reac.reactants, reac.products)
+        reacs[i].rate = reac.high_rate
+    elif isinstance(reac, ct.ThreeBodyReaction):
+        reacs[i] = ct.ElementaryReaction(reac.reactants, reac.products)
+        reacs[i].rate = reac.rate
+    elif isinstance(reac, ct.PlogReaction):
+        reacs[i] = ct.ElementaryReaction(reac.reactants, reac.products)
+        reacs[i].rate = reac.rates[-1][1]
+    elif isinstance(reac, ct.ChebyshevReaction):
+        # fit _something_ to it at a 10 bar, for 600-2200K
+        T = np.linspace(600, 2200, num=1000)
+        kf = np.zeros_like(T)
+        for i in range(T.size):
+            kf[i] = reac(T[i], 10 * 10*1e5)
+        A = 1e10
+        b = 1
+        Ea = 1200 * ct.gas_constant
+        (A, b, Ea), _ = curve_fit(
+            arrhenius, T, kf,
+            p0=[A, b, Ea],
+            bounds=[(0, -2, 0), (np.inf, 5, np.inf)],
+            maxfev=1e6,
+            xtol=1e-15)
+        if False:
+            plt.plot(T, kf, color='b', linestyle='-', label='cheb')
[…]

### 2018-05-23 scripts/minimum_spanning_set.py
first pass at a run

@@ -1,164 +1,0 @@
-import os
-
-import cantera as ct
-import numpy as np
-from scipy.optimize import curve_fit
-import matplotlib.pyplot as plt
-
-file = os.path.abspath(os.path.dirname(__file__))
-gas = ct.Solution(os.path.join(file, os.pardir, 'mech_data',
-                               'Sarathy_ic5_mech_rev.cti'))
-
-
-def arrhenius(T, A, b, Ea, print_stats=False):
-    vals = A * np.power(T, b) * np.exp(-Ea / (ct.gas_constant * T))
-    if print_stats:
-        print(A, b, Ea, np.max(np.abs(vals - kf) / kf))
-    return vals
-
-
-# first, convert all reactions to a single rate form
-reacs = gas.reactions()[:]
-for i, reac in enumerate(reacs):
-    if isinstance(reac, ct.ChemicallyActivatedReaction):
-        reacs[i] = ct.ElementaryReaction(reac.reactants, reac.products)
-        reacs[i].rate = reac.high_rate
-    elif isinstance(reac, ct.FalloffReaction):
-        reacs[i] = ct.ElementaryReaction(reac.reactants, reac.products)
-        reacs[i].rate = reac.high_rate
-    elif isinstance(reac, ct.ThreeBodyReaction):
-        reacs[i] = ct.ElementaryReaction(reac.reactants, reac.products)
-        reacs[i].rate = reac.rate
-    elif isinstance(reac, ct.PlogReaction):
-        reacs[i] = ct.ElementaryReaction(reac.reactants, reac.products)
-        reacs[i].rate = reac.rates[-1][1]
-    elif isinstance(reac, ct.ChebyshevReaction):
-        # fit _something_ to it at a 10 bar, for 600-2200K
-        T = np.linspace(600, 2200, num=1000)
-        kf = np.zeros_like(T)
-        for i in range(T.size):
-            kf[i] = reac(T[i], 10 * 10*1e5)
-        A = 1e10
-        b = 1
-        Ea = 1200 * ct.gas_constant
-        (A, b, Ea), _ = curve_fit(
-            arrhenius, T, kf,
-            p0=[A, b, Ea],
-            bounds=[(0, -2, 0), (np.inf, 5, np.inf)],
-            maxfev=1e6,
-            xtol=1e-15)
-        if False:
-            plt.plot(T, kf, color='b', linestyle='-', label='cheb')
[…]

### 2018-05-23 scripts/reaction_scaling_test.py

@@ -1,0 +1,300 @@
+import os
+from argparse import ArgumentParser
+import subprocess
+
+import cantera as ct
+import numpy as np
+from scipy.optimize import curve_fit
+import matplotlib.pyplot as plt
+
+from pyjac import create_jacobian
+from pyjac.libgen import build_type, generate_library
+from pyjac.tests.test_utils import data_bin_writer as dbw
+
+path = os.path.abspath(os.path.dirname(__file__))
+
+
+def run(gas, interval, num_states, work_dir, repeats=10):
+    def arrhenius(T, A, b, Ea, print_stats=False):
+        vals = A * np.power(T, b) * np.exp(-Ea / (ct.gas_constant * T))
+        if print_stats:
+            print(A, b, Ea, np.max(np.abs(vals - kf) / kf))
+        return vals
+
+    # first, convert all reactions to a single rate form
+    reacs = gas.reactions()[:]
+    for i, reac in enumerate(reacs):
+        if isinstance(reac, ct.ChemicallyActivatedReaction):
+            reacs[i] = ct.ElementaryReaction(reac.reactants, reac.products)
+            reacs[i].rate = reac.high_rate
+        elif isinstance(reac, ct.FalloffReaction):
+            reacs[i] = ct.ElementaryReaction(reac.reactants, reac.products)
+            reacs[i].rate = reac.high_rate
+        elif isinstance(reac, ct.ThreeBodyReaction):
+            reacs[i] = ct.ElementaryReaction(reac.reactants, reac.products)
+            reacs[i].rate = reac.rate
+        elif isinstance(reac, ct.PlogReaction):
+            reacs[i] = ct.ElementaryReaction(reac.reactants, reac.products)
+            reacs[i].rate = reac.rates[-1][1]
+        elif isinstance(reac, ct.ChebyshevReaction):
+            # fit _something_ to it at a 10 bar, for 600-2200K
+            T = np.linspace(600, 2200, num=1000)
+            kf = np.zeros_like(T)
+            for i in range(T.size):
+                kf[i] = reac(T[i], 10 * 10*1e5)
+            A = 1e10
+            b = 1
+            Ea = 1200 * ct.gas_constant
+            (A, b, Ea), _ = curve_fit(
+                arrhenius, T, kf,
+                p0=[A, b, Ea],
+                bounds=[(0, -2, 0), (np.inf, 5, np.inf)],
+                maxfev=1e6,
+                xtol=1e-15)
+            if False:
[…]

### 2018-05-23 scripts/reaction_scaling_test.py
fixes

@@ -8,6 +8,7 @@ from scipy.optimize import curve_fit
 import matplotlib.pyplot as plt
 
 from pyjac import create_jacobian
+from pyjac.utils import create_dir
 from pyjac.libgen import build_type, generate_library
 from pyjac.tests.test_utils import data_bin_writer as dbw
 
@@ -61,18 +62,19 @@ def run(gas, interval, num_states, work_dir, repeats=10):
             reacs[i].rate = ct.Arrhenius(A, b, Ea)
 
     # next, order the reactions by the number of distinct species
-    def get_reac_and_spec_maps(gas):
+    def get_reac_and_spec_maps(mygas):
         # first, create rxn->species maps
-        reac_nu = gas.reactant_stoich_coeffs()
-        prod_nu = gas.product_stoich_coeffs()
+        reac_nu = mygas.reactant_stoich_coeffs()
+        prod_nu = mygas.product_stoich_coeffs()
         # species -> rxn mappings
-        spec_masks = np.zeros((gas.n_species, gas.n_reactions), dtype=np.bool)
+        spec_masks = np.zeros((mygas.n_species, mygas.n_reactions),
+                              dtype=np.bool)
         # the reaction -> species mappings
         reac_masks = []
-        for i, reac in enumerate(reacs):
+        for i, reac in enumerate(mygas.reactions()):
             for spec in set(list(reac.reactants.keys()) + list(
                                  reac.products.keys())):
-                j = gas.species_index(spec)
+                j = mygas.species_index(spec)
                 if prod_nu[j, i] - reac_nu[j, i]:
                     # non-zero species
                     spec_masks[j, i] = True
@@ -210,23 +212,31 @@ def run(gas, interval, num_states, work_dir, repeats=10):
                 vectype, platform, rate_spec,
                 split, num_cores, conp) + '.txt'
 
+    build = os.path.join(path, 'out')
+    obj = os.path.join(path, 'obj')
+    lib = os.path.join(path, 'lib')
+    create_dir(build)
+    create_dir(obj)
+    create_dir(lib)
+
     for wide in [True, False]:
         vsize = vecsize if wide else None
         # now, go through the various generated reactions lists and run
         # the test on each
-        build = os.path.join(path, 'out')
-        obj = os.path.join(path, 'obj')
-        lib = os.path.join(path, 'lib')
         for reac_list in saved_reaction_lists:
+            subdir = os.path.join(work_dir, str(active_reactions(reac_list)))
[…]

### 2018-05-23 scripts/reaction_scaling_test.py
fix wrong filenames

@@ -244,7 +244,7 @@ def run(gas, interval, num_states, work_dir, repeats=10):
                                       btype=build_type.species_rates,
                                       as_executable=True)
 
-            outname = '{}.txt'.format(active_reactions(reac_list))
+            outname = get_filename(wide)
             # and do runs
             with open(os.path.join(subdir, outname), 'w') as file:
                 for i in range(repeats):

### 2018-05-23 scripts/reaction_scaling_test.py
clean temporaries

@@ -11,6 +11,7 @@ from pyjac import create_jacobian
 from pyjac.utils import create_dir
 from pyjac.libgen import build_type, generate_library
 from pyjac.tests.test_utils import data_bin_writer as dbw
+from pyjac.tests.test_utils import clean_dir
 
 path = os.path.abspath(os.path.dirname(__file__))
 
@@ -215,15 +216,17 @@ def run(gas, interval, num_states, work_dir, repeats=10):
     build = os.path.join(path, 'out')
     obj = os.path.join(path, 'obj')
     lib = os.path.join(path, 'lib')
-    create_dir(build)
-    create_dir(obj)
-    create_dir(lib)
 
     for wide in [True, False]:
         vsize = vecsize if wide else None
         # now, go through the various generated reactions lists and run
         # the test on each
         for reac_list in saved_reaction_lists:
+            # clean
+            clean_dir(build, remove=False)
+            clean_dir(obj, remove=False)
+            clean_dir(lib, remove=False)
+
             subdir = os.path.join(work_dir, str(active_reactions(reac_list)))
             create_dir(subdir)
             # generate the source rate evaluation

### 2018-05-23 scripts/reaction_scaling_test.py
sp

@@ -223,9 +223,9 @@ def run(gas, interval, num_states, work_dir, repeats=10):
         # the test on each
         for reac_list in saved_reaction_lists:
             # clean
-            clean_dir(build, remove=False)
-            clean_dir(obj, remove=False)
-            clean_dir(lib, remove=False)
+            clean_dir(build, remove_dir=False)
+            clean_dir(obj, remove_dir=False)
+            clean_dir(lib, remove_dir=False)
 
             subdir = os.path.join(work_dir, str(active_reactions(reac_list)))
             create_dir(subdir)

### 2018-05-24 scripts/reaction_scaling_test.py
rudimentary restarts

@@ -213,6 +213,22 @@ def run(gas, interval, num_states, work_dir, repeats=10):
                 vectype, platform, rate_spec,
                 split, num_cores, conp) + '.txt'
 
+    def check_file(file):
+        if not os.path.exists(file):
+            return repeats
+        with open(file, 'r') as f:
+            lines = f.readlines()
+        import re
+        todo = repeats
+        for line in lines:
+            line = line.split(',')
+            if len(line) > 1 and sum(
+                    1 if re.search(r'(?:\d+(?:\.\d+e[+-]\d+))', l) else 0
+                    for l in line) == 4:
+                # four doubles -> good line
+                todo -= 1
+        return todo
+
     build = os.path.join(path, 'out')
     obj = os.path.join(path, 'obj')
     lib = os.path.join(path, 'lib')
@@ -222,6 +238,8 @@ def run(gas, interval, num_states, work_dir, repeats=10):
         # now, go through the various generated reactions lists and run
         # the test on each
         for reac_list in saved_reaction_lists:
+            outname = get_filename(wide)
+            todo = check_file(outname)
             # clean
             clean_dir(build, remove_dir=False)
             clean_dir(obj, remove_dir=False)
@@ -247,11 +265,10 @@ def run(gas, interval, num_states, work_dir, repeats=10):
                                       btype=build_type.species_rates,
                                       as_executable=True)
 
-            outname = get_filename(wide)
             # and do runs
-            with open(os.path.join(subdir, outname), 'w') as file:
-                for i in range(repeats):
-                    print(i, "/", repeats)
+            with open(os.path.join(subdir, outname), 'a+') as file:
+                for i in range(todo):
+                    print(i, "/", todo)
                     subprocess.check_call([os.path.join(lib, tester),
                                            str(num_states), str(1)],
                                           stdout=file)

### 2018-05-24 scripts/reaction_scaling_test.py
make sure the gas uses the converted reactions

@@ -61,6 +61,10 @@ def run(gas, interval, num_states, work_dir, repeats=10):
                 plt.show()
             reacs[i] = ct.ElementaryReaction(reac.reactants, reac.products)
             reacs[i].rate = ct.Arrhenius(A, b, Ea)
+    # and convert gas
+    gas = ct.Solution(thermo='IdealGas', kinetics='GasKinetics',
+                      reacs=reacs,
+                      species=gas.species())
 
     # next, order the reactions by the number of distinct species
     def get_reac_and_spec_maps(mygas):

### 2018-05-24 scripts/reaction_scaling_test.py
fixes

@@ -42,8 +42,8 @@ def run(gas, interval, num_states, work_dir, repeats=10):
             # fit _something_ to it at a 10 bar, for 600-2200K
             T = np.linspace(600, 2200, num=1000)
             kf = np.zeros_like(T)
-            for i in range(T.size):
-                kf[i] = reac(T[i], 10 * 10*1e5)
+            for j in range(T.size):
+                kf[j] = reac(T[j], 10 * 10*1e5)
             A = 1e10
             b = 1
             Ea = 1200 * ct.gas_constant
@@ -63,7 +63,7 @@ def run(gas, interval, num_states, work_dir, repeats=10):
             reacs[i].rate = ct.Arrhenius(A, b, Ea)
     # and convert gas
     gas = ct.Solution(thermo='IdealGas', kinetics='GasKinetics',
-                      reacs=reacs,
+                      reactions=reacs,
                       species=gas.species())
 
     # next, order the reactions by the number of distinct species

### 2018-05-25 scripts/data_parser.py
first pass at reaction scaling plot

@@ -55,7 +55,8 @@ def dirs_in(path):
     return __listdir(path, False)
 
 
-def parse_data(rebuild=False, directory=os.path.join(script_dir, 'performance')):
+def parse_data(rebuild=False, directory=os.path.join(script_dir, 'performance'),
+               use_descriptor=None):
     # find all mechs
     mechs = [x for x in os.listdir(directory)
              if os.path.isdir(os.path.join(directory, x))]
@@ -67,9 +68,17 @@ def parse_data(rebuild=False, directory=os.path.join(script_dir, 'performance'))
         # read mechanism info
         with open(mech_meta_file, 'r') as f:
             mech_info = mechdata(**yaml.load(f))
-        for desc in dirs_in(path):
+        if use_descriptor:
+            descriptors = [path]
+        else:
+            descriptors = dirs_in(path)
+
+        for desc in descriptors:
             # get decriptor
-            descriptor = os.path.basename(os.path.normpath(desc))
+            if use_descriptor is None:
+                descriptor = os.path.basename(os.path.normpath(desc))
+            else:
+                descriptor = use_descriptor
             for file in files_in(desc):
                 # check for pickled data
                 pickle_file = file.replace('.txt', '.pickle')

### 2018-05-25 scripts/reaction_scaling_test.py

@@ -289,7 +289,7 @@ if __name__ == '__main__':
                         help='The mechanism to test, defaults to the Sarathy '
                              'isopentanol model.')
     parser.add_argument('-i', '--test_interval',
-                        default=50,
+                        default=200,
                         type=int,
                         help='The interval (in # of reactions) at which to '
                              'run the performance tests to produce the '

### 2018-05-25 scripts/reaction_scaling_test.py
force all reactions to reversible for consistency

@@ -61,6 +61,9 @@ def run(gas, interval, num_states, work_dir, repeats=10):
                 plt.show()
             reacs[i] = ct.ElementaryReaction(reac.reactants, reac.products)
             reacs[i].rate = ct.Arrhenius(A, b, Ea)
+        # set all to reversible
+        reacs[i].reversible = True
+
     # and convert gas
     gas = ct.Solution(thermo='IdealGas', kinetics='GasKinetics',
                       reactions=reacs,

### 2018-05-31 paper.tex
add draft of scaling test appendix

@@
 
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+
 \usepackage{xcolor}
@@
 \usepackage{macros}
+% macro for small underscores for variable names in appendix
+\usepackage{xspace}
+\usepackage{relsize}
+\newcommand{\smallunderscore}{\textscale{.5}{\textunderscore}\xspace}
 
@@
 \author[1]{Nicholas Curtis\corref{corr}}
+\author[2]{Kyle E.~Niemeyer}
 \ead{nicholas.curtis@uconn.edu}
@@
 \cortext[corr]{Corresponding author}
+\address[2]{School of Mechanical, Industrial, and Manufacturing Engineering, Oregon State University, Corvallis, OR 97331, USA}
 
@@
 \subsection{Testing platforms}
+\label{s:test_platforms}
 
@@
 \caption{The Intel CPU configurations used in this study.
-	 The vector-width reported in this table is reported in (ideal) number of double operations per SIMD instruction, as this will be used in measuring SIMD-efficiency; for reference, the vector-width in bits of the \sse/ and \avx/ machines are \SI{128}{\bit} and \SI{256}{\bit}, respectively.
+	 The vector-width reported in this table is reported in (ideal) number of double operations per SIMD instruction, as this will be used in measuring SIMD efficiency; for reference, the vector-width in bits of the \sse/ and \avx/ machines are \SI{128}{\bit} and \SI{256}{\bit}, respectively.
 	 The identifier field will be used as a shorthand descriptor in the performance plots to quickly identify the CPU type.}
@@
 No speedup is achieved for either vectorization type on either CPU---\add{though we stress that Intel OpenCL runtime achieved vectorization using the exact same code and hardware}---indeed, the OpenMP case is faster on both CPUs.
-The reasons for this lack of vectorization are quite technical,~\delete{and somewhat outside the scope of this work}however \add{personal communications~\cite{pocl_communication} with the developers of POCL revealed that, briefly speaking, substantial changes are required to the way POCL prepares LLVM intermediate-representation code (specifically, to ensure LLVM properly recognizes uniform vectorization loop-bounds), use of relaxed math (i.e., ``\texttt{-cl-fast-relaxed-math}'') to allow LLVM's loop-vectorizer to prove a vector instruction is identical to the corresponding scalar version, LLVM's handling of branches (potentially within POCL instructions themselves) and memory access patterns in order to achieve vectorization.}
+The reasons for this lack of vectorization are quite technical,~\delete{and somewhat outside the scope of this work}however \add{personal communications~\cite{pocl_communication} with the developers of POCL revealed that, briefly speaking, substantial changes are required to the way POCL prepares LLVM intermediate-representation code (specifically, to ensure LLVM properly recognizes uniform vectorization loop-bounds), LLVM's handling of branches (potentially within POCL instructions themselves) and memory access patterns and potentially use of relaxed math (i.e., ``\texttt{-cl-fast-relaxed-math}'') to allow LLVM's loop-vectorizer to prove a vector instruction is identical to the corresponding scalar version, in order to achieve vectorization.}
 As will be discussed in~\cref{S:future}, \add{it is hoped that use of explicit vector-types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a validation tool.
@@
 \Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source-term evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
-In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for both Intel OpenCL and OpenMP, likely resulting from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD instructions\slash multiple threads.
+In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for both Intel OpenCL and OpenMP, likely resulting from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them) \add{as demonstrated in~\cref{S:SIMD_scaling}}; i.e., there is simply not enough work to fully utilize the SIMD instructions\slash multiple threads.
 Additionally, OpenMP tends to have higher scalings than the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores, respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.
@@
 This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.
-\Cref{F:source_simd_scaling} shows the SIMD efficiency of source-term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models (isopentanol and USC-Mech II) have higher SIMD-efficiency of \numrange{0.76}{0.78}, and the smaller models (\ce{H2}\slash\ce{CO}, GRI-Mech \num{3.0}) have lower SIMD-efficiency, \numrange{0.6}{0.66}.
+\Cref{F:source_simd_scaling} shows the SIMD efficiency of source-term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models (isopentanol and USC-Mech II) have higher SIMD efficiency of \numrange{0.76}{0.78}, and the smaller models (\ce{H2}\slash\ce{CO}, GRI-Mech \num{3.0}) have lower SIMD efficiency, \numrange{0.6}{0.66}.
 This again demonstrates that the SIMD-vectorization becomes efficient as the amount of work to perform increases (with increasing model size).
@@
 The strong parallel scaling efficiency was also examined for source-term evaluation; the parallel OpenMP code generally had better efficiencies, but when compared to a shallow-vectorized code with roughly equivalent performance---e.g., source-term shallow-vectorized evaluation on \num{4} cores to \num{16} OpenMP cores---the scalings were similar.
-Examination of SIMD-efficiency revealed increasing speedups with chemical model-size for source-term evaluation on the \avx/ CPU, while on the \sse/ CPU speedups larger than the nominal vector-width were achieved.
+Examination of SIMD efficiency revealed increasing speedups with chemical model-size for source-term evaluation on the \avx/ CPU, while on the \sse/ CPU speedups larger than the nominal vector-width were achieved.
 
@@
 
+\section{SIMD efficiency Scaling Example}
+\label{S:SIMD_scaling}
+\add{This simple example demonstrates the dependency of SIMD efficiency for shallow-vectorized OpenCL source-term evaluation on the size of the chemical model in question, i.e., the amount of computational work per source-term evaluation.}
+\add{The base chemical model used in this example was the isopentanol model~\cite{Sarathy:2013jr} used throughout this work, and the same thermo-chemical state database described in~\cref{s:test_platforms} was used for source-term evaluation.}
+\add{As in~\cref{S:results} all reported results are based on \num{10} individual runs and in this example all cases were run on the \avx/ machine using the Intel OpenCL runtime.}
+
+\begin{algorithm}[htbp]
+\algnewcommand\algorithmicinput{\textbf{Input:}}
+\algnewcommand\INPUT{\item[\algorithmicinput]}
+\begin{algorithmic}[0]
+  \caption{A greedy selection algorithm to remove reactions from a base chemical model $M$, while preserving the number of active species.}
+  \label{a:model_gen}
+  \INPUT{Base chemical model $M$ with reactions $R$ and species $S$}
+  \Function{Determine Species Count}{\text{active}}
+    \For {Species $S_k$ in model $M$}
+      \State  $\text{species\smallunderscore{}rxn\smallunderscore{}count}\left[k\right] \gets 0$
+      \For {Reaction $R_j$ in model $M$}
+	\If{$\text{active}[j]$ and $\left(\left\lvert \nu_{k, j}^{\prime}\right\rvert + \left\lvert \nu_{k, j}^{\prime\prime} \right\rvert \right) > 0$}
+	  \State $\text{species\smallunderscore{}rxn\smallunderscore{}count}[k] \gets \text{species\smallunderscore{}rxn\smallunderscore{}count}[k] + 1$
+	\EndIf
+      \EndFor
+    \EndFor
+    \Return \text{species\smallunderscore{}rxn\smallunderscore{}count}
+  \EndFunction
+  \Procedure{Model Generation}{$M$}
+  \State $\text{active}[j] \gets \text{True}$ for all reactions $R_j$ in $M$
+  \State $\text{species\smallunderscore{}rxn\smallunderscore{}count}\gets \Call{Determine Species Count}{active}$
+  \While {$\operatorname{min}\left(species\smallunderscore rxn\smallunderscore count\right) \ge 1$}
+     \State $\text{species\smallunderscore{}rxn\smallunderscore{}count}\gets \Call{Determine Species Count}{active}$
+     \For {Reaction $R_j$ in model $M$}
+	\State $\text{rxn\smallunderscore{}count}[j] \gets \min_{\forall S_k \in R_j}{\left(\text{species\smallunderscore{}rxn\smallunderscore{}count}\left[k\right]\right)}$
+     \EndFor
+     \State $\text{remove\smallunderscore{}at} \gets \operatorname{argmax}\left(\text{rxn\smallunderscore{}count}\right)$
+     \State $\text{active}[\text{remove\smallunderscore{}at}] \gets \text{False}$
+  \EndWhile
+  \EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+\add{First, the reactions in the isopentanol model were converted to simple reversible Arrhenius reactions by either simply dropping third-body efficiency calculations (third-body enhanced reactions), using the high pressure-limit coefficients (falloff\slash chemically-activated and P-Log reactions) or fitting Arrenhius parameters to the calculated rate constant at a fixed pressure (chebyshev reactions).}
+\add{This conversion made the cost of reaction rate evaluation roughly equivalent between all reactions in model, separating the effect of chemical model size from computational intensities of different reaction types on the SIMD efficiency.}
+\add{Next, a greedy reaction removal algorithm (\cref{a:model_gen}) was utilized to generated a number of models of varying size ranging from \numrange{2100}{186} reactions, in increments of 200 reactions (except the final increment from \num{200} to \num{186} reactions).}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\linewidth]{simd_efficiency_scaling}
+\caption{The effect on SIMD efficiency of varying chemical model size}
+\label{F:simd_vs_rxns}
+\end{figure}
+
+\add{In order to discern the effect of varying chemical model size on the SIMD efficiency, shallow-vectorized and unvectorized source-term evaluation performance tests were run.}
+\add{As demonstrated in~\cref{F:simd_vs_rxns} the SIMD efficiency is strongly dependent on the generated model size and and thus, amount of computational work per thermo-chemical state.}
+\add{In addition, the range of SIMD efficiency in this example (\numrange{0.56}{0.91}) is is larger than the SIMD efficiencies calculated for real chemical models, as seen in~\cref{F:source_simd_scaling}.}
+\add{For smaller models---e.g., \ce{H2}\slash\ce{CO} which had a SIMD efficiency of \num{0.6}---this suggests that the presence of more expensive fall-off\slash chemically activated reactions in model can increase the SIMD efficiency.}
+\add{However, the isopentanol model achieved a SIMD efficiency of only \num{0.78} in~\cref{S:source_results}, suggesting that more work could be done to optimize the source-term evaluations.}
+\add{In particular, it's likely that a reaction sorting method, such as suggested by Sewerin and Rigopoulos~\cite{Sewerin20151375}, would be particularly beneficial to reduce the number of vector-gather\slash scatter operations incurred during source-term evaluation.}
+
 \pagebreak

### 2018-05-31 derivations.tex
mark as supplimental

@@
 
-\title{A System of Equations and Derivation of Sparse Constant-Pressure\slash Constant-Volume Chemical Kinetic Jacobians for \texttt{pyJac}}
+\title{Supplimental: A System of Equations and Derivation of Sparse Constant-Pressure\slash Constant-Volume Chemical Kinetic Jacobians for \texttt{pyJac}}
 
@@
 \begin{abstract} % not to exceed 200 words
-This document presents the system of chemical kinetic equations evaluated by the new version of code-generation platform, \texttt{pyJac-v2}.
+This document is a suppliment to the work, \textit{SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source-Term Evaluation} and presents the system of chemical kinetic equations evaluated by the new version of code-generation platform, \texttt{pyJac-v2}.
 First, the thermo-chemical source terms for fixed-mass, adiabatic constant-pressure\slash constant-volume (\conp/\slash\conv/) reactors are given, followed by the derivation of the sparse analytical chemical kinetic Jacobian for both reactors.
@@
 \section{Introduction}
-This document contains the derivations for \texttt{pyJac} v2.0, which generates code to evaluate the chemical source terms and analytical chemical kinetic Jacobians of constant-pressure\slash constant-volume, fixed-mass, adiabatic reactors.
+This document is a suppliment to the paper---\textit{SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source-Term Evaluation}---and is provided as a reference guide to the equations evaluated by \texttt{pyJac} v2.0, a code-generation platform for the evaluatation of chemical source terms and analytical chemical kinetic Jacobians of constant-pressure\slash constant-volume, fixed-mass, adiabatic reactors.
 Note that equations specific to the constant-pressure or constant-volume derivation will be marked with \conp/ or \conv/ respectively.

### 2018-05-31 paper.tex
add in section header was causing issues in the generated PDF sections

@@
 
-\subsection{\add{Derive and validate a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity}}
+\subsection{Derive and validate a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity}
 A new formulation resulted in highly sparse (\SIrange{28.6}{88.5}{\percent}) chemical kinetic Jacobians, with options to futher increase sparsity (to \SIrange{31.66}{92.0}{\percent}) at the expense of strict correctness.
 
-\subsection{\add{Enable cross-platform SIMD\slash SIMT vectorization for the CPU, GPU and other accelerators}}
+\subsection{Enable cross-platform SIMD\slash SIMT vectorization for the CPU, GPU and other accelerators}
 \add{Shallow-vectorized evaluation of \texttt{pyJac}'s chemical kinetic source-term and analytical Jacobian was demonstrated on both the CPU and GPU via OpenCL.}
@@
 
-\subsection{\add{Investigate the performance for a wide range chemical kinetic models, specifically looking at speedups due to SIMD\slash SIMT-vectorization as well as relative to the previous version of \texttt{pyJac}}}
+\subsection{Investigate the performance for a wide range chemical kinetic models, specifically looking at speedups due to SIMD\slash SIMT-vectorization as well as relative to the previous version of \texttt{pyJac}}
 
@@
 
-\subsection{\add{Future extensions to this work and practical notes on OpenCL usage}}
+\subsection{Future extensions to this work and practical notes on OpenCL usage}
 \label{S:future}

### 2018-06-01 paper.tex
tweak wording

@@
 In contrast,~\cref{F:pocl_source} shows the mean run time per condition of deep, shallow and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
-No speedup is achieved for either vectorization type on either CPU---\add{though we stress that Intel OpenCL runtime achieved vectorization using the exact same code and hardware}---indeed, the OpenMP case is faster on both CPUs.
 The reasons for this lack of vectorization are quite technical,~\delete{and somewhat outside the scope of this work}however \add{personal communications~\cite{pocl_communication} with the developers of POCL revealed that, briefly speaking, substantial changes are required to the way POCL prepares LLVM intermediate-representation code (specifically, to ensure LLVM properly recognizes uniform vectorization loop-bounds), LLVM's handling of branches (potentially within POCL instructions themselves) and memory access patterns and potentially use of relaxed math (i.e., ``\texttt{-cl-fast-relaxed-math}'') to allow LLVM's loop-vectorizer to prove a vector instruction is identical to the corresponding scalar version, in order to achieve vectorization.}
+No speedup is achieved for either vectorization type on either CPU---\add{though we stress that Intel OpenCL runtime achieved vectorization using the same code and hardware}---indeed, the OpenMP case is faster on both CPUs.
 As will be discussed in~\cref{S:future}, \add{it is hoped that use of explicit vector-types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a validation tool.

### 2018-06-01 paper.tex
make explanation match sharelatex

@@
 In contrast,~\cref{F:pocl_source} shows the mean run time per condition of deep, shallow and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
-The reasons for this lack of vectorization are quite technical,~\delete{and somewhat outside the scope of this work}however \add{personal communications~\cite{pocl_communication} with the developers of POCL revealed that, briefly speaking, substantial changes are required to the way POCL prepares LLVM intermediate-representation code (specifically, to ensure LLVM properly recognizes uniform vectorization loop-bounds), LLVM's handling of branches (potentially within POCL instructions themselves) and memory access patterns and potentially use of relaxed math (i.e., ``\texttt{-cl-fast-relaxed-math}'') to allow LLVM's loop-vectorizer to prove a vector instruction is identical to the corresponding scalar version, in order to achieve vectorization.}
 No speedup is achieved for either vectorization type on either CPU---\add{though we stress that Intel OpenCL runtime achieved vectorization using the same code and hardware}---indeed, the OpenMP case is faster on both CPUs.
+The reasons for this lack of vectorization are quite technical,~\delete{and somewhat outside the scope of this work}however \add{personal communications~\cite{pocl_communication} with the developers of POCL revealed that in order to achieve vectorization, changes are required to both the way POCL prepares LLVM \cite{Lattner:2004:LCF:977395.977673} IR-code---specifically, to specialize boundary cases to ensure LLVM recognizes uniform vectorization loop-bounds, even if said bounds \textit{are} uniform in practice---as well as improvements to LLVM's loop-vectorizer: e.g., for proofs of a vector instruction's ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns.}
 As will be discussed in~\cref{S:future}, \add{it is hoped that use of explicit vector-types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a validation tool.

### 2018-06-01 paper.tex
reword

@@
 \Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source-term evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
-In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for both Intel OpenCL and OpenMP, likely resulting from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them) \add{as demonstrated in~\cref{S:SIMD_scaling}}; i.e., there is simply not enough work to fully utilize the SIMD instructions\slash multiple threads.
+In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for both Intel OpenCL and OpenMP, likely resulting from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them); \add{as demonstrated in~\cref{S:SIMD_scaling} the amount of computational work required per thermo-chemical state plays a critical role in the full utilization of SIMD instructions\slash multiple threads}.
 Additionally, OpenMP tends to have higher scalings than the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores, respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.

### 2018-06-01 paper.tex
add note about reaction sorting

@@
 It is also possible, particularly for the Intel OpenCL\slash POCL runtimes, that better performance\slash stability might be achieved using so-called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}; \add{specifically this change is hoped to enable vectorization on the POCL runtime} and might also enable deep-vectorization on the Intel OpenCL runtime.
-\add{Finally, future extensions of this work will include additional target-languages (vectorized OpenMP, CUDA, ISPC, etc.) to improve ease of use and reliability, as well as improvements to the existing OpenCL targets, e.g., to enable meaningful deep-vectorization.}
+\add{Finally, future extensions of this work will include additional target-languages (vectorized OpenMP, CUDA, ISPC, etc.) to improve ease of use and reliability, as well as improvements to the existing OpenCL targets, e.g., to enable meaningful deep-vectorized evaluation, as well the implementation of reaction sorting~\cite{Sewerin20151375} to improve SIMD-efficiency (\cref{S:SIMD_scaling}).}
 

### 2018-06-01 paper.tex
fixes for scaling example

@@
 \label{S:SIMD_scaling}
-\add{This simple example demonstrates the dependency of SIMD efficiency for shallow-vectorized OpenCL source-term evaluation on the size of the chemical model in question, i.e., the amount of computational work per source-term evaluation.}
-\add{The base chemical model used in this example was the isopentanol model~\cite{Sarathy:2013jr} used throughout this work, and the same thermo-chemical state database described in~\cref{s:test_platforms} was used for source-term evaluation.}
+\add{This simple example demonstrates the dependency of the SIMD efficiency of shallow-vectorized OpenCL source-term evaluation on the size of the chemical model in question, i.e., the amount of computational work per source-term evaluation.}
+\add{The base chemical model for this example was the isopentanol model~\cite{Sarathy:2013jr} used throughout this work, and the same thermo-chemical state database described in~\cref{s:test_platforms} was used for source-term evaluation.}
 \add{As in~\cref{S:results} all reported results are based on \num{10} individual runs and in this example all cases were run on the \avx/ machine using the Intel OpenCL runtime.}
@@
   \State $\text{active}[j] \gets \text{True}$ for all reactions $R_j$ in $M$
-  \State $\text{species\smallunderscore{}rxn\smallunderscore{}count}\gets \Call{Determine Species Count}{active}$
-  \While {$\operatorname{min}\left(species\smallunderscore rxn\smallunderscore count\right) \ge 1$}
-     \State $\text{species\smallunderscore{}rxn\smallunderscore{}count}\gets \Call{Determine Species Count}{active}$
+  \State $\text{species\smallunderscore{}rxn\smallunderscore{}count}\gets \Call{Determine Species Count}{\text{active}}$
+  \While {$\operatorname{min}\left(\text{species\smallunderscore rxn\smallunderscore count}\right) \ge 1$}
+     \State $\text{species\smallunderscore{}rxn\smallunderscore{}count}\gets \Call{Determine Species Count}{\text{active}}$
      \For {Reaction $R_j$ in model $M$}
@@
 \add{This conversion made the cost of reaction rate evaluation roughly equivalent between all reactions in model, separating the effect of chemical model size from computational intensities of different reaction types on the SIMD efficiency.}
-\add{Next, a greedy reaction removal algorithm (\cref{a:model_gen}) was utilized to generated a number of models of varying size ranging from \numrange{2100}{186} reactions, in increments of 200 reactions (except the final increment from \num{200} to \num{186} reactions).}
+\add{Next, a greedy reaction removal algorithm (\cref{a:model_gen}) was utilized to generate a number of models ranging from \numrange{2100}{186} reactions, in increments of 200 reactions (except the final increment from \num{200} to \num{186} reactions).}
 
@@
 \add{As demonstrated in~\cref{F:simd_vs_rxns} the SIMD efficiency is strongly dependent on the generated model size and and thus, amount of computational work per thermo-chemical state.}
-\add{In addition, the range of SIMD efficiency in this example (\numrange{0.56}{0.91}) is is larger than the SIMD efficiencies calculated for real chemical models, as seen in~\cref{F:source_simd_scaling}.}
-\add{For smaller models---e.g., \ce{H2}\slash\ce{CO} which had a SIMD efficiency of \num{0.6}---this suggests that the presence of more expensive fall-off\slash chemically activated reactions in model can increase the SIMD efficiency.}
-\add{However, the isopentanol model achieved a SIMD efficiency of only \num{0.78} in~\cref{S:source_results}, suggesting that more work could be done to optimize the source-term evaluations.}
-\add{In particular, it's likely that a reaction sorting method, such as suggested by Sewerin and Rigopoulos~\cite{Sewerin20151375}, would be particularly beneficial to reduce the number of vector-gather\slash scatter operations incurred during source-term evaluation.}
+\add{In addition, the range of SIMD efficiency in this example (\numrange{0.56}{0.91}) is larger than the range of SIMD efficiencies calculated for real chemical models, as seen in~\cref{F:source_simd_scaling}.}
+\add{For smaller models---e.g., \ce{H2}\slash\ce{CO} which had a SIMD efficiency of \num{0.6} on the \avx/ CPU---this suggests that the presence of more computationally intensive fall-off\slash chemically activated reactions in model can increase the SIMD efficiency.}
+\add{However, the base isopentanol model achieved a SIMD efficiency of only \num{0.78} on the \avx/ machine in~\cref{S:source_results}, suggesting that more work could be done to optimize the source-term evaluations.}
+\add{In particular, it's likely that a reaction sorting method, such as suggested by Sewerin and Rigopoulos~\cite{Sewerin20151375}, would be particularly beneficial to reduce the number of vector-gather\slash scatter\slash masking operations incurred during source-term evaluation.}
 

### 2018-06-04 paper.tex
update grant statement
edits through page 3

@@
 \begin{abstract} % not to exceed 200 words
-A code generation platform, \texttt{pyJac}, for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized constant-pressure (\conp/)\slash constant-volume (\conv/) thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation has been developed.
+A code generation platform, \texttt{pyJac}, for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized constant-pressure\slash constant-volume thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation has been developed.% Can you introduce this a bit more?
+good to start with at least one sentence of introduction.
+also, this is passive and quite long\slash complex
 Selected chemical kinetic models covering the range of sizes typically used in reactive-flow simulations were used for demonstration.
-A new formulation of the chemical kinetic governing equations was derived and validated, resulting Jacobian sparsities of \SIrange{28.6}{92.0}{$\percent$} for the tested models.
-Speedups of \SIrange{3.40}{4.08}{$\times$} were found for shallow-vectorized OpenCL source-rate evaluation compared a parallel OpenMP code on an \avx/ central processing unit (CPU), increasing to \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.03}{4.23}{$\times$} for sparse and dense chemical kinetic Jacobian evaluation, respectively.
-In addition, the performance of a NVIDIA \gpunew/ graphics processing unit (GPU) was compared to a NVIDIA \gpuold/ GPU finding speedups of \SIrange{1.40}{1.88}{$\times$}, \SIrange{1.10}{1.59}{$\times$}, and \SIrange{1.36}{3.0}{$\times$} for source-rate, sparse and dense Jacobian evaluation, respectively.
-Furthermore, the effect of data-ordering was investigated and an storage pattern specifically formulated for vectorized evaluation was proposed; as well, the effect of the \conp/\slash\conv/ assumptions and varying vector widths were studied on source-term evaluation performance.
-Speedups over a first-order finite-difference approach reached \SIrange{17.60}{45.13}{$\times$} and \SIrange{55.11}{245.63}{$\times$} for sparse\slash dense evaluation on the CPU\slash GPU respectively, while dense Jacobian evaluation was up to \SI{19.56}{$\times$} and \SI{2.84}{$\times$} times faster than a previous version of \texttt{pyJac} on a \sse/ CPU and \gpuold/ GPU, respectively.
+A new formulation of the chemical kinetic governing equations was derived and validated, resulting in Jacobian sparsities of \SIrange{28.6}{92.0}{$\percent$} for the tested models.
+Speedups of \SIrange{3.40}{4.08}{$\times$} were found for shallow-vectorized OpenCL source-rate evaluation compared with a parallel OpenMP code on an \avx/ central processing unit (CPU), increasing to \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.03}{4.23}{$\times$} for sparse and dense chemical kinetic Jacobian evaluation, respectively.
+In addition, the performance of a NVIDIA \gpunew/ graphics processing unit (GPU) was compared to a NVIDIA \gpuold/ GPU finding speedups of \SIrange{1.40}{1.88}{$\times$}, \SIrange{1.10}{1.59}{$\times$}, and \SIrange{1.36}{3.0}{$\times$} for source-rate, sparse, and dense Jacobian evaluation, respectively.
+%is this sentence necessary in the abstract?
+Furthermore, the effect of data-ordering was investigated and a storage pattern specifically formulated for vectorized evaluation was proposed; as well, the effect of the constant pressure\slash volume assumptions and varying vector widths were studied on source-term evaluation performance.
+Speedups over a first-order finite-difference approach reached \SIrange{17.60}{45.13}{$\times$} and \SIrange{55.11}{245.63}{$\times$} for sparse\slash dense evaluation on the CPU \slash GPU respectively,%not sure what numbers refer to- sparse, dense, CPU, GPU?
+while dense Jacobian evaluation was up to \SI{19.56}{$\times$} and \SI{2.84}{$\times$} times faster than a previous version of \texttt{pyJac} on a CPU and GPU, respectively.
 Finally, future directions for vectorized chemical kinetic evaluation and sparse linear-algebra techniques were discussed.
@@
 \end{frontmatter}
+\todo[inline]{abstract comments in source}
 
@@
 
-As the importance of detailed chemical kinetics for predictive reactive-flow simulations has become recognized~\cite{LU2009192}, the size and complexity of chemical kinetic models describing current and next-generation fuels relevant to transportation and power generation have increased greatly; e.g., a biodiesel model~\cite{WESTBROOK2011742} consists of \textasciitilde\num{3500} chemical species, and over \num{17000} reactions.
-Moreover, the cost of evaluating the chemical source-terms scales linearly with the size of the model, while evaluation and factorization of the chemical kinetic Jacobian---if naively implemented via finite-difference method---scale quadratically and cubically with the number of species in the model~\cite{LU2009192}.
-These factors often make the use of detailed chemical kinetics prohibitive in practice; e.g., in a direct numerical simulation~\cite{Spafford:2010aa} utilizing a 22-species model, evaluation of reaction rates consumed around half of the total run time.
-In addition, chemical kinetic Jacobian evaluation and factorization---utilized by most common implicit integration techniques to deal with numerical stiffness---is a bottleneck for even modestly-sized chemical models in realistic reactive flow simulations, necessitating use of other cost-reduction strategies~\cite{LU2009192}.
-
-A host of techniques have been developed to lessen the computational demand of chemical kinetic calculations while maintaining fidelity, these fall broadly into a few categories: removal of unimportant species and reactions~\cite{Lu:2006bb,Pepiot-Desjardins:2008,Hiremath:2010jw,Niemeyer:2010bt,Curtis:2015}, lumping of species with similar thermo-chemical properties~\cite{Lu:2007,Ahmed:2007fa,Pepiot:2008kq}, and time-scale methods that reduce numerical stiffness~\cite{Maas:1992ws,Lam:1994ws,Lu:2001ve,Gou:2010}.
-The interested reader is referred to the recent review of Tur\'anyi and Tomlin~\cite{turanyi2016analysis} for a comprehensive overview.
+As the combustion and reacting-flows community has recognized the
+importance of detailed chemical kinetics for predictive reactive-flow simulations~\cite{LU2009192},
+chemical kinetic models have grown in size and complexity to describe current and next-generation fuels relevant to transportation and power generation.
+For example, a recent biodiesel model~\cite{WESTBROOK2011742} consists of \textasciitilde\num{3500} chemical species and over \num{17000} reactions.
+Moreover, the cost of evaluating the chemical source-terms scales linearly with the size of the model, while evaluating and factorizing the chemical kinetic Jacobian respectively scale quadratically and cubically with the number of species in the model, if naively implemented via a finite-difference method~\cite{LU2009192}.
+These factors often prohibit using detailed chemical kinetics in practice; e.g., in a direct numerical simulation~\cite{Spafford:2010aa} using a 22-species model, evaluating reaction rates consumed around half of the total run time.
+In addition, most common implicit integration techniques need to evaluate and factorize the Jacobian matrix to deal with stiffness.
+As a result, these operations are bottlenecks when using even moderately sized chemical models in realistic reactive-flow simulations, necessitating other cost-reduction strategies~\cite{LU2009192}.
+
+A host of techniques have been developed to lessen the computational demand of chemical kinetic calculations while maintaining fidelity, falling broadly into a few categories: removal of unimportant species and reactions~\cite{Lu:2006bb,Pepiot-Desjardins:2008,Hiremath:2010jw,Niemeyer:2010bt,Curtis:2015}, lumping of species with similar thermo-chemical properties~\cite{Lu:2007,Ahmed:2007fa,Pepiot:2008kq}, and time-scale methods that reduce numerical stiffness~\cite{Maas:1992ws,Lam:1994ws,Lu:2001ve,Gou:2010}.
+We refer interested readers to the recent review by Tur\'anyi and Tomlin~\cite{turanyi2016analysis} for a comprehensive overview.
 
 In addition to the previously mentioned cost reduction methods, effort has gone into improving the integration algorithms and codes that evaluate the chemical kinetics~\cite{Gou:2010,SCHWER2002270,Niemeyer:2016aa,GAO2015287}.
-In particular, a carefully derived analytical formulation of the Jacobian can greatly increase the sparsity of the matrix~\cite{SCHWER2002270} and drop the cost of Jacobian evaluation to a to a linear dependence on the number of species in the model~\cite{LU2009192}; Jacobian factorization can be accelerated via sparse-matrix techniques~\cite{superlu99}.
-Use of Single-Instruction, Multiple-Data (SIMD) and the related Single-Instruction, Multiple-Thread (SIMT) processors to accelerate chemical kinetic simulations has also been studied, e.g., as in~\cite{CurtisGPU:2017,Sewerin20151375,Shi:2012aa,Niemeyer:2016aa}.
+In particular, a carefully derived analytical formulation of the Jacobian matrix can greatly increase sparsity~\cite{SCHWER2002270} and drop the cost of Jacobian evaluation to linearly depend on the number of species in the model~\cite{LU2009192}; Jacobian factorization can then be accelerated via sparse-matrix techniques~\cite{superlu99}.
+In addition, studies have shown that Single-Instruction, Multiple-Data (SIMD) and the related Single-Instruction, Multiple-Thread (SIMT) processors can accelerate chemical kinetic simulations~\cite{Shi:2012aa,Niemeyer:2014aa,Sewerin20151375,Niemeyer:2016aa,CurtisGPU:2017,stone2018}.
 
 SIMD and SIMT programming are two important vector-processing paradigms used increasingly in scientific computing.
-Traditional multi-core parallelism is often used to increase central processing unit (CPU) performance, however as exponential growth in processing power---colloquially known as Moore's law---has slowed~\cite{khan2018science}, SIMD processors as well as SIMT processors, e.g., in the form of graphics processing units (GPUs), have gained recognition due to their increased floating operation throughput.
-The parallel programming standard OpenCL~\cite{stone2010opencl} has further enabled adoption of vector-processing based codes in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems, e.g., CPU, GPU, or Intel's Many Integrated Core (MIC) architecture, etc.
-In this discussion, we will largely use OpenCL terminology to describe these processing paradigms, as it provides a convenient way to classify otherwise disparate processor type (e.g., CPUs and GPUs), however the concepts discussed herein are broadly applicable to SIMD\slash SIMT processing and not tied to OpenCL specifically.
+Traditional multi-core parallelism is often used to increase central processing unit (CPU) performance; however, as exponential growth in processing power---colloquially known as Moore's law---has slowed~\cite{khan2018science}, SIMD processors as well as SIMT processors, e.g., in the form of graphics processing units (GPUs), have gained recognition due to their increased floating operation throughput.\todo{this isn't quite right; multicore parallelism is how CPUs have bypassed Moore's}{}
+The parallel programming standard OpenCL~\cite{stone2010opencl} has further enabled adoption of vector processing in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems, e.g., CPU, GPU, or Intel's Many Integrated Core (MIC) architecture.
+Here we will largely use OpenCL terminology to describe these processing paradigms, as it provides a convenient way to classify otherwise disparate processor types (e.g., CPUs and GPUs).
+However, the concepts discussed herein broadly apply to SIMD\slash SIMT processing.
 
@@
 
-A typical modern CPU has many compute units (i.e., cores), each with specialized vector processing units capable of running SIMD instructions (\cref{F:SIMD}).
-A SIMD instruction utilizes the vector processor to execute the same floating point operation (e.g., multiplication, division, etc.) on a several different data concurrently; the number of concurrent operations possible is known as the vector-width\footnote{OpenCL allows for use of vector-widths different from the actual hardware vector-width via implicit conversion, and may provide some performance benefit as will be studied in Sec.~\ref{S:results}} and is typically around 2--4 double precision operations.
+A typical modern CPU contains multiple compute units (i.e., cores), each with specialized vector processing units capable of running SIMD instructions, as \cref{F:SIMD} depicts.
+A SIMD instruction uses the vector processor to execute the same floating-point operation (e.g., multiplication, division) on different data concurrently;
+the number of concurrent operations possible is known as the vector-width\footnote{OpenCL allows for use of vector-widths different from the actual hardware vector-width via implicit conversion, and may provide some performance benefit as will be studied in Sec.~\ref{S:results}} and is typically around 2--4 double precision operations.
 Additionally, specialized hardware accelerators, e.g., Intel's Xeon Phi co-processor (MIC architecture) have been developed that have tens of cores with very wide vector-widths (e.g., \numrange{4}{8} double precision operations); these very wide vector-widths are also found on cutting-edge and forthcoming Intel CPUs; i.e., the Skylake Xeon and Cannon Lake architectures, respectively.
@@
 Kroshko and Spiteri~\cite{kroshko2013efficient} implemented a SIMD-vectorized third-order stiff Rosenbrock integrator for atmospheric chemistry on the CBE and found a speedup of \SI{1.89}{$\times$} (a parallel scaling efficiency of \SI{94}{$\percent$}) over a serial version of the same code.
-Stone and Niemeyer~\cite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
-SIMD-vectorization improved the integrator performance over an OpenMP baseline (which was vectorized by simple compiler hints (i.e., \texttt{\#pragmas}) by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance was only \SIrange{1.4}{1.6}{$\times$} faster than the OpenMP baseline due to thread-divergence concerns~\cite{stone2016}.
+Stone and Niemeyer~\cite{stone2018} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
+SIMD-vectorization improved the integrator performance over an OpenMP baseline (which was vectorized by simple compiler hints (i.e., \texttt{\#pragmas}) by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance was only \SIrange{1.4}{1.6}{$\times$} faster than the OpenMP baseline due to thread-divergence concerns~\cite{stone2018}.
 
@@
 \section{Acknowledgments}
-This research was funded by the National Science Foundation under grant ACI-1534688.
+This material is based upon work supported by the National Science Foundation
+under grants ACI-1534688 (Curtis and Sung) and ACI-1535065 (Niemeyer).
 

### 2018-06-04 paper.tex
changes up through section 1.1; added color for my revision

@@
 %======================================================================
-\usepackage{todonotes}
+
 \usepackage{graphicx}
-\usepackage[binary-units]{siunitx}
-\usepackage{gensymb}
-\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}, e.g., \ce{H2SO4}
 \usepackage{booktabs,multicol} %better tables
@@
 \usepackage{textcomp}
+\usepackage{gensymb}
 \usepackage{ulem}
@@
 
-\usepackage{xcolor}
-\definecolor{darkgreen}{rgb}{0,0.5,0}
-
-% \usepackage[xcolor]{changebar}
-% \newcommand{\add}[1]{{\sloppy\cbcolor{teal}\textcolor{teal}{\cbstart {#1}\cbend}}}  % add
-% \newcommand{\delete}[1]{\sloppy\cbcolor{red}\textcolor{red}{\cbdelete\sout{#1}}}
-
-\newcommand{\add}[1]{{\sloppy\textcolor{teal}{#1}}}  % add
-\newcommand{\delete}[1]{\sloppy\textcolor{red}{\sout{#1}}} %delete with no strikeout
+\usepackage[usenames,dvipsnames]{xcolor}
+
+\usepackage[xcolor]{changebar}
+\newcommand{\add}[1]{{\sloppy\cbcolor{teal}\textcolor{teal}{\cbstart {#1}\cbend}}}  % add
+\newcommand{\delete}[1]{\sloppy\cbcolor{red}\textcolor{red}{\cbdelete\sout{#1}}}
+\newcommand{\revise}[1]{{\sloppy\textcolor{Fuchsia}{#1}}}  % add
+
+\newcommand{\addnl}[1]{{\sloppy\textcolor{teal}{#1}}}  % add
+\newcommand{\deletenl}[1]{\sloppy\textcolor{red}{\sout{#1}}} %delete with no strikeout
 
@@
 
+\usepackage[version=4]{mhchem} % Formula subscripts using \ce{}, e.g., \ce{H2SO4}
 \usepackage{amsmath,eqparbox,xparse,mathtools}
+\usepackage[binary-units]{siunitx}
+
+\usepackage{todonotes}
 
@@
 
-\title{SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source-Term Evaluation}
+\title{Using SIMD and SIMT vectorization to evaluate sparse chemical kinetic Jacobian matrices and thermochemical source terms}
 
@@
 \begin{abstract} % not to exceed 200 words
-A code generation platform, \texttt{pyJac}, for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized constant-pressure\slash constant-volume thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation has been developed.% Can you introduce this a bit more?
+A code generation platform, \texttt{pyJac}, for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized \revise{constant-pressure\slash volume} thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation has been developed.% Can you introduce this a bit more?
 good to start with at least one sentence of introduction.
@@
 Selected chemical kinetic models covering the range of sizes typically used in reactive-flow simulations were used for demonstration.
-A new formulation of the chemical kinetic governing equations was derived and validated, resulting in Jacobian sparsities of \SIrange{28.6}{92.0}{$\percent$} for the tested models.
-Speedups of \SIrange{3.40}{4.08}{$\times$} were found for shallow-vectorized OpenCL source-rate evaluation compared with a parallel OpenMP code on an \avx/ central processing unit (CPU), increasing to \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.03}{4.23}{$\times$} for sparse and dense chemical kinetic Jacobian evaluation, respectively.
-In addition, the performance of a NVIDIA \gpunew/ graphics processing unit (GPU) was compared to a NVIDIA \gpuold/ GPU finding speedups of \SIrange{1.40}{1.88}{$\times$}, \SIrange{1.10}{1.59}{$\times$}, and \SIrange{1.36}{3.0}{$\times$} for source-rate, sparse, and dense Jacobian evaluation, respectively.
+A new formulation of the chemical kinetic governing equations was derived and validated, \revise{resulting in} Jacobian sparsities of \SIrange{28.6}{92.0}{$\percent$} for the tested models.
+Speedups of \SIrange{3.40}{4.08}{$\times$} were found for shallow-vectorized OpenCL source-rate evaluation \revise{compared with} a parallel OpenMP code on an \avx/ central processing unit (CPU), increasing to \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.03}{4.23}{$\times$} for sparse and dense chemical kinetic Jacobian evaluation, respectively.
+In addition, the performance of a Nvidia \gpunew/ graphics processing unit (GPU) was compared to a Nvidia \gpuold/ GPU finding speedups of \SIrange{1.40}{1.88}{$\times$}, \SIrange{1.10}{1.59}{$\times$}, and \SIrange{1.36}{3.0}{$\times$} for source-rate, sparse, and dense Jacobian evaluation, respectively.
 %is this sentence necessary in the abstract?
-Furthermore, the effect of data-ordering was investigated and a storage pattern specifically formulated for vectorized evaluation was proposed; as well, the effect of the constant pressure\slash volume assumptions and varying vector widths were studied on source-term evaluation performance.
+Furthermore, the effect of data-ordering was investigated and \revise{a} storage pattern specifically formulated for vectorized evaluation was proposed; as well, the effect of the \revise{constant pressure\slash volume} assumptions and varying vector widths were studied on source-term evaluation performance.
 Speedups over a first-order finite-difference approach reached \SIrange{17.60}{45.13}{$\times$} and \SIrange{55.11}{245.63}{$\times$} for sparse\slash dense evaluation on the CPU \slash GPU respectively,%not sure what numbers refer to- sparse, dense, CPU, GPU?
@@
 
-As the combustion and reacting-flows community has recognized the
+\revise{As the combustion and reacting-flows community has recognized the
 importance of detailed chemical kinetics for predictive reactive-flow simulations~\cite{LU2009192},
 chemical kinetic models have grown in size and complexity to describe current and next-generation fuels relevant to transportation and power generation.
-For example, a recent biodiesel model~\cite{WESTBROOK2011742} consists of \textasciitilde\num{3500} chemical species and over \num{17000} reactions.
-Moreover, the cost of evaluating the chemical source-terms scales linearly with the size of the model, while evaluating and factorizing the chemical kinetic Jacobian respectively scale quadratically and cubically with the number of species in the model, if naively implemented via a finite-difference method~\cite{LU2009192}.
-These factors often prohibit using detailed chemical kinetics in practice; e.g., in a direct numerical simulation~\cite{Spafford:2010aa} using a 22-species model, evaluating reaction rates consumed around half of the total run time.
-In addition, most common implicit integration techniques need to evaluate and factorize the Jacobian matrix to deal with stiffness.
-As a result, these operations are bottlenecks when using even moderately sized chemical models in realistic reactive-flow simulations, necessitating other cost-reduction strategies~\cite{LU2009192}.
-
-A host of techniques have been developed to lessen the computational demand of chemical kinetic calculations while maintaining fidelity, falling broadly into a few categories: removal of unimportant species and reactions~\cite{Lu:2006bb,Pepiot-Desjardins:2008,Hiremath:2010jw,Niemeyer:2010bt,Curtis:2015}, lumping of species with similar thermo-chemical properties~\cite{Lu:2007,Ahmed:2007fa,Pepiot:2008kq}, and time-scale methods that reduce numerical stiffness~\cite{Maas:1992ws,Lam:1994ws,Lu:2001ve,Gou:2010}.
-We refer interested readers to the recent review by Tur\'anyi and Tomlin~\cite{turanyi2016analysis} for a comprehensive overview.
+For example,} a recent biodiesel model~\cite{WESTBROOK2011742} consists of \textasciitilde\num{3500} chemical species and over \num{17000} reactions.
+Moreover, the cost of evaluating the chemical source-terms scales linearly with the size of the model, while evaluating and factorizing the chemical kinetic Jacobian respectively scale quadratically and cubically with the number of species in the model, \revise{if naively implemented via a finite-difference method~\cite{LU2009192}}.
+These factors often \revise{prohibit using} detailed chemical kinetics in practice; e.g., in a direct numerical simulation \revise{using} a 22-species model, \revise{evaluating} reaction rates consumed around half of the total run time~\cite{Spafford:2010aa}.
+\revise{In addition, most common implicit integration techniques need to evaluate and factorize the Jacobian matrix to deal with stiffness.
+As a result, these operations are bottlenecks when using even moderately sized chemical models in realistic reactive-flow simulations, necessitating other cost-reduction strategies~\cite{LU2009192}.}
+
+A host of techniques have been developed to lessen the computational demand of chemical kinetic calculations while maintaining fidelity, \revise{falling} broadly into \revise{three} categories: removal of unimportant species and reactions~\cite{Lu:2006bb,Pepiot-Desjardins:2008,Hiremath:2010jw,Niemeyer:2010bt,Curtis:2015},
+lumping of species with similar thermo-chemical properties~\cite{Lu:2007,Ahmed:2007fa,Pepiot:2008kq},
+and time-scale methods that reduce numerical stiffness~\cite{Maas:1992ws,Lam:1994ws,Lu:2001ve,Gou:2010}.
+\revise{We refer interested readers} to the recent review by Tur\'anyi and Tomlin~\cite{turanyi2016analysis} for a comprehensive overview.
 
 In addition to the previously mentioned cost reduction methods, effort has gone into improving the integration algorithms and codes that evaluate the chemical kinetics~\cite{Gou:2010,SCHWER2002270,Niemeyer:2016aa,GAO2015287}.
-In particular, a carefully derived analytical formulation of the Jacobian matrix can greatly increase sparsity~\cite{SCHWER2002270} and drop the cost of Jacobian evaluation to linearly depend on the number of species in the model~\cite{LU2009192}; Jacobian factorization can then be accelerated via sparse-matrix techniques~\cite{superlu99}.
-In addition, studies have shown that Single-Instruction, Multiple-Data (SIMD) and the related Single-Instruction, Multiple-Thread (SIMT) processors can accelerate chemical kinetic simulations~\cite{Shi:2012aa,Niemeyer:2014aa,Sewerin20151375,Niemeyer:2016aa,CurtisGPU:2017,stone2018}.
+In particular, a carefully derived analytical formulation of the Jacobian \revise{matrix can greatly increase sparsity}~\cite{SCHWER2002270} and drop the cost of Jacobian evaluation to \revise{linearly depend} on the number of species in the model~\cite{LU2009192}; \revise{sparse-matrix techniques can then reduce the cost of Jacobian factorization}~\cite{superlu99}.
+\revise{In addition, studies have shown that} Single-Instruction, Multiple-Data (SIMD) and the related Single-Instruction, Multiple-Thread (SIMT) processors can accelerate chemical kinetic simulations~\cite{Shi:2012aa,Niemeyer:2014aa,Sewerin20151375,Niemeyer:2016aa,CurtisGPU:2017,stone2018}.
 
@@
 The parallel programming standard OpenCL~\cite{stone2010opencl} has further enabled adoption of vector processing in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems, e.g., CPU, GPU, or Intel's Many Integrated Core (MIC) architecture.
-Here we will largely use OpenCL terminology to describe these processing paradigms, as it provides a convenient way to classify otherwise disparate processor types (e.g., CPUs and GPUs).
-However, the concepts discussed herein broadly apply to SIMD\slash SIMT processing.
+\revise{Here we will} largely use OpenCL terminology to describe these processing paradigms, as it provides a convenient way to classify otherwise disparate processor types (e.g., CPUs and GPUs).
+However, the concepts discussed herein \revise{broadly apply} to SIMD\slash SIMT processing.
 
@@
 
-A typical modern CPU contains multiple compute units (i.e., cores), each with specialized vector processing units capable of running SIMD instructions, as \cref{F:SIMD} depicts.
-A SIMD instruction uses the vector processor to execute the same floating-point operation (e.g., multiplication, division) on different data concurrently;
-the number of concurrent operations possible is known as the vector-width\footnote{OpenCL allows for use of vector-widths different from the actual hardware vector-width via implicit conversion, and may provide some performance benefit as will be studied in Sec.~\ref{S:results}} and is typically around 2--4 double precision operations.
-Additionally, specialized hardware accelerators, e.g., Intel's Xeon Phi co-processor (MIC architecture) have been developed that have tens of cores with very wide vector-widths (e.g., \numrange{4}{8} double precision operations); these very wide vector-widths are also found on cutting-edge and forthcoming Intel CPUs; i.e., the Skylake Xeon and Cannon Lake architectures, respectively.
-
-SIMT processing---the foundation of modern GPUs~\cite{lindholm2008nvidia}---is a related computing paradigm that hosts large numbers of threads on a single compute element (a streaming multiprocessor in NVIDIA terminology).
-A group of threads---typically \num{32}, known as a warp on NVIDIA GPUs---execute the same SIMD instruction on multiple data concurrently (\cref{F:SIMT}).
-If some threads must execute a different instruction---e.g., due to if\slash then branching or predication---they are forced to wait and execute later.
-This phenomenon, known as thread-divergence, is a key consideration for SIMT-processing and can cause serious performance degradation for complicated algorithms~\cite{CurtisGPU:2017}.
-
-\subsection{Previous works and goals of this study}
-\label{S:Goals}
-
-Recognizing the need to accelerate Jacobian evaluation and factorization, a number of recent works have been published on analytical chemical kinetic Jacobian evaluation; although as will be discussed at the end of this section, this work offers several key improvements over past efforts.
+A typical modern CPU \revise{contains multiple} compute units (i.e., cores), each with specialized vector processing units capable of running SIMD instructions, \revise{as \cref{F:SIMD} depicts}.
+A SIMD instruction uses the vector processor to execute the same floating-point operation (e.g., multiplication, division) \revise{on different data} concurrently.
+\revise{The vector-width is the number of possible concurrent operations, typically around two to four in double precision.}\footnote{OpenCL allows for use of vector-widths different from the actual hardware vector-width via implicit conversion, and may provide some performance benefit as Sec.~\ref{S:results} \revise{discusses.}}
+Specialized hardware accelerators \revise{have also been developed, like Intel's Xeon Phi co-processor (i.e., the MIC architecture)}, that have tens of cores with wide vector-widths (e.g., \numrange{4}{8} double-precision operations).
+\revise{Cutting-edge and forthcoming Intel CPUs also include these wide vector-widths, like the Skylake Xeon and Cannon Lake architectures.}
+
+\revise{Modern GPUs rely on the related computing paradigm of SIMT processing, where a single compute element hosts large
+numbers of threads (a streaming multiprocessor in Nvidia terminology)~\cite{lindholm2008Nvidia}.}
+\revise{Figure~\ref{F:SIMT} depicts a SIMT compute unit, where a group of threads---typically \num{32}, known as a warp on Nvidia GPUs---execute the same SIMT instruction on multiple data concurrently}.
+If some threads must execute a different instruction, they are forced to wait and execute later; \revise{this may occur due to if\slash then branching or predication.}
+This phenomenon, known as thread-divergence, is a key consideration for SIMT processing and can cause serious performance degradation for complicated algorithms~\cite{CurtisGPU:2017}.
+
+\subsection{Related work}
+\label{S:related}
+
+Recognizing the need to accelerate chemical-kinetic Jacobian evaluation and factorization, a number of recent works have been published on \revise{constructing analytical Jacobian matrices}; although as will be discussed at the end of this section, \revise{here we offer} several key improvements over past efforts.
 \add{Schwer et al.~\cite{SCHWER2002270} were among the first to recognize the critical importance of a sparse analytical Jacobian to accelerate chemical kinetic simulations.}
@@
 
+\subsection{Goals of this study}
+\label{S:Goals}
+
 This work will build upon our previous analytical chemical kinetic Jacobian code, \texttt{pyJac}~\cite{Niemeyer:2016aa}, in order to:
@@
   \end{minipage}
-  \caption{\add{Simple data-layout patterns for 2-D arrays}}
+  \caption{\addnl{Simple data-layout patterns for 2-D arrays}}
 \end{figure}
@@
 If the data is stored in ``F''-order, the SIMD-lanes\slash SIMT-thread accessing $[C]_{1, j}\ldots[C]_{N_v, j}$---the concentration of species $j$ for states $1, 2,\ldots N_v$, where $N_v$ is the SIMD vector-width or the number of threads in a SIMT warp---will load sequential locations in memory.
-The first $(j+1)$th species concentration, $[C]_{1, j+1}$, will be $N_{\text{state}}$ memory locations away; this increases the likelihood of cache-misses on the CPU~\cite{gray2000rules}, but conversely is well suited to coalesced memory accesses on the GPU~\cite{NVIDIA:2018}.
+The first $(j+1)$th species concentration, $[C]_{1, j+1}$, will be $N_{\text{state}}$ memory locations away; this increases the likelihood of cache-misses on the CPU~\cite{gray2000rules}, but conversely is well suited to coalesced memory accesses on the GPU~\cite{Nvidia:2018}.
 
@@
 Additionally, \texttt{v5.0} of the LLVM\slash\texttt{clang}~\cite{Lattner:2004:LCF:977395.977673} compiler chain was installed on all machines to enable use of POCL.
-\Cref{t:gpus} lists the model, number of CUDA cores and NVIDIA driver of each of the GPUs used in this work.
-NVIDIA's OpenCL runtime is bundled with the NVIDIA driver~\cite{NVIDIA:2018}, hence the driver version is used to specify the OpenCL runtime version.
+\Cref{t:gpus} lists the model, number of CUDA cores and Nvidia driver of each of the GPUs used in this work.
+Nvidia's OpenCL runtime is bundled with the Nvidia driver~\cite{Nvidia:2018}, hence the driver version is used to specify the OpenCL runtime version.
 
@@
 \toprule
-NVIDIA Model   & Tesla C2075    & Tesla K40m    \\
+Nvidia Model   & Tesla C2075    & Tesla K40m    \\
 \midrule
@@
 \end{tabular}
-\caption{The NVIDIA GPU configurations used in this study.  NVIDIA's OpenCL runtime is provided with the graphics driver, rather than any specific version of CUDA.
+\caption{The Nvidia GPU configurations used in this study.  Nvidia's OpenCL runtime is provided with the graphics driver, rather than any specific version of CUDA.
 Again, the identifier field will be used as a shorthand descriptor during analysis of the performance results.
@@
 \end{table}
-\footnotetext{Due to a driver implementation issue~\cite{nvidia_memory}, the total memory used on the \gpuold/ and \gpunew/ GPUs were to limited to \SI{4}{\giga\byte} and \SI{10}{\giga\byte}, respectively.\addtocounter{footnote}{1}}
+\footnotetext{Due to a driver implementation issue~\cite{Nvidia_memory}, the total memory used on the \gpuold/ and \gpunew/ GPUs were to limited to \SI{4}{\giga\byte} and \SI{10}{\giga\byte}, respectively.\addtocounter{footnote}{1}}
 
 \Cref{t:platforms} lists the platforms and vectorization\slash execution patterns that they are capable of running.
-The Intel and NVIDIA OpenCL runtimes lack implementations of atomic operations on double precision variables; these are currently necessary for \texttt{pyJac} to run deep-vectorized code.
+The Intel and Nvidia OpenCL runtimes lack implementations of atomic operations on double precision variables; these are currently necessary for \texttt{pyJac} to run deep-vectorized code.
 On the other hand, POCL is an open source OpenCL runtime that works on all CPU types tested in this work, and does implement these atomic operations.
@@
 Intel OpenCL & X & X & -- \\
-NVIDIA OpenCL & -- & X & -- \\
+Nvidia OpenCL & -- & X & -- \\
 \bottomrule
@@
 In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source-terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean run time per condition; the run time continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
-\Cref{F:gpu_source_speedup} shows the speedup in source-term evaluations that the \gpunew/ GPU achieves over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} with two different vector-widths (a.k.a., the block size in NVIDIA terminology), \num{64} and \num{128}; the fastest \gpunew/ case with a vector-width of 128 is \SIrange{1.40}{1.88}{$\times$} faster than the slowest case (\gpuold/ with a vector-width of \num{64}) depending on the chemical model in question.
+\Cref{F:gpu_source_speedup} shows the speedup in source-term evaluations that the \gpunew/ GPU achieves over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} with two different vector-widths (a.k.a., the block size in Nvidia terminology), \num{64} and \num{128}; the fastest \gpunew/ case with a vector-width of 128 is \SIrange{1.40}{1.88}{$\times$} faster than the slowest case (\gpuold/ with a vector-width of \num{64}) depending on the chemical model in question.
 Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector-width has minimal effect on performance for most of the \gpunew/ and \gpuold/ cases; the GRI-Mech 3.0 and USC-Mech II models show the largest improvements with a vector-width of \num{128}, \textasciitilde\SIrange{10}{18}{\percent} for both GPUs.
-This is likely caused by higher occupancy on the GPUs, however it is unclear exactly how NVIDIA's OpenCL runtime balances the registers\slash warps per streaming-multiprocessor, as controls occupancy in CUDA~\cite{occupancy}; use of even larger vector-widths (i.e., \num{256}\texttt{+}) on the GPU is a topic that merits future investigation.
+This is likely caused by higher occupancy on the GPUs, however it is unclear exactly how Nvidia's OpenCL runtime balances the registers\slash warps per streaming-multiprocessor, as controls occupancy in CUDA~\cite{occupancy}; use of even larger vector-widths (i.e., \num{256}\texttt{+}) on the GPU is a topic that merits future investigation.
 \begin{figure}[htbp]
@@
   \end{subfigure}
-  \caption{\texttt{pyJac} source-term evaluation performance on the NVIDIA GPUs}
+  \caption{\texttt{pyJac} source-term evaluation performance on the Nvidia GPUs}
   \label{F:gpu_source}
@@
       \includegraphics[width=\textwidth]{source_gpu_cvsf.pdf}
-      \caption{The affect on performance of ``C'' vs ``F''-ordering for the shallow-vectorized NVIDIA OpenCL code on both GPUs with a vector-width of \num{128}.}
+      \caption{The affect on performance of ``C'' vs ``F''-ordering for the shallow-vectorized Nvidia OpenCL code on both GPUs with a vector-width of \num{128}.}
       \label{F:source_gpu_cvsf}
@@
 It is important to note that while OpenCL provides an simple interface to enable cross-platform execution, and significant speedups were achieved via shallow-vectorized OpenCL code in this work, there are some serious potential pit-falls in its use.
-The closed source OpenCL runtimes tested in this work (Intel and NVIDIA) contain several bugs that result in compilation failures, simply incorrect vectorized machine code or even segmentation faults.
-Further, these runtimes (in our experience) tend to be less responsive to fixing said bugs, with relatively infrequent new releases (or in NVIDIA's case, lack of changelogs\slash public records of bug-fixes).
+The closed source OpenCL runtimes tested in this work (Intel and Nvidia) contain several bugs that result in compilation failures, simply incorrect vectorized machine code or even segmentation faults.
+Further, these runtimes (in our experience) tend to be less responsive to fixing said bugs, with relatively infrequent new releases (or in Nvidia's case, lack of changelogs\slash public records of bug-fixes).
 On the other hand, the open-source OpenCL runtime used in this work, POCL, has (in our experience) far fewer implementation bugs, and when issues arise the community is very responsive to bug reports and user outreach; however, POCL fails to achieve vectorization as noted in~\cref{S:source_results}.
 
-As demonstration of the type of issue discussed, a minimum working example has been created~\cite{nvidia_mwe} that demonstrates a failure of NVIDIA's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla \gpunew/ GPU; by simply changing to another runtime (e.g., Intel)---with no code changes or recompilation---the correct result can be found.
+As demonstration of the type of issue discussed, a minimum working example has been created~\cite{Nvidia_mwe} that demonstrates a failure of Nvidia's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla \gpunew/ GPU; by simply changing to another runtime (e.g., Intel)---with no code changes or recompilation---the correct result can be found.
 This provides a particularly vexing problem for the programmer, as there is often little that can be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
@@
 Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler \texttt{\#pragma}s) is a key priority going forwards as OpenMP is a standard library on most machines.
-In addition, CUDA~\cite{NVIDIA:2018} has been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017}, while Intel's open-source OpenCL alternative, ISPC~\cite{pharr2012ispc}, has been relatively stable and easy to work with during preliminary efforts with the unit-testing discussed in~\cref{s:unittest}.
+In addition, CUDA~\cite{Nvidia:2018} has been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017}, while Intel's open-source OpenCL alternative, ISPC~\cite{pharr2012ispc}, has been relatively stable and easy to work with during preliminary efforts with the unit-testing discussed in~\cref{s:unittest}.
 The current deep-vectorization formulation would be executable for both CUDA and ISPC targets, as these languages implement double-precision atomic operations, further recommending their use.
@@
 This section gives more detail on the results presented in~\cref{S:jac_valid}, breaking down the reported error statistics per test-platform\slash language.
-The error of the Intel OpenCL runtime is presented in~\cref{T:intel_error}, the Portable OpenCL (POCL) runtime in~\cref{T:pocl_error}, OpenMP in~\cref{T:omp_error} and the NVIDIA OpenCL runtime in~\cref{T:nv_error}.
-POCL and OpenMP tend to have the smallest error norms, while NVIDIA tends to have the largest; in particular the stringent filtered error norm $E_{\mathcal{C} = 10^{20}}$ is two orders of magnitude larger for the NVIDIA runtime than the other test platforms with the \ce{H2}\slash\ce{CO} and USC-Mech II models.
+The error of the Intel OpenCL runtime is presented in~\cref{T:intel_error}, the Portable OpenCL (POCL) runtime in~\cref{T:pocl_error}, OpenMP in~\cref{T:omp_error} and the Nvidia OpenCL runtime in~\cref{T:nv_error}.
+POCL and OpenMP tend to have the smallest error norms, while Nvidia tends to have the largest; in particular the stringent filtered error norm $E_{\mathcal{C} = 10^{20}}$ is two orders of magnitude larger for the Nvidia runtime than the other test platforms with the \ce{H2}\slash\ce{CO} and USC-Mech II models.
 
@@
 \end{tabular}
-\caption{Summary of Jacobian matrix validation results for NVIDIA OpenCL execution.
+\caption{Summary of Jacobian matrix validation results for Nvidia OpenCL execution.
 The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/ and sparse\slash dense Jacobians.

### 2018-06-04 paper.tex
edits through page 7

@@
 \begin{abstract} % not to exceed 200 words
-A code generation platform, \texttt{pyJac}, for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized \revise{constant-pressure\slash volume} thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation has been developed.% Can you introduce this a bit more?
+A code generation platform, \texttt{pyJac}, for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized \revise{constant-pressure\slash volume} thermochemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation has been developed.% Can you introduce this a bit more?
 good to start with at least one sentence of introduction.
@@
 A host of techniques have been developed to lessen the computational demand of chemical kinetic calculations while maintaining fidelity, \revise{falling} broadly into \revise{three} categories: removal of unimportant species and reactions~\cite{Lu:2006bb,Pepiot-Desjardins:2008,Hiremath:2010jw,Niemeyer:2010bt,Curtis:2015},
-lumping of species with similar thermo-chemical properties~\cite{Lu:2007,Ahmed:2007fa,Pepiot:2008kq},
+lumping of species with similar thermochemical properties~\cite{Lu:2007,Ahmed:2007fa,Pepiot:2008kq},
 and time-scale methods that reduce numerical stiffness~\cite{Maas:1992ws,Lam:1994ws,Lu:2001ve,Gou:2010}.
@@
 \add{Schwer et al.~\cite{SCHWER2002270} were among the first to recognize the critical importance of a sparse analytical Jacobian to accelerate chemical kinetic simulations.}
-The \texttt{TChem} software package~\cite{Safta:2011vn} was one of the first codes developed providing analytical Jacobian evaluation, but has several limitations including: incompatibilities with modern reaction types---i.e., pressure dependent Arrhenius (or P-Log) and Chebyshev reactions---and lack of thread-safety to enable parallel execution~\cite{Curtis2017:tchem}.
+\revise{Later, Safta et al.~\cite{Safta:2011vn} developed the \texttt{TChem} software package}, which was one of the first developed that provides analytical Jacobian evaluation.
+However, \texttt{TChem} has several limitations, including incompatibility with modern reaction types---i.e., pressure dependent Arrhenius (or P-Log) and Chebyshev reactions---and its lack of thread-safety to enable parallel execution~\cite{Curtis2017:tchem}.
 Youssefi~\cite{Youssefi:2011tm} explored the importance of analytical Jacobian matrices for time-scale analysis techniques as well as their effect on computational efficiency in zero-dimensional homogeneous reactor simulations.
-Bisetti~\cite{Bisetti:2012jw} developed an isothermal, isobaric analytical Jacobian code-generation utility; although the chosen isothermal assumption is not typical in most combustion simulations, a significant increase in Jacobian sparsity was exhibited.
-Additionally, in the same work Bisetti provided a novel way to compute dense matrix-vector multiplications resulting from a change of system variables without need for storage of the full dense Jacobian.
-Perini et al.~\cite{Perini:2012gy} developed an analytical Jacobian code for constant-volume combustion, with additional options to increase sparsity (at expense of strict correctness), and enable tabulation of temperature dependent properties; an \SI{80}{\percent} speedup over a finite-difference-based Jacobian was reported when used in a multidimensional reactive-flow simulation.
-Dijkmans et al.~\cite{Dijkmans:2014bb} developed a GPU-based analytical Jacobian code with optional tabulation of temperature-dependent properties, and found speedups up to \SI{120}{$\times$} for zero-dimensional chemical kinetic integration with large chemical models (\textasciitilde\num{3000} species) using a Tesla C2075 GPU.
-\add{Bauer et al.~\cite{Bauer:2014} used warp-specialization to improve GPU-vectorization over a standard data-parallel vectorization approach; speedups up to \SIrange{2.81}{3.75}{$\times$}, \SIrange{1.91}{2.58}{$\times$}, and \SIrange{1.4}{1.5}{$\times$} were achieved on a Telsa K20c GPU\todo{keep hardware spec?} for evaluation of viscosity, species diffusion and chemical source terms, respectively}.
-Niemeyer et al.~\cite{Niemeyer:2016aa} created and validated the open-source analytical chemical kinetic Jacobian code-generator, \texttt{pyJac}, supporting parallel execution on the CPU and SIMT execution on the GPU; a speedup of \SIrange{3}{7.5}{$\times$} over a finite-difference Jacobian was found on the CPU.
-Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but it was not validated outside the context of use with an implicit-integration technique; in addition, since the Jacobian was based on an over-constrained system~\cite{HANSEN2018257}, the effect on strict conservation of mass\slash energy was not studied.
-\add{Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermo-chemical state-vectors and their resulting effect on consistency and errors in conserved properties such as mass and energy; further, the effect on implicit\slash linearly-implicit integration techniques and analysis techniques such as chemical explosive mode analysis~\cite{lu_yoo_chen_law_2010} were characterized.}
-
-\todo[inline]{I don't know that I actually want to remove the processor information alone here -- it doesn't save us much space.
-It would be more valuable to re-write this entirely, but I'm not sure that's necessary either}
-A number of recent works have investigated the use of high-performance SIMT-devices to accelerate reactive-flow and chemical kinetics simulations.
-Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup \delete{on a Tesla C1060} compared to a serial implementation \add{CPU}\delete{on a AMD-Opteron processor}.
-Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order of magnitude speedup for homogeneous reactor simulations of large chemical models\delete{on a Tesla 2050 GPU} over a serial\delete{Intel i7 930} CPU implementation.
+Bisetti~\cite{Bisetti:2012jw} developed an isothermal, isobaric analytical Jacobian code-generation utility;
+\revise{this approach significantly increases Jacobian sparsity, although the chosen isothermal assumption is not typical in most combustion simulations.}
+In the same work Bisetti \revise{also} provided a novel way to compute dense matrix-vector multiplications resulting from a change of system variables without \revise{storing} the full Jacobian.
+Perini et al.~\cite{Perini:2012gy} developed an analytical Jacobian code for constant-volume combustion, with additional options to increase sparsity (at the expense of strict correctness) and \revise{tabulate} temperature-dependent properties; \revise{they reported} an \SI{80}{\percent} speedup over a finite-difference-based Jacobian when used in a multidimensional reactive-flow simulation.
+Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but did not validate it outside the context of use with an implicit-integration technique.
+In addition, since the Jacobian was based on an over-constrained system~\cite{HANSEN2018257}, the effect on strict conservation of mass\slash energy was not studied.
+
+\revise{Recently, some groups have developed frameworks for constructing analytical Jacobians for evaluation on modern SIMD or SIMT processors.}
+Dijkmans et al.~\cite{Dijkmans:2014bb} developed a GPU-based analytical Jacobian code with optional tabulation of temperature-dependent properties, and \revise{showed} speedups up to \SI{120}{$\times$} for zero-dimensional chemical kinetic integration with large chemical models (\textasciitilde\num{3000} species).
+\add{Bauer et al.~\cite{Bauer:2014} used warp-specialization to improve GPU-vectorization over a standard data-parallel vectorization approach; they achieved speedups of up to \SIrange{2.81}{3.75}{$\times$}, \SIrange{1.91}{2.58}{$\times$}, and \SIrange{1.4}{1.5}{$\times$} for evaluating viscosity, species diffusion, and chemical source terms, respectively}.
+Niemeyer et al.~\cite{Niemeyer:2016aa} created and validated the open-source analytical chemical kinetic Jacobian code-generator, \texttt{pyJac}, which supports parallel execution on CPUs and SIMT execution on GPUs; \revise{\texttt{pyJac} enables} a speedup of \SIrange{3}{7.5}{$\times$} over a finite-difference Jacobian on the CPU.
+
+\add{Relevant to all of the aforementioned efforts, Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermochemical state vectors and the resulting effect on consistency and errors in conserved properties such as mass and energy.
+They also characterized how the choice of state vector affects implicit\slash linearly implicit integration algorithms and chemical mode analysis techniques.
+Overall they found that while many literature Jacobian formulations are not strictly correct or over-specified, such flaws negligibly affect Newton--Krylov methods---perhaps because the incorrect Jacobian reasonably approximates the true Jacobian.
+On the other hand, linearly implicit algorithms like Rosenbrock methods and analysis techniques like chemical explosive mode analysis~\cite{lu_yoo_chen_law_2010} need accurate and correct Jacobians.
+}
+
+A number of recent works have investigated \revise{using} high-performance SIMT devices \revise{like GPUs} to accelerate reactive-flow and chemical kinetics simulations.
+Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup \delete{on a Tesla C1060} compared to a \add{CPU-based} serial implementation \delete{on a AMD-Opteron processor}.
+Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order-of-magnitude speedup for homogeneous reactor simulations of large chemical models\delete{on a Tesla 2050 GPU} over a serial\delete{Intel i7 930} CPU implementation.
 Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential \add{CPU} code\delete{on an Intel Xeon X5650 CPU}.
-Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, a traditional CPU-based implicit integrator handled the stiffest computational cells while a stabilized-explicit GPU-solver\delete{on a Tesla C2050 GPU} solved the less stiff cells; this hybrid solution technique was \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
-Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code\delete{ on a Tesla C2070} as compared to a sequential \add{CPU} version of the same\delete{on an Intel Xeon X5650 CPU}.
-Stone and Davis~\cite{Stone:2013aa} investigated a GPU-version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order of magnitude speedup over a serial CPU implementation.
+Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, where a traditional implicit integrator handled the stiffest computational cells on a CPU and a stabilized-explicit solver \delete{on a Tesla C2050 GPU} solved the less-stiff cells on a GPU; this hybrid solution technique \revise{performs} \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
+Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code\delete{ on a Tesla C2070} as compared with a sequential \add{CPU} version of the same\delete{on an Intel Xeon X5650 CPU}.
+Stone and Davis~\cite{Stone:2013aa} investigated a GPU-based version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order-of-magnitude speedup over a serial CPU implementation.
 Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup\delete{on a Tesla C2075 GPU} over a multithreaded VODE solver on a six-core\delete{Intel X5650} CPU.
-Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge-Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
-Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs)\delete{on an Intel Xeon E5-2680} \add{both the CPU} and\delete{a Tesla K20} GPU.
-In evaluation of the thermo-chemical properties, speedups between \SIrange{8}{13}{$\times$} were found (over a serial version of the same code) while running on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU, respectively.
-Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa}\delete{on a Tesla C2075} \add{on the GPU and CPU}; the \add{GPU-based} implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38}\delete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models with an integration time step of \SI{e-6}{$\sec$}.
+Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge--Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
+Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates for solving partial differential equations (PDEs) \delete{on an Intel Xeon E5-2680} \add{both the CPU} and \delete{a Tesla K20} GPU.
+In evaluating the thermochemical properties, speedups between \SIrange{8}{13}{$\times$} were found (over a serial version of the same code) while running on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU, respectively.\todo{revise this sentence, fairly awkward}{}
+Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa}\delete{on a Tesla C2075} \add{on the GPU and CPU}.
+The \add{GPU-based} implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38}\delete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models \delete{with an integration time step of $10^{-6}$ s}.
 
@@
 \Cref{F:mem} depicts an example data array with $K$ rows and $N$ columns, and index ($i$, $j$) corresponding to the $i$th row and $j$th column.
-For example, the concentration of species $j$ for the $i$th thermo-chemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{state}$ (the number of thermo-chemical states considered for evaluation), and $1 \le j \le \ns/$ (the number of species in the model).
+For example, the concentration of species $j$ for the $i$th thermochemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{state}$ (the number of thermochemical states considered for evaluation), and $1 \le j \le \ns/$ (the number of species in the model).
 The stored concentrations would then have $K = N_{\text{state}}$ rows, and $N = N_{\text{sp}}$ columns.
 
-In the ``C'' (C row-major) format, the concentrations of all species for a single thermo-chemical condition $i$ would be stored sequentially in memory, i.e., with  $[C]_{1, 1}$ in index \num{1}, $[C]_{1, 2}$ in index \num{2}, etc., as shown in \cref{F:row_major}.
-Conversely in the ``F'' (Fortran column-major) format, the concentrations of a single species $j$ over all thermo-chemical states are adjacent in memory, corresponding to storing $[C]_{1, 1}$ in index \num{1}, $[C]_{2, 1}$ in index \num{2} and so on, as shown in \cref{F:column_major}.
+In the ``C'' (C row-major) format, the concentrations of all species for a single thermochemical condition $i$ would be stored sequentially in memory, i.e., with  $[C]_{1, 1}$ in index \num{1}, $[C]_{1, 2}$ in index \num{2}, etc., as shown in \cref{F:row_major}.
+Conversely in the ``F'' (Fortran column-major) format, the concentrations of a single species $j$ over all thermochemical states are adjacent in memory, corresponding to storing $[C]_{1, 1}$ in index \num{1}, $[C]_{2, 1}$ in index \num{2} and so on, as shown in \cref{F:column_major}.
 This ordering---along with the device (CPU, GPU, etc.\@) and vectorization pattern in question---have a large effect on the performance of SIMD\slash SIMT-vectorized algorithms.
 
-In a \textit{shallow}-SIMD\slash SIMT vectorization (also referred to as ``per-thread'' in previous works~\cite{Stone:2013aa} using GPUs), each SIMD-lane (or SIMT-thread) in a compute-unit evaluates the thermo-chemical source-terms\slash chemical kinetic Jacobian for a different thermo-chemical state.
+In a \textit{shallow}-SIMD\slash SIMT vectorization (also referred to as ``per-thread'' in previous works~\cite{Stone:2013aa} using GPUs), each SIMD-lane (or SIMT-thread) in a compute-unit evaluates the thermochemical source-terms\slash chemical kinetic Jacobian for a different thermochemical state.
 If the data is stored in ``F''-order, the SIMD-lanes\slash SIMT-thread accessing $[C]_{1, j}\ldots[C]_{N_v, j}$---the concentration of species $j$ for states $1, 2,\ldots N_v$, where $N_v$ is the SIMD vector-width or the number of threads in a SIMT warp---will load sequential locations in memory.
@@
 
-In a \textit{deep}-SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works (e.g.,~\cite{Stone:2013aa,CurtisGPU:2017}), a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermo-chemical source-terms for a single thermo-chemical state; thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + N_v}$---the species concentrations for state 1, species $j \ldots j + N_v$---will access sequential memory locations if the data is stored in ``C''-order.
-Further, in ``C''-ordering the maximum distance between any two species concentrations within the same thermo-chemical state is $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU, but may lead to uncoalesced memory-accesses on the GPU.
+In a \textit{deep}-SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works (e.g.,~\cite{Stone:2013aa,CurtisGPU:2017}), a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermochemical source-terms for a single thermochemical state; thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + N_v}$---the species concentrations for state 1, species $j \ldots j + N_v$---will access sequential memory locations if the data is stored in ``C''-order.
+Further, in ``C''-ordering the maximum distance between any two species concentrations within the same thermochemical state is $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU, but may lead to uncoalesced memory-accesses on the GPU.
 In addition, a deep vectorization requires both synchronization between SIMD-lanes\slash SIMT-threads via memory-fences\slash barriers (an expensive operation), and may result in SIMD-waste\slash SIMT thread-divergence, caused by different SIMD-lanes\slash SIMT-threads executing different instructions (e.g., resulting from different if\slash then branches).
@@
 
-\subsection{Thermo-chemical Source-Terms and Jacobian}
-This new version of \texttt{pyJac} is capable of evaluating the thermo-chemical source-terms for using both the constant-pressure (\conp/) or constant-volume (\conv/) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
+\subsection{thermochemical Source-Terms and Jacobian}
+This new version of \texttt{pyJac} is capable of evaluating the thermochemical source-terms for using both the constant-pressure (\conp/) or constant-volume (\conv/) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
 In this section, we will only outline a brief summary of the system evaluated by \texttt{pyJac}; the interested reader is referred to the supplemental material for the complete derivations.
 
-The thermo-chemical state vector derivation consists of the temperature, a non-constant thermodynamic state parameter (volume or pressure for \conp/ and \conv/, respectively), and the number of moles of all species except the last species in the chemical model which is typically taken to be the bath-gas (e.g., \ce{N2}):
+The thermochemical state vector derivation consists of the temperature, a non-constant thermodynamic state parameter (volume or pressure for \conp/ and \conv/, respectively), and the number of moles of all species except the last species in the chemical model which is typically taken to be the bath-gas (e.g., \ce{N2}):
 \loadeq{state}
@@
 
-The evolution of the thermo-chemical state vector is described by a set of chemical kinetic ordinary differential equations:
+The evolution of the thermochemical state vector is described by a set of chemical kinetic ordinary differential equations:
 \loadeq{dstate}
@@
 The reasons for this choice are explained fully in~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with minimal additional programming effort.
-The discrepancy between the analytical and automatic differentiated Jacobians for thermo-chemical state $k$, denoted by $\mathcal{J}_k$ and $\hat{\mathcal{J}}_k$ respectively, is determined by the relative error Frobenius norm over all Jacobian indices $i, j$:
+The discrepancy between the analytical and automatic differentiated Jacobians for thermochemical state $k$, denoted by $\mathcal{J}_k$ and $\hat{\mathcal{J}}_k$ respectively, is determined by the relative error Frobenius norm over all Jacobian indices $i, j$:
 \begin{equation}
@@
 
-To avoid large relative errors of very small nonzero Jacobian elements due to accumulation of floating point error, the Frobenius norm of the automatically differentiated Jacobian is calculated over all thermo-chemical states $k$:
+To avoid large relative errors of very small nonzero Jacobian elements due to accumulation of floating point error, the Frobenius norm of the automatically differentiated Jacobian is calculated over all thermochemical states $k$:
 \begin{equation}
@@
 The error statistics reported in this section are then based only on matrix elements where $\mathcal{J}_{ijk} \ge \frac{\mathcal{T}}{\mathcal{C}}$, where $\mathcal{C}$ is the threshold parameter; this filtered form of~\cref{e:jac_error_base} is denoted $E_{\mathcal{C},k}$.
-Finally, the Frobenius norm is calculated over all the states in the PaSR thermo-chemical condition database:
+Finally, the Frobenius norm is calculated over all the states in the PaSR thermochemical condition database:
 \begin{equation}
@@
 \Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source-term evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
-In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for both Intel OpenCL and OpenMP, likely resulting from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them); \add{as demonstrated in~\cref{S:SIMD_scaling} the amount of computational work required per thermo-chemical state plays a critical role in the full utilization of SIMD instructions\slash multiple threads}.
+In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for both Intel OpenCL and OpenMP, likely resulting from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them); \add{as demonstrated in~\cref{S:SIMD_scaling} the amount of computational work required per thermochemical state plays a critical role in the full utilization of SIMD instructions\slash multiple threads}.
 Additionally, OpenMP tends to have higher scalings than the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores, respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.
@@
 In this case, the isopentanol model has the worst scaling for both codes.
-Tt is noted that due to the sheer size of the dense isopentanol Jacobian---storing the dense matrix for a single thermo-chemical state takes over \SI{1}{\mega\byte} of data---the total number of states for the dense isopentanol Jacobian evaluation was limited to \num{50000} (which requires over \SI{50}{\giga\byte} of memory); this greatly drops the computation cost for this case, and adversely affects the scaling efficiency as was discussed in~\cref{S:source_results}.
+Tt is noted that due to the sheer size of the dense isopentanol Jacobian---storing the dense matrix for a single thermochemical state takes over \SI{1}{\mega\byte} of data---the total number of states for the dense isopentanol Jacobian evaluation was limited to \num{50000} (which requires over \SI{50}{\giga\byte} of memory); this greatly drops the computation cost for this case, and adversely affects the scaling efficiency as was discussed in~\cref{S:source_results}.
 Excluding the isopentanol model, the dense shallow-vectorized OpenCL scalings are slightly higher than for source-term evaluation, e.g., \numrange{0.70}{0.78} and \numrange{0.52}{0.61} for \num{4} and \num{8} cores, respectively (compared to \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for shallow-vectorized OpenCL source-term on the same \num{4} and \num{8} cores); this is due to the higher computational cost of Jacobian evaluation, and hence more available work per CPU-core.
@@
 \section{Conclusions}
-In this work, automatically generated OpenCL codes for SIMD\slash SIMT-vectorized thermo-chemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation were developed and validated for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
+In this work, automatically generated OpenCL codes for SIMD\slash SIMT-vectorized thermochemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation were developed and validated for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
 \add{Reviewing the goals of this work, as discussed in~\cref{S:Goals}}:
@@
 \add{This simple example demonstrates the dependency of the SIMD efficiency of shallow-vectorized OpenCL source-term evaluation on the size of the chemical model in question, i.e., the amount of computational work per source-term evaluation.}
-\add{The base chemical model for this example was the isopentanol model~\cite{Sarathy:2013jr} used throughout this work, and the same thermo-chemical state database described in~\cref{s:test_platforms} was used for source-term evaluation.}
+\add{The base chemical model for this example was the isopentanol model~\cite{Sarathy:2013jr} used throughout this work, and the same thermochemical state database described in~\cref{s:test_platforms} was used for source-term evaluation.}
 \add{As in~\cref{S:results} all reported results are based on \num{10} individual runs and in this example all cases were run on the \avx/ machine using the Intel OpenCL runtime.}
@@
 \add{In order to discern the effect of varying chemical model size on the SIMD efficiency, shallow-vectorized and unvectorized source-term evaluation performance tests were run.}
-\add{As demonstrated in~\cref{F:simd_vs_rxns} the SIMD efficiency is strongly dependent on the generated model size and and thus, amount of computational work per thermo-chemical state.}
+\add{As demonstrated in~\cref{F:simd_vs_rxns} the SIMD efficiency is strongly dependent on the generated model size and and thus, amount of computational work per thermochemical state.}
 \add{In addition, the range of SIMD efficiency in this example (\numrange{0.56}{0.91}) is larger than the range of SIMD efficiencies calculated for real chemical models, as seen in~\cref{F:source_simd_scaling}.}

### 2018-06-04 paper.tex
a few more edits in page 7

@@
 Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates for solving partial differential equations (PDEs) \delete{on an Intel Xeon E5-2680} \add{both the CPU} and \delete{a Tesla K20} GPU.
-In evaluating the thermochemical properties, speedups between \SIrange{8}{13}{$\times$} were found (over a serial version of the same code) while running on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU, respectively.\todo{revise this sentence, fairly awkward}{}
-Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa}\delete{on a Tesla C2075} \add{on the GPU and CPU}.
-The \add{GPU-based} implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38}\delete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models \delete{with an integration time step of $10^{-6}$ s}.
+In evaluating the thermochemical properties, speedups between \SIrange{8}{13}{$\times$} were found (over a serial version of the same code) while running on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; they showed speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} when simulating a partially premixed methanol flame PDE on 16 CPU cores and the GPU, respectively.\todo{revise this sentence, fairly awkward}{}
+Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} \delete{on a Tesla C2075} \add{on the GPU and CPU}.
+The \add{GPU-based} implicit Runge--Kutta method performs equivalently to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38} \delete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models \delete{with an integration time step of $10^{-6}$ s}.
 

### 2018-06-05 paper.tex
edits through end of section 1

@@
 
-In contrast, SIMD-based chemical kinetics evaluation\slash integration have been far less studied.
-Linford et al.~\cite{Linford:2011} implemented a three-stage, second-order Rosenbrock integrator for atmospheric chemical kinetics on the CPU, GPU and cell croadband engine (CBE)---a specially designed vector processor---and found speedups regularly exceeding~\SI{25}{$\times$} over a serial CPU implementation.
+In contrast, SIMD-based chemical kinetics evaluation\slash integration have been studied far less.
+Linford et al.~\cite{Linford:2011} implemented a three-stage, second-order Rosenbrock integrator for atmospheric chemical kinetics on the CPU, GPU, and cell broadband engine (CBE)---a specially designed vector processor---and found speedups regularly exceeding~\SI{25}{$\times$} over a serial CPU implementation.
 Kroshko and Spiteri~\cite{kroshko2013efficient} implemented a SIMD-vectorized third-order stiff Rosenbrock integrator for atmospheric chemistry on the CBE and found a speedup of \SI{1.89}{$\times$} (a parallel scaling efficiency of \SI{94}{$\percent$}) over a serial version of the same code.
-Stone and Niemeyer~\cite{stone2018} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
-SIMD-vectorization improved the integrator performance over an OpenMP baseline (which was vectorized by simple compiler hints (i.e., \texttt{\#pragmas}) by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance was only \SIrange{1.4}{1.6}{$\times$} faster than the OpenMP baseline due to thread-divergence concerns~\cite{stone2018}.
+Stone et al.~\cite{stone2018} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
+They found that SIMD vectorization improves integrator performance over an OpenMP baseline vectorized by simple compiler hints (i.e., \texttt{\#pragmas}) by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performs only \SIrange{1.4}{1.6}{$\times$} faster than the OpenMP baseline due to thread divergence~\cite{stone2018}.
 
@@
 
-This work will build upon our previous analytical chemical kinetic Jacobian code, \texttt{pyJac}~\cite{Niemeyer:2016aa}, in order to:
+This work builds upon our previous analytical chemical kinetic Jacobian code, \texttt{pyJac}~\cite{Niemeyer:2016aa}.
+In this article we
 \begin{itemize}
- \item Derive and validate a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity
- \item Enable cross-platform SIMD\slash SIMT vectorization for the CPU, GPU and other accelerators
- \item Investigate the performance for a wide range chemical kinetic models, specifically looking at speedups due to SIMD\slash SIMT-vectorization as well as relative to the previous version of \texttt{pyJac}, and finally
+ \item Derive and validate a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity;
+ \item Enable cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
+ \item Investigate the performance for a wide range chemical kinetic models, specifically looking at speedups due to SIMD\slash SIMT-vectorization as well as relative to the previous version of \texttt{pyJac}; and finally
  \item Discuss future extensions to this work as well as several promising directions for SIMD\slash SIMT vectorization in reactive-flow simulations.
 \end{itemize}
-To our knowledge, this is the first \add{open-source, validated} effort that enables vectorized chemical-kinetic source-term and Jacobian evaluation for any chemical model on a wide-selection of platforms, i.e., the CPU, GPU and other accelerators.
+To our knowledge, this is the first \add{open-source, validated} effort that vectorizes the evaluation of chemical-kinetic source terms and Jacobian matrices for any chemical model on a wide selection of platforms.
 

### 2018-06-05 paper.tex
edits onto page 9

@@
 numbers of threads (a streaming multiprocessor in Nvidia terminology)~\cite{lindholm2008Nvidia}.}
-\revise{Figure~\ref{F:SIMT} depicts a SIMT compute unit, where a group of threads---typically \num{32}, known as a warp on Nvidia GPUs---execute the same SIMT instruction on multiple data concurrently}.
+\revise{\Cref{F:SIMT} depicts a SIMT compute unit, where a group of threads---typically \num{32}, known as a warp on Nvidia GPUs---execute the same SIMT instruction on multiple data concurrently}.
 If some threads must execute a different instruction, they are forced to wait and execute later; \revise{this may occur due to if\slash then branching or predication.}
@@
 When storing arrays for a chemical kinetic model, the data-storage layout and vectorization patterns are critical to achieving high-performance code.
-\Cref{F:mem} depicts an example data array with $K$ rows and $N$ columns, and index ($i$, $j$) corresponding to the $i$th row and $j$th column.
-For example, the concentration of species $j$ for the $i$th thermochemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{state}$ (the number of thermochemical states considered for evaluation), and $1 \le j \le \ns/$ (the number of species in the model).
-The stored concentrations would then have $K = N_{\text{state}}$ rows, and $N = N_{\text{sp}}$ columns.
-
-In the ``C'' (C row-major) format, the concentrations of all species for a single thermochemical condition $i$ would be stored sequentially in memory, i.e., with  $[C]_{1, 1}$ in index \num{1}, $[C]_{1, 2}$ in index \num{2}, etc., as shown in \cref{F:row_major}.
-Conversely in the ``F'' (Fortran column-major) format, the concentrations of a single species $j$ over all thermochemical states are adjacent in memory, corresponding to storing $[C]_{1, 1}$ in index \num{1}, $[C]_{2, 1}$ in index \num{2} and so on, as shown in \cref{F:column_major}.
-This ordering---along with the device (CPU, GPU, etc.\@) and vectorization pattern in question---have a large effect on the performance of SIMD\slash SIMT-vectorized algorithms.
-
-In a \textit{shallow}-SIMD\slash SIMT vectorization (also referred to as ``per-thread'' in previous works~\cite{Stone:2013aa} using GPUs), each SIMD-lane (or SIMT-thread) in a compute-unit evaluates the thermochemical source-terms\slash chemical kinetic Jacobian for a different thermochemical state.
+\Cref{F:mem} depicts an example data array with $K$ rows and $N$ columns \revise{where} index ($i$, $j$) \revise{corresponds} to the $i$th row and $j$th column.
+For example, the concentration of species $j$ for the $i$th thermochemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{\text{state}}$ (the number of thermochemical states considered for evaluation) and $1 \le j \le \ns/$ (the number of species in the model).
+The stored concentrations would then have $K = N_{\text{state}}$ rows and $N = \ns/$ columns.
+
+The ``C'' (C row-major) format stores the concentrations of all species for a single thermochemical condition $i$  sequentially in memory, i.e., with  $[C]_{1, 1}$ in index 1, $[C]_{1, 2}$ in index 2, and so on, as shown in \cref{F:row_major}.\todo{wouldn't these start at index 0?}{}
+Conversely, in the ``F'' (Fortran column-major) format the concentrations of a single species $j$ over all thermochemical states lie adjacent in memory, corresponding to storing $[C]_{1, 1}$ in index 1, $[C]_{2, 1}$ in index 2, and so on, as shown in \cref{F:column_major}.
+This ordering \revise{strongly affects} the performance of SIMD\slash SIMT-vectorized algorithms, \revise{as does the device (CPU, GPU, etc.\@) and vectorization pattern in question.}
+
+In a \textit{shallow}-SIMD\slash SIMT vectorization (also referred to as ``per-thread'' in previous works using GPUs~\cite{Stone:2013aa}), each SIMD lane (or SIMT thread) in a compute unit evaluates the source terms or Jacobian for a different thermochemical state.
 If the data is stored in ``F''-order, the SIMD-lanes\slash SIMT-thread accessing $[C]_{1, j}\ldots[C]_{N_v, j}$---the concentration of species $j$ for states $1, 2,\ldots N_v$, where $N_v$ is the SIMD vector-width or the number of threads in a SIMT warp---will load sequential locations in memory.

### 2018-06-05 paper.tex
edits through section 2.2

@@
 \revise{Later, Safta et al.~\cite{Safta:2011vn} developed the \texttt{TChem} software package}, which was one of the first developed that provides analytical Jacobian evaluation.
-However, \texttt{TChem} has several limitations, including incompatibility with modern reaction types---i.e., pressure dependent Arrhenius (or P-Log) and Chebyshev reactions---and its lack of thread-safety to enable parallel execution~\cite{Curtis2017:tchem}.
+However, \texttt{TChem} has several limitations, including incompatibility with modern reaction types---i.e., pressure-dependent Arrhenius (or P-Log) and Chebyshev reactions---and its lack of thread-safety to enable parallel execution~\cite{Curtis2017:tchem}.
 Youssefi~\cite{Youssefi:2011tm} explored the importance of analytical Jacobian matrices for time-scale analysis techniques as well as their effect on computational efficiency in zero-dimensional homogeneous reactor simulations.
@@
 Stone and Davis~\cite{Stone:2013aa} investigated a GPU-based version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order-of-magnitude speedup over a serial CPU implementation.
-Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup\delete{on a Tesla C2075 GPU} over a multithreaded VODE solver on a six-core\delete{Intel X5650} CPU.
+Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup\delete{on a Tesla C2075 GPU} over a multithreaded VODE solver on a six-core \delete{Intel X5650} CPU.
 Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge--Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
@@
 \begin{itemize}
- \item Derive and validate a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity;
+ \item Derive and validate a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity: \revise{\texttt{pyJac v2}};
  \item Enable cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
@@
 
-In a \textit{shallow}-SIMD\slash SIMT vectorization (also referred to as ``per-thread'' in previous works using GPUs~\cite{Stone:2013aa}), each SIMD lane (or SIMT thread) in a compute unit evaluates the source terms or Jacobian for a different thermochemical state.
-If the data is stored in ``F''-order, the SIMD-lanes\slash SIMT-thread accessing $[C]_{1, j}\ldots[C]_{N_v, j}$---the concentration of species $j$ for states $1, 2,\ldots N_v$, where $N_v$ is the SIMD vector-width or the number of threads in a SIMT warp---will load sequential locations in memory.
-The first $(j+1)$th species concentration, $[C]_{1, j+1}$, will be $N_{\text{state}}$ memory locations away; this increases the likelihood of cache-misses on the CPU~\cite{gray2000rules}, but conversely is well suited to coalesced memory accesses on the GPU~\cite{Nvidia:2018}.
-
-In a \textit{deep}-SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works (e.g.,~\cite{Stone:2013aa,CurtisGPU:2017}), a compute-unit utilizes its SIMD-lanes\slash SIMT-threads cooperatively to evaluate the thermochemical source-terms for a single thermochemical state; thus SIMD-lanes loading $[C]_{1, j} \ldots [C]_{1, j + N_v}$---the species concentrations for state 1, species $j \ldots j + N_v$---will access sequential memory locations if the data is stored in ``C''-order.
-Further, in ``C''-ordering the maximum distance between any two species concentrations within the same thermochemical state is $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU, but may lead to uncoalesced memory-accesses on the GPU.
-In addition, a deep vectorization requires both synchronization between SIMD-lanes\slash SIMT-threads via memory-fences\slash barriers (an expensive operation), and may result in SIMD-waste\slash SIMT thread-divergence, caused by different SIMD-lanes\slash SIMT-threads executing different instructions (e.g., resulting from different if\slash then branches).
-It is noted that generally speaking, shallow-vectorizations may also experience SIMD-waste\slash SIMT thread-divergence, e.g., in chemical kinetic integration due to varying internal solver time-step sizes~\cite{CurtisGPU:2017}; however, in this work shallow-vectorizations are largely unaffected by this concern \add{as the only major code-paths that differ between vector-lanes are high\slash low-temperature polynomial evaluations and differing pressures for P-Log reactions, a relatively small concern as compared to differing internal ODE integration time-steps~\cite{CurtisGPU:2017}}.
+In a \textit{shallow} SIMD\slash SIMT vectorization (also referred to as ``per-thread'' in previous works using GPUs~\cite{Stone:2013aa}), each SIMD lane or SIMT thread in a compute unit evaluates the source terms or Jacobian for a different thermochemical state.
+If the data is stored in ``F''-order, the SIMD lanes\slash SIMT thread accessing $[C]_{1, j}\ldots[C]_{N_v, j}$ will load sequential locations in memory, \revise{where $[C]_{i, j}$ is the concentration of species $j$ for state $i$ and $N_v$ is the SIMD vector-width or the number of threads in a SIMT warp}.
+The first $(j+1)$th species concentration, $[C]_{1, j+1}$, will be $N_{\text{state}}$ memory locations away; this increases the likelihood of cache misses on the CPU~\cite{gray2000rules}, but conversely well matches the pattern of coalesced memory access on the GPU~\cite{Nvidia:2018}.
+
+In a \textit{deep} SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works~\cite{Stone:2013aa,CurtisGPU:2017}), a compute unit uses its SIMD lanes\slash SIMT threads cooperatively to evaluate the thermochemical source terms for a single thermochemical state; thus SIMD lanes loading $[C]_{1, j} \ldots [C]_{1, j + N_v}$ will access sequential memory locations if the data is stored in ``C''-order.
+Further, in ``C'' ordering \revise{any two species concentrations within the same thermochemical state lie at most $N_{\text{sp}}$ locations away}, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data locality increases the chances of a cache hit on the CPU, but may lead to uncoalesced memory accesses on the GPU.
+In addition, deep vectorization requires both synchronization between SIMD lanes\slash SIMT threads via memory fences\slash barriers.
+\revise{This is an expensive operation, may result} in SIMD waste or SIMT thread divergence caused by different lanes\slash threads executing different instructions (e.g., resulting from different if\slash then branches).
+Shallow vectorization may also experience SIMD waste or SIMT thread divergence, e.g., in chemical kinetic integration due to varying internal solver time-step sizes~\cite{CurtisGPU:2017}.
+\revise{However, here this largely unaffects shallow vectorization} \add{as the only major code paths that differ between vector lanes are high\slash low-temperature polynomial evaluations and differing pressures for P-Log reactions, which cause far fewer issues compared with differing internal ODE integration time-steps~\cite{CurtisGPU:2017}}.
 
@@
   \end{minipage}
-  \caption{Vectorized data ordering patterns}
+  \caption{Vectorized data-ordering patterns}
   \label{F:vector_data}
@@
 Finally,~\cref{F:vector_data} shows a vectorized data-ordering that improves the caching patterns of a shallow, ``C''-ordered SIMD vectorization on the CPU (\cref{F:row_major_split}) and a deep, ``F''-ordered SIMT vectorization on the GPU (\cref{F:column_major_split}).
-This is accomplished by splitting the slower-varying axis of the data array---columns for ``C''-ordering, and rows for ``F''-ordering---into chunks of size $N_v$ (the SIMD vector-width, or SIMT warp-size) and laying these data out contiguously in memory.
-For example, using the shallow-vectorized ``C''-ordering pictured in~\cref{F:row_major_split}, the concentration of species $j$ for states $i\ldots i+N_v$, $[C]_{i, j} \ldots [C]_{i + N_v, j}$, are contiguous in memory and are followed by the concentration of species $j + 1$ for the same states, $[C]_{i, j + 1} \ldots [C]_{i + N_v, j + 1}$.
-This pattern, similar to OpenCL's native vector data-types---e.g., \texttt{double\num{8}} which treats \num{8} contiguous double precision floating point numbers as a single vector datum---ensures that any SIMD operation occurs on data that is contiguous in memory, greatly improving caching and SIMD-throughput.
-Conversely, the data-ordering in~\cref{F:column_major_split} is designed to enable coalesced memory accesses for ``F''-ordered, deep-SIMT vectorization on the GPU.
-The effects of these various data-ordering and vectorization patterns on performance will be studied in~\cref{S:results}.
-
-\subsection{thermochemical Source-Terms and Jacobian}
+\revise{We accomplish this} by splitting the slower-varying axis of the data array---columns for ``C''-ordering, and rows for ``F''-ordering---into chunks of size $N_v$ (the SIMD vector width or SIMT warp size) and laying these data out contiguously in memory.
+For example, using the shallow-vectorized ``C''-ordering pictured in~\cref{F:row_major_split}, the concentrations of species $j$ for states $i$ to $i+N_v$ ($[C]_{i, j}, \ldots, [C]_{i + N_v, j}$) lie contiguously in memory and are followed by the concentrations of species $j + 1$ for the same states ($[C]_{i, j + 1}, \ldots, [C]_{i + N_v, j + 1}$).
+This pattern ensures that any SIMD operation occurs on data contiguous in memory, which greatly improves caching and SIMD throughput;
+\revise{it is also similar to OpenCL's native vector data-types, e.g., \texttt{double\num{8}} treats eight contiguous double-precision floating-point numbers as a single vector datum}.
+Conversely, the data-ordering in~\cref{F:column_major_split} \revise{enables} coalesced memory accesses for ``F''-ordered, deep SIMT vectorization on the GPU.
+\revise{We will discuss the} effects of these various data-ordering and vectorization patterns on performance in~\cref{S:results}.
+
+\subsection{Thermochemical Source Terms and Jacobian}
 This new version of \texttt{pyJac} is capable of evaluating the thermochemical source-terms for using both the constant-pressure (\conp/) or constant-volume (\conv/) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}

### 2018-06-05 derivations.tex
fixes to SM

--- 2018-03-21 derivations.tex
@@
 
-\title{A System of Equations and Derivation of Sparse Constant-Pressure\slash Constant-Volume Chemical Kinetic Jacobians for \texttt{pyJac}}
+\title{Supplemental: A System of Equations and Derivation of Sparse Constant-Pressure\slash Constant-Volume Chemical Kinetic Jacobians for \texttt{pyJac}}
 
 \author[1]{Nicholas Curtis\corref{corr}}
+\author[2]{Kyle E.~Niemeyer}
 \ead{nicholas.curtis@uconn.edu}
@@
 \begin{abstract} % not to exceed 200 words
-This document presents the system of chemical kinetic equations evaluated by the new version of code-generation platform, \texttt{pyJac-v2}.
+This document supplements the work, \textit{SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source-Term Evaluation} and presents the system of chemical kinetic equations evaluated by the new version of code-generation platform, \texttt{pyJac-v2}.
 First, the thermo-chemical source terms for fixed-mass, adiabatic constant-pressure\slash constant-volume (\conp/\slash\conv/) reactors are given, followed by the derivation of the sparse analytical chemical kinetic Jacobian for both reactors.
@@
 % sessioning/indexing purposes.)
-\begin{keyword}
-    Chemical Kinetics\sep SIMD\sep SIMT\sep Sparse\sep Jacobian
-\end{keyword}
+% \begin{keyword}
+%     Chemical Kinetics\sep SIMD\sep SIMT\sep Sparse\sep Jacobian
+% \end{keyword}
 
@@
 \section{Introduction}
-This document contains the derivations for \texttt{pyJac} v2.0, which generates code to evaluate the chemical source terms and analytical chemical kinetic Jacobians of constant-pressure\slash constant-volume, fixed-mass, adiabatic reactors.
+This document is a supplement to the paper---\textit{SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source-Term Evaluation}---and is provided as a reference guide to the equations evaluated by \texttt{pyJac} v2.0, a code-generation platform for the evaluatation of chemical source terms and analytical chemical kinetic Jacobians of constant-pressure\slash constant-volume, fixed-mass, adiabatic reactors.
 Note that equations specific to the constant-pressure or constant-volume derivation will be marked with \conp/ or \conv/ respectively.

### 2018-06-05 paper.tex
edits to beginning of section 2.3

@@
 \subsection{Thermochemical Source Terms and Jacobian}
-This new version of \texttt{pyJac} is capable of evaluating the thermochemical source-terms for using both the constant-pressure (\conp/) or constant-volume (\conv/) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
-In this section, we will only outline a brief summary of the system evaluated by \texttt{pyJac}; the interested reader is referred to the supplemental material for the complete derivations.
-
-The thermochemical state vector derivation consists of the temperature, a non-constant thermodynamic state parameter (volume or pressure for \conp/ and \conv/, respectively), and the number of moles of all species except the last species in the chemical model which is typically taken to be the bath-gas (e.g., \ce{N2}):
+This new version of \texttt{pyJac} is capable of evaluating the thermochemical source-terms for using the constant-pressure (\conp/) or constant-volume (\conv/) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
+In this section, we will outline a brief summary of the system evaluated by \texttt{pyJac}; \revise{the supplemental material contains the complete---and lengthy---derivations}.
+
+The thermochemical state vector consists of the temperature, a non-constant thermodynamic state parameter (volume or pressure for \conp/ and \conv/, respectively), and the number of moles of all species except the last species in the chemical model, typically taken to be the bath gas (e.g., \ce{N2}):
 \loadeq{state}
-where $T$ is the temperature, $V$ and $P$ the volume and pressure, respectively, and $n_j$ the number of moles of the $j$th species in the model (containing $\ns/$ total species).
-
-This state vector---inspired by~\cite{SCHWER2002270}---has a number of features that recommend its use.
-First, the state vector results in highly sparse chemical kinetic Jacobians as will be detailed in~\cref{S:sparsity}.
-Second, mass is explicitly conserved in this formulation owing to the fact that the moles of the final species is calculated from the ideal gas law, and the rate of change of the final species from conservation of mass (see \cref{deriv-s:state,deriv-s:source} of the supplemental material, respectively); additionally, system is not over-constrained~\cite{HANSEN2018257} and does not require use of a more complicated differential algebraic equation solver (as compared to an ODE integrator) for integration.
-Finally, the chemical kinetic Jacobian changes relatively little between the \conp/ and \conv/ forms, making maintaining the code-base much simpler.
-Although most current combustion codes do not use species moles as a state variable, conversion to\slash from more common mass\slash mole-fractions and moles is straightforward, and once inside integration of an chemical kinetic initial value problem (IVP), the choice of variables no longer matters.
+where $T$ is the temperature, $V$ and $P$ the volume and pressure respectively, and $n_j$ the number of moles of the $j$th species in the model (containing $\ns/$ total species).
+
+This state vector---inspired by Schwer et al.~\cite{SCHWER2002270}---has a number of \revise{beneficial features}.
+First, the state vector results in highly sparse chemical kinetic Jacobians, as will be detailed in~\cref{S:sparsity}.
+Second, \revise{this formulation explicitly conserves mass, because the number of moles and rate of change of the final species are calculated from the ideal gas law and conservation of mass, respectively}; \revise{see the supplemental material for the full details of the governing equations}.
+\revise{The system is} not over-constrained~\cite{HANSEN2018257} and does not require use of a more-complicated differential algebraic equation solver (as compared to an ODE integrator) for integration.
+Finally, the chemical kinetic Jacobian \revise{for this formulation} changes relatively little between the \conp/ and \conv/ forms, making maintaining the codebase much simpler.
+Although most current combustion codes do not use species moles as a state variable, conversion to\slash from the more-common mass\slash mole fractions and moles is straightforward, and \revise{the choice of variables no longer matters once inside the integration} of an chemical kinetic initial-value problem (IVP).
 
@@
 \loadeq{dstate}
-
-For both \conp/ and \conv/, the molar source-terms are~\cite{TurnsStephenR2012Aitc}:
+For both \conp/ and \conv/, the molar source terms are~\cite{TurnsStephenR2012Aitc}:
 \loadeq{dmolar}
@@
 \begin{equation}
- \dot{\omega}_{k} = \sum_{i=1}^{\nr/} \nu_{k, i} R_{i} c_{i}
+ \dot{\omega}_{k} = \sum_{i=1}^{\nr/} \nu_{k, i} R_{i} c_{i} \;,
 \end{equation}
-with $\nu_{k, i}$ the net stoichiometic coefficient of species $k$ in reaction $i$, $R_{i}$ the net rate of progress of reaction $i$ (described in~\cref{deriv-s:rop_net} of the supplemental material), and $c_{i}$ the pressure-dependent modification term, i.e., for third-body (see~\cref{deriv-s:thdbody} of the supplemental material), or falloff\slash chemically-activated reactions (\cref{deriv-s:falloff} of the same); \texttt{pyJac} is capable of evaluating all modern reaction types, e.g., P-Log and Chebyshev reactions.
-In addition, $N_{reac}$ is the total number of reactions in the chemical kinetic model.
+$\nu_{k, i}$ is the net stoichiometric coefficient of species $k$ in reaction $i$, $N_{reac}$ is the total number of reactions, $R_{i}$ is the net rate of progress of reaction $i$, and $c_{i}$ is the pressure-dependent modification term, i.e., for third-body or falloff\slash chemically-activated reactions.
+\texttt{pyJac} is capable of evaluating all modern reaction types, e.g., P-Log and Chebyshev reactions.
 
@@
 \loadeq{dtemp}
-where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure specific heat and constant-volume specific heat of species $k$ in molar units respectively, while $[C]_{k}$ is the concentration, given by:
+where $H_k$, $U_k$, $C_{p,k}$, and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure specific heat, and constant-volume specific heat of species $k$ in molar units, respectively, while $[C]_{k}$ is the concentration, given by
 \begin{equation}
- [C]_{k} = \frac{n_{k}}{V}
+ [C]_{k} = \frac{n_{k}}{V} \;.
 \end{equation}
-By differentiating the ideal gas law:
+By differentiating the ideal gas law, given by
 \begin{equation}
- PV = n\mathcal{R}T \;,
+ PV = n\mathcal{R}T
 \end{equation}
-where $\mathcal{R}$ is the ideal gas-constant in molar units, the volume and pressure source-terms can be found (where $W_k$ and $W_{\ns/}$ are the molecular weights of species $k$ and $\ns/$, respectively):
+where $\mathcal{R}$ is the ideal-gas constant in molar units, \revise{we find} the volume and pressure source terms (where $W_k$ and $W_{\ns/}$ are the molecular weights of species $k$ and $\ns/$, respectively):
 \loadeq{dparam}
 
-The Jacobian entries computed by \texttt{pyJac} are arranged such that entry ($i$, $j$) corresponds to the partial derivative of the $i$th source-term in~\cref{e:source_terms} by the $j$th state variable in~\cref{e:state}:
+\revise{\texttt{pyJac} arranges the computed Jacobian entries} such that entry ($i$, $j$) corresponds to the partial derivative of the $i$th source-term in~\cref{e:source_terms} by the $j$th state variable in~\cref{e:state}:
 \loadeq{jacgeneral}
@@
 \delete{The derivative of the temperature source-term with respect to itself is:\ldots}
-\add{The complete derivation of the Jacobian for \texttt{pyJac} is attached as supplemental material for the interested reader.}
+\add{The supplemental material for this article contains the complete derivation of the Jacobian used by \texttt{pyJac}, for interested readers.}
 

### 2018-06-05 paper.tex
edits through end of section 2

@@
 \label{s:unittest}
-Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates pseudo-code and data to OpenCL\slash C-code.
-As the name implies, \texttt{loo.py} generates code based on for-loops; this differs from the previous version of \texttt{pyJac}~\cite{pyjac16} which generates static code (i.e., fully unrolled loops, with thermodynamic\slash reaction parameters written directly in code rather than stored in arrays).
-In our previous work~\cite{Niemeyer:2016aa}, this static code generation caused some issues with large file sizes, long compilation times and even occasionally breaking the \texttt{gcc} and \texttt{nvcc} compilers (the latter issue necessitated breaking the Jacobian\slash source-term evaluations into separate files).
-The implications of this change will be studied in~\cref{S:jacobian_results}, where the performance of the new version of~\texttt{pyJac} will be compared to the previous version.
-
-In addition, \texttt{loo.py} allows the user to more easily make changes to the structure of the generated program, e.g., the data ordering, vectorization, and threading patterns, as well as switching the target-language for code-generation (and simpler extension to additional languages, e.g., CUDA).
-Further, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source-terms.
-The source-terms and sub-components thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{hogan2014fast,adept-v11} is used to obtain reference answers for Jacobian sub-components.
-The Portable OpenCL (POCL) implementation~\cite{poclIJPP} and OpenMP~\cite{dagum1998openmp} are used to run OpenCL and C unit-testing, respectively, on the continuous integration framework Travis CI~\cite{travis:2018}.
-Validation of the complete (as opposed to the sub-component testing discussed here) generated source-terms and Jacobian codes will be discussed in detail in~\cref{s:validation}.
+
+\revise{The new version of \texttt{pyJac} uses the Python package \texttt{loo.py}~\cite{kloeckner_loopy_2014} for code generation}, which translates pseudo-code and data to OpenCL\slash C code.
+As the name implies, \texttt{loo.py} generates code using for loops; this differs from the previous version of \texttt{pyJac}~\cite{pyjac16} that generates static code---i.e., fully unrolled loops, with thermodynamic\slash reaction parameters written directly in code rather than stored in arrays.
+In our previous work~\cite{Niemeyer:2016aa}, this static code generation caused some issues with large file sizes, long compilation times, and even occasionally broke the \texttt{gcc} and \texttt{nvcc} compilers (the latter issue necessitated splitting the Jacobian\slash source-term evaluations into separate files).
+\revise{We will discuss the implications of this change} in~\cref{S:jacobian_results}, where the performance of the new version of~\texttt{pyJac} will be compared with the previous version.
+
+In addition, \texttt{loo.py} allows the user to more easily make changes to the structure of the generated program, e.g., the data ordering, vectorization, and threading patterns, as well as switch the target language for code generation (and more simply extend to additional languages, e.g., CUDA).
+Further, \texttt{loo.py} can execute developed subroutines from Python (natively for C code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL), enabling unit testing\slash validation for each component of the Jacobian or source terms.
+The source terms and sub-components thereof (e.g., rates of progress, pressure-modification terms) are directly compared with Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{hogan2014fast,adept-v11} provides reference values for Jacobian sub-components.
+\revise{We use the} Portable OpenCL (POCL) implementation~\cite{poclIJPP} and OpenMP~\cite{dagum1998openmp} to perform OpenCL and C unit testing, respectively, on the continuous-integration framework Travis CI~\cite{travis:2018}.
+\revise{We will discuss validation} of the complete (as opposed to the sub-component testing discussed here) generated source-terms and Jacobian codes in detail in~\cref{s:validation}.
 

### 2018-06-05 paper.tex
Revise sentence kyle left a note on

@@
 Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge--Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
-Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates for solving partial differential equations (PDEs) \delete{on an Intel Xeon E5-2680} \add{both the CPU} and \delete{a Tesla K20} GPU.
-In evaluating the thermochemical properties, speedups between \SIrange{8}{13}{$\times$} were found (over a serial version of the same code) while running on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; they showed speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} when simulating a partially premixed methanol flame PDE on 16 CPU cores and the GPU, respectively.\todo{revise this sentence, fairly awkward}{}
+Yonkee and Sutherland~\cite{Yonkee2016} implemented \delete{GPU-}accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates for solving partial differential equations (PDEs) \delete{on an Intel Xeon E5-2680} \add{both the CPU} and \delete{a Tesla K20} GPU, \revise{achieving speedups over serial evaluation of thermochemical properties between \SIrange{8}{13}{$\times$} on a 16 core CPU, and \SIrange{20}{40}{$\times$} on the GPU}.
+\revise{In addition, \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} speedups} were achieved for the simulation of a partially premixed methanol flame PDE on 16 CPU cores and the GPU, respectively.
 Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} \delete{on a Tesla C2075} \add{on the GPU and CPU}.

### 2018-06-05 paper.tex
add note about 1-indexed

@@
 
-The ``C'' (C row-major) format stores the concentrations of all species for a single thermochemical condition $i$  sequentially in memory, i.e., with  $[C]_{1, 1}$ in index 1, $[C]_{1, 2}$ in index 2, and so on, as shown in \cref{F:row_major}.\todo{wouldn't these start at index 0?}{}
+The ``C'' (C row-major) format stores the concentrations of all species for a single thermochemical condition $i$  sequentially in memory, i.e., with  $[C]_{1, 1}$ in index 1 (\revise{using one-based index notation}), $[C]_{1, 2}$ in index 2, and so on, as shown in \cref{F:row_major}.
 Conversely, in the ``F'' (Fortran column-major) format the concentrations of a single species $j$ over all thermochemical states lie adjacent in memory, corresponding to storing $[C]_{1, 1}$ in index 1, $[C]_{2, 1}$ in index 2, and so on, as shown in \cref{F:column_major}.

### 2018-06-05 paper.tex
revise moores law statement

@@
 SIMD and SIMT programming are two important vector-processing paradigms used increasingly in scientific computing.
-Traditional multi-core parallelism is often used to increase central processing unit (CPU) performance; however, as exponential growth in processing power---colloquially known as Moore's law---has slowed~\cite{khan2018science}, SIMD processors as well as SIMT processors, e.g., in the form of graphics processing units (GPUs), have gained recognition due to their increased floating operation throughput.\todo{this isn't quite right; multicore parallelism is how CPUs have bypassed Moore's}{}
+Traditional multi-core parallelism \revise{has} often \revise{been} used to increase central processing unit (CPU) performance, \revise{even as the exponential growth in processing power}---colloquially known as Moore's law---has slowed~\cite{khan2018science}.
+\revise{Recently, SIMD\slash SIMT processors, e.g., in the form of graphics processing units (GPUs), have gained recognition due to their increased floating operation throughput}.
 The parallel programming standard OpenCL~\cite{stone2010opencl} has further enabled adoption of vector processing in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems, e.g., CPU, GPU, or Intel's Many Integrated Core (MIC) architecture.

### 2018-06-06 paper.tex
address Kyle's comments in the abstract

@@
 \begin{frontmatter}
-
 %====================================================================
-\begin{abstract} % not to exceed 200 words
-A code generation platform, \texttt{pyJac}, for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized \revise{constant-pressure\slash volume} thermochemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation has been developed.% Can you introduce this a bit more?
-good to start with at least one sentence of introduction.
-also, this is passive and quite long\slash complex
+\begin{abstract}
+\revise{The accurate prediction of key combustion phenomena in reactive-flow simulations, e.g., lean blow-out, extinction\slash ignition limits and pollutant formation, necessitates the use of detailed chemical kinetics.}
+\revise{The large size and high levels of numerical stiffness typically present in chemical kinetic models relevant to transportation\slash power-generation applications make the efficient evaluation\slash factorization of the chemical kinetic Jacobian and thermochemical source-terms critical to the performance of reactive-flow codes.}
+\revise{Here we present} \texttt{pyJac}, \revise{a code generation platform} for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized \revise{constant-pressure\slash volume} thermochemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation.
 Selected chemical kinetic models covering the range of sizes typically used in reactive-flow simulations were used for demonstration.
@@
 Speedups of \SIrange{3.40}{4.08}{$\times$} were found for shallow-vectorized OpenCL source-rate evaluation \revise{compared with} a parallel OpenMP code on an \avx/ central processing unit (CPU), increasing to \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.03}{4.23}{$\times$} for sparse and dense chemical kinetic Jacobian evaluation, respectively.
-In addition, the performance of a Nvidia \gpunew/ graphics processing unit (GPU) was compared to a Nvidia \gpuold/ GPU finding speedups of \SIrange{1.40}{1.88}{$\times$}, \SIrange{1.10}{1.59}{$\times$}, and \SIrange{1.36}{3.0}{$\times$} for source-rate, sparse, and dense Jacobian evaluation, respectively.
-%is this sentence necessary in the abstract?
+%\delete{In addition, the performance of a Nvidia \gpunew/ graphics processing unit (GPU) was compared to a Nvidia \gpuold/ GPU finding speedups of \SIrange{1.40}{1.88}{$\times$}, \SIrange{1.10}{1.59}{$\times$}, and \SIrange{1.36}{3.0}{$\times$} for source-rate, sparse, and dense Jacobian evaluation, respectively.}
 Furthermore, the effect of data-ordering was investigated and \revise{a} storage pattern specifically formulated for vectorized evaluation was proposed; as well, the effect of the \revise{constant pressure\slash volume} assumptions and varying vector widths were studied on source-term evaluation performance.
-Speedups over a first-order finite-difference approach reached \SIrange{17.60}{45.13}{$\times$} and \SIrange{55.11}{245.63}{$\times$} for sparse\slash dense evaluation on the CPU \slash GPU respectively,%not sure what numbers refer to- sparse, dense, CPU, GPU?
-while dense Jacobian evaluation was up to \SI{19.56}{$\times$} and \SI{2.84}{$\times$} times faster than a previous version of \texttt{pyJac} on a CPU and GPU, respectively.
+Speedups over a first-order finite-difference \revise{Jacobian} approach \revise{reached up to} \SI{17.60}{$\times$} and \SI{45.13}{$\times$} for dense\slash sparse evaluation \revise{respectively on the GPU, and up to} \SI{55.11}{$\times$} and \SI{245.63}{$\times$} on the CPU.
+\revise{Further, }dense Jacobian evaluation was up to \SI{19.56}{$\times$} and \SI{2.84}{$\times$} times faster than a previous version of \texttt{pyJac} on a CPU and GPU, respectively.
 Finally, future directions for vectorized chemical kinetic evaluation and sparse linear-algebra techniques were discussed.

### 2018-06-06 paper.tex
s/reacting flows/reactive-flows/ for consistency

@@
 
-\revise{As the combustion and reacting-flows community has recognized the
-importance of detailed chemical kinetics for predictive reactive-flow simulations~\cite{LU2009192},
-chemical kinetic models have grown in size and complexity to describe current and next-generation fuels relevant to transportation and power generation.
-For example,} a recent biodiesel model~\cite{WESTBROOK2011742} consists of \textasciitilde\num{3500} chemical species and over \num{17000} reactions.
+\revise{As the combustion and reactive-flows community has recognized the importance of detailed chemical kinetics for predictive reactive-flow simulations~\cite{LU2009192}, chemical kinetic models have grown in size and complexity to describe current and next-generation fuels relevant to transportation and power generation.}
+\revise{For example,} a recent biodiesel model~\cite{WESTBROOK2011742} consists of \textasciitilde\num{3500} chemical species and over \num{17000} reactions.
 Moreover, the cost of evaluating the chemical source-terms scales linearly with the size of the model, while evaluating and factorizing the chemical kinetic Jacobian respectively scale quadratically and cubically with the number of species in the model, \revise{if naively implemented via a finite-difference method~\cite{LU2009192}}.
@@
 \Cref{F:cpu_source} explores the performance of the source-term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
-Source-term evaluations---critical in their own right for direct numerical simulations of reacting flows~\cite{Spafford:2010aa}, among other applications---also give a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation performance.
+Source-term evaluations---critical in their own right for direct numerical simulations of reactive-flows~\cite{Spafford:2010aa}, among other applications---also give a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation performance.
 

### 2018-06-07 paper.tex

@@
 Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge--Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
-Yonkee and Sutherland~\cite{Yonkee2016} implemented \delete{GPU-}accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates for solving partial differential equations (PDEs) \delete{on an Intel Xeon E5-2680} \add{both the CPU} and \delete{a Tesla K20} GPU, \revise{achieving speedups over serial evaluation of thermochemical properties between \SIrange{8}{13}{$\times$} on a 16 core CPU, and \SIrange{20}{40}{$\times$} on the GPU}.
-\revise{In addition, \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} speedups} were achieved for the simulation of a partially premixed methanol flame PDE on 16 CPU cores and the GPU, respectively.
+Yonkee and Sutherland~\cite{Yonkee2016} implemented \delete{GPU-}accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates \delete{on an Intel Xeon E5-2680} \add{on both the CPU} and \delete{a Tesla K20} GPU, \revise{achieving speedups over serial evaluation between \SIrange{8}{13}{$\times$} on a 16 core CPU, and \SIrange{20}{40}{$\times$} on the GPU}.
+\revise{In addition, \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} speedups} were achieved for the simulation of a partially premixed methanol flame for solving partial differential equations (PDE) on 16 CPU cores and the GPU, respectively.
 Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} \delete{on a Tesla C2075} \add{on the GPU and CPU}.

### 2018-06-07 paper.tex
revise

@@
 Furthermore, the effect of data-ordering was investigated and \revise{a} storage pattern specifically formulated for vectorized evaluation was proposed; as well, the effect of the \revise{constant pressure\slash volume} assumptions and varying vector widths were studied on source-term evaluation performance.
-Speedups over a first-order finite-difference \revise{Jacobian} approach \revise{reached up to} \SI{17.60}{$\times$} and \SI{45.13}{$\times$} for dense\slash sparse evaluation \revise{respectively on the GPU, and up to} \SI{55.11}{$\times$} and \SI{245.63}{$\times$} on the CPU.
+Speedups \revise{reached up to} \SI{17.60}{$\times$} and \SI{45.13}{$\times$} for dense and sparse evaluation \revise{on the GPU, and up to} \SI{55.11}{$\times$} and \SI{245.63}{$\times$} on the CPU \revise{over a first-order finite-difference  Jacobian approach}.
 \revise{Further, }dense Jacobian evaluation was up to \SI{19.56}{$\times$} and \SI{2.84}{$\times$} times faster than a previous version of \texttt{pyJac} on a CPU and GPU, respectively.
@@
 \end{frontmatter}
-\todo[inline]{abstract comments in source}
 

### 2018-06-07 paper.tex
cleanup

@@
 SIMD and SIMT programming are two important vector-processing paradigms used increasingly in scientific computing.
-Traditional multi-core parallelism \revise{has} often \revise{been} used to increase central processing unit (CPU) performance, \revise{even as the exponential growth in processing power}---colloquially known as Moore's law---has slowed~\cite{khan2018science}.
+\revise{Traditional multicore parallelism is now used to increase} central processing unit (CPU) performance, \revise{as the exponential growth in processing power}---colloquially known as Moore's law---has slowed~\cite{khan2018science}.
 \revise{Recently, SIMD\slash SIMT processors, e.g., in the form of graphics processing units (GPUs), have gained recognition due to their increased floating operation throughput}.

### 2018-06-08 paper.tex
make change colors more distinct
starting to revise in results

@@
 \usepackage[xcolor]{changebar}
-\newcommand{\add}[1]{{\sloppy\cbcolor{teal}\textcolor{teal}{\cbstart {#1}\cbend}}}  % add
-\newcommand{\delete}[1]{\sloppy\cbcolor{red}\textcolor{red}{\cbdelete\sout{#1}}}
-\newcommand{\revise}[1]{{\sloppy\textcolor{Fuchsia}{#1}}}  % add
-
-\newcommand{\addnl}[1]{{\sloppy\textcolor{teal}{#1}}}  % add
-\newcommand{\deletenl}[1]{\sloppy\textcolor{red}{\sout{#1}}} %delete with no strikeout
+\newcommand{\add}[1]{{\sloppy\cbcolor{OliveGreen}\textcolor{OliveGreen}{\cbstart {#1}\cbend}}}  % add
+\newcommand{\delete}[1]{\sloppy\cbcolor{RedOrange}\textcolor{RedOrange}{\cbdelete\sout{#1}}}
+\newcommand{\revise}[1]{{\sloppy\textcolor{RoyalPurple}{#1}}}  % add
+\newcommand{\revisedelete}[1]{\sloppy\textcolor{RoyalPurple}{\sout{#1}}} %delete
+
+\newcommand{\addnl}[1]{{\sloppy\textcolor{OliveGreen}{#1}}}  % add
+\newcommand{\deletenl}[1]{\sloppy\textcolor{RedOrange}{\sout{#1}}} %delete with no strikeout
 
@@
 Instruction Set  & SSE4.2 	   & AVX2 	    \\
-Vector Width     & \SI{2}{\doubles}&\SI{4}{\doubles}\\
+Vector Width     & two doubles & four doubles \\
 Cores            & $2 \times 6$    & $2 \times 12$  \\
@@
 \caption{The Intel CPU configurations used in this study.
-	 The vector-width reported in this table is reported in (ideal) number of double operations per SIMD instruction, as this will be used in measuring SIMD efficiency; for reference, the vector-width in bits of the \sse/ and \avx/ machines are \SI{128}{\bit} and \SI{256}{\bit}, respectively.
+	 The vector widths are reported in (ideal) number of double operations per SIMD instruction, as this will be used in measuring SIMD efficiency; for reference, the vector widths of the \sse/ and \avx/ machines are \SI{128}{\bit} and \SI{256}{\bit}, respectively.
 	 The identifier field will be used as a shorthand descriptor in the performance plots to quickly identify the CPU type.}

### 2018-06-11 paper.tex
changed deletes to revisedelete in related work section

@@
 A number of recent works have investigated \revise{using} high-performance SIMT devices \revise{like GPUs} to accelerate reactive-flow and chemical kinetics simulations.
-Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup \delete{on a Tesla C1060} compared to a \add{CPU-based} serial implementation \delete{on a AMD-Opteron processor}.
-Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order-of-magnitude speedup for homogeneous reactor simulations of large chemical models\delete{on a Tesla 2050 GPU} over a serial\delete{Intel i7 930} CPU implementation.
-Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential \add{CPU} code\delete{on an Intel Xeon X5650 CPU}.
-Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, where a traditional implicit integrator handled the stiffest computational cells on a CPU and a stabilized-explicit solver \delete{on a Tesla C2050 GPU} solved the less-stiff cells on a GPU; this hybrid solution technique \revise{performs} \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
-Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code\delete{ on a Tesla C2070} as compared with a sequential \add{CPU} version of the same\delete{on an Intel Xeon X5650 CPU}.
+Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup \revisedelete{on a Tesla C1060} compared to a \revise{CPU-based} serial implementation \revisedelete{on a AMD-Opteron processor}.
+Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order-of-magnitude speedup for homogeneous reactor simulations of large chemical models \revisedelete{on a Tesla 2050 GPU} over a serial\revisedelete{Intel i7 930} CPU implementation.
+Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential \revise{CPU} code\revisedelete{on an Intel Xeon X5650 CPU}.
+Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, where a traditional implicit integrator handled the stiffest computational cells on a CPU and a stabilized-explicit solver \revisedelete{on a Tesla C2050 GPU} solved the less-stiff cells on a GPU; this hybrid solution technique \revise{performs} \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
+Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code \revisedelete{on a Tesla C2070} as compared with a sequential \revise{CPU} version of the same \revisedelete{on an Intel Xeon X5650 CPU}.
 Stone and Davis~\cite{Stone:2013aa} investigated a GPU-based version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order-of-magnitude speedup over a serial CPU implementation.
-Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup\delete{on a Tesla C2075 GPU} over a multithreaded VODE solver on a six-core \delete{Intel X5650} CPU.
+Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup \revisedelete{on a Tesla C2075 GPU} over a multithreaded VODE solver on a six-core \revisedelete{Intel X5650} CPU.
 Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge--Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
-Yonkee and Sutherland~\cite{Yonkee2016} implemented \delete{GPU-}accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates \delete{on an Intel Xeon E5-2680} \add{on both the CPU} and \delete{a Tesla K20} GPU, \revise{achieving speedups over serial evaluation between \SIrange{8}{13}{$\times$} on a 16 core CPU, and \SIrange{20}{40}{$\times$} on the GPU}.
+Yonkee and Sutherland~\cite{Yonkee2016} implemented \revisedelete{GPU-}accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates \revisedelete{on an Intel Xeon E5-2680} \revise{on both the CPU} and \revisedelete{a Tesla K20} GPU, \revise{achieving speedups over serial evaluation between \SIrange{8}{13}{$\times$} on a 16 core CPU, and \SIrange{20}{40}{$\times$} on the GPU}.
 \revise{In addition, \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} speedups} were achieved for the simulation of a partially premixed methanol flame for solving partial differential equations (PDE) on 16 CPU cores and the GPU, respectively.
-Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} \delete{on a Tesla C2075} \add{on the GPU and CPU}.
-The \add{GPU-based} implicit Runge--Kutta method performs equivalently to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38} \delete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models \delete{with an integration time step of $10^{-6}$ s}.
+Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} \revisedelete{on a Tesla C2075} \revise{on the GPU and CPU}.
+The \revise{GPU-based} implicit Runge--Kutta method performs equivalently to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38} \revisedelete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models \revisedelete{with an integration time step of $10^{-6}$ s}.
 

### 2018-06-12 paper.tex
changes through table 6

@@
 \usepackage{textcomp}
-\usepackage{gensymb}
+\usepackage{gensymb,amssymb}
 \usepackage{ulem}
@@
 Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge--Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
-Yonkee and Sutherland~\cite{Yonkee2016} implemented \revisedelete{GPU-}accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates \revisedelete{on an Intel Xeon E5-2680} \revise{on both the CPU} and \revisedelete{a Tesla K20} GPU, \revise{achieving speedups over serial evaluation between \SIrange{8}{13}{$\times$} on a 16 core CPU, and \SIrange{20}{40}{$\times$} on the GPU}.
+Yonkee and Sutherland~\cite{Yonkee2016} implemented \revisedelete{GPU-}accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates \revisedelete{on an Intel Xeon E5-2680} \revise{on both the CPU} and \revisedelete{a Tesla K20} GPU, \revise{achieving speedups over serial evaluation between \SIrange{8}{13}{$\times$} on a 16-core CPU and \SIrange{20}{40}{$\times$} on the GPU}.
 \revise{In addition, \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} speedups} were achieved for the simulation of a partially premixed methanol flame for solving partial differential equations (PDE) on 16 CPU cores and the GPU, respectively.
@@
 
-The performance and validation studies for this work were run on a variety of CPU and GPU platforms.
-\Cref{t:cpus} shows the number of cores, vector instruction set and model of the CPUs used in this work; each CPU had both \texttt{v16.1.1} of the Intel OpenCL runtime~\cite{intelopencl:2018} and \texttt{v1.0} of the POCL~\cite{poclIJPP} runtime installed, each enabling OpenCL \texttt{v1.2} execution.
+\revise{We ran the} performance and validation studies for this work on a variety of CPU and GPU platforms.
+\Cref{t:cpus} shows the number of cores, vector instruction set, and model of the CPUs used in this work; each CPU had both \texttt{v16.1.1} of the Intel OpenCL runtime~\cite{intelopencl:2018} and \texttt{v1.0} of the POCL~\cite{poclIJPP} runtime installed, both enabling OpenCL \texttt{v1.2} execution.
 Additionally, \texttt{v5.0} of the LLVM\slash\texttt{clang}~\cite{Lattner:2004:LCF:977395.977673} compiler chain was installed on all machines to enable use of POCL.
-\Cref{t:gpus} lists the model, number of CUDA cores and Nvidia driver of each of the GPUs used in this work.
+\Cref{t:gpus} lists the model, number of CUDA cores, and Nvidia driver of each GPU \revise{we used}.
 Nvidia's OpenCL runtime is bundled with the Nvidia driver~\cite{Nvidia:2018}, hence the driver version is used to specify the OpenCL runtime version.
@@
 OpenCL Version & \num{1.1}	& \num{1.2}	\\
-\addtocounter{footnote}{1}
-Memory\footnotemark[\thefootnote] & \SI{6}{\giga\byte} & \SI{12}{\giga\byte} \\
+\addtocounter{footnote}{1}Memory\footnotemark[\thefootnote] & \SI{6}{\giga\byte} & \SI{12}{\giga\byte}  \\
 \bottomrule
@@
 \caption{The Nvidia GPU configurations used in this study.  Nvidia's OpenCL runtime is provided with the graphics driver, rather than any specific version of CUDA.
-Again, the identifier field will be used as a shorthand descriptor during analysis of the performance results.
+The identifier field will be used as a shorthand descriptor during analysis of the performance results.
 }
@@
 \end{table}
-\footnotetext{Due to a driver implementation issue~\cite{Nvidia_memory}, the total memory used on the \gpuold/ and \gpunew/ GPUs were to limited to \SI{4}{\giga\byte} and \SI{10}{\giga\byte}, respectively.\addtocounter{footnote}{1}}
+\footnotetext{A driver implementation issue limited total memory to \SI{4}{\giga\byte} and \SI{10}{\giga\byte} on the \gpuold/ and \gpunew/ GPUs, respectively~\cite{Nvidia_memory}.\addtocounter{footnote}{1}}
 
 \Cref{t:platforms} lists the platforms and vectorization\slash execution patterns that they are capable of running.
-The Intel and Nvidia OpenCL runtimes lack implementations of atomic operations on double precision variables; these are currently necessary for \texttt{pyJac} to run deep-vectorized code.
-On the other hand, POCL is an open source OpenCL runtime that works on all CPU types tested in this work, and does implement these atomic operations.
-However, POCL's implicit vectorization module---which utilizes the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to translate OpenCL code to vectorized machine code---typically fails to achieve much speedup, if any.
-Thus POCL is a particularly useful tool for validation, but not necessarily for performance studies.
-This discussion will be expanded upon in~\cref{S:future} to highlight future directions.
+The Intel and Nvidia OpenCL runtimes lack implementations of atomic operations on double-precision variables; \revise{\texttt{pyJac} currently needs these} to run deep-vectorized code.
+On the other hand, POCL is an open-source OpenCL runtime that works on all CPU types tested here, and does implement these atomic operations.
+However, POCL's implicit vectorization module---which uses the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to translate OpenCL code to vectorized machine code---typically fails to achieve much, if any, speedup.
+\revise{Thus POCL is useful for validation but not necessarily for performance studies.}\todo{should we emphasize here that it isn't needed for pyJac to function properly?}{}
+\revise{We will expand upon this discussion} in~\cref{S:future} to highlight future directions.
 
@@
 \midrule
-OpenMP & X & -- & -- \\
-POCL OpenCL & X & X & X \\
-Intel OpenCL & X & X & -- \\
-Nvidia OpenCL & -- & X & -- \\
+OpenMP & \checkmark & -- & -- \\
+POCL OpenCL & \checkmark & \checkmark & \checkmark \\
+Intel OpenCL & \checkmark & \checkmark & -- \\
+Nvidia OpenCL & -- & \checkmark & -- \\
 \bottomrule
 \end{tabular}
-\caption{The various platforms used in this study and the execution \slash vectorization patterns that they are capable of running}
+\caption{The platforms used in this study and the execution \slash vectorization patterns that they are capable of running.}
 \label{t:platforms}
@@
 Finally, \cref{t:models} displays the chemical kinetic models used in this work, as well as number of partially stirred reaction conditions (PaSR) used in the condition database for each.
-Creation of the PaSR databases is described in detail in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
+\revise{Our previous works describe the creation of the PaSR databases in detail works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.}
 
@@
 \ce{H2}\slash\ce{CO} & \num{900900} & \cite{Burke:2011fh} \\
-GRI-Mech.~3.0         & \num{450900} & \cite{smith_gri-mech_30} \\
+GRI-Mech 3.0         & \num{450900} & \cite{smith_gri-mech_30} \\
 USC-Mech II           & \num{91800}  & \cite{Wang:2007} \\
@@
 \label{s:validation}
-The reaction rates of progress (ROP), species and temperature rates in this study are validated by comparison with Cantera~\cite{Cantera}.
-However special care must be taken due floating point arithmetic issues.
-
-For a direct comparison, a relative error norm of a quantity $X_{ij}$ over all states $j$, and reactions (or species) $i$ was computed using the $L^{\infty}$ norm:
+\revise{We validated the reaction rates of progress (ROP), species production rates, and temperature rates in this study by comparing with values calculated using Cantera~\cite{Cantera}.}
+However, special care must be taken due to floating-point arithmetic issues.
+
+For a direct comparison, a relative error norm of a quantity $X_{ij}$ over all states $j$ and reactions (or species) $i$ was computed using the $L^{\infty}$ norm:
 \begin{equation}
-E_{X} = \left\lVert \frac{\left\lvert X_{ij,\text{CT}} - X_{ij}\right\rvert}{\num{e-10} + \num{e-6} * \left\lvert X_{ij,\text{CT}} \right\rvert} \right\rVert_{i,j,\infty}
+E_{X} = \left\lVert \frac{\left\lvert X_{ij,\text{CT}} - X_{ij}\right\rvert}{\num{e-10} + \num{e-6} \times \left\lvert X_{ij,\text{CT}} \right\rvert} \right\rVert_{i,j,\infty} \;,
 \label{e:rel_err}
 \end{equation}
-where the \text{CT} subscript indicates values from Cantera~\cite{Cantera}
-
-However, in computing the net ROP of reaction $i$ for state $j$ from the forward and reverse ROP: $R_{ij} = R_{ij}^{\prime} - R_{ij}^{\prime\prime}$, accuracy can easily be lost as the net ROP may be---particularly near chemical equilibrium---many orders of magnitude smaller than the forward and\slash or reverse rates.
-To quantify this phenomena, the error in forward ROP is first defined as:
+where the \text{CT} subscript indicates values from Cantera~\cite{Cantera}.
+\todo[inline]{the $i,j$ subscript for the overall norm looks a little strange to me, is that typical?}
+
+However, computing the net ROP of reaction $i$ for state $j$ from the forward and reverse ROP, i.e., $R_{ij} = R_{ij}^{\prime} - R_{ij}^{\prime\prime}$, can easily lose accuracy as the net ROP may be many orders of magnitude smaller than the forward and\slash or reverse rates---particularly near chemical equilibrium.
+To quantify this phenomena, we first define the error in forward ROP as
 \begin{equation}
-\varepsilon^{\prime}_{ij} = \left\lvert R_{ij}^{\prime} - R_{ij,\text{CT}}^{\prime} \right\rvert
+\varepsilon^{\prime}_{ij} = \left\lvert R_{ij}^{\prime} - R_{ij,\text{CT}}^{\prime} \right\rvert \;,
 \end{equation}
 while the error in reverse ROP, $\varepsilon^{\prime\prime}_{ij}$, can be defined analogously.
-Finally, for the reaction $i^{*}$ and the state $j^{*}$ that result in the largest error in net ROP, i.e.
-$E_{R}$, an estimate of the error attributable to floating point error accumulation from the forward and reverse ROPs can be obtained as:
+Finally, for the reaction $i^{*}$ and the state $j^{*}$ that result in the largest error in net ROP, i.e., $E_{R}$, an estimate of the error attributable to floating-point error accumulation from the forward and reverse ROPs can be obtained using
 \begin{equation}
-E_{\varepsilon} = \frac{\max(\varepsilon^{\prime}_{i^{*}j^{*}}\text{, }\varepsilon^{\prime\prime}_{i^{*}j^{*}})}{\num{e-10} + \num{e-6} * \left\lvert R_{i^{*}j^{*},\text{CT}} \right\rvert}
+E_{\varepsilon} = \frac{\max(\varepsilon^{\prime}_{i^{*}j^{*}}\text{, }\varepsilon^{\prime\prime}_{i^{*}j^{*}})}{\num{e-10} + \num{e-6} \times \left\lvert R_{i^{*}j^{*},\text{CT}} \right\rvert} \;.
 \end{equation}
-This estimate allows for direct comparison of the error in forward or reverse ROPs to the value of the net ROP itself, if they are of similar magnitude the error in net ROP will be large.
+This estimate allows for directly comparing the error in forward or reverse ROPs with the value of the net ROP itself; the error in net ROP will be large if these are similar in magnitude.
 
@@
 \centering
-\begin{tabular}{@{}l S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1]@{}}
+\begin{tabular}{@{}l S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] @{}}
 \toprule
-\multicolumn{1}{l}{Model} & \multicolumn{1}{l}{\ce{H2}\slash\ce{CO}} & \multicolumn{1}{l}{GRI-Mech.~3.0} & \multicolumn{1}{l}{USC-Mech II} & \multicolumn{1}{l}{\ce{iC5H11OH}} \\
+{Model} & {\ce{H2}\slash\ce{CO}} & {GRI-Mech.~3.0} & {USC-Mech II} & {\ce{iC5H11OH}} \\
 \midrule
-\multicolumn{1}{c}{$E_{R^{\prime}}$}                    & 1.56e-8 & 2.95e-8 & 9.42e-8 & 4.86e-4 \\
-\multicolumn{1}{c}{$E_{R^{\prime\prime}}$}              & 6.92e-8 & 6.53e-8 & 1.20e-7 & 5.07e-4 \\
-\multicolumn{1}{c}{$E_{R}$}                             & 1.49e1  & 1.11e0  & 2.80e0  & 4.82e-1 \\
-\multicolumn{1}{c}{$E_{\varepsilon}$}                   & 1.48e1  & 1.13e0  & 2.93e0  & 5.03e-1 \\
-\multicolumn{1}{c}{$E_{\frac{\text{d} n}{\text{d} t}}$} & 2.53e1  & 2.60e0  & 7.62e0  & 1.58e1 \\
-\multicolumn{1}{c}{$E_{\frac{\text{d} T}{\text{d} t}}$} & 3.94e5  & 3.35e8  & 3.95e6  & 7.11e7 \\
-\multicolumn{1}{c}{$E_{\frac{\text{d} S}{\text{d} t}}$} & 3.52e12 & 3.46e12 & 3.44e12 & 3.38e12 \\
+$E_{R^{\prime}}$                    & 1.56e-8 & 2.95e-8 & 9.42e-8 & 4.86e-4 \\
+$E_{R^{\prime\prime}}$              & 6.92e-8 & 6.53e-8 & 1.20e-7 & 5.07e-4 \\
+$E_{R}$                             & 1.49e1  & 1.11e0  & 2.80e0  & 4.82e-1 \\
+$E_{\varepsilon}$                   & 1.48e1  & 1.13e0  & 2.93e0  & 5.03e-1 \\
+$E_{\frac{\text{d} n}{\text{d} t}}$ & 2.53e1  & 2.60e0  & 7.62e0  & 1.58e1 \\
+$E_{\frac{\text{d} T}{\text{d} t}}$ & 3.94e5  & 3.35e8  & 3.95e6  & 7.11e7 \\
+$E_{\frac{\text{d} S}{\text{d} t}}$ & 3.52e12 & 3.46e12 & 3.44e12 & 3.38e12 \\
 \bottomrule
 \end{tabular}
-\caption{Summary of rate of progress, species, temperature and thermodynamic state-parameter rate correctness.
+\caption{Summary of errors in rates of progress, species, temperature, and thermodynamic state-parameter rate \revise{compared with Cantera}.
 Error statistics are based on the infinity-norm of the relative error detailed in~\cref{e:rel_err} for each quantity.
@@
 
-In Table~\ref{T:source_error}, the results of this code's source-term evaluations are compared to Cantera~\cite{Cantera} on the library of PaSR conditions (\cref{t:models}).
-Very close agreement is observed for the forward and reverse ROPs for all models, however the error norm is \textasciitilde\numrange{3}{4} orders of magnitude larger for the isopentanol model.
-This results from differences in evaluation of P-Log reactions between \texttt{pyJac} and Cantera; namely \texttt{pyJac} computes the logarithm of the upper and lower reaction Arrhenius rates (see supplemental~\cref{deriv-s:pdep}) analytically while Cantera evaluates this term numerically.
-If the error from P-Log reactions are not considered in~\cref{e:rel_err}, the errors for the forward and reverse ROPs fall to \num{5.44e-08} and \num{1.59e-07}, respectively.
-We note that this discrepancy does not imply any incorrectness in either \texttt{pyJac} or Cantera---in fact, the error is still well within the proscribed tolerances in~\cref{e:rel_err}---but merely emphasizes the effect small code-changes can have on the accumulation of floating point errors.
-
-This point is further underscored by the error in the net ROP, which is \textasciitilde\numrange{3}{9} orders of magnitude (or \numrange{7}{9} orders of magnitude for the P-Log excluded error) larger than the error in either the forward or reverse ROP.
-In~\cref{T:source_error}, it is seen that $E_{\varepsilon}$ is very similar in magnitude to $E_R$ in all cases, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
-The molar species production rate error norm is similar in magnitude to that of the net ROP, however the temperature and state-parameter rate errors again increase as the error in net species production rates is amplified by multiplication with thermodynamic properties.
-Again, we note that the above discussion does not imply that discrepancies in computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROPs are nearly equal, implying the reaction is in near-equilibrium.
-
-\subsection{Jacobian validation}
+Table~\ref{T:source_error} compares \revise{\texttt{pyJac} v2's} source-term evaluations with Cantera's~\cite{Cantera} using the data set of PaSR conditions (\cref{t:models}).
+\revise{The forward and reverse ROPs agree closely} for all models, though the error norm is \textasciitilde\numrange{3}{4} orders of magnitude larger for the isopentanol model.
+This \revise{discrepancy} results from differences in evaluation of P-Log reactions between \texttt{pyJac} and Cantera: \texttt{pyJac} computes the logarithm of the upper and lower reaction Arrhenius rates analytically (see supplemental material) while Cantera evaluates this term numerically.
+If we neglect the errors from P-Log reactions in~\cref{e:rel_err}, the errors for the forward and reverse ROPs fall to \num{5.44e-08} and \num{1.59e-07}, respectively.
+This discrepancy does not imply any actual error in either \texttt{pyJac} or Cantera---in fact, the error still lies well within the proscribed tolerances in~\cref{e:rel_err}---but merely emphasizes
+\revise{how even small code changes can affect the accumulation of floating-point errors.}
+
+\revise{The error in the net ROP further underscores this point: it is}
+\textasciitilde\numrange{3}{9} orders of magnitude (or \numrange{7}{9} orders of magnitude when including P-Log reaction contributions) larger than the error in forward or reverse ROP.
+\revise{\Cref{T:source_error} shows that the magnitudes of $E_{\varepsilon}$ and $E_R$ agree} in all cases, indicating that the accumulation of floating-point error from the forward and reverse ROPs causes this large increase in error as previously discussed.
+\revise{The magnitudes of the errors in molar species production rate and net ROP agree, but
+thermodynamic properties amplify the error in net species production rates and lead to high discrepancies in temperature and state-parameter rates.}
+Again, these discrepancies in net ROP will not necessarily cause errors when integrating the chemical kinetics---either in \texttt{pyJac} or Cantera---as this loss of accuracy only occurs when the forward and reverse ROPs are nearly equal \revise{(i.e., near equilibrium)}.
+
+\subsection{Jacobian verification}
 \label{S:jac_valid}
 
-As in our previous work~\cite{Niemeyer:2016aa}, the correctness of the Jacobian was determined by comparison to the Jacobian obtained by automatic differentiation of the source-terms kernel generated by \texttt{pyJac} via the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
-The reasons for this choice are explained fully in~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with minimal additional programming effort.
-The discrepancy between the analytical and automatic differentiated Jacobians for thermochemical state $k$, denoted by $\mathcal{J}_k$ and $\hat{\mathcal{J}}_k$ respectively, is determined by the relative error Frobenius norm over all Jacobian indices $i, j$:
+As in our previous work~\cite{Niemeyer:2016aa}, we determined Jacobian matrix correctness by comparing with that obtained by automatic differentiation of the \texttt{pyJac}-generated source term, using the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
+We previously explained this choice fully~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with minimal additional programming effort.
+\revise{(In contrast, it is challenging to obtain robust, accurate Jacobians using finite differences.)}
+The discrepancy between the analytical and automatic-differentiation Jacobians for thermochemical state $k$, denoted by $\mathcal{J}_k$ and $\hat{\mathcal{J}}_k$ respectively, is determined by the relative error Frobenius norm over all Jacobian indices $i, j$:
 \begin{equation}
  \label{e:jac_error_base}
- E_{\text{rel}, k} = \left\lVert \frac{\hat{\mathcal{J}}_{ijk} - \mathcal{J}_{ijk}}{\hat{\mathcal{J}}_{ijk}} \right\rVert_{i,j,F}
+ E_{\text{rel}, k} = \left\lVert \frac{\hat{\mathcal{J}}_{ij,k} - \mathcal{J}_{ij,k}}{\hat{\mathcal{J}}_{ij,k}} \right\rVert_{ij,F} \;.
 \end{equation}
 
-To avoid large relative errors of very small nonzero Jacobian elements due to accumulation of floating point error, the Frobenius norm of the automatically differentiated Jacobian is calculated over all thermochemical states $k$:
+To avoid large relative discrepancies in small nonzero Jacobian elements due to accumulation of floating-point error, the Frobenius norm of the automatically differentiated Jacobian is calculated over all thermochemical states $k$:
 \begin{equation}
  \label{e:thresh}
- \mathcal{T} = \left\lVert \mathcal{\hat{J}} \right\rVert_{k, F}
+ \mathcal{T} = \left\lVert \mathcal{\hat{J}} \right\rVert_{k, F} \;.
 \end{equation}
-The error statistics reported in this section are then based only on matrix elements where $\mathcal{J}_{ijk} \ge \frac{\mathcal{T}}{\mathcal{C}}$, where $\mathcal{C}$ is the threshold parameter; this filtered form of~\cref{e:jac_error_base} is denoted $E_{\mathcal{C},k}$.
+The error statistics reported in this section are then based only on matrix elements where $\mathcal{J}_{ijk} \ge \frac{\mathcal{T}}{\mathcal{C}}$, where $\mathcal{C}$ is \revise{a tunable} threshold parameter; this filtered form of~\cref{e:jac_error_base} is denoted $E_{\mathcal{C},k}$.
 Finally, the Frobenius norm is calculated over all the states in the PaSR thermochemical condition database:
@@
  \label{e:thresholded_error}
- E_{\mathcal{C}} = \left\lVert E_{\mathcal{C},k} \right\rVert_{k, F}
+ E_{\mathcal{C}} = \left\lVert E_{\mathcal{C},k} \right\rVert_{k, F} \;.
 \end{equation}
 
-It is noted that this error norm is quite different from the relative error Frobenius norm suggested by Anderson et al.~\cite{Anderson:1999aa} for quantifying the error of matrices in LAPACK:
+This error norm is quite different from the relative error Frobenius norm suggested by Anderson et al.~\cite{Anderson:1999aa} for quantifying the error of matrices in LAPACK:
 \begin{align}
  \label{e:error_lapack}
- \lhs{lapack}{E_{\mathcal{L}, k}} &=  \rhs{lapack}{\frac{\left\lVert \hat{\mathcal{J}}_{ijk} - \mathcal{J}_{ijk} \right\rVert_{i,j,F}}{\left\lVert \hat{\mathcal{J}}_{ijk} \right\rVert_{i,j,F}}} \nonumber \\
- \lhs{lapack}{E_{\mathcal{L}}} &= \rhs{lapack}{\left\lVert  E_{\mathcal{L}, k} \right\rVert_{k, F}}[c]
+ \lhs{lapack}{E_{\mathcal{L}, k}} &=  \rhs{lapack}{\frac{\left\lVert \hat{\mathcal{J}}_{ijk} - \mathcal{J}_{ijk} \right\rVert_{ij,F}}{\left\lVert \hat{\mathcal{J}}_{ijk} \right\rVert_{i,j,F}}} \nonumber \quad \text{and} \\
+ \lhs{lapack}{E_{\mathcal{L}}} &= \rhs{lapack}{\left\lVert  E_{\mathcal{L}, k} \right\rVert_{k, F}}[c] \;.
 \end{align}
-In our experience, the LAPACK error norm is often dominated by the accuracy of the larger elements in the Jacobian, while the filtered error norm is useful for identifying errors in both large and small Jacobian entries.
-Further, with a tunable threshold parameter, $\mathcal{C}$, the error of different ranges of of element sizes can be assessed, and the effects of floating point error can be isolated.
-For reference, both our own error norm and the LAPACK error norm will be reported.
+In our experience, the accuracy of larger elements in a Jacobian often dominates the LAPACK error norm, while the filtered error norm can identify errors in both large and small Jacobian entries.
+Further, with the tunable threshold parameter $\mathcal{C}$, we can assess the error of different ranges of element sizes and isolate the effects of floating-point error.
+For reference, both our error norm and the LAPACK error norm will be reported.
 
@@
 \end{tabular}
-\caption{Summary of Jacobian matrix validation results.
-The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all test platforms, vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/ and sparse\slash dense Jacobians.
-It is noted that the Frobenius norm described in~\cref{e:thresh} varies slightly between the \conp/ and \conv/ cases; the reported $\bar{\mathcal{T}}$ is the average of the two, however the appropriate value was used during calculations of the error statistics.
+\caption{Summary of Jacobian matrix verification results.
+The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all test platforms, vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/, and sparse\slash dense Jacobians.
+The Frobenius norm described in~\cref{e:thresh} varies slightly between the \conp/ and \conv/ cases; the reported $\bar{\mathcal{T}}$ is the average of the two, with the appropriate value used during calculations of the error statistics.
 }
@@
 
-\Cref{T:error} reports the maximum $E_{\mathcal{C}}$ and $E_{\mathcal{L}}$ values over all test platforms\slash vectorization patterns (see~\cref{t:platforms}), sparse\slash dense (see~\cref{S:sparsity}) and \conp/\slash\conv/ Jacobians.
+\Cref{T:error} reports the maximum $E_{\mathcal{C}}$ and $E_{\mathcal{L}}$ values over all test platforms and vectorization patterns (see~\cref{t:platforms}), sparse and dense (see~\cref{S:sparsity}), as well as \conp/ and \conv/ formulations.
 The most stringent filtered error norm ($\mathcal{C} = 10^{20}$) ranges from \numrange[retain-zero-exponent=true]{e-7}{e0}; the largest error is for the \ce{H2}\slash\ce{CO} model.

### 2018-06-13 paper.tex
edits through section 3.5.2

@@
 
-\usepackage{todonotes}
+\usepackage[textsize=footnotesize,textwidth=1.5\linewidth]{todonotes}
 
@@
 For this model, $\mathcal{T}$ is smaller than the tolerance of \num{e20}, and hence the error norm considers Jacobian entries smaller than $\mathcal{O}(1)$.
-It is noted that for GRI-Mech 3.0, which has a $\mathcal{T}$ roughly an order of magnitude larger, the stringent error norm is significantly smaller: $\mathcal{O}(\num{e-7})$.
-Given the intricacy of floating point error evaluation, the use of different languages and OpenCL platforms (the effect of these differences will be explored in \cref{A:per_platform}), and for the general complexity of \texttt{pyJac} it would be exceedingly difficult to pinpoint an exact cause for this phenomenon, as was done in~\cref{s:validation}.
-To ensure there is no bug or error present in the Jacobians generated by \texttt{pyJac}, a relaxed filtered error norm ($\mathcal{C} = 10^{15}$) is also presented for each model in~\cref{T:error}.
-This relaxed norm is smaller by \numrange{2}{5} orders of magnitude for all models---except GRI-Mech 3.0, where the stringent case has very small error as previously discussed---indicating that accuracy is higher when the smaller Jacobian entries are excluded; this suggests floating point error accumulation is the primary driver of the stringent filtered error norm.
+GRI-Mech 3.0 has a $\mathcal{T}$ roughly an order of magnitude larger and so the stringent error norm is significantly smaller: $\mathcal{O}(\num{e-7})$.
+Given the intricacy of floating-point error evaluation, the use of different languages and OpenCL platforms (the effect of these differences will be explored in \cref{A:per_platform}), and the general complexity of \texttt{pyJac} it would be exceedingly difficult to pinpoint an exact cause for this phenomenon, as was done in~\cref{s:validation}.
+To ensure no bugs or errors present in the Jacobians generated by \texttt{pyJac}, relaxed filtered error norms ($\mathcal{C} = 10^{15}$) are also presented for each model in~\cref{T:error}.
+This relaxed norm is smaller by \numrange{2}{5} orders of magnitude for all models---except GRI-Mech 3.0, where the stringent case already has small error as previously discussed---which indicates that accuracy is higher when the smaller Jacobian entries are excluded.
+\revise{This result suggests that floating-point error accumulation controls the stringent filtered error norm.}
 
 The relative LAPACK error norm---ranging from \textasciitilde\numrange{e-10}{e-14}---re-enforces this finding, as it indicates roughly \numrange{10}{14} digits of accuracy~\cite{Anderson:1999aa}.
-The \ce{iC5H11OH} model has the largest LAPACK error norm, likely due to the presence of P-Log\slash Chebyshev reactions and associated complicated derivatives with many logarithms, exponetiations and summations.
-Further, the LAPACK error norm does not correlate well with the stringent filtered error norm, e.g., USC-Mech II has the smallest LAPACK error norm (\num{1.137e-14}), but the second largest stringent filtered error norm (\num{1.119e-2}).
-Conversely, the model with the largest LAPACK error norm, \ce{iC5H11OH} (with $E_{\mathcal{L}} = \num{1.227e-10}$), has the second smallest stringent filtered error norm (\num{1.363e-3}) , again suggesting that the stringent error norm is influenced by floating point error accumulation.
-These findings, along with the individual unit-testing of all chemical source-terms and Jacobian sub-components described in~\cref{s:unittest}, gives high confidence in the correctness of this new version of \texttt{pyJac}.
+The \ce{iC5H11OH} model has the largest LAPACK error norm, likely due to the presence of P-Log\slash Chebyshev reactions and the resulting complicated derivatives with many logarithms, exponetiations, and summations.
+Further, the LAPACK error norm does not correlate well with the stringent filtered error norm, e.g., USC-Mech II has the smallest LAPACK error norm (\num{1.137e-14}) but the second-largest stringent filtered error norm (\num{1.119e-2}).
+\revise{Conversely, the model with the largest LAPACK error norm, \ce{iC5H11OH} has the second smallest stringent filtered error norm: $E_{\mathcal{L}} = \num{1.227e-10}$ and $E_{\mathcal{C} = 10^{20}} = \num{1.363e-3}$, again suggesting that floating-point error accumulation influences the stringent error norm.}
+These findings, along with the individual unit-testing of all chemical source-terms and Jacobian sub-components described in~\cref{s:unittest}, gives high confidence in the correctness of \revise{\texttt{pyJac} v2}.
 
@@
 
-In general, the Jacobian matrices generated by \texttt{pyJac} are largely sparse with non-zero entries corresponding mainly to species that participate in the same reaction (or non-default efficiency third-body species in a reaction), and mostly dense rows\slash columns corresponding to the temperature and thermodynamic state parameter.
+In general, the Jacobian matrices generated by \texttt{pyJac} are largely sparse with non-zero entries corresponding to species that participate in the same reaction or non-default efficiency third-body species in a reaction, with dense rows\slash columns corresponding to temperature and the thermodynamic state parameter.
 However, the explicit-mass conservation formulation of \texttt{pyJac} can introduce additional non-zero entries in two ways.
-First, if the last-species in the model (i.e., the bath-gas) participates directly in a reaction, the derivative of the forward or reverse rate of progress with respect to any other species is non-zero (see~\cref{deriv-s:dri_dnj} of the supplemental material).
-Similarly, if the last-species has a third-body efficiency not equal to the default (unity), this will again force all the derivative of the pressure-modification term with respect to other species to be non-zero (see~\cref{deriv-s:dci_thd} of the same).
-Either case will result in an undesirable fully dense Jacobian row for all species that have a non-zero net stoichiometic coefficient in such a reaction.
-
-However, the current \texttt{pyJac} allows the user to simply ignore these derivatives (via a command-line switch) and avoid any adverse effects on the Jacobian sparsity.
-The rationale behind this choice is that many common implicit integration techniques (e.g., CVODE~\cite{Hindmarsh:2005}) used in solution of chemical kinetic initial value problems are formulated around the assumption that the supplied Jacobian is approximate; this allows the Jacobian and its LU factorization to be reused for multiple internal integration time-steps, accelerating the solution process.
-For such solvers, there is no need for the exact form of the Jacobian and the so-called ``approximate'' form is preferable.
-\add{Though this might be used as a crude form of preconditioning for such solvers, the primary purpose is merely to increase the resulting sparsity of the Jacobian; the reader is directed to~\cite{MCNENLY2015581} for a more thorough discussion of preconditioners.}
+First, if the last species in the model (i.e., the bath gas) participates directly in any reaction, the derivative of its forward or reverse rate of progress \revise{is non-zero with respect to other species participating in that reaction.}\todo{Nick: check this}{}
+Similarly, if the last species has a third-body efficiency not equal to the default (one), \revise{this will again create nonzero derivatives for the pressure-modification term with respect to other species (see the supplemental material).}
+Either case will result in a fully dense Jacobian row for all species with a non-zero net stoichiometric coefficient in such a reaction.
+
+However, \texttt{pyJac} v2 allows the user to ignore these derivatives (via a command-line switch) and avoid the adverse effects on Jacobian sparsity.\footnote{Alternatively, one may choose the last species as one that does not participate in any reactions.}
+The rationale behind this choice is that many common implicit integration techniques (e.g., CVODE~\cite{Hindmarsh:2005}) used to solve chemical-kinetic initial-value problems are formulated around the assumption that the supplied Jacobian is approximate; this allows the Jacobian and its LU factorization to be reused for multiple internal integration time steps, accelerating the solution process.
+\revise{Such solvers do not need the} exact form of the Jacobian and thus the so-called ``approximate'' form is preferable.
+\add{Though this might be used as a crude form of preconditioning for such solvers, the primary purpose is merely to increase Jacobian sparsity; McNenly et al.~\cite{MCNENLY2015581} more thorough investigated preconditioners.}
 Hence, in this section we will detail the sparsity of both forms of the Jacobian for the chemical models tested.
 
-In~\cref{F:jac_sparsity}, a graphical representation of the sparsity pattern for GRI-Mech 3.0 is displayed.
-In particular we note that~\cref{F:jac_sparse_approx} is has several rows that are no longer fully-dense, as result of its approximate form; these rows correspond to species directly interacting with \ce{N2}, largely in GRI-Mech 3.0's nitrogen chemistry reactions.
+\Cref{F:jac_sparsity} graphically represents the Jacobian sparsity of GRI-Mech 3.0.
+In particular we note that~\cref{F:jac_sparse_approx} has several rows that are no longer fully dense, as result of its approximate form; these rows correspond to species directly interacting with \ce{N2}, largely in GRI-Mech 3.0's nitrogen chemistry reactions.
 \cref{T:jac_sparsity} shows the density of the exact and approximate Jacobians for all chemical kinetic models tested in this work.
-The smallest model, \ce{H2}\slash\ce{CO} is very dense with \SI{71.4}{\percent} of the exact Jacobian entries non-zero, however this drops to \SI{56.7}{\percent} for GRI-Mech 3.0, and continues to decrease to \SI{28.2}{\percent} for USC-Mech II and just \SI{11.5}{\percent} for the isopentanol model.
+The smallest model, \ce{H2}\slash\ce{CO}, is very dense with \SI{71.4}{\percent} of the exact Jacobian entries non-zero; this drops to \SI{56.7}{\percent} for GRI-Mech 3.0, continues to decrease to \SI{28.2}{\percent} for USC-Mech II, and is just \SI{11.5}{\percent} for the isopentanol model.
 The approximate Jacobian assumption drops the density of Jacobian by \textasciitilde\SIrange{3}{7}{\percent} for all models.
@@
   \label{F:jac_sparsity}
-\end{figure}
-
-Currently, \texttt{pyJac} can utilize two common sparse matrix storage formats~\cite{netlib_templates}, compressed row\slash column storage (CRS\slash CCS), used for ``C'' and ``F''-ordered data respectively.
-For convenience we will outline only the CRS format but the CCS format is very similar~\cite{netlib_templates}.
-An $N\times N$ CRS matrix is composed of three vectors, a value vector of length $NNZ$---the number of non-zero elements in the Jacobian---which stores the elements of the Jacobian, a row pointer vector---of length $N$---which stores the locations in the value vector that begin a row, and a column index vector---of length $NNZ$---which stores the column indices of the elements in the value vector.
+  \todo[inline]{the figure on the left is noticeably blurry compared to the other... can we fix that?}
+\end{figure}
+
+Currently, \texttt{pyJac} can use two common sparse-matrix storage formats~\cite{netlib_templates}: compressed row storage (CRS) and compressed column storage (CCS), used for ``C'' and ``F''-ordered data respectively.
+For brevity we will outline only the CRS format but the CCS format is similar~\cite{netlib_templates}.
+An $N\times N$ CRS matrix is stored using three vectors: a value vector of length $N_{NZ}$ (the number of non-zero elements in the Jacobian) that stores the elements of the Jacobian, a row pointer vector of length $N$ that stores the locations in the value vector that begin a row, and a column index vector length $N_{NZ}$ that stores the column indices of the elements in the value vector.
 
@@
 \toprule
-Model                 & \multicolumn{1}{c}{Exact Jacobian Density} & \multicolumn{1}{c}{Approximate Jacobian Density} \\
+Model                 & {Exact Jacobian Density} & {Approximate Jacobian Density} \\
 \midrule
@@
 
-The Jacobian access pattern used by \texttt{pyJac} is fairly irregular; for simplicity we will only discuss looping-structure of species derivatives calculations as these form the bulk of the computation, and have the most challenging Jacobian access patterns.
+The Jacobian access pattern used by \texttt{pyJac} is fairly irregular; for simplicity we will only discuss looping-structure of species derivatives calculations since these form the bulk of the computation and have the most challenging Jacobian access patterns.
 In general, an outer loop iterates over all reactions of a certain type (e.g., falloff reactions) and calculates the relevant Jacobian subproducts---independent of any particular species---for the reaction (e.g., the derivative of the falloff pressure modification term).
 Two inner loops then iterate over the species in a reaction, updating the Jacobian entries for these species as appropriate.
-This pattern leads to fairly easily vectorizable code, and efficient Jacobian evaluation as the bulk of the computation is dependent only on the reaction in question, as discussed in our previous work~\cite{Niemeyer:2016aa}.
+This pattern leads to fairly easily vectorizable code and efficient Jacobian evaluation, since the bulk of the computation depends only on the reaction in question, as discussed in our previous work~\cite{Niemeyer:2016aa}.
 Generally, this means that a lookup operation is required to find the sparse Jacobian index for any pair of state variables; in some cases this can be avoided, e.g., the rows corresponding to derivatives of the temperature and thermodynamic state parameter source-terms are fully dense in \texttt{pyJac}, and hence no lookup is necessary.
-This lookup operation is currently implemented as a simple for-loop, e.g., for a sparse lookup of a pair of indices $(i, j)$ in a CRS matrix, the lookup function searches the column index vector between the values $\texttt{row\_pointer}[i] \ldots \texttt{row\_pointer}[i + 1]$ for $j$, and returns the offset from $\texttt{row\_pointer}[i]$ (or $\num{-1}$ if not found).
-As will be explored in~\cref{S:jacobian_results}, this slows down sparse Jacobian evaluation, and might be improved---at the cost of increased constant data usage, a limitation for OpenCL---by a static mapping of the full Jacobian indices to the sparse index (or some null value if the entry is empty).
-Additionally, this might be an excellent usage of OpenCL's Image memory type (similar to texture memory in CUDA terminology) and both of these sparse indexing techniques merit future investigation.
+This lookup operation is currently implemented as a simple ``for'' loop, e.g., for a sparse lookup of a pair of indices $(i, j)$ in a CRS matrix, the lookup function searches the column index vector between the values $\texttt{row\_pointer}[i], \ldots, \texttt{row\_pointer}[i + 1]$ for $j$, and returns the offset from $\texttt{row\_pointer}[i]$ (or $\num{-1}$ if not found).
+As will be explored in~\cref{S:jacobian_results}, this slows down sparse Jacobian evaluation, and might be improved by a static mapping of the full Jacobian indices to the sparse index (or some null value if the entry is empty).
+\revise{However, this would require increased constant-data usage, a limitation for OpenCL.}
+Additionally, this might be an excellent usage of OpenCL's Image memory type (similar to texture memory in CUDA terminology).
+Both of these sparse indexing techniques merit future investigation.
 
@@
 The performance studies in this work were run on the platforms listed in~\cref{t:cpus,t:gpus}.
-Run times in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
-The OpenMP Jacobian\slash source-term kernels, as well as the OpenMP\slash OpenCL wrapping code---responsible for initializing\slash transferring memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms, and \texttt{gcc v4.8.5} on the \sse/\slash\gpuold/ machines.
-The optimization level ``\texttt{-O3 -mtune=native}'' was used, and no ``fast-math'' OpenCL optimizations were enabled.
+Run times in each case were averaged over ten runs, each using the same set of PaSR conditions used in validation.
+The OpenMP Jacobian\slash source-term kernels, as well as the OpenMP\slash OpenCL wrapping code (responsible for initializing\slash transferring memory, reading input, etc.) was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms and \texttt{gcc v4.8.5} on the \sse/\slash\gpuold/ machines.
+The optimization level ``\texttt{-O3 -mtune=native}'' was used and no ``fast math'' OpenCL optimizations were enabled.
 Additionally, the exact form of the Jacobian (as opposed the the ``approximate'' form discussed in~\cref{S:sparsity}) was used in all cases.
-Finally, unless stated otherwise: the performance results utilized a single CPU core, the \conp/ assumption, a vector-width of \num{8}\slash\num{128} and ``C''\slash ``F''-ordered data for the CPU\slash GPU cases, respectively; the run times reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
-The effects of choice of vector-width, data-ordering and differences between \conp/ and \conv/ evaluations on the CPU\slash GPU will be explored in~\cref{S:source_results}, while parallel scaling for multiple CPU cores will be exampled in~\cref{S:source_results,S:jacobian_results}.
+Finally, unless stated otherwise: the performance results used a single CPU core, the \conp/ assumption, a vector width of \num{8}\slash\num{128}, and ``C''\slash ``F''-ordered data for the CPU\slash GPU cases, respectively; the run times reported are for the number of conditions specified in~\cref{T:error} and include data-transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
+The effects of choice of vector width, data ordering, and differences between \conp/ and \conv/ evaluations on the CPU\slash GPU will be explored in~\cref{S:source_results}, while parallel scaling for multiple CPU cores will be exampled in~\cref{S:source_results,S:jacobian_results}.\todo{is ``exampled'' really the word you want here?}{}
 
@@
 \Cref{F:cpu_source} explores the performance of the source-term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
-Source-term evaluations---critical in their own right for direct numerical simulations of reactive-flows~\cite{Spafford:2010aa}, among other applications---also give a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation performance.
-
-In~\cref{F:intel_source_nonorm}, the mean run time per initial condition is shown for both the~\avx/ and~\sse/ CPUs, using both Intel OpenCL and OpenMP.
+Source-term evaluations---critical in their own right for direct numerical simulations of reactive-flows~\cite{Spafford:2010aa}, among other applications---also provide a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation performance.
+
+\Cref{F:intel_source_nonorm} shows the mean run time per initial condition for both the~\avx/ and~\sse/ CPUs, using Intel OpenCL and OpenMP.
 This normalization of the run time by the number of initial conditions tested is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{t:models}).
-For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the unvectorized (strictly parallel) Intel OpenCL code is slightly faster than OpenMP in all cases.
-As expected, the \avx/ machine is faster than the \sse/ CPU for the strictly parallel cases, \SIrange{1.82}{2.13}{$\times$} for the unvectorized OpenCL case, and \SIrange{1.72}{1.85}{$\times$} for OpenMP.
-Additionally, the shallow-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.
+For both CPUs, the OpenMP implementation is the slowest for all models; interestingly, the unvectorized (i.e., strictly parallel) Intel OpenCL code is slightly faster than OpenMP in all cases.
+As expected, the \avx/ machine is faster than the \sse/ CPU for the strictly parallel cases, performing \SIrange{1.82}{2.13}{$\times$} and \SIrange{1.72}{1.85}{$\times$} faster for the unvectorized OpenCL case and OpenMP, respectively.
+Additionally, the shallow-vectorized OpenCL code performs significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.
 
@@
 
-In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is calculated per-machine, e.g., the \avx/ shallow-vectorized code speedup is relative to OpenMP on the same CPU such that the speedup of the shallow-vectorized OpenCL code can be readily assessed.
-On both machines, the unvectorized OpenCL code is faster than the baseline parallel OpenMP code; by \SIrange{1.30}{1.51}{$\times$} on the \avx/ CPU, and \SIrange{1.25}{1.27}{$\times$} on the \sse/ machine.
+\Cref{F:intel_source} details the extent of this speedup; the speedup displayed is calculated per-machine, e.g., the \avx/ shallow-vectorized code speedup is relative to OpenMP on the same CPU\revisedelete{ such that the speedup of the shallow-vectorized OpenCL code can be readily assessed}.
+On both machines, the unvectorized OpenCL code is faster than the baseline parallel OpenMP code, by \SIrange{1.30}{1.51}{$\times$} on the \avx/ CPU and \SIrange{1.25}{1.27}{$\times$} on the \sse/ machine.
 Additionally, the shallow-vectorized OpenCL code is \SIrange{2.53}{2.92}{$\times$} and \SIrange{3.40}{4.08}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines, respectively.
 
-In contrast,~\cref{F:pocl_source} shows the mean run time per condition of deep, shallow and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
-No speedup is achieved for either vectorization type on either CPU---\add{though we stress that Intel OpenCL runtime achieved vectorization using the same code and hardware}---indeed, the OpenMP case is faster on both CPUs.
-The reasons for this lack of vectorization are quite technical,~\delete{and somewhat outside the scope of this work}however \add{personal communications~\cite{pocl_communication} with the developers of POCL revealed that in order to achieve vectorization, changes are required to both the way POCL prepares LLVM \cite{Lattner:2004:LCF:977395.977673} IR-code---specifically, to specialize boundary cases to ensure LLVM recognizes uniform vectorization loop-bounds, even if said bounds \textit{are} uniform in practice---as well as improvements to LLVM's loop-vectorizer: e.g., for proofs of a vector instruction's ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns.}
-As will be discussed in~\cref{S:future}, \add{it is hoped that use of explicit vector-types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a validation tool.
-
-\Cref{F:gpu_source} shows the performance of the source-term evaluation on the GPUs listed in~\cref{t:gpus}.
-In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source-terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean run time per condition; the run time continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
-\Cref{F:gpu_source_speedup} shows the speedup in source-term evaluations that the \gpunew/ GPU achieves over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} with two different vector-widths (a.k.a., the block size in Nvidia terminology), \num{64} and \num{128}; the fastest \gpunew/ case with a vector-width of 128 is \SIrange{1.40}{1.88}{$\times$} faster than the slowest case (\gpuold/ with a vector-width of \num{64}) depending on the chemical model in question.
-Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector-width has minimal effect on performance for most of the \gpunew/ and \gpuold/ cases; the GRI-Mech 3.0 and USC-Mech II models show the largest improvements with a vector-width of \num{128}, \textasciitilde\SIrange{10}{18}{\percent} for both GPUs.
-This is likely caused by higher occupancy on the GPUs, however it is unclear exactly how Nvidia's OpenCL runtime balances the registers\slash warps per streaming-multiprocessor, as controls occupancy in CUDA~\cite{occupancy}; use of even larger vector-widths (i.e., \num{256}\texttt{+}) on the GPU is a topic that merits future investigation.
+In contrast,~\cref{F:pocl_source} shows the mean run time per condition of deep, shallow, and unvectorized OpenCL codes using the POCL runtime, as compared with OpenMP parallelization.
+No speedup is achieved for either vectorization type on either CPU---indeed, the OpenMP case is faster on both CPUs\add{, though we stress that Intel OpenCL runtime achieves vectorization using the same code and hardware}.
+The reasons for this lack of vectorization with POCL are quite technical~\delete{ and somewhat outside the scope of this work}; however, \add{personal communication with the developers of POCL revealed that to achieve vectorization, changes are required to both the way POCL prepares LLVM \cite{Lattner:2004:LCF:977395.977673} IR-code---specifically, to specialize boundary cases to ensure LLVM recognizes uniform vectorization loop bounds, even if said bounds \textit{are} uniform in practice---as well as improvements to LLVM's loop-vectorizer: e.g., for proofs of a vector instruction's ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns~\cite{pocl_communication}.}\todo{this is one super long runon sentence---needs to be broken up and possibly simplified.}{}
+As will be discussed in~\cref{S:future}, \add{we hope that using explicit vector types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a validation tool.
+
+\Cref{F:gpu_source} shows the performance of evaluating source terms on the GPUs listed in~\cref{t:gpus}.
+\revise{\Cref{F:gpu_source_scaling} investigates how the number of initial conditions evaluated affects the mean run time per condition on the \gpunew/ GPU};
+the run time decreases until around \textasciitilde\num{e4} conditions for all chemical models, at which point the GPU becomes saturated and performance levels off.
+The performance plateaus slightly later for the \ce{H2}\slash\ce{CO} model compared with the others.
+\Cref{F:gpu_source_speedup} shows the speedup in source-term evaluation that the \gpunew/ GPU achieves over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling} with two vector widths (i.e., GPU block size), \num{64} and \num{128}.
+The best \gpunew/ case with a vector width of 128 is \SIrange{1.40}{1.88}{$\times$} faster than the slowest case (\gpuold/ with a vector-width of 64) depending on the chemical model in question.
+\Cref{F:gpu_source_speedup} also shows that varying the vector-width minimally affects performance for most of the \gpunew/ and \gpuold/ cases; the GRI-Mech 3.0 and USC-Mech II models show the largest improvements with a vector width of 128: \textasciitilde\SIrange{10}{18}{\percent} for both GPUs.
+This is likely caused by higher occupancy on the GPUs, but it is unclear exactly how Nvidia's OpenCL runtime balances the registers\slash warps per streaming-multiprocessor, as controls occupancy in CUDA~\cite{occupancy};\todo{this last fragment is missing something}{}
+use of even larger vector widths (i.e., $\geq$\num{256}) on the GPU is a topic that merits future investigation.\todo[inline]{I feel like we should pull this last bit, since someone could ask why you didn't just run with 256}
+
 \begin{figure}[htbp]
@@
 
-\Cref{f:source_permutate} shows the effect of changing data-ordering patterns, the \conp/ or \conv/ formulation, and the CPU vector-width size on the performance of source-term evaluation in \texttt{pyJac}.
-In~\cref{F:source_conpvsconv}, it is seen that choice of the \conp/ or \conv/ formulation has little to no effect on run time for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx/ machine.
-Generally speaking, the difference between the \conp/ and \conv/ formulations has only a marginal effect on performance regardless of CPU\slash GPU choice.
+\Cref{f:source_permutate} shows how changing data-ordering patterns, the \conp/ or \conv/ formulation, and the CPU vector width affect the performance of source-term evaluation in \texttt{pyJac}.
+Per~\cref{F:source_conpvsconv}, we see that the choice of \conp/ or \conv/ formulation has little to no effect on run time for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx/ machine.
+Generally speaking, the difference between the \conp/ and \conv/ formulations only marginally affects performance regardless of CPU\slash GPU choice.
 
@@
       \caption{The speedup achieved by ``C'' ordering over ``F'' ordering for Intel OpenCL and OpenMP on the \avx/ CPU.
-The speedup presented is calculated per-language (OpenMP and OpenCL) to better assess the effect of the data-ordering.}
+The speedup presented is calculated per-language (OpenMP and OpenCL) to better assess the effect of the data ordering.}
       \label{F:source_cvsf}
@@
   \caption{The effect of the \conp/ and \conv/ formulations, ``C'' and ``F'' data-ordering, and CPU vector-width on source-term evaluation performance in \texttt{pyJac}.
-	   It is noted the shallow-vectorized ``C'' ordered OpenCL cases correspond to the vectorized data ordering described in~\cref{S:data}}
+  The shallow-vectorized ``C''-ordered OpenCL cases correspond to the vectorized data ordering described in~\cref{S:data}.}
   \label{f:source_permutate}
@@
 
-In contrast,~\cref{F:source_cvsf} shows significant speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, e.g., the \SIrange{1.35}{2.09}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-ordered OpenMP baseline.
-Additionally, it is noted that the ``C''-ordered shallow-vectorization in~\cref{F:source_cvsf} (as well as the other shallow-vectorized CPU data shown in \cref{S:source_results,S:jacobian_results}) utilizes the vectorized-data ordering described in~\cref{S:data}; this case achieves speedups of \SIrange{2.14}{2.58}{$\times$} over the ``F''-ordered shallow-vectorization, demonstrating the value of the vectorized data ordering for CPU execution.
-
-\Cref{F:source_gpu_cvsf} shows the performance effect of the ``C'' versus ``F''-ordering on source-term evaluation performance on both GPUs, with the speedup presented per-GPU; the ``C'' and ``F''-ordered shallow-vectorizations are nearly equivalent on both GPUs, with less than a \SI{10}{\percent} difference in run time between data-orderings.
-It is noted that for the isopentanol model, ``C''-ordered data is \textasciitilde\SI{1.08}{$\times$} faster on both GPUs (while the trend is less clear for the other models).
-The rough equivalence between the ``C'' and ``F''-ordered codes on GPUs is counter to what one might expect, as typically speaking coalesced memory access in a shallow-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
-However, the vectorized-data ordering here ensures that memory accesses are aligned to the vector-width size and thus, likely well-coalesced.
-
-\Cref{F:source_vector_width} shows the effect of changing vector-width size on source-term performance the \avx/ CPU.
-The vector-width of size \num{8} was the fastest tested, while the larger vector-width of \num{16} is slightly slower due to increased register pressure~\cite{intel_vectypes}.
-It is unclear why the vector-width of \num{4} results in no speedup at all (it is in fact, the slowest case); Intel's vectorization guide~\cite{intel_vecknobs} mentions that a heuristic determines the optimal vector-width (in this case, it appears from compiler output to be \num{8}), so it is possible using vector-width smaller than the heuristic breaks the implicit vectorizer.
-It is noted that this issue does not occur for a vector-width of \num{4} on the \sse/ CPU.
+In contrast, \cref{F:source_cvsf} shows significant speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, e.g., the \SIrange{1.35}{2.09}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-ordered OpenMP baseline.
+Additionally, the ``C''-ordered shallow-vectorization in~\cref{F:source_cvsf} and the other shallow-vectorized CPU data shown in \cref{S:source_results,S:jacobian_results} use the vectorized-data ordering described in~\cref{S:data}; this case achieves speedups of \SIrange{2.14}{2.58}{$\times$} over the ``F''-ordered shallow-vectorization, demonstrating the value of the vectorized-data ordering for CPU execution.
+
+\Cref{F:source_gpu_cvsf} shows how the ``C''- and ``F''-ordering affect the performance of source-term evaluation on both GPUs, with the speedup presented per-GPU.
+The ``C''- and ``F''-ordered shallow-vectorizations perform almost equivalently on both GPUs, with less than a \SI{10}{\percent} difference in run time between data orderings.
+For the isopentanol model, ``C''-ordered data is \textasciitilde\SI{1.08}{$\times$} faster on both GPUs (while the trend is less clear for the other models).
+The roughly equivalent performance between the ``C'' and ``F''-ordered approaches on GPUs counters what one might expect: typically speaking, coalesced memory access in a shallow vectorization is easier to achieve with ``F''-ordering (see~\cref{S:data}).
+However, the vectorized-data ordering here ensures that memory storage is aligned to the vector width and, thus, encourages coalesced accesses.
+
+\Cref{F:source_vector_width} shows how changing vector width affects source-term evaluation performance on the \avx/ CPU.
+The vector width of 8 \revise{performs the fastest (out of those tested)}, while the larger vector width of 16 is slightly slower due to increased register pressure~\cite{intel_vectypes}.
+It is unclear why the vector width of 4 results in no speedup at all (in fact, it is the slowest case).
+Intel's vectorization guide~\cite{intel_vecknobs} mentions that a heuristic determines the optimal vector width (in this case, it appears from compiler output to be 8), so it is possible that using a vector width smaller than the heuristic breaks the implicit vectorizer.
+This issue does not occur for a vector width of 4 on the \sse/ CPU.
 
 Finally,~\cref{F:source_scaling} displays the (strong) parallel scaling efficiency and SIMD efficiency for the CPU platforms.
-The strong parallel scaling efficiency is defined as:
+The strong parallel scaling efficiency is defined as
 \begin{equation}
  \label{e:strong_scaling}
- \varepsilon = \frac{\bar{t}_{1}}{N \bar{t}_{N}}
+ \varepsilon = \frac{\bar{t}_{1}}{N \bar{t}_{N}} \;,
 \end{equation}
-where $\bar{t}_{N}$ is the mean run time per condition on $N$ CPU cores, and $\bar{t}_{1}$ the same on a single CPU core.
-The strong parallel scaling efficiency measures the speedup due to the use of additional CPU cores, as a fraction of linear speedup; strong scaling tends to decrease with the number of processors used due to memory bandwidth limitations and decreasing computation work allocated per CPU core~\cite{strong_scaling}.
+where $\bar{t}_{N}$ is the mean run time per condition on $N$ CPU cores and $\bar{t}_{1}$ the same on a single CPU core.
+The strong parallel scaling efficiency measures the speedup due to the use of additional CPU cores as a fraction of linear speedup; strong scaling tends to decrease with the number of processors used due to memory-bandwidth limitations and decreasing computation work allocated per CPU core~\cite{strong_scaling}.
 
@@
 \Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source-term evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
-In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for both Intel OpenCL and OpenMP, likely resulting from both its relatively small size, and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them); \add{as demonstrated in~\cref{S:SIMD_scaling} the amount of computational work required per thermochemical state plays a critical role in the full utilization of SIMD instructions\slash multiple threads}.
-Additionally, OpenMP tends to have higher scalings than the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for \num{4} and \num{8} CPU cores, respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorizated code.
+In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for both Intel OpenCL and OpenMP, likely resulting from both its relatively small size and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them).
+\add{As demonstrated in~\cref{S:SIMD_scaling}, the amount of computational work required per thermochemical state plays a critical role in fully utilizing SIMD instructions\slash multiple threads}.
+Additionally, OpenMP tends to scale better than the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for four and eight CPU cores, respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL.
+Though not pictured (to keep the figure readable), the unvectorized Intel OpenCL code scales only slightly worse than the OpenMP code, hence the poorer scaling is unique to the shallow-vectorized code.
 This is due in large part to the superior performance of the shallow-vectorized OpenCL code, coupled with the relatively small amounts of work associated with source-term evaluation.
-To illustrate this,~\cref{F:source_crossover} shows the mean run time per-condition of the OpenMP source-term evaluations for \numrange{2}{16} cores, compared to the shallow-vectorized OpenCL code on \num{1} (solid line) and \num{4} (dashed line) cores on the \avx/ machine.
-For all chemical models, the mean run time per-condition (and hence the computational work allocated per-core, one of the key-drivers of parallel scaling efficiency~\cite{strong_scaling}) of OpenMP running on \num{4} cores is roughly equal to that of the shallow-vectorized OpenCL code on a single core.
+To illustrate this,~\cref{F:source_crossover} shows the mean run time per-condition of the OpenMP source-term evaluations for \numrange{2}{16} cores, compared to the shallow-vectorized OpenCL code on one (solid line) and four (dashed line) cores on the \avx/ machine.
+For all chemical models, the mean run time per-condition (and hence the computational work allocated per-core, one of the key-drivers of parallel scaling efficiency~\cite{strong_scaling}) of OpenMP running on four cores is roughly equal to that of the shallow-vectorized OpenCL code on a single core.
 Similarly, OpenMP running on \num{16} cores is roughly equivalent to the OpenCL code on \num{4} cores.
-Therefore, a more fair comparison of parallel scaling efficiencies is to compare OpenCL running on \num{4} cores, to OpenMP on \num{16}; the OpenMP code's parallel efficiency drops to \textasciitilde\num{0.64} for \num{16} cores, similar to OpenCL's parallel scaling efficiency of \numrange{0.66}{0.72} at \num{4} cores.
-Indeed, as will be seen in~\cref{S:jacobian_results} the strong parallel scaling efficiency for sparse Jacobian evaluation---the most computationally intensive task in this work---are far more similar between Intel OpenCL and OpenMP.
-
-The SIMD efficiency is defined as:
+Therefore, a more fair comparison of parallel scaling efficiencies is to compare OpenCL running on \num{4} cores with OpenMP on \num{16}; the OpenMP code's parallel efficiency drops to \textasciitilde\num{0.64} for \num{16} cores, similar to OpenCL's parallel scaling efficiency of \numrange{0.66}{0.72} at \num{4} cores.
+Indeed, as will be seen in~\cref{S:jacobian_results}, \revise{sparse Jacobian evaluation---the most computationally intensive task in this work---exhibits similar strong-scaling efficiency on Intel OpenCL and OpenMP.}
+
+The SIMD efficiency is defined as
 \begin{equation}
  \label{e:simd_efficiency}
- \varepsilon_{\text{SIMD}} = \frac{\bar{t}_{\text{unvec}}}{W \bar{t}_{\text{shallow}}}
+ \varepsilon_{\text{SIMD}} = \frac{\bar{t}_{\text{unvec}}}{W \bar{t}_{\text{shallow}}} \;,
 \end{equation}
-where $\bar{t}_{\text{unvec}}$ is the mean run time per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$, the vector-width reported in number of double operations in~\cref{t:cpus}.
-This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.
-\Cref{F:source_simd_scaling} shows the SIMD efficiency of source-term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models (isopentanol and USC-Mech II) have higher SIMD efficiency of \numrange{0.76}{0.78}, and the smaller models (\ce{H2}\slash\ce{CO}, GRI-Mech \num{3.0}) have lower SIMD efficiency, \numrange{0.6}{0.66}.
-This again demonstrates that the SIMD-vectorization becomes efficient as the amount of work to perform increases (with increasing model size).
-Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one; this is likely caused by a combination of use of an OpenCL vector-width greater than the native CPU vector-width (i.e., \num{8} versus \num{2}), and the improved data-locality for the vectorized-data ordering, as discussed in~\cref{S:data}, and results in a modest \SIrange{7}{17}{\percent} improvement over the nominal vector-width.
+where $\bar{t}_{\text{unvec}}$ is the mean run time per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$ is the vector width reported in number of double operations (see~\cref{t:cpus}).
+This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.\todo{this sentence needs work, see alternates in src}{}
+%% you have both ``compares'' and ``compared'' in this sentence... would the following be correct?
+% This measure compares the actual speedup due to shallow vectorization with the ideal speedup based on the nominal vector width of the machine.
+\Cref{F:source_simd_scaling} shows the SIMD efficiency of source-term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models (isopentanol and USC-Mech II) have higher SIMD efficiencies of \numrange{0.76}{0.78}, and the smaller models (\ce{H2}\slash\ce{CO}, GRI-Mech \num{3.0}) have lower SIMD efficiencies of \numrange{0.6}{0.66}.
+This again demonstrates that the SIMD vectorization becomes more efficient with increasing amounts of work to perform (i.e., with increasing model size).
+Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one.
+This is likely caused by a combination of using an OpenCL vector width greater than the native CPU vector width (i.e., eight versus two) and improved data locality for the vectorized-data ordering as discussed in~\cref{S:data}, and results in a modest \SIrange{7}{17}{\percent} improvement over the nominal vector width.
 
@@
 \label{S:jacobian_results}
-
-\begin{figure}[htbp]
-   \centering
-  \begin{subfigure}[t]{0.48\linewidth}
-      \includegraphics[width=\textwidth]{sparse_vs_dense.pdf}
-      \caption{The mean run time per condition of sparse and dense Jacobian evaluations for Intel OpenCL\slash OpenMP on the \avx/ machine.}
-      \label{F:sparse_vs_dense}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.48\linewidth}
-      \includegraphics[width=\textwidth]{sparse_vs_dense_speedup.pdf}
-      \caption{The speedup of the sparse and dense Jacobian evaluations for Intel OpenCL over the sparse\slash dense OpenMP baseline on the \avx/ machine.}
-      \label{F:sparse_vs_dense_speedup}
-  \end{subfigure}
-  \\
-  \begin{subfigure}[t]{0.48\linewidth}
-      \includegraphics[width=\textwidth]{sparse_vs_dense_sse.pdf}
-      \caption{The mean run time per condition of sparse and dense Jacobian evaluations for Intel OpenCL\slash OpenMP on the \sse/ machine.}
-      \label{F:sparse_vs_dense_sse}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.48\linewidth}
-      \includegraphics[width=\textwidth]{sparse_vs_dense_see_speedup.pdf}
-      \caption{The speedup of the sparse and dense Jacobian evaluations for Intel OpenCL over the sparse\slash dense OpenMP baseline on the \sse/ machine.}
-      \label{F:sparse_vs_dense_sse_speedup}
-  \end{subfigure}
-  \caption{Performance of sparse and dense Jacobian evaluations on the CPU platforms in \texttt{pyJac}.}
-  \label{F:jacobian_perfomance}
-\end{figure}
 
@@
 This speedup increases to \SIrange{6.63}{9.44}{$\times$} for the sparse Jacobian.
+
+\begin{figure}[htbp]
+   \centering
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{sparse_vs_dense.pdf}
+      \caption{The mean run time per condition of sparse and dense Jacobian evaluations for Intel OpenCL\slash OpenMP on the \avx/ machine.}
+      \label{F:sparse_vs_dense}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{sparse_vs_dense_speedup.pdf}
+      \caption{The speedup of the sparse and dense Jacobian evaluations for Intel OpenCL over the sparse\slash dense OpenMP baseline on the \avx/ machine.}
+      \label{F:sparse_vs_dense_speedup}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{sparse_vs_dense_sse.pdf}
+      \caption{The mean run time per condition of sparse and dense Jacobian evaluations for Intel OpenCL\slash OpenMP on the \sse/ machine.}
+      \label{F:sparse_vs_dense_sse}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+      \includegraphics[width=\textwidth]{sparse_vs_dense_see_speedup.pdf}
+      \caption{The speedup of the sparse and dense Jacobian evaluations for Intel OpenCL over the sparse\slash dense OpenMP baseline on the \sse/ machine.}
+      \label{F:sparse_vs_dense_sse_speedup}
+  \end{subfigure}
+  \caption{Performance of sparse and dense Jacobian evaluations on the CPU platforms in \texttt{pyJac}.}
+  \label{F:jacobian_perfomance}
+\end{figure}
 

### 2018-06-13 paper.tex
edits through line 972

@@
 In~\cref{F:sparse_vs_dense}, the mean run time per condition is presented for the shallow-vectorized Intel OpenCL and OpenMP codes on the \avx/ CPU.
-The sparse Jacobian evaluations are slower on both Intel OpenCL and OpenMP due to indirect lookup indexing requirements, as discussed in~\cref{S:sparsity}.
-Interestingly the shallow-vectorized OpenCL code is less negatively impacted by the indirect lookup; the sparse OpenMP code is \SIrange{2.47}{10.42}{$\times$} slower than the dense OpenMP evaluation, while the sparse shallow-vectorized OpenCL code is just \SIrange{1.41}{3.34}{$\times$} slower than its dense counterpart.
-As a result, the shallow-vectorized sparse OpenCL code is as fast, or faster than the dense OpenMP code in all cases on the \avx/ machine (\cref{F:sparse_vs_dense}).
-\Cref{F:sparse_vs_dense_speedup} shows the speedup of the sparse\slash dense shallow-vectorized OpenCL over the same sparse\slash dense Jacobian format on OpenMP; the dense OpenCL code is \SIrange{3.03}{4.23}{$\times$} faster than the corresponding dense OpenMP code.
+The sparse Jacobian evaluates slower on both Intel OpenCL and OpenMP due to indirect lookup indexing requirements, as discussed in~\cref{S:sparsity}.
+Interestingly, indirect lookup less-negatively impacts the shallow-vectorized OpenCL code: the sparse OpenMP code is \SIrange{2.47}{10.42}{$\times$} slower than the dense OpenMP evaluation, while the sparse shallow-vectorized OpenCL code is just \SIrange{1.41}{3.34}{$\times$} slower than its dense counterpart.
+As a result, the shallow-vectorized sparse OpenCL code performs as fast or faster than the dense OpenMP code in all cases on the \avx/ machine (\cref{F:sparse_vs_dense}).
+\Cref{F:sparse_vs_dense_speedup} shows the speedup of the shallow-vectorized OpenCL, sparse and dense Jacobian evaluations over the same on OpenMP; the dense OpenCL code is \SIrange{3.03}{4.23}{$\times$} faster than the corresponding dense OpenMP code.
 This speedup increases to \SIrange{6.63}{9.44}{$\times$} for the sparse Jacobian.
@@
 
-On the \sse/ machine,~\cref{F:sparse_vs_dense_sse} shows similar results; the sparse OpenMP code is the slowest in all cases, and the shallow-vectorized OpenCL code is nearly as fast as the dense OpenMP code.
-Once again, the sparse OpenCL code is far less negatively impacted by the indirect lookup, and is only \SIrange{1.76}{3.33}{$\times$} slower than its dense counterpart, while the sparse OpenMP code is significantly (\SIrange{2.25}{8.72}{$\times$}) slower than the dense version.
-In~\cref{F:sparse_vs_dense_sse_speedup}, the speedup of the sparse and dense shallow-vectorized OpenCL codes are compared to their OpenMP version; the dense OpenCL code is \SIrange{1.92}{2.47}{$\times$} faster, while the sparse shallow-vectorization achieves a speedup of \SIrange{3.14}{5.03}{$\times$}.
+On the \sse/ machine,~\cref{F:sparse_vs_dense_sse} shows similar results: the sparse OpenMP code is the slowest in all cases, and the shallow-vectorized OpenCL code is nearly as fast as the dense OpenMP code.
+Once again, indirect lookup less-negatively impacts the sparse OpenCL code, which is only \SIrange{1.76}{3.33}{$\times$} slower than its dense counterpart, while the sparse OpenMP code is significantly (\SIrange{2.25}{8.72}{$\times$}) slower than the dense version.
+In~\cref{F:sparse_vs_dense_sse_speedup}, the speedup of the sparse and dense shallow-vectorized OpenCL codes are compared with their OpenMP versions; the dense OpenCL code is \SIrange{1.92}{2.47}{$\times$} faster while the sparse shallow-vectorization achieves a speedup of \SIrange{3.14}{5.03}{$\times$}.
 
@@
   \end{subfigure}
-  \caption{The parallel scaling efficiencies of sparse and dense Jacobian evaluations for the shallow-vectorized Intel OpenCL and OpenMP codes on the \avx/ CPU.}
-\end{figure}
-
-In~\cref{F:sparse_jac_scaling}, the strong parallel scaling efficiency of the sparse shallow-vectorized OpenCL is compared to the sparse OpenMP code.
-The plot is a bit difficult to read, as most of the data-points are clustered together, however the key takeaway is that the parallel scaling efficiency of the shallow-vectorized OpenCL code is very similar to that of the OpenMP code, in contrast to the parallel scaling efficiency of source-term evaluation (\cref{F:source_parallel_scaling}).
-The \ce{H2}\slash\ce{CO} model has the worst scaling for both codes, ranging from \numrange{0.94}{0.54} and \numrange{0.99}{0.82} for OpenCL and OpenMP, respectively on \numrange{2}{16} cores.
+  \caption{Strong parallel scaling efficiencies of sparse and dense Jacobian evaluations for the shallow-vectorized Intel OpenCL and OpenMP codes on the \avx/ CPU.}
+\end{figure}
+
+\Cref{F:sparse_jac_scaling} compares the strong parallel scaling efficiency of the sparse shallow-vectorized OpenCL with the sparse OpenMP code.
+Although the plot is challenging to read since most of the data are clustered together, it shows that the shallow-vectorized OpenCL code scales similarly to the OpenMP code, in contrast to the parallel scaling efficiency of source-term evaluation (\cref{F:source_parallel_scaling}).
+The \ce{H2}\slash\ce{CO} model scales the worst for both codes, ranging from \numrange{0.94}{0.54} and \numrange{0.99}{0.82} efficiency for OpenCL and OpenMP respectively on \numrange{2}{16} cores.
 As the model size increases, the efficiency of the OpenCL code improves dramatically, reaching \numrange{0.997}{0.84} for the isopentanol model.
-\Cref{F:dense_jac_scaling} shows scaling for the shallow-vectorized dense Jacobian OpenCL code.
-In this case, the isopentanol model has the worst scaling for both codes.
-Tt is noted that due to the sheer size of the dense isopentanol Jacobian---storing the dense matrix for a single thermochemical state takes over \SI{1}{\mega\byte} of data---the total number of states for the dense isopentanol Jacobian evaluation was limited to \num{50000} (which requires over \SI{50}{\giga\byte} of memory); this greatly drops the computation cost for this case, and adversely affects the scaling efficiency as was discussed in~\cref{S:source_results}.
-Excluding the isopentanol model, the dense shallow-vectorized OpenCL scalings are slightly higher than for source-term evaluation, e.g., \numrange{0.70}{0.78} and \numrange{0.52}{0.61} for \num{4} and \num{8} cores, respectively (compared to \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for shallow-vectorized OpenCL source-term on the same \num{4} and \num{8} cores); this is due to the higher computational cost of Jacobian evaluation, and hence more available work per CPU-core.
-As in~\cref{S:source_results}, we note that the shallow-vectorized dense Jacobian code running on \num{1} and \num{4} cores of the \avx/ machine is roughly as fast as the OpenMP code on \num{4} and \num{16} cores, respectively; the parallel scaling efficiency of OpenMP at \num{16} cores (excluding isopentanol as previously) is \numrange{0.63}{0.71}, similar to the shallow-vectorization's efficiency of \numrange{0.70}{0.78} for \num{4} cores.
+
+\Cref{F:dense_jac_scaling} shows scaling for the shallow-vectorized dense Jacobian OpenCL code.\todo{OpenCP and OpenMP?}{}
+In this case, the isopentanol model scales the worst for both cases.
+The sheer size of the dense isopentanol Jacobian limited the total number of states for the dense isopentanol Jacobian evaluation to \num{50000}---storing the dense matrix for a single thermochemical state takes over \SI{1}{\mega\byte} of data, so \num{50000} states requires over \SI{50}{\giga\byte} of memory); this greatly drops the computation cost for this case, and adversely affects the scaling efficiency as discussed in~\cref{S:source_results}.
+Excluding the isopentanol model, the dense shallow-vectorized OpenCL code scales slightly better than for source-term evaluation: \numrange{0.70}{0.78} and \numrange{0.52}{0.61} for four and eight cores, respectively (compared with \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for shallow-vectorized OpenCL source-term evaluations).
+This results from the higher computational cost of Jacobian evaluation, and hence more available work per CPU core.
+As in~\cref{S:source_results}, the shallow-vectorized dense Jacobian code running on \num{1} and \num{4} cores of the \avx/ machine performs roughly as fast as the OpenMP code on \num{4} and \num{16} cores, respectively.
+The parallel scaling efficiency of OpenMP on \num{16} cores (excluding isopentanol) is \numrange{0.63}{0.71}, similar to the shallow vectorization's efficiency of \numrange{0.70}{0.78} for \num{4} cores.
 
@@
 
-The mean run time per condition for the sparse and dense Jacobian evaluations on the \gpunew/ GPU is plotted in~\cref{F:gpu_sparse_vs_dense}.
+\Cref{F:gpu_sparse_vs_dense} plots the mean run time per condition for the sparse and dense Jacobian evaluations on the \gpunew/ GPU.
 As in the species evaluation case, the mean run time per condition drops steadily, for both cases becoming roughly constant after just \num{3e3} states (except for \ce{H2}\slash\ce{CO}, which levels off near \num{e4} conditions).
-Sparse Jacobian evaluation is significantly slower for all models before the GPU becomes saturated (due to the indirect indexing lookup), however past the saturation point the performance gap between the sparse and dense evaluations narrows.
-This is likely due to the ability to fit many more sparse Jacobian matrices in the \gpunew/'s memory, as well as improved data-locality\slash caching due to the smaller size of the sparse Jacobian.
+Sparse Jacobian evaluation is significantly slower for all models before the GPU becomes saturated (due to the indirect indexing lookup), but the performance gap between sparse and dense evaluations narrows past the saturation point.
+This is likely due to the ability to fit many more sparse Jacobian matrices in the \gpunew/'s memory, as well as improved data locality\slash caching due to the smaller size of the sparse Jacobian.
 \Cref{F:gpu_jacobian_speedup} presents the speedup of the \gpunew/ over the \gpuold/ GPU for sparse\slash dense Jacobian evaluation; sparse evaluation on the \gpunew/ is \SIrange{1.10}{1.59}{$\times$} faster than on the \gpuold/, while dense evaluation is \SIrange{1.36}{3.0}{$\times$} faster.
-The speedup on the \gpunew/ appears to decrease with increasing mechanism size for the sparse evaluation, but increases for the larger models in dense evaluation; this is likely due to the larger available memory of the \gpunew/, and hence fewer data-transfer operations to\slash from the GPU.
+The speedup on the \gpunew/ decreases with increasing model size for the sparse formulation, but increases for larger models when dense; this likely results from the larger available memory of the \gpunew/, and hence fewer data-transfer operations to\slash from the GPU.
 

### 2018-06-13 paper.tex
edits through end of paper

@@
 \crefname{appendix}{}{}
+
+\usepackage{placeins}
 
@@
 
-In~\cref{F:fd_vs_analytical}, the performance of the sparse analytical-Jacobian is compared to a sparse first-order finite-difference Jacobian on both the \avx/ CPU and the \gpuold/ \gpunew/ GPUs.
-\Cref{F:fd_vs_analytical_cpu} shows large speedups for both the OpenMP code and shallow-vectorized OpenCL code; the analytical OpenMP Jacobian is \SIrange{3.92}{8.67}{$\times$} faster, while the analytical OpenCL Jacobian achieves speedups of \SIrange{17.22}{55.11}{$\times$}.
-It is noted that the isopentanol case was excluded here, as a single run of the sparse finite-difference Jacobian using either OpenCL or OpenMP took over \num{12} hours of run time.
-In addition the current finite-difference formulation breaks Intel's auto-vectorizer, hence the OpenCL speedup comparison is made against the unvectorized OpenCL code; it is not a high priority to fix this issue, as the finite-difference Jacobian was implemented for comparison purposes only.
-Although we do not display the dense finite-difference Jacobian speedup in~\cref{F:fd_vs_analytical_cpu}, the dense OpenCL and OpenMP analytical Jacobian codes out-perform the finite-difference variants by even larger margins; \SIrange{24.44}{245.63}{$\times$} for OpenCL, and \SIrange{9.68}{112.73}{$\times$} for OpenMP (these data do include the isopentanol model, though limited to \num{50000} conditions as discussed previously).
-A comparison of the sparse analytical and finite-difference Jacobians on the GPUs is presented in~\cref{F:fd_vs_analytical_gpu}.
-The analytical Jacobian on the \gpunew/ and \gpuold/ shows large speedups, increasing with chemical model size, in the range of \SIrange{3.81}{17.60}{$\times$}; the \gpunew/ has a larger speedup than the \gpuold/ (\SI{17.60}{$\times$} v.s., \SI{14.75}{$\times$}) for the isopentanol model due to its larger available memory.
-Although not pictured, the dense analytical-Jacobian on the \gpunew/ GPU has larger speedups compared with the dense finite-difference Jacobian: \SIrange{4.04}{45.13}{$\times$}; the \gpunew/ GPU again has a significantly larger speedup than the \gpuold/ for the isopentanol model (\SI{45.13}{$\times$} versus \SI{23.85}{$\times$}), further underscoring the effect of the larger available memory on the \gpunew/.
+\Cref{F:fd_vs_analytical} compares the performance of the sparse analytical Jacobian with a sparse first-order finite-difference Jacobian on both the \avx/ CPU and \gpuold/\slash \gpunew/ GPUs.
+\Cref{F:fd_vs_analytical_cpu} shows large speedups for both OpenMP and shallow-vectorized OpenCL; the analytical OpenMP Jacobian is \SIrange{3.92}{8.67}{$\times$} faster, while the analytical OpenCL Jacobian achieves speedups of \SIrange{17.22}{55.11}{$\times$}.
+We excluded the isopentanol case here, since a single run of the sparse finite-difference Jacobian using either OpenCL or OpenMP took over 12 hours of run time.
+In addition the current finite-difference formulation breaks Intel's auto-vectorizer, hence we compared OpenCL against the unvectorized OpenCL code (we did not prioritize fixing this issue, since we implemented the finite-difference Jacobian for comparison purposes only).
+Although we do not display the dense finite-difference Jacobian speedup in~\cref{F:fd_vs_analytical_cpu}, the dense OpenCL and OpenMP analytical Jacobian codes outperform the finite-difference variants by even larger margins: \SIrange{24.44}{245.63}{$\times$} for OpenCL and \SIrange{9.68}{112.73}{$\times$} for OpenMP (these data do include the isopentanol model, though limited to \num{50000} conditions as discussed previously).
+\Cref{F:fd_vs_analytical_gpu} compares the sparse analytical and finite-difference Jacobians on the GPUs.
+The analytical Jacobian on the \gpunew/ and \gpuold/ shows speedups of \SIrange{3.81}{17.60}{$\times$} that increase with chemical model size; the \gpunew/ has a larger speedup than the \gpuold/ for the isopentanol model (\SI{17.60}{$\times$} vs.\ \SI{14.75}{$\times$}) due to its larger available memory.
+Although not pictured, the dense analytical Jacobian on the \gpunew/ GPU has larger speedups compared with the dense finite-difference Jacobian: \SIrange{4.04}{45.13}{$\times$}.
+The \gpunew/ GPU again shows significantly larger speedups over the \gpuold/ for the isopentanol model (\SI{45.13}{$\times$} vs.\ \SI{23.85}{$\times$}), further underscoring the effect of more available memory on the \gpunew/.
 
@@
       \includegraphics[width=\textwidth]{finite_difference_vs_analytical.pdf}
-      \caption{Speedup of the sparse analytical Jacobian versus finite-difference Jacobian evaluation on the \avx/ CPU, normalized per-language}
+      \caption{Speedup of the sparse analytical Jacobian versus finite-difference Jacobian evaluation on the \avx/ CPU, normalized per-language.}
       \label{F:fd_vs_analytical_cpu}
@@
       \includegraphics[width=\textwidth]{finite_difference_vs_analytical_gpu.pdf}
-      \caption{Speedup of the analytical versus finite-difference Jacobian evaluation on both the \gpunew/ and \gpuold/ GPUs, normalized per-GPU}
+      \caption{Speedup of the analytical versus finite-difference Jacobian evaluation on both the \gpunew/ and \gpuold/ GPUs, normalized per-GPU.}
       \label{F:fd_vs_analytical_gpu}
@@
 
-\Cref{F:v1_vs_v2} compares the performance for evaluation of the dense-analytical Jacobian of the new version of \texttt{pyJac}, \texttt{pyJac-v2}, to the previous version, \texttt{pyJac-v1} \cite{pyjac16}, on the \sse/ CPU and \gpuold/ GPU; dense Jacobian evaluation was selected as it was the only type implemented in the previous version of \texttt{pyJac}.
-On the \sse/ CPU, the \texttt{pyJac-v2} is faster than \texttt{pyJac-v1} for OpenMP evaluation for the larger chemical models; the static OpenMP code generated by \texttt{pyJac-v1} (see~\cref{s:unittest}) is \SI{1.79}{$\times$} faster for the \ce{H2}\slash\ce{CO} model, and only \SI{1.09}{$\times$} slower for the GRI-Mech 3.0 model.
+\Cref{F:v1_vs_v2} compares the performance of evaluating the dense analytical Jacobian of \texttt{pyJac-v2} with that of the previous version, \texttt{pyJac-v1}~\cite{pyjac16}, on the \sse/ CPU and \gpuold/ GPU.
+(We selected dense Jacobian evaluation for this comparison since it was the only type implemented in the previous version of \texttt{pyJac}.)
+On the \sse/ CPU, the \texttt{pyJac-v2} evaluates faster than \texttt{pyJac-v1} for OpenMP for the larger chemical models; the static OpenMP code generated by \texttt{pyJac-v1} (see~\cref{s:unittest}) is \SI{1.79}{$\times$} faster for the \ce{H2}\slash\ce{CO} model, and only \SI{1.09}{$\times$} slower for the GRI-Mech 3.0 model.
 In contrast, the loop-based OpenMP code of \texttt{pyJac-v2} is \SIrange{3.37}{10.19}{$\times$} faster than \texttt{pyJac-v1} for the USC-Mech II and isopentanol models.
-The shallow-vectorized OpenCL code is faster than \texttt{pyJac-v1} in all cases, achieving speedups of \SIrange{1.37}{19.56}{$\times$} increasing with model size.
-In~\cref{F:v1_vs_v2_gpu}, the performance of the \texttt{pyJac-v2} dense-analytical Jacobian on the \gpuold/ GPU is compared to \texttt{pyJac-v1} on the same.
-As with the CPU, the static code of \texttt{pyJac-v1} is slightly faster for the \ce{H2}\slash\ce{CO} model, however \texttt{pyJac-v2} outperforms \texttt{pyJac-v1} by \SIrange{1.25}{2.84}{$\times$} for the other models; it is noted that the decrease in speedup for the isopentanol model is likely due to the limited number of conditions in dense-evaluation for \texttt{pyJac-v2}, as noted earlier.
+The shallow-vectorized OpenCL \texttt{pyJac-v2} code is faster than \texttt{pyJac-v1} in all cases, achieving speedups of \SIrange{1.37}{19.56}{$\times$} that increase with model size.
+\Cref{F:v1_vs_v2_gpu} compares the performance of the \texttt{pyJac-v2} with \texttt{pyJac-v1} for evaluating dense analytical Jacobians on the \gpuold/ GPU.
+As with the CPU, the static code of \texttt{pyJac-v1} is slightly faster for the \ce{H2}\slash\ce{CO} model, but \texttt{pyJac-v2} outperforms \texttt{pyJac-v1} by \SIrange{1.25}{2.84}{$\times$} for the other models.
+\revise{Performance likely drops for the isopentanol model} due to the limited number of conditions in dense evaluation for \texttt{pyJac-v2}, as noted earlier.
 
@@
       \includegraphics[width=\textwidth]{v1_vs_v2.pdf}
-      \caption{Speedup of dense Jacobian evaluation in \texttt{pyJac-v2} over \texttt{pyJac-v1} on the \sse/ CPU}
+      \caption{\sse/ CPU}
       \label{F:v1_vs_v2_cpu}
@@
       \includegraphics[width=\textwidth]{v1_vs_v2_gpu.pdf}
-      \caption{Speedup of dense Jacobian evaluation in \texttt{pyJac-v2} over \texttt{pyJac-v1} on the \gpuold/ GPU}
+      \caption{\gpuold/ GPU}
       \label{F:v1_vs_v2_gpu}
   \end{subfigure}
-  \caption{Performance comparison of the new version of \texttt{pyJac} (\texttt{v2}) and the previous version, \texttt{v1}~\cite{pyjac16}.}
+  \caption{Performance comparison of dense Jacobian evaluation using \texttt{pyJac-v2} and \texttt{pyJac-v1}~\cite{pyjac16}.}
   \label{F:v1_vs_v2}
@@
 \section{Conclusions}
-In this work, automatically generated OpenCL codes for SIMD\slash SIMT-vectorized thermochemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation were developed and validated for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
-\add{Reviewing the goals of this work, as discussed in~\cref{S:Goals}}:
+
+In this work, we developed automatically generated OpenCL codes for SIMD and SIMT-vectorized evaluations of thermochemical source-term and sparse\slash dense chemical kinetic Jacobian matrices.
+We also verified and investigated the performance of source-term and Jacobian evaluation for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
+
+\todo[inline]{So I think some of the rest of this is really long and overly detailed---you don't need to rewrite all of the results, but really just the overall contributions and takeaways.
+I sorta think subsections are not appropriate (and help encourage overwriting...)}
+\add{The main contributions of this work are:}:
 
 \subsection{Derive and validate a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity}
-A new formulation resulted in highly sparse (\SIrange{28.6}{88.5}{\percent}) chemical kinetic Jacobians, with options to futher increase sparsity (to \SIrange{31.66}{92.0}{\percent}) at the expense of strict correctness.
+A new formulation resulted in highly sparse (\SIrange{28.6}{88.5}{\percent}) chemical kinetic Jacobians, with options to further increase sparsity (to \SIrange{31.66}{92.0}{\percent}) at the expense of strict correctness.
 
@@
 \add{Shallow-vectorized evaluation of \texttt{pyJac}'s chemical kinetic source-term and analytical Jacobian was demonstrated on both the CPU and GPU via OpenCL.}
-Deep-vectorization, though possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, did not yield any performance benefit as POCL did not achieve vectorizations for any execution pattern studied; deep-vectorization deserves further study with other platforms (e.g., CUDA).
+Deep vectorization, though possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, did not yield any performance benefit as POCL did not achieve vectorizations for any execution pattern studied.
+Deep vectorization deserves further study with other platforms (e.g., CUDA).
 \add{In addition, parallel OpenMP evaluation was demonstrated on the CPU.}
@@
 In addition, the choice of \conp/\slash\conv/ formulation was determined to be negligible on source-term evaluations, while varying the OpenCL vector-width had only modest (\textasciitilde\SI{10}{$\percent$}) performance effects on both the CPU and GPU.
-The strong parallel scaling efficiency was also examined for source-term evaluation; the parallel OpenMP code generally had better efficiencies, but when compared to a shallow-vectorized code with roughly equivalent performance---e.g., source-term shallow-vectorized evaluation on \num{4} cores to \num{16} OpenMP cores---the scalings were similar.
+The strong parallel scaling efficiency was also examined for source-term evaluation; the parallel OpenMP code generally had better efficiencies, but when compared to a shallow-vectorized code with roughly equivalent performance---e.g., source-term shallow-vectorized evaluation on \num{4} to \num{16} OpenMP cores---the scalings were similar.
 Examination of SIMD efficiency revealed increasing speedups with chemical model-size for source-term evaluation on the \avx/ CPU, while on the \sse/ CPU speedups larger than the nominal vector-width were achieved.
@@
 
-It is important to note that while OpenCL provides an simple interface to enable cross-platform execution, and significant speedups were achieved via shallow-vectorized OpenCL code in this work, there are some serious potential pit-falls in its use.
-The closed source OpenCL runtimes tested in this work (Intel and Nvidia) contain several bugs that result in compilation failures, simply incorrect vectorized machine code or even segmentation faults.
+\revisedelete{It is important to note that while }While OpenCL provides a simple interface to enable cross-platform execution, and significant speedups were achieved via shallow-vectorized OpenCL code in this work, there are some serious potential pitfalls in its use.
+The closed-source OpenCL runtimes tested in this work (Intel and Nvidia) contain several bugs that result in compilation failures, simply incorrect vectorized machine code, or even segmentation faults.
 Further, these runtimes (in our experience) tend to be less responsive to fixing said bugs, with relatively infrequent new releases (or in Nvidia's case, lack of changelogs\slash public records of bug-fixes).
-On the other hand, the open-source OpenCL runtime used in this work, POCL, has (in our experience) far fewer implementation bugs, and when issues arise the community is very responsive to bug reports and user outreach; however, POCL fails to achieve vectorization as noted in~\cref{S:source_results}.
-
+On the other hand, the open-source OpenCL runtime used in this work, POCL, has far fewer implementation bugs, and when issues arise the community is very responsive to bug reports and user outreach; however, POCL fails to achieve vectorization as noted in~\cref{S:source_results}.
+
+\todo[inline]{Does the following really belong in the conclusions?}
 As demonstration of the type of issue discussed, a minimum working example has been created~\cite{Nvidia_mwe} that demonstrates a failure of Nvidia's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla \gpunew/ GPU; by simply changing to another runtime (e.g., Intel)---with no code changes or recompilation---the correct result can be found.
@@
 Further, as noted throughout this work certain code-generation patterns can break OpenCL execution\slash vectorization, e.g., the failure of POCL to achieve vectorization or Intel OpenCL's failure to vectorize the finite-difference Jacobian.
-Indeed the vectorization process attempted by most OpenCL runtimes (and thus, reasons for incorrect\slash unvectorized code-output) is obscured from the user, making reasoning about errors or performance trends quite difficult.
+Indeed the vectorization process attempted by most OpenCL runtimes (and thus, reasons for incorrect\slash unvectorized code output) is obscured from the user, making reasoning about errors or performance trends quite difficult.
 
 Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change requires building of the wrapper that initializes\slash transfers memory and calls the source-rate\slash Jacobian kernel.
-Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler \texttt{\#pragma}s) is a key priority going forwards as OpenMP is a standard library on most machines.
+Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler \texttt{\#pragma}s) is a key priority going forwards as OpenMP is a standard library on most machines.
 In addition, CUDA~\cite{Nvidia:2018} has been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017}, while Intel's open-source OpenCL alternative, ISPC~\cite{pharr2012ispc}, has been relatively stable and easy to work with during preliminary efforts with the unit-testing discussed in~\cref{s:unittest}.
@@
 It is also possible, particularly for the Intel OpenCL\slash POCL runtimes, that better performance\slash stability might be achieved using so-called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}; \add{specifically this change is hoped to enable vectorization on the POCL runtime} and might also enable deep-vectorization on the Intel OpenCL runtime.
-\add{Finally, future extensions of this work will include additional target-languages (vectorized OpenMP, CUDA, ISPC, etc.) to improve ease of use and reliability, as well as improvements to the existing OpenCL targets, e.g., to enable meaningful deep-vectorized evaluation, as well the implementation of reaction sorting~\cite{Sewerin20151375} to improve SIMD-efficiency (\cref{S:SIMD_scaling}).}
-
-One key-component missing in this current version of \texttt{pyJac} is vectorized sparse\slash dense linear-algebra subroutines to maximize the performance of LU-factorization and matrix-vector multiplication (commonly used in implicit-integration techniques).
-Third-party\slash open-source options exist for some target-languages, e.g., cuBLAS~\cite{cublas}, clBLAS\slash clSPARSE~\cite{clmath} or SuperLU~\cite{superlu99}, but these do not necessarily cover all targets\slash required linear-algebra operations and, in the case of the CUDA\slash OpenCL libraries, are often optimized operation on one large matrix instead of many (relatively) smaller matrices.
-The extent to which these existing programs may be used needs to be assessed and missing operations should be implemented in \texttt{loo.py} to ensure easy switching between target-languages\slash vectorization types.
+\add{Finally, future extensions of this work will include additional target languages (e.g., vectorized OpenMP, CUDA, ISPC) to improve ease of use and reliability, as well as improvements to the existing OpenCL targets, e.g., to enable meaningful deep-vectorized evaluation, as well the implementation of reaction sorting~\cite{Sewerin20151375} to improve SIMD-efficiency (\cref{S:SIMD_scaling}).}
+
+One key component missing in the current version of \texttt{pyJac} is vectorized sparse\slash dense linear-algebra subroutines to maximize the performance of LU-factorization and matrix-vector multiplication (commonly used in implicit-integration techniques).
+Third-party\slash open-source options exist for some target languages, e.g., cuBLAS~\cite{cublas}, clBLAS\slash clSPARSE~\cite{clmath} or SuperLU~\cite{superlu99}, but these do not necessarily cover all targets\slash required linear-algebra operations and, in the case of the CUDA\slash OpenCL libraries, are often optimized operation on one large matrix instead of many (relatively) smaller matrices.
+The extent to which these existing programs may be used needs to be assessed and missing operations should be implemented in \texttt{loo.py} to ensure easy switching between target languages and vectorization types.
 
@@
 
-\section{Jacobian error statistics per test-platform}
+\section{Jacobian error statistics per test platform}
 \label{A:per_platform}
 
-This section gives more detail on the results presented in~\cref{S:jac_valid}, breaking down the reported error statistics per test-platform\slash language.
-The error of the Intel OpenCL runtime is presented in~\cref{T:intel_error}, the Portable OpenCL (POCL) runtime in~\cref{T:pocl_error}, OpenMP in~\cref{T:omp_error} and the Nvidia OpenCL runtime in~\cref{T:nv_error}.
+This section gives more detail on the results presented in~\cref{S:jac_valid}, breaking down the reported error statistics per test platform\slash language.
+The error of the Intel OpenCL runtime is presented in~\cref{T:intel_error}, the Portable OpenCL (POCL) runtime in~\cref{T:pocl_error}, OpenMP in~\cref{T:omp_error}, and the Nvidia OpenCL runtime in~\cref{T:nv_error}.
 POCL and OpenMP tend to have the smallest error norms, while Nvidia tends to have the largest; in particular the stringent filtered error norm $E_{\mathcal{C} = 10^{20}}$ is two orders of magnitude larger for the Nvidia runtime than the other test platforms with the \ce{H2}\slash\ce{CO} and USC-Mech II models.
@@
 \caption{Summary of Jacobian matrix validation results for the Intel OpenCL runtime.
-The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/ and sparse\slash dense Jacobians.
+The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}), \conp/\slash \conv/, and sparse\slash dense Jacobians.
 The threshold for the filtered relative error is the same as reported in~\cref{S:jac_valid}.
@@
 \caption{Summary of Jacobian matrix validation results for the Portable OpenCL (POCL) runtime.
-The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/ and sparse\slash dense Jacobians.
+The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}), \conp/\slash \conv/, and sparse\slash dense Jacobians.
 The threshold for the filtered relative error is the same as reported in~\cref{S:jac_valid}.
@@
 \caption{Summary of Jacobian matrix validation results for OpenMP execution.
-The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/ and sparse\slash dense Jacobians.
+The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/, and sparse\slash dense Jacobians.
 The threshold for the filtered relative error is the same as reported in~\cref{S:jac_valid}.
@@
 \caption{Summary of Jacobian matrix validation results for Nvidia OpenCL execution.
-The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/ and sparse\slash dense Jacobians.
+The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/, and sparse\slash dense Jacobians.
 The threshold for the filtered relative error is the same as reported in~\cref{S:jac_valid}.
@@
 
+\FloatBarrier
+
 \section{SIMD efficiency Scaling Example}
 \label{S:SIMD_scaling}
-\add{This simple example demonstrates the dependency of the SIMD efficiency of shallow-vectorized OpenCL source-term evaluation on the size of the chemical model in question, i.e., the amount of computational work per source-term evaluation.}
-\add{The base chemical model for this example was the isopentanol model~\cite{Sarathy:2013jr} used throughout this work, and the same thermochemical state database described in~\cref{s:test_platforms} was used for source-term evaluation.}
-\add{As in~\cref{S:results} all reported results are based on \num{10} individual runs and in this example all cases were run on the \avx/ machine using the Intel OpenCL runtime.}
+
+\add{This simple example demonstrates how the SIMD efficiency of shallow-vectorized OpenCL source-term evaluation depends on the size of the chemical model in question, i.e., the amount of computational work per source-term evaluation.
+The base chemical model for this example was the isopentanol model~\cite{Sarathy:2013jr} used throughout this article, and the same thermochemical state database described in~\cref{s:test_platforms} was used for source-term evaluation.
+As in~\cref{S:results} all reported results are based on \num{10} individual runs and in this example all cases were run on the \avx/ machine using the Intel OpenCL runtime.}
 
@@
 
-\add{First, the reactions in the isopentanol model were converted to simple reversible Arrhenius reactions by either simply dropping third-body efficiency calculations (third-body enhanced reactions), using the high pressure-limit coefficients (falloff\slash chemically-activated and P-Log reactions) or fitting Arrenhius parameters to the calculated rate constant at a fixed pressure (chebyshev reactions).}
-\add{This conversion made the cost of reaction rate evaluation roughly equivalent between all reactions in model, separating the effect of chemical model size from computational intensities of different reaction types on the SIMD efficiency.}
-\add{Next, a greedy reaction removal algorithm (\cref{a:model_gen}) was utilized to generate a number of models ranging from \numrange{2100}{186} reactions, in increments of 200 reactions (except the final increment from \num{200} to \num{186} reactions).}
+\add{First, the reactions in the isopentanol model were converted to simple reversible Arrhenius reactions by either simply dropping third-body efficiency calculations (third-body enhanced reactions), using the high-pressure-limit coefficients (falloff\slash chemically-activated and P-Log reactions) or fitting Arrhenius parameters to the calculated rate constant at a fixed pressure (Chebyshev reactions).
+This conversion made the cost of reaction rate evaluation roughly equivalent between all reactions in model, separating the effect of chemical model size from computational intensities of different reaction types on the SIMD efficiency.
+Next, a greedy reaction removal algorithm (\cref{a:model_gen}) generated a number of models ranging from \numrange{2100}{186} reactions, in increments of 200 reactions (except the final increment from \num{200} to \num{186} reactions).}
 
@@
 \includegraphics[width=0.5\linewidth]{simd_efficiency_scaling}
-\caption{The effect on SIMD efficiency of varying chemical model size}
+\caption{The effect on SIMD efficiency of varying chemical model size.}
 \label{F:simd_vs_rxns}
@@
 
-\add{In order to discern the effect of varying chemical model size on the SIMD efficiency, shallow-vectorized and unvectorized source-term evaluation performance tests were run.}
-\add{As demonstrated in~\cref{F:simd_vs_rxns} the SIMD efficiency is strongly dependent on the generated model size and and thus, amount of computational work per thermochemical state.}
-\add{In addition, the range of SIMD efficiency in this example (\numrange{0.56}{0.91}) is larger than the range of SIMD efficiencies calculated for real chemical models, as seen in~\cref{F:source_simd_scaling}.}
-\add{For smaller models---e.g., \ce{H2}\slash\ce{CO} which had a SIMD efficiency of \num{0.6} on the \avx/ CPU---this suggests that the presence of more computationally intensive fall-off\slash chemically activated reactions in model can increase the SIMD efficiency.}
-\add{However, the base isopentanol model achieved a SIMD efficiency of only \num{0.78} on the \avx/ machine in~\cref{S:source_results}, suggesting that more work could be done to optimize the source-term evaluations.}
-\add{In particular, it's likely that a reaction sorting method, such as suggested by Sewerin and Rigopoulos~\cite{Sewerin20151375}, would be particularly beneficial to reduce the number of vector-gather\slash scatter\slash masking operations incurred during source-term evaluation.}
+\add{To discern the effect of varying chemical model size on the SIMD efficiency, shallow-vectorized and unvectorized source-term evaluation performance tests were run.
+As demonstrated in~\cref{F:simd_vs_rxns} the SIMD efficiency strongly depends on the generated model size and thus the amount of computational work per thermochemical state.
+In addition, the range of SIMD efficiency in this example (\numrange{0.56}{0.91}) is larger than the range of SIMD efficiencies calculated for real chemical models, as seen in~\cref{F:source_simd_scaling}.
+For smaller models---e.g., \ce{H2}\slash\ce{CO} which had a SIMD efficiency of \num{0.6} on the \avx/ CPU---this suggests that the presence of more computationally intensive fall-off\slash chemically activated reactions in model can increase the SIMD efficiency.
+However, the base isopentanol model achieved a SIMD efficiency of only \num{0.78} on the \avx/ machine in~\cref{S:source_results}, suggesting that more work could be done to optimize the source-term evaluations.
+In particular, it is likely that a reaction sorting method, such as suggested by Sewerin and Rigopoulos~\cite{Sewerin20151375}, would be particularly beneficial to reduce the number of vector gather\slash scatter\slash masking operations incurred during source-term evaluation.}
 

### 2018-06-13 paper.tex
add availability appendix

@@
 \renewcommand*{\thesection}{\appendixname~\Alph{section}}
+
+\section{Availability of material}
+\label{A:availability}
+
+\add{The results for this article were obtained using \texttt{pyJac v2.x.x}.
+The most recent version of \texttt{pyJac} can be found at its GitHub repository:
+\url{https://github.com/SLACKHA/pyJac}.
+All figures, and the data and plotting scripts necessary to reproduce them,
+are available openly under the CC-BY license.}\todo{add citations for software and figures etc.\ archive}
 

### 2018-06-13 scripts/sparse_plotter.py
switch to png for sparsity figures to fix weird rendering on mac's and okular at different zoom levels

@@ -32,9 +32,9 @@ def plot(mech, jac_type, out=''):
     if out:
         if out.endswith('.pdf'):
             out = out[:out.index('.pdf')]
-        out += '_' + str(jac_type)[str(jac_type).index('.') + 1:] + '.pdf'
+        out += '_' + str(jac_type)[str(jac_type).index('.') + 1:] + '.png'
         plt.tight_layout()
-        plt.savefig(out)
+        plt.savefig(out, dpi=1000)
     else:
         plt.show()
 

### 2018-06-13 paper.tex
switch to png for sparsity figures to fix weird rendering on mac's and okular at different zoom levels
remove note
s/exampled/examined

@@
 However, the explicit-mass conservation formulation of \texttt{pyJac} can introduce additional non-zero entries in two ways.
-First, if the last species in the model (i.e., the bath gas) participates directly in any reaction, the derivative of its forward or reverse rate of progress \revise{is non-zero with respect to other species participating in that reaction.}\todo{Nick: check this}{}
-Similarly, if the last species has a third-body efficiency not equal to the default (one), \revise{this will again create nonzero derivatives for the pressure-modification term with respect to other species (see the supplemental material).}
+First, if the last species in the model (i.e., the bath gas) participates directly in any reaction, the derivative of its forward or reverse rate of progress \revise{is non-zero with respect to all other species in the model, regardless of whether the other species participate in that reaction or not.}
+Similarly, if the last species has a third-body efficiency not equal to the default (one), \revise{this will again create nonzero derivatives for the pressure-modification term with respect to all other species (see the supplemental material).}
 Either case will result in a fully dense Jacobian row for all species with a non-zero net stoichiometric coefficient in such a reaction.
@@
   \begin{subfigure}[t]{0.48\linewidth}
-      \includegraphics[width=\textwidth]{ch4_sparsity_exact.pdf}
+      \includegraphics[width=\textwidth]{ch4_sparsity_exact.png}
       \caption{The ``exact'' Jacobian.}
@@
   \begin{subfigure}[t]{0.48\linewidth}
-      \includegraphics[width=\textwidth]{ch4_sparsity_approximate.pdf}
+      \includegraphics[width=\textwidth]{ch4_sparsity_approximate.png}
       \caption{The ``approximate'' Jacobian.}
@@
   \label{F:jac_sparsity}
-  \todo[inline]{the figure on the left is noticeably blurry compared to the other... can we fix that?}
 \end{figure}
@@
 Finally, unless stated otherwise: the performance results used a single CPU core, the \conp/ assumption, a vector width of \num{8}\slash\num{128}, and ``C''\slash ``F''-ordered data for the CPU\slash GPU cases, respectively; the run times reported are for the number of conditions specified in~\cref{T:error} and include data-transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
-The effects of choice of vector width, data ordering, and differences between \conp/ and \conv/ evaluations on the CPU\slash GPU will be explored in~\cref{S:source_results}, while parallel scaling for multiple CPU cores will be exampled in~\cref{S:source_results,S:jacobian_results}.\todo{is ``exampled'' really the word you want here?}{}
+The effects of choice of vector width, data ordering, and differences between \conp/ and \conv/ evaluations on the CPU\slash GPU will be explored in~\cref{S:source_results}, while parallel scaling for multiple CPU cores will be examined in~\cref{S:source_results,S:jacobian_results}.
 

### 2018-06-13 paper.tex
revise and put specific changes in footnote

@@
 No speedup is achieved for either vectorization type on either CPU---indeed, the OpenMP case is faster on both CPUs\add{, though we stress that Intel OpenCL runtime achieves vectorization using the same code and hardware}.
-The reasons for this lack of vectorization with POCL are quite technical~\delete{ and somewhat outside the scope of this work}; however, \add{personal communication with the developers of POCL revealed that to achieve vectorization, changes are required to both the way POCL prepares LLVM \cite{Lattner:2004:LCF:977395.977673} IR-code---specifically, to specialize boundary cases to ensure LLVM recognizes uniform vectorization loop bounds, even if said bounds \textit{are} uniform in practice---as well as improvements to LLVM's loop-vectorizer: e.g., for proofs of a vector instruction's ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns~\cite{pocl_communication}.}\todo{this is one super long runon sentence---needs to be broken up and possibly simplified.}{}
+The reasons for this lack of vectorization with POCL are quite technical~\delete{ and somewhat outside the scope of this work}; however, \add{personal communication with the developers of POCL revealed that to achieve vectorization changes are required to both the way POCL prepares LLVM IR-code as well as improvements to LLVM's loop-vectorizer itself~\cite{pocl_communication}.\footnote{Specifically improvements such as: ensuring LLVM recognizes uniform vectorization loop bounds (even if said bounds \textit{are} uniform in practice), proving vector instruction's ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns.}}
 As will be discussed in~\cref{S:future}, \add{we hope that using explicit vector types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a validation tool.

### 2018-06-13 paper.tex
remove 256 fragment

@@
 \Cref{F:gpu_source_speedup} also shows that varying the vector-width minimally affects performance for most of the \gpunew/ and \gpuold/ cases; the GRI-Mech 3.0 and USC-Mech II models show the largest improvements with a vector width of 128: \textasciitilde\SIrange{10}{18}{\percent} for both GPUs.
-This is likely caused by higher occupancy on the GPUs, but it is unclear exactly how Nvidia's OpenCL runtime balances the registers\slash warps per streaming-multiprocessor, as controls occupancy in CUDA~\cite{occupancy};\todo{this last fragment is missing something}{}
-use of even larger vector widths (i.e., $\geq$\num{256}) on the GPU is a topic that merits future investigation.\todo[inline]{I feel like we should pull this last bit, since someone could ask why you didn't just run with 256}
+This is likely caused by higher occupancy on the GPUs, but it is unclear exactly how Nvidia's OpenCL runtime balances the registers\slash warps per streaming-multiprocessor, as controls occupancy in CUDA~\cite{occupancy}.
 

### 2018-06-13 paper.tex
use kyle's wording for simd efficiency

@@
 where $\bar{t}_{\text{unvec}}$ is the mean run time per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$ is the vector width reported in number of double operations (see~\cref{t:cpus}).
-This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.\todo{this sentence needs work, see alternates in src}{}
-%% you have both ``compares'' and ``compared'' in this sentence... would the following be correct?
-% This measure compares the actual speedup due to shallow vectorization with the ideal speedup based on the nominal vector width of the machine.
+This measure compares the actual speedup due to shallow vectorization with the ideal speedup based on the nominal vector width of the machine.
 \Cref{F:source_simd_scaling} shows the SIMD efficiency of source-term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models (isopentanol and USC-Mech II) have higher SIMD efficiencies of \numrange{0.76}{0.78}, and the smaller models (\ce{H2}\slash\ce{CO}, GRI-Mech \num{3.0}) have lower SIMD efficiencies of \numrange{0.6}{0.66}.

### 2018-06-13 paper.tex
add reference to OpenMP code

@@
 
-\Cref{F:dense_jac_scaling} shows scaling for the shallow-vectorized dense Jacobian OpenCL code.\todo{OpenCP and OpenMP?}{}
+\Cref{F:dense_jac_scaling} shows scaling for the OpenMP and shallow-vectorized dense Jacobian OpenCL codes.
 In this case, the isopentanol model scales the worst for both cases.

### 2018-06-13 paper.tex
revise pocl req.

@@
 However, POCL's implicit vectorization module---which uses the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to translate OpenCL code to vectorized machine code---typically fails to achieve much, if any, speedup.
-\revise{Thus POCL is useful for validation but not necessarily for performance studies.}\todo{should we emphasize here that it isn't needed for pyJac to function properly?}{}
+\revise{Thus POCL is useful for validation but not necessarily for performance studies; it is noted that although POCL is currently used for unit-testing purposes, it is not required by \texttt{pyJac}.}
 \revise{We will expand upon this discussion} in~\cref{S:future} to highlight future directions.

### 2018-06-13 paper.tex
add note about pocl not required

--- 2018-06-13 paper.tex
@@
 However, POCL's implicit vectorization module---which uses the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to translate OpenCL code to vectorized machine code---typically fails to achieve much, if any, speedup.
-\revise{Thus POCL is useful for validation but not necessarily for performance studies.}\todo{should we emphasize here that it isn't needed for pyJac to function properly?}{}
+\revise{Thus POCL is useful for validation but not necessarily for performance studies; it is noted that while POCL is currently used by \texttt{pyJac} for unit-testing purposes, it is not a required.}
 \revise{We will expand upon this discussion} in~\cref{S:future} to highlight future directions.

### 2018-06-13 paper.tex
remove norm indicies

@@
 \begin{equation}
-E_{X} = \left\lVert \frac{\left\lvert X_{ij,\text{CT}} - X_{ij}\right\rvert}{\num{e-10} + \num{e-6} \times \left\lvert X_{ij,\text{CT}} \right\rvert} \right\rVert_{i,j,\infty} \;,
+E_{X} = \left\lVert \frac{\left\lvert X_{ij,\text{CT}} - X_{ij}\right\rvert}{\num{e-10} + \num{e-6} \times \left\lvert X_{ij,\text{CT}} \right\rvert} \right\rVert_{\infty} \;,
 \label{e:rel_err}
@@
  \label{e:jac_error_base}
- E_{\text{rel}, k} = \left\lVert \frac{\hat{\mathcal{J}}_{ij,k} - \mathcal{J}_{ij,k}}{\hat{\mathcal{J}}_{ij,k}} \right\rVert_{ij,F} \;.
+ E_{\text{rel}, k} = \left\lVert \frac{\hat{\mathcal{J}}_{ij,k} - \mathcal{J}_{ij,k}}{\hat{\mathcal{J}}_{ij,k}} \right\rVert_{F} \;.
 \end{equation}
@@
  \label{e:thresh}
- \mathcal{T} = \left\lVert \mathcal{\hat{J}} \right\rVert_{k, F} \;.
+ \mathcal{T} = \left\lVert \mathcal{\hat{J}} \right\rVert_{F} \;.
 \end{equation}
 The error statistics reported in this section are then based only on matrix elements where $\mathcal{J}_{ijk} \ge \frac{\mathcal{T}}{\mathcal{C}}$, where $\mathcal{C}$ is \revise{a tunable} threshold parameter; this filtered form of~\cref{e:jac_error_base} is denoted $E_{\mathcal{C},k}$.
-Finally, the Frobenius norm is calculated over all the states in the PaSR thermochemical condition database:
+Finally, the Frobenius norm is calculated over all the states $k$ in the PaSR thermochemical condition database:
 \begin{equation}
  \label{e:thresholded_error}
- E_{\mathcal{C}} = \left\lVert E_{\mathcal{C},k} \right\rVert_{k, F} \;.
+ E_{\mathcal{C}} = \left\lVert E_{\mathcal{C},k} \right\rVert_{F} \;.
 \end{equation}
 
-This error norm is quite different from the relative error Frobenius norm suggested by Anderson et al.~\cite{Anderson:1999aa} for quantifying the error of matrices in LAPACK:
+This error norm is quite different from the relative error Frobenius norm suggested by Anderson et al.~\cite{Anderson:1999aa} for quantifying the error of matrices in LAPACK, \revise{e.g., over Jacobian indices $i, j$}:
 \begin{align}
  \label{e:error_lapack}
- \lhs{lapack}{E_{\mathcal{L}, k}} &=  \rhs{lapack}{\frac{\left\lVert \hat{\mathcal{J}}_{ijk} - \mathcal{J}_{ijk} \right\rVert_{ij,F}}{\left\lVert \hat{\mathcal{J}}_{ijk} \right\rVert_{i,j,F}}} \nonumber \quad \text{and} \\
- \lhs{lapack}{E_{\mathcal{L}}} &= \rhs{lapack}{\left\lVert  E_{\mathcal{L}, k} \right\rVert_{k, F}}[c] \;.
+ \lhs{lapack}{E_{\mathcal{L}, k}} &=  \rhs{lapack}{\frac{\left\lVert \hat{\mathcal{J}}_{ijk} - \mathcal{J}_{ijk} \right\rVert_{F}}{\left\lVert \hat{\mathcal{J}}_{ijk} \right\rVert_{F}}} \nonumber \\
+\intertext{and states $k$:}
+ \lhs{lapack}{E_{\mathcal{L}}} &= \rhs{lapack}{\left\lVert  E_{\mathcal{L}, k} \right\rVert_{F}}[c] \;.
 \end{align}

### 2018-06-13 paper.tex
reword pocl communications

@@
 No speedup is achieved for either vectorization type on either CPU---indeed, the OpenMP case is faster on both CPUs\add{, though we stress that Intel OpenCL runtime achieves vectorization using the same code and hardware}.
-The reasons for this lack of vectorization with POCL are quite technical~\delete{ and somewhat outside the scope of this work}; however, \add{personal communication with the developers of POCL revealed that to achieve vectorization changes are required to both the way POCL prepares LLVM IR-code as well as improvements to LLVM's loop-vectorizer itself~\cite{pocl_communication}.\footnote{Specifically improvements such as: ensuring LLVM recognizes uniform vectorization loop bounds (even if said bounds \textit{are} uniform in practice), proving vector instruction's ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns.}}
-As will be discussed in~\cref{S:future}, \add{we hope that using explicit vector types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a validation tool.
+The reasons for this lack of vectorization with POCL are quite technical~\delete{ and somewhat outside the scope of this work}; however, \add{personal communication with the developers of POCL revealed that to achieve vectorization, changes are required to both the way POCL prepares LLVM IR-code, as well as improvements to LLVM's loop-vectorizer itself~\cite{pocl_communication}.\footnote{Specifically improvements such as: ensuring LLVM recognizes uniform vectorization loop bounds (even if said bounds \textit{are} uniform in practice), proving vector instructions' ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns.}}
+As will be discussed in~\cref{S:opencl}, \add{we hope that using explicit vector types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a validation tool.
 

### 2018-06-13 paper.tex
first pass at revised conclusion

@@
 
-\section{Conclusions}
-
-In this work, we developed automatically generated OpenCL codes for SIMD and SIMT-vectorized evaluations of thermochemical source-term and sparse\slash dense chemical kinetic Jacobian matrices.
-We also verified and investigated the performance of source-term and Jacobian evaluation for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
-
-\todo[inline]{So I think some of the rest of this is really long and overly detailed---you don't need to rewrite all of the results, but really just the overall contributions and takeaways.
-I sorta think subsections are not appropriate (and help encourage overwriting...)}
-\add{The main contributions of this work are:}:
-
-\subsection{Derive and validate a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity}
-A new formulation resulted in highly sparse (\SIrange{28.6}{88.5}{\percent}) chemical kinetic Jacobians, with options to further increase sparsity (to \SIrange{31.66}{92.0}{\percent}) at the expense of strict correctness.
-
-\subsection{Enable cross-platform SIMD\slash SIMT vectorization for the CPU, GPU and other accelerators}
-\add{Shallow-vectorized evaluation of \texttt{pyJac}'s chemical kinetic source-term and analytical Jacobian was demonstrated on both the CPU and GPU via OpenCL.}
-Deep vectorization, though possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, did not yield any performance benefit as POCL did not achieve vectorizations for any execution pattern studied.
-Deep vectorization deserves further study with other platforms (e.g., CUDA).
-\add{In addition, parallel OpenMP evaluation was demonstrated on the CPU.}
-
-\subsection{Investigate the performance for a wide range chemical kinetic models, specifically looking at speedups due to SIMD\slash SIMT-vectorization as well as relative to the previous version of \texttt{pyJac}}
-
-Significant speedups in shallow SIMD-vectorized source-term evaluation over a parallel OpenMP code were demonstrated, up to \SIrange{2.53}{2.92}{$\times$} and \SIrange{3.40}{4.08}{$\times$} for \sse/ and \avx/ capable CPUs; in addition, source-term evaluation was \SIrange{1.40}{1.88}{$\times$} faster on a \gpunew/ GPU compared to a \gpuold/ GPU.
-Vectorized data-orderings specifically formulated for shallow\slash deep SIMD\slash SIMT vectorizations were proposed, resulting in a shallow-vectorized speedup of \SIrange{2.14}{2.58}{$\times$} over a more standard ``F''-ordered (column major) shallow-vectorization; when compared to the corresponding speedup of \SIrange{1.35}{2.09}{$\times$} for the ``C''-ordered (row major) OpenMP code over an ``F''-ordered equivalent, the value of the vectorized data-ordering is evident.
-In addition, the choice of \conp/\slash\conv/ formulation was determined to be negligible on source-term evaluations, while varying the OpenCL vector-width had only modest (\textasciitilde\SI{10}{$\percent$}) performance effects on both the CPU and GPU.
-The strong parallel scaling efficiency was also examined for source-term evaluation; the parallel OpenMP code generally had better efficiencies, but when compared to a shallow-vectorized code with roughly equivalent performance---e.g., source-term shallow-vectorized evaluation on \num{4} to \num{16} OpenMP cores---the scalings were similar.
-Examination of SIMD efficiency revealed increasing speedups with chemical model-size for source-term evaluation on the \avx/ CPU, while on the \sse/ CPU speedups larger than the nominal vector-width were achieved.
-
-The performance of sparse and dense Jacobian evaluation was also characterized on the CPU\slash GPU platforms.
-The sparse shallow-vectorized OpenCL code was \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.14}{5.03}{$\times$} faster than OpenMP evaluation on the \avx/ and \sse/ CPUs, respectively, while dense shallow-vectorized evaluation was \SIrange{3.03}{4.23}{$\times$} and \SIrange{1.92}{2.47}{$\times$} faster on the same.
-Sparse Jacobian evaluation was slower than dense evaluation on all CPU\slash GPU platforms due to indirect lookup requirements in array indexing, however the shallow-vectorized OpenCL code was less adversely affected---a slowdown of \SIrange{1.41}{3.34}{$\times$} on the \avx/ CPU---than OpenMP, which was \SIrange{2.47}{10.42}{$\times$} slower on the same.
-On the \gpunew/ GPU, sparse evaluation was \SIrange{1.10}{1.59}{$\times$} faster than on the \gpuold/ GPU, while dense evaluation was \SIrange{1.36}{3.0}{$\times$} faster; the use of OpenCL Image\slash CUDA Texture memory to accelerate the indirect lookup for sparse matrix evaluation should be investigated in the future.
-Analytical Jacobian evaluation was significantly faster than a first-order finite-difference based approach on all platforms, achieving speedups of \SIrange{3.81}{17.60}{$\times$} and \SIrange{4.04}{45.13}{$\times$} for sparse and dense evaluation (respectively) on the GPUs; on the \avx/ CPU analytical evaluation was \SIrange{3.92}{55.11}{$\times$} and \SIrange{9.68}{245.63}{$\times$} for sparse and dense Jacobian evaluation (respectively) with OpenMP\slash shallow-vectorized OpenCL.
-Finally, the performance of dense analytical-Jacobian evaluation in this new version of \texttt{pyJac} was compared to the previous version~\cite{pyjac16} on the \sse/\slash\gpuold/ CPU\slash GPU; OpenMP evaluation on the \sse/ CPU was \SI{1.79}{$\times$} slower for the smallest chemical model, but \SIrange{3.37}{10.19}{$\times$} faster for the larger models, while the shallow-vectorized OpenCL code was \SIrange{1.37}{19.56}{$\times$} faster than the previous version over all chemical models.
-Similarly, evaluation on the \gpuold/ GPU was \SI{1.89}{$\times$} slower for the smallest model but \SIrange{1.25}{2.84}{$\times$} faster for the larger models.
-
-\subsection{Future extensions to this work and practical notes on OpenCL usage}
-\label{S:future}
-
+\section{Practical notes on OpenCL Usage}
+\label{S:opencl}
 \revisedelete{It is important to note that while }While OpenCL provides a simple interface to enable cross-platform execution, and significant speedups were achieved via shallow-vectorized OpenCL code in this work, there are some serious potential pitfalls in its use.
@@
 
-\todo[inline]{Does the following really belong in the conclusions?}
 As demonstration of the type of issue discussed, a minimum working example has been created~\cite{Nvidia_mwe} that demonstrates a failure of Nvidia's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla \gpunew/ GPU; by simply changing to another runtime (e.g., Intel)---with no code changes or recompilation---the correct result can be found.
@@
 Indeed the vectorization process attempted by most OpenCL runtimes (and thus, reasons for incorrect\slash unvectorized code output) is obscured from the user, making reasoning about errors or performance trends quite difficult.
-
 Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change requires building of the wrapper that initializes\slash transfers memory and calls the source-rate\slash Jacobian kernel.
-Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler \texttt{\#pragma}s) is a key priority going forwards as OpenMP is a standard library on most machines.
+
+\section{Conclusions}
+\label{S:future}
+
+In this work, we developed automatically generated OpenCL codes for SIMD and SIMT-vectorized evaluations of thermochemical source-term and sparse\slash dense chemical kinetic Jacobian matrices.
+\revise{The main contributions of this work are:}
+\begin{itemize}
+ \item \revise{the derivation and validation of a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity.}
+ \item \revise{the extension of \texttt{pyJac} to enable vectorized execution on the CPU, GPU and other accelerators}
+ \item \revise{achievement of significant speedups over a strictly parallel Jacobian and source-term evaluation on SIMD-enabled processors (e.g., CPUs)}
+\end{itemize}
+
+\revise{The new formulation for \texttt{pyJac}, including both constant-volume and constant-pressure assumptions, resulted in highly sparse chemical kinetic Jacobians.}
+\revise{Though the Jacobian approximation discussed in~\cref{S:sparsity} is not a key feature of this work, future efforts to develop more sophisticated Jacobian approximations would be a worthwhile endevour.}
+
+We also demonstrated source-term and Jacobian evaluation for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
+\revise{In addition to parallel OpenMP evaluation on the CPU, the new version of \texttt{pyJac} enables the shallow-vectorized evaluation of \texttt{pyJac}'s chemical kinetic source-term and analytical Jacobian on both the CPU and GPU via OpenCL.}
+Deep vectorization, though possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, did not yield any performance benefit as POCL did not achieve vectorizations for any execution pattern studied.
+Deep vectorization deserves further study with other platforms (e.g., CUDA).
+
+Significant speedups in shallow SIMD-vectorized execution over a parallel OpenMP code were demonstrated for evaluations of the chemical kinetic source-term and sparse\slash dense Jacobian.
+Vectorized data-orderings specifically formulated for shallow\slash deep SIMD\slash SIMT vectorizations were proposed, and were shown to yield larger speedups in a shallow-vectorized execution as compared to standard data-ordering layouts in a parallel OpenMP code.
+In addition, the choice of \conp/\slash\conv/ formulation was determined to be negligible on source-term evaluations, while varying the OpenCL vector-width had only modest (\textasciitilde\SI{10}{$\percent$}) performance effects on both the CPU and GPU.
+The strong parallel scaling efficiency was also examined for source-term evaluation; the parallel OpenMP code generally had better efficiencies, but when compared to a shallow-vectorized code with roughly equivalent performance---e.g., source-term shallow-vectorized evaluation on \num{4} to \num{16} OpenMP cores---the scalings were similar.
+Examination of SIMD efficiency revealed increasing speedups with chemical model-size for source-term evaluation on the \avx/ CPU, while on the \sse/ CPU speedups larger than the nominal vector-width were achieved.
+
+Sparse Jacobian evaluation was slower than dense evaluation on all CPU\slash GPU platforms due to indirect lookup requirements in array indexing, however the shallow-vectorized OpenCL code was less adversely affected than OpenMP.
+Analytical Jacobian evaluation was significantly faster than a first-order finite-difference based approach on all platforms.
+Finally, the performance of dense analytical-Jacobian evaluation in this new version of \texttt{pyJac} was compared to the previous version~\cite{pyjac16}; OpenMP evaluation on the CPU was slightly slower for the smallest chemical model, but significantly faster for the larger models, while the shallow-vectorized OpenCL code was faster than the previous version over all chemical models.
+
+The OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler \texttt{\#pragma}s) is a key priority going forwards as OpenMP is a standard library on most machines.
 In addition, CUDA~\cite{Nvidia:2018} has been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017}, while Intel's open-source OpenCL alternative, ISPC~\cite{pharr2012ispc}, has been relatively stable and easy to work with during preliminary efforts with the unit-testing discussed in~\cref{s:unittest}.
+The use of OpenCL Image\slash CUDA Texture memory to accelerate the indirect lookup for sparse matrix evaluation should be investigated as well.
 The current deep-vectorization formulation would be executable for both CUDA and ISPC targets, as these languages implement double-precision atomic operations, further recommending their use.
 It is also possible, particularly for the Intel OpenCL\slash POCL runtimes, that better performance\slash stability might be achieved using so-called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}; \add{specifically this change is hoped to enable vectorization on the POCL runtime} and might also enable deep-vectorization on the Intel OpenCL runtime.
-\add{Finally, future extensions of this work will include additional target languages (e.g., vectorized OpenMP, CUDA, ISPC) to improve ease of use and reliability, as well as improvements to the existing OpenCL targets, e.g., to enable meaningful deep-vectorized evaluation, as well the implementation of reaction sorting~\cite{Sewerin20151375} to improve SIMD-efficiency (\cref{S:SIMD_scaling}).}
+Finally, future extensions of this work will include additional target languages (e.g., vectorized OpenMP, CUDA, ISPC) to improve ease of use and reliability, as well as improvements to the existing OpenCL targets, e.g., to enable meaningful deep-vectorized evaluation, \add{as well the implementation of reaction sorting~\cite{Sewerin20151375} to improve SIMD-efficiency (\cref{S:SIMD_scaling}).}
 

### 2018-06-13 paper.tex
cut more

@@
 Significant speedups in shallow SIMD-vectorized execution over a parallel OpenMP code were demonstrated for evaluations of the chemical kinetic source-term and sparse\slash dense Jacobian.
-Vectorized data-orderings specifically formulated for shallow\slash deep SIMD\slash SIMT vectorizations were proposed, and were shown to yield larger speedups in a shallow-vectorized execution as compared to standard data-ordering layouts in a parallel OpenMP code.
-In addition, the choice of \conp/\slash\conv/ formulation was determined to be negligible on source-term evaluations, while varying the OpenCL vector-width had only modest (\textasciitilde\SI{10}{$\percent$}) performance effects on both the CPU and GPU.
-The strong parallel scaling efficiency was also examined for source-term evaluation; the parallel OpenMP code generally had better efficiencies, but when compared to a shallow-vectorized code with roughly equivalent performance---e.g., source-term shallow-vectorized evaluation on \num{4} to \num{16} OpenMP cores---the scalings were similar.
-Examination of SIMD efficiency revealed increasing speedups with chemical model-size for source-term evaluation on the \avx/ CPU, while on the \sse/ CPU speedups larger than the nominal vector-width were achieved.
-
 Sparse Jacobian evaluation was slower than dense evaluation on all CPU\slash GPU platforms due to indirect lookup requirements in array indexing, however the shallow-vectorized OpenCL code was less adversely affected than OpenMP.
-Analytical Jacobian evaluation was significantly faster than a first-order finite-difference based approach on all platforms.
+Further, analytical Jacobian evaluation was significantly faster than a first-order finite-difference based approach on all platforms.
 Finally, the performance of dense analytical-Jacobian evaluation in this new version of \texttt{pyJac} was compared to the previous version~\cite{pyjac16}; OpenMP evaluation on the CPU was slightly slower for the smallest chemical model, but significantly faster for the larger models, while the shallow-vectorized OpenCL code was faster than the previous version over all chemical models.

### 2018-06-14 paper.tex
move
fixed location of paper info

@@
 
+\usepackage[hyphens]{url}
 \usepackage[breaklinks=true, linkcolor=blue, citecolor=blue, colorlinks=true]{hyperref}
@@
 
+\begin{document}
+
+\begin{frontmatter}
+
 \title{Using SIMD and SIMT vectorization to evaluate sparse chemical kinetic Jacobian matrices and thermochemical source terms}
 
-\author[1]{Nicholas Curtis\corref{corr}}
+\author[1]{Nicholas J.~Curtis\corref{corr}}
+\ead{nicholas.curtis@uconn.edu}
 \author[2]{Kyle E.~Niemeyer}
-\ead{nicholas.curtis@uconn.edu}
 \author[1]{Chih-Jen Sung}
@@
 
-\begin{document}
-
-\begin{frontmatter}
 %====================================================================
 \begin{abstract}
-\revise{The accurate prediction of key combustion phenomena in reactive-flow simulations, e.g., lean blow-out, extinction\slash ignition limits and pollutant formation, necessitates the use of detailed chemical kinetics.}
+\revise{Accurately predicting key combustion phenomena in reactive-flow simulations, e.g., lean blow-out, extinction\slash ignition limits and pollutant formation, necessitates the use of detailed chemical kinetics.}
 \revise{The large size and high levels of numerical stiffness typically present in chemical kinetic models relevant to transportation\slash power-generation applications make the efficient evaluation\slash factorization of the chemical kinetic Jacobian and thermochemical source-terms critical to the performance of reactive-flow codes.}
@@
 In addition, CUDA~\cite{Nvidia:2018} has been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017}, while Intel's open-source OpenCL alternative, ISPC~\cite{pharr2012ispc}, has been relatively stable and easy to work with during preliminary efforts with the unit-testing discussed in~\cref{s:unittest}.
-The use of OpenCL Image\slash CUDA Texture memory to accelerate the indirect lookup for sparse matrix evaluation should be investigated as well.
 The current deep-vectorization formulation would be executable for both CUDA and ISPC targets, as these languages implement double-precision atomic operations, further recommending their use.
 It is also possible, particularly for the Intel OpenCL\slash POCL runtimes, that better performance\slash stability might be achieved using so-called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}; \add{specifically this change is hoped to enable vectorization on the POCL runtime} and might also enable deep-vectorization on the Intel OpenCL runtime.
+The use of OpenCL Image\slash CUDA Texture memory to accelerate the indirect lookup for sparse matrix evaluation should be investigated as well.
 Finally, future extensions of this work will include additional target languages (e.g., vectorized OpenMP, CUDA, ISPC) to improve ease of use and reliability, as well as improvements to the existing OpenCL targets, e.g., to enable meaningful deep-vectorized evaluation, \add{as well the implementation of reaction sorting~\cite{Sewerin20151375} to improve SIMD-efficiency (\cref{S:SIMD_scaling}).}

### 2018-06-14 paper.tex
removed old todonote
edits to conclusions

@@
 where the \text{CT} subscript indicates values from Cantera~\cite{Cantera}.
-\todo[inline]{the $i,j$ subscript for the overall norm looks a little strange to me, is that typical?}
 
@@
 \begin{itemize}
- \item \revise{the derivation and validation of a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity.}
- \item \revise{the extension of \texttt{pyJac} to enable vectorized execution on the CPU, GPU and other accelerators}
- \item \revise{achievement of significant speedups over a strictly parallel Jacobian and source-term evaluation on SIMD-enabled processors (e.g., CPUs)}
+ \item \revise{Deriving and verifying a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity;}
+ \item \revise{Extending \texttt{pyJac} to enable vectorized execution on the CPU, GPU, and other accelerators; and}
+ \item \revise{Achieving significant speedups over a strictly parallel Jacobian and source-term evaluation on SIMD-enabled processors (e.g., CPUs).}
 \end{itemize}
 
-\revise{The new formulation for \texttt{pyJac}, including both constant-volume and constant-pressure assumptions, resulted in highly sparse chemical kinetic Jacobians.}
-\revise{Though the Jacobian approximation discussed in~\cref{S:sparsity} is not a key feature of this work, future efforts to develop more sophisticated Jacobian approximations would be a worthwhile endevour.}
+\revise{The new formulation for \texttt{pyJac}, including both constant-volume and constant-pressure assumptions, resulted in highly sparse chemical kinetic Jacobians.
+Though the Jacobian approximation discussed in~\cref{S:sparsity} is not a key feature of this work, future efforts to develop more sophisticated Jacobian approximations would be a worthwhile endevour.}
 

### 2018-06-14 paper.tex
edit in footnote

@@
 No speedup is achieved for either vectorization type on either CPU---indeed, the OpenMP case is faster on both CPUs\add{, though we stress that Intel OpenCL runtime achieves vectorization using the same code and hardware}.
-The reasons for this lack of vectorization with POCL are quite technical~\delete{ and somewhat outside the scope of this work}; however, \add{personal communication with the developers of POCL revealed that to achieve vectorization, changes are required to both the way POCL prepares LLVM IR-code, as well as improvements to LLVM's loop-vectorizer itself~\cite{pocl_communication}.\footnote{Specifically improvements such as: ensuring LLVM recognizes uniform vectorization loop bounds (even if said bounds \textit{are} uniform in practice), proving vector instructions' ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns.}}
+The reasons for this lack of vectorization with POCL are quite technical~\delete{ and somewhat outside the scope of this work}; however, \add{personal communication with the developers of POCL revealed that to achieve vectorization, changes are required to both the way POCL prepares LLVM IR-code, as well as improvements to LLVM's loop-vectorizer itself~\cite{pocl_communication}.\footnote{Specifically, improvements such as ensuring LLVM recognizes uniform vectorization loop bounds (even if said bounds \textit{are} uniform in practice), proving vector instructions' ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns.}}
 As will be discussed in~\cref{S:opencl}, \add{we hope that using explicit vector types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a validation tool.

### 2018-06-14 paper.tex
add data reference

@@
 \url{https://github.com/SLACKHA/pyJac}.
-All figures, and the data and plotting scripts necessary to reproduce them,
+All figures, and the data~\cite{data} and plotting scripts necessary to reproduce them,
 are available openly under the CC-BY license.}\todo{add citations for software and figures etc.\ archive}

### 2018-06-14 paper.tex
IR -> intermediate representation

@@
 No speedup is achieved for either vectorization type on either CPU---indeed, the OpenMP case is faster on both CPUs\add{, though we stress that Intel OpenCL runtime achieves vectorization using the same code and hardware}.
-The reasons for this lack of vectorization with POCL are quite technical~\delete{ and somewhat outside the scope of this work}; however, \add{personal communication with the developers of POCL revealed that to achieve vectorization, changes are required to both the way POCL prepares LLVM IR-code, as well as improvements to LLVM's loop-vectorizer itself~\cite{pocl_communication}.\footnote{Specifically, improvements such as ensuring LLVM recognizes uniform vectorization loop bounds (even if said bounds \textit{are} uniform in practice), proving vector instructions' ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns.}}
+The reasons for this lack of vectorization with POCL are quite technical~\delete{ and somewhat outside the scope of this work}; however, \add{personal communication with the developers of POCL revealed that to achieve vectorization, changes are required to both the way POCL prepares LLVM intermediate representation code, as well as improvements to LLVM's loop-vectorizer itself~\cite{pocl_communication}.\footnote{Specifically, improvements such as ensuring LLVM recognizes uniform vectorization loop bounds (even if said bounds \textit{are} uniform in practice), proving vector instructions' ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns.}}
 As will be discussed in~\cref{S:opencl}, \add{we hope that using explicit vector types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a validation tool.

### 2018-06-14 paper.tex
added pyjac-v2 citation

@@
 
-\add{The results for this article were obtained using \texttt{pyJac v2.x.x}.
+\add{The results for this article were obtained using \texttt{pyJac v2.x.x}~\cite{pyjac2}.
 The most recent version of \texttt{pyJac} can be found at its GitHub repository:
 \url{https://github.com/SLACKHA/pyJac}.
-All figures, and the data~\cite{data} and plotting scripts necessary to reproduce them,
-are available openly under the CC-BY license.}\todo{add citations for software and figures etc.\ archive}
+All figures, and the data and plotting scripts necessary to reproduce them,
+are available openly under the CC-BY license~\cite{data}.}
 

### 2018-06-15 paper.tex
fix pyjac version
sp

@@
 \revise{The new formulation for \texttt{pyJac}, including both constant-volume and constant-pressure assumptions, resulted in highly sparse chemical kinetic Jacobians.
-Though the Jacobian approximation discussed in~\cref{S:sparsity} is not a key feature of this work, future efforts to develop more sophisticated Jacobian approximations would be a worthwhile endevour.}
+Though the Jacobian approximation discussed in~\cref{S:sparsity} is not a key feature of this work, future efforts to develop more sophisticated Jacobian approximations would be a worthwhile endeavor.}
 
@@
 
-\add{The results for this article were obtained using \texttt{pyJac v2.x.x}~\cite{pyjac2}.
+\add{The results for this article were obtained using \texttt{pyJac v2.0.0b0}~\cite{pyjac2}.
 The most recent version of \texttt{pyJac} can be found at its GitHub repository:

### 2018-06-15 paper.tex
s/valid/verify

@@
 Selected chemical kinetic models covering the range of sizes typically used in reactive-flow simulations were used for demonstration.
-A new formulation of the chemical kinetic governing equations was derived and validated, \revise{resulting in} Jacobian sparsities of \SIrange{28.6}{92.0}{$\percent$} for the tested models.
+A new formulation of the chemical kinetic governing equations was derived and verified, \revise{resulting in} Jacobian sparsities of \SIrange{28.6}{92.0}{$\percent$} for the tested models.
 Speedups of \SIrange{3.40}{4.08}{$\times$} were found for shallow-vectorized OpenCL source-rate evaluation \revise{compared with} a parallel OpenMP code on an \avx/ central processing unit (CPU), increasing to \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.03}{4.23}{$\times$} for sparse and dense chemical kinetic Jacobian evaluation, respectively.
@@
 Perini et al.~\cite{Perini:2012gy} developed an analytical Jacobian code for constant-volume combustion, with additional options to increase sparsity (at the expense of strict correctness) and \revise{tabulate} temperature-dependent properties; \revise{they reported} an \SI{80}{\percent} speedup over a finite-difference-based Jacobian when used in a multidimensional reactive-flow simulation.
-Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but did not validate it outside the context of use with an implicit-integration technique.
+Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but did not verify it outside the context of use with an implicit-integration technique.
 In addition, since the Jacobian was based on an over-constrained system~\cite{HANSEN2018257}, the effect on strict conservation of mass\slash energy was not studied.
@@
 \add{Bauer et al.~\cite{Bauer:2014} used warp-specialization to improve GPU-vectorization over a standard data-parallel vectorization approach; they achieved speedups of up to \SIrange{2.81}{3.75}{$\times$}, \SIrange{1.91}{2.58}{$\times$}, and \SIrange{1.4}{1.5}{$\times$} for evaluating viscosity, species diffusion, and chemical source terms, respectively}.
-Niemeyer et al.~\cite{Niemeyer:2016aa} created and validated the open-source analytical chemical kinetic Jacobian code-generator, \texttt{pyJac}, which supports parallel execution on CPUs and SIMT execution on GPUs; \revise{\texttt{pyJac} enables} a speedup of \SIrange{3}{7.5}{$\times$} over a finite-difference Jacobian on the CPU.
+Niemeyer et al.~\cite{Niemeyer:2016aa} created and verified the open-source analytical chemical kinetic Jacobian code-generator, \texttt{pyJac}, which supports parallel execution on CPUs and SIMT execution on GPUs; \revise{\texttt{pyJac} enables} a speedup of \SIrange{3}{7.5}{$\times$} over a finite-difference Jacobian on the CPU.
 
@@
 \begin{itemize}
- \item Derive and validate a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity: \revise{\texttt{pyJac v2}};
+ \item Derive and verify a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity: \revise{\texttt{pyJac v2}};
  \item Enable cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
@@
 \end{itemize}
-To our knowledge, this is the first \add{open-source, validated} effort that vectorizes the evaluation of chemical-kinetic source terms and Jacobian matrices for any chemical model on a wide selection of platforms.
+To our knowledge, this is the first \add{open-source, verified} effort that vectorizes the evaluation of chemical-kinetic source terms and Jacobian matrices for any chemical model on a wide selection of platforms.
 
@@
 In addition, \texttt{loo.py} allows the user to more easily make changes to the structure of the generated program, e.g., the data ordering, vectorization, and threading patterns, as well as switch the target language for code generation (and more simply extend to additional languages, e.g., CUDA).
-Further, \texttt{loo.py} can execute developed subroutines from Python (natively for C code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL), enabling unit testing\slash validation for each component of the Jacobian or source terms.
+Further, \texttt{loo.py} can execute developed subroutines from Python (natively for C code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL), enabling unit testing\slash verification for each component of the Jacobian or source terms.
 The source terms and sub-components thereof (e.g., rates of progress, pressure-modification terms) are directly compared with Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{hogan2014fast,adept-v11} provides reference values for Jacobian sub-components.
 \revise{We use the} Portable OpenCL (POCL) implementation~\cite{poclIJPP} and OpenMP~\cite{dagum1998openmp} to perform OpenCL and C unit testing, respectively, on the continuous-integration framework Travis CI~\cite{travis:2018}.
-\revise{We will discuss validation} of the complete (as opposed to the sub-component testing discussed here) generated source-terms and Jacobian codes in detail in~\cref{s:validation}.
+\revise{We will discuss verification} of the complete (as opposed to the sub-component testing discussed here) generated source-terms and Jacobian codes in detail in~\cref{s:verification}.
 
@@
 
-\revise{We ran the} performance and validation studies for this work on a variety of CPU and GPU platforms.
+\revise{We ran the} performance and verification studies for this work on a variety of CPU and GPU platforms.
 \Cref{t:cpus} shows the number of cores, vector instruction set, and model of the CPUs used in this work; each CPU had both \texttt{v16.1.1} of the Intel OpenCL runtime~\cite{intelopencl:2018} and \texttt{v1.0} of the POCL~\cite{poclIJPP} runtime installed, both enabling OpenCL \texttt{v1.2} execution.
@@
 However, POCL's implicit vectorization module---which uses the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to translate OpenCL code to vectorized machine code---typically fails to achieve much, if any, speedup.
-\revise{Thus POCL is useful for validation but not necessarily for performance studies; it is noted that while POCL is currently used by \texttt{pyJac} for unit-testing purposes, it is not a required.}
+\revise{Thus POCL is useful for verification but not necessarily for performance studies; it is noted that while POCL is currently used by \texttt{pyJac} for unit-testing purposes, it is not a required.}
 \revise{We will expand upon this discussion} in~\cref{S:future} to highlight future directions.
@@
 
-\subsection{Source-term validation}
-\label{s:validation}
-\revise{We validated the reaction rates of progress (ROP), species production rates, and temperature rates in this study by comparing with values calculated using Cantera~\cite{Cantera}.}
+\subsection{Source-term verification}
+\label{s:verification}
+\revise{We verified the reaction rates of progress (ROP), species production rates, and temperature rates in this study by comparing with values calculated using Cantera~\cite{Cantera}.}
 However, special care must be taken due to floating-point arithmetic issues.
@@
 GRI-Mech 3.0 has a $\mathcal{T}$ roughly an order of magnitude larger and so the stringent error norm is significantly smaller: $\mathcal{O}(\num{e-7})$.
-Given the intricacy of floating-point error evaluation, the use of different languages and OpenCL platforms (the effect of these differences will be explored in \cref{A:per_platform}), and the general complexity of \texttt{pyJac} it would be exceedingly difficult to pinpoint an exact cause for this phenomenon, as was done in~\cref{s:validation}.
+Given the intricacy of floating-point error evaluation, the use of different languages and OpenCL platforms (the effect of these differences will be explored in \cref{A:per_platform}), and the general complexity of \texttt{pyJac} it would be exceedingly difficult to pinpoint an exact cause for this phenomenon, as was done in~\cref{s:verification}.
 To ensure no bugs or errors present in the Jacobians generated by \texttt{pyJac}, relaxed filtered error norms ($\mathcal{C} = 10^{15}$) are also presented for each model in~\cref{T:error}.
@@
 The performance studies in this work were run on the platforms listed in~\cref{t:cpus,t:gpus}.
-Run times in each case were averaged over ten runs, each using the same set of PaSR conditions used in validation.
+Run times in each case were averaged over ten runs, each using the same set of PaSR conditions used in verification.
 The OpenMP Jacobian\slash source-term kernels, as well as the OpenMP\slash OpenCL wrapping code (responsible for initializing\slash transferring memory, reading input, etc.) was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms and \texttt{gcc v4.8.5} on the \sse/\slash\gpuold/ machines.
@@
 The reasons for this lack of vectorization with POCL are quite technical~\delete{ and somewhat outside the scope of this work}; however, \add{personal communication with the developers of POCL revealed that to achieve vectorization, changes are required to both the way POCL prepares LLVM intermediate representation code, as well as improvements to LLVM's loop-vectorizer itself~\cite{pocl_communication}.\footnote{Specifically, improvements such as ensuring LLVM recognizes uniform vectorization loop bounds (even if said bounds \textit{are} uniform in practice), proving vector instructions' ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns.}}
-As will be discussed in~\cref{S:opencl}, \add{we hope that using explicit vector types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a validation tool.
+As will be discussed in~\cref{S:opencl}, \add{we hope that using explicit vector types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a verification tool.
 
@@
 \end{tabular}
-\caption{Summary of Jacobian matrix validation results for the Intel OpenCL runtime.
+\caption{Summary of Jacobian matrix verification results for the Intel OpenCL runtime.
 The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}), \conp/\slash \conv/, and sparse\slash dense Jacobians.
@@
 \end{tabular}
-\caption{Summary of Jacobian matrix validation results for the Portable OpenCL (POCL) runtime.
+\caption{Summary of Jacobian matrix verification results for the Portable OpenCL (POCL) runtime.
 The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}), \conp/\slash \conv/, and sparse\slash dense Jacobians.
@@
 \end{tabular}
-\caption{Summary of Jacobian matrix validation results for OpenMP execution.
+\caption{Summary of Jacobian matrix verification results for OpenMP execution.
 The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/, and sparse\slash dense Jacobians.
@@
 \end{tabular}
-\caption{Summary of Jacobian matrix validation results for Nvidia OpenCL execution.
+\caption{Summary of Jacobian matrix verification results for Nvidia OpenCL execution.
 The reported error statistics are the maximum filtered relative error $E_\mathcal{C}$ and LAPACK error $E_{\mathcal{L}}$ over all vectorization patterns (\cref{t:platforms}),  \conp/\slash \conv/, and sparse\slash dense Jacobians.

### 2018-06-15 paper.tex
put a *few* numbers back in the conclusions

@@
 
-Significant speedups in shallow SIMD-vectorized execution over a parallel OpenMP code were demonstrated for evaluations of the chemical kinetic source-term and sparse\slash dense Jacobian.
-Sparse Jacobian evaluation was slower than dense evaluation on all CPU\slash GPU platforms due to indirect lookup requirements in array indexing, however the shallow-vectorized OpenCL code was less adversely affected than OpenMP.
-Further, analytical Jacobian evaluation was significantly faster than a first-order finite-difference based approach on all platforms.
-Finally, the performance of dense analytical-Jacobian evaluation in this new version of \texttt{pyJac} was compared to the previous version~\cite{pyjac16}; OpenMP evaluation on the CPU was slightly slower for the smallest chemical model, but significantly faster for the larger models, while the shallow-vectorized OpenCL code was faster than the previous version over all chemical models.
+\todo{kyle double check that this paragraph reads well}
+\revise{Significant speedups in shallow SIMD-vectorized execution over a parallel OpenMP code were demonstrated for evaluations of the chemical kinetic source-term and sparse\slash dense Jacobian: up to \SI{4.09}{$\times$}, \SI{4.23}{$\times$} and \SI{9.44}{$\times$}, respectively, on an \avx/-capable CPU.}
+\revise{Sparse Jacobian evaluation was slower than dense evaluation on all CPU\slash GPU platforms due to indirect lookup requirements in array indexing, however the shallow-vectorized OpenCL code was less adversely affected than OpenMP.}
+\revise{Further, analytical Jacobian evaluation was significantly faster than a first-order finite-difference based approach on all platforms.}
+\revise{Finally, the performance of dense analytical-Jacobian evaluation in this new version of \texttt{pyJac} was compared to the previous version~\cite{pyjac16}; OpenMP evaluation on the CPU was moderately slower for the smallest chemical model (e.g., \SI{1.79}{$\times$} on the \avx/ CPU), but significantly faster (up to \SI{10.19}{$\times$}) for the larger models, while the shallow-vectorized OpenCL code was faster than the previous version over all chemical models, reaching speedups of \SI{19.56}{$\times$}.}
 

### 2018-06-15 paper.tex
minor edits

@@
 
-\add{Relevant to all of the aforementioned efforts, Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermochemical state vectors and the resulting effect on consistency and errors in conserved properties such as mass and energy.
+\revise{Relevant to all of the aforementioned efforts, Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermochemical state vectors and the resulting effect on consistency and errors in conserved properties such as mass and energy.
 They also characterized how the choice of state vector affects implicit\slash linearly implicit integration algorithms and chemical mode analysis techniques.
@@
 A number of recent works have investigated \revise{using} high-performance SIMT devices \revise{like GPUs} to accelerate reactive-flow and chemical kinetics simulations.
-Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup \revisedelete{on a Tesla C1060} compared to a \revise{CPU-based} serial implementation \revisedelete{on a AMD-Opteron processor}.
-Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order-of-magnitude speedup for homogeneous reactor simulations of large chemical models \revisedelete{on a Tesla 2050 GPU} over a serial\revisedelete{Intel i7 930} CPU implementation.
+Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup \revisedelete{on a Tesla C1060} compared to a \revise{CPU-based} serial implementation\revisedelete{on a AMD-Opteron processor}.
+Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order-of-magnitude speedup for homogeneous reactor simulations of large chemical models \revisedelete{on a Tesla 2050 GPU} over a serial \revisedelete{Intel i7 930} CPU implementation.
 Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential \revise{CPU} code\revisedelete{on an Intel Xeon X5650 CPU}.
 Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, where a traditional implicit integrator handled the stiffest computational cells on a CPU and a stabilized-explicit solver \revisedelete{on a Tesla C2050 GPU} solved the less-stiff cells on a GPU; this hybrid solution technique \revise{performs} \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
-Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code \revisedelete{on a Tesla C2070} as compared with a sequential \revise{CPU} version of the same \revisedelete{on an Intel Xeon X5650 CPU}.
+Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code \revisedelete{on a Tesla C2070} as compared with a sequential \revise{CPU} version of the same\revisedelete{on an Intel Xeon X5650 CPU}.
 Stone and Davis~\cite{Stone:2013aa} investigated a GPU-based version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order-of-magnitude speedup over a serial CPU implementation.
@@
 Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} \revisedelete{on a Tesla C2075} \revise{on the GPU and CPU}.
-The \revise{GPU-based} implicit Runge--Kutta method performs equivalently to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38} \revisedelete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models \revisedelete{with an integration time step of $10^{-6}$ s}.
+The \revise{GPU-based} implicit Runge--Kutta method performed equivalently to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38} \revisedelete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models \revisedelete{with an integration time step of $10^{-6}$ s}.
 
@@
 Further, in ``C'' ordering \revise{any two species concentrations within the same thermochemical state lie at most $N_{\text{sp}}$ locations away}, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data locality increases the chances of a cache hit on the CPU, but may lead to uncoalesced memory accesses on the GPU.
-In addition, deep vectorization requires both synchronization between SIMD lanes\slash SIMT threads via memory fences\slash barriers.
-\revise{This is an expensive operation, may result} in SIMD waste or SIMT thread divergence caused by different lanes\slash threads executing different instructions (e.g., resulting from different if\slash then branches).
+\revisedelete{In addition,} Deep vectorization requires synchronization between SIMD lanes\slash SIMT threads via memory fences\slash barriers, \revise{a potentially expensive operation}.
+\revise{In addition, deep vectorization may result} in SIMD waste or SIMT thread divergence caused by different lanes\slash threads executing different instructions (e.g., resulting from different if\slash then branches).
 Shallow vectorization may also experience SIMD waste or SIMT thread divergence, e.g., in chemical kinetic integration due to varying internal solver time-step sizes~\cite{CurtisGPU:2017}.
-\revise{However, here this largely unaffects shallow vectorization} \add{as the only major code paths that differ between vector lanes are high\slash low-temperature polynomial evaluations and differing pressures for P-Log reactions, which cause far fewer issues compared with differing internal ODE integration time-steps~\cite{CurtisGPU:2017}}.
+\revise{However, in this work shallow vectorization is largely unaffected by this concern} \add{as the only major code paths that differ between vector lanes are high\slash low-temperature polynomial evaluations and differing pressures for P-Log reactions, which cause far fewer issues compared with differing internal ODE integration time-steps~\cite{CurtisGPU:2017}}.
 
@@
 However, POCL's implicit vectorization module---which uses the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to translate OpenCL code to vectorized machine code---typically fails to achieve much, if any, speedup.
-\revise{Thus POCL is useful for verification but not necessarily for performance studies; it is noted that while POCL is currently used by \texttt{pyJac} for unit-testing purposes, it is not a required.}
+\revise{Thus POCL is useful for verification but not necessarily for performance studies; it is noted that while POCL is currently used by \texttt{pyJac} for unit-testing purposes, it is not required to use \texttt{pyJac}.}
 \revise{We will expand upon this discussion} in~\cref{S:future} to highlight future directions.

### 2018-06-15 paper.tex
fixes

@@
 \revise{Such solvers do not need the} exact form of the Jacobian and thus the so-called ``approximate'' form is preferable.
-\add{Though this might be used as a crude form of preconditioning for such solvers, the primary purpose is merely to increase Jacobian sparsity; McNenly et al.~\cite{MCNENLY2015581} more thorough investigated preconditioners.}
+\add{Though this might be used as a crude form of preconditioning for such solvers, the primary purpose is merely to increase Jacobian sparsity; McNenly et al.~\cite{MCNENLY2015581} more thoroughly investigated preconditioners.}
 Hence, in this section we will detail the sparsity of both forms of the Jacobian for the chemical models tested.
@@
 \todo{kyle double check that this paragraph reads well}
-\revise{Significant speedups in shallow SIMD-vectorized execution over a parallel OpenMP code were demonstrated for evaluations of the chemical kinetic source-term and sparse\slash dense Jacobian: up to \SI{4.09}{$\times$}, \SI{4.23}{$\times$} and \SI{9.44}{$\times$}, respectively, on an \avx/-capable CPU.}
+\revise{Significant speedups in shallow SIMD-vectorized execution over a parallel OpenMP code were demonstrated for evaluations of the chemical kinetic source-term and sparse\slash dense Jacobian: up to \SI{4.09}{$\times$}, \SI{9.44}{$\times$} and \SI{4.23}{$\times$}, respectively, on an \avx/-capable CPU.}
 \revise{Sparse Jacobian evaluation was slower than dense evaluation on all CPU\slash GPU platforms due to indirect lookup requirements in array indexing, however the shallow-vectorized OpenCL code was less adversely affected than OpenMP.}

### 2018-06-15 derivations.tex
correct paper name in supplimental

@@
 \begin{abstract} % not to exceed 200 words
-This document supplements the work, \textit{SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source-Term Evaluation} and presents the system of chemical kinetic equations evaluated by the new version of code-generation platform, \texttt{pyJac-v2}.
+This document supplements the work, \textit{Using SIMD and SIMT vectorization to evaluate sparse chemical kinetic Jacobian matrices and thermochemical source terms} and presents the system of chemical kinetic equations evaluated by the new version of code-generation platform, \texttt{pyJac-v2}.
 First, the thermo-chemical source terms for fixed-mass, adiabatic constant-pressure\slash constant-volume (\conp/\slash\conv/) reactors are given, followed by the derivation of the sparse analytical chemical kinetic Jacobian for both reactors.
@@
 \section{Introduction}
-This document is a supplement to the paper---\textit{SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source-Term Evaluation}---and is provided as a reference guide to the equations evaluated by \texttt{pyJac} v2.0, a code-generation platform for the evaluatation of chemical source terms and analytical chemical kinetic Jacobians of constant-pressure\slash constant-volume, fixed-mass, adiabatic reactors.
+This document is a supplement to the paper---\textit{Using SIMD and SIMT vectorization to evaluate sparse chemical kinetic Jacobian matrices and thermochemical source terms}---and is provided as a reference guide to the equations evaluated by \texttt{pyJac} v2.0, a code-generation platform for the evaluatation of chemical source terms and analytical chemical kinetic Jacobians of constant-pressure\slash constant-volume, fixed-mass, adiabatic reactors.
 Note that equations specific to the constant-pressure or constant-volume derivation will be marked with \conp/ or \conv/ respectively.

### 2018-06-15 paper.tex
edit in conclusions

@@
 
-\revise{The new formulation for \texttt{pyJac}, including both constant-volume and constant-pressure assumptions, resulted in highly sparse chemical kinetic Jacobians.
-Though the Jacobian approximation discussed in~\cref{S:sparsity} is not a key feature of this work, future efforts to develop more sophisticated Jacobian approximations would be a worthwhile endeavor.}
+\revise{The new molar-concentration-based formulation for \texttt{pyJac} resulted in highly sparse chemical kinetic Jacobians,
+and allows selection of either the constant-volume or constant-pressure approximation.
+In addition, sparsity can be increased further by eliminating components associated with the bath gas, as discussed in \cref{S:sparsity};
+this approximation is not a key feature of this work, and future efforts to develop more sophisticated Jacobian approximations based on the \texttt{pyJac} software would be a worthwhile endeavor.}
 

### 2018-06-15 paper.tex
more conclusion edits

@@
 We also demonstrated source-term and Jacobian evaluation for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
-\revise{In addition to parallel OpenMP evaluation on the CPU, the new version of \texttt{pyJac} enables the shallow-vectorized evaluation of \texttt{pyJac}'s chemical kinetic source-term and analytical Jacobian on both the CPU and GPU via OpenCL.}
-Deep vectorization, though possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, did not yield any performance benefit as POCL did not achieve vectorizations for any execution pattern studied.
+\revise{In addition to parallel OpenMP evaluation on the CPU, the new version of \texttt{pyJac} enables the shallow-vectorized evaluation of the chemical-kinetic source terms and analytical Jacobian on both the CPU and GPU via OpenCL.}
+Deep vectorization is possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, but it yields no performance benefit as POCL did not achieve vectorizations for any execution pattern studied.
 Deep vectorization deserves further study with other platforms (e.g., CUDA).
@@
 \todo{kyle double check that this paragraph reads well}
-\revise{Significant speedups in shallow SIMD-vectorized execution over a parallel OpenMP code were demonstrated for evaluations of the chemical kinetic source-term and sparse\slash dense Jacobian: up to \SI{4.09}{$\times$}, \SI{9.44}{$\times$} and \SI{4.23}{$\times$}, respectively, on an \avx/-capable CPU.}
+\revise{Significant speedups in shallow SIMD-vectorized execution over a parallel OpenMP code were demonstrated for evaluations of the chemical kinetic source-term and sparse\slash dense Jacobian: up to \SI{4.09}{$\times$}, \SI{9.44}{$\times$}, and \SI{4.23}{$\times$}, respectively, on an \avx/-capable CPU.}
 \revise{Sparse Jacobian evaluation was slower than dense evaluation on all CPU\slash GPU platforms due to indirect lookup requirements in array indexing, however the shallow-vectorized OpenCL code was less adversely affected than OpenMP.}

### 2018-06-15 paper.tex
edited new conclusion paragraph, removed note

@@
 
-\todo{kyle double check that this paragraph reads well}
-\revise{Significant speedups in shallow SIMD-vectorized execution over a parallel OpenMP code were demonstrated for evaluations of the chemical kinetic source-term and sparse\slash dense Jacobian: up to \SI{4.09}{$\times$}, \SI{9.44}{$\times$}, and \SI{4.23}{$\times$}, respectively, on an \avx/-capable CPU.}
-\revise{Sparse Jacobian evaluation was slower than dense evaluation on all CPU\slash GPU platforms due to indirect lookup requirements in array indexing, however the shallow-vectorized OpenCL code was less adversely affected than OpenMP.}
-\revise{Further, analytical Jacobian evaluation was significantly faster than a first-order finite-difference based approach on all platforms.}
-\revise{Finally, the performance of dense analytical-Jacobian evaluation in this new version of \texttt{pyJac} was compared to the previous version~\cite{pyjac16}; OpenMP evaluation on the CPU was moderately slower for the smallest chemical model (e.g., \SI{1.79}{$\times$} on the \avx/ CPU), but significantly faster (up to \SI{10.19}{$\times$}) for the larger models, while the shallow-vectorized OpenCL code was faster than the previous version over all chemical models, reaching speedups of \SI{19.56}{$\times$}.}
+\revise{We demonstrated significant speedups in shallow SIMD-vectorized execution over a parallel OpenMP code for evaluating the chemical-kinetic source terms and sparse\slash dense Jacobian: up to \SI{4.09}{$\times$}, \SI{9.44}{$\times$}, and \SI{4.23}{$\times$}, respectively, on an \avx/-capable CPU.
+Sparse Jacobians evaluate more slowly than dense Jacobians on all CPU\slash GPU platforms due to indirect lookup requirements in array indexing, but this adversely affects the shallow-vectorized OpenCL code less than OpenMP.
+Further, analytical Jacobians evaluate significantly faster than a first-order finite-difference-based approach on all platforms.
+Finally, we compared the performance of evaluating dense, analytical Jacobians in this new version of \texttt{pyJac} with the previous version~\cite{pyjac16}.
+The OpenMP version evaluates moderately slower on the CPU for the smallest chemical model (e.g., \SI{1.79}{$\times$} on the \avx/ CPU), but significantly faster for the larger models---up to \SI{10.19}{$\times$}.
+The shallow-vectorized OpenCL code runs faster than the previous version over all chemical models, reaching speedups of \SI{19.56}{$\times$}.}
 

### 2018-06-15 paper.tex
final edits to conclusions

@@
 
-The OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler \texttt{\#pragma}s) is a key priority going forwards as OpenMP is a standard library on most machines.
+The OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler \texttt{\#pragma}s) is a key priority going forwards since OpenMP is a standard library on most machines.
 In addition, CUDA~\cite{Nvidia:2018} has been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017}, while Intel's open-source OpenCL alternative, ISPC~\cite{pharr2012ispc}, has been relatively stable and easy to work with during preliminary efforts with the unit-testing discussed in~\cref{s:unittest}.
 The current deep-vectorization formulation would be executable for both CUDA and ISPC targets, as these languages implement double-precision atomic operations, further recommending their use.
-It is also possible, particularly for the Intel OpenCL\slash POCL runtimes, that better performance\slash stability might be achieved using so-called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}; \add{specifically this change is hoped to enable vectorization on the POCL runtime} and might also enable deep-vectorization on the Intel OpenCL runtime.
-The use of OpenCL Image\slash CUDA Texture memory to accelerate the indirect lookup for sparse matrix evaluation should be investigated as well.
-Finally, future extensions of this work will include additional target languages (e.g., vectorized OpenMP, CUDA, ISPC) to improve ease of use and reliability, as well as improvements to the existing OpenCL targets, e.g., to enable meaningful deep-vectorized evaluation, \add{as well the implementation of reaction sorting~\cite{Sewerin20151375} to improve SIMD-efficiency (\cref{S:SIMD_scaling}).}
+It is also possible, particularly for the Intel OpenCL\slash POCL runtimes, that better performance\slash stability might be achieved using so-called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}.
+\add{Specifically, this change could enable vectorization on the POCL runtime} and might also enable deep vectorization on the Intel OpenCL runtime.
+Using OpenCL Image\slash CUDA Texture memory to accelerate the indirect lookup for sparse matrix evaluation should be investigated as well.
+Finally, future extensions of this work will include extending to additional target languages (e.g., vectorized OpenMP, CUDA, ISPC) to improve ease of use and reliability, improving the existing OpenCL targets (e.g., to enable meaningful deep-vectorized evaluation), \add{and implementing reaction sorting~\cite{Sewerin20151375} to improve SIMD efficiency (\cref{S:SIMD_scaling}).}
 

### 2018-06-15 paper.tex
edits in section 4

@@
 \label{S:opencl}
-\revisedelete{It is important to note that while }While OpenCL provides a simple interface to enable cross-platform execution, and significant speedups were achieved via shallow-vectorized OpenCL code in this work, there are some serious potential pitfalls in its use.
+
+\revisedelete{It is important to note that while }
+\revise{While OpenCL provides a simple interface to enable cross-platform execution, and significant speedups were achieved via shallow-vectorized OpenCL code in this work, there are some serious potential pitfalls in its use.
 The closed-source OpenCL runtimes tested in this work (Intel and Nvidia) contain several bugs that result in compilation failures, simply incorrect vectorized machine code, or even segmentation faults.
 Further, these runtimes (in our experience) tend to be less responsive to fixing said bugs, with relatively infrequent new releases (or in Nvidia's case, lack of changelogs\slash public records of bug-fixes).
-On the other hand, the open-source OpenCL runtime used in this work, POCL, has far fewer implementation bugs, and when issues arise the community is very responsive to bug reports and user outreach; however, POCL fails to achieve vectorization as noted in~\cref{S:source_results}.
-
-As demonstration of the type of issue discussed, a minimum working example has been created~\cite{Nvidia_mwe} that demonstrates a failure of Nvidia's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla \gpunew/ GPU; by simply changing to another runtime (e.g., Intel)---with no code changes or recompilation---the correct result can be found.
-This provides a particularly vexing problem for the programmer, as there is often little that can be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
-Further, as noted throughout this work certain code-generation patterns can break OpenCL execution\slash vectorization, e.g., the failure of POCL to achieve vectorization or Intel OpenCL's failure to vectorize the finite-difference Jacobian.
+On the other hand, the open-source OpenCL runtime used in this work, POCL, has far fewer implementation bugs, and when issues arise the community is very responsive to bug reports and user outreach; however, POCL fails to achieve vectorization as noted in~\cref{S:source_results}.}
+
+\revise{To demonstrate the type of issue discussed, we created a minimum working example~\cite{Nvidia_mwe} that shows a failure of Nvidia's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla \gpunew/ GPU; simply changing to another runtime (e.g., Intel) produces the correct result---with no code changes or recompilation.
+This provides a particularly vexing problem for the programmer, since little can often be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
+Further, as noted throughout this work, certain code-generation patterns can break OpenCL execution\slash vectorization, e.g., the failure of POCL to achieve vectorization or Intel OpenCL's failure to vectorize the finite-difference Jacobian.
 Indeed the vectorization process attempted by most OpenCL runtimes (and thus, reasons for incorrect\slash unvectorized code output) is obscured from the user, making reasoning about errors or performance trends quite difficult.
-Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change requires building of the wrapper that initializes\slash transfers memory and calls the source-rate\slash Jacobian kernel.
+Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change requires building of the wrapper that initializes\slash transfers memory and calls the source-rate\slash Jacobian kernel.}
 

### 2018-06-16 paper.tex
minor tweaks

@@
 
-\revise{The new molar-concentration-based formulation for \texttt{pyJac} resulted in highly sparse chemical kinetic Jacobians,
+\revise{The new molar-based formulation for \texttt{pyJac} resulted in highly sparse chemical kinetic Jacobians,
 and allows selection of either the constant-volume or constant-pressure approximation.

### 2018-06-18 paper.tex

@@
 
+% https://tex.stackexchange.com/a/5896/56227
+\usepackage{etoolbox}
+\newtoggle{markup}
+\togglefalse{markup}
+%\toggletrue{markup}
+
 \usepackage[xcolor]{changebar}
-\newcommand{\add}[1]{{\sloppy\cbcolor{OliveGreen}\textcolor{OliveGreen}{\cbstart {#1}\cbend}}}  % add
-\newcommand{\delete}[1]{\sloppy\cbcolor{RedOrange}\textcolor{RedOrange}{\cbdelete\sout{#1}}}
-\newcommand{\revise}[1]{{\sloppy\textcolor{RoyalPurple}{#1}}}  % add
-\newcommand{\revisedelete}[1]{\sloppy\textcolor{RoyalPurple}{\sout{#1}}} %delete
-
-\newcommand{\addnl}[1]{{\sloppy\textcolor{OliveGreen}{#1}}}  % add
-\newcommand{\deletenl}[1]{\sloppy\textcolor{RedOrange}{\sout{#1}}} %delete with no strikeout
-
-% shared equations in separate file
-% https://yatb.giacomodrago.com/en/post/3/latex-loading-equations-from-an-external-file.html
-\usepackage{catchfilebetweentags}
-
-% external references https://tex.stackexchange.com/a/14365/56227
-\usepackage{xr-hyper}
-
-\usepackage[version=4]{mhchem} % Formula subscripts using \ce{}, e.g., \ce{H2SO4}
-\usepackage{amsmath,eqparbox,xparse,mathtools}
-\usepackage[binary-units]{siunitx}
-
-\usepackage[textsize=footnotesize,textwidth=1.5\linewidth]{todonotes}
-
-% load common macros / commands
-\usepackage{macros}
+
 % macro for small underscores for variable names in appendix
@@
 \newcommand{\smallunderscore}{\textscale{.5}{\textunderscore}\xspace}
+
+\iftoggle{markup}{
+\newcommand{\add}[1]{{\sloppy\cbcolor{OliveGreen}\textcolor{OliveGreen}{\cbstart {#1}\cbend}}\xspace}  % add
+\newcommand{\delete}[1]{\sloppy\cbcolor{RedOrange}\textcolor{RedOrange}{\cbdelete\sout{#1}}\xspace}
+\newcommand{\revise}[1]{{\sloppy\textcolor{RoyalPurple}{#1}}\xspace}  % add
+\newcommand{\revisedelete}[1]{\sloppy\textcolor{RoyalPurple}{\sout{#1}}\xspace} %delete
+
+\newcommand{\addnl}[1]{{\sloppy\textcolor{OliveGreen}{#1}}\xspace}  % add
+\newcommand{\deletenl}[1]{\sloppy\textcolor{RedOrange}{\sout{#1}}\xspace} %delete with no strikeout
+}{
+\newcommand{\add}[1]{#1\xspace}  % dummy add
+\newcommand{\delete}[1]{\xspace} % dummy delete
+\newcommand{\revise}[1]{#1\xspace}  % dummy add
+\newcommand{\revisedelete}[1]{\xspace} %dummy delete
+
+\newcommand{\addnl}[1]{#1\xspace}  % dummy add
+\newcommand{\deletenl}[1]{\xspace} % dummy delete
+}
+
+% shared equations in separate file
+% https://yatb.giacomodrago.com/en/post/3/latex-loading-equations-from-an-external-file.html
+\usepackage{catchfilebetweentags}
+
+% external references https://tex.stackexchange.com/a/14365/56227
+\usepackage{xr-hyper}
+
+\usepackage[version=4]{mhchem} % Formula subscripts using \ce{}, e.g., \ce{H2SO4}
+\usepackage{amsmath,eqparbox,xparse,mathtools}
+\usepackage[binary-units]{siunitx}
+
+\usepackage[textsize=footnotesize,textwidth=1.5\linewidth]{todonotes}
+
+% load common macros / commands
+\usepackage{macros}
 
@@
 A number of recent works have investigated \revise{using} high-performance SIMT devices \revise{like GPUs} to accelerate reactive-flow and chemical kinetics simulations.
-Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup \revisedelete{on a Tesla C1060} compared to a \revise{CPU-based} serial implementation\revisedelete{on a AMD-Opteron processor}.
-Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order-of-magnitude speedup for homogeneous reactor simulations of large chemical models \revisedelete{on a Tesla 2050 GPU} over a serial \revisedelete{Intel i7 930} CPU implementation.
-Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential \revise{CPU} code\revisedelete{on an Intel Xeon X5650 CPU}.
-Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, where a traditional implicit integrator handled the stiffest computational cells on a CPU and a stabilized-explicit solver \revisedelete{on a Tesla C2050 GPU} solved the less-stiff cells on a GPU; this hybrid solution technique \revise{performs} \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
-Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code \revisedelete{on a Tesla C2070} as compared with a sequential \revise{CPU} version of the same\revisedelete{on an Intel Xeon X5650 CPU}.
+Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup\revisedelete{ on a Tesla C1060} compared to a\revise{ CPU-based} serial implementation\revisedelete{on a AMD-Opteron processor}.
+Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order-of-magnitude speedup for homogeneous reactor simulations of large chemical models \revisedelete{on a Tesla 2050 GPU} over a serial\revisedelete{ Intel i7 930} CPU implementation.
+Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a\revisedelete{ Tesla C2075} GPU, achieving a speedup of nearly two orders of magnitude over a sequential \revise{CPU} code\revisedelete{on an Intel Xeon X5650 CPU}.
+Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, where a traditional implicit integrator handled the stiffest computational cells on a CPU and a stabilized-explicit solver\revisedelete{ on a Tesla C2050 GPU} solved the less-stiff cells on a GPU; this hybrid solution technique \revise{performs} \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
+Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code\revisedelete{ on a Tesla C2070} as compared with a sequential\revise{ CPU} version of the same\revisedelete{on an Intel Xeon X5650 CPU}.
 Stone and Davis~\cite{Stone:2013aa} investigated a GPU-based version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order-of-magnitude speedup over a serial CPU implementation.
-Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup \revisedelete{on a Tesla C2075 GPU} over a multithreaded VODE solver on a six-core \revisedelete{Intel X5650} CPU.
+Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup\revisedelete{ on a Tesla C2075 GPU} over a multithreaded VODE solver on a six-core\revisedelete{ Intel X5650} CPU.
 Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge--Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
-Yonkee and Sutherland~\cite{Yonkee2016} implemented \revisedelete{GPU-}accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates \revisedelete{on an Intel Xeon E5-2680} \revise{on both the CPU} and \revisedelete{a Tesla K20} GPU, \revise{achieving speedups over serial evaluation between \SIrange{8}{13}{$\times$} on a 16-core CPU and \SIrange{20}{40}{$\times$} on the GPU}.
+Yonkee and Sutherland~\cite{Yonkee2016} implemented \revisedelete{GPU-}accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates\revisedelete{ on an Intel Xeon E5-2680}\revise{ on both the CPU} and\revisedelete{ a Tesla K20} GPU, \revise{achieving speedups over serial evaluation between \SIrange{8}{13}{$\times$} on a 16-core CPU and \SIrange{20}{40}{$\times$} on the GPU}.
 \revise{In addition, \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} speedups} were achieved for the simulation of a partially premixed methanol flame for solving partial differential equations (PDE) on 16 CPU cores and the GPU, respectively.
-Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} \revisedelete{on a Tesla C2075} \revise{on the GPU and CPU}.
-The \revise{GPU-based} implicit Runge--Kutta method performed equivalently to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38} \revisedelete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models \revisedelete{with an integration time step of $10^{-6}$ s}.
+Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa}\revisedelete{ on a Tesla C2075}\revise{ on the GPU and CPU}.
+The \revise{GPU-based} implicit Runge--Kutta method performed equivalently to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38}\revisedelete{ Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models \revisedelete{with an integration time step of $10^{-6}$ s}.
 
@@
 No speedup is achieved for either vectorization type on either CPU---indeed, the OpenMP case is faster on both CPUs\add{, though we stress that Intel OpenCL runtime achieves vectorization using the same code and hardware}.
-The reasons for this lack of vectorization with POCL are quite technical~\delete{ and somewhat outside the scope of this work}; however, \add{personal communication with the developers of POCL revealed that to achieve vectorization, changes are required to both the way POCL prepares LLVM intermediate representation code, as well as improvements to LLVM's loop-vectorizer itself~\cite{pocl_communication}.\footnote{Specifically, improvements such as ensuring LLVM recognizes uniform vectorization loop bounds (even if said bounds \textit{are} uniform in practice), proving vector instructions' ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns.}}
+The reasons for this lack of vectorization with POCL are quite technical\delete{ and somewhat outside the scope of this work}; however, \add{personal communication with the developers of POCL revealed that to achieve vectorization, changes are required to both the way POCL prepares LLVM intermediate representation code, as well as improvements to LLVM's loop-vectorizer itself~\cite{pocl_communication}.\footnote{Specifically, improvements such as ensuring LLVM recognizes uniform vectorization loop bounds (even if said bounds \textit{are} uniform in practice), proving vector instructions' ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns.}}
 As will be discussed in~\cref{S:opencl}, \add{we hope that using explicit vector types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a verification tool.
@@
   \State $\text{species\smallunderscore{}rxn\smallunderscore{}count}\gets \Call{Determine Species Count}{\text{active}}$
-  \While {$\operatorname{min}\left(\text{species\smallunderscore rxn\smallunderscore count}\right) \ge 1$}
+  \While {$\operatorname{min}\left(\text{species\smallunderscore{}rxn\smallunderscore{}count}\right) \ge 1$}
      \State $\text{species\smallunderscore{}rxn\smallunderscore{}count}\gets \Call{Determine Species Count}{\text{active}}$

### 2018-07-26 paper.tex
get rid of old revision marks

@@
 \newtoggle{markup}
-\togglefalse{markup}
-%\toggletrue{markup}
+%\togglefalse{markup}
+\toggletrue{markup}
 
@@
 \begin{abstract}
-\revise{Accurately predicting key combustion phenomena in reactive-flow simulations, e.g., lean blow-out, extinction\slash ignition limits and pollutant formation, necessitates the use of detailed chemical kinetics.}
-\revise{The large size and high levels of numerical stiffness typically present in chemical kinetic models relevant to transportation\slash power-generation applications make the efficient evaluation\slash factorization of the chemical kinetic Jacobian and thermochemical source-terms critical to the performance of reactive-flow codes.}
-\revise{Here we present} \texttt{pyJac}, \revise{a code generation platform} for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized \revise{constant-pressure\slash volume} thermochemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation.
+Accurately predicting key combustion phenomena in reactive-flow simulations, e.g., lean blow-out, extinction\slash ignition limits and pollutant formation, necessitates the use of detailed chemical kinetics.
+The large size and high levels of numerical stiffness typically present in chemical kinetic models relevant to transportation\slash power-generation applications make the efficient evaluation\slash factorization of the chemical kinetic Jacobian and thermochemical source-terms critical to the performance of reactive-flow codes.
+Here we present \texttt{pyJac}, a code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized constant-pressure\slash volume thermochemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation.
 Selected chemical kinetic models covering the range of sizes typically used in reactive-flow simulations were used for demonstration.
-A new formulation of the chemical kinetic governing equations was derived and verified, \revise{resulting in} Jacobian sparsities of \SIrange{28.6}{92.0}{$\percent$} for the tested models.
-Speedups of \SIrange{3.40}{4.08}{$\times$} were found for shallow-vectorized OpenCL source-rate evaluation \revise{compared with} a parallel OpenMP code on an \avx/ central processing unit (CPU), increasing to \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.03}{4.23}{$\times$} for sparse and dense chemical kinetic Jacobian evaluation, respectively.
-%\delete{In addition, the performance of a Nvidia \gpunew/ graphics processing unit (GPU) was compared to a Nvidia \gpuold/ GPU finding speedups of \SIrange{1.40}{1.88}{$\times$}, \SIrange{1.10}{1.59}{$\times$}, and \SIrange{1.36}{3.0}{$\times$} for source-rate, sparse, and dense Jacobian evaluation, respectively.}
-Furthermore, the effect of data-ordering was investigated and \revise{a} storage pattern specifically formulated for vectorized evaluation was proposed; as well, the effect of the \revise{constant pressure\slash volume} assumptions and varying vector widths were studied on source-term evaluation performance.
-Speedups \revise{reached up to} \SI{17.60}{$\times$} and \SI{45.13}{$\times$} for dense and sparse evaluation \revise{on the GPU, and up to} \SI{55.11}{$\times$} and \SI{245.63}{$\times$} on the CPU \revise{over a first-order finite-difference  Jacobian approach}.
-\revise{Further, }dense Jacobian evaluation was up to \SI{19.56}{$\times$} and \SI{2.84}{$\times$} times faster than a previous version of \texttt{pyJac} on a CPU and GPU, respectively.
+A new formulation of the chemical kinetic governing equations was derived and verified, resulting in Jacobian sparsities of \SIrange{28.6}{92.0}{$\percent$} for the tested models.
+Speedups of \SIrange{3.40}{4.08}{$\times$} were found for shallow-vectorized OpenCL source-rate evaluation compared with a parallel OpenMP code on an \avx/ central processing unit (CPU), increasing to \SIrange{6.63}{9.44}{$\times$} and \SIrange{3.03}{4.23}{$\times$} for sparse and dense chemical kinetic Jacobian evaluation, respectively.
+Furthermore, the effect of data-ordering was investigated and a storage pattern specifically formulated for vectorized evaluation was proposed; as well, the effect of the constant pressure\slash volume assumptions and varying vector widths were studied on source-term evaluation performance.
+Speedups reached up to \SI{17.60}{$\times$} and \SI{45.13}{$\times$} for dense and sparse evaluation on the GPU, and up to \SI{55.11}{$\times$} and \SI{245.63}{$\times$} on the CPU over a first-order finite-difference  Jacobian approach.
+Further, dense Jacobian evaluation was up to \SI{19.56}{$\times$} and \SI{2.84}{$\times$} times faster than a previous version of \texttt{pyJac} on a CPU and GPU, respectively.
 Finally, future directions for vectorized chemical kinetic evaluation and sparse linear-algebra techniques were discussed.
@@
 
-\revise{As the combustion and reactive-flows community has recognized the importance of detailed chemical kinetics for predictive reactive-flow simulations~\cite{LU2009192}, chemical kinetic models have grown in size and complexity to describe current and next-generation fuels relevant to transportation and power generation.}
-\revise{For example,} a recent biodiesel model~\cite{WESTBROOK2011742} consists of \textasciitilde\num{3500} chemical species and over \num{17000} reactions.
-Moreover, the cost of evaluating the chemical source-terms scales linearly with the size of the model, while evaluating and factorizing the chemical kinetic Jacobian respectively scale quadratically and cubically with the number of species in the model, \revise{if naively implemented via a finite-difference method~\cite{LU2009192}}.
-These factors often \revise{prohibit using} detailed chemical kinetics in practice; e.g., in a direct numerical simulation \revise{using} a 22-species model, \revise{evaluating} reaction rates consumed around half of the total run time~\cite{Spafford:2010aa}.
-\revise{In addition, most common implicit integration techniques need to evaluate and factorize the Jacobian matrix to deal with stiffness.
-As a result, these operations are bottlenecks when using even moderately sized chemical models in realistic reactive-flow simulations, necessitating other cost-reduction strategies~\cite{LU2009192}.}
-
-A host of techniques have been developed to lessen the computational demand of chemical kinetic calculations while maintaining fidelity, \revise{falling} broadly into \revise{three} categories: removal of unimportant species and reactions~\cite{Lu:2006bb,Pepiot-Desjardins:2008,Hiremath:2010jw,Niemeyer:2010bt,Curtis:2015},
+As the combustion and reactive-flows community has recognized the importance of detailed chemical kinetics for predictive reactive-flow simulations~\cite{LU2009192}, chemical kinetic models have grown in size and complexity to describe current and next-generation fuels relevant to transportation and power generation.
+For example, a recent biodiesel model~\cite{WESTBROOK2011742} consists of \textasciitilde\num{3500} chemical species and over \num{17000} reactions.
+Moreover, the cost of evaluating the chemical source-terms scales linearly with the size of the model, while evaluating and factorizing the chemical kinetic Jacobian respectively scale quadratically and cubically with the number of species in the model, if naively implemented via a finite-difference method~\cite{LU2009192}.
+These factors often prohibit using detailed chemical kinetics in practice; e.g., in a direct numerical simulation using a 22-species model, evaluating reaction rates consumed around half of the total run time~\cite{Spafford:2010aa}.
+In addition, most common implicit integration techniques need to evaluate and factorize the Jacobian matrix to deal with stiffness.
+As a result, these operations are bottlenecks when using even moderately sized chemical models in realistic reactive-flow simulations, necessitating other cost-reduction strategies~\cite{LU2009192}.
+
+A host of techniques have been developed to lessen the computational demand of chemical kinetic calculations while maintaining fidelity, falling broadly into three categories: removal of unimportant species and reactions~\cite{Lu:2006bb,Pepiot-Desjardins:2008,Hiremath:2010jw,Niemeyer:2010bt,Curtis:2015},
 lumping of species with similar thermochemical properties~\cite{Lu:2007,Ahmed:2007fa,Pepiot:2008kq},
 and time-scale methods that reduce numerical stiffness~\cite{Maas:1992ws,Lam:1994ws,Lu:2001ve,Gou:2010}.
-\revise{We refer interested readers} to the recent review by Tur\'anyi and Tomlin~\cite{turanyi2016analysis} for a comprehensive overview.
+We refer interested readers to the recent review by Tur\'anyi and Tomlin~\cite{turanyi2016analysis} for a comprehensive overview.
 
 In addition to the previously mentioned cost reduction methods, effort has gone into improving the integration algorithms and codes that evaluate the chemical kinetics~\cite{Gou:2010,SCHWER2002270,Niemeyer:2016aa,GAO2015287}.
-In particular, a carefully derived analytical formulation of the Jacobian \revise{matrix can greatly increase sparsity}~\cite{SCHWER2002270} and drop the cost of Jacobian evaluation to \revise{linearly depend} on the number of species in the model~\cite{LU2009192}; \revise{sparse-matrix techniques can then reduce the cost of Jacobian factorization}~\cite{superlu99}.
-\revise{In addition, studies have shown that} Single-Instruction, Multiple-Data (SIMD) and the related Single-Instruction, Multiple-Thread (SIMT) processors can accelerate chemical kinetic simulations~\cite{Shi:2012aa,Niemeyer:2014aa,Sewerin20151375,Niemeyer:2016aa,CurtisGPU:2017,stone2018}.
+In particular, a carefully derived analytical formulation of the Jacobian matrix can greatly increase sparsity~\cite{SCHWER2002270} and drop the cost of Jacobian evaluation to linearly depend on the number of species in the model~\cite{LU2009192}; sparse-matrix techniques can then reduce the cost of Jacobian factorization~\cite{superlu99}.
+In addition, studies have shown that Single-Instruction, Multiple-Data (SIMD) and the related Single-Instruction, Multiple-Thread (SIMT) processors can accelerate chemical kinetic simulations~\cite{Shi:2012aa,Niemeyer:2014aa,Sewerin20151375,Niemeyer:2016aa,CurtisGPU:2017,stone2018}.
 
 SIMD and SIMT programming are two important vector-processing paradigms used increasingly in scientific computing.
-\revise{Traditional multicore parallelism is now used to increase} central processing unit (CPU) performance, \revise{as the exponential growth in processing power}---colloquially known as Moore's law---has slowed~\cite{khan2018science}.
-\revise{Recently, SIMD\slash SIMT processors, e.g., in the form of graphics processing units (GPUs), have gained recognition due to their increased floating operation throughput}.
+Traditional multicore parallelism is now used to increase central processing unit (CPU) performance, as the exponential growth in processing power---colloquially known as Moore's law---has slowed~\cite{khan2018science}.
+Recently, SIMD\slash SIMT processors, e.g., in the form of graphics processing units (GPUs), have gained recognition due to their increased floating operation throughput.
 The parallel programming standard OpenCL~\cite{stone2010opencl} has further enabled adoption of vector processing in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems, e.g., CPU, GPU, or Intel's Many Integrated Core (MIC) architecture.
-\revise{Here we will} largely use OpenCL terminology to describe these processing paradigms, as it provides a convenient way to classify otherwise disparate processor types (e.g., CPUs and GPUs).
-However, the concepts discussed herein \revise{broadly apply} to SIMD\slash SIMT processing.
+Here we will largely use OpenCL terminology to describe these processing paradigms, as it provides a convenient way to classify otherwise disparate processor types (e.g., CPUs and GPUs).
+However, the concepts discussed herein broadly apply to SIMD\slash SIMT processing.
 
@@
 
-A typical modern CPU \revise{contains multiple} compute units (i.e., cores), each with specialized vector processing units capable of running SIMD instructions, \revise{as \cref{F:SIMD} depicts}.
-A SIMD instruction uses the vector processor to execute the same floating-point operation (e.g., multiplication, division) \revise{on different data} concurrently.
-\revise{The vector-width is the number of possible concurrent operations, typically around two to four in double precision.}\footnote{OpenCL allows for use of vector-widths different from the actual hardware vector-width via implicit conversion, and may provide some performance benefit as Sec.~\ref{S:results} \revise{discusses.}}
-Specialized hardware accelerators \revise{have also been developed, like Intel's Xeon Phi co-processor (i.e., the MIC architecture)}, that have tens of cores with wide vector-widths (e.g., \numrange{4}{8} double-precision operations).
-\revise{Cutting-edge and forthcoming Intel CPUs also include these wide vector-widths, like the Skylake Xeon and Cannon Lake architectures.}
-
-\revise{Modern GPUs rely on the related computing paradigm of SIMT processing, where a single compute element hosts large
-numbers of threads (a streaming multiprocessor in Nvidia terminology)~\cite{lindholm2008Nvidia}.}
-\revise{\Cref{F:SIMT} depicts a SIMT compute unit, where a group of threads---typically \num{32}, known as a warp on Nvidia GPUs---execute the same SIMT instruction on multiple data concurrently}.
-If some threads must execute a different instruction, they are forced to wait and execute later; \revise{this may occur due to if\slash then branching or predication.}
+A typical modern CPU contains multiple compute units (i.e., cores), each with specialized vector processing units capable of running SIMD instructions, as \cref{F:SIMD} depicts.
+A SIMD instruction uses the vector processor to execute the same floating-point operation (e.g., multiplication, division) on different data concurrently.
+The vector-width is the number of possible concurrent operations, typically around two to four in double precision.\footnote{OpenCL allows for use of vector-widths different from the actual hardware vector-width via implicit conversion, and may provide some performance benefit as Sec.~\ref{S:results} discusses.}
+Specialized hardware accelerators have also been developed, like Intel's Xeon Phi co-processor (i.e., the MIC architecture), that have tens of cores with wide vector-widths (e.g., \numrange{4}{8} double-precision operations).
+Cutting-edge and forthcoming Intel CPUs also include these wide vector-widths, like the Skylake Xeon and Cannon Lake architectures.
+
+Modern GPUs rely on the related computing paradigm of SIMT processing, where a single compute element hosts large
+numbers of threads (a streaming multiprocessor in Nvidia terminology)~\cite{lindholm2008Nvidia}.
+\Cref{F:SIMT} depicts a SIMT compute unit, where a group of threads---typically \num{32}, known as a warp on Nvidia GPUs---execute the same SIMT instruction on multiple data concurrently.
+If some threads must execute a different instruction, they are forced to wait and execute later; this may occur due to if\slash then branching or predication.
 This phenomenon, known as thread-divergence, is a key consideration for SIMT processing and can cause serious performance degradation for complicated algorithms~\cite{CurtisGPU:2017}.
@@
 
-Recognizing the need to accelerate chemical-kinetic Jacobian evaluation and factorization, a number of recent works have been published on \revise{constructing analytical Jacobian matrices}; although as will be discussed at the end of this section, \revise{here we offer} several key improvements over past efforts.
-\add{Schwer et al.~\cite{SCHWER2002270} were among the first to recognize the critical importance of a sparse analytical Jacobian to accelerate chemical kinetic simulations.}
-\revise{Later, Safta et al.~\cite{Safta:2011vn} developed the \texttt{TChem} software package}, which was one of the first developed that provides analytical Jacobian evaluation.
+Recognizing the need to accelerate chemical-kinetic Jacobian evaluation and factorization, a number of recent works have been published on constructing analytical Jacobian matrices; although as will be discussed at the end of this section, here we offer several key improvements over past efforts.
+Schwer et al.~\cite{SCHWER2002270} were among the first to recognize the critical importance of a sparse analytical Jacobian to accelerate chemical kinetic simulations.
+Later, Safta et al.~\cite{Safta:2011vn} developed the \texttt{TChem} software package, which was one of the first developed that provides analytical Jacobian evaluation.
 However, \texttt{TChem} has several limitations, including incompatibility with modern reaction types---i.e., pressure-dependent Arrhenius (or P-Log) and Chebyshev reactions---and its lack of thread-safety to enable parallel execution~\cite{Curtis2017:tchem}.
@@
 Bisetti~\cite{Bisetti:2012jw} developed an isothermal, isobaric analytical Jacobian code-generation utility;
-\revise{this approach significantly increases Jacobian sparsity, although the chosen isothermal assumption is not typical in most combustion simulations.}
-In the same work Bisetti \revise{also} provided a novel way to compute dense matrix-vector multiplications resulting from a change of system variables without \revise{storing} the full Jacobian.
-Perini et al.~\cite{Perini:2012gy} developed an analytical Jacobian code for constant-volume combustion, with additional options to increase sparsity (at the expense of strict correctness) and \revise{tabulate} temperature-dependent properties; \revise{they reported} an \SI{80}{\percent} speedup over a finite-difference-based Jacobian when used in a multidimensional reactive-flow simulation.
+this approach significantly increases Jacobian sparsity, although the chosen isothermal assumption is not typical in most combustion simulations.
+In the same work Bisetti also provided a novel way to compute dense matrix-vector multiplications resulting from a change of system variables without storing the full Jacobian.
+Perini et al.~\cite{Perini:2012gy} developed an analytical Jacobian code for constant-volume combustion, with additional options to increase sparsity (at the expense of strict correctness) and tabulate temperature-dependent properties; they reported an \SI{80}{\percent} speedup over a finite-difference-based Jacobian when used in a multidimensional reactive-flow simulation.
 Gao et al.~\cite{GAO2015287} derived a sparse analytical Jacobian, but did not verify it outside the context of use with an implicit-integration technique.
@@
 
-\revise{Recently, some groups have developed frameworks for constructing analytical Jacobians for evaluation on modern SIMD or SIMT processors.}
-Dijkmans et al.~\cite{Dijkmans:2014bb} developed a GPU-based analytical Jacobian code with optional tabulation of temperature-dependent properties, and \revise{showed} speedups up to \SI{120}{$\times$} for zero-dimensional chemical kinetic integration with large chemical models (\textasciitilde\num{3000} species).
-\add{Bauer et al.~\cite{Bauer:2014} used warp-specialization to improve GPU-vectorization over a standard data-parallel vectorization approach; they achieved speedups of up to \SIrange{2.81}{3.75}{$\times$}, \SIrange{1.91}{2.58}{$\times$}, and \SIrange{1.4}{1.5}{$\times$} for evaluating viscosity, species diffusion, and chemical source terms, respectively}.
-Niemeyer et al.~\cite{Niemeyer:2016aa} created and verified the open-source analytical chemical kinetic Jacobian code-generator, \texttt{pyJac}, which supports parallel execution on CPUs and SIMT execution on GPUs; \revise{\texttt{pyJac} enables} a speedup of \SIrange{3}{7.5}{$\times$} over a finite-difference Jacobian on the CPU.
-
-\revise{Relevant to all of the aforementioned efforts, Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermochemical state vectors and the resulting effect on consistency and errors in conserved properties such as mass and energy.
+Recently, some groups have developed frameworks for constructing analytical Jacobians for evaluation on modern SIMD or SIMT processors.
+Dijkmans et al.~\cite{Dijkmans:2014bb} developed a GPU-based analytical Jacobian code with optional tabulation of temperature-dependent properties, and showed speedups up to \SI{120}{$\times$} for zero-dimensional chemical kinetic integration with large chemical models (\textasciitilde\num{3000} species).
+Bauer et al.~\cite{Bauer:2014} used warp-specialization to improve GPU-vectorization over a standard data-parallel vectorization approach; they achieved speedups of up to \SIrange{2.81}{3.75}{$\times$}, \SIrange{1.91}{2.58}{$\times$}, and \SIrange{1.4}{1.5}{$\times$} for evaluating viscosity, species diffusion, and chemical source terms, respectively.
+Niemeyer et al.~\cite{Niemeyer:2016aa} created and verified the open-source analytical chemical kinetic Jacobian code-generator, \texttt{pyJac}, which supports parallel execution on CPUs and SIMT execution on GPUs; \texttt{pyJac} enables a speedup of \SIrange{3}{7.5}{$\times$} over a finite-difference Jacobian on the CPU.
+
+Relevant to all of the aforementioned efforts, Hansen and Sutherland~\cite{HANSEN2018257} explored the choice of thermochemical state vectors and the resulting effect on consistency and errors in conserved properties such as mass and energy.
 They also characterized how the choice of state vector affects implicit\slash linearly implicit integration algorithms and chemical mode analysis techniques.
@@
 On the other hand, linearly implicit algorithms like Rosenbrock methods and analysis techniques like chemical explosive mode analysis~\cite{lu_yoo_chen_law_2010} need accurate and correct Jacobians.
-}
-
-A number of recent works have investigated \revise{using} high-performance SIMT devices \revise{like GPUs} to accelerate reactive-flow and chemical kinetics simulations.
-Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup\revisedelete{ on a Tesla C1060} compared to a\revise{ CPU-based} serial implementation\revisedelete{on a AMD-Opteron processor}.
-Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order-of-magnitude speedup for homogeneous reactor simulations of large chemical models \revisedelete{on a Tesla 2050 GPU} over a serial\revisedelete{ Intel i7 930} CPU implementation.
-Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a\revisedelete{ Tesla C2075} GPU, achieving a speedup of nearly two orders of magnitude over a sequential \revise{CPU} code\revisedelete{on an Intel Xeon X5650 CPU}.
-Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, where a traditional implicit integrator handled the stiffest computational cells on a CPU and a stabilized-explicit solver\revisedelete{ on a Tesla C2050 GPU} solved the less-stiff cells on a GPU; this hybrid solution technique \revise{performs} \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
-Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code\revisedelete{ on a Tesla C2070} as compared with a sequential\revise{ CPU} version of the same\revisedelete{on an Intel Xeon X5650 CPU}.
+
+A number of recent works have investigated using high-performance SIMT devices like GPUs to accelerate reactive-flow and chemical kinetics simulations.
+Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup compared to a CPU-based serial implementation.
+Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order-of-magnitude speedup for homogeneous reactor simulations of large chemical models  over a serial CPU implementation.
+Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a GPU, achieving a speedup of nearly two orders of magnitude over a sequential CPU code.
+Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, where a traditional implicit integrator handled the stiffest computational cells on a CPU and a stabilized-explicit solver solved the less-stiff cells on a GPU; this hybrid solution technique performs \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
+Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code as compared with a sequential CPU version of the same.
 Stone and Davis~\cite{Stone:2013aa} investigated a GPU-based version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order-of-magnitude speedup over a serial CPU implementation.
-Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup\revisedelete{ on a Tesla C2075 GPU} over a multithreaded VODE solver on a six-core\revisedelete{ Intel X5650} CPU.
+Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup over a multithreaded VODE solver on a six-core CPU.
 Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge--Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
-Yonkee and Sutherland~\cite{Yonkee2016} implemented \revisedelete{GPU-}accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates\revisedelete{ on an Intel Xeon E5-2680}\revise{ on both the CPU} and\revisedelete{ a Tesla K20} GPU, \revise{achieving speedups over serial evaluation between \SIrange{8}{13}{$\times$} on a 16-core CPU and \SIrange{20}{40}{$\times$} on the GPU}.
-\revise{In addition, \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} speedups} were achieved for the simulation of a partially premixed methanol flame for solving partial differential equations (PDE) on 16 CPU cores and the GPU, respectively.
-Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa}\revisedelete{ on a Tesla C2075}\revise{ on the GPU and CPU}.
-The \revise{GPU-based} implicit Runge--Kutta method performed equivalently to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38}\revisedelete{ Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models \revisedelete{with an integration time step of $10^{-6}$ s}.
+Yonkee and Sutherland~\cite{Yonkee2016} implemented accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates on both the CPU and GPU, achieving speedups over serial evaluation between \SIrange{8}{13}{$\times$} on a 16-core CPU and \SIrange{20}{40}{$\times$} on the GPU.
+\In addition, \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} speedups were achieved for the simulation of a partially premixed methanol flame for solving partial differential equations (PDE) on 16 CPU cores and the GPU, respectively.
+Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on the GPU and CPU.
+The GPU-based implicit Runge--Kutta method performed equivalently to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38} CPU cores for two relatively small chemical models with an integration time step of $10^{-6}$ s.
 
@@
 \begin{itemize}
- \item Derive and verify a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity: \revise{\texttt{pyJac v2}};
+ \item Derive and verify a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity: \texttt{pyJac v2};
  \item Enable cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
@@
 \end{itemize}
-To our knowledge, this is the first \add{open-source, verified} effort that vectorizes the evaluation of chemical-kinetic source terms and Jacobian matrices for any chemical model on a wide selection of platforms.
+To our knowledge, this is the first open-source, verified effort that vectorizes the evaluation of chemical-kinetic source terms and Jacobian matrices for any chemical model on a wide selection of platforms.
 
@@
   \end{minipage}
-  \caption{\addnl{Simple data-layout patterns for 2-D arrays}}
+  \caption{Simple data-layout patterns for 2-D arrays}
 \end{figure}
@@
 When storing arrays for a chemical kinetic model, the data-storage layout and vectorization patterns are critical to achieving high-performance code.
-\Cref{F:mem} depicts an example data array with $K$ rows and $N$ columns \revise{where} index ($i$, $j$) \revise{corresponds} to the $i$th row and $j$th column.
+\Cref{F:mem} depicts an example data array with $K$ rows and $N$ columns where index ($i$, $j$) corresponds to the $i$th row and $j$th column.
 For example, the concentration of species $j$ for the $i$th thermochemical state would be stored in $[C]_{i, j}$ with $1 \le i \le N_{\text{state}}$ (the number of thermochemical states considered for evaluation) and $1 \le j \le \ns/$ (the number of species in the model).
@@
 
-The ``C'' (C row-major) format stores the concentrations of all species for a single thermochemical condition $i$  sequentially in memory, i.e., with  $[C]_{1, 1}$ in index 1 (\revise{using one-based index notation}), $[C]_{1, 2}$ in index 2, and so on, as shown in \cref{F:row_major}.
+The ``C'' (C row-major) format stores the concentrations of all species for a single thermochemical condition $i$  sequentially in memory, i.e., with  $[C]_{1, 1}$ in index 1 (using one-based index notation), $[C]_{1, 2}$ in index 2, and so on, as shown in \cref{F:row_major}.
 Conversely, in the ``F'' (Fortran column-major) format the concentrations of a single species $j$ over all thermochemical states lie adjacent in memory, corresponding to storing $[C]_{1, 1}$ in index 1, $[C]_{2, 1}$ in index 2, and so on, as shown in \cref{F:column_major}.
-This ordering \revise{strongly affects} the performance of SIMD\slash SIMT-vectorized algorithms, \revise{as does the device (CPU, GPU, etc.\@) and vectorization pattern in question.}
+This ordering strongly affects the performance of SIMD\slash SIMT-vectorized algorithms, as does the device (CPU, GPU, etc.\@) and vectorization pattern in question.
 
 In a \textit{shallow} SIMD\slash SIMT vectorization (also referred to as ``per-thread'' in previous works using GPUs~\cite{Stone:2013aa}), each SIMD lane or SIMT thread in a compute unit evaluates the source terms or Jacobian for a different thermochemical state.
-If the data is stored in ``F''-order, the SIMD lanes\slash SIMT thread accessing $[C]_{1, j}\ldots[C]_{N_v, j}$ will load sequential locations in memory, \revise{where $[C]_{i, j}$ is the concentration of species $j$ for state $i$ and $N_v$ is the SIMD vector-width or the number of threads in a SIMT warp}.
+If the data is stored in ``F''-order, the SIMD lanes\slash SIMT thread accessing $[C]_{1, j}\ldots[C]_{N_v, j}$ will load sequential locations in memory, where $[C]_{i, j}$ is the concentration of species $j$ for state $i$ and $N_v$ is the SIMD vector-width or the number of threads in a SIMT warp.
 The first $(j+1)$th species concentration, $[C]_{1, j+1}$, will be $N_{\text{state}}$ memory locations away; this increases the likelihood of cache misses on the CPU~\cite{gray2000rules}, but conversely well matches the pattern of coalesced memory access on the GPU~\cite{Nvidia:2018}.
@@
 In a \textit{deep} SIMD\slash SIMT vectorization (also referred to as ``per-block'' in previous GPU works~\cite{Stone:2013aa,CurtisGPU:2017}), a compute unit uses its SIMD lanes\slash SIMT threads cooperatively to evaluate the thermochemical source terms for a single thermochemical state; thus SIMD lanes loading $[C]_{1, j} \ldots [C]_{1, j + N_v}$ will access sequential memory locations if the data is stored in ``C''-order.
-Further, in ``C'' ordering \revise{any two species concentrations within the same thermochemical state lie at most $N_{\text{sp}}$ locations away}, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data locality increases the chances of a cache hit on the CPU, but may lead to uncoalesced memory accesses on the GPU.
-\revisedelete{In addition,} Deep vectorization requires synchronization between SIMD lanes\slash SIMT threads via memory fences\slash barriers, \revise{a potentially expensive operation}.
-\revise{In addition, deep vectorization may result} in SIMD waste or SIMT thread divergence caused by different lanes\slash threads executing different instructions (e.g., resulting from different if\slash then branches).
+Further, in ``C'' ordering any two species concentrations within the same thermochemical state lie at most $N_{\text{sp}}$ locations away, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data locality increases the chances of a cache hit on the CPU, but may lead to uncoalesced memory accesses on the GPU.
+ Deep vectorization requires synchronization between SIMD lanes\slash SIMT threads via memory fences\slash barriers, a potentially expensive operation.
+In addition, deep vectorization may result in SIMD waste or SIMT thread divergence caused by different lanes\slash threads executing different instructions (e.g., resulting from different if\slash then branches).
 Shallow vectorization may also experience SIMD waste or SIMT thread divergence, e.g., in chemical kinetic integration due to varying internal solver time-step sizes~\cite{CurtisGPU:2017}.
-\revise{However, in this work shallow vectorization is largely unaffected by this concern} \add{as the only major code paths that differ between vector lanes are high\slash low-temperature polynomial evaluations and differing pressures for P-Log reactions, which cause far fewer issues compared with differing internal ODE integration time-steps~\cite{CurtisGPU:2017}}.
+However, in this work shallow vectorization is largely unaffected by this concern as the only major code paths that differ between vector lanes are high\slash low-temperature polynomial evaluations and differing pressures for P-Log reactions, which cause far fewer issues compared with differing internal ODE integration time-steps~\cite{CurtisGPU:2017}.
 
@@
 Finally,~\cref{F:vector_data} shows a vectorized data-ordering that improves the caching patterns of a shallow, ``C''-ordered SIMD vectorization on the CPU (\cref{F:row_major_split}) and a deep, ``F''-ordered SIMT vectorization on the GPU (\cref{F:column_major_split}).
-\revise{We accomplish this} by splitting the slower-varying axis of the data array---columns for ``C''-ordering, and rows for ``F''-ordering---into chunks of size $N_v$ (the SIMD vector width or SIMT warp size) and laying these data out contiguously in memory.
+We accomplish this by splitting the slower-varying axis of the data array---columns for ``C''-ordering, and rows for ``F''-ordering---into chunks of size $N_v$ (the SIMD vector width or SIMT warp size) and laying these data out contiguously in memory.
 For example, using the shallow-vectorized ``C''-ordering pictured in~\cref{F:row_major_split}, the concentrations of species $j$ for states $i$ to $i+N_v$ ($[C]_{i, j}, \ldots, [C]_{i + N_v, j}$) lie contiguously in memory and are followed by the concentrations of species $j + 1$ for the same states ($[C]_{i, j + 1}, \ldots, [C]_{i + N_v, j + 1}$).
 This pattern ensures that any SIMD operation occurs on data contiguous in memory, which greatly improves caching and SIMD throughput;
-\revise{it is also similar to OpenCL's native vector data-types, e.g., \texttt{double\num{8}} treats eight contiguous double-precision floating-point numbers as a single vector datum}.
-Conversely, the data-ordering in~\cref{F:column_major_split} \revise{enables} coalesced memory accesses for ``F''-ordered, deep SIMT vectorization on the GPU.
-\revise{We will discuss the} effects of these various data-ordering and vectorization patterns on performance in~\cref{S:results}.
+it is also similar to OpenCL's native vector data-types, e.g., \texttt{double\num{8}} treats eight contiguous double-precision floating-point numbers as a single vector datum.
+Conversely, the data-ordering in~\cref{F:column_major_split} enables coalesced memory accesses for ``F''-ordered, deep SIMT vectorization on the GPU.
+We will discuss the effects of these various data-ordering and vectorization patterns on performance in~\cref{S:results}.
 
@@
 This new version of \texttt{pyJac} is capable of evaluating the thermochemical source-terms for using the constant-pressure (\conp/) or constant-volume (\conv/) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
-In this section, we will outline a brief summary of the system evaluated by \texttt{pyJac}; \revise{the supplemental material contains the complete---and lengthy---derivations}.
+In this section, we will outline a brief summary of the system evaluated by \texttt{pyJac}; the supplemental material contains the complete---and lengthy---derivations.
 
@@
 
-This state vector---inspired by Schwer et al.~\cite{SCHWER2002270}---has a number of \revise{beneficial features}.
+This state vector---inspired by Schwer et al.~\cite{SCHWER2002270}---has a number of beneficial features.
 First, the state vector results in highly sparse chemical kinetic Jacobians, as will be detailed in~\cref{S:sparsity}.
-Second, \revise{this formulation explicitly conserves mass, because the number of moles and rate of change of the final species are calculated from the ideal gas law and conservation of mass, respectively}; \revise{see the supplemental material for the full details of the governing equations}.
-\revise{The system is} not over-constrained~\cite{HANSEN2018257} and does not require use of a more-complicated differential algebraic equation solver (as compared to an ODE integrator) for integration.
-Finally, the chemical kinetic Jacobian \revise{for this formulation} changes relatively little between the \conp/ and \conv/ forms, making maintaining the codebase much simpler.
-Although most current combustion codes do not use species moles as a state variable, conversion to\slash from the more-common mass\slash mole fractions and moles is straightforward, and \revise{the choice of variables no longer matters once inside the integration} of an chemical kinetic initial-value problem (IVP).
+Second, this formulation explicitly conserves mass, because the number of moles and rate of change of the final species are calculated from the ideal gas law and conservation of mass, respectively; see the supplemental material for the full details of the governing equations.
+The system is not over-constrained~\cite{HANSEN2018257} and does not require use of a more-complicated differential algebraic equation solver (as compared to an ODE integrator) for integration.
+Finally, the chemical kinetic Jacobian for this formulation changes relatively little between the \conp/ and \conv/ forms, making maintaining the codebase much simpler.
+Although most current combustion codes do not use species moles as a state variable, conversion to\slash from the more-common mass\slash mole fractions and moles is straightforward, and the choice of variables no longer matters once inside the integration of an chemical kinetic initial-value problem (IVP).
 
@@
 \end{equation}
-where $\mathcal{R}$ is the ideal-gas constant in molar units, \revise{we find} the volume and pressure source terms (where $W_k$ and $W_{\ns/}$ are the molecular weights of species $k$ and $\ns/$, respectively):
+where $\mathcal{R}$ is the ideal-gas constant in molar units, we find the volume and pressure source terms (where $W_k$ and $W_{\ns/}$ are the molecular weights of species $k$ and $\ns/$, respectively):
 \loadeq{dparam}
 
-\revise{\texttt{pyJac} arranges the computed Jacobian entries} such that entry ($i$, $j$) corresponds to the partial derivative of the $i$th source-term in~\cref{e:source_terms} by the $j$th state variable in~\cref{e:state}:
+\texttt{pyJac} arranges the computed Jacobian entries such that entry ($i$, $j$) corresponds to the partial derivative of the $i$th source-term in~\cref{e:source_terms} by the $j$th state variable in~\cref{e:state}:
 \loadeq{jacgeneral}
 
-\delete{The derivative of the temperature source-term with respect to itself is:\ldots}
-\add{The supplemental material for this article contains the complete derivation of the Jacobian used by \texttt{pyJac}, for interested readers.}
+The supplemental material for this article contains the complete derivation of the Jacobian used by \texttt{pyJac}, for interested readers.
 
@@
 
-\revise{The new version of \texttt{pyJac} uses the Python package \texttt{loo.py}~\cite{kloeckner_loopy_2014} for code generation}, which translates pseudo-code and data to OpenCL\slash C code.
+The new version of \texttt{pyJac} uses the Python package \texttt{loo.py}~\cite{kloeckner_loopy_2014} for code generation, which translates pseudo-code and data to OpenCL\slash C code.
 As the name implies, \texttt{loo.py} generates code using for loops; this differs from the previous version of \texttt{pyJac}~\cite{pyjac16} that generates static code---i.e., fully unrolled loops, with thermodynamic\slash reaction parameters written directly in code rather than stored in arrays.
 In our previous work~\cite{Niemeyer:2016aa}, this static code generation caused some issues with large file sizes, long compilation times, and even occasionally broke the \texttt{gcc} and \texttt{nvcc} compilers (the latter issue necessitated splitting the Jacobian\slash source-term evaluations into separate files).
-\revise{We will discuss the implications of this change} in~\cref{S:jacobian_results}, where the performance of the new version of~\texttt{pyJac} will be compared with the previous version.
+We will discuss the implications of this change in~\cref{S:jacobian_results}, where the performance of the new version of~\texttt{pyJac} will be compared with the previous version.
 
@@
 The source terms and sub-components thereof (e.g., rates of progress, pressure-modification terms) are directly compared with Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{hogan2014fast,adept-v11} provides reference values for Jacobian sub-components.
-\revise{We use the} Portable OpenCL (POCL) implementation~\cite{poclIJPP} and OpenMP~\cite{dagum1998openmp} to perform OpenCL and C unit testing, respectively, on the continuous-integration framework Travis CI~\cite{travis:2018}.
-\revise{We will discuss verification} of the complete (as opposed to the sub-component testing discussed here) generated source-terms and Jacobian codes in detail in~\cref{s:verification}.
+We use the Portable OpenCL (POCL) implementation~\cite{poclIJPP} and OpenMP~\cite{dagum1998openmp} to perform OpenCL and C unit testing, respectively, on the continuous-integration framework Travis CI~\cite{travis:2018}.
+We will discuss verification of the complete (as opposed to the sub-component testing discussed here) generated source-terms and Jacobian codes in detail in~\cref{s:verification}.
 
@@
 
-\revise{We ran the} performance and verification studies for this work on a variety of CPU and GPU platforms.
+We ran the performance and verification studies for this work on a variety of CPU and GPU platforms.
 \Cref{t:cpus} shows the number of cores, vector instruction set, and model of the CPUs used in this work; each CPU had both \texttt{v16.1.1} of the Intel OpenCL runtime~\cite{intelopencl:2018} and \texttt{v1.0} of the POCL~\cite{poclIJPP} runtime installed, both enabling OpenCL \texttt{v1.2} execution.
 Additionally, \texttt{v5.0} of the LLVM\slash\texttt{clang}~\cite{Lattner:2004:LCF:977395.977673} compiler chain was installed on all machines to enable use of POCL.
-\Cref{t:gpus} lists the model, number of CUDA cores, and Nvidia driver of each GPU \revise{we used}.
+\Cref{t:gpus} lists the model, number of CUDA cores, and Nvidia driver of each GPU we used.
 Nvidia's OpenCL runtime is bundled with the Nvidia driver~\cite{Nvidia:2018}, hence the driver version is used to specify the OpenCL runtime version.
@@
 \Cref{t:platforms} lists the platforms and vectorization\slash execution patterns that they are capable of running.
-The Intel and Nvidia OpenCL runtimes lack implementations of atomic operations on double-precision variables; \revise{\texttt{pyJac} currently needs these} to run deep-vectorized code.
+The Intel and Nvidia OpenCL runtimes lack implementations of atomic operations on double-precision variables; \texttt{pyJac} currently needs these to run deep-vectorized code.
 On the other hand, POCL is an open-source OpenCL runtime that works on all CPU types tested here, and does implement these atomic operations.
 However, POCL's implicit vectorization module---which uses the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to translate OpenCL code to vectorized machine code---typically fails to achieve much, if any, speedup.
-\revise{Thus POCL is useful for verification but not necessarily for performance studies; it is noted that while POCL is currently used by \texttt{pyJac} for unit-testing purposes, it is not required to use \texttt{pyJac}.}
-\revise{We will expand upon this discussion} in~\cref{S:future} to highlight future directions.
+Thus POCL is useful for verification but not necessarily for performance studies; it is noted that while POCL is currently used by \texttt{pyJac} for unit-testing purposes, it is not required to use \texttt{pyJac}.
+We will expand upon this discussion in~\cref{S:future} to highlight future directions.
 
@@
 Finally, \cref{t:models} displays the chemical kinetic models used in this work, as well as number of partially stirred reaction conditions (PaSR) used in the condition database for each.
-\revise{Our previous works describe the creation of the PaSR databases in detail works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.}
+Our previous works describe the creation of the PaSR databases in detail works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
 
@@
 \label{s:verification}
-\revise{We verified the reaction rates of progress (ROP), species production rates, and temperature rates in this study by comparing with values calculated using Cantera~\cite{Cantera}.}
+We verified the reaction rates of progress (ROP), species production rates, and temperature rates in this study by comparing with values calculated using Cantera~\cite{Cantera}.
 However, special care must be taken due to floating-point arithmetic issues.
@@
 \end{tabular}
-\caption{Summary of errors in rates of progress, species, temperature, and thermodynamic state-parameter rate \revise{compared with Cantera}.
+\caption{Summary of errors in rates of progress, species, temperature, and thermodynamic state-parameter rate compared with Cantera.
 Error statistics are based on the infinity-norm of the relative error detailed in~\cref{e:rel_err} for each quantity.
@@
 
-Table~\ref{T:source_error} compares \revise{\texttt{pyJac} v2's} source-term evaluations with Cantera's~\cite{Cantera} using the data set of PaSR conditions (\cref{t:models}).
-\revise{The forward and reverse ROPs agree closely} for all models, though the error norm is \textasciitilde\numrange{3}{4} orders of magnitude larger for the isopentanol model.
-This \revise{discrepancy} results from differences in evaluation of P-Log reactions between \texttt{pyJac} and Cantera: \texttt{pyJac} computes the logarithm of the upper and lower reaction Arrhenius rates analytically (see supplemental material) while Cantera evaluates this term numerically.
+Table~\ref{T:source_error} compares \texttt{pyJac} v2's source-term evaluations with Cantera's~\cite{Cantera} using the data set of PaSR conditions (\cref{t:models}).
+The forward and reverse ROPs agree closely for all models, though the error norm is \textasciitilde\numrange{3}{4} orders of magnitude larger for the isopentanol model.
+This discrepancy results from differences in evaluation of P-Log reactions between \texttt{pyJac} and Cantera: \texttt{pyJac} computes the logarithm of the upper and lower reaction Arrhenius rates analytically (see supplemental material) while Cantera evaluates this term numerically.
 If we neglect the errors from P-Log reactions in~\cref{e:rel_err}, the errors for the forward and reverse ROPs fall to \num{5.44e-08} and \num{1.59e-07}, respectively.
 This discrepancy does not imply any actual error in either \texttt{pyJac} or Cantera---in fact, the error still lies well within the proscribed tolerances in~\cref{e:rel_err}---but merely emphasizes
-\revise{how even small code changes can affect the accumulation of floating-point errors.}
-
-\revise{The error in the net ROP further underscores this point: it is}
+how even small code changes can affect the accumulation of floating-point errors.
+
+The error in the net ROP further underscores this point: it is
 \textasciitilde\numrange{3}{9} orders of magnitude (or \numrange{7}{9} orders of magnitude when including P-Log reaction contributions) larger than the error in forward or reverse ROP.
-\revise{\Cref{T:source_error} shows that the magnitudes of $E_{\varepsilon}$ and $E_R$ agree} in all cases, indicating that the accumulation of floating-point error from the forward and reverse ROPs causes this large increase in error as previously discussed.
-\revise{The magnitudes of the errors in molar species production rate and net ROP agree, but
-thermodynamic properties amplify the error in net species production rates and lead to high discrepancies in temperature and state-parameter rates.}
-Again, these discrepancies in net ROP will not necessarily cause errors when integrating the chemical kinetics---either in \texttt{pyJac} or Cantera---as this loss of accuracy only occurs when the forward and reverse ROPs are nearly equal \revise{(i.e., near equilibrium)}.
+\Cref{T:source_error} shows that the magnitudes of $E_{\varepsilon}$ and $E_R$ agree in all cases, indicating that the accumulation of floating-point error from the forward and reverse ROPs causes this large increase in error as previously discussed.
+The magnitudes of the errors in molar species production rate and net ROP agree, but
+thermodynamic properties amplify the error in net species production rates and lead to high discrepancies in temperature and state-parameter rates.
+Again, these discrepancies in net ROP will not necessarily cause errors when integrating the chemical kinetics---either in \texttt{pyJac} or Cantera---as this loss of accuracy only occurs when the forward and reverse ROPs are nearly equal (i.e., near equilibrium).
 
@@
 We previously explained this choice fully~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with minimal additional programming effort.
-\revise{(In contrast, it is challenging to obtain robust, accurate Jacobians using finite differences.)}
+(In contrast, it is challenging to obtain robust, accurate Jacobians using finite differences.)
 The discrepancy between the analytical and automatic-differentiation Jacobians for thermochemical state $k$, denoted by $\mathcal{J}_k$ and $\hat{\mathcal{J}}_k$ respectively, is determined by the relative error Frobenius norm over all Jacobian indices $i, j$:
@@
 \end{equation}
-The error statistics reported in this section are then based only on matrix elements where $\mathcal{J}_{ijk} \ge \frac{\mathcal{T}}{\mathcal{C}}$, where $\mathcal{C}$ is \revise{a tunable} threshold parameter; this filtered form of~\cref{e:jac_error_base} is denoted $E_{\mathcal{C},k}$.
+The error statistics reported in this section are then based only on matrix elements where $\mathcal{J}_{ijk} \ge \frac{\mathcal{T}}{\mathcal{C}}$, where $\mathcal{C}$ is a tunable threshold parameter; this filtered form of~\cref{e:jac_error_base} is denoted $E_{\mathcal{C},k}$.
 Finally, the Frobenius norm is calculated over all the states $k$ in the PaSR thermochemical condition database:
@@
 
-This error norm is quite different from the relative error Frobenius norm suggested by Anderson et al.~\cite{Anderson:1999aa} for quantifying the error of matrices in LAPACK, \revise{e.g., over Jacobian indices $i, j$}:
+This error norm is quite different from the relative error Frobenius norm suggested by Anderson et al.~\cite{Anderson:1999aa} for quantifying the error of matrices in LAPACK, e.g., over Jacobian indices $i, j$:
 \begin{align}
@@
 This relaxed norm is smaller by \numrange{2}{5} orders of magnitude for all models---except GRI-Mech 3.0, where the stringent case already has small error as previously discussed---which indicates that accuracy is higher when the smaller Jacobian entries are excluded.
-\revise{This result suggests that floating-point error accumulation controls the stringent filtered error norm.}
+This result suggests that floating-point error accumulation controls the stringent filtered error norm.
 
@@
 Further, the LAPACK error norm does not correlate well with the stringent filtered error norm, e.g., USC-Mech II has the smallest LAPACK error norm (\num{1.137e-14}) but the second-largest stringent filtered error norm (\num{1.119e-2}).
-\revise{Conversely, the model with the largest LAPACK error norm, \ce{iC5H11OH} has the second smallest stringent filtered error norm: $E_{\mathcal{L}} = \num{1.227e-10}$ and $E_{\mathcal{C} = 10^{20}} = \num{1.363e-3}$, again suggesting that floating-point error accumulation influences the stringent error norm.}
-These findings, along with the individual unit-testing of all chemical source-terms and Jacobian sub-components described in~\cref{s:unittest}, gives high confidence in the correctness of \revise{\texttt{pyJac} v2}.
+Conversely, the model with the largest LAPACK error norm, \ce{iC5H11OH} has the second smallest stringent filtered error norm: $E_{\mathcal{L}} = \num{1.227e-10}$ and $E_{\mathcal{C} = 10^{20}} = \num{1.363e-3}$, again suggesting that floating-point error accumulation influences the stringent error norm.
+These findings, along with the individual unit-testing of all chemical source-terms and Jacobian sub-components described in~\cref{s:unittest}, gives high confidence in the correctness of \texttt{pyJac} v2.
 
@@
 However, the explicit-mass conservation formulation of \texttt{pyJac} can introduce additional non-zero entries in two ways.
-First, if the last species in the model (i.e., the bath gas) participates directly in any reaction, the derivative of its forward or reverse rate of progress \revise{is non-zero with respect to all other species in the model, regardless of whether the other species participate in that reaction or not.}
-Similarly, if the last species has a third-body efficiency not equal to the default (one), \revise{this will again create nonzero derivatives for the pressure-modification term with respect to all other species (see the supplemental material).}
+First, if the last species in the model (i.e., the bath gas) participates directly in any reaction, the derivative of its forward or reverse rate of progress is non-zero with respect to all other species in the model, regardless of whether the other species participate in that reaction or not.
+Similarly, if the last species has a third-body efficiency not equal to the default (one), this will again create nonzero derivatives for the pressure-modification term with respect to all other species (see the supplemental material).
 Either case will result in a fully dense Jacobian row for all species with a non-zero net stoichiometric coefficient in such a reaction.
@@
 The rationale behind this choice is that many common implicit integration techniques (e.g., CVODE~\cite{Hindmarsh:2005}) used to solve chemical-kinetic initial-value problems are formulated around the assumption that the supplied Jacobian is approximate; this allows the Jacobian and its LU factorization to be reused for multiple internal integration time steps, accelerating the solution process.
-\revise{Such solvers do not need the} exact form of the Jacobian and thus the so-called ``approximate'' form is preferable.
-\add{Though this might be used as a crude form of preconditioning for such solvers, the primary purpose is merely to increase Jacobian sparsity; McNenly et al.~\cite{MCNENLY2015581} more thoroughly investigated preconditioners.}
+Such solvers do not need the exact form of the Jacobian and thus the so-called ``approximate'' form is preferable.
+Though this might be used as a crude form of preconditioning for such solvers, the primary purpose is merely to increase Jacobian sparsity; McNenly et al.~\cite{MCNENLY2015581} more thoroughly investigated preconditioners.
 Hence, in this section we will detail the sparsity of both forms of the Jacobian for the chemical models tested.
@@
 As will be explored in~\cref{S:jacobian_results}, this slows down sparse Jacobian evaluation, and might be improved by a static mapping of the full Jacobian indices to the sparse index (or some null value if the entry is empty).
-\revise{However, this would require increased constant-data usage, a limitation for OpenCL.}
+However, this would require increased constant-data usage, a limitation for OpenCL.
 Additionally, this might be an excellent usage of OpenCL's Image memory type (similar to texture memory in CUDA terminology).
@@
 
-\Cref{F:intel_source} details the extent of this speedup; the speedup displayed is calculated per-machine, e.g., the \avx/ shallow-vectorized code speedup is relative to OpenMP on the same CPU\revisedelete{ such that the speedup of the shallow-vectorized OpenCL code can be readily assessed}.
+\Cref{F:intel_source} details the extent of this speedup; the speedup displayed is calculated per-machine, e.g., the \avx/ shallow-vectorized code speedup is relative to OpenMP on the same CPU.
 On both machines, the unvectorized OpenCL code is faster than the baseline parallel OpenMP code, by \SIrange{1.30}{1.51}{$\times$} on the \avx/ CPU and \SIrange{1.25}{1.27}{$\times$} on the \sse/ machine.
@@
 In contrast,~\cref{F:pocl_source} shows the mean run time per condition of deep, shallow, and unvectorized OpenCL codes using the POCL runtime, as compared with OpenMP parallelization.
-No speedup is achieved for either vectorization type on either CPU---indeed, the OpenMP case is faster on both CPUs\add{, though we stress that Intel OpenCL runtime achieves vectorization using the same code and hardware}.
-The reasons for this lack of vectorization with POCL are quite technical\delete{ and somewhat outside the scope of this work}; however, \add{personal communication with the developers of POCL revealed that to achieve vectorization, changes are required to both the way POCL prepares LLVM intermediate representation code, as well as improvements to LLVM's loop-vectorizer itself~\cite{pocl_communication}.\footnote{Specifically, improvements such as ensuring LLVM recognizes uniform vectorization loop bounds (even if said bounds \textit{are} uniform in practice), proving vector instructions' ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns.}}
-As will be discussed in~\cref{S:opencl}, \add{we hope that using explicit vector types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue}, but for the moment POCL is still quite useful as a verification tool.
+No speedup is achieved for either vectorization type on either CPU---indeed, the OpenMP case is faster on both CPUs, though we stress that Intel OpenCL runtime achieves vectorization using the same code and hardware.
+The reasons for this lack of vectorization with POCL are quite technical; however, personal communication with the developers of POCL revealed that to achieve vectorization, changes are required to both the way POCL prepares LLVM intermediate representation code, as well as improvements to LLVM's loop-vectorizer itself~\cite{pocl_communication}.\footnote{Specifically, improvements such as ensuring LLVM recognizes uniform vectorization loop bounds (even if said bounds \textit{are} uniform in practice), proving vector instructions' ability to handle all edge cases identically to the corresponding scalar instruction, handing of branches and conditionals (potentially within POCL instructions themselves), and handling of memory access\slash vector-element extraction patterns.}
+As will be discussed in~\cref{S:opencl}, we hope that using explicit vector types (to lessen demands on the LLVM-vectorization module) in combination with some of these changes might solve this issue, but for the moment POCL is still quite useful as a verification tool.
 
 \Cref{F:gpu_source} shows the performance of evaluating source terms on the GPUs listed in~\cref{t:gpus}.
-\revise{\Cref{F:gpu_source_scaling} investigates how the number of initial conditions evaluated affects the mean run time per condition on the \gpunew/ GPU};
+\Cref{F:gpu_source_scaling} investigates how the number of initial conditions evaluated affects the mean run time per condition on the \gpunew/ GPU;
 the run time decreases until around \textasciitilde\num{e4} conditions for all chemical models, at which point the GPU becomes saturated and performance levels off.
@@
 \Cref{F:source_vector_width} shows how changing vector width affects source-term evaluation performance on the \avx/ CPU.
-The vector width of 8 \revise{performs the fastest (out of those tested)}, while the larger vector width of 16 is slightly slower due to increased register pressure~\cite{intel_vectypes}.
+The vector width of 8 performs the fastest (out of those tested), while the larger vector width of 16 is slightly slower due to increased register pressure~\cite{intel_vectypes}.
 It is unclear why the vector width of 4 results in no speedup at all (in fact, it is the slowest case).
@@
 In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for both Intel OpenCL and OpenMP, likely resulting from both its relatively small size and few falloff\slash chemically activated reaction (in particular, the additional expensive logarithm and exponential evaluations that accompany them).
-\add{As demonstrated in~\cref{S:SIMD_scaling}, the amount of computational work required per thermochemical state plays a critical role in fully utilizing SIMD instructions\slash multiple threads}.
+As demonstrated in~\cref{S:SIMD_scaling}, the amount of computational work required per thermochemical state plays a critical role in fully utilizing SIMD instructions\slash multiple threads.
 Additionally, OpenMP tends to scale better than the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.9} and \textasciitilde\num{0.75} for four and eight CPU cores, respectively, compared to just \numrange{0.66}{0.72} and \numrange{0.44}{0.48} for OpenCL.
@@
 Therefore, a more fair comparison of parallel scaling efficiencies is to compare OpenCL running on \num{4} cores with OpenMP on \num{16}; the OpenMP code's parallel efficiency drops to \textasciitilde\num{0.64} for \num{16} cores, similar to OpenCL's parallel scaling efficiency of \numrange{0.66}{0.72} at \num{4} cores.
-Indeed, as will be seen in~\cref{S:jacobian_results}, \revise{sparse Jacobian evaluation---the most computationally intensive task in this work---exhibits similar strong-scaling efficiency on Intel OpenCL and OpenMP.}
+Indeed, as will be seen in~\cref{S:jacobian_results}, sparse Jacobian evaluation---the most computationally intensive task in this work---exhibits similar strong-scaling efficiency on Intel OpenCL and OpenMP.
 
@@
 As with the CPU, the static code of \texttt{pyJac-v1} is slightly faster for the \ce{H2}\slash\ce{CO} model, but \texttt{pyJac-v2} outperforms \texttt{pyJac-v1} by \SIrange{1.25}{2.84}{$\times$} for the other models.
-\revise{Performance likely drops for the isopentanol model} due to the limited number of conditions in dense evaluation for \texttt{pyJac-v2}, as noted earlier.
+Performance likely drops for the isopentanol model due to the limited number of conditions in dense evaluation for \texttt{pyJac-v2}, as noted earlier.
 
@@
 
-\revisedelete{It is important to note that while }
-\revise{While OpenCL provides a simple interface to enable cross-platform execution, and significant speedups were achieved via shallow-vectorized OpenCL code in this work, there are some serious potential pitfalls in its use.
+While OpenCL provides a simple interface to enable cross-platform execution, and significant speedups were achieved via shallow-vectorized OpenCL code in this work, there are some serious potential pitfalls in its use.
 The closed-source OpenCL runtimes tested in this work (Intel and Nvidia) contain several bugs that result in compilation failures, simply incorrect vectorized machine code, or even segmentation faults.
 Further, these runtimes (in our experience) tend to be less responsive to fixing said bugs, with relatively infrequent new releases (or in Nvidia's case, lack of changelogs\slash public records of bug-fixes).
-On the other hand, the open-source OpenCL runtime used in this work, POCL, has far fewer implementation bugs, and when issues arise the community is very responsive to bug reports and user outreach; however, POCL fails to achieve vectorization as noted in~\cref{S:source_results}.}
-
-\revise{To demonstrate the type of issue discussed, we created a minimum working example~\cite{Nvidia_mwe} that shows a failure of Nvidia's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla \gpunew/ GPU; simply changing to another runtime (e.g., Intel) produces the correct result---with no code changes or recompilation.
+On the other hand, the open-source OpenCL runtime used in this work, POCL, has far fewer implementation bugs, and when issues arise the community is very responsive to bug reports and user outreach; however, POCL fails to achieve vectorization as noted in~\cref{S:source_results}.
+
+To demonstrate the type of issue discussed, we created a minimum working example~\cite{Nvidia_mwe} that shows a failure of Nvidia's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla \gpunew/ GPU; simply changing to another runtime (e.g., Intel) produces the correct result---with no code changes or recompilation.
 This provides a particularly vexing problem for the programmer, since little can often be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
@@
 Indeed the vectorization process attempted by most OpenCL runtimes (and thus, reasons for incorrect\slash unvectorized code output) is obscured from the user, making reasoning about errors or performance trends quite difficult.
-Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change requires building of the wrapper that initializes\slash transfers memory and calls the source-rate\slash Jacobian kernel.}
+Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change requires building of the wrapper that initializes\slash transfers memory and calls the source-rate\slash Jacobian kernel.
 
@@
 In this work, we developed automatically generated OpenCL codes for SIMD and SIMT-vectorized evaluations of thermochemical source-term and sparse\slash dense chemical kinetic Jacobian matrices.
-\revise{The main contributions of this work are:}
+The main contributions of this work are:
 \begin{itemize}
- \item \revise{Deriving and verifying a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity;}
- \item \revise{Extending \texttt{pyJac} to enable vectorized execution on the CPU, GPU, and other accelerators; and}
- \item \revise{Achieving significant speedups over a strictly parallel Jacobian and source-term evaluation on SIMD-enabled processors (e.g., CPUs).}
+ \item Deriving and verifying a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity;
+ \item Extending \texttt{pyJac} to enable vectorized execution on the CPU, GPU, and other accelerators; and
+ \item Achieving significant speedups over a strictly parallel Jacobian and source-term evaluation on SIMD-enabled processors (e.g., CPUs).
 \end{itemize}
 
-\revise{The new molar-based formulation for \texttt{pyJac} resulted in highly sparse chemical kinetic Jacobians,
+The new molar-based formulation for \texttt{pyJac} resulted in highly sparse chemical kinetic Jacobians,
 and allows selection of either the constant-volume or constant-pressure approximation.
 In addition, sparsity can be increased further by eliminating components associated with the bath gas, as discussed in \cref{S:sparsity};
-this approximation is not a key feature of this work, and future efforts to develop more sophisticated Jacobian approximations based on the \texttt{pyJac} software would be a worthwhile endeavor.}
+this approximation is not a key feature of this work, and future efforts to develop more sophisticated Jacobian approximations based on the \texttt{pyJac} software would be a worthwhile endeavor.
 
 We also demonstrated source-term and Jacobian evaluation for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
-\revise{In addition to parallel OpenMP evaluation on the CPU, the new version of \texttt{pyJac} enables the shallow-vectorized evaluation of the chemical-kinetic source terms and analytical Jacobian on both the CPU and GPU via OpenCL.}
+In addition to parallel OpenMP evaluation on the CPU, the new version of \texttt{pyJac} enables the shallow-vectorized evaluation of the chemical-kinetic source terms and analytical Jacobian on both the CPU and GPU via OpenCL.
 Deep vectorization is possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, but it yields no performance benefit as POCL did not achieve vectorizations for any execution pattern studied.
@@
 
-\revise{We demonstrated significant speedups in shallow SIMD-vectorized execution over a parallel OpenMP code for evaluating the chemical-kinetic source terms and sparse\slash dense Jacobian: up to \SI{4.09}{$\times$}, \SI{9.44}{$\times$}, and \SI{4.23}{$\times$}, respectively, on an \avx/-capable CPU.
+We demonstrated significant speedups in shallow SIMD-vectorized execution over a parallel OpenMP code for evaluating the chemical-kinetic source terms and sparse\slash dense Jacobian: up to \SI{4.09}{$\times$}, \SI{9.44}{$\times$}, and \SI{4.23}{$\times$}, respectively, on an \avx/-capable CPU.
 Sparse Jacobians evaluate more slowly than dense Jacobians on all CPU\slash GPU platforms due to indirect lookup requirements in array indexing, but this adversely affects the shallow-vectorized OpenCL code less than OpenMP.
@@
 The OpenMP version evaluates moderately slower on the CPU for the smallest chemical model (e.g., \SI{1.79}{$\times$} on the \avx/ CPU), but significantly faster for the larger models---up to \SI{10.19}{$\times$}.
-The shallow-vectorized OpenCL code runs faster than the previous version over all chemical models, reaching speedups of \SI{19.56}{$\times$}.}
+The shallow-vectorized OpenCL code runs faster than the previous version over all chemical models, reaching speedups of \SI{19.56}{$\times$}.
 
@@
 It is also possible, particularly for the Intel OpenCL\slash POCL runtimes, that better performance\slash stability might be achieved using so-called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}.
-\add{Specifically, this change could enable vectorization on the POCL runtime} and might also enable deep vectorization on the Intel OpenCL runtime.
+Specifically, this change could enable vectorization on the POCL runtime and might also enable deep vectorization on the Intel OpenCL runtime.
 Using OpenCL Image\slash CUDA Texture memory to accelerate the indirect lookup for sparse matrix evaluation should be investigated as well.
-Finally, future extensions of this work will include extending to additional target languages (e.g., vectorized OpenMP, CUDA, ISPC) to improve ease of use and reliability, improving the existing OpenCL targets (e.g., to enable meaningful deep-vectorized evaluation), \add{and implementing reaction sorting~\cite{Sewerin20151375} to improve SIMD efficiency (\cref{S:SIMD_scaling}).}
+Finally, future extensions of this work will include extending to additional target languages (e.g., vectorized OpenMP, CUDA, ISPC) to improve ease of use and reliability, improving the existing OpenCL targets (e.g., to enable meaningful deep-vectorized evaluation), and implementing reaction sorting~\cite{Sewerin20151375} to improve SIMD efficiency (\cref{S:SIMD_scaling}).
 
@@
 
-\add{The results for this article were obtained using \texttt{pyJac v2.0.0b0}~\cite{pyjac2}.
+The results for this article were obtained using \texttt{pyJac v2.0.0b0}~\cite{pyjac2}.
 The most recent version of \texttt{pyJac} can be found at its GitHub repository:
@@
 All figures, and the data and plotting scripts necessary to reproduce them,
-are available openly under the CC-BY license~\cite{data}.}
+are available openly under the CC-BY license~\cite{data}.
 
@@
 
-\add{This simple example demonstrates how the SIMD efficiency of shallow-vectorized OpenCL source-term evaluation depends on the size of the chemical model in question, i.e., the amount of computational work per source-term evaluation.
+This simple example demonstrates how the SIMD efficiency of shallow-vectorized OpenCL source-term evaluation depends on the size of the chemical model in question, i.e., the amount of computational work per source-term evaluation.
 The base chemical model for this example was the isopentanol model~\cite{Sarathy:2013jr} used throughout this article, and the same thermochemical state database described in~\cref{s:test_platforms} was used for source-term evaluation.
-As in~\cref{S:results} all reported results are based on \num{10} individual runs and in this example all cases were run on the \avx/ machine using the Intel OpenCL runtime.}
+As in~\cref{S:results} all reported results are based on \num{10} individual runs and in this example all cases were run on the \avx/ machine using the Intel OpenCL runtime.
 
@@
 
-\add{First, the reactions in the isopentanol model were converted to simple reversible Arrhenius reactions by either simply dropping third-body efficiency calculations (third-body enhanced reactions), using the high-pressure-limit coefficients (falloff\slash chemically-activated and P-Log reactions) or fitting Arrhenius parameters to the calculated rate constant at a fixed pressure (Chebyshev reactions).
+First, the reactions in the isopentanol model were converted to simple reversible Arrhenius reactions by either simply dropping third-body efficiency calculations (third-body enhanced reactions), using the high-pressure-limit coefficients (falloff\slash chemically-activated and P-Log reactions) or fitting Arrhenius parameters to the calculated rate constant at a fixed pressure (Chebyshev reactions).
 This conversion made the cost of reaction rate evaluation roughly equivalent between all reactions in model, separating the effect of chemical model size from computational intensities of different reaction types on the SIMD efficiency.
-Next, a greedy reaction removal algorithm (\cref{a:model_gen}) generated a number of models ranging from \numrange{2100}{186} reactions, in increments of 200 reactions (except the final increment from \num{200} to \num{186} reactions).}
+Next, a greedy reaction removal algorithm (\cref{a:model_gen}) generated a number of models ranging from \numrange{2100}{186} reactions, in increments of 200 reactions (except the final increment from \num{200} to \num{186} reactions).
 
@@
 
-\add{To discern the effect of varying chemical model size on the SIMD efficiency, shallow-vectorized and unvectorized source-term evaluation performance tests were run.
+To discern the effect of varying chemical model size on the SIMD efficiency, shallow-vectorized and unvectorized source-term evaluation performance tests were run.
 As demonstrated in~\cref{F:simd_vs_rxns} the SIMD efficiency strongly depends on the generated model size and thus the amount of computational work per thermochemical state.
@@
 However, the base isopentanol model achieved a SIMD efficiency of only \num{0.78} on the \avx/ machine in~\cref{S:source_results}, suggesting that more work could be done to optimize the source-term evaluations.
-In particular, it is likely that a reaction sorting method, such as suggested by Sewerin and Rigopoulos~\cite{Sewerin20151375}, would be particularly beneficial to reduce the number of vector gather\slash scatter\slash masking operations incurred during source-term evaluation.}
+In particular, it is likely that a reaction sorting method, such as suggested by Sewerin and Rigopoulos~\cite{Sewerin20151375}, would be particularly beneficial to reduce the number of vector gather\slash scatter\slash masking operations incurred during source-term evaluation.
 

### 2018-07-26 paper.tex
add notes for reviewer #1

@@
  \item Derive and verify a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity: \texttt{pyJac v2};
- \item Enable cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
+ \item \revisedelete{Enable}\add{Detail the implementation of} cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
  \item Investigate the performance for a wide range chemical kinetic models, specifically looking at speedups due to SIMD\slash SIMT-vectorization as well as relative to the previous version of \texttt{pyJac}; and finally
@@
 In addition, \texttt{loo.py} allows the user to more easily make changes to the structure of the generated program, e.g., the data ordering, vectorization, and threading patterns, as well as switch the target language for code generation (and more simply extend to additional languages, e.g., CUDA).
-Further, \texttt{loo.py} can execute developed subroutines from Python (natively for C code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL), enabling unit testing\slash verification for each component of the Jacobian or source terms.
+Further, \texttt{loo.py} can execute developed subroutines from Python (natively for C code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL), enabling unit testing\slash verification for each component of the Jacobian or source terms; \add{the unit testing suite also helps ensure that bugs are not present in less commonly used code-paths}.
 The source terms and sub-components thereof (e.g., rates of progress, pressure-modification terms) are directly compared with Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{hogan2014fast,adept-v11} provides reference values for Jacobian sub-components.
@@
 Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change requires building of the wrapper that initializes\slash transfers memory and calls the source-rate\slash Jacobian kernel.
+
+\add{These issues make it critical to provide adequate implementation details (e.g., runtime version, platform, etc., as given in~\cref{s:test_platforms}) for codes utilizing OpenCL in order to enable reproducible results.}
 

### 2018-08-09 paper.tex
add 'reproducilby'
removed buggy / and unnecessary hyphen

@@
 Yonkee and Sutherland~\cite{Yonkee2016} implemented accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates on both the CPU and GPU, achieving speedups over serial evaluation between \SIrange{8}{13}{$\times$} on a 16-core CPU and \SIrange{20}{40}{$\times$} on the GPU.
-\In addition, \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} speedups were achieved for the simulation of a partially premixed methanol flame for solving partial differential equations (PDE) on 16 CPU cores and the GPU, respectively.
+In addition, \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} speedups were achieved for the simulation of a partially premixed methanol flame for solving partial differential equations (PDE) on 16 CPU cores and the GPU, respectively.
 Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} on the GPU and CPU.
@@
 Kroshko and Spiteri~\cite{kroshko2013efficient} implemented a SIMD-vectorized third-order stiff Rosenbrock integrator for atmospheric chemistry on the CBE and found a speedup of \SI{1.89}{$\times$} (a parallel scaling efficiency of \SI{94}{$\percent$}) over a serial version of the same code.
-Stone et al.~\cite{stone2018} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
+Stone et al.~\cite{stone2018} implemented a linearly implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
 They found that SIMD vectorization improves integrator performance over an OpenMP baseline vectorized by simple compiler hints (i.e., \texttt{\#pragmas}) by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performs only \SIrange{1.4}{1.6}{$\times$} faster than the OpenMP baseline due to thread divergence~\cite{stone2018}.
@@
 \begin{itemize}
- \item Derive and verify a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity: \texttt{pyJac v2};
+ \item Derive and \add{reproducibly} verify a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity: \texttt{pyJac v2};
  \item \revisedelete{Enable}\add{Detail the implementation of} cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;

### 2018-08-09 paper.tex
revise abstract to emphasize implementation

@@
 The large size and high levels of numerical stiffness typically present in chemical kinetic models relevant to transportation\slash power-generation applications make the efficient evaluation\slash factorization of the chemical kinetic Jacobian and thermochemical source-terms critical to the performance of reactive-flow codes.
-Here we present \texttt{pyJac}, a code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized constant-pressure\slash volume thermochemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation.
+\add{Here we investigate the performance of vectorized evaluation of constant-pressure/volume thermochemical source-term and sparse/dense chemical kinetic Jacobians using single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) paradigms.
+These are implemented in pyJac, an open-source, reproducible code generation platform.}
+\delete{Here we present \texttt{pyJac}, an open-source, a code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized constant-pressure\slash volume thermochemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation.}
 Selected chemical kinetic models covering the range of sizes typically used in reactive-flow simulations were used for demonstration.

### 2018-08-09 paper.tex
possible goals revision

@@
 
-This work builds upon our previous analytical chemical kinetic Jacobian code, \texttt{pyJac}~\cite{Niemeyer:2016aa}.
 In this article we
 \begin{itemize}
- \item Derive and \add{reproducibly} verify a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity: \texttt{pyJac v2};
+ \item Derive and \add{reproducibly} verify a new Jacobian formulation that greatly increases sparsity\revisedelete{: \texttt{pyJac v2}};
  \item \revisedelete{Enable}\add{Detail the implementation of} cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
- \item Investigate the performance for a wide range chemical kinetic models, specifically looking at speedups due to SIMD\slash SIMT-vectorization as well as relative to the previous version of \texttt{pyJac}; and finally
+ \item Investigate the performance for a wide range chemical kinetic models, specifically looking at speedups due to SIMD\slash SIMT-vectorization as well as relative to a \add{previous version of our analytical chemical kinetic Jacobian code~\cite{Niemeyer:2016aa}}; and finally
  \item Discuss future extensions to this work as well as several promising directions for SIMD\slash SIMT vectorization in reactive-flow simulations.
 \end{itemize}
-To our knowledge, this is the first open-source, verified effort that vectorizes the evaluation of chemical-kinetic source terms and Jacobian matrices for any chemical model on a wide selection of platforms.
+This work builds upon our previous analytical chemical kinetic Jacobian code, \texttt{pyJac}~\cite{Niemeyer:2016aa}\add{, using the new formulation, \texttt{pyJac v2}, to achieve these goals}.
+To our knowledge is the first open-source, verified effort that vectorizes the evaluation of chemical-kinetic source terms and Jacobian matrices for any chemical model on a wide selection of platforms.
 

### 2018-08-09 paper.tex
possible conclusion revisions

@@
 \begin{itemize}
- \item Deriving and verifying a new Jacobian formulation for \texttt{pyJac} that greatly increases sparsity;
- \item Extending \texttt{pyJac} to enable vectorized execution on the CPU, GPU, and other accelerators; and
+ \item Deriving and \add{reproducibly} verifying a new Jacobian formulation \revisedelete{for \texttt{pyJac}} that greatly increases sparsity;
+ \item \revisedelete{Extending \texttt{pyJac} to enable}\add{Enabling} vectorized execution on the CPU, GPU, and other accelerators; and
  \item Achieving significant speedups over a strictly parallel Jacobian and source-term evaluation on SIMD-enabled processors (e.g., CPUs).
@@
 
-The new molar-based formulation for \texttt{pyJac} resulted in highly sparse chemical kinetic Jacobians,
-and allows selection of either the constant-volume or constant-pressure approximation.
-In addition, sparsity can be increased further by eliminating components associated with the bath gas, as discussed in \cref{S:sparsity};
-this approximation is not a key feature of this work, and future efforts to develop more sophisticated Jacobian approximations based on the \texttt{pyJac} software would be a worthwhile endeavor.
+\add{These achievements are made publicly available (see~\cref{A:availability}) via the open-source, high-performance, chemical kinetics code \texttt{pyJac}}.
+The new molar-based formulation \revisedelete{for \texttt{pyJac}} resulted in highly sparse chemical kinetic Jacobians, and allows selection of either the constant-volume or constant-pressure approximation.
+In addition, sparsity can be increased further by eliminating components associated with the bath gas, as discussed in \cref{S:sparsity}; this approximation is not a key feature of this work, and future efforts to develop more sophisticated Jacobian approximations based on the \texttt{pyJac} software would be a worthwhile endeavor.
 

### 2018-08-09 paper.tex
more aggressive revisions of conclusions

@@
 \begin{itemize}
- \item Deriving and \add{reproducibly} verifying a new Jacobian formulation \revisedelete{for \texttt{pyJac}} that greatly increases sparsity;
- \item \revisedelete{Extending \texttt{pyJac} to enable}\add{Enabling} vectorized execution on the CPU, GPU, and other accelerators; and
+ \item Deriving and \add{reproducibly} verifying a new Jacobian formulation \delete{for \texttt{pyJac}} that greatly increases sparsity;
+ \item \delete{Extending \texttt{pyJac} to enable}\add{Enabling} vectorized execution on the CPU, GPU, and other accelerators; and
  \item Achieving significant speedups over a strictly parallel Jacobian and source-term evaluation on SIMD-enabled processors (e.g., CPUs).
@@
 \add{These achievements are made publicly available (see~\cref{A:availability}) via the open-source, high-performance, chemical kinetics code \texttt{pyJac}}.
-The new molar-based formulation \revisedelete{for \texttt{pyJac}} resulted in highly sparse chemical kinetic Jacobians, and allows selection of either the constant-volume or constant-pressure approximation.
-In addition, sparsity can be increased further by eliminating components associated with the bath gas, as discussed in \cref{S:sparsity}; this approximation is not a key feature of this work, and future efforts to develop more sophisticated Jacobian approximations based on the \texttt{pyJac} software would be a worthwhile endeavor.
+The new molar-based formulation \delete{for \texttt{pyJac}} resulted in highly sparse chemical kinetic Jacobians, and allows selection of either the constant-volume or constant-pressure approximation.
+In addition, sparsity can be increased further by eliminating components associated with the bath gas, as discussed in \cref{S:sparsity}; this approximation is not a key feature of this work, and future efforts to \revisedelete{develop}\revise{incorporate} more sophisticated Jacobian approximations\delete{ based on the \texttt{pyJac} software} would be a worthwhile endeavor.
 
 We also demonstrated source-term and Jacobian evaluation for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
-In addition to parallel OpenMP evaluation on the CPU, the new version of \texttt{pyJac} enables the shallow-vectorized evaluation of the chemical-kinetic source terms and analytical Jacobian on both the CPU and GPU via OpenCL.
+In addition to parallel OpenMP evaluation on the CPU, \revise{this work}\delete{the new version of \texttt{pyJac}} enabled the shallow-vectorized evaluation of the chemical-kinetic source terms and analytical Jacobian on both the CPU and GPU via OpenCL.
 Deep vectorization is possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, but it yields no performance benefit as POCL did not achieve vectorizations for any execution pattern studied.
@@
 
-The OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler \texttt{\#pragma}s) is a key priority going forwards since OpenMP is a standard library on most machines.
+The OpenMP code-generation \delete{built into \texttt{pyJac}} is currently only capable of parallel execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler \texttt{\#pragma}s) is a key priority going forwards since OpenMP is a standard library on most machines.
 In addition, CUDA~\cite{Nvidia:2018} has been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017}, while Intel's open-source OpenCL alternative, ISPC~\cite{pharr2012ispc}, has been relatively stable and easy to work with during preliminary efforts with the unit-testing discussed in~\cref{s:unittest}.
@@
 
-One key component missing in the current version of \texttt{pyJac} is vectorized sparse\slash dense linear-algebra subroutines to maximize the performance of LU-factorization and matrix-vector multiplication (commonly used in implicit-integration techniques).
+One key component missing in \add{this work}\delete{ the current version of \texttt{pyJac}} is vectorized sparse\slash dense linear-algebra subroutines to maximize the performance of LU-factorization and matrix-vector multiplication (commonly used in implicit-integration techniques).
 Third-party\slash open-source options exist for some target languages, e.g., cuBLAS~\cite{cublas}, clBLAS\slash clSPARSE~\cite{clmath} or SuperLU~\cite{superlu99}, but these do not necessarily cover all targets\slash required linear-algebra operations and, in the case of the CUDA\slash OpenCL libraries, are often optimized operation on one large matrix instead of many (relatively) smaller matrices.

### 2018-08-09 paper.tex
kyle's feedback on goals revision

@@
  \item \revisedelete{Enable}\add{Detail the implementation of} cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
- \item Investigate the performance for a wide range chemical kinetic models, specifically looking at speedups due to SIMD\slash SIMT-vectorization as well as relative to a \add{previous version of our analytical chemical kinetic Jacobian code~\cite{Niemeyer:2016aa}}; and finally
+ \item \revise{Investigate the performance of SIMD\slash SIMT-vectorization for a wide of range chemical kinetic models, and compare with the previous version of our analytical chemical kinetic Jacobian code~\cite{Niemeyer:2016aa}}; and finally}
  \item Discuss future extensions to this work as well as several promising directions for SIMD\slash SIMT vectorization in reactive-flow simulations.

### 2018-08-09 paper.tex
s/achievements/efforts

@@
 
-\add{These achievements are made publicly available (see~\cref{A:availability}) via the open-source, high-performance, chemical kinetics code \texttt{pyJac}}.
+\add{These efforts are made publicly available (see~\cref{A:availability}) via the open-source, high-performance, chemical kinetics code \texttt{pyJac}}.
 The new molar-based formulation \delete{for \texttt{pyJac}} resulted in highly sparse chemical kinetic Jacobians, and allows selection of either the constant-volume or constant-pressure approximation.

### 2018-08-09 paper.tex
fixed error, and replace revises with adds

@@
 \begin{itemize}
- \item Derive and \add{reproducibly} verify a new Jacobian formulation that greatly increases sparsity\revisedelete{: \texttt{pyJac v2}};
- \item \revisedelete{Enable}\add{Detail the implementation of} cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
- \item \revise{Investigate the performance of SIMD\slash SIMT-vectorization for a wide of range chemical kinetic models, and compare with the previous version of our analytical chemical kinetic Jacobian code~\cite{Niemeyer:2016aa}}; and finally}
+ \item Derive and \add{reproducibly} verify a new Jacobian formulation that greatly increases sparsity\delete{: \texttt{pyJac v2}};
+ \item \delete{Enable}\add{Detail the implementation of} cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
+ \item \add{Investigate the performance of SIMD\slash SIMT-vectorization for a wide of range chemical kinetic models, and compare with the previous version of our analytical chemical kinetic Jacobian code~\cite{Niemeyer:2016aa}; and finally}
  \item Discuss future extensions to this work as well as several promising directions for SIMD\slash SIMT vectorization in reactive-flow simulations.
@@
 The new molar-based formulation \delete{for \texttt{pyJac}} resulted in highly sparse chemical kinetic Jacobians, and allows selection of either the constant-volume or constant-pressure approximation.
-In addition, sparsity can be increased further by eliminating components associated with the bath gas, as discussed in \cref{S:sparsity}; this approximation is not a key feature of this work, and future efforts to \revisedelete{develop}\revise{incorporate} more sophisticated Jacobian approximations\delete{ based on the \texttt{pyJac} software} would be a worthwhile endeavor.
+In addition, sparsity can be increased further by eliminating components associated with the bath gas, as discussed in \cref{S:sparsity}; this approximation is not a key feature of this work, and future efforts to \delete{develop}\add{incorporate} more sophisticated Jacobian approximations\delete{ based on the \texttt{pyJac} software} would be a worthwhile endeavor.
 
 We also demonstrated source-term and Jacobian evaluation for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
-In addition to parallel OpenMP evaluation on the CPU, \revise{this work}\delete{the new version of \texttt{pyJac}} enabled the shallow-vectorized evaluation of the chemical-kinetic source terms and analytical Jacobian on both the CPU and GPU via OpenCL.
+In addition to parallel OpenMP evaluation on the CPU, \add{this work}\delete{the new version of \texttt{pyJac}} enabled the shallow-vectorized evaluation of the chemical-kinetic source terms and analytical Jacobian on both the CPU and GPU via OpenCL.
 Deep vectorization is possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, but it yields no performance benefit as POCL did not achieve vectorizations for any execution pattern studied.

### 2018-08-09 paper.tex
made all section titles sentence-case

@@
 
-\subsection{Thermochemical Source Terms and Jacobian}
+\subsection{Thermochemical source terms and Jacobian}
 This new version of \texttt{pyJac} is capable of evaluating the thermochemical source-terms for using the constant-pressure (\conp/) or constant-volume (\conv/) assumption\footnote{Note: in this context, the ``constant-pressure'' and ``constant-volume'' assumptions refer to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure or constant-volume reactive-flow simulation.}
@@
 
-\subsection{Code Generation and Testing Infrastructure}
+\subsection{Code generation and testing infrastructure}
 \label{s:unittest}
@@
 
-\section{Results and Discussion}
+\section{Results and discussion}
 \subsection{Testing platforms}
@@
 
-\section{Practical notes on OpenCL Usage}
+\section{Practical notes on OpenCL use}
 \label{S:opencl}

### 2018-08-10 paper.tex
add note about future code changes

@@
 In addition, \texttt{loo.py} allows the user to more easily make changes to the structure of the generated program, e.g., the data ordering, vectorization, and threading patterns, as well as switch the target language for code generation (and more simply extend to additional languages, e.g., CUDA).
-Further, \texttt{loo.py} can execute developed subroutines from Python (natively for C code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL), enabling unit testing\slash verification for each component of the Jacobian or source terms; \add{the unit testing suite also helps ensure that bugs are not present in less commonly used code-paths}.
+Further, \texttt{loo.py} can execute developed subroutines from Python (natively for C code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL), enabling unit testing\slash verification for each component of the Jacobian or source terms; \add{the unit testing suite also helps ensure that bugs are not present in less commonly used code-paths, or introduced by future code changes}.
 The source terms and sub-components thereof (e.g., rates of progress, pressure-modification terms) are directly compared with Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{hogan2014fast,adept-v11} provides reference values for Jacobian sub-components.

### 2018-08-12 paper.tex
dr. sung edits

--- 2018-08-09 paper.tex
@@
 \begin{itemize}
- \item Derive and \add{reproducibly} verify a new Jacobian formulation that greatly increases sparsity\delete{: \texttt{pyJac v2}};
+ \item Derive and verify a new Jacobian formulation that greatly increases sparsity\delete{: \texttt{pyJac v2}};
  \item \delete{Enable}\add{Detail the implementation of} cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
- \item \add{Investigate the performance of SIMD\slash SIMT-vectorization for a wide of range chemical kinetic models, and compare with the previous version of our analytical chemical kinetic Jacobian code~\cite{Niemeyer:2016aa}; and finally}
+ \item \add{Investigate the performance of SIMD\slash SIMT-vectorization for a wide range of chemical kinetic models, and compare with the previous version of our analytical chemical kinetic Jacobian code~\cite{Niemeyer:2016aa}; and finally}
  \item Discuss future extensions to this work as well as several promising directions for SIMD\slash SIMT vectorization in reactive-flow simulations.
@@
 In addition, \texttt{loo.py} allows the user to more easily make changes to the structure of the generated program, e.g., the data ordering, vectorization, and threading patterns, as well as switch the target language for code generation (and more simply extend to additional languages, e.g., CUDA).
-Further, \texttt{loo.py} can execute developed subroutines from Python (natively for C code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL), enabling unit testing\slash verification for each component of the Jacobian or source terms; \add{the unit testing suite also helps ensure that bugs are not present in less commonly used code-paths}.
+Further, \texttt{loo.py} can execute developed subroutines from Python (natively for C code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL), enabling unit testing\slash verification for each component of the Jacobian or source terms; \add{the unit testing suite also helps ensure that bugs are not present in less commonly used code-paths, or are introduced by future code changes}.
 The source terms and sub-components thereof (e.g., rates of progress, pressure-modification terms) are directly compared with Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{hogan2014fast,adept-v11} provides reference values for Jacobian sub-components.
@@
 Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change requires building of the wrapper that initializes\slash transfers memory and calls the source-rate\slash Jacobian kernel.
-
 \add{These issues make it critical to provide adequate implementation details (e.g., runtime version, platform, etc., as given in~\cref{s:test_platforms}) for codes utilizing OpenCL in order to enable reproducible results.}
@@
 \begin{itemize}
- \item Deriving and \add{reproducibly} verifying a new Jacobian formulation \delete{for \texttt{pyJac}} that greatly increases sparsity;
+ \item Deriving and verifying a new Jacobian formulation \delete{for \texttt{pyJac}} that greatly increases sparsity;
  \item \delete{Extending \texttt{pyJac} to enable}\add{Enabling} vectorized execution on the CPU, GPU, and other accelerators; and

### 2018-08-13 paper.tex
remove review markers to avoid spacing issues

@@
 \newtoggle{markup}
-%\togglefalse{markup}
-\toggletrue{markup}
+\togglefalse{markup}
+%\toggletrue{markup}
 
@@
 The large size and high levels of numerical stiffness typically present in chemical kinetic models relevant to transportation\slash power-generation applications make the efficient evaluation\slash factorization of the chemical kinetic Jacobian and thermochemical source-terms critical to the performance of reactive-flow codes.
-\add{Here we investigate the performance of vectorized evaluation of constant-pressure/volume thermochemical source-term and sparse/dense chemical kinetic Jacobians using single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) paradigms.
-These are implemented in pyJac, an open-source, reproducible code generation platform.}
-\delete{Here we present \texttt{pyJac}, an open-source, a code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized constant-pressure\slash volume thermochemical source-term and sparse\slash dense chemical kinetic Jacobian evaluation.}
+Here we investigate the performance of vectorized evaluation of constant-pressure/volume thermochemical source-term and sparse/dense chemical kinetic Jacobians using single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) paradigms.
+These are implemented in pyJac, an open-source, reproducible code generation platform.
 Selected chemical kinetic models covering the range of sizes typically used in reactive-flow simulations were used for demonstration.
@@
  \item Derive and verify a new Jacobian formulation that greatly increases sparsity\delete{: \texttt{pyJac v2}};
- \item \delete{Enable}\add{Detail the implementation of} cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
- \item \add{Investigate the performance of SIMD\slash SIMT-vectorization for a wide range of chemical kinetic models, and compare with the previous version of our analytical chemical kinetic Jacobian code~\cite{Niemeyer:2016aa}; and finally}
+ \item Detail the implementation of cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
+ \item Investigate the performance of SIMD\slash SIMT-vectorization for a wide range of chemical kinetic models, and compare with the previous version of our analytical chemical kinetic Jacobian code~\cite{Niemeyer:2016aa}; and finally
  \item Discuss future extensions to this work as well as several promising directions for SIMD\slash SIMT vectorization in reactive-flow simulations.
 \end{itemize}
-This work builds upon our previous analytical chemical kinetic Jacobian code, \texttt{pyJac}~\cite{Niemeyer:2016aa}\add{, using the new formulation, \texttt{pyJac v2}, to achieve these goals}.
+This work builds upon our previous analytical chemical kinetic Jacobian code, \texttt{pyJac}~\cite{Niemeyer:2016aa}, using the new formulation, \texttt{pyJac v2}, to achieve these goals.
 To our knowledge is the first open-source, verified effort that vectorizes the evaluation of chemical-kinetic source terms and Jacobian matrices for any chemical model on a wide selection of platforms.
@@
 In addition, \texttt{loo.py} allows the user to more easily make changes to the structure of the generated program, e.g., the data ordering, vectorization, and threading patterns, as well as switch the target language for code generation (and more simply extend to additional languages, e.g., CUDA).
-Further, \texttt{loo.py} can execute developed subroutines from Python (natively for C code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL), enabling unit testing\slash verification for each component of the Jacobian or source terms; \add{the unit testing suite also helps ensure that bugs are not present in less commonly used code-paths, or are introduced by future code changes}.
+Further, \texttt{loo.py} can execute developed subroutines from Python (natively for C code, or via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL), enabling unit testing\slash verification for each component of the Jacobian or source terms; the unit testing suite also helps ensure that bugs are not present in less commonly used code-paths, or are introduced by future code changes.
 The source terms and sub-components thereof (e.g., rates of progress, pressure-modification terms) are directly compared with Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{hogan2014fast,adept-v11} provides reference values for Jacobian sub-components.
@@
 Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change requires building of the wrapper that initializes\slash transfers memory and calls the source-rate\slash Jacobian kernel.
-\add{These issues make it critical to provide adequate implementation details (e.g., runtime version, platform, etc., as given in~\cref{s:test_platforms}) for codes utilizing OpenCL in order to enable reproducible results.}
+These issues make it critical to provide adequate implementation details (e.g., runtime version, platform, etc., as given in~\cref{s:test_platforms}) for codes utilizing OpenCL in order to enable reproducible results.
 
@@
 \begin{itemize}
- \item Deriving and verifying a new Jacobian formulation \delete{for \texttt{pyJac}} that greatly increases sparsity;
- \item \delete{Extending \texttt{pyJac} to enable}\add{Enabling} vectorized execution on the CPU, GPU, and other accelerators; and
+ \item Deriving and verifying a new Jacobian formulation that greatly increases sparsity;
+ \item Enabling vectorized execution on the CPU, GPU, and other accelerators; and
  \item Achieving significant speedups over a strictly parallel Jacobian and source-term evaluation on SIMD-enabled processors (e.g., CPUs).
@@
 
-\add{These efforts are made publicly available (see~\cref{A:availability}) via the open-source, high-performance, chemical kinetics code \texttt{pyJac}}.
-The new molar-based formulation \delete{for \texttt{pyJac}} resulted in highly sparse chemical kinetic Jacobians, and allows selection of either the constant-volume or constant-pressure approximation.
-In addition, sparsity can be increased further by eliminating components associated with the bath gas, as discussed in \cref{S:sparsity}; this approximation is not a key feature of this work, and future efforts to \delete{develop}\add{incorporate} more sophisticated Jacobian approximations\delete{ based on the \texttt{pyJac} software} would be a worthwhile endeavor.
+These efforts are made publicly available (see~\cref{A:availability}) via the open-source, high-performance, chemical kinetics code \texttt{pyJac}.
+The new molar-based formulation resulted in highly sparse chemical kinetic Jacobians, and allows selection of either the constant-volume or constant-pressure approximation.
+In addition, sparsity can be increased further by eliminating components associated with the bath gas, as discussed in \cref{S:sparsity}; this approximation is not a key feature of this work, and future efforts to incorporate more sophisticated Jacobian approximations would be a worthwhile endeavor.
 
 We also demonstrated source-term and Jacobian evaluation for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
-In addition to parallel OpenMP evaluation on the CPU, \add{this work}\delete{the new version of \texttt{pyJac}} enabled the shallow-vectorized evaluation of the chemical-kinetic source terms and analytical Jacobian on both the CPU and GPU via OpenCL.
+In addition to parallel OpenMP evaluation on the CPU, this work enabled the shallow-vectorized evaluation of the chemical-kinetic source terms and analytical Jacobian on both the CPU and GPU via OpenCL.
 Deep vectorization is possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, but it yields no performance benefit as POCL did not achieve vectorizations for any execution pattern studied.
@@
 
-The OpenMP code-generation \delete{built into \texttt{pyJac}} is currently only capable of parallel execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler \texttt{\#pragma}s) is a key priority going forwards since OpenMP is a standard library on most machines.
+The OpenMP code-generation is currently only capable of parallel execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler \texttt{\#pragma}s) is a key priority going forwards since OpenMP is a standard library on most machines.
 In addition, CUDA~\cite{Nvidia:2018} has been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017}, while Intel's open-source OpenCL alternative, ISPC~\cite{pharr2012ispc}, has been relatively stable and easy to work with during preliminary efforts with the unit-testing discussed in~\cref{s:unittest}.
@@
 
-One key component missing in \add{this work}\delete{ the current version of \texttt{pyJac}} is vectorized sparse\slash dense linear-algebra subroutines to maximize the performance of LU-factorization and matrix-vector multiplication (commonly used in implicit-integration techniques).
+One key component missing in this work is vectorized sparse\slash dense linear-algebra subroutines to maximize the performance of LU-factorization and matrix-vector multiplication (commonly used in implicit-integration techniques).
 Third-party\slash open-source options exist for some target languages, e.g., cuBLAS~\cite{cublas}, clBLAS\slash clSPARSE~\cite{clmath} or SuperLU~\cite{superlu99}, but these do not necessarily cover all targets\slash required linear-algebra operations and, in the case of the CUDA\slash OpenCL libraries, are often optimized operation on one large matrix instead of many (relatively) smaller matrices.

### 2018-09-04 paper.tex
removed missed delete mark
remove duplicate

@@
 \begin{itemize}
- \item Derive and verify a new Jacobian formulation that greatly increases sparsity\delete{: \texttt{pyJac v2}};
+ \item Derive and verify a new Jacobian formulation that greatly increases sparsity;
  \item Detail the implementation of cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
@@
 The optimization level ``\texttt{-O3 -mtune=native}'' was used and no ``fast math'' OpenCL optimizations were enabled.
-Additionally, the exact form of the Jacobian (as opposed the the ``approximate'' form discussed in~\cref{S:sparsity}) was used in all cases.
+Additionally, the exact form of the Jacobian (as opposed the ``approximate'' form discussed in~\cref{S:sparsity}) was used in all cases.
 Finally, unless stated otherwise: the performance results used a single CPU core, the \conp/ assumption, a vector width of \num{8}\slash\num{128}, and ``C''\slash ``F''-ordered data for the CPU\slash GPU cases, respectively; the run times reported are for the number of conditions specified in~\cref{T:error} and include data-transfer overhead to\slash from internal buffers used in \texttt{pyJac}.

### 2018-09-04 derivations.tex
remove dupl

@@
 
-Derivatives of the falloff blending factor---except for for all but the Lindemann falloff blending factor, which is not a function of the reduced pressure---rely upon the reduced pressure derivatives developed in~\cref{s:dpr}.
+Derivatives of the falloff blending factor---except for all but the Lindemann falloff blending factor, which is not a function of the reduced pressure---rely upon the reduced pressure derivatives developed in~\cref{s:dpr}.
 First, the blending factor for each falloff reaction type will be differentiated along with any subcomponents (e.g., the $F_{cent}$ term for Troe falloff reactions).
